### 2015-09-04 Manuscript/Yeakel_Redner_2015.tex
manuscript files moved here

\documentclass[11pt]{article}
\usepackage[margin=1in,footskip=0.25in]{geometry}

%\usepackage{helvet}
%\renewcommand{\familydefault}{\sfdefault}
\renewcommand\refname{\vskip -1cm}
%\renewcommand{\rmdefault}{phv} % Arial
%\renewcommand{\sfdefault}{phv} % Arial
\usepackage{setspace}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage{bm}
\usepackage{wasysym}
\usepackage{placeins}
\usepackage{multirow}
\usepackage[T1]{fontenc}
\usepackage[super]{natbib}
\usepackage{framed}
\usepackage{caption}
\usepackage{longtable}

\begin{document}

\title{The effect of starvation on the dynamics of consumer populations}
\maketitle

\section{Introduction}
%Focus on the tradeoff between REPRODUCTION AND SURVIVAL AS A FUNCTION OF ENERGETIC STATE

The behaviorial ecology of most, if not all, organisms is influenced by the energetic state of individuals.
An individual's energetic state, which can range from starving (energetically defficient) to satiated (energetially replete), directly influences how it invests its stores in an uncertain environment.
Such state-dependent behaviors are generally manifested as trade-offs.
A major trade-off that we consider in detail concerns reproductive vs. non-reproductive energetic investment strategies.
As has been shown in [examples], individuals that are energetically deficient are unable to invest energy stores towards reproduction.
Among sexually reproducing species, even when reproduction occurs, the offspring is often naturally aborted if there is not sufficient caloric resources available to the parent.

Here we explore how reproductive trade-offs, which occur at the level of the individual, may influence the dynamics of a population of individuals.
We first establish a simple stage-structured population model that captures the essential elements of energetic reproductive tradeoffs, and explore how the rate of starvation impacts dynamics at the level of the population.
We then develop a more general model in order to understand if there are common attriburtes of these dynamics that have implications beyond the simple stage-structured model.
Finally, we compare our population-dynamic approach with a stochastic dynamic program that incorporates different individual-level behaviors as a function of energetic state.

\section{Methods}
\subsection{Model description}

We first integrate energetics into the dynamics of a consumer-resource system by assuming that the consumer population can be divided into discrete energetic states, the occupation of each being contingent on the consumption of a single resource $R$.
Throughout, we assume that the resource has logistic growth with a linear growth rate $\alpha$, a carrying capacity of unity, and an extrinsic mortality due to consumption at rate $\epsilon$.
In the simplest case, there are only two energetic states for the consumer population: \emph{i}) an energetically replete state $Y$, where the consumer reproduces at rate $\delta \epsilon$ and suffers mortality at rate $\mu_y$, and \emph{ii}) an energetically deficient state $X$, where reproduction is suppressed, and mortality occurs at rate $\mu_x$.
Because an energetically defficient state is generally more risky, we assume that $\mu_x > \mu_y$.
Consumers transition from state $Y$ to state $X$ by starvation at rate $p$, which is proportional to $1-R$.
Conversely, consumers recover from state $X$ to the energetically replete state $Y$ in proportion to the density of resources consumed $\epsilon R$.
Accordingly, the system of equations is written

\begin{align}
\frac{\rm d}{\rm dt} R &= \alpha R(1-R) - \epsilon R(X+Y), \nonumber \\ \nonumber \\
\frac{\rm d}{\rm dt} X &= \sigma Y(1-R) - \rho \epsilon XR - \mu X, \nonumber \\ \nonumber \\
\frac{\rm d}{\rm dt} Y &= \gamma \epsilon YR + \rho \epsilon XR - \sigma Y(1-R)
\end{align}

There are four steady states for the 2-stage consumer-resource system: two trivial steady states at $(R^*=0,X^*=0,Y^*=0)$ and $(R^*=1,X^*=0,Y^*=0)$, one non-trivial external (negative) steady state, and one non-trivial internal steady state where $(R^*>0,X^*>0,Y^*>0)$.
The latter steady state is the one of interest here, and although there is a tractible solution, it is cannot be written in a compressed form.
Moreover, because there is only one internal steady state, as long as it is stable the population trajectories will be globally attracted to it for any set of initial conditions greater than zero.

Analysis of the stability of the consumer-resource system is explored with respect to the local stability of the internal steady state, which is the only feasible steady state as long as both the consumer and resource have non-zero, positive, values.
In a multidimensional system, linear stability is determined with respect to the Jacobian Matrix $\bf J$, which is a matrix where each element is defined by the partial derivative of each equation with respect to each variable.
In the case of the 2-stage consumer model, the Jacobian is written

\begin{equation}
\mathbf{J}|_* = 
\left(
\begin{array}{ccc}
\alpha(1-2 R^*)-\epsilon (X^*-Y^*) & -\epsilon R^* & -\epsilon R^* \\
-\rho \epsilon X^*-\sigma Y^* & -\epsilon \rho R^*-\mu & \sigma(1-R^*) \\
\gamma \epsilon Y^*+\sigma Y^* +\rho \epsilon X^* & \rho \epsilon R^* & \gamma \epsilon R^* - \sigma(1-R^*) \\
\end{array}
\right).
\end{equation}

Whether the internal steady state is attracting or repelling is conditioned on the values of the parameters in the system.
If the parameters of the Jacobian matrix (at the internal steady state) are such that its eigenvalues have negative real values, then it is stable.
The stability of the system changes if a Saddle-Node bifurcation is crossed, such that a single eigenvalue obtains a positive real part, defined by the condition ${\rm Det}({\bf J})=0$.
A single Saddle-Node bifurcation always exists for $\alpha = 0$, such that $\alpha > 0$ for the internal state to be stable.
%Although a closed-form solution is not possible, a second stability condition generally requires $\delta > a \exp \{\frac{b}{c+\epsilon}\}$ assuming low values of $\mu_x$ and $\mu_y$ (where $a$, $b$, and $c$ are some combination of the model parameters), which simplifies to $\delta > \frac{\mu_y}{\epsilon}$ when $p=0$.
%This condition places a lower bound on the consumer's rate of growth as a function of its rate of predation on the resource.
%As the rate of consumption decreases, system is more likely to become unstable if the consumer's rate of reproduction declines.

Of more interest here is the existence of parameter regions that permit the existence of cyclic behavior.
Cycles arise when a pair of complex conjugate eigenvalues cross the imaginary axis and attain positive real parts.
This condition is called a Hopf bifurcation, and is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian matrix.
Although the Hopf condition cannot be solved analytically for the specific 2-stage model, it can be explored numerically.

\subsection{Generalized model}

$\xi$ ~~ $\zeta$

\subsection{Stochastic dynamic program}

\section{Results}

To determine how the consumer-resource system impacted by different rates of starvation, we analyze the systems with respect to $p$.

\section{Discussion}

%The reproductive tradeoff as a strategy

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{fig_Hopf.pdf}
	\caption{
	Saddle Node bifurcation
	}
	\label{SN}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
	\caption{
	Probability that the resource value is less than threshold = 0.05
	}
	\label{SN}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
	\caption{
	Hopf bifurcation
	}
	\label{SN}
\end{figure}

\end{document}

### 2015-09-04 Manuscript/version1.tex
manuscript files moved here

\documentclass[11pt]{iopart}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{cite}
\usepackage{float}
\makeatletter
\makeatother

\expandafter\let\csname equation*\endcsname\relax
\expandafter\let\csname endequation*\endcsname\relax
\usepackage{amsmath}
\newcommand{\agt}{\mathrel{\raise.3ex\hbox{$>$\kern-.75em\lower1ex\hbox{$\sim$}}}}
\begin{document}

\title{Role of Starvation on Foraging Dynamics}

\author{J. D. Yeakel} \address{Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, New
  Mexico 87501, USA}

\author{S. Redner} \address{Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, New
  Mexico 87501, USA}

\begin{abstract}

  We introduce a minimal model of foraging in which the interplay between
  reproduction and starvation of the foragers, and the regeneration of
  resources, controls the dynamics.  The system consists of a resource that
  renews by logistic growth, as well as two classes of foragers---satiated
  and starving.  Satiated foragers reproduce at a fixed rate and are not
  vulnerable to mortality.  However, a satiated forager can starve when
  resources are scarce; conversely, a starving forager can become satiated
  when the resource is abundant.  Starving foragers do not reproduce and they
  die at a rate $\mu$.  We investigate the dynamics of this system, both in
  the well-mixed (mean-field) limit and by stochastic simulations in two
  dimensions.  For parameters that correspond to a benign environment, the
  system settles into a steady state.  However, the forager populations
  oscillate in time with an increasing amplitude as the environment becomes
  more harsh.  {\bf needs closing sentence(s).}

\end{abstract}

%\maketitle
\section{Introduction}
%Focus on the tradeoff between REPRODUCTION AND SURVIVAL AS A FUNCTION OF ENERGETIC STATE

Foraging for food (resources) is an essential task of many living
organisms~\cite{B91,OB90,KM01,ASD97,KR85,SK86,Vea96,LKW88,BLMV11}.  Our focus
here is on foraging in an environment where the consumed resource can
regenerate, and with the new feature that the energetic state of the foragers
affects their behavior.  This state---either starving (energetically
deficient) or satiated (energetically replete)---determines how the forager
behaves in this time-dependent environment.  We will specifically investigate
the influence of a forager's energetic state and its reproductive behavior on
the long-time dynamics of the population.  As has been shown in [examples],
individuals that are energetically deficient are unlikely to engage is
reproduction.  Even when such individuals do reproduce, the offspring are
often naturally aborted if there is insufficient caloric resources available
to the parent.  On the other hand, satiated foragers do reproduce, which
thereby exerts increased pressure on resources.  In turn, this can lead to a
population collapse when the resources are subsequently depleted by the
earlier growing population.  Our goal is to understand this interplay between
the foragers' energetic state and environmental resources on the population
dynamics.

In our modeling, foragers look for food by wandering in this changing
environment.  If such a search is successful, the forager is satiated and it
can engage in the essential activity of reproduction.  However, if the search
is unsuccessful for a sufficiently long period, the forager ``starves''.
Such a forager can do nothing else but forage, until it either finds food and
again becomes satiated or it dies when it goes too long without finding
nourishment.  These rules are reminiscent of the ``starving random walk''
model, where a single random walk can take $\mathcal{S}$ steps without
encountering food before starving to death.  Moreover, the resource does not
regenerate, so that the forager ultimately starves to death.  For this
idealized model, it was found that the average lifetime of the forager scales
algebraically with $\mathcal{S}$ in $d\!=\!1$ and $d\!=\!2$ dimensions, and
as $\exp(-A\mathcal{S}^\omega)$ for $d>2$.  Here the exponent
$\omega\approx \frac{1}{2}$ for $d=3$, while $\omega\to 1$ only as
$d\to \infty$, with the latter behavior corresponding to the mean-field
limit.  As we will discuss, regeneration of the resource, together with the
behavioral change between starving and satiated foragers leads to still much
richer dynamical behavior.

We will begin by defining our stage-structured population model that
encapsulates the trade-off between regeneration, reproduction, and
starvation.  We will first explore how the rates of these processes impact
the population dynamics in the mean-field limit, where resources and foragers
are always perfectly mixed.  Under this perfect mixing condition, the species
densities can be described by rate equations, and we will investigate the
dynamical behaviors that arise from these equations.  We will also compare
this continuous rate-equation approach with stochastic simulations that
account for discreteness effects due to the finiteness of the population.
Next, we will extend these stochastic simulations to the physical situation
where the foragers perform random walks in two dimensions.  Intriguing
spatial organization effects arise, in which waves of satiated foragers
reproduce and deplete the environment, followed by satiated foragers
subsequently become starving foragers, many of whom eventually die.  This
allows for a later recovery of the environment and ultimately to repeated
cycles of boom and bust.  As we shall discuss, this quasi-cyclic behavior
arises when the environment is ``stressed'', because the consumption and
reproduction rates are high.  Conversely, for smaller environmental stress,
the system converges to a steady state.

Finally, we will develop a more general model to understand if there are
common attributes of the dynamics outlined above that have implications
beyond our simple stage-structured model.

\section{The Model}

We incorporate energetics into the model dynamics by assuming that the
foragers can exist in two discrete energetic states---satiated and starving.
Satiated (or full) foragers $F$ are those that have just encountered and
consumed a unit of resource $R$.  On the other hand, a satiated forager that
does not encounter a resource as it wanders is converted to a starving
forager $S$ with rate $\sigma$.  Whenever a forager, either satiating or
starving, encounters resources, one unit of the resource is consumed.  If the
forager was starving, it is recruited into the satiated population with rate
$\rho$.  During the time that a forager is starving, it dies with a fixed
mortality rate $\mu$, while satiated foragers do not experience mortality
risk.  Furthermore, satiated foragers reproduce with rate $\lambda$.
Finally, we assume that, in the absence of foragers, the underlying resource
undergoes logistic growth, with growth rate $\alpha$ and carrying capacity
equal to one.

According to these processes and also under the assumption that the densities
of satiated foragers, starving foragers, and resources (also denoted by $F$,
$S$, and $R$, respectively) are perfectly mixed, they evolve according to the
rate equations:
\begin{align}
  \label{RE}
\begin{split}
\dot F &= \lambda F + \rho  RS - \sigma (1-R)F\,,\\
\dot S &= \sigma (1-R)F - \rho RS - \mu S\,, \\
\dot R &= \alpha R(1-R) -  R(F+S),\\
\end{split}
\end{align}
where the overdot denotes time derivative.

This system of equations admits four fixed points.  Two of them, at
$(F^*\!=\!0,S^*\!=\!0,R^*\!=\!0)$ and $(F^*\!=\!0,S^*\!=\!0,R^*\!=\!1)$
correspond to an environment without foragers.  In this case, the latter
fixed point is stable.  There is also an unphysical fixed point at negative
densities and a non-trivial fixed point in which all populations are
represented, $(F^*,S^*,R^*\!>\!0)$, corresponding to an ecologically active
steady state.  If this unique steady state is stable, the species populations
will be globally attracted to it for any non-zero initial conditions.

To determine the stability of the non-trivial fixed point, we compute the
Jacobian Matrix $\bf J$, whose elements are given by
$J_{ij}=\partial X_i/\partial X_j$, where $X_i$, with $i=1,2,3$, refers to
the densities $F,S,R$, respectively.  At the non-trivial fixed point, the
Jacobian matrix is
\begin{equation}
\mathbf{J}|_* = 
\left(
\begin{array}{lll}
  \lambda  - \sigma(1-R^*)  & \rho R^* & \sigma(S^*+ S^*) \\
  \sigma(1-R^*) & - \rho R^*-\mu &   -\rho  (S^*+ F^*) \\
  - R^* & - R^* & \alpha(1-2 R^*)-(S^*+F^*) \\
\end{array}
\right).
\end{equation}

Whether the internal steady state is attracting or repelling is conditioned
on the values of the parameters in the system.  If the parameters of the
Jacobian matrix (at the internal steady state) are such that its eigenvalues
have negative real values, then it is stable.  The stability of the system
changes if a Saddle-Node bifurcation is crossed, such that a single
eigenvalue obtains a positive real part, defined by the condition
${\rm Det}({\bf J})=0$.  A single Saddle-Node bifurcation always exists for
$\alpha = 0$, such that $\alpha > 0$ for the internal state to be stable.
% Although a closed-form solution is not possible, a second stability
% condition generally requires $\delta > a \exp \{\frac{b}{c+\epsilon}\}$
% assuming low values of $\mu_x$ and $\mu_y$ (where $a$, $b$, and $c$ are
% some combination of the model parameters), which simplifies to
% $\delta > \frac{\mu_y}{\epsilon}$ when $p=0$.  This condition places a
% lower bound on the consumer's rate of growth as a function of its rate of
% predation on the resource.  As the rate of consumption decreases, system is
% more likely to become unstable if the consumer's rate of reproduction
% declines.

Of more interest here is the existence of parameter regions that permit the
existence of cyclic behavior.  Cycles arise when a pair of complex conjugate
eigenvalues cross the imaginary axis and attain positive real parts.  This
condition is called a Hopf bifurcation, and is defined by
${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is
composed of the coefficients of the characteristic polynomial describing the
Jacobian matrix.  Although the Hopf condition cannot be solved analytically
for the specific 2-stage model, it can be explored numerically.

\subsection{Generalized model}

$\xi$ ~~ $\zeta$

\subsection{Stochastic dynamic program}

\section{Results}

To determine how the consumer-resource system impacted by different rates of starvation, we analyze the systems with respect to $p$.

\section{Discussion}

%The reproductive tradeoff as a strategy

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{fig_Hopf.pdf}
	\caption{
	Saddle Node bifurcation
	}
	\label{SN}
\end{figure}

% \begin{figure}[h]
% 	\centering
% 	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
% 	\caption{
% 	Probability that the resource value is less than threshold = 0.05
% 	}
% 	\label{SN}
% \end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
	\caption{
	Hopf bifurcation
	}
	\label{SN1}
\end{figure}

JY thanks......  SR thanks NSF Grant No.\ DMR-1205797 (SR) and a grant from
the John S. Templeton Foundation) for partial support of this research.

\begin{thebibliography}{99}

\bibitem{B91} J. W. Bell, Searching Behaviour, the Behavioural Ecology of
  Finding Resources, Animal Behaviour Series (Chapman and Hall, London,
  1991).

\bibitem{OB90} W. J. O'Brien, H. I. Browman, and B. I. Evans, Am.\ Sci.\ {\bf
    78}, 152 (1990).

\bibitem{KM01} L. D. Kramer and R. L. McLaughlin, Am.\ Zool.\ {\bf 41}, 137 (2001).

\bibitem{ASD97} J. P. Anderson, D. W. Stephens, and S. R. Dunbar, Behav.\
  Ecol.\ {\bf 8}, 307 (1997).

\bibitem{KR85} P. Knoppien and J. Reddingius, J. Theor.\ Biol.\ {\bf 114}, 273 (1985).

\bibitem{SK86} D. W. Stephens and J. R. Krebs, Foraging Theory (Princeton
  University Press, Princeton, NJ, 1986).

\bibitem{Vea96} G. M. Viswanathan et al., Nature (London) {\bf 381}, 413
  (1996); G. M. Viswanathan et al., Nature (London) {\bf 401}, 911 (1999).

\bibitem{LKW88} M. Levandowsky, J. Klafter, and B. S. White, Bull.\ Mar.\
  Sci.\ {\bf 43}, 758 (1988).

\bibitem{BLMV11} O. B\'enichou, C. Loverdo, M. Moreau, and R. Voituriez,
  Rev.\ Mod.\ Phys.\ {\bf 83}, 03 (2011).

\bibitem{R85} S. A. Rice, in Diffusion-Limited Reactions, Comprehensive
  Chemical Kinetics Vol.\ 25, edited by C. H. Bamford, C. F. H. Tipper, and
  R. G. Compton (Elsevier, New York, 1985).

\bibitem{BB76} O. G. Berg and C. Blomberg, Biophys.\ Chem.\ {\bf 4}, 367
  (1976).

\bibitem{GDB14} T. Gueudr\'e, A. Dobrinevski, and J.-P. Bouchaud, Phys.\
  Rev.\ Lett.\ {\bf 112}, 050602 (2014).

\bibitem{T80} M. Tachiya, Chem.\ Phys.\ Lett.\ {\bf 69}, 605 (1980)

\bibitem{BNHW87} R. F. Bonner, R. Nossal, S. Havlin, and G. H. Weiss, J. Opt.\
  Soc.\ Am.\ A {\bf 4}, 423 (1987).

\bibitem{KF00} A. B. Kolomeisky and M. E. Fisher, Physica (Amsterdam) {\bf
    279A}, 1 (2000).

\bibitem{YAL13} S. B. Yuste, E. Abad, and K. Lindenberg, Phys.\ Rev.\
  Lett. {\bf 110}, 220603 (2013).

\bibitem{PW97} M. Perman and W. Werner, Probab.\ Theory Relat.\ Fields {\bf
    108}, 357 (1997).

\bibitem{D99} B. Davis,  Probab.\ Theory Relat.\ Fields {\bf 113}, 501 (1999).

\bibitem{BW03} I. Benjamini and D. B. Wilson, Electr.\ Commun.\ Probab.\ {\bf
    8}, 86 (2003).

\bibitem{ABV03} O. Angel O, I. Benjamini, and B. Virag, Electr.\ Commun.\
  Probab.\ {\bf 8}, 66 (2003).

\bibitem{Z04} M. P. W. Zerner, Probab.\ Theory Relat.\ Fields {\bf 133}, 98 (2005).

\bibitem{AR05} T. Antal and S. Redner, J. Phys.\ A {\bf 38}, 2555 (2005).

\bibitem{W94} G. H. Weiss, \textit{Aspects and Applications of the Random Walk}
(North-Holland, Amsterdam, 1994).

\bibitem{H95} B. D. Hughes, Random Walks and Random Environments (Clarendon,
  Oxford, 1995), Vol. 1.

\bibitem{R01} S. Redner, A Guide to First-Passage Processes (Cambridge
  University Press, Cambridge, England, 2001).

\bibitem{future} O. B\'enichou and S. Redner, unpublished.

\end{thebibliography}

\end{document}

### 2015-09-04 starvingRW_sim.R

@@ -243,7 +243,7 @@ for (lambda in lambdaseq) {
   #   c_burn_l[[toc]] <- res_burn_s
 }
 data <- list(); data[[1]] <- pop_r_traj;data[[2]] <- pop_c_traj
-save(data,file="/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/lambda_sigma_traj.RData")
+#save(data,file="/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/lambda_sigma_traj.RData")
 
 library(plotrix)
 mfreq <- matrix(0,length(lambdaseq),length(sigmaseq))

### 2016-01-12 Manuscript/simulation.tex
julia file

\documentclass[11pt]{iopart}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{cite}
\usepackage{float}
\makeatletter
\makeatother

\expandafter\let\csname equation*\endcsname\relax
\expandafter\let\csname endequation*\endcsname\relax
\usepackage{amsmath}
\newcommand{\agt}{\mathrel{\raise.3ex\hbox{$>$\kern-.75em\lower1ex\hbox{$\sim$}}}}
\begin{document}

\title{Role of Starvation on Foraging Dynamics}

\author{J. D. Yeakel} \address{Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, New
  Mexico 87501, USA}

\author{S. Redner} \address{Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, New
  Mexico 87501, USA}

\begin{abstract}

  We outline an event-driven simulation approach for the foraging model,
  which involves a resource that renews by logistic growth, as well as two
  classes of foragers---full and hungry.  Full foragers reproduce at a fixed
  rate and are not vulnerable to mortality.  However, a full forager can
  starve when resources are scarce; conversely, a hungry forager can become
  full when the resource is abundant.  Hungry foragers do not reproduce and
  die at rate $\mu$.
\end{abstract}

%\maketitle

\section{The Model}

We assume that foragers can exist in two discrete states---full and hungry.
Full foragers $F$ are those that have just encountered and consumed a unit of
resource $R$.  On the other hand, a full forager that does not encounter
a resource as it wanders is converted into a hungry forager $H$ with rate
$\sigma$.  Whenever a forager, either full or hungry, encounters resources,
one unit of the resource is consumed.  If the forager was hungry, it is
recruited into the full population with rate $\rho$.  During the time
that a forager is hungry, it dies with a fixed mortality rate $\mu$, while
full foragers do not experience mortality risk.  Furthermore, full
foragers reproduce with rate $\lambda$.  Finally, we assume that, in the
absence of foragers, the underlying resource undergoes logistic growth, with
growth rate $\alpha$ and carrying capacity equal to one.

According to these processes and also under the assumption that the densities
of full foragers, hungry foragers, and resources (also denoted by $F$,
$H$, and $R$, respectively) are perfectly mixed, they evolve according to the
rate equations:
\begin{align}
  \label{RE}
\begin{split}
\dot F &= \lambda F + \rho  RH - \sigma (1-R)F\,,\\
\dot H &= \sigma (1-R)F - \rho RH - \mu H\,, \\
\dot R &= \alpha R(1-R) -  R(F+H),\\
\end{split}
\end{align}
where the overdot denotes time derivative.

We now outline our event-driven algorithm to simulate these rate equations.
We will also generalize to the situation where the full and hungry foragers
undergo diffusion with possibly different diffusion coefficients.  Suppose
that the system at some time consists of $N_F$ individuals of type $F$, $N_H$
individuals of type $H$, and $N_R$ individuals of type $R$.  The total number
of particles $N=N_F+N_H+N_R$.

In each simulation step an individual is picked: a full individual is picked
with probability $N_F/N$, a hungry individual is picked with probability
$N_H/N$, and a individual resource is picked with probability $N_R/N$.  If an
$F$ is picked, it reproduces at rate $\lambda$ and becomes hungry with rate
$\sigma(1-R)$.  If an $H$ is picked, it becomes full with rate $\rho R$ and
dies with rate $\mu$.  Finally, if an $R$ is picked, it grows with rate
$\alpha(1-R)$ and is eaten with rate $(F+H)$.  Thus the total rate for each
individual event is
\begin{equation}
\mathcal{R}=F\big[\lambda +\sigma(1-R)\big]+ H\big[\rho R+\mu\big]+
R\big[\alpha(1-r)+(F=H)\big]\,.
\end{equation}

Now let's look at how the system evolves according to the various constituent
processes.  If an $F$ is picked, then it may either reproduce or go hungry
according to the processes outlined above.
That is:
\begin{align*}
&\mathrm{growth,~ prob.\ } \lambda/\mathcal{R}\hskip 0.72 in N_F\to N_F\!+\!1\,,\\
&\mathrm{starve,~ prob.\ } \sigma(1-R)\mathcal{R} \hskip 0.35 in N_F\to N_F\!-\!1,\,
  N_S\to N_S\!+\!1\,.
\end{align*}
Similarly, if an $H$ is picked, it may either become full or die following the
processes given above.  That is:
\begin{align*}
&\mathrm{become\ full,~  prob.\ } \rho R/\mathcal{R}\hskip 0.3in N_H\to N_H\!-\!1,\,
  N_F\to N_F\!+\!1\,,\\
&\mathrm{die,~ prob.\ } \mu/\mathcal{R}\hskip 1 in N_H\to N_H\!-\!1\,.
\end{align*}
Finally, if an $R$ is picked, it may either grow or be eaten following the
processes given above.  That is:
\begin{align*}
&\mathrm{grow,~ prob.\ } \alpha(1\!-\!R)/\mathcal{R}\hskip 0.4in N_R\to N_R\!+\!1\,,\\
&\mathrm{eaten,~ prob.\ } (F\!+\!S)/\mathcal{R}\hskip 0.43in  N_R\to N_R\!-\!1\,.
\end{align*}

With these steps, let's now determine the change in the expected number of
full individuals in a single event.  This change is
\begin{subequations}
\begin{equation}
  \Delta N_F =\left[\frac{N_F}{N}\left(\lambda-\sigma(1-R)\right)+\frac{N_S}{N}\rho
      R\right]/\mathcal{R}\,.
\end{equation}
Consequently, the change in the density of full individuals simply is
\begin{equation}
  \Delta F =\left[\frac{N_F}{N}\left(\lambda-\sigma(1-R)\right)+\frac{N_S}{N}\rho
      R\right]/N\mathcal{R}\,.
\end{equation}
\end{subequations}
Thus is we take the time step to be $\Delta t = (N\mathcal{R})^{-1}$, the
above reduces to the rate equation \eqref{RE} for $F$.

In a similar fashion, the change in the expected number of hungry individuals
in a single event is given by 
\begin{subequations}
\begin{equation}
  \Delta N_H =\left[-\frac{N_H}{N}\left(\mu+\rho R\right)+\frac{N_F}{N}\sigma
    (1-R)\right]/\mathcal{R}\,,
\end{equation}
so that the change in the density of hungry individuals simply is
\begin{equation}
  \Delta H =\left[-\frac{N_H}{N}\left(\mu+\rho R\right)+\frac{N_F}{N}\sigma (1-R)\right]/N\mathcal{R}\,.
\end{equation}
\end{subequations}
Finally, the change in the expected number of individual resources in a
single event is given by
\begin{subequations}
\begin{equation}
  \Delta N_R =\left[\frac{N_R}{N}\left(\alpha(1-R)\right)-\frac{N_F}{N}-\frac{N_H}{N}\right]/\mathcal{R}\,,
\end{equation}
so that the change in the density of resources is
\begin{equation}
  \Delta R ==\left[\frac{N_R}{N}\left(\alpha(1-R)\right)-\frac{N_F}{N}-\frac{N_H}{N}\right]/N\mathcal{R}\,.
\end{equation}
\end{subequations}

\end{document}

### 2016-01-12 src/starvingforager_eventNM.jl

@@ -1,0 +1,82 @@
+
+#Read in packages/function
+#ipbc :: torus movement
+include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
+
+
+#Initiate values
+#Lattice dimension
+L = 20;
+dim = 2;
+size = (L-2)^dim;
+t_term = 500;
+
+alpha = 0.5;
+K = 1;
+sigma = 0.25;
+rho = 0.25;
+lambda = 0.3;
+mu = 0.1;
+
+#Starting number of foragers
+initsize = 10;
+#Initial state values
+ind_vec = rand(collect(0:2),initsize);
+#Initial location values
+loc_vec = rand(collect(1:size),initsize);
+
+
+t = 0;
+
+#Initial count of how many resouces, starvers, and full in this timestep??
+#Count the number of individual R + S + F
+tot = length(ind_vec);
+R = length(find(x->x==0,ind_vec));
+S = length(find(x->x==1,ind_vec));
+F = initsize - (R + S);
+
+
+tic = 1;
+while t < (t_term-1)
+
+  #Construct probability lines, which are a function of R, S, F
+
+  #Grow <-----> Consumed
+  R_pr_line = (alpha*(K-R))/((alpha*(K-R)) + (F + S));
+
+  #Recover <-----> Mortality
+  S_pr_line = (rho*R)/(rho*R + mu);
+
+  #Grow <-----> Starve
+  F_pr_line = lambda/(lambda+sigma*(K-R));
+
+  #Randomly select an individual (R,S,F) with probability 1/N
+  #ind thus represents the POSITION of the individual
+  #Update total number of individuals
+
+  #Update the total
+  tot = length(ind_vec);
+  id = rand(collect(1:tot))
+
+  #If randomly selected individual is a resource
+  if ind_vec(id) == 0
+    #Append a new resource to the END of the vector
+    push!(ind_vec,state);
+    #Append the resource's location to the END of the vector
+    push!(loc_vec,location);
+    #Update Tally
+    R = R + 1;
+
+  end
+
+  #If randomly selected individual is a starver
+  if ind_vec(id) == 1
+
+
+  end
+
+  #If randomly selected individual is a full
+  if ind_vec(id) == 2
+
+
+  end

### 2016-01-12 starvingRW_sim.R

@@ -7,16 +7,16 @@ source("R/ipbc.R")
 sourceCpp("src/starvingforager_eventNM.cpp")
 
 #Initiate starting conditions
-L <- 20
+L <- 50
 size <- (L-2)^2
-t_term <- 50000
+t_term <- 10
 
 #Parameters
-alpha <- 0.5
-K <- 1
+alpha <- 1
+K <- size
 sigma <- 0.5
 rho <- 0.2
-lambda <- 0.2
+lambda <- 0.3
 mu <- 0.2
 #D <- 1
 

### 2016-01-13 Manuscript/simulation.tex
new julia file with correct methodology

@@
 
-\title{Role of Starvation on Foraging Dynamics}
-
-\author{J. D. Yeakel} \address{Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, New
-  Mexico 87501, USA}
-
-\author{S. Redner} \address{Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, New
-  Mexico 87501, USA}
+\title{Event-Driven Simulation of Foraging Dynamics}
 
@@
 
-  We outline an event-driven simulation approach for the foraging model,
-  which involves a resource that renews by logistic growth, as well as two
-  classes of foragers---full and hungry.  Full foragers reproduce at a fixed
-  rate and are not vulnerable to mortality.  However, a full forager can
-  starve when resources are scarce; conversely, a hungry forager can become
-  full when the resource is abundant.  Hungry foragers do not reproduce and
-  die at rate $\mu$.
+  We outline an event-driven simulation for the foraging model, which
+  involves a resource that renews by logistic growth, as well as two classes
+  of foragers---full and hungry.  Full foragers reproduce at a fixed rate and
+  are not vulnerable to mortality.  However, a full forager can become hungry
+  when resources are scarce; conversely, a hungry forager can become full
+  when the resource is abundant.  Hungry foragers do not reproduce and die at
+  a fixed rate.
 \end{abstract}
@@
 Full foragers $F$ are those that have just encountered and consumed a unit of
-resource $R$.  On the other hand, a full forager that does not encounter
-a resource as it wanders is converted into a hungry forager $H$ with rate
-$\sigma$.  Whenever a forager, either full or hungry, encounters resources,
-one unit of the resource is consumed.  If the forager was hungry, it is
-recruited into the full population with rate $\rho$.  During the time
-that a forager is hungry, it dies with a fixed mortality rate $\mu$, while
-full foragers do not experience mortality risk.  Furthermore, full
-foragers reproduce with rate $\lambda$.  Finally, we assume that, in the
-absence of foragers, the underlying resource undergoes logistic growth, with
-growth rate $\alpha$ and carrying capacity equal to one.
+resource $R$.  On the other hand, a full forager that does not encounter a
+resource as it wanders converts to a hungry forager $H$ with rate equal to
+the product of a parameter $\sigma$ and the density of non-resources.
+Whenever a forager, either full or hungry, encounters a resource, one unit of
+the resource is consumed.  If the forager was hungry, it turns into a full
+forager with rate $\rho$.  During the time that a forager is hungry, it dies
+with mortality rate $\mu$, while full foragers do not experience mortality
+risk.  Furthermore, full foragers reproduce with rate $\lambda$.  Finally, we
+assume that, in the absence of foragers, the underlying resource undergoes
+logistic growth, with growth rate $\alpha$ and carrying capacity equal to
+one.
 
@@
 
-We now outline our event-driven algorithm to simulate these rate equations.
-We will also generalize to the situation where the full and hungry foragers
-undergo diffusion with possibly different diffusion coefficients.  Suppose
-that the system at some time consists of $N_F$ individuals of type $F$, $N_H$
-individuals of type $H$, and $N_R$ individuals of type $R$.  The total number
-of particles $N=N_F+N_H+N_R$.
+We now outline an event-driven algorithm that mimic these rate equations.  We
+will also generalize to the situation where the full and hungry foragers
+undergo diffusion with possibly different diffusion coefficients on a
+finite-dimensional lattice.  Suppose that the system at some time consists of
+$N_F$ full foragers, $N_H$ hungry foragers, and $N_R$ individual resources.
+The total number of particles $N=N_F+N_H+N_R$.
 
-In each simulation step an individual is picked: a full individual is picked
-with probability $N_F/N$, a hungry individual is picked with probability
-$N_H/N$, and a individual resource is picked with probability $N_R/N$.  If an
-$F$ is picked, it reproduces at rate $\lambda$ and becomes hungry with rate
-$\sigma(1-R)$.  If an $H$ is picked, it becomes full with rate $\rho R$ and
-dies with rate $\mu$.  Finally, if an $R$ is picked, it grows with rate
-$\alpha(1-R)$ and is eaten with rate $(F+H)$.  Thus the total rate for each
-individual event is
+In each elemental event an individual is picked: a full forager with
+probability $N_F/N$, a hungry forager with probability $N_H/N$, and a
+resource with probability $N_R/N$.  If an $F$ is picked, it reproduces at
+rate $\lambda$ and becomes hungry with rate $\sigma(1-R)$.  If an $H$ is
+picked, it becomes full with rate $\rho R$ and dies with rate $\mu$.
+Finally, if an $R$ is picked, it grows with rate $\alpha(1-R)$ and is eaten
+with rate $(F+H)$.  Thus the total rate for each elemental event is
 \begin{equation}
 \mathcal{R}=F\big[\lambda +\sigma(1-R)\big]+ H\big[\rho R+\mu\big]+
-R\big[\alpha(1-r)+(F=H)\big]\,.
+R\big[\alpha(1-R)+(F+H)\big]\,.
 \end{equation}
 
-Now let's look at how the system evolves according to the various constituent
+Now let's look at how the system evolves according to its constituent
 processes.  If an $F$ is picked, then it may either reproduce or go hungry
-according to the processes outlined above.
-That is:
+according to the steps outlined above.  That is:
 \begin{align*}
-&\mathrm{growth,~ prob.\ } \lambda/\mathcal{R}\hskip 0.72 in N_F\to N_F\!+\!1\,,\\
-&\mathrm{starve,~ prob.\ } \sigma(1-R)\mathcal{R} \hskip 0.35 in N_F\to N_F\!-\!1,\,
-  N_S\to N_S\!+\!1\,.
+&\mathrm{growth,~ prob.\ } \lambda/\mathcal{R}\hskip 0.82 in N_F\to N_F\!+\!1\,,\\
+&\mathrm{starve,~ prob.\ } \sigma(1-R)/\mathcal{R} \hskip 0.35 in N_F\to N_F\!-\!1,\,
+  N_H\to N_H\!+\!1\,.
 \end{align*}
@@
 \begin{align*}
-&\mathrm{become\ full,~  prob.\ } \rho R/\mathcal{R}\hskip 0.3in N_H\to N_H\!-\!1,\,
+&\mathrm{become\ full,~  prob.\ } \rho R/\mathcal{R}\hskip 0.4in N_H\to N_H\!-\!1,\,
   N_F\to N_F\!+\!1\,,\\
-&\mathrm{die,~ prob.\ } \mu/\mathcal{R}\hskip 1 in N_H\to N_H\!-\!1\,.
+&\mathrm{die,~ prob.\ } \mu/\mathcal{R}\hskip 1.12 in N_H\to N_H\!-\!1\,.
 \end{align*}
@@
 \begin{align*}
-&\mathrm{grow,~ prob.\ } \alpha(1\!-\!R)/\mathcal{R}\hskip 0.4in N_R\to N_R\!+\!1\,,\\
-&\mathrm{eaten,~ prob.\ } (F\!+\!S)/\mathcal{R}\hskip 0.43in  N_R\to N_R\!-\!1\,.
+&\mathrm{grow,~ prob.\ } \alpha(1\!-\!R)/\mathcal{R}\hskip 0.8in N_R\to N_R\!+\!1\,,\phantom{~~~~~~~~~~}\\
+&\mathrm{eaten,~ prob.\ } (F\!+\!H)/\mathcal{R}\hskip 0.82in  N_R\to N_R\!-\!1\,.\phantom{~~~~~~~~~~}
 \end{align*}
 
-With these steps, let's now determine the change in the expected number of
-full individuals in a single event.  This change is
+We now determine the change in the expected number of individuals of each
+type in a single event that are consistent with the above processes.  For
+full foragers, this change is
 \begin{subequations}
 \begin{equation}
-  \Delta N_F =\left[\frac{N_F}{N}\left(\lambda-\sigma(1-R)\right)+\frac{N_S}{N}\rho
+  \Delta N_F =\left[\frac{N_F}{N}\big(\lambda-\sigma(1-R)\big)+\frac{N_H}{N}\rho
       R\right]/\mathcal{R}\,.
 \end{equation}
-Consequently, the change in the density of full individuals simply is
+The term proportional to $N_F$ comes from processes in which a full forager
+is picked while the term proportional to $N_H$ comes from processes in which
+a hungry forager is picked and it converted to a full forager.  Consequently,
+the change in the density of full foragers simply is
 \begin{equation}
-  \Delta F =\left[\frac{N_F}{N}\left(\lambda-\sigma(1-R)\right)+\frac{N_S}{N}\rho
+  \Delta F =\left[F\big(\lambda-\sigma(1-R)\big)+H\rho
     R\right]/N\mathcal{R}\,.
@@
 \end{subequations}
-Thus is we take the time step to be $\Delta t = (N\mathcal{R})^{-1}$, the
-above reduces to the rate equation \eqref{RE} for $F$.
+Thus if we take the time step for each event to be
+$\Delta t = (N\mathcal{R})^{-1}$, the above reduces to the rate equation
+\eqref{RE} for $F$.  Thus in each microscopic event of the model, the time
+should be advanced by $\Delta t = (N\mathcal{R})^{-1}$.
 
-In a similar fashion, the change in the expected number of hungry individuals
-in a single event is given by 
+In a similar fashion, the change in the expected number of hungry foragers in
+a single event is given by
 \begin{subequations}
 \begin{equation}
-  \Delta N_H =\left[-\frac{N_H}{N}\left(\mu+\rho R\right)+\frac{N_F}{N}\sigma
+  \Delta N_H =\left[-\frac{N_H}{N}(\rho R+\mu)+\frac{N_F}{N}\sigma
     (1-R)\right]/\mathcal{R}\,,
 \end{equation}
-so that the change in the density of hungry individuals simply is
+so that the change in the density of hungry foragers simply is
 \begin{equation}
-  \Delta H =\left[-\frac{N_H}{N}\left(\mu+\rho R\right)+\frac{N_F}{N}\sigma (1-R)\right]/N\mathcal{R}\,.
+  \Delta H =\left[-H(\rho R+\mu)+F\sigma (1-R)\right]/N\mathcal{R}\,.
 \end{equation}
@@
 \begin{equation}
-  \Delta N_R =\left[\frac{N_R}{N}\left(\alpha(1-R)\right)-\frac{N_F}{N}-\frac{N_H}{N}\right]/\mathcal{R}\,,
+  \Delta N_R =\left[\frac{N_R}{N}\big(\alpha(1-R)\big)-\frac{(N_F+N_H)}{N}\right]/\mathcal{R}\,,
 \end{equation}
@@
 \begin{equation}
-  \Delta R ==\left[\frac{N_R}{N}\left(\alpha(1-R)\right)-\frac{N_F}{N}-\frac{N_H}{N}\right]/N\mathcal{R}\,.
+  \Delta R =\left[R\big(\alpha(1-R)\big)-(F+H)\right]/N\mathcal{R}\,.
 \end{equation}
 \end{subequations}
+The equations for $\Delta F$, $\Delta H$, and $\Delta R$ then reproduce the
+original rate equations \eqref{RE} when the time step for an elemental event
+is taken to be $\Delta t = (N\mathcal{R})^{-1}$.
+
+We can straightforwardly generalize the above approach to the situation where
+the foragers are diffusing on a lattice.  Suppose that the full and hungry
+foragers diffuse with respective diffusion coefficients $D_F$ and $D_H$.  In
+this case, the rate equations \eqref{RE} generalize to the set of partial
+differential equations
+\begin{align}
+  \label{pde}
+\begin{split}
+\frac{\partial F}{\partial t} &= \lambda F + \rho  RH - \sigma (1-R)F+D_F\nabla^2F\,,\\[0.125in]
+\frac{\partial H}{\partial t}  &= \sigma (1-R)F - \rho RH - \mu H+D_H\nabla^2H\,, \\[0.125in]
+\frac{\partial R}{\partial t}  &= \alpha R(1-R) -  R(F+H).\\
+\end{split}
+\end{align}
+Here the densities $F,H,R$ are now functions of space and time,
+$F=F(\mathbf{r},t)$ and similarly for $H$ and $R$.
+
+We now determine the total rate for an elemental event.  If an $F$ is picked,
+it reproduces at rate $\lambda$, becomes hungry with rate $\sigma(1-R)$, and
+moves with rate $D_F$.  If an $H$ is picked, it becomes full with rate
+$\rho R$, dies with rate $\mu$, and moves with rate $D_H$.  As before, if an
+$R$ is picked, it grows with rate $\alpha(1-R)$ and is eaten with rate
+$(F+H)$.  Thus the total rate for each elemental event is
+\begin{equation}
+\mathcal{R}_d=F\big[\lambda +\sigma(1-R)+D_F\big]+ H\big[\rho R+\mu+D_H\big]+
+R\big[\alpha(1-r)+(F+H)\big]\,.
+\end{equation}
+Here the subscript $d$ indicates that this rate pertains to the foraging
+model with diffusing foragers on a $d$-dimensional lattice.
+
+As before, we determine at how the system evolves according to the various
+constituent processes.  If an $F$ is picked, it may either reproduce, go
+hungry, or move.  That is:
+\begin{align*}
+&\mathrm{growth,~ prob.\ } \lambda/\mathcal{R}_d\hskip 0.82 in N_F\to N_F\!+\!1\,,\\
+&\mathrm{starve,~ prob.\ } \sigma(1-R)/\mathcal{R}_d \hskip 0.35 in N_F\to N_F\!-\!1,\,\\
+&\mathrm{move,~ prob.\ } D_F/\mathcal{R}_d \hskip 0.8 in N_F\to N_F\,.
+\end{align*}
+Similarly, if an $H$ is picked, it may become full, die, or move following
+the processes given above.  That is:
+\begin{align*}
+&\mathrm{become\ full,~  prob.\ } \rho R/\mathcal{R}_d\hskip 0.4in N_H\to N_H\!-\!1,\,
+  N_F\to N_F\!+\!1\,,\\
+&\mathrm{die,~ prob.\ } \mu/\mathcal{R}_d\hskip 1.12 in N_H\to N_H\!-\!1\,,\\
+&\mathrm{move,~ prob.\ } D_H/\mathcal{R}_d\hskip 0.82 in N_H\to N_H\,.
+\end{align*}
+The evolution of the resource is the same as the mean-field case.  Following
+the same logic as in the mean-field case, the evolution of the densities will
+be described by the evolution equations \eqref{pde} when the time step of
+each elemental event is taken to be $(N\mathcal{R}_d)^{-1}$.
+
+There are several additional detailed subtleties of the simulation that are
+worth mentioning.  First, it seems that there does not need to be any
+constraint on the number of foragers on any site, so I think it's best to let
+this number be arbitrary.  However, the carrying capacity of the resource has
+been set to 1, so that the number of individual resources at any site should
+be either 0 or 1.
+
+Now let's turn to the mechanics of the various update events.  In the
+mean-field limit, when a full forager reproduces, its offspring should be
+placed anywhere.  For a finite-dimensional lattice, it seems reasonable to
+place the offspring at the same site as the parent.  Diffusion of a full
+forager on a lattice can be modeled by allowing it to hop to a nearest
+neighbor; similarly for the motion of hungry foragers.
+
+The mechanics of the resource evolution is a bit subtle.  In the mean-field
+limit, if an individual resource is picked a new resource is created with
+probability proportional to $1-R$, where $R$ is the global density of
+resource, which is necessarily less than 1.  If a new resource is created, it
+should be place randomly on any one of the sites that does not contain a
+resource, but this site could be occupied by a forager.  On a lattice, the
+creation of a new resource should be proportional to a local version of
+$1-R$.  I think it's reasonable to consider the $z$ nearest-neighbor sites of
+the current resource and put a new individual resource at one of the vacant
+nearest neighbors with probability $1-{n}/{z}$, where $n$ is the number
+of individual resources in the local neighborhood of $z$ sites.  The
+disappearance of an individual resource by consumption is straightforward.
 

### 2016-01-13 src/starvingforager_event.jl

@@ -1,0 +1,201 @@
+
+#Read in packages/function
+#ipbc :: torus movement
+include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
+
+
+#Initiate values
+#Lattice dimension
+L = 20;
+dim = 2;
+size = (L-2)^dim;
+t_term = 500;
+
+alpha = 0.5;
+K = 1;
+sigma = 0.25;
+rho = 0.25;
+lambda = 0.3;
+mu = 0.1;
+DF = 0.2;
+DH = 0.2;
+
+#Starting number of foragers
+initsize = 10;
+#Initial state values
+ind_vec = rand(collect(0:2),initsize);
+#Initial location values
+loc_vec = rand(collect(1:size),initsize);
+
+#NEED TO ENSURE THAT RESOURCES ARE PLACED 1 PER SITE
+#HAVE A noresourcesites vector that accounts for sites WITHOUT resources
+#Updated each timestep
+
+
+t = 0;
+
+#Initial count of how many resouces, starvers, and full in this timestep??
+#Count the number of individual R + S + F
+tot = length(ind_vec);
+NR = length(find(x->x==0,ind_vec));
+NH = length(find(x->x==1,ind_vec));
+NF = initsize - (R + S);
+
+
+tic = 1;
+while t < (t_term-1)
+
+  N = NF + NH + NR;
+  F = NF/N;
+  H = NH/N;
+  R = NR/N;
+
+
+  #Calculate Rate
+  Rate = F*(lambda + sigma*(1-R) + DF) + H*(rho*R + mu + DH) + R*(alpha*(1-R) + (F+H));
+  dt = 1/(Rate*N);
+
+
+  #Construct probability lines, which are a function of R, S, F
+
+  #Grow <-----> Starve <-----> Diffuse
+  F_pr_line = Array(Float64,2);
+  F_pr_line[1] = lambda/(lambda+sigma*(K-R) + DF);
+  F_Pr_line[2] = F_pr_line[1] + (sigma*(K-R))/(lambda+sigma*(K-R) + DF);
+
+  #Recover <-----> Mortality <-----> Diffuse
+  H_pr_line = Array(Float64,2);
+  H_pr_line[1] = (rho*R)/(rho*R + mu + DH);
+  H_pr_line[2] = mu/(rho*R + mu + DH);
+
+  #Grow <-----> Consumed
[…]

### 2016-01-13 src/starvingforager_eventNM.jl

@@ -60,23 +60,93 @@ while t < (t_term-1)
 
   #If randomly selected individual is a resource
   if ind_vec(id) == 0
-    #Append a new resource to the END of the vector
-    push!(ind_vec,state);
-    #Append the resource's location to the END of the vector
-    push!(loc_vec,location);
-    #Update Tally
-    R = R + 1;
+    state = 0;
+    location = loc_vec[id];
+
+    #Draw a random event
+    #Grow, become consumed or move?
+    draw_event = rand();
+
+    #GROW
+    if draw_event < R_Pr_line
+
+      #Append a new resource to the END of the vector
+      push!(ind_vec,state);
+      #Append the resource's location to the END of the vector
+      push!(loc_vec,location);
+      #Update Tally
+      R = R + 1;
+
+    end
+
+    #BECOME CONSUMED!
+    if draw_event >= R_pr_line && draw_event < 1
+      deleteat!(ind_vec,id);
+      deleteat!(loc_vec,id);
+      R = R-1;
+    end
+
 
   end
 
   #If randomly selected individual is a starver
   if ind_vec(id) == 1
+    state = 1;
+    location = loc_vec[id];
 
+    #Draw a random event
+    #Recover, die, or move??
+    draw_event = rand();
+
+    #Recover
+    if draw_event < S_pr_line
+      loc_vec[id] = 2;
+      S = S-1;
+      F = F+1;
+    end
+
+    #Die
+    if draw_event >= S_pr_line && draw_event < 1
+      deleteat!(ind_vec,id);
+      deleteat!(loc_vec,id);
+      S = S-1;
+    end
 
   end
 
   #If randomly selected individual is a full
   if ind_vec(id) == 2
+    state = 2;
+    location = loc_vec[id];
 
+    draw_event = rand();
 
+    #GROW
+    if draw_event < F_pr_line
+      push!(ind_vec,state);
+      push!(loc_vec,location);
+      F = F+1;
+    end
+
+    #STARVE
+    if draw_event >= F_pr_line && draw_event < 1
+      ind_vec[id] = 1;
+
+      S = S+1;
+      F = F-1;
+    end
   end
+
+  #NEED TO DEFINE dt
+
+  #Advance time
+  t = t + dt;
+
+  #Update output
+
+
+
+end #end while loop over t
+
+
+#Posthoc analysis

### 2016-01-13 src/starvingforager_event.jl
finished code but undetermined issues with updating vectors

@@ -1,15 +1,17 @@
 
 #Read in packages/function
 #ipbc :: torus movement
-include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
+using StatsBase
+using Gadfly
+#include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
 
 
 #Initiate values
 #Lattice dimension
-L = 20;
+L = 10;
 dim = 2;
 size = (L-2)^dim;
-t_term = 500;
+t_term = 50;
 
 alpha = 0.5;
 K = 1;
@@ -20,51 +22,68 @@ mu = 0.1;
 DF = 0.2;
 DH = 0.2;
 
-#Starting number of foragers
+#Starting number of foragers & resources
 initsize = 10;
 #Initial state values
-ind_vec = rand(collect(0:2),initsize);
+ind_vec = [rand(collect(1:2),Int(2*round(initsize/3)));zeros(Int64,Int(round(initsize/3)))];
 #Initial location values
-loc_vec = rand(collect(1:size),initsize);
+#2/3 of initsize will be foragers; 1/3 will be resources
+resourceloc_vec = sample(collect(1:size),Int(round(initsize/3)),replace=false);
+loc_vec = [rand(collect(1:size),Int(2*round(initsize/3)));resourceloc_vec];
+
+#Re-establish initsize to account for rounding errors
+initsize = length(ind_vec);
+
+#Arrays for output
+ind_out = Array(Array{Int64},1);
+loc_out = Array(Array{Int64},1);
+time_out = Array(Float64,1);
+
+ind_out[1] = ind_vec;
+loc_out[1] = loc_vec;
+time_out[1] = 0;
 
 #NEED TO ENSURE THAT RESOURCES ARE PLACED 1 PER SITE
 #HAVE A noresourcesites vector that accounts for sites WITHOUT resources
 #Updated each timestep
-
+noresourcesites = collect(1:size);
+deleteat!(noresourcesites,sort(resourceloc_vec));
 
 t = 0;
 
 #Initial count of how many resouces, starvers, and full in this timestep??
 #Count the number of individual R + S + F
-tot = length(ind_vec);
-NR = length(find(x->x==0,ind_vec));
+#tot = length(ind_vec);
+NF = length(find(x->x==2,ind_vec));
 NH = length(find(x->x==1,ind_vec));
[…]

### 2016-01-14 foraging_simulations.jl
imbedded code in function

@@ -1,0 +1,31 @@
+include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event.jl")
+using StatsBase
+using Gadfly
+
+
+L = 10;
+dim = 2;
+t_term = 50;
+alpha = 0.5;
+K = 1;
+sigma = 0.25;
+rho = 0.25;
+lambda = 0.3;
+mu = 0.1;
+DF = 0.2;
+DH = 0.2;
+
+ind_out = starvingforager_event(L,dim,t_term,alpha,K,sigma,rho,lambda,mu,DF,DH);
+
+
+
+#Posthoc analysis
+steps = tic;
+pop_F = zeros(steps);
+pop_H = zeros(steps);
+pop_R = zeros(steps);
+for i=1:steps
+  pop_F[i] = length(find(x->x==2,ind_out[i]));
+  pop_H[i] = length(find(x->x==1,ind_out[i]));
+  pop_R[i] = length(find(x->x==0,ind_out[i]));
+end

### 2016-01-14 src/starvingforager_event.jl

@@ -1,245 +1,226 @@
-
-#Read in packages/function
-#ipbc :: torus movement
-using StatsBase
-using Gadfly
-#include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
+function starvingforager_event(L,dim,t_term,alpha,K,sigma,rho,lambda,mu,DF,DH)
+  #Read in packages/function
+  #ipbc :: torus movement
+  #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
 
 
-#Initiate values
-#Lattice dimension
-L = 10;
-dim = 2;
-size = (L-2)^dim;
-t_term = 50;
+  #Initiate values
+  #Lattice dimension
 
-alpha = 0.5;
-K = 1;
-sigma = 0.25;
-rho = 0.25;
-lambda = 0.3;
-mu = 0.1;
-DF = 0.2;
-DH = 0.2;
-
-#Starting number of foragers & resources
-initsize = 10;
-#Initial state values
-ind_vec = [rand(collect(1:2),Int(2*round(initsize/3)));zeros(Int64,Int(round(initsize/3)))];
-#Initial location values
-#2/3 of initsize will be foragers; 1/3 will be resources
-resourceloc_vec = sample(collect(1:size),Int(round(initsize/3)),replace=false);
-loc_vec = [rand(collect(1:size),Int(2*round(initsize/3)));resourceloc_vec];
-
-#Re-establish initsize to account for rounding errors
-initsize = length(ind_vec);
-
-#Arrays for output
-ind_out = Array(Array{Int64},1);
-loc_out = Array(Array{Int64},1);
-time_out = Array(Float64,1);
-
-ind_out[1] = ind_vec;
-loc_out[1] = loc_vec;
-time_out[1] = 0;
-
-#NEED TO ENSURE THAT RESOURCES ARE PLACED 1 PER SITE
-#HAVE A noresourcesites vector that accounts for sites WITHOUT resources
-#Updated each timestep
-noresourcesites = collect(1:size);
-deleteat!(noresourcesites,sort(resourceloc_vec));
-
-t = 0;
-
-#Initial count of how many resouces, starvers, and full in this timestep??
-#Count the number of individual R + S + F
-#tot = length(ind_vec);
[…]

### 2016-02-02 foraging_simulations.jl
I think the sim code is right now - though resources still mismatching. Going to try with resource translocation

@@ -16,7 +16,7 @@ Fsimstar = zeros(length(sigmavec));
 Hsimstar = zeros(length(sigmavec));
 Rsimstar = zeros(length(sigmavec));
 for i = 1:length(sigmavec)
-
+  println("sigma = ",sigmavec[i])
   L = 50;
   dim = 2;
   prop_fill = 0.5
@@ -28,8 +28,8 @@ for i = 1:length(sigmavec)
   rho = 0.2;
   lambda = 0.2;
   mu = 0.2;
-  DF = 0;
-  DH = 0;
+  DF = 1;
+  DH = 1;
 
 
 
@@ -58,8 +58,9 @@ layer(x=sigmavec,y=Hstar,Geom.line,Theme(default_color=colorant"orange")),
 layer(x=sigmavec,y=Hsimstar,Geom.point,Theme(default_color=colorant"orange",default_point_size=3pt)),
 layer(x=sigmavec,y=Rstar,Geom.line,Theme(default_color=colorant"blue")),
 layer(x=sigmavec,y=Rsimstar,Geom.point,Theme(default_color=colorant"blue",default_point_size=3pt)),
-)
+Guide.XLabel("sigma"),Guide.YLabel("Steady state; F=green; H=orange; R=blue"));
 
+draw(PDF("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/figs/fig_comparison.pdf", 8inch, 5inch), comparison)
 
 F = prop_out[1,:];
 H = prop_out[2,:];

### 2016-02-02 src/starvingforager_event.jl

@@ -7,7 +7,7 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
   #Initiate values
   #Lattice dimension
 
-  size = (L-2)^dim;
+  S = (L-2)^dim;
 
 
   #Initial state values
@@ -19,8 +19,8 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
   #Initial location values
   #2/3 of initsize will be foragers; 1/3 will be resources
   #replace = false because want only one resource per site
-  resourceloc_vec = sample(collect(1:size),Int(round(initsize/3)),replace=false);
-  loc_vec = [rand(collect(1:size),Int(2*round(initsize/3)));resourceloc_vec];
+  resourceloc_vec = sample(collect(1:S),Int(round(initsize/3)),replace=false);
+  loc_vec = [rand(collect(1:S),Int(2*round(initsize/3)));resourceloc_vec];
 
   #Re-establish initsize to account for rounding errors
   initsize = length(ind_vec);
@@ -40,7 +40,7 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
   #NEED TO ENSURE THAT RESOURCES ARE PLACED 1 PER SITE
   #HAVE A noresourcesites vector that accounts for sites WITHOUT resources
   #Updated each timestep
-  noresourcesites = collect(1:size);
+  noresourcesites = collect(1:S);
   #Deletes locations of sites with resources from the list of open sites (noresourcesites)
   deleteat!(noresourcesites,sort(resourceloc_vec));
 
@@ -56,9 +56,9 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
   N = NF + NH + NR;
 
   #Initial densities
-  F = NF/size;
-  H = NH/size;
-  R = NR/size;
+  F = NF/S;
+  H = NH/S;
+  R = NR/S;
   prop_out = Array{Float64}(3,1);
   prop_out[:,1] = [F,H,R];
   push!(N_out,N);
@@ -85,9 +85,13 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
     #Calculate Rate
     Rate = F*(lambda + sigma*(K-R) + DF) + H*(rho*R + mu + DH) + R*(alpha*(K-R) + (F+H));
 
+    # TESTING
[…]

### 2016-02-03 foraging_simulations.jl
two version... diff has resource diffusion and if DF=DH=DR=1 the sim results match model perfectly

@@ -28,8 +28,9 @@ for i = 1:length(sigmavec)
   rho = 0.2;
   lambda = 0.2;
   mu = 0.2;
-  DF = 1;
-  DH = 1;
+  DF = 0;
+  DH = 0;
+  DR = 0;
 
 
 
@@ -39,7 +40,7 @@ for i = 1:length(sigmavec)
   Rstar[i] = (mu*(-lambda+sigma))/(lambda*rho+mu*sigma);
 
   #The simulation
-  time_out, prop_out, N_out = starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu,DF,DH);
+  time_out, prop_out, N_out = starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu,DF,DH,DR);
   F = prop_out[1,:];
   H = prop_out[2,:];
   R = prop_out[3,:];

### 2016-02-03 src/starvingforager_eventdiff.jl

@@ -83,7 +83,7 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
     # end
 
     #Calculate Rate
-    Rate = (F + H + R) + F*(lambda + sigma*(K-R) + DF) + H*(rho*R + mu + DH) + R*(alpha*(K-R) + (F+H));
+    Rate = F*(lambda + sigma*(K-R) + DF) + H*(rho*R + mu + DH) + R*(alpha*(K-R) + (F+H) + DR);
 
     # TESTING
     # Rate = (NF/N)*(lambda + sigma*(K-R) + DF) + (NH/N)*(rho*R + mu + DH) + (NR/N)*(alpha*(K-R) + (F+H));
@@ -110,7 +110,7 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
 
     #Grow <-----> Consumed
     R_pr_line[1] = (alpha*(K-R))/Rate;
-    # R_pr_line[2] = R_pr_line[1] + (F+H)/Rate;
+    R_pr_line[2] = R_pr_line[1] + (F+H)/Rate;
 
 
 
@@ -226,7 +226,7 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
       end
 
       #BECOME CONSUMED!
-      if draw_event >= R_pr_line[1]
+      if draw_event >= R_pr_line[1] && draw_event < R_pr_line[2]
 
         #Update the no resource site by adding the position that is now empty
         #Needs to be done BEFORE the ind/loc information is deleted.
@@ -240,6 +240,23 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
 
 
       end
+
+      #Resource Movement
+      if draw_event >= R_pr_line[2]
+        old_location = copy(loc_vec[id]);
+        #Choose random position on the noresourcesite list
+        newresourcepos = rand(collect(1:length(noresourcesites)));
+        #Define the new position as the empty site for resource movement
+        location = noresourcesites[newresourcepos];
+        #Change the current resource location
+        loc_vec[id] = location;
+
+        #Update the no resource site by deleting the position that has been filled
+        deleteat!(noresourcesites,newresourcepos);
+        #AND adding the position that the resource has moved FROM!
+        push!(noresourcesites,old_location)
+      end
+
 
     end
 

### 2016-02-04 foraging_simulations.jl
Changes to rates but still no cigar

@@ -1,11 +1,14 @@
-#include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event.jl")
-include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event.jl")
-# include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_spatial.jl")
-
 
 using StatsBase
 using Gadfly
 using Cairo
+
+
+#include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event.jl")
+include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_nodiff.jl")
+# include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_spatial.jl")
+
+
 
 
 sigmavec = collect(0.4:0.1:1.0);
@@ -28,10 +31,6 @@ for i = 1:length(sigmavec)
   rho = 0.2;
   lambda = 0.2;
   mu = 0.2;
-  DF = 0;
-  DH = 0;
-  DR = 0;
-
 
 
   eta = copy(sigma);
@@ -40,7 +39,7 @@ for i = 1:length(sigmavec)
   Rstar[i] = (mu*(-lambda+sigma))/(lambda*rho+mu*sigma);
 
   #The simulation
-  time_out, prop_out, N_out = starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu,DF,DH,DR);
+  time_out, prop_out, N_out = starvingforager_event_nodiff(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu);
   F = prop_out[1,:];
   H = prop_out[2,:];
   R = prop_out[3,:];

### 2016-02-04 src/starvingforager_event_nodiff.jl

@@ -1,4 +1,4 @@
-function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu,DF,DH,DR)
+function starvingforager_event_nodiff(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu)
   #Read in packages/function
   #ipbc :: torus movement
   #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
@@ -66,11 +66,11 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
 
   tic = 0;
 
-  F_pr_line = Array(Float64,2);
-  H_pr_line = Array(Float64,2);
-  R_pr_line = Array(Float64,2);
+  # F_pr_line = 0.0;
+  # H_pr_line = 0.0;
+  # R_pr_line = 0.0;
 
-  id = Array(Float64,1);
+  # id = Array(Float64,1);
 
   while t < (t_term-1)
     tic = tic + 1;
@@ -83,13 +83,10 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
     # end
 
     #Calculate Rate
-    Rate = (1 - (N/S)) + F*(lambda + sigma*(K-R) + DF) + H*(rho*R + mu + DH) + R*(alpha*(K-R) + (F+H)); #
+    Rate = F*(lambda + sigma*(K-R)) + H*(rho*R + mu) + R*(alpha*(K-R) + (F+H)); # (1 - (N/S)) +
 
     # TESTING
-    # Rate = (NF/N)*(lambda + sigma*(K-R) + DF) + (NH/N)*(rho*R + mu + DH) + (NR/N)*(alpha*(K-R) + (F+H));
-
-    # TESTING!
-    # Rate = F*(lambda + sigma*(K-R) + DF) + H*(rho*R + mu + DH) + R*(alpha*(K-R) + (F+H));
+    # Rate = (NF/N)*(lambda + sigma*(K-R)) + (NH/N)*(rho*R + mu) + (NR/N)*(alpha*(K-R) + (F+H));
 
     dt = 1/(Rate*N);
     if Rate == 0
@@ -100,16 +97,16 @@ function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu
[…]

### 2016-02-05 Manuscript/simulation-2.tex
revising the simulation as event_rate.jl

\documentclass[11pt]{iopart}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{cite}
\usepackage{float}
\makeatletter
\makeatother

\expandafter\let\csname equation*\endcsname\relax
\expandafter\let\csname endequation*\endcsname\relax
\usepackage{amsmath}
\newcommand{\agt}{\mathrel{\raise.3ex\hbox{$>$\kern-.75em\lower1ex\hbox{$\sim$}}}}
\begin{document}

\title{Event-Driven Simulation of Foraging Dynamics}

\begin{abstract}

  We outline an event-driven simulation for the foraging model, which
  involves a resource that renews by logistic growth, as well as two classes
  of foragers---full and hungry.  Full foragers reproduce at a fixed rate and
  are not vulnerable to mortality.  However, a full forager can become hungry
  when resources are scarce; conversely, a hungry forager can become full
  when the resource is abundant.  Hungry foragers do not reproduce and die at
  a fixed rate.
\end{abstract}

%\maketitle

\section{The Model}

We assume that foragers can exist in two discrete states---full and hungry.
Full foragers $F$ are those that have just encountered and consumed a unit of
resource $R$.  On the other hand, a full forager that does not encounter a
resource as it wanders converts to a hungry forager $H$ with rate equal to
the product of a parameter $\sigma$ and the density of non-resources.
Whenever a forager, either full or hungry, encounters a resource, one unit of
the resource is consumed.  If the forager was hungry, it turns into a full
forager with rate $\rho$.  During the time that a forager is hungry, it dies
with mortality rate $\mu$, while full foragers do not experience mortality
risk.  Furthermore, full foragers reproduce with rate $\lambda$.  Finally, we
assume that, in the absence of foragers, the underlying resource undergoes
logistic growth, with growth rate $\alpha$ and carrying capacity equal to
one.

According to these processes and also under the assumption that the densities
of full foragers, hungry foragers, and resources (also denoted by $F$,
$H$, and $R$, respectively) are perfectly mixed, they evolve according to the
rate equations:
\begin{align}
  \label{RE}
\begin{split}
\dot F &= \lambda F + \rho  RH - \sigma (1-R)F\,,\\
\dot H &= \sigma (1-R)F - \rho RH - \mu H\,, \\
\dot R &= \alpha R(1-R) -  R(F+H),\\
\end{split}
\end{align}
where the overdot denotes time derivative.  An important feature is the
assumption that the carrying capacity is set equal to 1.  This means that
when the system is completely occupied by resources there can be no further
generation of the resource.  The approach described below can be generalized
to carrying capacity less than 1, but one can take the carrying capacity
equal to 1 without loss of generality.

We now outline an event-driven algorithm that mimic these rate equations.
Suppose that the system at a given time consists of $N_F$ full foragers,
$N_H$ hungry foragers, and $N_R$ individual resources.  The total number of
particles $N=N_F+N_H+N_R$.  There are six basic processes embodied by the
rate equations \eqref{RE}:
\begin{enumerate}
\item Reproduction:  $F\to F+1$
\item Starvation:  $F\to H$
\item Recruitment: $H\to F$
\item Death: $H\to H-1$
\item Resource growth: $R\to R+1$
\item Resource consumption: $R\to R-1$
\end{enumerate}

In an event-driven simulation, one of the above events is picked with the
appropriate probability (to be defined below), the selected event is
implemented, and then the time is updated accordingly.  The total rate for
any event is proportional to
\begin{equation}
\mathcal{R}=F\big[\lambda +\sigma(1-R)\big]+ H\big[\rho R+\mu\big]+
R\big[\alpha(1-R)+(F+H)\big]\,.
\end{equation}
Here $F=N_F/V$, $H=N_H/V$, and $R=N_R/V$, where $N_i$ is the total number of
particles of type $i$, and $V$ is the total number of lattice sites.  Thus $F$,
$H$, and $R$ are the densities of each type of entity in the system.  The
total rate is defined only up to an overall constant, but this constant is
immaterial because the time step is also proportional to this same constant.

Now we define the steps of an event-driven simulation.  
\begin{enumerate}
\item With probability $\lambda F/\mathcal{R}$ implement the reproduction step.
  This step consists of picking one of the full foragers and allowing it to
  reproduce.
\item With probability $\sigma(1-R)F/\mathcal{R}$ implement the starvation
  step; a full forager is picked and it changes to a hungry forager.
\item With probaibility $\rho HR/\mathcal{R}$ implement the recruitment step;
  a hungry forager is picked and becomes full.
\item With probability $\mu H/\mathcal{R}$ implement the death step; a hungry
  forager is picked and is removed. 
\item With probability $\alpha R(1-R)/\mathcal{R}$ implement the growth step;
  a unit of resource is created.
\item With probability $(F+H)R/\mathcal{R}$ implement the consumption step; a
  unit of resource is removed.
\end{enumerate}
By explicit construction, the probabilities for the above events sum to 1.

We now determine the change in the expected number of individuals of each
type in a single event.  For
full foragers, the change in its number $N_F$ is
\begin{subequations}
\begin{equation}
  \Delta N_F =\left[F\big(\lambda-\sigma(1-R)\big)+\rho H
    R\right]/\mathcal{R}\,.
\end{equation}
The term proportional to $F$ comes from processes in which a full forager is
picked, while the term proportional to $N_H$ comes from processes in which a
hungry forager is picked and converted to a full forager.  Thus the change in
the density of full foragers is
\begin{equation}
  \Delta F =\frac{\Delta N_F}{V} =\left[F\big(\lambda-\sigma(1-R)\big)+H\rho
    R\right]/V\mathcal{R}\,.
\end{equation}
\end{subequations}
Thus if we take the time step for each event to be
$\Delta t = (V\mathcal{R})^{-1}$, the above reduces to the rate equation
\eqref{RE} for $F$.  Thus in each microscopic event of the model, the time
should be advanced by $\Delta t = (V\mathcal{R})^{-1}$.

In a similar fashion, the change in the expected number of hungry foragers in
a single event is given by
\begin{subequations}
\begin{equation}
  \Delta N_H =\left[-H(\rho R+\mu)+F\sigma
    (1-R)\right]/\mathcal{R}\,,
\end{equation}
so that the change in the density of hungry foragers simply is
\begin{equation}
  \Delta H =\left[-H(\rho R+\mu)+F\sigma (1-R)\right]/V\mathcal{R}\,.
\end{equation}
\end{subequations}
Finally, the change in the expected number of individual resources in a
single event is given by
\begin{subequations}
\begin{equation}
  \Delta N_R =\left[R\big(\alpha(1-R)\big)-(F+H)\right]/\mathcal{R}\,,
\end{equation}
so that the change in the resource density is
\begin{equation}
  \Delta R =\left[R\big(\alpha(1-R)\big)-(F+H)\right]/V\mathcal{R}\,.
\end{equation}
\end{subequations}
The equations for $\Delta F$, $\Delta H$, and $\Delta R$ would then reproduce
the original rate equations \eqref{RE} when the time step for an elemental
event is taken to be $\Delta t = (V\mathcal{R})^{-1}$.  To summarize, pick an
event according to the probabilities enumerated above and update the
densities according (3b), (4b), and (5b), and then increment the time by
$(N\mathcal{R})^{-1}$ after each event.

There are several details of the simulation that need to be specified.
First, there is no constraint on the number of foragers on any site, but
given that we are describing a harsh environment, the parameters should be
chosen so that this number is not large.  The carrying capacity of the
resource has been set to 1, so that the number of individual resources at any
site should be either 0 or 1.  Finally, when a full forager reproduces, its
offspring should be placed anywhere.  With these details, a simulation that
is based on the above sprocedure should reproduce the predictions of the rate
equations \eqref{RE}.

We now generalize the above approach to the situation where the foragers are
diffusing on a lattice.  Suppose that the full and hungry foragers diffuse
with respective diffusion coefficients $D_F$ and $D_H$.  In this case, the
rate equations \eqref{RE} generalize to the set of partial differential
equations
\begin{align}
  \label{pde}
\begin{split}
\frac{\partial F}{\partial t} &= \lambda F + \rho  RH - \sigma (1-R)F+D_F\nabla^2F\,,\\[0.125in]
\frac{\partial H}{\partial t}  &= \sigma (1-R)F - \rho RH - \mu H+D_H\nabla^2H\,, \\[0.125in]
\frac{\partial R}{\partial t}  &= \alpha R(1-R) -  R(F+H).\\
\end{split}
\end{align}
Here the densities $F,H,R$ are now functions of space and time,
$F=F(\mathbf{r},t)$ and similarly for $H$ and $R$.

To construct an event-driven simulation that mimic these rate equations, we
have to include two additional processes:
\begin{enumerate}
\item[vii] Full forager diffusion
\item[viii] Hungry forager diffusion
\end{enumerate}
In an event-driven simulation, one of the eight possible events is picked
with the appropriate probability (see below), the selected event is
implemented, and the time is updated accordingly.

The total rate for any event is
\begin{equation}
\mathcal{R}=F\big[\lambda +\sigma(1-R)+D_F\big]+ H\big[\rho R+\mu+D_H\big]+
R\big[\alpha(1-R)+(F+H)\big]\,.
\end{equation}
Now we define the steps of an event-driven simulation.  
\begin{enumerate}
\item With probability $\lambda F/\mathcal{R}$ implement the reproduction step.
  This step consists of picking one of the full foragers and allowing it to
  reproduce.
\item With probability $\sigma(1-R)F/\mathcal{R}$ implement the starvation
  step; a full forager is picked and it changes to a hungry forager.
\item With probaibility $\rho HR/\mathcal{R}$ implement the recruitment step;
  a hungry forager is picked and becomes full.
\item With probability $\mu H/\mathcal{R}$ implement the death step; a hungry
  forager is picked and is removed. 
\item With probability $\alpha R(1-R)/\mathcal{R}$ implement the growth step;
  a unit of resource is created.
\item With probability $(F+H)R/\mathcal{R}$ implement the consumption step; a
  unit of resource is removed.
\item With probability $D_F F/\mathcal{R}$, a full forager moves; pick one
  such forager at random and move it.
\item With probability $D_H H/\mathcal{R}$, a full forager moves; pick one
  such forager at random and move it.
\end{enumerate}
By explicit construction, the probabilities for the above events sum to 1.

We now determine the change in the expected number of individuals of each
type at a given site $\mathbf{R}$ in a single event.  For full foragers, the
change in its number $N_F(\mathbf{R})$ at site $\mathbf{R}$ is
\begin{subequations}
\begin{equation}
  \Delta N_F(\mathbf{R}) =\left[F(\mathbf{R})\big(\lambda-\sigma(1-R(\mathbf{R}))\big)+\rho H
    R\right]/\mathcal{R}\,.
\end{equation}
The term proportional to $F$ comes from processes in which a full forager is
picked, while the term proportional to $N_H$ comes from processes in which a
hungry forager is picked and converted to a full forager.  Thus the change in
the density of full foragers is
\begin{equation}
  \Delta F =\frac{\Delta N_F}{V} =\left[F\big(\lambda-\sigma(1-R)\big)+H\rho
    R\right]/V\mathcal{R}\,.
\end{equation}
\end{subequations}
Thus if we take the time step for each event to be
$\Delta t = (V\mathcal{R})^{-1}$, the above reduces to the rate equation
\eqref{RE} for $F$.  Thus in each microscopic event of the model, the time
should be advanced by $\Delta t = (V\mathcal{R})^{-1}$.

In a similar fashion, the change in the expected number of hungry foragers in
a single event is given by
\begin{subequations}
\begin{equation}
  \Delta N_H =\left[-H(\rho R+\mu)+F\sigma
    (1-R)\right]/\mathcal{R}\,,
\end{equation}
so that the change in the density of hungry foragers simply is
\begin{equation}
  \Delta H =\left[-H(\rho R+\mu)+F\sigma (1-R)\right]/V\mathcal{R}\,.
\end{equation}
\end{subequations}
Finally, the change in the expected number of individual resources in a
single event is given by
\begin{subequations}
\begin{equation}
  \Delta N_R =\left[R\big(\alpha(1-R)\big)-(F+H)\right]/\mathcal{R}\,,
\end{equation}
so that the change in the resource density is
\begin{equation}
  \Delta R =\left[R\big(\alpha(1-R)\big)-(F+H)\right]/V\mathcal{R}\,.
\end{equation}
\end{subequations}
The equations for $\Delta F$, $\Delta H$, and $\Delta R$ would then reproduce
the original rate equations \eqref{RE} when the time step for an elemental
event is taken to be $\Delta t = (V\mathcal{R})^{-1}$.  To summarize, pick an
event according to the probabilities enumerated above and update the
densities according (3b), (4b), and (5b), and then increment the time by
$(N\mathcal{R})^{-1}$ after each event.

There are several details of the simulation that need to be specified.
First, there is no constraint on the number of foragers on any site, but
given that we are describing a harsh environment, the parameters should be
chosen so that this number is not large.  The carrying capacity of the
resource has been set to 1, so that the number of individual resources at any
site should be either 0 or 1.  Finally, when a full forager reproduces, its
offspring should be placed anywhere.  With these details, a simulation that
is based on the above sprocedure should reproduce the predictions of the rate
equations \eqref{RE}.

\end{document}

### 2016-02-05 foraging_simulations_diff.jl

@@ -5,7 +5,7 @@ using Cairo
 
 
 #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event.jl")
-include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_nodiff.jl")
+include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_eventdiff.jl")
 # include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_spatial.jl")
 
 
@@ -24,14 +24,16 @@ for i = 1:length(sigmavec)
   dim = 2;
   prop_fill = 0.5
   initsize = convert(Int64,round(((L-2)^dim)*prop_fill));
-  t_term = 500;
+  t_term = 120;
   alpha = 0.5;
   K = 1;
   sigma = sigmavec[i];
   rho = 0.2;
   lambda = 0.2;
   mu = 0.2;
-
+  DF = 1;
+  DH = 1;
+  DR = 1;
 
   eta = copy(sigma);
   Fstar[i] = (alpha*lambda*mu*(mu + rho))/((lambda + mu)*(lambda*rho + mu*sigma));
@@ -39,7 +41,7 @@ for i = 1:length(sigmavec)
   Rstar[i] = (mu*(-lambda+sigma))/(lambda*rho+mu*sigma);
 
   #The simulation
-  time_out, prop_out, N_out = starvingforager_event_nodiff(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu);
+  time_out, prop_out, N_out = starvingforager_eventdiff(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu,DF,DH,DR);
   F = prop_out[1,:];
   H = prop_out[2,:];
   R = prop_out[3,:];
@@ -51,7 +53,7 @@ for i = 1:length(sigmavec)
 
 end
 
-comparison = plot(
+comparisondiff = plot(
 layer(x=sigmavec,y=Fstar,Geom.line,Theme(default_color=colorant"green")),
 layer(x=sigmavec,y=Fsimstar,Geom.point,Theme(default_color=colorant"green",default_point_size=3pt)),
[…]

### 2016-02-05 src/starvingforager_event_rate.jl

@@ -69,7 +69,7 @@ function starvingforager_event_nodiff(L,dim,initsize,t_term,alpha,K,sigma,rho,la
   # F_pr_line = 0.0;
   # H_pr_line = 0.0;
   # R_pr_line = 0.0;
-
+  pr_line = zeros(6);
   # id = Array(Float64,1);
 
   while t < (t_term-1)
@@ -83,7 +83,7 @@ function starvingforager_event_nodiff(L,dim,initsize,t_term,alpha,K,sigma,rho,la
     # end
 
     #Calculate Rate
-    Rate = (1 - (N/S)) + F*(lambda + sigma*(K-R)) + H*(rho*R + mu) + R*(alpha*(K-R) + (F+H)); # (1 - (N/S)) +
+    Rate = F*(lambda + sigma*(K-R)) + H*(rho*R + mu) + R*(alpha*(K-R) + (F+H)); # (1 - (N/S)) +
 
     # TESTING
     # Rate = (NF/N)*(lambda + sigma*(K-R)) + (NH/N)*(rho*R + mu) + (NR/N)*(alpha*(K-R) + (F+H));
@@ -94,27 +94,6 @@ function starvingforager_event_nodiff(L,dim,initsize,t_term,alpha,K,sigma,rho,la
       break
     end
 
-    #Construct probability lines, which are a function of R, S, F
-    #Grow <-----> Starve <-----> Diffuse
-    #
-    # F_pr_line = lambda/Rate;
-    # #F_pr_line[2] = copy(F_pr_line[1]) + (sigma*(K-R))/Rate;
-    #
-    # #Recover <-----> Mortality <-----> Diffuse
-    #
-    # H_pr_line = (rho*R)/Rate;
-    # #H_pr_line[2] = copy(H_pr_line[1]) + mu/Rate;
-    #
-    # #Grow <-----> Consumed
-    # R_pr_line = (alpha*(K-R))/Rate;
-    # # R_pr_line[2] = copy(R_pr_line[1]) + (F+H)/Rate;
-
-    # 2/4/2016
-    # The probabilities for each event sequence need to add up to one...
-    F_pr_line = lambda/(lambda + sigma*(1-R));
-    H_pr_line = (rho*R)/(rho*R + mu);
-    R_pr_line = (alpha*(1-R))/(alpha*(1-R) + (H + F));
-
 
 
     #Randomly select an individual (R,S,F) with probability 1/N
[…]

### 2016-02-05 src/starvingforager_eventdiff.jl

@@ -1,4 +1,4 @@
-function starvingforager_event(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu,DF,DH,DR)
+function starvingforager_eventdiff(L,dim,initsize,t_term,alpha,K,sigma,rho,lambda,mu,DF,DH,DR)
   #Read in packages/function
   #ipbc :: torus movement
   #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")

### 2016-02-23 Manuscript/simulation-3.tex
major makeover to include correct rate eqns

\documentclass[11pt]{iopart}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{cite}
\usepackage{float}
\makeatletter
\makeatother

\expandafter\let\csname equation*\endcsname\relax
\expandafter\let\csname endequation*\endcsname\relax
\usepackage{amsmath}
\newcommand{\agt}{\mathrel{\raise.3ex\hbox{$>$\kern-.75em\lower1ex\hbox{$\sim$}}}}
\begin{document}

\title{Event-Driven Simulation of Foraging Dynamics}

\begin{abstract}

  We outline an event-driven simulation for the foraging model, which
  involves a resource that renews by logistic growth, as well as two classes
  of foragers---full and hungry.  Full foragers reproduce at a fixed rate and
  are not vulnerable to mortality.  However, a full forager can become hungry
  when resources are scarce; conversely, a hungry forager can become full
  when the resource is abundant.  Hungry foragers do not reproduce and die at
  a fixed rate.
\end{abstract}

%\maketitle

\section{The Model}

We assume that foragers can exist in two states---full and hungry.  When a
full forager $F$ encounters a resource $R$, the resource is consumed with
rate $m$ so that the forager maintains its fullness.  A full forager that
does not encounter a resource as it wanders converts to a hungry forager $H$
with rate equal to the product of the starvation rate $\sigma$ and the
density of non-resources.  Full foragers also reproduce with rate
$\lambda$. When a hungry forager encounters a resource, the resource is again
consumed and the forager becomes full with rate $\rho$.  A hungry forager
dies with mortality rate $\mu$, while full foragers have no mortality risk.
We assume that in the absence of foragers the underlying resource undergoes
logistic growth, with growth rate $\alpha$ and carrying capacity equal to 1.

Based on these processes and also under the assumption that the densities of
full foragers, hungry foragers, and resources (also denoted by $F$, $H$, and
$R$, respectively) are perfectly mixed, they evolve according to the rate
equations:
\begin{align}
  \label{RE}
\begin{split}
\dot F &= \lambda F + \rho  RH - \sigma (1-R)F\,,\\
\dot H &= \sigma (1-R)F - \rho RH - \mu H\,, \\
\dot R &= \alpha R(1-R) -  R(mF+\rho H),\\
\end{split}
\end{align}
where the overdot denotes time derivative.  An important feature is the
assumption that the carrying capacity is set equal to 1.  This means that
when the system is completely occupied by resources there can be no further
generation of the resource.  The approach described below can be generalized
to carrying capacity less than 1, but we take the carrying capacity equal to
1 without loss of generality.

We now outline an event-driven algorithm that mimic these rate equations.
Suppose that the system at a given time consists of $N_F$ full foragers,
$N_H$ hungry foragers, and $N_R$ resource units.  The total number of
particles is $N=N_F+N_H+N_R$.  There are six elemental processes that are
embodied by the rate equations \eqref{RE}:
\newpage
\begin{enumerate}
\item Reproduction:  $F\to 2F$
\item Starvation:  $F\to H$
\item Consumption: $F\to F$ and $R\to 0$
\item Death: $H\to 0$
\item Recruitment: $H\to F$ and $R\to 0$
\item Growth: $R\to 2R$
\end{enumerate}
For specificity, we define ``consumption'' as the consumption of the resource
by a full forager and the maintenance of the forager in its full state, while
``recruitment'' is defined as the consumption of the resource by a hungry
forager and conversion of the latter to the full state.

In an event-driven simulation, one of the above events is picked with the
appropriate probability to be defined below, the selected event is
implemented, and then the time is updated accordingly.  Reading off from the
steps (i)--(vi) listed above, the total rate for any event is proportional to
\begin{equation}
\mathcal{R}=F\big[\lambda +\sigma(1-R)+mR\big]+ H\big[\mu+\rho R\big]+
\alpha R(1-R)\,.
\end{equation}
Here $F=N_F/V$, $H=N_H/V$, and $R=N_R/V$, where $N_i$ is the total number of
particles of type $i$, and $V$ is the total number of lattice sites.  Thus
$F$, $H$, and $R$ are the densities of each type of entity.  The total rate
is defined up to an overall constant, but this constant is immaterial because
the time step is also proportional to this same constant.

In the most simple version of a simulation in the mean-field limit, one
merely tracks the three variables $N_F$, $N_H$, and $N_R$ and updates these
numbers according to the steps outlined below.  However, we want a simulation
that can be directly adapted to treat foragers that diffuse on a lattice.
The algorithm outlined below is based on populating a lattice with the three
types of entities and updating the occupancy on the lattice sites according
to the following event-driven simulation:
\begin{enumerate}
\item Reproduction.  With probability $\lambda F/\mathcal{R}$, create a new
  full forager anywhere in the system equiprobably, consistent with the
  absence of spatial structure.
\item Starvation.  With probability $\sigma(1-R)F/\mathcal{R}$, pick a random
  full forager and change it to hungry.  The factor $1-R$ is the fraction of
  resource-free space.
\item Consumption.  With probability $mFR/\mathcal{R}$, pick and remove a random
  resource unit.
\item Death.  With probability $\mu H/\mathcal{R}$, remove a random
  hungry forager.
\item Recruitment.  With probability $\rho HR/\mathcal{R}$, change a random
  hungry forager into a full forager and remove a random resource unit.
\item Growth.  With probability $\alpha R(1-R)/\mathcal{R}$, create another
  resource unit anywhere in the system.
\end{enumerate}
By explicit construction, the probabilities for the above events sum to 1.
It is important to appreciate that one can perform this simulation either by
populating a lattice with the three types of entities or merely be keeping
track of the three variables $N_F$, $N_H$, and $N_R$.  A good test of the
lattice-based simulation is that it should match that of simulating the above
three variables.

Let's determine the change in the expected number of individuals of each type
in a single event.  For full foragers, the change in its number $N_F$ is
\begin{subequations}
\begin{equation}
  \Delta N_F =\left[F\big(\lambda-\sigma(1-R)\big)+\rho H
    R\right]/\mathcal{R}\,.
\end{equation}
The term proportional to $F$ describes processes in which a full forager is
picked, while the term proportional to $H$ describes processes in which a
hungry forager is picked and converted to a full forager.  Thus the change in
the density of full foragers is
\begin{equation}
\label{dF}
  \Delta F =\frac{\Delta N_F}{V} =\left[F\big(\lambda-\sigma(1-R)\big)+\rho H
    R\right]/V\mathcal{R}\,.
\end{equation}
\end{subequations}
If we take the time step for each event to be
$\Delta t = (V\mathcal{R})^{-1}$, the above reduces to the rate equation
\eqref{RE} for $F$.  Thus for each microscopic update event, the time should
be advanced by $\Delta t = (V\mathcal{R})^{-1}$.

In a similar fashion, the change in the expected density of hungry foragers in
a single event is given by
%\begin{subequations}
%\begin{equation}
%  \Delta N_H =\left[-H(\rho R+\mu)+\sigma F
%    (1-R)\right]/\mathcal{R}\,,
%\end{equation}
%so that the change in the density of hungry foragers simply is
\begin{equation}
\label{dH}
  \Delta H =\left[-H(\rho R+\mu)+\sigma F(1-R)\right]/V\mathcal{R}\,,
\end{equation}
%\end{subequations}
and the change in the density of individual resources in a single event is
%\begin{subequations}
%\begin{equation}
%  \Delta N_R =\left[R\big(\alpha(1-R)\big)-(mF+\rho H)\right]/\mathcal{R}\,,
%\end{equation}
%so that the change in the resource density is
\begin{equation}
\label{dR}
  \Delta R =\left[R\big(\alpha(1-R)\big)-(mF+\rho H)\right]/V\mathcal{R}\,.
\end{equation}
%\end{subequations}
The expressions for $\Delta F/\Delta t$, $\Delta H/\Delta t$, and
$\Delta R/\Delta t$ reproduce the original rate equations \eqref{RE} when the
time step for an elemental update event is $\Delta t = (V\mathcal{R})^{-1}$.
To summarize, pick an event according to the probabilities above, update the
densities according \eqref{dF}, \eqref{dH}, and \eqref{dR}, and then
increment the time by $(V\mathcal{R})^{-1}$.  A simulation that is based on
the above steps should reproduce the predictions of the rate equations
\eqref{RE}.  In the three-variable simulation, keep only $N_F$, $N_H$, and
$N_R$.  In an update, pick an event according to the above probabilities, and
then update $N_F$, $N_H$, and $N_R$ and the time appropriately.

In implementing this algorithm for a lattice-based simulation, several points
deserve emphasis.  First, there is no constraint on the number of foragers on
any site, but if we are describing a harsh environment, the parameters should
be chosen so that this number is not large.  Since the carrying capacity of
the resource has been set to 1, $N_R$ cannot exceed $V$.  However, the number
of resource units on any given site can exceed 1.

We now generalize the above approach to the situation where the foragers are
diffusing on a lattice.  Suppose that the full and hungry foragers diffuse
with respective diffusion coefficients $D_F$ and $D_H$.  In this case, the
rate equations \eqref{RE} generalize to the reaction-diffusion equations
\begin{align}
  \label{pde}
\begin{split}
\frac{\partial F}{\partial t} &= \lambda F + \rho  RH - \sigma (1-R)F+D_F\nabla^2F\,,\\[0.125in]
\frac{\partial H}{\partial t}  &= \sigma (1-R)F - \rho RH - \mu H+D_H\nabla^2H\,, \\[0.125in]
\frac{\partial R}{\partial t}  &= \alpha R(1-R) -  R(m F+\rho H)\,.\\
\end{split}
\end{align}
Here the densities $F,H,R$ are now functions of space and time,
$F=F(\mathbf{r},t)$ and similarly for $H$ and $R$.

To construct an event-driven simulation that mimic these rate equations, we
have to include two additional processes:
\begin{enumerate}
\item[(vii)] Diffusion of full foragers
\item[(viii)] Diffusion of hungry foragers
\end{enumerate}
In an event-driven simulation, one of the eight possible events is picked
with the appropriate probability (see below), the selected event is
implemented, and the time is updated accordingly.
The total rate for any event is
\begin{align}
\label{R}
\mathcal{R}=&\frac{1}{V}\sum_{\mathbf{r}}\Big\{F(\mathbf{r})\big[\lambda
  +\sigma(1-R(\mathbf{r}))+D_F\big]
+ H(\mathbf{r})\big[\mu+\rho R(\mathbf{r})+D_H\big]\Big.\nonumber \\[-0.05in]
&\Big.\hskip 0.5in
  +\alpha R(\mathbf{r})(1-R(\mathbf{r'}))\Big\}\,\nonumber
  \\[0.1in]
&=F\big[\lambda +\sigma(1-R)+D_F\big]
+ H\big[\mu+\rho R+D_H\big]+\alpha R(1-R)\,,
\end{align}
in which the role of diffusion is now included.  Notice also the presence of
a two-particle correlation in the term $R(\mathbf{r})(1-R(\mathbf{r}'))$,
where $\mathbf{r'}$ is a neighbor of $\mathbf{r}$.  Thus going to the second
line of the above equation is not quite kosher, but below I give a
prescription to deal with this issue in a way that avoids using a correlation
function.

The steps of an event-driven simulation are:
\begin{enumerate}
\item Reproduction.  With probability $\lambda F/\mathcal{R}$, pick a full
  forager at random and let it reproduce.  The offspring should be placed
  close to the parent---either on the same site or on a neighboring site.
\item Starvation.  With probability $\sigma(1-R)F/\mathcal{R}$, pick a random
  full forager that is on a resource-free site and change the forager from
  full to hungry.  Note the two-particle nature of this rule: we need both
  the forager type and the resource state on a given site.  To avoid keeping
  track of sites that are simultaneously resource free and occupied by a full
  forager, a simpler alternative is:
\item[(ii')] Starvation'.  With probability $\sigma F/\mathcal{R}$, pick a
  random full forager.  If the site is resource free, change the forager from
  full to hungry.  If not, nothing happens.
\item Consumption.  With probability $mFR/\mathcal{R}$, pick a random on a
  site that is also by a full forager and remove the resource.  To avoid
  tracking sites that simultaneously contain a resource and a full forager,
  perform the following alternative:
\item[(iii$^\prime$)] Consumption$^\prime$.  With probability
  $R/\mathcal{R}$, pick a random resource.  If the site also contains a
  forager, remove the resource with probability $m$.  Note the implicit
  assumption that the density of foragers is not large because the rate of
  resource removal does not increase if multiple foragers occupy the site.
  If the full forager density is large then the following works:
\item[(iii$^{\prime\prime}$)] Consumption$^{\prime\prime}$.  With probability
  $m F/\mathcal{R}$, pick a random full forager.  If the site also contains a
  resource, remove it.  This prescription works if the forager density is
  small or large (as long as the resource density is less than 1) so it's
  best to use this last alternative.
\item Death.  With probability $\mu H/\mathcal{R}$, pick and remove a random
  hungry forager.
\item Recruitment.  With probability $\rho HR/\mathcal{R}$, pick a random
  hungry forager on a site that also contains a resource and change the
  forager from hungry to full and remove the resource.  This process also
  involves a two-body reaction and a simpler way to implement this two-body
  feature is:
\item[(v$^\prime$)] Recruitment$^\prime$.  With probability
  $\rho H/\mathcal{R}$, pick a random hungry forager.  If the site contains a
  resource, change the forager from hungry to full and remove the resource.
\item Growth.  With probability $\alpha R(1-R)/\mathcal{R}$, pick a random
  neighboring pair of sites in which one contains a resource and the neighbor
  does not.  Create a resource on the site without the resource.  A
  computationally simpler alternative is:
\item[(vi$^\prime$)] Growth$^\prime$.  With probability
  $\alpha R/\mathcal{R}$, pick a random resource unit.  A new resource unit
  is created on one of the neighboring sites with probability $1-m/z$, where
  $z$ is the number of nearest-neighbors and $m$ is the number of these
  neighbors that are occupied by resource.  If the local neighborhood is
  devoid of resources, growth necessarily occurs at one of these neighbors,
  while if the local neighborhood is full of resources, there is no growth.
  This rule ensures that the growth rate is proportional to
  $(1-R)_{\rm local}$.
\item Diffusion.  With probability $D_F\, F/\mathcal{R}$, pick a random full
  forager and move it to a neighboring site.
\item Diffusion.  With probability $D_H\, H/\mathcal{R}$, pick a random
  hungry forager and move it to a neighboring site.
\end{enumerate}
By explicit construction, the probabilities for the above events sum to 1.
Note also that the primed steps are computationally simpler than the
corresponding unprimed steps, but have the possibility of a null step; this
introduces some inefficiency.  A true event-driven simulation avoids all null
steps, but perhaps the computational tradeoff in bookkeeping simplicity is
worth the loss of efficiency in allowing for null steps.  This is somewhat a
matter of user taste and testing.

We now determine the change in the expected number of individuals of each
type at a given site $\mathbf{r}$ in a single event.  For full foragers, the
change in its number $N_F(\mathbf{r})$ at site $\mathbf{r}$ is
\begin{subequations}
\begin{equation}
  \Delta N_F(\mathbf{r})  =\Big\{F(\mathbf{r})\big(\lambda-\sigma(1-R(\mathbf{r}))\big)+
\rho H(\mathbf{r})  R(\mathbf{r})
+D_F\big[\sum_{\mathbf{r}'}F(\mathbf{r}')-zF(\mathbf{r})\big]\Big\}/\mathcal{R}\,.
\end{equation}
The terms proportional to $F$ accounts for processes in which a full forager
is picked, while the term proportional to $H$ accounts for processes in which
a hungry forager is picked.  The terms proportional to $D_F$ account for
nearest-neighbor hopping: $F(\mathbf{r})$ increases because full foragers at
neighboring sites $\mathbf{r}'$ hop to $\mathbf{r}$, while $F(\mathbf{r})$
decreases because full foragers at $\mathbf{r}$ hop to one of the $z$
neighboring sites.  Thus the change in the density of full foragers at site
$\mathbf{r}$ is (after taking the continuum limit for nearest-neighbor
hopping)
\begin{equation}
  \Delta F(\mathbf{r}) =\frac{\Delta N_F(\mathbf{r})}{V} 
  =\left[F(\mathbf{r})\big(\lambda-\sigma(1-R(\mathbf{r}))\big)+\rho H(\mathbf{r})
    R(\mathbf{r})+\nabla^2F(\mathbf{r})\right]/V\mathcal{R}\,.
\end{equation}
\end{subequations}
If we take the time step for each event to be
$\Delta t = (V\mathcal{R})^{-1}$, the above reduces to the partial
differential evolution equation \eqref{pde} for $F(\mathbf{r})$.  In each
microscopic event of the model, the time should be advanced by
$\Delta t = (V\mathcal{R})^{-1}$.

In a similar fashion, the change in the density of hungry foragers is
\begin{equation}
  \Delta H(\mathbf{r}) =\left[-H(\mathbf{r})(\rho R(\mathbf{r})+\mu)+F(\mathbf{r})\sigma (1-R(\mathbf{r}))+\nabla^2H(\mathbf{r})\right]/V\mathcal{R}\,.
\end{equation}
Finally, the change in the resource density is
\begin{equation}
  \Delta R(\mathbf{r}) =\left[R(\mathbf{r})\big(\alpha(1-R(\mathbf{r}))\big)-(F(\mathbf{r})+H(\mathbf{r}))\right]/V\mathcal{R}\,.
\end{equation}
The equations for $\Delta F$, $\Delta H$, and $\Delta R$ would then reproduce
the partial differential equations \eqref{pde} when the time step for an
elemental event is taken to be $\Delta t = (V\mathcal{R})^{-1}$.

In this algorithm, notice that single-particle and two-particle reactions are
treated differently.  In the former case, a particle of the right type is
picked at random.  In the latter case, a random \emph{pair} of particles of
the appropriate types is picked.  Again, there is no constraint on the number
of foragers on any site, but given that we are describing a harsh
environment, the parameters should be chosen so that this number is not
large.  The carrying capacity of the resource has been set to 1, so that the
number of individual resources at any site must be either 0 or 1.

It would be nice if there was a simple way to adjust parameters in the above
lattice simulation so as to recover the mean-field limit.  I describe a
simple modification of the above lattice simulation steps that should recover
the mean-field limit.  We use the same rate as in Eq.~\eqref{R}, but the
actual update steps are slightly modified compared to the rules of the
lattice simulation above.  The differences are highlighted in italics.  I'm
also assuming that the simulation with steps (ii$^\prime$),
(iii$^\prime\prime$), (v$^\prime$), and (vi$^\prime$) are being used.
\begin{enumerate}
\item Reproduction.  With probability $\lambda F/\mathcal{R}$, pick a random
  full forager and let it reproduce.  \emph{Place the offspring anywhere.}
\item Starvation.  With probability $\sigma F/\mathcal{R}$, \emph{change a
    random full forager to hungry with probability $(1-R)$.}
\item Consumption.  With probability $m F/\mathcal{R}$, pick a random full
  forager.  If the site is also occupied by a resource, remove it.
\item Death.  With probability $\mu H/\mathcal{R}$, pick and remove a random
  hungry forager.
\item Recruitment.  With probability $\rho H/\mathcal{R}$, pick a random
  hungry forager. \emph{Change it to full with probability $R$ and also
    remove a random resource.}
\item Growth.  With probability $\alpha R/\mathcal{R}$, pick a random
  resource.  Create a resource anywhere at random with probability $1-R$.
\item $F$ Diffusion.  With probability $D_F\, F/\mathcal{R}$, pick a random full
  forager and move it to a neighboring site.
\item $H$ Diffusion.  With probability $D_H\, H/\mathcal{R}$, pick a random
  hungry forager and move it to a neighboring site.
\end{enumerate}
By explicit construction, the probabilities for the above events sum to 1.

It is important to appreciate that there are multiple ways to simulate the
elemental reaction steps that involve a tradeoff between bookkeeping
complexity and computational efficiency and still reproduce the same behavior
as in the rate equations \eqref{RE} or the reaction-diffusion equations
\eqref{pde} in the limit of a large system.

\end{document}

### 2016-02-23 foraging_simulations.jl

@@ -5,7 +5,7 @@ using Cairo
 
 
 #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event.jl")
-include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_nodiff.jl")
+include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_rate.jl")
 # include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_spatial.jl")
 
 
@@ -24,12 +24,12 @@ for i = 1:length(sigmavec)
   dim = 2;
   prop_fill = 0.5
   initsize = convert(Int64,round(((L-2)^dim)*prop_fill));
-  t_term = 500;
+  t_term = 80;
   alpha = 0.5;
   K = 1;
   sigma = sigmavec[i];
   rho = 0.2;
-  B = 0.2;
+  m = 0.2;
   lambda = 0.2;
   mu = 0.2;
 
@@ -40,7 +40,7 @@ for i = 1:length(sigmavec)
   Rstar[i] = (mu*(-lambda+sigma))/(lambda*rho + mu*sigma);
 
   #The simulation
-  time_out, prop_out, N_out = starvingforager_event_nodiff(L,dim,initsize,t_term,alpha,K,sigma,rho,B,lambda,mu);
+  time_out, prop_out, N_out = starvingforager_event_rate(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu);
   F = prop_out[1,:];
   H = prop_out[2,:];
   R = prop_out[3,:];

### 2016-02-23 src/starvingforager_event_rate.jl

@@ -1,4 +1,4 @@
-function starvingforager_event_nodiff(L,dim,initsize,t_term,alpha,K,sigma,rho,B,lambda,mu)
+function starvingforager_event_rate(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu)
   #Read in packages/function
   #ipbc :: torus movement
   #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
@@ -9,13 +9,16 @@ function starvingforager_event_nodiff(L,dim,initsize,t_term,alpha,K,sigma,rho,B,
 
   S = (L-2)^dim;
 
+  #intisize after rounding
+  r_initsize = Int(round(initsize/3));
 
   #Initial state values
-  Find_vec = zeros(Int64,Int(round(initsize/3))) + 2;
-  Hind_vec = zeros(Int64,Int(round(initsize/3))) + 1;
-  Rind_vec = zeros(Int64,Int(round(initsize/3)));
+  Find_vec = zeros(Int64,r_initsize) + 2;
+  Hind_vec = zeros(Int64,r_initsize) + 1;
+  Rind_vec = zeros(Int64,r_initsize);
 
-  ind_vec = [Find_vec;Hind_vec;Rind_vec];
+  #Join the three types... not sure if I will need to use this?
+  #ind_vec = [Find_vec;Hind_vec;Rind_vec];
   #OR Pure Resource Start (for testing)
   #ind_vec = zeros(Int64,initsize);
 
@@ -24,14 +27,12 @@ function starvingforager_event_nodiff(L,dim,initsize,t_term,alpha,K,sigma,rho,B,
   #2/3 of initsize will be foragers; 1/3 will be resources
   #replace = false because want only one resource per site
 
-  Floc_vec = sample(collect(1:S),Int(round(initsize/3)));
-  Hloc_vec = sample(collect(1:S),Int(round(initsize/3)));
-  Rloc_vec = sample(collect(1:S),Int(round(initsize/3)),replace=false);
+  Floc_vec = sample(collect(1:S),r_initsize);
+  Hloc_vec = sample(collect(1:S),r_initsize);
+  #Use replace=false to ensure that no location is chosen twice for resources only
+  Rloc_vec = sample(collect(1:S),r_initsize,replace=false);
 
-  loc_vec = [Floc_vec;Hloc_vec;Rloc_vec];
-
[…]

### 2016-02-23 foraging_simulations_spatial.jl
added spatial code - not ready - and organized old code

@@ -5,7 +5,7 @@ using Cairo
 
 
 #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event.jl")
-include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_rate.jl")
+include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_rate_spatial.jl")
 # include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_spatial.jl")
 
 
@@ -29,9 +29,9 @@ for i = 1:length(sigmavec)
   K = 1;
   sigma = sigmavec[i];
   rho = 0.2;
-  m = 0.8;
+  m = 0.2;
   lambda = 0.2;
-  mu = 0.1;
+  mu = 0.2;
 
 
   eta = copy(sigma);
@@ -40,7 +40,7 @@ for i = 1:length(sigmavec)
   Rstar[i] = (mu*(-lambda+sigma))/(lambda*rho + mu*sigma);
 
   #The simulation
-  time_out, prop_out, N_out = starvingforager_event_rate(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu);
+  time_out, prop_out, N_out = starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu);
   F = prop_out[1,:];
   H = prop_out[2,:];
   R = prop_out[3,:];

### 2016-02-23 src/starvingforager_event_rate_spatial.jl

@@ -1,0 +1,266 @@
+function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu)
+  #Read in packages/function
+  #ipbc :: torus movement
+  #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
+
+
+  #Initiate values
+  #Lattice dimension
+
+  S = (L-2)^dim;
+
+  #intisize after rounding
+  r_initsize = Int(round(initsize/3));
+
+  #Initial state values
+  Find_vec = zeros(Int64,r_initsize) + 2;
+  Hind_vec = zeros(Int64,r_initsize) + 1;
+  Rind_vec = zeros(Int64,r_initsize);
+
+  #Join the three types... not sure if I will need to use this?
+  #ind_vec = [Find_vec;Hind_vec;Rind_vec];
+  #OR Pure Resource Start (for testing)
+  #ind_vec = zeros(Int64,initsize);
+
+
+  #Initial location values
+  #2/3 of initsize will be foragers; 1/3 will be resources
+  #replace = false because want only one resource per site
+
+  Floc_vec = sample(collect(1:S),r_initsize);
+  Hloc_vec = sample(collect(1:S),r_initsize);
+  #Use replace=false to ensure that no location is chosen twice for resources only
+  Rloc_vec = sample(collect(1:S),r_initsize,replace=false);
+
+  #loc_vec = [Floc_vec;Hloc_vec;Rloc_vec];
+
+  #Arrays for output (start out empty)
+  # ind_out = (Array{Int64,1})[];
+  # loc_out = (Array{Int64,1})[];
+  time_out = Array(Float64,0);
+  N_out = Array(Int64,0);
+  #prop_out = (Array{Float64,1})[];
+
+  #Copy so that ind_vec is independent of ind_out
+  # push!(ind_out, copy(ind_vec));
+  # push!(loc_out, copy(loc_vec));
+  push!(time_out, 0);
+
+  #NEED TO ENSURE THAT RESOURCES ARE PLACED 1 PER SITE
+  #HAVE A noresourcesites vector that accounts for sites WITHOUT resources
+  #Updated each timestep
+  noresourcesites = collect(1:S);
+  #Deletes locations of sites with resources from the list of open sites (noresourcesites)
+  deleteat!(noresourcesites,sort(Rloc_vec));
+
+  t = 0;
[…]

### 2016-02-23 foraging_simulations.jl
eliminated the ind_vec vectors to speed things up

@@ -5,7 +5,7 @@ using Cairo
 
 
 #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event.jl")
-include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_rate.jl")
+include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_rate_spatial.jl")
 # include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_spatial.jl")
 
 
@@ -40,7 +40,7 @@ for i = 1:length(sigmavec)
   Rstar[i] = (mu*(-lambda+sigma))/(lambda*rho + mu*sigma);
 
   #The simulation
-  time_out, prop_out, N_out = starvingforager_event_rate(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu);
+  time_out, prop_out, N_out = starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu);
   F = prop_out[1,:];
   H = prop_out[2,:];
   R = prop_out[3,:];

### 2016-02-23 foraging_simulations_spatial.jl

@@ -29,9 +29,9 @@ for i = 1:length(sigmavec)
   K = 1;
   sigma = sigmavec[i];
   rho = 0.2;
-  m = 0.2;
+  m = 0.8;
   lambda = 0.2;
-  mu = 0.2;
+  mu = 0.1;
 
 
   eta = copy(sigma);

### 2016-02-23 src/starvingforager_event_rate.jl

@@ -1,4 +1,4 @@
-function starvingforager_event_rate(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu)
+function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu)
   #Read in packages/function
   #ipbc :: torus movement
   #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
@@ -13,14 +13,27 @@ function starvingforager_event_rate(L,dim,initsize,t_term,alpha,K,sigma,rho,m,la
   r_initsize = Int(round(initsize/3));
 
   #Initial state values
-  Find_vec = zeros(Int64,r_initsize) + 2;
-  Hind_vec = zeros(Int64,r_initsize) + 1;
-  Rind_vec = zeros(Int64,r_initsize);
+  # Find_vec = zeros(Int64,r_initsize) + 2;
+  # Hind_vec = zeros(Int64,r_initsize) + 1;
+  # Rind_vec = zeros(Int64,r_initsize);
 
-  #Join the three types... not sure if I will need to use this?
-  #ind_vec = [Find_vec;Hind_vec;Rind_vec];
-  #OR Pure Resource Start (for testing)
-  #ind_vec = zeros(Int64,initsize);
+  time_out = Array(Float64,0);
+  N_out = Array(Int64,0);
+
+  NF = r_initsize;
+  NH = r_initsize;
+  NR = r_initsize;
+
+  N = NF + NH + NR;
+
+  #Initial densities
+  F = NF/S;
+  H = NH/S;
+  R = NR/S;
+  prop_out = Array{Float64}(3,1);
+  prop_out[:,1] = [F,H,R];
+  push!(N_out,N);
+
 
 
   #Initial location values
@@ -32,18 +45,7 @@ function starvingforager_event_rate(L,dim,initsize,t_term,alpha,K,sigma,rho,m,la
   #Use replace=false to ensure that no location is chosen twice for resources only
   Rloc_vec = sample(collect(1:S),r_initsize,replace=false);
 
-  #loc_vec = [Floc_vec;Hloc_vec;Rloc_vec];
 
-  #Arrays for output (start out empty)
-  # ind_out = (Array{Int64,1})[];
-  # loc_out = (Array{Int64,1})[];
-  time_out = Array(Float64,0);
-  N_out = Array(Int64,0);
-  #prop_out = (Array{Float64,1})[];
-
[…]

### 2016-02-23 src/starvingforager_event_rate_spatial.jl

@@ -13,14 +13,27 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
   r_initsize = Int(round(initsize/3));
 
   #Initial state values
-  Find_vec = zeros(Int64,r_initsize) + 2;
-  Hind_vec = zeros(Int64,r_initsize) + 1;
-  Rind_vec = zeros(Int64,r_initsize);
+  # Find_vec = zeros(Int64,r_initsize) + 2;
+  # Hind_vec = zeros(Int64,r_initsize) + 1;
+  # Rind_vec = zeros(Int64,r_initsize);
 
-  #Join the three types... not sure if I will need to use this?
-  #ind_vec = [Find_vec;Hind_vec;Rind_vec];
-  #OR Pure Resource Start (for testing)
-  #ind_vec = zeros(Int64,initsize);
+  time_out = Array(Float64,0);
+  N_out = Array(Int64,0);
+
+  NF = r_initsize;
+  NH = r_initsize;
+  NR = r_initsize;
+
+  N = NF + NH + NR;
+
+  #Initial densities
+  F = NF/S;
+  H = NH/S;
+  R = NR/S;
+  prop_out = Array{Float64}(3,1);
+  prop_out[:,1] = [F,H,R];
+  push!(N_out,N);
+
 
 
   #Initial location values
@@ -32,18 +45,7 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
   #Use replace=false to ensure that no location is chosen twice for resources only
   Rloc_vec = sample(collect(1:S),r_initsize,replace=false);
 
-  #loc_vec = [Floc_vec;Hloc_vec;Rloc_vec];
 
-  #Arrays for output (start out empty)
-  # ind_out = (Array{Int64,1})[];
-  # loc_out = (Array{Int64,1})[];
-  time_out = Array(Float64,0);
-  N_out = Array(Int64,0);
-  #prop_out = (Array{Float64,1})[];
-
-  #Copy so that ind_vec is independent of ind_out
-  # push!(ind_out, copy(ind_vec));
-  # push!(loc_out, copy(loc_vec));
   push!(time_out, 0);
 
   #NEED TO ENSURE THAT RESOURCES ARE PLACED 1 PER SITE
@@ -56,21 +58,7 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
   t = 0;
   next_time_int = t + 1;
 
[…]

### 2016-02-24 foraging_simulations_spatial.jl
changes to spatial

@@ -32,6 +32,8 @@ for i = 1:length(sigmavec)
   m = 0.8;
   lambda = 0.2;
   mu = 0.1;
+  Df = 0.2;
+  Dh = 0.2;
 
 
   eta = copy(sigma);
@@ -40,7 +42,7 @@ for i = 1:length(sigmavec)
   Rstar[i] = (mu*(-lambda+sigma))/(lambda*rho + mu*sigma);
 
   #The simulation
-  time_out, prop_out, N_out = starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu);
+  time_out, prop_out, N_out = starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu,Df,Dh);
   F = prop_out[1,:];
   H = prop_out[2,:];
   R = prop_out[3,:];

### 2016-02-24 src/starvingforager_event_rate_spatial.jl

@@ -1,4 +1,4 @@
-function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu)
+function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu,Df,Dh)
   #Read in packages/function
   #ipbc :: torus movement
   #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
@@ -79,7 +79,7 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
     # end
 
     #Calculate Rate
-    Rate = F*(lambda + sigma*(K-R)) + H*(rho*R + mu) + R*(alpha*(K-R) + (rho*H + m*F)); # (1 - (N/S)) +
+    Rate = F*(lambda + sigma*(K-R) + Df) + H*(rho*R + mu + Dh) + R*(alpha*(K-R) + (rho*H + m*F)); # (1 - (N/S)) +
 
     # TESTING
     # Rate = (NF/N)*(lambda + sigma*(K-R)) + (NH/N)*(rho*R + mu) + (NR/N)*(alpha*(K-R) + (F+H));
@@ -98,17 +98,22 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
     pr_line = zeros(6);
     #Update the total
     #Events
+    #1 F growth
     pr_line[1] = (lambda*F)/Rate;
     #2  Starvation
     pr_line[2] = pr_line[1] + (sigma*(K-R)*F)/Rate;
+    #3  F Movement
+    pr_line[3] = pr_line[2] + (Df*F)/Rate;
     #3  Recruitment
-    pr_line[3] = pr_line[2] + (rho*H*R)/Rate;
+    pr_line[4] = pr_line[3] + (rho*H*R)/Rate;
     #4  Death
-    pr_line[4] = pr_line[3] + (mu*H)/Rate;
+    pr_line[5] = pr_line[4] + (mu*H)/Rate;
+    #5  H Movement
+    pr_line[6] = pr_line[5] + (Dh*H)/Rate;
     #5  Resource Growth
-    pr_line[5] = pr_line[4] + (alpha*R*(K-R))/Rate;
[…]

### 2016-02-25 foraging_simulations.jl
first draft of spatial sim

@@ -5,7 +5,7 @@ using Cairo
 
 
 #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event.jl")
-include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_rate_spatial.jl")
+include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_rateV.jl")
 # include("$(homedir())/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/starvingforager_event_spatial.jl")
 
 
@@ -33,14 +33,12 @@ for i = 1:length(sigmavec)
   lambda = 0.2;
   mu = 0.1;
 
-
-  eta = copy(sigma);
   Fstar[i] = (alpha*lambda*mu*(mu + rho))/((lambda*rho + m*mu)*(lambda*rho + sigma*mu));
   Hstar[i] = (alpha*lambda^2*(mu + rho))/((lambda*rho + m*mu)*(lambda*rho + sigma*mu));
   Rstar[i] = (mu*(-lambda+sigma))/(lambda*rho + mu*sigma);
 
   #The simulation
-  time_out, prop_out, N_out = starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu);
+  time_out, prop_out, N_out = starvingforager_event_rateV(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu);
   F = prop_out[1,:];
   H = prop_out[2,:];
   R = prop_out[3,:];

### 2016-02-25 src/starvingforager_event_rateV.jl

@@ -1,4 +1,4 @@
-function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu,Df,Dh)
+function starvingforager_event_rateV(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu)
   #Read in packages/function
   #ipbc :: torus movement
   #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
@@ -45,8 +45,8 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
   #Use replace=false to ensure that no location is chosen twice for resources only
   Rloc_vec = sample(collect(1:S),r_initsize,replace=false);
 
-  #Record the location of F/H/R in this vector, where each element is a site.
-  #F and H can have more than one element in a site; R can only have 1.
+
+  #Record the location of resources in this vector, where each element is a site.
   Fsite_vec = zeros(S);
   Hsite_vec = zeros(S);
   for i=1:r_initsize
@@ -58,6 +58,13 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
   Rsite_vec[Rloc_vec] = 1;
 
   push!(time_out, 0);
+
+  #NEED TO ENSURE THAT RESOURCES ARE PLACED 1 PER SITE
+  #HAVE A noresourcesites vector that accounts for sites WITHOUT resources
+  #Updated each timestep
+  noresourcesites = collect(1:S);
+  #Deletes locations of sites with resources from the list of open sites (noresourcesites)
+  deleteat!(noresourcesites,sort(Rloc_vec));
 
   t = 0;
   next_time_int = t + 1;
@@ -83,7 +90,7 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
     # end
 
     #Calculate Rate
-    Rate = F*(lambda + sigma*(K-R) + Df) + H*(rho*R + mu + Dh) + R*(alpha*(K-R) + (rho*H + m*F)); # (1 - (N/S)) +
+    Rate = F*(lambda + sigma*(K-R)) + H*(rho*R + mu) + R*(alpha*(K-R) + (rho*H + m*F)); # (1 - (N/S)) +
 
     # TESTING
[…]

### 2016-02-25 src/starvingforager_event_rate_spatial.jl
problem with ipbc... make sure this is working correctly

@@ -1,7 +1,7 @@
 function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,rho,m,lambda,mu,Df,Dh)
   #Read in packages/function
   #ipbc :: torus movement
-  #include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
+  include("/Users/justinyeakel/Dropbox/PostDoc/2014_DiffusingForager/DiffusingForager/src/ipbc.jl")
 
 
   #Initiate values
@@ -59,6 +59,14 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
 
   push!(time_out, 0);
 
+  #NEED TO ENSURE THAT RESOURCES ARE PLACED 1 PER SITE
+  #HAVE A noresourcesites vector that accounts for sites WITHOUT resources
+  #Updated each timestep
+  noresourcesites = collect(1:S);
+  #Deletes locations of sites with resources from the list of open sites (noresourcesites)
+  deleteat!(noresourcesites,sort(Rloc_vec));
+
+
   t = 0;
   next_time_int = t + 1;
 
@@ -99,7 +107,7 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
     #Randomly select an individual (R,S,F) with probability 1/N
     #ind thus represents the POSITION of the individual
     #Update total number of individuals
-    pr_line = zeros(6);
+    pr_line = zeros(8);
     #Update the total
     #Events
     #1 F growth
@@ -149,7 +157,7 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
 
       #update site vectors
       Fsite_vec[location] -= 1;
-      Hsite_vec[locaiton] += 1;
+      Hsite_vec[location] += 1;
 
       #Update
       NF = NF - 1;
@@ -162,13 +170,13 @@ function starvingforager_event_rate_spatial(L,dim,initsize,t_term,alpha,K,sigma,
       id = rand(collect(1:NF));
       location = Floc_vec[id];
 
-      nn = ipbc(location,(L-2));
+      nn = ipbc(location,L);
       new_location = rand(nn);
       Floc_vec[id] = new_location;
 
       #update site vectors
       Fsite_vec[location] -= 1;
-      Fsite_vec[new_locaiton] += 1;
[…]

### 2016-03-23 Manuscript/Yeakel_Redner_2015.tex
hopf figures updates

@@
 
-The behaviorial ecology of most, if not all, organisms is influenced by the energetic state of individuals.
+The behavioral ecology of most, if not all, organisms is influenced by the energetic state of individuals from which the population is composed.
 An individual's energetic state, which can range from starving (energetically defficient) to satiated (energetially replete), directly influences how it invests its stores in an uncertain environment.
@@
 We first integrate energetics into the dynamics of a consumer-resource system by assuming that the consumer population can be divided into discrete energetic states, the occupation of each being contingent on the consumption of a single resource $R$.
-Throughout, we assume that the resource has logistic growth with a linear growth rate $\alpha$, a carrying capacity of unity, and an extrinsic mortality due to consumption at rate $\epsilon$.
-In the simplest case, there are only two energetic states for the consumer population: \emph{i}) an energetically replete state $Y$, where the consumer reproduces at rate $\delta \epsilon$ and suffers mortality at rate $\mu_y$, and \emph{ii}) an energetically deficient state $X$, where reproduction is suppressed, and mortality occurs at rate $\mu_x$.
-Because an energetically defficient state is generally more risky, we assume that $\mu_x > \mu_y$.
-Consumers transition from state $Y$ to state $X$ by starvation at rate $p$, which is proportional to $1-R$.
-Conversely, consumers recover from state $X$ to the energetically replete state $Y$ in proportion to the density of resources consumed $\epsilon R$.
+In the simplest case, there are only two energetic states for the consumer population: \emph{i}) an energetically replete state $F$, where the consumer reproduces at rate $\lambda$, and \emph{ii}) an energetically deficient state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
+Consumers transition from state $F$ to state $H$ by starvation at rate $\sigma$ and in proportion to the lack of food $1-R$.
+Conversely, consumers recover from state $H$ to the energetically replete state $F$ at rate $\rho$ and in proportion to the density of resources consumed $R$.
+The resource has logistic growth with a linear growth rate $a$ and a carrying capacity of unity.
+Resources are eliminated by the consumer in either state: by energetically deficient consumers at rate $\rho$, and by energetically replete consumers at rate $b$.
 Accordingly, the system of equations is written
@@
 \begin{align}
-\frac{\rm d}{\rm dt} R &= \alpha R(1-R) - \epsilon R(X+Y), \nonumber \\ \nonumber \\
-\frac{\rm d}{\rm dt} X &= \sigma Y(1-R) - \rho \epsilon XR - \mu X, \nonumber \\ \nonumber \\
-\frac{\rm d}{\rm dt} Y &= \gamma \epsilon YR + \rho \epsilon XR - \sigma Y(1-R)
+\frac{\rm d}{\rm dt} F &= \lambda F + \rho HR - \sigma (1-R)F, \nonumber \\
+\frac{\rm d}{\rm dt} H &= \sigma (1-R)F - \rho RH - \mu H, \nonumber \\
+\frac{\rm d}{\rm dt} R &= a R(1-R) - R(\rho H+ b F).
 \end{align}
 
-There are four steady states for the 2-stage consumer-resource system: two trivial steady states at $(R^*=0,X^*=0,Y^*=0)$ and $(R^*=1,X^*=0,Y^*=0)$, one non-trivial external (negative) steady state, and one non-trivial internal steady state where $(R^*>0,X^*>0,Y^*>0)$.
-The latter steady state is the one of interest here, and although there is a tractible solution, it is cannot be written in a compressed form.
-Moreover, because there is only one internal steady state, as long as it is stable the population trajectories will be globally attracted to it for any set of initial conditions greater than zero.
+There are three steady states for the 2-stage consumer-resource system: two trivial steady states at $(R^*=0,H^*=0,F^*=0)$ and $(R^*=1,H^*=0,F^*=0)$, and one non-trivial internal steady state where $(R^*>0,H^*>0,F^*>0)$.
+The latter steady state is the one of chief ecological interest, where
+
+\begin{align}
+F^* &= \frac{a  \lambda  \mu  (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  m)}, \nonumber \\
+H^* &= \frac{a  \lambda ^2 (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  m)}, \nonumber \\
+R^* &= \frac{\mu  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma }.	
+\end{align}
+
+\noindent Because there is only one internal steady state, as long as it is stable the population trajectories will be globally attracted to it for any set of initial conditions greater than zero.
 
@@
 In a multidimensional system, linear stability is determined with respect to the Jacobian Matrix $\bf J$, which is a matrix where each element is defined by the partial derivative of each equation with respect to each variable.
-In the case of the 2-stage consumer model, the Jacobian is written
+In the case of the 2-stage consumer model, the Jacobian evaluated at the internal steady state (denoted by $|_*$) is written
 
@@
 \mathbf{J}|_* =
+%\left(
+%\begin{array}{ccc}
+% -F^* m+(1-R^*) a -R^* a -H^* \rho  & -R^* \rho  & -m R^* \\
+% -H^* \rho -F^* \sigma  & -\mu -R^* \rho  & (1-R^*) \sigma  \\
+% H^* \rho +F^* \sigma  & R^* \rho  & \lambda -(1-R^*) \sigma  \\
+%\end{array}
+%\right).
 \left(
 \begin{array}{ccc}
-\alpha(1-2 R^*)-\epsilon (X^*-Y^*) & -\epsilon R^* & -\epsilon R^* \\
--\rho \epsilon X^*-\sigma Y^* & -\epsilon \rho R^*-\mu & \sigma(1-R^*) \\
-\gamma \epsilon Y^*+\sigma Y^* +\rho \epsilon X^* & \rho \epsilon R^* & \gamma \epsilon R^* - \sigma(1-R^*) \\
+ \frac{\lambda  \rho  (\lambda -\sigma )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
+ \frac{\lambda  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\mu  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
+ \frac{m \mu  (\lambda -\sigma )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\lambda -\sigma )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \mu  (\lambda -\sigma )}{\lambda  \rho +\mu  \sigma } \\
 \end{array}
@@
 
-Whether the internal steady state is attracting or repelling is conditioned on the values of the parameters in the system.
-If the parameters of the Jacobian matrix (at the internal steady state) are such that its eigenvalues have negative real values, then it is stable.
-The stability of the system changes if a Saddle-Node bifurcation is crossed, such that a single eigenvalue obtains a positive real part, defined by the condition ${\rm Det}({\bf J})=0$.
-A single Saddle-Node bifurcation always exists for $\alpha = 0$, such that $\alpha > 0$ for the internal state to be stable.
-%Although a closed-form solution is not possible, a second stability condition generally requires $\delta > a \exp \{\frac{b}{c+\epsilon}\}$ assuming low values of $\mu_x$ and $\mu_y$ (where $a$, $b$, and $c$ are some combination of the model parameters), which simplifies to $\delta > \frac{\mu_y}{\epsilon}$ when $p=0$.
-%This condition places a lower bound on the consumer's rate of growth as a function of its rate of predation on the resource.
-%As the rate of consumption decreases, system is more likely to become unstable if the consumer's rate of reproduction declines.
+%Transcritical bifurcation
+If the parameters of the Jacobian matrix at the internal steady state are such that its leading eigenvalue is $<0$, then the system is stable to small pulse perturbations, conditioned on the value of the starvation rate $\sigma$ relative to the value of the consumer reproduction rate $\lambda$.
+As $\sigma$ nears and becomes lower than a given $\lambda$, the resource steady state $R^*$ crosses the origin and exchanges stability to become unstable.
+As such, a transcritical bifurcation exists at $\lambda = \sigma$, such that the existence of an internal stable fixed point is dependent on the condition that $\sigma > \lambda$.
+Biologically, this means that the rate of starvation is greater (operating on a smaller timescale) than the rate of consumer reproduction (operating on a relatively longer timescale).
+As will be shown in Section XX, this general expectation will hold for most classes of organisms, while the exact difference in timescales between reproduction and starvation can be derived using allometric scaling relationships.
+%Kempes (REF) has shown that population growth rates scale to $M^{\alpha-1}$ (where $M$ is the reproductive mass of an organism and $\alpha$ is the scaling exponent), whereas starvation is necessarily bound to shorter timescales, which will be elaborated on below.
 
-Of more interest here is the existence of parameter regions that permit the existence of cyclic behavior.
+%Hopf bifurcation
+Oscillating, or cyclic, dynamics presents additional risks to a population.
+If the cycles are large, or have regions where low values are attained,
+
+Of additional interest is the existence of parameter regions that permit the existence of oscillatory, or cyclic, dynamics.
 Cycles arise when a pair of complex conjugate eigenvalues cross the imaginary axis and attain positive real parts.
 This condition is called a Hopf bifurcation, and is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian matrix.
-Although the Hopf condition cannot be solved analytically for the specific 2-stage model, it can be explored numerically.
+Although the Hopf condition for the specific 2-stage model cannot be easily solved, it can be explored using a symbolic computational language such as \emph{Mathematica}.
 
-\subsection{Generalized model}
+%What parameters is the Hopf bifurcation primarily dependent on?
+
+%The natural ordering of the scaled rates
+%lambda < rho < sigma
+%if lambda is 0.2, and sigma is 0.8, then rho is between 0.2 and 0.8.
+%In other words, lambda has the SLOWEST timescale, sigma has the fastest timescale, and rho has a timescale intermediate to both.
+
+\subsection{Analysis of a generalized model}
 
@@
 
-\subsection{Stochastic dynamic program}
+\subsection{Allometric scaling relationships}
 
@@
 %The reproductive tradeoff as a strategy
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth]{fig_Hopf.pdf}
-	\caption{
-	Saddle Node bifurcation
-	}
-	\label{SN}
-\end{figure}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
-	\caption{
-	Probability that the resource value is less than threshold = 0.05
-	}
-	\label{SN}
-\end{figure}
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
-	\caption{
-	Hopf bifurcation
-	}
-	\label{SN}
-\end{figure}
+%
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=0.5\textwidth]{fig_Hopf.pdf}
+% 	\caption{
+% 	Saddle Node bifurcation
+% 	}
+% 	\label{SN}
+% \end{figure}
+%
+%
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
+% 	\caption{
+% 	Probability that the resource value is less than threshold = 0.05
+% 	}
+% 	\label{SN}
+% \end{figure}
+%
+%
+%
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
+% 	\caption{
+% 	Hopf bifurcation
+% 	}
+% 	\label{SN}
+% \end{figure}
 

### 2016-03-23 foraging_simulations.jl

@@ -20,18 +20,18 @@ Hsimstar = zeros(length(sigmavec));
 Rsimstar = zeros(length(sigmavec));
 for i = 1:length(sigmavec)
   println("sigma = ",sigmavec[i])
-  L = 50;
+  L = 100;
   dim = 2;
   prop_fill = 0.5
   initsize = convert(Int64,round(((L-2)^dim)*prop_fill));
-  t_term = 80;
+  t_term = 100;
   alpha = 0.5;
   K = 1;
   sigma = sigmavec[i];
-  rho = 0.5;
-  m = 0.2;
+  rho = 0.8;
+  m = 0.1;
   lambda = 0.2;
-  mu = 0.2;
+  mu = 0.3;
 
   Fstar[i] = (alpha*lambda*mu*(mu + rho))/((lambda*rho + m*mu)*(lambda*rho + sigma*mu));
   Hstar[i] = (alpha*lambda^2*(mu + rho))/((lambda*rho + m*mu)*(lambda*rho + sigma*mu));

### 2016-03-23 src/starvingforager_event_rate.jl

@@ -248,7 +248,7 @@ function starvingforager_event_rate(L,dim,initsize,t_term,alpha,K,sigma,rho,m,la
       break
     end
     #Break loop if resources go extinct and the other populations run away
-    if NR == 0 && (NF + NH) > S
+    if NR == 0 && (NF + NH) > S*2
       println("Runaway growth has occured at t=",round(t,2)," and loop ",tic)
       break
     end

### 2016-04-13 Manuscript/Yeakel_Redner_2015.tex
manuscript updates

@@
-\documentclass[11pt]{article}
-\usepackage[margin=1in,footskip=0.25in]{geometry}
+\documentclass[12pt,english]{article}
+%\usepackage[margin=1in,footskip=0.25in]{geometry}
 
@@
 \usepackage[T1]{fontenc}
-\usepackage[super]{natbib}
+\usepackage[comma]{natbib}
 \usepackage{framed}
@@
 \usepackage{longtable}
+\usepackage{geometry}
+\usepackage{lineno}
+
+\geometry{verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm}
 
 \begin{document}
+\begin{spacing}{1.9}
 
 \title{The effect of starvation on the dynamics of consumer populations}
+\author{Yeakel Kempes Redner}
 \maketitle
+
+\linenumbers
+%\begin{flushleft}
 
@@
 
-The behavioral ecology of most, if not all, organisms is influenced by the energetic state of individuals from which the population is composed.
-An individual's energetic state, which can range from starving (energetically defficient) to satiated (energetially replete), directly influences how it invests its stores in an uncertain environment.
-Such state-dependent behaviors are generally manifested as trade-offs.
-A major trade-off that we consider in detail concerns reproductive vs. non-reproductive energetic investment strategies.
-As has been shown in [examples], individuals that are energetically deficient are unable to invest energy stores towards reproduction.
-Among sexually reproducing species, even when reproduction occurs, the offspring is often naturally aborted if there is not sufficient caloric resources available to the parent.
+The behavioral ecology of most, if not all, organisms is influenced by the energetic state of individuals.
+An individual's energetic state directly influences how it invests its stores in an uncertain environment.
+Such behaviors are generally manifested as trade-offs, which often concern investing in individual maintenance and growth vs.
+producing offspring [Mangel], among a host of other behavioral duties [REFS].
+The timing of these behaviors is often important and is under strong selective pressure, as they tend to have large effects on the future fitness of the organism [Mangel].
+%For example, rotifers that reproduce have lower survival rates than those that don't when they are under nutritional stress.
+To what extent, and when, organisms invest in these two necessary biological functions -- growth and maintenance vs.
+reproduction -- may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, or even resource limitation.
+The influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to delays or shifts in reproduction. %and the subsequent loss of nutritional reserves
+%Such tradeoffs are also bound to have population-level consequences, although this has received little theoretical or empirical attention.
 
-Here we explore how reproductive trade-offs, which occur at the level of the individual, may influence the dynamics of a population of individuals.
-We first establish a simple stage-structured population model that captures the essential elements of energetic reproductive tradeoffs, and explore how the rate of starvation impacts dynamics at the level of the population.
-We then develop a more general model in order to understand if there are common attriburtes of these dynamics that have implications beyond the simple stage-structured model.
-Finally, we compare our population-dynamic approach with a stochastic dynamic program that incorporates different individual-level behaviors as a function of energetic state.
+%The reproductive vs. storage tradeoff
+Maximizing fitness between growth and maintenance activities vs.
+reproductive behaviors in large part structures the life-history of species, and this can be achieved by a variety of potential mechanisms all of which, to some extent, depend on resource availability:
+%Reproduction can be separated from somatic growth/maintenance by a variety of potential mechanisms:
+%Multiple mechanisms contribute to the differentiation of these behaviors:
+\emph{i}) Behavioral: The investment of time and energy towards reproductive and parental behaviors depends on resource availability \citep{Morris:1987eo}.
+%, while stress can generally lead to reprodusuppression of reproductive behaviors \citep{Morgan:1999do}. 
+For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \citep{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for an entire season \citep{Barboza:2002in}.
+Resource limitation can also alter the behaviors of species in well-mixed environments: freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress \citep{Threlkeld:1976ih}, while those that do reproduce evince lower survival rates \citep{Kirk:1997cc}.
+Similarly, artificially induced stress has been observed to decrease reproductive success in Atlantic cod \citep{Morgan:1999do}.
+\emph{ii}) Physiological: Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation), often timing with accumulation of nutritional reserves \citep{Mead:1989dt,Sandell:1990kw}.
+%, including multiple species of Grizzly bears (\emph{Ursus horribilis}) and the European badger (\emph{Meles melesc}) \citep{Woodroffe:1995id
+Furthermore, many mammals (including humans) suffer irregular menstrual cycling and higher rates of spontaneous abortion during periods of nutritional stress \citep{Bulik:1999eo,Trites:2003cc}.
+\emph{iii}) Spatio-temporal: Organisms may also separate maintenance/growth from reproduction over space and time. 
+For example, many salmonids, birds, and some mammals return to migratory breeding sites to reproduce after one or multiple seasons in alternative environments spent accumulating body mass and nutritional reserves \citep{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+%great tits invest more energy into broods when resource abundance is greater, while food availability is generally seen to correlate with clutch size among many bird species \citep{Daan:1988va}.
+%The mechanisms that control reproductive effort as a function of energetic reserves differ among species.
+%Even humans alter their growth trajectories and have higher chances of fetal abortion during periods of starvation.
+%Thus, some organisms partition reproductive vs.
+foraging effort over both space and time, whereas other rely on physiological mechanisms to optimize reproductive success against survival when adequate resources are unavailable.
+The existence of so many independently evolved mechanisms across such a diverse suite of organisms points to the importance and universality of the fundamental tradeoff between spending energy on ones own tissues and spending energy on passing down genetic material.
+
+%To population-level effects
+Organisms employ different strategies to avoid reproduction during times of nutritional stress, and how this is achieved has received tremendous empirical and theoretical attention, owing to the importance of these activities in shaping life-history [REF].
+Less well understood is how resource limitation affect dynamics at the level of the population when organisms are able to avoid the risk of spending energy on reproduction when starvation is likely.
+Traditional Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus implicitly incorporating the requirement of resource availability for reproduction.
+However, resource limitation and the subsequent effects of starvation can also be accounted for \emph{explicitly}, such that reproductive growth of a population is only allowed to occur if individuals have sufficient energetic reserves.
+Incorporating energetic dynamics that occur on an individual level [REF] into a population-based framework \citep{Kooi2000,Sousa:2010ez}, though straightforward conceptually, has many challenges that arise mathematically \citep{Diekmann:2010da}, and this has limited simple theoretical models aiding our understanding of larger-scale dynamics.
+
+%What we did...
+Here we explore how reproductive trade-offs, which occur at the level of the individual, may influence the dynamics of a population.
+We first establish a simple stage-structured population model that captures the essential dynamics of energetic reproductive tradeoffs, and explore the impact of different fluxes on system-level stability.
+By relating different rate constants to allometric relationships, we uncover important constraints in the timescales of different physiological processes that determine dynamics and investigate how organisms of alternative body sizes and taxonomic affinities are expected to evince contrasting population-level dynamics.
+We then develop a more general dynamic model in order to understand if there are common attributes of these energetically-constrained systems that have implications beyond assumptions inherent to the specific model.
+We show [WHAT]?
+%Finally, we compare our population-dynamic approach with a stochastic dynamic program that incorporates different individual-level behaviors as a function of energetic state.
+
+%We devise the simplest energetic-compartment system to investigate the effects of a reproductive tradeoff with starvation on population dynamics.
+
+%FITNESS and ENERGETICS... the model incorporates fitness consequences of energetic loss...
 
@@
 We first integrate energetics into the dynamics of a consumer-resource system by assuming that the consumer population can be divided into discrete energetic states, the occupation of each being contingent on the consumption of a single resource $R$.
-In the simplest case, there are only two energetic states for the consumer population: \emph{i}) an energetically replete state $F$, where the consumer reproduces at rate $\lambda$, and \emph{ii}) an energetically deficient state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
+In the simplest case, there are only two energetic states for the consumer population: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
 Consumers transition from state $F$ to state $H$ by starvation at rate $\sigma$ and in proportion to the lack of food $1-R$.
-Conversely, consumers recover from state $H$ to the energetically replete state $F$ at rate $\rho$ and in proportion to the density of resources consumed $R$.
+Conversely, consumers recover from state $H$ to the full state $F$ at rate $\rho$ and in proportion to the density of resources consumed $R$.
 The resource has logistic growth with a linear growth rate $a$ and a carrying capacity of unity.
@@
 \frac{\rm d}{\rm dt} R &= a R(1-R) - R(\rho H+ b F).
+\label{eq_specific}
 \end{align}
@@
 \begin{array}{ccc}
- \frac{\lambda  \rho  (\lambda -\sigma )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
+ -\frac{\lambda  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
  \frac{\lambda  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\mu  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
- \frac{m \mu  (\lambda -\sigma )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\lambda -\sigma )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \mu  (\lambda -\sigma )}{\lambda  \rho +\mu  \sigma } \\
+ -\frac{m \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } & -\frac{\mu  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } \\
 \end{array}
@@
 %Hopf bifurcation
-Oscillating, or cyclic, dynamics presents additional risks to a population.
-If the cycles are large, or have regions where low values are attained,
+Oscillating, or cyclic, dynamics present additional risks to populations.
+If cycles are large, stochastic effects may result in extinction.
+In continuous-time systems, cycles arise when a pair of complex conjugate eigenvalues cross the imaginary axis and attain positive real parts.
+%Of additional interest is the existence of parameter regions that permit the existence of oscillatory, or cyclic, dynamics.
+This condition is called a Hopf bifurcation, and is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian matrix.
+Although the Hopf condition for the specific 2-stage model cannot be easily written, the analytical solution can be explored using a symbolic computational language such as \emph{Mathematica}.
 
-Of additional interest is the existence of parameter regions that permit the existence of oscillatory, or cyclic, dynamics.
-Cycles arise when a pair of complex conjugate eigenvalues cross the imaginary axis and attain positive real parts.
-This condition is called a Hopf bifurcation, and is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian matrix.
-Although the Hopf condition for the specific 2-stage model cannot be easily solved, it can be explored using a symbolic computational language such as \emph{Mathematica}.
+\subsection{Analysis of a generalized model}
 
-%What parameters is the Hopf bifurcation primarily dependent on?
+We may gain additional insight by that we do not know the specific rate functions from our 2-stage consumer resource model presented in Eq. \ref{eq_specific}.
+For example, we assume a linear mortality term for hungry foragers $\mu H$, though we may wish to assert that our knowledge of consumer mortality involves those that are energetically deficient, but nothing else.
+In this case, we would assume only that the rate of mortality is governed by the function $M(H)$.
+Substituting general functions for all rate laws from Eq. \ref{eq_specific}, we obtain the generalized ODE system
+
+\begin{align}
+\frac{\rm d}{\rm dt} F &= G(F) + S(R,H) - K(R,F), \nonumber \\
+\frac{\rm d}{\rm dt} H &= K(R,F) - S(R,H) - M(H), \nonumber \\
+\frac{\rm d}{\rm dt} R &= P(R) - L(R,H,F).
+\label{eq_gen}
+\end{align}
+
+\noindent where $G(F)$ determines consumer growth, $S(R,H)$ and $K(R,F)$ are the recovery and starvation functions, respectively, $M(H)$ determines consumer mortality, and $P(R)$ and $L(R,H,F)$ are functions describing the growth and consumption-loss of resources, respectively.
+
+If the system is written in this generalized manner, we cannot solve for the steady state solution $(F^*,~H^*,~R^*)$, however we can normalize the system to the unknown steady states.
+We denote normalized variables and functions in lowercase, such that $f = F/F^*$, $h = H/H^*$, $r = R/R^*$, and for example the normalized mortality function $m(h) = M(H)/M^*$, where $M^*$ is shorthand for $M(H^*)$.
+Additional rearrangements of terms under equilibrium conditions allows us to define two additional sets of scaling parameters with intuitive biological properties: the turnover rates of full foragers, hungry foragers, and the resource ($\alpha_f$, $\alpha_h$, $\alpha_r)$, and the proportional branching biomass through different compartments of the model, generally designated by the parameter $\beta$.
+For instance, $\beta_f$ is the proportion of full consumer growth due to reproduction, whereas $(1-\beta_f)$ is the proportion of full consumer growth due to recruitment \emph{from} the hungry forager class.
+Similarly, $\beta_h$ is the proportion of hungry consumer loss due to mortality, whereas $(1-\beta_h)$ is the proportion of hungry consumer loss due to recruitment \emph{into} the full consumer class (see SUPP for a detailed derivation).
+Substituting the normalized variables and functions into Eq. \ref{eq_gen}, we obtain
+
+\begin{align}
+\dot{f} &= \alpha_f \left[ \beta_f g(f) + (1-\beta_f)s(r,h) - k(r,f)\right], \nonumber \\
+\dot{h} &= \alpha_h \left[ k(r,f) - (1-\beta_h)s(r,h) - \beta_h m(h) \right], \nonumber \\
+\dot{r} &= \alpha_r \left[ p(r) - l(r,h,f) \right],
+\label{eq_gen}
+\end{align}
+
+\noindent and linearization of this normalized, general ODE system yields the Jacobian matrix
+
+\begin{equation}
+\mathbf{J}_{\rm gen}|_* =
+\left(
+\begin{array}{ccc}
+ \alpha_f \left(\beta_f \frac{\partial g}{\partial f} - \frac{\partial k}{\partial f} \right) & \alpha_f(1-\beta_f)\frac{\partial s}{\partial h} & \alpha_f \left( (1-\beta_f)\frac{\partial s}{\partial r} - \frac{\partial k}{\partial r} \right) \\
+ \alpha_h \frac{\partial k}{\partial f} & -\alpha_h \left(\beta_h \frac{\partial m}{\partial h} + (1-\beta_h) \frac{\partial s}{\partial h} \right) & \alpha_h \left(\frac{\partial k}{\partial r} - (1 - \beta_h)\frac{\partial s}{\partial r} \right) \\
+ -\alpha_r \frac{\partial l}{\partial f} & -\alpha_r \frac{\partial l}{\partial h} & \alpha_r \left( \frac{\partial p}{\partial r} - \frac{\partial l}{\partial r} \right) \\
+\end{array}
+\right).
+\label{eq_Jacgen}
+\end{equation}
+
+%Importance of the elasticities
+Due to the normalization procedure, the partial derivatives in Eq. \ref{eq_Jacgen} have tangible biological meaning.
+Because, for example, the partial derivative (containing functions and variables normalized to the unknown steady states) $\partial g / \partial f = \partial \log G / \partial \log F$, it scales in such a way that it represents the percent change in consumer growth (governed by $G(F)$) relative to a percent change in the density of full consumers $F$, more commonly known as a functional elasticity.
+For example, if growth is a linear function (e.g.
+$G(F) = \lambda F$), $\partial g / \partial f = 1$; if growth is a quadratic function (e.g.
+$G(F) = \lambda F^2$), $\partial g / \partial f = 2$, while more complex functions may depend on the value of the steady state.
+For example, if consumer growth is modeled as Holling Type II growth, such that $G(F) = c_1 F^2/(c_2 + F^2)$, where $c_1$ and $c_2$ are unknown constants, then its elasticity will vary between $0$ and $2$, depending on the steady state value $F^*$, which is unknown in the generalized system.
+
+%Why do this?
+%The normalization of the generalized system described above results in the inclusion of functional elasticities into the generalized Jacobian matrix. 
+Deriving a Jacobian in terms of the normalized ODE system is useful because it allows us to place strict constraints on the values of the unknown variables (the turnover rates, biomass branching parameters, and the functional elasticities), without assuming detailed knowledge of the functions controlling different rates within the system.
+% retaining the fundamental architecture of the 2-class consumer-resource system.
+%Simplifications
+In addition, we can now insert a number of assumptions that will align our generalized Jacobian more closely with the original 2-stage consumer resource system.
+We will assume the following:
+\emph{i}) both the consumer and resource suffers linear mortality,
+\emph{ii}) resource growth is logistic,
+\emph{iii}) recovery and starvation are linear with respect to both full and hungry consumer densities, and
+\emph{iv}) consumers and resources have equivalent turnover rates scaled to unity.
+These assumptions lead to the following simplifications
+
+\begin{align}
+	i)~&\frac{\partial m}{\partial h} = 1, ~~~~~~~~ \frac{\partial l}{\partial r} = 1, ~~~~~~~~ \frac{\partial l}{\partial f} = 1 - \frac{\partial l}{\partial h}, \nonumber \\
+	ii)~&\frac{\partial k}{\partial r} = \left(1 - \frac{1}{R^*} \right)^{-1}, \nonumber \\
+	iii)~&\frac{\partial s}{\partial r} = \frac{\partial s}{\partial h} = 1, ~ \frac{\partial k}{\partial f} = 1, \nonumber \\
+	iv)~&\alpha_f = \alpha_r = 1,
+\end{align}
+
+\noindent where $R^*$ ranges from 0 to 1.
+The remaining free parameters include the timescale of hungry foragers $\alpha_h$, the branching parameters $\beta_f$ and $\beta_h$, the elasticity of consumer growth with respect to full consumer densities $\partial g / \partial f$, the elasticity of resource growth with respect to resource density $\partial p / \partial r$, the elasticity of resource loss with respect to full consumer density $\partial l / \partial f$, and the elasticity of starvation with respect to resource density $\partial k / \partial r$.
+See Table XX for a list of the free parameters in the generalized model, as well as the ranges of potential values for each.
+These substitutions result in the simplified Jacobian matrix
+
+\begin{equation}
+\mathbf{J}_{\rm gen}|_* =
+\left(
+\begin{array}{ccc}
+ \left(\beta_f \frac{\partial g}{\partial f} - 1 \right) & (1-\beta_f) & \left( (1-\beta_f)\frac{\partial s}{\partial r} - 1 \right) \\
+ \alpha_h & -\alpha_h & -\alpha_h \left(1-\frac{\partial k}{\partial r} - \beta_h  \right) \\
+ \frac{\partial l}{\partial h}-1 & -\frac{\partial l}{\partial h} &  \left( \frac{\partial p}{\partial r} - 1 \right) \\
+\end{array}
+\right).
+\label{eq_Jacgen}
+\end{equation}
+
+%Testing stability
+Because the remaining free parameters have known ranges, but not specific values, we wish to assess the correlations of each to system stability.
+By randomly drawing values from uniform distributions bounded by the known ranges of each free parameter, we obtain an ensemble of potential Jacobian matrices whereupon the stability of each is determined by numerically calculating the real part of the leading eigenvalue.
+If the real part of the leading eigenvalue is $<1\times10^{-6}$, we assumed the system to be stable.
+Replicating this procedure on the order of $10^7$ times allowed us to determine the correlation of each free parameter with stability to the extent that variance was negligible.
+
+\subsection{Allometric scaling relationships}
+
+Nearly all of the rates described in the specific model, and generalized upon in the generalized model, are to some extent governed by the body size of the consumer.
+The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ plays a central role in other scaling relationships.
+Organismal metabolic rate $B$ is known to scale as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to be 3/4 for metazoans, etc.
+Kempes et al. [REF] show how the population-level growth rate also can be related to body size as $\lambda = \lambda_0 M^{\eta-1}$.
+
+Population growth requires that individuals
+
+%\begin{align}
+%\lamdba &=  \nonumber \\
+%\sigma &= \nonumber \\	
+%\end{align}
+
+\section{Results}
+
+\subsection{2-stage consumer resource model}
+
+Analysis of the continuous-time 2-stage consumer resource model shows that the equilibrial states of both populations are highly sensitive to changes in  starvation and recovery rates of the consumer.
+The consumer and resource population densities vary inveresely: when the consumer densities are high, resource densities are low, and vice versa.
+High starvation and low recovery rates result in low consumer densities and high resource densities. %whereas the opposite is true for low starvation and high recovery rates.
+If starvation rates are low, resources have a fixed point near zero for any value of the recovery rate.
+Full and hungry consumer stages tend have fixed points that are tightly correlated, the extent to which is driven by the similarity of consumer growth and mortality rates; if $\lambda = \mu$, then $F^* = H^*$.
+
+%Transcritical & rate law ranges
+A transcritical bifurcation exists at $\lambda = \sigma$, such that the condition $\sigma > \lambda$ is required for biologically reasonable dynamics.
+The TC bifurcation occurs in this model because we have assumed that all energetically replete individuals reproduce at the same rate, whereas the traditional Lotka-Volterra dynamic assumes that the reproductive rate of the consumer is scaled to resource density, such that the growth function would be $G(F,R) = \lambda R F$.
+Thus, the Lotka-Volterra dynamic \emph{implicitly} accounts for starvation in reducing the reproductive rate of the consumer.
+However, our 2-stage model \emph{explicitly} accounts for starvation as well as recovery, such that individuals who are not starved should adopt a reproductive rate independent of resource density.
+
+The existence of the transcritical bifurcation reveals important biological insight.
+Reproduction requires maintenance and growth of biological tissues, both of which have strong scaling relationships with body size.
+Recent work by Kempes et al. [REF] derived the timescale of reproduction in terms of allometric considerations, where $t_\lambda \propto M^{1-\eta}$ (REF).
+Starvation is the loss of energy required for maintenance, and we have shown it to have a timescale $t_\sigma \propto \log(M)$.
+Accordingly, the timescale of reproduction is always larger than the timescale of starvation, such that $\lambda < \sigma$.
+A third important parameter in our framework is the rate of recovery.
+The recovery timescale $t_\rho$ controls the rate at which individuals move from the hungry class to the full class, and this requires not only tissue maintenance, but growth, such that it is bounded on the short side by $t_\sigma$.
+Moreover, [why is recovery timescale bounded on the high side], such that it is bounded on the long side by $t_\lambda$.
+
+We have used scaling relationships between tissue turnover and growth to strictly constrain 5/6 population-level parameters in our 2-stage consumer resource model (including the mortality rate $t_\mu$, which we have shown is just a xxx of $t_\sigma$).
+This exercise accomplishes two goals:
+1) it allows us to constrain the plausible parameter space of the two-stage model, and 
+2)
+
+This allows us to derive many aspects of the system in terms of consumer body mass $M$ and the allometric scaling exponent $\eta$.
+
+\subsection{Generalized analysis}
+Constraining the model with allometric scaling relationships yields important insights regarding the tradeoff between [fitness and energy]
 
@@
 %In other words, lambda has the SLOWEST timescale, sigma has the fastest timescale, and rho has a timescale intermediate to both.
-
-\subsection{Analysis of a generalized model}
-
-$\xi$ ~~ $\zeta$
-
-\subsection{Allometric scaling relationships}
-
-\section{Results}
-
-To determine how the consumer-resource system impacted by different rates of starvation, we analyze the systems with respect to $p$.
 
@@
 
+\bibliographystyle{ieeetr} % for Science,
+\bibliography{aa_starving}
+
+%\end{flushleft}
+\end{spacing}
 \end{document}

### 2016-04-14 Manuscript/Yeakel_Redner_2015.tex
changes to paramteres in mathematica as well as some figure changes as well. Updates to the Results section in ms

@@
 Maximizing fitness between growth and maintenance activities vs.
-reproductive behaviors in large part structures the life-history of species, and this can be achieved by a variety of potential mechanisms all of which, to some extent, depend on resource availability:
+reproductive behaviors in large part structures the life-history of species, and this can be achieved by a variety of potential mechanisms all of which, to some extent, depend on resource availability.
 %Reproduction can be separated from somatic growth/maintenance by a variety of potential mechanisms:
 %Multiple mechanisms contribute to the differentiation of these behaviors:
-\emph{i}) Behavioral: The investment of time and energy towards reproductive and parental behaviors depends on resource availability \citep{Morris:1987eo}.
+Behavioral: The investment of time and energy towards reproductive and parental behaviors depends on resource availability \citep{Morris:1987eo}.
 %, while stress can generally lead to reprodusuppression of reproductive behaviors \citep{Morgan:1999do}. 
-For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \citep{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for an entire season \citep{Barboza:2002in}.
+For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \citep{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for a breeding season \citep{Barboza:2002in}.
 Resource limitation can also alter the behaviors of species in well-mixed environments: freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress \citep{Threlkeld:1976ih}, while those that do reproduce evince lower survival rates \citep{Kirk:1997cc}.
 Similarly, artificially induced stress has been observed to decrease reproductive success in Atlantic cod \citep{Morgan:1999do}.
-\emph{ii}) Physiological: Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation), often timing with accumulation of nutritional reserves \citep{Mead:1989dt,Sandell:1990kw}.
+Physiological: Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation), often timing with accumulation of nutritional reserves \citep{Mead:1989dt,Sandell:1990kw}.
 %, including multiple species of Grizzly bears (\emph{Ursus horribilis}) and the European badger (\emph{Meles melesc}) \citep{Woodroffe:1995id
 Furthermore, many mammals (including humans) suffer irregular menstrual cycling and higher rates of spontaneous abortion during periods of nutritional stress \citep{Bulik:1999eo,Trites:2003cc}.
-\emph{iii}) Spatio-temporal: Organisms may also separate maintenance/growth from reproduction over space and time. 
+Spatio-temporal: Organisms may also separate maintenance/growth from reproduction over space and time. 
 For example, many salmonids, birds, and some mammals return to migratory breeding sites to reproduce after one or multiple seasons in alternative environments spent accumulating body mass and nutritional reserves \citep{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
@@
 foraging effort over both space and time, whereas other rely on physiological mechanisms to optimize reproductive success against survival when adequate resources are unavailable.
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms points to the importance and universality of the fundamental tradeoff between spending energy on ones own tissues and spending energy on passing down genetic material.
+The existence of so many independently evolved mechanisms across such a diverse suite of organisms points to the importance and universality of the fundamental tradeoff between spending energy on tissues vs.
+spending energy on passing down genetic material.
 
@@
 Organisms employ different strategies to avoid reproduction during times of nutritional stress, and how this is achieved has received tremendous empirical and theoretical attention, owing to the importance of these activities in shaping life-history [REF].
-Less well understood is how resource limitation affect dynamics at the level of the population when organisms are able to avoid the risk of spending energy on reproduction when starvation is likely.
+Less well understood is how resource limitation and these behavioral/physiological tradeoffs affect dynamics at the level of the population.
+% when organisms are able to avoid the risk of spending energy on reproduction when starvation is likely.
 Traditional Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus implicitly incorporating the requirement of resource availability for reproduction.
-However, resource limitation and the subsequent effects of starvation can also be accounted for \emph{explicitly}, such that reproductive growth of a population is only allowed to occur if individuals have sufficient energetic reserves.
-Incorporating energetic dynamics that occur on an individual level [REF] into a population-based framework \citep{Kooi2000,Sousa:2010ez}, though straightforward conceptually, has many challenges that arise mathematically \citep{Diekmann:2010da}, and this has limited simple theoretical models aiding our understanding of larger-scale dynamics.
+Although this implicit dependence connects resource limitation to lower consumer growth rates, the following biological realities are not taken into account:
+\emph{i}) some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation,
+\emph{ii}) the portion of the population that is not nutritionally stressed reproduces at a near-constant rate, constrained by allometric relationships, and
+\emph{iii}) the rate at which individuals transition from nutritionally poor to replete states (and vice versa) are governed by the organism's metabolism, which is also strongly constrained by allometry.
+To incorporate these important elements, resource limitation and the subsequent effects of starvation can be accounted for \emph{explicitly}, such that reproductive growth of a population is only allowed to occur for individuals with sufficient energetic reserves.
+Incorporating energetic dynamics that occur at an individual level [REF] into a population-based framework \citep{Kooi2000,Sousa:2010ez}, though straightforward conceptually, has many challenges that arise mathematically \citep{Diekmann:2010da}, and this has limited simple theoretical models aiding our understanding of larger-scale dynamics.
 
@@
 In the simplest case, there are only two energetic states for the consumer population: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
-Consumers transition from state $F$ to state $H$ by starvation at rate $\sigma$ and in proportion to the lack of food $1-R$.
+Consumers transition from state $F$ to state $H$ by starvation at rate $\sigma$ and in proportion to the lack of food $(1-R)$.
 Conversely, consumers recover from state $H$ to the full state $F$ at rate $\rho$ and in proportion to the density of resources consumed $R$.
 The resource has logistic growth with a linear growth rate $a$ and a carrying capacity of unity.
-Resources are eliminated by the consumer in either state: by energetically deficient consumers at rate $\rho$, and by energetically replete consumers at rate $b$.
+Resources are eliminated by the consumer in both states: by energetically deficient consumers at rate $\rho$, and by energetically replete consumers at rate $b$.
 Accordingly, the system of equations is written
@@
 
-\subsection{2-stage consumer resource model}
-
-Analysis of the continuous-time 2-stage consumer resource model shows that the equilibrial states of both populations are highly sensitive to changes in  starvation and recovery rates of the consumer.
-The consumer and resource population densities vary inveresely: when the consumer densities are high, resource densities are low, and vice versa.
+Analysis of the 2-stage consumer resource model shows that the equilibrial states of both populations are highly sensitive to changes in  starvation and recovery rates of the consumer.
+The consumer and resource population densities vary inversely: when the consumer densities are high, resource densities are low, and vice versa.
 High starvation and low recovery rates result in low consumer densities and high resource densities. %whereas the opposite is true for low starvation and high recovery rates.
@@
 A transcritical bifurcation exists at $\lambda = \sigma$, such that the condition $\sigma > \lambda$ is required for biologically reasonable dynamics.
-The TC bifurcation occurs in this model because we have assumed that all energetically replete individuals reproduce at the same rate, whereas the traditional Lotka-Volterra dynamic assumes that the reproductive rate of the consumer is scaled to resource density, such that the growth function would be $G(F,R) = \lambda R F$.
-Thus, the Lotka-Volterra dynamic \emph{implicitly} accounts for starvation in reducing the reproductive rate of the consumer.
-However, our 2-stage model \emph{explicitly} accounts for starvation as well as recovery, such that individuals who are not starved should adopt a reproductive rate independent of resource density.
-
-The existence of the transcritical bifurcation reveals important biological insight.
+The TC bifurcation occurs in this model because we have assumed that the portion of the population that is not starved reproduces at a constant rate.
+Because the process of starvation is incorporated explicitly, the consumer's rate of reproduction is not dependent on the density of resources.
+In fact, the existence of the TC bifurcation at $\lambda = \sigma$ reveals an important biological insight. 
 Reproduction requires maintenance and growth of biological tissues, both of which have strong scaling relationships with body size.
@@
 Starvation is the loss of energy required for maintenance, and we have shown it to have a timescale $t_\sigma \propto \log(M)$.
-Accordingly, the timescale of reproduction is always larger than the timescale of starvation, such that $\lambda < \sigma$.
+Accordingly, the timescale of reproduction is always larger than the timescale of starvation, such that $\lambda$ must be less than $\sigma$ by definition.
 A third important parameter in our framework is the rate of recovery.
@@
 Moreover, [why is recovery timescale bounded on the high side], such that it is bounded on the long side by $t_\lambda$.
+Thus, incorporating allometric considerations shows us that $\lambda < \rho < \sigma$ (alternatively $t_\lambda > t_\rho > t_\sigma$).
+
+%Hopf bifurcation and oscillations sigma vs. rho
+The 2-stage consumer resource model exhibits two qualitatively different behavioral regimes.
+Because portions of the consumer population exist in either full or hungry states simultaneously, the internal fixed point can either be a stable equilibrium, or exhibit sustained oscillatory behavior, depending on the rates of starvation $\sigma$ and recovery $\rho$.
+The transition from stable non-oscillatory dynamics to oscillatory dynamics occurs at the Hopf bifurcation condition where two complex conjugate eigenvalues cross the imaginary axis and attain positive real parts.
+Although there is an analytical solution to the Hopf bifurcation condition, it cannot be written efficiently
+When the starvation rate is low, oscillatory dynamics are more likely to occur for a given value of the recovery rate.
+This can be understood intuitively: for low starvation rates, resources are depressed by an infusion of full consumers, which subsequently starve thereby allowing the resource to recover and continuing the cycle.
+When the starvation rate is high, the response of consumer growth to resource abundance is muted, such that oscillations tend to decay over time.
+Thus, higher starvation rates $\sigma$ desensitizes changes in the consumer population to changes in resource density, and lower rates of recovery $\rho$ amplifies this effect.
+
+%Hopf bifurcation as a function of b
+Both full and hungry consumers remove resources at rates $b$ and $\rho$, respectively.
+As the rate of resource consumption by full consumers increases, the Hopf condition changes from a convex to a concave function over $\sigma$, limiting the potential for oscillatory dynamics.
+These rates are considered separately because full consumers need only to maintain their tissues, whereas hungry consumers require both growth and maintenance, such that $t_\rho > t_b$, or $\rho < b$.
+If this constraint is enacted, the likelihood of oscillatory dynamics is reduced for a given
+
+If instead the consumer's growth was proportional to resource abundance, such that the effects of starvation on reproduction were incorporated explicitly (where the 2-stage consumer resource model collapses to the Lotka-Volterra consumer-resource model with logistic growth of the resource), the TC bifurcation exists only for $\lambda = \mu$, such that the rate of mortality cannot exceed the intrinsic birth rate.
+
+ whereas the traditional Lotka-Volterra dynamic assumes that the reproductive rate of the consumer is scaled to resource density, such that the growth function would be $G(F,R) = \lambda R F$.
+Thus, the Lotka-Volterra dynamic \emph{implicitly} accounts for starvation in reducing the reproductive rate of the consumer.
+However, our 2-stage model \emph{explicitly} accounts for starvation as well as recovery, such that individuals who are not starved should adopt a reproductive rate independent of resource density.
 
@@
 
+%Starving Random Walk
+In our modeling, foragers look for food by wandering in this changing
+environment.  If such a search is successful, the forager is satiated and it
+can engage in the essential activity of reproduction.  However, if the search
+is unsuccessful for a sufficiently long period, the forager ``starves''.
+Such a forager can do nothing else but forage, until it either finds food and
+again becomes satiated or it dies when it goes too long without finding
+nourishment.  These rules are reminiscent of the ``starving random walk''
+model, where a single random walk can take $\mathcal{S}$ steps without
+encountering food before starving to death.  Moreover, the resource does not
+regenerate, so that the forager ultimately starves to death.  For this
+idealized model, it was found that the average lifetime of the forager scales
+algebraically with $\mathcal{S}$ in $d\!=\!1$ and $d\!=\!2$ dimensions, and
+as $\exp(-A\mathcal{S}^\omega)$ for $d>2$.  Here the exponent
+$\omega\approx \frac{1}{2}$ for $d=3$, while $\omega\to 1$ only as
+$d\to \infty$, with the latter behavior corresponding to the mean-field
+limit.  As we will discuss, regeneration of the resource, together with the
+behavioral change between starving and satiated foragers leads to still much
+richer dynamical behavior.
+
 %The reproductive tradeoff as a strategy
 %
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=0.5\textwidth]{fig_Hopf.pdf}
-% 	\caption{
-% 	Saddle Node bifurcation
-% 	}
-% 	\label{SN}
-% \end{figure}
+
+ \begin{figure}[h]
+ 	\centering
+ 	\includegraphics[width=1\textwidth]{fig_HopfFP.jpg}
+ 	\caption{
+ 	}
+ 	\label{Hopfb}
+ \end{figure}
+
+ \begin{figure}[h]
+ 	\centering
+ 	\includegraphics[width=0.5\textwidth]{fig_HopfPlotb.pdf}
+ 	\caption{
+ 	}
+ 	\label{Hopfb}
+ \end{figure}
 %

### 2016-06-15 Manuscript/Yeakel_Kempes_Redner-5-16-2016.tex
New analyses with allometric rates

--- 2016-04-14 Manuscript/Yeakel_Redner_2015.tex
@@
 \title{The effect of starvation on the dynamics of consumer populations}
-\author{Yeakel Kempes Redner}
+\author{Yeakel, Kempes, \& Redner}
 \maketitle
@@
 An individual's energetic state directly influences how it invests its stores in an uncertain environment.
-Such behaviors are generally manifested as trade-offs, which often concern investing in individual maintenance and growth vs.
-producing offspring [Mangel], among a host of other behavioral duties [REFS].
-The timing of these behaviors is often important and is under strong selective pressure, as they tend to have large effects on the future fitness of the organism [Mangel].
+Such behaviors are generally manifested as trade-offs, which often concern investing in individual maintenance and growth or allocating energy towards reproduction \citep{Kirk:1997cc,Kempes:2012hy}.
+%among a host of other behavioral duties [REFS].
+The timing of these behaviors is often important and is under strong selective pressure, as they tend to have large effects on the future fitness of the organism \citep{Mangel:1988uaa}.
 %For example, rotifers that reproduce have lower survival rates than those that don't when they are under nutritional stress.
 To what extent, and when, organisms invest in these two necessary biological functions -- growth and maintenance vs.
-reproduction -- may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, or even resource limitation.
-The influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to delays or shifts in reproduction. %and the subsequent loss of nutritional reserves
+reproduction -- may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, as well as resource limitation.
+Importantly, the influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
+%and the subsequent loss of nutritional reserves
 %Such tradeoffs are also bound to have population-level consequences, although this has received little theoretical or empirical attention.
@@
 Maximizing fitness between growth and maintenance activities vs.
-reproductive behaviors in large part structures the life-history of species, and this can be achieved by a variety of potential mechanisms all of which, to some extent, depend on resource availability.
+reproductive behaviors in large part structures the life-history of species, and this can be achieved by a variety of potential mechanisms, all of which, to some extent, depend on resource availability.
+[In extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the individual regulates all aspects of the cell cycle.]
 %Reproduction can be separated from somatic growth/maintenance by a variety of potential mechanisms:
@@
 For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \citep{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for a breeding season \citep{Barboza:2002in}.
-Resource limitation can also alter the behaviors of species in well-mixed environments: freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress \citep{Threlkeld:1976ih}, while those that do reproduce evince lower survival rates \citep{Kirk:1997cc}.
-Similarly, artificially induced stress has been observed to decrease reproductive success in Atlantic cod \citep{Morgan:1999do}.
-Physiological: Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation), often timing with accumulation of nutritional reserves \citep{Mead:1989dt,Sandell:1990kw}.
+%Resource limitation can also alter the behaviors of species in well-mixed environments: 
+Freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress \citep{Threlkeld:1976ih}, with those that do reproduce evincing lower survival rates \citep{Kirk:1997cc}, while artificially induced stress has been observed to decrease reproductive success in Atlantic cod \citep{Morgan:1999do}.
+Physiological: Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation) to time with accumulation of nutritional reserves \citep{Mead:1989dt,Sandell:1990kw}.
 %, including multiple species of Grizzly bears (\emph{Ursus horribilis}) and the European badger (\emph{Meles melesc}) \citep{Woodroffe:1995id
@@
 foraging effort over both space and time, whereas other rely on physiological mechanisms to optimize reproductive success against survival when adequate resources are unavailable.
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms points to the importance and universality of the fundamental tradeoff between spending energy on tissues vs.
-spending energy on passing down genetic material.
+The existence of so many independently evolved mechanisms across such a diverse suite of organisms points to the importance and universality of the fundamental tradeoff between energetic expenditures on the growth and maintenance of tissues vs.
+reproductive behaviors.
 
 %To population-level effects
-Organisms employ different strategies to avoid reproduction during times of nutritional stress, and how this is achieved has received tremendous empirical and theoretical attention, owing to the importance of these activities in shaping life-history [REF].
+The different strategies employed to avoid reproduction during times of nutritional stress, and how this is achieved among different organisms, has received tremendous empirical and theoretical attention owing to the importance of these activities in shaping life-history [REF].
 Less well understood is how resource limitation and these behavioral/physiological tradeoffs affect dynamics at the level of the population.
 % when organisms are able to avoid the risk of spending energy on reproduction when starvation is likely.
-Traditional Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus implicitly incorporating the requirement of resource availability for reproduction.
+Traditional Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction.
 Although this implicit dependence connects resource limitation to lower consumer growth rates, the following biological realities are not taken into account: 
-\emph{i}) some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation,
-\emph{ii}) the portion of the population that is not nutritionally stressed reproduces at a near-constant rate, constrained by allometric relationships, and
-\emph{iii}) the rate at which individuals transition from nutritionally poor to replete states (and vice versa) are governed by the organism's metabolism, which is also strongly constrained by allometry.
-To incorporate these important elements, resource limitation and the subsequent effects of starvation can be accounted for \emph{explicitly}, such that reproductive growth of a population is only allowed to occur for individuals with sufficient energetic reserves.
-Incorporating energetic dynamics that occur at an individual level [REF] into a population-based framework \citep{Kooi2000,Sousa:2010ez}, though straightforward conceptually, has many challenges that arise mathematically \citep{Diekmann:2010da}, and this has limited simple theoretical models aiding our understanding of larger-scale dynamics.
+\begin{itemize}
+\item some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation,
+\item the portion of the population that is not nutritionally stressed is expected to reproduce at a near-constant rate, and
+\item the rates at which individuals transition from nutritionally poor to replete states and back are governed by the organism's metabolic capabilities of growing and maintaining tissue, which is allometrically constrained \citep{Kempes:2012hy}.
+\end{itemize}
+[The neglect of these features has important dynamic shortcomings:]
+[Here we use body size and allometries to capture a wide variety of taxonomic diversity]
+[There isn't a deep understanding with how models vary across different life histories, traits, etc... how do equations change with allometric constraints for different class of organisms? 
+i.e. Use body size as a generating parameter to capture a wide variety of diversity.]
+
+To incorporate these elements, resource limitation and the subsequent effects of starvation may be accounted for \emph{explicitly}, such that reproductive growth of a population is permitted only for individuals with sufficient energetic reserves.
+Incorporating energetic dynamics that occur at an individual level \citep{Kooi2000} into a population-based framework \citep{Kooi2000,Sousa:2010ez}, though straightforward conceptually, quickly runs into mathematical obstacles \citep{Diekmann:2010da}.
+[An alternative approach is to start with the macroscale patterns in order to understand certain tradeoffs etc]
+The complexities associated with scaling individual-level energetics to that of a population has thus limited the development of theoretical models that may aid our understanding of the effects of such tradeoffs on population dynamics.
 
 %What we did...
-Here we explore how reproductive trade-offs, which occur at the level of the individual, may influence the dynamics of a population.
-We first establish a simple stage-structured population model that captures the essential dynamics of energetic reproductive tradeoffs, and explore the impact of different fluxes on system-level stability.
-By relating different rate constants to allometric relationships, we uncover important constraints in the timescales of different physiological processes that determine dynamics and investigate how organisms of alternative body sizes and taxonomic affinities are expected to evince contrasting population-level dynamics.
-We then develop a more general dynamic model in order to understand if there are common attributes of these energetically-constrained systems that have implications beyond assumptions inherent to the specific model.
+Here we explore how the energetic tradeoff between maintaining and building tissue and reproduction, which occurs at the level of the individual, may influence the dynamics of a population.
+We first establish a simple nutritional state-structured population model that captures the essential dynamics of energetic reproductive tradeoffs, and explore the impact of different fluxes on system-level stability.
+By relating different rate constants to allometric constraints, we uncover important relationships between the timescales of physiological and reproductive processes, and show how organisms of different body sizes and taxonomic affinities are expected may be prone to different dynamics.
+The general rate laws that control both reproductive and energetic processes of the population, we show that
+
+%We then develop a more general model to understand if there are common attributes of these energetically-constrained systems that have implications beyond assumptions inherent to the specific model.
 We show [WHAT]?
@@
 \begin{align}
-\frac{\rm d}{\rm dt} F &= \lambda F + \rho HR - \sigma (1-R)F, \nonumber \\
-\frac{\rm d}{\rm dt} H &= \sigma (1-R)F - \rho RH - \mu H, \nonumber \\
-\frac{\rm d}{\rm dt} R &= a R(1-R) - R(\rho H+ b F).
+\begin{split}
+\frac{\rm d}{\rm dt} F &= \lambda F + \rho RH - \sigma (1-R)F,  \\
+\frac{\rm d}{\rm dt} H &= \sigma (1-R)F - \rho RH - \mu H,  \\
+\frac{\rm d}{\rm dt} R &= \alpha R(1-R) - R(\rho H+ \beta F).
 \label{eq_specific}
+\end{split}
 \end{align}
+
+[NSM = Nutritional State Model]
+[LVM = Lotka Volterra Model]
 
@@
 \begin{align}
-F^* &= \frac{a  \lambda  \mu  (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  m)}, \nonumber \\
-H^* &= \frac{a  \lambda ^2 (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  m)}, \nonumber \\
+\begin{split}
+F^* &= \frac{a  \lambda  \mu  (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  m)}, \\
+H^* &= \frac{a  \lambda ^2 (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  m)}, \\
 R^* &= \frac{\mu  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma }.
+\end{split}
 \end{align}
@@
 As $\sigma$ nears and becomes lower than a given $\lambda$, the resource steady state $R^*$ crosses the origin and exchanges stability to become unstable.
-As such, a transcritical bifurcation exists at $\lambda = \sigma$, such that the existence of an internal stable fixed point is dependent on the condition that $\sigma > \lambda$.
-Biologically, this means that the rate of starvation is greater (operating on a smaller timescale) than the rate of consumer reproduction (operating on a relatively longer timescale).
-As will be shown in Section XX, this general expectation will hold for most classes of organisms, while the exact difference in timescales between reproduction and starvation can be derived using allometric scaling relationships.
+[Explain what a transcritical bifurcation means in terms of the dynamics] From a regime where resources are negative and consumer populations become infinite to one that is biological feasable]
+%As such, a transcritical bifurcation exists at $\lambda = \sigma$, such that the existence of an internal stable fixed point is dependent on the condition that $\sigma > \lambda$.
+This means that the rate of starvation must be greater (operating on a smaller timescale) than the rate of consumer reproduction (operating on a relatively longer timescale).
+This general expectation will hold for most classes of organisms, which we will show by deriving allometric scaling relationship that serve to constrain the covariation of parameter values.
+% while the exact difference in timescales between reproduction and starvation can be derived using allometric scaling relationships.
 %Kempes (REF) has shown that population growth rates scale to $M^{\alpha-1}$ (where $M$ is the reproductive mass of an organism and $\alpha$ is the scaling exponent), whereas starvation is necessarily bound to shorter timescales, which will be elaborated on below.
@@
 %Hopf bifurcation
-Oscillating, or cyclic, dynamics present additional risks to populations.
+[In addition to the hard bound defined by the TC bifurcation, oscillating or cyclic dynamics represent an implicit constraint by putting populations at risk...
+also present certain constraints to the feasibility of populations]
+%Oscillating, or cyclic, dynamics present additional risks to populations.
 If cycles are large, stochastic effects may result in extinction.
@@
 This condition is called a Hopf bifurcation, and is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian matrix.
-Although the Hopf condition for the specific 2-stage model cannot be easily written, the analytical solution can be explored using a symbolic computational language such as \emph{Mathematica}.
+%Although the Hopf condition for the specific 2-stage model cannot be easily written, the analytical solution can be explored using a symbolic computational language such as \emph{Mathematica}.
 
@@
 
-Nearly all of the rates described in the specific model, and generalized upon in the generalized model, are to some extent governed by the body size of the consumer.
-The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ plays a central role in other scaling relationships.
-Organismal metabolic rate $B$ is known to scale as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to be 3/4 for metazoans, etc.
-Kempes et al. [REF] show how the population-level growth rate also can be related to body size as $\lambda = \lambda_0 M^{\eta-1}$.
+[Link Allometry stuff to our model - what does it provide to the story]
+[Introduce biological and linked constraints]
+[Allometries capture vast amounts of diversity via a single parameter --- body size]
+[Allometries have captured int terest etc across multiple scales and biological classes of organisms]
 
-Population growth requires that individuals
+Nearly all of the rates described in the specific model, and later in the generalized model, are to some extent governed by the body size of the consumer.
+For example, the scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \cite{} and plays a central role in a variety of scaling relationships.
+Organismal metabolic rate $B$ is known to scale as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to be $3/4$ for metazoans, and varies in unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \cite{delong}.
+Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \cite{kempes}.
+More specifically, the interspecific trends in growth rate can be approximated by $\lambda = \lambda_0 M^{\eta-1}$.
+This relationship is derived from the simple balance
+ \begin{equation}
+ B_{0}m^{\alpha}=E_{m}\frac{dm}{dt}+B_{m}m
+ \label{balance}
+ \end{equation}
+ [a and b notation --- these parameters are easily measured bioenergetic parameters which are often approximately invariant across organisms of vastly different size.
+Our notation seeks to illustrate that the allometric model fundamentally depends on a small number of free parameters.]
+ where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is the metabolic rate to support an existing unit of mass, and $m$ is the mass at any point in development.
+It is useful to explicitly write this balance because it can also be modified to understand the rates of both starvation and recovery from starvation.
+[Spell out the connection to nutritional state more explicitly] [As we will see it is possible to derive both sigma and rho from this balance]
+
+For the rate of starvation, we make the simple assumption that an organism must meet its maintenance requirements using digested mass as the sole energy source.
+This assumption implies the simple metabolic balance
+\begin{equation}
+\frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
+\end{equation}
+where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing body mass which may differ from  $E_{m}$, the energy required to synthesis a unit of biomass.
+Give the adult mass,  $M$, of an organism this energy balance prescribes the mass trajectory of a starving organism:
+\begin{equation}
+m\left(t\right)=Me^{-B_{m}t/E_{m}^{\prime}}.
+\end{equation}
+Considering that only certain tissues can be digested for energy, for example the brain cannot be degraded to fuel metabolism, we define the rate for starvation and death by the timescales required to reach specific fractions of normal adult mass.
+We define $m_{starve}=\epsilon M$ where it could be the case that organisms have a systematic size-dependent requirement for essential tissues, such as the minimal bone or brain mass.
+For example, considering the observation that body fat in mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$, and assuming that once this mass is fully digested the organism begins to starve, would imply that $\epsilon=1-f_{0}M^{\gamma}/M$.
+Taken together the time scale for starvation is given by
+\begin{equation}
+t_{\sigma}=-\frac{E_{m}\log\left(\epsilon\right)}{B_{m}}.
+\end{equation}
+The starvation rate is $\sigma=1/t_{s}$, which implies that $\sigma$ is independent of adult mass if $\epsilon$ is a constant, and if $\epsilon$ does scale with mass, then $\sigma$ will have a factor of $1/\log\left(1-f_{0}M^{\gamma}/M\right)$.
+In either case $\sigma$ does not have a simple scaling with $\lambda$ which is important for the dynamics that we later discuss.
+
+The time to death should follow a similar relationship, but defined by a lower fraction of adult mass, $m_{death}=\epsilon^{\prime} M$.
+Consider, for example, that an organism dies once it has digested all fat and muscle tissues, and that muscle tissue scales with body mass according to $M_{mm}=mm_{0}M^{\zeta}$, then $\epsilon^{\prime}=1-\left(f_{0}M^{\gamma}+mm_{0}M^{\zeta}\right)/M$.
+Muscle mass has been shown to be roughly proportional to body mass \cite{} in mammals and thus  $\epsilon^{\prime}$ is effectively $\epsilon$ minus a constant.
+Thus
+\begin{equation}
+t_{d}=-\frac{E_{m}\log\left(\epsilon^{\prime}\right)}{B_{m}}
+\end{equation}
+and $\mu=1/t_{d}$.
+
+It should be noted that we have thus far used mammals an example to describe the size-based relationships for growth, starvation, and death.
+However, our presentation is general, and other functional forms for $\epsilon$, for example, could be determined for other classes of organisms.
+Considering bacteria, we might expect that starvation or death is defined by the complete digestion of proteins, and in Table \ref{parameter-values} we provide all parameter values for bacteria which we later use as a comparison in our analysis.
+
+Completing our allometric analysis, for the rate of becoming full, we again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass.
+The general solution to Equation \ref{balance} is given by
+\begin{equation}
+m\left(t\right)=\left[1-\left(1-\frac{b}{a}m_{0}^{1-\alpha}\right)e^{-b\left(1-\alpha\right)t}\right]^{1/\left(1-\alpha\right)}\left(\frac{a}{b}\right)^{1/\left(1-\alpha\right)}
+\end{equation}
+with $a=B_{0}/E_{m}$ and $b=B_{m}/E_{m}$.
+We are then interested in the timescale, $t_{f}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, which has the final form of
+\begin{equation}
+t_{f}=\frac{\log \left(1-\left(M \left(\frac{a}{b}\right)^{\frac{1}{\alpha -1}}\right)^{1-\alpha }\right)-\log \left(1-\left(M \epsilon  \left(\frac{a}{b}\right)^{\frac{1}{\alpha
+   -1}}\right)^{1-\alpha }\right)}{(\alpha -1) b}.
+\end{equation}
+
+\begin{table}[h]
+\caption{Parameter Values For Various Classes of Organisms}
+\label{parameter-values}
+    \begin{center}
+    \small
+     \begin{tabular}{| p{1.2cm}| p{3.2cm} | p{2.6cm} | p{3.2cm} | }
+     \hline
+     & {\bf Mammals} & {\bf Unicellular Eukaryotes} & {\bf Bacteria} \\
+     \hline
+   $\eta$ & $3/4$ & & $1.70$ \\ 
+   $E_{m}$ & $10695$ (J gram$^{-1}$) & & $10695$ (J gram$^{-1}$) \\ 
+   $E_{m}^{\prime}$ & $\approx E_{m}$ & & $\approx E_{m}$ \\ 
+   $B_{0}$ & $0.019$ (W gram$^{-\alpha}$) & & $1.96\times10^{17}$ \\
+   $B_{m}$ & $0.025$ (W gram$^{-1}$)   & & $0.025$ (W gram$^{-1}$)\\
+   $a$ & $1.78\times10^{-6}$ & & $1.83\times10^{13}$ \\ 
+   $b$ & $2.29\times10^{-6}$ & & $2.29\times10^{-6}$ \\  
+   $\eta-1$ & $-0.21$ & & $0.73$ \\ 
+   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) & & $56493$ \\ 
+   $\gamma$ & $1.19$ & & $0.68$ \\ 
+   $f_{0}$ & $0.02$ & & $1.30\times10^{-5}$ \\ 
+   $\zeta$ & $1.01$ & & \\ 
+   $mm_{0}$ & $0.32$ & & \\
+
+   \hline
+    \end{tabular}
+    \end{center}
+   \end{table}
+
+%Population growth requires that individuals
 
@@
 The transition from stable non-oscillatory dynamics to oscillatory dynamics occurs at the Hopf bifurcation condition where two complex conjugate eigenvalues cross the imaginary axis and attain positive real parts.
-Although there is an analytical solution to the Hopf bifurcation condition, it cannot be written efficiently
+Although there is an analytical solution to the Hopf bifurcation condition, it cannot be written efficiently.
+
 When the starvation rate is low, oscillatory dynamics are more likely to occur for a given value of the recovery rate.
@@
 %
+\newpage
 
@@
  \end{figure}
+
+  \begin{figure}[h]
+ 	\centering
+ 	\includegraphics[width=0.5\textwidth]{fig_TS.pdf}
+ 	\caption{
+ 	}
+ 	\label{Hopfb}
+ \end{figure}
+
 %
@@
 
+\newpage
+
+\section*{References}
+
 \bibliographystyle{ieeetr} % for Science,

### 2016-07-05 Manuscript/Yeakel_Kempes_Redner-7-5-2016.tex
New allometric extinction fig and manuscript updates

--- 2016-06-15 Manuscript/Yeakel_Kempes_Redner-5-16-2016.tex
@@
 \usepackage[T1]{fontenc}
-\usepackage[comma]{natbib}
+\usepackage[authoryear]{natbib}
+
+%\usepackage[style=authoryear,sorting=ynt]{biblatex}
+
 \usepackage{framed}
@@
 
-\section{Introduction}
 %Focus on the tradeoff between REPRODUCTION AND SURVIVAL AS A FUNCTION OF ENERGETIC STATE
@@
 Although this implicit dependence connects resource limitation to lower consumer growth rates, the following biological realities are not taken into account: 
-\begin{itemize}
-\item some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation,
-\item the portion of the population that is not nutritionally stressed is expected to reproduce at a near-constant rate, and
-\item the rates at which individuals transition from nutritionally poor to replete states and back are governed by the organism's metabolic capabilities of growing and maintaining tissue, which is allometrically constrained \citep{Kempes:2012hy}.
-\end{itemize}
-[The neglect of these features has important dynamic shortcomings:]
-[Here we use body size and allometries to capture a wide variety of taxonomic diversity]
-[There isn't a deep understanding with how models vary across different life histories, traits, etc... how do equations change with allometric constraints for different class of organisms? 
-i.e. Use body size as a generating parameter to capture a wide variety of diversity.]
+\emph{i}) some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation;
+\emph{ii} the portion of the population that is not nutritionally stressed is expected to reproduce at a near-constant rate and this is -- on average -- determined by body size;
+\emph{iii} the rates at which individuals transition from nutritionally poor to replete states and back are governed by metabolic constraints \citep{Kempes:2012hy}.
+Importantly, the neglect of additional constraints may have important dynamic shortcomings, masking the effects of resource limitation on consumer population dynamics.
 
-To incorporate these elements, resource limitation and the subsequent effects of starvation may be accounted for \emph{explicitly}, such that reproductive growth of a population is permitted only for individuals with sufficient energetic reserves.
-Incorporating energetic dynamics that occur at an individual level \citep{Kooi2000} into a population-based framework \citep{Kooi2000,Sousa:2010ez}, though straightforward conceptually, quickly runs into mathematical obstacles \citep{Diekmann:2010da}.
-[An alternative approach is to start with the macroscale patterns in order to understand certain tradeoffs etc]
+Resource limitation and the subsequent effects of starvation may be alternatively accounted for \emph{explicitly}, such that reproduction is permitted only for individuals with sufficient energetic reserves.
+Incorporating energetic dynamics that occur at an individual level \citep{Kooi2000} into a population-based framework \citep{Kooi2000,Sousa:2010ez}, though straightforward conceptually, quickly runs into mathematical obstacles \citep{Diekmann:2010da}, and suffers from over-parameterization.
 The complexities associated with scaling individual-level energetics to that of a population has thus limited the development of theoretical models that may aid our understanding of the effects of such tradeoffs on population dynamics.
+An alternative approach to individual-based frameworks is to concentrate on the macroscale relationships to examine the primary effects of reproductive vs. maintenance tradeoffs on populations.
+
+%[Here we use body size and allometries to capture a wide variety of taxonomic diversity]
+%[There isn't a deep understanding with how models vary across different life histories, traits, etc... how do equations change with allometric constraints for different class of organisms? 
+%i.e. Use body size as a generating parameter to capture a wide variety of diversity.]
 
 %What we did...
-Here we explore how the energetic tradeoff between maintaining and building tissue and reproduction, which occurs at the level of the individual, may influence the dynamics of a population.
-We first establish a simple nutritional state-structured population model that captures the essential dynamics of energetic reproductive tradeoffs, and explore the impact of different fluxes on system-level stability.
-By relating different rate constants to allometric constraints, we uncover important relationships between the timescales of physiological and reproductive processes, and show how organisms of different body sizes and taxonomic affinities are expected may be prone to different dynamics.
-The general rate laws that control both reproductive and energetic processes of the population, we show that
+Here we explore how the energetic tradeoff between maintaining and building tissue vs.
+reproduction can influence the dynamics of populations, and how these dynamics may be constrained by allometric relationships.
+We begin by establishing a simple nutritional state-structured population model, where consumer starvation, and cessation of reproduction, is the consequence of resource limitation.
+Similarly, recovery from the starved state -- and resumption of reproduction -- increases with resource density.
+%that captures the essential dynamics of energetic vs. reproductive tradeoffs
+%and explore the impact of different fluxes on system-level stability.
+Importantly, the rate at which consumers starve and recover from a starved state is wholly constrained by metabolism, which in turn is largely determined by body size, or allometry.
+By relating different rate constants to allometric constraints, we uncover important relationships between the timescales of physiological and reproductive processes, and show how organisms of different body sizes and taxonomic affinities may be prone to alternative dynamics.
 
-%We then develop a more general model to understand if there are common attributes of these energetically-constrained systems that have implications beyond assumptions inherent to the specific model.
-We show [WHAT]?
-%Finally, we compare our population-dynamic approach with a stochastic dynamic program that incorporates different individual-level behaviors as a function of energetic state.
+%By deriving allometric constraints for the starvation, recovery, and reproductive timescales of the consumer-resource system, 
+We show that rates of starvation and recovery (parameterized for endothermic organisms) always result in systems with stable, non-cyclic fixed points.
+Moreover, larger consumer body size results in rates that are less prone to cyclic dynamics than are rates for smaller organisms, pointing to potential empirical verification of the starvation dynamics framework.
+Finally, we show that rates of starvation and recovery appear to be constrained to a parameter range where both transient and equilibrial population dynamics result in the lowest risk of extinction for the consumer.
+This surprising result suggests that the risks associated with different fluxes of consumers in and out of a starved (non-reproductive) state may serve as an important selective driver over evolutionary time.
 
-%We devise the simplest energetic-compartment system to investigate the effects of a reproductive tradeoff with starvation on population dynamics.
-
-%FITNESS and ENERGETICS... the model incorporates fitness consequences of energetic loss...
-
-\section{Methods}
-\subsection{Model description}
-
-We first integrate energetics into the dynamics of a consumer-resource system by assuming that the consumer population can be divided into discrete energetic states, the occupation of each being contingent on the consumption of a single resource $R$.
-In the simplest case, there are only two energetic states for the consumer population: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
+\noindent {\bf Starvation dynamics} \\ \nonumber
+We integrate energetics into the dynamics of a consumer-resource system by assuming that the consumer population can be divided into discrete energetic states, the occupation of each being contingent on the consumption of a single resource $R$.
+In the Nutritional State Model (NSM) there are only two energetic states for the consumer population: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
 Consumers transition from state $F$ to state $H$ by starvation at rate $\sigma$ and in proportion to the lack of food $(1-R)$.
 Conversely, consumers recover from state $H$ to the full state $F$ at rate $\rho$ and in proportion to the density of resources consumed $R$.
-The resource has logistic growth with a linear growth rate $a$ and a carrying capacity of unity.
-Resources are eliminated by the consumer in both states: by energetically deficient consumers at rate $\rho$, and by energetically replete consumers at rate $b$.
+The resource has logistic growth with a linear growth rate $\alpha$ and a carrying capacity of unity.
+Resources are eliminated by the consumer in both states: by energetically deficient consumers at rate $\rho$, and by energetically replete consumers at rate $\beta$.
 Accordingly, the system of equations is written
@@
 
-[NSM = Nutritional State Model]
-[LVM = Lotka Volterra Model]
+%[NSM = Nutritional State Model]
+%[LVM = Lotka Volterra Model]
 
-There are three steady states for the 2-stage consumer-resource system: two trivial steady states at $(R^*=0,H^*=0,F^*=0)$ and $(R^*=1,H^*=0,F^*=0)$, and one non-trivial internal steady state where $(R^*>0,H^*>0,F^*>0)$.
+There are three steady states for the NSM: two trivial steady states at $(R^*=0,H^*=0,F^*=0)$ and $(R^*=1,H^*=0,F^*=0)$, and one non-trivial internal steady state where $(R^*>0,H^*>0,F^*>0)$.
 The latter steady state is the one of chief ecological interest, where
@@
 
-Analysis of the stability of the consumer-resource system is explored with respect to the local stability of the internal steady state, which is the only feasible steady state as long as both the consumer and resource have non-zero, positive, values.
+Analysis of the stability of the NSM is explored with respect to the local stability of the internal steady state, which is the only feasible steady state as long as both the consumer and resource have non-zero, positive, values.
 In a multidimensional system, linear stability is determined with respect to the Jacobian Matrix $\bf J$, which is a matrix where each element is defined by the partial derivative of each equation with respect to each variable.
-In the case of the 2-stage consumer model, the Jacobian evaluated at the internal steady state (denoted by $|_*$) is written
-
-\begin{equation}
-\mathbf{J}|_* =
+%In the case of the 2-stage consumer model, the Jacobian evaluated at the internal steady state (denoted by $|_*$) is written
+%
+%\begin{equation}
+%\mathbf{J}|_* =
+%%\left(
+%%\begin{array}{ccc}
+%% -F^* m+(1-R^*) a -R^* a -H^* \rho  & -R^* \rho  & -m R^* \\
+%% -H^* \rho -F^* \sigma  & -\mu -R^* \rho  & (1-R^*) \sigma  \\
+%% H^* \rho +F^* \sigma  & R^* \rho  & \lambda -(1-R^*) \sigma  \\
+%%\end{array}
+%%\right).
 %\left(
 %\begin{array}{ccc}
-% -F^* m+(1-R^*) a -R^* a -H^* \rho  & -R^* \rho  & -m R^* \\
-% -H^* \rho -F^* \sigma  & -\mu -R^* \rho  & (1-R^*) \sigma  \\
-% H^* \rho +F^* \sigma  & R^* \rho  & \lambda -(1-R^*) \sigma  \\
+% -\frac{\lambda  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
+% \frac{\lambda  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\mu  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
+% -\frac{m \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } & -\frac{\mu  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } \\
 %\end{array}
 %\right).
-\left(
-\begin{array}{ccc}
- -\frac{\lambda  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
- \frac{\lambda  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\mu  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
- -\frac{m \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } & -\frac{\mu  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } \\
-\end{array}
-\right).
-\end{equation}
+%\end{equation}
 
@@
 If the parameters of the Jacobian matrix at the internal steady state are such that its leading eigenvalue is $<0$, then the system is stable to small pulse perturbations, conditioned on the value of the starvation rate $\sigma$ relative to the value of the consumer reproduction rate $\lambda$.
-As $\sigma$ nears and becomes lower than a given $\lambda$, the resource steady state $R^*$ crosses the origin and exchanges stability to become unstable.
-[Explain what a transcritical bifurcation means in terms of the dynamics] From a regime where resources are negative and consumer populations become infinite to one that is biological feasable]
+As $\sigma$ becomes less than $\lambda$, the resource steady state $R^*$ crosses the origin and becomes unstable.
+The condition $\sigma = \lambda$  marks a transcritical bifurcation, such that if a consumer's starvation rate is lower than it's reproductive rate, the consumer population's growth becomes exponential and unphysical.
+Thus, the timescale of starvation must be less than the timescale of reproduction, an expectation that will hold for most classes of organisms, which we will show by deriving allometric scaling relationship that serve to constrain the covariation of parameter values.
+%[Explain what a transcritical bifurcation means in terms of the dynamics. From a regime where resources are negative and consumer populations become infinite to one that is biological feasable]
 %As such, a transcritical bifurcation exists at $\lambda = \sigma$, such that the existence of an internal stable fixed point is dependent on the condition that $\sigma > \lambda$.
-This means that the rate of starvation must be greater (operating on a smaller timescale) than the rate of consumer reproduction (operating on a relatively longer timescale).
-This general expectation will hold for most classes of organisms, which we will show by deriving allometric scaling relationship that serve to constrain the covariation of parameter values.
+%This means that the rate of starvation must be greater (operating on a smaller timescale) than the rate of consumer reproduction (operating on a relatively longer timescale).
+%This general expectation will hold for most classes of organisms, which we will show by deriving allometric scaling relationship that serve to constrain the covariation of parameter values.
 % while the exact difference in timescales between reproduction and starvation can be derived using allometric scaling relationships.
@@
 
-\subsection{Analysis of a generalized model}
-
-We may gain additional insight by that we do not know the specific rate functions from our 2-stage consumer resource model presented in Eq. \ref{eq_specific}.
-For example, we assume a linear mortality term for hungry foragers $\mu H$, though we may wish to assert that our knowledge of consumer mortality involves those that are energetically deficient, but nothing else.
-In this case, we would assume only that the rate of mortality is governed by the function $M(H)$.
-Substituting general functions for all rate laws from Eq. \ref{eq_specific}, we obtain the generalized ODE system
-
-\begin{align}
-\frac{\rm d}{\rm dt} F &= G(F) + S(R,H) - K(R,F), \nonumber \\
-\frac{\rm d}{\rm dt} H &= K(R,F) - S(R,H) - M(H), \nonumber \\
-\frac{\rm d}{\rm dt} R &= P(R) - L(R,H,F).
-\label{eq_gen}
-\end{align}
-
-\noindent where $G(F)$ determines consumer growth, $S(R,H)$ and $K(R,F)$ are the recovery and starvation functions, respectively, $M(H)$ determines consumer mortality, and $P(R)$ and $L(R,H,F)$ are functions describing the growth and consumption-loss of resources, respectively.
-
-If the system is written in this generalized manner, we cannot solve for the steady state solution $(F^*,~H^*,~R^*)$, however we can normalize the system to the unknown steady states.
-We denote normalized variables and functions in lowercase, such that $f = F/F^*$, $h = H/H^*$, $r = R/R^*$, and for example the normalized mortality function $m(h) = M(H)/M^*$, where $M^*$ is shorthand for $M(H^*)$.
-Additional rearrangements of terms under equilibrium conditions allows us to define two additional sets of scaling parameters with intuitive biological properties: the turnover rates of full foragers, hungry foragers, and the resource ($\alpha_f$, $\alpha_h$, $\alpha_r)$, and the proportional branching biomass through different compartments of the model, generally designated by the parameter $\beta$.
-For instance, $\beta_f$ is the proportion of full consumer growth due to reproduction, whereas $(1-\beta_f)$ is the proportion of full consumer growth due to recruitment \emph{from} the hungry forager class.
-Similarly, $\beta_h$ is the proportion of hungry consumer loss due to mortality, whereas $(1-\beta_h)$ is the proportion of hungry consumer loss due to recruitment \emph{into} the full consumer class (see SUPP for a detailed derivation).
-Substituting the normalized variables and functions into Eq. \ref{eq_gen}, we obtain
-
-\begin{align}
-\dot{f} &= \alpha_f \left[ \beta_f g(f) + (1-\beta_f)s(r,h) - k(r,f)\right], \nonumber \\
-\dot{h} &= \alpha_h \left[ k(r,f) - (1-\beta_h)s(r,h) - \beta_h m(h) \right], \nonumber \\
-\dot{r} &= \alpha_r \left[ p(r) - l(r,h,f) \right],
-\label{eq_gen}
-\end{align}
-
-\noindent and linearization of this normalized, general ODE system yields the Jacobian matrix
-
-\begin{equation}
-\mathbf{J}_{\rm gen}|_* =
-\left(
-\begin{array}{ccc}
- \alpha_f \left(\beta_f \frac{\partial g}{\partial f} - \frac{\partial k}{\partial f} \right) & \alpha_f(1-\beta_f)\frac{\partial s}{\partial h} & \alpha_f \left( (1-\beta_f)\frac{\partial s}{\partial r} - \frac{\partial k}{\partial r} \right) \\
- \alpha_h \frac{\partial k}{\partial f} & -\alpha_h \left(\beta_h \frac{\partial m}{\partial h} + (1-\beta_h) \frac{\partial s}{\partial h} \right) & \alpha_h \left(\frac{\partial k}{\partial r} - (1 - \beta_h)\frac{\partial s}{\partial r} \right) \\
- -\alpha_r \frac{\partial l}{\partial f} & -\alpha_r \frac{\partial l}{\partial h} & \alpha_r \left( \frac{\partial p}{\partial r} - \frac{\partial l}{\partial r} \right) \\
-\end{array}
-\right).
-\label{eq_Jacgen}
-\end{equation}
-
-%Importance of the elasticities
-Due to the normalization procedure, the partial derivatives in Eq. \ref{eq_Jacgen} have tangible biological meaning.
-Because, for example, the partial derivative (containing functions and variables normalized to the unknown steady states) $\partial g / \partial f = \partial \log G / \partial \log F$, it scales in such a way that it represents the percent change in consumer growth (governed by $G(F)$) relative to a percent change in the density of full consumers $F$, more commonly known as a functional elasticity.
-For example, if growth is a linear function (e.g.
-$G(F) = \lambda F$), $\partial g / \partial f = 1$; if growth is a quadratic function (e.g.
-$G(F) = \lambda F^2$), $\partial g / \partial f = 2$, while more complex functions may depend on the value of the steady state.
-For example, if consumer growth is modeled as Holling Type II growth, such that $G(F) = c_1 F^2/(c_2 + F^2)$, where $c_1$ and $c_2$ are unknown constants, then its elasticity will vary between $0$ and $2$, depending on the steady state value $F^*$, which is unknown in the generalized system.
-
-%Why do this?
-%The normalization of the generalized system described above results in the inclusion of functional elasticities into the generalized Jacobian matrix. 
-Deriving a Jacobian in terms of the normalized ODE system is useful because it allows us to place strict constraints on the values of the unknown variables (the turnover rates, biomass branching parameters, and the functional elasticities), without assuming detailed knowledge of the functions controlling different rates within the system.
-% retaining the fundamental architecture of the 2-class consumer-resource system.
-%Simplifications
-In addition, we can now insert a number of assumptions that will align our generalized Jacobian more closely with the original 2-stage consumer resource system.
-We will assume the following:
-\emph{i}) both the consumer and resource suffers linear mortality,
-\emph{ii}) resource growth is logistic,
-\emph{iii}) recovery and starvation are linear with respect to both full and hungry consumer densities, and
-\emph{iv}) consumers and resources have equivalent turnover rates scaled to unity.
-These assumptions lead to the following simplifications
-
-\begin{align}
-	i)~&\frac{\partial m}{\partial h} = 1, ~~~~~~~~ \frac{\partial l}{\partial r} = 1, ~~~~~~~~ \frac{\partial l}{\partial f} = 1 - \frac{\partial l}{\partial h}, \nonumber \\
-	ii)~&\frac{\partial k}{\partial r} = \left(1 - \frac{1}{R^*} \right)^{-1}, \nonumber \\
-	iii)~&\frac{\partial s}{\partial r} = \frac{\partial s}{\partial h} = 1, ~ \frac{\partial k}{\partial f} = 1, \nonumber \\
-	iv)~&\alpha_f = \alpha_r = 1,
-\end{align}
-
-\noindent where $R^*$ ranges from 0 to 1.
-The remaining free parameters include the timescale of hungry foragers $\alpha_h$, the branching parameters $\beta_f$ and $\beta_h$, the elasticity of consumer growth with respect to full consumer densities $\partial g / \partial f$, the elasticity of resource growth with respect to resource density $\partial p / \partial r$, the elasticity of resource loss with respect to full consumer density $\partial l / \partial f$, and the elasticity of starvation with respect to resource density $\partial k / \partial r$.
-See Table XX for a list of the free parameters in the generalized model, as well as the ranges of potential values for each.
-These substitutions result in the simplified Jacobian matrix
-
-\begin{equation}
-\mathbf{J}_{\rm gen}|_* =
-\left(
-\begin{array}{ccc}
- \left(\beta_f \frac{\partial g}{\partial f} - 1 \right) & (1-\beta_f) & \left( (1-\beta_f)\frac{\partial s}{\partial r} - 1 \right) \\
- \alpha_h & -\alpha_h & -\alpha_h \left(1-\frac{\partial k}{\partial r} - \beta_h  \right) \\
- \frac{\partial l}{\partial h}-1 & -\frac{\partial l}{\partial h} &  \left( \frac{\partial p}{\partial r} - 1 \right) \\
-\end{array}
-\right).
-\label{eq_Jacgen}
-\end{equation}
-
-%Testing stability
-Because the remaining free parameters have known ranges, but not specific values, we wish to assess the correlations of each to system stability.
-By randomly drawing values from uniform distributions bounded by the known ranges of each free parameter, we obtain an ensemble of potential Jacobian matrices whereupon the stability of each is determined by numerically calculating the real part of the leading eigenvalue.
-If the real part of the leading eigenvalue is $<1\times10^{-6}$, we assumed the system to be stable.
-Replicating this procedure on the order of $10^7$ times allowed us to determine the correlation of each free parameter with stability to the extent that variance was negligible.
-
-\subsection{Allometric scaling relationships}
-
+\noindent {\bf Allometric constraints} \\ \nonumber
 [Link Allometry stuff to our model - what does it provide to the story]
@@
 
-\section{Results}
+\noindent {\bf Allometry leads to stability} \\ \noindent
+[First describe the dynamics of the model and HOW IT WORKS]
+[Then Allometry and stability]
 
@@
 
-\subsection{Generalized analysis}
-Constraining the model with allometric scaling relationships yields important insights regarding the tradeoff between [fitness and energy]
+\noindent {\bf Minimizing extinction risk} \\ \nonumber
 
-%The natural ordering of the scaled rates
-%lambda < rho < sigma
-%if lambda is 0.2, and sigma is 0.8, then rho is between 0.2 and 0.8.
-%In other words, lambda has the SLOWEST timescale, sigma has the fastest timescale, and rho has a timescale intermediate to both.
-
-\section{Discussion}
-
-%Starving Random Walk
-In our modeling, foragers look for food by wandering in this changing
-environment.  If such a search is successful, the forager is satiated and it
-can engage in the essential activity of reproduction.  However, if the search
-is unsuccessful for a sufficiently long period, the forager ``starves''.
-Such a forager can do nothing else but forage, until it either finds food and
-again becomes satiated or it dies when it goes too long without finding
-nourishment.  These rules are reminiscent of the ``starving random walk''
-model, where a single random walk can take $\mathcal{S}$ steps without
-encountering food before starving to death.  Moreover, the resource does not
-regenerate, so that the forager ultimately starves to death.  For this
-idealized model, it was found that the average lifetime of the forager scales
-algebraically with $\mathcal{S}$ in $d\!=\!1$ and $d\!=\!2$ dimensions, and
-as $\exp(-A\mathcal{S}^\omega)$ for $d>2$.  Here the exponent
-$\omega\approx \frac{1}{2}$ for $d=3$, while $\omega\to 1$ only as
-$d\to \infty$, with the latter behavior corresponding to the mean-field
-limit.  As we will discuss, regeneration of the resource, together with the
-behavioral change between starving and satiated foragers leads to still much
-richer dynamical behavior.
+%
+%
+%%Starving Random Walk
+%In our modeling, foragers look for food by wandering in this changing
+%environment.  If such a search is successful, the forager is satiated and it
+%can engage in the essential activity of reproduction.  However, if the search
+%is unsuccessful for a sufficiently long period, the forager ``starves''.
+%Such a forager can do nothing else but forage, until it either finds food and
+%again becomes satiated or it dies when it goes too long without finding
+%nourishment.  These rules are reminiscent of the ``starving random walk''
+%model, where a single random walk can take $\mathcal{S}$ steps without
+%encountering food before starving to death.  Moreover, the resource does not
+%regenerate, so that the forager ultimately starves to death.  For this
+%idealized model, it was found that the average lifetime of the forager scales
+%algebraically with $\mathcal{S}$ in $d\!=\!1$ and $d\!=\!2$ dimensions, and
+%as $\exp(-A\mathcal{S}^\omega)$ for $d>2$.  Here the exponent
+%$\omega\approx \frac{1}{2}$ for $d=3$, while $\omega\to 1$ only as
+%$d\to \infty$, with the latter behavior corresponding to the mean-field
+%limit.  As we will discuss, regeneration of the resource, together with the
+%behavioral change between starving and satiated foragers leads to still much
+%richer dynamical behavior.
 
@@
  	\centering
- 	\includegraphics[width=0.5\textwidth]{fig_HopfPlotb.pdf}
+ 	\includegraphics[width=1\textwidth]{fig_DataHopf.pdf}
  	\caption{
@@
  	\centering
- 	\includegraphics[width=0.5\textwidth]{fig_TS.pdf}
+ 	\includegraphics[width=1\textwidth]{fig_ExtinctionAllometric.pdf}
  	\caption{
@@
 
-\bibliographystyle{ieeetr} % for Science,
+\bibliographystyle{apalike} % for Science,
 \bibliography{aa_starving}

### 2016-07-09 Manuscript/Yeakel_Kempes_Redner-7-5-2016.tex
allometric extinction sims and fp analysis

@@
 An individual's energetic state directly influences how it invests its stores in an uncertain environment.
-Such behaviors are generally manifested as trade-offs, which often concern investing in individual maintenance and growth or allocating energy towards reproduction \citep{Kirk:1997cc,Kempes:2012hy}.
+Such behaviors are generally manifested as trade-offs, which often concern investing in individual maintenance and growth (somatic effort) or allocating energy towards reproduction (reproductive effort) \citep{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
 %among a host of other behavioral duties [REFS].
@@
 %For example, rotifers that reproduce have lower survival rates than those that don't when they are under nutritional stress.
-To what extent, and when, organisms invest in these two necessary biological functions -- growth and maintenance vs.
-reproduction -- may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, as well as resource limitation.
+To what extent, and when, organisms invest in somatic vs.
+reproductive expenditures may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, as well as resource limitation.
 Importantly, the influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
@@
 Maximizing fitness between growth and maintenance activities vs.
-reproductive behaviors in large part structures the life-history of species, and this can be achieved by a variety of potential mechanisms, all of which, to some extent, depend on resource availability.
-[In extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the individual regulates all aspects of the cell cycle.]
+reproductive efforts in large part structures the life-history of species, and this can be achieved by alternative behavioral strategies or via physiological switches, both of which are sensitive to resource availability.
 %Reproduction can be separated from somatic growth/maintenance by a variety of potential mechanisms:
 %Multiple mechanisms contribute to the differentiation of these behaviors:
-Behavioral: The investment of time and energy towards reproductive and parental behaviors depends on resource availability \citep{Morris:1987eo}.
+Behavioral changes in somatic or reproductive investment occur in response to limited resources \citep{Morris:1987eo}.
 %, while stress can generally lead to reprodusuppression of reproductive behaviors \citep{Morgan:1999do}. 
-For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \citep{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for a breeding season \citep{Barboza:2002in}.
+For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \citep{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for a breeding season \citep{Martin:1987dl,Barboza:2002in}.
+Such breeding behaviors is generally referred to as \emph{capital breeding}
 %Resource limitation can also alter the behaviors of species in well-mixed environments: 
 Freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress \citep{Threlkeld:1976ih}, with those that do reproduce evincing lower survival rates \citep{Kirk:1997cc}, while artificially induced stress has been observed to decrease reproductive success in Atlantic cod \citep{Morgan:1999do}.
-Physiological: Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation) to time with accumulation of nutritional reserves \citep{Mead:1989dt,Sandell:1990kw}.
+Organisms may also separate maintenance and growth from reproduction over space and time. 
+Many salmonids, birds, and some mammals return to migratory breeding sites to reproduce after one or multiple seasons in alternative environments spent accumulating body mass and nutritional reserves \citep{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+
+Physiological mechanisms also play an important role in regulating reproductive expenditures during periods of high-risk.
+Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation) to time with accumulation of nutritional reserves \citep{Mead:1989dt,Sandell:1990kw}.
 %, including multiple species of Grizzly bears (\emph{Ursus horribilis}) and the European badger (\emph{Meles melesc}) \citep{Woodroffe:1995id
-Furthermore, many mammals (including humans) suffer irregular menstrual cycling and higher rates of spontaneous abortion during periods of nutritional stress \citep{Bulik:1999eo,Trites:2003cc}.
-Spatio-temporal: Organisms may also separate maintenance/growth from reproduction over space and time. 
-For example, many salmonids, birds, and some mammals return to migratory breeding sites to reproduce after one or multiple seasons in alternative environments spent accumulating body mass and nutritional reserves \citep{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+Furthermore, many mammals (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress \citep{Bulik:1999eo,Trites:2003cc}.
+In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the individual regulates all aspects of the cell cycle \citep{Glazier:2009hq}.
 %great tits invest more energy into broods when resource abundance is greater, while food availability is generally seen to correlate with clutch size among many bird species \citep{Daan:1988va}.
@@
 foraging effort over both space and time, whereas other rely on physiological mechanisms to optimize reproductive success against survival when adequate resources are unavailable.
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms points to the importance and universality of the fundamental tradeoff between energetic expenditures on the growth and maintenance of tissues vs.
-reproductive behaviors.
+The existence of so many independently evolved mechanisms across such a diverse suite of organisms points to the importance and universality of the fundamental tradeoff between somatic and reproductive investment, however the dynamic implications of these constraints are unknown.
 
 %To population-level effects
-The different strategies employed to avoid reproduction during times of nutritional stress, and how this is achieved among different organisms, has received tremendous empirical and theoretical attention owing to the importance of these activities in shaping life-history [REF].
+The mechanisms by which different organisms avoid or delay reproduction during times of nutritional stress has received tremendous empirical and theoretical attention owing to the importance of these activities in shaping life-history \citep{Cody:1966fk,Martin:1987dl,Mangel:1988uaa}.
 Less well understood is how resource limitation and these behavioral/physiological tradeoffs affect dynamics at the level of the population.
 % when organisms are able to avoid the risk of spending energy on reproduction when starvation is likely.
-Traditional Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction.
-Although this implicit dependence connects resource limitation to lower consumer growth rates, the following biological realities are not taken into account: 
+Traditional Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction \citep{murdoch:2003}.
+Although this implicit dependence connects resource limitation to lower consumer growth rates, the following biological realities are not included: 
 \emph{i}) some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation;
-\emph{ii} the portion of the population that is not nutritionally stressed is expected to reproduce at a near-constant rate and this is -- on average -- determined by body size;
-\emph{iii} the rates at which individuals transition from nutritionally poor to replete states and back are governed by metabolic constraints \citep{Kempes:2012hy}.
-Importantly, the neglect of additional constraints may have important dynamic shortcomings, masking the effects of resource limitation on consumer population dynamics.
+\emph{ii}) the portion of the population that is not nutritionally stressed is expected to reproduce at a near-constant rate and this is -- averaged across species -- determined strongly by body size \citep{Kempes:2012hy};
+\emph{iii}) the rates that individuals transition from nutritionally poor to replete states and back have different metabolically-constrained timescales that can lead to reproductive lags.
+Importantly, the exclusion of these biological details may have important dynamic shortcomings, masking the effects of starvation on consumer population dynamics.
 
 Resource limitation and the subsequent effects of starvation may be alternatively accounted for \emph{explicitly}, such that reproduction is permitted only for individuals with sufficient energetic reserves.
-Incorporating energetic dynamics that occur at an individual level \citep{Kooi2000} into a population-based framework \citep{Kooi2000,Sousa:2010ez}, though straightforward conceptually, quickly runs into mathematical obstacles \citep{Diekmann:2010da}, and suffers from over-parameterization.
-The complexities associated with scaling individual-level energetics to that of a population has thus limited the development of theoretical models that may aid our understanding of the effects of such tradeoffs on population dynamics.
-An alternative approach to individual-based frameworks is to concentrate on the macroscale relationships to examine the primary effects of reproductive vs. maintenance tradeoffs on populations.
+Though straightforward conceptually, incorporating the energetic dynamics of individuals \citep{Kooi2000} into a population-level framework \citep{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles \citep[in particular a lack of smoothness, or differentiability;][]{Diekmann:2010da}, and often suffers from over-fitting due to an overabundance of parameters.
+These issue have limited the development of theoretical models that may aid our understanding of the effects of such tradeoffs on population dynamics.
+An alternative approach to individual-to-population frameworks is redirecting focus to macroscale relationships guiding somatic vs. reproductive investment in a consumer-resource system.
 
@@
 %What we did...
-Here we explore how the energetic tradeoff between maintaining and building tissue vs.
-reproduction can influence the dynamics of populations, and how these dynamics may be constrained by allometric relationships.
-We begin by establishing a simple nutritional state-structured population model, where consumer starvation, and cessation of reproduction, is the consequence of resource limitation.
-Similarly, recovery from the starved state -- and resumption of reproduction -- increases with resource density.
+Here we explore how the energetic tradeoff between maintaining and building somatic tissue vs.
+reproduction can influence the dynamics of populations, and how such dynamics may be determined by allometric constraints.
+We begin by establishing a simple Nutritional State-structured population Model (NSM), where consumer starvation, and cessation of reproduction, is the consequence of resource limitation.
+Similarly, recovery from the starved to a reproductive state increases with resource density.
 %that captures the essential dynamics of energetic vs. reproductive tradeoffs
 %and explore the impact of different fluxes on system-level stability.
-Importantly, the rate at which consumers starve and recover from a starved state is wholly constrained by metabolism, which in turn is largely determined by body size, or allometry.
+Importantly, the rate at which consumers decline and recover from a starved state is wholly constrained by metabolism.
 By relating different rate constants to allometric constraints, we uncover important relationships between the timescales of physiological and reproductive processes, and show how organisms of different body sizes and taxonomic affinities may be prone to alternative dynamics.
@@
 %By deriving allometric constraints for the starvation, recovery, and reproductive timescales of the consumer-resource system, 
-We show that rates of starvation and recovery (parameterized for endothermic organisms) always result in systems with stable, non-cyclic fixed points.
-Moreover, larger consumer body size results in rates that are less prone to cyclic dynamics than are rates for smaller organisms, pointing to potential empirical verification of the starvation dynamics framework.
+We show that rates of starvation and recovery tend to result in systems with stable, non-cyclic fixed points and that the ratio between the rate of starvation and recovery may -- in part -- contribute to lower size limitations.
+Moreover, larger consumer body size results in rates that are less prone to cyclic dynamics than are rates for smaller organisms, pointing to potential empirical verification of the NSM framework.
 Finally, we show that rates of starvation and recovery appear to be constrained to a parameter range where both transient and equilibrial population dynamics result in the lowest risk of extinction for the consumer.
@@
 \noindent {\bf Starvation dynamics} \\ \nonumber
-We integrate energetics into the dynamics of a consumer-resource system by assuming that the consumer population can be divided into discrete energetic states, the occupation of each being contingent on the consumption of a single resource $R$.
-In the Nutritional State Model (NSM) there are only two energetic states for the consumer population: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
-Consumers transition from state $F$ to state $H$ by starvation at rate $\sigma$ and in proportion to the lack of food $(1-R)$.
-Conversely, consumers recover from state $H$ to the full state $F$ at rate $\rho$ and in proportion to the density of resources consumed $R$.
+We integrate the somatic/reproductive tradeoff into the dynamics of a consumer-resource system by dividing the consumer population into discrete energetic states, the occupation of each being contingent on the consumption of a single resource $R$.
+In the NSM there are only two energetic states for the consumer population: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
+Consumers transition from state $F$ to state $H$ by starvation at rate $\sigma$ and in proportion to the lack of resources $(1-R)$.
+Conversely, consumers recover from state $H$ to the full state $F$ at rate $\rho$ and in proportion to $R$.
 The resource has logistic growth with a linear growth rate $\alpha$ and a carrying capacity of unity.
@@
 \begin{split}
-\frac{\rm d}{\rm dt} F &= \lambda F + \rho RH - \sigma (1-R)F,  \\
-\frac{\rm d}{\rm dt} H &= \sigma (1-R)F - \rho RH - \mu H,  \\
-\frac{\rm d}{\rm dt} R &= \alpha R(1-R) - R(\rho H+ \beta F).
+\dot{F} &= \lambda F + \rho RH - \sigma (1-R)F,  \\
+\dot{H} &= \sigma (1-R)F - \rho RH - \mu H,  \\
+\dot{R} &= \alpha R(1-R) - R(\rho H+ \beta F).
 \label{eq_specific}
@@
 
-There are three steady states for the NSM: two trivial steady states at $(R^*=0,H^*=0,F^*=0)$ and $(R^*=1,H^*=0,F^*=0)$, and one non-trivial internal steady state where $(R^*>0,H^*>0,F^*>0)$.
-The latter steady state is the one of chief ecological interest, where
+There are three steady states for the NSM: two trivial fixed points at $(R^*=0,H^*=0,F^*=0)$ and $(R^*=1,H^*=0,F^*=0)$, and one non-trivial internal fixed point at
 
@@
 \begin{split}
-F^* &= \frac{a  \lambda  \mu  (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  m)}, \\
-H^* &= \frac{a  \lambda ^2 (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  m)}, \\
+F^* &= \frac{\alpha  \lambda  \mu  (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  \beta)}, \\
+H^* &= \frac{\alpha  \lambda ^2 (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  \beta)}, \\
 R^* &= \frac{\mu  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma }.
@@
 
-\noindent Because there is only one internal steady state, as long as it is stable the population trajectories will be globally attracted to it for any set of initial conditions greater than zero.
-
-Analysis of the stability of the NSM is explored with respect to the local stability of the internal steady state, which is the only feasible steady state as long as both the consumer and resource have non-zero, positive, values.
-In a multidimensional system, linear stability is determined with respect to the Jacobian Matrix $\bf J$, which is a matrix where each element is defined by the partial derivative of each equation with respect to each variable.
+\noindent Because there is only one internal fixed point, as long as it is stable the population trajectories will serve as a global attractor for any set of positive initial conditions.
+%Analysis of the stability of the NSM is explored with respect to the local stability of the internal steady state, which is the only feasible steady state as long as both the consumer and resource have non-zero, positive, values.
+In a multidimensional system, linear stability is determined with respect to the Jacobian Matrix $\bf J|_*$ (where $|_*$ denotes evaluation at the internal steady state), where each element of the matrix is defined by the partial derivative of each differential equation with respect to each variable.
 %In the case of the 2-stage consumer model, the Jacobian evaluated at the internal steady state (denoted by $|_*$) is written
@@
 %Transcritical bifurcation
-If the parameters of the Jacobian matrix at the internal steady state are such that its leading eigenvalue is $<0$, then the system is stable to small pulse perturbations, conditioned on the value of the starvation rate $\sigma$ relative to the value of the consumer reproduction rate $\lambda$.
-As $\sigma$ becomes less than $\lambda$, the resource steady state $R^*$ crosses the origin and becomes unstable.
-The condition $\sigma = \lambda$  marks a transcritical bifurcation, such that if a consumer's starvation rate is lower than it's reproductive rate, the consumer population's growth becomes exponential and unphysical.
-Thus, the timescale of starvation must be less than the timescale of reproduction, an expectation that will hold for most classes of organisms, which we will show by deriving allometric scaling relationship that serve to constrain the covariation of parameter values.
+If the parameters of $\bf J|_*$ are such that the real part of its leading eigenvalue is $<0$, then the system is stable to small pulse perturbations. 
+In the case of the NSM, stability is conditioned on the consumer's starvation rate $\sigma$ relative to the its reproduction rate $\lambda$.
+As $\sigma$ becomes less than $\lambda$, the $R^*$ crosses the origin and the fixed point becomes unstable.
+This condition $\sigma = \lambda$  marks a transcritical (TC) bifurcation, thus marking a hard boundary below which the system becomes unphysical due to the unregulated growth of the consumer population.
+That the timescale of reproduction is larger than the timescale of starvation is intuitive for macroscopic organisms, as the rate at which one loses tissue due to lack of resources is generally much faster than reproduction, though we will show using allometric principles that this relationship always holds.
+%such that if a consumer's starvation rate is lower than it's reproductive rate, the consumer population's growth becomes exponential and unphysical.
+%Thus, the timescale of starvation must be less than the timescale of reproduction, an expectation that will hold for most classes of organisms, which we will show by deriving allometric scaling relationship that serve to constrain the covariation of parameter values.
 %[Explain what a transcritical bifurcation means in terms of the dynamics. From a regime where resources are negative and consumer populations become infinite to one that is biological feasable]
@@
 %Hopf bifurcation
-[In addition to the hard bound defined by the TC bifurcation, oscillating or cyclic dynamics represent an implicit constraint by putting populations at risk...
-also present certain constraints to the feasibility of populations]
+In addition to the hard bound defined by the TC bifurcation, oscillating or cyclic dynamics present an implicit constraint to persistence by increasing the risk of extinction.
+%also present certain constraints to the feasibility of populations.
 %Oscillating, or cyclic, dynamics present additional risks to populations.
-If cycles are large, stochastic effects may result in extinction.
-In continuous-time systems, cycles arise when a pair of complex conjugate eigenvalues cross the imaginary axis and attain positive real parts.
+If cycles are large, stochastic effects may result in either the consumer or resource population.
+In continuous-time systems, a stable limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts \citep{GuckHolmes}.
 %Of additional interest is the existence of parameter regions that permit the existence of oscillatory, or cyclic, dynamics.
-This condition is called a Hopf bifurcation, and is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian matrix.
+This condition is called a Hopf bifurcation, and is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian \citep{Gross:2004p2428}.
+In addition to the onset of stable cycles, as a system with a stable fixed point nears the Hopf bifurcation, transient or decaying cycles can grow in magnitude.
+Given that ecological systems exist in a state of constant perturbation \citep{Hastings:2001jh}, even the onset of transient cycles that decay over time can increase the risk of extinction \citep{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}, such that the distance of a system from the Hopf bifurcation, even in the stable region, is relevant to persistence.
+
 %Although the Hopf condition for the specific 2-stage model cannot be easily written, the analytical solution can be explored using a symbolic computational language such as \emph{Mathematica}.
@@
 \noindent {\bf Allometric constraints} \\ \nonumber
-[Link Allometry stuff to our model - what does it provide to the story]
-[Introduce biological and linked constraints]
-[Allometries capture vast amounts of diversity via a single parameter --- body size]
-[Allometries have captured int terest etc across multiple scales and biological classes of organisms]
+%[Link Allometry stuff to our model - what does it provide to the story]
+%[Introduce biological and linked constraints]
+%[Allometries capture vast amounts of diversity via a single parameter --- body size]
+%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
+Chris - Linking paragraph here -- what does allometry provide to the story? Biological, linked constraints... vast amounts of diversity in a small number of parameters. Etc.
 
-Nearly all of the rates described in the specific model, and later in the generalized model, are to some extent governed by the body size of the consumer.
-For example, the scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \cite{} and plays a central role in a variety of scaling relationships.
-Organismal metabolic rate $B$ is known to scale as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to be $3/4$ for metazoans, and varies in unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \cite{delong}.
-Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \cite{kempes}.
+%Second paragraph
+Nearly all of the rates described in the NSM are to some extent governed by consumer metabolism, and thus can be estimated based on known allometric constraints. 
+The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \citep{West:2002it} and plays a central role in a variety of scaling relationships.
+Organismal metabolic rate $B$ is known to scale as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to be $3/4$ for metazoans, and varies in unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \citep{DeLong:2010dy}.
+Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \citep{Kempes:2012hy}.
 More specifically, the interspecific trends in growth rate can be approximated by $\lambda = \lambda_0 M^{\eta-1}$.
@@
  \begin{equation}
- B_{0}m^{\alpha}=E_{m}\frac{dm}{dt}+B_{m}m
+ B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m
  \label{balance}
@@
 \end{equation}
-The starvation rate is $\sigma=1/t_{s}$, which implies that $\sigma$ is independent of adult mass if $\epsilon$ is a constant, and if $\epsilon$ does scale with mass, then $\sigma$ will have a factor of $1/\log\left(1-f_{0}M^{\gamma}/M\right)$.
+The starvation rate is $\sigma=1/t_{\sigma}$, which implies that $\sigma$ is independent of adult mass if $\epsilon$ is a constant, and if $\epsilon$ does scale with mass, then $\sigma$ will have a factor of $1/\log\left(1-f_{0}M^{\gamma}/M\right)$.
 In either case $\sigma$ does not have a simple scaling with $\lambda$ which is important for the dynamics that we later discuss.
@@
 Consider, for example, that an organism dies once it has digested all fat and muscle tissues, and that muscle tissue scales with body mass according to $M_{mm}=mm_{0}M^{\zeta}$, then $\epsilon^{\prime}=1-\left(f_{0}M^{\gamma}+mm_{0}M^{\zeta}\right)/M$.
-Muscle mass has been shown to be roughly proportional to body mass \cite{} in mammals and thus  $\epsilon^{\prime}$ is effectively $\epsilon$ minus a constant.
+Muscle mass has been shown to be roughly proportional to body mass \cite{muscle} in mammals and thus  $\epsilon^{\prime}$ is effectively $\epsilon$ minus a constant.
 Thus
 \begin{equation}
-t_{d}=-\frac{E_{m}\log\left(\epsilon^{\prime}\right)}{B_{m}}
+t_{\mu}=-\frac{E_{m}\log\left(\epsilon^{\prime}\right)}{B_{m}}
 \end{equation}
-and $\mu=1/t_{d}$.
+and $\mu=1/t_{\mu}$.
 
-It should be noted that we have thus far used mammals an example to describe the size-based relationships for growth, starvation, and death.
+It should be noted that we have thus far used mammalian allometry to describe the size-based relationships for growth, starvation, and death.
 However, our presentation is general, and other functional forms for $\epsilon$, for example, could be determined for other classes of organisms.
@@
 
-Completing our allometric analysis, for the rate of becoming full, we again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass.
+The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues tissue from the starving state to the full state.
+We again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass.
 The general solution to Equation \ref{balance} is given by
@@
 with $a=B_{0}/E_{m}$ and $b=B_{m}/E_{m}$.
-We are then interested in the timescale, $t_{f}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, which has the final form of
+We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, which has the final form of
 \begin{equation}
-t_{f}=\frac{\log \left(1-\left(M \left(\frac{a}{b}\right)^{\frac{1}{\alpha -1}}\right)^{1-\alpha }\right)-\log \left(1-\left(M \epsilon  \left(\frac{a}{b}\right)^{\frac{1}{\alpha
+t_{\rho}=\frac{\log \left(1-\left(M \left(\frac{a}{b}\right)^{\frac{1}{\alpha -1}}\right)^{1-\alpha }\right)-\log \left(1-\left(M \epsilon  \left(\frac{a}{b}\right)^{\frac{1}{\alpha
    -1}}\right)^{1-\alpha }\right)}{(\alpha -1) b}.
 \end{equation}
-
-\begin{table}[h]
-\caption{Parameter Values For Various Classes of Organisms}
-\label{parameter-values}
-    \begin{center}
-    \small
-     \begin{tabular}{| p{1.2cm}| p{3.2cm} | p{2.6cm} | p{3.2cm} | }
-     \hline
-     & {\bf Mammals} & {\bf Unicellular Eukaryotes} & {\bf Bacteria} \\
-     \hline
-   $\eta$ & $3/4$ & & $1.70$ \\ 
-   $E_{m}$ & $10695$ (J gram$^{-1}$) & & $10695$ (J gram$^{-1}$) \\ 
-   $E_{m}^{\prime}$ & $\approx E_{m}$ & & $\approx E_{m}$ \\ 
-   $B_{0}$ & $0.019$ (W gram$^{-\alpha}$) & & $1.96\times10^{17}$ \\
-   $B_{m}$ & $0.025$ (W gram$^{-1}$)   & & $0.025$ (W gram$^{-1}$)\\
-   $a$ & $1.78\times10^{-6}$ & & $1.83\times10^{13}$ \\ 
-   $b$ & $2.29\times10^{-6}$ & & $2.29\times10^{-6}$ \\  
-   $\eta-1$ & $-0.21$ & & $0.73$ \\ 
-   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) & & $56493$ \\ 
-   $\gamma$ & $1.19$ & & $0.68$ \\ 
-   $f_{0}$ & $0.02$ & & $1.30\times10^{-5}$ \\ 
-   $\zeta$ & $1.01$ & & \\ 
-   $mm_{0}$ & $0.32$ & & \\
-
-   \hline
-    \end{tabular}
-    \end{center}
-   \end{table}
 
@@
 
-\noindent {\bf Allometry leads to stability} \\ \noindent
+\noindent {\bf The stabilizing effects of allometric constraints} \\ \noindent
+Although starvation leads to mortality risk for the individual, a moderate amount promotes persistence of both consumer and resource populations.
+%Sigma and rho vs. Hopf and TC
+Analysis of the NSM shows that (non-cyclic) stability of the fixed point generally requires a higher starvation rate $\sigma$ relative to the recovery rate $\rho$.
+The intuition behind this is that transition to the hungry (non-reproductive) state permits the resource to recover and transient dynamics to subside, whereas a low $\sigma$ overloads the system with energetically-replete (reproducing) individuals, thus producing maintained oscillations between consumer and resource (Fig.
+\ref{Hopfb}).
+However if $\sigma$ is too large, mortality due to starvation depletes the consumer population, resulting in a lower steady state density, and an opposingly higher steady state density for the resource.
+
+Whereas the rate of consumer growth defines a hard bound of biological feasibility (the TC bifurcation), the rate of starvation determines the sensitivity of the consumer population to changes in resource density.
+While higher rates of starvation result in lower steady state population size -- increasing the risk of stochastic extinction -- lower rates of starvation result in a system poised near either the TC or Hopf bifurcation (or both), which will lead to extinction of the resource or the development of cyclic oscillations, respectively.
+Which bifurcation is approached is wholly dependent on the rate of recovery: if it is high, then cyclic dynamics will develop; if it is low, resource extinction becomes more likely.
+
+%Bifurcation space as a function of mass
+As the allometric derivations of NSM rate laws reveal, $\sigma$ and $\rho$ are not independent parameters, such that the bifurcation space shown in Fig.
+\ref{Hopfb} cannot be freely navigated if assuming biologically reasonable parameterizations.
+Given the parameterization for mammals shown in Table \ref{xx} with mass $M$ as the only free parameter, rates of starvation and recovery are constrained to a fairly small window of potential values.
+Importantly, for larger organismal masses, the distance between $\sigma(M)$ and $\rho(M)$ compared to both the TC and Hopf bifurcations is increased.
+This suggests that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than large mammals, a general prediction that appears to qualitatively hold.
+
+%Effect noise
+
+%NOTES 7/8/16
+%Calder has a hypothesis that period ~ Mass^(1/4).
+%Can we get "transient period" from Jacobian?
+%Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.
+
+both the location of TC and Hopf bifurcations over $\sigma$ and $\rho$ relative to the allometrically-determined starvation and recovery rates of species across different body sizes 
+%Mass changes the values of the fixed points as well
+
 [First describe the dynamics of the model and HOW IT WORKS]
@@
 %
-\newpage
+\clearpage
+
+\begin{table}[h]
+\caption{Parameter Values For Various Classes of Organisms}
+\label{parameter-values}
+    \begin{center}
+    \small
+     \begin{tabular}{| p{1.2cm}| p{3.2cm} | p{2.6cm} | p{3.2cm} | }
+     \hline
+     & {\bf Mammals} & {\bf Unicellular Eukaryotes} & {\bf Bacteria} \\
+     \hline
+   $\eta$ & $3/4$ & & $1.70$ \\ 
+   $E_{m}$ & $10695$ (J gram$^{-1}$) & & $10695$ (J gram$^{-1}$) \\ 
+   $E_{m}^{\prime}$ & $\approx E_{m}$ & & $\approx E_{m}$ \\ 
+   $B_{0}$ & $0.019$ (W gram$^{-\alpha}$) & & $1.96\times10^{17}$ \\
+   $B_{m}$ & $0.025$ (W gram$^{-1}$)   & & $0.025$ (W gram$^{-1}$)\\
+   $a$ & $1.78\times10^{-6}$ & & $1.83\times10^{13}$ \\ 
+   $b$ & $2.29\times10^{-6}$ & & $2.29\times10^{-6}$ \\  
+   $\eta-1$ & $-0.21$ & & $0.73$ \\ 
+   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) & & $56493$ \\ 
+   $\gamma$ & $1.19$ & & $0.68$ \\ 
+   $f_{0}$ & $0.02$ & & $1.30\times10^{-5}$ \\ 
+   $\zeta$ & $1.01$ & & \\ 
+   $mm_{0}$ & $0.32$ & & \\
+
+   \hline
+    \end{tabular}
+    \end{center}
+   \end{table}
+
+\clearpage
 
@@
  	\centering
- 	\includegraphics[width=1\textwidth]{fig_HopfFP.jpg}
+ 	\includegraphics[width=0.75\textwidth]{fig_FixedPoint.pdf}
  	\caption{
@@
  	\centering
- 	\includegraphics[width=1\textwidth]{fig_DataHopf.pdf}
+ 	\includegraphics[width=0.8\textwidth]{fig_DataHopf.pdf}
  	\caption{
  	}
- 	\label{Hopfb}
+ 	\label{DataHopf}
  \end{figure}
@@
  	\centering
- 	\includegraphics[width=1\textwidth]{fig_ExtinctionAllometric.pdf}
+ 	\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
  	\caption{
  	}
- 	\label{Hopfb}
+ 	\label{Asymp}
+ \end{figure}
+
+  \begin{figure}[h]
+ 	\centering
+ 	\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometric.pdf}
+ 	\caption{
+ 	}
+ 	\label{Ext}
  \end{figure}
@@
 
-\newpage
+\clearpage
 

### 2016-07-09 Manuscript/Yeakel_Kempes_Redner-7-5-2016.tex
manuscript updates

@@
 Importantly, for larger organismal masses, the distance between $\sigma(M)$ and $\rho(M)$ compared to both the TC and Hopf bifurcations is increased.
-This suggests that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than large mammals, a general prediction that appears to qualitatively hold.
+Uncertainty in allometric parameters (20\% variation around the mean; Fig. \ref{xx}) results in little difference in the position of both bifurcations as well as consumer energetic rates.
+This suggests 
+1)that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than large mammals, and
+2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
+Although the prediction of oscillations with smaller body sizes  generally holds, empirical observations of large animal population cycles are plagued by long generation times and the influence of top-down effects [REFS].
 

### 2016-07-13 Manuscript/Yeakel_Kempes_Redner-7-10-2016.tex
new stuff

--- 2016-07-09 Manuscript/Yeakel_Kempes_Redner-7-5-2016.tex
@@
 To what extent, and when, organisms invest in somatic vs.
-reproductive expenditures may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, as well as resource limitation.
+reproductive expenditures may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, the distribution of resources.
 Importantly, the influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
@@
 %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-Chris - Linking paragraph here -- what does allometry provide to the story? Biological, linked constraints... vast amounts of diversity in a small number of parameters. Etc.
+Our model explores a wide variety of possible dynamics yet we have not described the realistic regimes occupied by organisms where the challenge is to constrain the covariation of rates.
+Allometric scaling relationships highlight common constraints and average trends across large ranges in body size and species diversity.
+Many of these relationships can be derived from a small set of assumptions and below we describe a framework for the covariation of timescales and rates across the range of mammals for each of the key parameters of our model.
+We are able to define the regime of dynamics occupied by the entire class of mammals along with the key differences between the largest and smallest mammals.
 
@@
 \begin{equation}
-m\left(t\right)=\left[1-\left(1-\frac{b}{a}m_{0}^{1-\alpha}\right)e^{-b\left(1-\alpha\right)t}\right]^{1/\left(1-\alpha\right)}\left(\frac{a}{b}\right)^{1/\left(1-\alpha\right)}
+m\left(t\right)=\left[1-\left(1-\frac{b}{a}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
 \end{equation}
@@
 \begin{equation}
-t_{\rho}=\frac{\log \left(1-\left(M \left(\frac{a}{b}\right)^{\frac{1}{\alpha -1}}\right)^{1-\alpha }\right)-\log \left(1-\left(M \epsilon  \left(\frac{a}{b}\right)^{\frac{1}{\alpha
-   -1}}\right)^{1-\alpha }\right)}{(\alpha -1) b}.
+t_{\rho}=\frac{\log \left(1-\left(M \left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}\right)^{1-\eta }\right)-\log \left(1-\left(M \epsilon  \left(\frac{a}{b}\right)^{\frac{1}{\eta
+   -1}}\right)^{1-\alpha }\right)}{(\eta -1) b}.
 \end{equation}
@@
  	\centering
+ 	\includegraphics[width=0.75\textwidth]{Growth-trajectory-diagram.pdf}
+ 	\caption{
+ 	}
+ 	\label{growth-diagram}
+ \end{figure}
+
+\clearpage
+
+ \begin{figure}[h]
+ 	\centering
  	\includegraphics[width=0.75\textwidth]{fig_FixedPoint.pdf}

### 2016-07-17 Manuscript/Yeakel_Kempes_Redner-7-10-2016.tex
changes to text

@@
 \ref{Hopfb}).
-However if $\sigma$ is too large, mortality due to starvation depletes the consumer population, resulting in a lower steady state density, and an opposingly higher steady state density for the resource.
+However if $\sigma$ is too large, mortality due to starvation depletes the consumer population, resulting in a lower steady state density for the consumer and an opposingly higher steady state density for the resource.
 
 Whereas the rate of consumer growth defines a hard bound of biological feasibility (the TC bifurcation), the rate of starvation determines the sensitivity of the consumer population to changes in resource density.
-While higher rates of starvation result in lower steady state population size -- increasing the risk of stochastic extinction -- lower rates of starvation result in a system poised near either the TC or Hopf bifurcation (or both), which will lead to extinction of the resource or the development of cyclic oscillations, respectively.
-Which bifurcation is approached is wholly dependent on the rate of recovery: if it is high, then cyclic dynamics will develop; if it is low, resource extinction becomes more likely.
+While higher rates of starvation result in lower steady state population size -- increasing the risk of stochastic extinction -- lower rates of starvation result in a system poised near either the TC or Hopf bifurcation (or both), which will lead to elimination of the resource or the development of cyclic oscillations, respectively.
+Which bifurcation is approached is wholly dependent on the rate of recovery: if it is high, then cyclic dynamics will develop; if it is low, resource extinction becomes increasingly likely.
 
@@
 \ref{Hopfb} cannot be freely navigated if assuming biologically reasonable parameterizations.
-Given the parameterization for mammals shown in Table \ref{xx} with mass $M$ as the only free parameter, rates of starvation and recovery are constrained to a fairly small window of potential values.
-Importantly, for larger organismal masses, the distance between $\sigma(M)$ and $\rho(M)$ compared to both the TC and Hopf bifurcations is increased.
-Uncertainty in allometric parameters (20\% variation around the mean; Fig. \ref{xx}) results in little difference in the position of both bifurcations as well as consumer energetic rates.
-This suggests 
-1)that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than large mammals, and
+Given the parameterization for terrestrial endotherms shown in Table \ref{xx} with mass $M$ as the only free parameter, rates of starvation and recovery are constrained to a fairly small window of potential values.
+We find that the allometric constraints given for terrestrial endothermic organisms confine dynamics to the steady state regime across all reasonable body size classes, which for mammals ranges from ca.
+1 gram (the Etruscan shrew) to ca.
+$10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
+Moreover, for larger organismal mass, the distance increases between the allometric values for $\sigma(M)$ and $\rho(M)$ relative to the Hopf bifurcation, while uncertainty in allometric parameters (20\% variation around the mean; Fig.
+\ref{xx}) results in little difference in the position of the TC and Hopf bifurcations as well as consumer energetic rates.
+This result suggests 
+1) that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than large mammals, and
 2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
-Although the prediction of oscillations with smaller body sizes  generally holds, empirical observations of large animal population cycles are plagued by long generation times and the influence of top-down effects [REFS].
+Although the prediction of larger oscillations for animals with smaller body sizes generally holds, empirical observations of large animal population cycles are plagued by long generation times and the influence of top-down effects [REFS] making direct empirical observation problematic.
 
-%Effect noise
+%Extinction to the left and the right
+Higher rates of starvation results in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of death, however it is the relative size of the rate of starvation to the rate of recovery that determines the long-term dynamics of the system.
+By looking at the ratio $\sigma/\rho$, we can assess the competing effects of cyclic dynamics and decreasing rates of extinction
+
+The ratio $\sigma/\rho$
+
+%Results are insensitive to alpha but possibly sensitive to beta (eff parameter)
 

### 2016-07-17 Manuscript/Yeakel_Kempes_Redner-7-10-2016.tex
some more text

@@
 Higher rates of starvation results in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of death, however it is the relative size of the rate of starvation to the rate of recovery that determines the long-term dynamics of the system.
-By looking at the ratio $\sigma/\rho$, we can assess the competing effects of cyclic dynamics and decreasing rates of extinction
+By examining the ratio $\sigma/\rho$, we can understand the competing effects of cyclic dynamics and population density on the probability of extinction.
+We computed the probability of extinction, which we defined as $H(t)+F(t)=10$ at any instant across all values of $t$, for 1000 replicates of the continuous-time system shown in Eq.
+XX with random starting conditions within the interval $(R^*,H^*,F^*)\pm10\%$ (check) for an organism of $M=100{\rm g}$.
+By allowing the rate of starvation to vary, we assessed extinction risk across a range of the ratio $\sigma/\rho$ varying between XXX and XXX, thus examining a vertical cross-ection of Fig \ref{Hopfb}.
+As we expected, we found higher rates of extinction for both low and high values of $\sigma/\rho$; for low values the higher extinction risk is the consequence of transient cycles with larger amplitudes as the system nears the Hopf bifurcation.
+For high values of $\sigma/\rho$, higher extinction risk is due to the steady decrease in the steady state consumer population density.
+This interplay creates an `extinction refuge', such that for a relatively small range of $\sigma/\rho$ extinction probabilities are minimized.
+
+As has been described, the $\sigma$ vs. $\rho$ space cannot be freely traversed, such that not all values of $\sigma/\rho$ are biologically feasible.
+We observe that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around the expectation) fall squarely within the extinction refugia, such that the are close enough to the Hopf bifurcation to avoid low steady state densities, but far enough away to avoid large-amplitude transient cycles.
+The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window is suggestive of a selective mechanism driving these rates within populations, and suggests that the NSM system provides insight to a general phenomena that may both drive -- and constrain -- natural animal populations.
 

### 2016-07-19 Manuscript/Yeakel_Kempes_Redner-7-10-2016.tex
manuscript edits along with invasion FP changes and analysis of asymptote for 10-8 masses

--- 2016-07-17 Manuscript/Yeakel_Kempes_Redner-7-10-2016.tex
@@
 In a multidimensional system, linear stability is determined with respect to the Jacobian Matrix $\bf J|_*$ (where $|_*$ denotes evaluation at the internal steady state), where each element of the matrix is defined by the partial derivative of each differential equation with respect to each variable.
+%General statement on stability
+If the parameters of $\bf J|_*$ are such that the real part of its leading eigenvalue is $<0$, then the system is stable to small pulse perturbations.
+
 %In the case of the 2-stage consumer model, the Jacobian evaluated at the internal steady state (denoted by $|_*$) is written
@@
 %\end{equation}
-
-%Transcritical bifurcation
-If the parameters of $\bf J|_*$ are such that the real part of its leading eigenvalue is $<0$, then the system is stable to small pulse perturbations. 
-In the case of the NSM, stability is conditioned on the consumer's starvation rate $\sigma$ relative to the its reproduction rate $\lambda$.
-As $\sigma$ becomes less than $\lambda$, the $R^*$ crosses the origin and the fixed point becomes unstable.
-This condition $\sigma = \lambda$  marks a transcritical (TC) bifurcation, thus marking a hard boundary below which the system becomes unphysical due to the unregulated growth of the consumer population.
-That the timescale of reproduction is larger than the timescale of starvation is intuitive for macroscopic organisms, as the rate at which one loses tissue due to lack of resources is generally much faster than reproduction, though we will show using allometric principles that this relationship always holds.
-%such that if a consumer's starvation rate is lower than it's reproductive rate, the consumer population's growth becomes exponential and unphysical.
-%Thus, the timescale of starvation must be less than the timescale of reproduction, an expectation that will hold for most classes of organisms, which we will show by deriving allometric scaling relationship that serve to constrain the covariation of parameter values.
-%[Explain what a transcritical bifurcation means in terms of the dynamics. From a regime where resources are negative and consumer populations become infinite to one that is biological feasable]
-%As such, a transcritical bifurcation exists at $\lambda = \sigma$, such that the existence of an internal stable fixed point is dependent on the condition that $\sigma > \lambda$.
-%This means that the rate of starvation must be greater (operating on a smaller timescale) than the rate of consumer reproduction (operating on a relatively longer timescale).
-%This general expectation will hold for most classes of organisms, which we will show by deriving allometric scaling relationship that serve to constrain the covariation of parameter values.
-% while the exact difference in timescales between reproduction and starvation can be derived using allometric scaling relationships.
-%Kempes (REF) has shown that population growth rates scale to $M^{\alpha-1}$ (where $M$ is the reproductive mass of an organism and $\alpha$ is the scaling exponent), whereas starvation is necessarily bound to shorter timescales, which will be elaborated on below.
-
-%Hopf bifurcation
-In addition to the hard bound defined by the TC bifurcation, oscillating or cyclic dynamics present an implicit constraint to persistence by increasing the risk of extinction.
-%also present certain constraints to the feasibility of populations.
-%Oscillating, or cyclic, dynamics present additional risks to populations.
-If cycles are large, stochastic effects may result in either the consumer or resource population.
-In continuous-time systems, a stable limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts \citep{GuckHolmes}.
-%Of additional interest is the existence of parameter regions that permit the existence of oscillatory, or cyclic, dynamics.
-This condition is called a Hopf bifurcation, and is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian \citep{Gross:2004p2428}.
-In addition to the onset of stable cycles, as a system with a stable fixed point nears the Hopf bifurcation, transient or decaying cycles can grow in magnitude.
-Given that ecological systems exist in a state of constant perturbation \citep{Hastings:2001jh}, even the onset of transient cycles that decay over time can increase the risk of extinction \citep{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}, such that the distance of a system from the Hopf bifurcation, even in the stable region, is relevant to persistence.
-
-%Although the Hopf condition for the specific 2-stage model cannot be easily written, the analytical solution can be explored using a symbolic computational language such as \emph{Mathematica}.
 
@@
 \begin{equation}
+\label{eqsigma}
 t_{\sigma}=-\frac{E_{m}\log\left(\epsilon\right)}{B_{m}}.
@@
 t_{\rho}=\frac{\log \left(1-\left(M \left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}\right)^{1-\eta }\right)-\log \left(1-\left(M \epsilon  \left(\frac{a}{b}\right)^{\frac{1}{\eta
-   -1}}\right)^{1-\alpha }\right)}{(\eta -1) b}.
+   -1}}\right)^{1-\eta }\right)}{(\eta -1) b}.
 \end{equation}
+Although these rate equations are general and will hold for any organism, here we focus on parameterizations specifically for terrestrial-bound endotherms, specifically mammals, which range from $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to $M\approx10^7$ grams (the late Eocene to early Miocene Indricotheriinae).
+Investigating other classes of organisms requires only substituting the energetic and scale parameters shown in Table \ref{param}, though this is beyond the current scope of our analysis, and we henceforth focus our examination on mammalian species.
 
@@
 \noindent {\bf The stabilizing effects of allometric constraints} \\ \noindent
+%Transcritical bifurcation
+Stability in the NSM is conditioned on the consumer's starvation rate $\sigma$ relative to the its reproduction rate $\lambda$.
+As $\sigma$ becomes less than $\lambda$, the resource steady state density becomes negative and extinction is inevitable.
+This condition $\sigma = \lambda$ marks a transcritical (TC) bifurcation, thus marking a hard boundary below which the system becomes unphysical due to the unregulated growth of the consumer population.
+That the timescale of reproduction is larger than the timescale of starvation is intuitive for macroscopic organisms, as the rate at which one loses tissue due to a lack of resources is generally much faster than reproduction.
+In fact, allometric derivations for both reproduction \citep{Kempes:2012hy} and starvation (Eq. \ref{eqsigma}) show that this relationship always holds for organisms within observed body size ranges.
+We note that the asymptote for the starvation rate at $M \approx 8.3\times10^8$ defines the mass at which fat tissue accounts for 100\% of organismal weight, thereby placing a hard scaling boundary on our derivation for the starvation rate.
+%According to this derivation, the endotherm (the blue whale \emph{}) is 74\% fat, which accords with empirical estimates REF.
+
+%
+%%Transcritical & rate law ranges
+%The TC bifurcation occurs in this model because we have assumed that the portion of the population that is not starved reproduces at a constant rate.
+%Because the process of starvation is incorporated explicitly, the consumer's rate of reproduction is not dependent on the density of resources.
+%In fact, the existence of the TC bifurcation at $\lambda = \sigma$ reveals an important biological insight. 
+%Reproduction requires maintenance and growth of biological tissues, both of which have strong scaling relationships with body size.
+%Recent work by Kempes et al. [REF] derived the timescale of reproduction in terms of allometric considerations, where $t_\lambda \propto M^{1-\eta}$ (REF).
+%Starvation is the loss of energy required for maintenance, and we have shown it to have a timescale $t_\sigma \propto \log(M)$.
+%
+%A third important parameter in our framework is the rate of recovery.
+%The recovery timescale $t_\rho$ controls the rate at which individuals move from the hungry class to the full class, and this requires not only tissue maintenance, but growth, such that it is bounded on the short side by $t_\sigma$.
+%Moreover, [why is recovery timescale bounded on the high side], such that it is bounded on the long side by $t_\lambda$.
+%Thus, incorporating allometric considerations shows us that $\lambda < \rho < \sigma$ (alternatively $t_\lambda > t_\rho > t_\sigma$).
+
+%General Hopf bifurcation info
+In addition to the hard bound defined by the TC bifurcation, oscillating or cyclic dynamics present an implicit constraint to persistence by increasing the risk of extinction.
+%also present certain constraints to the feasibility of populations.
+%Oscillating, or cyclic, dynamics present additional risks to populations.
+If cycles are large, stochastic effects may result in extinction in either the consumer or resource population.
+In continuous-time systems, a stable limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts \citep{GuckHolmes}.
+%Of additional interest is the existence of parameter regions that permit the existence of oscillatory, or cyclic, dynamics.
+This condition, known as a Hopf bifurcation, is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian \citep{Gross:2004p2428}.
+
+%%Hopf bifurcation and oscillations sigma vs. rho
+%The NSM exhibits both non-cyclic as well as cyclic dynamics, and which behavior is predominant depends strongly on the rate of starvation relative to the rate of recovery.
+%When the starvation rate is low, cyclic dynamics are more likely to occur for a given value of the recovery rate.
+%This can be understood intuitively: for low starvation rates, resources are depressed by an infusion of full consumers, which subsequently starve thereby allowing the resource to recover and continuing the cycle.
+%When the starvation rate is high, the response of consumer growth to resource abundance is muted, such that oscillations tend to decay over time.
+%Thus, higher starvation rates $\sigma$ dampens changes in the consumer population to changes in resource density, and lower rates of recovery $\rho$ amplifies this effect.
+
+%General transient dynamics statement
+In addition to the onset of stable cycles, as a system with a stable fixed point nears the Hopf bifurcation, transient or decaying cycles can grow in magnitude, despite the existence of a positive, non-cyclic, steady state density.
+Given that ecological systems exist in a state of constant perturbation \citep{Hastings:2001jh}, even the onset of transient cycles that decay over time can increase the risk of extinction \citep{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}, such that the distance of a system from the Hopf bifurcation, even in the stable region, is relevant to persistence.
+
+%Although the Hopf condition for the specific 2-stage model cannot be easily written, the analytical solution can be explored using a symbolic computational language such as \emph{Mathematica}.
+
+The NSM exhibits both non-cyclic as well as cyclic dynamics, and which behavior dominates depends strongly on the rate of starvation $\sigma$ relative to the rate of recovery $\rho$.
 Although starvation leads to mortality risk for the individual, a moderate amount promotes persistence of both consumer and resource populations.
 %Sigma and rho vs. Hopf and TC
-Analysis of the NSM shows that (non-cyclic) stability of the fixed point generally requires a higher starvation rate $\sigma$ relative to the recovery rate $\rho$.
+Non-cyclic stability of the fixed point generally requires a higher starvation rate $\sigma$ relative to the recovery rate $\rho$.
 The intuition behind this is that transition to the hungry (non-reproductive) state permits the resource to recover and transient dynamics to subside, whereas a low $\sigma$ overloads the system with energetically-replete (reproducing) individuals, thus producing maintained oscillations between consumer and resource (Fig.
@@
 Which bifurcation is approached is wholly dependent on the rate of recovery: if it is high, then cyclic dynamics will develop; if it is low, resource extinction becomes increasingly likely.
+
+%
+%
+%
+%%Hopf bifurcation as a function of b
+%Both full and hungry consumers remove resources at rates $b$ and $\rho$, respectively.
+%As the rate of resource consumption by full consumers increases, the Hopf condition changes from a convex to a concave function over $\sigma$, limiting the potential for oscillatory dynamics.
+%These rates are considered separately because full consumers need only to maintain their tissues, whereas hungry consumers require both growth and maintenance, such that $t_\rho > t_b$, or $\rho < b$.
+%If this constraint is enacted, the likelihood of oscillatory dynamics is reduced for a given 
+%
+%
+%
+%If instead the consumer's growth was proportional to resource abundance, such that the effects of starvation on reproduction were incorporated explicitly (where the 2-stage consumer resource model collapses to the Lotka-Volterra consumer-resource model with logistic growth of the resource), the TC bifurcation exists only for $\lambda = \mu$, such that the rate of mortality cannot exceed the intrinsic birth rate.
+%
+%
+% whereas the traditional Lotka-Volterra dynamic assumes that the reproductive rate of the consumer is scaled to resource density, such that the growth function would be $G(F,R) = \lambda R F$.
+%Thus, the Lotka-Volterra dynamic \emph{implicitly} accounts for starvation in reducing the reproductive rate of the consumer.
+%However, our 2-stage model \emph{explicitly} accounts for starvation as well as recovery, such that individuals who are not starved should adopt a reproductive rate independent of resource density.
+%
+%
+%
+%We have used scaling relationships between tissue turnover and growth to strictly constrain 5/6 population-level parameters in our 2-stage consumer resource model (including the mortality rate $t_\mu$, which we have shown is just a xxx of $t_\sigma$).
+%This exercise accomplishes two goals:
+%1) it allows us to constrain the plausible parameter space of the two-stage model, and 
+%2) 
+%
+%This allows us to derive many aspects of the system in terms of consumer body mass $M$ and the allometric scaling exponent $\eta$.
 
@@
 %Extinction to the left and the right
-Higher rates of starvation results in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of death, however it is the relative size of the rate of starvation to the rate of recovery that determines the long-term dynamics of the system.
-By looking at the ratio $\sigma/\rho$, we can assess the competing effects of cyclic dynamics and decreasing rates of extinction
+Higher rates of starvation results in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of death, however it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{Hopfb}).
+By examining the ratio $\sigma/\rho$, we can understand the competing effects of cyclic dynamics vs. population density on the probability of extinction.
+We computed the probability of extinction, which we defined as $H(t)+F(t)=10$ at any instant across all values of $t>100$ for 1000 replicates of the continuous-time system shown in Eq.
+XX with random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$ (check) for an organism of $M=100{\rm g}$.
+By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for the ratio $\sigma/\rho$ varying between $10^-2$ to $2.5$, thus examining a vertical cross-ection of Fig.
+\ref{Hopfb}.
+As we expected, we found higher rates of extinction for both low and high values of $\sigma/\rho$; for low values the higher extinction risk is the consequence of transient cycles with larger amplitudes as the system nears the Hopf bifurcation.
+For high values of $\sigma/\rho$, higher extinction risk is due to the steady decrease in the steady state consumer population density.
+This interplay creates an `extinction refuge', such that for a relatively small range of $\sigma/\rho$, extinction probabilities are minimized.
 
-The ratio $\sigma/\rho$
+As has been described, the $\sigma$ vs. $\rho$ space cannot be freely traversed, such that not all values of $\sigma/\rho$ are biologically feasible.
+We observe that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around the mean for the energetic parameters used to determine each rate) fall within the extinction refugia, such that they are close enough to the Hopf bifurcation to avoid low steady state densities, though far enough away to avoid large-amplitude transient cycles.
+The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving observed starvation and recovery rates within populations.
+Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
+Despite the possible mechanisms, our finding that allometrically-determined energetic rates place the system within this low probability of extinction region suggests that the NSM system provides general insight to a phenomena that may both drive -- and constrain -- natural animal populations.
 
@@
 %Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.
-
-both the location of TC and Hopf bifurcations over $\sigma$ and $\rho$ relative to the allometrically-determined starvation and recovery rates of species across different body sizes 
-%Mass changes the values of the fixed points as well
-
-[First describe the dynamics of the model and HOW IT WORKS]
-[Then Allometry and stability]
-
-Analysis of the 2-stage consumer resource model shows that the equilibrial states of both populations are highly sensitive to changes in  starvation and recovery rates of the consumer.
-The consumer and resource population densities vary inversely: when the consumer densities are high, resource densities are low, and vice versa.
-High starvation and low recovery rates result in low consumer densities and high resource densities. %whereas the opposite is true for low starvation and high recovery rates.
-If starvation rates are low, resources have a fixed point near zero for any value of the recovery rate.
-Full and hungry consumer stages tend have fixed points that are tightly correlated, the extent to which is driven by the similarity of consumer growth and mortality rates; if $\lambda = \mu$, then $F^* = H^*$.
-
-%Transcritical & rate law ranges
-A transcritical bifurcation exists at $\lambda = \sigma$, such that the condition $\sigma > \lambda$ is required for biologically reasonable dynamics.
-The TC bifurcation occurs in this model because we have assumed that the portion of the population that is not starved reproduces at a constant rate.
-Because the process of starvation is incorporated explicitly, the consumer's rate of reproduction is not dependent on the density of resources.
-In fact, the existence of the TC bifurcation at $\lambda = \sigma$ reveals an important biological insight. 
-Reproduction requires maintenance and growth of biological tissues, both of which have strong scaling relationships with body size.
-Recent work by Kempes et al. [REF] derived the timescale of reproduction in terms of allometric considerations, where $t_\lambda \propto M^{1-\eta}$ (REF).
-Starvation is the loss of energy required for maintenance, and we have shown it to have a timescale $t_\sigma \propto \log(M)$.
-Accordingly, the timescale of reproduction is always larger than the timescale of starvation, such that $\lambda$ must be less than $\sigma$ by definition.
-A third important parameter in our framework is the rate of recovery.
-The recovery timescale $t_\rho$ controls the rate at which individuals move from the hungry class to the full class, and this requires not only tissue maintenance, but growth, such that it is bounded on the short side by $t_\sigma$.
-Moreover, [why is recovery timescale bounded on the high side], such that it is bounded on the long side by $t_\lambda$.
-Thus, incorporating allometric considerations shows us that $\lambda < \rho < \sigma$ (alternatively $t_\lambda > t_\rho > t_\sigma$).
-
-%Hopf bifurcation and oscillations sigma vs. rho
-The 2-stage consumer resource model exhibits two qualitatively different behavioral regimes.
-Because portions of the consumer population exist in either full or hungry states simultaneously, the internal fixed point can either be a stable equilibrium, or exhibit sustained oscillatory behavior, depending on the rates of starvation $\sigma$ and recovery $\rho$.
-The transition from stable non-oscillatory dynamics to oscillatory dynamics occurs at the Hopf bifurcation condition where two complex conjugate eigenvalues cross the imaginary axis and attain positive real parts.
-Although there is an analytical solution to the Hopf bifurcation condition, it cannot be written efficiently.
-
-When the starvation rate is low, oscillatory dynamics are more likely to occur for a given value of the recovery rate.
-This can be understood intuitively: for low starvation rates, resources are depressed by an infusion of full consumers, which subsequently starve thereby allowing the resource to recover and continuing the cycle.
-When the starvation rate is high, the response of consumer growth to resource abundance is muted, such that oscillations tend to decay over time.
-Thus, higher starvation rates $\sigma$ desensitizes changes in the consumer population to changes in resource density, and lower rates of recovery $\rho$ amplifies this effect.
-
-%Hopf bifurcation as a function of b
-Both full and hungry consumers remove resources at rates $b$ and $\rho$, respectively.
-As the rate of resource consumption by full consumers increases, the Hopf condition changes from a convex to a concave function over $\sigma$, limiting the potential for oscillatory dynamics.
-These rates are considered separately because full consumers need only to maintain their tissues, whereas hungry consumers require both growth and maintenance, such that $t_\rho > t_b$, or $\rho < b$.
-If this constraint is enacted, the likelihood of oscillatory dynamics is reduced for a given
-
-If instead the consumer's growth was proportional to resource abundance, such that the effects of starvation on reproduction were incorporated explicitly (where the 2-stage consumer resource model collapses to the Lotka-Volterra consumer-resource model with logistic growth of the resource), the TC bifurcation exists only for $\lambda = \mu$, such that the rate of mortality cannot exceed the intrinsic birth rate.
-
- whereas the traditional Lotka-Volterra dynamic assumes that the reproductive rate of the consumer is scaled to resource density, such that the growth function would be $G(F,R) = \lambda R F$.
-Thus, the Lotka-Volterra dynamic \emph{implicitly} accounts for starvation in reducing the reproductive rate of the consumer.
-However, our 2-stage model \emph{explicitly} accounts for starvation as well as recovery, such that individuals who are not starved should adopt a reproductive rate independent of resource density.
-
-We have used scaling relationships between tissue turnover and growth to strictly constrain 5/6 population-level parameters in our 2-stage consumer resource model (including the mortality rate $t_\mu$, which we have shown is just a xxx of $t_\sigma$).
-This exercise accomplishes two goals:
-1) it allows us to constrain the plausible parameter space of the two-stage model, and 
-2)
-
-This allows us to derive many aspects of the system in terms of consumer body mass $M$ and the allometric scaling exponent $\eta$.
 
@@
 \caption{Parameter Values For Various Classes of Organisms}
-\label{parameter-values}
+\label{param}
     \begin{center}
@@
 		 \centering
- 	\includegraphics[width=0.75\textwidth]{Growth-trajectory-diagram.pdf}
+		 \includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
 		 \caption{
@@
 
-\clearpage
+	 \begin{figure}[h]
+		 \centering
+		 \includegraphics[width=0.5\textwidth]{fig_GvS.pdf}
+		 \caption{
+		 }
+		 \label{GvS}
+	 \end{figure}
 
@@
 		\centering
+	 	\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometric.pdf}
+	 	\caption{
+	 	}
+	 	\label{Ext}
+	\end{figure}
+
+	\begin{figure}[h]
+	 	\centering
 	 	\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
@@
 	 	\label{Asymp}
- \end{figure}
-
-  \begin{figure}[h]
- 	\centering
- 	\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometric.pdf}
- 	\caption{
- 	}
- 	\label{Ext}
 	\end{figure}

### 2016-07-19 Manuscript/Yeakel_Kempes-Redner_PNAS_7-19.tex
PNAS styling and significant text modifications

--- 2016-07-19 Manuscript/Yeakel_Kempes_Redner-7-10-2016.tex
@@
-\documentclass[12pt,english]{article}
-%\usepackage[margin=1in,footskip=0.25in]{geometry}
+%\usepackage{scicite}
 
-%\usepackage{helvet}
-%\renewcommand{\familydefault}{\sfdefault}
-\renewcommand\refname{\vskip -1cm}
-%\renewcommand{\rmdefault}{phv} % Arial
-%\renewcommand{\sfdefault}{phv} % Arial
-\usepackage{setspace}
-\usepackage{wrapfig}
-\usepackage{amsmath}
-\usepackage{amssymb}
+%\usepackage{times}
+%\usepackage{graphicx}
+%\usepackage{verbatim}
+%\usepackage{subfigure}
+%\usepackage{tikz}
+%\usepackage{amsmath}
+%\usepackage{amssymb}
+%\usepackage{array}
+%\usepackage{tabularx}
+%\usepackage{multirow}
+%\usepackage{multicol}
+%\usepackage{multibox}
+%\usepackage{rotating}
+%\usepackage{color}
+%\usepackage{layout}
+%\usepackage{epstopdf}
+%\usepackage{afterpage}
+%\usepackage[left]{lineno}
+
+%% PNAStmpl.tex
+%% Template file to use for PNAS articles prepared in LaTeX
+%% Version: Apr 14, 2008
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% BASIC CLASS FILE
+
+\documentclass{pnastwo}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% OPTIONAL GRAPHICS STYLE FILE
+
+\usepackage[pdftex]{graphicx}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% OPTIONAL POSTSCRIPT FONT FILES
+
+\usepackage{pnastwof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% ADDITIONAL OPTIONAL STYLE FILES
+
+\usepackage{amssymb,amsfonts,amsmath}
+\usepackage[font=large]{caption}
+\usepackage{subfigure}
+\usepackage{epstopdf}
 \usepackage{graphicx}
-\usepackage{mathrsfs}
-\usepackage{bm}
-\usepackage{wasysym}
-\usepackage{placeins}
+\usepackage{verbatim}
+\usepackage{array}
+\usepackage{tabularx}
 \usepackage{multirow}
-\usepackage[T1]{fontenc}
-\usepackage[authoryear]{natbib}
+%\usepackage{multibox}
+\usepackage{rotating}
+%\usepackage{color}
+%\usepackage{lineno}
+\usepackage[left]{lineno}
 
-%\usepackage[style=authoryear,sorting=ynt]{biblatex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% OPTIONAL MACRO FILES
 
-\usepackage{framed}
-\usepackage{caption}
-\usepackage{longtable}
-\usepackage{geometry}
-\usepackage{lineno}
+\bibliographystyle{pnas}
 
-\geometry{verbose,letterpaper,tmargin=2.54cm,bmargin=2.54cm,lmargin=2.54cm,rmargin=2.54cm}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Don't type in anything in the following section:
+%%%%%%%%%%%%
+%% For PNAS Only:
+\contributor{Submitted to Proceedings
+of the National Academy of Sciences of the United States of America}
+\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
+\copyrightyear{2011}
+\issuedate{Issue Date}
+\volume{Volume}
+\issuenumber{Issue Number}
+%%%%%%%%%%%%
 
 \begin{document}
-\begin{spacing}{1.9}
+
+\linenumbers
+\setlength\linenumbersep{3pt}
 
 \title{The effect of starvation on the dynamics of consumer populations}
-\author{Yeakel, Kempes, \& Redner}
+
+\author
+{Justin D.
+Yeakel\affil{1}{School of Natural Science, University of California Merced, Merced, CA} \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{To whom correspondence should be addressed: jdyeakel@gmail.com}, Christopher P.
+Kempes \affil{2}{The Santa Fe Institute, Santa Fe, NM}, \and Sidney Redner \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{3}{Department of Physics, Boston University, Boston MA}
+}
+
+\contributor{Submitted to Proceedings of the National Academy of Sciences
+of the United States of America}
+
 \maketitle
 
-\linenumbers
-%\begin{flushleft}
+\begin{article}
 
-%Focus on the tradeoff between REPRODUCTION AND SURVIVAL AS A FUNCTION OF ENERGETIC STATE
+\begin{abstract}
+This is the abstract.
+\end{abstract}
 
-The behavioral ecology of most, if not all, organisms is influenced by the energetic state of individuals.
-An individual's energetic state directly influences how it invests its stores in an uncertain environment.
-Such behaviors are generally manifested as trade-offs, which often concern investing in individual maintenance and growth (somatic effort) or allocating energy towards reproduction (reproductive effort) \citep{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
+%\section*{Introduction}
+The behavioral ecology of most, if not all, organisms is influenced by the energetic state of individuals, which directly influences how an organism invests its reserves in an uncertain environment.
+Such behaviors are generally manifested as trade-offs between investing in individual maintenance and growth (somatic effort) or allocating energy towards reproduction (reproductive effort) \cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
 %among a host of other behavioral duties [REFS].
-The timing of these behaviors is often important and is under strong selective pressure, as they tend to have large effects on the future fitness of the organism \citep{Mangel:1988uaa}.
+The timing of these behaviors is often important and is under strong selective pressure, as they tend to have large effects on the future fitness of the organism \cite{Mangel:1988uaa}.
 %For example, rotifers that reproduce have lower survival rates than those that don't when they are under nutritional stress.
 To what extent, and when, organisms invest in somatic vs.
-reproductive expenditures may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, the distribution of resources.
+reproductive expenditures may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, and the distribution of resources.
 Importantly, the influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
@@
 Maximizing fitness between growth and maintenance activities vs.
-reproductive efforts in large part structures the life-history of species, and this can be achieved by alternative behavioral strategies or via physiological switches, both of which are sensitive to resource availability.
+reproductive efforts in large part structures the life-history of species, and this can be achieved by alternative behavioral strategies or via changes in physiology, both of which are sensitive to resource availability.
 %Reproduction can be separated from somatic growth/maintenance by a variety of potential mechanisms:
 %Multiple mechanisms contribute to the differentiation of these behaviors:
-Behavioral changes in somatic or reproductive investment occur in response to limited resources \citep{Morris:1987eo}.
-%, while stress can generally lead to reprodusuppression of reproductive behaviors \citep{Morgan:1999do}. 
-For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \citep{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for a breeding season \citep{Martin:1987dl,Barboza:2002in}.
-Such breeding behaviors is generally referred to as \emph{capital breeding}
+Behavioral changes in somatic or reproductive investment often occur in response to limited resources \cite{Morris:1987eo}.
+%, while stress can generally lead to reprodusuppression of reproductive behaviors \cite{Morgan:1999do}. 
+For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \cite{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for a breeding season (capital breeding) \cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
+%Such breeding behaviors is generally referred to as \emph{capital breeding}
 %Resource limitation can also alter the behaviors of species in well-mixed environments: 
-Freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress \citep{Threlkeld:1976ih}, with those that do reproduce evincing lower survival rates \citep{Kirk:1997cc}, while artificially induced stress has been observed to decrease reproductive success in Atlantic cod \citep{Morgan:1999do}.
-Organisms may also separate maintenance and growth from reproduction over space and time. 
-Many salmonids, birds, and some mammals return to migratory breeding sites to reproduce after one or multiple seasons in alternative environments spent accumulating body mass and nutritional reserves \citep{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress \cite{Threlkeld:1976ih}, with those that do reproduce evincing lower survival rates \cite{Kirk:1997cc}, while artificially induced stress has been observed to decrease reproductive success in Atlantic cod \cite{Morgan:1999do}.
+Organisms may also separate maintenance and growth from reproduction over space and time: 
+many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in alternative environments spent accumulating body mass and nutritional reserves \cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
 
-Physiological mechanisms also play an important role in regulating reproductive expenditures during periods of high-risk.
-Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation) to time with accumulation of nutritional reserves \citep{Mead:1989dt,Sandell:1990kw}.
-%, including multiple species of Grizzly bears (\emph{Ursus horribilis}) and the European badger (\emph{Meles melesc}) \citep{Woodroffe:1995id
-Furthermore, many mammals (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress \citep{Bulik:1999eo,Trites:2003cc}.
-In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the individual regulates all aspects of the cell cycle \citep{Glazier:2009hq}.
-%great tits invest more energy into broods when resource abundance is greater, while food availability is generally seen to correlate with clutch size among many bird species \citep{Daan:1988va}.
+Physiological mechanisms also play an important role in regulating reproductive expenditures during periods of resource limitation.
+Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation) to time with accumulation of nutritional reserves \cite{Mead:1989dt,Sandell:1990kw}.
+%, including multiple species of Grizzly bears (\emph{Ursus horribilis}) and the European badger (\emph{Meles melesc}) \cite{Woodroffe:1995id
+Furthermore, many mammals (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress \cite{Bulik:1999eo,Trites:2003cc}.
+In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.
+%great tits invest more energy into broods when resource abundance is greater, while food availability is generally seen to correlate with clutch size among many bird species \cite{Daan:1988va}.
 %The mechanisms that control reproductive effort as a function of energetic reserves differ among species.
@@
 
-%To population-level effects
-The mechanisms by which different organisms avoid or delay reproduction during times of nutritional stress has received tremendous empirical and theoretical attention owing to the importance of these activities in shaping life-history \citep{Cody:1966fk,Martin:1987dl,Mangel:1988uaa}.
-Less well understood is how resource limitation and these behavioral/physiological tradeoffs affect dynamics at the level of the population.
-% when organisms are able to avoid the risk of spending energy on reproduction when starvation is likely.
-Traditional Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction \citep{murdoch:2003}.
-Although this implicit dependence connects resource limitation to lower consumer growth rates, the following biological realities are not included: 
-\emph{i}) some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation;
-\emph{ii}) the portion of the population that is not nutritionally stressed is expected to reproduce at a near-constant rate and this is -- averaged across species -- determined strongly by body size \citep{Kempes:2012hy};
-\emph{iii}) the rates that individuals transition from nutritionally poor to replete states and back have different metabolically-constrained timescales that can lead to reproductive lags.
-Importantly, the exclusion of these biological details may have important dynamic shortcomings, masking the effects of starvation on consumer population dynamics.
+%%To population-level effects
+%The mechanisms by which different organisms avoid or delay reproduction during times of nutritional stress has received tremendous empirical and theoretical attention owing to the importance of these activities in shaping life-history \cite{Cody:1966fk,Martin:1987dl,Mangel:1988uaa}.
+%Less well understood is how resource limitation and these behavioral/physiological tradeoffs affect dynamics at the level of the population.
+%% when organisms are able to avoid the risk of spending energy on reproduction when starvation is likely.
+%The influence of energetic constraints -- and their consequent effects on reproduction -- on population dynamics is not well understood, despite considerable effort in 
+%Traditional Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction \cite{murdoch:2003}.
+%Although this implicit dependence connects resource limitation to lower consumer growth rates, the following biological realities are not included: 
+%\emph{i}) some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation;
+%\emph{ii}) the portion of the population that is not nutritionally stressed is expected to reproduce at a near-constant rate and this is -- averaged across species -- determined strongly by body size \cite{Kempes:2012hy};
+%\emph{iii}) the rates that individuals transition from nutritionally poor to replete states and back have different metabolically-constrained timescales that can lead to reproductive lags.
+%Importantly, the exclusion of these biological details may have important dynamic shortcomings, masking the effects of starvation on consumer population dynamics.
 
+Though straightforward conceptually, incorporating the energetic dynamics of individuals \cite{Kooi2000} into a population-level framework \cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles (in particular a lack of smoothness or differentiability \cite{Diekmann:2010da}), and often suffers from over-fitting due to an overabundance of parameters.
+%These issues have limited the development of theoretical models that may aid our understanding of the effects of such tradeoffs on population dynamics.
+An alternative approach involves modeling the macroscale relationships that guide somatic vs.
+reproductive investment in a consumer-resource system.
+%to individual-to-population frameworks is redirecting focus to
+Macroscale Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction \cite{murdoch:2003}.
 Resource limitation and the subsequent effects of starvation may be alternatively accounted for \emph{explicitly}, such that reproduction is permitted only for individuals with sufficient energetic reserves.
-Though straightforward conceptually, incorporating the energetic dynamics of individuals \citep{Kooi2000} into a population-level framework \citep{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles \citep[in particular a lack of smoothness, or differentiability;][]{Diekmann:2010da}, and often suffers from over-fitting due to an overabundance of parameters.
-These issue have limited the development of theoretical models that may aid our understanding of the effects of such tradeoffs on population dynamics.
-An alternative approach to individual-to-population frameworks is redirecting focus to macroscale relationships guiding somatic vs. reproductive investment in a consumer-resource system.
+This permits us to incorporate 
+\emph{i}) the reproductive time lag that is sensitive to changing rates of starvation and recovery, and 
+\emph{ii}) the idea that reproduction is strongly allometrically constrained \cite{Kempes:2012hy}, and is not generally a linear function of resource density.
+%Incorporating explicit starvation introduces
+%\emph{i}) some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation;
+%\emph{ii}) the portion of the population that is not nutritionally stressed is expected to reproduce at a near-constant rate and this is -- averaged across species -- determined strongly by body size \cite{Kempes:2012hy};
+%\emph{iii}) the rates that individuals transition from nutritionally poor to replete states and back have different metabolically-constrained timescales that can lead to reproductive lags.
 
@@
 
-%What we did...
-Here we explore how the energetic tradeoff between maintaining and building somatic tissue vs.
-reproduction can influence the dynamics of populations, and how such dynamics may be determined by allometric constraints.
-We begin by establishing a simple Nutritional State-structured population Model (NSM), where consumer starvation, and cessation of reproduction, is the consequence of resource limitation.
-Similarly, recovery from the starved to a reproductive state increases with resource density.
-%that captures the essential dynamics of energetic vs. reproductive tradeoffs
-%and explore the impact of different fluxes on system-level stability.
-Importantly, the rate at which consumers decline and recover from a starved state is wholly constrained by metabolism.
-By relating different rate constants to allometric constraints, we uncover important relationships between the timescales of physiological and reproductive processes, and show how organisms of different body sizes and taxonomic affinities may be prone to alternative dynamics.
-
-%By deriving allometric constraints for the starvation, recovery, and reproductive timescales of the consumer-resource system, 
-We show that rates of starvation and recovery tend to result in systems with stable, non-cyclic fixed points and that the ratio between the rate of starvation and recovery may -- in part -- contribute to lower size limitations.
-Moreover, larger consumer body size results in rates that are less prone to cyclic dynamics than are rates for smaller organisms, pointing to potential empirical verification of the NSM framework.
-Finally, we show that rates of starvation and recovery appear to be constrained to a parameter range where both transient and equilibrial population dynamics result in the lowest risk of extinction for the consumer.
-This surprising result suggests that the risks associated with different fluxes of consumers in and out of a starved (non-reproductive) state may serve as an important selective driver over evolutionary time.
+%CONCLUSIONS
+%We show that rates of starvation and recovery tend to result in systems with stable, non-cyclic fixed points and that the ratio between the rate of starvation and recovery may -- in part -- contribute to lower size limitations.
+%Moreover, larger consumer body size results in rates that are less prone to cyclic dynamics than are rates for smaller organisms, pointing to potential empirical verification of the NSM framework.
+%Finally, we show that rates of starvation and recovery appear to be constrained to a parameter range where both transient and equilibrial population dynamics result in the lowest risk of extinction for the consumer.
+%This surprising result suggests that the risks associated with different fluxes of consumers in and out of a starved (non-reproductive) state may serve as an important selective driver over evolutionary time.
 
 \noindent {\bf Starvation dynamics} \\ \nonumber
-We integrate the somatic/reproductive tradeoff into the dynamics of a consumer-resource system by dividing the consumer population into discrete energetic states, the occupation of each being contingent on the consumption of a single resource $R$.
-In the NSM there are only two energetic states for the consumer population: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
+%What we did...
+We explore how the energetic tradeoff between maintaining and building somatic tissue vs. reproduction can influence the dynamics of populations, and how such dynamics may be constrained by allometry.
+We begin by establishing a minimal Nutritional State-structured population Model (NSM), where the consumer population is divided into two energetic states: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
+%consumer starvation and recovery is contingent on resource density, the starved portion of the population does not reproduce, and the rate at which consumers decline and recover from starvation are functions of metabolic rate.
+%By relating different rate constants to allometry, we uncover important relationships between the timescales of physiological and reproductive processes, and show how organisms of different body sizes may be prone to alternative dynamic regimes.
+%We integrate the somatic/reproductive tradeoff into the dynamics of a consumer-resource system by dividing the consumer population into two discrete energetic states, the occupation of each being contingent on the consumption of a single resource $R$.
+%In the NSM there are only two energetic states for the consumer population: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
+The resource $R$ has logistic growth with a linear growth rate $\alpha$ and carrying capacity of unity. %(throughout we set $\kappa=1$ such that $R$ has units of $1/\kappa$).
 Consumers transition from state $F$ to state $H$ by starvation at rate $\sigma$ and in proportion to the lack of resources $(1-R)$.
 Conversely, consumers recover from state $H$ to the full state $F$ at rate $\rho$ and in proportion to $R$.
-The resource has logistic growth with a linear growth rate $\alpha$ and a carrying capacity of unity.
 Resources are eliminated by the consumer in both states: by energetically deficient consumers at rate $\rho$, and by energetically replete consumers at rate $\beta$.
@@
 
-\begin{align}
-\begin{split}
+\begin{equation} \label{eq_specific}
+\begin{aligned}
 \dot{F} &= \lambda F + \rho RH - \sigma (1-R)F,  \\
@@
 \dot{R} &= \alpha R(1-R) - R(\rho H+ \beta F).
-\label{eq_specific}
-\end{split}
-\end{align}
+\end{aligned}
+\end{equation}
 
@@
 
-\begin{align}
-\begin{split}
+\begin{equation}
+\begin{aligned}
 F^* &= \frac{\alpha  \lambda  \mu  (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  \beta)}, \\
@@
 R^* &= \frac{\mu  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma }.
-\end{split}
-\end{align}
+\end{aligned}
+\end{equation}
 
@@
 
-\noindent {\bf Allometric constraints} \\ \nonumber
+\vspace{2mm}
+\noindent {\bf Allometric relationships} \\ \nonumber
 %[Link Allometry stuff to our model - what does it provide to the story]
@@
 Nearly all of the rates described in the NSM are to some extent governed by consumer metabolism, and thus can be estimated based on known allometric constraints. 
-The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \citep{West:2002it} and plays a central role in a variety of scaling relationships.
-Organismal metabolic rate $B$ is known to scale as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to be $3/4$ for metazoans, and varies in unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \citep{DeLong:2010dy}.
-Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \citep{Kempes:2012hy}.
+The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \cite{West:2002it} and plays a central role in a variety of scaling relationships.
+Organismal metabolic rate $B$ is known to scale as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to be $3/4$ for metazoans, and varies in unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \cite{DeLong:2010dy}.
+Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \cite{Kempes:2012hy}.
 More specifically, the interspecific trends in growth rate can be approximated by $\lambda = \lambda_0 M^{\eta-1}$.
@@
 Taken together the time scale for starvation is given by
-\begin{equation}
-\label{eqsigma}
+\begin{equation}\label{eq:sigma}
 t_{\sigma}=-\frac{E_{m}\log\left(\epsilon\right)}{B_{m}}.
@@
 \begin{equation}
-m\left(t\right)=\left[1-\left(1-\frac{b}{a}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
+m\left(t\right)=c\left[1-\left(1-\frac{b}{a}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
 \end{equation}
-with $a=B_{0}/E_{m}$ and $b=B_{m}/E_{m}$.
+with $a=B_{0}/E_{m}$, $b=B_{m}/E_{m}$, and $c=(a/b)^{1/(\eta-1)}$.
 We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, which has the final form of
 \begin{equation}
-t_{\rho}=\frac{\log \left(1-\left(M \left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}\right)^{1-\eta }\right)-\log \left(1-\left(M \epsilon  \left(\frac{a}{b}\right)^{\frac{1}{\eta
-   -1}}\right)^{1-\eta }\right)}{(\eta -1) b}.
+t_{\rho}=\frac{\log \left(1-\left(cM \right)^{1-\eta }\right)-\log \left(1-\left(c\epsilon M \right)^{1-\eta }\right)}{(\eta -1) b}.
 \end{equation}
-Although these rate equations are general and will hold for any organism, here we focus on parameterizations specifically for terrestrial-bound endotherms, specifically mammals, which range from $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to $M\approx10^7$ grams (the late Eocene to early Miocene Indricotheriinae).
-Investigating other classes of organisms requires only substituting the energetic and scale parameters shown in Table \ref{param}, though this is beyond the current scope of our analysis, and we henceforth focus our examination on mammalian species.
+%\left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}
+%where $c=(a/b)^{1/(\eta-1)}$.
+Although these rate equations are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to $M\approx10^7$ grams (the late Eocene to early Miocene Indricotheriinae).
+Investigating other classes of organisms requires only substituting the energetic and scale parameters shown in Table \ref{param}.
+Moreover, we emphasize that our allometric equations describe the mean allometric , and do not account for the (sometimes considerable) variance associated with individual species. 
+%, though this is beyond the current scope of our analysis. %, and we henceforth focus our examination on mammalian species.
 
@@
 
+\vspace{2mm}
 \noindent {\bf The stabilizing effects of allometric constraints} \\ \noindent
 %Transcritical bifurcation
-Stability in the NSM is conditioned on the consumer's starvation rate $\sigma$ relative to the its reproduction rate $\lambda$.
-As $\sigma$ becomes less than $\lambda$, the resource steady state density becomes negative and extinction is inevitable.
-This condition $\sigma = \lambda$ marks a transcritical (TC) bifurcation, thus marking a hard boundary below which the system becomes unphysical due to the unregulated growth of the consumer population.
+Stability in the NSM is conditioned on the consumer's starvation rate $\sigma$ relative to its reproduction rate $\lambda$.
+If $\sigma<\lambda$, the resource steady state density is negative and extinction is inevitable.
+The condition $\sigma = \lambda$ is a transcritical (TC) bifurcation, thus marking a hard boundary below which the system is unphysical due to the unregulated growth of the consumer population.
 That the timescale of reproduction is larger than the timescale of starvation is intuitive for macroscopic organisms, as the rate at which one loses tissue due to a lack of resources is generally much faster than reproduction.
-In fact, allometric derivations for both reproduction \citep{Kempes:2012hy} and starvation (Eq. \ref{eqsigma}) show that this relationship always holds for organisms within observed body size ranges.
-We note that the asymptote for the starvation rate at $M \approx 8.3\times10^8$ defines the mass at which fat tissue accounts for 100\% of organismal weight, thereby placing a hard scaling boundary on our derivation for the starvation rate.
+In fact, allometric derivations for both reproduction \cite{Kempes:2012hy} and starvation (Eq.
+\ref{eq:sigma}) show that this relationship always holds for organisms within observed body size ranges (Fig.
+\ref{fig:gvs}).
+We note that the asymptote for the starvation rate at $M \approx 6.16\times10^7$ defines the mass at which fat and muscle tissue accounts for 100\% of organismal weight, thereby placing a hard scaling boundary on our derivation for the starvation rate.
 %According to this derivation, the endotherm (the blue whale \emph{}) is 74\% fat, which accords with empirical estimates REF.
@@
 If cycles are large, stochastic effects may result in extinction in either the consumer or resource population.
-In continuous-time systems, a stable limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts \citep{GuckHolmes}.
+In continuous-time systems, a stable limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts \cite{GuckHolmes}.
 %Of additional interest is the existence of parameter regions that permit the existence of oscillatory, or cyclic, dynamics.
-This condition, known as a Hopf bifurcation, is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian \citep{Gross:2004p2428}.
+This condition, known as a Hopf bifurcation, is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian \cite{Gross:2004p2428}.
+%General transient dynamics statement
+Moreover, as a non-cyclic stable system nears the Hopf bifurcation, transient or decaying cycles can grow in magnitude, despite the existence of a positive, non-cyclic, steady state density.
+Given that ecological systems exist in a state of constant perturbation \cite{Hastings:2001jh}, even the onset of transient cycles that decay over time can increase the risk of extinction \cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}, such that the distance of a system from the Hopf bifurcation, even in the stable region, is relevant to persistence.
 
@@
 %Thus, higher starvation rates $\sigma$ dampens changes in the consumer population to changes in resource density, and lower rates of recovery $\rho$ amplifies this effect.
-
-%General transient dynamics statement
-In addition to the onset of stable cycles, as a system with a stable fixed point nears the Hopf bifurcation, transient or decaying cycles can grow in magnitude, despite the existence of a positive, non-cyclic, steady state density.
-Given that ecological systems exist in a state of constant perturbation \citep{Hastings:2001jh}, even the onset of transient cycles that decay over time can increase the risk of extinction \citep{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}, such that the distance of a system from the Hopf bifurcation, even in the stable region, is relevant to persistence.
 
@@
 By examining the ratio $\sigma/\rho$, we can understand the competing effects of cyclic dynamics vs. population density on the probability of extinction.
-We computed the probability of extinction, which we defined as $H(t)+F(t)=10$ at any instant across all values of $t>100$ for 1000 replicates of the continuous-time system shown in Eq.
-XX with random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$ (check) for an organism of $M=100{\rm g}$.
+We computed the probability of extinction, which we defined as $H(t)+F(t)=10$ at any instant across all values of $10^3<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq.
+XX for an organism of $M=100{\rm g}$, assuming random initial conditions around the steady state solution.
+%random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
 By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for the ratio $\sigma/\rho$ varying between $10^-2$ to $2.5$, thus examining a vertical cross-ection of Fig.
@@
 
+\vspace{2mm}
 \noindent {\bf Minimizing extinction risk} \\ \nonumber
@@
 %The reproductive tradeoff as a strategy
+%
+
+%
+%
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
+% 	\caption{
+% 	Probability that the resource value is less than threshold = 0.05
+% 	}
+% 	\label{SN}
+% \end{figure}
+%
+%
+%
+% \begin{figure}[h]
+% 	\centering
+% 	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
+% 	\caption{
+% 	Hopf bifurcation
+% 	}
+% 	\label{SN}
+% \end{figure}
+
+\bibliography{aa_starving}
+
+\begin{acknowledgments}
+C.P.K was supported by the Omidyar Fellowship at the Santa Fe Institute
+\end{acknowledgments}
+
+%\onecolumn
+%\let\cleardoublepage\clearpage
+
+%FIGURE 1
+
+\end{article}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
+\caption{ % 58 words
+}
+\label{fig:growth}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.5\textwidth]{fig_GvS.pdf}
+\caption{ % 58 words
+}
+\label{fig:gvs}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.5\textwidth]{fig_FixedPoint.pdf}
+\caption{ % 58 words
+}
+\label{fig:fp}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.5\textwidth]{fig_DataHopf.pdf}
+\caption{ % 58 words
+}
+\label{fig:hopf}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.75\textwidth]{fig_ExtinctionAllometric.pdf}
+\caption{ % 58 words
+}
+\label{fig:ext}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
+\caption{ % 58 words
+}
+\label{fig:mass}
+\end{figure*}
+
+% 
+% 
+%	 \begin{figure}[h]
+%		 \centering
+%		 \includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
+%		 \caption{
+%		 }
+%		 \label{growth-diagram}
+%	 \end{figure}
+%	 
+%	 \begin{figure}[h]
+%		 \centering
+%		 \includegraphics[width=0.25\textwidth]{fig_GvS.pdf}
+%		 \caption{
+%		 }
+%		 \label{GvS}
+%	 \end{figure}
+%
+%
+%	\begin{figure}[h]
+%	 	\centering
+%	 	\includegraphics[width=0.25\textwidth]{fig_FixedPoint.pdf}
+%	 	\caption{
+%	 	}
+%		\label{Hopfb}
+%	\end{figure}
+%
+%
+%	\begin{figure}[h]
+%		\centering
+%		\includegraphics[width=0.8\textwidth]{fig_DataHopf.pdf}
+%		\caption{
+%		}
+%		\label{DataHopf}
+%	\end{figure}
+%	 
+% 
+%	\begin{figure}[h]
+%		\centering
+%	 	\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometric.pdf}
+%	 	\caption{
+%	 	}
+%	 	\label{Ext}
+%	\end{figure}
+%	
+%	\begin{figure}[h]
+%	 	\centering
+%	 	\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
+%	 	\caption{
+%	 	}
+%	 	\label{Asymp}
+%	\end{figure}
 %
@@
 
-\clearpage
-
-	 \begin{figure}[h]
-		 \centering
-		 \includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-		 \caption{
-		 }
-		 \label{growth-diagram}
-	 \end{figure}
-
-	 \begin{figure}[h]
-		 \centering
-		 \includegraphics[width=0.5\textwidth]{fig_GvS.pdf}
-		 \caption{
-		 }
-		 \label{GvS}
-	 \end{figure}
-
-	\begin{figure}[h]
-	 	\centering
-	 	\includegraphics[width=0.75\textwidth]{fig_FixedPoint.pdf}
-	 	\caption{
-	 	}
-		\label{Hopfb}
-	\end{figure}
-
-	\begin{figure}[h]
-		\centering
-		\includegraphics[width=0.8\textwidth]{fig_DataHopf.pdf}
-		\caption{
-		}
-		\label{DataHopf}
-	\end{figure}
-
-	\begin{figure}[h]
-		\centering
-	 	\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometric.pdf}
-	 	\caption{
-	 	}
-	 	\label{Ext}
-	\end{figure}
-
-	\begin{figure}[h]
-	 	\centering
-	 	\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
-	 	\caption{
-	 	}
-	 	\label{Asymp}
-	\end{figure}
-
-%
-%
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
-% 	\caption{
-% 	Probability that the resource value is less than threshold = 0.05
-% 	}
-% 	\label{SN}
-% \end{figure}
-%
-%
-%
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
-% 	\caption{
-% 	Hopf bifurcation
-% 	}
-% 	\label{SN}
-% \end{figure}
-
-\clearpage
-
-\section*{References}
-
-\bibliographystyle{apalike} % for Science,
-\bibliography{aa_starving}
-
-%\end{flushleft}
-\end{spacing}
 \end{document}

### 2016-07-20 Manuscript/Yeakel_Kempes-Redner_PNAS_7-19.tex
Final first draft complete

@@
 %\section*{Introduction}
-The behavioral ecology of most, if not all, organisms is influenced by the energetic state of individuals, which directly influences how an organism invests its reserves in an uncertain environment.
-Such behaviors are generally manifested as trade-offs between investing in individual maintenance and growth (somatic effort) or allocating energy towards reproduction (reproductive effort) \cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
+The behavioral ecology of most, if not all, organisms is influenced by the energetic state of individuals, which directly influences how organisms invest reserves in uncertain environments.
+Such behaviors are generally manifested as trade-offs between investing in somatic maintenance and growth or allocating energy towards reproduction \cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
 %among a host of other behavioral duties [REFS].
-The timing of these behaviors is often important and is under strong selective pressure, as they tend to have large effects on the future fitness of the organism \cite{Mangel:1988uaa}.
+The timing of these behaviors is often important and is under strong selective pressure, as it tends to directly impact future fitness \cite{Mangel:1988uaa}.
 %For example, rotifers that reproduce have lower survival rates than those that don't when they are under nutritional stress.
-To what extent, and when, organisms invest in somatic vs.
+%To what extent, and when, organisms invest in somatic vs.
 reproductive expenditures may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, and the distribution of resources.
@@
 Maximizing fitness between growth and maintenance activities vs.
-reproductive efforts in large part structures the life-history of species, and this can be achieved by alternative behavioral strategies or via changes in physiology, both of which are sensitive to resource availability.
+reproduction structures the life-history of many species, and this can be achieved by alternative behavioral strategies conditioned on resource availability \cite{Morris:1987eo}.
+%or by changes in physiology, both of which are sensitive to
 %Reproduction can be separated from somatic growth/maintenance by a variety of potential mechanisms:
 %Multiple mechanisms contribute to the differentiation of these behaviors:
-Behavioral changes in somatic or reproductive investment often occur in response to limited resources \cite{Morris:1987eo}.
+%Behavioral changes in somatic or reproductive investment often occur in response to limited resources \cite{Morris:1987eo}.
 %, while stress can generally lead to reprodusuppression of reproductive behaviors \cite{Morgan:1999do}. 
-For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \cite{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for a breeding season (capital breeding) \cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
+For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \cite{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for a breeding season \cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
 %Such breeding behaviors is generally referred to as \emph{capital breeding}
 %Resource limitation can also alter the behaviors of species in well-mixed environments: 
-Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress \cite{Threlkeld:1976ih}, with those that do reproduce evincing lower survival rates \cite{Kirk:1997cc}, while artificially induced stress has been observed to decrease reproductive success in Atlantic cod \cite{Morgan:1999do}.
+Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress \cite{Threlkeld:1976ih}, with those that do reproduce have lower survival rates \cite{Kirk:1997cc}.
+%, while artificially induced stress has been observed to decrease reproductive success in Atlantic cod \cite{Morgan:1999do}.
 Organisms may also separate maintenance and growth from reproduction over space and time: 
-many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in alternative environments spent accumulating body mass and nutritional reserves \cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in alternative environments spent accumulating nutritional reserves \cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
 
 Physiological mechanisms also play an important role in regulating reproductive expenditures during periods of resource limitation.
-Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation) to time with accumulation of nutritional reserves \cite{Mead:1989dt,Sandell:1990kw}.
-%, including multiple species of Grizzly bears (\emph{Ursus horribilis}) and the European badger (\emph{Meles melesc}) \cite{Woodroffe:1995id
-Furthermore, many mammals (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress \cite{Bulik:1999eo,Trites:2003cc}.
+Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation) to time with accumulation of nutritional reserves \cite{Mead:1989dt,Sandell:1990kw}, while many others (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress \cite{Bulik:1999eo,Trites:2003cc}.
 In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.
@@
 
-Though straightforward conceptually, incorporating the energetic dynamics of individuals \cite{Kooi2000} into a population-level framework \cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles (in particular a lack of smoothness or differentiability \cite{Diekmann:2010da}), and often suffers from over-fitting due to an overabundance of parameters.
+Though straightforward conceptually, incorporating the energetic dynamics of individuals \cite{Kooi2000} into a population-level framework \cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles (in particular a lack of smoothness or differentiability \cite{Diekmann:2010da}), and is prone to over-fitting.
 %These issues have limited the development of theoretical models that may aid our understanding of the effects of such tradeoffs on population dynamics.
@@
 Resource limitation and the subsequent effects of starvation may be alternatively accounted for \emph{explicitly}, such that reproduction is permitted only for individuals with sufficient energetic reserves.
-This permits us to incorporate 
-\emph{i}) the reproductive time lag that is sensitive to changing rates of starvation and recovery, and 
-\emph{ii}) the idea that reproduction is strongly allometrically constrained \cite{Kempes:2012hy}, and is not generally a linear function of resource density.
+Such a dynamic introduces 
+\emph{i}) the reproductive time lag associated with changing rates of starvation and recovery, and 
+\emph{ii}) the idea that reproduction is strongly allometrically constrained \cite{Kempes:2012hy}, and not generally linearly related to resource density.
 %Incorporating explicit starvation introduces
@@
 
-\noindent {\bf Starvation dynamics} \\ \nonumber
+\noindent {\bf Nutritional state-structured model (NSM)} \\ \nonumber
 %What we did...
-We explore how the energetic tradeoff between maintaining and building somatic tissue vs. reproduction can influence the dynamics of populations, and how such dynamics may be constrained by allometry.
+We explore how the energetic tradeoff between somatic maintenance and growth vs. reproduction can influence population dynamics, and how such dynamics may be constrained by allometry.
 We begin by establishing a minimal Nutritional State-structured population Model (NSM), where the consumer population is divided into two energetic states: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
@@
 %In the NSM there are only two energetic states for the consumer population: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
-The resource $R$ has logistic growth with a linear growth rate $\alpha$ and carrying capacity of unity. %(throughout we set $\kappa=1$ such that $R$ has units of $1/\kappa$).
+The resource $R$ has logistic growth with an intrinsic growth rate $\alpha$ and carrying capacity of unity. %(throughout we set $\kappa=1$ such that $R$ has units of $1/\kappa$).
 Consumers transition from state $F$ to state $H$ by starvation at rate $\sigma$ and in proportion to the lack of resources $(1-R)$.
@@
 
-\begin{equation} \label{eq_specific}
+\begin{equation} \label{eq:system}
 \begin{aligned}
@@
 
-There are three steady states for the NSM: two trivial fixed points at $(R^*=0,H^*=0,F^*=0)$ and $(R^*=1,H^*=0,F^*=0)$, and one non-trivial internal fixed point at
+There are three fixed points associated with the NSM: two trivial fixed points at $(R^*=0,H^*=0,F^*=0)$ and $(R^*=1,H^*=0,F^*=0)$, and one internal fixed point at
 
-\begin{equation}
+\begin{equation} \label{eq:ss}
 \begin{aligned}
@@
 
-\noindent Because there is only one internal fixed point, as long as it is stable the population trajectories will serve as a global attractor for any set of positive initial conditions.
+\noindent Because there is only one internal fixed point, as long as it is stable the population trajectories will serve as a global attractor for any set of initial conditions $>0$.
 %Analysis of the stability of the NSM is explored with respect to the local stability of the internal steady state, which is the only feasible steady state as long as both the consumer and resource have non-zero, positive, values.
-In a multidimensional system, linear stability is determined with respect to the Jacobian Matrix $\bf J|_*$ (where $|_*$ denotes evaluation at the internal steady state), where each element of the matrix is defined by the partial derivative of each differential equation with respect to each variable.
+Stability is determined with respect to the Jacobian Matrix $\bf J|_*$ (where $|_*$ denotes evaluation at the internal steady state), where each element of the matrix is defined by the partial derivative of each differential equation with respect to each variable.
 %General statement on stability
@@
 
+%CHRIS - we need to define beta in this section...
+
 \vspace{2mm}
-\noindent {\bf Allometric relationships} \\ \nonumber
+\noindent {\bf Allometric rates} \\ \nonumber
 %[Link Allometry stuff to our model - what does it provide to the story]
@@
 %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-Our model explores a wide variety of possible dynamics yet we have not described the realistic regimes occupied by organisms where the challenge is to constrain the covariation of rates.
+The parameters in the NSM cannot be freely navigated in biologically-reasonable settings, and our first challenge is to constrain the covariation of rates in a principled and biologically meaningful manner.
+% we have not described the realistic regimes occupied by organisms where the
 Allometric scaling relationships highlight common constraints and average trends across large ranges in body size and species diversity.
-Many of these relationships can be derived from a small set of assumptions and below we describe a framework for the covariation of timescales and rates across the range of mammals for each of the key parameters of our model.
+Many of these relationships can be derived from a small set of assumptions and below we describe a framework for the covariation of timescales and rates across the range of mammals for each of the key parameters of our model (cf.
+\cite{Yodzis:1992hg}).
 We are able to define the regime of dynamics occupied by the entire class of mammals along with the key differences between the largest and smallest mammals.
@@
 
-It should be noted that we have thus far used mammalian allometry to describe the size-based relationships for growth, starvation, and death.
+%CHRIS - MAYBE WE CUT THIS OUT SINCE WE STICK TO MAMMALS IN THIS PAPER
+%It should be noted that we have thus far used mammalian allometry to describe the size-based relationships for growth, starvation, and death.
 However, our presentation is general, and other functional forms for $\epsilon$, for example, could be determined for other classes of organisms.
@@
 Investigating other classes of organisms requires only substituting the energetic and scale parameters shown in Table \ref{param}.
-Moreover, we emphasize that our allometric equations describe the mean allometric , and do not account for the (sometimes considerable) variance associated with individual species. 
+Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species. 
 %, though this is beyond the current scope of our analysis. %, and we henceforth focus our examination on mammalian species.
@@
 \ref{fig:gvs}).
-We note that the asymptote for the starvation rate at $M \approx 6.16\times10^7$ defines the mass at which fat and muscle tissue accounts for 100\% of organismal weight, thereby placing a hard scaling boundary on our derivation for the starvation rate.
+We note that the asymptote for the starvation rate at $M \approx 8\times10^8$ defines the mass at which body fat accounts for 100\% of organismal weight, thereby placing a scaling bound on our derivation for the starvation rate.
 %According to this derivation, the endotherm (the blue whale \emph{}) is 74\% fat, which accords with empirical estimates REF.
@@
 %General Hopf bifurcation info
-In addition to the hard bound defined by the TC bifurcation, oscillating or cyclic dynamics present an implicit constraint to persistence by increasing the risk of extinction.
+In addition to the hard bound defined by the TC bifurcation, oscillating or cyclic dynamics present an implicit constraint to persistence by increasing extinction risk due to stochastic effects.
 %also present certain constraints to the feasibility of populations.
 %Oscillating, or cyclic, dynamics present additional risks to populations.
-If cycles are large, stochastic effects may result in extinction in either the consumer or resource population.
 In continuous-time systems, a stable limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts \cite{GuckHolmes}.
@@
 Moreover, as a non-cyclic stable system nears the Hopf bifurcation, transient or decaying cycles can grow in magnitude, despite the existence of a positive, non-cyclic, steady state density.
-Given that ecological systems exist in a state of constant perturbation \cite{Hastings:2001jh}, even the onset of transient cycles that decay over time can increase the risk of extinction \cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}, such that the distance of a system from the Hopf bifurcation, even in the stable region, is relevant to persistence.
+Given that ecological systems exist in a state of constant perturbation \cite{Hastings:2001jh}, even the onset of transient cycles that decay over time can increase the risk of extinction \cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}, such that the distance of a system from the Hopf bifurcation is relevant to persistence.
 
@@
 Non-cyclic stability of the fixed point generally requires a higher starvation rate $\sigma$ relative to the recovery rate $\rho$.
-The intuition behind this is that transition to the hungry (non-reproductive) state permits the resource to recover and transient dynamics to subside, whereas a low $\sigma$ overloads the system with energetically-replete (reproducing) individuals, thus producing maintained oscillations between consumer and resource (Fig.
-\ref{Hopfb}).
-However if $\sigma$ is too large, mortality due to starvation depletes the consumer population, resulting in a lower steady state density for the consumer and an opposingly higher steady state density for the resource.
+The intuition behind this is that transition to the hungry (non-reproductive) state permits the resource to recover and transient dynamics to subside, whereas a low $\sigma$ overloads the system with energetically-replete (reproducing) individuals, thus maintaining oscillations between consumer and resource (Fig.
+\ref{fig:fp}).
+However if $\sigma$ is too large, mortality due to starvation depletes the consumer population, resulting in a lower steady state density for the consumer and a higher steady state density for the resource.
 
@@
 \ref{Hopfb} cannot be freely navigated if assuming biologically reasonable parameterizations.
-Given the parameterization for terrestrial endotherms shown in Table \ref{xx} with mass $M$ as the only free parameter, rates of starvation and recovery are constrained to a fairly small window of potential values.
-We find that the allometric constraints given for terrestrial endothermic organisms confine dynamics to the steady state regime across all reasonable body size classes, which for mammals ranges from ca.
-1 gram (the Etruscan shrew) to ca.
-$10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
-Moreover, for larger organismal mass, the distance increases between the allometric values for $\sigma(M)$ and $\rho(M)$ relative to the Hopf bifurcation, while uncertainty in allometric parameters (20\% variation around the mean; Fig.
-\ref{xx}) results in little difference in the position of the TC and Hopf bifurcations as well as consumer energetic rates.
+Given the parameterization for terrestrial endotherms shown in Table \ref{param} with mass $M$ as the only free parameter, rates of starvation and recovery are constrained to a fairly small window of potential values (Fig.
+\ref{fig:hopf}), thus confining dynamics to the steady state regime for all reasonable body size classes.
+%, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
+Moreover, for larger $M$, the distance increases between the allometric values for $\sigma(M)$ and $\rho(M)$ relative to the Hopf bifurcation, while uncertainty in allometric parameters (20\% variation around the mean; Fig.
+\ref{fig:hopf}) results in little difference in the position of the TC and Hopf bifurcations as well as consumer energetic rates.
 This result suggests 
@@
 2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
-Although the prediction of larger oscillations for animals with smaller body sizes generally holds, empirical observations of large animal population cycles are plagued by long generation times and the influence of top-down effects [REFS] making direct empirical observation problematic.
+
+Allometric constraints have been invoked to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, such that period $\propto M^{0.25}$, however this relationship seems to hold only for some species \cite{Hendriks:2012fc} and competing explanations exist \cite{Kendall:1999iy,Hogstedt:2005cr}.
+%It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic \cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
+Though statistically significant support for the existence of population cycles among mammals is based on data for smaller body sized organisms \cite{Kendall:1998hl}, we acknowledge that absence of evidence for cyclic dynamics for longer-lived species is not necessarily evidence of absence.
+We thus obtain a specific prediction from our model: population cycles should be less common among larger species and more common for smaller species, particularly in environments where resources are limiting, emphasizing the role of starvation on dynamics.
 
 %Extinction to the left and the right
-Higher rates of starvation results in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of death, however it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{Hopfb}).
-By examining the ratio $\sigma/\rho$, we can understand the competing effects of cyclic dynamics vs. population density on the probability of extinction.
-We computed the probability of extinction, which we defined as $H(t)+F(t)=10$ at any instant across all values of $10^3<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq.
-XX for an organism of $M=100{\rm g}$, assuming random initial conditions around the steady state solution.
+Higher rates of starvation results in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality, however it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}).
+By examining the ratio $\sigma/\rho$, we can understand the competing effects of cyclic dynamics vs. steady state density on extinction risk.
+We computed the probability of extinction, where extinction is defined as $H(t)+F(t)=10$ at any instant across all values of $10^3<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq.
+\ref{eq:system} for an organism of $M=100$ grams, assuming random initial conditions around the steady state (Eq.
+\ref{eq:ss}).
 %random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
-By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for the ratio $\sigma/\rho$ varying between $10^-2$ to $2.5$, thus examining a vertical cross-ection of Fig.
-\ref{Hopfb}.
-As we expected, we found higher rates of extinction for both low and high values of $\sigma/\rho$; for low values the higher extinction risk is the consequence of transient cycles with larger amplitudes as the system nears the Hopf bifurcation.
-For high values of $\sigma/\rho$, higher extinction risk is due to the steady decrease in the steady state consumer population density.
-This interplay creates an `extinction refuge', such that for a relatively small range of $\sigma/\rho$, extinction probabilities are minimized.
+By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a horizontal cross-section of Fig.
+\ref{fig:fp}.
+As expected, higher rates of extinction correlated with both low and high values of $\sigma/\rho$; for low values the higher extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.
+\ref{fig:ext}).
+For large values of $\sigma/\rho$, higher extinction risk is due to the decrease in the steady state consumer population density.
+This interplay creates an `extinction refuge' as shown in Fig. \ref{fig:ext}, such that for a relatively small range of $\sigma/\rho$, extinction probabilities are minimized.
 
 As has been described, the $\sigma$ vs. $\rho$ space cannot be freely traversed, such that not all values of $\sigma/\rho$ are biologically feasible.
-We observe that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around the mean for the energetic parameters used to determine each rate) fall within the extinction refugia, such that they are close enough to the Hopf bifurcation to avoid low steady state densities, though far enough away to avoid large-amplitude transient cycles.
+We observe that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around energetic parameter means) fall within the extinction refugia, such that they are close enough to the Hopf bifurcation to avoid low steady state densities, though far enough away to avoid large-amplitude transient cycles.
 The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving observed starvation and recovery rates within populations.
 Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-Despite the possible mechanisms, our finding that allometrically-determined energetic rates place the system within this low probability of extinction region suggests that the NSM system provides general insight to a phenomena that may both drive -- and constrain -- natural animal populations.
+Nevertheless, our finding that allometrically-determined energetic rates place the system within this low probability of extinction region suggests that the NSM system provides general insight to a phenomena that may both drive -- and constrain -- natural animal populations.
 
@@
 \vspace{2mm}
-\noindent {\bf Minimizing extinction risk} \\ \nonumber
+\noindent {\bf Dynamic and energetic barriers on body size} \\ \nonumber
+Metabolite transportation constraints are widely thought to place strict boundaries on biological scaling \cite{Brown:1993p708,West:1997cg,Brown:2004wq}, leading to specific predictions on the minimum possible body size for organisms \cite{West:2002ud}.
+In the opposite vein, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits of larger body masses, particularly for mammals.
+The \emph{fasting endurance hypothesis} contends that larger body size, being energetically more efficient per unit biomass, may buffer organisms against environmental fluctuations in resource availability \cite{Millar:1990p923}.
+Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's Rule) \cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers may instigate a selective force towards an optimal upper-bound of ca.
+$10^7$ grams \cite{Alroy:1998p1594}, the value of which may be due to higher extinction risks for large taxa over longer timescales \cite{Clauset:2009fh}.
+These evolutionary trends are thought to be driven by a combination of climate and niche availability \cite{Saarinen:2014br}, however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, has not been explored, and we contend that the NSM provides a suitable framework to explore this problem.
+
+%Minimal mass bound
+The NSM correctly predicts that species with smaller masses have larger steady state population densities, however we observe that there is a sharp asymptote in both steady state densities as well as $\sigma/\rho$ at $M \approx 0.3$ grams  (Fig.
+\ref{fig:mass}b).
+Observation of the rates of starvation and recovery explain why: as mass decreases, the rate of starvation increases steadily, while the rate of recovery begins to decline precipitously.
+This exponential decline in $\rho$ occurs when body fat percentage is $1 - 1/(cM) \approx 2\%$, whereupon consumers have no eligible route out of starvation, leading to extinction.
+Compellingly, this dynamic bound determined by the rate of energetic recovery is close to the minimum observed mammalian body size of ca.
+1.3-2.5 grams (Fig.
+\ref{fig:mass}b,c), a range that occurs as the recovery rate begins to decline exponentially.
+In addition to barriers associated with metabolite transport constraints \cite{West:2002ud}, we suggest that an additional limitation stems from the dynamics of starvation.
+
+%Maximum mass bound
+Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
+Instead, we investigate population invasibility with respect to a mutated subset of the population (identified by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-0.5,0.5]$, thus altering rates of starvation, recovery, and maintenance $\beta$.
+Although there is not an internal fixed point where both residents and invaders coexist (except for the trivial state $\chi=0$), we can assess invasibility as a function of organismal mass by determining which consumer steady state is larger across values of modified body fat.
+We find that for $1\leq M<10^6$g, having additional body fat ($\chi > 0$) results in a higher steady state density for the invader population ($H^{\prime *}+F^{\prime *}>H^*+F^*$), such that it has an intrinsic advantage over the resident population.
+For $M>10^6$, however, there is an increasing range of $\chi < 0$ such that leaner individuals have the advantage, and this is due to the changing covariance between energetic rates as a function of modified body fat.
+
+The observed switch in invasibility as a function of $\chi$ at $M_{\rm opt}\approx10^6$ thus serves as an attractor, where over evolutionary time the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
+Moreover, $M_{\rm opt}$, which is entirely determined by the population-level consequences of energetic constraints is within an order of magnitude as that boserved in the North American mammalian fossil record \cite{Alroy:1998p1594} and as that predicted from an evolutionary model of body size evolution \cite{Clauset:2009fh}.
+Of course, the state of the environment in combination with the realized competitive landscape will determine whether specific body size classes are selected for or not \cite{Saarinen:2014br}, however the internal dynamic that we examine with the NSM may supply the fundamental momentum fueling the evolution of larger body size among terrestrial mammals.
+
+%Closing
+The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations \cite{Stearns:1989ip}.
+The NSM is a minimal and general model that incorporates the dynamics of starvation that are expected to occur in resource limited environments.
+By incorporating allometric relationships between the rates in the NSM, we find
+\emph{i}) different organismal masses are more or less prone to different population dynamic regimes,
+\emph{ii}) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and
+\emph{iii}) the dynamic consequences of these rates may place additional barriers on the evolution of minimum and maximum body size.
+We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.
+Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation.
+
+%In accordance with the fasting endurance hypothesis, 
+%
+%
+%Such energetic advantages, coupled with the 
+%Ove and tends to be selected for over evolutionary time.
+%The NSM is in accordance with the XXX, where species with larger masses are less likely to evince cyclic dynamics.
+%
+%
+%
+%
+%We offer an additional mechanism for the minimum and maximum limitation on mammalian body size operating on ecological timescales.
+%Such a mechanism would 
+%
+%
+%
+%
+%Although these physiological barriers are indeed of primary importance, our minimal model of starvation dynamics suggests that there may exist additional constraints due to dynamic and evolutionary effects at small and large body sizes, respectively.
+%
+%
+%
+
+%Looking Forward
 
@@
 \begin{acknowledgments}
-C.P.K was supported by the Omidyar Fellowship at the Santa Fe Institute
+C.P.K was supported by a Trump Fellowship from the American League of Conservatives.
 \end{acknowledgments}
@@
 \centering
-\includegraphics[width=0.75\textwidth]{fig_ExtinctionAllometric.pdf}
+\includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric.pdf}
 \caption{ % 58 words
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
+\includegraphics[width=0.45\textwidth]{fig_FPAllometric.pdf}
 \caption{ % 58 words
@@
 \label{fig:mass}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=1\textwidth]{fig_Invasion.pdf}
+\caption{ % 58 words
+}
+\label{fig:invasion}
 \end{figure*}

### 2016-07-21 Manuscript/Yeakel_Kempes-Redner_PNAS_7-19.tex
additional grammatical edits

@@
 \ref{fig:gvs}).
-We note that the asymptote for the starvation rate at $M \approx 8\times10^8$ defines the mass at which body fat accounts for 100\% of organismal weight, thereby placing a scaling bound on our derivation for the starvation rate.
+%We note that the asymptote for the starvation rate at $M \approx 8\times10^8$ defines the mass at which body fat accounts for 100\% of organismal weight, thereby placing a scaling bound on our derivation for the starvation rate.
 %According to this derivation, the endotherm (the blue whale \emph{}) is 74\% fat, which accords with empirical estimates REF.
@@
 The NSM exhibits both non-cyclic as well as cyclic dynamics, and which behavior dominates depends strongly on the rate of starvation $\sigma$ relative to the rate of recovery $\rho$.
-Although starvation leads to mortality risk for the individual, a moderate amount promotes persistence of both consumer and resource populations.
-%Sigma and rho vs. Hopf and TC
-Non-cyclic stability of the fixed point generally requires a higher starvation rate $\sigma$ relative to the recovery rate $\rho$.
-The intuition behind this is that transition to the hungry (non-reproductive) state permits the resource to recover and transient dynamics to subside, whereas a low $\sigma$ overloads the system with energetically-replete (reproducing) individuals, thus maintaining oscillations between consumer and resource (Fig.
+Although starvation leads to mortality risk, a moderate amount promotes persistence of both consumer and resource populations as non-cyclic stability of the fixed point generally requires a higher $\sigma$ relative to $\rho$.
+The intuition behind this is that transition to the hungry (non-reproductive) state permits the resource to recover and transient dynamics to subside, whereas a low $\sigma$ overloads the system with energetically-replete (reproducing) individuals, maintaining oscillations between consumer and resource (Fig.
 \ref{fig:fp}).
-However if $\sigma$ is too large, mortality due to starvation depletes the consumer population, resulting in a lower steady state density for the consumer and a higher steady state density for the resource.
+If $\sigma$ is too large, mortality due to starvation depletes the consumer population, resulting in a lower steady state density for the consumer and a higher steady state density for the resource.
 
-Whereas the rate of consumer growth defines a hard bound of biological feasibility (the TC bifurcation), the rate of starvation determines the sensitivity of the consumer population to changes in resource density.
-While higher rates of starvation result in lower steady state population size -- increasing the risk of stochastic extinction -- lower rates of starvation result in a system poised near either the TC or Hopf bifurcation (or both), which will lead to elimination of the resource or the development of cyclic oscillations, respectively.
+Whereas the rate of consumer growth defines a hard bound of biological feasibility (the TC bifurcation), the rate of starvation thus determines the sensitivity of the consumer population to changes in resource density.
+While higher rates of starvation result in lower steady state population size -- increasing the risk of stochastic extinction -- lower rates of starvation result in a system poised near either the TC or Hopf bifurcation, which will lead to elimination of the resource or the development of cyclic oscillations, respectively.
 Which bifurcation is approached is wholly dependent on the rate of recovery: if it is high, then cyclic dynamics will develop; if it is low, resource extinction becomes increasingly likely.
 
-%
-%
-%
-%%Hopf bifurcation as a function of b
-%Both full and hungry consumers remove resources at rates $b$ and $\rho$, respectively.
-%As the rate of resource consumption by full consumers increases, the Hopf condition changes from a convex to a concave function over $\sigma$, limiting the potential for oscillatory dynamics.
-%These rates are considered separately because full consumers need only to maintain their tissues, whereas hungry consumers require both growth and maintenance, such that $t_\rho > t_b$, or $\rho < b$.
-%If this constraint is enacted, the likelihood of oscillatory dynamics is reduced for a given 
-%
-%
-%
-%If instead the consumer's growth was proportional to resource abundance, such that the effects of starvation on reproduction were incorporated explicitly (where the 2-stage consumer resource model collapses to the Lotka-Volterra consumer-resource model with logistic growth of the resource), the TC bifurcation exists only for $\lambda = \mu$, such that the rate of mortality cannot exceed the intrinsic birth rate.
-%
-%
-% whereas the traditional Lotka-Volterra dynamic assumes that the reproductive rate of the consumer is scaled to resource density, such that the growth function would be $G(F,R) = \lambda R F$.
-%Thus, the Lotka-Volterra dynamic \emph{implicitly} accounts for starvation in reducing the reproductive rate of the consumer.
-%However, our 2-stage model \emph{explicitly} accounts for starvation as well as recovery, such that individuals who are not starved should adopt a reproductive rate independent of resource density.
-%
-%
-%
-%We have used scaling relationships between tissue turnover and growth to strictly constrain 5/6 population-level parameters in our 2-stage consumer resource model (including the mortality rate $t_\mu$, which we have shown is just a xxx of $t_\sigma$).
-%This exercise accomplishes two goals:
-%1) it allows us to constrain the plausible parameter space of the two-stage model, and 
-%2) 
-%
-%This allows us to derive many aspects of the system in terms of consumer body mass $M$ and the allometric scaling exponent $\eta$.
-
 %Bifurcation space as a function of mass
-As the allometric derivations of NSM rate laws reveal, $\sigma$ and $\rho$ are not independent parameters, such that the bifurcation space shown in Fig.
-\ref{Hopfb} cannot be freely navigated if assuming biologically reasonable parameterizations.
-Given the parameterization for terrestrial endotherms shown in Table \ref{param} with mass $M$ as the only free parameter, rates of starvation and recovery are constrained to a fairly small window of potential values (Fig.
-\ref{fig:hopf}), thus confining dynamics to the steady state regime for all reasonable body size classes.
+As the allometric derivations of NSM rate laws reveal, $\sigma$ and $\rho$ are not independent parameters, and the bifurcation space shown in Fig.
+\ref{fig:fp} cannot be freely navigated if assuming biologically reasonable parameterizations.
+Given the parameterization for terrestrial endotherms shown in Table \ref{param} with mass $M$ as a free parameter, we show that $\sigma$ and $\rho$ are constrained to a small window of potential values (Fig.
+\ref{fig:hopf}), thus confining dynamics to the steady state regime for all realized body size classes.
 %, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
-Moreover, for larger $M$, the distance increases between the allometric values for $\sigma(M)$ and $\rho(M)$ relative to the Hopf bifurcation, while uncertainty in allometric parameters (20\% variation around the mean; Fig.
-\ref{fig:hopf}) results in little difference in the position of the TC and Hopf bifurcations as well as consumer energetic rates.
-This result suggests 
-1) that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than large mammals, and
-2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
+Moreover, for larger $M$, the distance to the Hopf bifurcation increases, while uncertainty in allometric parameters (20\% variation around the mean; Fig.
+\ref{fig:hopf}) results in little qualitative difference.
+This suggests that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than mammals with larger body size. %, and
+%2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
 
@@
 %It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic \cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
-Though statistically significant support for the existence of population cycles among mammals is based on data for smaller body sized organisms \cite{Kendall:1998hl}, we acknowledge that absence of evidence for cyclic dynamics for longer-lived species is not necessarily evidence of absence.
-We thus obtain a specific prediction from our model: population cycles should be less common among larger species and more common for smaller species, particularly in environments where resources are limiting, emphasizing the role of starvation on dynamics.
+Statistically significant support for the existence of population cycles among mammals is predominantly based on time-series for smaller bodied mammals \cite{Kendall:1998hl}, though we acknowledge that longer generation times precludes similar quality data for larger organisms.
+%, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
+We thus obtain a specific prediction from our model: population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting.
+%, emphasizing the role of starvation on dynamics.
 
@@
 Higher rates of starvation results in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality, however it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}).
-By examining the ratio $\sigma/\rho$, we can understand the competing effects of cyclic dynamics vs. steady state density on extinction risk.
-We computed the probability of extinction, where extinction is defined as $H(t)+F(t)=10$ at any instant across all values of $10^3<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq.
+We examine the competing effects of cyclic dynamics vs. changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.
+We computed the probability of extinction, where extinction is defined as $H(t)+F(t)=10$ at any instant across all values of $10^2<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq.
 \ref{eq:system} for an organism of $M=100$ grams, assuming random initial conditions around the steady state (Eq.
@@
 %random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
-By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a horizontal cross-section of Fig.
+By allowing the rate of starvation to vary, we assessed extinction risk across a range of values of the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a horizontal cross-section of Fig.
 \ref{fig:fp}.
@@
 For large values of $\sigma/\rho$, higher extinction risk is due to the decrease in the steady state consumer population density.
-This interplay creates an `extinction refuge' as shown in Fig. \ref{fig:ext}, such that for a relatively small range of $\sigma/\rho$, extinction probabilities are minimized.
+This interplay creates an `extinction refuge' as shown in Fig. \ref{fig:ext}, such that for a relatively constrained range of $\sigma/\rho$, extinction probabilities are minimized.
 
 As has been described, the $\sigma$ vs. $\rho$ space cannot be freely traversed, such that not all values of $\sigma/\rho$ are biologically feasible.
-We observe that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around energetic parameter means) fall within the extinction refugia, such that they are close enough to the Hopf bifurcation to avoid low steady state densities, though far enough away to avoid large-amplitude transient cycles.
+We observe that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around energetic parameter means) fall within the extinction refuge, such that they are close enough to the Hopf bifurcation to avoid low steady state densities, though far enough away to avoid large-amplitude transient cycles.
 The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving observed starvation and recovery rates within populations.
 Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-Nevertheless, our finding that allometrically-determined energetic rates place the system within this low probability of extinction region suggests that the NSM system provides general insight to a phenomena that may both drive -- and constrain -- natural animal populations.
+Nevertheless, our finding that allometrically-determined energetic rates place the system within this low extinction probability region suggests that the NSM system provides general insight to a phenomena that may both drive -- and constrain -- natural animal populations.
 
@@
 \vspace{2mm}
-\noindent {\bf Dynamic and energetic barriers on body size} \\ \nonumber
+\noindent {\bf Dynamic and energetic barriers to body size} \\ \nonumber
 Metabolite transportation constraints are widely thought to place strict boundaries on biological scaling \cite{Brown:1993p708,West:1997cg,Brown:2004wq}, leading to specific predictions on the minimum possible body size for organisms \cite{West:2002ud}.
-In the opposite vein, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits of larger body masses, particularly for mammals.
+Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
 The \emph{fasting endurance hypothesis} contends that larger body size, being energetically more efficient per unit biomass, may buffer organisms against environmental fluctuations in resource availability \cite{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's Rule) \cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers may instigate a selective force towards an optimal upper-bound of ca.
-$10^7$ grams \cite{Alroy:1998p1594}, the value of which may be due to higher extinction risks for large taxa over longer timescales \cite{Clauset:2009fh}.
-These evolutionary trends are thought to be driven by a combination of climate and niche availability \cite{Saarinen:2014br}, however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, has not been explored, and we contend that the NSM provides a suitable framework to explore this problem.
+Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's Rule) \cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate a selective force towards an optimal upper-bound of ca.
+$10^7$ grams \cite{Alroy:1998p1594}, the value of which may arise from higher extinction risk for large taxa over evolutionary timescales \cite{Clauset:2009fh}.
+These trends are thought to be driven by a combination of climate change and niche availability \cite{Saarinen:2014br}, however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, has not been explored, and we contend that the NSM provides a suitable framework to explore this problem.
 
@@
 The NSM correctly predicts that species with smaller masses have larger steady state population densities, however we observe that there is a sharp asymptote in both steady state densities as well as $\sigma/\rho$ at $M \approx 0.3$ grams  (Fig.
-\ref{fig:mass}b).
-Observation of the rates of starvation and recovery explain why: as mass decreases, the rate of starvation increases steadily, while the rate of recovery begins to decline precipitously.
-This exponential decline in $\rho$ occurs when body fat percentage is $1 - 1/(cM) \approx 2\%$, whereupon consumers have no eligible route out of starvation, leading to extinction.
+\ref{fig:mass}a,b).
+Observation of the rates of starvation and recovery explain why: as mass decreases, the rate of starvation increases, while the rate of recovery declines super-exponentially.
+This decline in $\rho$ occurs when body fat percentage is $1 - 1/(cM) \approx 2\%$, whereupon consumers have no eligible route out of starvation.
 Compellingly, this dynamic bound determined by the rate of energetic recovery is close to the minimum observed mammalian body size of ca.
 1.3-2.5 grams (Fig.
-\ref{fig:mass}b,c), a range that occurs as the recovery rate begins to decline exponentially.
-In addition to barriers associated with metabolite transport constraints \cite{West:2002ud}, we suggest that an additional limitation stems from the dynamics of starvation.
+\ref{fig:mass}b,c), a range that occurs as the recovery rate begins its decline.
+In addition to known transport limitations \cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.
 
@@
 Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
-Instead, we investigate population invasibility with respect to a mutated subset of the population (identified by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-0.5,0.5]$, thus altering rates of starvation, recovery, and maintenance $\beta$.
-Although there is not an internal fixed point where both residents and invaders coexist (except for the trivial state $\chi=0$), we can assess invasibility as a function of organismal mass by determining which consumer steady state is larger across values of modified body fat.
+Instead, we examine a potential upper bound to body mass by assessing population invasibility with respect to a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-0.5,0.5]$, thus altering rates of starvation, recovery, and maintenance $\beta$.
+Although there is not an internal fixed point where both residents and invaders coexist (except for the trivial state $\chi=0$), we can assess invasibility as a function of organismal mass by determining which consumer steady state is larger over $\chi$.
 We find that for $1\leq M<10^6$g, having additional body fat ($\chi > 0$) results in a higher steady state density for the invader population ($H^{\prime *}+F^{\prime *}>H^*+F^*$), such that it has an intrinsic advantage over the resident population.
-For $M>10^6$, however, there is an increasing range of $\chi < 0$ such that leaner individuals have the advantage, and this is due to the changing covariance between energetic rates as a function of modified body fat.
+For $M>10^6$, however, there is an increasing range of $\chi < 0$ such that leaner individuals have the advantage, and this is due to the changing covariance between energetic rates as a function of modified energetic reserves.
 
 The observed switch in invasibility as a function of $\chi$ at $M_{\rm opt}\approx10^6$ thus serves as an attractor, where over evolutionary time the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
-Moreover, $M_{\rm opt}$, which is entirely determined by the population-level consequences of energetic constraints is within an order of magnitude as that boserved in the North American mammalian fossil record \cite{Alroy:1998p1594} and as that predicted from an evolutionary model of body size evolution \cite{Clauset:2009fh}.
-Of course, the state of the environment in combination with the realized competitive landscape will determine whether specific body size classes are selected for or not \cite{Saarinen:2014br}, however the internal dynamic that we examine with the NSM may supply the fundamental momentum fueling the evolution of larger body size among terrestrial mammals.
+Moreover, $M_{\rm opt}$, which is entirely determined by the population-level consequences of energetic constraints is within an order of magnitude as that observed in the North American mammalian fossil record \cite{Alroy:1998p1594} and as that predicted from an evolutionary model of body size evolution \cite{Clauset:2009fh}.
+While the state of the environment, as well as the competitive landscape, will determine whether specific body size sizes are selected for or against \cite{Saarinen:2014br}, we suggest that the starvation dynamic proposed here may supply the fundamental momentum fueling the evolution of larger body size among terrestrial mammals.
+%and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
 

### 2016-07-21 Manuscript/Yeakel_Kempes-Redner_PNAS_7-19.tex
small text edit and title change

@@
 
-\title{The effect of starvation on the dynamics of consumer populations}
+\title{Ecological and evolutionary implications of starvation and body size}
 
@@
 %Extinction to the left and the right
-Higher rates of starvation results in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality, however it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}).
+Higher rates of starvation result in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality, however it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}).
 We examine the competing effects of cyclic dynamics vs. changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.

### 2016-07-21 Manuscript/Yeakel_Kempes-Redner_PNAS_7-19.tex
final small edits before sending to Chris and Sid

@@
 Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
-The \emph{fasting endurance hypothesis} contends that larger body size, being energetically more efficient per unit biomass, may buffer organisms against environmental fluctuations in resource availability \cite{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's Rule) \cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate a selective force towards an optimal upper-bound of ca.
+The \emph{fasting endurance hypothesis} contends that larger body size, with lower metabolic rates and able to hold more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability \cite{Millar:1990p923}.
+Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's Rule) \cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper-bound of ca.
 $10^7$ grams \cite{Alroy:1998p1594}, the value of which may arise from higher extinction risk for large taxa over evolutionary timescales \cite{Clauset:2009fh}.
-These trends are thought to be driven by a combination of climate change and niche availability \cite{Saarinen:2014br}, however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, has not been explored, and we contend that the NSM provides a suitable framework to explore this problem.
+These trends are thought to be driven by a combination of climate change and niche availability \cite{Saarinen:2014br}, however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, has not been explored, and we contend that the NSM provides a suitable framework to explore these issues.
 

### 2016-07-21 Manuscript/Old versions/version1_JY.tex
cleaning up old files

--- 2015-09-04 Manuscript/version1.tex
@@
-\documentclass[11pt]{iopart}
+\documentclass[11pt]{article}
 \usepackage{amssymb}
-\usepackage{epsfig}
+%\usepackage{epsfig}
 \usepackage{subfigure}
@@
   dimensions.  For parameters that correspond to a benign environment, the
-  system settles into a steady state.  However, the forager populations
+  system settles into a fixed point.  However, the forager populations
   oscillate in time with an increasing amplitude as the environment becomes
@@
 reproduction rates are high.  Conversely, for smaller environmental stress,
-the system converges to a steady state.
+the system converges to a fixed point.
 
@@
 
-This system of equations admits four fixed points.  Two of them, at
+This system of equations admits three fixed points.  Two of them, at
 $(F^*\!=\!0,S^*\!=\!0,R^*\!=\!0)$ and $(F^*\!=\!0,S^*\!=\!0,R^*\!=\!1)$
 correspond to an environment without foragers.  In this case, the latter
-fixed point is stable.  There is also an unphysical fixed point at negative
-densities and a non-trivial fixed point in which all populations are
+fixed point is stable.  There is also a non-trivial fixed point in which all populations are
 represented, $(F^*,S^*,R^*\!>\!0)$, corresponding to an ecologically active
-steady state.  If this unique steady state is stable, the species populations
-will be globally attracted to it for any non-zero initial conditions.
+fixed point at
 
+\begin{align}
+\begin{split}
+F^* &= \frac{\alpha  \lambda  \mu  (\mu +\rho )}{(\lambda +\mu ) (\lambda  \rho +\mu  \sigma )}, \\
+S^* &= \frac{\alpha  \lambda ^2 (\mu +\rho )}{(\lambda +\mu ) (\lambda  \rho +\mu  \sigma )}, \\ 
+R^* &= 	\frac{\mu  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma }.
+\end{split}
+\end{align}
+
+\noindent If this unique fixed point is stable, the species populations
+will be globally attracted to it for any set of initial conditions with values greater than zero.
+
+%Linearization
 To determine the stability of the non-trivial fixed point, we compute the
@@
 \begin{array}{lll}
-  \lambda  - \sigma(1-R^*)  & \rho R^* & \sigma(S^*+ S^*) \\
+  \lambda  - \sigma(1-R^*)  & \rho R^* & \sigma(S^*+ F^*) \\
   \sigma(1-R^*) & - \rho R^*-\mu &   -\rho  (S^*+ F^*) \\
@@
 
-Whether the internal steady state is attracting or repelling is conditioned
+Whether the internal fixed point is attracting or repelling is conditioned
 on the values of the parameters in the system.  If the parameters of the
-Jacobian matrix (at the internal steady state) are such that its eigenvalues
+Jacobian matrix (at the internal fixed point) are such that its eigenvalues
 have negative real values, then it is stable.  The stability of the system
@@
 eigenvalue obtains a positive real part, defined by the condition
-${\rm Det}({\bf J})=0$.  A single Saddle-Node bifurcation always exists for
-$\alpha = 0$, such that $\alpha > 0$ for the internal state to be stable.
+${\rm Det}({\bf J})=0$.  A single Transcritical bifurcation exists at
+$\lambda = \sigma$, such that the condition $\sigma < \lambda$ is required for stability of the internal fixed point.
+
+In traditional consumer-resource models, the growth of the consumer scales to both itself as well as the resource.
+The reasoning behind this is sound: as the resource density declines, the growth of the consumer is likewise expected to decline.
+This dynamic implicitly incorporates the idea that resource depletion lowers consumer growth rates.
+However, because we model the effects of starvation by the explicit inclusion of a starving consumer class, we assume that the satiated component of the consumer population has a fixed reproductive rate, which is not dependent on resource density.
+Thus, as resource density declines, the recruitment of satiated consumers into the starving class increases, serving to lower consumer growth mechanistically.
+
+Introducing mechanistic starvation imposes an added requirement for the internal fixed point to be stable: the rate of starvation must be greater than the rate of consumer reproduction.
+If this is violated, the consumer class reproduces faster than it starves, irrespective of resource density, increasing exponentially.
+We suggest that the criterion $\sigma < \lambda$ is in fact biologically reasonable because starvation occurs on a much shorter timescale than reproduction.
+
 % Although a closed-form solution is not possible, a second stability
@@
 
-Of more interest here is the existence of parameter regions that permit the
+Of additional interest is the existence of parameter regions that permit the
 existence of cyclic behavior.  Cycles arise when a pair of complex conjugate
@@
 for the specific 2-stage model, it can be explored numerically.
-
-\subsection{Generalized model}
-
-$\xi$ ~~ $\zeta$
-
-\subsection{Stochastic dynamic program}
 
@@
 
-\begin{figure}[h]
+\subsection{Generalized starvation model}
+Our conclusions of the mean-field limit starvation model are of course contingent on the functional forms chosen for the birth, death, starvation, and recovery functions.
+To expand understanding of the starvation/reproduction tradeoff on stability, we examine a generalization of our model.
+
+To generalize the starvation model, we define the functions controlling reproduction, mortality, and starvation for the original rate equations, such that
+
+\begin{align}
+  \label{GE}
+\begin{split}
+\dot F &= \lambda F + \rho  RS - \sigma (1-R)F = G(F) + H(R,S) - K(R,F)\,,\\
+\dot S &= \sigma (1-R)F - \rho RS - \mu S = K(R,F) - H(R,S) - M(S)\,, \\
+\dot R &= \alpha R(1-R) -  R(F+S) = P(R) - L(R,S,F).\\
+\end{split}
+\end{align}
+
+\noindent Although we cannot solve for the fixed points of the generalized model, we can define them as $R^*$, $S^*$, and $F^*$ for the resource, starving consumer class, and satiated consumer class, respectively.
+We then define the variables normalized to the fixed point, where
+
+\begin{equation}
+r = \frac{R}{R^*},~~~ s = \frac{S}{S^*},~~~\mbox{and}~f = \frac{F}{F^*}.
+\end{equation}
+
+By substitution, we can rewrite the general model in terms of normalized variables and functions
+
+\begin{align}
+  \label{GNE}
+\begin{split}
+\dot f &= \frac{G^*}{F^*}g(f) + \frac{H^*}{F^*}h(r,s) - \frac{K^*}{F^*}k(r,f)\,,\\
+\dot s &= \frac{K^*}{S^*}k(r,f) - \frac{H^*}{S^*}h(r,s) - \frac{M^*}{S^*}m(s)\,, \\
+\dot r &= \frac{P^*}{R^*}p(r) - \frac{L^*}{R^*}l(r,s,f).\\
+\end{split}
+\end{align}
+
+\noindent where, for example, we use the shorthand $G^* = G(F^*)$.
+The normalized general model has the important property that at the fixed point $\dot f = \dot s = \dot r = 0$, then $g(f) = h(r,s) = k(r,f) = m(s) = p(r) = l(r,s,f) = 1$.
+This allows us to define the following \emph{scale parameters}.
+These include those associated with the timescales of the system: $a_f = G^*/F^* + H^*/F^* = K^*/F^*$, $a_s = K^*/S^* = H^*/S^* + M^*/S^*$, and $a_r = P^*/R^* = L^*/R^*$, respectively denoting the turnover time for satiated consumers, starving consumers, and resources.
+We also define those scale parameters associated with the branching of biomass through the system, where
+$\beta_f = a_f^{-1}G^*/F^*$, which denotes the proportion of gain to the satiated consumer class due to growth (and $1-\beta_f$ is the proportion of gain due to recovery from starvation), and
+$beta_s = a_s^{-1}M^*/S^*$, which denotes the proportion of loss to the starving consumer class due to natural mortality (and $1-\beta_s$ is the proportion of loss due to recovery).
+
+The normalized general model can then be written
+
+\begin{align}
+  \label{GNE}
+\begin{split}
+\dot f &= a_f\left[ \beta_f g(f) + (1-\beta_f) h(r,s) - k(r,f)\right]\,,\\
+\dot s &= a_s\left[ k(r,f) - (1-\beta_s)h(r,s) - \beta_sm(s)\right]\,, \\
+\dot r &= a_r\left[p(r) - l(r,s,f)\right].\\
+\end{split}
+\end{align}
+
+%A bit on elasticities
+%The Jacobian
+%The SN/T bifurcation
+%The Hopf bifurcation
+%Correlation simulations
+%What does this tell us?
+
+\begin{figure}[h!]
 	\centering
-	\includegraphics[width=0.5\textwidth]{fig_Hopf.pdf}
+	\includegraphics[width=0.5\textwidth]{fig_FP.pdf}
 	\caption{
-	Saddle Node bifurcation
+	Fixed points.
 	}
-	\label{SN}
+	\label{fig_FP}
 \end{figure}
 
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
-% 	\caption{
-% 	Probability that the resource value is less than threshold = 0.05
-% 	}
-% 	\label{SN}
-% \end{figure}
-
-\begin{figure}[h]
+\begin{figure}[h!]
+	\begin{subfigure}
 		\centering
-	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
+		\includegraphics[width=0.4\textwidth]{fig_Hopf.pdf}
+	\end{subfigure}
+	\begin{subfigure}
+		\centering
+		\includegraphics[width=0.4\textwidth]{fig_Hopf3D.pdf}
+	\end{subfigure}
 	\caption{
@@
 	}
-	\label{SN1}
+	\label{fig_Hopf}
+\end{figure}
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.5\textwidth]{fig_freqsim.pdf}
+	\caption{
+	Time series frequencies for spatial simulation model
+	}
+	\label{fig_FreqSim}
+\end{figure}
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.5\textwidth]{fig_GenCorr.pdf}
+	\caption{
+	Generalized modeling results
+	}
+	\label{fig_Gen}
 \end{figure}
@@
 the John S. Templeton Foundation) for partial support of this research.
+
+\newpage
 

### 2016-07-21 Manuscript/Simulation_description/simulation-2.tex
cleaning up old files

--- 2016-02-23 Manuscript/simulation-3.tex
@@
 
-We assume that foragers can exist in two states---full and hungry.  When a
-full forager $F$ encounters a resource $R$, the resource is consumed with
-rate $m$ so that the forager maintains its fullness.  A full forager that
-does not encounter a resource as it wanders converts to a hungry forager $H$
-with rate equal to the product of the starvation rate $\sigma$ and the
-density of non-resources.  Full foragers also reproduce with rate
-$\lambda$. When a hungry forager encounters a resource, the resource is again
-consumed and the forager becomes full with rate $\rho$.  A hungry forager
-dies with mortality rate $\mu$, while full foragers have no mortality risk.
-We assume that in the absence of foragers the underlying resource undergoes
-logistic growth, with growth rate $\alpha$ and carrying capacity equal to 1.
+We assume that foragers can exist in two discrete states---full and hungry.
+Full foragers $F$ are those that have just encountered and consumed a unit of
+resource $R$.  On the other hand, a full forager that does not encounter a
+resource as it wanders converts to a hungry forager $H$ with rate equal to
+the product of a parameter $\sigma$ and the density of non-resources.
+Whenever a forager, either full or hungry, encounters a resource, one unit of
+the resource is consumed.  If the forager was hungry, it turns into a full
+forager with rate $\rho$.  During the time that a forager is hungry, it dies
+with mortality rate $\mu$, while full foragers do not experience mortality
+risk.  Furthermore, full foragers reproduce with rate $\lambda$.  Finally, we
+assume that, in the absence of foragers, the underlying resource undergoes
+logistic growth, with growth rate $\alpha$ and carrying capacity equal to
+one.
 
-Based on these processes and also under the assumption that the densities of
-full foragers, hungry foragers, and resources (also denoted by $F$, $H$, and
-$R$, respectively) are perfectly mixed, they evolve according to the rate
-equations:
+According to these processes and also under the assumption that the densities
+of full foragers, hungry foragers, and resources (also denoted by $F$,
+$H$, and $R$, respectively) are perfectly mixed, they evolve according to the
+rate equations:
 \begin{align}
@@
 \dot H &= \sigma (1-R)F - \rho RH - \mu H\,, \\
-\dot R &= \alpha R(1-R) -  R(mF+\rho H),\\
+\dot R &= \alpha R(1-R) -  R(F+H),\\
 \end{split}
@@
 generation of the resource.  The approach described below can be generalized
-to carrying capacity less than 1, but we take the carrying capacity equal to
-1 without loss of generality.
+to carrying capacity less than 1, but one can take the carrying capacity
+equal to 1 without loss of generality.  Here is another subtle point worth
+mentioning: $\dot R$ does not involve the recruitment rate $\rho$.  From the
+perspective of the resource, either type of forager can consume it.  However,
+from the perspective of the hungry forager, once it has eaten a resource, it
+converts to a full forager with rate $\rho$.  One has to be careful to not
+overcount the events that involve interactions between a hungry forager and a
+resource.
 
@@
 Suppose that the system at a given time consists of $N_F$ full foragers,
-$N_H$ hungry foragers, and $N_R$ resource units.  The total number of
-particles is $N=N_F+N_H+N_R$.  There are six elemental processes that are
-embodied by the rate equations \eqref{RE}:
-\newpage
+$N_H$ hungry foragers, and $N_R$ individual resources.  The total number of
+particles $N=N_F+N_H+N_R$.  There are six basic processes embodied by the
+rate equations \eqref{RE}:
 \begin{enumerate}
@@
 \item Starvation:  $F\to H$
-\item Consumption: $F\to F$ and $R\to 0$
+\item Recruitment: $H\to F$ and $R\to 0$
 \item Death: $H\to 0$
-\item Recruitment: $H\to F$ and $R\to 0$
 \item Growth: $R\to 2R$
+\item Consumption: $R\to 0$
 \end{enumerate}
-For specificity, we define ``consumption'' as the consumption of the resource
-by a full forager and the maintenance of the forager in its full state, while
-``recruitment'' is defined as the consumption of the resource by a hungry
-forager and conversion of the latter to the full state.
+Note that recruitment also involves consumption of a resource, so we must be
+careful to not double count.  So I define ``Recruitment'' as the consumption
+of the resource by a hungry forager and ``Consumption'' as the consumption of
+the resource by a full forager.
 
 In an event-driven simulation, one of the above events is picked with the
-appropriate probability to be defined below, the selected event is
-implemented, and then the time is updated accordingly.  Reading off from the
-steps (i)--(vi) listed above, the total rate for any event is proportional to
+appropriate probability (to be defined below), the selected event is
+implemented, and then the time is updated accordingly.  The total rate for
+any event is proportional to
 \begin{equation}
-\mathcal{R}=F\big[\lambda +\sigma(1-R)+mR\big]+ H\big[\mu+\rho R\big]+
-\alpha R(1-R)\,.
+\mathcal{R}=F\big[\lambda +\sigma(1-R)\big]+ H\big[\rho R+\mu\big]+
+R\big[\alpha(1-R)+(F+H)\big]\,.
 \end{equation}
 Here $F=N_F/V$, $H=N_H/V$, and $R=N_R/V$, where $N_i$ is the total number of
-particles of type $i$, and $V$ is the total number of lattice sites.  Thus
-$F$, $H$, and $R$ are the densities of each type of entity.  The total rate
-is defined up to an overall constant, but this constant is immaterial because
-the time step is also proportional to this same constant.
+particles of type $i$, and $V$ is the total number of lattice sites.  Thus $F$,
+$H$, and $R$ are the densities of each type of entity.  The
+total rate is defined only up to an overall constant, but this constant is
+immaterial because the time step is also proportional to this same constant.
 
-In the most simple version of a simulation in the mean-field limit, one
-merely tracks the three variables $N_F$, $N_H$, and $N_R$ and updates these
-numbers according to the steps outlined below.  However, we want a simulation
-that can be directly adapted to treat foragers that diffuse on a lattice.
-The algorithm outlined below is based on populating a lattice with the three
-types of entities and updating the occupancy on the lattice sites according
-to the following event-driven simulation:
+At the simplest level one merely keeps three variables: $N_F$, $N_H$, and
+$N_R$ and updates these numbers according to the steps outlined below.
+However, we want a simulation that can be directly adapted to treat foragers
+that diffuse on a lattice.  Thus the algorithm outlined below is based on
+populating a lattice with the three types of entities and then updating the
+occupancy on the lattice sites according to the following event-driven
+simulation:
 \begin{enumerate}
-\item Reproduction.  With probability $\lambda F/\mathcal{R}$, create a new
-  full forager anywhere in the system equiprobably, consistent with the
-  absence of spatial structure.
-\item Starvation.  With probability $\sigma(1-R)F/\mathcal{R}$, pick a random
-  full forager and change it to hungry.  The factor $1-R$ is the fraction of
-  resource-free space.
-\item Consumption.  With probability $mFR/\mathcal{R}$, pick and remove a random
-  resource unit.
-\item Death.  With probability $\mu H/\mathcal{R}$, remove a random
-  hungry forager.
-\item Recruitment.  With probability $\rho HR/\mathcal{R}$, change a random
-  hungry forager into a full forager and remove a random resource unit.
-\item Growth.  With probability $\alpha R(1-R)/\mathcal{R}$, create another
-  resource unit anywhere in the system.
+\item Reproduction.  With probability $\lambda F/\mathcal{R}$, pick one of
+  the full foragers and allow it to reproduce.  The offspring should be
+  placed anywhere in the system; this is consistent with there being no
+  spatial structure.
+\item Starvation.  With probability $\sigma(1-R)F/\mathcal{R}$, pick a full
+  forager and change it to a hungry forager.  The factor $1-R$ is the
+  fraction of resource-free space.
+\item Recruitment.  With probability $\rho HR/\mathcal{R}$, pick a hungry
+  forager, turn it into a full forager, and remove the resource.
+\item Death.  With probability $\mu H/\mathcal{R}$, pick a hungry forager and
+  remove it.
+\item Growth.  With probability $\alpha R(1-R)/\mathcal{R}$, pick a resource
+  unit and create another unit of resource anywhere in the system.
+\item Consumption.  With probability $(F+H)R/\mathcal{R}$, pick a resource
+  unit and remove it.
 \end{enumerate}
 By explicit construction, the probabilities for the above events sum to 1.
-It is important to appreciate that one can perform this simulation either by
-populating a lattice with the three types of entities or merely be keeping
-track of the three variables $N_F$, $N_H$, and $N_R$.  A good test of the
-lattice-based simulation is that it should match that of simulating the above
-three variables.
+Note also that a good test of the correctness of the above lattice-based
+algorithm is to compare it with the three-variable update rule that is
+mentioned at the start of this paragraph.
 
-Let's determine the change in the expected number of individuals of each type
-in a single event.  For full foragers, the change in its number $N_F$ is
+We now determine the change in the expected number of individuals of each
+type in a single event.  For full foragers, the change in its number $N_F$ is
 \begin{subequations}
@@
 \end{equation}
-The term proportional to $F$ describes processes in which a full forager is
+The terms proportional to $F$ describes processes in which a full forager is
 picked, while the term proportional to $H$ describes processes in which a
@@
 \begin{equation}
-\label{dF}
-  \Delta F =\frac{\Delta N_F}{V} =\left[F\big(\lambda-\sigma(1-R)\big)+\rho H
+  \Delta F =\frac{\Delta N_F}{V} =\left[F\big(\lambda-\sigma(1-R)\big)+H\rho
     R\right]/V\mathcal{R}\,.
@@
 \end{subequations}
-If we take the time step for each event to be
+Thus if we take the time step for each event to be
 $\Delta t = (V\mathcal{R})^{-1}$, the above reduces to the rate equation
-\eqref{RE} for $F$.  Thus for each microscopic update event, the time should
-be advanced by $\Delta t = (V\mathcal{R})^{-1}$.
+\eqref{RE} for $F$.  Thus in each microscopic event of the model, the time
+should be advanced by $\Delta t = (V\mathcal{R})^{-1}$.
 
-In a similar fashion, the change in the expected density of hungry foragers in
+In a similar fashion, the change in the expected number of hungry foragers in
 a single event is given by
-%\begin{subequations}
-%\begin{equation}
-%  \Delta N_H =\left[-H(\rho R+\mu)+\sigma F
-%    (1-R)\right]/\mathcal{R}\,,
-%\end{equation}
-%so that the change in the density of hungry foragers simply is
+\begin{subequations}
 \begin{equation}
-\label{dH}
-  \Delta H =\left[-H(\rho R+\mu)+\sigma F(1-R)\right]/V\mathcal{R}\,,
+  \Delta N_H =\left[-H(\rho R+\mu)+F\sigma
+    (1-R)\right]/\mathcal{R}\,,
 \end{equation}
-%\end{subequations}
-and the change in the density of individual resources in a single event is
-%\begin{subequations}
-%\begin{equation}
-%  \Delta N_R =\left[R\big(\alpha(1-R)\big)-(mF+\rho H)\right]/\mathcal{R}\,,
-%\end{equation}
-%so that the change in the resource density is
+so that the change in the density of hungry foragers simply is
 \begin{equation}
-\label{dR}
-  \Delta R =\left[R\big(\alpha(1-R)\big)-(mF+\rho H)\right]/V\mathcal{R}\,.
+  \Delta H =\left[-H(\rho R+\mu)+F\sigma (1-R)\right]/V\mathcal{R}\,.
 \end{equation}
-%\end{subequations}
+\end{subequations}
+Finally, the change in the expected number of individual resources in a
+single event is given by
+\begin{subequations}
+\begin{equation}
+  \Delta N_R =\left[R\big(\alpha(1-R)\big)-(F+H)\right]/\mathcal{R}\,,
+\end{equation}
+so that the change in the resource density is
+\begin{equation}
+  \Delta R =\left[R\big(\alpha(1-R)\big)-(F+H)\right]/V\mathcal{R}\,.
+\end{equation}
+\end{subequations}
 The expressions for $\Delta F/\Delta t$, $\Delta H/\Delta t$, and
 $\Delta R/\Delta t$ reproduce the original rate equations \eqref{RE} when the
-time step for an elemental update event is $\Delta t = (V\mathcal{R})^{-1}$.
-To summarize, pick an event according to the probabilities above, update the
-densities according \eqref{dF}, \eqref{dH}, and \eqref{dR}, and then
-increment the time by $(V\mathcal{R})^{-1}$.  A simulation that is based on
-the above steps should reproduce the predictions of the rate equations
-\eqref{RE}.  In the three-variable simulation, keep only $N_F$, $N_H$, and
-$N_R$.  In an update, pick an event according to the above probabilities, and
-then update $N_F$, $N_H$, and $N_R$ and the time appropriately.
+time step for an elemental event is $\Delta t = (V\mathcal{R})^{-1}$.  To
+summarize, pick an event according to the probabilities enumerated above and
+update the densities according (3b), (4b), and (5b), and then increment the
+time by $(V\mathcal{R})^{-1}$ after each event.  In the three-variable
+simulation, we keep only $N_F$, $N_H$, and $N_R$.  In an update, pick an
+event according to the above probabilities, update $N_F$, $N_H$, and $N_R$
+appropriately,and then increment the time.
 
-In implementing this algorithm for a lattice-based simulation, several points
-deserve emphasis.  First, there is no constraint on the number of foragers on
-any site, but if we are describing a harsh environment, the parameters should
-be chosen so that this number is not large.  Since the carrying capacity of
-the resource has been set to 1, $N_R$ cannot exceed $V$.  However, the number
-of resource units on any given site can exceed 1.
+There are several aspects of this algorithm that need to be specified for a
+lattice simulation.  First, there is no constraint on the number of foragers
+on any site, but given that we are describing a harsh environment, it seems
+that the parameters should be chosen so that this number is not large.  Since
+the carrying capacity of the resource has been set to 1, this means that
+$N_R$ cannot exceed $V$.  However, the number of resource units on any given
+site can exceed 1.  With these details, a simulation that is based on the
+above steps should reproduce the predictions of the rate equations
+\eqref{RE}.
 
@@
 \frac{\partial H}{\partial t}  &= \sigma (1-R)F - \rho RH - \mu H+D_H\nabla^2H\,, \\[0.125in]
-\frac{\partial R}{\partial t}  &= \alpha R(1-R) -  R(m F+\rho H)\,.\\
+\frac{\partial R}{\partial t}  &= \alpha R(1-R) -  R(F+H)\,.\\
 \end{split}
@@
 implemented, and the time is updated accordingly.
+
 The total rate for any event is
@@
   +\sigma(1-R(\mathbf{r}))+D_F\big]
-+ H(\mathbf{r})\big[\mu+\rho R(\mathbf{r})+D_H\big]\Big.\nonumber \\[-0.05in]
++ H(\mathbf{r})\big[\rho R(\mathbf{r})+\mu+D_H\big]\Big.\nonumber \\[-0.05in]
 &\Big.\hskip 0.5in
-  +\alpha R(\mathbf{r})(1-R(\mathbf{r'}))\Big\}\,\nonumber
+  +R(\mathbf{r})\big[\alpha(1-R(\mathbf{r'}))+(F(\mathbf{r})+H(\mathbf{r}))\big]\Big\}\,\nonumber
   \\[0.1in]
 &=F\big[\lambda +\sigma(1-R)+D_F\big]
-+ H\big[\mu+\rho R+D_H\big]+\alpha R(1-R)\,,
++ H\big[\rho R+\mu+D_H\big]+R\big[\alpha(1-R)+(F+H)\big]\,,
 \end{align}
@@
 \begin{enumerate}
-\item Reproduction.  With probability $\lambda F/\mathcal{R}$, pick a full
-  forager at random and let it reproduce.  The offspring should be placed
-  close to the parent---either on the same site or on a neighboring site.
+\item Reproduction.  With probability $\lambda F/\mathcal{R}$, pick one of
+  the full foragers at random and let it reproduce.  Since the process occurs
+  on a finite-dimensional lattice, the offspring should be placed close to
+  the parent.  This could be either on the same site or a neighboring site.
 \item Starvation.  With probability $\sigma(1-R)F/\mathcal{R}$, pick a random
@@
   full to hungry.  Note the two-particle nature of this rule: we need both
-  the forager type and the resource state on a given site.  To avoid keeping
-  track of sites that are simultaneously resource free and occupied by a full
-  forager, a simpler alternative is:
+  the forager type and the resource state on a given site.  There is a way to
+  get around tracking sites that are simultaneously resource free and
+  occupied by a full forager.  This leads to the alternative rule:
 \item[(ii')] Starvation'.  With probability $\sigma F/\mathcal{R}$, pick a
@@
   full to hungry.  If not, nothing happens.
-\item Consumption.  With probability $mFR/\mathcal{R}$, pick a random on a
-  site that is also by a full forager and remove the resource.  To avoid
-  tracking sites that simultaneously contain a resource and a full forager,
-  perform the following alternative:
-\item[(iii$^\prime$)] Consumption$^\prime$.  With probability
-  $R/\mathcal{R}$, pick a random resource.  If the site also contains a
-  forager, remove the resource with probability $m$.  Note the implicit
-  assumption that the density of foragers is not large because the rate of
-  resource removal does not increase if multiple foragers occupy the site.
-  If the full forager density is large then the following works:
-\item[(iii$^{\prime\prime}$)] Consumption$^{\prime\prime}$.  With probability
-  $m F/\mathcal{R}$, pick a random full forager.  If the site also contains a
-  resource, remove it.  This prescription works if the forager density is
-  small or large (as long as the resource density is less than 1) so it's
-  best to use this last alternative.
+\item Recruitment.  With probability $\rho HR/\mathcal{R}$, pick a random
+  hungry forager on a site that contains a resource unit and change the
+  forager from hungry to full and remove the resource.  This process also
+  involves a two-body reaction and perhaps a simpler way to implement this
+  two-body feature is the following:
+\item[(iii')] Recruitment'.  With probability $\rho H/\mathcal{R}$, pick a
+  random hungry forager.  If the site contains a resource unit, change the
+  forager from hungry to full and remove the resource.
 \item Death.  With probability $\mu H/\mathcal{R}$, pick and remove a random
   hungry forager.
-\item Recruitment.  With probability $\rho HR/\mathcal{R}$, pick a random
-  hungry forager on a site that also contains a resource and change the
-  forager from hungry to full and remove the resource.  This process also
-  involves a two-body reaction and a simpler way to implement this two-body
-  feature is:
-\item[(v$^\prime$)] Recruitment$^\prime$.  With probability
-  $\rho H/\mathcal{R}$, pick a random hungry forager.  If the site contains a
-  resource, change the forager from hungry to full and remove the resource.
 \item Growth.  With probability $\alpha R(1-R)/\mathcal{R}$, pick a random
-  neighboring pair of sites in which one contains a resource and the neighbor
-  does not.  Create a resource on the site without the resource.  A
-  computationally simpler alternative is:
-\item[(vi$^\prime$)] Growth$^\prime$.  With probability
-  $\alpha R/\mathcal{R}$, pick a random resource unit.  A new resource unit
-  is created on one of the neighboring sites with probability $1-m/z$, where
-  $z$ is the number of nearest-neighbors and $m$ is the number of these
-  neighbors that are occupied by resource.  If the local neighborhood is
-  devoid of resources, growth necessarily occurs at one of these neighbors,
-  while if the local neighborhood is full of resources, there is no growth.
-  This rule ensures that the growth rate is proportional to
-  $(1-R)_{\rm local}$.
+  neighboring pair of sites in which one contains a resource unit and the
+  neighbor does not.  Create a resource on the site without a resource.  A
+  computationally simpler alternative is the following:
+\item[(v)'] Growth.  With probability $\alpha R/\mathcal{R}$, pick a random
+  resource unit.  A new resource unit is created on one of the neighboring
+  sites with probability $1-m/z$, where $z$ is the number of
+  nearest-neighbors and $m$ is the number of these neighbors that are
+  occupied by resource.  Thus if the local neighborhood is devoid of
+  resources, growth necessarily occurs at one of these neighbors, while if
+  the local neighborhood is full of resources, there is no growth.  This rule
+  ensures that the growth rate is proportional to $(1-R)_{\rm local}$.
+\item Consumption.  With probability $(F+H)\,R/\mathcal{R}$, pick a random
+  resource unit on a site that is also occupied by a forager of either type
+  and remove the resource  Alternatively:
+\item[(vi')] With probability $R/\mathcal{R}$, pick a random resource unit.
+  If the site also contains a forager, remove the resource.  Note the
+  implicit assumption that the density of foragers is not large because the
+  rate of resource removal does not increase if there are multiple foragers
+  on the site.  If this forager density is large then the following will
+  work:
+\item[(vi'')] With probability $(F+H)/\mathcal{R}$, pick a random forager
+  (either full or hungry).  If the site also contains a resource, remove the
+  resource.  This prescription works if the forager density is small or large
+  (as long as the resource density is less than 1) so perhaps it's best to
+  use this last alternative.
 \item Diffusion.  With probability $D_F\, F/\mathcal{R}$, pick a random full
@@
 By explicit construction, the probabilities for the above events sum to 1.
-Note also that the primed steps are computationally simpler than the
-corresponding unprimed steps, but have the possibility of a null step; this
-introduces some inefficiency.  A true event-driven simulation avoids all null
-steps, but perhaps the computational tradeoff in bookkeeping simplicity is
-worth the loss of efficiency in allowing for null steps.  This is somewhat a
-matter of user taste and testing.
+Note also that the steps (ii'), (iii'), (v'), and (vi'') are computationally
+simpler than the unprimed steps, but allow for the possibility of a null
+step.  A true event-driven simulation avoids all null steps, but perhaps the
+computational tradeoff in bookkeeping simplicity is worth the loss of
+efficiency in allowing for null steps.  This is somewhat a matter of user
+taste.
 
@@
 is picked, while the term proportional to $H$ accounts for processes in which
-a hungry forager is picked.  The terms proportional to $D_F$ account for
-nearest-neighbor hopping: $F(\mathbf{r})$ increases because full foragers at
-neighboring sites $\mathbf{r}'$ hop to $\mathbf{r}$, while $F(\mathbf{r})$
-decreases because full foragers at $\mathbf{r}$ hop to one of the $z$
-neighboring sites.  Thus the change in the density of full foragers at site
-$\mathbf{r}$ is (after taking the continuum limit for nearest-neighbor
-hopping)
+a hungry forager is picked and converts to a full forager.  The terms
+proportional to $D_F$ account for nearest-neighbor hopping: $F(\mathbf{r})$
+increases because full foragers at neighboring sites $\mathbf{r}'$ hop to
+$\mathbf{r}$, while $F(\mathbf{r})$ decreases because full foragers at
+$\mathbf{r}$ hop to one of the $z$ neighboring sites.  Thus the change in the
+density of full foragers at site $\mathbf{r}$ is (after taking the continuum
+limit for the nearest-neighbor hopping)
 \begin{equation}
@@
 \end{subequations}
-If we take the time step for each event to be
+Thus if we take the time step for each event to be
 $\Delta t = (V\mathcal{R})^{-1}$, the above reduces to the partial
-differential evolution equation \eqref{pde} for $F(\mathbf{r})$.  In each
-microscopic event of the model, the time should be advanced by
+differential evolution equation \eqref{pde} for $F(\mathbf{r})$.  Thus in
+each microscopic event of the model, the time should be advanced by
 $\Delta t = (V\mathcal{R})^{-1}$.
@@
 
-In this algorithm, notice that single-particle and two-particle reactions are
-treated differently.  In the former case, a particle of the right type is
-picked at random.  In the latter case, a random \emph{pair} of particles of
-the appropriate types is picked.  Again, there is no constraint on the number
-of foragers on any site, but given that we are describing a harsh
-environment, the parameters should be chosen so that this number is not
-large.  The carrying capacity of the resource has been set to 1, so that the
-number of individual resources at any site must be either 0 or 1.
+There are various aspects of the simulation that require discussion.  First,
+single-particle and two-particle reactions must be treated differently.  In
+the former case, a particle of the right type is picked at random.  In the
+latter case, a random \emph{pair} of particles of the appropriate types is
+picked.  Again, there is no constraint on the number of foragers on any site,
+but given that we are describing a harsh environment, the parameters should
+be chosen so that this number is not large.  The carrying capacity of the
+resource has been set to 1, so that the number of individual resources at any
+site must be either 0 or 1.
 
-It would be nice if there was a simple way to adjust parameters in the above
+It would be nice if there was a simple way to adjust parameters in the
 lattice simulation so as to recover the mean-field limit.  I describe a
-simple modification of the above lattice simulation steps that should recover
-the mean-field limit.  We use the same rate as in Eq.~\eqref{R}, but the
-actual update steps are slightly modified compared to the rules of the
-lattice simulation above.  The differences are highlighted in italics.  I'm
-also assuming that the simulation with steps (ii$^\prime$),
-(iii$^\prime\prime$), (v$^\prime$), and (vi$^\prime$) are being used.
+simple modification of the steps in the event-driven lattice simulation that
+I think should recover the mean-field limit.  We use the same rate as in
+Eq.~\eqref{R}, but the actual update steps are slightly modified compared to
+the rules of the lattice simulation above.  The differences are highlighted in
+italics.  I'm also assuming that the simulation with steps (ii'), (iii'),
+(v'), and (vi'') are being used.
 \begin{enumerate}
-\item Reproduction.  With probability $\lambda F/\mathcal{R}$, pick a random
-  full forager and let it reproduce.  \emph{Place the offspring anywhere.}
+\item Reproduction.  With probability $\lambda F/\mathcal{R}$, pick one of
+  the full foragers at random and let it reproduce.  \emph{Place the
+    offspring anywhere.}
 \item Starvation.  With probability $\sigma F/\mathcal{R}$, \emph{change a
     random full forager to hungry with probability $(1-R)$.}
-\item Consumption.  With probability $m F/\mathcal{R}$, pick a random full
-  forager.  If the site is also occupied by a resource, remove it.
+\item Recruitment.  With probability $\rho H/\mathcal{R}$, pick a random
+  hungry forager and \emph{change it to full with probability $R$.}
 \item Death.  With probability $\mu H/\mathcal{R}$, pick and remove a random
   hungry forager.
-\item Recruitment.  With probability $\rho H/\mathcal{R}$, pick a random
-  hungry forager. \emph{Change it to full with probability $R$ and also
-    remove a random resource.}
-\item Growth.  With probability $\alpha R/\mathcal{R}$, pick a random
-  resource.  Create a resource anywhere at random with probability $1-R$.
-\item $F$ Diffusion.  With probability $D_F\, F/\mathcal{R}$, pick a random full
+\item Growth.  With probability $\alpha R(1-R)/\mathcal{R}$, pick a random
+  neighboring pair of sites in which one contains a resource unit and the
+  neighbor does not contain a resource unit.  Create a resource on the site
+  without a resource (see below for a simpler alternative).
+\item With probability $(F+H)/\mathcal{R}$, pick a random forager
+  (either full or hungry).  \emph{With probability $R$ remove a randomly
+    chosen resource unity.}
+\item Consumption.  With probability $(F+H)\,R/\mathcal{R}$, pick a random
+  resource unit on a site that is also occupied by a forager of either type
+  and remove the resource.
+\item Diffusion.  With probability $D_F\, F/\mathcal{R}$, pick a random full
   forager and move it to a neighboring site.
-\item $H$ Diffusion.  With probability $D_H\, H/\mathcal{R}$, pick a random
+\item Diffusion.  With probability $D_H\, H/\mathcal{R}$, pick a random
   hungry forager and move it to a neighboring site.

### 2016-07-21 Manuscript/Simulation_description/simulation-new.tex
cleaning up old files

--- 2016-01-13 Manuscript/simulation.tex
@@
 \end{align}
-where the overdot denotes time derivative.
+where the overdot denotes time derivative.  An important feature is the
+assumption that the carrying capacity is set equal to 1.  This means that
+when the system is completely occupied by resources there can be no further
+generation of the resource.  The approach described below can be generalized
+to carrying capacity less than 1, but one can take the carrying capacity
+equal to 1 without loss of generality.
 
-We now outline an event-driven algorithm that mimic these rate equations.  We
-will also generalize to the situation where the full and hungry foragers
-undergo diffusion with possibly different diffusion coefficients on a
-finite-dimensional lattice.  Suppose that the system at some time consists of
-$N_F$ full foragers, $N_H$ hungry foragers, and $N_R$ individual resources.
-The total number of particles $N=N_F+N_H+N_R$.
+We now outline an event-driven algorithm that mimic these rate equations.
+This algorithm is based on picking \emph{sites} at random, rather than
+\emph{aprticles} at random.  However, we show how to recast this site-based
+simulation into a particle-based simulation with an concomitant increase in
+efficiency.  We will also generalize to the situation where the full and
+hungry foragers undergo diffusion with possibly different diffusion
+coefficients on a finite-dimensional lattice.
 
-In each elemental event an individual is picked: a full forager with
-probability $N_F/N$, a hungry forager with probability $N_H/N$, and a
-resource with probability $N_R/N$.  If an $F$ is picked, it reproduces at
-rate $\lambda$ and becomes hungry with rate $\sigma(1-R)$.  If an $H$ is
-picked, it becomes full with rate $\rho R$ and dies with rate $\mu$.
-Finally, if an $R$ is picked, it grows with rate $\alpha(1-R)$ and is eaten
-with rate $(F+H)$.  Thus the total rate for each elemental event is
+Suppose that the system at some time consists of $N_F$ full foragers, $N_H$
+hungry foragers, and $N_R$ individual resources.  The total number of
+particles $N=N_F+N_H+N_R$.  In each elemental event a lattice site is picked:
+on average, it is occupied by a full forager with probability $N_F/V$, a
+hungry forager with probability $N_H/V$, a resource with probability $N_R/V$,
+or, on average, the site may be empty with probability $1-\rho$, where
+$\rho=(N_F+N_H+N_R)/V$ is the total density of particles, on average and $V$
+is the total number of lattice sites..  I use the term ``on average'' because
+it is possible to have rare realizations where all particles are concentrated
+on just a few sites.  However, for large systems, the actually number of
+particles at a given site will not signficantly deviate from the average
+number.
+
+If a site that is occupied by an $F$ is picked, it reproduces at rate
+$\lambda$ and becomes hungry with rate $\sigma(1-R)$.  If a site that is
+occupied by an $H$ is picked, it becomes full with rate $\rho R$ and dies
+with rate $\mu$.  Finally, if a site that is occupied by an $R$ is picked, it
+grows with rate $\alpha(1-R)$ and is eaten with rate $(F+H)$.  Thus the total
+rate for each elemental event is
 \begin{equation}
 \mathcal{R}=F\big[\lambda +\sigma(1-R)\big]+ H\big[\rho R+\mu\big]+
-R\big[\alpha(1-R)+(F+H)\big]\,.
+R\big[\alpha(1-R)+(F+H)\big]+(1-\rho)\,.
 \end{equation}
@@
 \begin{equation}
-  \Delta N_F =\left[\frac{N_F}{N}\big(\lambda-\sigma(1-R)\big)+\frac{N_H}{N}\rho
+  \Delta N_F =\left[\frac{N_F}{V}\big(\lambda-\sigma(1-R)\big)+\frac{N_H}{V}\rho
       R\right]/\mathcal{R}\,.
@@
 \begin{equation}
-  \Delta F =\left[F\big(\lambda-\sigma(1-R)\big)+H\rho
-    R\right]/N\mathcal{R}\,.
+  \Delta F =\frac{\Delta N_F}{V} =\left[F\big(\lambda-\sigma(1-R)\big)+H\rho
+    R\right]/V\mathcal{R}\,.
 \end{equation}
@@
 Thus if we take the time step for each event to be
-$\Delta t = (N\mathcal{R})^{-1}$, the above reduces to the rate equation
+$\Delta t = (V\mathcal{R})^{-1}$, the above reduces to the rate equation
 \eqref{RE} for $F$.  Thus in each microscopic event of the model, the time
-should be advanced by $\Delta t = (N\mathcal{R})^{-1}$.
+should be advanced by $\Delta t = (V\mathcal{R})^{-1}$.
 
@@
 \begin{equation}
-  \Delta N_H =\left[-\frac{N_H}{N}(\rho R+\mu)+\frac{N_F}{N}\sigma
+  \Delta N_H =\left[-\frac{N_H}{V}(\rho R+\mu)+\frac{N_F}{V}\sigma
     (1-R)\right]/\mathcal{R}\,,
@@
 \begin{equation}
-  \Delta H =\left[-H(\rho R+\mu)+F\sigma (1-R)\right]/N\mathcal{R}\,.
+  \Delta H =\left[-H(\rho R+\mu)+F\sigma (1-R)\right]/V\mathcal{R}\,.
 \end{equation}
@@
 \begin{equation}
-  \Delta N_R =\left[\frac{N_R}{N}\big(\alpha(1-R)\big)-\frac{(N_F+N_H)}{N}\right]/\mathcal{R}\,,
+  \Delta N_R =\left[\frac{N_R}{V}\big(\alpha(1-R)\big)-\frac{(N_F+N_H)}{V}\right]/\mathcal{R}\,,
 \end{equation}
@@
 \begin{equation}
-  \Delta R =\left[R\big(\alpha(1-R)\big)-(F+H)\right]/N\mathcal{R}\,.
+  \Delta R =\left[R\big(\alpha(1-R)\big)-(F+H)\right]/V\mathcal{R}\,.
 \end{equation}
 \end{subequations}
-The equations for $\Delta F$, $\Delta H$, and $\Delta R$ then reproduce the
-original rate equations \eqref{RE} when the time step for an elemental event
-is taken to be $\Delta t = (N\mathcal{R})^{-1}$.
+The equations for $\Delta F$, $\Delta H$, and $\Delta R$ would then reproduce
+the original rate equations \eqref{RE} when the time step for an elemental
+event is taken to be $\Delta t = (V\mathcal{R})^{-1}$.
+
+However, this approach is inefficient if many of the lattice sites are empty
+because an empty site is likely to be picked and then nothing happens.
+Instead, we would rather pick \emph{actual particles} rather than
+\emph{lattice sites}.  If the particle density is $\rho=N/V$, then in picking
+a \emph{particle} at random always leads to a non-trivial update, while
+picking a \emph{site} at random leads to a non-trivial update with
+probability $\rho$.  Thus by picking \emph{particles} at random---an $F$ with
+probability $N_F/N$, an $H$ with probability $N_H/N$, or an $R$ with
+probability $N_R/N$---compared to picking \emph{sites} at random, leads to an
+overall speedup of the simulation by a factor of $1/\rho$.  Therefore, for
+the particle-based upate, time increment changes to
+$\Delta t = (1/\rho)\times (V\mathcal{R})^{-1}= (N\mathcal{R})^{-1}$.
+
+To summarize, pick a particle of type $i$ with $N_i/N$, and update the
+densities $N_i/V$ according (3b), (4b), and (5b), and then increment the time
+by $(N\mathcal{R})^{-1}$ after each event.  It is important to appreciate
+that this appoach implicitly assumes that the number of particles on each
+site is small.  This assumption is reasonable if we are describing a harsh
+environment where a non-negligible fraction of the foragers are either going
+hungry or starving.
 
@@
 \mathcal{R}_d=F\big[\lambda +\sigma(1-R)+D_F\big]+ H\big[\rho R+\mu+D_H\big]+
-R\big[\alpha(1-r)+(F+H)\big]\,.
+R\big[\alpha(1-r)+(F+H)\big]+(1-\rho)\,.
 \end{equation}
@@
 placed anywhere.  For a finite-dimensional lattice, it seems reasonable to
-place the offspring at the same site as the parent.  Diffusion of a full
-forager on a lattice can be modeled by allowing it to hop to a nearest
-neighbor; similarly for the motion of hungry foragers.
+place the offspring at the same site as the parent.  However, because the
+implicity assumption of low density, perhaps it would be better to put the
+newly born forager on the closest lattice site that is not occupied by
+another forager.  Diffusion of a full forager on a lattice can be modeled by
+allowing it to hop to a nearest neighbor; similarly for the motion of hungry
+foragers.
 

### 2016-07-22 Manuscript/1.tex
first edits from Sid

--- 2016-07-21 Manuscript/Yeakel_Kempes-Redner_PNAS_7-19.tex
@@
 %\usepackage{scicite}
-
 %\usepackage{times}
@@
 %\usepackage{rotating}
-%\usepackage{color}
 %\usepackage{layout}
@@
 
-\usepackage{amssymb,amsfonts,amsmath}
+%\usepackage{amssymb,amsfonts,amsmath}
+\usepackage{amsmath}
+
 \usepackage[font=large]{caption}
@@
 \usepackage{multirow}
+\usepackage{color}
 %\usepackage{multibox}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newcommand{\sid}[1]{\textcolor{red}{\bf [#1]}}
+
 %% Don't type in anything in the following section:
@@
 
+%\title{Survival of the Stable: Dynamic Consequence of Starvation and Reproduction}
+
+%\title{Size Constraints on Starvation/Reproduction Dynamics}
+
 \author
@@
 \begin{abstract}
-This is the abstract.
+This is the abstract.  \emph{ No it isn't.  It's merely a placeholder.}
 \end{abstract}
+\keywords{foraging | starvation|reproduction}
 
-%\section*{Introduction}
-The behavioral ecology of most, if not all, organisms is influenced by the energetic state of individuals, which directly influences how organisms invest reserves in uncertain environments.
-Such behaviors are generally manifested as trade-offs between investing in somatic maintenance and growth or allocating energy towards reproduction \cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
-%among a host of other behavioral duties [REFS].
-The timing of these behaviors is often important and is under strong selective pressure, as it tends to directly impact future fitness \cite{Mangel:1988uaa}.
-%For example, rotifers that reproduce have lower survival rates than those that don't when they are under nutritional stress.
-%To what extent, and when, organisms invest in somatic vs.
-reproductive expenditures may be driven by habitat, seasonality, evolutionary history, inter- or intra-specific interactions, and the distribution of resources.
-Importantly, the influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
-%and the subsequent loss of nutritional reserves
-%Such tradeoffs are also bound to have population-level consequences, although this has received little theoretical or empirical attention.
+\noindent {\bf Introduction} \\ \nonumber
+The behavioral ecology of most, if not all, organisms is influenced by the
+energetic state of individuals, which directly influences how organisms
+invest reserves in uncertain environments.  Such behaviors are generally
+manifested as trade-offs between investing in somatic maintenance and growth
+or allocating energy towards reproduction
+\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}. 
+The timing of these behaviors responds to selective pressure, as the choice
+of the investment impacts future fitness!\cite{Mangel:1988uaa}.
+The influence of resource limitation on an organism's ability to maintain its
+nutritional stores may lead to repeated delays or shifts in reproduction over
+the course of an organism's life.
 
-%The reproductive vs. storage tradeoff
-Maximizing fitness between growth and maintenance activities vs.
-reproduction structures the life-history of many species, and this can be achieved by alternative behavioral strategies conditioned on resource availability \cite{Morris:1987eo}.
-%or by changes in physiology, both of which are sensitive to
-%Reproduction can be separated from somatic growth/maintenance by a variety of potential mechanisms:
-%Multiple mechanisms contribute to the differentiation of these behaviors:
-%Behavioral changes in somatic or reproductive investment often occur in response to limited resources \cite{Morris:1987eo}.
-%, while stress can generally lead to reprodusuppression of reproductive behaviors \cite{Morgan:1999do}. 
-For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is poor) than in calves born after moderate winters \cite{Tveraa:2003fq}, whereas many bird species invest differently in broods during periods of resource scarcity \cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or foregoing reproduction for a breeding season \cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
-%Such breeding behaviors is generally referred to as \emph{capital breeding}
-%Resource limitation can also alter the behaviors of species in well-mixed environments: 
-Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress \cite{Threlkeld:1976ih}, with those that do reproduce have lower survival rates \cite{Kirk:1997cc}.
-%, while artificially induced stress has been observed to decrease reproductive success in Atlantic cod \cite{Morgan:1999do}.
-Organisms may also separate maintenance and growth from reproduction over space and time: 
-many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in alternative environments spent accumulating nutritional reserves \cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+The life history of most species is typically comprised of (a) somatic growth
+and maintenance and (b) reproduction.  The balance between these two
+activities is often conditioned on resource
+availability~\cite{Morris:1987eo}.  For example, reindeer invest less in
+calves born after harsh winters (when the mother's energetic state is
+depleted) than in calves born after moderate winters \cite{Tveraa:2003fq}.
+Many bird species invest differently in broods during periods of resource
+scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv},
+sometimes delaying or even foregoing reproduction for a breeding
+season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.  Even freshwater
+and marine zooplankton have been observed to avoid reproduction under
+nutritional stress \cite{Threlkeld:1976ih}, and those that do reproduce have
+lower survival rates~\cite{Kirk:1997cc}.  Organisms may also separate
+maintenance and growth from reproduction over space and time: many salmonids,
+birds, and some mammals return to migratory breeding grounds to reproduce
+after one or multiple seasons in resource-rich environments where they
+accumulate nutritional
+reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
 
-Physiological mechanisms also play an important role in regulating reproductive expenditures during periods of resource limitation.
-Diverse mammals (47 species in 10 families) exhibit delayed implantation whereby females postpone fetal development (blastocyst implantation) to time with accumulation of nutritional reserves \cite{Mead:1989dt,Sandell:1990kw}, while many others (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress \cite{Bulik:1999eo,Trites:2003cc}.
-In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.
-%great tits invest more energy into broods when resource abundance is greater, while food availability is generally seen to correlate with clutch size among many bird species \cite{Daan:1988va}.
-%The mechanisms that control reproductive effort as a function of energetic reserves differ among species.
-%Even humans alter their growth trajectories and have higher chances of fetal abortion during periods of starvation.
-%Thus, some organisms partition reproductive vs.
-foraging effort over both space and time, whereas other rely on physiological mechanisms to optimize reproductive success against survival when adequate resources are unavailable.
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms points to the importance and universality of the fundamental tradeoff between somatic and reproductive investment, however the dynamic implications of these constraints are unknown.
+Physiological mechanisms also play an important role in regulating
+reproductive expenditures during periods of resource limitation.  Diverse
+mammals (47 species in 10 families) exhibit delayed implantation, whereby
+females postpone fetal development (blastocyst implantation) until times
+where nutritional reserves can be
+accumulated~\cite{Mead:1989dt,Sandell:1990kw}.  Many other many species
+(including humans) suffer irregular menstrual cycling and higher abortion
+rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
+In the extreme case of unicellular organisms, nutrition is unavoidably linked
+to reproduction because the nutritional state of the cell regulates all
+aspects of the cell cycle \cite{Glazier:2009hq}.  The existence of so many
+independently evolved mechanisms across such a diverse suite of organisms
+highlights the importance and universality of the fundamental tradeoff
+between somatic and reproductive investment.  However the dynamic
+implications of these constraints are unknown.
 
-%%To population-level effects
-%The mechanisms by which different organisms avoid or delay reproduction during times of nutritional stress has received tremendous empirical and theoretical attention owing to the importance of these activities in shaping life-history \cite{Cody:1966fk,Martin:1987dl,Mangel:1988uaa}.
-%Less well understood is how resource limitation and these behavioral/physiological tradeoffs affect dynamics at the level of the population.
-%% when organisms are able to avoid the risk of spending energy on reproduction when starvation is likely.
-%The influence of energetic constraints -- and their consequent effects on reproduction -- on population dynamics is not well understood, despite considerable effort in 
-%Traditional Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction \cite{murdoch:2003}.
-%Although this implicit dependence connects resource limitation to lower consumer growth rates, the following biological realities are not included: 
-%\emph{i}) some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation;
-%\emph{ii}) the portion of the population that is not nutritionally stressed is expected to reproduce at a near-constant rate and this is -- averaged across species -- determined strongly by body size \cite{Kempes:2012hy};
-%\emph{iii}) the rates that individuals transition from nutritionally poor to replete states and back have different metabolically-constrained timescales that can lead to reproductive lags.
-%Importantly, the exclusion of these biological details may have important dynamic shortcomings, masking the effects of starvation on consumer population dynamics.
+Though straightforward conceptually, incorporating the energetic dynamics of
+individuals~\cite{Kooi2000} into a population-level
+framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical
+obstacles~\cite{Diekmann:2010da}).  An alternative approach involves modeling
+the macroscale relations that guide somatic versus reproductive investment in
+a consumer-resource system.  For example, macroscale Lotka-Volterra models
+assume that the growth rate of the consumer population depends on resource
+density, thus \emph{implicitly} incorporating the requirement of resource
+availability for reproduction \cite{murdoch:2003}.
 
-Though straightforward conceptually, incorporating the energetic dynamics of individuals \cite{Kooi2000} into a population-level framework \cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles (in particular a lack of smoothness or differentiability \cite{Diekmann:2010da}), and is prone to over-fitting.
-%These issues have limited the development of theoretical models that may aid our understanding of the effects of such tradeoffs on population dynamics.
-An alternative approach involves modeling the macroscale relationships that guide somatic vs.
-reproductive investment in a consumer-resource system.
-%to individual-to-population frameworks is redirecting focus to
-Macroscale Lotka-Volterra models assume a dependence of consumer population growth rates on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction \cite{murdoch:2003}.
-Resource limitation and the subsequent effects of starvation may be alternatively accounted for \emph{explicitly}, such that reproduction is permitted only for individuals with sufficient energetic reserves.
-Such a dynamic introduces 
-\emph{i}) the reproductive time lag associated with changing rates of starvation and recovery, and 
-\emph{ii}) the idea that reproduction is strongly allometrically constrained \cite{Kempes:2012hy}, and not generally linearly related to resource density.
-%Incorporating explicit starvation introduces
-%\emph{i}) some individuals experience nutritional stress at a given time and under a given set of external conditions, while others do not; those that do have multiple pathways enabling reproductive cessation;
-%\emph{ii}) the portion of the population that is not nutritionally stressed is expected to reproduce at a near-constant rate and this is -- averaged across species -- determined strongly by body size \cite{Kempes:2012hy};
-%\emph{iii}) the rates that individuals transition from nutritionally poor to replete states and back have different metabolically-constrained timescales that can lead to reproductive lags.
+In this work, we adopt an alternative approach in which resource limitation
+and the subsequent effect of starvation is accounted for \emph{explicitly}.
+Namely, only individuals with sufficient energetic reserves can reproduce.
+Such a constraint leads to reproductive time lags due to some members of the
+population starving and then recovering.  Additonally, we incorporate the
+idea that reproduction is strongly constrained
+allometrically~\cite{Kempes:2012hy}, and is not generally linearly related to
+resource density.  As we shall show, these constraints influence the ensuing
+population dynamics in dramatic ways.
+\\
 
-%[Here we use body size and allometries to capture a wide variety of taxonomic diversity]
-%[There isn't a deep understanding with how models vary across different life histories, traits, etc... how do equations change with allometric constraints for different class of organisms? 
-%i.e. Use body size as a generating parameter to capture a wide variety of diversity.]
+\noindent {\bf Nutritional-state-structured model (NSM)}
 
-%CONCLUSIONS
-%We show that rates of starvation and recovery tend to result in systems with stable, non-cyclic fixed points and that the ratio between the rate of starvation and recovery may -- in part -- contribute to lower size limitations.
-%Moreover, larger consumer body size results in rates that are less prone to cyclic dynamics than are rates for smaller organisms, pointing to potential empirical verification of the NSM framework.
-%Finally, we show that rates of starvation and recovery appear to be constrained to a parameter range where both transient and equilibrial population dynamics result in the lowest risk of extinction for the consumer.
-%This surprising result suggests that the risks associated with different fluxes of consumers in and out of a starved (non-reproductive) state may serve as an important selective driver over evolutionary time.
+We begin by defining a minimal Nutritional-State-structured population Model
+(NSM), where the consumer population is divided into two energetic states:
+(a) an energetically replete (full) state $F$, where the consumer reproduces
+at a constant rate $\lambda$, and (b) an energetically deficient (hungry)
+state $H$, where the consumer does not reproduce but dies at rate $\mu$.  The
+underlying resource $R$ evolves by logistic growth with an intrinsic growth
+rate $\alpha$ and a carrying capacity equal to one.  Consumers transition
+from the full state $F$ to the hungry state $H$ by starvation at rate
+$\sigma$ and also in proportion to the absence of resources $(1-R)$.
+Conversely, consumers recover from state $H$ to state $F$ at rate $\rho$ and
+in proportion to $R$.  Resources are also eaten by the consumers---at rate
+$\rho$ by hungry consumers and at rate $\beta<\rho$ by full consumers.  This
+inequality accounts for hungry consumers requiring more resources to rebuild
+body weight.
 
-\noindent {\bf Nutritional state-structured model (NSM)} \\ \nonumber
-%What we did...
-We explore how the energetic tradeoff between somatic maintenance and growth vs. reproduction can influence population dynamics, and how such dynamics may be constrained by allometry.
-We begin by establishing a minimal Nutritional State-structured population Model (NSM), where the consumer population is divided into two energetic states: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
-%consumer starvation and recovery is contingent on resource density, the starved portion of the population does not reproduce, and the rate at which consumers decline and recover from starvation are functions of metabolic rate.
-%By relating different rate constants to allometry, we uncover important relationships between the timescales of physiological and reproductive processes, and show how organisms of different body sizes may be prone to alternative dynamic regimes.
-%We integrate the somatic/reproductive tradeoff into the dynamics of a consumer-resource system by dividing the consumer population into two discrete energetic states, the occupation of each being contingent on the consumption of a single resource $R$.
-%In the NSM there are only two energetic states for the consumer population: \emph{i}) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$, and \emph{ii}) an energetically deficient (hungry) state $H$, where reproduction is suppressed, and mortality occurs at rate $\mu$.
-The resource $R$ has logistic growth with an intrinsic growth rate $\alpha$ and carrying capacity of unity. %(throughout we set $\kappa=1$ such that $R$ has units of $1/\kappa$).
-Consumers transition from state $F$ to state $H$ by starvation at rate $\sigma$ and in proportion to the lack of resources $(1-R)$.
-Conversely, consumers recover from state $H$ to the full state $F$ at rate $\rho$ and in proportion to $R$.
-Resources are eliminated by the consumer in both states: by energetically deficient consumers at rate $\rho$, and by energetically replete consumers at rate $\beta$.
-Accordingly, the system of equations is written
-
-\begin{equation} \label{eq:system}
-\begin{aligned}
+In the mean-field approximation, in which the consumers and resources are
+perfectly mixed, their densities evolve according to the rate equations
+\begin{eqnarray} 
+\label{eq:system}
+\begin{split}
 \dot{F} &= \lambda F + \rho RH - \sigma (1-R)F,  \\
@@
 \dot{R} &= \alpha R(1-R) - R(\rho H+ \beta F).
-\end{aligned}
-\end{equation}
+\end{split}
+\end{eqnarray}
+Notice that the total consumer density $F+H$ evolves according to
+$\dot{F}+\dot{H}=\lambda F-\mu H$.  This resmebles the equation of motion for
+the predator density in the classic Lotka-Volterra model, except that the
+resource density does not appear in the growth term.  As discussed above, the
+attributes of reproduction and mortality have been explicitly apportioned to
+the full and hungry consumers, respectively, so that the growth in the total
+density is decoupled from the resource density.
 
-%[NSM = Nutritional State Model]
-%[LVM = Lotka Volterra Model]
-
-There are three fixed points associated with the NSM: two trivial fixed points at $(R^*=0,H^*=0,F^*=0)$ and $(R^*=1,H^*=0,F^*=0)$, and one internal fixed point at
-
-\begin{equation} \label{eq:ss}
-\begin{aligned}
+Equation~\eqref{eq:system} has three fixed points: two trivial fixed points
+at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal
+fixed point at
+\begin{eqnarray}
+\label{eq:ss}
+\begin{split}
 F^* &= \frac{\alpha  \lambda  \mu  (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  \beta)}, \\
@@
 R^* &= \frac{\mu  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma }.
-\end{aligned}
-\end{equation}
+\end{split}
+\end{eqnarray}
+If this unique internal fixed point is stable, it will be the global
+attractor for all population trajectories for any set of initial conditions
+where the resource and consumer densities are both non zero.  Stability of
+the fixed point is determined by the Jacobian Matrix $\bf J$, where each
+matrix element $J_{ij}$ equals $\partial{\dot X_i}/\partial{X_j}$ when
+evaluated at the internal fixed point, and $\mathbf{X}$ is the vector
+$(F,H,R)$.  If the parameters in Eq.~\eqref{eq:system} are such that the real
+part of the largest eigenvalue of $\mathbf{J}$ is negative, then the system
+is stable with respect to small perturbations from the fixed point.
 
-\noindent Because there is only one internal fixed point, as long as it is stable the population trajectories will serve as a global attractor for any set of initial conditions $>0$.
-%Analysis of the stability of the NSM is explored with respect to the local stability of the internal steady state, which is the only feasible steady state as long as both the consumer and resource have non-zero, positive, values.
-Stability is determined with respect to the Jacobian Matrix $\bf J|_*$ (where $|_*$ denotes evaluation at the internal steady state), where each element of the matrix is defined by the partial derivative of each differential equation with respect to each variable.
-%General statement on stability
-If the parameters of $\bf J|_*$ are such that the real part of its leading eigenvalue is $<0$, then the system is stable to small pulse perturbations.
+A fundamental constraint on the NSM is that the reproduction rate $\lambda$
+must be less than the starvation rate $\sigma$.  If $\lambda>\sigma$, the
+fixed point value of $R^*$ has an unphysical negative value.  Moreover, when
+the resource density $R=0$, the rate equation for $F$ gives exponential
+growth for $\lambda>\sigma$.  The condition $\sigma = \lambda$ represents a
+transcritical (TC) bifurcation that demarcates the physical and unphysical
+regimes.  The biological meaning of the constraint $\lambda<\sigma$ is
+simple---the rate at which a macroscopic organism loses mass due to a lack of
+resources is generally much faster than the rate of reproduction.  As we will
+discuss below, this inequality is a natural consequence of allometric
+constraints~\cite{Kempes:2012hy} for organisms within observed body size
+ranges (Fig.~\ref{fig:gvs}).
+
+In the physical regime where $\lambda<\sigma$, the fixed point \eqref{eq:ss}
+may either be a stable node or a stable spiral node.
+
+In addition to the bound defined by the TC bifurcation, oscillating or cyclic
+dynamics present an implicit constraint to persistence by increasing
+extinction risk due to stochastic effects.  In continuous-time systems, a
+stable limit cycle arises when a pair of complex conjugate eigenvalues
+crosses the imaginary axis to attain positive real parts \cite{GuckHolmes}.
+This condition, known as a Hopf bifurcation, is defined by
+${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is
+composed of the coefficients of the characteristic polynomial describing the
+Jacobian \cite{Gross:2004p2428}.
+
+%General transient dynamics statement
+Moreover, as a non-cyclic stable system nears the Hopf bifurcation, transient
+or decaying cycles can grow in magnitude, despite the existence of a
+positive, non-cyclic, steady state density.  Given that ecological systems
+exist in a state of constant perturbation \cite{Hastings:2001jh}, even the
+onset of transient cycles that decay over time can increase the risk of
+extinction \cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}, such that the
+distance of a system from the Hopf bifurcation is relevant to persistence.
+
+The NSM exhibits both non-cyclic as well as cyclic dynamics, and which
+behavior dominates depends strongly on the rate of starvation $\sigma$
+relative to the rate of recovery $\rho$.  Although starvation leads to
+mortality risk, a moderate amount promotes persistence of both consumer and
+resource populations as non-cyclic stability of the fixed point generally
+requires a higher $\sigma$ relative to $\rho$.  The intuition behind this is
+that transition to the hungry (non-reproductive) state permits the resource
+to recover and transient dynamics to subside, whereas a low $\sigma$
+overloads the system with energetically-replete (reproducing) individuals,
+maintaining oscillations between consumer and resource (Fig. \ref{fig:fp}).
+If $\sigma$ is too large, mortality due to starvation depletes the consumer
+population, resulting in a lower steady state density for the consumer and a
+higher steady state density for the resource.
+
+Whereas the rate of consumer growth defines a hard bound of biological
+feasibility (the TC bifurcation), the rate of starvation thus determines the
+sensitivity of the consumer population to changes in resource density.  While
+higher rates of starvation result in lower steady state population size --
+increasing the risk of stochastic extinction -- lower rates of starvation
+result in a system poised near either the TC or Hopf bifurcation, which will
+lead to elimination of the resource or the development of cyclic
+oscillations, respectively.  Which bifurcation is approached is wholly
+dependent on the rate of recovery: if it is high, then cyclic dynamics will
+develop; if it is low, resource extinction becomes increasingly likely.
 

### 2016-07-25 Manuscript/1.tex
Kempes edits to v2

@@
 \end{eqnarray}
-If this unique internal fixed point is stable, it will be the global
-attractor for all population trajectories for any set of initial conditions
-where the resource and consumer densities are both non zero.  Stability of
-the fixed point is determined by the Jacobian Matrix $\bf J$, where each
+If this internal fixed point, which is unique, is stable, it will be the
+global attractor for all population trajectories for any initial condition
+where the resource and consumer densities are both non zero.  The stability
+of this fixed point is determined by the Jacobian Matrix $\bf J$, where each
 matrix element $J_{ij}$ equals $\partial{\dot X_i}/\partial{X_j}$ when
@@
 
-A fundamental constraint on the NSM is that the reproduction rate $\lambda$
-must be less than the starvation rate $\sigma$.  If $\lambda>\sigma$, the
-fixed point value of $R^*$ has an unphysical negative value.  Moreover, when
-the resource density $R=0$, the rate equation for $F$ gives exponential
-growth for $\lambda>\sigma$.  The condition $\sigma = \lambda$ represents a
-transcritical (TC) bifurcation that demarcates the physical and unphysical
-regimes.  The biological meaning of the constraint $\lambda<\sigma$ is
-simple---the rate at which a macroscopic organism loses mass due to a lack of
-resources is generally much faster than the rate of reproduction.  As we will
-discuss below, this inequality is a natural consequence of allometric
-constraints~\cite{Kempes:2012hy} for organisms within observed body size
-ranges (Fig.~\ref{fig:gvs}).
+From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the
+reproduction rate $\lambda$ must be less than the starvation rate $\sigma$,
+so that $R^*$ is positive.  Moreover, when the resource density $R=0$, the
+rate equation for $F$ gives exponential growth for $\lambda>\sigma$.  The
+condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation that
+demarcates the physical and unphysical regimes.  The biological implication
+of the constraint $\lambda<\sigma$ is simple---the rate at which a
+macroscopic organism loses mass due to lack of resources is generally much
+faster than the rate of reproduction.  As we will discuss below, this
+inequality is a natural consequence of allometric
+constraints~\cite{Kempes:2012hy} for organisms within empirically observed
+body size ranges (Fig.~\ref{fig:gvs}).
 
-In the physical regime where $\lambda<\sigma$, the fixed point \eqref{eq:ss}
-may either be a stable node or a stable spiral node.
+In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may
+either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).  In
+continuous-time systems, a limit cycle arises when a pair of complex
+conjugate eigenvalues crosses the imaginary axis to attain positive real
+parts~\cite{GuckHolmes}.  This Hopf bifurcation is defined by
+${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is
+composed of the coefficients of the characteristic polynomial of the Jacobian
+matrix~\cite{Gross:2004p2428}.  As the system parameters are tuned to be
+within the stable regime but close to the Hopf bifurcation, the amplitude of
+the transient but decaying cycles become large.  Given that ecological
+systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of
+transient cycles, even though they decay with time, can increase the
+extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.  Thus
+the distance of a system from the Hopf bifurcation provides a measure of its
+persistence.
 
-In addition to the bound defined by the TC bifurcation, oscillating or cyclic
-dynamics present an implicit constraint to persistence by increasing
-extinction risk due to stochastic effects.  In continuous-time systems, a
-stable limit cycle arises when a pair of complex conjugate eigenvalues
-crosses the imaginary axis to attain positive real parts \cite{GuckHolmes}.
-This condition, known as a Hopf bifurcation, is defined by
-${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is
-composed of the coefficients of the characteristic polynomial describing the
-Jacobian \cite{Gross:2004p2428}.
+When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the
+consumers are driven to the hungry non-reproductive state.  Because
+reproduction is inhibited, there is a low steady-state consumer density and a
+high steady-state resource density.  However, if $\sigma/\lambda$ approaches
+one, the population is overloaded with energetically-replete (reproducing)
+individuals, thereby promoting oscillations between the consumer and resource
+densities (Fig.~\ref{fig:fp}).
 
-%General transient dynamics statement
-Moreover, as a non-cyclic stable system nears the Hopf bifurcation, transient
-or decaying cycles can grow in magnitude, despite the existence of a
-positive, non-cyclic, steady state density.  Given that ecological systems
-exist in a state of constant perturbation \cite{Hastings:2001jh}, even the
-onset of transient cycles that decay over time can increase the risk of
-extinction \cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}, such that the
-distance of a system from the Hopf bifurcation is relevant to persistence.
-
-The NSM exhibits both non-cyclic as well as cyclic dynamics, and which
-behavior dominates depends strongly on the rate of starvation $\sigma$
-relative to the rate of recovery $\rho$.  Although starvation leads to
-mortality risk, a moderate amount promotes persistence of both consumer and
-resource populations as non-cyclic stability of the fixed point generally
-requires a higher $\sigma$ relative to $\rho$.  The intuition behind this is
-that transition to the hungry (non-reproductive) state permits the resource
-to recover and transient dynamics to subside, whereas a low $\sigma$
-overloads the system with energetically-replete (reproducing) individuals,
-maintaining oscillations between consumer and resource (Fig. \ref{fig:fp}).
-If $\sigma$ is too large, mortality due to starvation depletes the consumer
-population, resulting in a lower steady state density for the consumer and a
-higher steady state density for the resource.
-
-Whereas the rate of consumer growth defines a hard bound of biological
-feasibility (the TC bifurcation), the rate of starvation thus determines the
-sensitivity of the consumer population to changes in resource density.  While
-higher rates of starvation result in lower steady state population size --
-increasing the risk of stochastic extinction -- lower rates of starvation
-result in a system poised near either the TC or Hopf bifurcation, which will
-lead to elimination of the resource or the development of cyclic
-oscillations, respectively.  Which bifurcation is approached is wholly
-dependent on the rate of recovery: if it is high, then cyclic dynamics will
-develop; if it is low, resource extinction becomes increasingly likely.
+Whereas the consumer growth rate $\lambda$ defines an absolute bound of
+biological feasibility---the TC bifurcation---the starvation rate $\sigma$
+determines the sensitivity of the consumer population to changes in resource
+density.  When $\sigma\gg\lambda$, the steady-state population density is
+small, thereby increasing the risk of stochastic extinction.  Conversely, as
+$\sigma$ decreases, the system will ultimately be poised either near the TC
+or the Hopf bifurcation (Fig.~\ref{fig:fp}).  If the recovery rate $\rho$ is
+sufficiently small, the TC bifurcation is reached and the resource eventually
+is eliminated.  If $\rho$ exceeds a threshold value, cyclic dynamics will
+develop as the Hopf bifurcation is approached.
 
@@
 \vspace{2mm}
-\noindent {\bf Allometric rates} \\ \nonumber
+\noindent {\bf Role of allometry} \\ 
 %[Link Allometry stuff to our model - what does it provide to the story]
@@
 This relationship is derived from the simple balance
- \begin{equation}
+\begin{eqnarray}
+\label{balance}
 B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m
- \label{balance}
- \end{equation}
+\end{eqnarray}
  [a and b notation --- these parameters are easily measured bioenergetic parameters which are often approximately invariant across organisms of vastly different size.
@@
 This assumption implies the simple metabolic balance
-\begin{equation}
+\begin{eqnarray}
 \frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
-\end{equation}
+\end{eqnarray}
 where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing body mass which may differ from  $E_{m}$, the energy required to synthesis a unit of biomass.
 Give the adult mass, $M$, of an organism this energy balance prescribes the mass trajectory of a starving organism:
-\begin{equation}
+\begin{eqnarray}
 m\left(t\right)=Me^{-B_{m}t/E_{m}^{\prime}}.
-\end{equation}
+\end{eqnarray}
 Considering that only certain tissues can be digested for energy, for example the brain cannot be degraded to fuel metabolism, we define the rate for starvation and death by the timescales required to reach specific fractions of normal adult mass.
@@
 Thus
-\begin{equation}
+\begin{eqnarray}
 t_{\mu}=-\frac{E_{m}\log\left(\epsilon^{\prime}\right)}{B_{m}}
-\end{equation}
+\end{eqnarray}
 and $\mu=1/t_{\mu}$.
@@
 The general solution to Equation \ref{balance} is given by
-\begin{equation}
+\begin{eqnarray}
 m\left(t\right)=c\left[1-\left(1-\frac{b}{a}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
-\end{equation}
+\end{eqnarray}
 with $a=B_{0}/E_{m}$, $b=B_{m}/E_{m}$, and $c=(a/b)^{1/(\eta-1)}$.
 We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, which has the final form of
-\begin{equation}
+\begin{eqnarray}
 t_{\rho}=\frac{\log \left(1-\left(cM \right)^{1-\eta }\right)-\log \left(1-\left(c\epsilon M \right)^{1-\eta }\right)}{(\eta -1) b}.
-\end{equation}
+\end{eqnarray}
 %\left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}
@@
 
-%Population growth requires that individuals
-
-%\begin{align}
-%\lamdba &=  \nonumber \\
-%\sigma &= \nonumber \\	
-%\end{align}
-
 \vspace{2mm}
 \noindent {\bf The stabilizing effects of allometric constraints} \\ \noindent
-%Transcritical bifurcation
-Stability in the NSM is conditioned on the consumer's starvation rate $\sigma$ relative to its reproduction rate $\lambda$.
-If $\sigma<\lambda$, the resource steady state density is negative and extinction is inevitable.
-The condition $\sigma = \lambda$ is a transcritical (TC) bifurcation, thus marking a hard boundary below which the system is unphysical due to the unregulated growth of the consumer population.
-That the timescale of reproduction is larger than the timescale of starvation is intuitive for macroscopic organisms, as the rate at which one loses tissue due to a lack of resources is generally much faster than reproduction.
-In fact, allometric derivations for both reproduction \cite{Kempes:2012hy} and starvation (Eq.
-\ref{eq:sigma}) show that this relationship always holds for organisms within observed body size ranges (Fig.
-\ref{fig:gvs}).
-%We note that the asymptote for the starvation rate at $M \approx 8\times10^8$ defines the mass at which body fat accounts for 100\% of organismal weight, thereby placing a scaling bound on our derivation for the starvation rate.
-%According to this derivation, the endotherm (the blue whale \emph{}) is 74\% fat, which accords with empirical estimates REF.
-
-%
-%%Transcritical & rate law ranges
-%The TC bifurcation occurs in this model because we have assumed that the portion of the population that is not starved reproduces at a constant rate.
-%Because the process of starvation is incorporated explicitly, the consumer's rate of reproduction is not dependent on the density of resources.
-%In fact, the existence of the TC bifurcation at $\lambda = \sigma$ reveals an important biological insight. 
-%Reproduction requires maintenance and growth of biological tissues, both of which have strong scaling relationships with body size.
-%Recent work by Kempes et al. [REF] derived the timescale of reproduction in terms of allometric considerations, where $t_\lambda \propto M^{1-\eta}$ (REF).
-%Starvation is the loss of energy required for maintenance, and we have shown it to have a timescale $t_\sigma \propto \log(M)$.
-%
-%A third important parameter in our framework is the rate of recovery.
-%The recovery timescale $t_\rho$ controls the rate at which individuals move from the hungry class to the full class, and this requires not only tissue maintenance, but growth, such that it is bounded on the short side by $t_\sigma$.
-%Moreover, [why is recovery timescale bounded on the high side], such that it is bounded on the long side by $t_\lambda$.
-%Thus, incorporating allometric considerations shows us that $\lambda < \rho < \sigma$ (alternatively $t_\lambda > t_\rho > t_\sigma$).
-
-%General Hopf bifurcation info
-In addition to the hard bound defined by the TC bifurcation, oscillating or cyclic dynamics present an implicit constraint to persistence by increasing extinction risk due to stochastic effects.
-%also present certain constraints to the feasibility of populations.
-%Oscillating, or cyclic, dynamics present additional risks to populations.
-In continuous-time systems, a stable limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts \cite{GuckHolmes}.
-%Of additional interest is the existence of parameter regions that permit the existence of oscillatory, or cyclic, dynamics.
-This condition, known as a Hopf bifurcation, is defined by ${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial describing the Jacobian \cite{Gross:2004p2428}.
-%General transient dynamics statement
-Moreover, as a non-cyclic stable system nears the Hopf bifurcation, transient or decaying cycles can grow in magnitude, despite the existence of a positive, non-cyclic, steady state density.
-Given that ecological systems exist in a state of constant perturbation \cite{Hastings:2001jh}, even the onset of transient cycles that decay over time can increase the risk of extinction \cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}, such that the distance of a system from the Hopf bifurcation is relevant to persistence.
-
-%%Hopf bifurcation and oscillations sigma vs. rho
-%The NSM exhibits both non-cyclic as well as cyclic dynamics, and which behavior is predominant depends strongly on the rate of starvation relative to the rate of recovery.
-%When the starvation rate is low, cyclic dynamics are more likely to occur for a given value of the recovery rate.
-%This can be understood intuitively: for low starvation rates, resources are depressed by an infusion of full consumers, which subsequently starve thereby allowing the resource to recover and continuing the cycle.
-%When the starvation rate is high, the response of consumer growth to resource abundance is muted, such that oscillations tend to decay over time.
-%Thus, higher starvation rates $\sigma$ dampens changes in the consumer population to changes in resource density, and lower rates of recovery $\rho$ amplifies this effect.
-
-%Although the Hopf condition for the specific 2-stage model cannot be easily written, the analytical solution can be explored using a symbolic computational language such as \emph{Mathematica}.
-
-The NSM exhibits both non-cyclic as well as cyclic dynamics, and which behavior dominates depends strongly on the rate of starvation $\sigma$ relative to the rate of recovery $\rho$.
-Although starvation leads to mortality risk, a moderate amount promotes persistence of both consumer and resource populations as non-cyclic stability of the fixed point generally requires a higher $\sigma$ relative to $\rho$.
-The intuition behind this is that transition to the hungry (non-reproductive) state permits the resource to recover and transient dynamics to subside, whereas a low $\sigma$ overloads the system with energetically-replete (reproducing) individuals, maintaining oscillations between consumer and resource (Fig.
-\ref{fig:fp}).
-If $\sigma$ is too large, mortality due to starvation depletes the consumer population, resulting in a lower steady state density for the consumer and a higher steady state density for the resource.
-
-Whereas the rate of consumer growth defines a hard bound of biological feasibility (the TC bifurcation), the rate of starvation thus determines the sensitivity of the consumer population to changes in resource density.
-While higher rates of starvation result in lower steady state population size -- increasing the risk of stochastic extinction -- lower rates of starvation result in a system poised near either the TC or Hopf bifurcation, which will lead to elimination of the resource or the development of cyclic oscillations, respectively.
-Which bifurcation is approached is wholly dependent on the rate of recovery: if it is high, then cyclic dynamics will develop; if it is low, resource extinction becomes increasingly likely.
 

### 2016-07-25 Manuscript/2.tex
Kempes edits to v2

--- 2016-07-25 Manuscript/1.tex
@@
 %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-The parameters in the NSM cannot be freely navigated in biologically-reasonable settings, and our first challenge is to constrain the covariation of rates in a principled and biologically meaningful manner.
+The NSM describes a broad range of dynamics, yet organisms are likely unable to access most of the total parameter space.
+Here we use allometric scaling relationships to constrain the covariation of rates in a principled and biologically meaningful manner.
 % we have not described the realistic regimes occupied by organisms where the
@@
 %Second paragraph
-Nearly all of the rates described in the NSM are to some extent governed by consumer metabolism, and thus can be estimated based on known allometric constraints. 
-The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \cite{West:2002it} and plays a central role in a variety of scaling relationships.
-Organismal metabolic rate $B$ is known to scale as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to be $3/4$ for metazoans, and varies in unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \cite{DeLong:2010dy}.
-Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \cite{Kempes:2012hy}.
-More specifically, the interspecific trends in growth rate can be approximated by $\lambda = \lambda_0 M^{\eta-1}$.
+Nearly all of the rates described in the NSM are to some extent governed by consumer metabolism, which can be used to describe a variety of organism features \cite{}.
+The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \cite{West:2002it} and scales as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to vary around $2/3$ or $3/4$ for metazoans \cite{}, and has taxonomic shifts for unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \cite{DeLong:2010dy,Kempes:2012hy}.
+Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \cite{Kempes:2012hy,more}.
+More specifically, the interspecific trends in growth rate can be approximated by
+\begin{eqnarray}
+\lambda = \lambda_0 M^{\eta-1}.
+\label{lambda}
+\end{eqnarray}
 This relationship is derived from the simple balance
@@
 \end{eqnarray}
- [a and b notation --- these parameters are easily measured bioenergetic parameters which are often approximately invariant across organisms of vastly different size.
-Our notation seeks to illustrate that the allometric model fundamentally depends on a small number of free parameters.]
 where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is the metabolic rate to support an existing unit of mass, and $m$ is the mass at any point in development.
-It is useful to explicitly write this balance because it can also be modified to understand the rates of both starvation and recovery from starvation.
-[Spell out the connection to nutritional state more explicitly] [As we will see it is possible to derive both sigma and rho from this balance]
+It is useful to explicitly write this balance because it can also be modified to understand the timescales of both starvation and recovery from starvation as we show below.
 
-For the rate of starvation, we make the simple assumption that an organism must meet its maintenance requirements using digested mass as the sole energy source.
+For the rate of starvation, $\sigma$, we are interested in the time that it takes an organism to go from healthy adulthood to a smaller size where reproduction is impossible.
+This process is defined by the energy balance in Equation \ref{balance} where we make the simple assumption that an organism must meet its maintenance requirements using the digestion of existing mass as the sole energy source.
 This assumption implies the simple metabolic balance
@@
 \end{eqnarray}
-where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing body mass which may differ from  $E_{m}$, the energy required to synthesis a unit of biomass.
-Give the adult mass, $M$, of an organism this energy balance prescribes the mass trajectory of a starving organism:
+where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing body mass which differs from $E_{m}$ \cite{}, the energy required to synthesis a unit of biomass.
+Given the adult mass, $M$, of an organism this energy balance prescribes the mass trajectory of a starving organism:
 \begin{eqnarray}
@@
 Considering that only certain tissues can be digested for energy, for example the brain cannot be degraded to fuel metabolism, we define the rate for starvation and death by the timescales required to reach specific fractions of normal adult mass.
-We define $m_{starve}=\epsilon M$ where it could be the case that organisms have a systematic size-dependent requirement for essential tissues, such as the minimal bone or brain mass.
+We define $m_{starve}=\epsilon M$, where $\epsilon<1$ is the fraction of adult mass where reproduction ceases.
+This fraction will change if tissue composition systematically scales with adult mass.
 For example, considering the observation that body fat in mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$, and assuming that once this mass is fully digested the organism begins to starve, would imply that $\epsilon=1-f_{0}M^{\gamma}/M$.
 Taken together the time scale for starvation is given by
-\begin{equation}\label{eq:sigma}
+\begin{equation}
+\label{eq:sigma}
 t_{\sigma}=-\frac{E_{m}\log\left(\epsilon\right)}{B_{m}}.
 \end{equation}
-The starvation rate is $\sigma=1/t_{\sigma}$, which implies that $\sigma$ is independent of adult mass if $\epsilon$ is a constant, and if $\epsilon$ does scale with mass, then $\sigma$ will have a factor of $1/\log\left(1-f_{0}M^{\gamma}/M\right)$.
-In either case $\sigma$ does not have a simple scaling with $\lambda$ which is important for the dynamics that we later discuss.
+The starvation rate is then $\sigma=1/t_{\sigma}$, which will scale with adult mass following $1/\log\left(1-f_{0}M^{\gamma}/M\right)$.
+In either case $\sigma$ does not have a simple scaling with $\lambda$ (Equation \ref{lambda}) which is important for the dynamics that we later discuss.
 
@@
 The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues tissue from the starving state to the full state.
-We again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass.
-The general solution to Equation \ref{balance} is given by
+We again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass. The solution to Equation \ref{balance} is given by
 \begin{eqnarray}
-m\left(t\right)=c\left[1-\left(1-\frac{b}{a}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
+m\left(t\right)=\left(\frac{B_{0}}{B_{m}}\right)^{1/(\eta-1)}\left[1-\left(1-\frac{B_{m}}{B_{0}}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
 \end{eqnarray}
-with $a=B_{0}/E_{m}$, $b=B_{m}/E_{m}$, and $c=(a/b)^{1/(\eta-1)}$.
-We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, which has the final form of
+We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, or
 \begin{eqnarray}
-t_{\rho}=\frac{\log \left(1-\left(cM \right)^{1-\eta }\right)-\log \left(1-\left(c\epsilon M \right)^{1-\eta }\right)}{(\eta -1) b}.
+t_{\rho}=\frac{E_{m}\left(\log \left(1-\frac{B_{0}}{B_{m}}\left(M \right)^{1-\eta }\right)-\log \left(1-\frac{B_{0}}{B_{m}}\left(\epsilon M \right)^{1-\eta }\right)\right)}{(\eta -1) B_{m}}.
 \end{eqnarray}
@@
 Although these rate equations are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to $M\approx10^7$ grams (the late Eocene to early Miocene Indricotheriinae).
-Investigating other classes of organisms requires only substituting the energetic and scale parameters shown in Table \ref{param}.
+Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
 Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
@@
 As the allometric derivations of NSM rate laws reveal, $\sigma$ and $\rho$ are not independent parameters, and the bifurcation space shown in Fig.
-\ref{fig:fp} cannot be freely navigated if assuming biologically reasonable parameterizations.
-Given the parameterization for terrestrial endotherms shown in Table \ref{param} with mass $M$ as a free parameter, we show that $\sigma$ and $\rho$ are constrained to a small window of potential values (Fig.
-\ref{fig:hopf}), thus confining dynamics to the steady state regime for all realized body size classes.
+\ref{fig:fp} is navigated via covarying parameters.
+Given the parameterization for terrestrial endotherms we find that $\sigma$ and $\rho$ are constrained to a small window of potential values (Fig.
+\ref{fig:hopf}) for the known range of body sizes $M$.
+We find that the dynamics for all mammalian body sizes is confined to the steady-state regime which does not contain limit cycles.
 %, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
@@
 \ref{fig:hopf}) results in little qualitative difference.
-This suggests that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than mammals with larger body size. %, and
+These results suggest that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than mammals with larger body size.
+Our model predicts that population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting.%, and
 %2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
 
-Allometric constraints have been invoked to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, such that period $\propto M^{0.25}$, however this relationship seems to hold only for some species \cite{Hendriks:2012fc} and competing explanations exist \cite{Kendall:1999iy,Hogstedt:2005cr}.
+Previous studies have used allometric constraints explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$, however this relationship seems to hold only for some species \cite{Hendriks:2012fc} and competing explanations [related to ....] exist \cite{Kendall:1999iy,Hogstedt:2005cr}.
 %It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic \cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
-Statistically significant support for the existence of population cycles among mammals is predominantly based on time-series for smaller bodied mammals \cite{Kendall:1998hl}, though we acknowledge that longer generation times precludes similar quality data for larger organisms.
+Statistically significant support for the existence of population cycles among mammals is predominantly based on time-series for smaller bodied mammals \cite{Kendall:1998hl}, where we our model would predict much longer and more pronounced transient dynamics given how close these points are to the Hopf bifurcation.
+We acknowledge that longer generation times, and more difficult measurement procedures, precludes similar quality data for larger organisms.
 %, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
-We thus obtain a specific prediction from our model: population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting.
  %, emphasizing the role of starvation on dynamics.
@@
 %Extinction to the left and the right
-Higher rates of starvation result in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality, however it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}).
-We examine the competing effects of cyclic dynamics vs. changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.
+\paragraph{Extinction risk}
+Within our model higher rates of starvation result in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality.
+However, from the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}).
+We examine the competing effects of cyclic dynamics vs.
+changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.
 We computed the probability of extinction, where extinction is defined as $H(t)+F(t)=10$ at any instant across all values of $10^2<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq.
@@
 
-As has been described, the $\sigma$ vs. $\rho$ space cannot be freely traversed, such that not all values of $\sigma/\rho$ are biologically feasible.
-We observe that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around energetic parameter means) fall within the extinction refuge, such that they are close enough to the Hopf bifurcation to avoid low steady state densities, though far enough away to avoid large-amplitude transient cycles.
-The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving observed starvation and recovery rates within populations.
-Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-Nevertheless, our finding that allometrically-determined energetic rates place the system within this low extinction probability region suggests that the NSM system provides general insight to a phenomena that may both drive -- and constrain -- natural animal populations.
+We find that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around energetic parameter means) fall within the extinction refuge.
+These values are close enough to the Hopf bifurcation to avoid low steady state densities, and far enough away to avoid large-amplitude transient cycles.
+The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving observed starvation and recovery rates within populations.Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that avoid extinction, and this could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
+% OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
+Nevertheless, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive -- and constrain -- natural animal populations.
 
@@
 %Minimal mass bound
-The NSM correctly predicts that species with smaller masses have larger steady state population densities, however we observe that there is a sharp asymptote in both steady state densities as well as $\sigma/\rho$ at $M \approx 0.3$ grams  (Fig.
+A lower bound on mammalian body size is given by $\epsilon=1$, where mammals have no reserves and immediately starve, which occurs at a size of $M=value$.
+This calculation gives an extreme limit on size but does not account for the subtleties of starvation dynamics that may limit body size.
+The NSM correctly predicts that species with smaller masses have larger steady-state population densities, however we observe that there is a sharp asymptote in both steady state densities as well as $\sigma/\rho$ at $M \approx 0.3$ grams  (Fig.
 \ref{fig:mass}a,b).
 Observation of the rates of starvation and recovery explain why: as mass decreases, the rate of starvation increases, while the rate of recovery declines super-exponentially.
-This decline in $\rho$ occurs when body fat percentage is $1 - 1/(cM) \approx 2\%$, whereupon consumers have no eligible route out of starvation.
+This decline in $\rho$ occurs when body fat percentage is $1 - 1/(\left(B_{0}/B_{m}\right)^{1/\left(\eta-1\right)}M) \approx 2\%$, whereupon consumers have no eligible route out of starvation.
 Compellingly, this dynamic bound determined by the rate of energetic recovery is close to the minimum observed mammalian body size of ca.
@@
 In addition to known transport limitations \cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.
+This work mirrors other efforts where coincident limitations seem to limit the smallest possibilities for life within a particular class or organisms \cite{kempes}.
 
 %Maximum mass bound
-Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
-Instead, we examine a potential upper bound to body mass by assessing population invasibility with respect to a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-0.5,0.5]$, thus altering rates of starvation, recovery, and maintenance $\beta$.
-Although there is not an internal fixed point where both residents and invaders coexist (except for the trivial state $\chi=0$), we can assess invasibility as a function of organismal mass by determining which consumer steady state is larger over $\chi$.
+%Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
+We examine a potential upper bound to body mass by assessing population invasibility with respect to a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-0.5,0.5]$, thus altering rates of starvation, recovery, and maintenance $\beta$.
+Although there is not an internal fixed point where both residents and invaders coexist (except for the trivial state $\chi=0$), we assess invasibility, as a function of organismal mass, by determining which consumer has a higher steady state given $\chi$.
 We find that for $1\leq M<10^6$g, having additional body fat ($\chi > 0$) results in a higher steady state density for the invader population ($H^{\prime *}+F^{\prime *}>H^*+F^*$), such that it has an intrinsic advantage over the resident population.
-For $M>10^6$, however, there is an increasing range of $\chi < 0$ such that leaner individuals have the advantage, and this is due to the changing covariance between energetic rates as a function of modified energetic reserves.
+However, for $M>10^6$, leaner individuals ($\chi < 0$) have an advantage, and this is due to the changing covariance between energetic rates as a function of modified energetic reserves.
 
@@
 \centering
-\includegraphics[width=1\textwidth]{fig_Invasion.pdf}
+\includegraphics[width=0.5\textwidth]{fig_Invasion.pdf}
 \caption{ % 58 words

### 2016-07-25 Manuscript/2.tex
Sid first pass

@@
 changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.
-We computed the probability of extinction, where extinction is defined as $H(t)+F(t)=10$ at any instant across all values of $10^2<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq.
+We computed the probability of extinction, where extinction is assumed if the population trajectory goes below $0.2~\times$ the allometrically constrained steady state across all values of $10^2<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq.
 \ref{eq:system} for an organism of $M=100$ grams, assuming random initial conditions around the steady state (Eq.

### 2016-07-25 Manuscript/3.tex
Sid first pass

--- 2016-07-25 Manuscript/2.tex
@@
 \usepackage{rotating}
-%\usepackage{color}
+\usepackage{color}
 %\usepackage{lineno}
@@
 
-\title{Ecological and evolutionary implications of starvation and body size}
+%\title{Ecological and evolutionary implications of starvation and body size}
 
-%\title{Survival of the Stable: Dynamic Consequence of Starvation and Reproduction}
+%\title{Survival of the Stable: Dynamic Consequences of Starvation and Reproduction}
 
-%\title{Size Constraints on Starvation/Reproduction Dynamics}
+%\title{Size Constraints on Starvation \& Reproduction Dynamics}
+
+%\title{Eco-evolutionary Consequences of Starvation and Reproduction: Survival
+%of the Stable}
+
+\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
 
@@
 \begin{abstract}
-This is the abstract.  \emph{ No it isn't.  It's merely a placeholder.}
+This is the abstract.  \emph{ \sid{No it isn't.  It's merely a placeholder.}}
 \end{abstract}
@@
 
-\noindent {\bf Introduction} \\ \nonumber
-The behavioral ecology of most, if not all, organisms is influenced by the
-energetic state of individuals, which directly influences how organisms
-invest reserves in uncertain environments.  Such behaviors are generally
-manifested as trade-offs between investing in somatic maintenance and growth
-or allocating energy towards reproduction
-\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}. 
-The timing of these behaviors responds to selective pressure, as the choice
-of the investment impacts future fitness!\cite{Mangel:1988uaa}.
-The influence of resource limitation on an organism's ability to maintain its
-nutritional stores may lead to repeated delays or shifts in reproduction over
-the course of an organism's life.
+\noindent {\bf Introduction} \\ The behavioral ecology of most, if
+not all, organisms is influenced by the energetic state of individuals, which
+directly influences how they invest reserves in uncertain environments.  Such
+behaviors are generally manifested as trade-offs between investing in somatic
+maintenance and growth, or allocating energy towards
+reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.  The timing of
+these behaviors responds to selective pressure, as the choice of the
+investment impacts future fitness~\cite{Mangel:1988uaa}.  The influence of
+resource limitation on an organism's ability to maintain its nutritional
+stores may lead to repeated delays or shifts in reproduction over the course
+of an organism's life.
 
 The life history of most species is typically comprised of (a) somatic growth
-and maintenance and (b) reproduction.  The balance between these two
+and maintenance, and (b) reproduction.  The balance between these two
 activities is often conditioned on resource
@@
 calves born after harsh winters (when the mother's energetic state is
-depleted) than in calves born after moderate winters \cite{Tveraa:2003fq}.
+depleted) than in calves born after moderate winters~\cite{Tveraa:2003fq}.
 Many bird species invest differently in broods during periods of resource
@@
 and marine zooplankton have been observed to avoid reproduction under
-nutritional stress \cite{Threlkeld:1976ih}, and those that do reproduce have
+nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have
 lower survival rates~\cite{Kirk:1997cc}.  Organisms may also separate
@@
 Physiological mechanisms also play an important role in regulating
-reproductive expenditures during periods of resource limitation.  Diverse
-mammals (47 species in 10 families) exhibit delayed implantation, whereby
-females postpone fetal development (blastocyst implantation) until times
-where nutritional reserves can be
+reproductive expenditures during periods of resource limitation.  The data
+collected thus far has shown that diverse mammals (47 species in 10 families)
+exhibit delayed implantation, whereby females postpone fetal development
+(blastocyst implantation) until times where nutritional reserves can be
 accumulated~\cite{Mead:1989dt,Sandell:1990kw}.  Many other many species
-(including humans) suffer irregular menstrual cycling and higher abortion
-rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
-In the extreme case of unicellular organisms, nutrition is unavoidably linked
-to reproduction because the nutritional state of the cell regulates all
-aspects of the cell cycle \cite{Glazier:2009hq}.  The existence of so many
-independently evolved mechanisms across such a diverse suite of organisms
-highlights the importance and universality of the fundamental tradeoff
-between somatic and reproductive investment.  However the dynamic
-implications of these constraints are unknown.
+(including humans) suffer irregular menstrual cycling and higher spontaneous
+abortion rates during periods of nutritional
+stress~\cite{Bulik:1999eo,Trites:2003cc}.  In the extreme case of unicellular
+organisms, nutrition is unavoidably linked to reproduction because the
+nutritional state of the cell regulates all aspects of the cell cycle
+\cite{Glazier:2009hq}.  The existence of so many independently evolved
+mechanisms across such a diverse suite of organisms highlights the importance
+and universality of the fundamental tradeoff between somatic and reproductive
+investment.  However the dynamic implications of these constraints are
+unknown.
 
@@
 framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical
-obstacles~\cite{Diekmann:2010da}).  An alternative approach involves modeling
+obstacles~\cite{Diekmann:2010da}.  An alternative approach involves modeling
 the macroscale relations that guide somatic versus reproductive investment in
@@
 density, thus \emph{implicitly} incorporating the requirement of resource
-availability for reproduction \cite{murdoch:2003}.
+availability for reproduction~\cite{murdoch:2003}.
 
@@
 Such a constraint leads to reproductive time lags due to some members of the
-population starving and then recovering.  Additonally, we incorporate the
+population going hungry and then recovering.  Additionally, we incorporate the
 idea that reproduction is strongly constrained
@@
 
-\noindent {\bf Nutritional-state-structured model (NSM)}
-
+\noindent {\bf Nutritional-state-structured model (NSM)}\\
 We begin by defining a minimal Nutritional-State-structured population Model
@@
 (a) an energetically replete (full) state $F$, where the consumer reproduces
-at a constant rate $\lambda$, and (b) an energetically deficient (hungry)
-state $H$, where the consumer does not reproduce but dies at rate $\mu$.  The
-underlying resource $R$ evolves by logistic growth with an intrinsic growth
-rate $\alpha$ and a carrying capacity equal to one.  Consumers transition
-from the full state $F$ to the hungry state $H$ by starvation at rate
-$\sigma$ and also in proportion to the absence of resources $(1-R)$.
-Conversely, consumers recover from state $H$ to state $F$ at rate $\rho$ and
-in proportion to $R$.  Resources are also eaten by the consumers---at rate
-$\rho$ by hungry consumers and at rate $\beta<\rho$ by full consumers.  This
-inequality accounts for hungry consumers requiring more resources to rebuild
-body weight.
+at a constant rate $\lambda$ and has no mortality risk, and (b) an
+energetically deficient (hungry) state $H$, where the consumer does not
+reproduce but dies at rate $\mu$.  The underlying resource $R$ evolves by
+logistic growth with an intrinsic growth rate $\alpha$ and a carrying
+capacity equal to one.  Consumers transition from the full state $F$ to the
+hungry state $H$ by starvation at rate $\sigma$ and also in proportion to the
+absence of resources $(1-R)$.  Conversely, consumers recover from state $H$
+to state $F$ at rate $\rho$ and in proportion to $R$.  Resources are also
+eaten by the consumers---at rate $\rho$ by hungry consumers and at rate
+$\beta<\rho$ by full consumers.  This inequality accounts for hungry
+consumers requiring more resources to rebuild body weight.
 
@@
 Notice that the total consumer density $F+H$ evolves according to
-$\dot{F}+\dot{H}=\lambda F-\mu H$.  This resmebles the equation of motion for
+$\dot{F}+\dot{H}=\lambda F-\mu H$.  This resembles the equation of motion for
 the predator density in the classic Lotka-Volterra model, except that the
@@
 \end{eqnarray}
-If this internal fixed point, which is unique, is stable, it will be the
-global attractor for all population trajectories for any initial condition
-where the resource and consumer densities are both non zero.  The stability
-of this fixed point is determined by the Jacobian Matrix $\bf J$, where each
-matrix element $J_{ij}$ equals $\partial{\dot X_i}/\partial{X_j}$ when
-evaluated at the internal fixed point, and $\mathbf{X}$ is the vector
-$(F,H,R)$.  If the parameters in Eq.~\eqref{eq:system} are such that the real
-part of the largest eigenvalue of $\mathbf{J}$ is negative, then the system
-is stable with respect to small perturbations from the fixed point.
+The stability of this fixed point is determined by the Jacobian Matrix
+$\bf J$, where each matrix element $J_{ij}$ equals
+$\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed
+point, and $\mathbf{X}$ is the vector $(F,H,R)$.  The parameters in
+Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue
+of $\mathbf{J}$ is negative, so that the system is stable with respect to
+small perturbations from the fixed point.  Because this fixed point is
+unique, it is the global attractor for all population trajectories for any
+initial condition where the resource and consumer densities are both non
+zero.
 
@@
 reproduction rate $\lambda$ must be less than the starvation rate $\sigma$,
-so that $R^*$ is positive.  Moreover, when the resource density $R=0$, the
-rate equation for $F$ gives exponential growth for $\lambda>\sigma$.  The
-condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation that
-demarcates the physical and unphysical regimes.  The biological implication
-of the constraint $\lambda<\sigma$ is simple---the rate at which a
-macroscopic organism loses mass due to lack of resources is generally much
-faster than the rate of reproduction.  As we will discuss below, this
-inequality is a natural consequence of allometric
+so that $R^*$ is positive.  In fact, when the resource density $R=0$, the
+rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
+The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation
+that demarcates the physical and unphysical regimes\sid{give Ref}.  The
+biological implication of the constraint $\lambda<\sigma$ has a simple
+interpretation---the rate at which a macroscopic organism loses mass due to
+lack of resources is generally much faster than the rate of reproduction.  As
+we will discuss below, this inequality is a natural consequence of allometric
 constraints~\cite{Kempes:2012hy} for organisms within empirically observed
@@
 parts~\cite{GuckHolmes}.  This Hopf bifurcation is defined by
-${\rm Det}({\bf S}) = 0$, where $\bf S$ is the Sylvester matrix, which is
+${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is
 composed of the coefficients of the characteristic polynomial of the Jacobian
@@
 systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of
-transient cycles, even though they decay with time, can increase the
-extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.  Thus
-the distance of a system from the Hopf bifurcation provides a measure of its
-persistence.
+transient cycles, even though they decay with time in the mean-field
+description, can increase the extinction
+risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.  Thus the distance
+of a system from the Hopf bifurcation provides a measure of its persistence.
 
 When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the
-consumers are driven to the hungry non-reproductive state.  Because
+consumers are driven to the hungry non-reproducing state.  Because
 reproduction is inhibited, there is a low steady-state consumer density and a
-high steady-state resource density.  However, if $\sigma/\lambda$ approaches
-one, the population is overloaded with energetically-replete (reproducing)
+high steady-state resource density.  However, if $\sigma/\lambda\to 1$ from
+above, the population is overloaded with energetically-replete (reproducing)
 individuals, thereby promoting oscillations between the consumer and resource
@@
 
-Whereas the consumer growth rate $\lambda$ defines an absolute bound of
-biological feasibility---the TC bifurcation---the starvation rate $\sigma$
+Whereas the relation between consumer growth rate $\lambda$ and the
+starvation rate $\sigma$ defines an absolute bound of biological
+feasibility---the TC bifurcation---the starvation rate $\sigma$ also
 determines the sensitivity of the consumer population to changes in resource
 density.  When $\sigma\gg\lambda$, the steady-state population density is
-small, thereby increasing the risk of stochastic extinction.  Conversely, as
-$\sigma$ decreases, the system will ultimately be poised either near the TC
-or the Hopf bifurcation (Fig.~\ref{fig:fp}).  If the recovery rate $\rho$ is
-sufficiently small, the TC bifurcation is reached and the resource eventually
-is eliminated.  If $\rho$ exceeds a threshold value, cyclic dynamics will
-develop as the Hopf bifurcation is approached.
+small, thereby increasing the risk of stochastic extinction.  On the other
+hand, as $\sigma$ decreases, the system will ultimately be poised either near
+the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).  If the recovery rate
+$\rho$ is sufficiently small, the TC bifurcation is reached and the resource
+eventually is eliminated.  If $\rho$ exceeds a threshold value, cyclic
+dynamics will develop as the Hopf bifurcation is approached.
+\\
 
@@
 
-\vspace{2mm}
 \noindent {\bf Role of allometry} \\ 
@@
 %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-The NSM describes a broad range of dynamics, yet organisms are likely unable to access most of the total parameter space.
-Here we use allometric scaling relationships to constrain the covariation of rates in a principled and biologically meaningful manner.
-% we have not described the realistic regimes occupied by organisms where the
-Allometric scaling relationships highlight common constraints and average trends across large ranges in body size and species diversity.
-Many of these relationships can be derived from a small set of assumptions and below we describe a framework for the covariation of timescales and rates across the range of mammals for each of the key parameters of our model (cf.
-\cite{Yodzis:1992hg}).
-We are able to define the regime of dynamics occupied by the entire class of mammals along with the key differences between the largest and smallest mammals.
+The NSM describes a broad range of dynamics, yet organisms are likely unable
+to access most of the total parameter space. Here we use allometric scaling
+relations to constrain the covariation of rates in a principled and
+biologically meaningful manner.  Allometric scaling relations highlight
+common constraints and average trends across large ranges in body size and
+species diversity. Many of these relations can be derived from a small set of
+assumptions and below we describe a framework to determine the covariation of
+timescales and rates across the range of mammals for each of the key
+parameters of our model (cf.~\cite{Yodzis:1992hg}).  We are thereby able to
+define the regime of dynamics occupied by the entire class of mammals along
+with the key differences between the largest and smallest mammals.
 
-%Second paragraph
-Nearly all of the rates described in the NSM are to some extent governed by consumer metabolism, which can be used to describe a variety of organism features \cite{}.
-The scaling relationship between an organism's metabolic rate $B$ and its body size at reproductive maturity $M$ is well documented \cite{West:2002it} and scales as $B = B_0 M^\eta$, where $\eta$ is the scaling exponent, generally assumed to vary around $2/3$ or $3/4$ for metazoans \cite{}, and has taxonomic shifts for unicellular species between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria \cite{DeLong:2010dy,Kempes:2012hy}.
-Several efforts have shown how a partitioning of this metabolic rate between growth and maintenance purposes can be used to derive a general equation for the growth trajectories and growth rates of organisms ranging from bacteria to metazoans \cite{Kempes:2012hy,more}.
-More specifically, the interspecific trends in growth rate can be approximated by
+Nearly all of the rates described in the NSM are to some extent governed by
+consumer metabolism, which can be used to describe a variety of organismal
+features~\cite{}\sid{Ref}.  The scaling relation between an organism's
+metabolic rate $B$ and its body size at reproductive maturity $M$ is well
+documented~\cite{West:2002it} and scales as $B = B_0 M^\eta$, where $\eta$ is
+the scaling exponent, generally assumed to vary around $2/3$ or $3/4$ for
+metazoans~\cite{}\sid{Ref}, and has taxonomic shifts for unicellular species
+between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria
+\cite{DeLong:2010dy,Kempes:2012hy}. Several efforts have shown how a
+partitioning of this metabolic rate between growth and maintenance purposes
+can be used to derive a general equation for the growth trajectories and
+growth rates of organisms ranging from bacteria to metazoans
+\cite{Kempes:2012hy,more}\sid{fix Ref}. More specifically, the interspecific
+\sid{what does this word mean?}  trends in growth rate can be approximated by
 \begin{eqnarray}
@@
 \end{eqnarray}
-This relationship is derived from the simple balance
+This relation is derived from the simple balance condition
 \begin{eqnarray}
 \label{balance}
-B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m
+B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
 \end{eqnarray}
-where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is the metabolic rate to support an existing unit of mass, and $m$ is the mass at any point in development.
-It is useful to explicitly write this balance because it can also be modified to understand the timescales of both starvation and recovery from starvation as we show below.
+where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
+the metabolic rate to support an existing unit of mass, and $m$ is the mass
+at any point in development. It is useful to explicitly write this balance
+because it can also be modified to understand the timescales of both
+starvation and recovery from starvation as we show below.
 
-For the rate of starvation, $\sigma$, we are interested in the time that it takes an organism to go from healthy adulthood to a smaller size where reproduction is impossible.
-This process is defined by the energy balance in Equation \ref{balance} where we make the simple assumption that an organism must meet its maintenance requirements using the digestion of existing mass as the sole energy source.
-This assumption implies the simple metabolic balance
+To determine the starvation rate, $\sigma$, we are interested in the time
+required for an organism to go from its replete \sid{replete vs. adult} state
+to a reduced-size hungry state where reproduction is impossible.  This
+transition time can be inferred from the energy balance in Eq.~\ref{balance},
+where we make the basic assumption that an organism must meet its maintenance
+requirements using the digestion of existing mass as the sole energy source.
+This assumption implies the simple metabolic balance \sid{don't understand
+  this equation}
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing body mass which differs from $E_{m}$ \cite{}, the energy required to synthesis a unit of biomass.
-Given the adult mass, $M$, of an organism this energy balance prescribes the mass trajectory of a starving organism:
+where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
+body mass which differs from $E_{m}$~\cite{}\sid{Ref}, the energy required to
+synthesis a unit of biomass. Given the replete mass, $M$, of an organism, the
+above energy balance prescribes the mass trajectory of a non-consuming
+organism:
 \begin{eqnarray}
+\label{mt}
 m\left(t\right)=Me^{-B_{m}t/E_{m}^{\prime}}.
 \end{eqnarray}
-Considering that only certain tissues can be digested for energy, for example the brain cannot be degraded to fuel metabolism, we define the rate for starvation and death by the timescales required to reach specific fractions of normal adult mass.
-We define $m_{starve}=\epsilon M$, where $\epsilon<1$ is the fraction of adult mass where reproduction ceases.
-This fraction will change if tissue composition systematically scales with adult mass.
-For example, considering the observation that body fat in mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$, and assuming that once this mass is fully digested the organism begins to starve, would imply that $\epsilon=1-f_{0}M^{\gamma}/M$.
-Taken together the time scale for starvation is given by
+Since only certain tissues can be digested for energy (for example the brain
+cannot be degraded to fuel metabolism), we define the rate for starvation and
+death by the timescales required to reach specific fractions of normal adult
+mass. We define $m_{\rm starve}=\epsilon M$, where $\epsilon<1$ is the
+fraction of adult mass where reproduction ceases.  This fraction will be
+modified if tissue composition systematically scales with adult mass.  For
+example, making use of the observation that body fat in mammals scales with
+overall body size according to $M_{f}=f_{0}M^{\gamma}$ and assuming that once
+this mass is fully digested the organism begins to starve, this would imply
+that $\epsilon=1-f_{0}M^{\gamma}/M$.  Using this criterion in Eq.~\ref{mt},
+the time scale for starvation is given by
 \begin{equation}
 \label{eq:sigma}
-t_{\sigma}=-\frac{E_{m}\log\left(\epsilon\right)}{B_{m}}.
+t_{\sigma}=-\frac{E_{m}\ln\left(\epsilon\right)}{B_{m}}.
 \end{equation}
-The starvation rate is then $\sigma=1/t_{\sigma}$, which will scale with adult mass following $1/\log\left(1-f_{0}M^{\gamma}/M\right)$.
-In either case $\sigma$ does not have a simple scaling with $\lambda$ (Equation \ref{lambda}) which is important for the dynamics that we later discuss.
+The starvation rate is then $\sigma=1/t_{\sigma}$, which scales with replete
+mass as $1/\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important feature is
+that $\sigma$ does not have a simple scaling dependence on $\lambda$
+(Eq.~\ref{lambda}), which is important for the dynamics that we later
+discuss.
 
-The time to death should follow a similar relationship, but defined by a lower fraction of adult mass, $m_{death}=\epsilon^{\prime} M$.
-Consider, for example, that an organism dies once it has digested all fat and muscle tissues, and that muscle tissue scales with body mass according to $M_{mm}=mm_{0}M^{\zeta}$, then $\epsilon^{\prime}=1-\left(f_{0}M^{\gamma}+mm_{0}M^{\zeta}\right)/M$.
-Muscle mass has been shown to be roughly proportional to body mass \cite{muscle} in mammals and thus  $\epsilon^{\prime}$ is effectively $\epsilon$ minus a constant.
+The time to death should follow a similar relation, but defined by a lower
+fraction of adult mass, $m_{\rm death}=\epsilon^{\prime} M$.  Suppose, for
+example, that an organism dies once it has digested all fat and muscle
+tissues, and that muscle tissue scales with body mass according to
+$M_{mm}=mm_{0}M^{\zeta}$ \sid{what is $mm_0$?}.  This gives
+$\epsilon^{\prime}=1-\left(f_{0}M^{\gamma}+mm_{0}M^{\zeta}\right)/M$. Muscle
+mass has been shown to be roughly proportional to body mass~\cite{muscle} in
+mammals and thus $\epsilon^{\prime}$ is merely $\epsilon$ minus a constant.
 Thus
 \begin{eqnarray}
-t_{\mu}=-\frac{E_{m}\log\left(\epsilon^{\prime}\right)}{B_{m}}
+t_{\mu}=-\frac{E_{m}\ln\left(\epsilon^{\prime}\right)}{B_{m}}
 \end{eqnarray}
@@
 %CHRIS - MAYBE WE CUT THIS OUT SINCE WE STICK TO MAMMALS IN THIS PAPER
-%It should be noted that we have thus far used mammalian allometry to describe the size-based relationships for growth, starvation, and death.
+%It should be noted that we have thus far used mammalian allometry to describe the size-based relations for growth, starvation, and death.
 However, our presentation is general, and other functional forms for $\epsilon$, for example, could be determined for other classes of organisms.
@@
 
-The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues tissue from the starving state to the full state.
-We again use the balance given in Equation \ref{balance} to find the timescale to return to the mature mass from a given reduced starvation mass. The solution to Equation \ref{balance} is given by
+The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
+sufficient tissue to transition from the starving \sid{starving vs.\ hungry}
+state to the full state.  We again use the balance given in Eq.~\ref{balance}
+to find the timescale to return to the replete mass from a given reduced
+starvation mass.  From the solution to Eq.~\ref{balance}
 \begin{eqnarray}
-m\left(t\right)=\left(\frac{B_{0}}{B_{m}}\right)^{1/(\eta-1)}\left[1-\left(1-\frac{B_{m}}{B_{0}}m_{0}^{1-\eta}\right)e^{-b\left(1-\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}
+m\left(t\right)\!=\!\left(\!\frac{B_{0}}{B_{m}}\!\right)^{1/(\eta-1)}\!\!
+\left[1\!-\!\left(1\!-\!\frac{B_{m}}{B_{0}}m_{0}^{1\!-\!\eta}\right)e^{-b\left(1\!-\!\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}\,,
 \end{eqnarray}
-We are then interested in the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, or
+we require the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes
+to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, or
 \begin{eqnarray}
-t_{\rho}=\frac{E_{m}\left(\log \left(1-\frac{B_{0}}{B_{m}}\left(M \right)^{1-\eta }\right)-\log \left(1-\frac{B_{0}}{B_{m}}\left(\epsilon M \right)^{1-\eta }\right)\right)}{(\eta -1) B_{m}}.
+t_{\rho}=\frac{E_{m}\left\{\ln \left[1\!-\!\frac{B_{0}}{B_{m}}\left(M
+  \right)^{1\!-\!\eta }\right]
+-\ln \left[1\!-\!\frac{B_{0}}{B_{m}}\left(\epsilon M \right)^{1\!-\!\eta }\right]\right\}}{(\eta -1) B_{m}}.
 \end{eqnarray}
@@
 %where $c=(a/b)^{1/(\eta-1)}$.
-Although these rate equations are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to $M\approx10^7$ grams (the late Eocene to early Miocene Indricotheriinae).
-Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
-Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
+Although these rate equations are general, here we focus on parameterizations
+for terrestrial-bound endotherms, specifically mammals, which range from a
+minimum of $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to a
+maximum of $M\approx10^7$ grams (the late Eocene to early Miocene
+Indricotheriinae).  Investigating other classes of organisms would simply
+involve altering the metabolic exponents and scalings associate with
+$\epsilon$. Moreover, we emphasize that our allometric equations describe
+mean relationships, and do not account for the (sometimes considerable)
+variance associated with individual species.
+\\
 %, though this is beyond the current scope of our analysis. %, and we henceforth focus our examination on mammalian species.
 
-\vspace{2mm}
-\noindent {\bf The stabilizing effects of allometric constraints} \\ \noindent
-
-%Bifurcation space as a function of mass
-As the allometric derivations of NSM rate laws reveal, $\sigma$ and $\rho$ are not independent parameters, and the bifurcation space shown in Fig.
-\ref{fig:fp} is navigated via covarying parameters.
-Given the parameterization for terrestrial endotherms we find that $\sigma$ and $\rho$ are constrained to a small window of potential values (Fig.
-\ref{fig:hopf}) for the known range of body sizes $M$.
-We find that the dynamics for all mammalian body sizes is confined to the steady-state regime which does not contain limit cycles.
+\noindent {\bf  Stabilizing effects of allometric constraints} \\
+As the allometric derivations of the NSM rate laws reveal, $\sigma$ and
+$\rho$ are not independent parameters, and the bifurcation space shown in
+Fig.~\ref{fig:fp} is navigated via covarying parameters.  Given the
+parameters of terrestrial endotherms, we find that $\sigma$ and $\rho$ are
+constrained to lie within a small window of potential values
+(Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.  We thus find
+that the dynamics for all mammalian body sizes is confined to the
+steady-state regime of the NSM and that  limit-cycle behavior is precluded.
 %, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
-Moreover, for larger $M$, the distance to the Hopf bifurcation increases, while uncertainty in allometric parameters (20\% variation around the mean; Fig.
-\ref{fig:hopf}) results in little qualitative difference.
-These results suggest that small mammals are more prone to population oscillations -- including both stable limit cycles as well as transient cycles -- than mammals with larger body size.
-Our model predicts that population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting.%, and
+Moreover, for larger $M$, the distance to the Hopf bifurcation increases,
+while uncertainty in allometric parameters (20\% variation around the mean;
+Fig.~\ref{fig:hopf}) results in little qualitative difference in the distance
+to the the Hopf bifurcation. These results suggest that small mammals are
+more prone to population oscillations---including both stable limit cycles as
+well as transient cycles---than mammals with larger body size.  Thus our NSM
+model predicts that population cycles should be less common for larger
+species and more common for smaller species, particularly in environments
+where resources are limiting.%, and
 %2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
 
-Previous studies have used allometric constraints explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$, however this relationship seems to hold only for some species \cite{Hendriks:2012fc} and competing explanations [related to ....] exist \cite{Kendall:1999iy,Hogstedt:2005cr}.
-%It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic \cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
-Statistically significant support for the existence of population cycles among mammals is predominantly based on time-series for smaller bodied mammals \cite{Kendall:1998hl}, where we our model would predict much longer and more pronounced transient dynamics given how close these points are to the Hopf bifurcation.
-We acknowledge that longer generation times, and more difficult measurement procedures, precludes similar quality data for larger organisms.
+Previous studies have used allometric constraints to explain the periodicity
+of cyclic populations
+\cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period
+$\propto M^{0.25}$, however this relation seems to hold only for some species
+\cite{Hendriks:2012fc} and competing explanations [related to ....\sid{???}]
+exist~\cite{Kendall:1999iy,Hogstedt:2005cr}.
+%It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic~\cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
+Statistically significant support for the existence of population cycles
+among mammals is predominantly based on time series for small
+mammals~\cite{Kendall:1998hl}, where we our model would predict much longer
+and more pronounced transient dynamics, given how close these points are to
+the Hopf bifurcation.  On the other hand, the longer gestational times and
+the increased difficulty in measurements, precludes obtaining similar-quality
+data for larger organisms.
+\\
 %, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
@@
 %Extinction to the left and the right
-\paragraph{Extinction risk}
-Within our model higher rates of starvation result in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of mortality.
-However, from the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the system (Fig \ref{fig:fp}).
-We examine the competing effects of cyclic dynamics vs.
-changes in steady state density on extinction risk as a function of the ratio $\sigma/\rho$.
-We computed the probability of extinction, where extinction is defined as $H(t)+F(t)=10$ at any instant across all values of $10^2<t\leq 10^6$, for 1000 replicates of the continuous-time system shown in Eq.
-\ref{eq:system} for an organism of $M=100$ grams, assuming random initial conditions around the steady state (Eq.
-\ref{eq:ss}).
-%random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
-By allowing the rate of starvation to vary, we assessed extinction risk across a range of values of the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a horizontal cross-section of Fig.
-\ref{fig:fp}.
-As expected, higher rates of extinction correlated with both low and high values of $\sigma/\rho$; for low values the higher extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.
-\ref{fig:ext}).
-For large values of $\sigma/\rho$, higher extinction risk is due to the decrease in the steady state consumer population density.
-This interplay creates an `extinction refuge' as shown in Fig. \ref{fig:ext}, such that for a relatively constrained range of $\sigma/\rho$, extinction probabilities are minimized.
 
-We find that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around energetic parameter means) fall within the extinction refuge.
-These values are close enough to the Hopf bifurcation to avoid low steady state densities, and far enough away to avoid large-amplitude transient cycles.
-The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving observed starvation and recovery rates within populations.Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that avoid extinction, and this could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
+\noindent {\bf Extinction risk}\\
+Within our model, higher rates of starvation result in a larger flux of the
+population to the hungry state.  In this state reproduction is absent, thus
+increasing the likelihood of extinction.  However, from the perspective of
+population survival, it is the rate of starvation relative to the rate of
+recovery that determines the long-term dynamics of the system
+(Fig.~\ref{fig:fp}).  We now examine the competing effects of cyclic dynamics
+vs.\ changes in steady state density on extinction risk as a function of the
+ratio $\sigma/\rho$.  To this end, we computed the probability of extinction,
+where extinction is defined as the population trajectory going below
+$0.2~\times$ the allometrically constrained steady state for all times
+between $10^2$ and $\leq 10^6$.  This procedure is repeated for 1000
+replicates \sid{what is changing? isn't the system deterministic?} of the
+continuous-time system shown in Eq.~\ref{eq:system} for an organism of
+$M=100$ grams, assuming random initial conditions around the steady state
+(Eq.~\ref{eq:ss}). %random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
+By allowing the rate of starvation to vary, we assessed extinction risk
+across a range of values of the ratio $\sigma/\rho$ varying between $10^{-2}$
+to $2.5$, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.  As
+expected, higher rates of extinction correlated with both low and high values
+of $\sigma/\rho$.  For low values of $\sigma/\rho$, the increased extinction
+risk results from transient cycles with larger amplitudes as the system nears
+the Hopf bifurcation (Fig.~\ref{fig:ext}).  For large values of
+$\sigma/\rho$, higher extinction risk arises because of to the decrease in
+the steady state consumer population density.  This interplay creates an
+`extinction refuge' as shown in Fig.~\ref{fig:ext}, such that for a
+relatively constrained range of $\sigma/\rho$, extinction probabilities are
+minimized.
+
+We find that the allometrically constrained values of $\sigma/\rho$ (with
+$\pm 20\%$ variability around energetic parameter means) fall within the
+extinction refuge. These values are close enough to the Hopf bifurcation to
+avoid low steady state densities, and far enough away to avoid
+large-amplitude transient cycles. The fact that allometric values of $\sigma$
+and $\rho$ fall within this relatively small window supports the possibility
+that a selective mechanism has constrained the physiological conditions that
+drive observed starvation and recovery rates within populations.  Such a
+mechanism would select for organism physiology that generates appropriate
+$\sigma$ and $\rho$ values that avoid extinction.  This selection could occur
+via the tuning of body fat percentages, metabolic rates, and biomass
+maintenance efficiencies.
 % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-Nevertheless, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive -- and constrain -- natural animal populations.
+To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
 
@@
 
-\vspace{2mm}
-\noindent {\bf Dynamic and energetic barriers to body size} \\ \nonumber
-Metabolite transportation constraints are widely thought to place strict boundaries on biological scaling \cite{Brown:1993p708,West:1997cg,Brown:2004wq}, leading to specific predictions on the minimum possible body size for organisms \cite{West:2002ud}.
-Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
-The \emph{fasting endurance hypothesis} contends that larger body size, with lower metabolic rates and able to hold more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability \cite{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's Rule) \cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper-bound of ca.
-$10^7$ grams \cite{Alroy:1998p1594}, the value of which may arise from higher extinction risk for large taxa over evolutionary timescales \cite{Clauset:2009fh}.
-These trends are thought to be driven by a combination of climate change and niche availability \cite{Saarinen:2014br}, however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, has not been explored, and we contend that the NSM provides a suitable framework to explore these issues.
+\noindent {\bf Dynamic and energetic barriers to body size} \\
+Metabolite transport constraints are widely thought to place strict
+boundaries on biological
+scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to
+specific predictions on the minimum possible body size for
+organisms~\cite{West:2002ud}.  Above this bound, a number of energetic and
+evolutionary mechanisms have been explored to assess the costs and benefits
+associated with larger body masses, particularly for mammals.  One important
+such example is the \emph{fasting endurance hypothesis}, which contends that
+larger body size, with consequent lower metabolic rates and increased ability
+to maintain more endogenous energetic reserves, may buffer organisms against
+environmental fluctuations in resource availability~\cite{Millar:1990p923}.
+Over evolutionary time, terrestrial mammalian lineages show a significant
+trend towards larger body size (known as Cope's
+Rule)~\cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br},
+and it is thought that within-lineage drivers generate selection towards an
+optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, the value
+of which may arise from higher extinction risk for large taxa over
+evolutionary timescales~\cite{Clauset:2009fh}.  These trends are thought to
+be driven by a combination of climate change and niche
+availability~\cite{Saarinen:2014br}; however the underpinning energetic costs
+and benefits of larger body sizes, and how they influence dynamics over
+ecological timescales, have not been explored.  We argue that the NSM
+provides a suitable framework to explore these issues.
 
 %Minimal mass bound
-A lower bound on mammalian body size is given by $\epsilon=1$, where mammals have no reserves and immediately starve, which occurs at a size of $M=value$.
-This calculation gives an extreme limit on size but does not account for the subtleties of starvation dynamics that may limit body size.
-The NSM correctly predicts that species with smaller masses have larger steady-state population densities, however we observe that there is a sharp asymptote in both steady state densities as well as $\sigma/\rho$ at $M \approx 0.3$ grams  (Fig.
-\ref{fig:mass}a,b).
-Observation of the rates of starvation and recovery explain why: as mass decreases, the rate of starvation increases, while the rate of recovery declines super-exponentially.
-This decline in $\rho$ occurs when body fat percentage is $1 - 1/(\left(B_{0}/B_{m}\right)^{1/\left(\eta-1\right)}M) \approx 2\%$, whereupon consumers have no eligible route out of starvation.
-Compellingly, this dynamic bound determined by the rate of energetic recovery is close to the minimum observed mammalian body size of ca.
-1.3-2.5 grams (Fig.
-\ref{fig:mass}b,c), a range that occurs as the recovery rate begins its decline.
-In addition to known transport limitations \cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.
-This work mirrors other efforts where coincident limitations seem to limit the smallest possibilities for life within a particular class or organisms \cite{kempes}.
+A lower bound on mammalian body size is given by $\epsilon=1$, where mammals
+have no metabolic reserves and immediately starve; this occurs at a size of
+\sid{$M=value$}.  This calculation \sid{what calculation?} gives an extreme
+limit on size but does not account for the subtleties of starvation dynamics
+that may limit body size. The NSM correctly predicts that species with
+smaller masses have larger steady-state population densities.  However we
+observe that there is a sharp change in the mass dependence of both the
+steady-state densities and $\sigma/\rho$ at $M \approx 0.3$ grams
+(Fig.~\ref{fig:mass}a,b).  The dependence of the rates of starvation and
+recovery explain this phenomenon.  As the mass decreases, the rate of
+starvation increases, while the rate of recovery declines super-exponentially
+\sid{how do we know this?}.  This decline in $\rho$ occurs when the
+percentage of body fat is
+$1 - 1/\left[\big(B_{0}/B_{m}\big)^{1/\left(\eta-1\right)}M\right] \approx
+2\%$,
+whereupon consumers have no eligible route \sid{what does this mean?} out of
+starvation.  Compellingly, this dynamic bound determined by the rate of
+energetic recovery is close to the minimum observed mammalian body size of
+ca. 1.3-2.5 grams (Fig.~\ref{fig:mass}b,c), a range that occurs as the
+recovery rate begins its decline.  In addition to known transport
+limitations~\cite{West:2002ud}, we suggest that an additional constraint of
+lower body size stems from the dynamics of starvation.  This work \sid{which
+  work?} mirrors other efforts where coincident limitations seem to limit the
+smallest possibilities for life within a particular class or
+organisms~\cite{kempes}.
 
@@
 %Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
-We examine a potential upper bound to body mass by assessing population invasibility with respect to a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-0.5,0.5]$, thus altering rates of starvation, recovery, and maintenance $\beta$.
-Although there is not an internal fixed point where both residents and invaders coexist (except for the trivial state $\chi=0$), we assess invasibility, as a function of organismal mass, by determining which consumer has a higher steady state given $\chi$.
-We find that for $1\leq M<10^6$g, having additional body fat ($\chi > 0$) results in a higher steady state density for the invader population ($H^{\prime *}+F^{\prime *}>H^*+F^*$), such that it has an intrinsic advantage over the resident population.
-However, for $M>10^6$, leaner individuals ($\chi < 0$) have an advantage, and this is due to the changing covariance between energetic rates as a function of modified energetic reserves.
+We determine a potential upper bound to body mass by assessing the
+susceptibility of an otherwise homogeneous population to invasion by a
+mutated subset of the population (denoted by ${}^\prime$) where individuals
+have a modified proportion of body fat $M^\prime=M(1+\chi)$ where
+$\chi \in [-0.5,0.5]$, thus altering the rates of starvation $\sigma$,
+recovery $\rho$, and maintenance $\beta$.  There is no internal fixed point
+that correspond to a state where both original residents and invaders coexist
+(except for the trivial state $\chi=0$).  To assess the susceptibility to
+invasion as a function of the invader mass, we determine which consumer has a
+higher steady-state density for a given value of $\chi$.  We find that for
+$1\leq M<10^6$g, having additional body fat ($\chi > 0$) results in a higher
+steady-state invader population density
+($H^{\prime *}+F^{\prime *}>H^*+F^*$).  Thus the invader has an intrinsic
+advantage over the resident population.  However, for $M>10^6$, leaner
+individuals ($\chi < 0$) have the advantage, and this is due to the changing
+covariance between energetic rates as a function of modified energetic
+reserves \sid{I don't understand the phrase after the comma}.
 
-The observed switch in invasibility as a function of $\chi$ at $M_{\rm opt}\approx10^6$ thus serves as an attractor, where over evolutionary time the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
-Moreover, $M_{\rm opt}$, which is entirely determined by the population-level consequences of energetic constraints is within an order of magnitude as that observed in the North American mammalian fossil record \cite{Alroy:1998p1594} and as that predicted from an evolutionary model of body size evolution \cite{Clauset:2009fh}.
-While the state of the environment, as well as the competitive landscape, will determine whether specific body size sizes are selected for or against \cite{Saarinen:2014br}, we suggest that the starvation dynamic proposed here may supply the fundamental momentum fueling the evolution of larger body size among terrestrial mammals.
+The observed switch in susceptibility as a function of $\chi$ at
+$M_{\rm opt}\approx10^6$ thus serves as an attractor, where over evolutionary
+times the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and
+decrease if $M>M_{\rm opt}$.  Moreover, $M_{\rm opt}$, which is entirely
+determined by the population-level consequences of energetic constraints, is
+within an order of magnitude of the mass observed in the North American
+mammalian fossil record~\cite{Alroy:1998p1594} and also the mass predicted
+from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
+While the state of the environment, as well as the competitive landscape,
+will determine whether specific body sizes are selected for or
+against~\cite{Saarinen:2014br}, we suggest that the starvation dynamic
+proposed here may provide the driving mechanism for the evolution of larger
+body size among terrestrial mammals.
 %and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
@@
 %Closing
-The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations \cite{Stearns:1989ip}.
-The NSM is a minimal and general model that incorporates the dynamics of starvation that are expected to occur in resource limited environments.
-By incorporating allometric relationships between the rates in the NSM, we find
-\emph{i}) different organismal masses are more or less prone to different population dynamic regimes,
-\emph{ii}) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and
-\emph{iii}) the dynamic consequences of these rates may place additional barriers on the evolution of minimum and maximum body size.
-We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.
-Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation.
+The energetics associated with somatic maintenance, growth, and reproduction
+are important elements that influence the dynamics of all
+populations~\cite{Stearns:1989ip}.  The NSM is a minimal and general model
+that incorporates the dynamics of starvation that are expected to occur in
+resource-limited environments.  By incorporating allometric relations between
+the rates in the NSM, we find: (i) different organismal masses have distinct
+population dynamic regimes, (ii) allometrically-determined rates of
+starvation and recovery appear to minimize extinction risk, and (iii) the
+dynamic consequences of these rates may place additional barriers on the
+evolution of minimum and maximum body size.  We suggest that the NSM offers a
+means by which the dynamic consequences of energetic constraints can be
+assessed using macroscale interactions between and among species.  Future
+efforts will involve exploring the consequences of these dynamics in a
+spatially explicit framework, thus incorporating elements such as movement
+costs and spatial heterogeneity, which may elucidate additional tradeoffs
+associated with the dynamics of starvation.
 
@@
 \begin{acknowledgments}
-C.P.K was supported by a Trump Fellowship from the American League of Conservatives.
+  C.P.K was supported by a Trump Fellowship from the American League of
+  Conservatives.  S.R. was supported by grants DMR-1608211 and 1623243 from
+  the National Science Foundation, and by the John Templeton Foundation.
 \end{acknowledgments}

### 2016-07-26 Manuscript/3.tex
sid updates

@@
 dynamics will develop as the Hopf bifurcation is approached.
+
+\sid{mention  and refer to the starving random walk model somewhere.}
 \\
@@
 \cite{Kempes:2012hy,more}\sid{fix Ref}. More specifically, the interspecific
-\sid{what does this word mean?}  trends in growth rate can be approximated by
+\sid{what does interspecific mean?}  trends in growth rate can be
+approximated by
 \begin{eqnarray}
@@
 To determine the starvation rate, $\sigma$, we are interested in the time
-required for an organism to go from its replete \sid{replete vs. adult} state
-to a reduced-size hungry state where reproduction is impossible.  This
-transition time can be inferred from the energy balance in Eq.~\ref{balance},
-where we make the basic assumption that an organism must meet its maintenance
+required for an organism to go from a mature adult that reproduces at rate
+$\lambda$ (henceforth we term this state as the ``replete'' state), to a
+reduced-size hungry state where reproduction is impossible.  This transition
+time can be inferred from the energy balance in Eq.~\ref{balance}, where we
+make the basic assumption that an organism must meet its maintenance
 requirements using the digestion of existing mass as the sole energy source.
@@
 cannot be degraded to fuel metabolism), we define the rate for starvation and
-death by the timescales required to reach specific fractions of normal adult
-mass. We define $m_{\rm starve}=\epsilon M$, where $\epsilon<1$ is the
-fraction of adult mass where reproduction ceases.  This fraction will be
-modified if tissue composition systematically scales with adult mass.  For
-example, making use of the observation that body fat in mammals scales with
-overall body size according to $M_{f}=f_{0}M^{\gamma}$ and assuming that once
-this mass is fully digested the organism begins to starve, this would imply
-that $\epsilon=1-f_{0}M^{\gamma}/M$.  Using this criterion in Eq.~\ref{mt},
-the time scale for starvation is given by
+death by the timescales required to reach specific fractions of the
+replete-state mass.  We define $m_{\rm starve}=\epsilon M$, where
+$\epsilon<1$ is the fraction of replete-state mass where reproduction ceases.
+This fraction will be modified if tissue composition systematically scales
+with adult mass.  For example, making use of the observation that body fat in
+mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$
+and assuming that once this mass is fully digested the organism begins to
+starve, this would imply that $\epsilon=1-f_{0}M^{\gamma}/M$.  Using this
+criterion in Eq.~\ref{mt}, the time scale for starvation is given by
 \begin{equation}
@@
 \end{equation}
-The starvation rate is then $\sigma=1/t_{\sigma}$, which scales with replete
-mass as $1/\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important feature is
-that $\sigma$ does not have a simple scaling dependence on $\lambda$
-(Eq.~\ref{lambda}), which is important for the dynamics that we later
-discuss.
+The starvation rate is then $\sigma=1/t_{\sigma}$, which scales with
+replete-state mass as $1/\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important
+feature is that $\sigma$ does not have a simple scaling dependence on
+$\lambda$ (Eq.~\ref{lambda}), which is important for the dynamics that we
+later discuss.
 
 The time to death should follow a similar relation, but defined by a lower
-fraction of adult mass, $m_{\rm death}=\epsilon^{\prime} M$.  Suppose, for
-example, that an organism dies once it has digested all fat and muscle
-tissues, and that muscle tissue scales with body mass according to
+fraction of replete-state mass, $m_{\rm death}=\epsilon^{\prime} M$.
+Suppose, for example, that an organism dies once it has digested all fat and
+muscle tissues, and that muscle tissue scales with body mass according to
 $M_{mm}=mm_{0}M^{\zeta}$ \sid{what is $mm_0$?}.  This gives
@@
 state to the full state.  We again use the balance given in Eq.~\ref{balance}
-to find the timescale to return to the replete mass from a given reduced
-starvation mass.  From the solution to Eq.~\ref{balance}
+to find the timescale to return to the replete-state mass from a given
+reduced starvation mass.  From the solution to Eq.~\ref{balance}
 \begin{eqnarray}
@@
 between $10^2$ and $\leq 10^6$.  This procedure is repeated for 1000
-replicates \sid{what is changing? isn't the system deterministic?} of the
-continuous-time system shown in Eq.~\ref{eq:system} for an organism of
-$M=100$ grams, assuming random initial conditions around the steady state
-(Eq.~\ref{eq:ss}). %random initial conditions within the interval $(R^*,H^*,F^*)\pm100\%$.
-By allowing the rate of starvation to vary, we assessed extinction risk
-across a range of values of the ratio $\sigma/\rho$ varying between $10^{-2}$
-to $2.5$, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.  As
-expected, higher rates of extinction correlated with both low and high values
-of $\sigma/\rho$.  For low values of $\sigma/\rho$, the increased extinction
-risk results from transient cycles with larger amplitudes as the system nears
-the Hopf bifurcation (Fig.~\ref{fig:ext}).  For large values of
-$\sigma/\rho$, higher extinction risk arises because of to the decrease in
-the steady state consumer population density.  This interplay creates an
-`extinction refuge' as shown in Fig.~\ref{fig:ext}, such that for a
-relatively constrained range of $\sigma/\rho$, extinction probabilities are
-minimized.
+replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an
+organism of $M=100$ grams.  In each replicate the initial condition is
+distributed around the steady state (Eq.~\ref{eq:ss}).  Specifically the
+initial densities are chosen to be $A(F^*,H^*,R^*)$, with $A$ a random
+variable that is uniformly distribution in $[0,2]$.  By allowing the rate of
+starvation to vary, we assessed extinction risk across a range of values of
+the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a
+horizontal cross-section of Fig.~\ref{fig:fp}.  As expected, higher rates of
+extinction correlated with both low and high values of $\sigma/\rho$.  For
+low values of $\sigma/\rho$, the increased extinction risk results from
+transient cycles with larger amplitudes as the system nears the Hopf
+bifurcation (Fig.~\ref{fig:ext}).  For large values of $\sigma/\rho$, higher
+extinction risk arises because of to the decrease in the steady state
+consumer population density.  This interplay creates an `extinction refuge'
+as shown in Fig.~\ref{fig:ext}, such that for a relatively constrained range
+of $\sigma/\rho$, extinction probabilities are minimized.
 

### 2016-07-29 Manuscript/4.tex
Kempes edits

--- 2016-07-26 Manuscript/3.tex
@@
 capacity equal to one.  Consumers transition from the full state $F$ to the
-hungry state $H$ by starvation at rate $\sigma$ and also in proportion to the
-absence of resources $(1-R)$.  Conversely, consumers recover from state $H$
-to state $F$ at rate $\rho$ and in proportion to $R$.  Resources are also
-eaten by the consumers---at rate $\rho$ by hungry consumers and at rate
-$\beta<\rho$ by full consumers.  This inequality accounts for hungry
-consumers requiring more resources to rebuild body weight.
+hungry state $H$ at a rate $\sigma$---the starvation rate---and also in
+proportion to the absence of resources $(1-R)$.  Conversely, consumers
+recover from state $H$ to state $F$ at rate $\rho$ and in proportion to $R$.
+Resources are also eaten by the consumers---at rate $\rho$ by hungry
+consumers and at rate $\beta<\rho$ by full consumers.  This inequality
+accounts for hungry consumers requiring more resources to rebuild body
+weight.
 
@@
 at any point in development.  It is useful to explicitly write this balance
-because it can also be modified to understand the timescales of both
-starvation and recovery from starvation as we show below.
+because it can also be modified to understand the timescales of starvation as
+well as recovery from starvation as we show below.
 
@@
 $\lambda$ (henceforth we term this state as the ``replete'' state), to a
-reduced-size hungry state where reproduction is impossible.  This transition
+reduced-mass hungry state where reproduction is impossible.  This transition
 time can be inferred from the energy balance in Eq.~\ref{balance}, where we
@@
 mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$
-and assuming that once this mass is fully digested the organism begins to
-starve, this would imply that $\epsilon=1-f_{0}M^{\gamma}/M$.  Using this
-criterion in Eq.~\ref{mt}, the time scale for starvation is given by
+and assuming that once this mass is fully digested the organism starves, this
+would imply that $\epsilon=1-f_{0}M^{\gamma}/M$.  Using this criterion in
+Eq.~\ref{mt}, the time scale for starvation is given by
 \begin{equation}
@@
 The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
-sufficient tissue to transition from the starving \sid{starving vs.\ hungry}
-state to the full state.  We again use the balance given in Eq.~\ref{balance}
-to find the timescale to return to the replete-state mass from a given
-reduced starvation mass.  From the solution to Eq.~\ref{balance}
+sufficient tissue to transition from the hungry to the full state.  We again
+use the balance given in Eq.~\ref{balance} to find the timescale for an
+organism to return to the replete-state mass from a given reduced
+hungry-state mass.  From the solution to Eq.~\ref{balance}
 \begin{eqnarray}
@@
 2\%$,
-whereupon consumers have no eligible route \sid{what does this mean?} out of
+whereupon consumers have no eligible route \sid{what does this mean?} that avoids
 starvation.  Compellingly, this dynamic bound determined by the rate of
@@
 will determine whether specific body sizes are selected for or
-against~\cite{Saarinen:2014br}, we suggest that the starvation dynamic
+against~\cite{Saarinen:2014br}, we suggest that the starvation dynamics
 proposed here may provide the driving mechanism for the evolution of larger
@@
 \includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-\caption{ % 58 words
+\caption{
+The growth trajectory over absolute time of an individual organism as a function of body mass.
+Initial growth follows the red trajectory to an energetically replete adult mass $M$.
+Starvation follows the concave blue trajectory to $m_{\rm starve} < M$, whereas recovery follows the convex growth trajectory from $m_\sigma$ to $M$.
 }
@@
 \includegraphics[width=0.5\textwidth]{fig_GvS.pdf}
-\caption{ % 58 words
+\caption{
+Allometrically constrained starvation rate $\sigma$ (red) vs. reproductive rate $\lambda$ (yellow) as a function of mass $M$.
+The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
 }
@@
 \includegraphics[width=0.5\textwidth]{fig_FixedPoint.pdf}
-\caption{ % 58 words
+\caption{
+The transcritical (TC; dashed line) and Hopf bifurcation (solid line) as a function of the starvation rate $\sigma$ and recovery rate $\rho$.
+These bifurcation conditions separate parameter space into infeasible, cyclic, and steady state dynamic regimes.
+The color gradient shows the steady state densities for (A) the resource $R^*$ and the (B) energetically replete consumers $F^*$, with warm colors denoting higher densities and cool colors denoting lower densities.
+Steady state densities for the energetically deficient consumers $H^*$ are not shown because they closely mirror those for $F^*$.
 }
@@
 \includegraphics[width=0.5\textwidth]{fig_DataHopf.pdf}
-\caption{ % 58 words
+\caption{
+Transcritical (TC; vertical lines) and Hopf bifurcations (curved lines) with allometrically determined starvation $\sigma$ and recovery $\rho$ rates as a function of minimum and maximum mammalian body sizes: $1$ gram (blue) and $10^7$ grams (orange), respectively.
+Replicates show the influence of variation (20\% around the mean) on allometric parameters, which influences both the energetic rates as well as the position of the TC and Hopf bifurcations.
 }
@@
 \includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric.pdf}
-\caption{ % 58 words
+\caption{
+The probability of extinction for 1000 consumer population trajectories as a function of $\sigma/\rho$ within initial densities chosen to as $A(F^*,H^*,R^*)$, with $A$ a random
+variable that is uniformly distribution in $[0,2]$.
+Extinction is defined as the population trajectory going below
+$0.2~\times$ the allometrically constrained steady state for all times
+between $10^2$ and $\leq 10^6$.
+The values above the extinction plot are the allometrically constrained $\sigma/\rho$ with 20\% variation around the mean.
 }
@@
 \includegraphics[width=0.45\textwidth]{fig_FPAllometric.pdf}
-\caption{ % 58 words
+\caption{
+(\emph{A}) Consumer steady states as a function of body size, showing both energetically deficient and replete consumer states ($H^*$ and $F^*$, respectively). 
+Energetic rates as a function of body size, with the ratio $\sigma/\rho$ (\emph{B}) and both $\sigma$ (red) and $\rho$ (purple; \emph{C}) drawn separately with 20\% variation around the mean.
+Steady state densities decline sharply at $M=M_{\rm min}$ due to the super-exponential decrease in the rate of recovery.
+The minimum body size observed for mammals (the Etruscan shrew) is denoted by the orange shaded region at values marking the initial decline of the recovery rate.
 }
@@
 \includegraphics[width=0.5\textwidth]{fig_Invasion.pdf}
-\caption{ % 58 words
+\caption{
+Invasion feasibility for organisms with a proportional change in mass $\chi$ against a population with a resident body mass $M$.
+The blue region denotes which values of $\chi$ result in successful invasion.
+The red region denotes which values of $\chi$ result in a mass that is below the starvation threshold and is thus infeasible.
 }

### 2016-07-29 Manuscript/5.tex
Kempes edits

--- 2016-07-29 Manuscript/4.tex
@@
 growth rates of organisms ranging from bacteria to metazoans
-\cite{Kempes:2012hy,more}\sid{fix Ref}. More specifically, the interspecific
-\sid{what does interspecific mean?}  trends in growth rate can be
+\cite{Kempes:2012hy,more}\sid{fix Ref}. More specifically, the cross-species trends in growth rate can be
 approximated by
@@
 requirements using the digestion of existing mass as the sole energy source.
-This assumption implies the simple metabolic balance \sid{don't understand
-  this equation}
-\begin{eqnarray}
+This assumption implies a simple metabolic balance where the left hand side of Equation \ref{balance} becomes zero and we have \begin{eqnarray}
 \frac{dm}{dt}E_{m}^{\prime}=-B_{m}m

### 2016-08-02 Manuscript/6.tex
Kempes second round of edits

--- 2016-07-29 Manuscript/5.tex
@@
 consumer metabolism, which can be used to describe a variety of organismal
-features~\cite{}\sid{Ref}.  The scaling relation between an organism's
+features \cite{Brown:2004wq}.  The scaling relation between an organism's
 metabolic rate $B$ and its body size at reproductive maturity $M$ is well
@@
 the scaling exponent, generally assumed to vary around $2/3$ or $3/4$ for
-metazoans~\cite{}\sid{Ref}, and has taxonomic shifts for unicellular species
+metazoans (e.g. \cite{Brown:2004wq}), and has taxonomic shifts for unicellular species
 between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria
@@
 growth rates of organisms ranging from bacteria to metazoans
-\cite{Kempes:2012hy,more}\sid{fix Ref}. More specifically, the cross-species trends in growth rate can be
+\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. More specifically, the cross-species trends in growth rate can be
 approximated by
@@
 \end{eqnarray}
-This relation is derived from the simple balance condition
+This relation is derived from the simple balance condition \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}
 \begin{eqnarray}
@@
 the metabolic rate to support an existing unit of mass, and $m$ is the mass
-at any point in development.  It is useful to explicitly write this balance
-because it can also be modified to understand the timescales of starvation as
-well as recovery from starvation as we show below.
+at any point in development.  This balance has the general solution \cite{moses2008rmo, Kempes:2012hy}
+\begin{eqnarray}
+m\left(t\right)\!=\left[1\!-\!\left(1\!-\!\frac{b}{a}m_{0}^{1\!-\!\eta}\right)e^{-b\left(1\!-\!\eta\right)t}\right]^{1/\left(1-\eta\right)}\!\left(\!\frac{a}{b}\!\right)^{1/(1-\eta)}
+\end{eqnarray}
+where, $a=B_{0}/E_{m}$ and $b=B_{m}/E_{m}$ , andwhich we use to define the timescale of recovery from starvation. The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
+sufficient tissue to transition from the hungry to the full state.  Since only certain tissues can be digested for energy (for example the brain
+cannot be degraded to fuel metabolism), we define the rate for starvation,
+death, and recovery by the timescales required to reach, or return from, specific fractions of the
+replete-state mass. We define $m_{\rm starve}=\epsilon M$, where
+$\epsilon<1$ is the fraction of replete-state mass where reproduction ceases. This fraction will be modified if tissue composition systematically scales
+with adult mass.  For example, making use of the observation that body fat in
+mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$
+and assuming that once this mass is fully digested the organism starves, this
+would imply that $\epsilon=1-f_{0}M^{\gamma}/M$. It follows that the recovery timescale, $t_{\rho}=t_{2}-t_{1}$, is the time it takes
+to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, or
+\begin{eqnarray}
+t_{\rho}=\frac{\ln\left[1\!-\!\frac{a}{b}\left(M
+  \right)^{1\!-\!\eta }\right]
+-\ln \left[1\!-\!\frac{a}{b}\left(\epsilon M \right)^{1\!-\!\eta }\right]}{(\eta -1) b}.
+\end{eqnarray}
+It should be noted that more complicated ontogenetic models explicitly handle storage \cite{hou}, whereas this is implicitly covered by the body fat scaling in our framework.
 
@@
 $\lambda$ (henceforth we term this state as the ``replete'' state), to a
-reduced-mass hungry state where reproduction is impossible.  This transition
-time can be inferred from the energy balance in Eq.~\ref{balance}, where we
-make the basic assumption that an organism must meet its maintenance
-requirements using the digestion of existing mass as the sole energy source.
-This assumption implies a simple metabolic balance where the left hand side of Equation \ref{balance} becomes zero and we have \begin{eqnarray}
+reduced-mass hungry state where reproduction is impossible.
+For starving individuals we assume that an organism must meet its maintenance requirements using the digestion of existing mass as the sole energy source.
+This assumption implies the following simple metabolic balance 
+\begin{eqnarray}
 \frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
@@
 where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
-body mass which differs from $E_{m}$~\cite{}\sid{Ref}, the energy required to
-synthesis a unit of biomass. Given the replete mass, $M$, of an organism, the
+body mass which differs from $E_{m}$, the energy required to
+synthesis a unit of biomass \cite{hou}. Given the replete mass, $M$, of an organism, the
 above energy balance prescribes the mass trajectory of a non-consuming
@@
 \label{mt}
-m\left(t\right)=Me^{-B_{m}t/E_{m}^{\prime}}.
+m\left(t\right)=Me^{-b^{\prime}t}
 \end{eqnarray}
-Since only certain tissues can be digested for energy (for example the brain
-cannot be degraded to fuel metabolism), we define the rate for starvation and
-death by the timescales required to reach specific fractions of the
-replete-state mass.  We define $m_{\rm starve}=\epsilon M$, where
-$\epsilon<1$ is the fraction of replete-state mass where reproduction ceases.
-This fraction will be modified if tissue composition systematically scales
-with adult mass.  For example, making use of the observation that body fat in
-mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$
-and assuming that once this mass is fully digested the organism starves, this
-would imply that $\epsilon=1-f_{0}M^{\gamma}/M$.  Using this criterion in
-Eq.~\ref{mt}, the time scale for starvation is given by
+where $b^{\prime}=E_{m}^{\prime}/B_{m}$. The time time scale for starvation is given by the time it takes this decay to reach $\epsilon M$, which is
 \begin{equation}
 \label{eq:sigma}
-t_{\sigma}=-\frac{E_{m}\ln\left(\epsilon\right)}{B_{m}}.
+t_{\sigma}=-b^{\prime}\ln\left(\epsilon\right).
 \end{equation}
@@
 muscle tissues, and that muscle tissue scales with body mass according to
-$M_{mm}=mm_{0}M^{\zeta}$ \sid{what is $mm_0$?}.  This gives
-$\epsilon^{\prime}=1-\left(f_{0}M^{\gamma}+mm_{0}M^{\zeta}\right)/M$. Muscle
+$M_{m}=u_{0}M^{\zeta}$.  This gives
+$\epsilon^{\prime}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
 mass has been shown to be roughly proportional to body mass~\cite{muscle} in
@@
 \begin{eqnarray}
-t_{\mu}=-\frac{E_{m}\ln\left(\epsilon^{\prime}\right)}{B_{m}}
+t_{\mu}=-b^{\prime}\ln\left(\epsilon^{\prime}\right)
 \end{eqnarray}
@@
 
-The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
-sufficient tissue to transition from the hungry to the full state.  We again
-use the balance given in Eq.~\ref{balance} to find the timescale for an
-organism to return to the replete-state mass from a given reduced
-hungry-state mass.  From the solution to Eq.~\ref{balance}
-\begin{eqnarray}
-m\left(t\right)\!=\!\left(\!\frac{B_{0}}{B_{m}}\!\right)^{1/(\eta-1)}\!\!
-\left[1\!-\!\left(1\!-\!\frac{B_{m}}{B_{0}}m_{0}^{1\!-\!\eta}\right)e^{-b\left(1\!-\!\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}\,,
-\end{eqnarray}
-we require the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes
-to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, or
-\begin{eqnarray}
-t_{\rho}=\frac{E_{m}\left\{\ln \left[1\!-\!\frac{B_{0}}{B_{m}}\left(M
-  \right)^{1\!-\!\eta }\right]
--\ln \left[1\!-\!\frac{B_{0}}{B_{m}}\left(\epsilon M \right)^{1\!-\!\eta }\right]\right\}}{(\eta -1) B_{m}}.
-\end{eqnarray}
 %\left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}
@@
 percentage of body fat is
-$1 - 1/\left[\big(B_{0}/B_{m}\big)^{1/\left(\eta-1\right)}M\right] \approx
+$1 - 1/\left[\big(a/b\big)^{1/\left(\eta-1\right)}M\right] \approx
 2\%$,
@@
 smallest possibilities for life within a particular class or
-organisms~\cite{kempes}.
+organisms~\cite{Kempes:2012hy}.
 
@@
 \begin{acknowledgments}
+  J.D.Y. was supported by an Omidyar postdoctoral Fellowship at the Santa Fe Institute. 
   C.P.K was supported by a Trump Fellowship from the American League of
-  Conservatives.  S.R. was supported by grants DMR-1608211 and 1623243 from
+  Conservatives.  
+  S.R. was supported by grants DMR-1608211 and 1623243 from
   the National Science Foundation, and by the John Templeton Foundation.
@@
 %
-\clearpage
+%\clearpage
 
-\begin{table}[h]
-\caption{Parameter Values For Various Classes of Organisms}
-\label{param}
-    \begin{center}
-    \small
-     \begin{tabular}{| p{1.2cm}| p{3.2cm} | p{2.6cm} | p{3.2cm} | }
-     \hline
-     & {\bf Mammals} & {\bf Unicellular Eukaryotes} & {\bf Bacteria} \\
-     \hline
-   $\eta$ & $3/4$ & & $1.70$ \\ 
-   $E_{m}$ & $10695$ (J gram$^{-1}$) & & $10695$ (J gram$^{-1}$) \\ 
-   $E_{m}^{\prime}$ & $\approx E_{m}$ & & $\approx E_{m}$ \\ 
-   $B_{0}$ & $0.019$ (W gram$^{-\alpha}$) & & $1.96\times10^{17}$ \\
-   $B_{m}$ & $0.025$ (W gram$^{-1}$)   & & $0.025$ (W gram$^{-1}$)\\
-   $a$ & $1.78\times10^{-6}$ & & $1.83\times10^{13}$ \\ 
-   $b$ & $2.29\times10^{-6}$ & & $2.29\times10^{-6}$ \\  
-   $\eta-1$ & $-0.21$ & & $0.73$ \\ 
-   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) & & $56493$ \\ 
-   $\gamma$ & $1.19$ & & $0.68$ \\ 
-   $f_{0}$ & $0.02$ & & $1.30\times10^{-5}$ \\ 
-   $\zeta$ & $1.01$ & & \\ 
-   $mm_{0}$ & $0.32$ & & \\
-
-   \hline
-    \end{tabular}
-    \end{center}
-   \end{table}
+%\begin{table}[h]
+%\caption{Parameter Values For Various Classes of Organisms}
+%\label{param}
+%    \begin{center}
+%    \small
+%     \begin{tabular}{| p{1.2cm}| p{3.2cm} | p{2.6cm} | p{3.2cm} | }
+%     \hline
+%     & {\bf Mammals} & {\bf Unicellular Eukaryotes} & {\bf Bacteria} \\
+%     \hline
+%   $\eta$ & $3/4$ & & $1.70$ \\ 
+%   $E_{m}$ & $10695$ (J gram$^{-1}$) & & $10695$ (J gram$^{-1}$) \\ 
+%   $E_{m}^{\prime}$ & $\approx E_{m}$ & & $\approx E_{m}$ \\ 
+%   $B_{0}$ & $0.019$ (W gram$^{-\alpha}$) & & $1.96\times10^{17}$ \\
+%   $B_{m}$ & $0.025$ (W gram$^{-1}$)   & & $0.025$ (W gram$^{-1}$)\\
+%   $a$ & $1.78\times10^{-6}$ & & $1.83\times10^{13}$ \\ 
+%   $b$ & $2.29\times10^{-6}$ & & $2.29\times10^{-6}$ \\  
+%   $\eta-1$ & $-0.21$ & & $0.73$ \\ 
+%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) & & $56493$ \\ 
+%   $\gamma$ & $1.19$ & & $0.68$ \\ 
+%   $f_{0}$ & $0.02$ & & $1.30\times10^{-5}$ \\ 
+%   $\zeta$ & $1.01$ & & \\ 
+%   $mm_{0}$ & $0.32$ & & \\ 
+%   
+%      
+%   \hline
+%    \end{tabular}
+%    \end{center}
+%   \end{table}
 

### 2016-08-05 Manuscript/7.tex
Yeakel version of ms, and updated figures for ef

--- 2016-08-02 Manuscript/6.tex
@@
 %\usepackage{tikz}
-%\usepackage{amsmath}
-%\usepackage{amssymb}
 %\usepackage{array}
@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 
@@
 \begin{abstract}
-This is the abstract.  \emph{ \sid{No it isn't.  It's merely a placeholder.}}
+We introduce a minimal foraging model that incorporates two classes of
+foragers: nutritionally replete foragers that can procreate, and
+undernourished and non-procreating foragers that are susceptible to
+mortality.  As a function of the transition rates between these replete and
+undernourished states that are determined by the presence or absence of
+resources, as well as the replenishment rate of the underlying resource, the
+forager populations can either undergo cyclic dynamics or reach a steady
+state.  By appplying basic allometric considerations, we obtain strong
+constraints on these transition rates and thereby find that the population
+dynamics that is subject to these constraints is typically driven to a steady
+state.  As a further consequence of this viewpoint, we find that
+allometrically constrained rates fall within a `refuge' in parameter space,
+where the probability of extinction of the consumer population is minimized.
+Thus our model provides a mechanism that may both drive and constrain natural
+animal populations.  We also find that our model provides a natural framework
+to determine a natural size for animal populations by examining the relative
+stability of a otherwise homoegeneous population to a mutant population of a
+different size.  For body masses $\lesssim 10^6$g, a more massive invader
+dominates over the resident population, and vice versa for body mass $\gtrsim
+10^6$g, thus providing a simple principled mechanism for Cope's rule.
+
 \end{abstract}
@@
 percentage of body fat is
-$1 - 1/\left[\big(a/b\big)^{1/\left(\eta-1\right)}M\right] \approx
+$1 - \big(a/b\big)^{-1/\left(\eta-1\right)}M^{-1} \approx
 2\%$,
@@
 %and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
+
+A potential critique of our results, which show that larger mammals are less susceptible to extinction, is why we don't observe a greater number of large mammals in the modern world.
+However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which where also much more geographically widespread than today.
+These results combined with our findings suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
 
@@
 
-\bibliography{aa_starving}
+\bibliography{aa_starving2}
 
 \begin{acknowledgments}
-  J.D.Y. was supported by an Omidyar postdoctoral Fellowship at the Santa Fe Institute. 
-  C.P.K was supported by a Trump Fellowship from the American League of
-  Conservatives.  
-  S.R. was supported by grants DMR-1608211 and 1623243 from
-  the National Science Foundation, and by the John Templeton Foundation.
+  J.D.Y. was supported by an Omidyar postdoctoral Fellowship at the Santa Fe
+  Institute and by startup funds at the University of California, Merced.
+  C.P.K was supported by an Omidyar postdoctoral Fellowship at the Santa Fe
+  Institute.  S.R. was supported by grants DMR-1608211 and 1623243 from the
+  National Science Foundation, and by the John Templeton Foundation, all at
+  the Santa Fe Institute.
 \end{acknowledgments}

### 2016-08-05 Manuscript/8.tex
Yeakel version of ms, and updated figures for ef

--- 2016-08-02 Manuscript/6.tex
@@
 
-\begin{abstract}
-This is the abstract.  \emph{ \sid{No it isn't.  It's merely a placeholder.}}
+\begin{abstract} %246 words
+The eco-evolutionary dynamics of species is fundamentally linked to the energetic constraints of individuals.
+Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery in resource limited environments.
+We introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete consumers that reproduce, and
+undernourished, non-reproducing consumers that are susceptible to
+mortality.  
+As a function of the energetic rates between these replete and
+undernourished states that are determined by the presence or absence of
+resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
+We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships and find that population dynamics subject to these constraints can approach the cyclic regime but are typically driven to a steady state.
+Moreover, we find these rates fall within a `refuge' in parameter space, where the probability of extinction of the consumer population is minimized.
+Thus we introduce a potential mechanism that may both drive and constrain the dynamics of animal populations.
+Importantly, our model provides a natural framework that predicts maximum body size for mammals by examining the relative stability of an otherwise homogeneous population to a mutant population with altered percent body fat.
+For body masses $\lesssim 10^6$g, an invader with increased percent body fat dominates over the resident population, and vice versa for body mass $\gtrsim 10^6$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
 \end{abstract}
@@
 The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation
-that demarcates the physical and unphysical regimes\sid{give Ref}.  The
+that demarcates the physical and unphysical regimes.  The
 biological implication of the constraint $\lambda<\sigma$ has a simple
@@
 starvation rate $\sigma$ defines an absolute bound of biological
-feasibility---the TC bifurcation---the starvation rate $\sigma$ also
+feasibility---the TC bifurcation---$\sigma$ also
 determines the sensitivity of the consumer population to changes in resource
@@
 dynamics will develop as the Hopf bifurcation is approached.
-
-\sid{mention  and refer to the starving random walk model somewhere.}
 \\
@@
 data for larger organisms.
-\\
+\vspace{6mm}
 %, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
@@
 %Extinction to the left and the right
-
 \noindent {\bf Extinction risk}\\
@@
 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an
-organism of $M=100$ grams.  In each replicate the initial condition is
+organism of $M=100$ grams.  
+In each replicate the initial condition is
 distributed around the steady state (Eq.~\ref{eq:ss}).  Specifically the
@@
 the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a
-horizontal cross-section of Fig.~\ref{fig:fp}.  As expected, higher rates of
+horizontal cross-section of Fig.~\ref{fig:fp}.  
+As expected, higher rates of
 extinction correlated with both low and high values of $\sigma/\rho$.  For
@@
 
-\bibliography{aa_starving}
+\bibliography{aa_starving3}
 
 \begin{acknowledgments}
-  J.D.Y. was supported by an Omidyar postdoctoral Fellowship at the Santa Fe Institute. 
-  C.P.K was supported by a Trump Fellowship from the American League of
-  Conservatives.  
-  S.R. was supported by grants DMR-1608211 and 1623243 from
-  the National Science Foundation, and by the John Templeton Foundation.
+We thank Luis Bettancourt, Eric Libby, Seth Newsome, et al. for helpful discussions and comments that greatly increased the quality of the manuscript.
+  J.D.Y. was supported by startup funds at the University of California, Merced, and an Omidyar postdoctoral Fellowship at the Santa Fe Institute.
+  C.P.K. was supported by an Omidyar postdoctoral Fellowship at the Santa Fe
+  Institute.  
+  S.R. was supported by grants DMR-1608211 and 1623243 from the National Science Foundation, and by the John Templeton Foundation, all at the Santa Fe Institute.
 \end{acknowledgments}

### 2016-08-09 Manuscript/8.tex
Manuscript and body size invasion stuff edits JY

@@
 
-The life history of most species is typically comprised of (a) somatic growth
-and maintenance, and (b) reproduction.  The balance between these two
-activities is often conditioned on resource
+The balance between (a) somatic growth and maintenance, and (b) reproduction is often conditioned on resource
 availability~\cite{Morris:1987eo}.  For example, reindeer invest less in
@@
 exhibit delayed implantation, whereby females postpone fetal development
-(blastocyst implantation) until times where nutritional reserves can be
+(blastocyst implantation) until nutritional reserves can be
 accumulated~\cite{Mead:1989dt,Sandell:1990kw}.  Many other many species
@@
 
-\noindent {\bf Nutritional-state-structured model (NSM)}\\
-We begin by defining a minimal Nutritional-State-structured population Model
+\noindent {\bf Nutritional state-structured model (NSM)}\\
+We begin by defining a minimal Nutritional State-structured population Model
 (NSM), where the consumer population is divided into two energetic states:
 (a) an energetically replete (full) state $F$, where the consumer reproduces
-at a constant rate $\lambda$ and has no mortality risk, and (b) an
+at a constant rate $\lambda$ and does not die from starvation, and (b) an
 energetically deficient (hungry) state $H$, where the consumer does not
-reproduce but dies at rate $\mu$.  The underlying resource $R$ evolves by
+reproduce but dies by starvation at rate $\mu$.  The underlying resource $R$ evolves by
 logistic growth with an intrinsic growth rate $\alpha$ and a carrying
@@
 consumers and at rate $\beta<\rho$ by full consumers.  This inequality
-accounts for hungry consumers requiring more resources to rebuild body
-weight.
+accounts for hungry consumers requiring more resources to replace lost body
+tissue.
 
@@
 features \cite{Brown:2004wq}.  The scaling relation between an organism's
-metabolic rate $B$ and its body size at reproductive maturity $M$ is well
+metabolic rate $B$ and its body size at reproductive maturity $m=M$ (where $m$ denotes body sizes during growth and prior to reproductive maturity) is well
 documented~\cite{West:2002it} and scales as $B = B_0 M^\eta$, where $\eta$ is
@@
 between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria
-\cite{DeLong:2010dy,Kempes:2012hy}. Several efforts have shown how a
-partitioning of this metabolic rate between growth and maintenance purposes
+\cite{DeLong:2010dy,Kempes:2012hy}. 
+%Justin adding an explanation for beta
+An organism's metabolic rate is proportional to the cost of tissue maintenance without growth (i.e.
+at $m=M$) such that $\beta = \xi B$, where $\xi$ is related to the conversion efficiency of resource to consumer tissue (Supporting Information, section XX).
+
+Several efforts have shown how a
+partitioning of $B$ between growth and maintenance purposes
 can be used to derive a general equation for the growth trajectories and
@@
 \label{balance}
-B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
+B_{0}m^{\eta}=E_{m}\dot{m}+B_{m}m\,,
 \end{eqnarray}
@@
 \end{eqnarray}
-where, $a=B_{0}/E_{m}$ and $b=B_{m}/E_{m}$ , andwhich we use to define the timescale of recovery from starvation. The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
+where, $a=B_{0}/E_{m}$, $b=B_{m}/E_{m}$, and $m_0$ is mass at birth, which we use to define the timescale of recovery from starvation.
+The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
 sufficient tissue to transition from the hungry to the full state.  Since only certain tissues can be digested for energy (for example the brain
@@
 death, and recovery by the timescales required to reach, or return from, specific fractions of the
-replete-state mass. We define $m_{\rm starve}=\epsilon M$, where
+replete-state mass. 
+Here and henceforth, see Supporting Information, sections XX-XX for details on parameterizations.
+We define $m_{\rm starve}=\epsilon M$, where
 $\epsilon<1$ is the fraction of replete-state mass where reproduction ceases. This fraction will be modified if tissue composition systematically scales
 with adult mass.  For example, making use of the observation that body fat in
-mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$
+mammals scales with overall body size according to $M_{\rm fat}=f_{0}M^{\gamma}$
 and assuming that once this mass is fully digested the organism starves, this
@@
 required for an organism to go from a mature adult that reproduces at rate
-$\lambda$ (henceforth we term this state as the ``replete'' state), to a
+$\lambda$, to a
 reduced-mass hungry state where reproduction is impossible.
@@
 \begin{eqnarray}
-\frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
+\dot{m}E_{m}^{\prime}=-B_{m}m
 \end{eqnarray}
@@
 \end{eqnarray}
-where $b^{\prime}=E_{m}^{\prime}/B_{m}$. The time time scale for starvation is given by the time it takes this decay to reach $\epsilon M$, which is
+where $b^{\prime}=E_{m}^{\prime}/B_{m}$. The time scale for starvation is given by the time it takes this decay to reach $\epsilon M$, which is
 \begin{equation}
@@
 muscle tissues, and that muscle tissue scales with body mass according to
-$M_{m}=u_{0}M^{\zeta}$.  This gives
+$M_{\rm musc}=u_{0}M^{\zeta}$.  This gives
 $\epsilon^{\prime}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
-mass has been shown to be roughly proportional to body mass~\cite{muscle} in
+mass has been shown to be roughly proportional to body mass~\cite{Folland:2008ij} in
 mammals and thus $\epsilon^{\prime}$ is merely $\epsilon$ minus a constant.
@@
 \begin{eqnarray}
-t_{\mu}=-b^{\prime}\ln\left(\epsilon^{\prime}\right)
+t_{\mu}=-b^{\prime}\ln\left(\epsilon^{\prime}\right),
 \end{eqnarray}
@@
 \noindent {\bf  Stabilizing effects of allometric constraints} \\
-As the allometric derivations of the NSM rate laws reveal, $\sigma$ and
-$\rho$ are not independent parameters, and the bifurcation space shown in
+As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the bifurcation space shown in
 Fig.~\ref{fig:fp} is navigated via covarying parameters.  Given the
@@
 $\propto M^{0.25}$, however this relation seems to hold only for some species
-\cite{Hendriks:2012fc} and competing explanations [related to ....\sid{???}]
-exist~\cite{Kendall:1999iy,Hogstedt:2005cr}.
+\cite{Hendriks:2012fc}, and potential drivers range from predator and/or prey lifespan to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
 %It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic~\cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
@@
 among mammals is predominantly based on time series for small
-mammals~\cite{Kendall:1998hl}, where we our model would predict much longer
-and more pronounced transient dynamics, given how close these points are to
+mammals~\cite{Kendall:1998hl}, where our model would predict more pronounced transient dynamics, given how close these points are to
 the Hopf bifurcation.  On the other hand, the longer gestational times and
@@
 ratio $\sigma/\rho$.  To this end, we computed the probability of extinction,
-where extinction is defined as the population trajectory going below
+where extinction is defined as the population trajectory falling below
 $0.2~\times$ the allometrically constrained steady state for all times
-between $10^2$ and $\leq 10^6$.  This procedure is repeated for 1000
+between $10^6$ and $\leq 10^8$.  This procedure is repeated for 1000
 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an
 organism of $M=100$ grams.  
-In each replicate the initial condition is
-distributed around the steady state (Eq.~\ref{eq:ss}).  Specifically the
+%the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
+In each replicate the
 initial densities are chosen to be $A(F^*,H^*,R^*)$, with $A$ a random
-variable that is uniformly distribution in $[0,2]$.  By allowing the rate of
+variable that is uniformly distributed in $[0,2]$.  By allowing the rate of
 starvation to vary, we assessed extinction risk across a range of values of
-the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a
+the ratio $\sigma/\rho$ varying between ca. $10^{-3}$ to $5$, thus examining a
 horizontal cross-section of Fig.~\ref{fig:fp}.  
@@
 We find that the allometrically constrained values of $\sigma/\rho$ (with
-$\pm 20\%$ variability around energetic parameter means) fall within the
+$20\%$ variability around energetic parameter means) fall within the
 extinction refuge. These values are close enough to the Hopf bifurcation to
@@
 that a selective mechanism has constrained the physiological conditions that
-drive observed starvation and recovery rates within populations.  Such a
+drive starvation and recovery rates within populations.  Such a
 mechanism would select for organism physiology that generates appropriate
@@
 %Minimal mass bound
-A lower bound on mammalian body size is given by $\epsilon=1$, where mammals
-have no metabolic reserves and immediately starve; this occurs at a size of
-\sid{$M=value$}.  This calculation \sid{what calculation?} gives an extreme
-limit on size but does not account for the subtleties of starvation dynamics
-that may limit body size. The NSM correctly predicts that species with
-smaller masses have larger steady-state population densities.  However we
+%A theoretical lower bound on mammalian body size occurs when $\epsilon=0$ where mammals have no metabolic reserves and immediately starve; this occurs at a size of $M=0$, which is obviously infeasible.
+%Such prediction gives an extreme limit on size but does not account for the subtleties of starvation dynamics
+%that may limit body size (among other factors). 
+The NSM correctly predicts that species with
+smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}a).  However we
 observe that there is a sharp change in the mass dependence of both the
 steady-state densities and $\sigma/\rho$ at $M \approx 0.3$ grams
-(Fig.~\ref{fig:mass}a,b).  The dependence of the rates of starvation and
+(Fig.~\ref{fig:mass}b,c).  The dependence of the rates of starvation and
 recovery explain this phenomenon.  As the mass decreases, the rate of
-starvation increases, while the rate of recovery declines super-exponentially
-\sid{how do we know this?}.  This decline in $\rho$ occurs when the
+starvation increases, while the rate of recovery declines super-exponentially.  This decline in $\rho$ occurs when the
 percentage of body fat is
@@
 2\%$,
-whereupon consumers have no eligible route \sid{what does this mean?} that avoids
-starvation.  Compellingly, this dynamic bound determined by the rate of
+whereupon consumers are unable to recover from the hungry state.  Compellingly, this dynamic bound determined by the rate of
 energetic recovery is close to the minimum observed mammalian body size of
@@
 limitations~\cite{West:2002ud}, we suggest that an additional constraint of
-lower body size stems from the dynamics of starvation.  This work \sid{which
-  work?} mirrors other efforts where coincident limitations seem to limit the
-smallest possibilities for life within a particular class or
+lower body size stems from the dynamics of starvation.  This result mirrors other efforts where coincident limitations seem to constrain the
+smallest possibilities for life within a particular class of
 organisms~\cite{Kempes:2012hy}.
@@
 %Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
-We determine a potential upper bound to body mass by assessing the
+A theoretical upper bound on mammalian body size is given by $\epsilon=0$, where mammals are entirely composed of metabolic reserves, and this occurs at a size of $M=8.3\times 10^8$, or $4.5\times$ the mass of a blue whale.
+We determine a more realistic upper bound to body mass by assessing the
 susceptibility of an otherwise homogeneous population to invasion by a

### 2016-08-12 Manuscript/8.tex
Major changes to rates

@@
 $M_{\rm opt}\approx10^6$ thus serves as an attractor, where over evolutionary
-times the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and
+time the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and
 decrease if $M>M_{\rm opt}$.  Moreover, $M_{\rm opt}$, which is entirely

### 2016-08-12 Manuscript/Supplement.tex
Major changes to rates

--- 2016-07-26 Manuscript/3.tex
@@
 
-\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
+\title{Supporting Information for ``The dynamics of starvation and recovery''}%: Eco-evolutionary feedbacks}
 
@@
 
-\begin{abstract}
-This is the abstract.  \emph{ \sid{No it isn't.  It's merely a placeholder.}}
-\end{abstract}
-\keywords{foraging | starvation|reproduction}
+\section*{Parameter Values and Estimates}
 
-\noindent {\bf Introduction} \\ The behavioral ecology of most, if
-not all, organisms is influenced by the energetic state of individuals, which
-directly influences how they invest reserves in uncertain environments.  Such
-behaviors are generally manifested as trade-offs between investing in somatic
-maintenance and growth, or allocating energy towards
-reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.  The timing of
-these behaviors responds to selective pressure, as the choice of the
-investment impacts future fitness~\cite{Mangel:1988uaa}.  The influence of
-resource limitation on an organism's ability to maintain its nutritional
-stores may lead to repeated delays or shifts in reproduction over the course
-of an organism's life.
+Many of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
+Here we outline previous measurements and simple estimates of the parameters.
 
-The life history of most species is typically comprised of (a) somatic growth
-and maintenance, and (b) reproduction.  The balance between these two
-activities is often conditioned on resource
-availability~\cite{Morris:1987eo}.  For example, reindeer invest less in
-calves born after harsh winters (when the mother's energetic state is
-depleted) than in calves born after moderate winters~\cite{Tveraa:2003fq}.
-Many bird species invest differently in broods during periods of resource
-scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv},
-sometimes delaying or even foregoing reproduction for a breeding
-season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.  Even freshwater
-and marine zooplankton have been observed to avoid reproduction under
-nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have
-lower survival rates~\cite{Kirk:1997cc}.  Organisms may also separate
-maintenance and growth from reproduction over space and time: many salmonids,
-birds, and some mammals return to migratory breeding grounds to reproduce
-after one or multiple seasons in resource-rich environments where they
-accumulate nutritional
-reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+\subsection*{Standard synthesis and metabolic parameters}
 
-Physiological mechanisms also play an important role in regulating
-reproductive expenditures during periods of resource limitation.  The data
-collected thus far has shown that diverse mammals (47 species in 10 families)
-exhibit delayed implantation, whereby females postpone fetal development
-(blastocyst implantation) until times where nutritional reserves can be
-accumulated~\cite{Mead:1989dt,Sandell:1990kw}.  Many other many species
-(including humans) suffer irregular menstrual cycling and higher spontaneous
-abortion rates during periods of nutritional
-stress~\cite{Bulik:1999eo,Trites:2003cc}.  In the extreme case of unicellular
-organisms, nutrition is unavoidably linked to reproduction because the
-nutritional state of the cell regulates all aspects of the cell cycle
-\cite{Glazier:2009hq}.  The existence of so many independently evolved
-mechanisms across such a diverse suite of organisms highlights the importance
-and universality of the fundamental tradeoff between somatic and reproductive
-investment.  However the dynamic implications of these constraints are
-unknown.
+Metabolic rate has been often reported to follow an exponent close to $0.75$ (e.g.
+\cite{West:2001bv,moses2008rmo} and the supplement of \cite{hou}).
+We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \cite{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around all parameter values.
+It is important to note the exponent, because it defines the value of the metabolic normalization constant, $B_{0}$, given a set of data.
+For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$) \cite{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
+We employ the field metabolic rate for our NSM model which is appropriate for active mammals.
 
-Though straightforward conceptually, incorporating the energetic dynamics of
-individuals~\cite{Kooi2000} into a population-level
-framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical
-obstacles~\cite{Diekmann:2010da}.  An alternative approach involves modeling
-the macroscale relations that guide somatic versus reproductive investment in
-a consumer-resource system.  For example, macroscale Lotka-Volterra models
-assume that the growth rate of the consumer population depends on resource
-density, thus \emph{implicitly} incorporating the requirement of resource
-availability for reproduction~\cite{murdoch:2003}.
+The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g.
+\cite{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \cite{moses2008rmo}.
+The unit energy available during starvation, $E^{\prime}$, can be considered in several ways.
+The first is to assume that the total energy stored during ontogeny is returned during starvation which would give a value of $7000$ (J g$^{-1}$) \cite{hou}.
+However, since our model considers the consumption of all body fat as defining the transition to starving, it is more appropriate to consider the energetics of fat metabolism where we would expect $E^{\prime}=36,000$ (J g$^{-1}$) for palmitate\cite{stryer,hou}.
 
-In this work, we adopt an alternative approach in which resource limitation
-and the subsequent effect of starvation is accounted for \emph{explicitly}.
-Namely, only individuals with sufficient energetic reserves can reproduce.
-Such a constraint leads to reproductive time lags due to some members of the
-population going hungry and then recovering.  Additionally, we incorporate the
-idea that reproduction is strongly constrained
-allometrically~\cite{Kempes:2012hy}, and is not generally linearly related to
-resource density.  As we shall show, these constraints influence the ensuing
-population dynamics in dramatic ways.
-\\
+The energy required for maintaining an existing unit of mass is reported to be
 
-\noindent {\bf Nutritional-state-structured model (NSM)}\\
-We begin by defining a minimal Nutritional-State-structured population Model
-(NSM), where the consumer population is divided into two energetic states:
-(a) an energetically replete (full) state $F$, where the consumer reproduces
-at a constant rate $\lambda$ and has no mortality risk, and (b) an
-energetically deficient (hungry) state $H$, where the consumer does not
-reproduce but dies at rate $\mu$.  The underlying resource $R$ evolves by
-logistic growth with an intrinsic growth rate $\alpha$ and a carrying
-capacity equal to one.  Consumers transition from the full state $F$ to the
-hungry state $H$ by starvation at rate $\sigma$ and also in proportion to the
-absence of resources $(1-R)$.  Conversely, consumers recover from state $H$
-to state $F$ at rate $\rho$ and in proportion to $R$.  Resources are also
-eaten by the consumers---at rate $\rho$ by hungry consumers and at rate
-$\beta<\rho$ by full consumers.  This inequality accounts for hungry
-consumers requiring more resources to rebuild body weight.
+\onecolumn
 
-In the mean-field approximation, in which the consumers and resources are
-perfectly mixed, their densities evolve according to the rate equations
-\begin{eqnarray} 
-\label{eq:system}
-\begin{split}
-\dot{F} &= \lambda F + \rho RH - \sigma (1-R)F,  \\
-\dot{H} &= \sigma (1-R)F - \rho RH - \mu H,  \\
-\dot{R} &= \alpha R(1-R) - R(\rho H+ \beta F).
-\end{split}
-\end{eqnarray}
-Notice that the total consumer density $F+H$ evolves according to
-$\dot{F}+\dot{H}=\lambda F-\mu H$.  This resembles the equation of motion for
-the predator density in the classic Lotka-Volterra model, except that the
-resource density does not appear in the growth term.  As discussed above, the
-attributes of reproduction and mortality have been explicitly apportioned to
-the full and hungry consumers, respectively, so that the growth in the total
-density is decoupled from the resource density.
+\begin{table}[h]
+\caption{Parameter Values For Various Classes of Organisms}
+\label{param}
+    \begin{center}
+    \small
+     \begin{tabular}{| p{1.2cm}| p{4.2cm} | }
+     \hline
+     & {\bf Mammals}  \\
+     \hline
+   $\eta$ & $3/4$  \quad \quad  (e.g. \cite{West:2001bv,moses2008rmo,hou}) \\ 
+   $E_{m}$ & $5774$ (J gram$^{-1}$)  \quad \quad  \cite{moses2008rmo,West:2001bv,hou} \\ 
+   $E_{m}^{\prime}$ & $36,000$  \quad \quad \cite{stryer,hou} \\ 
+   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    \quad \quad \cite{hou}  \\
+   $B_{m}$ & $0.025$ (W gram$^{-1}$) \quad \quad \\
+   $a$ & $1.78\times10^{-6}$  \quad \quad \\ 
+   $b$ & $2.29\times10^{-6}$  \quad \quad \\  
+   $\eta-1$ & $-0.21$  \quad \quad \\ 
+   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\ 
+   $\gamma$ & $1.19$ \quad \quad \\ 
+   $f_{0}$ & $0.02$ \quad \quad \\ 
+   $\zeta$ & $1.01$  \quad \quad \\ 
+   $mm_{0}$ & $0.32$  \quad \quad \\
 
-Equation~\eqref{eq:system} has three fixed points: two trivial fixed points
-at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal
-fixed point at
-\begin{eqnarray}
-\label{eq:ss}
-\begin{split}
-F^* &= \frac{\alpha  \lambda  \mu  (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  \beta)}, \\
-H^* &= \frac{\alpha  \lambda ^2 (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  \beta)}, \\
-R^* &= \frac{\mu  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma }.
-\end{split}
-\end{eqnarray}
-The stability of this fixed point is determined by the Jacobian Matrix
-$\bf J$, where each matrix element $J_{ij}$ equals
-$\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed
-point, and $\mathbf{X}$ is the vector $(F,H,R)$.  The parameters in
-Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue
-of $\mathbf{J}$ is negative, so that the system is stable with respect to
-small perturbations from the fixed point.  Because this fixed point is
-unique, it is the global attractor for all population trajectories for any
-initial condition where the resource and consumer densities are both non
-zero.
+   \hline
+    \end{tabular}
+    \end{center}
+   \end{table}
 
-From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the
-reproduction rate $\lambda$ must be less than the starvation rate $\sigma$,
-so that $R^*$ is positive.  In fact, when the resource density $R=0$, the
-rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
-The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation
-that demarcates the physical and unphysical regimes\sid{give Ref}.  The
-biological implication of the constraint $\lambda<\sigma$ has a simple
-interpretation---the rate at which a macroscopic organism loses mass due to
-lack of resources is generally much faster than the rate of reproduction.  As
-we will discuss below, this inequality is a natural consequence of allometric
-constraints~\cite{Kempes:2012hy} for organisms within empirically observed
-body size ranges (Fig.~\ref{fig:gvs}).
+\bibliography{aa_starving_supplement}
 
-In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may
-either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).  In
-continuous-time systems, a limit cycle arises when a pair of complex
-conjugate eigenvalues crosses the imaginary axis to attain positive real
-parts~\cite{GuckHolmes}.  This Hopf bifurcation is defined by
-${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is
-composed of the coefficients of the characteristic polynomial of the Jacobian
-matrix~\cite{Gross:2004p2428}.  As the system parameters are tuned to be
-within the stable regime but close to the Hopf bifurcation, the amplitude of
-the transient but decaying cycles become large.  Given that ecological
-systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of
-transient cycles, even though they decay with time in the mean-field
-description, can increase the extinction
-risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.  Thus the distance
-of a system from the Hopf bifurcation provides a measure of its persistence.
-
-When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the
-consumers are driven to the hungry non-reproducing state.  Because
-reproduction is inhibited, there is a low steady-state consumer density and a
-high steady-state resource density.  However, if $\sigma/\lambda\to 1$ from
-above, the population is overloaded with energetically-replete (reproducing)
-individuals, thereby promoting oscillations between the consumer and resource
-densities (Fig.~\ref{fig:fp}).
-
-Whereas the relation between consumer growth rate $\lambda$ and the
-starvation rate $\sigma$ defines an absolute bound of biological
-feasibility---the TC bifurcation---the starvation rate $\sigma$ also
-determines the sensitivity of the consumer population to changes in resource
-density.  When $\sigma\gg\lambda$, the steady-state population density is
-small, thereby increasing the risk of stochastic extinction.  On the other
-hand, as $\sigma$ decreases, the system will ultimately be poised either near
-the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).  If the recovery rate
-$\rho$ is sufficiently small, the TC bifurcation is reached and the resource
-eventually is eliminated.  If $\rho$ exceeds a threshold value, cyclic
-dynamics will develop as the Hopf bifurcation is approached.
-
-\sid{mention  and refer to the starving random walk model somewhere.}
-\\
-
-%In the case of the 2-stage consumer model, the Jacobian evaluated at the internal steady state (denoted by $|_*$) is written
-%
-%\begin{equation}
-%\mathbf{J}|_* =
-%%\left(
-%%\begin{array}{ccc}
-%% -F^* m+(1-R^*) a -R^* a -H^* \rho  & -R^* \rho  & -m R^* \\
-%% -H^* \rho -F^* \sigma  & -\mu -R^* \rho  & (1-R^*) \sigma  \\
-%% H^* \rho +F^* \sigma  & R^* \rho  & \lambda -(1-R^*) \sigma  \\
-%%\end{array}
-%%\right).
-%\left(
-%\begin{array}{ccc}
-% -\frac{\lambda  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
-% \frac{\lambda  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\mu  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
-% -\frac{m \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } & -\frac{\mu  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } \\
-%\end{array}
-%\right).
-%\end{equation}
-
-%CHRIS - we need to define beta in this section...
-
-\noindent {\bf Role of allometry} \\ 
-%[Link Allometry stuff to our model - what does it provide to the story]
-%[Introduce biological and linked constraints]
-%[Allometries capture vast amounts of diversity via a single parameter --- body size]
-%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-The NSM describes a broad range of dynamics, yet organisms are likely unable
-to access most of the total parameter space. Here we use allometric scaling
-relations to constrain the covariation of rates in a principled and
-biologically meaningful manner.  Allometric scaling relations highlight
-common constraints and average trends across large ranges in body size and
-species diversity. Many of these relations can be derived from a small set of
-assumptions and below we describe a framework to determine the covariation of
-timescales and rates across the range of mammals for each of the key
-parameters of our model (cf.~\cite{Yodzis:1992hg}).  We are thereby able to
-define the regime of dynamics occupied by the entire class of mammals along
-with the key differences between the largest and smallest mammals.
-
-Nearly all of the rates described in the NSM are to some extent governed by
-consumer metabolism, which can be used to describe a variety of organismal
-features~\cite{}\sid{Ref}.  The scaling relation between an organism's
-metabolic rate $B$ and its body size at reproductive maturity $M$ is well
-documented~\cite{West:2002it} and scales as $B = B_0 M^\eta$, where $\eta$ is
-the scaling exponent, generally assumed to vary around $2/3$ or $3/4$ for
-metazoans~\cite{}\sid{Ref}, and has taxonomic shifts for unicellular species
-between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria
-\cite{DeLong:2010dy,Kempes:2012hy}. Several efforts have shown how a
-partitioning of this metabolic rate between growth and maintenance purposes
-can be used to derive a general equation for the growth trajectories and
-growth rates of organisms ranging from bacteria to metazoans
-\cite{Kempes:2012hy,more}\sid{fix Ref}. More specifically, the interspecific
-\sid{what does interspecific mean?}  trends in growth rate can be
-approximated by
-\begin{eqnarray}
-\lambda = \lambda_0 M^{\eta-1}.
-\label{lambda}
-\end{eqnarray}
-This relation is derived from the simple balance condition
-\begin{eqnarray}
-\label{balance}
-B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
-\end{eqnarray}
-where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
-the metabolic rate to support an existing unit of mass, and $m$ is the mass
-at any point in development. It is useful to explicitly write this balance
-because it can also be modified to understand the timescales of both
-starvation and recovery from starvation as we show below.
-
-To determine the starvation rate, $\sigma$, we are interested in the time
-required for an organism to go from a mature adult that reproduces at rate
-$\lambda$ (henceforth we term this state as the ``replete'' state), to a
-reduced-size hungry state where reproduction is impossible.  This transition
-time can be inferred from the energy balance in Eq.~\ref{balance}, where we
-make the basic assumption that an organism must meet its maintenance
-requirements using the digestion of existing mass as the sole energy source.
-This assumption implies the simple metabolic balance \sid{don't understand
-  this equation}
-\begin{eqnarray}
-\frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
-\end{eqnarray}
-where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
-body mass which differs from $E_{m}$~\cite{}\sid{Ref}, the energy required to
-synthesis a unit of biomass. Given the replete mass, $M$, of an organism, the
-above energy balance prescribes the mass trajectory of a non-consuming
-organism:
-\begin{eqnarray}
-\label{mt}
-m\left(t\right)=Me^{-B_{m}t/E_{m}^{\prime}}.
-\end{eqnarray}
-Since only certain tissues can be digested for energy (for example the brain
-cannot be degraded to fuel metabolism), we define the rate for starvation and
-death by the timescales required to reach specific fractions of the
-replete-state mass.  We define $m_{\rm starve}=\epsilon M$, where
-$\epsilon<1$ is the fraction of replete-state mass where reproduction ceases.
-This fraction will be modified if tissue composition systematically scales
-with adult mass.  For example, making use of the observation that body fat in
-mammals scales with overall body size according to $M_{f}=f_{0}M^{\gamma}$
-and assuming that once this mass is fully digested the organism begins to
-starve, this would imply that $\epsilon=1-f_{0}M^{\gamma}/M$.  Using this
-criterion in Eq.~\ref{mt}, the time scale for starvation is given by
-\begin{equation}
-\label{eq:sigma}
-t_{\sigma}=-\frac{E_{m}\ln\left(\epsilon\right)}{B_{m}}.
-\end{equation}
-The starvation rate is then $\sigma=1/t_{\sigma}$, which scales with
-replete-state mass as $1/\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important
-feature is that $\sigma$ does not have a simple scaling dependence on
-$\lambda$ (Eq.~\ref{lambda}), which is important for the dynamics that we
-later discuss.
-
-The time to death should follow a similar relation, but defined by a lower
-fraction of replete-state mass, $m_{\rm death}=\epsilon^{\prime} M$.
-Suppose, for example, that an organism dies once it has digested all fat and
-muscle tissues, and that muscle tissue scales with body mass according to
-$M_{mm}=mm_{0}M^{\zeta}$ \sid{what is $mm_0$?}.  This gives
-$\epsilon^{\prime}=1-\left(f_{0}M^{\gamma}+mm_{0}M^{\zeta}\right)/M$. Muscle
-mass has been shown to be roughly proportional to body mass~\cite{muscle} in
-mammals and thus $\epsilon^{\prime}$ is merely $\epsilon$ minus a constant.
-Thus
-\begin{eqnarray}
-t_{\mu}=-\frac{E_{m}\ln\left(\epsilon^{\prime}\right)}{B_{m}}
-\end{eqnarray}
-and $\mu=1/t_{\mu}$.
-
-%CHRIS - MAYBE WE CUT THIS OUT SINCE WE STICK TO MAMMALS IN THIS PAPER
-%It should be noted that we have thus far used mammalian allometry to describe the size-based relations for growth, starvation, and death.
-However, our presentation is general, and other functional forms for $\epsilon$, for example, could be determined for other classes of organisms.
-Considering bacteria, we might expect that starvation or death is defined by the complete digestion of proteins, and in Table \ref{parameter-values} we provide all parameter values for bacteria which we later use as a comparison in our analysis.
-
-The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
-sufficient tissue to transition from the starving \sid{starving vs.\ hungry}
-state to the full state.  We again use the balance given in Eq.~\ref{balance}
-to find the timescale to return to the replete-state mass from a given
-reduced starvation mass.  From the solution to Eq.~\ref{balance}
-\begin{eqnarray}
-m\left(t\right)\!=\!\left(\!\frac{B_{0}}{B_{m}}\!\right)^{1/(\eta-1)}\!\!
-\left[1\!-\!\left(1\!-\!\frac{B_{m}}{B_{0}}m_{0}^{1\!-\!\eta}\right)e^{-b\left(1\!-\!\eta\right)t}\right]^{1/\left(1-\eta\right)}%\left(\frac{a}{b}\right)^{1/\left(1-\eta\right)}\,,
-\end{eqnarray}
-we require the timescale, $t_{\rho}=t_{2}-t_{1}$, which is the time it takes
-to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, or
-\begin{eqnarray}
-t_{\rho}=\frac{E_{m}\left\{\ln \left[1\!-\!\frac{B_{0}}{B_{m}}\left(M
-  \right)^{1\!-\!\eta }\right]
--\ln \left[1\!-\!\frac{B_{0}}{B_{m}}\left(\epsilon M \right)^{1\!-\!\eta }\right]\right\}}{(\eta -1) B_{m}}.
-\end{eqnarray}
-%\left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}
-%where $c=(a/b)^{1/(\eta-1)}$.
-Although these rate equations are general, here we focus on parameterizations
-for terrestrial-bound endotherms, specifically mammals, which range from a
-minimum of $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to a
-maximum of $M\approx10^7$ grams (the late Eocene to early Miocene
-Indricotheriinae).  Investigating other classes of organisms would simply
-involve altering the metabolic exponents and scalings associate with
-$\epsilon$. Moreover, we emphasize that our allometric equations describe
-mean relationships, and do not account for the (sometimes considerable)
-variance associated with individual species.
-\\
-%, though this is beyond the current scope of our analysis. %, and we henceforth focus our examination on mammalian species.
-
-\noindent {\bf  Stabilizing effects of allometric constraints} \\
-As the allometric derivations of the NSM rate laws reveal, $\sigma$ and
-$\rho$ are not independent parameters, and the bifurcation space shown in
-Fig.~\ref{fig:fp} is navigated via covarying parameters.  Given the
-parameters of terrestrial endotherms, we find that $\sigma$ and $\rho$ are
-constrained to lie within a small window of potential values
-(Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.  We thus find
-that the dynamics for all mammalian body sizes is confined to the
-steady-state regime of the NSM and that  limit-cycle behavior is precluded.
-%, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
-Moreover, for larger $M$, the distance to the Hopf bifurcation increases,
-while uncertainty in allometric parameters (20\% variation around the mean;
-Fig.~\ref{fig:hopf}) results in little qualitative difference in the distance
-to the the Hopf bifurcation. These results suggest that small mammals are
-more prone to population oscillations---including both stable limit cycles as
-well as transient cycles---than mammals with larger body size.  Thus our NSM
-model predicts that population cycles should be less common for larger
-species and more common for smaller species, particularly in environments
-where resources are limiting.%, and
-%2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
-
-Previous studies have used allometric constraints to explain the periodicity
-of cyclic populations
-\cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period
-$\propto M^{0.25}$, however this relation seems to hold only for some species
-\cite{Hendriks:2012fc} and competing explanations [related to ....\sid{???}]
-exist~\cite{Kendall:1999iy,Hogstedt:2005cr}.
-%It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic~\cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
-Statistically significant support for the existence of population cycles
-among mammals is predominantly based on time series for small
-mammals~\cite{Kendall:1998hl}, where we our model would predict much longer
-and more pronounced transient dynamics, given how close these points are to
-the Hopf bifurcation.  On the other hand, the longer gestational times and
-the increased difficulty in measurements, precludes obtaining similar-quality
-data for larger organisms.
-\\
-%, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
- %, emphasizing the role of starvation on dynamics.
-
-%Extinction to the left and the right
-
-\noindent {\bf Extinction risk}\\
-Within our model, higher rates of starvation result in a larger flux of the
-population to the hungry state.  In this state reproduction is absent, thus
-increasing the likelihood of extinction.  However, from the perspective of
-population survival, it is the rate of starvation relative to the rate of
-recovery that determines the long-term dynamics of the system
-(Fig.~\ref{fig:fp}).  We now examine the competing effects of cyclic dynamics
-vs.\ changes in steady state density on extinction risk as a function of the
-ratio $\sigma/\rho$.  To this end, we computed the probability of extinction,
-where extinction is defined as the population trajectory going below
-$0.2~\times$ the allometrically constrained steady state for all times
-between $10^2$ and $\leq 10^6$.  This procedure is repeated for 1000
-replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an
-organism of $M=100$ grams.  In each replicate the initial condition is
-distributed around the steady state (Eq.~\ref{eq:ss}).  Specifically the
-initial densities are chosen to be $A(F^*,H^*,R^*)$, with $A$ a random
-variable that is uniformly distribution in $[0,2]$.  By allowing the rate of
-starvation to vary, we assessed extinction risk across a range of values of
-the ratio $\sigma/\rho$ varying between $10^{-2}$ to $2.5$, thus examining a
-horizontal cross-section of Fig.~\ref{fig:fp}.  As expected, higher rates of
-extinction correlated with both low and high values of $\sigma/\rho$.  For
-low values of $\sigma/\rho$, the increased extinction risk results from
-transient cycles with larger amplitudes as the system nears the Hopf
-bifurcation (Fig.~\ref{fig:ext}).  For large values of $\sigma/\rho$, higher
-extinction risk arises because of to the decrease in the steady state
-consumer population density.  This interplay creates an `extinction refuge'
-as shown in Fig.~\ref{fig:ext}, such that for a relatively constrained range
-of $\sigma/\rho$, extinction probabilities are minimized.
-
-We find that the allometrically constrained values of $\sigma/\rho$ (with
-$\pm 20\%$ variability around energetic parameter means) fall within the
-extinction refuge. These values are close enough to the Hopf bifurcation to
-avoid low steady state densities, and far enough away to avoid
-large-amplitude transient cycles. The fact that allometric values of $\sigma$
-and $\rho$ fall within this relatively small window supports the possibility
-that a selective mechanism has constrained the physiological conditions that
-drive observed starvation and recovery rates within populations.  Such a
-mechanism would select for organism physiology that generates appropriate
-$\sigma$ and $\rho$ values that avoid extinction.  This selection could occur
-via the tuning of body fat percentages, metabolic rates, and biomass
-maintenance efficiencies.
-% OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
-
-%Results are insensitive to alpha but possibly sensitive to beta (eff parameter)
-
-%NOTES 7/8/16
-%Calder has a hypothesis that period ~ Mass^(1/4).
-%Can we get "transient period" from Jacobian?
-%Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.
-
-\noindent {\bf Dynamic and energetic barriers to body size} \\
-Metabolite transport constraints are widely thought to place strict
-boundaries on biological
-scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to
-specific predictions on the minimum possible body size for
-organisms~\cite{West:2002ud}.  Above this bound, a number of energetic and
-evolutionary mechanisms have been explored to assess the costs and benefits
-associated with larger body masses, particularly for mammals.  One important
-such example is the \emph{fasting endurance hypothesis}, which contends that
-larger body size, with consequent lower metabolic rates and increased ability
-to maintain more endogenous energetic reserves, may buffer organisms against
-environmental fluctuations in resource availability~\cite{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant
-trend towards larger body size (known as Cope's
-Rule)~\cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br},
-and it is thought that within-lineage drivers generate selection towards an
-optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, the value
-of which may arise from higher extinction risk for large taxa over
-evolutionary timescales~\cite{Clauset:2009fh}.  These trends are thought to
-be driven by a combination of climate change and niche
-availability~\cite{Saarinen:2014br}; however the underpinning energetic costs
-and benefits of larger body sizes, and how they influence dynamics over
-ecological timescales, have not been explored.  We argue that the NSM
-provides a suitable framework to explore these issues.
-
-%Minimal mass bound
-A lower bound on mammalian body size is given by $\epsilon=1$, where mammals
-have no metabolic reserves and immediately starve; this occurs at a size of
-\sid{$M=value$}.  This calculation \sid{what calculation?} gives an extreme
-limit on size but does not account for the subtleties of starvation dynamics
-that may limit body size. The NSM correctly predicts that species with
-smaller masses have larger steady-state population densities.  However we
-observe that there is a sharp change in the mass dependence of both the
-steady-state densities and $\sigma/\rho$ at $M \approx 0.3$ grams
-(Fig.~\ref{fig:mass}a,b).  The dependence of the rates of starvation and
-recovery explain this phenomenon.  As the mass decreases, the rate of
-starvation increases, while the rate of recovery declines super-exponentially
-\sid{how do we know this?}.  This decline in $\rho$ occurs when the
-percentage of body fat is
-$1 - 1/\left[\big(B_{0}/B_{m}\big)^{1/\left(\eta-1\right)}M\right] \approx
-2\%$,
-whereupon consumers have no eligible route \sid{what does this mean?} out of
-starvation.  Compellingly, this dynamic bound determined by the rate of
-energetic recovery is close to the minimum observed mammalian body size of
-ca. 1.3-2.5 grams (Fig.~\ref{fig:mass}b,c), a range that occurs as the
-recovery rate begins its decline.  In addition to known transport
-limitations~\cite{West:2002ud}, we suggest that an additional constraint of
-lower body size stems from the dynamics of starvation.  This work \sid{which
-  work?} mirrors other efforts where coincident limitations seem to limit the
-smallest possibilities for life within a particular class or
-organisms~\cite{kempes}.
-
-%Maximum mass bound
-%Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
-We determine a potential upper bound to body mass by assessing the
-susceptibility of an otherwise homogeneous population to invasion by a
-mutated subset of the population (denoted by ${}^\prime$) where individuals
-have a modified proportion of body fat $M^\prime=M(1+\chi)$ where
-$\chi \in [-0.5,0.5]$, thus altering the rates of starvation $\sigma$,
-recovery $\rho$, and maintenance $\beta$.  There is no internal fixed point
-that correspond to a state where both original residents and invaders coexist
-(except for the trivial state $\chi=0$).  To assess the susceptibility to
-invasion as a function of the invader mass, we determine which consumer has a
-higher steady-state density for a given value of $\chi$.  We find that for
-$1\leq M<10^6$g, having additional body fat ($\chi > 0$) results in a higher
-steady-state invader population density
-($H^{\prime *}+F^{\prime *}>H^*+F^*$).  Thus the invader has an intrinsic
-advantage over the resident population.  However, for $M>10^6$, leaner
-individuals ($\chi < 0$) have the advantage, and this is due to the changing
-covariance between energetic rates as a function of modified energetic
-reserves \sid{I don't understand the phrase after the comma}.
-
-The observed switch in susceptibility as a function of $\chi$ at
-$M_{\rm opt}\approx10^6$ thus serves as an attractor, where over evolutionary
-times the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and
-decrease if $M>M_{\rm opt}$.  Moreover, $M_{\rm opt}$, which is entirely
-determined by the population-level consequences of energetic constraints, is
-within an order of magnitude of the mass observed in the North American
-mammalian fossil record~\cite{Alroy:1998p1594} and also the mass predicted
-from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
-While the state of the environment, as well as the competitive landscape,
-will determine whether specific body sizes are selected for or
-against~\cite{Saarinen:2014br}, we suggest that the starvation dynamic
-proposed here may provide the driving mechanism for the evolution of larger
-body size among terrestrial mammals.
-%and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
-
-%Closing
-The energetics associated with somatic maintenance, growth, and reproduction
-are important elements that influence the dynamics of all
-populations~\cite{Stearns:1989ip}.  The NSM is a minimal and general model
-that incorporates the dynamics of starvation that are expected to occur in
-resource-limited environments.  By incorporating allometric relations between
-the rates in the NSM, we find: (i) different organismal masses have distinct
-population dynamic regimes, (ii) allometrically-determined rates of
-starvation and recovery appear to minimize extinction risk, and (iii) the
-dynamic consequences of these rates may place additional barriers on the
-evolution of minimum and maximum body size.  We suggest that the NSM offers a
-means by which the dynamic consequences of energetic constraints can be
-assessed using macroscale interactions between and among species.  Future
-efforts will involve exploring the consequences of these dynamics in a
-spatially explicit framework, thus incorporating elements such as movement
-costs and spatial heterogeneity, which may elucidate additional tradeoffs
-associated with the dynamics of starvation.
-
-%In accordance with the fasting endurance hypothesis, 
-%
-%
-%Such energetic advantages, coupled with the 
-%Ove and tends to be selected for over evolutionary time.
-%The NSM is in accordance with the XXX, where species with larger masses are less likely to evince cyclic dynamics.
-%
-%
-%
-%
-%We offer an additional mechanism for the minimum and maximum limitation on mammalian body size operating on ecological timescales.
-%Such a mechanism would 
-%
-%
-%
-%
-%Although these physiological barriers are indeed of primary importance, our minimal model of starvation dynamics suggests that there may exist additional constraints due to dynamic and evolutionary effects at small and large body sizes, respectively.
-%
-%
-%
-
-%Looking Forward
-
-%
-%
-%%Starving Random Walk
-%In our modeling, foragers look for food by wandering in this changing
-%environment.  If such a search is successful, the forager is satiated and it
-%can engage in the essential activity of reproduction.  However, if the search
-%is unsuccessful for a sufficiently long period, the forager ``starves''.
-%Such a forager can do nothing else but forage, until it either finds food and
-%again becomes satiated or it dies when it goes too long without finding
-%nourishment.  These rules are reminiscent of the ``starving random walk''
-%model, where a single random walk can take $\mathcal{S}$ steps without
-%encountering food before starving to death.  Moreover, the resource does not
-%regenerate, so that the forager ultimately starves to death.  For this
-%idealized model, it was found that the average lifetime of the forager scales
-%algebraically with $\mathcal{S}$ in $d\!=\!1$ and $d\!=\!2$ dimensions, and
-%as $\exp(-A\mathcal{S}^\omega)$ for $d>2$.  Here the exponent
-%$\omega\approx \frac{1}{2}$ for $d=3$, while $\omega\to 1$ only as
-%$d\to \infty$, with the latter behavior corresponding to the mean-field
-%limit.  As we will discuss, regeneration of the resource, together with the
-%behavioral change between starving and satiated foragers leads to still much
-%richer dynamical behavior.
-
-%The reproductive tradeoff as a strategy
-%
-
-%
-%
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
-% 	\caption{
-% 	Probability that the resource value is less than threshold = 0.05
-% 	}
-% 	\label{SN}
-% \end{figure}
-%
-%
-%
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
-% 	\caption{
-% 	Hopf bifurcation
-% 	}
-% 	\label{SN}
-% \end{figure}
-
-\bibliography{aa_starving}
-
-\begin{acknowledgments}
-  C.P.K was supported by a Trump Fellowship from the American League of
-  Conservatives.  S.R. was supported by grants DMR-1608211 and 1623243 from
-  the National Science Foundation, and by the John Templeton Foundation.
-\end{acknowledgments}
-
-%\onecolumn
 %\let\cleardoublepage\clearpage
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-\caption{ % 58 words
-}
-\label{fig:growth}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_GvS.pdf}
-\caption{ % 58 words
-}
-\label{fig:gvs}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_FixedPoint.pdf}
-\caption{ % 58 words
-}
-\label{fig:fp}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_DataHopf.pdf}
-\caption{ % 58 words
-}
-\label{fig:hopf}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric.pdf}
-\caption{ % 58 words
-}
-\label{fig:ext}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_FPAllometric.pdf}
-\caption{ % 58 words
-}
-\label{fig:mass}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_Invasion.pdf}
-\caption{ % 58 words
-}
-\label{fig:invasion}
-\end{figure*}
-
-% 
-% 
-%	 \begin{figure}[h]
-%		 \centering
-%		 \includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-%		 \caption{
-%		 }
-%		 \label{growth-diagram}
-%	 \end{figure}
-%	 
-%	 \begin{figure}[h]
-%		 \centering
-%		 \includegraphics[width=0.25\textwidth]{fig_GvS.pdf}
-%		 \caption{
-%		 }
-%		 \label{GvS}
-%	 \end{figure}
-%
-%
-%	\begin{figure}[h]
-%	 	\centering
-%	 	\includegraphics[width=0.25\textwidth]{fig_FixedPoint.pdf}
-%	 	\caption{
-%	 	}
-%		\label{Hopfb}
-%	\end{figure}
-%
-%
-%	\begin{figure}[h]
-%		\centering
-%		\includegraphics[width=0.8\textwidth]{fig_DataHopf.pdf}
-%		\caption{
-%		}
-%		\label{DataHopf}
-%	\end{figure}
-%	 
-% 
-%	\begin{figure}[h]
-%		\centering
-%	 	\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometric.pdf}
-%	 	\caption{
-%	 	}
-%	 	\label{Ext}
-%	\end{figure}
-%	
-%	\begin{figure}[h]
-%	 	\centering
-%	 	\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
-%	 	\caption{
-%	 	}
-%	 	\label{Asymp}
-%	\end{figure}
-%
-\clearpage
-
-\begin{table}[h]
-\caption{Parameter Values For Various Classes of Organisms}
-\label{param}
-    \begin{center}
-    \small
-     \begin{tabular}{| p{1.2cm}| p{3.2cm} | p{2.6cm} | p{3.2cm} | }
-     \hline
-     & {\bf Mammals} & {\bf Unicellular Eukaryotes} & {\bf Bacteria} \\
-     \hline
-   $\eta$ & $3/4$ & & $1.70$ \\ 
-   $E_{m}$ & $10695$ (J gram$^{-1}$) & & $10695$ (J gram$^{-1}$) \\ 
-   $E_{m}^{\prime}$ & $\approx E_{m}$ & & $\approx E_{m}$ \\ 
-   $B_{0}$ & $0.019$ (W gram$^{-\alpha}$) & & $1.96\times10^{17}$ \\
-   $B_{m}$ & $0.025$ (W gram$^{-1}$)   & & $0.025$ (W gram$^{-1}$)\\
-   $a$ & $1.78\times10^{-6}$ & & $1.83\times10^{13}$ \\ 
-   $b$ & $2.29\times10^{-6}$ & & $2.29\times10^{-6}$ \\  
-   $\eta-1$ & $-0.21$ & & $0.73$ \\ 
-   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) & & $56493$ \\ 
-   $\gamma$ & $1.19$ & & $0.68$ \\ 
-   $f_{0}$ & $0.02$ & & $1.30\times10^{-5}$ \\ 
-   $\zeta$ & $1.01$ & & \\ 
-   $mm_{0}$ & $0.32$ & & \\
-
-   \hline
-    \end{tabular}
-    \end{center}
-   \end{table}
-
 \end{document}

### 2016-08-24 Manuscript/11.tex
Final analyses before first submission to PNAS

--- 2016-08-12 Manuscript/8.tex
@@
 \begin{abstract} %246 words
-The eco-evolutionary dynamics of species is fundamentally linked to the energetic constraints of individuals.
-Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery in resource limited environments.
-We introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete consumers that reproduce, and
-undernourished, non-reproducing consumers that are susceptible to
-mortality.  
-As a function of the energetic rates between these replete and
-undernourished states that are determined by the presence or absence of
-resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
-We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships and find that population dynamics subject to these constraints can approach the cyclic regime but are typically driven to a steady state.
-Moreover, we find these rates fall within a `refuge' in parameter space, where the probability of extinction of the consumer population is minimized.
-Thus we introduce a potential mechanism that may both drive and constrain the dynamics of animal populations.
-Importantly, our model provides a natural framework that predicts maximum body size for mammals by examining the relative stability of an otherwise homogeneous population to a mutant population with altered percent body fat.
-For body masses $\lesssim 10^6$g, an invader with increased percent body fat dominates over the resident population, and vice versa for body mass $\gtrsim 10^6$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
+  The eco-evolutionary dynamics of species is fundamentally linked to the
+  energetic constraints of its constituent individuals.  Of particular
+  importance are the tradeoffs between reproduction and the dynamics of
+  starvation and recovery in resource-limited environments.  To elucidate the
+  consequences of this tradeoff, we introduce a minimal nutritional
+  state-structured model that incorporates two classes of consumer:
+  nutritionally replete consumers that reproduce, and undernourished,
+  non-reproducing consumers that are susceptible to mortality.  As a function
+  of the transition rates between these replete and undernourished states
+  that are determined by the presence or absence of resources, the consumer
+  populations can either undergo cyclic dynamics or reach a steady state.  We
+  obtain strong constraints on starvation and recovery rates by deriving
+  allometric scaling relationships and find that population dynamics subject
+  to these constraints can approach the cyclic regime but are typically
+  driven to a steady state.  Moreover, we find that these rates fall within a
+  `refuge' in parameter space, where the probability of extinction of the
+  consumer population is minimized.  Thus we identify a potential mechanism
+  that may both drive and constrain the dynamics of animal populations.  Our
+  model provides a natural framework that predicts maximum body size for
+  mammals by determining the relative stability of an otherwise homogeneous
+  population to a mutant population with altered percent body fat.  For body
+  masses $\lesssim 10^7$g, a more apidose invader dominates over the resident
+  population, and vice versa for body mass $\gtrsim 10^7$g, thus providing a
+  principled mechanism for a within-lineage driver of Cope's rule.
 \end{abstract}
@@
 
-\noindent {\bf Introduction} \\ The behavioral ecology of most, if
-not all, organisms is influenced by the energetic state of individuals, which
-directly influences how they invest reserves in uncertain environments.  Such
-behaviors are generally manifested as trade-offs between investing in somatic
+\noindent {\bf Introduction} \\ The behavioral ecology of nearly all
+organisms is influenced by the energetic state of individuals, which directly
+influences how they invest reserves in uncertain environments.  Such
+behaviors are generally manifested as tradeoffs between investing in somatic
 maintenance and growth, or allocating energy towards
@@
 
-The balance between (a) somatic growth and maintenance, and (b) reproduction is often conditioned on resource
-availability~\cite{Morris:1987eo}.  For example, reindeer invest less in
-calves born after harsh winters (when the mother's energetic state is
-depleted) than in calves born after moderate winters~\cite{Tveraa:2003fq}.
-Many bird species invest differently in broods during periods of resource
-scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv},
-sometimes delaying or even foregoing reproduction for a breeding
+The balance between (a) somatic growth and maintenance, and (b) reproduction
+is often conditioned on resource availability~\cite{Morris:1987eo}.  For
+example, reindeer invest less in calves born after harsh winters (when the
+mother's energetic state is depleted) than in calves born after moderate
+winters~\cite{Tveraa:2003fq}.  Many bird species invest differently in broods
+during periods of resource scarcity compared to normal
+periods~\cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing
+reproduction for a breeding
 season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.  Even freshwater
@@
 
-Physiological mechanisms also play an important role in regulating
-reproductive expenditures during periods of resource limitation.  The data
-collected thus far has shown that diverse mammals (47 species in 10 families)
-exhibit delayed implantation, whereby females postpone fetal development
-(blastocyst implantation) until nutritional reserves can be
+Physiology also plays an important role in regulating reproductive
+expenditures during periods of resource limitation.  The data collected thus
+far has shown that diverse mammals (47 species in 10 families) exhibit
+delayed implantation, whereby females postpone fetal development (blastocyst
+implantation) until nutritional reserves can be
 accumulated~\cite{Mead:1989dt,Sandell:1990kw}.  Many other many species
@@
 
-In this work, we adopt an alternative approach in which resource limitation
-and the subsequent effect of starvation is accounted for \emph{explicitly}.
+In this work, we adopt an alternative approach in which we \emph{explicitly}
+acccount for resource limitation and the subsequent effect of starvation.
 Namely, only individuals with sufficient energetic reserves can reproduce.
 Such a constraint leads to reproductive time lags due to some members of the
-population going hungry and then recovering.  Additionally, we incorporate the
-idea that reproduction is strongly constrained
-allometrically~\cite{Kempes:2012hy}, and is not generally linearly related to
+population going hungry and then recovering.  Additionally, we incorporate
+the idea that reproduction is strongly constrained
+allometrically\cite{Kempes:2012hy}, and is not generally linearly related to
 resource density.  As we shall show, these constraints influence the ensuing
@@
 We begin by defining a minimal Nutritional State-structured population Model
-(NSM), where the consumer population is divided into two energetic states:
-(a) an energetically replete (full) state $F$, where the consumer reproduces
-at a constant rate $\lambda$ and does not die from starvation, and (b) an
+(NSM), where the consumer population is partitioned into two states: (a) an
+energetically replete (full) state $F$, where the consumer reproduces at a
+constant rate $\lambda$ and does not die from starvation, and (b) an
 energetically deficient (hungry) state $H$, where the consumer does not
-reproduce but dies by starvation at rate $\mu$.  The underlying resource $R$ evolves by
-logistic growth with an intrinsic growth rate $\alpha$ and a carrying
-capacity equal to one.  Consumers transition from the full state $F$ to the
-hungry state $H$ at a rate $\sigma$---the starvation rate---and also in
-proportion to the absence of resources $(1-R)$.  Conversely, consumers
+reproduce but dies by starvation at rate $\mu$.  The underlying resource $R$
+evolves by logistic growth with an intrinsic growth rate $\alpha$ and a
+carrying capacity equal to one.  Consumers transition from the full state $F$
+to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also
+in proportion to the absence of resources $(1-R)$.  Conversely, consumers
 recover from state $H$ to state $F$ at rate $\rho$ and in proportion to $R$.
@@
 accounts for hungry consumers requiring more resources to replace lost body
-tissue.
+tissue.  The NSM represents a fundamental extension of the idealized starving
+random walk model of foraging, which focuses on resource depletion, to
+include reproduction and resource
+replenishment~\cite{PhysRevLett.113.238101,2016arXiv160500892B,PhysRevE.93.032403}.
 
@@
 perfectly mixed, their densities evolve according to the rate equations
-\begin{eqnarray} 
+\begin{align} 
 \label{eq:system}
@@
 \end{split}
-\end{eqnarray}
+\end{align}
 Notice that the total consumer density $F+H$ evolves according to
 $\dot{F}+\dot{H}=\lambda F-\mu H$.  This resembles the equation of motion for
-the predator density in the classic Lotka-Volterra model, except that the
-resource density does not appear in the growth term.  As discussed above, the
-attributes of reproduction and mortality have been explicitly apportioned to
-the full and hungry consumers, respectively, so that the growth in the total
-density is decoupled from the resource density.
+the predator density in the classic Lotka-Volterra model~\cite{Murray:2002},
+except that the resource density does not appear in the growth term.  As
+discussed above, the attributes of reproduction and mortality have been
+explicitly apportioned to the full and hungry consumers, respectively, so
+that the growth in the total density is decoupled from the resource density.
 
@@
 fixed point at
-\begin{eqnarray}
+\begin{align}
 \label{eq:ss}
@@
 \end{split}
-\end{eqnarray}
-The stability of this fixed point is determined by the Jacobian Matrix
-$\bf J$, where each matrix element $J_{ij}$ equals
-$\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed
-point, and $\mathbf{X}$ is the vector $(F,H,R)$.  The parameters in
-Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue
-of $\mathbf{J}$ is negative, so that the system is stable with respect to
-small perturbations from the fixed point.  Because this fixed point is
-unique, it is the global attractor for all population trajectories for any
-initial condition where the resource and consumer densities are both non
-zero.
+\end{align}
+The stability of this fixed point is determined by the Jacobian matrix
+$\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$
+when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector
+$(F,H,R)$.  The parameters in Eq.~\eqref{eq:system} are such that the real
+part of the largest eigenvalue of $\mathbf{J}$ is negative, so that the
+system is stable with respect to small perturbations from the fixed point.
+Because this fixed point is unique, it is the global attractor for all
+population trajectories for any initial condition where the resource and
+consumer densities are both nonzero.
 
@@
 rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
-The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation
-that demarcates the physical and unphysical regimes.  The
+The condition $\sigma = \lambda$ represents a transcritical (TC)
+bifurcation~\cite{Strogatz:2014} that demarcates the physical regime from the
+unphysical regime where $F$ would grow exponentially with time.  The
 biological implication of the constraint $\lambda<\sigma$ has a simple
@@
 %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-The NSM describes a broad range of dynamics, yet organisms are likely unable
-to access most of the total parameter space. Here we use allometric scaling
-relations to constrain the covariation of rates in a principled and
-biologically meaningful manner.  Allometric scaling relations highlight
-common constraints and average trends across large ranges in body size and
-species diversity. Many of these relations can be derived from a small set of
-assumptions and below we describe a framework to determine the covariation of
-timescales and rates across the range of mammals for each of the key
-parameters of our model (cf.~\cite{Yodzis:1992hg}).  We are thereby able to
-define the regime of dynamics occupied by the entire class of mammals along
-with the key differences between the largest and smallest mammals.
+While there are no a priori constraints on the parameters in the NSM, most
+organisms correspond to restricted portions of the parameter space.  Here we
+use allometric scaling relations to constrain the covariation of rates in a
+principled and biologically meaningful manner.  Allometric scaling relations
+highlight common constraints and average trends across large ranges in body
+size and species diversity. Many of these relations can be derived from a
+small set of assumptions and below we describe a framework to determine the
+covariation of timescales and rates across the range of mammals for each of
+the key parameters of our model (cf.~\cite{Yodzis:1992hg}).  We are thereby
+able to define the regime of dynamics occupied by the entire class of mammals
+along with the key differences between the largest and smallest mammals.
 
-Nearly all of the rates described in the NSM are to some extent governed by
-consumer metabolism, which can be used to describe a variety of organismal
-features \cite{Brown:2004wq}.  The scaling relation between an organism's
-metabolic rate $B$ and its body size at reproductive maturity $m=M$ (where $m$ denotes body sizes during growth and prior to reproductive maturity) is well
-documented~\cite{West:2002it} and scales as $B = B_0 M^\eta$, where $\eta$ is
-the scaling exponent, generally assumed to vary around $2/3$ or $3/4$ for
-metazoans (e.g. \cite{Brown:2004wq}), and has taxonomic shifts for unicellular species
-between $\eta\approx 1$ in eukaryotes and $\eta\approx 1.76$ in bacteria
+Nearly all of the rates described in the NSM are determined by consumer
+metabolism, which can be used to describe a variety of organismal features
+\cite{Brown:2004wq}.  The scaling relation between an organism's metabolic
+rate $B$ and its body mass $M$ at reproductive maturity is known to scale as
+$B = B_0 M^\eta$~\cite{West:2002it}, where the scaling exponent $\eta$ is
+typically close to $2/3$ or $3/4$ for metazoans (e.g., \cite{Brown:2004wq}),
+and has taxonomic shifts for unicellular species between $\eta\approx 1$ in
+eukaryotes and $\eta\approx 1.76$ in bacteria
 \cite{DeLong:2010dy,Kempes:2012hy}.
 %Justin adding an explanation for beta
-An organism's metabolic rate is proportional to the cost of tissue maintenance without growth (i.e.
-at $m=M$) such that $\beta = \xi B$, where $\xi$ is related to the conversion efficiency of resource to consumer tissue (Supporting Information, section XX).
+An organism's metabolic rate $B$ is proportional to the cost of tissue
+maintenance in the absence of growth (i.e., when the body mass is $M$).  By
+definition $B=\beta/\xi$, where $\beta$ is the rate at which resources are
+consumed for full consumers (see Eq.~\eqref{eq:system}) and where $\xi$ is
+related to the conversion efficiency of resource to consumer tissue
+(Supporting Information, section XX).
 
-Several efforts have shown how a
-partitioning of $B$ between growth and maintenance purposes
-can be used to derive a general equation for the growth trajectories and
-growth rates of organisms ranging from bacteria to metazoans
-\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. More specifically, the cross-species trends in growth rate can be
-approximated by
-\begin{eqnarray}
-\lambda = \lambda_0 M^{\eta-1}.
-\label{lambda}
-\end{eqnarray}
-This relation is derived from the simple balance condition \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}
+Several efforts have shown how a partitioning of $B$ between growth and
+maintenance purposes can be used to derive a general equation for both the
+growth trajectories and growth rates of organisms ranging from bacteria to
+metazoans
+\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. This relation is derived from the simple balance condition \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}
 \begin{eqnarray}
@@
 the metabolic rate to support an existing unit of mass, and $m$ is the mass
-at any point in development.  This balance has the general solution \cite{moses2008rmo, Kempes:2012hy}
+of the organism at any point in its development.  This balance has the
+general solution \cite{bettencourt,Kempes:2012hy}
 \begin{eqnarray}
-m\left(t\right)\!=\left[1\!-\!\left(1\!-\!\frac{b}{a}m_{0}^{1\!-\!\eta}\right)e^{-b\left(1\!-\!\eta\right)t}\right]^{1/\left(1-\eta\right)}\!\left(\!\frac{a}{b}\!\right)^{1/(1-\eta)}
+\label{m1}
+\left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}}
 \end{eqnarray}
-where, $a=B_{0}/E_{m}$, $b=B_{m}/E_{m}$, and $m_0$ is mass at birth, which we use to define the timescale of recovery from starvation.
+where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth.
+We now use this solution to define the timescale (see \cite{moses2008rmo} for a detailed presentation of these timescales) of reproduction and recovery from starvation.
+The time that it takes to reach a particular mass $\epsilon M$ is given by the timescale
+\begin{equation}
+\label{t1}
+\tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)}
+\end{equation}
+where we will define values of $\epsilon$ to describe a set of rates within our model.
+For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$ \cite{West:2001bv}).
+The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$ this will scale like $\lambda\propto M^{\eta-1}$ for $M>>m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
+
 The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
-sufficient tissue to transition from the hungry to the full state.  Since only certain tissues can be digested for energy (for example the brain
-cannot be degraded to fuel metabolism), we define the rate for starvation,
-death, and recovery by the timescales required to reach, or return from, specific fractions of the
-replete-state mass. 
-Here and henceforth, see Supporting Information, sections XX-XX for details on parameterizations.
-We define $m_{\rm starve}=\epsilon M$, where
-$\epsilon<1$ is the fraction of replete-state mass where reproduction ceases. This fraction will be modified if tissue composition systematically scales
-with adult mass.  For example, making use of the observation that body fat in
-mammals scales with overall body size according to $M_{\rm fat}=f_{0}M^{\gamma}$
-and assuming that once this mass is fully digested the organism starves, this
-would imply that $\epsilon=1-f_{0}M^{\gamma}/M$. It follows that the recovery timescale, $t_{\rho}=t_{2}-t_{1}$, is the time it takes
-to go from $m\left(t_{1}\right)=\epsilon M$ to $m\left(t_{2}\right)=M$, or
+sufficient tissue to transition from the hungry to the full state.  Since
+only certain tissues can be digested for energy (for example the brain cannot
+be degraded to fuel metabolism), we define the rates for starvation, death,
+and recovery by the timescales required to reach, or return from, specific
+fractions of the replete-state mass (see the Supporting Information for details on these parameterizations).  We define
+$m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction of
+replete-state mass where reproduction ceases. This fraction will be modified
+if tissue composition systematically scales with adult mass.  For example,
+making use of the observation that body fat in mammals scales with overall
+body size according to $M_{\rm fat}=f_{0}M^{\gamma}$ and assuming that once
+this mass is fully digested the organism starves, this would imply that
+$\epsilon_{\sigma}=1-f_{0}M^{\gamma}/M$. It follows that the
+recovery timescale, $t_{\rho}$, is the time to go from
+$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$ (Figure \ref{fig:growth}).
+Using Eqs.~\eqref{m1} and \eqref{t1} this timescale is given by simply considering an adjusted starting mass of $m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
 \begin{eqnarray}
-t_{\rho}=\frac{\ln\left[1\!-\!\frac{a}{b}\left(M
-  \right)^{1\!-\!\eta }\right]
--\ln \left[1\!-\!\frac{a}{b}\left(\epsilon M \right)^{1\!-\!\eta }\right]}{(\eta -1) b}.
+t_{\rho}=\ln\left[\frac{1-\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}
 \end{eqnarray}
-It should be noted that more complicated ontogenetic models explicitly handle storage \cite{hou}, whereas this is implicitly covered by the body fat scaling in our framework.
+where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in the biosynthetic energetics during recovery (see Supporting Information).
+It should be noted that more complicated ontogenetic models explicitly handle
+storage \cite{hou}, whereas this feature is implicitly covered by the body
+fat scaling in our framework.
 
@@
 \end{eqnarray}
+or
+\begin{eqnarray}
+\dot{m}=-\frac{a^{\prime}}{M^{1-\eta}}m
+\end{eqnarray}
 where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
@@
 \label{mt}
-m\left(t\right)=Me^{-b^{\prime}t}
+m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}.
 \end{eqnarray}
-where $b^{\prime}=E_{m}^{\prime}/B_{m}$. The time scale for starvation is given by the time it takes this decay to reach $\epsilon M$, which is
+The time scale for starvation is
+given by the time it takes $m(t)$ to reach $\epsilon_{\sigma} M$, which gives
 \begin{equation}
 \label{eq:sigma}
-t_{\sigma}=-b^{\prime}\ln\left(\epsilon\right).
+t_{\sigma}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\sigma}\right).
 \end{equation}
 The starvation rate is then $\sigma=1/t_{\sigma}$, which scales with
-replete-state mass as $1/\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important
+replete-state mass as $1/M^{1-\eta}\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important
 feature is that $\sigma$ does not have a simple scaling dependence on
-$\lambda$ (Eq.~\ref{lambda}), which is important for the dynamics that we
+$\lambda$, which is important for the dynamics that we
 later discuss.
@@
 The time to death should follow a similar relation, but defined by a lower
-fraction of replete-state mass, $m_{\rm death}=\epsilon^{\prime} M$.
+fraction of replete-state mass, $m_{\mu}=\epsilon_{\mu} M$.
 Suppose, for example, that an organism dies once it has digested all fat and
@@
 $M_{\rm musc}=u_{0}M^{\zeta}$.  This gives
-$\epsilon^{\prime}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
+$\epsilon_{\mu}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
 mass has been shown to be roughly proportional to body mass~\cite{Folland:2008ij} in
-mammals and thus $\epsilon^{\prime}$ is merely $\epsilon$ minus a constant.
-Thus
+mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant. The time to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or
 \begin{eqnarray}
-t_{\mu}=-b^{\prime}\ln\left(\epsilon^{\prime}\right),
+t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
 \end{eqnarray}
@@
 %where $c=(a/b)^{1/(\eta-1)}$.
-Although these rate equations are general, here we focus on parameterizations
-for terrestrial-bound endotherms, specifically mammals, which range from a
-minimum of $M\approx1$ gram (the Etruscan shrew \emph{Suncus etruscus}) to a
-maximum of $M\approx10^7$ grams (the late Eocene to early Miocene
-Indricotheriinae).  Investigating other classes of organisms would simply
-involve altering the metabolic exponents and scalings associate with
-$\epsilon$. Moreover, we emphasize that our allometric equations describe
-mean relationships, and do not account for the (sometimes considerable)
-variance associated with individual species.
+Although the rate equations \eqref{eq:system} are general, here we focus on
+parameterizations for terrestrial-bound endotherms, specifically mammals,
+which range from a minimum of $M\approx1$ gram (the Etruscan shrew
+\emph{Suncus etruscus}) to a maximum of $M\approx10^7$ grams (the late Eocene
+to early Miocene Indricotheriinae).  Investigating other classes of organisms
+would simply involve altering the metabolic exponents and scalings associate
+with $\epsilon$. Moreover, we emphasize that our allometric equations
+describe mean relationships, and do not account for the (sometimes
+considerable) variance associated with individual species.
 \\
@@
 \noindent {\bf  Stabilizing effects of allometric constraints} \\
-As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the bifurcation space shown in
-Fig.~\ref{fig:fp} is navigated via covarying parameters.  Given the
-parameters of terrestrial endotherms, we find that $\sigma$ and $\rho$ are
-constrained to lie within a small window of potential values
-(Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.  We thus find
-that the dynamics for all mammalian body sizes is confined to the
-steady-state regime of the NSM and that  limit-cycle behavior is precluded.
+As the allometric derivations of the NSM rate laws reveal, starvation and
+recovery rates are not independent parameters, and the biologically relevant
+portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via
+covarying parameters.  Given the parameters of terrestrial endotherms, we
+find that $\sigma$ and $\rho$ are constrained to lie within a small window of
+potential values (Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.
+We thus find that the dynamics for all mammalian body sizes is confined to
+the steady-state regime of the NSM and that limit-cycle behavior is
+precluded.
 %, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
@@
 to the the Hopf bifurcation. These results suggest that small mammals are
-more prone to population oscillations---including both stable limit cycles as
-well as transient cycles---than mammals with larger body size.  Thus our NSM
-model predicts that population cycles should be less common for larger
-species and more common for smaller species, particularly in environments
-where resources are limiting.%, and
+more prone to population oscillations---both stable limit cycles and
+transient cycles---than mammals with larger body size.  Thus our NSM model
+predicts that population cycles should be less common for larger species and
+more common for smaller species, particularly in environments where resources
+are limiting.%, and
 %2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
 
-Previous studies have used allometric constraints to explain the periodicity
-of cyclic populations
+It should be noted that previous studies have used allometric constraints to explain the periodicity of cyclic populations
 \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period
-$\propto M^{0.25}$, however this relation seems to hold only for some species
-\cite{Hendriks:2012fc}, and potential drivers range from predator and/or prey lifespan to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
+$\propto M^{0.25}$.  However this relation seems to hold only for some
+species \cite{Hendriks:2012fc}, and potential drivers range from predator
+and/or prey lifespans to competitive
+dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
 %It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic~\cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
@@
 population to the hungry state.  In this state reproduction is absent, thus
-increasing the likelihood of extinction.  However, from the perspective of
-population survival, it is the rate of starvation relative to the rate of
-recovery that determines the long-term dynamics of the system
-(Fig.~\ref{fig:fp}).  We now examine the competing effects of cyclic dynamics
-vs.\ changes in steady state density on extinction risk as a function of the
-ratio $\sigma/\rho$.  To this end, we computed the probability of extinction,
-where extinction is defined as the population trajectory falling below
-$0.2~\times$ the allometrically constrained steady state for all times
-between $10^6$ and $\leq 10^8$.  This procedure is repeated for 1000
-replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an
-organism of $M=100$ grams.  
+increasing the likelihood of extinction.  From the perspective of population
+survival, it is the rate of starvation relative to the rate of recovery that
+determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
+We therefore examine the competing effects of cyclic dynamics vs.\ changes in
+steady state density on extinction risk as a function of the ratio
+$\sigma/\rho$.  To this end, we computed the probability of extinction, where
+we define extinction as a population trajectory falling below one fifth of
+the allometrically constrained steady state at any time between $10^6$ and
+$\leq 10^8$.  This procedure is repeated for 1000 replicates of the
+continuous-time system shown in Eq.~\ref{eq:system} for an organism of
+$M=100$ grams.
 %the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
-In each replicate the
-initial densities are chosen to be $A(F^*,H^*,R^*)$, with $A$ a random
-variable that is uniformly distributed in $[0,2]$.  By allowing the rate of
-starvation to vary, we assessed extinction risk across a range of values of
-the ratio $\sigma/\rho$ varying between ca. $10^{-3}$ to $5$, thus examining a
-horizontal cross-section of Fig.~\ref{fig:fp}.  
-As expected, higher rates of
-extinction correlated with both low and high values of $\sigma/\rho$.  For
+In each replicate the initial densities are chosen to be $A(F^*,H^*,R^*)$,
+with $A$ a random variable that is uniformly distributed in $[0,2]$.  By
+allowing the rate of starvation to vary, we assessed extinction risk across a
+range of values of $\sigma/\rho$ between ca.\ $10^{-3}$ to $5$, thus examining
+a horizontal cross-section of Fig.~\ref{fig:fp}.  As expected, higher rates
+of extinction correlate with both low and high values of $\sigma/\rho$.  For
 low values of $\sigma/\rho$, the increased extinction risk results from
@@
 bifurcation (Fig.~\ref{fig:ext}).  For large values of $\sigma/\rho$, higher
-extinction risk arises because of to the decrease in the steady state
-consumer population density.  This interplay creates an `extinction refuge'
-as shown in Fig.~\ref{fig:ext}, such that for a relatively constrained range
-of $\sigma/\rho$, extinction probabilities are minimized.
+extinction risk arises because of the decrease in the steady state consumer
+population density.  This interplay creates an `extinction refuge' as shown
+in Fig.~\ref{fig:ext}, such that for a constrained range of $\sigma/\rho$,
+extinction probabilities are minimized.
 
@@
 and it is thought that within-lineage drivers generate selection towards an
-optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, the value
-of which may arise from higher extinction risk for large taxa over
-evolutionary timescales~\cite{Clauset:2009fh}.  These trends are thought to
-be driven by a combination of climate change and niche
-availability~\cite{Saarinen:2014br}; however the underpinning energetic costs
-and benefits of larger body sizes, and how they influence dynamics over
-ecological timescales, have not been explored.  We argue that the NSM
-provides a suitable framework to explore these issues.
+optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, whose
+value may arise from higher extinction risk for large taxa over evolutionary
+timescales~\cite{Clauset:2009fh}.  These trends are thought to be driven by a
+combination of climate change and niche availability~\cite{Saarinen:2014br};
+however the underpinning energetic costs and benefits of larger body sizes,
+and how they influence dynamics over ecological timescales, have not been
+explored.  We argue that the NSM provides a suitable framework to explore
+these issues.
 
@@
 %that may limit body size (among other factors). 
-The NSM correctly predicts that species with
-smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}a).  However we
-observe that there is a sharp change in the mass dependence of both the
-steady-state densities and $\sigma/\rho$ at $M \approx 0.3$ grams
-(Fig.~\ref{fig:mass}b,c).  The dependence of the rates of starvation and
-recovery explain this phenomenon.  As the mass decreases, the rate of
-starvation increases, while the rate of recovery declines super-exponentially.  This decline in $\rho$ occurs when the
-percentage of body fat is
-$1 - \big(a/b\big)^{-1/\left(\eta-1\right)}M^{-1} \approx
-2\%$,
-whereupon consumers are unable to recover from the hungry state.  Compellingly, this dynamic bound determined by the rate of
-energetic recovery is close to the minimum observed mammalian body size of
-ca. 1.3-2.5 grams (Fig.~\ref{fig:mass}b,c), a range that occurs as the
-recovery rate begins its decline.  In addition to known transport
-limitations~\cite{West:2002ud}, we suggest that an additional constraint of
-lower body size stems from the dynamics of starvation.  This result mirrors other efforts where coincident limitations seem to constrain the
-smallest possibilities for life within a particular class of
-organisms~\cite{Kempes:2012hy}.
+The NSM correctly predicts that species with smaller masses have larger
+steady-state population densities (Fig.~\ref{fig:mass}a).
+It should also be noted that $R^{*}$ is decreasing for increasing body size.
+From the perspective of classic resource competition theory we would expect that the species that can live on the lowest resource abundance should outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
+The NSM then predicts that larger mammals should outcompete smaller ones, supporting copes rule.
+This is a natural consequence of the steady-state dynamics combined with the derived timescales.
+
+Furthermore, many previous studies have focused on the energy equivalence hypothesis which argues that the total energy use, $B_{tot}$, of a population is constant independent of species size (e.g.
+\cite{allen2002,enquist1998}).
+This hypothesis is based on observations showing that the abundance, $N$, of a species is proportional to the inverse of individual metabolism (e.g.
+$N\propto M^{-3/4}/B_{0}$)  (e.g.
+\cite{allen2002,enquist1998}).
+This is usually stated as $B_{tot}=NB=constant$ which has been shown to hold in mammalian and vascular plant communities (e.g.
+\cite{allen2002,enquist1998}).
+Figure \ref{} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{} shows that $F^{*}B$ is a constant over this same range.
+This result if remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to the energy equivalence result.
+Furthermore, the equivalence breaks down at distinct sizes at both the large and small end of mammals suggesting that these are hard limits on mammalian body sizes because organisms outside this range do not meet the constant efficiency obeyed by other populations.
+Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of mammals and $M\approx 6.5*10^7$ for the large end.
+Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is
+close to the minimum observed mammalian body size of ca.\ 1.3-2.5 grams
+(Fig.~\ref{fig:mass}b,c), a range that occurs as the recovery rate begins its
+decline.  In addition to known transport limitations~\cite{West:2002ud}, we
+suggest that an additional constraint of lower body size stems from the
+dynamics of starvation.  This result mirrors other efforts \cite{Kempes:2012hy,Kempes:2016}
+where at a given scale multiple limitations constrain the smallest possibilities for
+life within a class of organisms.
+
+%
+%However we observe
+%that there is a sharp change in the mass dependence of both the steady-state
+%densities and $\sigma/\rho$ at $M \approx 0.3$ grams
+%(Fig.~\ref{fig:mass}b,c).  The dependence of the rates of starvation and
+%recovery explain this phenomenon.  As the mass decreases, the rate of
+%starvation increases, while the rate of recovery declines precipitously.
+%This decline in $\rho$ occurs when the percentage of body fat is
+%$1 - \big(a/b\big)^{-1/\left(\eta-1\right)}M^{-1} \approx 2\%$, whereupon
+%consumers are unable to recover from the hungry state.
 
@@
 %Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
-A theoretical upper bound on mammalian body size is given by $\epsilon=0$, where mammals are entirely composed of metabolic reserves, and this occurs at a size of $M=8.3\times 10^8$, or $4.5\times$ the mass of a blue whale.
+A additional theoretical upper bound on mammalian body size is given by $\epsilon=0$, where mammals are entirely composed of metabolic reserves, and this occurs at a size of $M=8.3\times 10^8$, or $4.5\times$ the mass of a blue whale.
 We determine a more realistic upper bound to body mass by assessing the
@@
 advantage over the resident population.  However, for $M>10^6$, leaner
-individuals ($\chi < 0$) have the advantage, and this is due to the changing
-covariance between energetic rates as a function of modified energetic
-reserves \sid{I don't understand the phrase after the comma}.
+individuals ($\chi < 0$) have advantageous steady state densities.
+%, and this is due to the changing
+%covariance between energetic rates as a function of modified energetic
+%reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
 
 The observed switch in susceptibility as a function of $\chi$ at
-$M_{\rm opt}\approx10^6$ thus serves as an attractor, where over evolutionary
-time the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and
+$M_{\rm opt}\approx10^6$ thus serves as an attractor, or an uninvadible evolutionary stable state, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and
 decrease if $M>M_{\rm opt}$.  Moreover, $M_{\rm opt}$, which is entirely
@@
 against~\cite{Saarinen:2014br}, we suggest that the starvation dynamics
-proposed here may provide the driving mechanism for the evolution of larger
+outlined here may provide the driving mechanism for the evolution of larger
 body size among terrestrial mammals.
@@
 
-A potential critique of our results, which show that larger mammals are less susceptible to extinction, is why we don't observe a greater number of large mammals in the modern world.
-However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which where also much more geographically widespread than today.
-These results combined with our findings suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
+One might be concerned a greater number of large mammals are currently not
+observed in the modern world given that larger mammals are less susceptible
+to extinction.  However, recent research suggests that the pleistocene may
+have been much more populated with a significant diversity of very large
+mammals \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also
+much more geographically widespread than today. These results, combined with
+our findings, suggest that the modern diversity of mammals may not represent a
+true steady state the current distribution of nutrients and large seeds may
+be very different from the past
+\cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
 
@@
 resource-limited environments.  By incorporating allometric relations between
-the rates in the NSM, we find: (i) different organismal masses have distinct
+the rates in the NSM, we found: (i) different organismal masses have distinct
 population dynamic regimes, (ii) allometrically-determined rates of
@@
 \begin{acknowledgments}
-We thank Luis Bettancourt, Eric Libby, Seth Newsome, et al. for helpful discussions and comments that greatly increased the quality of the manuscript.
-  J.D.Y. was supported by startup funds at the University of California, Merced, and an Omidyar postdoctoral Fellowship at the Santa Fe Institute.
-  C.P.K. was supported by an Omidyar postdoctoral Fellowship at the Santa Fe
+  We thank Luis Bettancourt, Eric Libby, and Seth Newsome for helpful
+  discussions and comments on the manuscript.  J.D.Y. was supported by
+  startup funds at the University of California, Merced, and an Omidyar
+  Postdoctoral Fellowship at the Santa Fe Institute.  C.P.K. was supported by
+  an Omidyar Postdoctoral Fellowship at the Santa Fe Institute.  S.R. was
+  supported by grants DMR-1608211 and 1623243 from the National Science
+  Foundation, and by the John Templeton Foundation, all at the Santa Fe
   Institute.
-  S.R. was supported by grants DMR-1608211 and 1623243 from the National Science Foundation, and by the John Templeton Foundation, all at the Santa Fe Institute.
 \end{acknowledgments}
@@
 \includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-\caption{
-The growth trajectory over absolute time of an individual organism as a function of body mass.
-Initial growth follows the red trajectory to an energetically replete adult mass $M$.
-Starvation follows the concave blue trajectory to $m_{\rm starve} < M$, whereas recovery follows the convex growth trajectory from $m_\sigma$ to $M$.
-}
+\caption{ The growth trajectory over absolute time of an individual organism
+  as a function of body mass.  Initial growth follows the red trajectory to
+  an energetically replete adult mass $M$.  Starvation follows the concave
+  blue trajectory to $m_{\rm starve} < M$, whereas recovery follows the
+  convex growth trajectory from $m_\sigma$ to $M$.  }
 \label{fig:growth}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig_GvS.pdf}
+\includegraphics[width=0.5\textwidth]{fig_Rates.pdf}
 \caption{ 
-Allometrically constrained starvation rate $\sigma$ (red) vs. reproductive rate $\lambda$ (yellow) as a function of mass $M$.
+Allometrically constrained starvation rate $\sigma$ (red) and recovery rate $\rho$ relative to the reproductive rate $\lambda$ (yellow) as a function of body mass.  
 The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
+Mean values $\pm 20\%$ variation are shown by the shaded region for each rate.
 }
@@
 \includegraphics[width=0.5\textwidth]{fig_FixedPoint.pdf}
-\caption{
-The transcritical (TC; dashed line) and Hopf bifurcation (solid line) as a function of the starvation rate $\sigma$ and recovery rate $\rho$.
-These bifurcation conditions separate parameter space into infeasible, cyclic, and steady state dynamic regimes.
-The color gradient shows the steady state densities for (A) the resource $R^*$ and the (B) energetically replete consumers $F^*$, with warm colors denoting higher densities and cool colors denoting lower densities.
-Steady state densities for the energetically deficient consumers $H^*$ are not shown because they closely mirror those for $F^*$.
-}
+\caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
+  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100 gram consumer.  These
+  bifurcation conditions separate parameter space into infeasible, cyclic,
+  and steady state dynamic regimes.  The color gradient shows the steady
+  state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
+  replete consumers $F^*$, (warmer colors denote higher densities).  Steady
+  state densities for the energetically deficient consumers $H^*$ (not shown)
+  scale with those for $F^*$.  }
 \label{fig:fp}
@@
 \includegraphics[width=0.5\textwidth]{fig_DataHopf.pdf}
-\caption{
-Transcritical (TC; vertical lines) and Hopf bifurcations (curved lines) with allometrically determined starvation $\sigma$ and recovery $\rho$ rates as a function of minimum and maximum mammalian body sizes: $1$ gram (blue) and $10^7$ grams (orange), respectively.
-Replicates show the influence of variation (20\% around the mean) on allometric parameters, which influences both the energetic rates as well as the position of the TC and Hopf bifurcations.
+\caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
+  allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
+  a function of different mammalian body sizes: $M=A\times10^1$ gram (blue) and
+  $M=A\times10^6$ grams (green), where $A$ is a random uniform variable in $[1,9]$.
   }
@@
 \includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric.pdf}
-\caption{
-The probability of extinction for 1000 consumer population trajectories as a function of $\sigma/\rho$ within initial densities chosen to as $A(F^*,H^*,R^*)$, with $A$ a random
-variable that is uniformly distribution in $[0,2]$.
-Extinction is defined as the population trajectory going below
-$0.2~\times$ the allometrically constrained steady state for all times
-between $10^2$ and $\leq 10^6$.
-The values above the extinction plot are the allometrically constrained $\sigma/\rho$ with 20\% variation around the mean.
-}
+\caption{ Probability of extinction for a 100g consumer as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $A(F^*,H^*,R^*)$, with
+  $A$ being a random uniform variable in $[0,2]$.  Extinction is defined as the
+  population trajectory falling below $0.1~\times$ the allometrically
+  constrained steady state. The white point denotes the allometrically constrained starvation and recovery rate.}
 \label{fig:ext}
@@
 \includegraphics[width=0.45\textwidth]{fig_FPAllometric.pdf}
-\caption{
-(\emph{A}) Consumer steady states as a function of body size, showing both energetically deficient and replete consumer states ($H^*$ and $F^*$, respectively). 
-Energetic rates as a function of body size, with the ratio $\sigma/\rho$ (\emph{B}) and both $\sigma$ (red) and $\rho$ (purple; \emph{C}) drawn separately with 20\% variation around the mean.
-Steady state densities decline sharply at $M=M_{\rm min}$ due to the super-exponential decrease in the rate of recovery.
-The minimum body size observed for mammals (the Etruscan shrew) is denoted by the orange shaded region at values marking the initial decline of the recovery rate.
-}
+\caption{ (\emph{A}) Consumer steady states $F^*$ and $H^*$ as a function of
+  body mass. 
+  (\emph{B}) Resource steady state $R^*$ as a function of consumer body mass.
+  (\emph{C}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.}
 \label{fig:mass}
@@
 \includegraphics[width=0.5\textwidth]{fig_Invasion.pdf}
-\caption{
-Invasion feasibility for organisms with a proportional change in mass $\chi$ against a population with a resident body mass $M$.
-The blue region denotes which values of $\chi$ result in successful invasion.
-The red region denotes which values of $\chi$ result in a mass that is below the starvation threshold and is thus infeasible.
-}
+\caption{ Invasion feasibility for organisms with a proportional change in
+  mass $\chi$ against a population with a resident body mass $M$.  The blue
+  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
+  red region denotes values of $\chi$ that result in a mass that is below the
+  starvation threshold and is thus infeasible.
+  Arrows point to the predicted optimal mass $M_{\rm opt}=8.43\times 10^6$, which serves as the evolutionary stable state.}
 \label{fig:invasion}

### 2016-08-24 Manuscript/12.tex
Final analyses before first submission to PNAS

--- 2016-08-24 Manuscript/11.tex
@@
 \usepackage[left]{lineno}
+%\usepackage[comma,sort&compress]{natbib}
 
@@
 
-\begin{abstract} %246 words
-  The eco-evolutionary dynamics of species is fundamentally linked to the
-  energetic constraints of its constituent individuals.  Of particular
-  importance are the tradeoffs between reproduction and the dynamics of
-  starvation and recovery in resource-limited environments.  To elucidate the
-  consequences of this tradeoff, we introduce a minimal nutritional
-  state-structured model that incorporates two classes of consumer:
-  nutritionally replete consumers that reproduce, and undernourished,
-  non-reproducing consumers that are susceptible to mortality.  As a function
-  of the transition rates between these replete and undernourished states
-  that are determined by the presence or absence of resources, the consumer
-  populations can either undergo cyclic dynamics or reach a steady state.  We
-  obtain strong constraints on starvation and recovery rates by deriving
-  allometric scaling relationships and find that population dynamics subject
-  to these constraints can approach the cyclic regime but are typically
-  driven to a steady state.  Moreover, we find that these rates fall within a
-  `refuge' in parameter space, where the probability of extinction of the
-  consumer population is minimized.  Thus we identify a potential mechanism
-  that may both drive and constrain the dynamics of animal populations.  Our
-  model provides a natural framework that predicts maximum body size for
-  mammals by determining the relative stability of an otherwise homogeneous
-  population to a mutant population with altered percent body fat.  For body
-  masses $\lesssim 10^7$g, a more apidose invader dominates over the resident
-  population, and vice versa for body mass $\gtrsim 10^7$g, thus providing a
-  principled mechanism for a within-lineage driver of Cope's rule.
+\begin{abstract} %250 words
+The eco-evolutionary dynamics of species is fundamentally linked to the energetic constraints of its constituent individuals.  
+Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery in resource-limited environments.  
+To elucidate the consequences of this tradeoff, we introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete consumers that reproduce, and undernourished, non-reproducing consumers that are susceptible to mortality.
+As a function of the transition rates between these replete and undernourished states that are determined by the presence or absence of resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
+We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships and find that population dynamics subject to these constraints can approach the cyclic regime but are typically driven to a steady state.
+Moreover, we find that these rates fall within a `refuge' in parameter space, where the probability of extinction of the consumer population is minimized.  
+Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations.  
+Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percent body fat.
+For body masses $\lesssim 10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $\gtrsim 10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
 \end{abstract}
@@
 
-\noindent {\bf Introduction} \\ The behavioral ecology of nearly all
-organisms is influenced by the energetic state of individuals, which directly
-influences how they invest reserves in uncertain environments.  Such
-behaviors are generally manifested as tradeoffs between investing in somatic
-maintenance and growth, or allocating energy towards
-reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.  The timing of
-these behaviors responds to selective pressure, as the choice of the
-investment impacts future fitness~\cite{Mangel:1988uaa}.  The influence of
-resource limitation on an organism's ability to maintain its nutritional
-stores may lead to repeated delays or shifts in reproduction over the course
-of an organism's life.
+\noindent {\small {\bf Significance Statement} %105/120 words
+Energetic investment in somatic maintenance and growth vs. reproduction directly impacts the dynamics of populations among species.
+Here, we construct a Nutritional State-structured Model (NSM) to assess the population-level effects of starvation and recovery of a consumer population in a resource-limited environment, and use allometric scaling relationships for mammals to establish all timescales and rates.
+Our model reveals that mammalian energetic rates minimize the probability of stochastic extinction, establishes dynamic bounds on mammalian body size while providing independent theoretical support for the energy equivalence hypothesis, and provides a mechanistic driver for the evolutionary trend towards larger body size known as Cope's rule.}
 
-The balance between (a) somatic growth and maintenance, and (b) reproduction
-is often conditioned on resource availability~\cite{Morris:1987eo}.  For
-example, reindeer invest less in calves born after harsh winters (when the
-mother's energetic state is depleted) than in calves born after moderate
-winters~\cite{Tveraa:2003fq}.  Many bird species invest differently in broods
-during periods of resource scarcity compared to normal
-periods~\cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing
-reproduction for a breeding
-season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.  Even freshwater
-and marine zooplankton have been observed to avoid reproduction under
-nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have
-lower survival rates~\cite{Kirk:1997cc}. Organisms may also separate
-maintenance and growth from reproduction over space and time: many salmonids,
-birds, and some mammals return to migratory breeding grounds to reproduce
-after one or multiple seasons in resource-rich environments where they
-accumulate nutritional
-reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+\vspace{2mm}
 
-Physiology also plays an important role in regulating reproductive
-expenditures during periods of resource limitation.  The data collected thus
-far has shown that diverse mammals (47 species in 10 families) exhibit
-delayed implantation, whereby females postpone fetal development (blastocyst
-implantation) until nutritional reserves can be
-accumulated~\cite{Mead:1989dt,Sandell:1990kw}.  Many other many species
-(including humans) suffer irregular menstrual cycling and higher spontaneous
-abortion rates during periods of nutritional
-stress~\cite{Bulik:1999eo,Trites:2003cc}.  In the extreme case of unicellular
-organisms, nutrition is unavoidably linked to reproduction because the
-nutritional state of the cell regulates all aspects of the cell cycle
-\cite{Glazier:2009hq}.  The existence of so many independently evolved
-mechanisms across such a diverse suite of organisms highlights the importance
-and universality of the fundamental tradeoff between somatic and reproductive
-investment.  However the dynamic implications of these constraints are
-unknown.
+\noindent {\bf Introduction} \\ 
+The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly influences how they invest reserves in uncertain environments.  
+Such behaviors are generally manifested as tradeoffs between investing in somatic maintenance and growth, or allocating energy towards reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.  
+The timing of these behaviors responds to selective pressure, as the choice of the investment impacts future fitness~\cite{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.  
+The influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
 
-Though straightforward conceptually, incorporating the energetic dynamics of
-individuals~\cite{Kooi2000} into a population-level
-framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical
-obstacles~\cite{Diekmann:2010da}.  An alternative approach involves modeling
-the macroscale relations that guide somatic versus reproductive investment in
-a consumer-resource system.  For example, macroscale Lotka-Volterra models
-assume that the growth rate of the consumer population depends on resource
-density, thus \emph{implicitly} incorporating the requirement of resource
-availability for reproduction~\cite{murdoch:2003}.
+The balance between (a) somatic growth and maintenance, and (b) reproduction depends on resource availability~\cite{Morris:1987eo}.  
+For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\cite{Tveraa:2003fq}.  
+Many bird species invest differently in broods during periods of resource scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
+Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\cite{Kirk:1997cc}.
+Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate nutritional reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
 
-In this work, we adopt an alternative approach in which we \emph{explicitly}
-acccount for resource limitation and the subsequent effect of starvation.
+Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.  
+The data collected thus far has shown that diverse mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development (blastocyst implantation) until nutritional reserves can be accumulated~\cite{Mead:1989dt,Sandell:1990kw}.
+Many other many species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.  
+In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.  
+The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the importance and universality of the fundamental tradeoff between somatic and reproductive investment.
+However the general dynamic implications of these constraints are unknown.
+
+Though straightforward conceptually, incorporating the energetic dynamics of individuals~\cite{Kooi2000} into a population-level framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles~\cite{Diekmann:2010da}.
+An alternative approach involves modeling the macroscale relations that guide somatic versus reproductive investment in a consumer-resource system.
+For example, macroscale Lotka-Volterra models assume that the growth rate of the consumer population depends on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction~\cite{murdoch:2003}.
+
+In this work, we adopt an alternative approach in which we \emph{explicitly} account for resource limitation and the subsequent effects of starvation.
 Namely, only individuals with sufficient energetic reserves can reproduce.
-Such a constraint leads to reproductive time lags due to some members of the
-population going hungry and then recovering.  Additionally, we incorporate
-the idea that reproduction is strongly constrained
-allometrically\cite{Kempes:2012hy}, and is not generally linearly related to
-resource density.  As we shall show, these constraints influence the ensuing
-population dynamics in dramatic ways.
+Such a constraint leads to reproductive time lags due to some members of the population going hungry and then recovering.  
+Additionally, we incorporate the idea that reproduction is strongly constrained allometrically \cite{Kempes:2012hy}, and is not generally linearly related to resource density.  
+As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.
 \\
@@
 \noindent {\bf Nutritional state-structured model (NSM)}\\
-We begin by defining a minimal Nutritional State-structured population Model
-(NSM), where the consumer population is partitioned into two states: (a) an
-energetically replete (full) state $F$, where the consumer reproduces at a
-constant rate $\lambda$ and does not die from starvation, and (b) an
-energetically deficient (hungry) state $H$, where the consumer does not
-reproduce but dies by starvation at rate $\mu$.  The underlying resource $R$
-evolves by logistic growth with an intrinsic growth rate $\alpha$ and a
-carrying capacity equal to one.  Consumers transition from the full state $F$
-to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also
-in proportion to the absence of resources $(1-R)$.  Conversely, consumers
-recover from state $H$ to state $F$ at rate $\rho$ and in proportion to $R$.
-Resources are also eaten by the consumers---at rate $\rho$ by hungry
-consumers and at rate $\beta<\rho$ by full consumers.  This inequality
-accounts for hungry consumers requiring more resources to replace lost body
-tissue.  The NSM represents a fundamental extension of the idealized starving
-random walk model of foraging, which focuses on resource depletion, to
-include reproduction and resource
-replenishment~\cite{PhysRevLett.113.238101,2016arXiv160500892B,PhysRevE.93.032403}.
+We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$.
+The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity equal to one.  
+Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$.
+Conversely, consumers recover from state $H$ to state $F$ at rate $\rho$ and in proportion to $R$.
+Resources are also eaten by the consumers---at rate $\rho$ by hungry consumers and at rate $\beta<\rho$ by full consumers.  
+This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.  
+The NSM represents a fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}.
 
-In the mean-field approximation, in which the consumers and resources are
-perfectly mixed, their densities evolve according to the rate equations
+In the mean-field approximation, in which the consumers and resources are perfectly mixed, their densities evolve according to the rate equations
 \begin{align} 
@@
 \end{align}
-Notice that the total consumer density $F+H$ evolves according to
-$\dot{F}+\dot{H}=\lambda F-\mu H$.  This resembles the equation of motion for
-the predator density in the classic Lotka-Volterra model~\cite{Murray:2002},
-except that the resource density does not appear in the growth term.  As
-discussed above, the attributes of reproduction and mortality have been
-explicitly apportioned to the full and hungry consumers, respectively, so
-that the growth in the total density is decoupled from the resource density.
+Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.  
+This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\cite{murray2011mathematical}, except that the resource density does not appear in the growth term.  
+As discussed above, the attributes of reproduction and mortality have been explicitly apportioned to the full and hungry consumers, respectively, so that the growth in the total density is decoupled from the resource density.
 
-Equation~\eqref{eq:system} has three fixed points: two trivial fixed points
-at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal
-fixed point at
+Equation~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
 \begin{align}
@@
 \end{align}
-The stability of this fixed point is determined by the Jacobian matrix
-$\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$
-when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector
-$(F,H,R)$.  The parameters in Eq.~\eqref{eq:system} are such that the real
-part of the largest eigenvalue of $\mathbf{J}$ is negative, so that the
-system is stable with respect to small perturbations from the fixed point.
-Because this fixed point is unique, it is the global attractor for all
-population trajectories for any initial condition where the resource and
-consumer densities are both nonzero.
+The stability of this fixed point is determined by the Jacobian matrix $\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.
+The parameters in Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue of $\mathbf{J}$ is negative, so that the system is stable with respect to small perturbations from the fixed point.
+Because this fixed point is unique, it is the global attractor for all population trajectories for any initial condition where the resource and consumer densities are both nonzero.
 
-From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the
-reproduction rate $\lambda$ must be less than the starvation rate $\sigma$,
-so that $R^*$ is positive.  In fact, when the resource density $R=0$, the
-rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
-The condition $\sigma = \lambda$ represents a transcritical (TC)
-bifurcation~\cite{Strogatz:2014} that demarcates the physical regime from the
-unphysical regime where $F$ would grow exponentially with time.  The
-biological implication of the constraint $\lambda<\sigma$ has a simple
-interpretation---the rate at which a macroscopic organism loses mass due to
-lack of resources is generally much faster than the rate of reproduction.  As
-we will discuss below, this inequality is a natural consequence of allometric
-constraints~\cite{Kempes:2012hy} for organisms within empirically observed
-body size ranges (Fig.~\ref{fig:gvs}).
+From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that $R^*$ is positive.
+In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
+The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates the physical regime from the unphysical regime where $F$ would grow exponentially with time.
+The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
+As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges. %(Fig.~\ref{fig:gvs})
 
-In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may
-either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).  In
-continuous-time systems, a limit cycle arises when a pair of complex
-conjugate eigenvalues crosses the imaginary axis to attain positive real
-parts~\cite{GuckHolmes}.  This Hopf bifurcation is defined by
-${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is
-composed of the coefficients of the characteristic polynomial of the Jacobian
-matrix~\cite{Gross:2004p2428}.  As the system parameters are tuned to be
-within the stable regime but close to the Hopf bifurcation, the amplitude of
-the transient but decaying cycles become large.  Given that ecological
-systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of
-transient cycles, even though they decay with time in the mean-field
-description, can increase the extinction
-risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.  Thus the distance
-of a system from the Hopf bifurcation provides a measure of its persistence.
+In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).  
+In continuous-time systems, a limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts~\cite{GuckHolmes}.  
+This Hopf bifurcation is defined by ${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial of the Jacobian matrix~\cite{Gross:2004p2428}.
+As the system parameters are tuned to be within the stable regime but close to the Hopf bifurcation, the amplitude of the transient but decaying cycles become large.  
+Given that ecological systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase the extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
+Thus the distance of a system from the Hopf bifurcation provides a measure of its persistence.
 
-When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the
-consumers are driven to the hungry non-reproducing state.  Because
-reproduction is inhibited, there is a low steady-state consumer density and a
-high steady-state resource density.  However, if $\sigma/\lambda\to 1$ from
-above, the population is overloaded with energetically-replete (reproducing)
-individuals, thereby promoting oscillations between the consumer and resource
-densities (Fig.~\ref{fig:fp}).
+When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the consumers are driven to the hungry non-reproducing state.  
+Because reproduction is inhibited, there is a low steady-state consumer density and a high steady-state resource density.  
+However, if $\sigma/\lambda\to 1$ from above, the population is overloaded with energetically-replete (reproducing) individuals, thereby promoting oscillations between the consumer and resource densities (Fig.~\ref{fig:fp}).
 
-Whereas the relation between consumer growth rate $\lambda$ and the
-starvation rate $\sigma$ defines an absolute bound of biological
-feasibility---the TC bifurcation---$\sigma$ also
-determines the sensitivity of the consumer population to changes in resource
-density.  When $\sigma\gg\lambda$, the steady-state population density is
-small, thereby increasing the risk of stochastic extinction.  On the other
-hand, as $\sigma$ decreases, the system will ultimately be poised either near
-the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).  If the recovery rate
-$\rho$ is sufficiently small, the TC bifurcation is reached and the resource
-eventually is eliminated.  If $\rho$ exceeds a threshold value, cyclic
-dynamics will develop as the Hopf bifurcation is approached.
+Whereas the relation between consumer growth rate $\lambda$ and the starvation rate $\sigma$ defines an absolute bound of biological feasibility---the TC bifurcation---$\sigma$ also determines the sensitivity of the consumer population to changes in resource density.
+When $\sigma\gg\lambda$, the steady-state population density is small, thereby increasing the risk of stochastic extinction.  
+On the other hand, as $\sigma$ decreases, the system will ultimately be poised either near the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).  
+If the recovery rate $\rho$ is sufficiently small, the TC bifurcation is reached and the resource eventually is eliminated.  
+If $\rho$ exceeds a threshold value, cyclic dynamics will develop as the Hopf bifurcation is approached.
 \\
@@
 related to the conversion efficiency of resource to consumer tissue
-(Supporting Information, section XX).
+(Supporting Information).
 
@@
 where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth.
-We now use this solution to define the timescale (see \cite{moses2008rmo} for a detailed presentation of these timescales) of reproduction and recovery from starvation.
+We now use this solution to define the timescale of reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \cite{moses2008rmo} for a detailed presentation of these timescales).
 The time that it takes to reach a particular mass $\epsilon M$ is given by the timescale
@@
 and recovery by the timescales required to reach, or return from, specific
-fractions of the replete-state mass (see the Supporting Information for details on these parameterizations).  We define
+fractions of the replete-state mass (Fig. \ref{fig:gvs}; see Supporting Information, Table I for parameterizations).  We define
 $m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction of
@@
 recovery timescale, $t_{\rho}$, is the time to go from
-$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$ (Figure \ref{fig:growth}).
+$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$ (Fig.
+\ref{fig:growth}).
 Using Eqs.~\eqref{m1} and \eqref{t1} this timescale is given by simply considering an adjusted starting mass of $m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
@@
 feature is that $\sigma$ does not have a simple scaling dependence on
-$\lambda$, which is important for the dynamics that we
+$\lambda$ (Fig. \ref{fig:gvs}), which is important for the dynamics that we
 later discuss.
@@
 parameterizations for terrestrial-bound endotherms, specifically mammals,
-which range from a minimum of $M\approx1$ gram (the Etruscan shrew
-\emph{Suncus etruscus}) to a maximum of $M\approx10^7$ grams (the late Eocene
+which range from a minimum of $M\approx1$g (the Etruscan shrew
+\emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the late Eocene
 to early Miocene Indricotheriinae).  Investigating other classes of organisms
@@
 \noindent {\bf  Stabilizing effects of allometric constraints} \\
-As the allometric derivations of the NSM rate laws reveal, starvation and
-recovery rates are not independent parameters, and the biologically relevant
-portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via
-covarying parameters.  Given the parameters of terrestrial endotherms, we
-find that $\sigma$ and $\rho$ are constrained to lie within a small window of
-potential values (Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.
-We thus find that the dynamics for all mammalian body sizes is confined to
-the steady-state regime of the NSM and that limit-cycle behavior is
-precluded.
+As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
+Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small window of potential values (Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.
+We thus find that the dynamics for all mammalian body sizes is confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
 %, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
-Moreover, for larger $M$, the distance to the Hopf bifurcation increases,
-while uncertainty in allometric parameters (20\% variation around the mean;
-Fig.~\ref{fig:hopf}) results in little qualitative difference in the distance
-to the the Hopf bifurcation. These results suggest that small mammals are
-more prone to population oscillations---both stable limit cycles and
-transient cycles---than mammals with larger body size.  Thus our NSM model
-predicts that population cycles should be less common for larger species and
-more common for smaller species, particularly in environments where resources
-are limiting.%, and
+Moreover, for larger $M$, the distance to the Hopf bifurcation increases, while uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}) results in little qualitative difference in the distance to the the Hopf bifurcation.
+These results suggest that small mammals are more prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.  
+Thus our NSM model predicts that population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting.%, and
 %2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
 
-It should be noted that previous studies have used allometric constraints to explain the periodicity of cyclic populations
-\cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period
-$\propto M^{0.25}$.  However this relation seems to hold only for some
-species \cite{Hendriks:2012fc}, and potential drivers range from predator
-and/or prey lifespans to competitive
-dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
+It should be noted that previous studies have used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
+However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
 %It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic~\cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
-Statistically significant support for the existence of population cycles
-among mammals is predominantly based on time series for small
-mammals~\cite{Kendall:1998hl}, where our model would predict more pronounced transient dynamics, given how close these points are to
-the Hopf bifurcation.  On the other hand, the longer gestational times and
-the increased difficulty in measurements, precludes obtaining similar-quality
-data for larger organisms.
+Statistically significant support for the existence of population cycles among mammals is predominantly based on time series for small mammals~\cite{Kendall:1998hl}, in agreement with our predictions of more pronounced transient dynamics, given how close these points are to the Hopf bifurcation.
+On the other hand, the longer gestational times and the increased difficulty in measurements, precludes obtaining similar-quality data for larger organisms.
 \vspace{6mm}
@@
 \noindent {\bf Extinction risk}\\
-Within our model, higher rates of starvation result in a larger flux of the
-population to the hungry state.  In this state reproduction is absent, thus
-increasing the likelihood of extinction.  From the perspective of population
-survival, it is the rate of starvation relative to the rate of recovery that
-determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
-We therefore examine the competing effects of cyclic dynamics vs.\ changes in
-steady state density on extinction risk as a function of the ratio
-$\sigma/\rho$.  To this end, we computed the probability of extinction, where
-we define extinction as a population trajectory falling below one fifth of
-the allometrically constrained steady state at any time between $10^6$ and
-$\leq 10^8$.  This procedure is repeated for 1000 replicates of the
-continuous-time system shown in Eq.~\ref{eq:system} for an organism of
-$M=100$ grams.
+Within our model, higher rates of starvation result in a larger flux of the population to the hungry state.  
+In this state reproduction is absent, thus increasing the likelihood of extinction.
+From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
+We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady state density on extinction risk as a function of $\sigma$ and $\rho$.  
+To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one tenth of the allometrically constrained steady state at any time between $t=10^5$ and $t \leq 10^8$.
+This procedure is repeated for 1000 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an organism of $M=100$ grams.
 %the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
-In each replicate the initial densities are chosen to be $A(F^*,H^*,R^*)$,
-with $A$ a random variable that is uniformly distributed in $[0,2]$.  By
-allowing the rate of starvation to vary, we assessed extinction risk across a
-range of values of $\sigma/\rho$ between ca.\ $10^{-3}$ to $5$, thus examining
-a horizontal cross-section of Fig.~\ref{fig:fp}.  As expected, higher rates
-of extinction correlate with both low and high values of $\sigma/\rho$.  For
-low values of $\sigma/\rho$, the increased extinction risk results from
-transient cycles with larger amplitudes as the system nears the Hopf
-bifurcation (Fig.~\ref{fig:ext}).  For large values of $\sigma/\rho$, higher
-extinction risk arises because of the decrease in the steady state consumer
-population density.  This interplay creates an `extinction refuge' as shown
-in Fig.~\ref{fig:ext}, such that for a constrained range of $\sigma/\rho$,
-extinction probabilities are minimized.
+In each replicate the initial densities are chosen to be $A(F^*,H^*,R^*)$, with $A$ a random variable that is uniformly distributed in $[0,2]$.  
+By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-6}$ to $10^{-1}$.
+%, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
+As expected, higher rates of extinction correlate with both high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if $\sigma$ is small.  
+For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
+For high values of $\sigma$ and low values of $\rho$, higher extinction risk arises because of the decrease in the steady state consumer population density (Figs.
+\ref{fig:fp}\emph{B}, \ref{fig:ext}).
+This interplay creates an `extinction refuge', such that for a constrained range of $\sigma$ and $\rho$, extinction probabilities are minimized.
 
-We find that the allometrically constrained values of $\sigma/\rho$ (with
-$20\%$ variability around energetic parameter means) fall within the
-extinction refuge. These values are close enough to the Hopf bifurcation to
-avoid low steady state densities, and far enough away to avoid
-large-amplitude transient cycles. The fact that allometric values of $\sigma$
-and $\rho$ fall within this relatively small window supports the possibility
-that a selective mechanism has constrained the physiological conditions that
-drive starvation and recovery rates within populations.  Such a
-mechanism would select for organism physiology that generates appropriate
-$\sigma$ and $\rho$ values that avoid extinction.  This selection could occur
-via the tuning of body fat percentages, metabolic rates, and biomass
-maintenance efficiencies.
+We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge (Fig. \ref{fig:ext}, white point). 
+These values are close enough to the Hopf bifurcation to avoid low steady state densities, and far enough away to avoid large-amplitude transient cycles. 
+The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
+Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that serve to minimize extinction risk.  
+This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
 % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
@@
 \noindent {\bf Dynamic and energetic barriers to body size} \\
-Metabolite transport constraints are widely thought to place strict
-boundaries on biological
-scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to
-specific predictions on the minimum possible body size for
-organisms~\cite{West:2002ud}.  Above this bound, a number of energetic and
-evolutionary mechanisms have been explored to assess the costs and benefits
-associated with larger body masses, particularly for mammals.  One important
-such example is the \emph{fasting endurance hypothesis}, which contends that
-larger body size, with consequent lower metabolic rates and increased ability
-to maintain more endogenous energetic reserves, may buffer organisms against
-environmental fluctuations in resource availability~\cite{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant
-trend towards larger body size (known as Cope's
-Rule)~\cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br},
-and it is thought that within-lineage drivers generate selection towards an
-optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, whose
-value may arise from higher extinction risk for large taxa over evolutionary
-timescales~\cite{Clauset:2009fh}.  These trends are thought to be driven by a
-combination of climate change and niche availability~\cite{Saarinen:2014br};
-however the underpinning energetic costs and benefits of larger body sizes,
-and how they influence dynamics over ecological timescales, have not been
-explored.  We argue that the NSM provides a suitable framework to explore
-these issues.
+Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\cite{West:2002ud}.
+Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.  
+One important such example is the \emph{fasting endurance hypothesis}, which contends that larger body size, with consequent lower metabolic rates and increased ability to maintain more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability~\cite{Millar:1990p923}.
+Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's rule)~\cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\cite{Clauset:2009fh}.
+These trends are thought to be driven by a combination of climate change and niche availability~\cite{Saarinen:2014br}; however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, have not been explored.
+We argue that the NSM provides a suitable framework to explore these issues.
 
@@
 %that may limit body size (among other factors). 
-The NSM correctly predicts that species with smaller masses have larger
-steady-state population densities (Fig.~\ref{fig:mass}a).
-It should also be noted that $R^{*}$ is decreasing for increasing body size.
-From the perspective of classic resource competition theory we would expect that the species that can live on the lowest resource abundance should outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
-The NSM then predicts that larger mammals should outcompete smaller ones, supporting copes rule.
-This is a natural consequence of the steady-state dynamics combined with the derived timescales.
-
-Furthermore, many previous studies have focused on the energy equivalence hypothesis which argues that the total energy use, $B_{tot}$, of a population is constant independent of species size (e.g.
+The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
+Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis \cite{allen2002,enquist1998}.
+The energy equivalence hypothesis argues that the total energy use, $B_{tot}$, of a population is constant independent of species size (e.g.
 \cite{allen2002,enquist1998}).
+%Previous studies have focused on t
 This hypothesis is based on observations showing that the abundance, $N$, of a species is proportional to the inverse of individual metabolism (e.g.
@@
 \cite{allen2002,enquist1998}).
-This is usually stated as $B_{tot}=NB=constant$ which has been shown to hold in mammalian and vascular plant communities (e.g.
-\cite{allen2002,enquist1998}).
-Figure \ref{} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{} shows that $F^{*}B$ is a constant over this same range.
+This is usually stated as $B_{tot}=NB=C$ where $C$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{allen2002,enquist1998}. 
+Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is relatively constant over this same range.
 This result if remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to the energy equivalence result. 
-Furthermore, the equivalence breaks down at distinct sizes at both the large and small end of mammals suggesting that these are hard limits on mammalian body sizes because organisms outside this range do not meet the constant efficiency obeyed by other populations.
-Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of mammals and $M\approx 6.5*10^7$ for the large end.
-Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is
-close to the minimum observed mammalian body size of ca.\ 1.3-2.5 grams
-(Fig.~\ref{fig:mass}b,c), a range that occurs as the recovery rate begins its
-decline.  In addition to known transport limitations~\cite{West:2002ud}, we
-suggest that an additional constraint of lower body size stems from the
-dynamics of starvation.  This result mirrors other efforts \cite{Kempes:2012hy,Kempes:2016}
-where at a given scale multiple limitations constrain the smallest possibilities for
-life within a class of organisms.
+Our model shows that the equivalence breaks down at both the minimum and maximum observed body sizes for mammals, suggesting that these are hard limits where deviations outside of this range are energetically suboptimal.
+%, violating the energy equivalence.
+% set by deviations from constant efficiency obeyed by other populations.
+Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end. 
+%Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is close to the minimum observed mammalian body size of ca.\ 1.3-2.5g, a range that occurs as the recovery rate begins its decline.
+%In addition to known transport limitations~\cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.  
+%This result mirrors other efforts \cite{Kempes:2012hy,Kempes:2016} where at a given scale multiple limitations constrain the smallest possibilities for life within a class of organisms.
 
-%
-%However we observe
-%that there is a sharp change in the mass dependence of both the steady-state
-%densities and $\sigma/\rho$ at $M \approx 0.3$ grams
-%(Fig.~\ref{fig:mass}b,c).  The dependence of the rates of starvation and
-%recovery explain this phenomenon.  As the mass decreases, the rate of
-%starvation increases, while the rate of recovery declines precipitously.
-%This decline in $\rho$ occurs when the percentage of body fat is
-%$1 - \big(a/b\big)^{-1/\left(\eta-1\right)}M^{-1} \approx 2\%$, whereupon
-%consumers are unable to recover from the hungry state.
+%The combined steady-state dynamics and allometric timescales of the NSM predict that larger mammals should outcompete smaller ones, and this suggests that the model may provide a framework with which to explore the energetic mechanisms behind Cope's rule.
+%, which we explore further below.
 
-%Maximum mass bound
-%Although there are upper bounds to the rate equations (e.g. when percent body fat becomes unity), they are not biologically feasible and we do not discuss them further.
-A additional theoretical upper bound on mammalian body size is given by $\epsilon=0$, where mammals are entirely composed of metabolic reserves, and this occurs at a size of $M=8.3\times 10^8$, or $4.5\times$ the mass of a blue whale.
-We determine a more realistic upper bound to body mass by assessing the
-susceptibility of an otherwise homogeneous population to invasion by a
-mutated subset of the population (denoted by ${}^\prime$) where individuals
-have a modified proportion of body fat $M^\prime=M(1+\chi)$ where
-$\chi \in [-0.5,0.5]$, thus altering the rates of starvation $\sigma$,
-recovery $\rho$, and maintenance $\beta$.  There is no internal fixed point
-that correspond to a state where both original residents and invaders coexist
-(except for the trivial state $\chi=0$).  To assess the susceptibility to
-invasion as a function of the invader mass, we determine which consumer has a
-higher steady-state density for a given value of $\chi$.  We find that for
-$1\leq M<10^6$g, having additional body fat ($\chi > 0$) results in a higher
-steady-state invader population density
-($H^{\prime *}+F^{\prime *}>H^*+F^*$).  Thus the invader has an intrinsic
-advantage over the resident population.  However, for $M>10^6$, leaner
-individuals ($\chi < 0$) have advantageous steady state densities.
+We contend that the NSM provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size as well as the observed trend towards larger body size over evolutionary time.
+The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
+\ref{fig:mass}\emph{C}), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
+Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but does not offer a within-lineage mechanism by which larger body sizes are selected for.
+
+To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $4.5\times$ the mass of a blue whale.
+Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-linneage mechanism for Cope's rule. 
+We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-1,1]$, thus altering the rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and maintenance $\beta(M^\prime)$.
+There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).  
+To assess the susceptibility to invasion as a function of the invader mass, we determine which consumer has a higher steady-state density for a given value of $\chi$.  
+We find that for $1\leq M<8.43\times 10^6$g, having additional body fat ($\chi > 0$) results in a higher steady-state invader population density ($H^{\prime *}+F^{\prime *}>H^*+F^*$).  
+Thus the invader has an intrinsic advantage over the resident population.  
+However, for $M>8.43\times 10^6$, leaner individuals ($\chi < 0$) have advantageous steady state densities.
 %, and this is due to the changing
@@
 
-The observed switch in susceptibility as a function of $\chi$ at
-$M_{\rm opt}\approx10^6$ thus serves as an attractor, or an uninvadible evolutionary stable state, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and
-decrease if $M>M_{\rm opt}$.  Moreover, $M_{\rm opt}$, which is entirely
-determined by the population-level consequences of energetic constraints, is
-within an order of magnitude of the mass observed in the North American
-mammalian fossil record~\cite{Alroy:1998p1594} and also the mass predicted
-from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
-While the state of the environment, as well as the competitive landscape,
-will determine whether specific body sizes are selected for or
-against~\cite{Saarinen:2014br}, we suggest that the starvation dynamics
-outlined here may provide the driving mechanism for the evolution of larger
-body size among terrestrial mammals.
+The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}=8.43 \times 10^6$ thus serves as an attractor, or an uninvadible evolutionary stable state, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
+%This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
+Moreover, $M_{\rm opt}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
+While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we suggest that the dynamics of starvation and recovery described in the NSM may provide a general driving mechanism for the evolution of larger body size among terrestrial mammals.
 %and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
 
-One might be concerned a greater number of large mammals are currently not
-observed in the modern world given that larger mammals are less susceptible
-to extinction.  However, recent research suggests that the pleistocene may
-have been much more populated with a significant diversity of very large
-mammals \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also
-much more geographically widespread than today. These results, combined with
-our findings, suggest that the modern diversity of mammals may not represent a
-true steady state the current distribution of nutrients and large seeds may
-be very different from the past
-\cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
+%One might be concerned a greater number of large mammals are currently not observed in the modern world given that larger mammals are less susceptible to extinction.  
+%However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also much more geographically widespread than today.
+%These results, combined with our findings, suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
 
 %Closing
-The energetics associated with somatic maintenance, growth, and reproduction
-are important elements that influence the dynamics of all
-populations~\cite{Stearns:1989ip}.  The NSM is a minimal and general model
-that incorporates the dynamics of starvation that are expected to occur in
-resource-limited environments.  By incorporating allometric relations between
-the rates in the NSM, we found: (i) different organismal masses have distinct
-population dynamic regimes, (ii) allometrically-determined rates of
-starvation and recovery appear to minimize extinction risk, and (iii) the
-dynamic consequences of these rates may place additional barriers on the
-evolution of minimum and maximum body size.  We suggest that the NSM offers a
-means by which the dynamic consequences of energetic constraints can be
-assessed using macroscale interactions between and among species.  Future
-efforts will involve exploring the consequences of these dynamics in a
-spatially explicit framework, thus incorporating elements such as movement
-costs and spatial heterogeneity, which may elucidate additional tradeoffs
-associated with the dynamics of starvation.
+The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations~\cite{Stearns:1989ip}.  
+The NSM is a minimal and general model that incorporates the dynamics of starvation and recovery that are expected to occur in resource-limited environments.  
+By incorporating allometric relations between the rates in the NSM, we found: (i) different organismal masses have distinct population dynamic regimes, (ii) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and (iii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of minimum and maximum body size.
+We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.  
+Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.
 
@@
 \begin{acknowledgments}
-  We thank Luis Bettancourt, Eric Libby, and Seth Newsome for helpful
+  We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
   discussions and comments on the manuscript.  J.D.Y. was supported by
@@
 \centering
-\includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-\caption{ The growth trajectory over absolute time of an individual organism
-  as a function of body mass.  Initial growth follows the red trajectory to
-  an energetically replete adult mass $M$.  Starvation follows the concave
-  blue trajectory to $m_{\rm starve} < M$, whereas recovery follows the
-  convex growth trajectory from $m_\sigma$ to $M$.  }
-\label{fig:growth}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_Rates.pdf}
-\caption{ 
-Allometrically constrained starvation rate $\sigma$ (red) and recovery rate $\rho$ relative to the reproductive rate $\lambda$ (yellow) as a function of body mass.  
-The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
-Mean values $\pm 20\%$ variation are shown by the shaded region for each rate.
-}
-\label{fig:gvs}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_FixedPoint.pdf}
+\includegraphics[width=0.75\textwidth]{fig_FixedPoint.pdf}
 \caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
-  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100 gram consumer.  These
+  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
   bifurcation conditions separate parameter space into infeasible, cyclic,
@@
 \centering
+\includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
+\caption{ The growth trajectory over absolute time of an individual organism
+  as a function of body mass.  Initial growth follows the black trajectory to
+  an energetically replete reproductive adult mass $m=\epsilon_\lambda M$ which we assume is 95\% asymptotic mass $M$.  Starvation follows the red
+  trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$, and recovery follows the
+  green growth trajectory to the replete adult mass. Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}
+\label{fig:growth}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.5\textwidth]{fig_Rates.pdf}
+\caption{ 
+Allometrically constrained starvation rate $\sigma$ (red) and recovery rate $\rho$ (green) relative to the reproductive rate $\lambda$ (orange) as a function of body mass.  
+The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
+Mean values $\pm 20\%$ variation are shown by the shaded region for each rate.
+}
+\label{fig:gvs}
+\end{figure*}
+
+\begin{figure*}
+\centering
 \includegraphics[width=0.5\textwidth]{fig_DataHopf.pdf}
@@
   allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
-  a function of different mammalian body sizes: $M=A\times10^1$ gram (blue) and
-  $M=A\times10^6$ grams (green), where $A$ is a random uniform variable in $[1,9]$.
+  a function of different mammalian body sizes: $M=A\times10^1$g (blue) and
+  $M=A\times10^6$g (green), where $A$ is a random uniform variable in $[1,9]$.
+  Points denote realized values of $\sigma$ and $\rho$ given the drawn values for $M$.
   }
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric.pdf}
+\includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric.jpg}
 \caption{ Probability of extinction for a 100g consumer as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $A(F^*,H^*,R^*)$, with
@@
 \includegraphics[width=0.45\textwidth]{fig_FPAllometric.pdf}
-\caption{ (\emph{A}) Consumer steady states $F^*$ and $H^*$ as a function of
+\caption{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
   body mass. 
-  (\emph{B}) Resource steady state $R^*$ as a function of consumer body mass.
-  (\emph{C}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.}
+  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
+  (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.}
 \label{fig:mass}
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig_Invasion.pdf}
+\includegraphics[width=0.75\textwidth]{fig_Invasion.pdf}
 \caption{ Invasion feasibility for organisms with a proportional change in
@@
   starvation threshold and is thus infeasible.
-  Arrows point to the predicted optimal mass $M_{\rm opt}=8.43\times 10^6$, which serves as the evolutionary stable state.}
+  Arrows point to the predicted optimal mass $M_{\rm opt}=8.43\times 10^6$, which serves as the uninvadable, evolutionary stable state.}
 \label{fig:invasion}

### 2016-08-24 Manuscript/9.tex
Final analyses before first submission to PNAS

--- 2016-08-24 Manuscript/11.tex
@@
   to these constraints can approach the cyclic regime but are typically
-  driven to a steady state.  Moreover, we find that these rates fall within a
+  driven to a steady state.  Moreover, we find these rates fall within a
   `refuge' in parameter space, where the probability of extinction of the
@@
   population to a mutant population with altered percent body fat.  For body
-  masses $\lesssim 10^7$g, a more apidose invader dominates over the resident
-  population, and vice versa for body mass $\gtrsim 10^7$g, thus providing a
+  masses $\lesssim 10^6$g, a more apidose invader dominates over the resident
+  population, and vice versa for body mass $\gtrsim 10^6$g, thus providing a
   principled mechanism for a within-lineage driver of Cope's rule.
@@
 For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$ \cite{West:2001bv}).
-The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$ this will scale like $\lambda\propto M^{\eta-1}$ for $M>>m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
+The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced (see supplement), and for any constant value of $\epsilon_{\lambda}$ this will scale like $\lambda\propto M^{\eta-1}$ for $M>>m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
 
@@
 recovery timescale, $t_{\rho}$, is the time to go from
-$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$ (Figure \ref{fig:growth}).
+$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$.
 Using Eqs.~\eqref{m1} and \eqref{t1} this timescale is given by simply considering an adjusted starting mass of $m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
@@
 \end{eqnarray}
-where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in the biosynthetic energetics during recovery (see Supporting Information).
+where $a^{\prime}$ accounts for possible deviations in the biosynthetic energetics during recovery (see Supporting Information).
 It should be noted that more complicated ontogenetic models explicitly handle
@@
 
-It should be noted that previous studies have used allometric constraints to explain the periodicity of cyclic populations
+\sid{I don't see the point of the paragraph below.}
+Previous studies have used allometric constraints to explain the periodicity
+of cyclic populations
 \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period
@@
 Furthermore, the equivalence breaks down at distinct sizes at both the large and small end of mammals suggesting that these are hard limits on mammalian body sizes because organisms outside this range do not meet the constant efficiency obeyed by other populations.
-Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of mammals and $M\approx 6.5*10^7$ for the large end.
+Significant deviations from constant energy use occur at $Justin -VALUE$ at the small end of mammals and $Justin -VALUE$ for the large end.
 Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is
@@
 advantage over the resident population.  However, for $M>10^6$, leaner
-individuals ($\chi < 0$) have advantageous steady state densities.
-%, and this is due to the changing
-%covariance between energetic rates as a function of modified energetic
-%reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
+individuals ($\chi < 0$) have the advantage, and this is due to the changing
+covariance between energetic rates as a function of modified energetic
+reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
 
 The observed switch in susceptibility as a function of $\chi$ at
-$M_{\rm opt}\approx10^6$ thus serves as an attractor, or an uninvadible evolutionary stable state, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and
+$M_{\rm opt}\approx10^6$ thus serves as an attractor, where over evolutionary
+time the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and
 decrease if $M>M_{\rm opt}$.  Moreover, $M_{\rm opt}$, which is entirely
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig_Rates.pdf}
-\caption{ 
-Allometrically constrained starvation rate $\sigma$ (red) and recovery rate $\rho$ relative to the reproductive rate $\lambda$ (yellow) as a function of body mass.  
-The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
-Mean values $\pm 20\%$ variation are shown by the shaded region for each rate.
-}
+\includegraphics[width=0.5\textwidth]{fig_GvS.pdf}
+\caption{ Allometrically constrained starvation rate $\sigma$ (red) vs.\
+  reproductive rate $\lambda$ (yellow) as a function of mass $M$.  The rate
+  of starvation is greater than the rate of reproduction for all realized
+  terrestrial endotherm body sizes.  Each data point is determined by fixing
+  the mass $M$ and then choosing parameters in a $\pm 20$\% range about the
+  allometric values.}
 \label{fig:gvs}
@@
 \caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
-  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100 gram consumer.  These
+  function of the starvation rate $\sigma$ and recovery rate $\rho$.  These
   bifurcation conditions separate parameter space into infeasible, cyclic,
   and steady state dynamic regimes.  The color gradient shows the steady
-  state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
+  state densities for (A) the resource $R^*$ and the (B) energetically
   replete consumers $F^*$, (warmer colors denote higher densities).  Steady
   state densities for the energetically deficient consumers $H^*$ (not shown)
-  scale with those for $F^*$.  }
+  closely mirror those for $F^*$.  }
 \label{fig:fp}
@@
 \includegraphics[width=0.5\textwidth]{fig_DataHopf.pdf}
-\caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
+\caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) with
   allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
-  a function of different mammalian body sizes: $M=A\times10^1$ gram (blue) and
-  $M=A\times10^6$ grams (green), where $A$ is a random uniform variable in $[1,9]$.
-  }
+  a function of minimum and maximum mammalian body sizes: $1$ gram (blue) and
+  $10^7$ grams (orange), respectively.  Replicates show the role of the 20\%
+  variation about the mean allometric parameters, which influences both the
+  energetic rates as well as the position of the TC and Hopf bifurcations.  }
 \label{fig:hopf}
@@
 \includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric.pdf}
-\caption{ Probability of extinction for a 100g consumer as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $A(F^*,H^*,R^*)$, with
-  $A$ being a random uniform variable in $[0,2]$.  Extinction is defined as the
-  population trajectory falling below $0.1~\times$ the allometrically
-  constrained steady state. The white point denotes the allometrically constrained starvation and recovery rate.}
+\caption{ Probability of extinction for 1000 consumer population trajectories
+  as a function of $\sigma/\rho$ for initial densities $A(F^*,H^*,R^*)$, with
+  $A$ a random uniform variable in $[0,2]$.  Extinction is defined as the
+  population trajectory going below $0.2~\times$ the allometrically
+  constrained steady state at any time between $10^2$ and $\leq 10^6$.
+  \sid{Insert a distribution instead of a scatter plot above the main panel
+    and describe it precisely} The values above the extinction plot are the
+  allometrically constrained $\sigma/\rho$ with 20\% variation around the
+  mean.  }
 \label{fig:ext}
@@
 \caption{ (\emph{A}) Consumer steady states $F^*$ and $H^*$ as a function of
-  body mass. 
-  (\emph{B}) Resource steady state $R^*$ as a function of consumer body mass.
-  (\emph{C}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.}
+  body size.  Energetic rates as a function of body size, with the ratio
+  $\sigma/\rho$ (\emph{B}) and both $\sigma$ (red) and $\rho$ (purple;
+  \emph{C}) drawn separately with 20\% variation around the mean.  Steady
+  state densities decline sharply at $M=M_{\rm min}$ due to the precipitous
+  decrease in the recovery rate.  The minimum mammalian body mass (Etruscan
+  shrew) is denoted by the orange shaded region that coincides with the
+  decline of the recovery rate.  }
 \label{fig:mass}
@@
   mass $\chi$ against a population with a resident body mass $M$.  The blue
-  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
+  region denotes values of $\chi$ that result in successful invasion.  The
   red region denotes values of $\chi$ that result in a mass that is below the
-  starvation threshold and is thus infeasible.
-  Arrows point to the predicted optimal mass $M_{\rm opt}=8.43\times 10^6$, which serves as the evolutionary stable state.}
+  starvation threshold and is thus infeasible.  }
 \label{fig:invasion}

### 2016-08-24 Manuscript/Supplement.tex
Final analyses before first submission to PNAS

@@
 
-Metabolic rate has been often reported to follow an exponent close to $0.75$ (e.g.
+Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g.
 \cite{West:2001bv,moses2008rmo} and the supplement of \cite{hou}).
-We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \cite{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around all parameter values.
-It is important to note the exponent, because it defines the value of the metabolic normalization constant, $B_{0}$, given a set of data.
+We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \cite{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
+It is important to note the exponent, because it not only defines several scalings in our framework but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.
 For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$) \cite{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
-We employ the field metabolic rate for our NSM model which is appropriate for active mammals.
+We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).
 
@@
 \cite{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \cite{moses2008rmo}.
-The unit energy available during starvation, $E^{\prime}$, can be considered in several ways.
-The first is to assume that the total energy stored during ontogeny is returned during starvation which would give a value of $7000$ (J g$^{-1}$) \cite{hou}.
-However, since our model considers the consumption of all body fat as defining the transition to starving, it is more appropriate to consider the energetics of fat metabolism where we would expect $E^{\prime}=36,000$ (J g$^{-1}$) for palmitate\cite{stryer,hou}.
+The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \cite{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \cite{stryer,hou}.
+For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \cite{hou}.
 
-The energy required for maintaining an existing unit of mass is reported to be
+For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \cite{Dunbrack:1993ec}, $0.02M^{1.19}$ \cite{Lindstedt:1985hm}, and $0.026M^{1.14}$ \cite{Lindstedt:2002td}.
+We use the values from \cite{Lindstedt:1985hm} which falls in the middle of this range.
+Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \cite{Lindstedt:2002td}.
 
-\onecolumn
+The final parameters that we must consider connect the resource growth rate to the total metabolic rate of an organism.
+That is, we are interested in the relative rates of resource recovery and consumption by the total population.
+From \cite{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$).
+Considering an energy density of $18200$ (J g$^{-1}$) of grass \cite{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$).
+This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.
+
+%For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
+
+%More complicated models of fecundity (which, for example, account for the average length of adulthood and the number of individuals produced over this span) could be employed.
+However, the scaling of population growth rate has been studied in detail before and follows a relationship of $$ which matches the theory well for $\phi=.95$ and $=$.
+
+%In our calculations we include $20\%$ variation around this value which could account for differences in efficiency during
 
  \begin{table}[h]
-\caption{Parameter Values For Various Classes of Organisms}
+\caption{Parameter values for mammals}
 \label{param}
@@
     \small
-     \begin{tabular}{| p{1.2cm}| p{4.2cm} | }
+     \begin{tabular}{ p{1.2cm} p{3.2cm} l p{2.2cm}|}
      \hline
-     & {\bf Mammals}  \\
+     Parameter & Value & References  \\
      \hline
-   $\eta$ & $3/4$  \quad \quad  (e.g. \cite{West:2001bv,moses2008rmo,hou}) \\ 
-   $E_{m}$ & $5774$ (J gram$^{-1}$)  \quad \quad  \cite{moses2008rmo,West:2001bv,hou} \\ 
-   $E_{m}^{\prime}$ & $36,000$  \quad \quad \cite{stryer,hou} \\ 
-   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    \quad \quad \cite{hou}  \\
-   $B_{m}$ & $0.025$ (W gram$^{-1}$) \quad \quad \\
-   $a$ & $1.78\times10^{-6}$  \quad \quad \\ 
-   $b$ & $2.29\times10^{-6}$  \quad \quad \\  
-   $\eta-1$ & $-0.21$  \quad \quad \\ 
-   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\ 
-   $\gamma$ & $1.19$ \quad \quad \\ 
-   $f_{0}$ & $0.02$ \quad \quad \\ 
-   $\zeta$ & $1.01$  \quad \quad \\ 
-   $mm_{0}$ & $0.32$  \quad \quad \\
+   $\eta$ & $3/4$  &  (e.g. \cite{West:2001bv,moses2008rmo,hou}) \\ 
+   $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \cite{moses2008rmo,West:2001bv,hou} \\ 
+   $E_{m}^{\prime}$ & $36,000$  & \cite{stryer,hou} \\ 
+   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \cite{hou}  \\
+%   $a$ & $1.78\times10^{-6}$  \quad \quad \\ 
+%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\ 
+   $\gamma$ & $1.19$ & \cite{Lindstedt:1985hm} \\ 
+   $f_{0}$ & $0.02$ & \cite{Lindstedt:1985hm}\\ 
+   $\zeta$ & $1.00$  & \cite{Lindstedt:2002td} \\ 
+   $u_{0}$ & $0.38$  & \cite{Lindstedt:2002td} \\
 
@@
    \end{table}
+
+\subsection*{Rate equations for invaders with modified body mass}
+If an invading subset of the resident population of mass $M$ has an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $[-1,1]$ ($\chi<0$ denotes a leaner invader; $\chi > 0$ denotes an invader with more endogenous reserves), the invading population will have the following modified rates: $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$.
+Because we are assuming that the invading population is only modifying its endogenous energetic stores, we assume that the proportion of body mass that is non-adipose tissue remains the same as the resident population.
+This assumption leads to the following modified timescales:
+
+\begin{align}
+t_{\sigma^\prime} &= \frac{-M^{1/4}}{B_0/E_m^\prime}\log \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
+t_{\rho^\prime} &= \frac{-4 M^{1/4} }{B_0/E_m^\prime}\log \left(\frac{1-( \epsilon_\lambda(\chi +1))^{1/4}}{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}\right), \\ \nonumber
+t_{\beta^\prime} &= \xi B_0\left(M(\chi + 1)\right)^{3/4}.
+\end{align}
 

### 2017-02-09 Manuscript/12.tex
new yield constants

@@
 
-\usepackage{pnastwof}
+\usepackage{pnastwoF}
 
@@
 
-\bibliography{aa_starving3}
+%\bibliography{aa_starving3}
+
+\begin{thebibliography}{10}
+
+\bibitem{Martin:1987dl}
+Martin TE
+\newblock (1987) {Food as a Limit on Breeding Birds: A Life-History
+  Perspective}.
+\newblock \emph{Annu. Rev. Ecol. Syst.} 18:453--487.
+
+\bibitem{Kirk:1997cc}
+Kirk KL
+\newblock (1997) {Life-History Responses to Variable Environments: Starvation
+  and Reproduction in Planktonic Rotifers}.
+\newblock \emph{Ecology} 78:434--441.
+
+\bibitem{Kempes:2012hy}
+Kempes CP, Dutkiewicz S, Follows MJ
+\newblock (2012) {Growth, metabolic partitioning, and the size of
+  microorganisms.}
+\newblock \emph{PNAS} 109:495--500.
+
+\bibitem{Mangel:1988uaa}
+Mangel M, Clark CW
+\newblock (1988) \emph{{Dynamic Modeling in Behavioral Ecology}}
+\newblock (Princeton University Press, Princeton).
+
+\bibitem{Mangel:2014kz}
+Mangel M
+\newblock (2014) {Stochastic dynamic programming illuminates the link between
+  environment, physiology, and evolution}.
+\newblock \emph{B. Math. Biol.} 77:857--877.
+
+\bibitem{Yeakel:2013hi}
+Yeakel JD, Dominy NJ, Koch PL, Mangel M
+\newblock (2014) {Functional morphology, stable isotopes, and human evolution:
+  a model of consilience}.
+\newblock \emph{Evolution} 68:190--203.
+
+\bibitem{Morris:1987eo}
+Morris DW
+\newblock (1987) {Optimal Allocation of Parental Investment}.
+\newblock \emph{Oikos} 49:332.
+
+\bibitem{Tveraa:2003fq}
+Tveraa T, Fauchald P, Henaug C, Yoccoz NG
+\newblock (2003) {An examination of a compensatory relationship between food
+  limitation and predation in semi-domestic reindeer}.
+\newblock \emph{Oecologia} 137:370--376.
+
+\bibitem{Daan:1988va}
+Daan S, Dijkstra C, Drent R, Meijer T
+\newblock (1988) \emph{{Food supply and the annual timing of avian
+  reproduction}}.
+
+\bibitem{Jacot:2009dv}
+Jacot A, Valcu M, van Oers K, Kempenaers B
+\newblock (2009) {Experimental nest site limitation affects reproductive
+  strategies and parental investment in a hole-nesting passerine}.
+\newblock \emph{Animal Behaviour} 77:1075--1083.
+
+\bibitem{Stearns:1989ip}
+Stearns SC
+\newblock (1989) {Trade-Offs in Life-History Evolution}.
+\newblock \emph{Funct. Ecol.} 3:259.
+
+\bibitem{Barboza:2002in}
+Barboza P, Jorde D
+\newblock (2002) {Intermittent fasting during winter and spring affects body
+  composition and reproduction of a migratory duck}.
+\newblock \emph{J Comp Physiol B} 172:419--434.
+
+\bibitem{Threlkeld:1976ih}
+Threlkeld ST
+\newblock (1976) {Starvation and the size structure of zooplankton
+  communities}.
+\newblock \emph{Freshwater Biol.} 6:489--496.
+
+\bibitem{Weber:1998jg}
+Weber TP, Ens BJ, Houston AI
+\newblock (1998) {Optimal avian migration: A dynamic model of fuel stores and
+  site use}.
+\newblock \emph{Evolutionary Ecology} 12:377--401.
+
+\bibitem{Mduma:1999cp}
+Mduma SAR, Sinclair ARE, Hilborn R
+\newblock (1999) {Food regulates the Serengeti wildebeest: a 40-year record}.
+\newblock \emph{J. Anim. Ecol.} 68:1101--1122.
+
+\bibitem{Moore:2014hi}
+Moore JW, Yeakel JD, Peard D, Lough J, Beere M
+\newblock (2014) {Life-history diversity and its importance to population
+  stability and persistence of a migratory fish: steelhead in two large North
+  American watersheds}.
+\newblock \emph{J. Anim. Ecol.} 83:1035--1046.
+
+\bibitem{Mead:1989dt}
+Mead RA
+\newblock (1989) in \emph{Carnivore Behavior, Ecology, and Evolution}
+\newblock (Springer US, Boston, MA), pp 437--464.
+
+\bibitem{Sandell:1990kw}
+Sandell M
+\newblock (1990) {The Evolution of Seasonal Delayed Implantation}.
+\newblock \emph{The Quarterly Review of Biology} 65:23--42.
+
+\bibitem{Bulik:1999eo}
+Bulik CM, {et~al.}
+\newblock (1999) {Fertility and Reproduction in Women With Anorexia Nervosa}.
+\newblock \emph{J. Clin. Psychiatry} 60:130--135.
+
+\bibitem{Trites:2003cc}
+Trites AW, Donnelly CP
+\newblock (2003) {The decline of Steller sea lions Eumetopias jubatus in
+  Alaska: a review of the nutritional stress hypothesis}.
+\newblock \emph{Mammal Review} 33:3--28.
+
+\bibitem{Glazier:2009hq}
+Glazier DS
+\newblock (2009) {Metabolic level and size scaling of rates of respiration and
+  growth in unicellular organisms}.
+\newblock \emph{Funct. Ecol.} 23:963--968.
+
+\bibitem{Kooi2000}
+Kooijman SALM
+\newblock (2000) \emph{{Dynamic Energy and Mass Budgets in Biological Systems}}
+\newblock (Cambridge).
+
+\bibitem{Sousa:2010ez}
+Sousa T, Domingos T, Poggiale JC, Kooijman SALM
+\newblock (2010) {Dynamic energy budget theory restores coherence in biology}.
+\newblock \emph{Philos. T. Roy. Soc. B} 365:3413--3428.
+
+\bibitem{Diekmann:2010da}
+Diekmann O, Metz JAJ
+\newblock (2010) {How to lift a model for individual behaviour to the
+  population level?}
+\newblock \emph{Philos. T. Roy. Soc. B} 365:3523--3530.
+
+\bibitem{murdoch:2003}
+Murdoch WW, Briggs CJ, Nisbet RM
+\newblock (2003) \emph{{Consumer-resource Dynamics}}, Monographs in population
+  biology
+\newblock (Princeton University Press).
+
+\bibitem{Benichou:2014wu}
+Benichou O, Redner S
+\newblock (2014) {Depletion-Controlled Starvation of a Diffusing Forager}.
+\newblock \emph{arXiv}.
+
+\bibitem{Benichou:2016wl}
+B{\'e}nichou O, Chupeau M, Redner S
+\newblock (2016) {Role of Depletion on the Dynamics of a Diffusing Forager}.
+
+\bibitem{Chupeau:2016jf}
+Chupeau M, B{\'e}nichou O, Redner S
+\newblock (2016) {Universality classes of foraging with resource renewal}.
+\newblock \emph{Phys. Rev. E} 93:032403.
+
+\bibitem{murray2011mathematical}
+Murray JD
+\newblock (2011) \emph{{Mathematical Biology: I. An Introduction}},
+  Interdisciplinary Applied Mathematics
+\newblock (Springer New York, Melaka Manipal Medical College (Manipal Campus),
+  International Centre for Health Sciences, Madhav Nagar, Manipal, Udupi
+  District, Karnataka State, India. nayaksathish@yahoo.com) Vol.{} 110.
+
+\bibitem{Strogatz:2008wo}
+Strogatz SH
+\newblock (2008) \emph{{Nonlinear Dynamics and Chaos}}, With Applications to
+  Physics, Biology, Chemistry, and Engineering
+\newblock (Westview Press).
+
+\bibitem{GuckHolmes}
+Guckenheimer J, Holmes P
+\newblock (1983) \emph{{Nonlinear oscillations, dynamical systems, and
+  bifurcations of vector fields}}
+\newblock (Springer, New York).
+
+\bibitem{Gross:2004p2428}
+Gross T, Feudel U
+\newblock (2004) {Analytical search for bifurcation surfaces in parameter
+  space}.
+\newblock \emph{Physica D} 195:292--302.
+
+\bibitem{Hastings:2001jh}
+Hastings A
+\newblock (2001) {Transient dynamics and persistence of ecological systems}.
+\newblock \emph{Ecol. Lett.} 4:215--220.
+
+\bibitem{Neubert:1997wk}
+Neubert M, Caswell H
+\newblock (1997) {Alternatives to resilience for measuring the responses of
+  ecological systems to perturbations}.
+\newblock \emph{Ecology} 78:653--665.
+
+\bibitem{Caswell:2005eo}
+Caswell H, Neubert MG
+\newblock (2005) {Reactivity and transient dynamics of discrete-time ecological
+  systems}.
+\newblock \emph{Journal of Difference Equations and Applications} 11:295--310.
+
+\bibitem{Neubert:2009td}
+Neubert M, Caswell H
+\newblock (2009) {Detecting reactivity}.
+\newblock \emph{Ecology}.
+
+\bibitem{Yodzis:1992hg}
+Yodzis P, Innes S
+\newblock (1992) {Body Size and Consumer-Resource Dynamics}.
+\newblock \emph{Am. Nat.} 139:1151--1175.
+
+\bibitem{Brown:2004wq}
+Brown J, Gillooly J, Allen A, Savage V, West G
+\newblock (2004) {Toward a metabolic theory of ecology}.
+\newblock \emph{Ecology} 85:1771--1789.
+
+\bibitem{West:2002it}
+West GB, Woodruff WH, Brown JH
+\newblock (2002) {Allometric scaling of metabolic rate from molecules and
+  mitochondria to cells and mammals.}
+\newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.
+
+\bibitem{DeLong:2010dy}
+DeLong JP, Okie JG, Moses ME, Sibly RM, Brown JH
+\newblock (2010) {Shifts in metabolic scaling, production, and efficiency
+  across major evolutionary transitions of life.}
+\newblock \emph{PNAS} 107:12941--12945.
+
+\bibitem{West:2001bv}
+West GB, Brown JH, Enquist BJ
+\newblock (2001) {A general model for ontogenetic growth}.
+\newblock \emph{Nature} 413:628--631.
+
+\bibitem{moses2008rmo}
+Moses ME, {et~al.}
+\newblock (2008) {Revisiting a Model of Ontogenetic Growth: Estimating Model
+  Parameters from Theory and Data}.
+\newblock \emph{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}
+  171:632--645.
+
+\bibitem{gillooly2002esa}
+Gillooly JF, Charnov EL, West GB, Savage VM, Brown JH
+\newblock (2002) {Effects of size and temperature on developmental time}.
+\newblock \emph{Nature} 417:70--73.
+
+\bibitem{hou}
+Hou C, {et~al.}
+\newblock (2008) {Energy Uptake and Allocation During Ontogeny}.
+\newblock \emph{Science} 322:736--739.
+
+\bibitem{bettencourt}
+Bettencourt LMA, Lobo J, Helbing D, Kuhnert C, West GB
+\newblock (2007) {Growth, innovation, scaling, and the pace of life in cities}.
+\newblock \emph{Proc. Natl. Acad. Sci. USA} 104:7301--7306.
+
+\bibitem{Folland:2008ij}
+Folland JP, Mc~Cauley TM, Williams AG
+\newblock (2008) {Allometric scaling of strength measurements to body size}.
+\newblock \emph{Eur J Appl Physiol} 102:739--745.
+
+\bibitem{CalderIII:1983jd}
+Calder~III WA
+\newblock (1983) {An allometric approach to population cycles of mammals}.
+\newblock \emph{J. Theor. Biol.} 100:275--282.
+
+\bibitem{Peterson:1984hj}
+Peterson RO, PAGE RE, DODGE KM
+\newblock (1984) {Wolves, Moose, and the Allometry of Population Cycles}.
+\newblock \emph{Science} 224:1350--1352.
+
+\bibitem{Krukonis:1991fk}
+Krukonis G, Schaffer WM
+\newblock (1991) {Population cycles in mammals and birds: Does periodicity
+  scale with body size?}
+\newblock \emph{J. Theor. Biol.} 148:469--493.
+
+\bibitem{Hendriks:2012fc}
+Hendriks AJ, Mulder C
+\newblock (2012) {Delayed logistic and Rosenzweig{\textendash}MacArthur models
+  with allometric parameter setting estimate population cycles at lower trophic
+  levels well}.
+\newblock \emph{Ecological Complexity} 9:43--54.
+
+\bibitem{Kendall:1999iy}
+Kendall BE, {et~al.}
+\newblock (1999) {Why do populations cycle? A synthesis of statistical and
+  mechanistic modeling approaches}.
+\newblock \emph{Ecology} 80:1789--1805.
+
+\bibitem{Hogstedt:2005cr}
+H{\"o}gstedt G, Seldal T, Breistol A
+\newblock (2005) {Period length in cyclic animal populations}.
+\newblock \emph{Ecology} 86:373--378.
+
+\bibitem{Kendall:1998hl}
+Kendall, Prendergast, Bjornstad
+\newblock (1998) {The macroecology of population dynamics: taxonomic and
+  biogeographic patterns in population cycles}.
+\newblock \emph{Ecol. Lett.} 1:160--164.
+
+\bibitem{Brown:1993p708}
+Brown J, Marquet P, Taper M
+\newblock (1993) {Evolution of body size: consequences of an energetic
+  definition of fitness}.
+\newblock \emph{Am. Nat.} 142:573--584.
+
+\bibitem{West:1997cg}
+West GB, Brown JH, Enquist BJ
+\newblock (1997) {A General Model for the Origin of Allometric Scaling Laws in
+  Biology}.
+\newblock \emph{Science} 276:122--126.
+
+\bibitem{West:2002ud}
+West GB, Woodruff WH, Brown JH
+\newblock (2002) {Allometric scaling of metabolic rate from molecules and
+  mitochondria to cells and mammals}.
+\newblock \emph{Proc. Natl. Acad. Sci. USA} 99:2473--2478.
+
+\bibitem{Millar:1990p923}
+Millar J, Hickling G
+\newblock (1990) { Fasting Endurance and the Evolution of Mammalian Body Size}.
+\newblock \emph{Funct. Ecol.} 4:5--12.
+
+\bibitem{Alroy:1998p1594}
+Alroy J
+\newblock (1998) {Cope's rule and the dynamics of body mass evolution in North
+  American fossil mammals}.
+\newblock \emph{Science} 280:731.
+
+\bibitem{Clauset:2009fh}
+Clauset A, Redner S
+\newblock (2009) {Evolutionary Model of Species Body Mass Diversification}.
+\newblock \emph{Phys. Rev. Lett.} 102:038103.
+
+\bibitem{Smith:2010p3442}
+Smith F, Boyer A, Brown J, Costa D
+\newblock (2010) {The Evolution of Maximum Body Size of Terrestrial Mammals}.
+\newblock \emph{Science}.
+
+\bibitem{Saarinen:2014br}
+Saarinen JJ, {et~al.}
+\newblock (2014) {Patterns of maximum body size evolution in Cenozoic land
+  mammals: eco-evolutionary processes and abiotic forcing}.
+\newblock \emph{Proc Biol Sci} 281:20132049--20132049.
+
+\bibitem{allen2002}
+Allen AP, Brown JH, Gillooly JF
+\newblock (2002) {Global Biodiversity, Biochemical Kinetics, and the
+  Energetic-Equivalence Rule}.
+\newblock \emph{Science} 297:1545--1548.
+
+\bibitem{enquist1998}
+Enquist BJ, Brown JH, West GB
+\newblock (1998) {Allometric scaling of plant energetics and population density
+  : Abstract : Nature}.
+\newblock \emph{Nature} 395:163--165.
+
+\bibitem{tilman1981}
+Tilman D
+\newblock (1981) {Tests of Resource Competition Theory Using Four Species of
+  Lake Michigan Algae}.
+\newblock \emph{Ecology} 62:802--815.
+
+\bibitem{dutkiewicz2009}
+Dutkiewicz S, Follows MJ, Bragg JG
+\newblock (2009) {Modeling the coupling of ocean ecology and biogeochemistry}.
+\newblock \emph{Global Biogeochem. Cycles} 23:n/a--n/a.
+
+\bibitem{barton2010}
+Barton AD, Dutkiewicz S, Flierl G, Bragg J, Follows MJ
+\newblock (2010) {Patterns of Diversity in Marine Phytoplankton}.
+\newblock \emph{Science} 327:1509--1511.
+
+\end{thebibliography}
 

### 2017-02-09 Manuscript/Yield-Constants-Calculation-2.tex
new yield constants

--- 2016-08-24 Manuscript/Supplement.tex
@@
 
+\section*{Explicit Consumption Derivation}
+If I were to write out an explicit mass balance for our system in a dimensional system it would be:
+\begin{align} 
+\begin{split}
+\dot{F} &= \lambda\left(R\right) F + \rho\left(R\right)H - \sigma \left(1-\frac{R}{C}\right)F,  \\
+\dot{H} &= \sigma \left(1-\frac{R}{C}\right)F - \rho\left(R\right)H - \mu H,  \\
+\dot{R} &= \alpha R\left(1-\frac{R}{C}\right) -\left[\left(\frac{\rho\left(R\right)}{Y}+P\right)H+\left(\frac{\lambda\left(R\right)}{Y}+P\right)F\right]
+\end{split}
+\end{align}
+(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations), where $\lambda\left(R\right)$ and $\rho\left(R\right)$ are the growth and recovery functions dependent on resource availability.
+Typically these can be written as $\lambda\left(R\right)=\lambda_{max}S\left(R\right)$ where $\lambda_{max}$ is the maximum growth rate which changes with size, and $S\left(R\right)$ is a saturating function of resources, for example a Michealis-Menten or Monod function of the form $\frac{R}{k_{H}+R}$.
+In these equations $Y$ represents the quantity of resources required to build a unit of organism (e.g.
+gram of mammal produced per gram of grass consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
+If we pick $F$ and $H$ to have units of (g organisms $\cdot$ m$^{-2}$), then terms like $\frac{\rho\left(R\right)}{Y}H$ have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which is just the net primary productivity (NPP), a natural unit for $\dot{R}$, and this also gives us the units of $R$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is just the biomass density.
+In this system of units $\alpha$ is the growth rate of $R$.
+First, before diving into the constant values and a general nondimensionalization, let's look at the limits of the above equations in relationship to our two-state model: The saturation function should be defined by a linear regime proportional to $R$ when $R<<k$ and a constant value for $R>>k$.
+Thus for starving individuals, where $R<<k$, we have that $\rho\left(R\right)=\rho_{max}R/k$, and for the full class or organisms $\lambda\left(R\right)=\lambda_{max}$ such that the above relationships reduce to
+\begin{align} 
+\begin{split}
+\dot{F} &= \lambda_{max} F + \rho_{max}RH/k - \sigma \left(1-\frac{R}{C}\right)F,  \\
+\dot{H} &= \sigma \left(1-\frac{R}{C}\right)F - \rho_{max}R H/k - \mu H,  \\
+\dot{R} &= \alpha R\left(1-\frac{R}{C}\right) -\left[\left(\frac{\rho_{max}R}{Yk}+P\right)H+\beta F\right]
+\end{split}
+\end{align}
+where $\beta=\frac{\lambda_{max}}{Y}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$.
+
+Now consider the following nondimensionalization (ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),: $F^{*}=fF$, $H^{*}=fH$, $R^{*}=qR$, $t^{*}=st$ then we would have:
+\begin{align} 
+\begin{split}
+\dot{F}^{*} &= \frac{1}{s}\left[\lambda_{max} F^{*} + \rho_{max}\frac{R^{*}}{qk}H^{*} - \sigma \left(1-\frac{R^{*}}{qC}\right)F^{*}\right],  \\
+\dot{H}^{*} &= \frac{1}{s}\left[\sigma \left(1-\frac{R^{*}}{qC}\right)F^{*} - \rho_{max}\frac{R^{*}}{qk} H^{*} - \mu H^{*}\right],  \\
+\dot{R}^{*} &= \frac{1}{s}\left[\alpha R^{*}\left(1-\frac{R^{*}}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R^{*}}{Ykq}+P\right)H^{*}+\beta F^{*}\right]\right]
+\end{split}
+\end{align}
+Now there is a choice of factors that provides us with our original set of equations: if we pick $s=1$, $q=1/C$, and $f=1/Yk$, then we are left with 
+\begin{align} 
+\begin{split}
+\dot{F}^{*} &= \lambda_{max} F^{*} + \rho_{2}R^{*}H^{*} - \sigma \left(1-R^{*}\right)F^{*},  \\
+\dot{H}^{*} &= \sigma \left(1-R^{*}\right)F^{*} - \rho_{2} R^{*}H^{*} - \mu H^{*},  \\
+\dot{R}^{*} &= \alpha R^{*}\left(1-R^{*}\right) -\left(\rho_{2}R^{*}+P_{2}\right)H^{*}-\beta_{2} F^{*}
+\end{split}
+\end{align}
+where $\rho_{2}=C\rho_{max}/k$, $P_{2}=YkP/C$, and $\beta_{2}=Yk\beta/C$.
+Typically the value of $C/k$ should roughly be $2$.
+A summary of the parameters values is then: $\rho_{2}=$ 2 times our old $\rho$; $\lambda$, $\sigma$, and $\mu$ are all unchanged (the max notation can be dropped here and equated with the old values).
+Now $P=B_{0}m^{3/4}$, and $Y=18200$ (J g$^{-1}$) for grass \cite{estermann} which allows us to calculate $P_{2}=YB_{0}m^{3/4}/2$ and $\beta_{2}=Y/2$ (note again that $\beta=\frac{\lambda_{max}}{Y}+P$, and $\lambda_{max}$ is our old value for $\lambda$) .
+Finally, the value of $\alpha$ is given by the NPP divided by the corresponding biomass densities which I find to be $\alpha=2.10\times10^{-9}$ (s$^{-1}$) in a database that I have.
+
+%
+% term reduces to $\rho_{max}/k_{H}RH$ which is what we had before except that $\rho_{max}$ has an extra constant in it.
+For full individuals,  $R>>k$ and the $\lambda_{max}\frac{R}{k_{F}+R} F$ term becomes simply $\lambda_{max}F$ which differs from our previous version of the last term in the third equation.
+We are justified in taking both of these limits at the same time because of how we are binning and defining nutritional state (Note: we can simulate the full dynamics with the $\frac{R}{k+R}$ terms and see how different things are).
+
+%Now consider the following nondimensionalization (ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),: $F^{*}=fF$, $H^{*}=fH$, $R^{*}=qR$, $t^{*}=st$ then we would have:
+%\begin{align} 
+%\begin{split}
+%\frac{\partial F^{*}}{\partial t^{*}} &= \frac{1}{s}\left(\lambda_{max}\frac{R^{*}}{k_{F}q+R^{*}} F^{*} + \rho_{max}\frac{R^{*}}{k_{H}q+R^{*}}H^{*}\right),  \\
+%\frac{\partial H^{*}}{\partial t^{*}} &=  \frac{1}{s}\left(- \rho_{max}\frac{R^{*}}{k_{H}q+R^{*}}H^{*} - \mu H^{*}\right),  \\
+%\frac{\partial R^{*}}{\partial t^{*}} &=  \frac{1}{s}\left[\alpha R^{*}\left(1-\frac{R^{*}}{qC}\right) -\frac{q}{f}\left(\frac{\rho_{max}}{Y}\frac{R^{*}}{k_{H}q+R^{*}}H^{*}+\frac{\lambda_{max}}{Y}\frac{R^{*}}{k_{F}q+R^{*}}F^{*}\right)\right].
+%\end{split}
+%\end{align}
+
+%\begin{align} 
+%\begin{split}
+%\frac{\partial F^{*}}{\partial t^{*}} &=\lambda_{max}\frac{R^{*}}{k_{F}/C+R^{*}} F^{*} + \rho_{max}\frac{R^{*}}{k_{H}/C+R^{*}}H^{*},  \\
+%\frac{\partial H^{*}}{\partial t^{*}} &=  - \rho_{max}\frac{R^{*}}{k_{H}/C+R^{*}}H^{*} - \mu H^{*},  \\
+%\frac{\partial R^{*}}{\partial t^{*}} &=  \alpha R^{*}\left(1-R^{*}\right) -\left(\rho_{max}\frac{R^{*}}{k_{H}/C+R^{*}}H^{*}+\lambda_{max}\frac{R^{*}}{k_{F}/C+R^{*}}F^{*}\right).
+%\end{split}
+%\end{align}
+%and taking the above limits we have 
+%\begin{align} 
+%\begin{split}
+%\frac{\partial F^{*}}{\partial t^{*}} &=\lambda_{max}F^{*} + \frac{\rho_{max}C}{k_{H}}R^{*}H^{*},  \\
+%\frac{\partial H^{*}}{\partial t^{*}} &=  - \frac{\rho_{max}C}{k_{H}}R^{*}H^{*} - \mu H^{*},  \\
+%\frac{\partial R^{*}}{\partial t^{*}} &=  \alpha R^{*}\left(1-R^{*}\right) -\left(\frac{\rho_{max}C}{k_{H}}R^{*}H^{*}+\lambda_{max}F^{*}\right).
+%\end{split}
+%\end{align}
+%In summary, this is exactly our original system of equations (again with a $\lambda F$ instead of a $\lambda RF$ term in the last equation; and without dealing with the $\sigma (1-R)F$ term) and the only modification is that $\rho$ is adjusted by a factor of $C/k_{H}$ which should just be equal to $2$, and $F^{*}$, $H^{*}$, and $R^{*}$ are unites.
+
+%Now let us return to the third equation which is the last place where we need to define constants slightly more mechanistically. In the dimensional version we would have the following:
+%\begin{equation}
+%\label{Rprime}
+%\dot{R^{\prime}} = \alpha^{\prime} R^{\prime}\left(1-\frac{R^{\prime}}{C}\right) - \delta^{\prime} \left(\frac{\rho_{max}}{Y}\frac{R^{\prime}}{k_{H}^{\prime}+R^{\prime}}H+\frac{\lambda_{max}}{Y}\frac{R^{\prime}}{k_{F}^{\prime}+R^{\prime}}F\right).
+%\end{equation}
+%If we pick $R^{\prime}$ to have units of (g $\cdot$ m$^{-2}$) then $C$ is the maximum biomass density (g $\cdot$ m$^{-2}$), and $\alpha^{\prime}$ will have units of s$^{-1}$. Nondimensionalizing 
+%as $R=R^{\prime}/C$ then leads to $\alpha=\alpha^{\prime}C$ with units of (g $\cdot$ m$^{-2}$ s$^{-1}$) such that the $\alpha$ in our system of equations can simply be treated as the net primary productivity (NPP).
+The second term of Equation \ref{Rprime} is mostly unaffected by the nondimensionalization except that the $k\equiv k^{\prime}/C$.
+Now $\alpha$ in our model is just the externally provided NPP values ranging between $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) where we can take $3\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) as a typical value.
+
+%Now $\rho_{max}/Y$ and $\lambda_{max}/Y$ both have units of (g R $\cdot$ s$^{-1}$ $\cdot$ g Organism $^{-1}$) and thus the unit of $\delta^{\prime}$ must be (g organism $\cdot$ Individuals $^{-1}$ $\cdot$ m$^{-2}$)
+
 \section*{Parameter Values and Estimates}

### 2017-02-09 Manuscript/Yield-Constants-Calculation-3.tex
new yield constants

--- 2017-02-09 Manuscript/Yield-Constants-Calculation-2.tex
@@
 \dot{H} &= \sigma \left(1-\frac{R}{C}\right)F - \rho_{max}R H/k - \mu H,  \\
-\dot{R} &= \alpha R\left(1-\frac{R}{C}\right) -\left[\left(\frac{\rho_{max}R}{Yk}+P\right)H+\beta F\right]
+\dot{R} &= \alpha R\left(1-\frac{R}{C}\right) -\left[\left(\frac{\rho_{max}R}{Y_{H}k}+P\right)H+\beta F\right]
 \end{split}
 \end{align}
-where $\beta=\frac{\lambda_{max}}{Y}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$.
+where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$.
+A natural system of units for this system is that $R$ is a resource density (g m$^{-2}$) and $F$ and $H$ are the density of individuals (individuals m$^{-2}$).
+Given these definitions, $Y$ is the amount of resources required for an organism to grow to maturity (g resource individual$^{-1}$).
 
@@
 \dot{H}^{*} &= \frac{1}{s}\left[\sigma \left(1-\frac{R^{*}}{qC}\right)F^{*} - \rho_{max}\frac{R^{*}}{qk} H^{*} - \mu H^{*}\right],  \\
-\dot{R}^{*} &= \frac{1}{s}\left[\alpha R^{*}\left(1-\frac{R^{*}}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R^{*}}{Ykq}+P\right)H^{*}+\beta F^{*}\right]\right]
+\dot{R}^{*} &= \frac{1}{s}\left[\alpha R^{*}\left(1-\frac{R^{*}}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R^{*}}{Y_{H}kq}+P\right)H^{*}+\beta F^{*}\right]\right]
 \end{split}
 \end{align}
-Now there is a choice of factors that provides us with our original set of equations: if we pick $s=1$, $q=1/C$, and $f=1/Yk$, then we are left with 
+Now there is a choice of factors that provides us with our original set of equations: if we pick $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with 
 \begin{align} 
@@
 \end{align}
-where $\rho_{2}=C\rho_{max}/k$, $P_{2}=YkP/C$, and $\beta_{2}=Yk\beta/C$.
+where $\rho_{2}=C\rho_{max}/k$, $P_{2}=Y_{H}kP/C$, and $\beta_{2}=Y_{H}k\beta/C$.
 Typically the value of $C/k$ should roughly be $2$.
 A summary of the parameters values is then: $\rho_{2}=$ 2 times our old $\rho$; $\lambda$, $\sigma$, and $\mu$ are all unchanged (the max notation can be dropped here and equated with the old values).
-Now $P=B_{0}m^{3/4}$, and $Y=18200$ (J g$^{-1}$) for grass \cite{estermann} which allows us to calculate $P_{2}=YB_{0}m^{3/4}/2$ and $\beta_{2}=Y/2$ (note again that $\beta=\frac{\lambda_{max}}{Y}+P$, and $\lambda_{max}$ is our old value for $\lambda$) .
+Consider an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \cite{estermann} then the maintenance values is given by $P=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
+\begin{equation}
+B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+\end{equation}
+Similarly, the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\lambda}$ where
+\begin{equation}
+B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+\end{equation}
+which allows us to calculate $P_{2}=Y_{H}B_{0}M^{3/4}/2ME_{d}$ and $\beta_{2}=Y_{H}\beta/2$ (note again that $\beta=\frac{\lambda_{max}}{Y_{F}}+P$, and $\lambda_{max}$ is our old value for $\lambda$) .
 Finally, the value of $\alpha$ is given by the NPP divided by the corresponding biomass densities which I find to be $\alpha=2.10\times10^{-9}$ (s$^{-1}$) in a database that I have.

### 2017-02-09 Manuscript/Yield-Constants-Calculation.tex
new yield constants

--- 2016-08-24 Manuscript/Supplement.tex
@@
 
+\section*{Explicit Consumption Derivation}
+If I were to write out an explicit mass balance for our system in a dimensional system it would be:
+\begin{align} 
+\begin{split}
+\dot{F} &= \lambda_{max}\frac{R}{k_{F}+R} F + \rho_{max}\frac{R}{k_{H}+R}H - \sigma (1-R)F,  \\
+\dot{H} &= \sigma (1-R)F - \rho_{max}\frac{R}{k_{H}+R}H - \mu H,  \\
+\dot{R} &= \alpha R\left(1-\frac{R}{C}\right) -\left(\frac{\rho_{max}}{Y}\frac{R}{k_{H}+R}H+\frac{\lambda_{max}}{Y}\frac{R}{k_{F}+R}F\right).
+\end{split}
+\end{align}
+(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations), where the $\lambda_{max}$ and $\rho_{max}$ terms scale with size and represent the maximum rate under infinite resources, and the $k$ terms are define how fast rates saturate to the maximum.
+In these equations $Y$ represents the quantity of resources required to build a unit of organism (e.g.
+gram of mammal produced per gram of grass consumed) such that if we pick $F$ and $H$ to have units of (g organisms $\cdot$ m$^{-2}$), then terms like $\frac{\rho_{max}}{Y}\frac{R}{k_{H}+R}H$ have units of (g R $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$ $\cdot$ number of individuals) which is just the net primary productivity (NPP), a natural unit for $\dot{R}$, and this also gives us the units of $R$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is just the biomass density.
+In this system of units $\alpha$ is the growth rate of $R$.
+(Note: these equations ignore a constant maintenance rate which is easy to add in.) First, before diving into the constant values and a general nondimensionalization, let's look at the limits of the above equations in relationship to our two-state model: for starving individuals $R<<k$ and the $\rho_{max}\frac{R}{k_{H}+R}H$ term reduces to $\rho_{max}/k_{H}RH$ which is what we had before except that $\rho_{max}$ has an extra constant in it.
+For full individuals,  $R>>k$ and the $\lambda_{max}\frac{R}{k_{F}+R} F$ term becomes simply $\lambda_{max}F$ which differs from our previous version of the last term in the third equation.
+We are justified in taking both of these limits at the same time because of how we are binning and defining nutritional state (Note: we can simulate the full dynamics with the $\frac{R}{k+R}$ terms and see how different things are).
+
+Now consider the following nondimensionalization (ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),: $F^{*}=fF$, $H^{*}=fH$, $R^{*}=qR$, $t^{*}=st$ then we would have:
+\begin{align} 
+\begin{split}
+\frac{\partial F^{*}}{\partial t^{*}} &= \frac{1}{s}\left(\lambda_{max}\frac{R^{*}}{k_{F}q+R^{*}} F^{*} + \rho_{max}\frac{R^{*}}{k_{H}q+R^{*}}H^{*}\right),  \\
+\frac{\partial H^{*}}{\partial t^{*}} &=  \frac{1}{s}\left(- \rho_{max}\frac{R^{*}}{k_{H}q+R^{*}}H^{*} - \mu H^{*}\right),  \\
+\frac{\partial R^{*}}{\partial t^{*}} &=  \frac{1}{s}\left[\alpha R^{*}\left(1-\frac{R^{*}}{qC}\right) -\frac{q}{f}\left(\frac{\rho_{max}}{Y}\frac{R^{*}}{k_{H}q+R^{*}}H^{*}+\frac{\lambda_{max}}{Y}\frac{R^{*}}{k_{F}q+R^{*}}F^{*}\right)\right].
+\end{split}
+\end{align}
+Now there is a choice of factors that provides us with our original set of equations: if we pick $s=1$, $q=1/c$, and $f=q/Y$, then we are left with 
+\begin{align} 
+\begin{split}
+\frac{\partial F^{*}}{\partial t^{*}} &=\lambda_{max}\frac{R^{*}}{k_{F}/C+R^{*}} F^{*} + \rho_{max}\frac{R^{*}}{k_{H}/C+R^{*}}H^{*},  \\
+\frac{\partial H^{*}}{\partial t^{*}} &=  - \rho_{max}\frac{R^{*}}{k_{H}/C+R^{*}}H^{*} - \mu H^{*},  \\
+\frac{\partial R^{*}}{\partial t^{*}} &=  \alpha R^{*}\left(1-R^{*}\right) -\left(\rho_{max}\frac{R^{*}}{k_{H}/C+R^{*}}H^{*}+\lambda_{max}\frac{R^{*}}{k_{F}/C+R^{*}}F^{*}\right).
+\end{split}
+\end{align}
+and taking the above limits we have 
+\begin{align} 
+\begin{split}
+\frac{\partial F^{*}}{\partial t^{*}} &=\lambda_{max}F^{*} + \frac{\rho_{max}C}{k_{H}}R^{*}H^{*},  \\
+\frac{\partial H^{*}}{\partial t^{*}} &=  - \frac{\rho_{max}C}{k_{H}}R^{*}H^{*} - \mu H^{*},  \\
+\frac{\partial R^{*}}{\partial t^{*}} &=  \alpha R^{*}\left(1-R^{*}\right) -\left(\frac{\rho_{max}C}{k_{H}}R^{*}H^{*}+\lambda_{max}F^{*}\right).
+\end{split}
+\end{align}
+In summary, this is exactly our original system of equations (again with a $\lambda F$ instead of a $\lambda RF$ term in the last equation; and without dealing with the $\sigma (1-R)F$ term) and the only modification is that $\rho$ is adjusted by a factor of $C/k_{H}$ which should just be equal to $2$, and $F^{*}$, $H^{*}$, and $R^{*}$ are unites.
+
+%Now let us return to the third equation which is the last place where we need to define constants slightly more mechanistically. In the dimensional version we would have the following:
+%\begin{equation}
+%\label{Rprime}
+%\dot{R^{\prime}} = \alpha^{\prime} R^{\prime}\left(1-\frac{R^{\prime}}{C}\right) - \delta^{\prime} \left(\frac{\rho_{max}}{Y}\frac{R^{\prime}}{k_{H}^{\prime}+R^{\prime}}H+\frac{\lambda_{max}}{Y}\frac{R^{\prime}}{k_{F}^{\prime}+R^{\prime}}F\right).
+%\end{equation}
+%If we pick $R^{\prime}$ to have units of (g $\cdot$ m$^{-2}$) then $C$ is the maximum biomass density (g $\cdot$ m$^{-2}$), and $\alpha^{\prime}$ will have units of s$^{-1}$. Nondimensionalizing 
+%as $R=R^{\prime}/C$ then leads to $\alpha=\alpha^{\prime}C$ with units of (g $\cdot$ m$^{-2}$ s$^{-1}$) such that the $\alpha$ in our system of equations can simply be treated as the net primary productivity (NPP).
+The second term of Equation \ref{Rprime} is mostly unaffected by the nondimensionalization except that the $k\equiv k^{\prime}/C$.
+Now $\alpha$ in our model is just the externally provided NPP values ranging between $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) where we can take $3\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) as a typical value.
+
+%Now $\rho_{max}/Y$ and $\lambda_{max}/Y$ both have units of (g R $\cdot$ s$^{-1}$ $\cdot$ g Organism $^{-1}$) and thus the unit of $\delta^{\prime}$ must be (g organism $\cdot$ Individuals $^{-1}$ $\cdot$ m$^{-2}$)
+
 \section*{Parameter Values and Estimates}

### 2017-02-09 Manuscript/arxiv_version/YKR_arxiv.tex
new yield constants

--- 2016-08-24 Manuscript/12.tex
@@
-%\usepackage{scicite}
-%\usepackage{times}
-%\usepackage{graphicx}
-%\usepackage{verbatim}
-%\usepackage{subfigure}
-%\usepackage{tikz}
-%\usepackage{array}
-%\usepackage{tabularx}
-%\usepackage{multirow}
-%\usepackage{multicol}
-%\usepackage{multibox}
-%\usepackage{rotating}
-%\usepackage{layout}
-%\usepackage{epstopdf}
-%\usepackage{afterpage}
-%\usepackage[left]{lineno}
-
-%% PNAStmpl.tex
-%% Template file to use for PNAS articles prepared in LaTeX
-%% Version: Apr 14, 2008
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BASIC CLASS FILE
-
-\documentclass{pnastwo}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL GRAPHICS STYLE FILE
+\documentclass[11pt]{article}
+\usepackage{geometry}
+\geometry{margin=1in}
 
 \usepackage[pdftex]{graphicx}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL POSTSCRIPT FONT FILES
-
-\usepackage{pnastwof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% ADDITIONAL OPTIONAL STYLE FILES
 
@@
 %\usepackage[comma,sort&compress]{natbib}
+\usepackage{authblk}
+\usepackage{multicol}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL MACRO FILES
+\bibliographystyle{ieeetr}
 
-\bibliographystyle{pnas}
+%\linenumbers
+%\setlength\linenumbersep{3pt}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
+\author[1,2]{Justin D. Yeakel\thanks{To whom correspondence should be addressed: jdyeakel@gmail.com}}
+\author[2]{Christopher P. Kempes}
+\author[2,3]{Sidney Redner}
+\affil[1]{School of Natural Sciences, UC Merced, Merced, CA}
+\affil[2]{The Santa Fe Institute, Santa Fe, NM}
+\affil[3]{Department of Physics, Boston University, Boston, MA}
 
-\newcommand{\sid}[1]{\textcolor{red}{\bf [#1]}}
+%\author{Justin D. Yeakel${}^{1,2,*}$, Christopher P. Kempes${}^{2}$, \& Sidney Redner${}^{2,3}$ \\ \\
+%${}^1$School of Natural Science, University of California Merced, Merced, CA \\
+%${}^2$The Santa Fe Institute, Santa Fe, NM \\
+%${}^3$Department of Physics, Boston University, Boston MA \\
+%${}^*$To whom correspondence should be addressed: jdyeakel@gmail.com
+%}
 
-%% Don't type in anything in the following section:
-%%%%%%%%%%%%
-%% For PNAS Only:
-\contributor{Submitted to Proceedings
-of the National Academy of Sciences of the United States of America}
-\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
-\copyrightyear{2011}
-\issuedate{Issue Date}
-\volume{Volume}
-\issuenumber{Issue Number}
-%%%%%%%%%%%%
+\date{}
 
@@
 
-\linenumbers
-\setlength\linenumbersep{3pt}
-
-%\title{Ecological and evolutionary implications of starvation and body size}
-
-%\title{Survival of the Stable: Dynamic Consequences of Starvation and Reproduction}
-
-%\title{Size Constraints on Starvation \& Reproduction Dynamics}
-
-%\title{Eco-evolutionary Consequences of Starvation and Reproduction: Survival
-%of the Stable}
-
-\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
-
-\author
-{Justin D.
-Yeakel\affil{1}{School of Natural Science, University of California Merced, Merced, CA} \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{To whom correspondence should be addressed: jdyeakel@gmail.com}, Christopher P.
-Kempes \affil{2}{The Santa Fe Institute, Santa Fe, NM}, \and Sidney Redner \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{3}{Department of Physics, Boston University, Boston MA}
-}
-
-\contributor{Submitted to Proceedings of the National Academy of Sciences
-of the United States of America}
-
 \maketitle
 
-\begin{article}
-
-\begin{abstract} %250 words
+\noindent {\bf Abstract} \\ %250 words
 The eco-evolutionary dynamics of species is fundamentally linked to the energetic constraints of its constituent individuals.  
@@
 For body masses $\lesssim 10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $\gtrsim 10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
-\end{abstract}
-\keywords{foraging | starvation|reproduction}
 
-\noindent {\small {\bf Significance Statement} %105/120 words
-Energetic investment in somatic maintenance and growth vs. reproduction directly impacts the dynamics of populations among species.
-Here, we construct a Nutritional State-structured Model (NSM) to assess the population-level effects of starvation and recovery of a consumer population in a resource-limited environment, and use allometric scaling relationships for mammals to establish all timescales and rates.
-Our model reveals that mammalian energetic rates minimize the probability of stochastic extinction, establishes dynamic bounds on mammalian body size while providing independent theoretical support for the energy equivalence hypothesis, and provides a mechanistic driver for the evolutionary trend towards larger body size known as Cope's rule.}
+%\newpage
 
-\vspace{2mm}
+\begin{multicols}{2}
 
@@
 The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
-As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges. %(Fig.~\ref{fig:gvs})
+As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges.
 
@@
 
-%In the case of the 2-stage consumer model, the Jacobian evaluated at the internal steady state (denoted by $|_*$) is written
-%
-%\begin{equation}
-%\mathbf{J}|_* =
-%%\left(
-%%\begin{array}{ccc}
-%% -F^* m+(1-R^*) a -R^* a -H^* \rho  & -R^* \rho  & -m R^* \\
-%% -H^* \rho -F^* \sigma  & -\mu -R^* \rho  & (1-R^*) \sigma  \\
-%% H^* \rho +F^* \sigma  & R^* \rho  & \lambda -(1-R^*) \sigma  \\
-%%\end{array}
-%%\right).
-%\left(
-%\begin{array}{ccc}
-% -\frac{\lambda  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
-% \frac{\lambda  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\mu  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
-% -\frac{m \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } & -\frac{\mu  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } \\
-%\end{array}
-%\right).
-%\end{equation}
-
-%CHRIS - we need to define beta in this section...
-
 \noindent {\bf Role of allometry} \\ 
-%[Link Allometry stuff to our model - what does it provide to the story]
-%[Introduce biological and linked constraints]
-%[Allometries capture vast amounts of diversity via a single parameter --- body size]
-%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
 While there are no a priori constraints on the parameters in the NSM, most
@@
 \cite{DeLong:2010dy,Kempes:2012hy}.
-%Justin adding an explanation for beta
 An organism's metabolic rate $B$ is proportional to the cost of tissue
@@
 
-%CHRIS - MAYBE WE CUT THIS OUT SINCE WE STICK TO MAMMALS IN THIS PAPER
-%It should be noted that we have thus far used mammalian allometry to describe the size-based relations for growth, starvation, and death.
-However, our presentation is general, and other functional forms for $\epsilon$, for example, could be determined for other classes of organisms.
-Considering bacteria, we might expect that starvation or death is defined by the complete digestion of proteins, and in Table \ref{parameter-values} we provide all parameter values for bacteria which we later use as a comparison in our analysis.
-
-%\left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}
-%where $c=(a/b)^{1/(\eta-1)}$.
 Although the rate equations \eqref{eq:system} are general, here we focus on
@@
 \\
-%, though this is beyond the current scope of our analysis. %, and we henceforth focus our examination on mammalian species.
 
@@
 We thus find that the dynamics for all mammalian body sizes is confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
-%, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
 Moreover, for larger $M$, the distance to the Hopf bifurcation increases, while uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}) results in little qualitative difference in the distance to the the Hopf bifurcation.
 These results suggest that small mammals are more prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.  
-Thus our NSM model predicts that population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting.%, and
-%2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
+Thus our NSM model predicts that population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting.
 
@@
 However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
-%It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic~\cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
 Statistically significant support for the existence of population cycles among mammals is predominantly based on time series for small mammals~\cite{Kendall:1998hl}, in agreement with our predictions of more pronounced transient dynamics, given how close these points are to the Hopf bifurcation.
@@
 \vspace{6mm}
-%, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
- %, emphasizing the role of starvation on dynamics.
 
-%Extinction to the left and the right
 \noindent {\bf Extinction risk}\\
@@
 This procedure is repeated for 1000 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an organism of $M=100$ grams.
-%the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
 In each replicate the initial densities are chosen to be $A(F^*,H^*,R^*)$, with $A$ a random variable that is uniformly distributed in $[0,2]$.  
 By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-6}$ to $10^{-1}$. 
-%, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
 As expected, higher rates of extinction correlate with both high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if $\sigma$ is small.  
@@
 This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
-% OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
 To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
-
-%Results are insensitive to alpha but possibly sensitive to beta (eff parameter)
-
-%NOTES 7/8/16
-%Calder has a hypothesis that period ~ Mass^(1/4).
-%Can we get "transient period" from Jacobian?
-%Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.
 
@@
 
-%Minimal mass bound
-%A theoretical lower bound on mammalian body size occurs when $\epsilon=0$ where mammals have no metabolic reserves and immediately starve; this occurs at a size of $M=0$, which is obviously infeasible.
-%Such prediction gives an extreme limit on size but does not account for the subtleties of starvation dynamics
-%that may limit body size (among other factors). 
 The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
@@
 Our model shows that the equivalence breaks down at both the minimum and maximum observed body sizes for mammals, suggesting that these are hard limits where deviations outside of this range are energetically suboptimal.
-%, violating the energy equivalence.
-% set by deviations from constant efficiency obeyed by other populations.
 Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end.
-%Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is close to the minimum observed mammalian body size of ca.\ 1.3-2.5g, a range that occurs as the recovery rate begins its decline.
-%In addition to known transport limitations~\cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.  
-%This result mirrors other efforts \cite{Kempes:2012hy,Kempes:2016} where at a given scale multiple limitations constrain the smallest possibilities for life within a class of organisms.
-
-%The combined steady-state dynamics and allometric timescales of the NSM predict that larger mammals should outcompete smaller ones, and this suggests that the model may provide a framework with which to explore the energetic mechanisms behind Cope's rule.
-%, which we explore further below.
 
@@
 However, for $M>8.43\times 10^6$, leaner individuals ($\chi < 0$) have advantageous steady state densities.
-%, and this is due to the changing
-%covariance between energetic rates as a function of modified energetic
-%reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
 
 The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}=8.43 \times 10^6$ thus serves as an attractor, or an uninvadible evolutionary stable state, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
-%This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
 Moreover, $M_{\rm opt}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
 While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we suggest that the dynamics of starvation and recovery described in the NSM may provide a general driving mechanism for the evolution of larger body size among terrestrial mammals.
-%and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
-
-%One might be concerned a greater number of large mammals are currently not observed in the modern world given that larger mammals are less susceptible to extinction.  
-%However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also much more geographically widespread than today.
-%These results, combined with our findings, suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
 
@@
 
-%In accordance with the fasting endurance hypothesis, 
-%
-%
-%Such energetic advantages, coupled with the 
-%Ove and tends to be selected for over evolutionary time.
-%The NSM is in accordance with the XXX, where species with larger masses are less likely to evince cyclic dynamics.
-%
-%
-%
-%
-%We offer an additional mechanism for the minimum and maximum limitation on mammalian body size operating on ecological timescales.
-%Such a mechanism would 
-%
-%
-%
-%
-%Although these physiological barriers are indeed of primary importance, our minimal model of starvation dynamics suggests that there may exist additional constraints due to dynamic and evolutionary effects at small and large body sizes, respectively.
-%
-%
-%
+\vspace{2mm}
 
-%Looking Forward
-
-%
-%
-%%Starving Random Walk
-%In our modeling, foragers look for food by wandering in this changing
-%environment.  If such a search is successful, the forager is satiated and it
-%can engage in the essential activity of reproduction.  However, if the search
-%is unsuccessful for a sufficiently long period, the forager ``starves''.
-%Such a forager can do nothing else but forage, until it either finds food and
-%again becomes satiated or it dies when it goes too long without finding
-%nourishment.  These rules are reminiscent of the ``starving random walk''
-%model, where a single random walk can take $\mathcal{S}$ steps without
-%encountering food before starving to death.  Moreover, the resource does not
-%regenerate, so that the forager ultimately starves to death.  For this
-%idealized model, it was found that the average lifetime of the forager scales
-%algebraically with $\mathcal{S}$ in $d\!=\!1$ and $d\!=\!2$ dimensions, and
-%as $\exp(-A\mathcal{S}^\omega)$ for $d>2$.  Here the exponent
-%$\omega\approx \frac{1}{2}$ for $d=3$, while $\omega\to 1$ only as
-%$d\to \infty$, with the latter behavior corresponding to the mean-field
-%limit.  As we will discuss, regeneration of the resource, together with the
-%behavioral change between starving and satiated foragers leads to still much
-%richer dynamical behavior.
-
-%The reproductive tradeoff as a strategy
-%
-
-%
-%
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
-% 	\caption{
-% 	Probability that the resource value is less than threshold = 0.05
-% 	}
-% 	\label{SN}
-% \end{figure}
-%
-%
-%
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
-% 	\caption{
-% 	Hopf bifurcation
-% 	}
-% 	\label{SN}
-% \end{figure}
-
-\bibliography{aa_starving3}
-
-\begin{acknowledgments}
+\noindent {\bf Acknowledgments}
+{\small
   We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
@@
   Institute.
-\end{acknowledgments}
+}
 
-%\onecolumn
-%\let\cleardoublepage\clearpage
+%\bibliography{aa_starving3}
+\small{
+\begin{thebibliography}{10}
 
-%FIGURE 1
+\bibitem{Martin:1987dl}
+T.~E. Martin, ``{Food as a Limit on Breeding Birds: A Life-History
+  Perspective},'' {\em Annu. Rev. Ecol. Syst.}, vol.~18, pp.~453--487, Jan.
+  1987.
 
-\end{article}
+\bibitem{Kirk:1997cc}
+K.~L. Kirk, ``{Life-History Responses to Variable Environments: Starvation and
+  Reproduction in Planktonic Rotifers},'' {\em Ecology}, vol.~78, pp.~434--441,
+  Mar. 1997.
+
+\bibitem{Kempes:2012hy}
+C.~P. Kempes, S.~Dutkiewicz, and M.~J. Follows, ``{Growth, metabolic
+  partitioning, and the size of microorganisms.},'' {\em PNAS}, vol.~109,
+  pp.~495--500, Jan. 2012.
+
+\bibitem{Mangel:1988uaa}
+M.~Mangel and C.~W. Clark, {\em {Dynamic Modeling in Behavioral Ecology}}.
+\newblock Princeton: Princeton University Press, 1988.
+
+\bibitem{Mangel:2014kz}
+M.~Mangel, ``{Stochastic dynamic programming illuminates the link between
+  environment, physiology, and evolution},'' {\em B. Math. Biol.}, vol.~77,
+  no.~5, pp.~857--877, 2014.
+
+\bibitem{Yeakel:2013hi}
+J.~D. Yeakel, N.~J. Dominy, P.~L. Koch, and M.~Mangel, ``{Functional
+  morphology, stable isotopes, and human evolution: a model of consilience},''
+  {\em Evolution}, vol.~68, no.~1, pp.~190--203, 2014.
+
+\bibitem{Morris:1987eo}
+D.~W. Morris, ``{Optimal Allocation of Parental Investment},'' {\em Oikos},
+  vol.~49, p.~332, July 1987.
+
+\bibitem{Tveraa:2003fq}
+T.~Tveraa, P.~Fauchald, C.~Henaug, and N.~G. Yoccoz, ``{An examination of a
+  compensatory relationship between food limitation and predation in
+  semi-domestic reindeer},'' {\em Oecologia}, vol.~137, pp.~370--376, Nov.
+  2003.
+
+\bibitem{Daan:1988va}
+S.~Daan, C.~Dijkstra, R.~Drent, and T.~Meijer, ``{Food supply and the annual
+  timing of avian reproduction},'' in {\em Proceedings of the International
+  {\ldots}}, 1988.
+
+\bibitem{Jacot:2009dv}
+A.~Jacot, M.~Valcu, K.~van Oers, and B.~Kempenaers, ``{Experimental nest site
+  limitation affects reproductive strategies and parental investment in a
+  hole-nesting passerine},'' {\em Animal Behaviour}, vol.~77, pp.~1075--1083,
+  May 2009.
+
+\bibitem{Stearns:1989ip}
+S.~C. Stearns, ``{Trade-Offs in Life-History Evolution},'' {\em Funct. Ecol.},
+  vol.~3, no.~3, p.~259, 1989.
+
+\bibitem{Barboza:2002in}
+P.~Barboza and D.~Jorde, ``{Intermittent fasting during winter and spring
+  affects body composition and reproduction of a migratory duck},'' {\em J Comp
+  Physiol B}, vol.~172, pp.~419--434, July 2002.
+
+\bibitem{Threlkeld:1976ih}
+S.~T. Threlkeld, ``{Starvation and the size structure of zooplankton
+  communities},'' {\em Freshwater Biol.}, vol.~6, pp.~489--496, Dec. 1976.
+
+\bibitem{Weber:1998jg}
+T.~P. Weber, B.~J. Ens, and A.~I. Houston, ``{Optimal avian migration: A
+  dynamic model of fuel stores and site use},'' {\em Evolutionary Ecology},
+  vol.~12, pp.~377--401, May 1998.
+
+\bibitem{Mduma:1999cp}
+S.~A.~R. Mduma, A.~R.~E. Sinclair, and R.~Hilborn, ``{Food regulates the
+  Serengeti wildebeest: a 40-year record},'' {\em J. Anim. Ecol.}, vol.~68,
+  pp.~1101--1122, Nov. 1999.
+
+\bibitem{Moore:2014hi}
+J.~W. Moore, J.~D. Yeakel, D.~Peard, J.~Lough, and M.~Beere, ``{Life-history
+  diversity and its importance to population stability and persistence of a
+  migratory fish: steelhead in two large North American watersheds},'' {\em J.
+  Anim. Ecol.}, vol.~83, no.~5, pp.~1035--1046, 2014.
+
+\bibitem{Mead:1989dt}
+R.~A. Mead, ``{The Physiology and Evolution of Delayed Implantation in
+  Carnivores},'' in {\em Carnivore Behavior, Ecology, and Evolution},
+  pp.~437--464, Boston, MA: Springer US, 1989.
+
+\bibitem{Sandell:1990kw}
+M.~Sandell, ``{The Evolution of Seasonal Delayed Implantation},'' {\em The
+  Quarterly Review of Biology}, vol.~65, no.~1, pp.~23--42, 1990.
+
+\bibitem{Bulik:1999eo}
+C.~M. Bulik, P.~F. Sullivan, J.~L. Fear, A.~Pickering, A.~Dawn, and
+  M.~McCullin, ``{Fertility and Reproduction in Women With Anorexia Nervosa},''
+  {\em J. Clin. Psychiatry}, vol.~60, pp.~130--135, Feb. 1999.
+
+\bibitem{Trites:2003cc}
+A.~W. Trites and C.~P. Donnelly, ``{The decline of Steller sea lions Eumetopias
+  jubatus in Alaska: a review of the nutritional stress hypothesis},'' {\em
+  Mammal Review}, vol.~33, pp.~3--28, Mar. 2003.
+
+\bibitem{Glazier:2009hq}
+D.~S. Glazier, ``{Metabolic level and size scaling of rates of respiration and
+  growth in unicellular organisms},'' {\em Funct. Ecol.}, vol.~23,
+  pp.~963--968, Oct. 2009.
+
+\bibitem{Kooi2000}
+S.~A. L.~M. Kooijman, {\em {Dynamic Energy and Mass Budgets in Biological
+  Systems}}.
+\newblock 2000.
+
+\bibitem{Sousa:2010ez}
+T.~Sousa, T.~Domingos, J.~C. Poggiale, and S.~A. L.~M. Kooijman, ``{Dynamic
+  energy budget theory restores coherence in biology},'' {\em Philos. T. Roy.
+  Soc. B}, vol.~365, pp.~3413--3428, Oct. 2010.
+
+\bibitem{Diekmann:2010da}
+O.~Diekmann and J.~A.~J. Metz, ``{How to lift a model for individual behaviour
+  to the population level?},'' {\em Philos. T. Roy. Soc. B}, vol.~365,
+  pp.~3523--3530, Nov. 2010.
+
+\bibitem{murdoch:2003}
+W.~W. Murdoch, C.~J. Briggs, and R.~M. Nisbet, {\em {Consumer-resource
+  Dynamics}}.
+\newblock Monographs in population biology, Princeton University Press, 2003.
+
+\bibitem{Benichou:2014wu}
+O.~Benichou and S.~Redner, ``{Depletion-Controlled Starvation of a Diffusing
+  Forager},'' {\em arXiv}, May 2014.
+
+\bibitem{Benichou:2016wl}
+O.~B{\'e}nichou, M.~Chupeau, and S.~Redner, ``{Role of Depletion on the
+  Dynamics of a Diffusing Forager},'' May 2016.
+
+\bibitem{Chupeau:2016jf}
+M.~Chupeau, O.~B{\'e}nichou, and S.~Redner, ``{Universality classes of foraging
+  with resource renewal},'' {\em Phys. Rev. E}, vol.~93, p.~032403, Mar. 2016.
+
+\bibitem{murray2011mathematical}
+J.~D. Murray, {\em {Mathematical Biology: I. An Introduction}}, vol.~110 of
+  {\em Interdisciplinary Applied Mathematics}.
+\newblock Melaka Manipal Medical College (Manipal Campus), International Centre
+  for Health Sciences, Madhav Nagar, Manipal, Udupi District, Karnataka State,
+  India. nayaksathish@yahoo.com: Springer New York, 2011.
+
+\bibitem{Strogatz:2008wo}
+S.~H. Strogatz, {\em {Nonlinear Dynamics and Chaos}}.
+\newblock With Applications to Physics, Biology, Chemistry, and Engineering,
+  Westview Press, Aug. 2008.
+
+\bibitem{GuckHolmes}
+J.~Guckenheimer and P.~Holmes, {\em {Nonlinear oscillations, dynamical systems,
+  and bifurcations of vector fields}}.
+\newblock New York: Springer, 1983.
+
+\bibitem{Gross:2004p2428}
+T.~Gross and U.~Feudel, ``{Analytical search for bifurcation surfaces in
+  parameter space},'' {\em Physica D}, vol.~195, no.~3-4, pp.~292--302, 2004.
+
+\bibitem{Hastings:2001jh}
+A.~Hastings, ``{Transient dynamics and persistence of ecological systems},''
+  {\em Ecol. Lett.}, vol.~4, pp.~215--220, May 2001.
+
+\bibitem{Neubert:1997wk}
+M.~Neubert and H.~Caswell, ``{Alternatives to resilience for measuring the
+  responses of ecological systems to perturbations},'' {\em Ecology}, vol.~78,
+  no.~3, pp.~653--665, 1997.
+
+\bibitem{Caswell:2005eo}
+H.~Caswell and M.~G. Neubert, ``{Reactivity and transient dynamics of
+  discrete-time ecological systems},'' {\em Journal of Difference Equations and
+  Applications}, vol.~11, pp.~295--310, Apr. 2005.
+
+\bibitem{Neubert:2009td}
+M.~Neubert and H.~Caswell, ``{Detecting reactivity},'' {\em Ecology}, 2009.
+
+\bibitem{Yodzis:1992hg}
+P.~Yodzis and S.~Innes, ``{Body Size and Consumer-Resource Dynamics},'' {\em
+  Am. Nat.}, vol.~139, no.~6, pp.~1151--1175, 1992.
+
+\bibitem{Brown:2004wq}
+J.~Brown, J.~Gillooly, A.~Allen, V.~Savage, and G.~West, ``{Toward a metabolic
+  theory of ecology},'' {\em Ecology}, vol.~85, no.~7, pp.~1771--1789, 2004.
+
+\bibitem{West:2002it}
+G.~B. West, W.~H. Woodruff, and J.~H. Brown, ``{Allometric scaling of metabolic
+  rate from molecules and mitochondria to cells and mammals.},'' {\em Proc.
+  Natl. Acad. Sci. USA}, vol.~99 Suppl 1, pp.~2473--2478, Feb. 2002.
+
+\bibitem{DeLong:2010dy}
+J.~P. DeLong, J.~G. Okie, M.~E. Moses, R.~M. Sibly, and J.~H. Brown, ``{Shifts
+  in metabolic scaling, production, and efficiency across major evolutionary
+  transitions of life.},'' {\em PNAS}, vol.~107, pp.~12941--12945, July 2010.
+
+\bibitem{West:2001bv}
+G.~B. West, J.~H. Brown, and B.~J. Enquist, ``{A general model for ontogenetic
+  growth},'' {\em Nature}, vol.~413, pp.~628--631, Oct. 2001.
+
+\bibitem{moses2008rmo}
+M.~E. Moses, C.~Hou, W.~H. Woodruff, G.~B. West, J.~C. Nekola, W.~Zuo, and
+  J.~H. Brown, ``{Revisiting a Model of Ontogenetic Growth: Estimating Model
+  Parameters from Theory and Data},'' {\em
+  http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}, vol.~171, pp.~632--645,
+  May 2008.
+
+\bibitem{gillooly2002esa}
+J.~F. Gillooly, E.~L. Charnov, G.~B. West, V.~M. Savage, and J.~H. Brown,
+  ``{Effects of size and temperature on developmental time},'' {\em Nature},
+  vol.~417, pp.~70--73, May 2002.
+
+\bibitem{hou}
+C.~Hou, W.~Zuo, M.~E. Moses, W.~H. Woodruff, J.~H. Brown, and G.~B. West,
+  ``{Energy Uptake and Allocation During Ontogeny},'' {\em Science}, vol.~322,
+  pp.~736--739, Oct. 2008.
+
+\bibitem{bettencourt}
+L.~M.~A. Bettencourt, J.~Lobo, D.~Helbing, C.~Kuhnert, and G.~B. West,
+  ``{Growth, innovation, scaling, and the pace of life in cities},'' {\em Proc.
+  Natl. Acad. Sci. USA}, vol.~104, pp.~7301--7306, Apr. 2007.
+
+\bibitem{Folland:2008ij}
+J.~P. Folland, T.~M. Mc~Cauley, and A.~G. Williams, ``{Allometric scaling of
+  strength measurements to body size},'' {\em Eur J Appl Physiol}, vol.~102,
+  pp.~739--745, Jan. 2008.
+
+\bibitem{CalderIII:1983jd}
+W.~A. Calder~III, ``{An allometric approach to population cycles of mammals},''
+  {\em J. Theor. Biol.}, vol.~100, pp.~275--282, Jan. 1983.
+
+\bibitem{Peterson:1984hj}
+R.~O. Peterson, R.~E. PAGE, and K.~M. DODGE, ``{Wolves, Moose, and the
+  Allometry of Population Cycles},'' {\em Science}, vol.~224, pp.~1350--1352,
+  June 1984.
+
+\bibitem{Krukonis:1991fk}
+G.~Krukonis and W.~M. Schaffer, ``{Population cycles in mammals and birds: Does
+  periodicity scale with body size?},'' {\em J. Theor. Biol.}, vol.~148,
+  pp.~469--493, Feb. 1991.
+
+\bibitem{Hendriks:2012fc}
+A.~J. Hendriks and C.~Mulder, ``{Delayed logistic and
+  Rosenzweig{\textendash}MacArthur models with allometric parameter setting
+  estimate population cycles at lower trophic levels well},'' {\em Ecological
+  Complexity}, vol.~9, pp.~43--54, Feb. 2012.
+
+\bibitem{Kendall:1999iy}
+B.~E. Kendall, C.~J. Briggs, W.~W. Murdoch, P.~Turchin, S.~P. Ellner,
+  E.~McCauley, R.~M. Nisbet, and S.~N. Wood, ``{Why do populations cycle? A
+  synthesis of statistical and mechanistic modeling approaches},'' {\em
+  Ecology}, vol.~80, pp.~1789--1805, Sept. 1999.
+
+\bibitem{Hogstedt:2005cr}
+G.~H{\"o}gstedt, T.~Seldal, and A.~Breistol, ``{Period length in cyclic animal
+  populations},'' {\em Ecology}, vol.~86, pp.~373--378, Feb. 2005.
+
+\bibitem{Kendall:1998hl}
+Kendall, Prendergast, and Bjornstad, ``{The macroecology of population
+  dynamics: taxonomic and biogeographic patterns in population cycles},'' {\em
+  Ecol. Lett.}, vol.~1, pp.~160--164, Nov. 1998.
+
+\bibitem{Brown:1993p708}
+J.~Brown, P.~Marquet, and M.~Taper, ``{Evolution of body size: consequences of
+  an energetic definition of fitness},'' {\em Am. Nat.}, vol.~142, no.~4,
+  pp.~573--584, 1993.
+
+\bibitem{West:1997cg}
+G.~B. West, J.~H. Brown, and B.~J. Enquist, ``{A General Model for the Origin
+  of Allometric Scaling Laws in Biology},'' {\em Science}, vol.~276,
+  pp.~122--126, Apr. 1997.
+
+\bibitem{West:2002ud}
+G.~B. West, W.~H. Woodruff, and J.~H. Brown, ``{Allometric scaling of metabolic
+  rate from molecules and mitochondria to cells and mammals},'' {\em Proc.
+  Natl. Acad. Sci. USA}, vol.~99, no.~Suppl 1, pp.~2473--2478, 2002.
+
+\bibitem{Millar:1990p923}
+J.~Millar and G.~Hickling, ``{ Fasting Endurance and the Evolution of Mammalian
+  Body Size},'' {\em Funct. Ecol.}, vol.~4, no.~1, pp.~5--12, 1990.
+
+\bibitem{Alroy:1998p1594}
+J.~Alroy, ``{Cope's rule and the dynamics of body mass evolution in North
+  American fossil mammals},'' {\em Science}, vol.~280, no.~5364, p.~731, 1998.
+
+\bibitem{Clauset:2009fh}
+A.~Clauset and S.~Redner, ``{Evolutionary Model of Species Body Mass
+  Diversification},'' {\em Phys. Rev. Lett.}, vol.~102, p.~038103, Jan. 2009.
+
+\bibitem{Smith:2010p3442}
+F.~Smith, A.~Boyer, J.~Brown, and D.~Costa, ``{The Evolution of Maximum Body
+  Size of Terrestrial Mammals},'' {\em Science}, Jan. 2010.
+
+\bibitem{Saarinen:2014br}
+J.~J. Saarinen, A.~G. Boyer, J.~H. Brown, D.~P. Costa, S.~K.~M. Ernest, A.~R.
+  Evans, M.~Fortelius, J.~L. Gittleman, M.~J. Hamilton, L.~E. Harding,
+  K.~Lintulaakso, S.~K. Lyons, J.~G. Okie, R.~M. Sibly, P.~R. Stephens,
+  J.~Theodor, M.~D. Uhen, and F.~A. Smith, ``{Patterns of maximum body size
+  evolution in Cenozoic land mammals: eco-evolutionary processes and abiotic
+  forcing},'' {\em Proc Biol Sci}, vol.~281, pp.~20132049--20132049, June 2014.
+
+\bibitem{allen2002}
+A.~P. Allen, J.~H. Brown, and J.~F. Gillooly, ``{Global Biodiversity,
+  Biochemical Kinetics, and the Energetic-Equivalence Rule},'' {\em Science},
+  vol.~297, pp.~1545--1548, Aug. 2002.
+
+\bibitem{enquist1998}
+B.~J. Enquist, J.~H. Brown, and G.~B. West, ``{Allometric scaling of plant
+  energetics and population density : Abstract : Nature},'' {\em Nature},
+  vol.~395, pp.~163--165, Sept. 1998.
+
+\bibitem{tilman1981}
+D.~Tilman, ``{Tests of Resource Competition Theory Using Four Species of Lake
+  Michigan Algae},'' {\em Ecology}, vol.~62, pp.~802--815, June 1981.
+
+\bibitem{dutkiewicz2009}
+S.~Dutkiewicz, M.~J. Follows, and J.~G. Bragg, ``{Modeling the coupling of
+  ocean ecology and biogeochemistry},'' {\em Global Biogeochem. Cycles},
+  vol.~23, pp.~n/a--n/a, Dec. 2009.
+
+\bibitem{barton2010}
+A.~D. Barton, S.~Dutkiewicz, G.~Flierl, J.~Bragg, and M.~J. Follows,
+  ``{Patterns of Diversity in Marine Phytoplankton},'' {\em Science}, vol.~327,
+  pp.~1509--1511, Mar. 2010.
+
+\end{thebibliography}
+}
+
+\end{multicols}
+
+\clearpage
 
@@
 \centering
-\includegraphics[width=0.45\textwidth]{fig_FPAllometric.pdf}
+\includegraphics[width=0.4\textwidth]{fig_FPAllometric.pdf}
 \caption{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
@@
 \centering
-\includegraphics[width=0.75\textwidth]{fig_Invasion.pdf}
+\includegraphics[width=0.65\textwidth]{fig_Invasion.pdf}
 \caption{ Invasion feasibility for organisms with a proportional change in
@@
 
-% 
-% 
-%	 \begin{figure}[h]
-%		 \centering
-%		 \includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-%		 \caption{
-%		 }
-%		 \label{growth-diagram}
-%	 \end{figure}
-%	 
-%	 \begin{figure}[h]
-%		 \centering
-%		 \includegraphics[width=0.25\textwidth]{fig_GvS.pdf}
-%		 \caption{
-%		 }
-%		 \label{GvS}
-%	 \end{figure}
-%
-%
-%	\begin{figure}[h]
-%	 	\centering
-%	 	\includegraphics[width=0.25\textwidth]{fig_FixedPoint.pdf}
-%	 	\caption{
-%	 	}
-%		\label{Hopfb}
-%	\end{figure}
-%
-%
-%	\begin{figure}[h]
-%		\centering
-%		\includegraphics[width=0.8\textwidth]{fig_DataHopf.pdf}
-%		\caption{
-%		}
-%		\label{DataHopf}
-%	\end{figure}
-%	 
-% 
-%	\begin{figure}[h]
-%		\centering
-%	 	\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometric.pdf}
-%	 	\caption{
-%	 	}
-%	 	\label{Ext}
-%	\end{figure}
-%	
-%	\begin{figure}[h]
-%	 	\centering
-%	 	\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
-%	 	\caption{
-%	 	}
-%	 	\label{Asymp}
-%	\end{figure}
-%
-%\clearpage
-
-%\begin{table}[h]
-%\caption{Parameter Values For Various Classes of Organisms}
-%\label{param}
-%    \begin{center}
-%    \small
-%     \begin{tabular}{| p{1.2cm}| p{3.2cm} | p{2.6cm} | p{3.2cm} | }
-%     \hline
-%     & {\bf Mammals} & {\bf Unicellular Eukaryotes} & {\bf Bacteria} \\
-%     \hline
-%   $\eta$ & $3/4$ & & $1.70$ \\ 
-%   $E_{m}$ & $10695$ (J gram$^{-1}$) & & $10695$ (J gram$^{-1}$) \\ 
-%   $E_{m}^{\prime}$ & $\approx E_{m}$ & & $\approx E_{m}$ \\ 
-%   $B_{0}$ & $0.019$ (W gram$^{-\alpha}$) & & $1.96\times10^{17}$ \\
-%   $B_{m}$ & $0.025$ (W gram$^{-1}$)   & & $0.025$ (W gram$^{-1}$)\\
-%   $a$ & $1.78\times10^{-6}$ & & $1.83\times10^{13}$ \\ 
-%   $b$ & $2.29\times10^{-6}$ & & $2.29\times10^{-6}$ \\  
-%   $\eta-1$ & $-0.21$ & & $0.73$ \\ 
-%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) & & $56493$ \\ 
-%   $\gamma$ & $1.19$ & & $0.68$ \\ 
-%   $f_{0}$ & $0.02$ & & $1.30\times10^{-5}$ \\ 
-%   $\zeta$ & $1.01$ & & \\ 
-%   $mm_{0}$ & $0.32$ & & \\ 
-%   
-%      
-%   \hline
-%    \end{tabular}
-%    \end{center}
-%   \end{table}
-
 \end{document}

### 2017-02-09 Manuscript/arxiv_version/tmp.tex
new yield constants

--- 2017-02-09 Manuscript/arxiv_version/YKR_arxiv.tex
@@
-\documentclass[11pt]{article}
-\usepackage{geometry}
-\geometry{margin=1in}
+\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
+%\usepackage[pdftex]{graphicx}
 
-\usepackage[pdftex]{graphicx}
+\usepackage{amsmath,amsfonts,amssymb}
+\usepackage[english]{babel} 
+\usepackage[latin1]{inputenc} 
+\usepackage[T1]{fontenc}
+\usepackage{color}
+\usepackage{float}
+\usepackage{verbatim}
+\usepackage{graphicx}
+\usepackage{bm}
+\usepackage{mathtools}
+\usepackage{stmaryrd} 
+\usepackage{anyfontsize}
 
-%\usepackage{amssymb,amsfonts,amsmath}
-\usepackage{amsmath}
-\usepackage{amssymb}
-
-\usepackage[font=large]{caption}
-\usepackage{subfigure}
-\usepackage{epstopdf}
-\usepackage{graphicx}
-\usepackage{verbatim}
-\usepackage{array}
-\usepackage{tabularx}
-\usepackage{multirow}
+%\usepackage{epstopdf}
+%\usepackage{array}
+%\usepackage{tabularx}
+%\usepackage{multirow}
 \usepackage{color}
 %\usepackage{multibox}
-\usepackage{rotating}
-\usepackage{color}
+%\usepackage{rotating}
 %\usepackage{lineno}
-\usepackage[left]{lineno}
+%\usepackage[left]{lineno}
 %\usepackage[comma,sort&compress]{natbib}
-\usepackage{authblk}
-\usepackage{multicol}
+%\usepackage{authblk}
+%\usepackage{multicol}
 
-\bibliographystyle{ieeetr}
+%\bibliographystyle{ieeetr}
 
@@
 
+\begin{document}
+
+\author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
+  of California, Merced, Merced, CA 95340, USA}
+
+\author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
+  Park Road, Santa Fe, NM 87501, USA}
+
+\author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
+  Road, Santa Fe, NM 87501, USA}
+
 \title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
-\author[1,2]{Justin D. Yeakel\thanks{To whom correspondence should be addressed: jdyeakel@gmail.com}}
-\author[2]{Christopher P. Kempes}
-\author[2,3]{Sidney Redner}
-\affil[1]{School of Natural Sciences, UC Merced, Merced, CA}
-\affil[2]{The Santa Fe Institute, Santa Fe, NM}
-\affil[3]{Department of Physics, Boston University, Boston, MA}
 
@@
 
-\date{}
+\begin{abstract}
 
-\begin{document}
+The eco-evolutionary dynamics of species is fundamentally linked to the
+energetic constraints of its constituent individuals.  Of particular
+importance are the tradeoffs between reproduction and the dynamics of
+starvation and recovery in resource-limited environments.  To elucidate the
+consequences of this tradeoff, we introduce a minimal nutritional
+state-structured model that incorporates two classes of consumer:
+nutritionally replete consumers that reproduce, and undernourished,
+non-reproducing consumers that are susceptible to mortality.  As a function
+of the transition rates between these replete and undernourished states that
+are determined by the presence or absence of resources, the consumer
+populations can either undergo cyclic dynamics or reach a steady state.  We
+obtain strong constraints on starvation and recovery rates by deriving
+allometric scaling relationships and find that population dynamics subject to
+these constraints can approach the cyclic regime but are typically driven to
+a steady state.  Moreover, we find that these rates fall within a `refuge' in
+parameter space, where the probability of extinction of the consumer
+population is minimized.  Thus we identify a potential mechanism that may
+both drive and constrain the dynamics of animal populations.  Our model
+provides a natural framework that predicts maximum body size for mammals by
+determining the relative stability of an otherwise homogeneous population to
+a mutant population with altered percent body fat.  For body masses
+$\lesssim 10^7$g, individuals with increased energetic reserves can invade
+resident populations, and vice versa for body mass $\gtrsim 10^7$g, thus
+providing a principled mechanism for a within-lineage driver of Cope's rule.
+
+\end{abstract}
+
+%\newpage
+
+%\begin{multicols}{2}
 
@@
 
-\noindent {\bf Abstract} \\ %250 words
-The eco-evolutionary dynamics of species is fundamentally linked to the energetic constraints of its constituent individuals.  
-Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery in resource-limited environments.  
-To elucidate the consequences of this tradeoff, we introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete consumers that reproduce, and undernourished, non-reproducing consumers that are susceptible to mortality.
-As a function of the transition rates between these replete and undernourished states that are determined by the presence or absence of resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
-We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships and find that population dynamics subject to these constraints can approach the cyclic regime but are typically driven to a steady state.
-Moreover, we find that these rates fall within a `refuge' in parameter space, where the probability of extinction of the consumer population is minimized.  
-Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations.  
-Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percent body fat.
-For body masses $\lesssim 10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $\gtrsim 10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
+\section*{Introduction}
+The behavioral ecology of all organisms is influenced by the energetic state
+of individuals, which directly influences how they invest reserves in
+uncertain environments.  Such behaviors are generally manifested as tradeoffs
+between investing in somatic maintenance and growth, or allocating energy
+towards reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.  The
+timing of these behaviors responds to selective pressure, as the choice of
+the investment impacts future
+fitness~\cite{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.  The influence of
+resource limitation on an organism's ability to maintain its nutritional
+stores may lead to repeated delays or shifts in reproduction over the course
+of an organism's life.
 
-%\newpage
+The balance between (a) somatic growth and maintenance, and (b) reproduction
+depends on resource availability~\cite{Morris:1987eo}.  For example, reindeer
+invest less in calves born after harsh winters (when the mother's energetic
+state is depleted) than in calves born after moderate
+winters~\cite{Tveraa:2003fq}.  Many bird species invest differently in broods
+during periods of resource scarcity compared to normal
+periods~\cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing
+reproduction for a breeding
+season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.  Even freshwater
+and marine zooplankton have been observed to avoid reproduction under
+nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have
+lower survival rates~\cite{Kirk:1997cc}.  Organisms may also separate
+maintenance and growth from reproduction over space and time: many salmonids,
+birds, and some mammals return to migratory breeding grounds to reproduce
+after one or multiple seasons in resource-rich environments where they
+accumulate nutritional
+reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
 
-\begin{multicols}{2}
+Physiology also plays an important role in regulating reproductive
+expenditures during periods of resource limitation.  The data collected thus
+far has shown that diverse mammals (47 species in 10 families) exhibit
+delayed implantation, whereby females postpone fetal development (blastocyst
+implantation) until nutritional reserves can be
+accumulated~\cite{Mead:1989dt,Sandell:1990kw}.  Many other many species
+(including humans) suffer irregular menstrual cycling and higher abortion
+rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
+In the extreme case of unicellular organisms, nutrition is unavoidably linked
+to reproduction because the nutritional state of the cell regulates all
+aspects of the cell cycle \cite{Glazier:2009hq}.  The existence of so many
+independently evolved mechanisms across such a diverse suite of organisms
+highlights the importance and universality of the fundamental tradeoff
+between somatic and reproductive investment.  However the general dynamic
+implications of these constraints are unknown.
 
-\noindent {\bf Introduction} \\ 
-The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly influences how they invest reserves in uncertain environments.  
-Such behaviors are generally manifested as tradeoffs between investing in somatic maintenance and growth, or allocating energy towards reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.  
-The timing of these behaviors responds to selective pressure, as the choice of the investment impacts future fitness~\cite{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.  
-The influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
+Though straightforward conceptually, incorporating the energetic dynamics of
+individuals~\cite{Kooi2000} into a population-level
+framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical
+obstacles~\cite{Diekmann:2010da}.  An alternative approach involves modeling
+the macroscale relations that guide somatic versus reproductive investment in
+a consumer-resource system.  For example, macroscale Lotka-Volterra models
+assume that the growth rate of the consumer population depends on resource
+density, thus \emph{implicitly} incorporating the requirement of resource
+availability for reproduction~\cite{murdoch:2003}.
 
-The balance between (a) somatic growth and maintenance, and (b) reproduction depends on resource availability~\cite{Morris:1987eo}.  
-For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\cite{Tveraa:2003fq}.  
-Many bird species invest differently in broods during periods of resource scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
-Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\cite{Kirk:1997cc}.
-Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate nutritional reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+In this work, we adopt an alternative approach in which we \emph{explicitly}
+account for resource limitation and the subsequent effects of starvation.
+Namely, only individuals with sufficient energetic reserves can reproduce.
+Such a constraint leads to reproductive time lags due to some members of the
+population going hungry and then recovering.  Additionally, we incorporate
+the idea that reproduction is strongly constrained allometrically
+\cite{Kempes:2012hy}, and is not generally linearly related to resource
+density.  As we shall show, these constraints influence the ensuing
+population dynamics in dramatic ways.
 
-Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.  
-The data collected thus far has shown that diverse mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development (blastocyst implantation) until nutritional reserves can be accumulated~\cite{Mead:1989dt,Sandell:1990kw}.
-Many other many species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.  
-In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.  
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the importance and universality of the fundamental tradeoff between somatic and reproductive investment.
-However the general dynamic implications of these constraints are unknown.
-
-Though straightforward conceptually, incorporating the energetic dynamics of individuals~\cite{Kooi2000} into a population-level framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles~\cite{Diekmann:2010da}.
-An alternative approach involves modeling the macroscale relations that guide somatic versus reproductive investment in a consumer-resource system.
-For example, macroscale Lotka-Volterra models assume that the growth rate of the consumer population depends on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction~\cite{murdoch:2003}.
-
-In this work, we adopt an alternative approach in which we \emph{explicitly} account for resource limitation and the subsequent effects of starvation.
-Namely, only individuals with sufficient energetic reserves can reproduce.
-Such a constraint leads to reproductive time lags due to some members of the population going hungry and then recovering.  
-Additionally, we incorporate the idea that reproduction is strongly constrained allometrically \cite{Kempes:2012hy}, and is not generally linearly related to resource density.  
-As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.
-\\
-
-\noindent {\bf Nutritional state-structured model (NSM)}\\
+\section*{Nutritional state-structured model (NSM)}
 We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$.
@@
 The NSM represents a fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}.
+
+\begin{figure*}[ht]
+\centering
+\includegraphics[width=0.75\textwidth]{fig_FixedPoint.pdf}
+\caption{\small The transcritical (dashed) and Hopf bifurcation (solid) as a
+  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
+  bifurcation conditions separate parameter space into infeasible, cyclic,
+  and steady state dynamic regimes.  The color gradient shows the steady
+  state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
+  replete consumers $F^*$, (warmer colors denote higher densities).  Steady
+  state densities for the energetically deficient consumers $H^*$ (not shown)
+  scale with those for $F^*$.  }
+\label{fig:fp}
+\end{figure*}
 
@@
 \end{align}
-The stability of this fixed point is determined by the Jacobian matrix $\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.
-The parameters in Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue of $\mathbf{J}$ is negative, so that the system is stable with respect to small perturbations from the fixed point.
-Because this fixed point is unique, it is the global attractor for all population trajectories for any initial condition where the resource and consumer densities are both nonzero.
+The stability of this fixed point is determined by the Jacobian matrix
+$\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$
+when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector
+$(F,H,R)$.  The parameters in Eq.~\eqref{eq:system} are such that the real
+part of the largest eigenvalue of $\mathbf{J}$ is negative, so that the
+system is stable with respect to small perturbations from the fixed point.
+Because this fixed point is unique, it is the global attractor for all
+population trajectories for any initial condition where the resource and
+consumer densities are both nonzero.
 
-From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that $R^*$ is positive.
-In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
-The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates the physical regime from the unphysical regime where $F$ would grow exponentially with time.
-The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
-As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges.
+From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the
+reproduction rate $\lambda$ must be less than the starvation rate $\sigma$,
+so that $R^*$ is positive.  In fact, when the resource density $R=0$, the
+rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
+The condition $\sigma = \lambda$ represents a transcritical (TC)
+bifurcation~\cite{Strogatz:2008wo} that demarcates the physical regime from
+the unphysical regime where $F$ would grow exponentially with time.  The
+biological implication of the constraint $\lambda<\sigma$ has a simple
+interpretation---the rate at which a macroscopic organism loses mass due to
+lack of resources is generally much faster than the rate of reproduction.  As
+we will discuss below, this inequality is a natural consequence of allometric
+constraints~\cite{Kempes:2012hy} for organisms within empirically observed
+body size ranges.
 
-In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).  
-In continuous-time systems, a limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts~\cite{GuckHolmes}.  
-This Hopf bifurcation is defined by ${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial of the Jacobian matrix~\cite{Gross:2004p2428}.
-As the system parameters are tuned to be within the stable regime but close to the Hopf bifurcation, the amplitude of the transient but decaying cycles become large.  
-Given that ecological systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase the extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
-Thus the distance of a system from the Hopf bifurcation provides a measure of its persistence.
+In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may
+either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).  In
+continuous-time systems, a limit cycle arises when a pair of complex
+conjugate eigenvalues crosses the imaginary axis to attain positive real
+parts~\cite{GuckHolmes}.  This Hopf bifurcation is defined by
+${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is
+composed of the coefficients of the characteristic polynomial of the Jacobian
+matrix~\cite{Gross:2004p2428}.  As the system parameters are tuned to be
+within the stable regime but close to the Hopf bifurcation, the amplitude of
+the transient but decaying cycles become large.  Given that ecological
+systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of
+transient cycles, even though they decay with time in the mean-field
+description, can increase the extinction
+risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.  Thus the distance
+of a system from the Hopf bifurcation provides a measure of its persistence.
 
-When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the consumers are driven to the hungry non-reproducing state.  
-Because reproduction is inhibited, there is a low steady-state consumer density and a high steady-state resource density.  
-However, if $\sigma/\lambda\to 1$ from above, the population is overloaded with energetically-replete (reproducing) individuals, thereby promoting oscillations between the consumer and resource densities (Fig.~\ref{fig:fp}).
+When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the
+consumers are driven to the hungry non-reproducing state.  Because
+reproduction is inhibited, there is a low steady-state consumer density and a
+high steady-state resource density.  However, if $\sigma/\lambda\to 1$ from
+above, the population is overloaded with energetically-replete (reproducing)
+individuals, thereby promoting oscillations between the consumer and resource
+densities (Fig.~\ref{fig:fp}).
 
-Whereas the relation between consumer growth rate $\lambda$ and the starvation rate $\sigma$ defines an absolute bound of biological feasibility---the TC bifurcation---$\sigma$ also determines the sensitivity of the consumer population to changes in resource density.
-When $\sigma\gg\lambda$, the steady-state population density is small, thereby increasing the risk of stochastic extinction.  
-On the other hand, as $\sigma$ decreases, the system will ultimately be poised either near the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).  
-If the recovery rate $\rho$ is sufficiently small, the TC bifurcation is reached and the resource eventually is eliminated.  
-If $\rho$ exceeds a threshold value, cyclic dynamics will develop as the Hopf bifurcation is approached.
-\\
+Whereas the relation between consumer growth rate $\lambda$ and the
+starvation rate $\sigma$ defines an absolute bound of biological
+feasibility---the TC bifurcation---$\sigma$ also determines the sensitivity
+of the consumer population to changes in resource density.  When
+$\sigma\gg\lambda$, the steady-state population density is small, thereby
+increasing the risk of stochastic extinction.  On the other hand, as $\sigma$
+decreases, the system will ultimately be poised either near the TC or the
+Hopf bifurcation (Fig.~\ref{fig:fp}).  If the recovery rate $\rho$ is
+sufficiently small, the TC bifurcation is reached and the resource eventually
+is eliminated.  If $\rho$ exceeds a threshold value, cyclic dynamics will
+develop as the Hopf bifurcation is approached.
 
-\noindent {\bf Role of allometry} \\ 
+\section*{Role of allometry}
 While there are no a priori constraints on the parameters in the NSM, most
@@
 
+\begin{figure}
+\includegraphics[width=0.375\textwidth]{Growth-trajectory-diagram.pdf}
+\caption{\small The growth trajectory over absolute time of an individual
+  organism as a function of body mass.  Initial growth follows the black
+  trajectory to an energetically replete reproductive adult mass
+  $m=\epsilon_\lambda M$ which we assume is 95\% asymptotic mass $M$.
+  Starvation follows the red trajectory to
+  $m = \epsilon_\sigma \epsilon_\lambda M$, and recovery follows the green
+  growth trajectory to the replete adult mass. Alternatively, death from
+  starvation follows the blue trajectory to
+  $m=\epsilon_\mu \epsilon_\lambda M$.}
+\label{fig:growth}
+\end{figure}
+
 Several efforts have shown how a partitioning of $B$ between growth and
@@
 \end{eqnarray}
-where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth.
-We now use this solution to define the timescale of reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \cite{moses2008rmo} for a detailed presentation of these timescales).
-The time that it takes to reach a particular mass $\epsilon M$ is given by the timescale
+where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass,
+$a=B_{0}/E_{m}$, and $m_0$ is mass at birth.  We now use this solution to
+define the timescale of reproduction and recovery from starvation
+(Fig.~\ref{fig:growth}; see \cite{moses2008rmo} for a detailed presentation
+of these timescales). The time that it takes to reach a particular mass
+$\epsilon M$ is given by the timescale
 \begin{equation}
@@
 \end{equation}
-where we will define values of $\epsilon$ to describe a set of rates within our model.
-For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$ \cite{West:2001bv}).
-The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$ this will scale like $\lambda\propto M^{\eta-1}$ for $M>>m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
+where we will define values of $\epsilon$ to describe a set of rates within
+our model. For the time to reproduce,
+$t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
+is the fraction of the asymptotic mass where an organism is reproductively
+mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$
+\cite{West:2001bv}). The growth rate is then given by
+$\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number
+of offspring produced, and for any constant value of $\epsilon_{\lambda}$
+this will scale like $\lambda\propto M^{\eta-1}$ for $M>>m_{0}$
+\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
+
+\begin{figure}[ht]
+\includegraphics[width=0.45\textwidth]{fig_Rates.pdf}
+\caption{\small Allometrically constrained starvation rate $\sigma$ (red) and
+  recovery rate $\rho$ (green) relative to the reproductive rate $\lambda$
+  (orange) as a function of body mass.  The rate of starvation is greater
+  than the rate of reproduction for all realized terrestrial endotherm body
+  sizes.  Mean values $\pm 20\%$ variation are shown by the shaded region for
+  each rate.  }
+\label{fig:gvs}
+\end{figure}
 
@@
 and recovery by the timescales required to reach, or return from, specific
-fractions of the replete-state mass (Fig. \ref{fig:gvs}; see Supporting Information, Table I for parameterizations).  We define
-$m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction of
-replete-state mass where reproduction ceases. This fraction will be modified
-if tissue composition systematically scales with adult mass.  For example,
-making use of the observation that body fat in mammals scales with overall
-body size according to $M_{\rm fat}=f_{0}M^{\gamma}$ and assuming that once
-this mass is fully digested the organism starves, this would imply that
-$\epsilon_{\sigma}=1-f_{0}M^{\gamma}/M$. It follows that the
-recovery timescale, $t_{\rho}$, is the time to go from
-$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$ (Fig.
-\ref{fig:growth}).
-Using Eqs.~\eqref{m1} and \eqref{t1} this timescale is given by simply considering an adjusted starting mass of $m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
+fractions of the replete-state mass (Fig. \ref{fig:gvs}; see Supporting
+Information, Table I for parameterizations).  We define
+$m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction
+of replete-state mass where reproduction ceases. This fraction will be
+modified if tissue composition systematically scales with adult mass.  For
+example, making use of the observation that body fat in mammals scales with
+overall body size according to $M_{\rm fat}=f_{0}M^{\gamma}$ and assuming
+that once this mass is fully digested the organism starves, this would imply
+that $\epsilon_{\sigma}=1-f_{0}M^{\gamma}/M$. It follows that the recovery
+timescale, $t_{\rho}$, is the time to go from
+$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$
+(Fig. \ref{fig:growth}). Using Eqs.~\eqref{m1} and \eqref{t1} this timescale
+is given by simply considering an adjusted starting mass of
+$m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
 \begin{eqnarray}
@@
 feature is that $\sigma$ does not have a simple scaling dependence on
-$\lambda$ (Fig. \ref{fig:gvs}), which is important for the dynamics that we
+$\lambda$ (Fig.~\ref{fig:gvs}), which is important for the dynamics that we
 later discuss.
@@
 considerable) variance associated with individual species.
-\\
 
-\noindent {\bf  Stabilizing effects of allometric constraints} \\
+\section*{Stabilizing effects of allometric constraints}
 As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
@@
 
-It should be noted that previous studies have used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
-However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
-Statistically significant support for the existence of population cycles among mammals is predominantly based on time series for small mammals~\cite{Kendall:1998hl}, in agreement with our predictions of more pronounced transient dynamics, given how close these points are to the Hopf bifurcation.
-On the other hand, the longer gestational times and the increased difficulty in measurements, precludes obtaining similar-quality data for larger organisms.
-\vspace{6mm}
-
-\noindent {\bf Extinction risk}\\
-Within our model, higher rates of starvation result in a larger flux of the population to the hungry state.  
-In this state reproduction is absent, thus increasing the likelihood of extinction.
-From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
-We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady state density on extinction risk as a function of $\sigma$ and $\rho$.  
-To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one tenth of the allometrically constrained steady state at any time between $t=10^5$ and $t \leq 10^8$.
-This procedure is repeated for 1000 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an organism of $M=100$ grams.
-In each replicate the initial densities are chosen to be $A(F^*,H^*,R^*)$, with $A$ a random variable that is uniformly distributed in $[0,2]$.  
-By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-6}$ to $10^{-1}$. 
-As expected, higher rates of extinction correlate with both high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if $\sigma$ is small.  
-For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
-For high values of $\sigma$ and low values of $\rho$, higher extinction risk arises because of the decrease in the steady state consumer population density (Figs.
-\ref{fig:fp}\emph{B}, \ref{fig:ext}).
-This interplay creates an `extinction refuge', such that for a constrained range of $\sigma$ and $\rho$, extinction probabilities are minimized.
-
-We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge (Fig. \ref{fig:ext}, white point). 
-These values are close enough to the Hopf bifurcation to avoid low steady state densities, and far enough away to avoid large-amplitude transient cycles. 
-The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
-Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that serve to minimize extinction risk.  
-This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
-To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
-
-\noindent {\bf Dynamic and energetic barriers to body size} \\
-Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\cite{West:2002ud}.
-Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.  
-One important such example is the \emph{fasting endurance hypothesis}, which contends that larger body size, with consequent lower metabolic rates and increased ability to maintain more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability~\cite{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's rule)~\cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\cite{Clauset:2009fh}.
-These trends are thought to be driven by a combination of climate change and niche availability~\cite{Saarinen:2014br}; however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, have not been explored.
-We argue that the NSM provides a suitable framework to explore these issues.
-
-The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
-Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis \cite{allen2002,enquist1998}.
-The energy equivalence hypothesis argues that the total energy use, $B_{tot}$, of a population is constant independent of species size (e.g.
-\cite{allen2002,enquist1998}).
-%Previous studies have focused on t
-This hypothesis is based on observations showing that the abundance, $N$, of a species is proportional to the inverse of individual metabolism (e.g.
-$N\propto M^{-3/4}/B_{0}$)  (e.g.
-\cite{allen2002,enquist1998}).
-This is usually stated as $B_{tot}=NB=C$ where $C$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{allen2002,enquist1998}. 
-Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is relatively constant over this same range.
-This result if remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to the energy equivalence result. 
-Our model shows that the equivalence breaks down at both the minimum and maximum observed body sizes for mammals, suggesting that these are hard limits where deviations outside of this range are energetically suboptimal.
-Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end.
-
-We contend that the NSM provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size as well as the observed trend towards larger body size over evolutionary time.
-The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
-\ref{fig:mass}\emph{C}), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
-Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but does not offer a within-lineage mechanism by which larger body sizes are selected for.
-
-To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $4.5\times$ the mass of a blue whale.
-Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-linneage mechanism for Cope's rule. 
-We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-1,1]$, thus altering the rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and maintenance $\beta(M^\prime)$.
-There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).  
-To assess the susceptibility to invasion as a function of the invader mass, we determine which consumer has a higher steady-state density for a given value of $\chi$.  
-We find that for $1\leq M<8.43\times 10^6$g, having additional body fat ($\chi > 0$) results in a higher steady-state invader population density ($H^{\prime *}+F^{\prime *}>H^*+F^*$).  
-Thus the invader has an intrinsic advantage over the resident population.  
-However, for $M>8.43\times 10^6$, leaner individuals ($\chi < 0$) have advantageous steady state densities.
-
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}=8.43 \times 10^6$ thus serves as an attractor, or an uninvadible evolutionary stable state, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
-Moreover, $M_{\rm opt}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
-While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we suggest that the dynamics of starvation and recovery described in the NSM may provide a general driving mechanism for the evolution of larger body size among terrestrial mammals.
-
-%Closing
-The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations~\cite{Stearns:1989ip}.  
-The NSM is a minimal and general model that incorporates the dynamics of starvation and recovery that are expected to occur in resource-limited environments.  
-By incorporating allometric relations between the rates in the NSM, we found: (i) different organismal masses have distinct population dynamic regimes, (ii) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and (iii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of minimum and maximum body size.
-We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.  
-Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.
-
-\vspace{2mm}
-
-\noindent {\bf Acknowledgments}
-{\small
-  We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
-  discussions and comments on the manuscript.  J.D.Y. was supported by
-  startup funds at the University of California, Merced, and an Omidyar
-  Postdoctoral Fellowship at the Santa Fe Institute.  C.P.K. was supported by
-  an Omidyar Postdoctoral Fellowship at the Santa Fe Institute.  S.R. was
-  supported by grants DMR-1608211 and 1623243 from the National Science
-  Foundation, and by the John Templeton Foundation, all at the Santa Fe
-  Institute.
-}
-
-%\bibliography{aa_starving3}
-\small{
-\begin{thebibliography}{10}
-
-\bibitem{Martin:1987dl}
-T.~E. Martin, ``{Food as a Limit on Breeding Birds: A Life-History
-  Perspective},'' {\em Annu. Rev. Ecol. Syst.}, vol.~18, pp.~453--487, Jan.
-  1987.
-
-\bibitem{Kirk:1997cc}
-K.~L. Kirk, ``{Life-History Responses to Variable Environments: Starvation and
-  Reproduction in Planktonic Rotifers},'' {\em Ecology}, vol.~78, pp.~434--441,
-  Mar. 1997.
-
-\bibitem{Kempes:2012hy}
-C.~P. Kempes, S.~Dutkiewicz, and M.~J. Follows, ``{Growth, metabolic
-  partitioning, and the size of microorganisms.},'' {\em PNAS}, vol.~109,
-  pp.~495--500, Jan. 2012.
-
-\bibitem{Mangel:1988uaa}
-M.~Mangel and C.~W. Clark, {\em {Dynamic Modeling in Behavioral Ecology}}.
-\newblock Princeton: Princeton University Press, 1988.
-
-\bibitem{Mangel:2014kz}
-M.~Mangel, ``{Stochastic dynamic programming illuminates the link between
-  environment, physiology, and evolution},'' {\em B. Math. Biol.}, vol.~77,
-  no.~5, pp.~857--877, 2014.
-
-\bibitem{Yeakel:2013hi}
-J.~D. Yeakel, N.~J. Dominy, P.~L. Koch, and M.~Mangel, ``{Functional
-  morphology, stable isotopes, and human evolution: a model of consilience},''
-  {\em Evolution}, vol.~68, no.~1, pp.~190--203, 2014.
-
-\bibitem{Morris:1987eo}
-D.~W. Morris, ``{Optimal Allocation of Parental Investment},'' {\em Oikos},
-  vol.~49, p.~332, July 1987.
-
-\bibitem{Tveraa:2003fq}
-T.~Tveraa, P.~Fauchald, C.~Henaug, and N.~G. Yoccoz, ``{An examination of a
-  compensatory relationship between food limitation and predation in
-  semi-domestic reindeer},'' {\em Oecologia}, vol.~137, pp.~370--376, Nov.
-  2003.
-
-\bibitem{Daan:1988va}
-S.~Daan, C.~Dijkstra, R.~Drent, and T.~Meijer, ``{Food supply and the annual
-  timing of avian reproduction},'' in {\em Proceedings of the International
-  {\ldots}}, 1988.
-
-\bibitem{Jacot:2009dv}
-A.~Jacot, M.~Valcu, K.~van Oers, and B.~Kempenaers, ``{Experimental nest site
-  limitation affects reproductive strategies and parental investment in a
-  hole-nesting passerine},'' {\em Animal Behaviour}, vol.~77, pp.~1075--1083,
-  May 2009.
-
-\bibitem{Stearns:1989ip}
-S.~C. Stearns, ``{Trade-Offs in Life-History Evolution},'' {\em Funct. Ecol.},
-  vol.~3, no.~3, p.~259, 1989.
-
-\bibitem{Barboza:2002in}
-P.~Barboza and D.~Jorde, ``{Intermittent fasting during winter and spring
-  affects body composition and reproduction of a migratory duck},'' {\em J Comp
-  Physiol B}, vol.~172, pp.~419--434, July 2002.
-
-\bibitem{Threlkeld:1976ih}
-S.~T. Threlkeld, ``{Starvation and the size structure of zooplankton
-  communities},'' {\em Freshwater Biol.}, vol.~6, pp.~489--496, Dec. 1976.
-
-\bibitem{Weber:1998jg}
-T.~P. Weber, B.~J. Ens, and A.~I. Houston, ``{Optimal avian migration: A
-  dynamic model of fuel stores and site use},'' {\em Evolutionary Ecology},
-  vol.~12, pp.~377--401, May 1998.
-
-\bibitem{Mduma:1999cp}
-S.~A.~R. Mduma, A.~R.~E. Sinclair, and R.~Hilborn, ``{Food regulates the
-  Serengeti wildebeest: a 40-year record},'' {\em J. Anim. Ecol.}, vol.~68,
-  pp.~1101--1122, Nov. 1999.
-
-\bibitem{Moore:2014hi}
-J.~W. Moore, J.~D. Yeakel, D.~Peard, J.~Lough, and M.~Beere, ``{Life-history
-  diversity and its importance to population stability and persistence of a
-  migratory fish: steelhead in two large North American watersheds},'' {\em J.
-  Anim. Ecol.}, vol.~83, no.~5, pp.~1035--1046, 2014.
-
-\bibitem{Mead:1989dt}
-R.~A. Mead, ``{The Physiology and Evolution of Delayed Implantation in
-  Carnivores},'' in {\em Carnivore Behavior, Ecology, and Evolution},
-  pp.~437--464, Boston, MA: Springer US, 1989.
-
-\bibitem{Sandell:1990kw}
-M.~Sandell, ``{The Evolution of Seasonal Delayed Implantation},'' {\em The
-  Quarterly Review of Biology}, vol.~65, no.~1, pp.~23--42, 1990.
-
-\bibitem{Bulik:1999eo}
-C.~M. Bulik, P.~F. Sullivan, J.~L. Fear, A.~Pickering, A.~Dawn, and
-  M.~McCullin, ``{Fertility and Reproduction in Women With Anorexia Nervosa},''
-  {\em J. Clin. Psychiatry}, vol.~60, pp.~130--135, Feb. 1999.
-
-\bibitem{Trites:2003cc}
-A.~W. Trites and C.~P. Donnelly, ``{The decline of Steller sea lions Eumetopias
-  jubatus in Alaska: a review of the nutritional stress hypothesis},'' {\em
-  Mammal Review}, vol.~33, pp.~3--28, Mar. 2003.
-
-\bibitem{Glazier:2009hq}
-D.~S. Glazier, ``{Metabolic level and size scaling of rates of respiration and
-  growth in unicellular organisms},'' {\em Funct. Ecol.}, vol.~23,
-  pp.~963--968, Oct. 2009.
-
-\bibitem{Kooi2000}
-S.~A. L.~M. Kooijman, {\em {Dynamic Energy and Mass Budgets in Biological
-  Systems}}.
-\newblock 2000.
-
-\bibitem{Sousa:2010ez}
-T.~Sousa, T.~Domingos, J.~C. Poggiale, and S.~A. L.~M. Kooijman, ``{Dynamic
-  energy budget theory restores coherence in biology},'' {\em Philos. T. Roy.
-  Soc. B}, vol.~365, pp.~3413--3428, Oct. 2010.
-
-\bibitem{Diekmann:2010da}
-O.~Diekmann and J.~A.~J. Metz, ``{How to lift a model for individual behaviour
-  to the population level?},'' {\em Philos. T. Roy. Soc. B}, vol.~365,
-  pp.~3523--3530, Nov. 2010.
-
-\bibitem{murdoch:2003}
-W.~W. Murdoch, C.~J. Briggs, and R.~M. Nisbet, {\em {Consumer-resource
-  Dynamics}}.
-\newblock Monographs in population biology, Princeton University Press, 2003.
-
-\bibitem{Benichou:2014wu}
-O.~Benichou and S.~Redner, ``{Depletion-Controlled Starvation of a Diffusing
-  Forager},'' {\em arXiv}, May 2014.
-
-\bibitem{Benichou:2016wl}
-O.~B{\'e}nichou, M.~Chupeau, and S.~Redner, ``{Role of Depletion on the
-  Dynamics of a Diffusing Forager},'' May 2016.
-
-\bibitem{Chupeau:2016jf}
-M.~Chupeau, O.~B{\'e}nichou, and S.~Redner, ``{Universality classes of foraging
-  with resource renewal},'' {\em Phys. Rev. E}, vol.~93, p.~032403, Mar. 2016.
-
-\bibitem{murray2011mathematical}
-J.~D. Murray, {\em {Mathematical Biology: I. An Introduction}}, vol.~110 of
-  {\em Interdisciplinary Applied Mathematics}.
-\newblock Melaka Manipal Medical College (Manipal Campus), International Centre
-  for Health Sciences, Madhav Nagar, Manipal, Udupi District, Karnataka State,
-  India. nayaksathish@yahoo.com: Springer New York, 2011.
-
-\bibitem{Strogatz:2008wo}
-S.~H. Strogatz, {\em {Nonlinear Dynamics and Chaos}}.
-\newblock With Applications to Physics, Biology, Chemistry, and Engineering,
-  Westview Press, Aug. 2008.
-
-\bibitem{GuckHolmes}
-J.~Guckenheimer and P.~Holmes, {\em {Nonlinear oscillations, dynamical systems,
-  and bifurcations of vector fields}}.
-\newblock New York: Springer, 1983.
-
-\bibitem{Gross:2004p2428}
-T.~Gross and U.~Feudel, ``{Analytical search for bifurcation surfaces in
-  parameter space},'' {\em Physica D}, vol.~195, no.~3-4, pp.~292--302, 2004.
-
-\bibitem{Hastings:2001jh}
-A.~Hastings, ``{Transient dynamics and persistence of ecological systems},''
-  {\em Ecol. Lett.}, vol.~4, pp.~215--220, May 2001.
-
-\bibitem{Neubert:1997wk}
-M.~Neubert and H.~Caswell, ``{Alternatives to resilience for measuring the
-  responses of ecological systems to perturbations},'' {\em Ecology}, vol.~78,
-  no.~3, pp.~653--665, 1997.
-
-\bibitem{Caswell:2005eo}
-H.~Caswell and M.~G. Neubert, ``{Reactivity and transient dynamics of
-  discrete-time ecological systems},'' {\em Journal of Difference Equations and
-  Applications}, vol.~11, pp.~295--310, Apr. 2005.
-
-\bibitem{Neubert:2009td}
-M.~Neubert and H.~Caswell, ``{Detecting reactivity},'' {\em Ecology}, 2009.
-
-\bibitem{Yodzis:1992hg}
-P.~Yodzis and S.~Innes, ``{Body Size and Consumer-Resource Dynamics},'' {\em
-  Am. Nat.}, vol.~139, no.~6, pp.~1151--1175, 1992.
-
-\bibitem{Brown:2004wq}
-J.~Brown, J.~Gillooly, A.~Allen, V.~Savage, and G.~West, ``{Toward a metabolic
-  theory of ecology},'' {\em Ecology}, vol.~85, no.~7, pp.~1771--1789, 2004.
-
-\bibitem{West:2002it}
-G.~B. West, W.~H. Woodruff, and J.~H. Brown, ``{Allometric scaling of metabolic
-  rate from molecules and mitochondria to cells and mammals.},'' {\em Proc.
-  Natl. Acad. Sci. USA}, vol.~99 Suppl 1, pp.~2473--2478, Feb. 2002.
-
-\bibitem{DeLong:2010dy}
-J.~P. DeLong, J.~G. Okie, M.~E. Moses, R.~M. Sibly, and J.~H. Brown, ``{Shifts
-  in metabolic scaling, production, and efficiency across major evolutionary
-  transitions of life.},'' {\em PNAS}, vol.~107, pp.~12941--12945, July 2010.
-
-\bibitem{West:2001bv}
-G.~B. West, J.~H. Brown, and B.~J. Enquist, ``{A general model for ontogenetic
-  growth},'' {\em Nature}, vol.~413, pp.~628--631, Oct. 2001.
-
-\bibitem{moses2008rmo}
-M.~E. Moses, C.~Hou, W.~H. Woodruff, G.~B. West, J.~C. Nekola, W.~Zuo, and
-  J.~H. Brown, ``{Revisiting a Model of Ontogenetic Growth: Estimating Model
-  Parameters from Theory and Data},'' {\em
-  http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}, vol.~171, pp.~632--645,
-  May 2008.
-
-\bibitem{gillooly2002esa}
-J.~F. Gillooly, E.~L. Charnov, G.~B. West, V.~M. Savage, and J.~H. Brown,
-  ``{Effects of size and temperature on developmental time},'' {\em Nature},
-  vol.~417, pp.~70--73, May 2002.
-
-\bibitem{hou}
-C.~Hou, W.~Zuo, M.~E. Moses, W.~H. Woodruff, J.~H. Brown, and G.~B. West,
-  ``{Energy Uptake and Allocation During Ontogeny},'' {\em Science}, vol.~322,
-  pp.~736--739, Oct. 2008.
-
-\bibitem{bettencourt}
-L.~M.~A. Bettencourt, J.~Lobo, D.~Helbing, C.~Kuhnert, and G.~B. West,
-  ``{Growth, innovation, scaling, and the pace of life in cities},'' {\em Proc.
-  Natl. Acad. Sci. USA}, vol.~104, pp.~7301--7306, Apr. 2007.
-
-\bibitem{Folland:2008ij}
-J.~P. Folland, T.~M. Mc~Cauley, and A.~G. Williams, ``{Allometric scaling of
-  strength measurements to body size},'' {\em Eur J Appl Physiol}, vol.~102,
-  pp.~739--745, Jan. 2008.
-
-\bibitem{CalderIII:1983jd}
-W.~A. Calder~III, ``{An allometric approach to population cycles of mammals},''
-  {\em J. Theor. Biol.}, vol.~100, pp.~275--282, Jan. 1983.
-
-\bibitem{Peterson:1984hj}
-R.~O. Peterson, R.~E. PAGE, and K.~M. DODGE, ``{Wolves, Moose, and the
-  Allometry of Population Cycles},'' {\em Science}, vol.~224, pp.~1350--1352,
-  June 1984.
-
-\bibitem{Krukonis:1991fk}
-G.~Krukonis and W.~M. Schaffer, ``{Population cycles in mammals and birds: Does
-  periodicity scale with body size?},'' {\em J. Theor. Biol.}, vol.~148,
-  pp.~469--493, Feb. 1991.
-
-\bibitem{Hendriks:2012fc}
-A.~J. Hendriks and C.~Mulder, ``{Delayed logistic and
-  Rosenzweig{\textendash}MacArthur models with allometric parameter setting
-  estimate population cycles at lower trophic levels well},'' {\em Ecological
-  Complexity}, vol.~9, pp.~43--54, Feb. 2012.
-
-\bibitem{Kendall:1999iy}
-B.~E. Kendall, C.~J. Briggs, W.~W. Murdoch, P.~Turchin, S.~P. Ellner,
-  E.~McCauley, R.~M. Nisbet, and S.~N. Wood, ``{Why do populations cycle? A
-  synthesis of statistical and mechanistic modeling approaches},'' {\em
-  Ecology}, vol.~80, pp.~1789--1805, Sept. 1999.
-
-\bibitem{Hogstedt:2005cr}
-G.~H{\"o}gstedt, T.~Seldal, and A.~Breistol, ``{Period length in cyclic animal
-  populations},'' {\em Ecology}, vol.~86, pp.~373--378, Feb. 2005.
-
-\bibitem{Kendall:1998hl}
-Kendall, Prendergast, and Bjornstad, ``{The macroecology of population
-  dynamics: taxonomic and biogeographic patterns in population cycles},'' {\em
-  Ecol. Lett.}, vol.~1, pp.~160--164, Nov. 1998.
-
-\bibitem{Brown:1993p708}
-J.~Brown, P.~Marquet, and M.~Taper, ``{Evolution of body size: consequences of
-  an energetic definition of fitness},'' {\em Am. Nat.}, vol.~142, no.~4,
-  pp.~573--584, 1993.
-
-\bibitem{West:1997cg}
-G.~B. West, J.~H. Brown, and B.~J. Enquist, ``{A General Model for the Origin
-  of Allometric Scaling Laws in Biology},'' {\em Science}, vol.~276,
-  pp.~122--126, Apr. 1997.
-
-\bibitem{West:2002ud}
-G.~B. West, W.~H. Woodruff, and J.~H. Brown, ``{Allometric scaling of metabolic
-  rate from molecules and mitochondria to cells and mammals},'' {\em Proc.
-  Natl. Acad. Sci. USA}, vol.~99, no.~Suppl 1, pp.~2473--2478, 2002.
-
-\bibitem{Millar:1990p923}
-J.~Millar and G.~Hickling, ``{ Fasting Endurance and the Evolution of Mammalian
-  Body Size},'' {\em Funct. Ecol.}, vol.~4, no.~1, pp.~5--12, 1990.
-
-\bibitem{Alroy:1998p1594}
-J.~Alroy, ``{Cope's rule and the dynamics of body mass evolution in North
-  American fossil mammals},'' {\em Science}, vol.~280, no.~5364, p.~731, 1998.
-
-\bibitem{Clauset:2009fh}
-A.~Clauset and S.~Redner, ``{Evolutionary Model of Species Body Mass
-  Diversification},'' {\em Phys. Rev. Lett.}, vol.~102, p.~038103, Jan. 2009.
-
-\bibitem{Smith:2010p3442}
-F.~Smith, A.~Boyer, J.~Brown, and D.~Costa, ``{The Evolution of Maximum Body
-  Size of Terrestrial Mammals},'' {\em Science}, Jan. 2010.
-
-\bibitem{Saarinen:2014br}
-J.~J. Saarinen, A.~G. Boyer, J.~H. Brown, D.~P. Costa, S.~K.~M. Ernest, A.~R.
-  Evans, M.~Fortelius, J.~L. Gittleman, M.~J. Hamilton, L.~E. Harding,
-  K.~Lintulaakso, S.~K. Lyons, J.~G. Okie, R.~M. Sibly, P.~R. Stephens,
-  J.~Theodor, M.~D. Uhen, and F.~A. Smith, ``{Patterns of maximum body size
-  evolution in Cenozoic land mammals: eco-evolutionary processes and abiotic
-  forcing},'' {\em Proc Biol Sci}, vol.~281, pp.~20132049--20132049, June 2014.
-
-\bibitem{allen2002}
-A.~P. Allen, J.~H. Brown, and J.~F. Gillooly, ``{Global Biodiversity,
-  Biochemical Kinetics, and the Energetic-Equivalence Rule},'' {\em Science},
-  vol.~297, pp.~1545--1548, Aug. 2002.
-
-\bibitem{enquist1998}
-B.~J. Enquist, J.~H. Brown, and G.~B. West, ``{Allometric scaling of plant
-  energetics and population density : Abstract : Nature},'' {\em Nature},
-  vol.~395, pp.~163--165, Sept. 1998.
-
-\bibitem{tilman1981}
-D.~Tilman, ``{Tests of Resource Competition Theory Using Four Species of Lake
-  Michigan Algae},'' {\em Ecology}, vol.~62, pp.~802--815, June 1981.
-
-\bibitem{dutkiewicz2009}
-S.~Dutkiewicz, M.~J. Follows, and J.~G. Bragg, ``{Modeling the coupling of
-  ocean ecology and biogeochemistry},'' {\em Global Biogeochem. Cycles},
-  vol.~23, pp.~n/a--n/a, Dec. 2009.
-
-\bibitem{barton2010}
-A.~D. Barton, S.~Dutkiewicz, G.~Flierl, J.~Bragg, and M.~J. Follows,
-  ``{Patterns of Diversity in Marine Phytoplankton},'' {\em Science}, vol.~327,
-  pp.~1509--1511, Mar. 2010.
-
-\end{thebibliography}
-}
-
-\end{multicols}
-
-\clearpage
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.75\textwidth]{fig_FixedPoint.pdf}
-\caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
-  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-  bifurcation conditions separate parameter space into infeasible, cyclic,
-  and steady state dynamic regimes.  The color gradient shows the steady
-  state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
-  replete consumers $F^*$, (warmer colors denote higher densities).  Steady
-  state densities for the energetically deficient consumers $H^*$ (not shown)
-  scale with those for $F^*$.  }
-\label{fig:fp}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-\caption{ The growth trajectory over absolute time of an individual organism
-  as a function of body mass.  Initial growth follows the black trajectory to
-  an energetically replete reproductive adult mass $m=\epsilon_\lambda M$ which we assume is 95\% asymptotic mass $M$.  Starvation follows the red
-  trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$, and recovery follows the
-  green growth trajectory to the replete adult mass. Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}
-\label{fig:growth}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_Rates.pdf}
-\caption{ 
-Allometrically constrained starvation rate $\sigma$ (red) and recovery rate $\rho$ (green) relative to the reproductive rate $\lambda$ (orange) as a function of body mass.  
-The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
-Mean values $\pm 20\%$ variation are shown by the shaded region for each rate.
-}
-\label{fig:gvs}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_DataHopf.pdf}
-\caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
+\begin{figure}
+\includegraphics[width=0.425\textwidth]{fig_DataHopf.pdf}
+\caption{\small Transcritical (vertical lines) and Hopf bifurcations (curves) for
   allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
@@
 \label{fig:hopf}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric.jpg}
-\caption{ Probability of extinction for a 100g consumer as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $A(F^*,H^*,R^*)$, with
-  $A$ being a random uniform variable in $[0,2]$.  Extinction is defined as the
-  population trajectory falling below $0.1~\times$ the allometrically
-  constrained steady state. The white point denotes the allometrically constrained starvation and recovery rate.}
+It should be noted that previous studies have used allometric constraints to
+explain the periodicity of cyclic populations
+\cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period
+$\propto M^{0.25}$.  However this relation seems to hold only for some
+species \cite{Hendriks:2012fc}, and potential drivers range from predator
+and/or prey lifespans to competitive
+dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.  Statistically significant
+support for the existence of population cycles among mammals is predominantly
+based on time series for small mammals~\cite{Kendall:1998hl}, in agreement
+with our predictions of more pronounced transient dynamics, given how close
+these points are to the Hopf bifurcation.  On the other hand, the longer
+gestational times and the increased difficulty in measurements, precludes
+obtaining similar-quality data for larger organisms.
+
+\section*{Extinction risk}
+Within our model, higher rates of starvation result in a larger flux of the
+population to the hungry state.  In this state reproduction is absent, thus
+increasing the likelihood of extinction.  From the perspective of population
+survival, it is the rate of starvation relative to the rate of recovery that
+determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
+We therefore examine the competing effects of cyclic dynamics vs.\ changes in
+steady state density on extinction risk as a function of $\sigma$ and $\rho$.
+To this end, we computed the probability of extinction, where we define
+extinction as a population trajectory falling below one tenth of the
+allometrically constrained steady state at any time between $t=10^5$ and
+$t \leq 10^8$.  This procedure is repeated for 1000 replicates of the
+continuous-time system shown in Eq.~\ref{eq:system} for an organism of
+$M=100$ grams.  In each replicate the initial densities are chosen to be
+$A(F^*,H^*,R^*)$, with $A$ a random variable that is uniformly distributed in
+$[0,2]$.  By allowing the rate of starvation to vary, we assessed extinction
+risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-6}$
+to $10^{-1}$.  As expected, higher rates of extinction correlate with both
+high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if
+$\sigma$ is small.  For low values of $\sigma$ and high values of $\rho$, the
+increased extinction risk results from transient cycles with larger
+amplitudes as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
+For high values of $\sigma$ and low values of $\rho$, higher extinction risk
+arises because of the decrease in the steady state consumer population
+density (Figs.~\ref{fig:fp}\emph{B}, \ref{fig:ext}).  This interplay creates
+an `extinction refuge', such that for a constrained range of $\sigma$ and
+$\rho$, extinction probabilities are minimized.
+
+\begin{figure}
+\includegraphics[width=0.425\textwidth]{fig_ExtinctionAllometric.jpg}
+\caption{\small Probability of extinction for a 100g consumer as a function of the
+  starvation rate $\sigma$ and recovery rate $\rho$, where the initial
+  density is given as $A(F^*,H^*,R^*)$, with $A$ being a random uniform
+  variable in $[0,2]$.  Extinction is defined as the population trajectory
+  falling below $0.1~\times$ the allometrically constrained steady state. The
+  white point denotes the allometrically constrained starvation and recovery
+  rate.}
 \label{fig:ext}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
-\centering
+We find that the allometrically constrained values of $\sigma$ and $\rho$
+fall squarely within the extinction refuge (Fig. \ref{fig:ext}, white point).
+These values are close enough to the Hopf bifurcation to avoid low steady
+state densities, and far enough away to avoid large-amplitude transient
+cycles.  The fact that allometric values of $\sigma$ and $\rho$ fall within
+this relatively small window supports the possibility that a selective
+mechanism has constrained the physiological conditions that drive starvation
+and recovery rates within populations.  Such a mechanism would select for
+organism physiology that generates appropriate $\sigma$ and $\rho$ values
+that serve to minimize extinction risk.  This selection could occur via the
+tuning of body fat percentages, metabolic rates, and biomass maintenance
+efficiencies.  To summarize, our finding that the allometrically-determined
+parameters fall within this low extinction probability region suggests that
+the NSM dynamics may both drive---and constrain---natural animal populations.
+
+\section*{Dynamic and energetic barriers to body size}
+Metabolite transport constraints are widely thought to place strict
+boundaries on biological
+scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to
+specific predictions on the minimum possible body size for
+organisms~\cite{West:2002ud}.  Above this bound, a number of energetic and
+evolutionary mechanisms have been explored to assess the costs and benefits
+associated with larger body masses, particularly for mammals.  One important
+such example is the \emph{fasting endurance hypothesis}, which contends that
+larger body size, with consequent lower metabolic rates and increased ability
+to maintain more endogenous energetic reserves, may buffer organisms against
+environmental fluctuations in resource availability~\cite{Millar:1990p923}.
+Over evolutionary time, terrestrial mammalian lineages show a significant
+trend towards larger body size (known as Cope's
+rule)~\cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br},
+and it is thought that within-lineage drivers generate selection towards an
+optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, a value
+that is likely limited by higher extinction risk for large taxa over longer
+timescales~\cite{Clauset:2009fh}.  These trends are thought to be driven by a
+combination of climate change and niche availability~\cite{Saarinen:2014br};
+however the underpinning energetic costs and benefits of larger body sizes,
+and how they influence dynamics over ecological timescales, have not been
+explored.  We argue that the NSM provides a suitable framework to explore
+these issues.
+
+\begin{figure}
 \includegraphics[width=0.4\textwidth]{fig_FPAllometric.pdf}
-\caption{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-  body mass. 
-  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.}
+\caption{\small (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$
+  (orange) as a function of body mass.  (\emph{B}) Total energetic use
+  $B_{\rm tot}$ of consumer populations at the steady state as a function of
+  body mass.  (\emph{C}) Resource steady state $R^*$ as a function of
+  consumer body mass.}
 \label{fig:mass}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
-\centering
-\includegraphics[width=0.65\textwidth]{fig_Invasion.pdf}
-\caption{ Invasion feasibility for organisms with a proportional change in
-  mass $\chi$ against a population with a resident body mass $M$.  The blue
-  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
-  red region denotes values of $\chi$ that result in a mass that is below the
-  starvation threshold and is thus infeasible.
-  Arrows point to the predicted optimal mass $M_{\rm opt}=8.43\times 10^6$, which serves as the uninvadable, evolutionary stable state.}
+The NSM correctly predicts that species with smaller masses have larger
+steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).  Moreover,
+we show that the NSM provides independent theoretical support for the energy
+equivalence hypothesis \cite{allen2002,enquist1998}.  The energy equivalence
+hypothesis argues that the total energy use, $B_{tot}$, of a population is
+constant independent of species size (e.g.
+\cite{allen2002,enquist1998}). %Previous studies have focused on t
+This hypothesis is based on observations showing that the abundance, $N$, of
+a species is proportional to the inverse of individual metabolism
+(e.g. $N\propto M^{-3/4}/B_{0}$) (e.g. \cite{allen2002,enquist1998}).  This
+is usually stated as $B_{tot}=NB=C$ where $C$ is a constant, and has been
+shown to hold in both mammalian and vascular plant communities
+\cite{allen2002,enquist1998}.  Figure \ref{fig:mass}\emph{A} shows that both
+$F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism
+sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is relatively
+constant over this same range.  This result if remarkable because it
+illustrates that the steady state values of the NSM combined with the derived
+timescales naturally give rise to the energy equivalence result.  Our model
+shows that the equivalence breaks down at both the minimum and maximum
+observed body sizes for mammals, suggesting that these are hard limits where
+deviations outside of this range are energetically suboptimal.  Significant
+deviations from constant energy use occur at $M \lesssim 1$ at the small end
+of the mammalian range and $M\approx 6.5*10^7$ at the large end.
+
+We contend that the NSM provides a mechanistic understanding of the energetic
+dynamics that give rise to both observed limitations on mammalian body size
+as well as the observed trend towards larger body size over evolutionary
+time.  The NSM predicts that the steady state resource density $R^{*}$
+decreases with increasing body size of the consumer population
+(Fig. \ref{fig:mass}\emph{C}), and classic resource competition theory
+predicts that the species surviving on the lowest resource abundance will
+outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.  Thus, the
+combined NSM steady-state dynamics and allometric timescales predict that
+larger mammals have an intrinsic competitive advantage given a common
+resource, but does not offer a within-lineage mechanism by which larger body
+sizes are selected for.
+
+To examine whether the NSM could provide such a mechanism, we begin by noting
+that a theoretical upper bound on mammalian body size is given by
+$\epsilon_\sigma=0$, where mammals are entirely composed of metabolic
+reserves, and this occurs at $M=8.3\times 10^8$, or $4.5\times$ the mass of a
+blue whale.  Next we examine to what extent a more realistic upper bound to
+body mass may serve as an evolutionary attractor, thus providing a suitable
+within-lineage mechanism for Cope's rule.  We directly assess the
+susceptibility of an otherwise homogeneous population to invasion by a
+mutated subset of the population (denoted by ${}^\prime$) where individuals
+have a modified proportion of body fat $M^\prime=M(1+\chi)$ where
+$\chi \in [-1,1]$, thus altering the rates of starvation $\sigma(M^\prime)$,
+recovery $\rho(M^\prime)$, and maintenance $\beta(M^\prime)$.  There is no
+internal fixed point corresponding to a state where both original residents
+and invaders coexist (except for the trivial state $\chi=0$).  To assess the
+susceptibility to invasion as a function of the invader mass, we determine
+which consumer has a higher steady-state density for a given value of $\chi$.
+We find that for $1\leq M<8.43\times 10^6$g, having additional body fat
+($\chi > 0$) results in a higher steady-state invader population density
+($H^{\prime *}+F^{\prime *}>H^*+F^*$).  Thus the invader has an intrinsic
+advantage over the resident population.  However, for $M>8.43\times 10^6$,
+leaner individuals ($\chi < 0$) have advantageous steady state densities
+(Fig.~\ref{fig:invasion}).
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{fig_Invasion.pdf}
+\caption{\small Invasion feasibility for organisms with a proportional change
+  in mass $\chi$ against a population with a resident body mass $M$.  The
+  blue region denotes proportions of modified mass $\chi$ resulting in
+  successful invasion.  The red region denotes values of $\chi$ that result
+  in a mass that is below the starvation threshold and is thus infeasible.
+  Arrows point to the predicted optimal mass $M_{\rm opt}=8.43\times 10^6$,
+  which serves as the uninvadable, evolutionary stable state.}
 \label{fig:invasion}
-\end{figure*}
+\end{figure}
+
+The observed switch in susceptibility as a function of $\chi$ at
+$M_{\rm opt}=8.43 \times 10^6$ thus serves as an attractor, or an uninvadable
+evolutionary stable state, such that the NSM predicts organismal mass to
+increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.  Moreover,
+$M_{\rm opt}$, which is entirely determined by the population-level
+consequences of energetic constraints, is remarkably close to the maximum
+body size observed in the North American mammalian fossil
+record~\cite{Alroy:1998p1594} as well as the mass predicted from an
+evolutionary model of body size evolution~\cite{Clauset:2009fh}.  While the
+state of the environment, as well as the competitive landscape, will
+determine whether specific body sizes are selected for or
+against~\cite{Saarinen:2014br}, we suggest that the dynamics of starvation
+and recovery described in the NSM may provide a general driving mechanism for
+the evolution of larger body size among terrestrial mammals.
+
+%Closing
+The energetics associated with somatic maintenance, growth, and reproduction
+are important elements that influence the dynamics of all
+populations~\cite{Stearns:1989ip}.  The NSM is a minimal and general model
+that incorporates the dynamics of starvation and recovery that are expected
+to occur in resource-limited environments.  By incorporating allometric
+relations between the rates in the NSM, we found: (i) different organismal
+masses have distinct population dynamic regimes, (ii)
+allometrically-determined rates of starvation and recovery appear to minimize
+extinction risk, and (iii) the dynamic consequences of these rates may
+introduce additional drivers and hard boundaries on the evolution of minimum
+and maximum body size.  We suggest that the NSM offers a means by which the
+dynamic consequences of energetic constraints can be assessed using
+macroscale interactions between and among species.  Future efforts will
+involve exploring the consequences of these dynamics in a spatially explicit
+framework, thus incorporating elements such as movement costs and spatial
+heterogeneity, which may elucidate additional tradeoffs associated with the
+dynamics of starvation and recovery.
+
+\vspace{2mm}
+
+\noindent {\bf Acknowledgments} {\small We thank Luis Bettencourt, Jean
+  Philippe Gibert, Eric Libby, and Seth Newsome for helpful discussions and
+  comments on the manuscript.  J.D.Y. was supported by startup funds at the
+  University of California, Merced, and an Omidyar Postdoctoral Fellowship at
+  the Santa Fe Institute.  C.P.K. was supported by an Omidyar Postdoctoral
+  Fellowship at the Santa Fe Institute.  S.R. was supported by grants
+  DMR-1608211 and 1623243 from the National Science Foundation, and by the
+  John Templeton Foundation, all at the Santa Fe Institute.  }
+
+%\bibliography{aa_starving3}
+\small{
+\begin{thebibliography}{10}
+
+\bibitem{Martin:1987dl} T.~E. Martin, ``{Food as a Limit on Breeding Birds: A
+    Life-History Perspective},'' {\em Annu.\ Rev.\ Ecol.\ Syst.}, vol.~18,
+  pp.~453--487, Jan.\ 1987.
+
+\bibitem{Kirk:1997cc} K.~L. Kirk, ``{Life-History Responses to Variable
+    Environments: Starvation and Reproduction in Planktonic Rotifers},'' {\em
+    Ecology}, vol.~78, pp.~434--441, Mar.\ 1997.
+
+\bibitem{Kempes:2012hy} C.~P. Kempes, S.~Dutkiewicz, and M.~J. Follows,
+  ``{Growth, metabolic partitioning, and the size of microorganisms.},'' {\em
+    Proc.  Natl. Acad. Sci. USA}, vol.~109, pp.~495--500, Jan.\ 2012.
+
+\bibitem{Mangel:1988uaa} M.~Mangel and C.~W. Clark, {\em {Dynamic Modeling in
+      Behavioral Ecology}}.  \newblock Princeton: Princeton University Press,
+  1988.
+
+\bibitem{Mangel:2014kz} M.~Mangel, ``{Stochastic dynamic programming
+    illuminates the link between environment, physiology, and evolution},''
+  {\em Bull.\ Math.\ Biol.}, vol.~77, no.~5, pp.~857--877, 2014.
+
+\bibitem{Yeakel:2013hi} J.~D. Yeakel, N.~J. Dominy, P.~L. Koch, and
+  M.~Mangel, ``{Functional morphology, stable isotopes, and human evolution:
+    a model of consilience},'' {\em Evolution}, vol.~68, no.~1, pp.~190--203,
+  2014.
+
+\bibitem{Morris:1987eo} D.~W. Morris, ``{Optimal Allocation of Parental
+    Investment},'' {\em Oikos}, vol.~49, p.~332, July 1987.
+
+\bibitem{Tveraa:2003fq} T.~Tveraa, P.~Fauchald, C.~Henaug, and N.~G. Yoccoz,
+  ``{An examination of a compensatory relationship between food limitation
+    and predation in semi-domestic reindeer},'' {\em Oecologia}, vol.~137,
+  pp.~370--376, Nov.\ 2003.
+
+\bibitem{Daan:1988va} S.~Daan, C.~Dijkstra, R.~Drent, and T.~Meijer, ``{Food
+    supply and the annual timing of avian reproduction},'' in {\em
+    Proceedings of the International Ornithological Congress} vol.~19,
+    pp.~392-407, 1988.
+
+\bibitem{Jacot:2009dv} A.~Jacot, M.~Valcu, K.~van Oers, and B.~Kempenaers,
+  ``{Experimental nest site limitation affects reproductive strategies and
+    parental investment in a hole-nesting passerine},'' {\em Animal
+    Behaviour}, vol.~77, pp.~1075--1083, May 2009.
+
+\bibitem{Stearns:1989ip} S.~C. Stearns, ``{Trade-Offs in Life-History
+    Evolution},'' {\em Funct. Ecol.}, vol.~3, no.~3, p.~259, 1989.
+
+\bibitem{Barboza:2002in} P.~Barboza and D.~Jorde, ``{Intermittent fasting
+    during winter and spring affects body composition and reproduction of a
+    migratory duck},'' {\em J. Comp.\ Physiol.\ B}, vol.~172, pp.~419--434,
+  July 2002.
+
+\bibitem{Threlkeld:1976ih} S.~T. Threlkeld, ``{Starvation and the size
+    structure of zooplankton communities},'' {\em Freshwater Biol.}, vol.~6,
+  pp.~489--496, Dec.\ 1976.
+
+\bibitem{Weber:1998jg} T.~P. Weber, B.~J. Ens, and A.~I. Houston, ``{Optimal
+    avian migration: A dynamic model of fuel stores and site use},'' {\em
+    Evolutionary Ecology}, vol.~12, pp.~377--401, May 1998.
+
+\bibitem{Mduma:1999cp} S.~A.~R. Mduma, A.~R.~E. Sinclair, and R.~Hilborn,
+  ``{Food regulates the Serengeti wildebeest: a 40-year record},'' {\em
+    J. Anim.\ Ecol.}, vol.~68, pp.~1101--1122, Nov.\ 1999.
+
+\bibitem{Moore:2014hi} J.~W. Moore, J.~D. Yeakel, D.~Peard, J.~Lough, and
+  M.~Beere, ``{Life-history diversity and its importance to population
+    stability and persistence of a migratory fish: steelhead in two large
+    North American watersheds},'' {\em J.  Anim.\ Ecol.}, vol.~83, no.~5,
+  pp.~1035--1046, 2014.
+
+\bibitem{Mead:1989dt} R.~A. Mead, ``{The Physiology and Evolution of Delayed
+    Implantation in Carnivores},'' in {\em Carnivore Behavior, Ecology, and
+    Evolution}, pp.~437--464, Boston, MA: Springer US, 1989.
+
+\bibitem{Sandell:1990kw}
+M.~Sandell, ``{The Evolution of Seasonal Delayed Implantation},'' {\em The
+  Quarterly Review of Biology}, vol.~65, no.~1, pp.~23--42, 1990.
+
+\bibitem{Bulik:1999eo} C.~M. Bulik, P.~F. Sullivan, J.~L. Fear, A.~Pickering,
+  A.~Dawn, and M.~McCullin, ``{Fertility and Reproduction in Women With
+    Anorexia Nervosa},'' {\em J. Clin.\ Psychiatry}, vol.~60, pp.~130--135,
+  Feb.\ 1999.
+
+\bibitem{Trites:2003cc} A.~W. Trites and C.~P. Donnelly, ``{The decline of
+    Steller sea lions Eumetopias jubatus in Alaska: a review of the
+    nutritional stress hypothesis},'' {\em Mammal Review}, vol.~33,
+  pp.~3--28, Mar.\ 2003.
+
+\bibitem{Glazier:2009hq} D.~S. Glazier, ``{Metabolic level and size scaling
+    of rates of respiration and growth in unicellular organisms},'' {\em
+    Funct.\ Ecol.}, vol.~23, pp.~963--968, Oct.\ 2009.
+
+\bibitem{Kooi2000} S.~A. L.~M. Kooijman, {\em {Dynamic Energy and Mass
+      Budgets in Biological Systems}}.  \newblock 2000.
+
+\bibitem{Sousa:2010ez} T.~Sousa, T.~Domingos, J.~C. Poggiale, and
+  S.~A. L.~M. Kooijman, ``{Dynamic energy budget theory restores coherence in
+    biology},'' {\em Philos.\ Trans.\ Roy.\ Soc.\ B}, vol.~365,
+  pp.~3413--3428, Oct.\ 2010.
+
+\bibitem{Diekmann:2010da} O.~Diekmann and J.~A.~J. Metz, ``{How to lift a
+    model for individual behaviour to the population level?},'' {\em Philos.\
+    Trans.\ Roy.\ Soc.\ B}, vol.~365, pp.~3523--3530, Nov.\ 2010.
+
+\bibitem{murdoch:2003} W.~W. Murdoch, C.~J. Briggs, and R.~M. Nisbet, {\em
+    {Consumer-Resource Dynamics}}.  \newblock Monographs in population
+  biology, Princeton University Press, 2003.
+
+\bibitem{Benichou:2014wu} O.~Benichou and S.~Redner, ``{Depletion-Controlled
+    Starvation of a Diffusing Forager},'' {\em Phys.\ Rev.\ Lett.}, vol.~113,
+  p.~238101, Dec.\ 2014.
+
+\bibitem{Benichou:2016wl} O.~B{\'e}nichou, M.~Chupeau, and S.~Redner, ``{Role
+    of Depletion on the Dynamics of a Diffusing Forager},'' J. Phys.\ A:
+  Math.\ Theor.\ (in press).
+
+\bibitem{Chupeau:2016jf} M.~Chupeau, O.~B{\'e}nichou, and S.~Redner,
+  ``{Universality Classes of Foraging with Resource Renewal},'' {\em
+    Phys. Rev. E}, vol.~93, p.~032403, Mar.\ 2016.
+
+\bibitem{murray2011mathematical} J.~D. Murray, {\em {Mathematical Biology:
+      I. An Introduction}}, vol.~110 of {\em Interdisciplinary Applied
+    Mathematics}.  \newblock Melaka Manipal Medical College (Manipal Campus),
+  International Centre for Health Sciences, Madhav Nagar, Manipal, Udupi
+  District, Karnataka State, India. nayaksathish@yahoo.com: Springer New
+  York, 2011.
+
+\bibitem{Strogatz:2008wo} S.~H. Strogatz, {\em {Nonlinear Dynamics and
+      Chaos}}.  \newblock With Applications to Physics, Biology, Chemistry,
+  and Engineering, Westview Press, Aug.\ 2008.
+
+\bibitem{GuckHolmes} J.~Guckenheimer and P.~Holmes, {\em {Nonlinear
+      oscillations, dynamical systems, and bifurcations of vector fields}}.
+  \newblock New York: Springer, 1983.
+
+\bibitem{Gross:2004p2428} T.~Gross and U.~Feudel, ``{Analytical search for
+    bifurcation surfaces in parameter space},'' {\em Physica D}, vol.~195,
+  no.~3-4, pp.~292--302, 2004.
+
+\bibitem{Hastings:2001jh} A.~Hastings, ``{Transient dynamics and persistence
+    of ecological systems},'' {\em Ecol.\ Lett.}, vol.~4, pp.~215--220, May
+  2001.
+
+\bibitem{Neubert:1997wk} M.~Neubert and H.~Caswell, ``{Alternatives to
+    resilience for measuring the responses of ecological systems to
+    perturbations},'' {\em Ecology}, vol.~78, no.~3, pp.~653--665, 1997.
+
+\bibitem{Caswell:2005eo} H.~Caswell and M.~G. Neubert, ``{Reactivity and
+    transient dynamics of discrete-time ecological systems},'' {\em Journal
+    of Difference Equations and Applications}, vol.~11, pp.~295--310, Apr.\
+  2005.
+
+\bibitem{Neubert:2009td} M.~Neubert and H.~Caswell, ``{Detecting
+    reactivity},'' {\em Ecology}, 2009.
+
+\bibitem{Yodzis:1992hg} P.~Yodzis and S.~Innes, ``{Body Size and
+    Consumer-Resource Dynamics},'' {\em Am.\ Nat.}, vol.~139, no.~6,
+  pp.~1151--1175, 1992.
+
+\bibitem{Brown:2004wq} J.~Brown, J.~Gillooly, A.~Allen, V.~Savage, and
+  G.~West, ``{Toward a metabolic theory of ecology},'' {\em Ecology},
+  vol.~85, no.~7, pp.~1771--1789, 2004.
+
+\bibitem{West:2002it}
+G.~B. West, W.~H. Woodruff, and J.~H. Brown, ``{Allometric scaling of metabolic
+  rate from molecules and mitochondria to cells and mammals.},'' {\em Proc.\
+  Natl.\ Acad.\ Sci.\ USA}, vol.~99 Suppl.\ 1, pp.~2473--2478, Feb.\ 2002.
+
+\bibitem{DeLong:2010dy} J.~P. DeLong, J.~G. Okie, M.~E. Moses, R.~M. Sibly,
+  and J.~H. Brown, ``{Shifts in metabolic scaling, production, and efficiency
+    across major evolutionary transitions of life.},'' {\em Proc.\ Natl.\
+    Acad.\ Sci.\ USA}, vol.~107, pp.~12941--12945, July 2010.
+
+\bibitem{West:2001bv} G.~B. West, J.~H. Brown, and B.~J. Enquist, ``{A
+    general model for ontogenetic growth},'' {\em Nature}, vol.~413,
+  pp.~628--631, Oct.\ 2001.
+
+\bibitem{moses2008rmo} M.~E. Moses, C.~Hou, W.~H. Woodruff, G.~B. West,
+  J.~C. Nekola, W.~Zuo, and J.~H. Brown, ``{Revisiting a Model of Ontogenetic
+    Growth: Estimating Model Parameters from Theory and Data},'' {\em
+    http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}, vol.~171,
+  pp.~632--645, May 2008.
+
+\bibitem{gillooly2002esa} J.~F. Gillooly, E.~L. Charnov, G.~B. West,
+  V.~M. Savage, and J.~H. Brown, ``{Effects of size and temperature on
+    developmental time},'' {\em Nature}, vol.~417, pp.~70--73, May 2002.
+
+\bibitem{hou} C.~Hou, W.~Zuo, M.~E. Moses, W.~H. Woodruff, J.~H. Brown, and
+  G.~B. West, ``{Energy Uptake and Allocation During Ontogeny},'' {\em
+    Science}, vol.~322, pp.~736--739, Oct.\ 2008.
+
+\bibitem{bettencourt} L.~M.~A. Bettencourt, J.~Lobo, D.~Helbing, C.~Kuhnert,
+  and G.~B. West, ``{Growth, innovation, scaling, and the pace of life in
+    cities},'' {\em Proc.\ Natl.\ Acad.\ Sci.\ USA}, vol.~104,
+  pp.~7301--7306, Apr.\ 2007.
+
+\bibitem{Folland:2008ij} J.~P. Folland, T.~M. Mc~Cauley, and A.~G. Williams,
+  ``{Allometric scaling of strength measurements to body size},'' {\em Eur.\
+    J. Appl.\ Physiol.}, vol.~102, pp.~739--745, Jan.\ 2008.
+
+\bibitem{CalderIII:1983jd} W.~A. Calder~III, ``{An allometric approach to
+    population cycles of mammals},'' {\em J. Theor.\ Biol.}, vol.~100,
+  pp.~275--282, Jan.\ 1983.
+
+\bibitem{Peterson:1984hj}
+R.~O. Peterson, R.~E. Page, and K.~M. Dodge, ``{Wolves, Moose, and the
+  Allometry of Population Cycles},'' {\em Science}, vol.~224, pp.~1350--1352,
+  June 1984.
+
+\bibitem{Krukonis:1991fk} G.~Krukonis and W.~M. Schaffer, ``{Population
+    cycles in mammals and birds: Does periodicity scale with body size?},''
+  {\em J. Theor.\ Biol.}, vol.~148, pp.~469--493, Feb.\ 1991.
+
+\bibitem{Hendriks:2012fc}
+A.~J. Hendriks and C.~Mulder, ``{Delayed logistic and
+  Rosenzweig{\textendash}MacArthur models with allometric parameter setting
+  estimate population cycles at lower trophic levels well},'' {\em Ecological
+  Complexity}, vol.~9, pp.~43--54, Feb.\ 2012.
+
+\bibitem{Kendall:1999iy} B.~E. Kendall, C.~J. Briggs, W.~W. Murdoch,
+  P.~Turchin, S.~P. Ellner, E.~McCauley, R.~M. Nisbet, and S.~N. Wood, ``{Why
+    do populations cycle? A synthesis of statistical and mechanistic modeling
+    approaches},'' {\em Ecology}, vol.~80, pp.~1789--1805, Sept.\ 1999.
+
+\bibitem{Hogstedt:2005cr} G.~H{\"o}gstedt, T.~Seldal, and A.~Breistol,
+  ``{Period length in cyclic animal populations},'' {\em Ecology}, vol.~86,
+  pp.~373--378, Feb.\ 2005.
+
+\bibitem{Kendall:1998hl} Kendall, Prendergast, and Bjornstad, ``{The
+    macroecology of population dynamics: taxonomic and biogeographic patterns
+    in population cycles},'' {\em Ecol. Lett.}, vol.~1, pp.~160--164,
+  Nov.\ 1998.
+
+\bibitem{Brown:1993p708} J.~Brown, P.~Marquet, and M.~Taper, ``{Evolution of
+    body size: consequences of an energetic definition of fitness},'' {\em
+    Am.\ Nat.}, vol.~142, no.~4, pp.~573--584, 1993.
+
+\bibitem{West:1997cg} G.~B. West, J.~H. Brown, and B.~J. Enquist, ``{A
+    General Model for the Origin of Allometric Scaling Laws in Biology},''
+  {\em Science}, vol.~276, pp.~122--126, Apr.\ 1997.
+
+\bibitem{West:2002ud} G.~B. West, W.~H. Woodruff, and J.~H. Brown,
+  ``{Allometric scaling of metabolic rate from molecules and mitochondria to
+    cells and mammals},'' {\em Proc.\ Natl.\ Acad.\ Sci.\ USA}, vol.~99,
+  no.~Suppl 1, pp.~2473--2478, 2002.
+
+\bibitem{Millar:1990p923} J.~Millar and G.~Hickling, ``{ Fasting Endurance
+    and the Evolution of Mammalian Body Size},'' {\em Funct.\ Ecol.}, vol.~4,
+  no.~1, pp.~5--12, 1990.
+
+\bibitem{Alroy:1998p1594} J.~Alroy, ``{Cope's rule and the dynamics of body
+    mass evolution in North American fossil mammals},'' {\em Science},
+  vol.~280, no.~5364, p.~731, 1998.
+
+\bibitem{Clauset:2009fh} A.~Clauset and S.~Redner, ``{Evolutionary Model of
+    Species Body Mass Diversification},'' {\em Phys. Rev. Lett.}, vol.~102,
+  p.~038103, Jan.\ 2009.
+
+\bibitem{Smith:2010p3442} F.~Smith, A.~Boyer, J.~Brown, and D.~Costa, ``{The
+    Evolution of Maximum Body Size of Terrestrial Mammals},'' {\em Science},
+  Jan.\ 2010.
+
+\bibitem{Saarinen:2014br} J.~J. Saarinen, A.~G. Boyer, J.~H. Brown,
+  D.~P. Costa, S.~K.~M. Ernest, A.~R.  Evans, M.~Fortelius, J.~L. Gittleman,
+  M.~J. Hamilton, L.~E. Harding, K.~Lintulaakso, S.~K. Lyons, J.~G. Okie,
+  R.~M. Sibly, P.~R. Stephens, J.~Theodor, M.~D. Uhen, and F.~A. Smith,
+  ``{Patterns of maximum body size evolution in Cenozoic land mammals:
+    eco-evolutionary processes and abiotic forcing},'' {\em Proc.\
+    Biol\. Sci.}, vol.~281, pp.~20132049--20132049, June 2014.
+
+\bibitem{allen2002} A.~P. Allen, J.~H. Brown, and J.~F. Gillooly, ``{Global
+    Biodiversity, Biochemical Kinetics, and the Energetic-Equivalence
+    Rule},'' {\em Science}, vol.~297, pp.~1545--1548, Aug.\ 2002.
+
+\bibitem{enquist1998} B.~J. Enquist, J.~H. Brown, and G.~B. West,
+  ``{Allometric scaling of plant energetics and population density : Abstract
+    : Nature},'' {\em Nature}, vol.~395, pp.~163--165, Sept.\ 1998.
+
+\bibitem{tilman1981} D.~Tilman, ``{Tests of Resource Competition Theory Using
+    Four Species of Lake Michigan Algae},'' {\em Ecology}, vol.~62,
+  pp.~802--815, June 1981.
+
+\bibitem{dutkiewicz2009} S.~Dutkiewicz, M.~J. Follows, and J.~G. Bragg,
+  ``{Modeling the coupling of ocean ecology and biogeochemistry},'' {\em
+    Global Biogeochem.\ Cycles}, vol.~23, pp.~n/a--n/a, Dec.\ 2009.
+
+\bibitem{barton2010} A.~D. Barton, S.~Dutkiewicz, G.~Flierl, J.~Bragg, and
+  M.~J. Follows, ``{Patterns of Diversity in Marine Phytoplankton},'' {\em
+    Science}, vol.~327, pp.~1509--1511, Mar.\ 2010.
+
+\end{thebibliography}
+}
+
+%\end{multicols}
+
+\clearpage
+
+\section*{Supporting Information for ``The dynamics of starvation and recovery''}%: Eco-evolutionary feedbacks}
+
+\subsection*{Parameter Values and Estimates}
+
+Many of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
+Here we outline previous measurements and simple estimates of the parameters.
+
+\subsubsection*{Standard synthesis and metabolic parameters}
+
+Metabolic rate has been generally reported to follow an exponent close to
+$\eta=0.75$ (e.g. \cite{supp:West:2001bv,supp:moses2008rmo} and the
+supplement of \cite{supp:hou}). We make this assumption in the current paper,
+although alternate exponents, which are know to vary between roughly $0.25$
+and $1.5$ for single species \cite{supp:moses2008rmo}, could be easily
+incorporated into our framework, and this variation is effectively handled by
+the $20\%$ variations that we consider around mean trends. It is important to
+note the exponent, because it not only defines several scalings in our
+framework but also the value of the metabolic normalization constant,
+$B_{0}$, given a set of data.  For mammals the metabolic normalization
+constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and
+$0.047$ (W g$^{-0.75}$) \cite{supp:hou,supp:West:2001bv}, where the former value
+represents basal metabolic rate and the latter represents the field metabolic
+rate. We employ the field metabolic rate for our NSM model which is
+appropriate for active mammals (Table 1).
+
+The energy to synthesize a unit of biomass, $E_{m}$, has been reported to
+vary between $1800$ to $9500$ (J g$^{-1}$)
+(e.g. \cite{supp:West:2001bv,supp:moses2008rmo,hou}) in mammals with a mean
+value across many taxonomic groups of $5,774$ (J g$^{-1}$)
+\cite{supp:moses2008rmo}. The unit energy available during starvation,
+$E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the
+total energy stored during ontogeny \cite{hou} to a biochemical upper bound
+of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate
+\cite{supp:stryer,supp:hou}. For our calculations we use the measured value
+for bulk tissues of $7000$ which assumes that the energy stored during
+ontogeny is returned during starvation \cite{supp:hou}.
+
+For the scaling of body composition it has been shown that fat mass follows
+$M_{\rm fat}=f_{0}M^{\gamma}$, with measured relationships following
+$0.018M^{1.25}$ \cite{supp:Dunbrack:1993ec}, $0.02M^{1.19}$
+\cite{supp:Lindstedt:1985hm}, and $0.026M^{1.14}$
+\cite{supp:Lindstedt:2002td}. We use the values from
+\cite{supp:Lindstedt:1985hm} which falls in the middle of this
+range. Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with
+$u_{0}=0.383$ and $\zeta=1.00$ \cite{supp:Lindstedt:2002td}.
+
+The final parameters that we must consider connect the resource growth rate
+to the total metabolic rate of an organism. That is, we are interested in the
+relative rates of resource recovery and consumption by the total
+population. From \cite{supp:allen2002global} the total resource use of a
+population with an individual body size of $M$ is given by
+$B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$). Considering an energy density of
+$18200$ (J g$^{-1}$) of grass \cite{supp:estermann} and an NPP between and
+$1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give
+a range of resource rates between $0.029$ and $1.44$ (W m$^{-2}$). This gives
+a ratio of total resource consumption to supply rates between $0.00042$ and
+$0.021$, and we used a value of $0.002$ in our calculations and simulations.
+
+%For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
+
+%More complicated models of fecundity (which, for example, account for the average length of adulthood and the number of individuals produced over this span) could be employed.
+However, the scaling of population growth rate has been studied in detail before and follows a relationship of $$ which matches the theory well for $\phi=.95$ and $=$.
+
+%In our calculations we include $20\%$ variation around this value which could account for differences in efficiency during
+
+ \begin{table}[h]
+\caption{Parameter values for mammals}
+\label{param}
+    \begin{center}
+    \small
+     \begin{tabular}{ p{1.8cm} p{2.6cm} l p{2.2cm}|}
+     \hline
+     Parameter & Value & References  \\
+     \hline
+   $\eta$ & $3/4$  &  (e.g. \cite{supp:West:2001bv,supp:moses2008rmo,hou}) \\ 
+   $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \cite{supp:moses2008rmo,supp:West:2001bv,supp:hou} \\ 
+   $E_{m}^{\prime}$ & $36,000$  & \cite{supp:stryer,supp:hou} \\ 
+   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \cite{supp:hou}  \\
+%   $a$ & $1.78\times10^{-6}$  \quad \quad \\ 
+%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\ 
+   $\gamma$ & $1.19$ & \cite{supp:Lindstedt:1985hm} \\ 
+   $f_{0}$ & $0.02$ & \cite{supp:Lindstedt:1985hm}\\ 
+   $\zeta$ & $1.00$  & \cite{supp:Lindstedt:2002td} \\ 
+   $u_{0}$ & $0.38$  & \cite{supp:Lindstedt:2002td} \\
+
+   \hline
+    \end{tabular}
+    \end{center}
+   \end{table}
+
+\subsubsection*{Rate equations for invaders with modified body mass}
+If an invading subset of the resident population of mass $M$ has an altered
+mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $[-1,1]$ ($\chi<0$
+denotes a leaner invader; $\chi > 0$ denotes an invader with more endogenous
+reserves), the invading population will have the following modified rates:
+$\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$,
+$\beta^\prime = \beta(M^\prime)$.  Because we are assuming that the invading
+population is only modifying its endogenous energetic stores, we assume that
+the proportion of body mass that is non-adipose tissue remains the same as
+the resident population.  This assumption leads to the following modified
+timescales:
+
+\begin{align}
+t_{\sigma^\prime} &= \frac{-M^{1/4}}{B_0/E_m^\prime}\log \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
+t_{\rho^\prime} &= \frac{-4 M^{1/4} }{B_0/E_m^\prime}\log \left(\frac{1-( \epsilon_\lambda(\chi +1))^{1/4}}{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}\right), \\ \nonumber
+t_{\beta^\prime} &= \xi B_0\left(M(\chi + 1)\right)^{3/4}.
+\end{align}
+
+\begin{thebibliography}{9}
+
+\bibitem{supp:West:2001bv} G.~B. West, J.~H. Brown, and B.~J. Enquist, ``{A
+    general model for ontogenetic growth},'' {\em Nature}, vol.~413,
+  pp.~628--631, Oct.\ 2001.
+
+\bibitem{supp:moses2008rmo} M.~E. Moses, C.~Hou, W.~H. Woodruff, G.~B. West,
+  J.~C. Nekola, W.~Zuo, and J.~H. Brown, ``{Revisiting a Model of Ontogenetic
+    Growth: Estimating Model Parameters from Theory and Data},'' {\em
+    http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}, vol.~171,
+  pp.~632--645, May 2008.
+
+\bibitem{supp:hou} C.~Hou, W.~Zuo, M.~E. Moses, W.~H. Woodruff, J.~H. Brown, and
+  G.~B. West, ``{Energy Uptake and Allocation During Ontogeny},'' {\em
+    Science}, vol.~322, pp.~736--739, Oct.\ 2008.
+
+\bibitem{supp:stryer} L. Stryer, \emph{Biochemistry, Fourth Edition} (W.H. Freeman and
+  Company, New York, 1995).
+
+\bibitem{supp:Dunbrack:1993ec} R. L. Dunbrack and M. A. Ramsay, ``The Allometry of
+  Mammalian Adaptations to Seasonal Environments: A Critique of the Fasting
+  Endurance Hypothesis,'' \emph{Oikos} vol.~66, pp.~336 (1993).
+
+\bibitem{supp:Lindstedt:1985hm} S. L. Lindstedt and M. S. Boyce, ``Seasonality,
+  Fasting Endurance, and Body Size in Mammals,'' Am.\ Nat.\ vol.~125, pp.~873
+  (1985).
+
+\bibitem{supp:Lindstedt:2002td} S. L. Lindstedt and P. J. Schaeffer, ``Use of
+  allometry in predicting anatomical and physiological parameters of
+  mammals,'' Lab.\ Anim.\ vol.~36, pp.~1 (2002).
+
+\bibitem{supp:allen2002global} A.~P. Allen, J.~H. Brown, and J.~F. Gillooly, ``{Global
+    Biodiversity, Biochemical Kinetics, and the Energetic-Equivalence
+    Rule},'' {\em Science}, vol.~297, pp.~1545--1548, Aug.\ 2002.
+
+\bibitem{supp:estermann} B. L. Estermann, H.-R. Wettstein, F. Sutter, and
+  M. Kreuzer, ``Nutrient and energy conversion of grass-fed dairy and suckler
+  beef cattle kept indoors and on high altitude pasture,'' Anim.\ Res.\ vol.~50,
+  pp.~477 (2001).
+
+\end{thebibliography}
 

### 2017-02-25 Manuscript/13.tex
initial body size and new invasion analysis

--- 2017-02-09 Manuscript/12.tex
@@
 \begin{split}
-\dot{F} &= \lambda F + \rho RH - \sigma (1-R)F,  \\
-\dot{H} &= \sigma (1-R)F - \rho RH - \mu H,  \\
-\dot{R} &= \alpha R(1-R) - R(\rho H+ \beta F).
+\dot{F} &= \lambda F + \rho RH - \sigma \left(1-R\right)F,  \\
+\dot{H} &= \sigma \left(1-R\right)F - \rho RH - \mu H,  \\
+\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+P\right)H-\beta F
 \end{split}
 \end{align}
-Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.  
+{\color{red} This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (please see the supplement for a full derivation and the dimensional form).} Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
 This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\cite{murray2011mathematical}, except that the resource density does not appear in the growth term.  
@@
 Equation~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
+{\color{green}
 \begin{align}
@@
 \end{align}
+}
 The stability of this fixed point is determined by the Jacobian matrix $\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.
@@
 This is usually stated as $B_{tot}=NB=C$ where $C$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{allen2002,enquist1998}. 
-Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is relatively constant over this same range.
+Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
 This result if remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to the energy equivalence result.
-Our model shows that the equivalence breaks down at both the minimum and maximum observed body sizes for mammals, suggesting that these are hard limits where deviations outside of this range are energetically suboptimal.
+{\color{red} Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this is a hard limit where deviations outside of this range are energetically suboptimal.
 %, violating the energy equivalence.
 % set by deviations from constant efficiency obeyed by other populations.
-Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end. 
+A stronger version of this statement is that the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, occuring at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
+This asymptotic behavior represents an upper bound on mammalian body size and occurs at $M_{max}=VALUE$.
+Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
+
+%Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end. 
 %Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is close to the minimum observed mammalian body size of ca.\ 1.3-2.5g, a range that occurs as the recovery rate begins its decline.
@@
 There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).  
-To assess the susceptibility to invasion as a function of the invader mass, we determine which consumer has a higher steady-state density for a given value of $\chi$.  
-We find that for $1\leq M<8.43\times 10^6$g, having additional body fat ($\chi > 0$) results in a higher steady-state invader population density ($H^{\prime *}+F^{\prime *}>H^*+F^*$).  
+To assess the susceptibility to invasion as a function of the invader mass, we determine which consumer has a lower steady-state resource density for a given value of $\chi$.  
+We find that for $1\leq M<VALUE$g, having additional body fat ($\chi > 0$) results in a higher steady-state invader population density ($H^{\prime *}+F^{\prime *}>H^*+F^*$).  
 Thus the invader has an intrinsic advantage over the resident population.  
-However, for $M>8.43\times 10^6$, leaner individuals ($\chi < 0$) have advantageous steady state densities.
+However, for $M>VALUE$, leaner individuals ($\chi < 0$) have advantageous steady state densities.
+
 %, and this is due to the changing
@@
 
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}=8.43 \times 10^6$ thus serves as an attractor, or an uninvadible evolutionary stable state, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
+The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}=VALUE$ thus serves as an attractor, or an uninvadible evolutionary stable state, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
+This value is close to but smaller than the asymptotic upper bound for terrestrial mammal size predicted by the NSM and is also remarkable close to independent estimates of the largest land mammal \cite{Alroy:1998p1594,Clauset:2009fh}.
 %This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
-Moreover, $M_{\rm opt}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
+}
+
 While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we suggest that the dynamics of starvation and recovery described in the NSM may provide a general driving mechanism for the evolution of larger body size among terrestrial mammals.
+
 %and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.

### 2017-02-25 Manuscript/Revision_2017/14.tex
initial body size and new invasion analysis

--- 2017-02-25 Manuscript/13.tex
@@
 {Justin D.
-Yeakel\affil{1}{School of Natural Science, University of California Merced, Merced, CA} \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{To whom correspondence should be addressed: jdyeakel@gmail.com}, Christopher P.
-Kempes \affil{2}{The Santa Fe Institute, Santa Fe, NM}, \and Sidney Redner \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{3}{Department of Physics, Boston University, Boston MA}
+Yeakel\affil{1}{School of Natural Science, University of California Merced, Merced, CA} \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}\affil{5}{To whom correspondence should be addressed: jdyeakel@gmail.com}, Christopher P.
+Kempes \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}, \and Sidney Redner \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{3}{Department of Physics, Boston University, Boston MA}\affil{4}{Contributed equally}
 }
@@
 \begin{split}
-\dot{F} &= \lambda F + \rho RH - \sigma \left(1-R\right)F,  \\
-\dot{H} &= \sigma \left(1-R\right)F - \rho RH - \mu H,  \\
-\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+P\right)H-\beta F
+\dot{F} &= \lambda F + \rho \frac{c}{k} RH - \sigma \left(1-R\right)F,  \\
+\dot{H} &= \sigma \left(1-R\right)F - \rho \frac{c}{k} RH - \mu H,  \\
+\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
 \end{split}
 \end{align}
-{\color{red} This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (please see the supplement for a full derivation and the dimensional form).} Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
+{\color{red} This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth, where $c/k$ scales the somatic regeneration of body tissue from resources and is $\approx 2$ (please see the supplement for a full derivation and the dimensional form).
+}
+Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
 This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\cite{murray2011mathematical}, except that the resource density does not appear in the growth term.
@@
 Equation~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
-{\color{green}
 \begin{align}
@@
 \begin{split}
-F^* &= \frac{\alpha  \lambda  \mu  (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  \beta)}, \\
-H^* &= \frac{\alpha  \lambda ^2 (\mu +\rho )}{(\lambda  \rho +\mu  \sigma ) (\lambda  \rho +\mu  \beta)}, \\
-R^* &= \frac{\mu  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma }.
+F^* &= \frac{(\sigma -\lambda ) \alpha  \lambda  k \mu ^2  (c \rho +k \mu )}{A (\lambda  \rho B +k \mu  \sigma  [\beta  \mu +\lambda  (\delta +\rho )])}, \\
+H^* &= \frac{(\sigma -\lambda ) \alpha  \lambda ^2 k \mu  (c \rho +k \mu )}{A (\lambda  \rho  B+k \mu  \sigma  [\beta  \mu +\lambda  (\delta +\rho )])}, \\
+R^* &= (\sigma -\lambda )k \mu /A.
 \end{split}
 \end{align}
-}
+where $A=(c \lambda  \rho +k \mu  \sigma )$ and $B=(c\beta \mu + c \lambda \delta - k \lambda \mu)$.
 The stability of this fixed point is determined by the Jacobian matrix $\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.
@@
 We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady state density on extinction risk as a function of $\sigma$ and $\rho$.
-To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one tenth of the allometrically constrained steady state at any time between $t=10^5$ and $t \leq 10^8$.
+To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^5$ and $t \leq 10^8$.
 This procedure is repeated for 1000 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an organism of $M=100$ grams.
 %the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
-In each replicate the initial densities are chosen to be $A(F^*,H^*,R^*)$, with $A$ a random variable that is uniformly distributed in $[0,2]$.  
-By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-6}$ to $10^{-1}$.
+In each replicate the initial densities are chosen to be $X(F^*,H^*,R^*)$, with $X$ a random variable that is uniformly distributed in $[0,2]$.
+By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-7}$ to $10^{-3}$.
 %, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
@@
 For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
-For high values of $\sigma$ and low values of $\rho$, higher extinction risk arises because of the decrease in the steady state consumer population density (Figs.
-\ref{fig:fp}\emph{B}, \ref{fig:ext}).
+For high values of $\sigma$ and low values of $\rho$, higher extinction risk arises because of the decrease in the steady state consumer population density (Figs. \ref{fig:fp}\emph{B}, \ref{fig:ext}).
 This interplay creates an `extinction refuge', such that for a constrained range of $\sigma$ and $\rho$, extinction probabilities are minimized.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric.jpg}
+\includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric2.png}
 \caption{ Probability of extinction for a 100g consumer as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $A(F^*,H^*,R^*)$, with

### 2017-02-25 Manuscript/Revision_2017/14.tex
some ms edits on 14

@@
 Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis \cite{allen2002,enquist1998}.
-The energy equivalence hypothesis argues that the total energy use, $B_{tot}$, of a population is constant independent of species size (e.g.
+The energy equivalence hypothesis argues that the total energy use, $B_{tot}=N^*B_0M^\eta$ (where $N^*$ is the steady state population density), of a population is constant independent of species size (e.g.
 \cite{allen2002,enquist1998}).

### 2017-02-28 Manuscript/Revision_2017/15.tex
letter and chris ms

--- 2017-02-25 Manuscript/Revision_2017/14.tex
@@
 \usepackage[left]{lineno}
-%\usepackage[comma,sort&compress]{natbib}
+\usepackage[comma,sort&compress]{natbib}
 
@@
 \begin{abstract} %250 words
-The eco-evolutionary dynamics of species is fundamentally linked to the energetic constraints of its constituent individuals.
+The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of its constituent individuals.
 Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery in resource-limited environments.
@@
 The data collected thus far has shown that diverse mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development (blastocyst implantation) until nutritional reserves can be accumulated~\cite{Mead:1989dt,Sandell:1990kw}.
-Many other many species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
+Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
 In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.
@@
 We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$.
-The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity equal to one.
+The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity $C$.
+{\color{red}  The rate at which consumers transition between states and consume resources is dependent on their overall abundance, the abundance of resources, the efficiency of converting resources into metabolism, and how that metabolism is partitioned between maintenance and growth purposes.
+In the supplement we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple nondimensional form which we have described below (please see the supplement for a detailed derivation).
+
 Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$.
-Conversely, consumers recover from state $H$ to state $F$ at rate $\rho$ and in proportion to $R$.
-Resources are also eaten by the consumers---at rate $\rho$ by hungry consumers and at rate $\beta<\rho$ by full consumers.
-This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
-The NSM represents a fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}.
+Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a ratio between maximal resource consumption and the carrying capacity of the resource (please see the supplement).
+%and $\approx 2$.
+Resources are eaten by the consumers at rate $\rho R + \delta$, which accounts for somatic growth and maintenance for starving consumers, respectively, and at rate $\beta$ by full consumers which is a constant representing maximal maintenance and somatic growth (see supplement).}
+%This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
+The NSM represents a fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\cite{Persson:1998hz}.
 
@@
 \begin{split}
-\dot{F} &= \lambda F + \rho \frac{c}{k} RH - \sigma \left(1-R\right)F,  \\
-\dot{H} &= \sigma \left(1-R\right)F - \rho \frac{c}{k} RH - \mu H,  \\
+\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
+\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
 \dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
@@
 \end{align}
-{\color{red} This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth, where $c/k$ scales the somatic regeneration of body tissue from resources and is $\approx 2$ (please see the supplement for a full derivation and the dimensional form).
-}
+This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see Supplementary Materials for a full derivation and the dimensional form).
 Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
@@
 \begin{split}
-F^* &= \frac{(\sigma -\lambda ) \alpha  \lambda  k \mu ^2  (c \rho +k \mu )}{A (\lambda  \rho B +k \mu  \sigma  [\beta  \mu +\lambda  (\delta +\rho )])}, \\
-H^* &= \frac{(\sigma -\lambda ) \alpha  \lambda ^2 k \mu  (c \rho +k \mu )}{A (\lambda  \rho  B+k \mu  \sigma  [\beta  \mu +\lambda  (\delta +\rho )])}, \\
-R^* &= (\sigma -\lambda )k \mu /A.
+F^* &= \frac{(\sigma-\lambda) \alpha  \lambda  \mu ^2  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
+H^* &= \frac{(\sigma-\lambda) \alpha  \lambda ^2 \mu  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
+R^* &= (\sigma - \lambda)\frac{\mu  }{A}.
 \end{split}
 \end{align}
-where $A=(c \lambda  \rho +k \mu  \sigma )$ and $B=(c\beta \mu + c \lambda \delta - k \lambda \mu)$.
+where $A=(\lambda  \xi  \rho +\mu  \sigma )$ and $B=(\beta  \mu  \xi +\delta  \lambda  \xi -\lambda  \mu )$.
 The stability of this fixed point is determined by the Jacobian matrix $\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.
@@
 In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
-The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates the physical regime from the unphysical regime where $F$ would grow exponentially with time.
+The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates the physical regime from the unphysical regime where all densities become $<0$.
 The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
@@
 Given that ecological systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase the extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
-Thus the distance of a system from the Hopf bifurcation provides a measure of its persistence.
+%Thus the distance of a system from the Hopf bifurcation provides a measure of its persistence.
 
@@
 %Justin adding an explanation for beta
-An organism's metabolic rate $B$ is proportional to the cost of tissue
-maintenance in the absence of growth (i.e., when the body mass is $M$).  By
-definition $B=\beta/\xi$, where $\beta$ is the rate at which resources are
-consumed for full consumers (see Eq.~\eqref{eq:system}) and where $\xi$ is
-related to the conversion efficiency of resource to consumer tissue
-(Supporting Information).
 
@@
 \label{m1}
-\left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}}
+\left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}},
 \end{eqnarray}
-where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth.
+where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically as $9.7M^{0.92}$ {\bf cite?}.
 We now use this solution to define the timescale of reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \cite{moses2008rmo} for a detailed presentation of these timescales).
@@
 \label{t1}
-\tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)}
+\tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)},
 \end{equation}
@@
 The time to death should follow a similar relation, but defined by a lower
-fraction of replete-state mass, $m_{\mu}=\epsilon_{\mu} M$.
+fraction of replete-state mass, $m_{\mu}=\epsilon_{\mu} M$ where $\epsilon_\mu < \epsilon_\sigma$.
 Suppose, for example, that an organism dies once it has digested all fat and
@@
 Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small window of potential values (Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.
-We thus find that the dynamics for all mammalian body sizes is confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
+We thus find that the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
 %, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
-Moreover, for larger $M$, the distance to the Hopf bifurcation increases, while uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}) results in little qualitative difference in the distance to the the Hopf bifurcation.
-These results suggest that small mammals are more prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
-Thus our NSM model predicts that population cycles should be less common for larger species and more common for smaller species, particularly in environments where resources are limiting.%, and
+Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
+These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size, though starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation, suggesting that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
+%Thus our NSM model predicts that population cycles should be less common for smaller species and more common for larger species, particularly in environments where resources are limiting.%, and
 %2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
@@
 %It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic~\cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
-Statistically significant support for the existence of population cycles among mammals is predominantly based on time series for small mammals~\cite{Kendall:1998hl}, in agreement with our predictions of more pronounced transient dynamics, given how close these points are to the Hopf bifurcation.
-On the other hand, the longer gestational times and the increased difficulty in measurements, precludes obtaining similar-quality data for larger organisms.
-\vspace{6mm}
+Statistically significant support for the existence of population cycles among mammals is relatively rare though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
+However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms such that testing this particular result is not straightforward.\\ % in agreement with our predictions of more pronounced transient dynamics, given how close these points are to the Hopf bifurcation.On the other hand,
+% \vspace{6mm}
 %, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
@@
 To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^5$ and $t \leq 10^8$.
-This procedure is repeated for 1000 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an organism of $M=100$ grams.
+This procedure is repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an organism of $M=100$ grams.
 %the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
-In each replicate the initial densities are chosen to be $X(F^*,H^*,R^*)$, with $X$ a random variable that is uniformly distributed in $[0,2]$.
+In each replicate the initial densities are chosen to be $(XF^*,XH^*,R^*)$, with $X$ a random variable that is uniformly distributed in $[0,2]$.
 By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-7}$ to $10^{-3}$.
@@
 Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis \cite{allen2002,enquist1998}.
-The energy equivalence hypothesis argues that the total energy use, $B_{tot}$, of a population is constant independent of species size (e.g.
+The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size (e.g.
 \cite{allen2002,enquist1998}).
 %Previous studies have focused on t
-This hypothesis is based on observations showing that the abundance, $N$, of a species is proportional to the inverse of individual metabolism (e.g.
-$N\propto M^{-3/4}/B_{0}$)  (e.g.
+This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism (e.g.
+$N^*\propto M^{-3/4}/B_{0}$)  (e.g.
 \cite{allen2002,enquist1998}).
-This is usually stated as $B_{tot}=NB=C$ where $C$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{allen2002,enquist1998}.
+This is usually stated as $B_{\rm tot}=N^*B(M)=Q$ where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{allen2002,enquist1998}.
 Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
 This result if remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to the energy equivalence result.
-{\color{red} Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this is a hard limit where deviations outside of this range are energetically suboptimal.
+Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this is a hard limit where deviations outside of this range are energetically suboptimal.
 %, violating the energy equivalence.
 % set by deviations from constant efficiency obeyed by other populations.
-A stronger version of this statement is that the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, occuring at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
-This asymptotic behavior represents an upper bound on mammalian body size and occurs at $M_{max}=VALUE$.
+In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, occurring at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
+This asymptotic behavior represents an upper bound on mammalian body size and occurs at $M_{\rm max}=6.5\times10^7$.
 Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
+{\color{red} It should be noted that this prediction of an asymptotic limit on mammalian size resembles efforts on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here \cite{Kempes:2012hy,Kempes:2016}.}
 
@@
 
-To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $4.5\times$ the mass of a blue whale.
-Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-linneage mechanism for Cope's rule.
-We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-1,1]$, thus altering the rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and maintenance $\beta(M^\prime)$.
-There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
-To assess the susceptibility to invasion as a function of the invader mass, we determine which consumer has a lower steady-state resource density for a given value of $\chi$.
-We find that for $1\leq M<VALUE$g, having additional body fat ($\chi > 0$) results in a higher steady-state invader population density ($H^{\prime *}+F^{\prime *}>H^*+F^*$).
-Thus the invader has an intrinsic advantage over the resident population.
-However, for $M>VALUE$, leaner individuals ($\chi < 0$) have advantageous steady state densities.
+To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120\times$ the mass of a male African elephant.
+Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
+We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-f_0M^{\gamma-1},0.05]$ (where the lower bound of $\chi$ is the proportion of body mass that is not composed of metabolic reserves and the upper bound maintains the organism's reproductive mass), thus altering the rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
+Importantly, $\epsilon_\sigma$, which determines the point along the growth curve the defines the state of starved foragers, is assumed to remain unchanged for the invader population (see Supplemental Material for detailed derivations of invader rates).
+%There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
+
+To assess the susceptibility of the resident population to invasion, we determine which consumer has a lower steady-state resource density for a given value of $\chi$, again with the expectation that populations able to survive on lower resource densities have a competitive advantage \cite{tilman1981}.
+We find that for $M\leq 1.73\times10^7$g, having additional body fat ($\chi > 0$) results in lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
+However, for $M> 1.73\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
 
@@
 
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}=VALUE$ thus serves as an attractor, or an uninvadible evolutionary stable state, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
-This value is close to but smaller than the asymptotic upper bound for terrestrial mammal size predicted by the NSM and is also remarkable close to independent estimates of the largest land mammal \cite{Alroy:1998p1594,Clauset:2009fh}.
+The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.73\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
+This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca.
+$1.3\times10^7$g and the late Miocene \emph{Deinotherium} at ca.
+$1.74\times10^7$g ~\cite{Smith:2010p3442}.
+Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \cite{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals outlined by Alroy~\cite{Alroy:1998p1594}.
+%as well as  \cite{Alroy:1998p1594,Clauset:2009fh}.
 %This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
-}
-
-While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we suggest that the dynamics of starvation and recovery described in the NSM may provide a general driving mechanism for the evolution of larger body size among terrestrial mammals.
+While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM may provide a general within-linneage mechanism for the evolution of larger body size among terrestrial mammals.
 
@@
 The NSM is a minimal and general model that incorporates the dynamics of starvation and recovery that are expected to occur in resource-limited environments.
-By incorporating allometric relations between the rates in the NSM, we found: (i) different organismal masses have distinct population dynamic regimes, (ii) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and (iii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of minimum and maximum body size.
+By incorporating allometric relations between the rates in the NSM, we found: (i) different organismal masses have distinct population dynamic regimes, (ii) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and (iii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of maximum body size.
 We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.
@@
 
-%\bibliography{aa_starving3}
-
-\begin{thebibliography}{10}
-
-\bibitem{Martin:1987dl}
-Martin TE
-\newblock (1987) {Food as a Limit on Breeding Birds: A Life-History
-  Perspective}.
-\newblock \emph{Annu. Rev. Ecol. Syst.} 18:453--487.
-
-\bibitem{Kirk:1997cc}
-Kirk KL
-\newblock (1997) {Life-History Responses to Variable Environments: Starvation
-  and Reproduction in Planktonic Rotifers}.
-\newblock \emph{Ecology} 78:434--441.
-
-\bibitem{Kempes:2012hy}
-Kempes CP, Dutkiewicz S, Follows MJ
-\newblock (2012) {Growth, metabolic partitioning, and the size of
-  microorganisms.}
-\newblock \emph{PNAS} 109:495--500.
-
-\bibitem{Mangel:1988uaa}
-Mangel M, Clark CW
-\newblock (1988) \emph{{Dynamic Modeling in Behavioral Ecology}}
-\newblock (Princeton University Press, Princeton).
-
-\bibitem{Mangel:2014kz}
-Mangel M
-\newblock (2014) {Stochastic dynamic programming illuminates the link between
-  environment, physiology, and evolution}.
-\newblock \emph{B. Math. Biol.} 77:857--877.
-
-\bibitem{Yeakel:2013hi}
-Yeakel JD, Dominy NJ, Koch PL, Mangel M
-\newblock (2014) {Functional morphology, stable isotopes, and human evolution:
-  a model of consilience}.
-\newblock \emph{Evolution} 68:190--203.
-
-\bibitem{Morris:1987eo}
-Morris DW
-\newblock (1987) {Optimal Allocation of Parental Investment}.
-\newblock \emph{Oikos} 49:332.
-
-\bibitem{Tveraa:2003fq}
-Tveraa T, Fauchald P, Henaug C, Yoccoz NG
-\newblock (2003) {An examination of a compensatory relationship between food
-  limitation and predation in semi-domestic reindeer}.
-\newblock \emph{Oecologia} 137:370--376.
-
-\bibitem{Daan:1988va}
-Daan S, Dijkstra C, Drent R, Meijer T
-\newblock (1988) \emph{{Food supply and the annual timing of avian
-  reproduction}}.
-
-\bibitem{Jacot:2009dv}
-Jacot A, Valcu M, van Oers K, Kempenaers B
-\newblock (2009) {Experimental nest site limitation affects reproductive
-  strategies and parental investment in a hole-nesting passerine}.
-\newblock \emph{Animal Behaviour} 77:1075--1083.
-
-\bibitem{Stearns:1989ip}
-Stearns SC
-\newblock (1989) {Trade-Offs in Life-History Evolution}.
-\newblock \emph{Funct. Ecol.} 3:259.
-
-\bibitem{Barboza:2002in}
-Barboza P, Jorde D
-\newblock (2002) {Intermittent fasting during winter and spring affects body
-  composition and reproduction of a migratory duck}.
-\newblock \emph{J Comp Physiol B} 172:419--434.
-
-\bibitem{Threlkeld:1976ih}
-Threlkeld ST
-\newblock (1976) {Starvation and the size structure of zooplankton
-  communities}.
-\newblock \emph{Freshwater Biol.} 6:489--496.
-
-\bibitem{Weber:1998jg}
-Weber TP, Ens BJ, Houston AI
-\newblock (1998) {Optimal avian migration: A dynamic model of fuel stores and
-  site use}.
-\newblock \emph{Evolutionary Ecology} 12:377--401.
-
-\bibitem{Mduma:1999cp}
-Mduma SAR, Sinclair ARE, Hilborn R
-\newblock (1999) {Food regulates the Serengeti wildebeest: a 40-year record}.
-\newblock \emph{J. Anim. Ecol.} 68:1101--1122.
-
-\bibitem{Moore:2014hi}
-Moore JW, Yeakel JD, Peard D, Lough J, Beere M
-\newblock (2014) {Life-history diversity and its importance to population
-  stability and persistence of a migratory fish: steelhead in two large North
-  American watersheds}.
-\newblock \emph{J. Anim. Ecol.} 83:1035--1046.
-
-\bibitem{Mead:1989dt}
-Mead RA
-\newblock (1989) in \emph{Carnivore Behavior, Ecology, and Evolution}
-\newblock (Springer US, Boston, MA), pp 437--464.
-
-\bibitem{Sandell:1990kw}
-Sandell M
-\newblock (1990) {The Evolution of Seasonal Delayed Implantation}.
-\newblock \emph{The Quarterly Review of Biology} 65:23--42.
-
-\bibitem{Bulik:1999eo}
-Bulik CM, {et~al.}
-\newblock (1999) {Fertility and Reproduction in Women With Anorexia Nervosa}.
-\newblock \emph{J. Clin. Psychiatry} 60:130--135.
-
-\bibitem{Trites:2003cc}
-Trites AW, Donnelly CP
-\newblock (2003) {The decline of Steller sea lions Eumetopias jubatus in
-  Alaska: a review of the nutritional stress hypothesis}.
-\newblock \emph{Mammal Review} 33:3--28.
-
-\bibitem{Glazier:2009hq}
-Glazier DS
-\newblock (2009) {Metabolic level and size scaling of rates of respiration and
-  growth in unicellular organisms}.
-\newblock \emph{Funct. Ecol.} 23:963--968.
-
-\bibitem{Kooi2000}
-Kooijman SALM
-\newblock (2000) \emph{{Dynamic Energy and Mass Budgets in Biological Systems}}
-\newblock (Cambridge).
-
-\bibitem{Sousa:2010ez}
-Sousa T, Domingos T, Poggiale JC, Kooijman SALM
-\newblock (2010) {Dynamic energy budget theory restores coherence in biology}.
-\newblock \emph{Philos. T. Roy. Soc. B} 365:3413--3428.
-
-\bibitem{Diekmann:2010da}
-Diekmann O, Metz JAJ
-\newblock (2010) {How to lift a model for individual behaviour to the
-  population level?}
-\newblock \emph{Philos. T. Roy. Soc. B} 365:3523--3530.
-
-\bibitem{murdoch:2003}
-Murdoch WW, Briggs CJ, Nisbet RM
-\newblock (2003) \emph{{Consumer-resource Dynamics}}, Monographs in population
-  biology
-\newblock (Princeton University Press).
-
-\bibitem{Benichou:2014wu}
-Benichou O, Redner S
-\newblock (2014) {Depletion-Controlled Starvation of a Diffusing Forager}.
-\newblock \emph{arXiv}.
-
-\bibitem{Benichou:2016wl}
-B{\'e}nichou O, Chupeau M, Redner S
-\newblock (2016) {Role of Depletion on the Dynamics of a Diffusing Forager}.
-
-\bibitem{Chupeau:2016jf}
-Chupeau M, B{\'e}nichou O, Redner S
-\newblock (2016) {Universality classes of foraging with resource renewal}.
-\newblock \emph{Phys. Rev. E} 93:032403.
-
-\bibitem{murray2011mathematical}
-Murray JD
-\newblock (2011) \emph{{Mathematical Biology: I. An Introduction}},
-  Interdisciplinary Applied Mathematics
-\newblock (Springer New York, Melaka Manipal Medical College (Manipal Campus),
-  International Centre for Health Sciences, Madhav Nagar, Manipal, Udupi
-  District, Karnataka State, India. nayaksathish@yahoo.com) Vol.{} 110.
-
-\bibitem{Strogatz:2008wo}
-Strogatz SH
-\newblock (2008) \emph{{Nonlinear Dynamics and Chaos}}, With Applications to
-  Physics, Biology, Chemistry, and Engineering
-\newblock (Westview Press).
-
-\bibitem{GuckHolmes}
-Guckenheimer J, Holmes P
-\newblock (1983) \emph{{Nonlinear oscillations, dynamical systems, and
-  bifurcations of vector fields}}
-\newblock (Springer, New York).
-
-\bibitem{Gross:2004p2428}
-Gross T, Feudel U
-\newblock (2004) {Analytical search for bifurcation surfaces in parameter
-  space}.
-\newblock \emph{Physica D} 195:292--302.
-
-\bibitem{Hastings:2001jh}
-Hastings A
-\newblock (2001) {Transient dynamics and persistence of ecological systems}.
-\newblock \emph{Ecol. Lett.} 4:215--220.
-
-\bibitem{Neubert:1997wk}
-Neubert M, Caswell H
-\newblock (1997) {Alternatives to resilience for measuring the responses of
-  ecological systems to perturbations}.
-\newblock \emph{Ecology} 78:653--665.
-
-\bibitem{Caswell:2005eo}
-Caswell H, Neubert MG
-\newblock (2005) {Reactivity and transient dynamics of discrete-time ecological
-  systems}.
-\newblock \emph{Journal of Difference Equations and Applications} 11:295--310.
-
-\bibitem{Neubert:2009td}
-Neubert M, Caswell H
-\newblock (2009) {Detecting reactivity}.
-\newblock \emph{Ecology}.
-
-\bibitem{Yodzis:1992hg}
-Yodzis P, Innes S
-\newblock (1992) {Body Size and Consumer-Resource Dynamics}.
-\newblock \emph{Am. Nat.} 139:1151--1175.
-
-\bibitem{Brown:2004wq}
-Brown J, Gillooly J, Allen A, Savage V, West G
-\newblock (2004) {Toward a metabolic theory of ecology}.
-\newblock \emph{Ecology} 85:1771--1789.
-
-\bibitem{West:2002it}
-West GB, Woodruff WH, Brown JH
-\newblock (2002) {Allometric scaling of metabolic rate from molecules and
-  mitochondria to cells and mammals.}
-\newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.
-
-\bibitem{DeLong:2010dy}
-DeLong JP, Okie JG, Moses ME, Sibly RM, Brown JH
-\newblock (2010) {Shifts in metabolic scaling, production, and efficiency
-  across major evolutionary transitions of life.}
-\newblock \emph{PNAS} 107:12941--12945.
-
-\bibitem{West:2001bv}
-West GB, Brown JH, Enquist BJ
-\newblock (2001) {A general model for ontogenetic growth}.
-\newblock \emph{Nature} 413:628--631.
-
-\bibitem{moses2008rmo}
-Moses ME, {et~al.}
-\newblock (2008) {Revisiting a Model of Ontogenetic Growth: Estimating Model
-  Parameters from Theory and Data}.
-\newblock \emph{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}
-  171:632--645.
-
-\bibitem{gillooly2002esa}
-Gillooly JF, Charnov EL, West GB, Savage VM, Brown JH
-\newblock (2002) {Effects of size and temperature on developmental time}.
-\newblock \emph{Nature} 417:70--73.
-
-\bibitem{hou}
-Hou C, {et~al.}
-\newblock (2008) {Energy Uptake and Allocation During Ontogeny}.
-\newblock \emph{Science} 322:736--739.
-
-\bibitem{bettencourt}
-Bettencourt LMA, Lobo J, Helbing D, Kuhnert C, West GB
-\newblock (2007) {Growth, innovation, scaling, and the pace of life in cities}.
-\newblock \emph{Proc. Natl. Acad. Sci. USA} 104:7301--7306.
-
-\bibitem{Folland:2008ij}
-Folland JP, Mc~Cauley TM, Williams AG
-\newblock (2008) {Allometric scaling of strength measurements to body size}.
-\newblock \emph{Eur J Appl Physiol} 102:739--745.
-
-\bibitem{CalderIII:1983jd}
-Calder~III WA
-\newblock (1983) {An allometric approach to population cycles of mammals}.
-\newblock \emph{J. Theor. Biol.} 100:275--282.
-
-\bibitem{Peterson:1984hj}
-Peterson RO, PAGE RE, DODGE KM
-\newblock (1984) {Wolves, Moose, and the Allometry of Population Cycles}.
-\newblock \emph{Science} 224:1350--1352.
-
-\bibitem{Krukonis:1991fk}
-Krukonis G, Schaffer WM
-\newblock (1991) {Population cycles in mammals and birds: Does periodicity
-  scale with body size?}
-\newblock \emph{J. Theor. Biol.} 148:469--493.
-
-\bibitem{Hendriks:2012fc}
-Hendriks AJ, Mulder C
-\newblock (2012) {Delayed logistic and Rosenzweig{\textendash}MacArthur models
-  with allometric parameter setting estimate population cycles at lower trophic
-  levels well}.
-\newblock \emph{Ecological Complexity} 9:43--54.
-
-\bibitem{Kendall:1999iy}
-Kendall BE, {et~al.}
-\newblock (1999) {Why do populations cycle? A synthesis of statistical and
-  mechanistic modeling approaches}.
-\newblock \emph{Ecology} 80:1789--1805.
-
-\bibitem{Hogstedt:2005cr}
-H{\"o}gstedt G, Seldal T, Breistol A
-\newblock (2005) {Period length in cyclic animal populations}.
-\newblock \emph{Ecology} 86:373--378.
-
-\bibitem{Kendall:1998hl}
-Kendall, Prendergast, Bjornstad
-\newblock (1998) {The macroecology of population dynamics: taxonomic and
-  biogeographic patterns in population cycles}.
-\newblock \emph{Ecol. Lett.} 1:160--164.
-
-\bibitem{Brown:1993p708}
-Brown J, Marquet P, Taper M
-\newblock (1993) {Evolution of body size: consequences of an energetic
-  definition of fitness}.
-\newblock \emph{Am. Nat.} 142:573--584.
-
-\bibitem{West:1997cg}
-West GB, Brown JH, Enquist BJ
-\newblock (1997) {A General Model for the Origin of Allometric Scaling Laws in
-  Biology}.
-\newblock \emph{Science} 276:122--126.
-
-\bibitem{West:2002ud}
-West GB, Woodruff WH, Brown JH
-\newblock (2002) {Allometric scaling of metabolic rate from molecules and
-  mitochondria to cells and mammals}.
-\newblock \emph{Proc. Natl. Acad. Sci. USA} 99:2473--2478.
-
-\bibitem{Millar:1990p923}
-Millar J, Hickling G
-\newblock (1990) { Fasting Endurance and the Evolution of Mammalian Body Size}.
-\newblock \emph{Funct. Ecol.} 4:5--12.
-
-\bibitem{Alroy:1998p1594}
-Alroy J
-\newblock (1998) {Cope's rule and the dynamics of body mass evolution in North
-  American fossil mammals}.
-\newblock \emph{Science} 280:731.
-
-\bibitem{Clauset:2009fh}
-Clauset A, Redner S
-\newblock (2009) {Evolutionary Model of Species Body Mass Diversification}.
-\newblock \emph{Phys. Rev. Lett.} 102:038103.
-
-\bibitem{Smith:2010p3442}
-Smith F, Boyer A, Brown J, Costa D
-\newblock (2010) {The Evolution of Maximum Body Size of Terrestrial Mammals}.
-\newblock \emph{Science}.
-
-\bibitem{Saarinen:2014br}
-Saarinen JJ, {et~al.}
-\newblock (2014) {Patterns of maximum body size evolution in Cenozoic land
-  mammals: eco-evolutionary processes and abiotic forcing}.
-\newblock \emph{Proc Biol Sci} 281:20132049--20132049.
-
-\bibitem{allen2002}
-Allen AP, Brown JH, Gillooly JF
-\newblock (2002) {Global Biodiversity, Biochemical Kinetics, and the
-  Energetic-Equivalence Rule}.
-\newblock \emph{Science} 297:1545--1548.
-
-\bibitem{enquist1998}
-Enquist BJ, Brown JH, West GB
-\newblock (1998) {Allometric scaling of plant energetics and population density
-  : Abstract : Nature}.
-\newblock \emph{Nature} 395:163--165.
-
-\bibitem{tilman1981}
-Tilman D
-\newblock (1981) {Tests of Resource Competition Theory Using Four Species of
-  Lake Michigan Algae}.
-\newblock \emph{Ecology} 62:802--815.
-
-\bibitem{dutkiewicz2009}
-Dutkiewicz S, Follows MJ, Bragg JG
-\newblock (2009) {Modeling the coupling of ocean ecology and biogeochemistry}.
-\newblock \emph{Global Biogeochem. Cycles} 23:n/a--n/a.
-
-\bibitem{barton2010}
-Barton AD, Dutkiewicz S, Flierl G, Bragg J, Follows MJ
-\newblock (2010) {Patterns of Diversity in Marine Phytoplankton}.
-\newblock \emph{Science} 327:1509--1511.
-
-\end{thebibliography}
+\bibliography{aa_starving3}
+
+% \begin{thebibliography}{10}
+%
+% \bibitem{Martin:1987dl}
+% Martin TE
+% \newblock (1987) {Food as a Limit on Breeding Birds: A Life-History
+%   Perspective}.
+% \newblock \emph{Annu. Rev. Ecol. Syst.} 18:453--487.
+%
+% \bibitem{Kirk:1997cc}
+% Kirk KL
+% \newblock (1997) {Life-History Responses to Variable Environments: Starvation
+%   and Reproduction in Planktonic Rotifers}.
+% \newblock \emph{Ecology} 78:434--441.
+%
+% \bibitem{Kempes:2012hy}
+% Kempes CP, Dutkiewicz S, Follows MJ
+% \newblock (2012) {Growth, metabolic partitioning, and the size of
+%   microorganisms.}
+% \newblock \emph{PNAS} 109:495--500.
+%
+% \bibitem{Mangel:1988uaa}
+% Mangel M, Clark CW
+% \newblock (1988) \emph{{Dynamic Modeling in Behavioral Ecology}}
+% \newblock (Princeton University Press, Princeton).
+%
+% \bibitem{Mangel:2014kz}
+% Mangel M
+% \newblock (2014) {Stochastic dynamic programming illuminates the link between
+%   environment, physiology, and evolution}.
+% \newblock \emph{B. Math. Biol.} 77:857--877.
+%
+% \bibitem{Yeakel:2013hi}
+% Yeakel JD, Dominy NJ, Koch PL, Mangel M
+% \newblock (2014) {Functional morphology, stable isotopes, and human evolution:
+%   a model of consilience}.
+% \newblock \emph{Evolution} 68:190--203.
+%
+% \bibitem{Morris:1987eo}
+% Morris DW
+% \newblock (1987) {Optimal Allocation of Parental Investment}.
+% \newblock \emph{Oikos} 49:332.
+%
+% \bibitem{Tveraa:2003fq}
+% Tveraa T, Fauchald P, Henaug C, Yoccoz NG
+% \newblock (2003) {An examination of a compensatory relationship between food
+%   limitation and predation in semi-domestic reindeer}.
+% \newblock \emph{Oecologia} 137:370--376.
+%
+% \bibitem{Daan:1988va}
+% Daan S, Dijkstra C, Drent R, Meijer T
+% \newblock (1988) \emph{{Food supply and the annual timing of avian
+%   reproduction}}.
+%
+% \bibitem{Jacot:2009dv}
+% Jacot A, Valcu M, van Oers K, Kempenaers B
+% \newblock (2009) {Experimental nest site limitation affects reproductive
+%   strategies and parental investment in a hole-nesting passerine}.
+% \newblock \emph{Animal Behaviour} 77:1075--1083.
+%
+% \bibitem{Stearns:1989ip}
+% Stearns SC
+% \newblock (1989) {Trade-Offs in Life-History Evolution}.
+% \newblock \emph{Funct. Ecol.} 3:259.
+%
+% \bibitem{Barboza:2002in}
+% Barboza P, Jorde D
+% \newblock (2002) {Intermittent fasting during winter and spring affects body
+%   composition and reproduction of a migratory duck}.
+% \newblock \emph{J Comp Physiol B} 172:419--434.
+%
+% \bibitem{Threlkeld:1976ih}
+% Threlkeld ST
+% \newblock (1976) {Starvation and the size structure of zooplankton
+%   communities}.
+% \newblock \emph{Freshwater Biol.} 6:489--496.
+%
+% \bibitem{Weber:1998jg}
+% Weber TP, Ens BJ, Houston AI
+% \newblock (1998) {Optimal avian migration: A dynamic model of fuel stores and
+%   site use}.
+% \newblock \emph{Evolutionary Ecology} 12:377--401.
+%
+% \bibitem{Mduma:1999cp}
+% Mduma SAR, Sinclair ARE, Hilborn R
+% \newblock (1999) {Food regulates the Serengeti wildebeest: a 40-year record}.
+% \newblock \emph{J. Anim. Ecol.} 68:1101--1122.
+%
+% \bibitem{Moore:2014hi}
+% Moore JW, Yeakel JD, Peard D, Lough J, Beere M
+% \newblock (2014) {Life-history diversity and its importance to population
+%   stability and persistence of a migratory fish: steelhead in two large North
+%   American watersheds}.
+% \newblock \emph{J. Anim. Ecol.} 83:1035--1046.
+%
+% \bibitem{Mead:1989dt}
+% Mead RA
+% \newblock (1989) in \emph{Carnivore Behavior, Ecology, and Evolution}
+% \newblock (Springer US, Boston, MA), pp 437--464.
+%
+% \bibitem{Sandell:1990kw}
+% Sandell M
+% \newblock (1990) {The Evolution of Seasonal Delayed Implantation}.
+% \newblock \emph{The Quarterly Review of Biology} 65:23--42.
+%
+% \bibitem{Bulik:1999eo}
+% Bulik CM, {et~al.}
+% \newblock (1999) {Fertility and Reproduction in Women With Anorexia Nervosa}.
+% \newblock \emph{J. Clin. Psychiatry} 60:130--135.
+%
+% \bibitem{Trites:2003cc}
+% Trites AW, Donnelly CP
+% \newblock (2003) {The decline of Steller sea lions Eumetopias jubatus in
+%   Alaska: a review of the nutritional stress hypothesis}.
+% \newblock \emph{Mammal Review} 33:3--28.
+%
+% \bibitem{Glazier:2009hq}
+% Glazier DS
+% \newblock (2009) {Metabolic level and size scaling of rates of respiration and
+%   growth in unicellular organisms}.
+% \newblock \emph{Funct. Ecol.} 23:963--968.
+%
+% \bibitem{Kooi2000}
+% Kooijman SALM
+% \newblock (2000) \emph{{Dynamic Energy and Mass Budgets in Biological Systems}}
+% \newblock (Cambridge).
+%
+% \bibitem{Sousa:2010ez}
+% Sousa T, Domingos T, Poggiale JC, Kooijman SALM
+% \newblock (2010) {Dynamic energy budget theory restores coherence in biology}.
+% \newblock \emph{Philos. T. Roy. Soc. B} 365:3413--3428.
+%
+% \bibitem{Diekmann:2010da}
+% Diekmann O, Metz JAJ
+% \newblock (2010) {How to lift a model for individual behaviour to the
+%   population level?}
+% \newblock \emph{Philos. T. Roy. Soc. B} 365:3523--3530.
+%
+% \bibitem{murdoch:2003}
+% Murdoch WW, Briggs CJ, Nisbet RM
+% \newblock (2003) \emph{{Consumer-resource Dynamics}}, Monographs in population
+%   biology
+% \newblock (Princeton University Press).
+%
+% \bibitem{Benichou:2014wu}
+% Benichou O, Redner S
+% \newblock (2014) {Depletion-Controlled Starvation of a Diffusing Forager}.
+% \newblock \emph{arXiv}.
+%
+% \bibitem{Benichou:2016wl}
+% B{\'e}nichou O, Chupeau M, Redner S
+% \newblock (2016) {Role of Depletion on the Dynamics of a Diffusing Forager}.
+%
+% \bibitem{Chupeau:2016jf}
+% Chupeau M, B{\'e}nichou O, Redner S
+% \newblock (2016) {Universality classes of foraging with resource renewal}.
+% \newblock \emph{Phys. Rev. E} 93:032403.
+%
+% \bibitem{murray2011mathematical}
+% Murray JD
+% \newblock (2011) \emph{{Mathematical Biology: I. An Introduction}},
+%   Interdisciplinary Applied Mathematics
+% \newblock (Springer New York, Melaka Manipal Medical College (Manipal Campus),
+%   International Centre for Health Sciences, Madhav Nagar, Manipal, Udupi
+%   District, Karnataka State, India. nayaksathish@yahoo.com) Vol.{} 110.
+%
+% \bibitem{Strogatz:2008wo}
+% Strogatz SH
+% \newblock (2008) \emph{{Nonlinear Dynamics and Chaos}}, With Applications to
+%   Physics, Biology, Chemistry, and Engineering
+% \newblock (Westview Press).
+%
+% \bibitem{GuckHolmes}
+% Guckenheimer J, Holmes P
+% \newblock (1983) \emph{{Nonlinear oscillations, dynamical systems, and
+%   bifurcations of vector fields}}
+% \newblock (Springer, New York).
+%
+% \bibitem{Gross:2004p2428}
+% Gross T, Feudel U
+% \newblock (2004) {Analytical search for bifurcation surfaces in parameter
+%   space}.
+% \newblock \emph{Physica D} 195:292--302.
+%
+% \bibitem{Hastings:2001jh}
+% Hastings A
+% \newblock (2001) {Transient dynamics and persistence of ecological systems}.
+% \newblock \emph{Ecol. Lett.} 4:215--220.
+%
+% \bibitem{Neubert:1997wk}
+% Neubert M, Caswell H
+% \newblock (1997) {Alternatives to resilience for measuring the responses of
+%   ecological systems to perturbations}.
+% \newblock \emph{Ecology} 78:653--665.
+%
+% \bibitem{Caswell:2005eo}
+% Caswell H, Neubert MG
+% \newblock (2005) {Reactivity and transient dynamics of discrete-time ecological
+%   systems}.
+% \newblock \emph{Journal of Difference Equations and Applications} 11:295--310.
+%
+% \bibitem{Neubert:2009td}
+% Neubert M, Caswell H
+% \newblock (2009) {Detecting reactivity}.
+% \newblock \emph{Ecology}.
+%
+% \bibitem{Yodzis:1992hg}
+% Yodzis P, Innes S
+% \newblock (1992) {Body Size and Consumer-Resource Dynamics}.
+% \newblock \emph{Am. Nat.} 139:1151--1175.
+%
+% \bibitem{Brown:2004wq}
+% Brown J, Gillooly J, Allen A, Savage V, West G
+% \newblock (2004) {Toward a metabolic theory of ecology}.
+% \newblock \emph{Ecology} 85:1771--1789.
+%
+% \bibitem{West:2002it}
+% West GB, Woodruff WH, Brown JH
+% \newblock (2002) {Allometric scaling of metabolic rate from molecules and
+%   mitochondria to cells and mammals.}
+% \newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.
+%
+% \bibitem{DeLong:2010dy}
+% DeLong JP, Okie JG, Moses ME, Sibly RM, Brown JH
+% \newblock (2010) {Shifts in metabolic scaling, production, and efficiency
+%   across major evolutionary transitions of life.}
+% \newblock \emph{PNAS} 107:12941--12945.
+%
+% \bibitem{West:2001bv}
+% West GB, Brown JH, Enquist BJ
+% \newblock (2001) {A general model for ontogenetic growth}.
+% \newblock \emph{Nature} 413:628--631.
+%
+% \bibitem{moses2008rmo}
+% Moses ME, {et~al.}
+% \newblock (2008) {Revisiting a Model of Ontogenetic Growth: Estimating Model
+%   Parameters from Theory and Data}.
+% \newblock \emph{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}
+%   171:632--645.
+%
+% \bibitem{gillooly2002esa}
+% Gillooly JF, Charnov EL, West GB, Savage VM, Brown JH
+% \newblock (2002) {Effects of size and temperature on developmental time}.
+% \newblock \emph{Nature} 417:70--73.
+%
+% \bibitem{hou}
+% Hou C, {et~al.}
+% \newblock (2008) {Energy Uptake and Allocation During Ontogeny}.
+% \newblock \emph{Science} 322:736--739.
+%
+% \bibitem{bettencourt}
+% Bettencourt LMA, Lobo J, Helbing D, Kuhnert C, West GB
+% \newblock (2007) {Growth, innovation, scaling, and the pace of life in cities}.
+% \newblock \emph{Proc. Natl. Acad. Sci. USA} 104:7301--7306.
+%
+% \bibitem{Folland:2008ij}
+% Folland JP, Mc~Cauley TM, Williams AG
+% \newblock (2008) {Allometric scaling of strength measurements to body size}.
+% \newblock \emph{Eur J Appl Physiol} 102:739--745.
+%
+% \bibitem{CalderIII:1983jd}
+% Calder~III WA
+% \newblock (1983) {An allometric approach to population cycles of mammals}.
+% \newblock \emph{J. Theor. Biol.} 100:275--282.
+%
+% \bibitem{Peterson:1984hj}
+% Peterson RO, PAGE RE, DODGE KM
+% \newblock (1984) {Wolves, Moose, and the Allometry of Population Cycles}.
+% \newblock \emph{Science} 224:1350--1352.
+%
+% \bibitem{Krukonis:1991fk}
+% Krukonis G, Schaffer WM
+% \newblock (1991) {Population cycles in mammals and birds: Does periodicity
+%   scale with body size?}
+% \newblock \emph{J. Theor. Biol.} 148:469--493.
+%
+% \bibitem{Hendriks:2012fc}
+% Hendriks AJ, Mulder C
+% \newblock (2012) {Delayed logistic and Rosenzweig{\textendash}MacArthur models
+%   with allometric parameter setting estimate population cycles at lower trophic
+%   levels well}.
+% \newblock \emph{Ecological Complexity} 9:43--54.
+%
+% \bibitem{Kendall:1999iy}
+% Kendall BE, {et~al.}
+% \newblock (1999) {Why do populations cycle? A synthesis of statistical and
+%   mechanistic modeling approaches}.
+% \newblock \emph{Ecology} 80:1789--1805.
+%
+% \bibitem{Hogstedt:2005cr}
+% H{\"o}gstedt G, Seldal T, Breistol A
+% \newblock (2005) {Period length in cyclic animal populations}.
+% \newblock \emph{Ecology} 86:373--378.
+%
+% \bibitem{Kendall:1998hl}
+% Kendall, Prendergast, Bjornstad
+% \newblock (1998) {The macroecology of population dynamics: taxonomic and
+%   biogeographic patterns in population cycles}.
+% \newblock \emph{Ecol. Lett.} 1:160--164.
+%
+% \bibitem{Brown:1993p708}
+% Brown J, Marquet P, Taper M
+% \newblock (1993) {Evolution of body size: consequences of an energetic
+%   definition of fitness}.
+% \newblock \emph{Am. Nat.} 142:573--584.
+%
+% \bibitem{West:1997cg}
+% West GB, Brown JH, Enquist BJ
+% \newblock (1997) {A General Model for the Origin of Allometric Scaling Laws in
+%   Biology}.
+% \newblock \emph{Science} 276:122--126.
+%
+% \bibitem{West:2002ud}
+% West GB, Woodruff WH, Brown JH
+% \newblock (2002) {Allometric scaling of metabolic rate from molecules and
+%   mitochondria to cells and mammals}.
+% \newblock \emph{Proc. Natl. Acad. Sci. USA} 99:2473--2478.
+%
+% \bibitem{Millar:1990p923}
+% Millar J, Hickling G
+% \newblock (1990) { Fasting Endurance and the Evolution of Mammalian Body Size}.
+% \newblock \emph{Funct. Ecol.} 4:5--12.
+%
+% \bibitem{Alroy:1998p1594}
+% Alroy J
+% \newblock (1998) {Cope's rule and the dynamics of body mass evolution in North
+%   American fossil mammals}.
+% \newblock \emph{Science} 280:731.
+%
+% \bibitem{Clauset:2009fh}
+% Clauset A, Redner S
+% \newblock (2009) {Evolutionary Model of Species Body Mass Diversification}.
+% \newblock \emph{Phys. Rev. Lett.} 102:038103.
+%
+% \bibitem{Smith:2010p3442}
+% Smith F, Boyer A, Brown J, Costa D
+% \newblock (2010) {The Evolution of Maximum Body Size of Terrestrial Mammals}.
+% \newblock \emph{Science}.
+%
+% \bibitem{Saarinen:2014br}
+% Saarinen JJ, {et~al.}
+% \newblock (2014) {Patterns of maximum body size evolution in Cenozoic land
+%   mammals: eco-evolutionary processes and abiotic forcing}.
+% \newblock \emph{Proc Biol Sci} 281:20132049--20132049.
+%
+% \bibitem{allen2002}
+% Allen AP, Brown JH, Gillooly JF
+% \newblock (2002) {Global Biodiversity, Biochemical Kinetics, and the
+%   Energetic-Equivalence Rule}.
+% \newblock \emph{Science} 297:1545--1548.
+%
+% \bibitem{enquist1998}
+% Enquist BJ, Brown JH, West GB
+% \newblock (1998) {Allometric scaling of plant energetics and population density
+%   : Abstract : Nature}.
+% \newblock \emph{Nature} 395:163--165.
+%
+% \bibitem{tilman1981}
+% Tilman D
+% \newblock (1981) {Tests of Resource Competition Theory Using Four Species of
+%   Lake Michigan Algae}.
+% \newblock \emph{Ecology} 62:802--815.
+%
+% \bibitem{dutkiewicz2009}
+% Dutkiewicz S, Follows MJ, Bragg JG
+% \newblock (2009) {Modeling the coupling of ocean ecology and biogeochemistry}.
+% \newblock \emph{Global Biogeochem. Cycles} 23:n/a--n/a.
+%
+% \bibitem{barton2010}
+% Barton AD, Dutkiewicz S, Flierl G, Bragg J, Follows MJ
+% \newblock (2010) {Patterns of Diversity in Marine Phytoplankton}.
+% \newblock \emph{Science} 327:1509--1511.
+%
+% \end{thebibliography}
 
@@
 \centering
-\includegraphics[width=0.75\textwidth]{fig_FixedPoint.pdf}
+\includegraphics[width=0.5\textwidth]{fig_FixedPoint.pdf}
 \caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
@@
   allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
-  a function of different mammalian body sizes: $M=A\times10^1$g (blue) and
-  $M=A\times10^6$g (green), where $A$ is a random uniform variable in $[1,9]$.
+  a function of different mammalian body sizes: $M=P\times10^2$g (blue) and
+  $M=P\times10^6$g (green), where $P$ is a random uniform variable in $[1,9]$.
   Points denote realized values of $\sigma$ and $\rho$ given the drawn values for $M$.
@@
   starvation threshold and is thus infeasible.
-  Arrows point to the predicted optimal mass $M_{\rm opt}=8.43\times 10^6$, which serves as the uninvadable, evolutionary stable state.}
+  Arrows point to the predicted optimal mass $M_{\rm opt}=1.73\times 10^7$, which serves as an evolutionary attractor for body mass.
+  The black wedge points to the largest body mass known for terrestrial mammals at $7.4\times10^7$g~\cite{Smith:2010p3442}.}
 \label{fig:invasion}

### 2017-02-28 Manuscript/Revision_2017/Revised-Supplement-1.tex
letter and chris ms

--- 2016-08-24 Manuscript/Supplement.tex
@@
 
+\section*{Mechanisms of Starvation and Recovery}
+Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
+For the dynamics of growth, reproduction, and resource consumption past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and in turn dictates resource consumption (see Refs.
+\cite{kempes,kempes2} for a brief review of this perspective).
+This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis \cite{kempes,ALLOGM} and predictions of the steady-state resource abundance in communities of cells \cite{}.
+Here we can leverage these mechanisms with several additional concepts to define our nutritional state model.
+
+We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
+\begin{align} 
+\begin{split}
+\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
+\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\ 
+&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right]
+\end{split}
+\end{align}
+where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the nondimensional form which is what is later derived and presented in the main text).
+%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
+In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery functions as functions of the current resource availability.
+Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively (each which scale with body size as discussed later), and $S\left(R_{d}\right)$ is a saturating function of resources, for example a Michealis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
+A simplified version of the Michealis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
+In these equations $Y$ represents the yield coefficient \cite{} which is the quantity of resources required to build a unit of organism (e.g.
+gram of mammal produced per gram of grass consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
+If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP) a natural choice for $\dot{R_{d}}$.
+This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
+In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$ and $C$ is the carrying capacity or maximum density of $R_{d}$ in a particular environment.
+
+Before describing the values of each of these constants and a general nondimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations as this leads to a simplification.
+As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$ and a constant value for $R_{d}>>k$.
+Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$ such that the above relationships reduce to
+\begin{align} 
+\begin{split}
+\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
+\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
+& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
+\end{split}
+\end{align}
+%where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
+
+We can formally nondimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes 
+%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
+\begin{align} 
+\begin{split}
+\dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
+\dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
+\dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
+\end{split}
+\end{align}
+If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with 
+\begin{align} 
+\begin{split}
+\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
+\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
+\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
+\end{split}
+\end{align}
+where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$.
+The above equations represent the system of equations presented in the main text.
+
 \section*{Parameter Values and Estimates}
 
-Many of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
-Here we outline previous measurements and simple estimates of the parameters.
+%\subsection*{Standard synthesis and metabolic parameters}
 
-\subsection*{Standard synthesis and metabolic parameters}
+All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
+Below we outline previous measurements and simple estimates of the parameters.
 
@@
 
-The final parameters that we must consider connect the resource growth rate to the total metabolic rate of an organism.
+We also connect the resource growth rate to the total metabolic rate of an organism.
 That is, we are interested in the relative rates of resource recovery and consumption by the total population.
@@
 This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.
+
+Typically the value of $\xi=C/k$ should roughly be $2$.
+The value of $\rho$, $\lambda$, $\sigma$, and $\mu$ are all simple rates (note that we have not rescaled time in our nondimensionalization) as defined in the maintext.
+Given that our model considers transitions over entire stages of ontogeny or nutritional states the value of $Y$ must represent yields integrated over entire life stages.
+Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \cite{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
+\begin{equation}
+B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+\end{equation}
+Similarly, the maintenance for hungry individuals $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
+\begin{equation}
+B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+\end{equation}
+which, in combination, allows us to calculate $\delta$ and $\beta$.
+
+Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities.
+This has a range of $Value$ to $Value$ globally.
+It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter such that these steady states match the scale of know data from Damuth \cite{}.
+We find that the best fit is $\alpha=VALUE\times10^{-9}$ (s$^{-1}$).
+However, two points are important to note here: first, our overall framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, the asymptotic behavior of $F^{*}$ and $H^{*}$ is also independent of $\alpha$, that is, our prediction of the maximum mammal size does not depend on $\alpha$.
 

### 2017-03-02 Manuscript/Revision_2017/15.tex
eigenvalues

@@
 The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity $C$.
-{\color{red}  The rate at which consumers transition between states and consume resources is dependent on their overall abundance, the abundance of resources, the efficiency of converting resources into metabolism, and how that metabolism is partitioned between maintenance and growth purposes.
-In the supplement we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple nondimensional form which we have described below (please see the supplement for a detailed derivation).
-
-Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$.
-Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a ratio between maximal resource consumption and the carrying capacity of the resource (please see the supplement).
+The rate at which consumers transition between states and consume resources is dependent on their overall abundance, the abundance of resources, the efficiency of converting resources into metabolism, and how that metabolism is partitioned between maintenance and growth purposes.
+In the supplement we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we have described below (please see Supporting Information for a detailed derivation).
+
+Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$  (see Supporting Information for detailed a description and derivation of all rates).
+Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a ratio between maximal resource consumption and the carrying capacity of the resource.
 %and $\approx 2$.
-Resources are eaten by the consumers at rate $\rho R + \delta$, which accounts for somatic growth and maintenance for starving consumers, respectively, and at rate $\beta$ by full consumers which is a constant representing maximal maintenance and somatic growth (see supplement).}
+Resources are eaten by the consumers at rate $\rho R + \delta$, which accounts for somatic growth and maintenance for starving consumers, respectively, and at rate $\beta$ by full consumers which is a constant representing maximal maintenance and somatic growth.
 %This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
@@
 Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
-{\color{red} It should be noted that this prediction of an asymptotic limit on mammalian size resembles efforts on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here \cite{Kempes:2012hy,Kempes:2016}.}
+We note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\cite{Kempes:2012hy,Kempes:2016}.
+% reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
 
@@
   Arrows point to the predicted optimal mass $M_{\rm opt}=1.73\times 10^7$, which serves as an evolutionary attractor for body mass.
-  The black wedge points to the largest body mass known for terrestrial mammals at $7.4\times10^7$g~\cite{Smith:2010p3442}.}
+  The black wedge points to the largest body mass known for terrestrial mammals at $7.74\times10^7$g~\cite{Smith:2010p3442}.}
 \label{fig:invasion}

### 2017-03-02 Manuscript/Revision_2017/16.tex
eigenvalues

--- 2017-03-02 Manuscript/Revision_2017/15.tex
@@
 \end{eqnarray}
-where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically as $9.7M^{0.92}$ {\bf cite?}.
+where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically (see supplement).
 We now use this solution to define the timescale of reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \cite{moses2008rmo} for a detailed presentation of these timescales).
@@
 The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
-Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis \cite{allen2002,enquist1998}.
+Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{damuth1987interspecific,allen2002,enquist1998}.
 The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size (e.g.
-\cite{allen2002,enquist1998}).
+\cite{damuth1987interspecific,allen2002,enquist1998}).
 %Previous studies have focused on t
@@
 \cite{allen2002,enquist1998}).
-This is usually stated as $B_{\rm tot}=N^*B(M)=Q$ where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{allen2002,enquist1998}.
+This is usually stated as $B_{\rm tot}=N^*B(M)=Q$ where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{damuth1987interspecific,allen2002,enquist1998}.
 Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
@@
   (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.}
+  (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.
+The data are from Damuth \cite{damuth1987interspecific} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see supplement and Refs.
+\cite{West:2001bv,hou,moses2008rmo}).}
 \label{fig:mass}

### 2017-03-02 Manuscript/Revision_2017/Revised-Supplement-2.tex
eigenvalues

--- 2017-02-28 Manuscript/Revision_2017/Revised-Supplement-1.tex
@@
 \usepackage[left]{lineno}
+\usepackage[comma,sort&compress]{natbib}
 
@@
 For the dynamics of growth, reproduction, and resource consumption past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and in turn dictates resource consumption (see Refs.
-\cite{kempes,kempes2} for a brief review of this perspective).
-This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis \cite{kempes,ALLOGM} and predictions of the steady-state resource abundance in communities of cells \cite{}.
+\cite{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
+This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g.
+\cite{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions of the steady-state resource abundance in communities of cells \cite{kempes2014morphological}.
 Here we can leverage these mechanisms with several additional concepts to define our nutritional state model.
@@
 In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery functions as functions of the current resource availability.
-Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively (each which scale with body size as discussed later), and $S\left(R_{d}\right)$ is a saturating function of resources, for example a Michealis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
-A simplified version of the Michealis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
-In these equations $Y$ represents the yield coefficient \cite{} which is the quantity of resources required to build a unit of organism (e.g.
+Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively (each which scale with body size as discussed later), and $S\left(R_{d}\right)$ is a saturating function of resources, for example a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
+A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
+In these equations $Y$ represents the yield coefficient (e.g.
+\cite{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (e.g.
 gram of mammal produced per gram of grass consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
@@
 
+An important feature of our is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction.
+This starting size is known to follow an allometric relationship with adult mass of the form ...
+where estimates for ...
+range between ...
+\cite{}.
+We choose ....
+
 The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g.
@@
 This has a range of $Value$ to $Value$ globally.
-It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter such that these steady states match the scale of know data from Damuth \cite{}.
-We find that the best fit is $\alpha=VALUE\times10^{-9}$ (s$^{-1}$).
+It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter such that these steady states match the scale of know data from Damuth \cite{damuth1987interspecific}.
+We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$).
 However, two points are important to note here: first, our overall framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, the asymptotic behavior of $F^{*}$ and $H^{*}$ is also independent of $\alpha$, that is, our prediction of the maximum mammal size does not depend on $\alpha$.
@@
 \subsection*{Rate equations for invaders with modified body mass}
-If an invading subset of the resident population of mass $M$ has an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $[-1,1]$ ($\chi<0$ denotes a leaner invader; $\chi > 0$ denotes an invader with more endogenous reserves), the invading population will have the following modified rates: $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$.
-Because we are assuming that the invading population is only modifying its endogenous energetic stores, we assume that the proportion of body mass that is non-adipose tissue remains the same as the resident population.
-This assumption leads to the following modified timescales:
+We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $chi_{\rm min} <0$ and $chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional endogenous reserves.
+Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
+Thus $\chi_{\rm}$ is limited by the proportion of lean mass, such that $\chi_{\rm min} = -f_0M^{\gamma-1}$.
+Similarly, we assume that the invading organisms do not add endogenous reserves above the asymptyotic mass of the species $M$, such that $(1+\chi)\epsilon_\lambda M < M$, and $\chi_{\rm max} \approx 0.05$.
+
+Although the starved state of invading organisms remains unchanged, the rate of starvation from the modified full state to the starved state, the rate of recovery from the starved state to the modified full state, and the maintenance rates of both, will be different, such that $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$, $\delta^\prime = \delta(M^\prime)$.
+Rates of starvation and recovery for the invading population are easily derived by adjusting the starting or ending state before and after starvation and recovery, leading to the following timescales:
 
@@
 t_{\sigma^\prime} &= \frac{-M^{1/4}}{B_0/E_m^\prime}\log \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
-t_{\rho^\prime} &= \frac{-4 M^{1/4} }{B_0/E_m^\prime}\log \left(\frac{1-( \epsilon_\lambda(\chi +1))^{1/4}}{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}\right), \\ \nonumber
-t_{\beta^\prime} &= \xi B_0\left(M(\chi + 1)\right)^{3/4}.
+t_{\rho^\prime} &= \frac{-4 M^{1/4} }{B_0/E_m^\prime}\log \left(\frac{1-( \epsilon_\lambda(\chi +1))^{1/4}}{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}\right).
+\end{align}
+
+The maintenance rates for the invading population require more careful consideration.
+First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of endogenous reserves.
+Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \cite{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
+\begin{equation}
+B^{\prime}_{\lambda}=\int_{0}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
+\end{equation}
+where
+\begin{equation}
+t_{\lambda^\prime} = \frac{-4 M^{1/4} }{B_0/E_m}\log \left(\frac{1-(m_0/M)^{1/4}}{(1-\epsilon_\lambda (1+\chi))^{1/4}} \right).
+\end{equation}
+Note that we do not use this timescale to determine the reproductive rate of the invading consumer---which is assumed to remain the same as the resident population---but only to calulate the lifetime energy use.
+Similarly, the maintenance for hungry individuals $P^\prime_{H}=B_{0}(\epsilon_{\sigma}(1+\chi)M)^{3/4}/(\epsilon_{\sigma}(1+\chi)M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y^\prime_{H}=(1+\chi)ME_{d}/B^{\prime}_{\rho}$ where
+\begin{equation}
+B^{\prime}_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
+\end{equation}
+Finally, we can calculate the maintenance of the invaders as
+
+\begin{align}
+  \delta^\prime &= P^\prime_{H}Y^\prime_{H}/\xi \\ \nonumber
+  \beta^\prime &= \left(\frac{\lambda_{\rm max}}{Y^\prime_{F}}+P^\prime_{F} \right)Y^\prime_{H}/\xi
 \end{align}

### 2017-03-07 Manuscript/Revision_2017/17.tex
penultimate draft and figure updates

--- 2017-03-02 Manuscript/Revision_2017/16.tex
@@
 To elucidate the consequences of this tradeoff, we introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete consumers that reproduce, and undernourished, non-reproducing consumers that are susceptible to mortality.
-As a function of the transition rates between these replete and undernourished states that are determined by the presence or absence of resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
-We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships and find that population dynamics subject to these constraints can approach the cyclic regime but are typically driven to a steady state.
+As a function of the transition rates between these two states that are determined by the abundance of resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
+We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships between body size and a variety of traits and find that population dynamics subject to these constraints are typically driven to a steady state.
 Moreover, we find that these rates fall within a `refuge' in parameter space, where the probability of extinction of the consumer population is minimized.
 Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations.
-Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percent body fat.
+Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percentage of body fat.
 For body masses $\lesssim 10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $\gtrsim 10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
@@
 \noindent {\small {\bf Significance Statement} %105/120 words
-Energetic investment in somatic maintenance and growth vs. reproduction directly impacts the dynamics of populations among species.
+Energetic investment in somatic maintenance and growth vs$.$ reproduction directly impacts the dynamics of populations among species.
 Here, we construct a Nutritional State-structured Model (NSM) to assess the population-level effects of starvation and recovery of a consumer population in a resource-limited environment, and use allometric scaling relationships for mammals to establish all timescales and rates.
-Our model reveals that mammalian energetic rates minimize the probability of stochastic extinction, establishes dynamic bounds on mammalian body size while providing independent theoretical support for the energy equivalence hypothesis, and provides a mechanistic driver for the evolutionary trend towards larger body size known as Cope's rule.}
+Our model: i.
+reveals that mammalian energetic rates minimize the probability of stochastic extinction, ii.
+establishes dynamic bounds on mammalian body size while providing independent theoretical support for the energy equivalence hypothesis, and iii.
+provides a mechanistic driver for the evolutionary trend towards larger body size known as Cope's rule.}
 
@@
 The rate at which consumers transition between states and consume resources is dependent on their overall abundance, the abundance of resources, the efficiency of converting resources into metabolism, and how that metabolism is partitioned between maintenance and growth purposes.
-In the supplement we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we have described below (please see Supporting Information for a detailed derivation).
-
-Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$  (see Supporting Information for detailed a description and derivation of all rates).
+In the supplementary information (SI) we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we describe below.
+
+{\color{red} Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$  (see SI for detailed a description and derivation of all rates and dependencies).
 Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a ratio between maximal resource consumption and the carrying capacity of the resource.
 %and $\approx 2$.
-Resources are eaten by the consumers at rate $\rho R + \delta$, which accounts for somatic growth and maintenance for starving consumers, respectively, and at rate $\beta$ by full consumers which is a constant representing maximal maintenance and somatic growth.
+Resources are eaten by the hungry consumers at rate $\rho R + \delta$, that accounts for their somatic growth ($\rho R$) and  maintenance ($\delta$).
+Full consumers eat resources at a constant rate $\beta$ that accounts for maximal maintenance and somatic growth (please see SI for a mechanistic derivation of these terms from resource energetics).}
 %This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
-The NSM represents a fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\cite{Persson:1998hz}.
+The NSM represents an ecologically motivated fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\cite{Persson:1998hz}.
 
@@
 \begin{split}
-\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
-\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
+\frac{dF}{dt} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
+\frac{dH}{dt} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
+\frac{dR}{dt} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
 \end{split}
 \end{align}
-This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see Supplementary Materials for a full derivation and the dimensional form).
-Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
+This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see SI for a full derivation and the dimensional form).
+Notice that the total consumer density $F+H$ evolves according to $\frac{dF}{dt}+\frac{dH}{dt}=\lambda F-\mu H$.
 This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\cite{murray2011mathematical}, except that the resource density does not appear in the growth term.
@@
 In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
-The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates the physical regime from the unphysical regime where all densities become $<0$.
+The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates a physical  from an unphysical regime where all densities become less than zero.
 The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
@@
 This Hopf bifurcation is defined by ${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial of the Jacobian matrix~\cite{Gross:2004p2428}.
-As the system parameters are tuned to be within the stable regime but close to the Hopf bifurcation, the amplitude of the transient but decaying cycles become large.
+As the system parameters are tuned to be within the stable regime, but close to the Hopf bifurcation, the amplitude of the transient cycles become large.
 Given that ecological systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase the extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
@@
 size and species diversity. Many of these relations can be derived from a
-small set of assumptions and below we describe a framework to determine the
+small set of assumptions and below we describe our framework to determine the
 covariation of timescales and rates across the range of mammals for each of
 the key parameters of our model (cf.~\cite{Yodzis:1992hg}).  We are thereby
-able to define the regime of dynamics occupied by the entire class of mammals
-along with the key differences between the largest and smallest mammals.
+able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.
 
@@
 \label{balance}
-B_{0}m^{\eta}=E_{m}\dot{m}+B_{m}m\,,
+B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
 \end{eqnarray}
@@
 \end{eqnarray}
-where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically (see supplement).
-We now use this solution to define the timescale of reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \cite{moses2008rmo} for a detailed presentation of these timescales).
+where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically (see SI).
+We now use this solution to define the timescale for reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \cite{moses2008rmo} for a detailed presentation of these timescales).
 The time that it takes to reach a particular mass $\epsilon M$ is given by the timescale
@@
 For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$ \cite{West:2001bv}).
-The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$ this will scale like $\lambda\propto M^{\eta-1}$ for $M>>m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
+The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$, this rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
 
@@
 and recovery by the timescales required to reach, or return from, specific
-fractions of the replete-state mass (Fig. \ref{fig:gvs}; see Supporting Information, Table I for parameterizations).  We define
+fractions of the replete-state mass (Fig. \ref{fig:gvs}; see SI, Table I for parameterizations).  We define
 $m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction of
-replete-state mass where reproduction ceases. This fraction will be modified
+replete-state mass where reproduction ceases. This fraction will deviate from a constant
 if tissue composition systematically scales with adult mass.  For example,
@@
 \begin{eqnarray}
+\label{rhotimescale}
 t_{\rho}=\ln\left[\frac{1-\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}
 \end{eqnarray}
-where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in the biosynthetic energetics during recovery (see Supporting Information).
+where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in the biosynthetic energetics during recovery (see SI).
 It should be noted that more complicated ontogenetic models explicitly handle
@@
 reduced-mass hungry state where reproduction is impossible.
-For starving individuals we assume that an organism must meet its maintenance requirements using the digestion of existing mass as the sole energy source.
+For starving individuals we assume that an organism must meet its maintenance requirements by using the digestion of existing mass as the sole energy source.
 This assumption implies the following simple metabolic balance
 \begin{eqnarray}
-\dot{m}E_{m}^{\prime}=-B_{m}m
+\frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
 \end{eqnarray}
@@
 \begin{eqnarray}
-\dot{m}=-\frac{a^{\prime}}{M^{1-\eta}}m
+\frac{dm}{dt}=-\frac{a^{\prime}}{M^{1-\eta}}m
 \end{eqnarray}
 where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
-body mass which differs from $E_{m}$, the energy required to
+body mass, which differs from $E_{m}$, the energy required to
 synthesis a unit of biomass \cite{hou}. Given the replete mass, $M$, of an organism, the
@@
 \end{eqnarray}
-The time scale for starvation is
+The timescale for starvation is
 given by the time it takes $m(t)$ to reach $\epsilon_{\sigma} M$, which gives
@@
 mass has been shown to be roughly proportional to body mass~\cite{Folland:2008ij} in
-mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant. The time to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or
+mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant. The time to go from starvation to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or
 \begin{eqnarray}
+\label{mutimescale}
 t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
@@
 As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
-Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small window of potential values (Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.
+Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values (Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.
 We thus find that the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
@@
 Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
-These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size, though starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation, suggesting that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
+These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
+However, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation.
+This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
 %Thus our NSM model predicts that population cycles should be less common for smaller species and more common for larger species, particularly in environments where resources are limiting.%, and
@@
 
-It should be noted that previous studies have used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
-However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
+It should be noted that previous studies used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
+However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
 %It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic~\cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
-Statistically significant support for the existence of population cycles among mammals is relatively rare though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
-However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms such that testing this particular result is not straightforward.\\ % in agreement with our predictions of more pronounced transient dynamics, given how close these points are to the Hopf bifurcation.On the other hand,
+Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
+However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.\\ % in agreement with our predictions of more pronounced transient dynamics, given how close these points are to the Hopf bifurcation.On the other hand,
 % \vspace{6mm}
@@
 Within our model, higher rates of starvation result in a larger flux of the population to the hungry state.
-In this state reproduction is absent, thus increasing the likelihood of extinction.
+In this state, reproduction is absent, thus increasing the likelihood of extinction.
 From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
-We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady state density on extinction risk as a function of $\sigma$ and $\rho$.
+We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
 To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^5$ and $t \leq 10^8$.
@@
 For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
-For high values of $\sigma$ and low values of $\rho$, higher extinction risk arises because of the decrease in the steady state consumer population density (Figs. \ref{fig:fp}\emph{B}, \ref{fig:ext}).
+For high values of $\sigma$ and low values of $\rho$, increased extinction risk arises because of the decrease in the steady-state consumer population density (Figs.
+\ref{fig:fp}\emph{B}, \ref{fig:ext}).
 This interplay creates an `extinction refuge', such that for a constrained range of $\sigma$ and $\rho$, extinction probabilities are minimized.
@@
 We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge (Fig. \ref{fig:ext}, white point).
-These values are close enough to the Hopf bifurcation to avoid low steady state densities, and far enough away to avoid large-amplitude transient cycles.
-The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
+These values are close enough to the Hopf bifurcation to avoid low steady-state densities, and far enough away to avoid large-amplitude transient cycles.
+The feature that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
 Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that serve to minimize extinction risk.
@@
 % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
+{\it To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.}\\
 
@@
 Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{damuth1987interspecific,allen2002,enquist1998}.
-The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size (e.g.
-\cite{damuth1987interspecific,allen2002,enquist1998}).
+The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size (e.g., \cite{damuth1987interspecific,allen2002,enquist1998}).
 %Previous studies have focused on t
-This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism (e.g.
-$N^*\propto M^{-3/4}/B_{0}$)  (e.g.
-\cite{allen2002,enquist1998}).
-This is usually stated as $B_{\rm tot}=N^*B(M)=Q$ where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{damuth1987interspecific,allen2002,enquist1998}.
-Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale like $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
-This result if remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to the energy equivalence result.
-Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this is a hard limit where deviations outside of this range are energetically suboptimal.
+This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism ($N^*\propto M^{-3/4}/B_{0}$)  (e.g., \cite{allen2002,enquist1998}).
+This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{damuth1987interspecific,allen2002,enquist1998}.
+Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
+This result is remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to energy equivalence.
+Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
 %, violating the energy equivalence.
 % set by deviations from constant efficiency obeyed by other populations.
-In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, occurring at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
-This asymptotic behavior represents an upper bound on mammalian body size and occurs at $M_{\rm max}=6.5\times10^7$.
-Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
-We note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\cite{Kempes:2012hy,Kempes:2016}.
+In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
+{\color{red} This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero causing the timescales to become infinite (see Equation \ref{mutimescale}) and the rates $\mu$ and $\lambda$ to equal zero.
+The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$.
+This point predicts an upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.} Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
+{\color{red} It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters}.
+We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\cite{Kempes:2012hy,Kempes:2016}.
 % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
@@
 
-We contend that the NSM provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size as well as the observed trend towards larger body size over evolutionary time.
+We contend that the NSM provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time.
 The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
@@
 
-To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120\times$ the mass of a male African elephant.
-Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
-We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by ${}^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$ where $\chi \in [-f_0M^{\gamma-1},0.05]$ (where the lower bound of $\chi$ is the proportion of body mass that is not composed of metabolic reserves and the upper bound maintains the organism's reproductive mass), thus altering the rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
-Importantly, $\epsilon_\sigma$, which determines the point along the growth curve the defines the state of starved foragers, is assumed to remain unchanged for the invader population (see Supplemental Material for detailed derivations of invader rates).
+To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120$ times the mass of a male African elephant.
+Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.{\color{red}
+We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by $^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$.
+For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given Equation \ref{t1}, implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
+Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
+}The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
+Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the state of starved foragers, is assumed to remain unchanged for the invader population (see SI for detailed derivations of invader rates).
 %There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
@@
   function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-  bifurcation conditions separate parameter space into infeasible, cyclic,
-  and steady state dynamic regimes.  The color gradient shows the steady
+  bifurcation conditions separate parameter space into unphysical, cyclic,
+  and steady state dynamic regimes.  The colors show the steady
   state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
@@
   an energetically replete reproductive adult mass $m=\epsilon_\lambda M$ which we assume is 95\% asymptotic mass $M$.  Starvation follows the red
-  trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$, and recovery follows the
-  green growth trajectory to the replete adult mass. Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}
+  trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. Recovery follows the
+  green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires different energetics and a longer time to reach $\epsilon_\lambda M$.
+Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}
 \label{fig:growth}
@@
 \caption{
-Allometrically constrained starvation rate $\sigma$ (red) and recovery rate $\rho$ (green) relative to the reproductive rate $\lambda$ (orange) as a function of body mass.
+Allometrically constrained starvation rate $\sigma=1/t_{\sigma}$ (red), mortality $\mu=1/t_{\mu}$ (red) and recovery rate $\rho=1/t_{\rho}$ (green) relative to the reproductive rate $\lambda=1/t_{\lambda}$ (black) as a function of body mass (see Equations \ref{t1}, \ref{rhotimescale}, \ref{eq:sigma}, and \ref{mutimescale}).
 The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{fig_ExtinctionAllometric2.png}
-\caption{ Probability of extinction for a 100g consumer as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $A(F^*,H^*,R^*)$, with
-  $A$ being a random uniform variable in $[0,2]$.  Extinction is defined as the
-  population trajectory falling below $0.1~\times$ the allometrically
+\includegraphics[width=1\textwidth]{fig_ExtinctionAllometricComb3.png}
+\caption{ Probability of extinction for a 100g consumer as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, with $X$ being a random uniform variable in $[0,2]$.
+Extinction is defined as the
+  population trajectory falling below $0.1$ times the allometrically
   constrained steady state. The white point denotes the allometrically constrained starvation and recovery rate.}
@@
   (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.
-The data are from Damuth \cite{damuth1987interspecific} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see supplement and Refs.
+The data are from Damuth \cite{damuth1987interspecific} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
 \cite{West:2001bv,hou,moses2008rmo}).}

### 2017-03-07 Manuscript/Revision_2017/Revised-Supplement-2.tex
penultimate draft and figure updates

@@
 Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
-For the dynamics of growth, reproduction, and resource consumption past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and in turn dictates resource consumption (see Refs.
+For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
 \cite{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
 This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g.
-\cite{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions of the steady-state resource abundance in communities of cells \cite{kempes2014morphological}.
-Here we can leverage these mechanisms with several additional concepts to define our nutritional state model.
+\cite{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \cite{kempes2014morphological}.
+Here we leverage these mechanisms, combined with several additional concepts, to define our nutritional state model.
 
@@
 \end{align}
-where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the nondimensional form which is what is later derived and presented in the main text).
+where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the nondimensional form that was presented in the main text).
 %(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
-In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery functions as functions of the current resource availability.
-Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively (each which scale with body size as discussed later), and $S\left(R_{d}\right)$ is a saturating function of resources, for example a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
+In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
+Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
+The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
 A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
-In these equations $Y$ represents the yield coefficient (e.g.
-\cite{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (e.g.
-gram of mammal produced per gram of grass consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
-If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP) a natural choice for $\dot{R_{d}}$.
+
+In the above equations $Y$ represents the yield coefficient (e.g., \cite{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of grass consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
+If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
 This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
-In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$ and $C$ is the carrying capacity or maximum density of $R_{d}$ in a particular environment.
+In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment.
 
-Before describing the values of each of these constants and a general nondimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations as this leads to a simplification.
-As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$ and a constant value for $R_{d}>>k$.
-Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$ such that the above relationships reduce to
+Before describing the values of each of these constants, and a general nondimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
+As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
+Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
 \begin{align}
@@
 
-Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g.
-\cite{West:2001bv,moses2008rmo} and the supplement of \cite{hou}).
+Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., \cite{West:2001bv,moses2008rmo} and the supplement of \cite{hou}).
 We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \cite{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
-It is important to note the exponent, because it not only defines several scalings in our framework but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.
+The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.
 For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$) \cite{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
@@
 
-An important feature of our is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction.
-This starting size is known to follow an allometric relationship with adult mass of the form ...
-where estimates for ...
-range between ...
-\cite{}.
-We choose ....
+An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction.
+This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see \cite{peters1986ecological} for a review).
+We use $m_{0}=0.097M^{0.92}$ from \cite{blueweiss1978relationships} which encompasses the widest range of body sizes \cite{peters1986ecological}.
 
@@
 
-We also connect the resource growth rate to the total metabolic rate of an organism.
+%We also connect the resource growth rate to the total metabolic rate of an organism.
 That is, we are interested in the relative rates of resource recovery and consumption by the total population.
@@
 Given that our model considers transitions over entire stages of ontogeny or nutritional states the value of $Y$ must represent yields integrated over entire life stages.
-Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \cite{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
+Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \cite{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
 \begin{equation}
@@
 \end{equation}
-Similarly, the maintenance for hungry individuals $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
+Similarly, the maintenance for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
 \begin{equation}
@@
 \end{equation}
-which, in combination, allows us to calculate $\delta$ and $\beta$.
+Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.
 
 Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities.
-This has a range of $Value$ to $Value$ globally.
-It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter such that these steady states match the scale of know data from Damuth \cite{damuth1987interspecific}.
-We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$).
-However, two points are important to note here: first, our overall framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, the asymptotic behavior of $F^{*}$ and $H^{*}$ is also independent of $\alpha$, that is, our prediction of the maximum mammal size does not depend on $\alpha$.
+From the data in \cite{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
+It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \cite{damuth1987interspecific}.
+We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above.
+However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, the asymptotic behavior of $F^{*}$ and $H^{*}$ is also independent of $\alpha$, that is, our prediction of the maximum mammal size does not depend on $\alpha$.
 
@@
    $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \cite{moses2008rmo,West:2001bv,hou} \\
-   $E_{m}^{\prime}$ & $36,000$  & \cite{stryer,hou} \\
+   $E_{m}^{\prime}$ & $7000$  & \cite{stryer,hou} \\
    $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \cite{hou}  \\

### 2017-03-09 Manuscript/Revision_2017/17.tex
some ms changes

@@
 Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percentage of body fat.
-For body masses $\lesssim 10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $\gtrsim 10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
+For body masses $< 1.75\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.75\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
 \end{abstract}
@@
 
-{\color{red} Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$  (see SI for detailed a description and derivation of all rates and dependencies).
+Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$.
 Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a ratio between maximal resource consumption and the carrying capacity of the resource.
@@
 Resources are eaten by the hungry consumers at rate $\rho R + \delta$, that accounts for their somatic growth ($\rho R$) and  maintenance ($\delta$).
-Full consumers eat resources at a constant rate $\beta$ that accounts for maximal maintenance and somatic growth (please see SI for a mechanistic derivation of these terms from resource energetics).}
+Full consumers eat resources at a constant rate $\beta$ that accounts for maximal maintenance and somatic growth (see SI for mechanistic derivations of these rates from resource energetics).
 %This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
@@
 \frac{dH}{dt} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-\frac{dR}{dt} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
+\frac{dR}{dt} &= \alpha \left(1-R\right)R -\left(\rho R+\delta\right)H-\beta F
 \end{split}
@@
 \begin{split}
-F^* &= \frac{(\sigma-\lambda) \alpha  \lambda  \mu ^2  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
-H^* &= \frac{(\sigma-\lambda) \alpha  \lambda ^2 \mu  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
+F^* &= (\sigma-\lambda)\frac{ \alpha  \lambda  \mu ^2  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
+H^* &= (\sigma-\lambda)\frac{ \alpha  \lambda ^2 \mu  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
 R^* &= (\sigma - \lambda)\frac{\mu  }{A}.
@@
 
-From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that $R^*$ is positive.
-In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
-The condition $\sigma = \lambda$ represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates a physical  from an unphysical regime where all densities become less than zero.
+From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that the consumer and resource densities are positive.
+%In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
+The condition $\sigma = \lambda$ thus represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates a physical from an unphysical regime where all steady-state densities become negative after intersecting the trivial fixed point $(F^*,H^*,R^*)=(0,0,0)$.
 The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
@@
 This Hopf bifurcation is defined by ${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial of the Jacobian matrix~\cite{Gross:2004p2428}.
-As the system parameters are tuned to be within the stable regime, but close to the Hopf bifurcation, the amplitude of the transient cycles become large.
+As the system parameters are tuned to be within the stable regime, but close to the Hopf bifurcation, the amplitude of the transient cycles becomes large.
 Given that ecological systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase the extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
@@
 Because reproduction is inhibited, there is a low steady-state consumer density and a high steady-state resource density.
-However, if $\sigma/\lambda\to 1$ from above, the population is overloaded with energetically-replete (reproducing) individuals, thereby promoting oscillations between the consumer and resource densities (Fig.~\ref{fig:fp}).
-
-Whereas the relation between consumer growth rate $\lambda$ and the starvation rate $\sigma$ defines an absolute bound of biological feasibility---the TC bifurcation---$\sigma$ also determines the sensitivity of the consumer population to changes in resource density.
-When $\sigma\gg\lambda$, the steady-state population density is small, thereby increasing the risk of stochastic extinction.
-On the other hand, as $\sigma$ decreases, the system will ultimately be poised either near the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).
-If the recovery rate $\rho$ is sufficiently small, the TC bifurcation is reached and the resource eventually is eliminated.
-If $\rho$ exceeds a threshold value, cyclic dynamics will develop as the Hopf bifurcation is approached.
+However, if $\sigma/\lambda\to 1$ from above, the population is overloaded with energetically-replete (reproducing) individuals, thereby promoting transient oscillations between the consumer and resource densities (Fig.~\ref{fig:fp}).
+If the starvation rate is low enough that the Hopf bifurcation is crossed, these oscillations become stable over time.
+This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.
 \\
-
-%In the case of the 2-stage consumer model, the Jacobian evaluated at the internal steady state (denoted by $|_*$) is written
-%
-%\begin{equation}
-%\mathbf{J}|_* =
-%%\left(
-%%\begin{array}{ccc}
-%% -F^* m+(1-R^*) a -R^* a -H^* \rho  & -R^* \rho  & -m R^* \\
-%% -H^* \rho -F^* \sigma  & -\mu -R^* \rho  & (1-R^*) \sigma  \\
-%% H^* \rho +F^* \sigma  & R^* \rho  & \lambda -(1-R^*) \sigma  \\
-%%\end{array}
-%%\right).
-%\left(
-%\begin{array}{ccc}
-% -\frac{\lambda  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\mu  \rho  (\sigma -\lambda )}{\lambda  \rho +\mu  \sigma } & \frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
-% \frac{\lambda  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\mu  (\mu +\rho ) \sigma }{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \lambda  (\mu +\rho )}{m \mu +\lambda  \rho } \\
-% -\frac{m \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } & -\frac{\mu  \rho  (\sigma - \lambda )}{\lambda  \rho +\mu  \sigma } & -\frac{\alpha  \mu  (\sigma - \lambda)}{\lambda  \rho +\mu  \sigma } \\
-%\end{array}
-%\right).
-%\end{equation}
-
-%CHRIS - we need to define beta in this section...
+% Whereas the relation between consumer growth rate $\lambda$ and the starvation rate $\sigma$ defines an absolute bound of biological feasibility---the TC bifurcation---$\sigma$ also determines the sensitivity of the consumer population to changes in resource density.
+% When $\sigma\gg\lambda$, the steady-state population density is small, thereby increasing the risk of stochastic extinction.
+% On the other hand, as $\sigma$ decreases, the system will ultimately be poised either near the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).
+% If the recovery rate $\rho$ is sufficiently small, the TC bifurcation is reached and the resource eventually is eliminated.
+% If $\rho$ exceeds a threshold value, cyclic dynamics will develop as the Hopf bifurcation is approached.
 
@@
 covariation of timescales and rates across the range of mammals for each of
-the key parameters of our model (cf.~\cite{Yodzis:1992hg}).  We are thereby
+the key parameters of our model (cf. ref.~\citenum{Yodzis:1992hg}).  We are thereby
 able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.
@@
 $B = B_0 M^\eta$~\cite{West:2002it}, where the scaling exponent $\eta$ is
-typically close to $2/3$ or $3/4$ for metazoans (e.g., \cite{Brown:2004wq}),
+typically close to $2/3$ or $3/4$ for metazoans (e.g., ref. \citenum{Brown:2004wq}),
 and has taxonomic shifts for unicellular species between $\eta\approx 1$ in
@@
 where we will define values of $\epsilon$ to describe a set of rates within our model.
-For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$ \cite{West:2001bv}).
+For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$; \citenum{West:2001bv}).
 The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$, this rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
@@
 \label{rhotimescale}
-t_{\rho}=\ln\left[\frac{1-\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}
+t_{\rho}=\ln\left[\frac{1-\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)^{1-\eta}}{1-\epsilon_\lambda^{1-\eta}}\right]\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}
 \end{eqnarray}
@@
 %where $c=(a/b)^{1/(\eta-1)}$.
-Although the rate equations \eqref{eq:system} are general, here we focus on
-parameterizations for terrestrial-bound endotherms, specifically mammals,
-which range from a minimum of $M\approx1$g (the Etruscan shrew
-\emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the late Eocene
-to early Miocene Indricotheriinae).  Investigating other classes of organisms
-would simply involve altering the metabolic exponents and scalings associate
-with $\epsilon$. Moreover, we emphasize that our allometric equations
-describe mean relationships, and do not account for the (sometimes
-considerable) variance associated with individual species.
+Although the rate equations \eqref{eq:system} are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
+Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
+Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
 \\
@@
 In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
-{\color{red} This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero causing the timescales to become infinite (see Equation \ref{mutimescale}) and the rates $\mu$ and $\lambda$ to equal zero.
+This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero causing the timescales to become infinite (see Equation \ref{mutimescale}) and the rates $\mu$ and $\lambda$ to equal zero.
 The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$.
-This point predicts an upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.} Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
-{\color{red} It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters}.
+This point predicts a strong upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
+Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
+It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters.
 We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\cite{Kempes:2012hy,Kempes:2016}.
@@
 To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120$ times the mass of a male African elephant.
-Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.{\color{red}
+Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
 We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by $^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$.
@@
 Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
-}The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
+The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
 Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the state of starved foragers, is assumed to remain unchanged for the invader population (see SI for detailed derivations of invader rates).
@@
 To assess the susceptibility of the resident population to invasion, we determine which consumer has a lower steady-state resource density for a given value of $\chi$, again with the expectation that populations able to survive on lower resource densities have a competitive advantage \cite{tilman1981}.
-We find that for $M\leq 1.73\times10^7$g, having additional body fat ($\chi > 0$) results in lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
-However, for $M> 1.73\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
+We find that for $M\leq 1.75\times10^7$g, having additional body fat ($\chi > 0$) results in lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
+However, for $M> 1.75\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
 
@@
 
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.73\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
+The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.75\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
 This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca.
-$1.3\times10^7$g and the late Miocene \emph{Deinotherium} at ca.
+$1.5\times10^7$g and the late Miocene \emph{Deinotherium} at ca.
 $1.74\times10^7$g ~\cite{Smith:2010p3442}.
@@
 %This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
-While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM may provide a general within-linneage mechanism for the evolution of larger body size among terrestrial mammals.
+While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general within-lineage mechanism for the evolution of larger body size among terrestrial mammals.
 
@@
 The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
-Mean values $\pm 20\%$ variation are shown by the shaded region for each rate.
+Mean values $\pm 25\%$ variation are shown by the shaded region for each rate.
 }
@@
 \includegraphics[width=1\textwidth]{fig_ExtinctionAllometricComb3.png}
-\caption{ Probability of extinction for a 100g consumer as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, with $X$ being a random uniform variable in $[0,2]$.
-Extinction is defined as the
-  population trajectory falling below $0.1$ times the allometrically
-  constrained steady state. The white point denotes the allometrically constrained starvation and recovery rate.}
+\caption{ Probability of extinction for a consumer with (from left to right) $M=10^2,~10^4,~10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
+Note the change in scale for $M=10^4,~10^6$g.
+Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
+The white points denote the allometrically constrained starvation and recovery rate.}
 \label{fig:ext}
@@
   starvation threshold and is thus infeasible.
-  Arrows point to the predicted optimal mass $M_{\rm opt}=1.73\times 10^7$, which serves as an evolutionary attractor for body mass.
+  Arrows point to the predicted optimal mass $M_{\rm opt}=1.75\times 10^7$, which serves as an evolutionary attractor for body mass.
   The black wedge points to the largest body mass known for terrestrial mammals at $7.74\times10^7$g~\cite{Smith:2010p3442}.}

### 2017-03-20 Manuscript/Arxiv_Revision/YKR_Arxiv.tex
embedded figs

--- 2017-03-09 Manuscript/Revision_2017/17.tex
@@
-%\usepackage{scicite}
-%\usepackage{times}
-%\usepackage{graphicx}
-%\usepackage{verbatim}
-%\usepackage{subfigure}
-%\usepackage{tikz}
+\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
+%\usepackage[pdftex]{graphicx}
+
+\usepackage{amsmath,amsfonts,amssymb}
+\usepackage[english]{babel}
+\usepackage[latin1]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{color}
+\usepackage{float}
+\usepackage{verbatim}
+\usepackage{graphicx}
+\usepackage{bm}
+\usepackage{mathtools}
+\usepackage{stmaryrd}
+\usepackage{anyfontsize}
+
+%\usepackage{epstopdf}
 %\usepackage{array}
@@
 %\usepackage{multirow}
-%\usepackage{multicol}
+\usepackage{color}
 %\usepackage{multibox}
 %\usepackage{rotating}
-%\usepackage{layout}
-%\usepackage{epstopdf}
-%\usepackage{afterpage}
+%\usepackage{lineno}
 %\usepackage[left]{lineno}
-
-%% PNAStmpl.tex
-%% Template file to use for PNAS articles prepared in LaTeX
-%% Version: Apr 14, 2008
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BASIC CLASS FILE
-
-\documentclass{pnastwo}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL GRAPHICS STYLE FILE
-
-\usepackage[pdftex]{graphicx}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL POSTSCRIPT FONT FILES
-
-\usepackage{pnastwoF}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% ADDITIONAL OPTIONAL STYLE FILES
-
-%\usepackage{amssymb,amsfonts,amsmath}
-\usepackage{amsmath}
-\usepackage{amssymb}
-
-\usepackage[font=large]{caption}
-\usepackage{subfigure}
-\usepackage{epstopdf}
-\usepackage{graphicx}
-\usepackage{verbatim}
-\usepackage{array}
-\usepackage{tabularx}
-\usepackage{multirow}
-\usepackage{color}
-%\usepackage{multibox}
-\usepackage{rotating}
-\usepackage{color}
-%\usepackage{lineno}
-\usepackage[left]{lineno}
-\usepackage[comma,sort&compress]{natbib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL MACRO FILES
-
-\bibliographystyle{pnas}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\sid}[1]{\textcolor{red}{\bf [#1]}}
-
-%% Don't type in anything in the following section:
-%%%%%%%%%%%%
-%% For PNAS Only:
-\contributor{Submitted to Proceedings
-of the National Academy of Sciences of the United States of America}
-\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
-\copyrightyear{2011}
-\issuedate{Issue Date}
-\volume{Volume}
-\issuenumber{Issue Number}
-%%%%%%%%%%%%
+%\usepackage[comma,sort&compress]{natbib}
+%\usepackage{authblk}
+%\usepackage{multicol}
+
+%\bibliographystyle{ieeetr}
+
+\usepackage{bibunits}
+
+%\linenumbers
+%\setlength\linenumbersep{3pt}
 
@@
 
-\linenumbers
-\setlength\linenumbersep{3pt}
-
-%\title{Ecological and evolutionary implications of starvation and body size}
-
-%\title{Survival of the Stable: Dynamic Consequences of Starvation and Reproduction}
-
-%\title{Size Constraints on Starvation \& Reproduction Dynamics}
-
-%\title{Eco-evolutionary Consequences of Starvation and Reproduction: Survival
-%of the Stable}
+\author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
+  of California, Merced, Merced, CA 95340, USA}
+
+\author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
+  Park Road, Santa Fe, NM 87501, USA}
+
+\author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
+  Road, Santa Fe, NM 87501, USA}
 
 \title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
-
-\author
-{Justin D.
-Yeakel\affil{1}{School of Natural Science, University of California Merced, Merced, CA} \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}\affil{5}{To whom correspondence should be addressed: jdyeakel@gmail.com}, Christopher P.
-Kempes \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}, \and Sidney Redner \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{3}{Department of Physics, Boston University, Boston MA}\affil{4}{Contributed equally}
-}
-
-\contributor{Submitted to Proceedings of the National Academy of Sciences
-of the United States of America}
-
-\maketitle
-
-\begin{article}
 
@@
 Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percentage of body fat.
-For body masses $< 1.75\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.75\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
+For body masses $< 1.748\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.748\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
 \end{abstract}
-\keywords{foraging | starvation|reproduction}
-
-\noindent {\small {\bf Significance Statement} %105/120 words
-Energetic investment in somatic maintenance and growth vs$.$ reproduction directly impacts the dynamics of populations among species.
-Here, we construct a Nutritional State-structured Model (NSM) to assess the population-level effects of starvation and recovery of a consumer population in a resource-limited environment, and use allometric scaling relationships for mammals to establish all timescales and rates.
-Our model: i.
-reveals that mammalian energetic rates minimize the probability of stochastic extinction, ii.
-establishes dynamic bounds on mammalian body size while providing independent theoretical support for the energy equivalence hypothesis, and iii.
-provides a mechanistic driver for the evolutionary trend towards larger body size known as Cope's rule.}
-
-\vspace{2mm}
-
-\noindent {\bf Introduction} \\
+
+\maketitle
+
+\begin{bibunit}[unsrt]
+
+\section*{Introduction}
 The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly influences how they invest reserves in uncertain environments.
@@
 
-\noindent {\bf Nutritional state-structured model (NSM)}\\
+\section*{Nutritional state-structured model (NSM)}
 We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$.
@@
 The NSM represents an ecologically motivated fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\cite{Persson:1998hz}.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.45\textwidth]{fig_FixedPoint.png}
+\caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
+  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
+  bifurcation conditions separate parameter space into unphysical, cyclic,
+  and steady state dynamic regimes.  The colors show the steady
+  state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
+  replete consumers $F^*$, (warmer colors denote higher densities).  Steady
+  state densities for the energetically deficient consumers $H^*$ (not shown)
+  scale with those for $F^*$.  }
+\label{fig:fp}
+\end{figure}
 
@@
 From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that the consumer and resource densities are positive.
-%In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
 The condition $\sigma = \lambda$ thus represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates a physical from an unphysical regime where all steady-state densities become negative after intersecting the trivial fixed point $(F^*,H^*,R^*)=(0,0,0)$.
 The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
-As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges. %(Fig.~\ref{fig:gvs})
+As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges.
 
@@
 Given that ecological systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase the extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
-%Thus the distance of a system from the Hopf bifurcation provides a measure of its persistence.
 
@@
 This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.
-\\
-% Whereas the relation between consumer growth rate $\lambda$ and the starvation rate $\sigma$ defines an absolute bound of biological feasibility---the TC bifurcation---$\sigma$ also determines the sensitivity of the consumer population to changes in resource density.
-% When $\sigma\gg\lambda$, the steady-state population density is small, thereby increasing the risk of stochastic extinction.
-% On the other hand, as $\sigma$ decreases, the system will ultimately be poised either near the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).
-% If the recovery rate $\rho$ is sufficiently small, the TC bifurcation is reached and the resource eventually is eliminated.
-% If $\rho$ exceeds a threshold value, cyclic dynamics will develop as the Hopf bifurcation is approached.
-
-\noindent {\bf Role of allometry} \\
-%[Link Allometry stuff to our model - what does it provide to the story]
-%[Introduce biological and linked constraints]
-%[Allometries capture vast amounts of diversity via a single parameter --- body size]
-%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
+
+\section*{Role of allometry}
 While there are no a priori constraints on the parameters in the NSM, most
@@
 covariation of timescales and rates across the range of mammals for each of
-the key parameters of our model (cf. ref.~\citenum{Yodzis:1992hg}).  We are thereby
+the key parameters of our model (cf. ref.~\cite{Yodzis:1992hg}).  We are thereby
 able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.
@@
 $B = B_0 M^\eta$~\cite{West:2002it}, where the scaling exponent $\eta$ is
-typically close to $2/3$ or $3/4$ for metazoans (e.g., ref. \citenum{Brown:2004wq}),
+typically close to $2/3$ or $3/4$ for metazoans (e.g., ref. \cite{Brown:2004wq}),
 and has taxonomic shifts for unicellular species between $\eta\approx 1$ in
@@
 \cite{DeLong:2010dy,Kempes:2012hy}.
-%Justin adding an explanation for beta
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.pdf}
+\caption{ The growth trajectory over absolute time of an individual organism
+  as a function of body mass.  Initial growth follows the black trajectory to
+  an energetically replete reproductive adult mass $m=\epsilon_\lambda M$ which we assume is 95\% asymptotic mass $M$.  Starvation follows the red
+  trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. Recovery follows the
+  green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires different energetics and a longer time to reach $\epsilon_\lambda M$.
+Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}
+\label{fig:growth}
+\end{figure}
 
@@
 where we will define values of $\epsilon$ to describe a set of rates within our model.
-For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$; \citenum{West:2001bv}).
+For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$; \cite{West:2001bv}).
 The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$, this rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.45\textwidth]{fig_Rates.pdf}
+\caption{
+Allometrically constrained starvation rate $\sigma=1/t_{\sigma}$ (red), mortality $\mu=1/t_{\mu}$ (blue) and recovery rate $\rho=1/t_{\rho}$ (green) relative to the reproductive rate $\lambda=1/t_{\lambda}$ (black) as a function of body mass (see Equations \ref{t1}, \ref{rhotimescale}, \ref{eq:sigma}, and \ref{mutimescale}).
+The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
+Mean values $\pm 25\%$ variation are shown by the shaded region for each rate.
+}
+\label{fig:gvs}
+\end{figure}
 
@@
 
-%CHRIS - MAYBE WE CUT THIS OUT SINCE WE STICK TO MAMMALS IN THIS PAPER
-%It should be noted that we have thus far used mammalian allometry to describe the size-based relations for growth, starvation, and death.
-However, our presentation is general, and other functional forms for $\epsilon$, for example, could be determined for other classes of organisms.
-Considering bacteria, we might expect that starvation or death is defined by the complete digestion of proteins, and in Table \ref{parameter-values} we provide all parameter values for bacteria which we later use as a comparison in our analysis.
-
-%\left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}
-%where $c=(a/b)^{1/(\eta-1)}$.
 Although the rate equations \eqref{eq:system} are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
@@
 Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
-\\
-%, though this is beyond the current scope of our analysis. %, and we henceforth focus our examination on mammalian species.
-
-\noindent {\bf  Stabilizing effects of allometric constraints} \\
+
+\section*{Stabilizing effects of allometric constraints}
 As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
@@
 We thus find that the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
-%, which for mammals ranges from ca. 1 gram (the Etruscan shrew) to ca. $10^7$ grams (represented by the Indricotheriinae, a subfamily of mammals living from the mid-Eocene to early Miocene).
 Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
@@
 This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
-%Thus our NSM model predicts that population cycles should be less common for smaller species and more common for larger species, particularly in environments where resources are limiting.%, and
-%2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
-
-It should be noted that previous studies used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.425\textwidth]{fig_DataHopf.pdf}
+\caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
+  allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
+  a function of different mammalian body sizes: $M=P\times10^2$g (blue) and
+  $M=P\times10^6$g (green), where $P$ is a random uniform variable in $[1,9]$.
+  Points denote realized values of $\sigma$ and $\rho$ given the drawn values for $M$.
+  }
+\label{fig:hopf}
+\end{figure}
+
+Previous studies used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
 However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
-%It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic~\cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
 Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
-However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.\\ % in agreement with our predictions of more pronounced transient dynamics, given how close these points are to the Hopf bifurcation.On the other hand,
-% \vspace{6mm}
-%, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
- %, emphasizing the role of starvation on dynamics.
+However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.
+
+\begin{figure*}
+\centering
+\includegraphics[width=1\textwidth]{fig_ExtinctionAllometricComb3.png}
+\caption{ Probability of extinction for a consumer with (\emph{A}) $M=10^2$g, (\emph{B}) $M=10^4$g, and (\emph{C}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
+Note the change in scale for $M=10^4$ and $M=10^6$g.
+Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
+The white points denote the allometrically constrained starvation and recovery rate.}
+\label{fig:ext}
+\end{figure*}
 
 %Extinction to the left and the right
-\noindent {\bf Extinction risk}\\
+\section*{Extinction risk}
 Within our model, higher rates of starvation result in a larger flux of the population to the hungry state.
@@
 We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
-To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^5$ and $t \leq 10^8$.
-This procedure is repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for an organism of $M=100$ grams.
+To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
+This procedure is repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for organisms with mass ranging from $10^2$ to $10^6$ grams.
 %the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
@@
 
-We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge (Fig. \ref{fig:ext}, white point).
+We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge across a range of $M$ (Fig. \ref{fig:ext}\emph{A-C}, white points).
 These values are close enough to the Hopf bifurcation to avoid low steady-state densities, and far enough away to avoid large-amplitude transient cycles.
@@
 This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
+We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig.
+\ref{fig:ext}\emph{A-C}), suggesting that the population dynamics of larger organisms are more sensitive to smaller changes in physiological rates controlling starvation and recovery.
 % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
@@
 
-\noindent {\bf Dynamic and energetic barriers to body size} \\
+\section*{Dynamic and energetic barriers to body size}
 Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\cite{West:2002ud}.
@@
 
-%Minimal mass bound
-%A theoretical lower bound on mammalian body size occurs when $\epsilon=0$ where mammals have no metabolic reserves and immediately starve; this occurs at a size of $M=0$, which is obviously infeasible.
-%Such prediction gives an extreme limit on size but does not account for the subtleties of starvation dynamics
-%that may limit body size (among other factors).
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{fig_FPAllometric.pdf}
+\caption{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
+  body mass.
+  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
+  (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.
+The data are from Damuth \cite{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
+\cite{West:2001bv,hou,moses2008rmo}).}
+\label{fig:mass}
+\end{figure}
+
 The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
-Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{damuth1987interspecific,allen2002,enquist1998}.
-The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size (e.g., \cite{damuth1987interspecific,allen2002,enquist1998}).
-%Previous studies have focused on t
-This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism ($N^*\propto M^{-3/4}/B_{0}$)  (e.g., \cite{allen2002,enquist1998}).
-This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{damuth1987interspecific,allen2002,enquist1998}.
+Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{Damuth:1987kr,allen2002,enquist1998}.
+The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size~\cite{Damuth:1987kr,allen2002,enquist1998}.
+This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism, such that $N^*\propto M^{-3/4}/B_{0}$~\cite{allen2002,enquist1998}.
+This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{Damuth:1987kr,allen2002,enquist1998}.
 Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
@@
 Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
-%, violating the energy equivalence.
-% set by deviations from constant efficiency obeyed by other populations.
 In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
@@
 
-%Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end.
-%Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is close to the minimum observed mammalian body size of ca.\ 1.3-2.5g, a range that occurs as the recovery rate begins its decline.
-%In addition to known transport limitations~\cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.
-%This result mirrors other efforts \cite{Kempes:2012hy,Kempes:2016} where at a given scale multiple limitations constrain the smallest possibilities for life within a class of organisms.
-
-%The combined steady-state dynamics and allometric timescales of the NSM predict that larger mammals should outcompete smaller ones, and this suggests that the model may provide a framework with which to explore the energetic mechanisms behind Cope's rule.
-%, which we explore further below.
-
 We contend that the NSM provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time.
@@
 Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but does not offer a within-lineage mechanism by which larger body sizes are selected for.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.45\textwidth]{fig_Invasion.pdf}
+\caption{ Invasion feasibility for organisms with a proportional change in
+  mass $\chi$ against a population with a resident body mass $M$.  The blue
+  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
+  red region denotes values of $\chi$ that result in a mass that is below the
+  starvation threshold and are thus infeasible.
+  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
+  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}
+\label{fig:invasion}
+\end{figure}
 
@@
 The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
-Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the state of starved foragers, is assumed to remain unchanged for the invader population (see SI for detailed derivations of invader rates).
-%There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
+Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the body composition of starved foragers, is assumed to remain unchanged for the invader population (see SI for detailed derivations of invader rates).
 
 To assess the susceptibility of the resident population to invasion, we determine which consumer has a lower steady-state resource density for a given value of $\chi$, again with the expectation that populations able to survive on lower resource densities have a competitive advantage \cite{tilman1981}.
-We find that for $M\leq 1.75\times10^7$g, having additional body fat ($\chi > 0$) results in lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
-However, for $M> 1.75\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
-
-%, and this is due to the changing
-%covariance between energetic rates as a function of modified energetic
-%reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
-
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.75\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
+We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
+However, for $M> 1.748\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
+
+The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
 This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca.
@@
 Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \cite{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals outlined by Alroy~\cite{Alroy:1998p1594}.
-%as well as  \cite{Alroy:1998p1594,Clauset:2009fh}.
-%This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
 While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general within-lineage mechanism for the evolution of larger body size among terrestrial mammals.
 
-%and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
-
-%One might be concerned a greater number of large mammals are currently not observed in the modern world given that larger mammals are less susceptible to extinction.
-%However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also much more geographically widespread than today.
-%These results, combined with our findings, suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
-
-%Closing
 The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations~\cite{Stearns:1989ip}.
@@
 
-%In accordance with the fasting endurance hypothesis,
+% \putbib[aa_starving3]
+
+\begin{thebibliography}{10}
+
+\bibitem{Martin:1987dl}
+Thomas~E Martin.
+\newblock {Food as a limit on breeding birds: A life-history perspective}.
+\newblock {\em Annu. Rev. Ecol. Syst.}, 18:453--487, January 1987.
+
+\bibitem{Kirk:1997cc}
+Kevin~L Kirk.
+\newblock {Life-history responses to variable environments: Starvation and
+  reproduction in planktonic rotifers}.
+\newblock {\em Ecology}, 78(2):434--441, March 1997.
+
+\bibitem{Kempes:2012hy}
+C~P Kempes, S~Dutkiewicz, and M~J Follows.
+\newblock {Growth, metabolic partitioning, and the size of microorganisms}.
+\newblock {\em Proc. Natl. Acad. Sci. USA}, 109(2):495--500, January 2012.
+
+\bibitem{Mangel:1988uaa}
+M~Mangel and C.~W. Clark.
+\newblock {\em {Dynamic Modeling in Behavioral Ecology}}.
+\newblock Princeton University Press, Princeton, 1988.
+
+\bibitem{Mangel:2014kz}
+Marc Mangel.
+\newblock {Stochastic dynamic programming illuminates the link between
+  environment, physiology, and evolution}.
+\newblock {\em B. Math. Biol.}, 77(5):857--877, 2014.
+
+\bibitem{Yeakel:2013hi}
+Justin~D Yeakel, Nathaniel~J Dominy, Paul~L Koch, and Marc Mangel.
+\newblock {Functional morphology, stable isotopes, and human evolution: a model
+  of consilience}.
+\newblock {\em Evolution}, 68(1):190--203, 2014.
+
+\bibitem{Morris:1987eo}
+Douglas~W Morris.
+\newblock {Optimal allocation of parental investment}.
+\newblock {\em Oikos}, 49(3):332--339, July 1987.
+
+\bibitem{Tveraa:2003fq}
+Torkild Tveraa, Per Fauchald, Cathrine Henaug, and Nigel~G Yoccoz.
+\newblock {An examination of a compensatory relationship between food
+  limitation and predation in semi-domestic reindeer}.
+\newblock {\em Oecologia}, 137(3):370--376, November 2003.
+
+\bibitem{Daan:1988va}
+S~Daan, C~Dijkstra, R~Drent, and T~Meijer.
+\newblock {Food supply and the annual timing of avian reproduction}.
+\newblock In H~Ouellet, editor, {\em Acta XIX Congressus Internationalis
+  Ornithologici, Volume I: Proceedings XIX International Ornithological
+  Congress, 1986, Ottawa}, pages 392--407. Proceedings XIX International
+  Ornithological Congress, Ottawa, 1989.
+
+\bibitem{Jacot:2009dv}
+Alain Jacot, Mihai Valcu, Kees van Oers, and Bart Kempenaers.
+\newblock {Experimental nest site limitation affects reproductive strategies
+  and parental investment in a hole-nesting passerine}.
+\newblock {\em Animal Behaviour}, 77(5):1075--1083, May 2009.
+
+\bibitem{Stearns:1989ip}
+S~C Stearns.
+\newblock {Trade-offs in life-history evolution}.
+\newblock {\em Funct. Ecol.}, 3(3):259, 1989.
+
+\bibitem{Barboza:2002in}
+P~Barboza and D~Jorde.
+\newblock {Intermittent fasting during winter and spring affects body
+  composition and reproduction of a migratory duck}.
+\newblock {\em J Comp Physiol B}, 172(5):419--434, July 2002.
+
+\bibitem{Threlkeld:1976ih}
+Stephen~T Threlkeld.
+\newblock {Starvation and the size structure of zooplankton communities}.
+\newblock {\em Freshwater Biol.}, 6(6):489--496, December 1976.
+
+\bibitem{Weber:1998jg}
+Thomas~P Weber, Bruno~J Ens, and Alasdair~I Houston.
+\newblock {Optimal avian migration: A dynamic model of fuel stores and site
+  use}.
+\newblock {\em Evolutionary Ecology}, 12(4):377--401, May 1998.
+
+\bibitem{Mduma:1999cp}
+Simon A~R Mduma, A~R~E Sinclair, and Ray Hilborn.
+\newblock {Food regulates the Serengeti wildebeest: a 40-year record}.
+\newblock {\em J. Anim. Ecol.}, 68(6):1101--1122, November 1999.
+
+\bibitem{Moore:2014hi}
+Jonathan~W Moore, Justin~D Yeakel, Dean Peard, Jeff Lough, and Mark Beere.
+\newblock {Life-history diversity and its importance to population stability
+  and persistence of a migratory fish: steelhead in two large North American
+  watersheds}.
+\newblock {\em J. Anim. Ecol.}, 83(5):1035--1046, 2014.
+
+\bibitem{Mead:1989dt}
+Rodney~A Mead.
+\newblock {The Physiology and Evolution of Delayed Implantation in Carnivores}.
+\newblock In John~L Gittleman, editor, {\em Carnivore Behavior, Ecology, and
+  Evolution}, pages 437--464. Springer US, Ithaca, 1989.
+
+\bibitem{Sandell:1990kw}
+Mikael Sandell.
+\newblock {The evolution of seasonal delayed implantation}.
+\newblock {\em Q Rev Biol}, 65(1):23--42, 1990.
+
+\bibitem{Bulik:1999eo}
+Cynthia~M Bulik, Patrick~F Sullivan, Jennifer~L Fear, Alison Pickering, Aria
+  Dawn, and Mandy McCullin.
+\newblock {Fertility and reproduction in women with anorexia nervosa}.
+\newblock {\em J. Clin. Psychiat.}, 60(2):130--135, February 1999.
+
+\bibitem{Trites:2003cc}
+A~W Trites and C~P Donnelly.
+\newblock {The decline of Steller sea lions Eumetopias jubatus in Alaska: a
+  review of the nutritional stress hypothesis}.
+\newblock {\em Mammal Rev.}, 33(1):3--28, March 2003.
+
+\bibitem{Glazier:2009hq}
+Douglas~S Glazier.
+\newblock {Metabolic level and size scaling of rates of respiration and growth
+  in unicellular organisms}.
+\newblock {\em Funct. Ecol.}, 23(5):963--968, October 2009.
+
+\bibitem{Kooi2000}
+S~A L~M Kooijman.
+\newblock {\em {Dynamic Energy and Mass Budgets in Biological Systems}}.
+\newblock Cambridge, 2000.
+
+\bibitem{Sousa:2010ez}
+T~Sousa, T~Domingos, J~C Poggiale, and S~A L~M Kooijman.
+\newblock {Dynamic energy budget theory restores coherence in biology}.
+\newblock {\em Philos. T. Roy. Soc. B}, 365(1557):3413--3428, October 2010.
+
+\bibitem{Diekmann:2010da}
+O~Diekmann and J~A~J Metz.
+\newblock {How to lift a model for individual behaviour to the population
+  level?}
+\newblock {\em Philos. T. Roy. Soc. B}, 365(1557):3523--3530, November 2010.
+
+\bibitem{murdoch:2003}
+W~W Murdoch, C~J Briggs, and R~M Nisbet.
+\newblock {\em {Consumer-resource Dynamics}}, volume~36 of {\em Monographs in
+  population biology}.
+\newblock Princeton University Press, Princeton, 2003.
+
+\bibitem{Benichou:2014wu}
+Olivier Benichou and S~Redner.
+\newblock {Depletion-Controlled Starvation of a Diffusing Forager}.
+\newblock {\em arXiv}, pages 1--5, May 2014.
+
+\bibitem{Benichou:2016wl}
+O~B{\'e}nichou, M~Chupeau, and S~Redner.
+\newblock {Role of depletion on the dynamics of a diffusing forager}.
+\newblock {\em Journal of Physics A: {\ldots}}, 2016.
+
+\bibitem{Chupeau:2016jf}
+M~Chupeau, O~B{\'e}nichou, and S~Redner.
+\newblock {Universality classes of foraging with resource renewal}.
+\newblock {\em Phys. Rev. E}, 93(3):032403, March 2016.
+
+\bibitem{Persson:1998hz}
+Lennart Persson, Kjell Leonardsson, Andr{\'e}~M De~Roos, Mats Gyllenberg, and
+  Bent Christensen.
+\newblock {Ontogenetic scaling of foraging rates and the dynamics of a
+  size-structured consumer-resource model}.
+\newblock {\em Theor Popul Biol}, 54(3):270--293, December 1998.
+
+\bibitem{murray2011mathematical}
+J~D Murray.
+\newblock {\em {Mathematical Biology: I. An Introduction}}, volume 110 of {\em
+  Interdisciplinary Applied Mathematics}.
+\newblock Springer New York, 2011.
+
+\bibitem{Strogatz:2008wo}
+S~H Strogatz.
+\newblock {\em {Nonlinear Dynamics and Chaos: With Applications to Physics,
+  Biology, Chemistry, and Engineering}}.
+\newblock Studies in nonlinearity. Westview Press, Boulder, 2008.
+
+\bibitem{GuckHolmes}
+J~Guckenheimer and P~Holmes.
+\newblock {\em {Nonlinear Oscillations, Dynamical Systems, and Bifurcations of
+  Vector Fields}}.
+\newblock Springer, New York, 1983.
+
+\bibitem{Gross:2004p2428}
+T~Gross and U~Feudel.
+\newblock {Analytical search for bifurcation surfaces in parameter space}.
+\newblock {\em Physica D}, 195(3-4):292--302, 2004.
+
+\bibitem{Hastings:2001jh}
+Alan Hastings.
+\newblock {Transient dynamics and persistence of ecological systems}.
+\newblock {\em Ecol. Lett.}, 4(3):215--220, May 2001.
+
+\bibitem{Neubert:1997wk}
+MG~Neubert and H~Caswell.
+\newblock {Alternatives to resilience for measuring the responses of ecological
+  systems to perturbations}.
+\newblock {\em Ecology}, 78(3):653--665, 1997.
+
+\bibitem{Caswell:2005eo}
+Hal Caswell and Michael~G Neubert.
+\newblock {Reactivity and transient dynamics of discrete-time ecological
+  systems}.
+\newblock {\em J Differ Equ Appl}, 11(4-5):295--310, April 2005.
+
+\bibitem{Neubert:2009td}
+MG~Neubert and H~Caswell.
+\newblock {Detecting reactivity}.
+\newblock {\em Ecology}, 90(10):2683--2688, 2009.
+
+\bibitem{Yodzis:1992hg}
+P~Yodzis and S~Innes.
+\newblock {Body size and consumer-resource dynamics}.
+\newblock {\em Am. Nat.}, 139(6):1151--1175, 1992.
+
+\bibitem{Brown:2004wq}
+JH~Brown, JF~Gillooly, AP~Allen, VM~Savage, and GB~West.
+\newblock {Toward a metabolic theory of ecology}.
+\newblock {\em Ecology}, 85(7):1771--1789, 2004.
+
+\bibitem{West:2002it}
+Geoffrey~B West, William~H Woodruff, and James~H Brown.
+\newblock {Allometric scaling of metabolic rate from molecules and mitochondria
+  to cells and mammals.}
+\newblock {\em Proc. Natl. Acad. Sci. USA}, 99 Suppl 1(suppl 1):2473--2478,
+  February 2002.
+
+\bibitem{DeLong:2010dy}
+J~P DeLong, J~G Okie, M~E Moses, R~M Sibly, and J~H Brown.
+\newblock {Shifts in metabolic scaling, production, and efficiency across major
+  evolutionary transitions of life}.
+\newblock {\em Proc. Natl. Acad. Sci. USA}, 107(29):12941--12945, July 2010.
+
+\bibitem{West:2001bv}
+Geoffrey~B West, James~H Brown, and Brian~J Enquist.
+\newblock {A general model for ontogenetic growth}.
+\newblock {\em Nature}, 413(6856):628--631, October 2001.
+
+\bibitem{moses2008rmo}
+Melanie~E Moses, Chen Hou, William~H Woodruff, Geoffrey~B West, Jeffery~C
+  Nekola, Wenyun Zuo, and James~H Brown.
+\newblock {Revisiting a Model of Ontogenetic Growth: Estimating Model
+  Parameters from Theory and Data.}
+\newblock {\em Am. Nat.}, 171(5):632--645, 2008.
+
+\bibitem{gillooly2002esa}
+James~F Gillooly, Eric~L Charnov, Geoffrey~B West, Van~M Savage, and James~H
+  Brown.
+\newblock {Effects of size and temperature on developmental time}.
+\newblock {\em Nature}, 417(6884):70--73, May 2002.
+
+\bibitem{hou}
+Chen Hou, Wenyun Zuo, Melanie~E Moses, William~H Woodruff, James~H Brown, and
+  Geoffrey~B West.
+\newblock {Energy uptake and allocation during ontogeny}.
+\newblock {\em Science}, 322(5902):736--739, October 2008.
+
+\bibitem{bettencourt}
+L~M~A Bettencourt, J~Lobo, D~Helbing, C~Kuhnert, and G~B West.
+\newblock {Growth, innovation, scaling, and the pace of life in cities}.
+\newblock {\em Proc. Natl. Acad. Sci. USA}, 104(17):7301--7306, April 2007.
+
+\bibitem{Folland:2008ij}
+J~P Folland, T~M Mc~Cauley, and A~G Williams.
+\newblock {Allometric scaling of strength measurements to body size}.
+\newblock {\em Eur J Appl Physiol}, 102(6):739--745, January 2008.
+
+\bibitem{CalderIII:1983jd}
+W~A Calder~III.
+\newblock {An allometric approach to population cycles of mammals}.
+\newblock {\em J. Theor. Biol.}, 100(2):275--282, January 1983.
+
+\bibitem{Peterson:1984hj}
+R~O Peterson, R~E Page, and K~M Dodge.
+\newblock {Wolves, moose, and the allometry of population cycles}.
+\newblock {\em Science}, 224(4655):1350--1352, June 1984.
+
+\bibitem{Krukonis:1991fk}
+G~Krukonis and W~M Schaffer.
+\newblock {Population cycles in mammals and birds: Does periodicity scale with
+  body size?}
+\newblock {\em J. Theor. Biol.}, 148(4):469--493, February 1991.
+
+\bibitem{Hendriks:2012fc}
+A~Jan Hendriks and Christian Mulder.
+\newblock {Delayed logistic and Rosenzweig{\textendash}MacArthur models with
+  allometric parameter setting estimate population cycles at lower trophic
+  levels well}.
+\newblock {\em Ecol Complex}, 9:43--54, February 2012.
+
+\bibitem{Kendall:1999iy}
+Bruce~E Kendall, Cheryl~J Briggs, William~W Murdoch, Peter Turchin, Stephen~P
+  Ellner, Edward McCauley, Roger~M Nisbet, and Simon~N Wood.
+\newblock {Why do populations cycle? A synthesis of statistical and mechanistic
+  modeling approaches}.
+\newblock {\em Ecology}, 80(6):1789--1805, September 1999.
+
+\bibitem{Hogstedt:2005cr}
+G{\"o}ran H{\"o}gstedt, Tarald Seldal, and Arild Breistol.
+\newblock {Period length in cyclic animal populations}.
+\newblock {\em Ecology}, 86(2):373--378, February 2005.
+
+\bibitem{Kendall:1998hl}
+Kendall, Prendergast, and Bjornstad.
+\newblock {The macroecology of population dynamics: taxonomic and biogeographic
+  patterns in population cycles}.
+\newblock {\em Ecol. Lett.}, 1(3):160--164, November 1998.
+
+\bibitem{Brown:1993p708}
+JH~Brown, PA~Marquet, and ML~Taper.
+\newblock {Evolution of body size: consequences of an energetic definition of
+  fitness}.
+\newblock {\em Am. Nat.}, 142(4):573--584, 1993.
+
+\bibitem{West:1997cg}
+Geoffrey~B West, James~H Brown, and Brian~J Enquist.
+\newblock {A general model for the origin of allometric scaling laws in
+  biology}.
+\newblock {\em Science}, 276(5309):122--126, April 1997.
+
+\bibitem{West:2002ud}
+Geoffrey~B West, William~H Woodruff, and James~H Brown.
+\newblock {Allometric scaling of metabolic rate from molecules and mitochondria
+  to cells and mammals.}
+\newblock {\em Proc. Natl. Acad. Sci. USA}, 99 Suppl 1(Supplement
+  1):2473--2478, February 2002.
+
+\bibitem{Millar:1990p923}
+JS~Millar and G~Hickling.
+\newblock {Fasting endurance and the evolution of mammalian body size}.
+\newblock {\em Funct. Ecol.}, 4(1):5--12, 1990.
+
+\bibitem{Alroy:1998p1594}
+J~Alroy.
+\newblock {Cope's rule and the dynamics of body mass evolution in North
+  American fossil mammals}.
+\newblock {\em Science}, 280(5364):731--734, 1998.
+
+\bibitem{Clauset:2009fh}
+A~Clauset and S~Redner.
+\newblock {Evolutionary model of species body mass diversification}.
+\newblock {\em Phys. Rev. Lett.}, 102(3):038103, January 2009.
+
+\bibitem{Smith:2010p3442}
+Felisa~A Smith, Alison~G Boyer, James~H Brown, Daniel~P Costa, Tamar Dayan,
+  S~K~Morgan Ernest, Alistair~R Evans, Mikael Fortelius, John~L Gittleman,
+  Marcus~J Hamilton, Larisa~E Harding, Kari Lintulaakso, S~Kathleen Lyons,
+  Christy McCain, Jordan~G Okie, Juha~J Saarinen, Richard~M Sibly, Patrick~R
+  Stephens, Jessica Theodor, and Mark~D Uhen.
+\newblock {The evolution of maximum body size of terrestrial mammals}.
+\newblock {\em Science}, 330(6008):1216--1219, November 2010.
+
+\bibitem{Saarinen:2014br}
+Juha~J Saarinen, Alison~G Boyer, James~H Brown, Daniel~P Costa, S~K~Morgan
+  Ernest, Alistair~R Evans, Mikael Fortelius, John~L Gittleman, Marcus~J
+  Hamilton, Larisa~E Harding, Kari Lintulaakso, S~Kathleen Lyons, Jordan~G
+  Okie, Richard~M Sibly, Patrick~R Stephens, Jessica Theodor, Mark~D Uhen, and
+  Felisa~A Smith.
+\newblock {Patterns of maximum body size evolution in Cenozoic land mammals:
+  Eco-evolutionary processes and abiotic forcing}.
+\newblock {\em Proc Biol Sci}, 281(1784):20132049, June 2014.
+
+\bibitem{Damuth:1987kr}
+John Damuth.
+\newblock {Interspecific allometry of population density in mammals and other
+  animals: the independence of body mass and population energy-use}.
+\newblock {\em Biol. J. Linn. Soc.}, 31(3):193--246, July 1987.
+
+\bibitem{allen2002}
+Andrew~P Allen, James~H Brown, and James~F Gillooly.
+\newblock {Global biodiversity, biochemical kinetics, and the
+  energetic-equivalence rule}.
+\newblock {\em Science}, 297(5586):1545--1548, August 2002.
+
+\bibitem{enquist1998}
+Brian~J Enquist, James~H Brown, and Geoffrey~B West.
+\newblock {Allometric scaling of plant energetics and population density}.
+\newblock {\em Nature}, 395(6698):163--165, September 1998.
+
+\bibitem{Kempes:2016}
+Christopher~P Kempes, Lawrence Wang, Jan~P Amend, John Doyle, and Tori Hoehler.
+\newblock {Evolutionary tradeoffs in cellular composition across diverse
+  bacteria}.
+\newblock {\em ISME J}, 10(9):2145--2157, April 2016.
+
+\bibitem{tilman1981}
+David Tilman.
+\newblock {Tests of resource competition theory using four species of lake
+  michigan algae}.
+\newblock {\em Ecology}, 62(3):802--815, June 1981.
+
+\bibitem{dutkiewicz2009}
+S~Dutkiewicz, M~J Follows, and J~G Bragg.
+\newblock {Modeling the coupling of ocean ecology and biogeochemistry}.
+\newblock {\em Global Biogeochem. Cycles}, 23(4):1--15, December 2009.
+
+\bibitem{barton2010}
+Andrew~D Barton, Stephanie Dutkiewicz, Glenn Flierl, Jason Bragg, and Michael~J
+  Follows.
+\newblock {Patterns of diversity in marine phytoplankton}.
+\newblock {\em Science}, 327(5972):1509--1511, March 2010.
+
+\end{thebibliography}
+
+\end{bibunit}
+
+% \bibliography{aa_starving3}
+
 %
+% \begin{acknowledgments}
+%   We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
+%   discussions and comments on the manuscript.  J.D.Y. was supported by
+%   startup funds at the University of California, Merced, and an Omidyar
+%   Postdoctoral Fellowship at the Santa Fe Institute.  C.P.K. was supported by
+%   an Omidyar Postdoctoral Fellowship at the Santa Fe Institute.  S.R. was
+%   supported by grants DMR-1608211 and 1623243 from the National Science
+%   Foundation, and by the John Templeton Foundation, all at the Santa Fe
+%   Institute.
+% \end{acknowledgments}
 %
-%Such energetic advantages, coupled with the
-%Ove and tends to be selected for over evolutionary time.
-%The NSM is in accordance with the XXX, where species with larger masses are less likely to evince cyclic dynamics.
-%
-%
-%
-%
-%We offer an additional mechanism for the minimum and maximum limitation on mammalian body size operating on ecological timescales.
-%Such a mechanism would
-%
-%
-%
-%
-%Although these physiological barriers are indeed of primary importance, our minimal model of starvation dynamics suggests that there may exist additional constraints due to dynamic and evolutionary effects at small and large body sizes, respectively.
-%
-%
-%
-
-%Looking Forward
-
-%
-%
-%%Starving Random Walk
-%In our modeling, foragers look for food by wandering in this changing
-%environment.  If such a search is successful, the forager is satiated and it
-%can engage in the essential activity of reproduction.  However, if the search
-%is unsuccessful for a sufficiently long period, the forager ``starves''.
-%Such a forager can do nothing else but forage, until it either finds food and
-%again becomes satiated or it dies when it goes too long without finding
-%nourishment.  These rules are reminiscent of the ``starving random walk''
-%model, where a single random walk can take $\mathcal{S}$ steps without
-%encountering food before starving to death.  Moreover, the resource does not
-%regenerate, so that the forager ultimately starves to death.  For this
-%idealized model, it was found that the average lifetime of the forager scales
-%algebraically with $\mathcal{S}$ in $d\!=\!1$ and $d\!=\!2$ dimensions, and
-%as $\exp(-A\mathcal{S}^\omega)$ for $d>2$.  Here the exponent
-%$\omega\approx \frac{1}{2}$ for $d=3$, while $\omega\to 1$ only as
-%$d\to \infty$, with the latter behavior corresponding to the mean-field
-%limit.  As we will discuss, regeneration of the resource, together with the
-%behavioral change between starving and satiated foragers leads to still much
-%richer dynamical behavior.
-
-%The reproductive tradeoff as a strategy
-%
-
-%
-%
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=0.5\textwidth]{fig_resvuln.pdf}
-% 	\caption{
-% 	Probability that the resource value is less than threshold = 0.05
-% 	}
-% 	\label{SN}
-% \end{figure}
-%
-%
-%
-% \begin{figure}[h]
-% 	\centering
-% 	\includegraphics[width=0.5\textwidth]{fig_Competition.pdf}
-% 	\caption{
-% 	Hopf bifurcation
-% 	}
-% 	\label{SN}
-% \end{figure}
-
-\bibliography{aa_starving3}
-
-% \begin{thebibliography}{10}
-%
-% \bibitem{Martin:1987dl}
-% Martin TE
-% \newblock (1987) {Food as a Limit on Breeding Birds: A Life-History
-%   Perspective}.
-% \newblock \emph{Annu. Rev. Ecol. Syst.} 18:453--487.
-%
-% \bibitem{Kirk:1997cc}
-% Kirk KL
-% \newblock (1997) {Life-History Responses to Variable Environments: Starvation
-%   and Reproduction in Planktonic Rotifers}.
-% \newblock \emph{Ecology} 78:434--441.
-%
-% \bibitem{Kempes:2012hy}
-% Kempes CP, Dutkiewicz S, Follows MJ
-% \newblock (2012) {Growth, metabolic partitioning, and the size of
-%   microorganisms.}
-% \newblock \emph{PNAS} 109:495--500.
-%
-% \bibitem{Mangel:1988uaa}
-% Mangel M, Clark CW
-% \newblock (1988) \emph{{Dynamic Modeling in Behavioral Ecology}}
-% \newblock (Princeton University Press, Princeton).
-%
-% \bibitem{Mangel:2014kz}
-% Mangel M
-% \newblock (2014) {Stochastic dynamic programming illuminates the link between
-%   environment, physiology, and evolution}.
-% \newblock \emph{B. Math. Biol.} 77:857--877.
-%
-% \bibitem{Yeakel:2013hi}
-% Yeakel JD, Dominy NJ, Koch PL, Mangel M
-% \newblock (2014) {Functional morphology, stable isotopes, and human evolution:
-%   a model of consilience}.
-% \newblock \emph{Evolution} 68:190--203.
-%
-% \bibitem{Morris:1987eo}
-% Morris DW
-% \newblock (1987) {Optimal Allocation of Parental Investment}.
-% \newblock \emph{Oikos} 49:332.
-%
-% \bibitem{Tveraa:2003fq}
-% Tveraa T, Fauchald P, Henaug C, Yoccoz NG
-% \newblock (2003) {An examination of a compensatory relationship between food
-%   limitation and predation in semi-domestic reindeer}.
-% \newblock \emph{Oecologia} 137:370--376.
-%
-% \bibitem{Daan:1988va}
-% Daan S, Dijkstra C, Drent R, Meijer T
-% \newblock (1988) \emph{{Food supply and the annual timing of avian
-%   reproduction}}.
-%
-% \bibitem{Jacot:2009dv}
-% Jacot A, Valcu M, van Oers K, Kempenaers B
-% \newblock (2009) {Experimental nest site limitation affects reproductive
-%   strategies and parental investment in a hole-nesting passerine}.
-% \newblock \emph{Animal Behaviour} 77:1075--1083.
-%
-% \bibitem{Stearns:1989ip}
-% Stearns SC
-% \newblock (1989) {Trade-Offs in Life-History Evolution}.
-% \newblock \emph{Funct. Ecol.} 3:259.
-%
-% \bibitem{Barboza:2002in}
-% Barboza P, Jorde D
-% \newblock (2002) {Intermittent fasting during winter and spring affects body
-%   composition and reproduction of a migratory duck}.
-% \newblock \emph{J Comp Physiol B} 172:419--434.
-%
-% \bibitem{Threlkeld:1976ih}
-% Threlkeld ST
-% \newblock (1976) {Starvation and the size structure of zooplankton
-%   communities}.
-% \newblock \emph{Freshwater Biol.} 6:489--496.
-%
-% \bibitem{Weber:1998jg}
-% Weber TP, Ens BJ, Houston AI
-% \newblock (1998) {Optimal avian migration: A dynamic model of fuel stores and
-%   site use}.
-% \newblock \emph{Evolutionary Ecology} 12:377--401.
-%
-% \bibitem{Mduma:1999cp}
-% Mduma SAR, Sinclair ARE, Hilborn R
-% \newblock (1999) {Food regulates the Serengeti wildebeest: a 40-year record}.
-% \newblock \emph{J. Anim. Ecol.} 68:1101--1122.
-%
-% \bibitem{Moore:2014hi}
-% Moore JW, Yeakel JD, Peard D, Lough J, Beere M
-% \newblock (2014) {Life-history diversity and its importance to population
-%   stability and persistence of a migratory fish: steelhead in two large North
-%   American watersheds}.
-% \newblock \emph{J. Anim. Ecol.} 83:1035--1046.
-%
-% \bibitem{Mead:1989dt}
-% Mead RA
-% \newblock (1989) in \emph{Carnivore Behavior, Ecology, and Evolution}
-% \newblock (Springer US, Boston, MA), pp 437--464.
-%
-% \bibitem{Sandell:1990kw}
-% Sandell M
-% \newblock (1990) {The Evolution of Seasonal Delayed Implantation}.
-% \newblock \emph{The Quarterly Review of Biology} 65:23--42.
-%
-% \bibitem{Bulik:1999eo}
-% Bulik CM, {et~al.}
-% \newblock (1999) {Fertility and Reproduction in Women With Anorexia Nervosa}.
-% \newblock \emph{J. Clin. Psychiatry} 60:130--135.
-%
-% \bibitem{Trites:2003cc}
-% Trites AW, Donnelly CP
-% \newblock (2003) {The decline of Steller sea lions Eumetopias jubatus in
-%   Alaska: a review of the nutritional stress hypothesis}.
-% \newblock \emph{Mammal Review} 33:3--28.
-%
-% \bibitem{Glazier:2009hq}
-% Glazier DS
-% \newblock (2009) {Metabolic level and size scaling of rates of respiration and
-%   growth in unicellular organisms}.
-% \newblock \emph{Funct. Ecol.} 23:963--968.
-%
-% \bibitem{Kooi2000}
-% Kooijman SALM
-% \newblock (2000) \emph{{Dynamic Energy and Mass Budgets in Biological Systems}}
-% \newblock (Cambridge).
-%
-% \bibitem{Sousa:2010ez}
-% Sousa T, Domingos T, Poggiale JC, Kooijman SALM
-% \newblock (2010) {Dynamic energy budget theory restores coherence in biology}.
-% \newblock \emph{Philos. T. Roy. Soc. B} 365:3413--3428.
-%
-% \bibitem{Diekmann:2010da}
-% Diekmann O, Metz JAJ
-% \newblock (2010) {How to lift a model for individual behaviour to the
-%   population level?}
-% \newblock \emph{Philos. T. Roy. Soc. B} 365:3523--3530.
-%
-% \bibitem{murdoch:2003}
-% Murdoch WW, Briggs CJ, Nisbet RM
-% \newblock (2003) \emph{{Consumer-resource Dynamics}}, Monographs in population
-%   biology
-% \newblock (Princeton University Press).
-%
-% \bibitem{Benichou:2014wu}
-% Benichou O, Redner S
-% \newblock (2014) {Depletion-Controlled Starvation of a Diffusing Forager}.
-% \newblock \emph{arXiv}.
-%
-% \bibitem{Benichou:2016wl}
-% B{\'e}nichou O, Chupeau M, Redner S
-% \newblock (2016) {Role of Depletion on the Dynamics of a Diffusing Forager}.
-%
-% \bibitem{Chupeau:2016jf}
-% Chupeau M, B{\'e}nichou O, Redner S
-% \newblock (2016) {Universality classes of foraging with resource renewal}.
-% \newblock \emph{Phys. Rev. E} 93:032403.
-%
-% \bibitem{murray2011mathematical}
-% Murray JD
-% \newblock (2011) \emph{{Mathematical Biology: I. An Introduction}},
-%   Interdisciplinary Applied Mathematics
-% \newblock (Springer New York, Melaka Manipal Medical College (Manipal Campus),
-%   International Centre for Health Sciences, Madhav Nagar, Manipal, Udupi
-%   District, Karnataka State, India. nayaksathish@yahoo.com) Vol.{} 110.
-%
-% \bibitem{Strogatz:2008wo}
-% Strogatz SH
-% \newblock (2008) \emph{{Nonlinear Dynamics and Chaos}}, With Applications to
-%   Physics, Biology, Chemistry, and Engineering
-% \newblock (Westview Press).
-%
-% \bibitem{GuckHolmes}
-% Guckenheimer J, Holmes P
-% \newblock (1983) \emph{{Nonlinear oscillations, dynamical systems, and
-%   bifurcations of vector fields}}
-% \newblock (Springer, New York).
-%
-% \bibitem{Gross:2004p2428}
-% Gross T, Feudel U
-% \newblock (2004) {Analytical search for bifurcation surfaces in parameter
-%   space}.
-% \newblock \emph{Physica D} 195:292--302.
-%
-% \bibitem{Hastings:2001jh}
-% Hastings A
-% \newblock (2001) {Transient dynamics and persistence of ecological systems}.
-% \newblock \emph{Ecol. Lett.} 4:215--220.
-%
-% \bibitem{Neubert:1997wk}
-% Neubert M, Caswell H
-% \newblock (1997) {Alternatives to resilience for measuring the responses of
-%   ecological systems to perturbations}.
-% \newblock \emph{Ecology} 78:653--665.
-%
-% \bibitem{Caswell:2005eo}
-% Caswell H, Neubert MG
-% \newblock (2005) {Reactivity and transient dynamics of discrete-time ecological
-%   systems}.
-% \newblock \emph{Journal of Difference Equations and Applications} 11:295--310.
-%
-% \bibitem{Neubert:2009td}
-% Neubert M, Caswell H
-% \newblock (2009) {Detecting reactivity}.
-% \newblock \emph{Ecology}.
-%
-% \bibitem{Yodzis:1992hg}
-% Yodzis P, Innes S
-% \newblock (1992) {Body Size and Consumer-Resource Dynamics}.
-% \newblock \emph{Am. Nat.} 139:1151--1175.
-%
-% \bibitem{Brown:2004wq}
-% Brown J, Gillooly J, Allen A, Savage V, West G
-% \newblock (2004) {Toward a metabolic theory of ecology}.
-% \newblock \emph{Ecology} 85:1771--1789.
-%
-% \bibitem{West:2002it}
-% West GB, Woodruff WH, Brown JH
-% \newblock (2002) {Allometric scaling of metabolic rate from molecules and
-%   mitochondria to cells and mammals.}
-% \newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.
-%
-% \bibitem{DeLong:2010dy}
-% DeLong JP, Okie JG, Moses ME, Sibly RM, Brown JH
-% \newblock (2010) {Shifts in metabolic scaling, production, and efficiency
-%   across major evolutionary transitions of life.}
-% \newblock \emph{PNAS} 107:12941--12945.
-%
-% \bibitem{West:2001bv}
-% West GB, Brown JH, Enquist BJ
-% \newblock (2001) {A general model for ontogenetic growth}.
-% \newblock \emph{Nature} 413:628--631.
-%
-% \bibitem{moses2008rmo}
-% Moses ME, {et~al.}
-% \newblock (2008) {Revisiting a Model of Ontogenetic Growth: Estimating Model
-%   Parameters from Theory and Data}.
-% \newblock \emph{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}
-%   171:632--645.
-%
-% \bibitem{gillooly2002esa}
-% Gillooly JF, Charnov EL, West GB, Savage VM, Brown JH
-% \newblock (2002) {Effects of size and temperature on developmental time}.
-% \newblock \emph{Nature} 417:70--73.
-%
-% \bibitem{hou}
-% Hou C, {et~al.}
-% \newblock (2008) {Energy Uptake and Allocation During Ontogeny}.
-% \newblock \emph{Science} 322:736--739.
-%
-% \bibitem{bettencourt}
-% Bettencourt LMA, Lobo J, Helbing D, Kuhnert C, West GB
-% \newblock (2007) {Growth, innovation, scaling, and the pace of life in cities}.
-% \newblock \emph{Proc. Natl. Acad. Sci. USA} 104:7301--7306.
-%
-% \bibitem{Folland:2008ij}
-% Folland JP, Mc~Cauley TM, Williams AG
-% \newblock (2008) {Allometric scaling of strength measurements to body size}.
-% \newblock \emph{Eur J Appl Physiol} 102:739--745.
-%
-% \bibitem{CalderIII:1983jd}
-% Calder~III WA
-% \newblock (1983) {An allometric approach to population cycles of mammals}.
-% \newblock \emph{J. Theor. Biol.} 100:275--282.
-%
-% \bibitem{Peterson:1984hj}
-% Peterson RO, PAGE RE, DODGE KM
-% \newblock (1984) {Wolves, Moose, and the Allometry of Population Cycles}.
-% \newblock \emph{Science} 224:1350--1352.
-%
-% \bibitem{Krukonis:1991fk}
-% Krukonis G, Schaffer WM
-% \newblock (1991) {Population cycles in mammals and birds: Does periodicity
-%   scale with body size?}
-% \newblock \emph{J. Theor. Biol.} 148:469--493.
-%
-% \bibitem{Hendriks:2012fc}
-% Hendriks AJ, Mulder C
-% \newblock (2012) {Delayed logistic and Rosenzweig{\textendash}MacArthur models
-%   with allometric parameter setting estimate population cycles at lower trophic
-%   levels well}.
-% \newblock \emph{Ecological Complexity} 9:43--54.
-%
-% \bibitem{Kendall:1999iy}
-% Kendall BE, {et~al.}
-% \newblock (1999) {Why do populations cycle? A synthesis of statistical and
-%   mechanistic modeling approaches}.
-% \newblock \emph{Ecology} 80:1789--1805.
-%
-% \bibitem{Hogstedt:2005cr}
-% H{\"o}gstedt G, Seldal T, Breistol A
-% \newblock (2005) {Period length in cyclic animal populations}.
-% \newblock \emph{Ecology} 86:373--378.
-%
-% \bibitem{Kendall:1998hl}
-% Kendall, Prendergast, Bjornstad
-% \newblock (1998) {The macroecology of population dynamics: taxonomic and
-%   biogeographic patterns in population cycles}.
-% \newblock \emph{Ecol. Lett.} 1:160--164.
-%
-% \bibitem{Brown:1993p708}
-% Brown J, Marquet P, Taper M
-% \newblock (1993) {Evolution of body size: consequences of an energetic
-%   definition of fitness}.
-% \newblock \emph{Am. Nat.} 142:573--584.
-%
-% \bibitem{West:1997cg}
-% West GB, Brown JH, Enquist BJ
-% \newblock (1997) {A General Model for the Origin of Allometric Scaling Laws in
-%   Biology}.
-% \newblock \emph{Science} 276:122--126.
-%
-% \bibitem{West:2002ud}
-% West GB, Woodruff WH, Brown JH
-% \newblock (2002) {Allometric scaling of metabolic rate from molecules and
-%   mitochondria to cells and mammals}.
-% \newblock \emph{Proc. Natl. Acad. Sci. USA} 99:2473--2478.
-%
-% \bibitem{Millar:1990p923}
-% Millar J, Hickling G
-% \newblock (1990) { Fasting Endurance and the Evolution of Mammalian Body Size}.
-% \newblock \emph{Funct. Ecol.} 4:5--12.
-%
-% \bibitem{Alroy:1998p1594}
-% Alroy J
-% \newblock (1998) {Cope's rule and the dynamics of body mass evolution in North
-%   American fossil mammals}.
-% \newblock \emph{Science} 280:731.
-%
-% \bibitem{Clauset:2009fh}
-% Clauset A, Redner S
-% \newblock (2009) {Evolutionary Model of Species Body Mass Diversification}.
-% \newblock \emph{Phys. Rev. Lett.} 102:038103.
-%
-% \bibitem{Smith:2010p3442}
-% Smith F, Boyer A, Brown J, Costa D
-% \newblock (2010) {The Evolution of Maximum Body Size of Terrestrial Mammals}.
-% \newblock \emph{Science}.
-%
-% \bibitem{Saarinen:2014br}
-% Saarinen JJ, {et~al.}
-% \newblock (2014) {Patterns of maximum body size evolution in Cenozoic land
-%   mammals: eco-evolutionary processes and abiotic forcing}.
-% \newblock \emph{Proc Biol Sci} 281:20132049--20132049.
-%
-% \bibitem{allen2002}
-% Allen AP, Brown JH, Gillooly JF
-% \newblock (2002) {Global Biodiversity, Biochemical Kinetics, and the
-%   Energetic-Equivalence Rule}.
-% \newblock \emph{Science} 297:1545--1548.
-%
-% \bibitem{enquist1998}
-% Enquist BJ, Brown JH, West GB
-% \newblock (1998) {Allometric scaling of plant energetics and population density
-%   : Abstract : Nature}.
-% \newblock \emph{Nature} 395:163--165.
-%
-% \bibitem{tilman1981}
-% Tilman D
-% \newblock (1981) {Tests of Resource Competition Theory Using Four Species of
-%   Lake Michigan Algae}.
-% \newblock \emph{Ecology} 62:802--815.
-%
-% \bibitem{dutkiewicz2009}
-% Dutkiewicz S, Follows MJ, Bragg JG
-% \newblock (2009) {Modeling the coupling of ocean ecology and biogeochemistry}.
-% \newblock \emph{Global Biogeochem. Cycles} 23:n/a--n/a.
-%
-% \bibitem{barton2010}
-% Barton AD, Dutkiewicz S, Flierl G, Bragg J, Follows MJ
-% \newblock (2010) {Patterns of Diversity in Marine Phytoplankton}.
-% \newblock \emph{Science} 327:1509--1511.
-%
-% \end{thebibliography}
-
-\begin{acknowledgments}
-  We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
-  discussions and comments on the manuscript.  J.D.Y. was supported by
-  startup funds at the University of California, Merced, and an Omidyar
-  Postdoctoral Fellowship at the Santa Fe Institute.  C.P.K. was supported by
-  an Omidyar Postdoctoral Fellowship at the Santa Fe Institute.  S.R. was
-  supported by grants DMR-1608211 and 1623243 from the National Science
-  Foundation, and by the John Templeton Foundation, all at the Santa Fe
-  Institute.
-\end{acknowledgments}
-
-%\onecolumn
-%\let\cleardoublepage\clearpage
-
-%FIGURE 1
-
-\end{article}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_FixedPoint.pdf}
-\caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
-  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-  bifurcation conditions separate parameter space into unphysical, cyclic,
-  and steady state dynamic regimes.  The colors show the steady
-  state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
-  replete consumers $F^*$, (warmer colors denote higher densities).  Steady
-  state densities for the energetically deficient consumers $H^*$ (not shown)
-  scale with those for $F^*$.  }
-\label{fig:fp}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-\caption{ The growth trajectory over absolute time of an individual organism
-  as a function of body mass.  Initial growth follows the black trajectory to
-  an energetically replete reproductive adult mass $m=\epsilon_\lambda M$ which we assume is 95\% asymptotic mass $M$.  Starvation follows the red
-  trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. Recovery follows the
-  green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires different energetics and a longer time to reach $\epsilon_\lambda M$.
-Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}
-\label{fig:growth}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_Rates.pdf}
-\caption{
-Allometrically constrained starvation rate $\sigma=1/t_{\sigma}$ (red), mortality $\mu=1/t_{\mu}$ (red) and recovery rate $\rho=1/t_{\rho}$ (green) relative to the reproductive rate $\lambda=1/t_{\lambda}$ (black) as a function of body mass (see Equations \ref{t1}, \ref{rhotimescale}, \ref{eq:sigma}, and \ref{mutimescale}).
-The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
-Mean values $\pm 25\%$ variation are shown by the shaded region for each rate.
-}
-\label{fig:gvs}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_DataHopf.pdf}
-\caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
-  allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
-  a function of different mammalian body sizes: $M=P\times10^2$g (blue) and
-  $M=P\times10^6$g (green), where $P$ is a random uniform variable in $[1,9]$.
-  Points denote realized values of $\sigma$ and $\rho$ given the drawn values for $M$.
-  }
-\label{fig:hopf}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=1\textwidth]{fig_ExtinctionAllometricComb3.png}
-\caption{ Probability of extinction for a consumer with (from left to right) $M=10^2,~10^4,~10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
-Note the change in scale for $M=10^4,~10^6$g.
-Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
-The white points denote the allometrically constrained starvation and recovery rate.}
-\label{fig:ext}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_FPAllometric.pdf}
-\caption{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-  body mass.
-  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.
-The data are from Damuth \cite{damuth1987interspecific} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
-\cite{West:2001bv,hou,moses2008rmo}).}
-\label{fig:mass}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.75\textwidth]{fig_Invasion.pdf}
-\caption{ Invasion feasibility for organisms with a proportional change in
-  mass $\chi$ against a population with a resident body mass $M$.  The blue
-  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
-  red region denotes values of $\chi$ that result in a mass that is below the
-  starvation threshold and is thus infeasible.
-  Arrows point to the predicted optimal mass $M_{\rm opt}=1.75\times 10^7$, which serves as an evolutionary attractor for body mass.
-  The black wedge points to the largest body mass known for terrestrial mammals at $7.74\times10^7$g~\cite{Smith:2010p3442}.}
-\label{fig:invasion}
-\end{figure*}
-
-%
-%
-%	 \begin{figure}[h]
-%		 \centering
-%		 \includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-%		 \caption{
-%		 }
-%		 \label{growth-diagram}
-%	 \end{figure}
-%
-%	 \begin{figure}[h]
-%		 \centering
-%		 \includegraphics[width=0.25\textwidth]{fig_GvS.pdf}
-%		 \caption{
-%		 }
-%		 \label{GvS}
-%	 \end{figure}
-%
-%
-%	\begin{figure}[h]
-%	 	\centering
-%	 	\includegraphics[width=0.25\textwidth]{fig_FixedPoint.pdf}
-%	 	\caption{
-%	 	}
-%		\label{Hopfb}
-%	\end{figure}
-%
-%
-%	\begin{figure}[h]
-%		\centering
-%		\includegraphics[width=0.8\textwidth]{fig_DataHopf.pdf}
-%		\caption{
-%		}
-%		\label{DataHopf}
-%	\end{figure}
-%
-%
-%	\begin{figure}[h]
-%		\centering
-%	 	\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometric.pdf}
-%	 	\caption{
-%	 	}
-%	 	\label{Ext}
-%	\end{figure}
-%
-%	\begin{figure}[h]
-%	 	\centering
-%	 	\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
-%	 	\caption{
-%	 	}
-%	 	\label{Asymp}
-%	\end{figure}
-%
-%\clearpage
-
-%\begin{table}[h]
-%\caption{Parameter Values For Various Classes of Organisms}
-%\label{param}
-%    \begin{center}
-%    \small
-%     \begin{tabular}{| p{1.2cm}| p{3.2cm} | p{2.6cm} | p{3.2cm} | }
-%     \hline
-%     & {\bf Mammals} & {\bf Unicellular Eukaryotes} & {\bf Bacteria} \\
-%     \hline
-%   $\eta$ & $3/4$ & & $1.70$ \\
-%   $E_{m}$ & $10695$ (J gram$^{-1}$) & & $10695$ (J gram$^{-1}$) \\
-%   $E_{m}^{\prime}$ & $\approx E_{m}$ & & $\approx E_{m}$ \\
-%   $B_{0}$ & $0.019$ (W gram$^{-\alpha}$) & & $1.96\times10^{17}$ \\
-%   $B_{m}$ & $0.025$ (W gram$^{-1}$)   & & $0.025$ (W gram$^{-1}$)\\
-%   $a$ & $1.78\times10^{-6}$ & & $1.83\times10^{13}$ \\
-%   $b$ & $2.29\times10^{-6}$ & & $2.29\times10^{-6}$ \\
-%   $\eta-1$ & $-0.21$ & & $0.73$ \\
-%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) & & $56493$ \\
-%   $\gamma$ & $1.19$ & & $0.68$ \\
-%   $f_{0}$ & $0.02$ & & $1.30\times10^{-5}$ \\
-%   $\zeta$ & $1.01$ & & \\
-%   $mm_{0}$ & $0.32$ & & \\
-%
-%
-%   \hline
-%    \end{tabular}
-%    \end{center}
-%   \end{table}
+\clearpage
+
+\begin{bibunit}[unsrt]
+\section*{Supporting Information for ``The dynamics of starvation and recovery''}
+
+\subsection*{Mechanisms of Starvation and Recovery}
+Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
+For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
+\cite{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
+This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g.
+\cite{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \cite{kempes2014morphological}.
+Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).
+
+We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
+\begin{align}
+\begin{split}
+\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
+\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
+&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
+\end{split}
+\end{align}
+where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
+%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
+In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
+Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
+The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
+A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
+
+In the above equations $Y$ represents the yield coefficient (e.g., Refs.
+\cite{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
+If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
+This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
+In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment.
+
+Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
+As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
+Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
+\begin{align}
+\begin{split}
+\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
+\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
+& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
+\end{split}
+\end{align}
+%where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
+
+We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
+%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
+\begin{align}
+\begin{split}
+\dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
+\dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
+\dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
+\end{split}
+\end{align}
+If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
+\begin{align}
+\begin{split}
+\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
+\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
+\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
+\end{split}
+\end{align}
+where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$.
+The above equations represent the system of equations presented in the main text.
+
+\subsection*{Parameter Values and Estimates}
+
+%\subsection*{Standard synthesis and metabolic parameters}
+
+All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
+Below we outline previous measurements and simple estimates of the parameters.
+
+Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs.
+\cite{West:2001bv,moses2008rmo} and the supplement for Ref.
+\cite{hou}).
+We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \cite{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
+The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.
+For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs.
+\cite{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
+We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).
+
+An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction.
+This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref.
+\cite{peters1986ecological} for a review).
+We use $m_{0}=0.097M^{0.92}$ \cite{blueweiss1978relationships} which encompasses the widest range of body sizes \cite{peters1986ecological}.
+
+The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g.
+Refs.
+\cite{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \cite{moses2008rmo}.
+The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \cite{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \cite{stryer,hou}.
+For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \cite{hou}.
+
+For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \cite{Dunbrack:1993ec}, $0.02M^{1.19}$ \cite{Lindstedt:1985hm}, and $0.026M^{1.14}$ \cite{Lindstedt:2002td}.
+We use the values from \cite{Lindstedt:1985hm} which falls in the middle of this range.
+Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \cite{Lindstedt:2002td}.
+
+%We also connect the resource growth rate to the total metabolic rate of an organism.
+That is, we are interested in the relative rates of resource recovery and consumption by the total population.
+From \cite{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$).
+Considering an energy density of $18200$ (J g$^{-1}$) of grass \cite{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$).
+This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.
+
+Typically the value of $\xi=C/k$ should roughly be $2$.
+The value of $\rho$, $\lambda$, $\sigma$, and $\mu$ are all simple rates (note that we have not rescaled time in our non-dimensionalization) as defined in the maintext.
+Given that our model considers transitions over entire stages of ontogeny or nutritional states, the value of $Y$ must represent yields integrated over entire life stages.
+Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \cite{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
+\begin{equation}
+B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+\end{equation}
+Similarly, the maintenance for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
+\begin{equation}
+B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+\end{equation}
+Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.
+
+Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities.
+From the data in Ref.
+\cite{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
+It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \cite{damuth1987interspecific}.
+We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above.
+However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.
+
+%For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
+
+%More complicated models of fecundity (which, for example, account for the average length of adulthood and the number of individuals produced over this span) could be employed.
+However, the scaling of population growth rate has been studied in detail before and follows a relationship of $$ which matches the theory well for $\phi=.95$ and $=$.
+
+%In our calculations we include $20\%$ variation around this value which could account for differences in efficiency during
+
+ \begin{table}[h]
+\caption{Parameter values for mammals}
+\label{param}
+    \begin{center}
+    \small
+     \begin{tabular}{ p{1.2cm} p{3.2cm} l p{2.2cm}|}
+     \hline
+     Parameter & Value & References  \\
+     \hline
+   $\eta$ & $3/4$  &  (e.g. \cite{West:2001bv,moses2008rmo,hou}) \\
+   $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \cite{moses2008rmo,West:2001bv,hou} \\
+   $E_{m}^{\prime}$ & $7000$  & \cite{stryer,hou} \\
+   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \cite{hou}  \\
+%   $a$ & $1.78\times10^{-6}$  \quad \quad \\
+%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
+   $\gamma$ & $1.19$ & \cite{Lindstedt:1985hm} \\
+   $f_{0}$ & $0.02$ & \cite{Lindstedt:1985hm}\\
+   $\zeta$ & $1.00$  & \cite{Lindstedt:2002td} \\
+   $u_{0}$ & $0.38$  & \cite{Lindstedt:2002td} \\
+
+   \hline
+    \end{tabular}
+    \end{center}
+   \end{table}
+
+\subsection*{Rate equations for invaders with modified body mass}
+We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
+Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
+For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (see main text), implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
+Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
+
+Although the starved state of invading organisms remains unchanged, the rate of starvation from the modified full state to the starved state, the rate of recovery from the starved state to the modified full state, and the maintenance rates of both, will be different, such that $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$, $\delta^\prime = \delta(M^\prime)$.
+Rates of starvation and recovery for the invading population are easily derived by adjusting the starting or ending state before and after starvation and recovery, leading to the following timescales:
+
+\begin{align}
+t_{\sigma^\prime} &= -\frac{M^{1-\eta}}{a^{\prime}}\ln \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
+t_{\rho^\prime} &= \ln \left(\frac{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}{1-( \epsilon_\lambda(\chi +1))^{1/4}}\right)\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}.
+\end{align}
+
+The maintenance rates for the invading population require more careful consideration.
+First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of body fat reserves.
+Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \cite{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
+\begin{equation}
+B^{\prime}_{\lambda}=\int_{0}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
+\end{equation}
+where
+\begin{equation}
+t_{\lambda^\prime} = \frac{M^{1-\eta} }{a(1-\eta)}\ln \left(\frac{1-(m_0/M)^{1-\eta}}{1-(\epsilon_\lambda (1+\chi))^{1-\eta}} \right).
+\end{equation}
+Note that we do not use this timescale to determine the reproductive rate of the invading consumer---which is assumed to remain the same as the resident population---but only to calulate the lifetime energy use.
+Similarly, the maintenance for hungry individuals $P^\prime_{H}=B_{0}(\epsilon_{\sigma}(1+\chi)M)^{3/4}/(\epsilon_{\sigma}(1+\chi)M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y^\prime_{H}=(1+\chi)ME_{d}/B^{\prime}_{\rho}$ where
+\begin{equation}
+B^{\prime}_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
+\end{equation}
+Finally, we can calculate the maintenance of the invaders as
+
+\begin{align}
+  \delta^\prime &= P^\prime_{H}Y^\prime_{H}/\xi \\ \nonumber
+  \beta^\prime &= \left(\frac{\lambda_{\rm max}}{Y^\prime_{F}}+P^\prime_{F} \right)Y^\prime_{H}/\xi.
+\end{align}
+
+To determine whether or not the invader or resident population has an advantage, we compute $R^*(M)$ and $R^*(M^\prime=M(1+\chi))$ for values of $\chi \in (-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$, and the invading population is assumed to have an advantage over the resident population if $R^*(M^\prime)<R^*(M)$.
+
+% \putbib[aa_starving_supplement]
+
+\begin{thebibliography}{10}
+
+\bibitem{Kempes:2012hy}
+Christopher~P Kempes, Stephanie Dutkiewicz, and Michael~J Follows.
+\newblock {Growth, metabolic partitioning, and the size of microorganisms.}
+\newblock {\em PNAS}, 109(2):495--500, January 2012.
+
+\bibitem{kempes2014morphological}
+Christopher~P Kempes, Chinweike Okegbe, Zwoisaint Mears-Clarke, Michael~J
+  Follows, and Lars~EP Dietrich.
+\newblock Morphological optimization for access to dual oxidants in biofilms.
+\newblock {\em Proceedings of the National Academy of Sciences},
+  111(1):208--213, 2014.
+
+\bibitem{West:2001bv}
+Geoffrey~B West, James~H Brown, and Brian~J Enquist.
+\newblock {A general model for ontogenetic growth}.
+\newblock {\em Nature}, 413(6856):628--631, October 2001.
+
+\bibitem{moses2008rmo}
+Melanie~E Moses, Chen Hou, William~H Woodruff, Geoffrey~B West, Jeffery~C
+  Nekola, Wenyun Zuo, and James~H Brown.
+\newblock {Revisiting a model of ontogenetic growth: Estimating model
+  parameters from theory and data}.
+\newblock {\em http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735},
+  171(5):632--645, May 2008.
+
+\bibitem{hou}
+Chen Hou, Wenyun Zuo, Melanie~E Moses, William~H Woodruff, James~H Brown, and
+  Geoffrey~B West.
+\newblock {Energy uptake and allocation during ontogeny}.
+\newblock {\em Science}, 322(5902):736--739, October 2008.
+
+\bibitem{pirt}
+SJ~Pirt.
+\newblock The maintenance energy of bacteria in growing cultures.
+\newblock {\em Proceedings of the Royal Society of London B: Biological
+  Sciences}, 163(991):224--231, 1965.
+
+\bibitem{Heijnen}
+JJ~Heijnen and JA~Roels.
+\newblock A macroscopic model describing yield and maintenance relationships in
+  aerobic fermentation processes.
+\newblock {\em Biotechnology and Bioengineering}, 23(4):739--763, 1981.
+
+\bibitem{peters1986ecological}
+Robert~Henry Peters.
+\newblock {\em The Ecological Implications of Body Size}, volume~2.
+\newblock Cambridge University Press, Cambridge, 1986.
+
+\bibitem{blueweiss1978relationships}
+L~Blueweiss, H~Fox, V~Kudzma, D~Nakashima, R~Peters, and S~Sams.
+\newblock Relationships between body size and some life history parameters.
+\newblock {\em Oecologia}, 37(2):257--272, 1978.
+
+\bibitem{stryer}
+Lubert Stryer.
+\newblock {\em {Biochemistry, Fourth Edition}}.
+\newblock W.H. Freeman and Company, New York, 1995.
+
+\bibitem{Dunbrack:1993ec}
+Robert~L Dunbrack and Malcolm~A Ramsay.
+\newblock {The Allometry of Mammalian Adaptations to Seasonal Environments: A
+  Critique of the Fasting Endurance Hypothesis}.
+\newblock {\em Oikos}, 66(2):336--342, March 1993.
+
+\bibitem{Lindstedt:1985hm}
+Stan~L Lindstedt and Mark~S Boyce.
+\newblock {Seasonality, Fasting Endurance, and Body Size in Mammals}.
+\newblock {\em Am. Nat.}, 125(6):873--878, June 1985.
+
+\bibitem{Lindstedt:2002td}
+S~L Lindstedt and P~J Schaeffer.
+\newblock {Use of allometry in predicting anatomical and physiological
+  parameters of mammals.}
+\newblock {\em Lab. Anim.}, 36(1):1--19, January 2002.
+
+\bibitem{estermann}
+Beda~L Estermann, Hans-Rudolf Wettstein, Franz Sutter, and Michael Kreuzer.
+\newblock Nutrient and energy conversion of grass-fed dairy and suckler beef
+  cattle kept indoors and on high altitude pasture.
+\newblock {\em Animal Research}, 50(6):477--493, 2001.
+
+\bibitem{michaletz2014convergence}
+Sean~T Michaletz, Dongliang Cheng, Andrew~J Kerkhoff, and Brian~J Enquist.
+\newblock Convergence of terrestrial plant production across global climate
+  gradients.
+\newblock {\em Nature}, 512(7512):39--43, 2014.
+
+\bibitem{damuth1987interspecific}
+John Damuth.
+\newblock Interspecific allometry of population density in mammals and other
+  animals: the independence of body mass and population energy-use.
+\newblock {\em Biological Journal of the Linnean Society}, 31(3):193--246,
+  1987.
+
+\end{thebibliography}
+
+\end{bibunit}
 

### 2017-03-20 Manuscript/Revision_2017/18.tex
embedded figs

--- 2017-03-09 Manuscript/Revision_2017/17.tex
@@
 %% OPTIONAL MACRO FILES
+\setlength{\abovedisplayskip}{3pt}
+\setlength{\belowdisplayskip}{3pt}
+
+\usepackage[font=small,labelfont=bf]{caption}
 
@@
 Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percentage of body fat.
-For body masses $< 1.75\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.75\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
+For body masses $< 1.748\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.748\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
 \end{abstract}
@@
 \noindent {\small {\bf Significance Statement} %105/120 words
-Energetic investment in somatic maintenance and growth vs$.$ reproduction directly impacts the dynamics of populations among species.
+Energetic investment in somatic maintenance and growth vs.
+reproduction directly impacts the dynamics of populations among species.
 Here, we construct a Nutritional State-structured Model (NSM) to assess the population-level effects of starvation and recovery of a consumer population in a resource-limited environment, and use allometric scaling relationships for mammals to establish all timescales and rates.
@@
 In the supplementary information (SI) we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we describe below.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{fig_FixedPoint2.eps}
+\caption{\small{ The transcritical (dashed) and Hopf bifurcation (solid) as a
+  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
+  bifurcation conditions separate parameter space into unphysical, cyclic,
+  and steady state dynamic regimes.  The colors show the steady state densities for the energetically replete consumers $F^*$, (warmer colors denote higher densities).  
+  % Steady state densities for the energetically deficient consumers $H^*$ (not shown)
+  % scale with those for $F^*$.  
+  \vspace{-2mm}}}
+\label{fig:fp}
+\end{figure}
 
@@
 In the mean-field approximation, in which the consumers and resources are perfectly mixed, their densities evolve according to the rate equations
+
 \begin{align}
@@
 \end{align}
+
 This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see SI for a full derivation and the dimensional form).
@@
 % If $\rho$ exceeds a threshold value, cyclic dynamics will develop as the Hopf bifurcation is approached.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.eps}
+\caption{\small{ The growth trajectory over absolute time of an individual organism as a function of body mass.  
+Initial growth follows the black trajectory to an energetically replete reproductive adult mass $m=\epsilon_\lambda M$. %which we assume is 95\% asymptotic mass $M$.  
+Starvation follows the red trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. 
+Recovery follows the green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires a longer time to reach $\epsilon_\lambda M$.
+%different energetics and
+Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}}
+\label{fig:growth}
+\end{figure}
 
@@
 metazoans
-\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. This relation is derived from the simple balance condition \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}
-\begin{eqnarray}
-\label{balance}
-B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
-\end{eqnarray}
-where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
+\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. This relation is derived from the simple balance condition 
+$B_{0}m^{\eta}=E_{m}\dot{m}+B_{m}m\,,$
+% \begin{eqnarray}
+% \label{balance}
+% B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
+% \end{eqnarray}
+\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy} where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
 the metabolic rate to support an existing unit of mass, and $m$ is the mass
@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometricComb3.eps}
+\caption{\small{ Probability of extinction for a consumer with (\emph{A}) $M=10^2$g, (\emph{B}) $M=10^4$g, and (\emph{C}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
+Note the change in scale for $M=10^4$ and $M=10^6$g.
+Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
+The white points denote the allometrically constrained starvation and recovery rate.}}
+\label{fig:ext}
+\end{figure*}
+
 To determine the starvation rate, $\sigma$, we are interested in the time
@@
 This assumption implies the following simple metabolic balance
-\begin{eqnarray}
-\frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
-\end{eqnarray}
-or
-\begin{eqnarray}
-\frac{dm}{dt}=-\frac{a^{\prime}}{M^{1-\eta}}m
-\end{eqnarray}
+$\dot{m}E_{m}^{\prime}=-B_{m}m$ or $\dot{m}=-a^{\prime}m/M^{1-\eta}$
+% \begin{eqnarray}
+% \frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
+% \end{eqnarray}
+% or
+% \begin{eqnarray}
+% \frac{dm}{dt}=-\frac{a^{\prime}}{M^{1-\eta}}m
+% \end{eqnarray}
 where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
@@
 above energy balance prescribes the mass trajectory of a non-consuming
-organism:
-\begin{eqnarray}
-\label{mt}
-m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}.
-\end{eqnarray}
+organism:$m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}$.
+% \begin{eqnarray}
+% \label{mt}
+% m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}.
+% \end{eqnarray}
 The timescale for starvation is
@@
 
-%CHRIS - MAYBE WE CUT THIS OUT SINCE WE STICK TO MAMMALS IN THIS PAPER
-%It should be noted that we have thus far used mammalian allometry to describe the size-based relations for growth, starvation, and death.
-However, our presentation is general, and other functional forms for $\epsilon$, for example, could be determined for other classes of organisms.
-Considering bacteria, we might expect that starvation or death is defined by the complete digestion of proteins, and in Table \ref{parameter-values} we provide all parameter values for bacteria which we later use as a comparison in our analysis.
-
-%\left(\frac{a}{b}\right)^{\frac{1}{\eta -1}}
-%where $c=(a/b)^{1/(\eta-1)}$.
 Although the rate equations \eqref{eq:system} are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
 Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
-Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
-\\
-%, though this is beyond the current scope of our analysis. %, and we henceforth focus our examination on mammalian species.
+Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.\\
 
@@
 However, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation.
-This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
-%Thus our NSM model predicts that population cycles should be less common for smaller species and more common for larger species, particularly in environments where resources are limiting.%, and
-%2) the decreasing distance to the Hopf bifurcation with lower body sizes is suggestive of a dynamic barrier to the mass of endothermic organisms.
-
-It should be noted that previous studies used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
-However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
-%It is more difficult to ascertain whether a population is truly cycling according to an internal or external dynamic~\cite{Kendall:1999iy} or if fluctuations are due to stochasticity, and this is particularly true for longer-lived organisms.
-Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
-However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.\\ % in agreement with our predictions of more pronounced transient dynamics, given how close these points are to the Hopf bifurcation.On the other hand,
-% \vspace{6mm}
-%, we acknowledge that absence of evidence for cyclic dynamics among longer-lived species is not necessarily evidence of absence.
- %, emphasizing the role of starvation on dynamics.
+This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.\\
+
+% Previous studies used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
+% However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
+% Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
+% However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.\\
 
@@
 This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
+We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig.
+\ref{fig:ext}\emph{A-C}), suggesting that the population dynamics of larger organisms are more sensitive to smaller changes in physiological rates controlling starvation and recovery.
 % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-{\it To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.}\\
+To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.48\textwidth]{fig_FPAllometric2.eps}
+\caption{\small{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
+  body mass. Inset: Resource steady state $R^*$ as a function of consumer body mass.
+  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
+  The data are from Damuth \cite{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
+\citenum{West:2001bv,hou,moses2008rmo}).}}
+\label{fig:mass}
+\end{figure}
+
 %Minimal mass bound
@@
 The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
-Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{damuth1987interspecific,allen2002,enquist1998}.
-The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size (e.g., \cite{damuth1987interspecific,allen2002,enquist1998}).
+Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{Damuth:1987kr,allen2002,enquist1998}.
+The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size~\cite{Damuth:1987kr,allen2002,enquist1998}.
 %Previous studies have focused on t
-This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism ($N^*\propto M^{-3/4}/B_{0}$)  (e.g., \cite{allen2002,enquist1998}).
-This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{damuth1987interspecific,allen2002,enquist1998}.
+This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism, such that $N^*\propto M^{-3/4}/B_{0}$~\cite{allen2002,enquist1998}.
+This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{Damuth:1987kr,allen2002,enquist1998}.
 Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
@@
 The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
-Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the state of starved foragers, is assumed to remain unchanged for the invader population (see SI for detailed derivations of invader rates).
+Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the body composition of starved foragers, is assumed to remain unchanged for the invader population (see SI for detailed derivations of invader rates).
 %There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
@@
 To assess the susceptibility of the resident population to invasion, we determine which consumer has a lower steady-state resource density for a given value of $\chi$, again with the expectation that populations able to survive on lower resource densities have a competitive advantage \cite{tilman1981}.
-We find that for $M\leq 1.75\times10^7$g, having additional body fat ($\chi > 0$) results in lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
-However, for $M> 1.75\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
+We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
+However, for $M> 1.748\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
 
@@
 
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.75\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
+\caption{\small{ Invasion feasibility for organisms with a proportional change in
+  mass $\chi$ against a population with a resident body mass $M$.  The blue
+  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
+  red region denotes values of $\chi$ that result in a mass that is below the
+  starvation threshold and are thus infeasible.
+  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
+  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}}
+\label{fig:invasion}
+\end{figure}
+
+The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
 This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca.
@@
 
-\bibliography{aa_starving3}
-
-% \begin{thebibliography}{10}
-%
-% \bibitem{Martin:1987dl}
-% Martin TE
-% \newblock (1987) {Food as a Limit on Breeding Birds: A Life-History
-%   Perspective}.
-% \newblock \emph{Annu. Rev. Ecol. Syst.} 18:453--487.
-%
-% \bibitem{Kirk:1997cc}
-% Kirk KL
-% \newblock (1997) {Life-History Responses to Variable Environments: Starvation
-%   and Reproduction in Planktonic Rotifers}.
-% \newblock \emph{Ecology} 78:434--441.
-%
-% \bibitem{Kempes:2012hy}
-% Kempes CP, Dutkiewicz S, Follows MJ
-% \newblock (2012) {Growth, metabolic partitioning, and the size of
-%   microorganisms.}
-% \newblock \emph{PNAS} 109:495--500.
-%
-% \bibitem{Mangel:1988uaa}
-% Mangel M, Clark CW
-% \newblock (1988) \emph{{Dynamic Modeling in Behavioral Ecology}}
-% \newblock (Princeton University Press, Princeton).
-%
-% \bibitem{Mangel:2014kz}
-% Mangel M
-% \newblock (2014) {Stochastic dynamic programming illuminates the link between
-%   environment, physiology, and evolution}.
-% \newblock \emph{B. Math. Biol.} 77:857--877.
-%
-% \bibitem{Yeakel:2013hi}
-% Yeakel JD, Dominy NJ, Koch PL, Mangel M
-% \newblock (2014) {Functional morphology, stable isotopes, and human evolution:
-%   a model of consilience}.
-% \newblock \emph{Evolution} 68:190--203.
-%
-% \bibitem{Morris:1987eo}
-% Morris DW
-% \newblock (1987) {Optimal Allocation of Parental Investment}.
-% \newblock \emph{Oikos} 49:332.
-%
-% \bibitem{Tveraa:2003fq}
-% Tveraa T, Fauchald P, Henaug C, Yoccoz NG
-% \newblock (2003) {An examination of a compensatory relationship between food
-%   limitation and predation in semi-domestic reindeer}.
-% \newblock \emph{Oecologia} 137:370--376.
-%
-% \bibitem{Daan:1988va}
-% Daan S, Dijkstra C, Drent R, Meijer T
-% \newblock (1988) \emph{{Food supply and the annual timing of avian
-%   reproduction}}.
-%
-% \bibitem{Jacot:2009dv}
-% Jacot A, Valcu M, van Oers K, Kempenaers B
-% \newblock (2009) {Experimental nest site limitation affects reproductive
-%   strategies and parental investment in a hole-nesting passerine}.
-% \newblock \emph{Animal Behaviour} 77:1075--1083.
-%
-% \bibitem{Stearns:1989ip}
-% Stearns SC
-% \newblock (1989) {Trade-Offs in Life-History Evolution}.
-% \newblock \emph{Funct. Ecol.} 3:259.
-%
-% \bibitem{Barboza:2002in}
-% Barboza P, Jorde D
-% \newblock (2002) {Intermittent fasting during winter and spring affects body
-%   composition and reproduction of a migratory duck}.
-% \newblock \emph{J Comp Physiol B} 172:419--434.
-%
-% \bibitem{Threlkeld:1976ih}
-% Threlkeld ST
-% \newblock (1976) {Starvation and the size structure of zooplankton
-%   communities}.
-% \newblock \emph{Freshwater Biol.} 6:489--496.
-%
-% \bibitem{Weber:1998jg}
-% Weber TP, Ens BJ, Houston AI
-% \newblock (1998) {Optimal avian migration: A dynamic model of fuel stores and
-%   site use}.
-% \newblock \emph{Evolutionary Ecology} 12:377--401.
-%
-% \bibitem{Mduma:1999cp}
-% Mduma SAR, Sinclair ARE, Hilborn R
-% \newblock (1999) {Food regulates the Serengeti wildebeest: a 40-year record}.
-% \newblock \emph{J. Anim. Ecol.} 68:1101--1122.
-%
-% \bibitem{Moore:2014hi}
-% Moore JW, Yeakel JD, Peard D, Lough J, Beere M
-% \newblock (2014) {Life-history diversity and its importance to population
-%   stability and persistence of a migratory fish: steelhead in two large North
-%   American watersheds}.
-% \newblock \emph{J. Anim. Ecol.} 83:1035--1046.
-%
-% \bibitem{Mead:1989dt}
-% Mead RA
-% \newblock (1989) in \emph{Carnivore Behavior, Ecology, and Evolution}
-% \newblock (Springer US, Boston, MA), pp 437--464.
-%
-% \bibitem{Sandell:1990kw}
-% Sandell M
-% \newblock (1990) {The Evolution of Seasonal Delayed Implantation}.
-% \newblock \emph{The Quarterly Review of Biology} 65:23--42.
-%
-% \bibitem{Bulik:1999eo}
-% Bulik CM, {et~al.}
-% \newblock (1999) {Fertility and Reproduction in Women With Anorexia Nervosa}.
-% \newblock \emph{J. Clin. Psychiatry} 60:130--135.
-%
-% \bibitem{Trites:2003cc}
-% Trites AW, Donnelly CP
-% \newblock (2003) {The decline of Steller sea lions Eumetopias jubatus in
-%   Alaska: a review of the nutritional stress hypothesis}.
-% \newblock \emph{Mammal Review} 33:3--28.
-%
-% \bibitem{Glazier:2009hq}
-% Glazier DS
-% \newblock (2009) {Metabolic level and size scaling of rates of respiration and
-%   growth in unicellular organisms}.
-% \newblock \emph{Funct. Ecol.} 23:963--968.
-%
-% \bibitem{Kooi2000}
-% Kooijman SALM
-% \newblock (2000) \emph{{Dynamic Energy and Mass Budgets in Biological Systems}}
-% \newblock (Cambridge).
-%
-% \bibitem{Sousa:2010ez}
-% Sousa T, Domingos T, Poggiale JC, Kooijman SALM
-% \newblock (2010) {Dynamic energy budget theory restores coherence in biology}.
-% \newblock \emph{Philos. T. Roy. Soc. B} 365:3413--3428.
-%
-% \bibitem{Diekmann:2010da}
-% Diekmann O, Metz JAJ
-% \newblock (2010) {How to lift a model for individual behaviour to the
-%   population level?}
-% \newblock \emph{Philos. T. Roy. Soc. B} 365:3523--3530.
-%
-% \bibitem{murdoch:2003}
-% Murdoch WW, Briggs CJ, Nisbet RM
-% \newblock (2003) \emph{{Consumer-resource Dynamics}}, Monographs in population
-%   biology
-% \newblock (Princeton University Press).
-%
-% \bibitem{Benichou:2014wu}
-% Benichou O, Redner S
-% \newblock (2014) {Depletion-Controlled Starvation of a Diffusing Forager}.
-% \newblock \emph{arXiv}.
-%
-% \bibitem{Benichou:2016wl}
-% B{\'e}nichou O, Chupeau M, Redner S
-% \newblock (2016) {Role of Depletion on the Dynamics of a Diffusing Forager}.
-%
-% \bibitem{Chupeau:2016jf}
-% Chupeau M, B{\'e}nichou O, Redner S
-% \newblock (2016) {Universality classes of foraging with resource renewal}.
-% \newblock \emph{Phys. Rev. E} 93:032403.
-%
-% \bibitem{murray2011mathematical}
-% Murray JD
-% \newblock (2011) \emph{{Mathematical Biology: I. An Introduction}},
-%   Interdisciplinary Applied Mathematics
-% \newblock (Springer New York, Melaka Manipal Medical College (Manipal Campus),
-%   International Centre for Health Sciences, Madhav Nagar, Manipal, Udupi
-%   District, Karnataka State, India. nayaksathish@yahoo.com) Vol.{} 110.
-%
-% \bibitem{Strogatz:2008wo}
-% Strogatz SH
-% \newblock (2008) \emph{{Nonlinear Dynamics and Chaos}}, With Applications to
-%   Physics, Biology, Chemistry, and Engineering
-% \newblock (Westview Press).
-%
-% \bibitem{GuckHolmes}
-% Guckenheimer J, Holmes P
-% \newblock (1983) \emph{{Nonlinear oscillations, dynamical systems, and
-%   bifurcations of vector fields}}
-% \newblock (Springer, New York).
-%
-% \bibitem{Gross:2004p2428}
-% Gross T, Feudel U
-% \newblock (2004) {Analytical search for bifurcation surfaces in parameter
-%   space}.
-% \newblock \emph{Physica D} 195:292--302.
-%
-% \bibitem{Hastings:2001jh}
-% Hastings A
-% \newblock (2001) {Transient dynamics and persistence of ecological systems}.
-% \newblock \emph{Ecol. Lett.} 4:215--220.
-%
-% \bibitem{Neubert:1997wk}
-% Neubert M, Caswell H
-% \newblock (1997) {Alternatives to resilience for measuring the responses of
-%   ecological systems to perturbations}.
-% \newblock \emph{Ecology} 78:653--665.
-%
-% \bibitem{Caswell:2005eo}
-% Caswell H, Neubert MG
-% \newblock (2005) {Reactivity and transient dynamics of discrete-time ecological
-%   systems}.
-% \newblock \emph{Journal of Difference Equations and Applications} 11:295--310.
-%
-% \bibitem{Neubert:2009td}
-% Neubert M, Caswell H
-% \newblock (2009) {Detecting reactivity}.
-% \newblock \emph{Ecology}.
-%
-% \bibitem{Yodzis:1992hg}
-% Yodzis P, Innes S
-% \newblock (1992) {Body Size and Consumer-Resource Dynamics}.
-% \newblock \emph{Am. Nat.} 139:1151--1175.
-%
-% \bibitem{Brown:2004wq}
-% Brown J, Gillooly J, Allen A, Savage V, West G
-% \newblock (2004) {Toward a metabolic theory of ecology}.
-% \newblock \emph{Ecology} 85:1771--1789.
-%
-% \bibitem{West:2002it}
-% West GB, Woodruff WH, Brown JH
-% \newblock (2002) {Allometric scaling of metabolic rate from molecules and
-%   mitochondria to cells and mammals.}
-% \newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.
-%
-% \bibitem{DeLong:2010dy}
-% DeLong JP, Okie JG, Moses ME, Sibly RM, Brown JH
-% \newblock (2010) {Shifts in metabolic scaling, production, and efficiency
-%   across major evolutionary transitions of life.}
-% \newblock \emph{PNAS} 107:12941--12945.
-%
-% \bibitem{West:2001bv}
-% West GB, Brown JH, Enquist BJ
-% \newblock (2001) {A general model for ontogenetic growth}.
-% \newblock \emph{Nature} 413:628--631.
-%
-% \bibitem{moses2008rmo}
-% Moses ME, {et~al.}
-% \newblock (2008) {Revisiting a Model of Ontogenetic Growth: Estimating Model
-%   Parameters from Theory and Data}.
-% \newblock \emph{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}
-%   171:632--645.
-%
-% \bibitem{gillooly2002esa}
-% Gillooly JF, Charnov EL, West GB, Savage VM, Brown JH
-% \newblock (2002) {Effects of size and temperature on developmental time}.
-% \newblock \emph{Nature} 417:70--73.
-%
-% \bibitem{hou}
-% Hou C, {et~al.}
-% \newblock (2008) {Energy Uptake and Allocation During Ontogeny}.
-% \newblock \emph{Science} 322:736--739.
-%
-% \bibitem{bettencourt}
-% Bettencourt LMA, Lobo J, Helbing D, Kuhnert C, West GB
-% \newblock (2007) {Growth, innovation, scaling, and the pace of life in cities}.
-% \newblock \emph{Proc. Natl. Acad. Sci. USA} 104:7301--7306.
-%
-% \bibitem{Folland:2008ij}
-% Folland JP, Mc~Cauley TM, Williams AG
-% \newblock (2008) {Allometric scaling of strength measurements to body size}.
-% \newblock \emph{Eur J Appl Physiol} 102:739--745.
-%
-% \bibitem{CalderIII:1983jd}
-% Calder~III WA
-% \newblock (1983) {An allometric approach to population cycles of mammals}.
-% \newblock \emph{J. Theor. Biol.} 100:275--282.
-%
-% \bibitem{Peterson:1984hj}
-% Peterson RO, PAGE RE, DODGE KM
-% \newblock (1984) {Wolves, Moose, and the Allometry of Population Cycles}.
-% \newblock \emph{Science} 224:1350--1352.
-%
-% \bibitem{Krukonis:1991fk}
-% Krukonis G, Schaffer WM
-% \newblock (1991) {Population cycles in mammals and birds: Does periodicity
-%   scale with body size?}
-% \newblock \emph{J. Theor. Biol.} 148:469--493.
-%
-% \bibitem{Hendriks:2012fc}
-% Hendriks AJ, Mulder C
-% \newblock (2012) {Delayed logistic and Rosenzweig{\textendash}MacArthur models
-%   with allometric parameter setting estimate population cycles at lower trophic
-%   levels well}.
-% \newblock \emph{Ecological Complexity} 9:43--54.
-%
-% \bibitem{Kendall:1999iy}
-% Kendall BE, {et~al.}
-% \newblock (1999) {Why do populations cycle? A synthesis of statistical and
-%   mechanistic modeling approaches}.
-% \newblock \emph{Ecology} 80:1789--1805.
-%
-% \bibitem{Hogstedt:2005cr}
-% H{\"o}gstedt G, Seldal T, Breistol A
-% \newblock (2005) {Period length in cyclic animal populations}.
-% \newblock \emph{Ecology} 86:373--378.
-%
-% \bibitem{Kendall:1998hl}
-% Kendall, Prendergast, Bjornstad
-% \newblock (1998) {The macroecology of population dynamics: taxonomic and
-%   biogeographic patterns in population cycles}.
-% \newblock \emph{Ecol. Lett.} 1:160--164.
-%
-% \bibitem{Brown:1993p708}
-% Brown J, Marquet P, Taper M
-% \newblock (1993) {Evolution of body size: consequences of an energetic
-%   definition of fitness}.
-% \newblock \emph{Am. Nat.} 142:573--584.
-%
-% \bibitem{West:1997cg}
-% West GB, Brown JH, Enquist BJ
-% \newblock (1997) {A General Model for the Origin of Allometric Scaling Laws in
-%   Biology}.
-% \newblock \emph{Science} 276:122--126.
-%
-% \bibitem{West:2002ud}
-% West GB, Woodruff WH, Brown JH
-% \newblock (2002) {Allometric scaling of metabolic rate from molecules and
-%   mitochondria to cells and mammals}.
-% \newblock \emph{Proc. Natl. Acad. Sci. USA} 99:2473--2478.
-%
-% \bibitem{Millar:1990p923}
-% Millar J, Hickling G
-% \newblock (1990) { Fasting Endurance and the Evolution of Mammalian Body Size}.
-% \newblock \emph{Funct. Ecol.} 4:5--12.
-%
-% \bibitem{Alroy:1998p1594}
-% Alroy J
-% \newblock (1998) {Cope's rule and the dynamics of body mass evolution in North
-%   American fossil mammals}.
-% \newblock \emph{Science} 280:731.
-%
-% \bibitem{Clauset:2009fh}
-% Clauset A, Redner S
-% \newblock (2009) {Evolutionary Model of Species Body Mass Diversification}.
-% \newblock \emph{Phys. Rev. Lett.} 102:038103.
-%
-% \bibitem{Smith:2010p3442}
-% Smith F, Boyer A, Brown J, Costa D
-% \newblock (2010) {The Evolution of Maximum Body Size of Terrestrial Mammals}.
-% \newblock \emph{Science}.
-%
-% \bibitem{Saarinen:2014br}
-% Saarinen JJ, {et~al.}
-% \newblock (2014) {Patterns of maximum body size evolution in Cenozoic land
-%   mammals: eco-evolutionary processes and abiotic forcing}.
-% \newblock \emph{Proc Biol Sci} 281:20132049--20132049.
-%
-% \bibitem{allen2002}
-% Allen AP, Brown JH, Gillooly JF
-% \newblock (2002) {Global Biodiversity, Biochemical Kinetics, and the
-%   Energetic-Equivalence Rule}.
-% \newblock \emph{Science} 297:1545--1548.
-%
-% \bibitem{enquist1998}
-% Enquist BJ, Brown JH, West GB
-% \newblock (1998) {Allometric scaling of plant energetics and population density
-%   : Abstract : Nature}.
-% \newblock \emph{Nature} 395:163--165.
-%
-% \bibitem{tilman1981}
-% Tilman D
-% \newblock (1981) {Tests of Resource Competition Theory Using Four Species of
-%   Lake Michigan Algae}.
-% \newblock \emph{Ecology} 62:802--815.
-%
-% \bibitem{dutkiewicz2009}
-% Dutkiewicz S, Follows MJ, Bragg JG
-% \newblock (2009) {Modeling the coupling of ocean ecology and biogeochemistry}.
-% \newblock \emph{Global Biogeochem. Cycles} 23:n/a--n/a.
-%
-% \bibitem{barton2010}
-% Barton AD, Dutkiewicz S, Flierl G, Bragg J, Follows MJ
-% \newblock (2010) {Patterns of Diversity in Marine Phytoplankton}.
-% \newblock \emph{Science} 327:1509--1511.
-%
-% \end{thebibliography}
+% \bibliography{aa_starving3}
+\footnotesize{
+\begin{thebibliography}{10}
+
+\bibitem{Martin:1987dl}
+Martin TE
+\newblock (1987) {Food as a limit on breeding birds: A life-history
+  perspective}.
+\newblock \emph{Annu. Rev. Ecol. Syst.} 18:453--487.
+
+\bibitem{Kirk:1997cc}
+Kirk KL
+\newblock (1997) {Life-history responses to variable environments: Starvation
+  and reproduction in planktonic rotifers}.
+\newblock \emph{Ecology} 78:434--441.
+
+\bibitem{Kempes:2012hy}
+Kempes CP, Dutkiewicz S, Follows MJ
+\newblock (2012) {Growth, metabolic partitioning, and the size of
+  microorganisms}.
+\newblock \emph{Proc. Natl. Acad. Sci. USA} 109:495--500.
+
+\bibitem{Mangel:1988uaa}
+Mangel M, Clark CW
+\newblock (1988) \emph{{Dynamic Modeling in Behavioral Ecology}}
+\newblock (Princeton University Press, Princeton).
+
+\bibitem{Mangel:2014kz}
+Mangel M
+\newblock (2014) {Stochastic dynamic programming illuminates the link between
+  environment, physiology, and evolution}.
+\newblock \emph{B. Math. Biol.} 77:857--877.
+
+\bibitem{Yeakel:2013hi}
+Yeakel JD, Dominy NJ, Koch PL, Mangel M
+\newblock (2014) {Functional morphology, stable isotopes, and human evolution:
+  a model of consilience}.
+\newblock \emph{Evolution} 68:190--203.
+
+\bibitem{Morris:1987eo}
+Morris DW
+\newblock (1987) {Optimal allocation of parental investment}.
+\newblock \emph{Oikos} 49:332--339.
+
+\bibitem{Tveraa:2003fq}
+Tveraa T, Fauchald P, Henaug C, Yoccoz NG
+\newblock (2003) {An examination of a compensatory relationship between food
+  limitation and predation in semi-domestic reindeer}.
+\newblock \emph{Oecologia} 137:370--376.
+
+\bibitem{Daan:1988va}
+Daan S, Dijkstra C, Drent R, Meijer T
+\newblock (1989) in \emph{Acta XIX Congressus Internationalis Ornithologici,
+  Volume I: Proceedings XIX International Ornithological Congress, 1986,
+  Ottawa}, ed{} Ouellet H
+\newblock (Proceedings XIX International Ornithological Congress, Ottawa), pp
+  392--407.
+
+\bibitem{Jacot:2009dv}
+Jacot A, Valcu M, van Oers K, Kempenaers B
+\newblock (2009) {Experimental nest site limitation affects reproductive
+  strategies and parental investment in a hole-nesting passerine}.
+\newblock \emph{Animal Behaviour} 77:1075--1083.
+
+\bibitem{Stearns:1989ip}
+Stearns SC
+\newblock (1989) {Trade-offs in life-history evolution}.
+\newblock \emph{Funct. Ecol.} 3:259.
+
+\bibitem{Barboza:2002in}
+Barboza P, Jorde D
+\newblock (2002) {Intermittent fasting during winter and spring affects body
+  composition and reproduction of a migratory duck}.
+\newblock \emph{J Comp Physiol B} 172:419--434.
+
+\bibitem{Threlkeld:1976ih}
+Threlkeld ST
+\newblock (1976) {Starvation and the size structure of zooplankton
+  communities}.
+\newblock \emph{Freshwater Biol.} 6:489--496.
+
+\bibitem{Weber:1998jg}
+Weber TP, Ens BJ, Houston AI
+\newblock (1998) {Optimal avian migration: A dynamic model of fuel stores and
+  site use}.
+\newblock \emph{Evolutionary Ecology} 12:377--401.
+
+\bibitem{Mduma:1999cp}
+Mduma SAR, Sinclair ARE, Hilborn R
+\newblock (1999) {Food regulates the Serengeti wildebeest: a 40-year record}.
+\newblock \emph{J. Anim. Ecol.} 68:1101--1122.
+
+\bibitem{Moore:2014hi}
+Moore JW, Yeakel JD, Peard D, Lough J, Beere M
+\newblock (2014) {Life-history diversity and its importance to population
+  stability and persistence of a migratory fish: steelhead in two large North
+  American watersheds}.
+\newblock \emph{J. Anim. Ecol.} 83:1035--1046.
+
+\bibitem{Mead:1989dt}
+Mead RA
+\newblock (1989) in \emph{Carnivore Behavior, Ecology, and Evolution}, ed{}
+  Gittleman JL
+\newblock (Springer US, Ithaca), pp 437--464.
+
+\bibitem{Sandell:1990kw}
+Sandell M
+\newblock (1990) {The evolution of seasonal delayed implantation}.
+\newblock \emph{Q Rev Biol} 65:23--42.
+
+\bibitem{Bulik:1999eo}
+Bulik CM, {et~al.}
+\newblock (1999) {Fertility and reproduction in women with anorexia nervosa}.
+\newblock \emph{J. Clin. Psychiat.} 60:130--135.
+
+\bibitem{Trites:2003cc}
+Trites AW, Donnelly CP
+\newblock (2003) {The decline of Steller sea lions Eumetopias jubatus in
+  Alaska: a review of the nutritional stress hypothesis}.
+\newblock \emph{Mammal Rev.} 33:3--28.
+
+\bibitem{Glazier:2009hq}
+Glazier DS
+\newblock (2009) {Metabolic level and size scaling of rates of respiration and
+  growth in unicellular organisms}.
+\newblock \emph{Funct. Ecol.} 23:963--968.
+
+\bibitem{Kooi2000}
+Kooijman SALM
+\newblock (2000) \emph{{Dynamic Energy and Mass Budgets in Biological Systems}}
+\newblock (Cambridge).
+
+\bibitem{Sousa:2010ez}
+Sousa T, Domingos T, Poggiale JC, Kooijman SALM
+\newblock (2010) {Dynamic energy budget theory restores coherence in biology}.
+\newblock \emph{Philos. T. Roy. Soc. B} 365:3413--3428.
+
+\bibitem{Diekmann:2010da}
+Diekmann O, Metz JAJ
+\newblock (2010) {How to lift a model for individual behaviour to the
+  population level?}
+\newblock \emph{Philos. T. Roy. Soc. B} 365:3523--3530.
+
+\bibitem{murdoch:2003}
+Murdoch WW, Briggs CJ, Nisbet RM
+\newblock (2003) \emph{{Consumer-resource Dynamics}}, Monographs in population
+  biology
+\newblock (Princeton University Press, Princeton) Vol.{}~36.
+
+\bibitem{Benichou:2014wu}
+Benichou O, Redner S
+\newblock (2014) {Depletion-Controlled Starvation of a Diffusing Forager}.
+\newblock \emph{arXiv} pp 1--5.
+
+\bibitem{Benichou:2016wl}
+B{\'e}nichou O, Chupeau M, Redner S
+\newblock (2016) {Role of depletion on the dynamics of a diffusing forager}.
+\newblock \emph{Journal of Physics A: {\ldots}}.
+
+\bibitem{Chupeau:2016jf}
+Chupeau M, B{\'e}nichou O, Redner S
+\newblock (2016) {Universality classes of foraging with resource renewal}.
+\newblock \emph{Phys. Rev. E} 93:032403.
+
+\bibitem{Persson:1998hz}
+Persson L, Leonardsson K, De~Roos AM, Gyllenberg M, Christensen B
+\newblock (1998) {Ontogenetic scaling of foraging rates and the dynamics of a
+  size-structured consumer-resource model}.
+\newblock \emph{Theor Popul Biol} 54:270--293.
+
+\bibitem{murray2011mathematical}
+Murray JD
+\newblock (2011) \emph{{Mathematical Biology: I. An Introduction}},
+  Interdisciplinary Applied Mathematics
+\newblock (Springer New York) Vol.{} 110.
+
+\bibitem{Strogatz:2008wo}
+Strogatz SH
+\newblock (2008) \emph{{Nonlinear Dynamics and Chaos: With Applications to
+  Physics, Biology, Chemistry, and Engineering}}, Studies in nonlinearity
+\newblock (Westview Press, Boulder).
+
+\bibitem{GuckHolmes}
+Guckenheimer J, Holmes P
+\newblock (1983) \emph{{Nonlinear Oscillations, Dynamical Systems, and
+  Bifurcations of Vector Fields}}
+\newblock (Springer, New York).
+
+\bibitem{Gross:2004p2428}
+Gross T, Feudel U
+\newblock (2004) {Analytical search for bifurcation surfaces in parameter
+  space}.
+\newblock \emph{Physica D} 195:292--302.
+
+\bibitem{Hastings:2001jh}
+Hastings A
+\newblock (2001) {Transient dynamics and persistence of ecological systems}.
+\newblock \emph{Ecol. Lett.} 4:215--220.
+
+\bibitem{Neubert:1997wk}
+Neubert M, Caswell H
+\newblock (1997) {Alternatives to resilience for measuring the responses of
+  ecological systems to perturbations}.
+\newblock \emph{Ecology} 78:653--665.
+
+\bibitem{Caswell:2005eo}
+Caswell H, Neubert MG
+\newblock (2005) {Reactivity and transient dynamics of discrete-time ecological
+  systems}.
+\newblock \emph{J Differ Equ Appl} 11:295--310.
+
+\bibitem{Neubert:2009td}
+Neubert M, Caswell H
+\newblock (2009) {Detecting reactivity}.
+\newblock \emph{Ecology} 90:2683--2688.
+
+\bibitem{Yodzis:1992hg}
+Yodzis P, Innes S
+\newblock (1992) {Body size and consumer-resource dynamics}.
+\newblock \emph{Am. Nat.} 139:1151--1175.
+
+\bibitem{Brown:2004wq}
+Brown J, Gillooly J, Allen A, Savage V, West G
+\newblock (2004) {Toward a metabolic theory of ecology}.
+\newblock \emph{Ecology} 85:1771--1789.
+
+\bibitem{West:2002it}
+West GB, Woodruff WH, Brown JH
+\newblock (2002) {Allometric scaling of metabolic rate from molecules and
+  mitochondria to cells and mammals.}
+\newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.
+
+\bibitem{DeLong:2010dy}
+DeLong JP, Okie JG, Moses ME, Sibly RM, Brown JH
+\newblock (2010) {Shifts in metabolic scaling, production, and efficiency
+  across major evolutionary transitions of life}.
+\newblock \emph{Proc. Natl. Acad. Sci. USA} 107:12941--12945.
+
+\bibitem{West:2001bv}
+West GB, Brown JH, Enquist BJ
+\newblock (2001) {A general model for ontogenetic growth}.
+\newblock \emph{Nature} 413:628--631.
+
+\bibitem{moses2008rmo}
+Moses ME, {et~al.}
+\newblock (2008) {Revisiting a Model of Ontogenetic Growth: Estimating Model
+  Parameters from Theory and Data.}
+\newblock \emph{Am. Nat.} 171:632--645.
+
+\bibitem{gillooly2002esa}
+Gillooly JF, Charnov EL, West GB, Savage VM, Brown JH
+\newblock (2002) {Effects of size and temperature on developmental time}.
+\newblock \emph{Nature} 417:70--73.
+
+\bibitem{hou}
+Hou C, {et~al.}
+\newblock (2008) {Energy uptake and allocation during ontogeny}.
+\newblock \emph{Science} 322:736--739.
+
+\bibitem{bettencourt}
+Bettencourt LMA, Lobo J, Helbing D, Kuhnert C, West GB
+\newblock (2007) {Growth, innovation, scaling, and the pace of life in cities}.
+\newblock \emph{Proc. Natl. Acad. Sci. USA} 104:7301--7306.
+
+\bibitem{Folland:2008ij}
+Folland JP, Mc~Cauley TM, Williams AG
+\newblock (2008) {Allometric scaling of strength measurements to body size}.
+\newblock \emph{Eur J Appl Physiol} 102:739--745.
+
+\bibitem{CalderIII:1983jd}
+Calder~III WA
+\newblock (1983) {An allometric approach to population cycles of mammals}.
+\newblock \emph{J. Theor. Biol.} 100:275--282.
+
+\bibitem{Peterson:1984hj}
+Peterson RO, Page RE, Dodge KM
+\newblock (1984) {Wolves, moose, and the allometry of population cycles}.
+\newblock \emph{Science} 224:1350--1352.
+
+\bibitem{Krukonis:1991fk}
+Krukonis G, Schaffer WM
+\newblock (1991) {Population cycles in mammals and birds: Does periodicity
+  scale with body size?}
+\newblock \emph{J. Theor. Biol.} 148:469--493.
+
+\bibitem{Hendriks:2012fc}
+Hendriks AJ, Mulder C
+\newblock (2012) {Delayed logistic and Rosenzweig{\textendash}MacArthur models
+  with allometric parameter setting estimate population cycles at lower trophic
+  levels well}.
+\newblock \emph{Ecol Complex} 9:43--54.
+
+\bibitem{Kendall:1999iy}
+Kendall BE, {et~al.}
+\newblock (1999) {Why do populations cycle? A synthesis of statistical and
+  mechanistic modeling approaches}.
+\newblock \emph{Ecology} 80:1789--1805.
+
+\bibitem{Hogstedt:2005cr}
+H{\"o}gstedt G, Seldal T, Breistol A
+\newblock (2005) {Period length in cyclic animal populations}.
+\newblock \emph{Ecology} 86:373--378.
+
+\bibitem{Kendall:1998hl}
+Kendall, Prendergast, Bjornstad
+\newblock (1998) {The macroecology of population dynamics: taxonomic and
+  biogeographic patterns in population cycles}.
+\newblock \emph{Ecol. Lett.} 1:160--164.
+
+\bibitem{Brown:1993p708}
+Brown J, Marquet P, Taper M
+\newblock (1993) {Evolution of body size: consequences of an energetic
+  definition of fitness}.
+\newblock \emph{Am. Nat.} 142:573--584.
+
+\bibitem{West:1997cg}
+West GB, Brown JH, Enquist BJ
+\newblock (1997) {A general model for the origin of allometric scaling laws in
+  biology}.
+\newblock \emph{Science} 276:122--126.
+
+\bibitem{West:2002ud}
+West GB, Woodruff WH, Brown JH
+\newblock (2002) {Allometric scaling of metabolic rate from molecules and
+  mitochondria to cells and mammals.}
+\newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.
+
+\bibitem{Millar:1990p923}
+Millar J, Hickling G
+\newblock (1990) {Fasting endurance and the evolution of mammalian body size}.
+\newblock \emph{Funct. Ecol.} 4:5--12.
+
+\bibitem{Alroy:1998p1594}
+Alroy J
+\newblock (1998) {Cope's rule and the dynamics of body mass evolution in North
+  American fossil mammals}.
+\newblock \emph{Science} 280:731--734.
+
+\bibitem{Clauset:2009fh}
+Clauset A, Redner S
+\newblock (2009) {Evolutionary model of species body mass diversification}.
+\newblock \emph{Phys. Rev. Lett.} 102:038103.
+
+\bibitem{Smith:2010p3442}
+Smith FA, {et~al.}
+\newblock (2010) {The evolution of maximum body size of terrestrial mammals}.
+\newblock \emph{Science} 330:1216--1219.
+
+\bibitem{Saarinen:2014br}
+Saarinen JJ, {et~al.}
+\newblock (2014) {Patterns of maximum body size evolution in Cenozoic land
+  mammals: Eco-evolutionary processes and abiotic forcing}.
+\newblock \emph{Proc Biol Sci} 281:20132049.
+
+\bibitem{Damuth:1987kr}
+Damuth J
+\newblock (1987) {Interspecific allometry of population density in mammals and
+  other animals: the independence of body mass and population energy-use}.
+\newblock \emph{Biol. J. Linn. Soc.} 31:193--246.
+
+\bibitem{allen2002}
+Allen AP, Brown JH, Gillooly JF
+\newblock (2002) {Global biodiversity, biochemical kinetics, and the
+  energetic-equivalence rule}.
+\newblock \emph{Science} 297:1545--1548.
+
+\bibitem{enquist1998}
+Enquist BJ, Brown JH, West GB
+\newblock (1998) {Allometric scaling of plant energetics and population
+  density}.
+\newblock \emph{Nature} 395:163--165.
+
+\bibitem{Kempes:2016}
+Kempes CP, Wang L, Amend JP, Doyle J, Hoehler T
+\newblock (2016) {Evolutionary tradeoffs in cellular composition across diverse
+  bacteria}.
+\newblock \emph{ISME J} 10:2145--2157.
+
+\bibitem{tilman1981}
+Tilman D
+\newblock (1981) {Tests of resource competition theory using four species of
+  lake michigan algae}.
+\newblock \emph{Ecology} 62:802--815.
+
+\bibitem{dutkiewicz2009}
+Dutkiewicz S, Follows MJ, Bragg JG
+\newblock (2009) {Modeling the coupling of ocean ecology and biogeochemistry}.
+\newblock \emph{Global Biogeochem. Cycles} 23:1--15.
+
+\bibitem{barton2010}
+Barton AD, Dutkiewicz S, Flierl G, Bragg J, Follows MJ
+\newblock (2010) {Patterns of diversity in marine phytoplankton}.
+\newblock \emph{Science} 327:1509--1511.
+
+\end{thebibliography}
+}
 
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_FixedPoint.pdf}
-\caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
-  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-  bifurcation conditions separate parameter space into unphysical, cyclic,
-  and steady state dynamic regimes.  The colors show the steady
-  state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
-  replete consumers $F^*$, (warmer colors denote higher densities).  Steady
-  state densities for the energetically deficient consumers $H^*$ (not shown)
-  scale with those for $F^*$.  }
-\label{fig:fp}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-\caption{ The growth trajectory over absolute time of an individual organism
-  as a function of body mass.  Initial growth follows the black trajectory to
-  an energetically replete reproductive adult mass $m=\epsilon_\lambda M$ which we assume is 95\% asymptotic mass $M$.  Starvation follows the red
-  trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. Recovery follows the
-  green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires different energetics and a longer time to reach $\epsilon_\lambda M$.
-Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}
-\label{fig:growth}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_Rates.pdf}
-\caption{
-Allometrically constrained starvation rate $\sigma=1/t_{\sigma}$ (red), mortality $\mu=1/t_{\mu}$ (red) and recovery rate $\rho=1/t_{\rho}$ (green) relative to the reproductive rate $\lambda=1/t_{\lambda}$ (black) as a function of body mass (see Equations \ref{t1}, \ref{rhotimescale}, \ref{eq:sigma}, and \ref{mutimescale}).
-The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
-Mean values $\pm 25\%$ variation are shown by the shaded region for each rate.
-}
-\label{fig:gvs}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.5\textwidth]{fig_DataHopf.pdf}
-\caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
-  allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
-  a function of different mammalian body sizes: $M=P\times10^2$g (blue) and
-  $M=P\times10^6$g (green), where $P$ is a random uniform variable in $[1,9]$.
-  Points denote realized values of $\sigma$ and $\rho$ given the drawn values for $M$.
-  }
-\label{fig:hopf}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=1\textwidth]{fig_ExtinctionAllometricComb3.png}
-\caption{ Probability of extinction for a consumer with (from left to right) $M=10^2,~10^4,~10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
-Note the change in scale for $M=10^4,~10^6$g.
-Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
-The white points denote the allometrically constrained starvation and recovery rate.}
-\label{fig:ext}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_FPAllometric.pdf}
-\caption{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-  body mass.
-  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.
-The data are from Damuth \cite{damuth1987interspecific} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
-\cite{West:2001bv,hou,moses2008rmo}).}
-\label{fig:mass}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.75\textwidth]{fig_Invasion.pdf}
-\caption{ Invasion feasibility for organisms with a proportional change in
-  mass $\chi$ against a population with a resident body mass $M$.  The blue
-  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
-  red region denotes values of $\chi$ that result in a mass that is below the
-  starvation threshold and is thus infeasible.
-  Arrows point to the predicted optimal mass $M_{\rm opt}=1.75\times 10^7$, which serves as an evolutionary attractor for body mass.
-  The black wedge points to the largest body mass known for terrestrial mammals at $7.74\times10^7$g~\cite{Smith:2010p3442}.}
-\label{fig:invasion}
-\end{figure*}
+% 
+% \begin{figure*}
+% \centering
+% \includegraphics[width=0.5\textwidth]{fig_Rates.eps}
+% \caption{
+% Allometrically constrained starvation rate $\sigma=1/t_{\sigma}$ (red), mortality $\mu=1/t_{\mu}$ (blue) and recovery rate $\rho=1/t_{\rho}$ (green) relative to the reproductive rate $\lambda=1/t_{\lambda}$ (black) as a function of body mass (see Equations \ref{t1}, \ref{rhotimescale}, \ref{eq:sigma}, and \ref{mutimescale}).
+% The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
+% Mean values $\pm 25\%$ variation are shown by the shaded region for each rate.
+% }
+% \label{fig:gvs}
+% \end{figure*}
+
+% 
+% \begin{figure*}
+% \centering
+% \includegraphics[width=0.5\textwidth]{fig_DataHopf.eps}
+% \caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
+%   allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
+%   a function of different mammalian body sizes: $M=P\times10^2$g (blue) and
+%   $M=P\times10^6$g (green), where $P$ is a random uniform variable in $[1,9]$.
+%   Points denote realized values of $\sigma$ and $\rho$ given the drawn values for $M$.
+%   }
+% \label{fig:hopf}
+% \end{figure*}
 

### 2017-03-21 Manuscript/Revision_2017/18.tex
minor edits to figs and orientation

@@
 
-\bibitem{CalderIII:1983jd}
-Calder~III WA
-\newblock (1983) {An allometric approach to population cycles of mammals}.
-\newblock \emph{J. Theor. Biol.} 100:275--282.
-
-\bibitem{Peterson:1984hj}
-Peterson RO, Page RE, Dodge KM
-\newblock (1984) {Wolves, moose, and the allometry of population cycles}.
-\newblock \emph{Science} 224:1350--1352.
-
-\bibitem{Krukonis:1991fk}
-Krukonis G, Schaffer WM
-\newblock (1991) {Population cycles in mammals and birds: Does periodicity
-  scale with body size?}
-\newblock \emph{J. Theor. Biol.} 148:469--493.
-
-\bibitem{Hendriks:2012fc}
-Hendriks AJ, Mulder C
-\newblock (2012) {Delayed logistic and Rosenzweig{\textendash}MacArthur models
-  with allometric parameter setting estimate population cycles at lower trophic
-  levels well}.
-\newblock \emph{Ecol Complex} 9:43--54.
-
-\bibitem{Kendall:1999iy}
-Kendall BE, {et~al.}
-\newblock (1999) {Why do populations cycle? A synthesis of statistical and
-  mechanistic modeling approaches}.
-\newblock \emph{Ecology} 80:1789--1805.
-
-\bibitem{Hogstedt:2005cr}
-H{\"o}gstedt G, Seldal T, Breistol A
-\newblock (2005) {Period length in cyclic animal populations}.
-\newblock \emph{Ecology} 86:373--378.
-
-\bibitem{Kendall:1998hl}
-Kendall, Prendergast, Bjornstad
-\newblock (1998) {The macroecology of population dynamics: taxonomic and
-  biogeographic patterns in population cycles}.
-\newblock \emph{Ecol. Lett.} 1:160--164.
-
 \bibitem{Brown:1993p708}
@@
 \end{thebibliography}
+
 }

### 2017-03-24 Manuscript/Revision_2017/18.tex
final version of ms

@@
 
-\usepackage[font=large]{caption}
+% \usepackage[font=large]{caption}
 \usepackage{subfigure}
@@
 \usepackage[left]{lineno}
-\usepackage[comma,sort&compress]{natbib}
 
@@
 %% OPTIONAL MACRO FILES
-\setlength{\abovedisplayskip}{3pt}
-\setlength{\belowdisplayskip}{3pt}
-
-\usepackage[font=small,labelfont=bf]{caption}
+% \usepackage[comma,sort&compress]{natbib}
+% \setlength{\abovedisplayskip}{3pt}
+% \setlength{\belowdisplayskip}{3pt}
+% 
+% \usepackage[font=small,labelfont=bf]{caption}
+% 
+% \newcommand{\sid}[1]{\textcolor{red}{\bf [#1]}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\sid}[1]{\textcolor{red}{\bf [#1]}}
 
@@
   % scale with those for $F^*$.  
-  \vspace{-2mm}}}
-\label{fig:fp}
+}\label{fig:fp}}
 \end{figure}
@@
 \begin{split}
-\frac{dF}{dt} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
-\frac{dH}{dt} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-\frac{dR}{dt} &= \alpha \left(1-R\right)R -\left(\rho R+\delta\right)H-\beta F
+\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
+\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
+\dot{R} &= \alpha \left(1-R\right)R -\left(\rho R+\delta\right)H-\beta F
 \end{split}
@@
 This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see SI for a full derivation and the dimensional form).
-Notice that the total consumer density $F+H$ evolves according to $\frac{dF}{dt}+\frac{dH}{dt}=\lambda F-\mu H$.
+Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
 This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\cite{murray2011mathematical}, except that the resource density does not appear in the growth term.
@@
 %different energetics and
-Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}}
-\label{fig:growth}
+Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}\label{fig:growth}}
 \end{figure}
@@
 where we will define values of $\epsilon$ to describe a set of rates within our model.
-For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$; \citenum{West:2001bv}).
+For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$; \cite{West:2001bv}).
 The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$, this rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
@@
 and recovery by the timescales required to reach, or return from, specific
-fractions of the replete-state mass (Fig. \ref{fig:gvs}; see SI, Table I for parameterizations).  We define
+fractions of the replete-state mass (see SI, Table I for parameterizations).  We define
 $m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction of
@@
 Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
-The white points denote the allometrically constrained starvation and recovery rate.}}
-\label{fig:ext}
+The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
 \end{figure*}
@@
 feature is that $\sigma$ does not have a simple scaling dependence on
-$\lambda$ (Fig. \ref{fig:gvs}), which is important for the dynamics that we
+$\lambda$, which is important for the dynamics that we
 later discuss.
@@
 
-\noindent {\bf  Stabilizing effects of allometric constraints} \\
-As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
-Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values (Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.
-We thus find that the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
-Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
-These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
+% \noindent {\bf  Stabilizing effects of allometric constraints} \\
+% As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
+% Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
+% Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
+%Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
+% These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
 However, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation.
-This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.\\
+% This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.\\
 
@@
 \noindent {\bf Extinction risk}\\
-Within our model, higher rates of starvation result in a larger flux of the population to the hungry state.
+As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
+Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
+Indeed, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation. 
+This suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
+% Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
+
+Higher rates of starvation result in a larger flux of the population to the hungry state.
 In this state, reproduction is absent, thus increasing the likelihood of extinction.
 From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
-We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
+We therefore examine the competing effects of cyclic dynamics vs. changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
 To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
@@
   The data are from Damuth \cite{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
-\citenum{West:2001bv,hou,moses2008rmo}).}}
-\label{fig:mass}
+\cite{West:2001bv,hou,moses2008rmo}).}\label{fig:mass}}
 \end{figure}
@@
 The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
+Similar predictions have been made for carnivore populations using alternative consumer-resource models \cite{DeLong:2012kw}.
 Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{Damuth:1987kr,allen2002,enquist1998}.
@@
 This result is remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to energy equivalence.
+
 Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
@@
 The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
-\ref{fig:mass}\emph{C}), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
+\ref{fig:mass}\emph{A}, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
 Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but does not offer a within-lineage mechanism by which larger body sizes are selected for.
@@
 To assess the susceptibility of the resident population to invasion, we determine which consumer has a lower steady-state resource density for a given value of $\chi$, again with the expectation that populations able to survive on lower resource densities have a competitive advantage \cite{tilman1981}.
-We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
+We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population (Fig.
+\ref{fig:invasion}).
 However, for $M> 1.748\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
@@
   Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
-  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}}
-\label{fig:invasion}
+  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}\label{fig:invasion}}
 \end{figure}
@@
 %
-% \begin{figure}[h]
+% \begin{figure}[h][h]
 % 	\centering
@@
 
-% \bibliography{aa_starving3}
-\footnotesize{
-\begin{thebibliography}{10}
-
-\bibitem{Martin:1987dl}
-Martin TE
-\newblock (1987) {Food as a limit on breeding birds: A life-history
-  perspective}.
-\newblock \emph{Annu. Rev. Ecol. Syst.} 18:453--487.
-
-\bibitem{Kirk:1997cc}
-Kirk KL
-\newblock (1997) {Life-history responses to variable environments: Starvation
-  and reproduction in planktonic rotifers}.
-\newblock \emph{Ecology} 78:434--441.
-
-\bibitem{Kempes:2012hy}
-Kempes CP, Dutkiewicz S, Follows MJ
-\newblock (2012) {Growth, metabolic partitioning, and the size of
-  microorganisms}.
-\newblock \emph{Proc. Natl. Acad. Sci. USA} 109:495--500.
-
-\bibitem{Mangel:1988uaa}
-Mangel M, Clark CW
-\newblock (1988) \emph{{Dynamic Modeling in Behavioral Ecology}}
-\newblock (Princeton University Press, Princeton).
-
-\bibitem{Mangel:2014kz}
-Mangel M
-\newblock (2014) {Stochastic dynamic programming illuminates the link between
-  environment, physiology, and evolution}.
-\newblock \emph{B. Math. Biol.} 77:857--877.
-
-\bibitem{Yeakel:2013hi}
-Yeakel JD, Dominy NJ, Koch PL, Mangel M
-\newblock (2014) {Functional morphology, stable isotopes, and human evolution:
-  a model of consilience}.
-\newblock \emph{Evolution} 68:190--203.
-
-\bibitem{Morris:1987eo}
-Morris DW
-\newblock (1987) {Optimal allocation of parental investment}.
-\newblock \emph{Oikos} 49:332--339.
-
-\bibitem{Tveraa:2003fq}
-Tveraa T, Fauchald P, Henaug C, Yoccoz NG
-\newblock (2003) {An examination of a compensatory relationship between food
-  limitation and predation in semi-domestic reindeer}.
-\newblock \emph{Oecologia} 137:370--376.
-
-\bibitem{Daan:1988va}
-Daan S, Dijkstra C, Drent R, Meijer T
-\newblock (1989) in \emph{Acta XIX Congressus Internationalis Ornithologici,
-  Volume I: Proceedings XIX International Ornithological Congress, 1986,
-  Ottawa}, ed{} Ouellet H
-\newblock (Proceedings XIX International Ornithological Congress, Ottawa), pp
-  392--407.
-
-\bibitem{Jacot:2009dv}
-Jacot A, Valcu M, van Oers K, Kempenaers B
-\newblock (2009) {Experimental nest site limitation affects reproductive
-  strategies and parental investment in a hole-nesting passerine}.
-\newblock \emph{Animal Behaviour} 77:1075--1083.
-
-\bibitem{Stearns:1989ip}
-Stearns SC
-\newblock (1989) {Trade-offs in life-history evolution}.
-\newblock \emph{Funct. Ecol.} 3:259.
-
-\bibitem{Barboza:2002in}
-Barboza P, Jorde D
-\newblock (2002) {Intermittent fasting during winter and spring affects body
-  composition and reproduction of a migratory duck}.
-\newblock \emph{J Comp Physiol B} 172:419--434.
-
-\bibitem{Threlkeld:1976ih}
-Threlkeld ST
-\newblock (1976) {Starvation and the size structure of zooplankton
-  communities}.
-\newblock \emph{Freshwater Biol.} 6:489--496.
-
-\bibitem{Weber:1998jg}
-Weber TP, Ens BJ, Houston AI
-\newblock (1998) {Optimal avian migration: A dynamic model of fuel stores and
-  site use}.
-\newblock \emph{Evolutionary Ecology} 12:377--401.
-
-\bibitem{Mduma:1999cp}
-Mduma SAR, Sinclair ARE, Hilborn R
-\newblock (1999) {Food regulates the Serengeti wildebeest: a 40-year record}.
-\newblock \emph{J. Anim. Ecol.} 68:1101--1122.
-
-\bibitem{Moore:2014hi}
-Moore JW, Yeakel JD, Peard D, Lough J, Beere M
-\newblock (2014) {Life-history diversity and its importance to population
-  stability and persistence of a migratory fish: steelhead in two large North
-  American watersheds}.
-\newblock \emph{J. Anim. Ecol.} 83:1035--1046.
-
-\bibitem{Mead:1989dt}
-Mead RA
-\newblock (1989) in \emph{Carnivore Behavior, Ecology, and Evolution}, ed{}
-  Gittleman JL
-\newblock (Springer US, Ithaca), pp 437--464.
-
-\bibitem{Sandell:1990kw}
-Sandell M
-\newblock (1990) {The evolution of seasonal delayed implantation}.
-\newblock \emph{Q Rev Biol} 65:23--42.
-
-\bibitem{Bulik:1999eo}
-Bulik CM, {et~al.}
-\newblock (1999) {Fertility and reproduction in women with anorexia nervosa}.
-\newblock \emph{J. Clin. Psychiat.} 60:130--135.
-
-\bibitem{Trites:2003cc}
-Trites AW, Donnelly CP
-\newblock (2003) {The decline of Steller sea lions Eumetopias jubatus in
-  Alaska: a review of the nutritional stress hypothesis}.
-\newblock \emph{Mammal Rev.} 33:3--28.
-
-\bibitem{Glazier:2009hq}
-Glazier DS
-\newblock (2009) {Metabolic level and size scaling of rates of respiration and
-  growth in unicellular organisms}.
-\newblock \emph{Funct. Ecol.} 23:963--968.
-
-\bibitem{Kooi2000}
-Kooijman SALM
-\newblock (2000) \emph{{Dynamic Energy and Mass Budgets in Biological Systems}}
-\newblock (Cambridge).
-
-\bibitem{Sousa:2010ez}
-Sousa T, Domingos T, Poggiale JC, Kooijman SALM
-\newblock (2010) {Dynamic energy budget theory restores coherence in biology}.
-\newblock \emph{Philos. T. Roy. Soc. B} 365:3413--3428.
-
-\bibitem{Diekmann:2010da}
-Diekmann O, Metz JAJ
-\newblock (2010) {How to lift a model for individual behaviour to the
-  population level?}
-\newblock \emph{Philos. T. Roy. Soc. B} 365:3523--3530.
-
-\bibitem{murdoch:2003}
-Murdoch WW, Briggs CJ, Nisbet RM
-\newblock (2003) \emph{{Consumer-resource Dynamics}}, Monographs in population
-  biology
-\newblock (Princeton University Press, Princeton) Vol.{}~36.
-
-\bibitem{Benichou:2014wu}
-Benichou O, Redner S
-\newblock (2014) {Depletion-Controlled Starvation of a Diffusing Forager}.
-\newblock \emph{arXiv} pp 1--5.
-
-\bibitem{Benichou:2016wl}
-B{\'e}nichou O, Chupeau M, Redner S
-\newblock (2016) {Role of depletion on the dynamics of a diffusing forager}.
-\newblock \emph{Journal of Physics A: {\ldots}}.
-
-\bibitem{Chupeau:2016jf}
-Chupeau M, B{\'e}nichou O, Redner S
-\newblock (2016) {Universality classes of foraging with resource renewal}.
-\newblock \emph{Phys. Rev. E} 93:032403.
-
-\bibitem{Persson:1998hz}
-Persson L, Leonardsson K, De~Roos AM, Gyllenberg M, Christensen B
-\newblock (1998) {Ontogenetic scaling of foraging rates and the dynamics of a
-  size-structured consumer-resource model}.
-\newblock \emph{Theor Popul Biol} 54:270--293.
-
-\bibitem{murray2011mathematical}
-Murray JD
-\newblock (2011) \emph{{Mathematical Biology: I. An Introduction}},
-  Interdisciplinary Applied Mathematics
-\newblock (Springer New York) Vol.{} 110.
-
-\bibitem{Strogatz:2008wo}
-Strogatz SH
-\newblock (2008) \emph{{Nonlinear Dynamics and Chaos: With Applications to
-  Physics, Biology, Chemistry, and Engineering}}, Studies in nonlinearity
-\newblock (Westview Press, Boulder).
-
-\bibitem{GuckHolmes}
-Guckenheimer J, Holmes P
-\newblock (1983) \emph{{Nonlinear Oscillations, Dynamical Systems, and
-  Bifurcations of Vector Fields}}
-\newblock (Springer, New York).
-
-\bibitem{Gross:2004p2428}
-Gross T, Feudel U
-\newblock (2004) {Analytical search for bifurcation surfaces in parameter
-  space}.
-\newblock \emph{Physica D} 195:292--302.
-
-\bibitem{Hastings:2001jh}
-Hastings A
-\newblock (2001) {Transient dynamics and persistence of ecological systems}.
-\newblock \emph{Ecol. Lett.} 4:215--220.
-
-\bibitem{Neubert:1997wk}
-Neubert M, Caswell H
-\newblock (1997) {Alternatives to resilience for measuring the responses of
-  ecological systems to perturbations}.
-\newblock \emph{Ecology} 78:653--665.
-
-\bibitem{Caswell:2005eo}
-Caswell H, Neubert MG
-\newblock (2005) {Reactivity and transient dynamics of discrete-time ecological
-  systems}.
-\newblock \emph{J Differ Equ Appl} 11:295--310.
-
-\bibitem{Neubert:2009td}
-Neubert M, Caswell H
-\newblock (2009) {Detecting reactivity}.
-\newblock \emph{Ecology} 90:2683--2688.
-
-\bibitem{Yodzis:1992hg}
-Yodzis P, Innes S
-\newblock (1992) {Body size and consumer-resource dynamics}.
-\newblock \emph{Am. Nat.} 139:1151--1175.
-
-\bibitem{Brown:2004wq}
-Brown J, Gillooly J, Allen A, Savage V, West G
-\newblock (2004) {Toward a metabolic theory of ecology}.
-\newblock \emph{Ecology} 85:1771--1789.
-
-\bibitem{West:2002it}
-West GB, Woodruff WH, Brown JH
-\newblock (2002) {Allometric scaling of metabolic rate from molecules and
-  mitochondria to cells and mammals.}
-\newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.
-
-\bibitem{DeLong:2010dy}
-DeLong JP, Okie JG, Moses ME, Sibly RM, Brown JH
-\newblock (2010) {Shifts in metabolic scaling, production, and efficiency
-  across major evolutionary transitions of life}.
-\newblock \emph{Proc. Natl. Acad. Sci. USA} 107:12941--12945.
-
-\bibitem{West:2001bv}
-West GB, Brown JH, Enquist BJ
-\newblock (2001) {A general model for ontogenetic growth}.
-\newblock \emph{Nature} 413:628--631.
-
-\bibitem{moses2008rmo}
-Moses ME, {et~al.}
-\newblock (2008) {Revisiting a Model of Ontogenetic Growth: Estimating Model
-  Parameters from Theory and Data.}
-\newblock \emph{Am. Nat.} 171:632--645.
-
-\bibitem{gillooly2002esa}
-Gillooly JF, Charnov EL, West GB, Savage VM, Brown JH
-\newblock (2002) {Effects of size and temperature on developmental time}.
-\newblock \emph{Nature} 417:70--73.
-
-\bibitem{hou}
-Hou C, {et~al.}
-\newblock (2008) {Energy uptake and allocation during ontogeny}.
-\newblock \emph{Science} 322:736--739.
-
-\bibitem{bettencourt}
-Bettencourt LMA, Lobo J, Helbing D, Kuhnert C, West GB
-\newblock (2007) {Growth, innovation, scaling, and the pace of life in cities}.
-\newblock \emph{Proc. Natl. Acad. Sci. USA} 104:7301--7306.
-
-\bibitem{Folland:2008ij}
-Folland JP, Mc~Cauley TM, Williams AG
-\newblock (2008) {Allometric scaling of strength measurements to body size}.
-\newblock \emph{Eur J Appl Physiol} 102:739--745.
-
-\bibitem{Brown:1993p708}
-Brown J, Marquet P, Taper M
-\newblock (1993) {Evolution of body size: consequences of an energetic
-  definition of fitness}.
-\newblock \emph{Am. Nat.} 142:573--584.
-
-\bibitem{West:1997cg}
-West GB, Brown JH, Enquist BJ
-\newblock (1997) {A general model for the origin of allometric scaling laws in
-  biology}.
-\newblock \emph{Science} 276:122--126.
-
-\bibitem{West:2002ud}
-West GB, Woodruff WH, Brown JH
-\newblock (2002) {Allometric scaling of metabolic rate from molecules and
-  mitochondria to cells and mammals.}
-\newblock \emph{Proc. Natl. Acad. Sci. USA} 99 Suppl 1:2473--2478.
-
-\bibitem{Millar:1990p923}
-Millar J, Hickling G
-\newblock (1990) {Fasting endurance and the evolution of mammalian body size}.
-\newblock \emph{Funct. Ecol.} 4:5--12.
-
-\bibitem{Alroy:1998p1594}
-Alroy J
-\newblock (1998) {Cope's rule and the dynamics of body mass evolution in North
-  American fossil mammals}.
-\newblock \emph{Science} 280:731--734.
-
-\bibitem{Clauset:2009fh}
-Clauset A, Redner S
-\newblock (2009) {Evolutionary model of species body mass diversification}.
-\newblock \emph{Phys. Rev. Lett.} 102:038103.
-
-\bibitem{Smith:2010p3442}
-Smith FA, {et~al.}
-\newblock (2010) {The evolution of maximum body size of terrestrial mammals}.
-\newblock \emph{Science} 330:1216--1219.
-
-\bibitem{Saarinen:2014br}
-Saarinen JJ, {et~al.}
-\newblock (2014) {Patterns of maximum body size evolution in Cenozoic land
-  mammals: Eco-evolutionary processes and abiotic forcing}.
-\newblock \emph{Proc Biol Sci} 281:20132049.
-
-\bibitem{Damuth:1987kr}
-Damuth J
-\newblock (1987) {Interspecific allometry of population density in mammals and
-  other animals: the independence of body mass and population energy-use}.
-\newblock \emph{Biol. J. Linn. Soc.} 31:193--246.
-
-\bibitem{allen2002}
-Allen AP, Brown JH, Gillooly JF
-\newblock (2002) {Global biodiversity, biochemical kinetics, and the
-  energetic-equivalence rule}.
-\newblock \emph{Science} 297:1545--1548.
-
-\bibitem{enquist1998}
-Enquist BJ, Brown JH, West GB
-\newblock (1998) {Allometric scaling of plant energetics and population
-  density}.
-\newblock \emph{Nature} 395:163--165.
-
-\bibitem{Kempes:2016}
-Kempes CP, Wang L, Amend JP, Doyle J, Hoehler T
-\newblock (2016) {Evolutionary tradeoffs in cellular composition across diverse
-  bacteria}.
-\newblock \emph{ISME J} 10:2145--2157.
-
-\bibitem{tilman1981}
-Tilman D
-\newblock (1981) {Tests of resource competition theory using four species of
-  lake michigan algae}.
-\newblock \emph{Ecology} 62:802--815.
-
-\bibitem{dutkiewicz2009}
-Dutkiewicz S, Follows MJ, Bragg JG
-\newblock (2009) {Modeling the coupling of ocean ecology and biogeochemistry}.
-\newblock \emph{Global Biogeochem. Cycles} 23:1--15.
-
-\bibitem{barton2010}
-Barton AD, Dutkiewicz S, Flierl G, Bragg J, Follows MJ
-\newblock (2010) {Patterns of diversity in marine phytoplankton}.
-\newblock \emph{Science} 327:1509--1511.
-
-\end{thebibliography}
-
-}
+\bibliography{aa_starving3}
 
@@
 %
-%	 \begin{figure}[h]
+%	 \begin{figure}[h][h]
 %		 \centering
@@
 %
-%	\begin{figure}[h]
+%	\begin{figure}[h][h]
 %	 	\centering
@@
 %
-%	\begin{figure}[h]
+%	\begin{figure}[h][h]
 %		\centering

### 2017-06-09 Manuscript/Nature-Submission/14.tex
Nature submission

--- 2017-02-28 Manuscript/Revision_2017/15.tex
@@
 We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$.
-The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity $C$.
-{\color{red}  The rate at which consumers transition between states and consume resources is dependent on their overall abundance, the abundance of resources, the efficiency of converting resources into metabolism, and how that metabolism is partitioned between maintenance and growth purposes.
-In the supplement we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple nondimensional form which we have described below (please see the supplement for a detailed derivation).
-
+The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity $c$, which we non-dimensionalize to equal one.
 Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$.
-Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a ratio between maximal resource consumption and the carrying capacity of the resource (please see the supplement).
+Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ converts resources to somatic growth of consumer tissue.
 %and $\approx 2$.
-Resources are eaten by the consumers at rate $\rho R + \delta$, which accounts for somatic growth and maintenance for starving consumers, respectively, and at rate $\beta$ by full consumers which is a constant representing maximal maintenance and somatic growth (see supplement).}
+Resources are eaten by the consumers at rate $\rho R + \delta$, which accounts for somatic growth and maintenance for starving consumers, respectively, and at rate $\beta$ by full consumers.
 %This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
@@
 %Justin adding an explanation for beta
+{\color{blue}An organism's metabolic rate $B$ is proportional to the cost of tissue
+maintenance in the absence of growth (i.e., when the body mass is $M$).  By
+definition $B=\beta/\xi$, where $\beta$ is the rate at which resources are
+consumed for full consumers (see Eq.~\eqref{eq:system}) and where $\xi$ is
+related to the conversion efficiency of resource to consumer tissue
+(Supporting Information).
+}
 
@@
 Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is remarkably close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594} as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
-{\color{red} It should be noted that this prediction of an asymptotic limit on mammalian size resembles efforts on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here \cite{Kempes:2012hy,Kempes:2016}.}
 

### 2017-06-09 Manuscript/Nature-Submission/18.tex
Nature submission

--- 2017-03-21 Manuscript/Revision_2017/18.tex
@@
 mass has been shown to be roughly proportional to body mass~\cite{Folland:2008ij} in
-mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant. The time to go from starvation to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or
-\begin{eqnarray}
-\label{mutimescale}
-t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
-\end{eqnarray}
+mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant.
+The time to go from starvation to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or $t_{\mu}=-M^{1-\eta}\ln\left(\epsilon_{\mu}\right)/a^{\prime}-t_{\sigma}$,
+% \begin{eqnarray}
+% \label{mutimescale}
+% t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
+% \end{eqnarray}
 and $\mu=1/t_{\mu}$.
@@
 In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
-This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero causing the timescales to become infinite (see Equation \ref{mutimescale}) and the rates $\mu$ and $\lambda$ to equal zero.
+This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero, causing the timescales to become infinite and the rates $\mu$ and $\lambda$ to equal zero.
 The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$.
@@
 The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
-\ref{fig:mass}\emph{C}), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
+\ref{fig:mass}\emph{A}, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
 Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but does not offer a within-lineage mechanism by which larger body sizes are selected for.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
+\caption{\small{ Invasion feasibility for organisms with a proportional change in
+  mass $\chi$ against a population with a resident body mass $M$.  The blue
+  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
+  red region denotes values of $\chi$ that result in a mass that is below the
+  starvation threshold and are thus infeasible.
+  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
+  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}\label{fig:invasion}}
+\end{figure}
 
@@
 %reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
-\caption{\small{ Invasion feasibility for organisms with a proportional change in
-  mass $\chi$ against a population with a resident body mass $M$.  The blue
-  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
-  red region denotes values of $\chi$ that result in a mass that is below the
-  starvation threshold and are thus infeasible.
-  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
-  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}}
-\label{fig:invasion}
-\end{figure}
 
@@
 % \bibliography{aa_starving3}
-\footnotesize{
 \begin{thebibliography}{10}
@@
 
+\bibitem{DeLong:2012kw}
+DeLong JP, Vasseur DA
+\newblock (2012) {A dynamic explanation of size{\textendash}density scaling in
+  carnivores}.
+\newblock \emph{Ecology} 93:470--476.
+
 \bibitem{allen2002}
@@
 \end{thebibliography}
-
-}
 

### 2017-06-09 Manuscript/Nature-Submission/Nature-supp.tex
Nature submission

--- 2017-03-20 Manuscript/Arxiv_Revision/YKR_Arxiv.tex
@@
-\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
-%\usepackage[pdftex]{graphicx}
+%% Template for a preprint Letter or Article for submission
+%% to the journal Nature.
+%% Written by Peter Czoschke, 26 February 2004
+%%
 
-\usepackage{amsmath,amsfonts,amssymb}
-\usepackage[english]{babel}
-\usepackage[latin1]{inputenc}
-\usepackage[T1]{fontenc}
+\documentclass{nature-2}
+
+\usepackage{graphicx}
+\usepackage{verbatim}
+\usepackage{subfigure}
+\usepackage{tikz}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{array}
+\usepackage{tabularx}
+\usepackage{multirow}
+\usepackage{multicol}
+% \usepackage{multibox}
+\usepackage{rotating}
 \usepackage{color}
-\usepackage{float}
-\usepackage{verbatim}
-\usepackage{graphicx}
-\usepackage{bm}
-\usepackage{mathtools}
-\usepackage{stmaryrd}
-\usepackage{anyfontsize}
+\usepackage{layout}
+\usepackage{epstopdf}
+\usepackage{afterpage}
 
-%\usepackage{epstopdf}
-%\usepackage{array}
-%\usepackage{tabularx}
-%\usepackage{multirow}
-\usepackage{color}
-%\usepackage{multibox}
-%\usepackage{rotating}
-%\usepackage{lineno}
-%\usepackage[left]{lineno}
-%\usepackage[comma,sort&compress]{natbib}
-%\usepackage{authblk}
-%\usepackage{multicol}
+%% make sure you have the nature.cls and naturemag.bst files where
+%% LaTeX can find them
 
-%\bibliographystyle{ieeetr}
+\bibliographystyle{naturemag}
 
-\usepackage{bibunits}
+\title{Supporting Information for The dynamics of starvation and recovery}
 
-%\linenumbers
-%\setlength\linenumbersep{3pt}
+%% Notice placement of commas and superscripts and use of &
+%% in the author list
+
+\author{Justin D. Yeakel$^{1,2,3,4}$, Christopher P. Kempes$^{2,3}$ \& Sidney Redner$^{2,3}$}
 
@@
 
-\author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
-  of California, Merced, Merced, CA 95340, USA}
-
-\author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
-  Park Road, Santa Fe, NM 87501, USA}
-
-\author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
-  Road, Santa Fe, NM 87501, USA}
-
-\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
-
-\begin{abstract} %250 words
-The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of its constituent individuals.
-Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery in resource-limited environments.
-To elucidate the consequences of this tradeoff, we introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete consumers that reproduce, and undernourished, non-reproducing consumers that are susceptible to mortality.
-As a function of the transition rates between these two states that are determined by the abundance of resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
-We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships between body size and a variety of traits and find that population dynamics subject to these constraints are typically driven to a steady state.
-Moreover, we find that these rates fall within a `refuge' in parameter space, where the probability of extinction of the consumer population is minimized.
-Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations.
-Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percentage of body fat.
-For body masses $< 1.748\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.748\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
-\end{abstract}
-
 \maketitle
 
-\begin{bibunit}[unsrt]
+\begin{affiliations}
+ \item School of Natural Science, University of California Merced, Merced, CA, USA
+ \item The Santa Fe Institute, Santa Fe, NM, USA
+ \item Contributed equally
+ \item Corresponding author jdyeakel@gmail.com
+\end{affiliations}
 
-\section*{Introduction}
-The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly influences how they invest reserves in uncertain environments.
-Such behaviors are generally manifested as tradeoffs between investing in somatic maintenance and growth, or allocating energy towards reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
-The timing of these behaviors responds to selective pressure, as the choice of the investment impacts future fitness~\cite{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.
-The influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
-
-The balance between (a) somatic growth and maintenance, and (b) reproduction depends on resource availability~\cite{Morris:1987eo}.
-For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\cite{Tveraa:2003fq}.
-Many bird species invest differently in broods during periods of resource scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
-Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\cite{Kirk:1997cc}.
-Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate nutritional reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
-
-Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.
-The data collected thus far has shown that diverse mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development (blastocyst implantation) until nutritional reserves can be accumulated~\cite{Mead:1989dt,Sandell:1990kw}.
-Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
-In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the importance and universality of the fundamental tradeoff between somatic and reproductive investment.
-However the general dynamic implications of these constraints are unknown.
-
-Though straightforward conceptually, incorporating the energetic dynamics of individuals~\cite{Kooi2000} into a population-level framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles~\cite{Diekmann:2010da}.
-An alternative approach involves modeling the macroscale relations that guide somatic versus reproductive investment in a consumer-resource system.
-For example, macroscale Lotka-Volterra models assume that the growth rate of the consumer population depends on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction~\cite{murdoch:2003}.
-
-In this work, we adopt an alternative approach in which we \emph{explicitly} account for resource limitation and the subsequent effects of starvation.
-Namely, only individuals with sufficient energetic reserves can reproduce.
-Such a constraint leads to reproductive time lags due to some members of the population going hungry and then recovering.
-Additionally, we incorporate the idea that reproduction is strongly constrained allometrically \cite{Kempes:2012hy}, and is not generally linearly related to resource density.
-As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.
-\\
-
-\section*{Nutritional state-structured model (NSM)}
-We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$.
-The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity $C$.
-The rate at which consumers transition between states and consume resources is dependent on their overall abundance, the abundance of resources, the efficiency of converting resources into metabolism, and how that metabolism is partitioned between maintenance and growth purposes.
-In the supplementary information (SI) we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we describe below.
-
-Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$.
-Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a ratio between maximal resource consumption and the carrying capacity of the resource.
-%and $\approx 2$.
-Resources are eaten by the hungry consumers at rate $\rho R + \delta$, that accounts for their somatic growth ($\rho R$) and  maintenance ($\delta$).
-Full consumers eat resources at a constant rate $\beta$ that accounts for maximal maintenance and somatic growth (see SI for mechanistic derivations of these rates from resource energetics).
-%This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
-The NSM represents an ecologically motivated fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\cite{Persson:1998hz}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_FixedPoint.png}
-\caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
-  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-  bifurcation conditions separate parameter space into unphysical, cyclic,
-  and steady state dynamic regimes.  The colors show the steady
-  state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
-  replete consumers $F^*$, (warmer colors denote higher densities).  Steady
-  state densities for the energetically deficient consumers $H^*$ (not shown)
-  scale with those for $F^*$.  }
-\label{fig:fp}
-\end{figure}
-
-In the mean-field approximation, in which the consumers and resources are perfectly mixed, their densities evolve according to the rate equations
-\begin{align}
-\label{eq:system}
-\begin{split}
-\frac{dF}{dt} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
-\frac{dH}{dt} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-\frac{dR}{dt} &= \alpha \left(1-R\right)R -\left(\rho R+\delta\right)H-\beta F
-\end{split}
-\end{align}
-This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see SI for a full derivation and the dimensional form).
-Notice that the total consumer density $F+H$ evolves according to $\frac{dF}{dt}+\frac{dH}{dt}=\lambda F-\mu H$.
-This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\cite{murray2011mathematical}, except that the resource density does not appear in the growth term.
-As discussed above, the attributes of reproduction and mortality have been explicitly apportioned to the full and hungry consumers, respectively, so that the growth in the total density is decoupled from the resource density.
-
-Equation~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
-\begin{align}
-\label{eq:ss}
-\begin{split}
-F^* &= (\sigma-\lambda)\frac{ \alpha  \lambda  \mu ^2  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
-H^* &= (\sigma-\lambda)\frac{ \alpha  \lambda ^2 \mu  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
-R^* &= (\sigma - \lambda)\frac{\mu  }{A}.
-\end{split}
-\end{align}
-where $A=(\lambda  \xi  \rho +\mu  \sigma )$ and $B=(\beta  \mu  \xi +\delta  \lambda  \xi -\lambda  \mu )$.
-The stability of this fixed point is determined by the Jacobian matrix $\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.
-The parameters in Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue of $\mathbf{J}$ is negative, so that the system is stable with respect to small perturbations from the fixed point.
-Because this fixed point is unique, it is the global attractor for all population trajectories for any initial condition where the resource and consumer densities are both nonzero.
-
-From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that the consumer and resource densities are positive.
-The condition $\sigma = \lambda$ thus represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates a physical from an unphysical regime where all steady-state densities become negative after intersecting the trivial fixed point $(F^*,H^*,R^*)=(0,0,0)$.
-The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
-As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges.
-
-In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).
-In continuous-time systems, a limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts~\cite{GuckHolmes}.
-This Hopf bifurcation is defined by ${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial of the Jacobian matrix~\cite{Gross:2004p2428}.
-As the system parameters are tuned to be within the stable regime, but close to the Hopf bifurcation, the amplitude of the transient cycles becomes large.
-Given that ecological systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase the extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
-
-When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the consumers are driven to the hungry non-reproducing state.
-Because reproduction is inhibited, there is a low steady-state consumer density and a high steady-state resource density.
-However, if $\sigma/\lambda\to 1$ from above, the population is overloaded with energetically-replete (reproducing) individuals, thereby promoting transient oscillations between the consumer and resource densities (Fig.~\ref{fig:fp}).
-If the starvation rate is low enough that the Hopf bifurcation is crossed, these oscillations become stable over time.
-This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.
-
-\section*{Role of allometry}
-While there are no a priori constraints on the parameters in the NSM, most
-organisms correspond to restricted portions of the parameter space.  Here we
-use allometric scaling relations to constrain the covariation of rates in a
-principled and biologically meaningful manner.  Allometric scaling relations
-highlight common constraints and average trends across large ranges in body
-size and species diversity. Many of these relations can be derived from a
-small set of assumptions and below we describe our framework to determine the
-covariation of timescales and rates across the range of mammals for each of
-the key parameters of our model (cf. ref.~\cite{Yodzis:1992hg}).  We are thereby
-able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.
-
-Nearly all of the rates described in the NSM are determined by consumer
-metabolism, which can be used to describe a variety of organismal features
-\cite{Brown:2004wq}.  The scaling relation between an organism's metabolic
-rate $B$ and its body mass $M$ at reproductive maturity is known to scale as
-$B = B_0 M^\eta$~\cite{West:2002it}, where the scaling exponent $\eta$ is
-typically close to $2/3$ or $3/4$ for metazoans (e.g., ref. \cite{Brown:2004wq}),
-and has taxonomic shifts for unicellular species between $\eta\approx 1$ in
-eukaryotes and $\eta\approx 1.76$ in bacteria
-\cite{DeLong:2010dy,Kempes:2012hy}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.pdf}
-\caption{ The growth trajectory over absolute time of an individual organism
-  as a function of body mass.  Initial growth follows the black trajectory to
-  an energetically replete reproductive adult mass $m=\epsilon_\lambda M$ which we assume is 95\% asymptotic mass $M$.  Starvation follows the red
-  trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. Recovery follows the
-  green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires different energetics and a longer time to reach $\epsilon_\lambda M$.
-Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}
-\label{fig:growth}
-\end{figure}
-
-Several efforts have shown how a partitioning of $B$ between growth and
-maintenance purposes can be used to derive a general equation for both the
-growth trajectories and growth rates of organisms ranging from bacteria to
-metazoans
-\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. This relation is derived from the simple balance condition \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}
-\begin{eqnarray}
-\label{balance}
-B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
-\end{eqnarray}
-where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
-the metabolic rate to support an existing unit of mass, and $m$ is the mass
-of the organism at any point in its development.  This balance has the
-general solution \cite{bettencourt,Kempes:2012hy}
-\begin{eqnarray}
-\label{m1}
-\left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}},
-\end{eqnarray}
-where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically (see SI).
-We now use this solution to define the timescale for reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \cite{moses2008rmo} for a detailed presentation of these timescales).
-The time that it takes to reach a particular mass $\epsilon M$ is given by the timescale
-\begin{equation}
-\label{t1}
-\tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)},
-\end{equation}
-where we will define values of $\epsilon$ to describe a set of rates within our model.
-For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$; \cite{West:2001bv}).
-The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$, this rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_Rates.pdf}
-\caption{
-Allometrically constrained starvation rate $\sigma=1/t_{\sigma}$ (red), mortality $\mu=1/t_{\mu}$ (blue) and recovery rate $\rho=1/t_{\rho}$ (green) relative to the reproductive rate $\lambda=1/t_{\lambda}$ (black) as a function of body mass (see Equations \ref{t1}, \ref{rhotimescale}, \ref{eq:sigma}, and \ref{mutimescale}).
-The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
-Mean values $\pm 25\%$ variation are shown by the shaded region for each rate.
-}
-\label{fig:gvs}
-\end{figure}
-
-The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
-sufficient tissue to transition from the hungry to the full state.  Since
-only certain tissues can be digested for energy (for example the brain cannot
-be degraded to fuel metabolism), we define the rates for starvation, death,
-and recovery by the timescales required to reach, or return from, specific
-fractions of the replete-state mass (Fig. \ref{fig:gvs}; see SI, Table I for parameterizations).  We define
-$m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction of
-replete-state mass where reproduction ceases. This fraction will deviate from a constant
-if tissue composition systematically scales with adult mass.  For example,
-making use of the observation that body fat in mammals scales with overall
-body size according to $M_{\rm fat}=f_{0}M^{\gamma}$ and assuming that once
-this mass is fully digested the organism starves, this would imply that
-$\epsilon_{\sigma}=1-f_{0}M^{\gamma}/M$. It follows that the
-recovery timescale, $t_{\rho}$, is the time to go from
-$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$ (Fig.
-\ref{fig:growth}).
-Using Eqs.~\eqref{m1} and \eqref{t1} this timescale is given by simply considering an adjusted starting mass of $m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
-\begin{eqnarray}
-\label{rhotimescale}
-t_{\rho}=\ln\left[\frac{1-\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)^{1-\eta}}{1-\epsilon_\lambda^{1-\eta}}\right]\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}
-\end{eqnarray}
-where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in the biosynthetic energetics during recovery (see SI).
-It should be noted that more complicated ontogenetic models explicitly handle
-storage \cite{hou}, whereas this feature is implicitly covered by the body
-fat scaling in our framework.
-
-To determine the starvation rate, $\sigma$, we are interested in the time
-required for an organism to go from a mature adult that reproduces at rate
-$\lambda$, to a
-reduced-mass hungry state where reproduction is impossible.
-For starving individuals we assume that an organism must meet its maintenance requirements by using the digestion of existing mass as the sole energy source.
-This assumption implies the following simple metabolic balance
-\begin{eqnarray}
-\frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
-\end{eqnarray}
-or
-\begin{eqnarray}
-\frac{dm}{dt}=-\frac{a^{\prime}}{M^{1-\eta}}m
-\end{eqnarray}
-where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
-body mass, which differs from $E_{m}$, the energy required to
-synthesis a unit of biomass \cite{hou}. Given the replete mass, $M$, of an organism, the
-above energy balance prescribes the mass trajectory of a non-consuming
-organism:
-\begin{eqnarray}
-\label{mt}
-m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}.
-\end{eqnarray}
-The timescale for starvation is
-given by the time it takes $m(t)$ to reach $\epsilon_{\sigma} M$, which gives
-\begin{equation}
-\label{eq:sigma}
-t_{\sigma}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\sigma}\right).
-\end{equation}
-The starvation rate is then $\sigma=1/t_{\sigma}$, which scales with
-replete-state mass as $1/M^{1-\eta}\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important
-feature is that $\sigma$ does not have a simple scaling dependence on
-$\lambda$ (Fig. \ref{fig:gvs}), which is important for the dynamics that we
-later discuss.
-
-The time to death should follow a similar relation, but defined by a lower
-fraction of replete-state mass, $m_{\mu}=\epsilon_{\mu} M$ where $\epsilon_\mu < \epsilon_\sigma$.
-Suppose, for example, that an organism dies once it has digested all fat and
-muscle tissues, and that muscle tissue scales with body mass according to
-$M_{\rm musc}=u_{0}M^{\zeta}$.  This gives
-$\epsilon_{\mu}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
-mass has been shown to be roughly proportional to body mass~\cite{Folland:2008ij} in
-mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant. The time to go from starvation to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or
-\begin{eqnarray}
-\label{mutimescale}
-t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
-\end{eqnarray}
-and $\mu=1/t_{\mu}$.
-
-Although the rate equations \eqref{eq:system} are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
-Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
-Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
-
-\section*{Stabilizing effects of allometric constraints}
-As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
-Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values (Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.
-We thus find that the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
-Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
-These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
-However, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation.
-This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.425\textwidth]{fig_DataHopf.pdf}
-\caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
-  allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
-  a function of different mammalian body sizes: $M=P\times10^2$g (blue) and
-  $M=P\times10^6$g (green), where $P$ is a random uniform variable in $[1,9]$.
-  Points denote realized values of $\sigma$ and $\rho$ given the drawn values for $M$.
-  }
-\label{fig:hopf}
-\end{figure}
-
-Previous studies used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
-However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
-Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
-However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.
-
-\begin{figure*}
-\centering
-\includegraphics[width=1\textwidth]{fig_ExtinctionAllometricComb3.png}
-\caption{ Probability of extinction for a consumer with (\emph{A}) $M=10^2$g, (\emph{B}) $M=10^4$g, and (\emph{C}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
-Note the change in scale for $M=10^4$ and $M=10^6$g.
-Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
-The white points denote the allometrically constrained starvation and recovery rate.}
-\label{fig:ext}
-\end{figure*}
-
-%Extinction to the left and the right
-\section*{Extinction risk}
-Within our model, higher rates of starvation result in a larger flux of the population to the hungry state.
-In this state, reproduction is absent, thus increasing the likelihood of extinction.
-From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
-We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
-To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
-This procedure is repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for organisms with mass ranging from $10^2$ to $10^6$ grams.
-%the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
-In each replicate the initial densities are chosen to be $(XF^*,XH^*,R^*)$, with $X$ a random variable that is uniformly distributed in $[0,2]$.
-By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-7}$ to $10^{-3}$.
-%, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
-As expected, higher rates of extinction correlate with both high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if $\sigma$ is small.
-For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
-For high values of $\sigma$ and low values of $\rho$, increased extinction risk arises because of the decrease in the steady-state consumer population density (Figs.
-\ref{fig:fp}\emph{B}, \ref{fig:ext}).
-This interplay creates an `extinction refuge', such that for a constrained range of $\sigma$ and $\rho$, extinction probabilities are minimized.
-
-We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge across a range of $M$ (Fig. \ref{fig:ext}\emph{A-C}, white points).
-These values are close enough to the Hopf bifurcation to avoid low steady-state densities, and far enough away to avoid large-amplitude transient cycles.
-The feature that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
-Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that serve to minimize extinction risk.
-This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
-We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig.
-\ref{fig:ext}\emph{A-C}), suggesting that the population dynamics of larger organisms are more sensitive to smaller changes in physiological rates controlling starvation and recovery.
-% OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-{\it To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.}\\
-
-%Results are insensitive to alpha but possibly sensitive to beta (eff parameter)
-
-%NOTES 7/8/16
-%Calder has a hypothesis that period ~ Mass^(1/4).
-%Can we get "transient period" from Jacobian?
-%Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.
-
-\section*{Dynamic and energetic barriers to body size}
-Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\cite{West:2002ud}.
-Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
-One important such example is the \emph{fasting endurance hypothesis}, which contends that larger body size, with consequent lower metabolic rates and increased ability to maintain more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability~\cite{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's rule)~\cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\cite{Clauset:2009fh}.
-These trends are thought to be driven by a combination of climate change and niche availability~\cite{Saarinen:2014br}; however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, have not been explored.
-We argue that the NSM provides a suitable framework to explore these issues.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{fig_FPAllometric.pdf}
-\caption{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-  body mass.
-  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.
-The data are from Damuth \cite{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
-\cite{West:2001bv,hou,moses2008rmo}).}
-\label{fig:mass}
-\end{figure}
-
-The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
-Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{Damuth:1987kr,allen2002,enquist1998}.
-The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size~\cite{Damuth:1987kr,allen2002,enquist1998}.
-This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism, such that $N^*\propto M^{-3/4}/B_{0}$~\cite{allen2002,enquist1998}.
-This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{Damuth:1987kr,allen2002,enquist1998}.
-Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
-This result is remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to energy equivalence.
-Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
-In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
-This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero causing the timescales to become infinite (see Equation \ref{mutimescale}) and the rates $\mu$ and $\lambda$ to equal zero.
-The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$.
-This point predicts a strong upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
-Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
-It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters.
-We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\cite{Kempes:2012hy,Kempes:2016}.
-% reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
-
-We contend that the NSM provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time.
-The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
-\ref{fig:mass}\emph{C}), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
-Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but does not offer a within-lineage mechanism by which larger body sizes are selected for.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_Invasion.pdf}
-\caption{ Invasion feasibility for organisms with a proportional change in
-  mass $\chi$ against a population with a resident body mass $M$.  The blue
-  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
-  red region denotes values of $\chi$ that result in a mass that is below the
-  starvation threshold and are thus infeasible.
-  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
-  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}
-\label{fig:invasion}
-\end{figure}
-
-To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120$ times the mass of a male African elephant.
-Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
-We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by $^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$.
-For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given Equation \ref{t1}, implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
-Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
-The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
-Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the body composition of starved foragers, is assumed to remain unchanged for the invader population (see SI for detailed derivations of invader rates).
-
-To assess the susceptibility of the resident population to invasion, we determine which consumer has a lower steady-state resource density for a given value of $\chi$, again with the expectation that populations able to survive on lower resource densities have a competitive advantage \cite{tilman1981}.
-We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
-However, for $M> 1.748\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
-
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
-This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca.
-$1.5\times10^7$g and the late Miocene \emph{Deinotherium} at ca.
-$1.74\times10^7$g ~\cite{Smith:2010p3442}.
-Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \cite{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals outlined by Alroy~\cite{Alroy:1998p1594}.
-While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general within-lineage mechanism for the evolution of larger body size among terrestrial mammals.
-
-The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations~\cite{Stearns:1989ip}.
-The NSM is a minimal and general model that incorporates the dynamics of starvation and recovery that are expected to occur in resource-limited environments.
-By incorporating allometric relations between the rates in the NSM, we found: (i) different organismal masses have distinct population dynamic regimes, (ii) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and (iii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of maximum body size.
-We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.
-Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.
-
-% \putbib[aa_starving3]
-
-\begin{thebibliography}{10}
-
-\bibitem{Martin:1987dl}
-Thomas~E Martin.
-\newblock {Food as a limit on breeding birds: A life-history perspective}.
-\newblock {\em Annu. Rev. Ecol. Syst.}, 18:453--487, January 1987.
-
-\bibitem{Kirk:1997cc}
-Kevin~L Kirk.
-\newblock {Life-history responses to variable environments: Starvation and
-  reproduction in planktonic rotifers}.
-\newblock {\em Ecology}, 78(2):434--441, March 1997.
-
-\bibitem{Kempes:2012hy}
-C~P Kempes, S~Dutkiewicz, and M~J Follows.
-\newblock {Growth, metabolic partitioning, and the size of microorganisms}.
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 109(2):495--500, January 2012.
-
-\bibitem{Mangel:1988uaa}
-M~Mangel and C.~W. Clark.
-\newblock {\em {Dynamic Modeling in Behavioral Ecology}}.
-\newblock Princeton University Press, Princeton, 1988.
-
-\bibitem{Mangel:2014kz}
-Marc Mangel.
-\newblock {Stochastic dynamic programming illuminates the link between
-  environment, physiology, and evolution}.
-\newblock {\em B. Math. Biol.}, 77(5):857--877, 2014.
-
-\bibitem{Yeakel:2013hi}
-Justin~D Yeakel, Nathaniel~J Dominy, Paul~L Koch, and Marc Mangel.
-\newblock {Functional morphology, stable isotopes, and human evolution: a model
-  of consilience}.
-\newblock {\em Evolution}, 68(1):190--203, 2014.
-
-\bibitem{Morris:1987eo}
-Douglas~W Morris.
-\newblock {Optimal allocation of parental investment}.
-\newblock {\em Oikos}, 49(3):332--339, July 1987.
-
-\bibitem{Tveraa:2003fq}
-Torkild Tveraa, Per Fauchald, Cathrine Henaug, and Nigel~G Yoccoz.
-\newblock {An examination of a compensatory relationship between food
-  limitation and predation in semi-domestic reindeer}.
-\newblock {\em Oecologia}, 137(3):370--376, November 2003.
-
-\bibitem{Daan:1988va}
-S~Daan, C~Dijkstra, R~Drent, and T~Meijer.
-\newblock {Food supply and the annual timing of avian reproduction}.
-\newblock In H~Ouellet, editor, {\em Acta XIX Congressus Internationalis
-  Ornithologici, Volume I: Proceedings XIX International Ornithological
-  Congress, 1986, Ottawa}, pages 392--407. Proceedings XIX International
-  Ornithological Congress, Ottawa, 1989.
-
-\bibitem{Jacot:2009dv}
-Alain Jacot, Mihai Valcu, Kees van Oers, and Bart Kempenaers.
-\newblock {Experimental nest site limitation affects reproductive strategies
-  and parental investment in a hole-nesting passerine}.
-\newblock {\em Animal Behaviour}, 77(5):1075--1083, May 2009.
-
-\bibitem{Stearns:1989ip}
-S~C Stearns.
-\newblock {Trade-offs in life-history evolution}.
-\newblock {\em Funct. Ecol.}, 3(3):259, 1989.
-
-\bibitem{Barboza:2002in}
-P~Barboza and D~Jorde.
-\newblock {Intermittent fasting during winter and spring affects body
-  composition and reproduction of a migratory duck}.
-\newblock {\em J Comp Physiol B}, 172(5):419--434, July 2002.
-
-\bibitem{Threlkeld:1976ih}
-Stephen~T Threlkeld.
-\newblock {Starvation and the size structure of zooplankton communities}.
-\newblock {\em Freshwater Biol.}, 6(6):489--496, December 1976.
-
-\bibitem{Weber:1998jg}
-Thomas~P Weber, Bruno~J Ens, and Alasdair~I Houston.
-\newblock {Optimal avian migration: A dynamic model of fuel stores and site
-  use}.
-\newblock {\em Evolutionary Ecology}, 12(4):377--401, May 1998.
-
-\bibitem{Mduma:1999cp}
-Simon A~R Mduma, A~R~E Sinclair, and Ray Hilborn.
-\newblock {Food regulates the Serengeti wildebeest: a 40-year record}.
-\newblock {\em J. Anim. Ecol.}, 68(6):1101--1122, November 1999.
-
-\bibitem{Moore:2014hi}
-Jonathan~W Moore, Justin~D Yeakel, Dean Peard, Jeff Lough, and Mark Beere.
-\newblock {Life-history diversity and its importance to population stability
-  and persistence of a migratory fish: steelhead in two large North American
-  watersheds}.
-\newblock {\em J. Anim. Ecol.}, 83(5):1035--1046, 2014.
-
-\bibitem{Mead:1989dt}
-Rodney~A Mead.
-\newblock {The Physiology and Evolution of Delayed Implantation in Carnivores}.
-\newblock In John~L Gittleman, editor, {\em Carnivore Behavior, Ecology, and
-  Evolution}, pages 437--464. Springer US, Ithaca, 1989.
-
-\bibitem{Sandell:1990kw}
-Mikael Sandell.
-\newblock {The evolution of seasonal delayed implantation}.
-\newblock {\em Q Rev Biol}, 65(1):23--42, 1990.
-
-\bibitem{Bulik:1999eo}
-Cynthia~M Bulik, Patrick~F Sullivan, Jennifer~L Fear, Alison Pickering, Aria
-  Dawn, and Mandy McCullin.
-\newblock {Fertility and reproduction in women with anorexia nervosa}.
-\newblock {\em J. Clin. Psychiat.}, 60(2):130--135, February 1999.
-
-\bibitem{Trites:2003cc}
-A~W Trites and C~P Donnelly.
-\newblock {The decline of Steller sea lions Eumetopias jubatus in Alaska: a
-  review of the nutritional stress hypothesis}.
-\newblock {\em Mammal Rev.}, 33(1):3--28, March 2003.
-
-\bibitem{Glazier:2009hq}
-Douglas~S Glazier.
-\newblock {Metabolic level and size scaling of rates of respiration and growth
-  in unicellular organisms}.
-\newblock {\em Funct. Ecol.}, 23(5):963--968, October 2009.
-
-\bibitem{Kooi2000}
-S~A L~M Kooijman.
-\newblock {\em {Dynamic Energy and Mass Budgets in Biological Systems}}.
-\newblock Cambridge, 2000.
-
-\bibitem{Sousa:2010ez}
-T~Sousa, T~Domingos, J~C Poggiale, and S~A L~M Kooijman.
-\newblock {Dynamic energy budget theory restores coherence in biology}.
-\newblock {\em Philos. T. Roy. Soc. B}, 365(1557):3413--3428, October 2010.
-
-\bibitem{Diekmann:2010da}
-O~Diekmann and J~A~J Metz.
-\newblock {How to lift a model for individual behaviour to the population
-  level?}
-\newblock {\em Philos. T. Roy. Soc. B}, 365(1557):3523--3530, November 2010.
-
-\bibitem{murdoch:2003}
-W~W Murdoch, C~J Briggs, and R~M Nisbet.
-\newblock {\em {Consumer-resource Dynamics}}, volume~36 of {\em Monographs in
-  population biology}.
-\newblock Princeton University Press, Princeton, 2003.
-
-\bibitem{Benichou:2014wu}
-Olivier Benichou and S~Redner.
-\newblock {Depletion-Controlled Starvation of a Diffusing Forager}.
-\newblock {\em arXiv}, pages 1--5, May 2014.
-
-\bibitem{Benichou:2016wl}
-O~B{\'e}nichou, M~Chupeau, and S~Redner.
-\newblock {Role of depletion on the dynamics of a diffusing forager}.
-\newblock {\em Journal of Physics A: {\ldots}}, 2016.
-
-\bibitem{Chupeau:2016jf}
-M~Chupeau, O~B{\'e}nichou, and S~Redner.
-\newblock {Universality classes of foraging with resource renewal}.
-\newblock {\em Phys. Rev. E}, 93(3):032403, March 2016.
-
-\bibitem{Persson:1998hz}
-Lennart Persson, Kjell Leonardsson, Andr{\'e}~M De~Roos, Mats Gyllenberg, and
-  Bent Christensen.
-\newblock {Ontogenetic scaling of foraging rates and the dynamics of a
-  size-structured consumer-resource model}.
-\newblock {\em Theor Popul Biol}, 54(3):270--293, December 1998.
-
-\bibitem{murray2011mathematical}
-J~D Murray.
-\newblock {\em {Mathematical Biology: I. An Introduction}}, volume 110 of {\em
-  Interdisciplinary Applied Mathematics}.
-\newblock Springer New York, 2011.
-
-\bibitem{Strogatz:2008wo}
-S~H Strogatz.
-\newblock {\em {Nonlinear Dynamics and Chaos: With Applications to Physics,
-  Biology, Chemistry, and Engineering}}.
-\newblock Studies in nonlinearity. Westview Press, Boulder, 2008.
-
-\bibitem{GuckHolmes}
-J~Guckenheimer and P~Holmes.
-\newblock {\em {Nonlinear Oscillations, Dynamical Systems, and Bifurcations of
-  Vector Fields}}.
-\newblock Springer, New York, 1983.
-
-\bibitem{Gross:2004p2428}
-T~Gross and U~Feudel.
-\newblock {Analytical search for bifurcation surfaces in parameter space}.
-\newblock {\em Physica D}, 195(3-4):292--302, 2004.
-
-\bibitem{Hastings:2001jh}
-Alan Hastings.
-\newblock {Transient dynamics and persistence of ecological systems}.
-\newblock {\em Ecol. Lett.}, 4(3):215--220, May 2001.
-
-\bibitem{Neubert:1997wk}
-MG~Neubert and H~Caswell.
-\newblock {Alternatives to resilience for measuring the responses of ecological
-  systems to perturbations}.
-\newblock {\em Ecology}, 78(3):653--665, 1997.
-
-\bibitem{Caswell:2005eo}
-Hal Caswell and Michael~G Neubert.
-\newblock {Reactivity and transient dynamics of discrete-time ecological
-  systems}.
-\newblock {\em J Differ Equ Appl}, 11(4-5):295--310, April 2005.
-
-\bibitem{Neubert:2009td}
-MG~Neubert and H~Caswell.
-\newblock {Detecting reactivity}.
-\newblock {\em Ecology}, 90(10):2683--2688, 2009.
-
-\bibitem{Yodzis:1992hg}
-P~Yodzis and S~Innes.
-\newblock {Body size and consumer-resource dynamics}.
-\newblock {\em Am. Nat.}, 139(6):1151--1175, 1992.
-
-\bibitem{Brown:2004wq}
-JH~Brown, JF~Gillooly, AP~Allen, VM~Savage, and GB~West.
-\newblock {Toward a metabolic theory of ecology}.
-\newblock {\em Ecology}, 85(7):1771--1789, 2004.
-
-\bibitem{West:2002it}
-Geoffrey~B West, William~H Woodruff, and James~H Brown.
-\newblock {Allometric scaling of metabolic rate from molecules and mitochondria
-  to cells and mammals.}
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 99 Suppl 1(suppl 1):2473--2478,
-  February 2002.
-
-\bibitem{DeLong:2010dy}
-J~P DeLong, J~G Okie, M~E Moses, R~M Sibly, and J~H Brown.
-\newblock {Shifts in metabolic scaling, production, and efficiency across major
-  evolutionary transitions of life}.
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 107(29):12941--12945, July 2010.
-
-\bibitem{West:2001bv}
-Geoffrey~B West, James~H Brown, and Brian~J Enquist.
-\newblock {A general model for ontogenetic growth}.
-\newblock {\em Nature}, 413(6856):628--631, October 2001.
-
-\bibitem{moses2008rmo}
-Melanie~E Moses, Chen Hou, William~H Woodruff, Geoffrey~B West, Jeffery~C
-  Nekola, Wenyun Zuo, and James~H Brown.
-\newblock {Revisiting a Model of Ontogenetic Growth: Estimating Model
-  Parameters from Theory and Data.}
-\newblock {\em Am. Nat.}, 171(5):632--645, 2008.
-
-\bibitem{gillooly2002esa}
-James~F Gillooly, Eric~L Charnov, Geoffrey~B West, Van~M Savage, and James~H
-  Brown.
-\newblock {Effects of size and temperature on developmental time}.
-\newblock {\em Nature}, 417(6884):70--73, May 2002.
-
-\bibitem{hou}
-Chen Hou, Wenyun Zuo, Melanie~E Moses, William~H Woodruff, James~H Brown, and
-  Geoffrey~B West.
-\newblock {Energy uptake and allocation during ontogeny}.
-\newblock {\em Science}, 322(5902):736--739, October 2008.
-
-\bibitem{bettencourt}
-L~M~A Bettencourt, J~Lobo, D~Helbing, C~Kuhnert, and G~B West.
-\newblock {Growth, innovation, scaling, and the pace of life in cities}.
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 104(17):7301--7306, April 2007.
-
-\bibitem{Folland:2008ij}
-J~P Folland, T~M Mc~Cauley, and A~G Williams.
-\newblock {Allometric scaling of strength measurements to body size}.
-\newblock {\em Eur J Appl Physiol}, 102(6):739--745, January 2008.
-
-\bibitem{CalderIII:1983jd}
-W~A Calder~III.
-\newblock {An allometric approach to population cycles of mammals}.
-\newblock {\em J. Theor. Biol.}, 100(2):275--282, January 1983.
-
-\bibitem{Peterson:1984hj}
-R~O Peterson, R~E Page, and K~M Dodge.
-\newblock {Wolves, moose, and the allometry of population cycles}.
-\newblock {\em Science}, 224(4655):1350--1352, June 1984.
-
-\bibitem{Krukonis:1991fk}
-G~Krukonis and W~M Schaffer.
-\newblock {Population cycles in mammals and birds: Does periodicity scale with
-  body size?}
-\newblock {\em J. Theor. Biol.}, 148(4):469--493, February 1991.
-
-\bibitem{Hendriks:2012fc}
-A~Jan Hendriks and Christian Mulder.
-\newblock {Delayed logistic and Rosenzweig{\textendash}MacArthur models with
-  allometric parameter setting estimate population cycles at lower trophic
-  levels well}.
-\newblock {\em Ecol Complex}, 9:43--54, February 2012.
-
-\bibitem{Kendall:1999iy}
-Bruce~E Kendall, Cheryl~J Briggs, William~W Murdoch, Peter Turchin, Stephen~P
-  Ellner, Edward McCauley, Roger~M Nisbet, and Simon~N Wood.
-\newblock {Why do populations cycle? A synthesis of statistical and mechanistic
-  modeling approaches}.
-\newblock {\em Ecology}, 80(6):1789--1805, September 1999.
-
-\bibitem{Hogstedt:2005cr}
-G{\"o}ran H{\"o}gstedt, Tarald Seldal, and Arild Breistol.
-\newblock {Period length in cyclic animal populations}.
-\newblock {\em Ecology}, 86(2):373--378, February 2005.
-
-\bibitem{Kendall:1998hl}
-Kendall, Prendergast, and Bjornstad.
-\newblock {The macroecology of population dynamics: taxonomic and biogeographic
-  patterns in population cycles}.
-\newblock {\em Ecol. Lett.}, 1(3):160--164, November 1998.
-
-\bibitem{Brown:1993p708}
-JH~Brown, PA~Marquet, and ML~Taper.
-\newblock {Evolution of body size: consequences of an energetic definition of
-  fitness}.
-\newblock {\em Am. Nat.}, 142(4):573--584, 1993.
-
-\bibitem{West:1997cg}
-Geoffrey~B West, James~H Brown, and Brian~J Enquist.
-\newblock {A general model for the origin of allometric scaling laws in
-  biology}.
-\newblock {\em Science}, 276(5309):122--126, April 1997.
-
-\bibitem{West:2002ud}
-Geoffrey~B West, William~H Woodruff, and James~H Brown.
-\newblock {Allometric scaling of metabolic rate from molecules and mitochondria
-  to cells and mammals.}
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 99 Suppl 1(Supplement
-  1):2473--2478, February 2002.
-
-\bibitem{Millar:1990p923}
-JS~Millar and G~Hickling.
-\newblock {Fasting endurance and the evolution of mammalian body size}.
-\newblock {\em Funct. Ecol.}, 4(1):5--12, 1990.
-
-\bibitem{Alroy:1998p1594}
-J~Alroy.
-\newblock {Cope's rule and the dynamics of body mass evolution in North
-  American fossil mammals}.
-\newblock {\em Science}, 280(5364):731--734, 1998.
-
-\bibitem{Clauset:2009fh}
-A~Clauset and S~Redner.
-\newblock {Evolutionary model of species body mass diversification}.
-\newblock {\em Phys. Rev. Lett.}, 102(3):038103, January 2009.
-
-\bibitem{Smith:2010p3442}
-Felisa~A Smith, Alison~G Boyer, James~H Brown, Daniel~P Costa, Tamar Dayan,
-  S~K~Morgan Ernest, Alistair~R Evans, Mikael Fortelius, John~L Gittleman,
-  Marcus~J Hamilton, Larisa~E Harding, Kari Lintulaakso, S~Kathleen Lyons,
-  Christy McCain, Jordan~G Okie, Juha~J Saarinen, Richard~M Sibly, Patrick~R
-  Stephens, Jessica Theodor, and Mark~D Uhen.
-\newblock {The evolution of maximum body size of terrestrial mammals}.
-\newblock {\em Science}, 330(6008):1216--1219, November 2010.
-
-\bibitem{Saarinen:2014br}
-Juha~J Saarinen, Alison~G Boyer, James~H Brown, Daniel~P Costa, S~K~Morgan
-  Ernest, Alistair~R Evans, Mikael Fortelius, John~L Gittleman, Marcus~J
-  Hamilton, Larisa~E Harding, Kari Lintulaakso, S~Kathleen Lyons, Jordan~G
-  Okie, Richard~M Sibly, Patrick~R Stephens, Jessica Theodor, Mark~D Uhen, and
-  Felisa~A Smith.
-\newblock {Patterns of maximum body size evolution in Cenozoic land mammals:
-  Eco-evolutionary processes and abiotic forcing}.
-\newblock {\em Proc Biol Sci}, 281(1784):20132049, June 2014.
-
-\bibitem{Damuth:1987kr}
-John Damuth.
-\newblock {Interspecific allometry of population density in mammals and other
-  animals: the independence of body mass and population energy-use}.
-\newblock {\em Biol. J. Linn. Soc.}, 31(3):193--246, July 1987.
-
-\bibitem{allen2002}
-Andrew~P Allen, James~H Brown, and James~F Gillooly.
-\newblock {Global biodiversity, biochemical kinetics, and the
-  energetic-equivalence rule}.
-\newblock {\em Science}, 297(5586):1545--1548, August 2002.
-
-\bibitem{enquist1998}
-Brian~J Enquist, James~H Brown, and Geoffrey~B West.
-\newblock {Allometric scaling of plant energetics and population density}.
-\newblock {\em Nature}, 395(6698):163--165, September 1998.
-
-\bibitem{Kempes:2016}
-Christopher~P Kempes, Lawrence Wang, Jan~P Amend, John Doyle, and Tori Hoehler.
-\newblock {Evolutionary tradeoffs in cellular composition across diverse
-  bacteria}.
-\newblock {\em ISME J}, 10(9):2145--2157, April 2016.
-
-\bibitem{tilman1981}
-David Tilman.
-\newblock {Tests of resource competition theory using four species of lake
-  michigan algae}.
-\newblock {\em Ecology}, 62(3):802--815, June 1981.
-
-\bibitem{dutkiewicz2009}
-S~Dutkiewicz, M~J Follows, and J~G Bragg.
-\newblock {Modeling the coupling of ocean ecology and biogeochemistry}.
-\newblock {\em Global Biogeochem. Cycles}, 23(4):1--15, December 2009.
-
-\bibitem{barton2010}
-Andrew~D Barton, Stephanie Dutkiewicz, Glenn Flierl, Jason Bragg, and Michael~J
-  Follows.
-\newblock {Patterns of diversity in marine phytoplankton}.
-\newblock {\em Science}, 327(5972):1509--1511, March 2010.
-
-\end{thebibliography}
-
-\end{bibunit}
-
-% \bibliography{aa_starving3}
-
-%
-% \begin{acknowledgments}
-%   We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
-%   discussions and comments on the manuscript.  J.D.Y. was supported by
-%   startup funds at the University of California, Merced, and an Omidyar
-%   Postdoctoral Fellowship at the Santa Fe Institute.  C.P.K. was supported by
-%   an Omidyar Postdoctoral Fellowship at the Santa Fe Institute.  S.R. was
-%   supported by grants DMR-1608211 and 1623243 from the National Science
-%   Foundation, and by the John Templeton Foundation, all at the Santa Fe
-%   Institute.
-% \end{acknowledgments}
-%
-\clearpage
-
-\begin{bibunit}[unsrt]
-\section*{Supporting Information for ``The dynamics of starvation and recovery''}
-
-\subsection*{Mechanisms of Starvation and Recovery}
+\section*{Mechanisms of Starvation and Recovery}
 Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
@@
 \begin{split}
-\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
-\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
-&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
-\end{split}
-\end{align}
-where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
-%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
-In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
-Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
-The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
-A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
-
-In the above equations $Y$ represents the yield coefficient (e.g., Refs.
-\cite{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
-If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
-This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
-In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment.
-
-Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
-As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
-Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
-\begin{align}
-\begin{split}
 \dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
@@
 \end{align}
+where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
+In the above equations $Y$ represents the yield coefficient (e.g., Refs.
+\citenum{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
+If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
+This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
+In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
+%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
+
+%\begin{align}
+%\begin{split}
+%\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+%\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
+%\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
+%&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
+%\end{split}
+%\end{align}
+%
+%In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
+Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
+The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
+%A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
+%
+%Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
+As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
+Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
+
 %where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
@@
 
-\subsection*{Parameter Values and Estimates}
+\section*{Parameter Values and Estimates}
 
@@
 Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs.
-\cite{West:2001bv,moses2008rmo} and the supplement for Ref.
-\cite{hou}).
+\citenum{West:2001bv,moses2008rmo} and the supplement for Ref.
+\citenum{hou}).
 We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \cite{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
@@
 For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs.
-\cite{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
+\citenum{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
 We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).
@@
 This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref.
-\cite{peters1986ecological} for a review).
+\citenum{peters1986ecological} for a review).
 We use $m_{0}=0.097M^{0.92}$ \cite{blueweiss1978relationships} which encompasses the widest range of body sizes \cite{peters1986ecological}.
@@
 Refs.
-\cite{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \cite{moses2008rmo}.
+\citenum{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \cite{moses2008rmo}.
 The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \cite{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \cite{stryer,hou}.
@@
 From the data in Ref.
-\cite{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
+\citenum{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
 It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \cite{damuth1987interspecific}.
@@
 
-% \putbib[aa_starving_supplement]
-
-\begin{thebibliography}{10}
-
-\bibitem{Kempes:2012hy}
-Christopher~P Kempes, Stephanie Dutkiewicz, and Michael~J Follows.
-\newblock {Growth, metabolic partitioning, and the size of microorganisms.}
-\newblock {\em PNAS}, 109(2):495--500, January 2012.
-
-\bibitem{kempes2014morphological}
-Christopher~P Kempes, Chinweike Okegbe, Zwoisaint Mears-Clarke, Michael~J
-  Follows, and Lars~EP Dietrich.
-\newblock Morphological optimization for access to dual oxidants in biofilms.
-\newblock {\em Proceedings of the National Academy of Sciences},
-  111(1):208--213, 2014.
-
-\bibitem{West:2001bv}
-Geoffrey~B West, James~H Brown, and Brian~J Enquist.
-\newblock {A general model for ontogenetic growth}.
-\newblock {\em Nature}, 413(6856):628--631, October 2001.
-
-\bibitem{moses2008rmo}
-Melanie~E Moses, Chen Hou, William~H Woodruff, Geoffrey~B West, Jeffery~C
-  Nekola, Wenyun Zuo, and James~H Brown.
-\newblock {Revisiting a model of ontogenetic growth: Estimating model
-  parameters from theory and data}.
-\newblock {\em http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735},
-  171(5):632--645, May 2008.
-
-\bibitem{hou}
-Chen Hou, Wenyun Zuo, Melanie~E Moses, William~H Woodruff, James~H Brown, and
-  Geoffrey~B West.
-\newblock {Energy uptake and allocation during ontogeny}.
-\newblock {\em Science}, 322(5902):736--739, October 2008.
-
-\bibitem{pirt}
-SJ~Pirt.
-\newblock The maintenance energy of bacteria in growing cultures.
-\newblock {\em Proceedings of the Royal Society of London B: Biological
-  Sciences}, 163(991):224--231, 1965.
-
-\bibitem{Heijnen}
-JJ~Heijnen and JA~Roels.
-\newblock A macroscopic model describing yield and maintenance relationships in
-  aerobic fermentation processes.
-\newblock {\em Biotechnology and Bioengineering}, 23(4):739--763, 1981.
-
-\bibitem{peters1986ecological}
-Robert~Henry Peters.
-\newblock {\em The Ecological Implications of Body Size}, volume~2.
-\newblock Cambridge University Press, Cambridge, 1986.
-
-\bibitem{blueweiss1978relationships}
-L~Blueweiss, H~Fox, V~Kudzma, D~Nakashima, R~Peters, and S~Sams.
-\newblock Relationships between body size and some life history parameters.
-\newblock {\em Oecologia}, 37(2):257--272, 1978.
-
-\bibitem{stryer}
-Lubert Stryer.
-\newblock {\em {Biochemistry, Fourth Edition}}.
-\newblock W.H. Freeman and Company, New York, 1995.
-
-\bibitem{Dunbrack:1993ec}
-Robert~L Dunbrack and Malcolm~A Ramsay.
-\newblock {The Allometry of Mammalian Adaptations to Seasonal Environments: A
-  Critique of the Fasting Endurance Hypothesis}.
-\newblock {\em Oikos}, 66(2):336--342, March 1993.
-
-\bibitem{Lindstedt:1985hm}
-Stan~L Lindstedt and Mark~S Boyce.
-\newblock {Seasonality, Fasting Endurance, and Body Size in Mammals}.
-\newblock {\em Am. Nat.}, 125(6):873--878, June 1985.
-
-\bibitem{Lindstedt:2002td}
-S~L Lindstedt and P~J Schaeffer.
-\newblock {Use of allometry in predicting anatomical and physiological
-  parameters of mammals.}
-\newblock {\em Lab. Anim.}, 36(1):1--19, January 2002.
-
-\bibitem{estermann}
-Beda~L Estermann, Hans-Rudolf Wettstein, Franz Sutter, and Michael Kreuzer.
-\newblock Nutrient and energy conversion of grass-fed dairy and suckler beef
-  cattle kept indoors and on high altitude pasture.
-\newblock {\em Animal Research}, 50(6):477--493, 2001.
-
-\bibitem{michaletz2014convergence}
-Sean~T Michaletz, Dongliang Cheng, Andrew~J Kerkhoff, and Brian~J Enquist.
-\newblock Convergence of terrestrial plant production across global climate
-  gradients.
-\newblock {\em Nature}, 512(7512):39--43, 2014.
-
-\bibitem{damuth1987interspecific}
-John Damuth.
-\newblock Interspecific allometry of population density in mammals and other
-  animals: the independence of body mass and population energy-use.
-\newblock {\em Biological Journal of the Linnean Society}, 31(3):193--246,
-  1987.
-
-\end{thebibliography}
-
-\end{bibunit}
+\bibliography{aa_starving_supplement}
 

### 2017-06-09 Manuscript/Nature-Submission/Nature-temp.tex
Nature submission

--- 2017-03-24 Manuscript/Revision_2017/18.tex
@@
-%\usepackage{scicite}
-%\usepackage{times}
-%\usepackage{graphicx}
-%\usepackage{verbatim}
-%\usepackage{subfigure}
-%\usepackage{tikz}
-%\usepackage{array}
-%\usepackage{tabularx}
-%\usepackage{multirow}
-%\usepackage{multicol}
-%\usepackage{multibox}
-%\usepackage{rotating}
-%\usepackage{layout}
-%\usepackage{epstopdf}
-%\usepackage{afterpage}
-%\usepackage[left]{lineno}
+%% Template for a preprint Letter or Article for submission
+%% to the journal Nature.
+%% Written by Peter Czoschke, 26 February 2004
+%%
 
-%% PNAStmpl.tex
-%% Template file to use for PNAS articles prepared in LaTeX
-%% Version: Apr 14, 2008
+\documentclass{nature-2}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BASIC CLASS FILE
-
-\documentclass{pnastwo}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL GRAPHICS STYLE FILE
-
-\usepackage[pdftex]{graphicx}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL POSTSCRIPT FONT FILES
-
-\usepackage{pnastwoF}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% ADDITIONAL OPTIONAL STYLE FILES
-
-%\usepackage{amssymb,amsfonts,amsmath}
+\usepackage{graphicx}
+\usepackage{verbatim}
+\usepackage{subfigure}
+\usepackage{tikz}
 \usepackage{amsmath}
 \usepackage{amssymb}
-
-% \usepackage[font=large]{caption}
-\usepackage{subfigure}
-\usepackage{epstopdf}
-\usepackage{graphicx}
-\usepackage{verbatim}
 \usepackage{array}
@@
 \usepackage{multirow}
-\usepackage{color}
-%\usepackage{multibox}
+\usepackage{multicol}
+% \usepackage{multibox}
 \usepackage{rotating}
 \usepackage{color}
-%\usepackage{lineno}
-\usepackage[left]{lineno}
+\usepackage{layout}
+\usepackage{epstopdf}
+\usepackage{afterpage}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL MACRO FILES
-% \usepackage[comma,sort&compress]{natbib}
-% \setlength{\abovedisplayskip}{3pt}
-% \setlength{\belowdisplayskip}{3pt}
-% 
-% \usepackage[font=small,labelfont=bf]{caption}
-% 
-% \newcommand{\sid}[1]{\textcolor{red}{\bf [#1]}}
+%% make sure you have the nature.cls and naturemag.bst files where
+%% LaTeX can find them
 
-\bibliographystyle{pnas}
+\bibliographystyle{naturemag}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{The dynamics of starvation and recovery}
 
-%% Don't type in anything in the following section:
-%%%%%%%%%%%%
-%% For PNAS Only:
-\contributor{Submitted to Proceedings
-of the National Academy of Sciences of the United States of America}
-\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
-\copyrightyear{2011}
-\issuedate{Issue Date}
-\volume{Volume}
-\issuenumber{Issue Number}
-%%%%%%%%%%%%
+%% Notice placement of commas and superscripts and use of &
+%% in the author list
+
+\author{Justin D. Yeakel$^{1,2,3,4}$, Christopher P. Kempes$^{2,3}$ \& Sidney Redner$^{2,3}$}
 
@@
 
-\linenumbers
-\setlength\linenumbersep{3pt}
-
-%\title{Ecological and evolutionary implications of starvation and body size}
-
-%\title{Survival of the Stable: Dynamic Consequences of Starvation and Reproduction}
-
-%\title{Size Constraints on Starvation \& Reproduction Dynamics}
-
-%\title{Eco-evolutionary Consequences of Starvation and Reproduction: Survival
-%of the Stable}
-
-\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
-
-\author
-{Justin D.
-Yeakel\affil{1}{School of Natural Science, University of California Merced, Merced, CA} \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}\affil{5}{To whom correspondence should be addressed: jdyeakel@gmail.com}, Christopher P.
-Kempes \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}, \and Sidney Redner \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{3}{Department of Physics, Boston University, Boston MA}\affil{4}{Contributed equally}
-}
-
-\contributor{Submitted to Proceedings of the National Academy of Sciences
-of the United States of America}
-
 \maketitle
 
-\begin{article}
+\begin{affiliations}
+ \item School of Natural Science, University of California Merced, Merced, CA, USA
+ \item The Santa Fe Institute, Santa Fe, NM, USA
+ \item Contributed equally
+ \item Corresponding author jdyeakel@gmail.com
+\end{affiliations}
 
-\begin{abstract} %250 words
+\begin{abstract}
 The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of its constituent individuals.
@@
 \end{abstract}
-\keywords{foraging | starvation|reproduction}
 
-\noindent {\small {\bf Significance Statement} %105/120 words
-Energetic investment in somatic maintenance and growth vs.
+%\noindent {\small {\bf Significance Statement} %105/120 words
+%Energetic investment in somatic maintenance and growth vs.
 reproduction directly impacts the dynamics of populations among species.
@@
 provides a mechanistic driver for the evolutionary trend towards larger body size known as Cope's rule.}
-
-\vspace{2mm}
 
@@
 In the supplementary information (SI) we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we describe below.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{fig_FixedPoint2.eps}
-\caption{\small{ The transcritical (dashed) and Hopf bifurcation (solid) as a
-  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-  bifurcation conditions separate parameter space into unphysical, cyclic,
-  and steady state dynamic regimes.  The colors show the steady state densities for the energetically replete consumers $F^*$, (warmer colors denote higher densities).  
-  % Steady state densities for the energetically deficient consumers $H^*$ (not shown)
-  % scale with those for $F^*$.  
-}\label{fig:fp}}
-\end{figure}
 
@@
 % If $\rho$ exceeds a threshold value, cyclic dynamics will develop as the Hopf bifurcation is approached.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.eps}
-\caption{\small{ The growth trajectory over absolute time of an individual organism as a function of body mass.  
-Initial growth follows the black trajectory to an energetically replete reproductive adult mass $m=\epsilon_\lambda M$. %which we assume is 95\% asymptotic mass $M$.  
-Starvation follows the red trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. 
-Recovery follows the green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires a longer time to reach $\epsilon_\lambda M$.
-%different energetics and
-Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}\label{fig:growth}}
-\end{figure}
 
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometricComb3.eps}
-\caption{\small{ Probability of extinction for a consumer with (\emph{A}) $M=10^2$g, (\emph{B}) $M=10^4$g, and (\emph{C}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
-Note the change in scale for $M=10^4$ and $M=10^6$g.
-Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
-The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
-\end{figure*}
-
 To determine the starvation rate, $\sigma$, we are interested in the time
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.48\textwidth]{fig_FPAllometric2.eps}
-\caption{\small{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-  body mass. Inset: Resource steady state $R^*$ as a function of consumer body mass.
-  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  The data are from Damuth \cite{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
-\cite{West:2001bv,hou,moses2008rmo}).}\label{fig:mass}}
-\end{figure}
-
 %Minimal mass bound
@@
 %reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
-\caption{\small{ Invasion feasibility for organisms with a proportional change in
-  mass $\chi$ against a population with a resident body mass $M$.  The blue
-  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
-  red region denotes values of $\chi$ that result in a mass that is below the
-  starvation threshold and are thus infeasible.
-  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
-  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}\label{fig:invasion}}
-\end{figure}
 
@@
 
-\begin{acknowledgments}
-  We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
+% \bibliography{ADD}
+
+\begin{addendum}
+ \item We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
   discussions and comments on the manuscript.  J.D.Y. was supported by
@@
   Institute.
-\end{acknowledgments}
+ \item[Competing Interests] The authors declare that they have no
+competing financial interests.
+ \item[Correspondence] Correspondence and requests for materials
+should be addressed to J.D.Y.~(email:  jdyeakel@gmail.com).
+\end{addendum}
 
-%\onecolumn
-%\let\cleardoublepage\clearpage
+\clearpage
+\newpage
+
+%FIGURES
+
+\renewcommand{\figurename}{{\bf Figure}}
+\renewcommand{\thefigure}{{\bf \arabic{figure}}}
 
@@
 
-\end{article}
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{fig_FixedPoint2.eps}
+\caption{\small{ The transcritical (dashed) and Hopf bifurcation (solid) as a
+  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
+  bifurcation conditions separate parameter space into unphysical, cyclic,
+  and steady state dynamic regimes.  The colors show the steady state densities for the energetically replete consumers $F^*$, (warmer colors denote higher densities).  
+  % Steady state densities for the energetically deficient consumers $H^*$ (not shown)
+  % scale with those for $F^*$.  
+}\label{fig:fp}}
+\end{figure}
 
-% 
-% \begin{figure*}
-% \centering
-% \includegraphics[width=0.5\textwidth]{fig_Rates.eps}
-% \caption{
-% Allometrically constrained starvation rate $\sigma=1/t_{\sigma}$ (red), mortality $\mu=1/t_{\mu}$ (blue) and recovery rate $\rho=1/t_{\rho}$ (green) relative to the reproductive rate $\lambda=1/t_{\lambda}$ (black) as a function of body mass (see Equations \ref{t1}, \ref{rhotimescale}, \ref{eq:sigma}, and \ref{mutimescale}).
-% The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
-% Mean values $\pm 25\%$ variation are shown by the shaded region for each rate.
-% }
-% \label{fig:gvs}
-% \end{figure*}
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{Growth-trajectory-diagram.eps}
+\caption{\small{ The growth trajectory over absolute time of an individual organism as a function of body mass.  
+Initial growth follows the black trajectory to an energetically replete reproductive adult mass $m=\epsilon_\lambda M$. %which we assume is 95\% asymptotic mass $M$.  
+Starvation follows the red trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. 
+Recovery follows the green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires a longer time to reach $\epsilon_\lambda M$.
+%different energetics and
+Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}\label{fig:growth}}
+\end{figure}
 
-% 
-% \begin{figure*}
-% \centering
-% \includegraphics[width=0.5\textwidth]{fig_DataHopf.eps}
-% \caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
-%   allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
-%   a function of different mammalian body sizes: $M=P\times10^2$g (blue) and
-%   $M=P\times10^6$g (green), where $P$ is a random uniform variable in $[1,9]$.
-%   Points denote realized values of $\sigma$ and $\rho$ given the drawn values for $M$.
-%   }
-% \label{fig:hopf}
-% \end{figure*}
+\begin{figure*}
+\centering
+\includegraphics[width=0.9\textwidth]{fig_ExtinctionAllometricComb3.eps}
+\caption{\small{ Probability of extinction for a consumer with (\emph{A}) $M=10^2$g, (\emph{B}) $M=10^4$g, and (\emph{C}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
+Note the change in scale for $M=10^4$ and $M=10^6$g.
+Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
+The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
+\end{figure*}
 
-%
-%
-%	 \begin{figure}[h][h]
-%		 \centering
-%		 \includegraphics[width=0.5\textwidth]{Growth-trajectory-diagram.pdf}
-%		 \caption{
-%		 }
-%		 \label{growth-diagram}
-%	 \end{figure}
-%
-%	 \begin{figure}[h][h]
-%		 \centering
-%		 \includegraphics[width=0.25\textwidth]{fig_GvS.pdf}
-%		 \caption{
-%		 }
-%		 \label{GvS}
-%	 \end{figure}
-%
-%
-%	\begin{figure}[h][h]
-%	 	\centering
-%	 	\includegraphics[width=0.25\textwidth]{fig_FixedPoint.pdf}
-%	 	\caption{
-%	 	}
-%		\label{Hopfb}
-%	\end{figure}
-%
-%
-%	\begin{figure}[h][h]
-%		\centering
-%		\includegraphics[width=0.8\textwidth]{fig_DataHopf.pdf}
-%		\caption{
-%		}
-%		\label{DataHopf}
-%	\end{figure}
-%
-%
-%	\begin{figure}[h][h]
-%		\centering
-%	 	\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometric.pdf}
-%	 	\caption{
-%	 	}
-%	 	\label{Ext}
-%	\end{figure}
-%
-%	\begin{figure}[h][h]
-%	 	\centering
-%	 	\includegraphics[width=0.5\textwidth]{fig_FPAllometric.pdf}
-%	 	\caption{
-%	 	}
-%	 	\label{Asymp}
-%	\end{figure}
-%
-%\clearpage
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{fig_FPAllometric2.eps}
+\caption{\small{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
+  body mass. Inset: Resource steady state $R^*$ as a function of consumer body mass.
+  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
+  The data are from Damuth \cite{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
+\cite{West:2001bv,hou,moses2008rmo}).}\label{fig:mass}}
+\end{figure}
 
-%\begin{table}[h]
-%\caption{Parameter Values For Various Classes of Organisms}
-%\label{param}
-%    \begin{center}
-%    \small
-%     \begin{tabular}{| p{1.2cm}| p{3.2cm} | p{2.6cm} | p{3.2cm} | }
-%     \hline
-%     & {\bf Mammals} & {\bf Unicellular Eukaryotes} & {\bf Bacteria} \\
-%     \hline
-%   $\eta$ & $3/4$ & & $1.70$ \\
-%   $E_{m}$ & $10695$ (J gram$^{-1}$) & & $10695$ (J gram$^{-1}$) \\
-%   $E_{m}^{\prime}$ & $\approx E_{m}$ & & $\approx E_{m}$ \\
-%   $B_{0}$ & $0.019$ (W gram$^{-\alpha}$) & & $1.96\times10^{17}$ \\
-%   $B_{m}$ & $0.025$ (W gram$^{-1}$)   & & $0.025$ (W gram$^{-1}$)\\
-%   $a$ & $1.78\times10^{-6}$ & & $1.83\times10^{13}$ \\
-%   $b$ & $2.29\times10^{-6}$ & & $2.29\times10^{-6}$ \\
-%   $\eta-1$ & $-0.21$ & & $0.73$ \\
-%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) & & $56493$ \\
-%   $\gamma$ & $1.19$ & & $0.68$ \\
-%   $f_{0}$ & $0.02$ & & $1.30\times10^{-5}$ \\
-%   $\zeta$ & $1.01$ & & \\
-%   $mm_{0}$ & $0.32$ & & \\
-%
-%
-%   \hline
-%    \end{tabular}
-%    \end{center}
-%   \end{table}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.6\textwidth]{fig_Invasion.eps}
+\caption{\small{ Invasion feasibility for organisms with a proportional change in
+  mass $\chi$ against a population with a resident body mass $M$.  The blue
+  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
+  red region denotes values of $\chi$ that result in a mass that is below the
+  starvation threshold and are thus infeasible.
+  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
+  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}\label{fig:invasion}}
+\end{figure}
+
+%\pagestyle{empty}
+
+% EXTENDED FIGURES
+
+%\setcounter{figure}{0}
+%\renewcommand{\figurename}{{\bf Extended Figure}}
+%\renewcommand{\thefigure}{{\bf \arabic{figure}}}
 

### 2017-06-09 Manuscript/Nature-Submission/Revised-Supplement-1.tex
Nature submission

--- 2017-02-28 Manuscript/Revision_2017/Revised-Supplement-1.tex
@@
 It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter such that these steady states match the scale of know data from Damuth \cite{}.
-We find that the best fit is $\alpha=VALUE\times10^{-9}$ (s$^{-1}$).
+We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$).
 However, two points are important to note here: first, our overall framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, the asymptotic behavior of $F^{*}$ and $H^{*}$ is also independent of $\alpha$, that is, our prediction of the maximum mammal size does not depend on $\alpha$.

### 2017-06-09 Manuscript/Nature-Submission/Revised-Supplement-2.tex
Nature submission

--- 2017-03-20 Manuscript/Arxiv_Revision/YKR_Arxiv.tex
@@
-\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
-%\usepackage[pdftex]{graphicx}
-
-\usepackage{amsmath,amsfonts,amssymb}
-\usepackage[english]{babel}
-\usepackage[latin1]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{color}
-\usepackage{float}
-\usepackage{verbatim}
-\usepackage{graphicx}
-\usepackage{bm}
-\usepackage{mathtools}
-\usepackage{stmaryrd}
-\usepackage{anyfontsize}
-
-%\usepackage{epstopdf}
+%\usepackage{scicite}
+%\usepackage{times}
+%\usepackage{graphicx}
+%\usepackage{verbatim}
+%\usepackage{subfigure}
+%\usepackage{tikz}
 %\usepackage{array}
@@
 %\usepackage{multirow}
+%\usepackage{multicol}
+%\usepackage{multibox}
+%\usepackage{rotating}
+%\usepackage{layout}
+%\usepackage{epstopdf}
+%\usepackage{afterpage}
+%\usepackage[left]{lineno}
+
+%% PNAStmpl.tex
+%% Template file to use for PNAS articles prepared in LaTeX
+%% Version: Apr 14, 2008
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% BASIC CLASS FILE
+
+\documentclass{pnastwo}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% OPTIONAL GRAPHICS STYLE FILE
+
+\usepackage[pdftex]{graphicx}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% OPTIONAL POSTSCRIPT FONT FILES
+
+\usepackage{pnastwof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% ADDITIONAL OPTIONAL STYLE FILES
+
+%\usepackage{amssymb,amsfonts,amsmath}
+\usepackage{amsmath}
+\usepackage{amssymb}
+
+\usepackage[font=large]{caption}
+\usepackage{subfigure}
+\usepackage{epstopdf}
+\usepackage{graphicx}
+\usepackage{verbatim}
+\usepackage{array}
+\usepackage{tabularx}
+\usepackage{multirow}
 \usepackage{color}
 %\usepackage{multibox}
-%\usepackage{rotating}
+\usepackage{rotating}
+\usepackage{color}
 %\usepackage{lineno}
-%\usepackage[left]{lineno}
-%\usepackage[comma,sort&compress]{natbib}
-%\usepackage{authblk}
-%\usepackage{multicol}
+\usepackage[left]{lineno}
+\usepackage[comma,sort&compress]{natbib}
 
-%\bibliographystyle{ieeetr}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% OPTIONAL MACRO FILES
 
-\usepackage{bibunits}
+\bibliographystyle{pnas}
 
-%\linenumbers
-%\setlength\linenumbersep{3pt}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newcommand{\sid}[1]{\textcolor{red}{\bf [#1]}}
+
+%% Don't type in anything in the following section:
+%%%%%%%%%%%%
+%% For PNAS Only:
+\contributor{Submitted to Proceedings
+of the National Academy of Sciences of the United States of America}
+\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
+\copyrightyear{2011}
+\issuedate{Issue Date}
+\volume{Volume}
+\issuenumber{Issue Number}
+%%%%%%%%%%%%
 
@@
 
-\author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
-  of California, Merced, Merced, CA 95340, USA}
+\linenumbers
+\setlength\linenumbersep{3pt}
 
-\author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
-  Park Road, Santa Fe, NM 87501, USA}
+%\title{Ecological and evolutionary implications of starvation and body size}
 
-\author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
-  Road, Santa Fe, NM 87501, USA}
+%\title{Survival of the Stable: Dynamic Consequences of Starvation and Reproduction}
 
-\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
+%\title{Size Constraints on Starvation \& Reproduction Dynamics}
 
-\begin{abstract} %250 words
-The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of its constituent individuals.
-Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery in resource-limited environments.
-To elucidate the consequences of this tradeoff, we introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete consumers that reproduce, and undernourished, non-reproducing consumers that are susceptible to mortality.
-As a function of the transition rates between these two states that are determined by the abundance of resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
-We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships between body size and a variety of traits and find that population dynamics subject to these constraints are typically driven to a steady state.
-Moreover, we find that these rates fall within a `refuge' in parameter space, where the probability of extinction of the consumer population is minimized.
-Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations.
-Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percentage of body fat.
-For body masses $< 1.748\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.748\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
-\end{abstract}
+%\title{Eco-evolutionary Consequences of Starvation and Reproduction: Survival
+%of the Stable}
+
+\title{Supporting Information for ``The dynamics of starvation and recovery''}%: Eco-evolutionary feedbacks}
+
+\author
+{Justin D.
+Yeakel\affil{1}{School of Natural Science, University of California Merced, Merced, CA} \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}\affil{5}{To whom correspondence should be addressed: jdyeakel@gmail.com}, Christopher P.
+Kempes \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}, \and Sidney Redner \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{3}{Department of Physics, Boston University, Boston MA}\affil{4}{Contributed equally}
+}
+
+\contributor{Submitted to Proceedings of the National Academy of Sciences
+of the United States of America}
 
@@
 
-\begin{bibunit}[unsrt]
+\begin{article}
 
-\section*{Introduction}
-The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly influences how they invest reserves in uncertain environments.
-Such behaviors are generally manifested as tradeoffs between investing in somatic maintenance and growth, or allocating energy towards reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
-The timing of these behaviors responds to selective pressure, as the choice of the investment impacts future fitness~\cite{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.
-The influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
-
-The balance between (a) somatic growth and maintenance, and (b) reproduction depends on resource availability~\cite{Morris:1987eo}.
-For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\cite{Tveraa:2003fq}.
-Many bird species invest differently in broods during periods of resource scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
-Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\cite{Kirk:1997cc}.
-Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate nutritional reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
-
-Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.
-The data collected thus far has shown that diverse mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development (blastocyst implantation) until nutritional reserves can be accumulated~\cite{Mead:1989dt,Sandell:1990kw}.
-Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
-In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the importance and universality of the fundamental tradeoff between somatic and reproductive investment.
-However the general dynamic implications of these constraints are unknown.
-
-Though straightforward conceptually, incorporating the energetic dynamics of individuals~\cite{Kooi2000} into a population-level framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles~\cite{Diekmann:2010da}.
-An alternative approach involves modeling the macroscale relations that guide somatic versus reproductive investment in a consumer-resource system.
-For example, macroscale Lotka-Volterra models assume that the growth rate of the consumer population depends on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction~\cite{murdoch:2003}.
-
-In this work, we adopt an alternative approach in which we \emph{explicitly} account for resource limitation and the subsequent effects of starvation.
-Namely, only individuals with sufficient energetic reserves can reproduce.
-Such a constraint leads to reproductive time lags due to some members of the population going hungry and then recovering.
-Additionally, we incorporate the idea that reproduction is strongly constrained allometrically \cite{Kempes:2012hy}, and is not generally linearly related to resource density.
-As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.
-\\
-
-\section*{Nutritional state-structured model (NSM)}
-We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$.
-The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity $C$.
-The rate at which consumers transition between states and consume resources is dependent on their overall abundance, the abundance of resources, the efficiency of converting resources into metabolism, and how that metabolism is partitioned between maintenance and growth purposes.
-In the supplementary information (SI) we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we describe below.
-
-Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$.
-Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a ratio between maximal resource consumption and the carrying capacity of the resource.
-%and $\approx 2$.
-Resources are eaten by the hungry consumers at rate $\rho R + \delta$, that accounts for their somatic growth ($\rho R$) and  maintenance ($\delta$).
-Full consumers eat resources at a constant rate $\beta$ that accounts for maximal maintenance and somatic growth (see SI for mechanistic derivations of these rates from resource energetics).
-%This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
-The NSM represents an ecologically motivated fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\cite{Persson:1998hz}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_FixedPoint.png}
-\caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
-  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-  bifurcation conditions separate parameter space into unphysical, cyclic,
-  and steady state dynamic regimes.  The colors show the steady
-  state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
-  replete consumers $F^*$, (warmer colors denote higher densities).  Steady
-  state densities for the energetically deficient consumers $H^*$ (not shown)
-  scale with those for $F^*$.  }
-\label{fig:fp}
-\end{figure}
-
-In the mean-field approximation, in which the consumers and resources are perfectly mixed, their densities evolve according to the rate equations
-\begin{align}
-\label{eq:system}
-\begin{split}
-\frac{dF}{dt} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
-\frac{dH}{dt} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-\frac{dR}{dt} &= \alpha \left(1-R\right)R -\left(\rho R+\delta\right)H-\beta F
-\end{split}
-\end{align}
-This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see SI for a full derivation and the dimensional form).
-Notice that the total consumer density $F+H$ evolves according to $\frac{dF}{dt}+\frac{dH}{dt}=\lambda F-\mu H$.
-This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\cite{murray2011mathematical}, except that the resource density does not appear in the growth term.
-As discussed above, the attributes of reproduction and mortality have been explicitly apportioned to the full and hungry consumers, respectively, so that the growth in the total density is decoupled from the resource density.
-
-Equation~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
-\begin{align}
-\label{eq:ss}
-\begin{split}
-F^* &= (\sigma-\lambda)\frac{ \alpha  \lambda  \mu ^2  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
-H^* &= (\sigma-\lambda)\frac{ \alpha  \lambda ^2 \mu  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
-R^* &= (\sigma - \lambda)\frac{\mu  }{A}.
-\end{split}
-\end{align}
-where $A=(\lambda  \xi  \rho +\mu  \sigma )$ and $B=(\beta  \mu  \xi +\delta  \lambda  \xi -\lambda  \mu )$.
-The stability of this fixed point is determined by the Jacobian matrix $\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.
-The parameters in Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue of $\mathbf{J}$ is negative, so that the system is stable with respect to small perturbations from the fixed point.
-Because this fixed point is unique, it is the global attractor for all population trajectories for any initial condition where the resource and consumer densities are both nonzero.
-
-From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that the consumer and resource densities are positive.
-The condition $\sigma = \lambda$ thus represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates a physical from an unphysical regime where all steady-state densities become negative after intersecting the trivial fixed point $(F^*,H^*,R^*)=(0,0,0)$.
-The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
-As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges.
-
-In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).
-In continuous-time systems, a limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts~\cite{GuckHolmes}.
-This Hopf bifurcation is defined by ${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial of the Jacobian matrix~\cite{Gross:2004p2428}.
-As the system parameters are tuned to be within the stable regime, but close to the Hopf bifurcation, the amplitude of the transient cycles becomes large.
-Given that ecological systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase the extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
-
-When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the consumers are driven to the hungry non-reproducing state.
-Because reproduction is inhibited, there is a low steady-state consumer density and a high steady-state resource density.
-However, if $\sigma/\lambda\to 1$ from above, the population is overloaded with energetically-replete (reproducing) individuals, thereby promoting transient oscillations between the consumer and resource densities (Fig.~\ref{fig:fp}).
-If the starvation rate is low enough that the Hopf bifurcation is crossed, these oscillations become stable over time.
-This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.
-
-\section*{Role of allometry}
-While there are no a priori constraints on the parameters in the NSM, most
-organisms correspond to restricted portions of the parameter space.  Here we
-use allometric scaling relations to constrain the covariation of rates in a
-principled and biologically meaningful manner.  Allometric scaling relations
-highlight common constraints and average trends across large ranges in body
-size and species diversity. Many of these relations can be derived from a
-small set of assumptions and below we describe our framework to determine the
-covariation of timescales and rates across the range of mammals for each of
-the key parameters of our model (cf. ref.~\cite{Yodzis:1992hg}).  We are thereby
-able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.
-
-Nearly all of the rates described in the NSM are determined by consumer
-metabolism, which can be used to describe a variety of organismal features
-\cite{Brown:2004wq}.  The scaling relation between an organism's metabolic
-rate $B$ and its body mass $M$ at reproductive maturity is known to scale as
-$B = B_0 M^\eta$~\cite{West:2002it}, where the scaling exponent $\eta$ is
-typically close to $2/3$ or $3/4$ for metazoans (e.g., ref. \cite{Brown:2004wq}),
-and has taxonomic shifts for unicellular species between $\eta\approx 1$ in
-eukaryotes and $\eta\approx 1.76$ in bacteria
-\cite{DeLong:2010dy,Kempes:2012hy}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.pdf}
-\caption{ The growth trajectory over absolute time of an individual organism
-  as a function of body mass.  Initial growth follows the black trajectory to
-  an energetically replete reproductive adult mass $m=\epsilon_\lambda M$ which we assume is 95\% asymptotic mass $M$.  Starvation follows the red
-  trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. Recovery follows the
-  green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires different energetics and a longer time to reach $\epsilon_\lambda M$.
-Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}
-\label{fig:growth}
-\end{figure}
-
-Several efforts have shown how a partitioning of $B$ between growth and
-maintenance purposes can be used to derive a general equation for both the
-growth trajectories and growth rates of organisms ranging from bacteria to
-metazoans
-\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. This relation is derived from the simple balance condition \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}
-\begin{eqnarray}
-\label{balance}
-B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
-\end{eqnarray}
-where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
-the metabolic rate to support an existing unit of mass, and $m$ is the mass
-of the organism at any point in its development.  This balance has the
-general solution \cite{bettencourt,Kempes:2012hy}
-\begin{eqnarray}
-\label{m1}
-\left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}},
-\end{eqnarray}
-where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically (see SI).
-We now use this solution to define the timescale for reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \cite{moses2008rmo} for a detailed presentation of these timescales).
-The time that it takes to reach a particular mass $\epsilon M$ is given by the timescale
-\begin{equation}
-\label{t1}
-\tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)},
-\end{equation}
-where we will define values of $\epsilon$ to describe a set of rates within our model.
-For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$; \cite{West:2001bv}).
-The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$, this rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_Rates.pdf}
-\caption{
-Allometrically constrained starvation rate $\sigma=1/t_{\sigma}$ (red), mortality $\mu=1/t_{\mu}$ (blue) and recovery rate $\rho=1/t_{\rho}$ (green) relative to the reproductive rate $\lambda=1/t_{\lambda}$ (black) as a function of body mass (see Equations \ref{t1}, \ref{rhotimescale}, \ref{eq:sigma}, and \ref{mutimescale}).
-The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
-Mean values $\pm 25\%$ variation are shown by the shaded region for each rate.
-}
-\label{fig:gvs}
-\end{figure}
-
-The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
-sufficient tissue to transition from the hungry to the full state.  Since
-only certain tissues can be digested for energy (for example the brain cannot
-be degraded to fuel metabolism), we define the rates for starvation, death,
-and recovery by the timescales required to reach, or return from, specific
-fractions of the replete-state mass (Fig. \ref{fig:gvs}; see SI, Table I for parameterizations).  We define
-$m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction of
-replete-state mass where reproduction ceases. This fraction will deviate from a constant
-if tissue composition systematically scales with adult mass.  For example,
-making use of the observation that body fat in mammals scales with overall
-body size according to $M_{\rm fat}=f_{0}M^{\gamma}$ and assuming that once
-this mass is fully digested the organism starves, this would imply that
-$\epsilon_{\sigma}=1-f_{0}M^{\gamma}/M$. It follows that the
-recovery timescale, $t_{\rho}$, is the time to go from
-$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$ (Fig.
-\ref{fig:growth}).
-Using Eqs.~\eqref{m1} and \eqref{t1} this timescale is given by simply considering an adjusted starting mass of $m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
-\begin{eqnarray}
-\label{rhotimescale}
-t_{\rho}=\ln\left[\frac{1-\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)^{1-\eta}}{1-\epsilon_\lambda^{1-\eta}}\right]\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}
-\end{eqnarray}
-where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in the biosynthetic energetics during recovery (see SI).
-It should be noted that more complicated ontogenetic models explicitly handle
-storage \cite{hou}, whereas this feature is implicitly covered by the body
-fat scaling in our framework.
-
-To determine the starvation rate, $\sigma$, we are interested in the time
-required for an organism to go from a mature adult that reproduces at rate
-$\lambda$, to a
-reduced-mass hungry state where reproduction is impossible.
-For starving individuals we assume that an organism must meet its maintenance requirements by using the digestion of existing mass as the sole energy source.
-This assumption implies the following simple metabolic balance
-\begin{eqnarray}
-\frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
-\end{eqnarray}
-or
-\begin{eqnarray}
-\frac{dm}{dt}=-\frac{a^{\prime}}{M^{1-\eta}}m
-\end{eqnarray}
-where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
-body mass, which differs from $E_{m}$, the energy required to
-synthesis a unit of biomass \cite{hou}. Given the replete mass, $M$, of an organism, the
-above energy balance prescribes the mass trajectory of a non-consuming
-organism:
-\begin{eqnarray}
-\label{mt}
-m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}.
-\end{eqnarray}
-The timescale for starvation is
-given by the time it takes $m(t)$ to reach $\epsilon_{\sigma} M$, which gives
-\begin{equation}
-\label{eq:sigma}
-t_{\sigma}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\sigma}\right).
-\end{equation}
-The starvation rate is then $\sigma=1/t_{\sigma}$, which scales with
-replete-state mass as $1/M^{1-\eta}\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important
-feature is that $\sigma$ does not have a simple scaling dependence on
-$\lambda$ (Fig. \ref{fig:gvs}), which is important for the dynamics that we
-later discuss.
-
-The time to death should follow a similar relation, but defined by a lower
-fraction of replete-state mass, $m_{\mu}=\epsilon_{\mu} M$ where $\epsilon_\mu < \epsilon_\sigma$.
-Suppose, for example, that an organism dies once it has digested all fat and
-muscle tissues, and that muscle tissue scales with body mass according to
-$M_{\rm musc}=u_{0}M^{\zeta}$.  This gives
-$\epsilon_{\mu}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
-mass has been shown to be roughly proportional to body mass~\cite{Folland:2008ij} in
-mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant. The time to go from starvation to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or
-\begin{eqnarray}
-\label{mutimescale}
-t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
-\end{eqnarray}
-and $\mu=1/t_{\mu}$.
-
-Although the rate equations \eqref{eq:system} are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
-Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
-Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
-
-\section*{Stabilizing effects of allometric constraints}
-As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
-Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values (Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.
-We thus find that the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
-Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
-These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
-However, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation.
-This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.425\textwidth]{fig_DataHopf.pdf}
-\caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
-  allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
-  a function of different mammalian body sizes: $M=P\times10^2$g (blue) and
-  $M=P\times10^6$g (green), where $P$ is a random uniform variable in $[1,9]$.
-  Points denote realized values of $\sigma$ and $\rho$ given the drawn values for $M$.
-  }
-\label{fig:hopf}
-\end{figure}
-
-Previous studies used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
-However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
-Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
-However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.
-
-\begin{figure*}
-\centering
-\includegraphics[width=1\textwidth]{fig_ExtinctionAllometricComb3.png}
-\caption{ Probability of extinction for a consumer with (\emph{A}) $M=10^2$g, (\emph{B}) $M=10^4$g, and (\emph{C}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
-Note the change in scale for $M=10^4$ and $M=10^6$g.
-Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
-The white points denote the allometrically constrained starvation and recovery rate.}
-\label{fig:ext}
-\end{figure*}
-
-%Extinction to the left and the right
-\section*{Extinction risk}
-Within our model, higher rates of starvation result in a larger flux of the population to the hungry state.
-In this state, reproduction is absent, thus increasing the likelihood of extinction.
-From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
-We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
-To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
-This procedure is repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for organisms with mass ranging from $10^2$ to $10^6$ grams.
-%the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
-In each replicate the initial densities are chosen to be $(XF^*,XH^*,R^*)$, with $X$ a random variable that is uniformly distributed in $[0,2]$.
-By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-7}$ to $10^{-3}$.
-%, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
-As expected, higher rates of extinction correlate with both high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if $\sigma$ is small.
-For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
-For high values of $\sigma$ and low values of $\rho$, increased extinction risk arises because of the decrease in the steady-state consumer population density (Figs.
-\ref{fig:fp}\emph{B}, \ref{fig:ext}).
-This interplay creates an `extinction refuge', such that for a constrained range of $\sigma$ and $\rho$, extinction probabilities are minimized.
-
-We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge across a range of $M$ (Fig. \ref{fig:ext}\emph{A-C}, white points).
-These values are close enough to the Hopf bifurcation to avoid low steady-state densities, and far enough away to avoid large-amplitude transient cycles.
-The feature that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
-Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that serve to minimize extinction risk.
-This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
-We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig.
-\ref{fig:ext}\emph{A-C}), suggesting that the population dynamics of larger organisms are more sensitive to smaller changes in physiological rates controlling starvation and recovery.
-% OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-{\it To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.}\\
-
-%Results are insensitive to alpha but possibly sensitive to beta (eff parameter)
-
-%NOTES 7/8/16
-%Calder has a hypothesis that period ~ Mass^(1/4).
-%Can we get "transient period" from Jacobian?
-%Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.
-
-\section*{Dynamic and energetic barriers to body size}
-Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\cite{West:2002ud}.
-Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
-One important such example is the \emph{fasting endurance hypothesis}, which contends that larger body size, with consequent lower metabolic rates and increased ability to maintain more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability~\cite{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's rule)~\cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\cite{Clauset:2009fh}.
-These trends are thought to be driven by a combination of climate change and niche availability~\cite{Saarinen:2014br}; however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, have not been explored.
-We argue that the NSM provides a suitable framework to explore these issues.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{fig_FPAllometric.pdf}
-\caption{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-  body mass.
-  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.
-The data are from Damuth \cite{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
-\cite{West:2001bv,hou,moses2008rmo}).}
-\label{fig:mass}
-\end{figure}
-
-The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
-Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{Damuth:1987kr,allen2002,enquist1998}.
-The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size~\cite{Damuth:1987kr,allen2002,enquist1998}.
-This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism, such that $N^*\propto M^{-3/4}/B_{0}$~\cite{allen2002,enquist1998}.
-This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{Damuth:1987kr,allen2002,enquist1998}.
-Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
-This result is remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to energy equivalence.
-Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
-In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
-This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero causing the timescales to become infinite (see Equation \ref{mutimescale}) and the rates $\mu$ and $\lambda$ to equal zero.
-The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$.
-This point predicts a strong upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
-Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
-It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters.
-We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\cite{Kempes:2012hy,Kempes:2016}.
-% reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
-
-We contend that the NSM provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time.
-The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
-\ref{fig:mass}\emph{C}), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
-Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but does not offer a within-lineage mechanism by which larger body sizes are selected for.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_Invasion.pdf}
-\caption{ Invasion feasibility for organisms with a proportional change in
-  mass $\chi$ against a population with a resident body mass $M$.  The blue
-  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
-  red region denotes values of $\chi$ that result in a mass that is below the
-  starvation threshold and are thus infeasible.
-  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
-  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}
-\label{fig:invasion}
-\end{figure}
-
-To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120$ times the mass of a male African elephant.
-Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
-We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by $^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$.
-For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given Equation \ref{t1}, implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
-Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
-The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
-Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the body composition of starved foragers, is assumed to remain unchanged for the invader population (see SI for detailed derivations of invader rates).
-
-To assess the susceptibility of the resident population to invasion, we determine which consumer has a lower steady-state resource density for a given value of $\chi$, again with the expectation that populations able to survive on lower resource densities have a competitive advantage \cite{tilman1981}.
-We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
-However, for $M> 1.748\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
-
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
-This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca.
-$1.5\times10^7$g and the late Miocene \emph{Deinotherium} at ca.
-$1.74\times10^7$g ~\cite{Smith:2010p3442}.
-Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \cite{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals outlined by Alroy~\cite{Alroy:1998p1594}.
-While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general within-lineage mechanism for the evolution of larger body size among terrestrial mammals.
-
-The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations~\cite{Stearns:1989ip}.
-The NSM is a minimal and general model that incorporates the dynamics of starvation and recovery that are expected to occur in resource-limited environments.
-By incorporating allometric relations between the rates in the NSM, we found: (i) different organismal masses have distinct population dynamic regimes, (ii) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and (iii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of maximum body size.
-We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.
-Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.
-
-% \putbib[aa_starving3]
-
-\begin{thebibliography}{10}
-
-\bibitem{Martin:1987dl}
-Thomas~E Martin.
-\newblock {Food as a limit on breeding birds: A life-history perspective}.
-\newblock {\em Annu. Rev. Ecol. Syst.}, 18:453--487, January 1987.
-
-\bibitem{Kirk:1997cc}
-Kevin~L Kirk.
-\newblock {Life-history responses to variable environments: Starvation and
-  reproduction in planktonic rotifers}.
-\newblock {\em Ecology}, 78(2):434--441, March 1997.
-
-\bibitem{Kempes:2012hy}
-C~P Kempes, S~Dutkiewicz, and M~J Follows.
-\newblock {Growth, metabolic partitioning, and the size of microorganisms}.
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 109(2):495--500, January 2012.
-
-\bibitem{Mangel:1988uaa}
-M~Mangel and C.~W. Clark.
-\newblock {\em {Dynamic Modeling in Behavioral Ecology}}.
-\newblock Princeton University Press, Princeton, 1988.
-
-\bibitem{Mangel:2014kz}
-Marc Mangel.
-\newblock {Stochastic dynamic programming illuminates the link between
-  environment, physiology, and evolution}.
-\newblock {\em B. Math. Biol.}, 77(5):857--877, 2014.
-
-\bibitem{Yeakel:2013hi}
-Justin~D Yeakel, Nathaniel~J Dominy, Paul~L Koch, and Marc Mangel.
-\newblock {Functional morphology, stable isotopes, and human evolution: a model
-  of consilience}.
-\newblock {\em Evolution}, 68(1):190--203, 2014.
-
-\bibitem{Morris:1987eo}
-Douglas~W Morris.
-\newblock {Optimal allocation of parental investment}.
-\newblock {\em Oikos}, 49(3):332--339, July 1987.
-
-\bibitem{Tveraa:2003fq}
-Torkild Tveraa, Per Fauchald, Cathrine Henaug, and Nigel~G Yoccoz.
-\newblock {An examination of a compensatory relationship between food
-  limitation and predation in semi-domestic reindeer}.
-\newblock {\em Oecologia}, 137(3):370--376, November 2003.
-
-\bibitem{Daan:1988va}
-S~Daan, C~Dijkstra, R~Drent, and T~Meijer.
-\newblock {Food supply and the annual timing of avian reproduction}.
-\newblock In H~Ouellet, editor, {\em Acta XIX Congressus Internationalis
-  Ornithologici, Volume I: Proceedings XIX International Ornithological
-  Congress, 1986, Ottawa}, pages 392--407. Proceedings XIX International
-  Ornithological Congress, Ottawa, 1989.
-
-\bibitem{Jacot:2009dv}
-Alain Jacot, Mihai Valcu, Kees van Oers, and Bart Kempenaers.
-\newblock {Experimental nest site limitation affects reproductive strategies
-  and parental investment in a hole-nesting passerine}.
-\newblock {\em Animal Behaviour}, 77(5):1075--1083, May 2009.
-
-\bibitem{Stearns:1989ip}
-S~C Stearns.
-\newblock {Trade-offs in life-history evolution}.
-\newblock {\em Funct. Ecol.}, 3(3):259, 1989.
-
-\bibitem{Barboza:2002in}
-P~Barboza and D~Jorde.
-\newblock {Intermittent fasting during winter and spring affects body
-  composition and reproduction of a migratory duck}.
-\newblock {\em J Comp Physiol B}, 172(5):419--434, July 2002.
-
-\bibitem{Threlkeld:1976ih}
-Stephen~T Threlkeld.
-\newblock {Starvation and the size structure of zooplankton communities}.
-\newblock {\em Freshwater Biol.}, 6(6):489--496, December 1976.
-
-\bibitem{Weber:1998jg}
-Thomas~P Weber, Bruno~J Ens, and Alasdair~I Houston.
-\newblock {Optimal avian migration: A dynamic model of fuel stores and site
-  use}.
-\newblock {\em Evolutionary Ecology}, 12(4):377--401, May 1998.
-
-\bibitem{Mduma:1999cp}
-Simon A~R Mduma, A~R~E Sinclair, and Ray Hilborn.
-\newblock {Food regulates the Serengeti wildebeest: a 40-year record}.
-\newblock {\em J. Anim. Ecol.}, 68(6):1101--1122, November 1999.
-
-\bibitem{Moore:2014hi}
-Jonathan~W Moore, Justin~D Yeakel, Dean Peard, Jeff Lough, and Mark Beere.
-\newblock {Life-history diversity and its importance to population stability
-  and persistence of a migratory fish: steelhead in two large North American
-  watersheds}.
-\newblock {\em J. Anim. Ecol.}, 83(5):1035--1046, 2014.
-
-\bibitem{Mead:1989dt}
-Rodney~A Mead.
-\newblock {The Physiology and Evolution of Delayed Implantation in Carnivores}.
-\newblock In John~L Gittleman, editor, {\em Carnivore Behavior, Ecology, and
-  Evolution}, pages 437--464. Springer US, Ithaca, 1989.
-
-\bibitem{Sandell:1990kw}
-Mikael Sandell.
-\newblock {The evolution of seasonal delayed implantation}.
-\newblock {\em Q Rev Biol}, 65(1):23--42, 1990.
-
-\bibitem{Bulik:1999eo}
-Cynthia~M Bulik, Patrick~F Sullivan, Jennifer~L Fear, Alison Pickering, Aria
-  Dawn, and Mandy McCullin.
-\newblock {Fertility and reproduction in women with anorexia nervosa}.
-\newblock {\em J. Clin. Psychiat.}, 60(2):130--135, February 1999.
-
-\bibitem{Trites:2003cc}
-A~W Trites and C~P Donnelly.
-\newblock {The decline of Steller sea lions Eumetopias jubatus in Alaska: a
-  review of the nutritional stress hypothesis}.
-\newblock {\em Mammal Rev.}, 33(1):3--28, March 2003.
-
-\bibitem{Glazier:2009hq}
-Douglas~S Glazier.
-\newblock {Metabolic level and size scaling of rates of respiration and growth
-  in unicellular organisms}.
-\newblock {\em Funct. Ecol.}, 23(5):963--968, October 2009.
-
-\bibitem{Kooi2000}
-S~A L~M Kooijman.
-\newblock {\em {Dynamic Energy and Mass Budgets in Biological Systems}}.
-\newblock Cambridge, 2000.
-
-\bibitem{Sousa:2010ez}
-T~Sousa, T~Domingos, J~C Poggiale, and S~A L~M Kooijman.
-\newblock {Dynamic energy budget theory restores coherence in biology}.
-\newblock {\em Philos. T. Roy. Soc. B}, 365(1557):3413--3428, October 2010.
-
-\bibitem{Diekmann:2010da}
-O~Diekmann and J~A~J Metz.
-\newblock {How to lift a model for individual behaviour to the population
-  level?}
-\newblock {\em Philos. T. Roy. Soc. B}, 365(1557):3523--3530, November 2010.
-
-\bibitem{murdoch:2003}
-W~W Murdoch, C~J Briggs, and R~M Nisbet.
-\newblock {\em {Consumer-resource Dynamics}}, volume~36 of {\em Monographs in
-  population biology}.
-\newblock Princeton University Press, Princeton, 2003.
-
-\bibitem{Benichou:2014wu}
-Olivier Benichou and S~Redner.
-\newblock {Depletion-Controlled Starvation of a Diffusing Forager}.
-\newblock {\em arXiv}, pages 1--5, May 2014.
-
-\bibitem{Benichou:2016wl}
-O~B{\'e}nichou, M~Chupeau, and S~Redner.
-\newblock {Role of depletion on the dynamics of a diffusing forager}.
-\newblock {\em Journal of Physics A: {\ldots}}, 2016.
-
-\bibitem{Chupeau:2016jf}
-M~Chupeau, O~B{\'e}nichou, and S~Redner.
-\newblock {Universality classes of foraging with resource renewal}.
-\newblock {\em Phys. Rev. E}, 93(3):032403, March 2016.
-
-\bibitem{Persson:1998hz}
-Lennart Persson, Kjell Leonardsson, Andr{\'e}~M De~Roos, Mats Gyllenberg, and
-  Bent Christensen.
-\newblock {Ontogenetic scaling of foraging rates and the dynamics of a
-  size-structured consumer-resource model}.
-\newblock {\em Theor Popul Biol}, 54(3):270--293, December 1998.
-
-\bibitem{murray2011mathematical}
-J~D Murray.
-\newblock {\em {Mathematical Biology: I. An Introduction}}, volume 110 of {\em
-  Interdisciplinary Applied Mathematics}.
-\newblock Springer New York, 2011.
-
-\bibitem{Strogatz:2008wo}
-S~H Strogatz.
-\newblock {\em {Nonlinear Dynamics and Chaos: With Applications to Physics,
-  Biology, Chemistry, and Engineering}}.
-\newblock Studies in nonlinearity. Westview Press, Boulder, 2008.
-
-\bibitem{GuckHolmes}
-J~Guckenheimer and P~Holmes.
-\newblock {\em {Nonlinear Oscillations, Dynamical Systems, and Bifurcations of
-  Vector Fields}}.
-\newblock Springer, New York, 1983.
-
-\bibitem{Gross:2004p2428}
-T~Gross and U~Feudel.
-\newblock {Analytical search for bifurcation surfaces in parameter space}.
-\newblock {\em Physica D}, 195(3-4):292--302, 2004.
-
-\bibitem{Hastings:2001jh}
-Alan Hastings.
-\newblock {Transient dynamics and persistence of ecological systems}.
-\newblock {\em Ecol. Lett.}, 4(3):215--220, May 2001.
-
-\bibitem{Neubert:1997wk}
-MG~Neubert and H~Caswell.
-\newblock {Alternatives to resilience for measuring the responses of ecological
-  systems to perturbations}.
-\newblock {\em Ecology}, 78(3):653--665, 1997.
-
-\bibitem{Caswell:2005eo}
-Hal Caswell and Michael~G Neubert.
-\newblock {Reactivity and transient dynamics of discrete-time ecological
-  systems}.
-\newblock {\em J Differ Equ Appl}, 11(4-5):295--310, April 2005.
-
-\bibitem{Neubert:2009td}
-MG~Neubert and H~Caswell.
-\newblock {Detecting reactivity}.
-\newblock {\em Ecology}, 90(10):2683--2688, 2009.
-
-\bibitem{Yodzis:1992hg}
-P~Yodzis and S~Innes.
-\newblock {Body size and consumer-resource dynamics}.
-\newblock {\em Am. Nat.}, 139(6):1151--1175, 1992.
-
-\bibitem{Brown:2004wq}
-JH~Brown, JF~Gillooly, AP~Allen, VM~Savage, and GB~West.
-\newblock {Toward a metabolic theory of ecology}.
-\newblock {\em Ecology}, 85(7):1771--1789, 2004.
-
-\bibitem{West:2002it}
-Geoffrey~B West, William~H Woodruff, and James~H Brown.
-\newblock {Allometric scaling of metabolic rate from molecules and mitochondria
-  to cells and mammals.}
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 99 Suppl 1(suppl 1):2473--2478,
-  February 2002.
-
-\bibitem{DeLong:2010dy}
-J~P DeLong, J~G Okie, M~E Moses, R~M Sibly, and J~H Brown.
-\newblock {Shifts in metabolic scaling, production, and efficiency across major
-  evolutionary transitions of life}.
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 107(29):12941--12945, July 2010.
-
-\bibitem{West:2001bv}
-Geoffrey~B West, James~H Brown, and Brian~J Enquist.
-\newblock {A general model for ontogenetic growth}.
-\newblock {\em Nature}, 413(6856):628--631, October 2001.
-
-\bibitem{moses2008rmo}
-Melanie~E Moses, Chen Hou, William~H Woodruff, Geoffrey~B West, Jeffery~C
-  Nekola, Wenyun Zuo, and James~H Brown.
-\newblock {Revisiting a Model of Ontogenetic Growth: Estimating Model
-  Parameters from Theory and Data.}
-\newblock {\em Am. Nat.}, 171(5):632--645, 2008.
-
-\bibitem{gillooly2002esa}
-James~F Gillooly, Eric~L Charnov, Geoffrey~B West, Van~M Savage, and James~H
-  Brown.
-\newblock {Effects of size and temperature on developmental time}.
-\newblock {\em Nature}, 417(6884):70--73, May 2002.
-
-\bibitem{hou}
-Chen Hou, Wenyun Zuo, Melanie~E Moses, William~H Woodruff, James~H Brown, and
-  Geoffrey~B West.
-\newblock {Energy uptake and allocation during ontogeny}.
-\newblock {\em Science}, 322(5902):736--739, October 2008.
-
-\bibitem{bettencourt}
-L~M~A Bettencourt, J~Lobo, D~Helbing, C~Kuhnert, and G~B West.
-\newblock {Growth, innovation, scaling, and the pace of life in cities}.
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 104(17):7301--7306, April 2007.
-
-\bibitem{Folland:2008ij}
-J~P Folland, T~M Mc~Cauley, and A~G Williams.
-\newblock {Allometric scaling of strength measurements to body size}.
-\newblock {\em Eur J Appl Physiol}, 102(6):739--745, January 2008.
-
-\bibitem{CalderIII:1983jd}
-W~A Calder~III.
-\newblock {An allometric approach to population cycles of mammals}.
-\newblock {\em J. Theor. Biol.}, 100(2):275--282, January 1983.
-
-\bibitem{Peterson:1984hj}
-R~O Peterson, R~E Page, and K~M Dodge.
-\newblock {Wolves, moose, and the allometry of population cycles}.
-\newblock {\em Science}, 224(4655):1350--1352, June 1984.
-
-\bibitem{Krukonis:1991fk}
-G~Krukonis and W~M Schaffer.
-\newblock {Population cycles in mammals and birds: Does periodicity scale with
-  body size?}
-\newblock {\em J. Theor. Biol.}, 148(4):469--493, February 1991.
-
-\bibitem{Hendriks:2012fc}
-A~Jan Hendriks and Christian Mulder.
-\newblock {Delayed logistic and Rosenzweig{\textendash}MacArthur models with
-  allometric parameter setting estimate population cycles at lower trophic
-  levels well}.
-\newblock {\em Ecol Complex}, 9:43--54, February 2012.
-
-\bibitem{Kendall:1999iy}
-Bruce~E Kendall, Cheryl~J Briggs, William~W Murdoch, Peter Turchin, Stephen~P
-  Ellner, Edward McCauley, Roger~M Nisbet, and Simon~N Wood.
-\newblock {Why do populations cycle? A synthesis of statistical and mechanistic
-  modeling approaches}.
-\newblock {\em Ecology}, 80(6):1789--1805, September 1999.
-
-\bibitem{Hogstedt:2005cr}
-G{\"o}ran H{\"o}gstedt, Tarald Seldal, and Arild Breistol.
-\newblock {Period length in cyclic animal populations}.
-\newblock {\em Ecology}, 86(2):373--378, February 2005.
-
-\bibitem{Kendall:1998hl}
-Kendall, Prendergast, and Bjornstad.
-\newblock {The macroecology of population dynamics: taxonomic and biogeographic
-  patterns in population cycles}.
-\newblock {\em Ecol. Lett.}, 1(3):160--164, November 1998.
-
-\bibitem{Brown:1993p708}
-JH~Brown, PA~Marquet, and ML~Taper.
-\newblock {Evolution of body size: consequences of an energetic definition of
-  fitness}.
-\newblock {\em Am. Nat.}, 142(4):573--584, 1993.
-
-\bibitem{West:1997cg}
-Geoffrey~B West, James~H Brown, and Brian~J Enquist.
-\newblock {A general model for the origin of allometric scaling laws in
-  biology}.
-\newblock {\em Science}, 276(5309):122--126, April 1997.
-
-\bibitem{West:2002ud}
-Geoffrey~B West, William~H Woodruff, and James~H Brown.
-\newblock {Allometric scaling of metabolic rate from molecules and mitochondria
-  to cells and mammals.}
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 99 Suppl 1(Supplement
-  1):2473--2478, February 2002.
-
-\bibitem{Millar:1990p923}
-JS~Millar and G~Hickling.
-\newblock {Fasting endurance and the evolution of mammalian body size}.
-\newblock {\em Funct. Ecol.}, 4(1):5--12, 1990.
-
-\bibitem{Alroy:1998p1594}
-J~Alroy.
-\newblock {Cope's rule and the dynamics of body mass evolution in North
-  American fossil mammals}.
-\newblock {\em Science}, 280(5364):731--734, 1998.
-
-\bibitem{Clauset:2009fh}
-A~Clauset and S~Redner.
-\newblock {Evolutionary model of species body mass diversification}.
-\newblock {\em Phys. Rev. Lett.}, 102(3):038103, January 2009.
-
-\bibitem{Smith:2010p3442}
-Felisa~A Smith, Alison~G Boyer, James~H Brown, Daniel~P Costa, Tamar Dayan,
-  S~K~Morgan Ernest, Alistair~R Evans, Mikael Fortelius, John~L Gittleman,
-  Marcus~J Hamilton, Larisa~E Harding, Kari Lintulaakso, S~Kathleen Lyons,
-  Christy McCain, Jordan~G Okie, Juha~J Saarinen, Richard~M Sibly, Patrick~R
-  Stephens, Jessica Theodor, and Mark~D Uhen.
-\newblock {The evolution of maximum body size of terrestrial mammals}.
-\newblock {\em Science}, 330(6008):1216--1219, November 2010.
-
-\bibitem{Saarinen:2014br}
-Juha~J Saarinen, Alison~G Boyer, James~H Brown, Daniel~P Costa, S~K~Morgan
-  Ernest, Alistair~R Evans, Mikael Fortelius, John~L Gittleman, Marcus~J
-  Hamilton, Larisa~E Harding, Kari Lintulaakso, S~Kathleen Lyons, Jordan~G
-  Okie, Richard~M Sibly, Patrick~R Stephens, Jessica Theodor, Mark~D Uhen, and
-  Felisa~A Smith.
-\newblock {Patterns of maximum body size evolution in Cenozoic land mammals:
-  Eco-evolutionary processes and abiotic forcing}.
-\newblock {\em Proc Biol Sci}, 281(1784):20132049, June 2014.
-
-\bibitem{Damuth:1987kr}
-John Damuth.
-\newblock {Interspecific allometry of population density in mammals and other
-  animals: the independence of body mass and population energy-use}.
-\newblock {\em Biol. J. Linn. Soc.}, 31(3):193--246, July 1987.
-
-\bibitem{allen2002}
-Andrew~P Allen, James~H Brown, and James~F Gillooly.
-\newblock {Global biodiversity, biochemical kinetics, and the
-  energetic-equivalence rule}.
-\newblock {\em Science}, 297(5586):1545--1548, August 2002.
-
-\bibitem{enquist1998}
-Brian~J Enquist, James~H Brown, and Geoffrey~B West.
-\newblock {Allometric scaling of plant energetics and population density}.
-\newblock {\em Nature}, 395(6698):163--165, September 1998.
-
-\bibitem{Kempes:2016}
-Christopher~P Kempes, Lawrence Wang, Jan~P Amend, John Doyle, and Tori Hoehler.
-\newblock {Evolutionary tradeoffs in cellular composition across diverse
-  bacteria}.
-\newblock {\em ISME J}, 10(9):2145--2157, April 2016.
-
-\bibitem{tilman1981}
-David Tilman.
-\newblock {Tests of resource competition theory using four species of lake
-  michigan algae}.
-\newblock {\em Ecology}, 62(3):802--815, June 1981.
-
-\bibitem{dutkiewicz2009}
-S~Dutkiewicz, M~J Follows, and J~G Bragg.
-\newblock {Modeling the coupling of ocean ecology and biogeochemistry}.
-\newblock {\em Global Biogeochem. Cycles}, 23(4):1--15, December 2009.
-
-\bibitem{barton2010}
-Andrew~D Barton, Stephanie Dutkiewicz, Glenn Flierl, Jason Bragg, and Michael~J
-  Follows.
-\newblock {Patterns of diversity in marine phytoplankton}.
-\newblock {\em Science}, 327(5972):1509--1511, March 2010.
-
-\end{thebibliography}
-
-\end{bibunit}
-
-% \bibliography{aa_starving3}
-
-%
-% \begin{acknowledgments}
-%   We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
-%   discussions and comments on the manuscript.  J.D.Y. was supported by
-%   startup funds at the University of California, Merced, and an Omidyar
-%   Postdoctoral Fellowship at the Santa Fe Institute.  C.P.K. was supported by
-%   an Omidyar Postdoctoral Fellowship at the Santa Fe Institute.  S.R. was
-%   supported by grants DMR-1608211 and 1623243 from the National Science
-%   Foundation, and by the John Templeton Foundation, all at the Santa Fe
-%   Institute.
-% \end{acknowledgments}
-%
-\clearpage
-
-\begin{bibunit}[unsrt]
-\section*{Supporting Information for ``The dynamics of starvation and recovery''}
-
-\subsection*{Mechanisms of Starvation and Recovery}
+\section*{Mechanisms of Starvation and Recovery}
 Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
@@
 In the above equations $Y$ represents the yield coefficient (e.g., Refs.
-\cite{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
+\citenum{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
 If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
@@
 
-% \putbib[aa_starving_supplement]
-
+% \bibliography{aa_starving_supplement}
 \begin{thebibliography}{10}
@@
 \bibitem{Kempes:2012hy}
-Christopher~P Kempes, Stephanie Dutkiewicz, and Michael~J Follows.
-\newblock {Growth, metabolic partitioning, and the size of microorganisms.}
-\newblock {\em PNAS}, 109(2):495--500, January 2012.
+Kempes CP, Dutkiewicz S, Follows MJ
+\newblock (2012) {Growth, metabolic partitioning, and the size of
+  microorganisms.}
+\newblock \emph{PNAS} 109:495--500.
 
 \bibitem{kempes2014morphological}
-Christopher~P Kempes, Chinweike Okegbe, Zwoisaint Mears-Clarke, Michael~J
-  Follows, and Lars~EP Dietrich.
-\newblock Morphological optimization for access to dual oxidants in biofilms.
-\newblock {\em Proceedings of the National Academy of Sciences},
-  111(1):208--213, 2014.
+Kempes CP, Okegbe C, Mears-Clarke Z, Follows MJ, Dietrich LE
+\newblock (2014) Morphological optimization for access to dual oxidants in
+  biofilms.
+\newblock \emph{Proceedings of the National Academy of Sciences} 111:208--213.
 
 \bibitem{West:2001bv}
-Geoffrey~B West, James~H Brown, and Brian~J Enquist.
-\newblock {A general model for ontogenetic growth}.
-\newblock {\em Nature}, 413(6856):628--631, October 2001.
+West GB, Brown JH, Enquist BJ
+\newblock (2001) {A general model for ontogenetic growth}.
+\newblock \emph{Nature} 413:628--631.
 
 \bibitem{moses2008rmo}
-Melanie~E Moses, Chen Hou, William~H Woodruff, Geoffrey~B West, Jeffery~C
-  Nekola, Wenyun Zuo, and James~H Brown.
-\newblock {Revisiting a model of ontogenetic growth: Estimating model
+Moses ME, {et~al.}
+\newblock (2008) {Revisiting a model of ontogenetic growth: Estimating model
   parameters from theory and data}.
-\newblock {\em http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735},
-  171(5):632--645, May 2008.
+\newblock \emph{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}
+  171:632--645.
 
 \bibitem{hou}
-Chen Hou, Wenyun Zuo, Melanie~E Moses, William~H Woodruff, James~H Brown, and
-  Geoffrey~B West.
-\newblock {Energy uptake and allocation during ontogeny}.
-\newblock {\em Science}, 322(5902):736--739, October 2008.
+Hou C, {et~al.}
+\newblock (2008) {Energy uptake and allocation during ontogeny}.
+\newblock \emph{Science} 322:736--739.
 
 \bibitem{pirt}
-SJ~Pirt.
-\newblock The maintenance energy of bacteria in growing cultures.
-\newblock {\em Proceedings of the Royal Society of London B: Biological
-  Sciences}, 163(991):224--231, 1965.
+Pirt S
+\newblock (1965) The maintenance energy of bacteria in growing cultures.
+\newblock \emph{Proceedings of the Royal Society of London B: Biological
+  Sciences} 163:224--231.
 
 \bibitem{Heijnen}
-JJ~Heijnen and JA~Roels.
-\newblock A macroscopic model describing yield and maintenance relationships in
-  aerobic fermentation processes.
-\newblock {\em Biotechnology and Bioengineering}, 23(4):739--763, 1981.
+Heijnen J, Roels J
+\newblock (1981) A macroscopic model describing yield and maintenance
+  relationships in aerobic fermentation processes.
+\newblock \emph{Biotechnology and Bioengineering} 23:739--763.
 
 \bibitem{peters1986ecological}
-Robert~Henry Peters.
-\newblock {\em The Ecological Implications of Body Size}, volume~2.
-\newblock Cambridge University Press, Cambridge, 1986.
+Peters RH
+\newblock (1986) \emph{The Ecological Implications of Body Size}
+\newblock (Cambridge University Press, Cambridge) Vol.{}~2.
 
 \bibitem{blueweiss1978relationships}
-L~Blueweiss, H~Fox, V~Kudzma, D~Nakashima, R~Peters, and S~Sams.
-\newblock Relationships between body size and some life history parameters.
-\newblock {\em Oecologia}, 37(2):257--272, 1978.
+Blueweiss L, {et~al.}
+\newblock (1978) Relationships between body size and some life history
+  parameters.
+\newblock \emph{Oecologia} 37:257--272.
 
 \bibitem{stryer}
-Lubert Stryer.
-\newblock {\em {Biochemistry, Fourth Edition}}.
-\newblock W.H. Freeman and Company, New York, 1995.
+Stryer L
+\newblock (1995) \emph{{Biochemistry, Fourth Edition}}
+\newblock (W.H. Freeman and Company, New York), pp 608--611.
 
 \bibitem{Dunbrack:1993ec}
-Robert~L Dunbrack and Malcolm~A Ramsay.
-\newblock {The Allometry of Mammalian Adaptations to Seasonal Environments: A
-  Critique of the Fasting Endurance Hypothesis}.
-\newblock {\em Oikos}, 66(2):336--342, March 1993.
+Dunbrack RL, Ramsay MA
+\newblock (1993) {The Allometry of Mammalian Adaptations to Seasonal
+  Environments: A Critique of the Fasting Endurance Hypothesis}.
+\newblock \emph{Oikos} 66:336--342.
 
 \bibitem{Lindstedt:1985hm}
-Stan~L Lindstedt and Mark~S Boyce.
-\newblock {Seasonality, Fasting Endurance, and Body Size in Mammals}.
-\newblock {\em Am. Nat.}, 125(6):873--878, June 1985.
+Lindstedt SL, Boyce MS
+\newblock (1985) {Seasonality, Fasting Endurance, and Body Size in Mammals}.
+\newblock \emph{Am. Nat.} 125:873--878.
 
 \bibitem{Lindstedt:2002td}
-S~L Lindstedt and P~J Schaeffer.
-\newblock {Use of allometry in predicting anatomical and physiological
+Lindstedt SL, Schaeffer PJ
+\newblock (2002) {Use of allometry in predicting anatomical and physiological
   parameters of mammals.}
-\newblock {\em Lab. Anim.}, 36(1):1--19, January 2002.
+\newblock \emph{Lab. Anim.} 36:1--19.
 
 \bibitem{estermann}
-Beda~L Estermann, Hans-Rudolf Wettstein, Franz Sutter, and Michael Kreuzer.
-\newblock Nutrient and energy conversion of grass-fed dairy and suckler beef
-  cattle kept indoors and on high altitude pasture.
-\newblock {\em Animal Research}, 50(6):477--493, 2001.
+Estermann BL, Wettstein HR, Sutter F, Kreuzer M
+\newblock (2001) Nutrient and energy conversion of grass-fed dairy and suckler
+  beef cattle kept indoors and on high altitude pasture.
+\newblock \emph{Animal Research} 50:477--493.
 
 \bibitem{michaletz2014convergence}
-Sean~T Michaletz, Dongliang Cheng, Andrew~J Kerkhoff, and Brian~J Enquist.
-\newblock Convergence of terrestrial plant production across global climate
-  gradients.
-\newblock {\em Nature}, 512(7512):39--43, 2014.
+Michaletz ST, Cheng D, Kerkhoff AJ, Enquist BJ
+\newblock (2014) Convergence of terrestrial plant production across global
+  climate gradients.
+\newblock \emph{Nature} 512:39--43.
 
 \bibitem{damuth1987interspecific}
-John Damuth.
-\newblock Interspecific allometry of population density in mammals and other
-  animals: the independence of body mass and population energy-use.
-\newblock {\em Biological Journal of the Linnean Society}, 31(3):193--246,
-  1987.
+Damuth J
+\newblock (1987) Interspecific allometry of population density in mammals and
+  other animals: the independence of body mass and population energy-use.
+\newblock \emph{Biological Journal of the Linnean Society} 31:193--246.
 
@@
 
-\end{bibunit}
+%\let\cleardoublepage\clearpage
+
+%FIGURE 1
+
+\end{article}
 

### 2017-06-09 Manuscript/Nature-Submission/Revised-Supplement-3.tex
Nature submission

--- 2017-06-09 Manuscript/Nature-Submission/Revised-Supplement-2.tex
@@
 \begin{split}
-\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
-\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
-&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
-\end{split}
-\end{align}
-where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
-%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
-In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
-Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
-The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
-A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
-
-In the above equations $Y$ represents the yield coefficient (e.g., Refs.
-\citenum{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
-If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
-This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
-In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment.
-
-Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
-As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
-Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
-\begin{align}
-\begin{split}
 \dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\

### 2017-06-10 Manuscript/Nature-Submission/Nature-temp.tex
nature comm

@@
 \begin{abstract}
-The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of its constituent individuals.
-Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery in resource-limited environments.
+The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly influences how they invest reserves in uncertain environments.
+Such behaviors are generally manifested as a tradeoff between investing in somatic maintenance and growth, or allocating energy towards reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
+% Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery in resource-limited environments. 
 To elucidate the consequences of this tradeoff, we introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete consumers that reproduce, and undernourished, non-reproducing consumers that are susceptible to mortality.
-As a function of the transition rates between these two states that are determined by the abundance of resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
+% As a function of the transition rates between these two states that are determined by the abundance of resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
 We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships between body size and a variety of traits and find that population dynamics subject to these constraints are typically driven to a steady state.
 Moreover, we find that these rates fall within a `refuge' in parameter space, where the probability of extinction of the consumer population is minimized. 
-Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations.
+% Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations. 
 Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percentage of body fat.
-For body masses $< 1.748\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.748\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
+% For body masses $< 1.748\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.748\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
 \end{abstract}
 
-%\noindent {\small {\bf Significance Statement} %105/120 words
+% \noindent {\small {\bf Summary} %105/120 words
 %Energetic investment in somatic maintenance and growth vs.
@@
 
-\noindent {\bf Introduction} \\
-The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly influences how they invest reserves in uncertain environments.
-Such behaviors are generally manifested as tradeoffs between investing in somatic maintenance and growth, or allocating energy towards reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
-The timing of these behaviors responds to selective pressure, as the choice of the investment impacts future fitness~\cite{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.
-The influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
+% \noindent {\bf Introduction} \\
+
+% The timing of these behaviors responds to selective pressure, as the choice of the investment impacts future fitness~\cite{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.
+% The influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
 

### 2017-06-10 Manuscript/NatureComm/YKR_NatureComm.tex
nature comm

--- 2017-03-20 Manuscript/Arxiv_Revision/YKR_Arxiv.tex
@@
 
-%\bibliographystyle{ieeetr}
+\bibliographystyle{naturemag}
 
@@
 \author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
-  of California, Merced, Merced, CA 95340, USA}
+  of California, Merced, Merced, CA 95340, USA}\affiliation{The Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author: jdyeakel@gmail.com}
 
 \author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
-  Park Road, Santa Fe, NM 87501, USA}
+  Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}
 
 \author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
-  Road, Santa Fe, NM 87501, USA}
+  Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}
 
@@
 The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of its constituent individuals. 
-Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery in resource-limited environments.
-To elucidate the consequences of this tradeoff, we introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete consumers that reproduce, and undernourished, non-reproducing consumers that are susceptible to mortality.
-As a function of the transition rates between these two states that are determined by the abundance of resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
-We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships between body size and a variety of traits and find that population dynamics subject to these constraints are typically driven to a steady state.
+Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery. % in resource-limited environments
+We introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete, reproducing consumers, and undernourished, non-reproducing consumers.
+% that are susceptible to mortality
+% As a function of the transition rates between these two states that are determined by the abundance of resources, the consumer populations can either undergo cyclic dynamics or reach a steady state.
+To elucidate the consequences of this tradeoff, w
+We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships and find that population dynamics are typically driven to a steady state.
+%subject to these constraints between body size and a variety of traits
 Moreover, we find that these rates fall within a `refuge' in parameter space, where the probability of extinction of the consumer population is minimized. 
-Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations.
-Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percentage of body fat.
-For body masses $< 1.748\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.748\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
+% Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations. 
+Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percent body fat, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
+% For body masses $< 1.748\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.748\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
 \end{abstract}
@@
 
-\section*{Introduction}
 The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly influences how they invest reserves in uncertain environments.
@@
 
-\section*{Nutritional state-structured model (NSM)}
+The balance between (a) somatic growth and maintenance, and (b) reproduction depends on resource availability~\cite{Morris:1987eo}.
+For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\cite{Tveraa:2003fq}.
+Many bird species invest differently in broods during periods of resource scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
+Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\cite{Kirk:1997cc}.
+Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate nutritional reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+
+Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.
+The data collected thus far has shown that diverse mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development (blastocyst implantation) until nutritional reserves can be accumulated~\cite{Mead:1989dt,Sandell:1990kw}.
+Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
+In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.
+The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the importance and universality of the fundamental tradeoff between somatic and reproductive investment.
+However the general dynamic implications of these constraints are unknown.
+
+Though straightforward conceptually, incorporating the energetic dynamics of individuals~\cite{Kooi2000} into a population-level framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles~\cite{Diekmann:2010da}.
+An alternative approach involves modeling the macroscale relations that guide somatic versus reproductive investment in a consumer-resource system.
+For example, macroscale Lotka-Volterra models assume that the growth rate of the consumer population depends on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction~\cite{murdoch:2003}.
+
+In this work, we adopt an alternative approach in which we \emph{explicitly} account for resource limitation and the subsequent effects of starvation.
+Namely, only individuals with sufficient energetic reserves can reproduce.
+Such a constraint leads to reproductive time lags due to some members of the population going hungry and then recovering.
+Additionally, we incorporate the idea that reproduction is strongly constrained allometrically \cite{Kempes:2012hy}, and is not generally linearly related to resource density.
+As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.
+\\
+
+\noindent {\bf Nutritional state-structured model (NSM)}\\
 We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$.
@@
 In the supplementary information (SI) we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we describe below.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{fig_FixedPoint2.eps}
+\caption{\small{ The transcritical (dashed) and Hopf bifurcation (solid) as a
+  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
+  bifurcation conditions separate parameter space into unphysical, cyclic,
+  and steady state dynamic regimes.  The colors show the steady state densities for the energetically replete consumers $F^*$, (warmer colors denote higher densities).  
+  % Steady state densities for the energetically deficient consumers $H^*$ (not shown)
+  % scale with those for $F^*$.  
+}\label{fig:fp}}
+\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_FixedPoint.png}
-\caption{ The transcritical (dashed) and Hopf bifurcation (solid) as a
-  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-  bifurcation conditions separate parameter space into unphysical, cyclic,
-  and steady state dynamic regimes.  The colors show the steady
-  state densities for (\emph{A}) the resource $R^*$ and the (\emph{B}) energetically
-  replete consumers $F^*$, (warmer colors denote higher densities).  Steady
-  state densities for the energetically deficient consumers $H^*$ (not shown)
-  scale with those for $F^*$.  }
-\label{fig:fp}
-\end{figure}
+In the mean-field approximation, in which the consumers and resources are perfectly mixed, their densities evolve according to the rate equations
 
-In the mean-field approximation, in which the consumers and resources are perfectly mixed, their densities evolve according to the rate equations
 \begin{align}
@@
 \begin{split}
-\frac{dF}{dt} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
-\frac{dH}{dt} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-\frac{dR}{dt} &= \alpha \left(1-R\right)R -\left(\rho R+\delta\right)H-\beta F
+\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
+\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
+\dot{R} &= \alpha \left(1-R\right)R -\left(\rho R+\delta\right)H-\beta F
 \end{split}
 \end{align}
+
 This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see SI for a full derivation and the dimensional form).
-Notice that the total consumer density $F+H$ evolves according to $\frac{dF}{dt}+\frac{dH}{dt}=\lambda F-\mu H$.
+Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
 This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\cite{murray2011mathematical}, except that the resource density does not appear in the growth term.
@@
 From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that the consumer and resource densities are positive.
+%In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
 The condition $\sigma = \lambda$ thus represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates a physical from an unphysical regime where all steady-state densities become negative after intersecting the trivial fixed point $(F^*,H^*,R^*)=(0,0,0)$.
 The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
-As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges.
+As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges. %(Fig.~\ref{fig:gvs})
 
@@
 Given that ecological systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase the extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
+%Thus the distance of a system from the Hopf bifurcation provides a measure of its persistence.
 
@@
 This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.
+\\
+% Whereas the relation between consumer growth rate $\lambda$ and the starvation rate $\sigma$ defines an absolute bound of biological feasibility---the TC bifurcation---$\sigma$ also determines the sensitivity of the consumer population to changes in resource density.
+% When $\sigma\gg\lambda$, the steady-state population density is small, thereby increasing the risk of stochastic extinction.
+% On the other hand, as $\sigma$ decreases, the system will ultimately be poised either near the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).
+% If the recovery rate $\rho$ is sufficiently small, the TC bifurcation is reached and the resource eventually is eliminated.
+% If $\rho$ exceeds a threshold value, cyclic dynamics will develop as the Hopf bifurcation is approached.
 
-\section*{Role of allometry}
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.eps}
+\caption{\small{ The growth trajectory over absolute time of an individual organism as a function of body mass.  
+Initial growth follows the black trajectory to an energetically replete reproductive adult mass $m=\epsilon_\lambda M$. %which we assume is 95\% asymptotic mass $M$.  
+Starvation follows the red trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. 
+Recovery follows the green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires a longer time to reach $\epsilon_\lambda M$.
+%different energetics and
+Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}\label{fig:growth}}
+\end{figure}
+
+\noindent {\bf Role of allometry} \\
+%[Link Allometry stuff to our model - what does it provide to the story]
+%[Introduce biological and linked constraints]
+%[Allometries capture vast amounts of diversity via a single parameter --- body size]
+%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
 While there are no a priori constraints on the parameters in the NSM, most
@@
 \cite{DeLong:2010dy,Kempes:2012hy}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.pdf}
-\caption{ The growth trajectory over absolute time of an individual organism
-  as a function of body mass.  Initial growth follows the black trajectory to
-  an energetically replete reproductive adult mass $m=\epsilon_\lambda M$ which we assume is 95\% asymptotic mass $M$.  Starvation follows the red
-  trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. Recovery follows the
-  green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires different energetics and a longer time to reach $\epsilon_\lambda M$.
-Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}
-\label{fig:growth}
-\end{figure}
+%Justin adding an explanation for beta
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.45\textwidth]{fig_Rates.pdf}
-\caption{
-Allometrically constrained starvation rate $\sigma=1/t_{\sigma}$ (red), mortality $\mu=1/t_{\mu}$ (blue) and recovery rate $\rho=1/t_{\rho}$ (green) relative to the reproductive rate $\lambda=1/t_{\lambda}$ (black) as a function of body mass (see Equations \ref{t1}, \ref{rhotimescale}, \ref{eq:sigma}, and \ref{mutimescale}).
-The rate of starvation is greater than the rate of reproduction for all realized terrestrial endotherm body sizes.
-Mean values $\pm 25\%$ variation are shown by the shaded region for each rate.
-}
-\label{fig:gvs}
-\end{figure}
-
 The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometricComb3.eps}
+\caption{\small{ Probability of extinction for a consumer with (\emph{A}) $M=10^2$g, (\emph{B}) $M=10^4$g, and (\emph{C}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
+Note the change in scale for $M=10^4$ and $M=10^6$g.
+Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
+The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
+\end{figure*}
+
 To determine the starvation rate, $\sigma$, we are interested in the time
@@
 Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
-Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
+Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.\\
 
-\section*{Stabilizing effects of allometric constraints}
-As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
-Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values (Fig.~\ref{fig:hopf}) for the known range of body sizes $M$.
-We thus find that the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
-Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
-These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
+% \noindent {\bf  Stabilizing effects of allometric constraints} \\
+% As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
+% Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
+% Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
+%Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
+% These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
 However, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation.
-This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
+% This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.\\
 
-\begin{figure}
-\centering
-\includegraphics[width=0.425\textwidth]{fig_DataHopf.pdf}
-\caption{ Transcritical (vertical lines) and Hopf bifurcations (curves) for
-  allometrically determined starvation $\sigma$ and recovery $\rho$ rates as
-  a function of different mammalian body sizes: $M=P\times10^2$g (blue) and
-  $M=P\times10^6$g (green), where $P$ is a random uniform variable in $[1,9]$.
-  Points denote realized values of $\sigma$ and $\rho$ given the drawn values for $M$.
-  }
-\label{fig:hopf}
-\end{figure}
-
-Previous studies used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
-However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
-Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
-However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.
-
-\begin{figure*}
-\centering
-\includegraphics[width=1\textwidth]{fig_ExtinctionAllometricComb3.png}
-\caption{ Probability of extinction for a consumer with (\emph{A}) $M=10^2$g, (\emph{B}) $M=10^4$g, and (\emph{C}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
-Note the change in scale for $M=10^4$ and $M=10^6$g.
-Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
-The white points denote the allometrically constrained starvation and recovery rate.}
-\label{fig:ext}
-\end{figure*}
+% Previous studies used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
+% However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
+% Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
+% However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.\\
 
 %Extinction to the left and the right
-\section*{Extinction risk}
-Within our model, higher rates of starvation result in a larger flux of the population to the hungry state.
+\noindent {\bf Extinction risk}\\
+As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
+Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
+Indeed, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation. 
+This suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
+% Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
+
+Higher rates of starvation result in a larger flux of the population to the hungry state.
 In this state, reproduction is absent, thus increasing the likelihood of extinction.
 From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
-We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
+We therefore examine the competing effects of cyclic dynamics vs. changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
 To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
@@
 % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-{\it To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.}\\
+To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
 
@@
 
-\section*{Dynamic and energetic barriers to body size}
+\noindent {\bf Dynamic and energetic barriers to body size} \\
 Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\cite{West:2002ud}.
@@
 \centering
-\includegraphics[width=0.4\textwidth]{fig_FPAllometric.pdf}
-\caption{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-  body mass.
+\includegraphics[width=0.48\textwidth]{fig_FPAllometric2.eps}
+\caption{\small{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
+  body mass. Inset: Resource steady state $R^*$ as a function of consumer body mass.
   (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  (\emph{C}) Resource steady state $R^*$ as a function of consumer body mass.
   The data are from Damuth \cite{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
-\cite{West:2001bv,hou,moses2008rmo}).}
-\label{fig:mass}
+\cite{West:2001bv,hou,moses2008rmo}).}\label{fig:mass}}
 \end{figure}
 
+%Minimal mass bound
+%A theoretical lower bound on mammalian body size occurs when $\epsilon=0$ where mammals have no metabolic reserves and immediately starve; this occurs at a size of $M=0$, which is obviously infeasible.
+%Such prediction gives an extreme limit on size but does not account for the subtleties of starvation dynamics
+%that may limit body size (among other factors).
 The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
+Similar predictions have been made for carnivore populations using alternative consumer-resource models \cite{DeLong:2012kw}.
 Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{Damuth:1987kr,allen2002,enquist1998}.
 The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size~\cite{Damuth:1987kr,allen2002,enquist1998}.
+%Previous studies have focused on t
 This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism, such that $N^*\propto M^{-3/4}/B_{0}$~\cite{allen2002,enquist1998}.
@@
 This result is remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to energy equivalence.
+
 Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
+%, violating the energy equivalence.
+% set by deviations from constant efficiency obeyed by other populations.
 In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
-This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero causing the timescales to become infinite (see Equation \ref{mutimescale}) and the rates $\mu$ and $\lambda$ to equal zero.
+This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero, causing the timescales to become infinite and the rates $\mu$ and $\lambda$ to equal zero.
 The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$.
@@
 
+%Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end.
+%Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is close to the minimum observed mammalian body size of ca.\ 1.3-2.5g, a range that occurs as the recovery rate begins its decline.
+%In addition to known transport limitations~\cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.
+%This result mirrors other efforts \cite{Kempes:2012hy,Kempes:2016} where at a given scale multiple limitations constrain the smallest possibilities for life within a class of organisms.
+
+%The combined steady-state dynamics and allometric timescales of the NSM predict that larger mammals should outcompete smaller ones, and this suggests that the model may provide a framework with which to explore the energetic mechanisms behind Cope's rule.
+%, which we explore further below.
+
 We contend that the NSM provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time.
 The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
-\ref{fig:mass}\emph{C}), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
+\ref{fig:mass}\emph{A}, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
 Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but does not offer a within-lineage mechanism by which larger body sizes are selected for.
 
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=0.45\textwidth]{fig_Invasion.pdf}
-\caption{ Invasion feasibility for organisms with a proportional change in
+\includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
+\caption{\small{ Invasion feasibility for organisms with a proportional change in
   mass $\chi$ against a population with a resident body mass $M$.  The blue
@@
   Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
-  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}
-\label{fig:invasion}
+  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}\label{fig:invasion}}
 \end{figure}
@@
 Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the body composition of starved foragers, is assumed to remain unchanged for the invader population (see SI for detailed derivations of invader rates).
+%There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
 
 To assess the susceptibility of the resident population to invasion, we determine which consumer has a lower steady-state resource density for a given value of $\chi$, again with the expectation that populations able to survive on lower resource densities have a competitive advantage \cite{tilman1981}.
-We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population.
+We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population (Fig.
+\ref{fig:invasion}).
 However, for $M> 1.748\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
+
+%, and this is due to the changing
+%covariance between energetic rates as a function of modified energetic
+%reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
 
@@
 Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \cite{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals outlined by Alroy~\cite{Alroy:1998p1594}.
+%as well as  \cite{Alroy:1998p1594,Clauset:2009fh}.
+%This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
 While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general within-lineage mechanism for the evolution of larger body size among terrestrial mammals.
 
+%and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
+
+%One might be concerned a greater number of large mammals are currently not observed in the modern world given that larger mammals are less susceptible to extinction.
+%However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also much more geographically widespread than today.
+%These results, combined with our findings, suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
+
+%Closing
 The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations~\cite{Stearns:1989ip}.
@@
 
-% \putbib[aa_starving3]
-
-\begin{thebibliography}{10}
-
-\bibitem{Martin:1987dl}
-Thomas~E Martin.
-\newblock {Food as a limit on breeding birds: A life-history perspective}.
-\newblock {\em Annu. Rev. Ecol. Syst.}, 18:453--487, January 1987.
-
-\bibitem{Kirk:1997cc}
-Kevin~L Kirk.
-\newblock {Life-history responses to variable environments: Starvation and
-  reproduction in planktonic rotifers}.
-\newblock {\em Ecology}, 78(2):434--441, March 1997.
-
-\bibitem{Kempes:2012hy}
-C~P Kempes, S~Dutkiewicz, and M~J Follows.
-\newblock {Growth, metabolic partitioning, and the size of microorganisms}.
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 109(2):495--500, January 2012.
-
-\bibitem{Mangel:1988uaa}
-M~Mangel and C.~W. Clark.
-\newblock {\em {Dynamic Modeling in Behavioral Ecology}}.
-\newblock Princeton University Press, Princeton, 1988.
-
-\bibitem{Mangel:2014kz}
-Marc Mangel.
-\newblock {Stochastic dynamic programming illuminates the link between
-  environment, physiology, and evolution}.
-\newblock {\em B. Math. Biol.}, 77(5):857--877, 2014.
-
-\bibitem{Yeakel:2013hi}
-Justin~D Yeakel, Nathaniel~J Dominy, Paul~L Koch, and Marc Mangel.
-\newblock {Functional morphology, stable isotopes, and human evolution: a model
-  of consilience}.
-\newblock {\em Evolution}, 68(1):190--203, 2014.
-
-\bibitem{Morris:1987eo}
-Douglas~W Morris.
-\newblock {Optimal allocation of parental investment}.
-\newblock {\em Oikos}, 49(3):332--339, July 1987.
-
-\bibitem{Tveraa:2003fq}
-Torkild Tveraa, Per Fauchald, Cathrine Henaug, and Nigel~G Yoccoz.
-\newblock {An examination of a compensatory relationship between food
-  limitation and predation in semi-domestic reindeer}.
-\newblock {\em Oecologia}, 137(3):370--376, November 2003.
-
-\bibitem{Daan:1988va}
-S~Daan, C~Dijkstra, R~Drent, and T~Meijer.
-\newblock {Food supply and the annual timing of avian reproduction}.
-\newblock In H~Ouellet, editor, {\em Acta XIX Congressus Internationalis
-  Ornithologici, Volume I: Proceedings XIX International Ornithological
-  Congress, 1986, Ottawa}, pages 392--407. Proceedings XIX International
-  Ornithological Congress, Ottawa, 1989.
-
-\bibitem{Jacot:2009dv}
-Alain Jacot, Mihai Valcu, Kees van Oers, and Bart Kempenaers.
-\newblock {Experimental nest site limitation affects reproductive strategies
-  and parental investment in a hole-nesting passerine}.
-\newblock {\em Animal Behaviour}, 77(5):1075--1083, May 2009.
-
-\bibitem{Stearns:1989ip}
-S~C Stearns.
-\newblock {Trade-offs in life-history evolution}.
-\newblock {\em Funct. Ecol.}, 3(3):259, 1989.
-
-\bibitem{Barboza:2002in}
-P~Barboza and D~Jorde.
-\newblock {Intermittent fasting during winter and spring affects body
-  composition and reproduction of a migratory duck}.
-\newblock {\em J Comp Physiol B}, 172(5):419--434, July 2002.
-
-\bibitem{Threlkeld:1976ih}
-Stephen~T Threlkeld.
-\newblock {Starvation and the size structure of zooplankton communities}.
-\newblock {\em Freshwater Biol.}, 6(6):489--496, December 1976.
-
-\bibitem{Weber:1998jg}
-Thomas~P Weber, Bruno~J Ens, and Alasdair~I Houston.
-\newblock {Optimal avian migration: A dynamic model of fuel stores and site
-  use}.
-\newblock {\em Evolutionary Ecology}, 12(4):377--401, May 1998.
-
-\bibitem{Mduma:1999cp}
-Simon A~R Mduma, A~R~E Sinclair, and Ray Hilborn.
-\newblock {Food regulates the Serengeti wildebeest: a 40-year record}.
-\newblock {\em J. Anim. Ecol.}, 68(6):1101--1122, November 1999.
-
-\bibitem{Moore:2014hi}
-Jonathan~W Moore, Justin~D Yeakel, Dean Peard, Jeff Lough, and Mark Beere.
-\newblock {Life-history diversity and its importance to population stability
-  and persistence of a migratory fish: steelhead in two large North American
-  watersheds}.
-\newblock {\em J. Anim. Ecol.}, 83(5):1035--1046, 2014.
-
-\bibitem{Mead:1989dt}
-Rodney~A Mead.
-\newblock {The Physiology and Evolution of Delayed Implantation in Carnivores}.
-\newblock In John~L Gittleman, editor, {\em Carnivore Behavior, Ecology, and
-  Evolution}, pages 437--464. Springer US, Ithaca, 1989.
-
-\bibitem{Sandell:1990kw}
-Mikael Sandell.
-\newblock {The evolution of seasonal delayed implantation}.
-\newblock {\em Q Rev Biol}, 65(1):23--42, 1990.
-
-\bibitem{Bulik:1999eo}
-Cynthia~M Bulik, Patrick~F Sullivan, Jennifer~L Fear, Alison Pickering, Aria
-  Dawn, and Mandy McCullin.
-\newblock {Fertility and reproduction in women with anorexia nervosa}.
-\newblock {\em J. Clin. Psychiat.}, 60(2):130--135, February 1999.
-
-\bibitem{Trites:2003cc}
-A~W Trites and C~P Donnelly.
-\newblock {The decline of Steller sea lions Eumetopias jubatus in Alaska: a
-  review of the nutritional stress hypothesis}.
-\newblock {\em Mammal Rev.}, 33(1):3--28, March 2003.
-
-\bibitem{Glazier:2009hq}
-Douglas~S Glazier.
-\newblock {Metabolic level and size scaling of rates of respiration and growth
-  in unicellular organisms}.
-\newblock {\em Funct. Ecol.}, 23(5):963--968, October 2009.
-
-\bibitem{Kooi2000}
-S~A L~M Kooijman.
-\newblock {\em {Dynamic Energy and Mass Budgets in Biological Systems}}.
-\newblock Cambridge, 2000.
-
-\bibitem{Sousa:2010ez}
-T~Sousa, T~Domingos, J~C Poggiale, and S~A L~M Kooijman.
-\newblock {Dynamic energy budget theory restores coherence in biology}.
-\newblock {\em Philos. T. Roy. Soc. B}, 365(1557):3413--3428, October 2010.
-
-\bibitem{Diekmann:2010da}
-O~Diekmann and J~A~J Metz.
-\newblock {How to lift a model for individual behaviour to the population
-  level?}
-\newblock {\em Philos. T. Roy. Soc. B}, 365(1557):3523--3530, November 2010.
-
-\bibitem{murdoch:2003}
-W~W Murdoch, C~J Briggs, and R~M Nisbet.
-\newblock {\em {Consumer-resource Dynamics}}, volume~36 of {\em Monographs in
-  population biology}.
-\newblock Princeton University Press, Princeton, 2003.
-
-\bibitem{Benichou:2014wu}
-Olivier Benichou and S~Redner.
-\newblock {Depletion-Controlled Starvation of a Diffusing Forager}.
-\newblock {\em arXiv}, pages 1--5, May 2014.
-
-\bibitem{Benichou:2016wl}
-O~B{\'e}nichou, M~Chupeau, and S~Redner.
-\newblock {Role of depletion on the dynamics of a diffusing forager}.
-\newblock {\em Journal of Physics A: {\ldots}}, 2016.
-
-\bibitem{Chupeau:2016jf}
-M~Chupeau, O~B{\'e}nichou, and S~Redner.
-\newblock {Universality classes of foraging with resource renewal}.
-\newblock {\em Phys. Rev. E}, 93(3):032403, March 2016.
-
-\bibitem{Persson:1998hz}
-Lennart Persson, Kjell Leonardsson, Andr{\'e}~M De~Roos, Mats Gyllenberg, and
-  Bent Christensen.
-\newblock {Ontogenetic scaling of foraging rates and the dynamics of a
-  size-structured consumer-resource model}.
-\newblock {\em Theor Popul Biol}, 54(3):270--293, December 1998.
-
-\bibitem{murray2011mathematical}
-J~D Murray.
-\newblock {\em {Mathematical Biology: I. An Introduction}}, volume 110 of {\em
-  Interdisciplinary Applied Mathematics}.
-\newblock Springer New York, 2011.
-
-\bibitem{Strogatz:2008wo}
-S~H Strogatz.
-\newblock {\em {Nonlinear Dynamics and Chaos: With Applications to Physics,
-  Biology, Chemistry, and Engineering}}.
-\newblock Studies in nonlinearity. Westview Press, Boulder, 2008.
-
-\bibitem{GuckHolmes}
-J~Guckenheimer and P~Holmes.
-\newblock {\em {Nonlinear Oscillations, Dynamical Systems, and Bifurcations of
-  Vector Fields}}.
-\newblock Springer, New York, 1983.
-
-\bibitem{Gross:2004p2428}
-T~Gross and U~Feudel.
-\newblock {Analytical search for bifurcation surfaces in parameter space}.
-\newblock {\em Physica D}, 195(3-4):292--302, 2004.
-
-\bibitem{Hastings:2001jh}
-Alan Hastings.
-\newblock {Transient dynamics and persistence of ecological systems}.
-\newblock {\em Ecol. Lett.}, 4(3):215--220, May 2001.
-
-\bibitem{Neubert:1997wk}
-MG~Neubert and H~Caswell.
-\newblock {Alternatives to resilience for measuring the responses of ecological
-  systems to perturbations}.
-\newblock {\em Ecology}, 78(3):653--665, 1997.
-
-\bibitem{Caswell:2005eo}
-Hal Caswell and Michael~G Neubert.
-\newblock {Reactivity and transient dynamics of discrete-time ecological
-  systems}.
-\newblock {\em J Differ Equ Appl}, 11(4-5):295--310, April 2005.
-
-\bibitem{Neubert:2009td}
-MG~Neubert and H~Caswell.
-\newblock {Detecting reactivity}.
-\newblock {\em Ecology}, 90(10):2683--2688, 2009.
-
-\bibitem{Yodzis:1992hg}
-P~Yodzis and S~Innes.
-\newblock {Body size and consumer-resource dynamics}.
-\newblock {\em Am. Nat.}, 139(6):1151--1175, 1992.
-
-\bibitem{Brown:2004wq}
-JH~Brown, JF~Gillooly, AP~Allen, VM~Savage, and GB~West.
-\newblock {Toward a metabolic theory of ecology}.
-\newblock {\em Ecology}, 85(7):1771--1789, 2004.
-
-\bibitem{West:2002it}
-Geoffrey~B West, William~H Woodruff, and James~H Brown.
-\newblock {Allometric scaling of metabolic rate from molecules and mitochondria
-  to cells and mammals.}
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 99 Suppl 1(suppl 1):2473--2478,
-  February 2002.
-
-\bibitem{DeLong:2010dy}
-J~P DeLong, J~G Okie, M~E Moses, R~M Sibly, and J~H Brown.
-\newblock {Shifts in metabolic scaling, production, and efficiency across major
-  evolutionary transitions of life}.
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 107(29):12941--12945, July 2010.
-
-\bibitem{West:2001bv}
-Geoffrey~B West, James~H Brown, and Brian~J Enquist.
-\newblock {A general model for ontogenetic growth}.
-\newblock {\em Nature}, 413(6856):628--631, October 2001.
-
-\bibitem{moses2008rmo}
-Melanie~E Moses, Chen Hou, William~H Woodruff, Geoffrey~B West, Jeffery~C
-  Nekola, Wenyun Zuo, and James~H Brown.
-\newblock {Revisiting a Model of Ontogenetic Growth: Estimating Model
-  Parameters from Theory and Data.}
-\newblock {\em Am. Nat.}, 171(5):632--645, 2008.
-
-\bibitem{gillooly2002esa}
-James~F Gillooly, Eric~L Charnov, Geoffrey~B West, Van~M Savage, and James~H
-  Brown.
-\newblock {Effects of size and temperature on developmental time}.
-\newblock {\em Nature}, 417(6884):70--73, May 2002.
-
-\bibitem{hou}
-Chen Hou, Wenyun Zuo, Melanie~E Moses, William~H Woodruff, James~H Brown, and
-  Geoffrey~B West.
-\newblock {Energy uptake and allocation during ontogeny}.
-\newblock {\em Science}, 322(5902):736--739, October 2008.
-
-\bibitem{bettencourt}
-L~M~A Bettencourt, J~Lobo, D~Helbing, C~Kuhnert, and G~B West.
-\newblock {Growth, innovation, scaling, and the pace of life in cities}.
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 104(17):7301--7306, April 2007.
-
-\bibitem{Folland:2008ij}
-J~P Folland, T~M Mc~Cauley, and A~G Williams.
-\newblock {Allometric scaling of strength measurements to body size}.
-\newblock {\em Eur J Appl Physiol}, 102(6):739--745, January 2008.
-
-\bibitem{CalderIII:1983jd}
-W~A Calder~III.
-\newblock {An allometric approach to population cycles of mammals}.
-\newblock {\em J. Theor. Biol.}, 100(2):275--282, January 1983.
-
-\bibitem{Peterson:1984hj}
-R~O Peterson, R~E Page, and K~M Dodge.
-\newblock {Wolves, moose, and the allometry of population cycles}.
-\newblock {\em Science}, 224(4655):1350--1352, June 1984.
-
-\bibitem{Krukonis:1991fk}
-G~Krukonis and W~M Schaffer.
-\newblock {Population cycles in mammals and birds: Does periodicity scale with
-  body size?}
-\newblock {\em J. Theor. Biol.}, 148(4):469--493, February 1991.
-
-\bibitem{Hendriks:2012fc}
-A~Jan Hendriks and Christian Mulder.
-\newblock {Delayed logistic and Rosenzweig{\textendash}MacArthur models with
-  allometric parameter setting estimate population cycles at lower trophic
-  levels well}.
-\newblock {\em Ecol Complex}, 9:43--54, February 2012.
-
-\bibitem{Kendall:1999iy}
-Bruce~E Kendall, Cheryl~J Briggs, William~W Murdoch, Peter Turchin, Stephen~P
-  Ellner, Edward McCauley, Roger~M Nisbet, and Simon~N Wood.
-\newblock {Why do populations cycle? A synthesis of statistical and mechanistic
-  modeling approaches}.
-\newblock {\em Ecology}, 80(6):1789--1805, September 1999.
-
-\bibitem{Hogstedt:2005cr}
-G{\"o}ran H{\"o}gstedt, Tarald Seldal, and Arild Breistol.
-\newblock {Period length in cyclic animal populations}.
-\newblock {\em Ecology}, 86(2):373--378, February 2005.
-
-\bibitem{Kendall:1998hl}
-Kendall, Prendergast, and Bjornstad.
-\newblock {The macroecology of population dynamics: taxonomic and biogeographic
-  patterns in population cycles}.
-\newblock {\em Ecol. Lett.}, 1(3):160--164, November 1998.
-
-\bibitem{Brown:1993p708}
-JH~Brown, PA~Marquet, and ML~Taper.
-\newblock {Evolution of body size: consequences of an energetic definition of
-  fitness}.
-\newblock {\em Am. Nat.}, 142(4):573--584, 1993.
-
-\bibitem{West:1997cg}
-Geoffrey~B West, James~H Brown, and Brian~J Enquist.
-\newblock {A general model for the origin of allometric scaling laws in
-  biology}.
-\newblock {\em Science}, 276(5309):122--126, April 1997.
-
-\bibitem{West:2002ud}
-Geoffrey~B West, William~H Woodruff, and James~H Brown.
-\newblock {Allometric scaling of metabolic rate from molecules and mitochondria
-  to cells and mammals.}
-\newblock {\em Proc. Natl. Acad. Sci. USA}, 99 Suppl 1(Supplement
-  1):2473--2478, February 2002.
-
-\bibitem{Millar:1990p923}
-JS~Millar and G~Hickling.
-\newblock {Fasting endurance and the evolution of mammalian body size}.
-\newblock {\em Funct. Ecol.}, 4(1):5--12, 1990.
-
-\bibitem{Alroy:1998p1594}
-J~Alroy.
-\newblock {Cope's rule and the dynamics of body mass evolution in North
-  American fossil mammals}.
-\newblock {\em Science}, 280(5364):731--734, 1998.
-
-\bibitem{Clauset:2009fh}
-A~Clauset and S~Redner.
-\newblock {Evolutionary model of species body mass diversification}.
-\newblock {\em Phys. Rev. Lett.}, 102(3):038103, January 2009.
-
-\bibitem{Smith:2010p3442}
-Felisa~A Smith, Alison~G Boyer, James~H Brown, Daniel~P Costa, Tamar Dayan,
-  S~K~Morgan Ernest, Alistair~R Evans, Mikael Fortelius, John~L Gittleman,
-  Marcus~J Hamilton, Larisa~E Harding, Kari Lintulaakso, S~Kathleen Lyons,
-  Christy McCain, Jordan~G Okie, Juha~J Saarinen, Richard~M Sibly, Patrick~R
-  Stephens, Jessica Theodor, and Mark~D Uhen.
-\newblock {The evolution of maximum body size of terrestrial mammals}.
-\newblock {\em Science}, 330(6008):1216--1219, November 2010.
-
-\bibitem{Saarinen:2014br}
-Juha~J Saarinen, Alison~G Boyer, James~H Brown, Daniel~P Costa, S~K~Morgan
-  Ernest, Alistair~R Evans, Mikael Fortelius, John~L Gittleman, Marcus~J
-  Hamilton, Larisa~E Harding, Kari Lintulaakso, S~Kathleen Lyons, Jordan~G
-  Okie, Richard~M Sibly, Patrick~R Stephens, Jessica Theodor, Mark~D Uhen, and
-  Felisa~A Smith.
-\newblock {Patterns of maximum body size evolution in Cenozoic land mammals:
-  Eco-evolutionary processes and abiotic forcing}.
-\newblock {\em Proc Biol Sci}, 281(1784):20132049, June 2014.
-
-\bibitem{Damuth:1987kr}
-John Damuth.
-\newblock {Interspecific allometry of population density in mammals and other
-  animals: the independence of body mass and population energy-use}.
-\newblock {\em Biol. J. Linn. Soc.}, 31(3):193--246, July 1987.
-
-\bibitem{allen2002}
-Andrew~P Allen, James~H Brown, and James~F Gillooly.
-\newblock {Global biodiversity, biochemical kinetics, and the
-  energetic-equivalence rule}.
-\newblock {\em Science}, 297(5586):1545--1548, August 2002.
-
-\bibitem{enquist1998}
-Brian~J Enquist, James~H Brown, and Geoffrey~B West.
-\newblock {Allometric scaling of plant energetics and population density}.
-\newblock {\em Nature}, 395(6698):163--165, September 1998.
-
-\bibitem{Kempes:2016}
-Christopher~P Kempes, Lawrence Wang, Jan~P Amend, John Doyle, and Tori Hoehler.
-\newblock {Evolutionary tradeoffs in cellular composition across diverse
-  bacteria}.
-\newblock {\em ISME J}, 10(9):2145--2157, April 2016.
-
-\bibitem{tilman1981}
-David Tilman.
-\newblock {Tests of resource competition theory using four species of lake
-  michigan algae}.
-\newblock {\em Ecology}, 62(3):802--815, June 1981.
-
-\bibitem{dutkiewicz2009}
-S~Dutkiewicz, M~J Follows, and J~G Bragg.
-\newblock {Modeling the coupling of ocean ecology and biogeochemistry}.
-\newblock {\em Global Biogeochem. Cycles}, 23(4):1--15, December 2009.
-
-\bibitem{barton2010}
-Andrew~D Barton, Stephanie Dutkiewicz, Glenn Flierl, Jason Bragg, and Michael~J
-  Follows.
-\newblock {Patterns of diversity in marine phytoplankton}.
-\newblock {\em Science}, 327(5972):1509--1511, March 2010.
-
-\end{thebibliography}
+\putbib[aa_starving3]
 
@@
 \begin{bibunit}[unsrt]
-\section*{Supporting Information for ``The dynamics of starvation and recovery''}
+{\bf Supporting Information for ``The dynamics of starvation and recovery''}\\
 
-\subsection*{Mechanisms of Starvation and Recovery}
+{\bf Mechanisms of Starvation and Recovery}
 Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
@@
 
-\subsection*{Parameter Values and Estimates}
-
-%\subsection*{Standard synthesis and metabolic parameters}
-
+{\bf Parameter Values and Estimates}
 All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
@@
 
-\subsection*{Rate equations for invaders with modified body mass}
+{\bf Rate equations for invaders with modified body mass}
 We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
@@
 
-% \putbib[aa_starving_supplement]
-
-\begin{thebibliography}{10}
-
-\bibitem{Kempes:2012hy}
-Christopher~P Kempes, Stephanie Dutkiewicz, and Michael~J Follows.
-\newblock {Growth, metabolic partitioning, and the size of microorganisms.}
-\newblock {\em PNAS}, 109(2):495--500, January 2012.
-
-\bibitem{kempes2014morphological}
-Christopher~P Kempes, Chinweike Okegbe, Zwoisaint Mears-Clarke, Michael~J
-  Follows, and Lars~EP Dietrich.
-\newblock Morphological optimization for access to dual oxidants in biofilms.
-\newblock {\em Proceedings of the National Academy of Sciences},
-  111(1):208--213, 2014.
-
-\bibitem{West:2001bv}
-Geoffrey~B West, James~H Brown, and Brian~J Enquist.
-\newblock {A general model for ontogenetic growth}.
-\newblock {\em Nature}, 413(6856):628--631, October 2001.
-
-\bibitem{moses2008rmo}
-Melanie~E Moses, Chen Hou, William~H Woodruff, Geoffrey~B West, Jeffery~C
-  Nekola, Wenyun Zuo, and James~H Brown.
-\newblock {Revisiting a model of ontogenetic growth: Estimating model
-  parameters from theory and data}.
-\newblock {\em http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735},
-  171(5):632--645, May 2008.
-
-\bibitem{hou}
-Chen Hou, Wenyun Zuo, Melanie~E Moses, William~H Woodruff, James~H Brown, and
-  Geoffrey~B West.
-\newblock {Energy uptake and allocation during ontogeny}.
-\newblock {\em Science}, 322(5902):736--739, October 2008.
-
-\bibitem{pirt}
-SJ~Pirt.
-\newblock The maintenance energy of bacteria in growing cultures.
-\newblock {\em Proceedings of the Royal Society of London B: Biological
-  Sciences}, 163(991):224--231, 1965.
-
-\bibitem{Heijnen}
-JJ~Heijnen and JA~Roels.
-\newblock A macroscopic model describing yield and maintenance relationships in
-  aerobic fermentation processes.
-\newblock {\em Biotechnology and Bioengineering}, 23(4):739--763, 1981.
-
-\bibitem{peters1986ecological}
-Robert~Henry Peters.
-\newblock {\em The Ecological Implications of Body Size}, volume~2.
-\newblock Cambridge University Press, Cambridge, 1986.
-
-\bibitem{blueweiss1978relationships}
-L~Blueweiss, H~Fox, V~Kudzma, D~Nakashima, R~Peters, and S~Sams.
-\newblock Relationships between body size and some life history parameters.
-\newblock {\em Oecologia}, 37(2):257--272, 1978.
-
-\bibitem{stryer}
-Lubert Stryer.
-\newblock {\em {Biochemistry, Fourth Edition}}.
-\newblock W.H. Freeman and Company, New York, 1995.
-
-\bibitem{Dunbrack:1993ec}
-Robert~L Dunbrack and Malcolm~A Ramsay.
-\newblock {The Allometry of Mammalian Adaptations to Seasonal Environments: A
-  Critique of the Fasting Endurance Hypothesis}.
-\newblock {\em Oikos}, 66(2):336--342, March 1993.
-
-\bibitem{Lindstedt:1985hm}
-Stan~L Lindstedt and Mark~S Boyce.
-\newblock {Seasonality, Fasting Endurance, and Body Size in Mammals}.
-\newblock {\em Am. Nat.}, 125(6):873--878, June 1985.
-
-\bibitem{Lindstedt:2002td}
-S~L Lindstedt and P~J Schaeffer.
-\newblock {Use of allometry in predicting anatomical and physiological
-  parameters of mammals.}
-\newblock {\em Lab. Anim.}, 36(1):1--19, January 2002.
-
-\bibitem{estermann}
-Beda~L Estermann, Hans-Rudolf Wettstein, Franz Sutter, and Michael Kreuzer.
-\newblock Nutrient and energy conversion of grass-fed dairy and suckler beef
-  cattle kept indoors and on high altitude pasture.
-\newblock {\em Animal Research}, 50(6):477--493, 2001.
-
-\bibitem{michaletz2014convergence}
-Sean~T Michaletz, Dongliang Cheng, Andrew~J Kerkhoff, and Brian~J Enquist.
-\newblock Convergence of terrestrial plant production across global climate
-  gradients.
-\newblock {\em Nature}, 512(7512):39--43, 2014.
-
-\bibitem{damuth1987interspecific}
-John Damuth.
-\newblock Interspecific allometry of population density in mammals and other
-  animals: the independence of body mass and population energy-use.
-\newblock {\em Biological Journal of the Linnean Society}, 31(3):193--246,
-  1987.
-
-\end{thebibliography}
+\putbib[aa_starving_supplement]
 

### 2017-06-11 Manuscript/NatureComm/YKR_NatureComm.tex
revised organization for nature comm and some added text

@@
 %\usepackage[left]{lineno}
-%\usepackage[comma,sort&compress]{natbib}
+\usepackage[comma,sort&compress]{natbib}
+\bibpunct{}{}{,}{s}{}{;}
 %\usepackage{authblk}
@@
 % Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations. 
-Our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percent body fat, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
+We also show that our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percent body fat, providing a principled mechanism for a within-lineage driver of Cope's rule.
 % For body masses $< 1.748\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.748\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
@@
 The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly influences how they invest reserves in uncertain environments.
-Such behaviors are generally manifested as tradeoffs between investing in somatic maintenance and growth, or allocating energy towards reproduction~\cite{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
-The timing of these behaviors responds to selective pressure, as the choice of the investment impacts future fitness~\cite{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.
+Such behaviors are generally manifested as tradeoffs between investing in somatic maintenance and growth, or allocating energy towards reproduction~\citep{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
+The timing of these behaviors responds to selective pressure, as the choice of the investment impacts future fitness~\citep{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.
 The influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
 
-The balance between (a) somatic growth and maintenance, and (b) reproduction depends on resource availability~\cite{Morris:1987eo}.
-For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\cite{Tveraa:2003fq}.
-Many bird species invest differently in broods during periods of resource scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
-Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\cite{Kirk:1997cc}.
-Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate nutritional reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+The balance between (a) somatic growth and maintenance, and (b) reproduction depends on resource availability~\citep{Morris:1987eo}.
+For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\citep{Tveraa:2003fq}.
+Many bird species invest differently in broods during periods of resource scarcity compared to normal periods~\citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\citep{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
+Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\citep{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\citep{Kirk:1997cc}.
+Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate nutritional reserves~\citep{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
 
 Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.
-The data collected thus far has shown that diverse mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development (blastocyst implantation) until nutritional reserves can be accumulated~\cite{Mead:1989dt,Sandell:1990kw}.
-Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
-In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.
+The data collected thus far has shown that diverse mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development (blastocyst implantation) until nutritional reserves can be accumulated~\citep{Mead:1989dt,Sandell:1990kw}.
+Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\citep{Bulik:1999eo,Trites:2003cc}.
+In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \citep{Glazier:2009hq}.
 The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the importance and universality of the fundamental tradeoff between somatic and reproductive investment.
@@
 
-Though straightforward conceptually, incorporating the energetic dynamics of individuals~\cite{Kooi2000} into a population-level framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles~\cite{Diekmann:2010da}.
+Though straightforward conceptually, incorporating the energetic dynamics of individuals~\citep{Kooi2000} into a population-level framework~\citep{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles~\citep{Diekmann:2010da}.
 An alternative approach involves modeling the macroscale relations that guide somatic versus reproductive investment in a consumer-resource system.
-For example, macroscale Lotka-Volterra models assume that the growth rate of the consumer population depends on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction~\cite{murdoch:2003}.
+For example, macroscale Lotka-Volterra models assume that the growth rate of the consumer population depends on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction~\citep{murdoch:2003}.
 
@@
 Such a constraint leads to reproductive time lags due to some members of the population going hungry and then recovering.
-Additionally, we incorporate the idea that reproduction is strongly constrained allometrically \cite{Kempes:2012hy}, and is not generally linearly related to resource density.
-As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.
-\\
-
-The balance between (a) somatic growth and maintenance, and (b) reproduction depends on resource availability~\cite{Morris:1987eo}.
-For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\cite{Tveraa:2003fq}.
-Many bird species invest differently in broods during periods of resource scarcity compared to normal periods~\cite{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\cite{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
-Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\cite{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\cite{Kirk:1997cc}.
-Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate nutritional reserves~\cite{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
-
-Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.
-The data collected thus far has shown that diverse mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development (blastocyst implantation) until nutritional reserves can be accumulated~\cite{Mead:1989dt,Sandell:1990kw}.
-Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\cite{Bulik:1999eo,Trites:2003cc}.
-In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \cite{Glazier:2009hq}.
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the importance and universality of the fundamental tradeoff between somatic and reproductive investment.
-However the general dynamic implications of these constraints are unknown.
-
-Though straightforward conceptually, incorporating the energetic dynamics of individuals~\cite{Kooi2000} into a population-level framework~\cite{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles~\cite{Diekmann:2010da}.
-An alternative approach involves modeling the macroscale relations that guide somatic versus reproductive investment in a consumer-resource system.
-For example, macroscale Lotka-Volterra models assume that the growth rate of the consumer population depends on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction~\cite{murdoch:2003}.
-
-In this work, we adopt an alternative approach in which we \emph{explicitly} account for resource limitation and the subsequent effects of starvation.
-Namely, only individuals with sufficient energetic reserves can reproduce.
-Such a constraint leads to reproductive time lags due to some members of the population going hungry and then recovering.
-Additionally, we incorporate the idea that reproduction is strongly constrained allometrically \cite{Kempes:2012hy}, and is not generally linearly related to resource density.
-As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.
-\\
+Additionally, we incorporate the idea that reproduction is strongly constrained allometrically \citep{Kempes:2012hy}, and is not generally linearly related to resource density.
+As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.\\
 
@@
 The rate at which consumers transition between states and consume resources is dependent on their overall abundance, the abundance of resources, the efficiency of converting resources into metabolism, and how that metabolism is partitioned between maintenance and growth purposes.
-In the supplementary information (SI) we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we describe below.
+In the Supplementary Information we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we describe below.
 
@@
 \includegraphics[width=0.4\textwidth]{fig_FixedPoint2.eps}
-\caption{\small{ The transcritical (dashed) and Hopf bifurcation (solid) as a
+\caption{\small{ The transcritical (TC; dashed line) and Hopf bifurcation (solid line) as a
   function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-  bifurcation conditions separate parameter space into unphysical, cyclic,
-  and steady state dynamic regimes.  The colors show the steady state densities for the energetically replete consumers $F^*$, (warmer colors denote higher densities).  
+  bifurcation conditions separate parameter space into unphysical (left of the TC), cyclic,
+  and steady state dynamic regimes.  The colors show the steady state densities for the energetically replete consumers $F^*$.  
   % Steady state densities for the energetically deficient consumers $H^*$ (not shown)
@@
 Resources are eaten by the hungry consumers at rate $\rho R + \delta$, that accounts for their somatic growth ($\rho R$) and  maintenance ($\delta$).
-Full consumers eat resources at a constant rate $\beta$ that accounts for maximal maintenance and somatic growth (see SI for mechanistic derivations of these rates from resource energetics).
+Full consumers eat resources at a constant rate $\beta$ that accounts for maximal maintenance and somatic growth (see Supplementary Information for mechanistic derivations of these rates from resource energetics).
 %This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
-The NSM represents an ecologically motivated fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\cite{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\cite{Persson:1998hz}.
+The NSM represents an ecologically motivated fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\citep{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\citep{Persson:1998hz}.
 
@@
 
-This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see SI for a full derivation and the dimensional form).
+This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see Supplementary Information for a full derivation and the dimensional form).
 Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
-This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\cite{murray2011mathematical}, except that the resource density does not appear in the growth term.
-As discussed above, the attributes of reproduction and mortality have been explicitly apportioned to the full and hungry consumers, respectively, so that the growth in the total density is decoupled from the resource density.
+This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\citep{murray2011mathematical}, except that the resource density does not appear in the growth term.
+% As discussed above, the attributes of reproduction and mortality have been explicitly apportioned to the full and hungry consumers, respectively, so that the growth in the total density is decoupled from the resource density.
+The rate of reproduction is independent of resource density because it is assumed that the satiated state of the full consumer allows it to partition a constant amount of energy towards reproduction, whereas a starved consumer partitions no energy towards reproduction.
+The rate of reproduction for the total consumer density \emph{is} dependent on resource density, which determines the size of the full and starved portions of the consumer population.
+Similarly, the consumer maintenance terms ($\delta H$ and $\beta F$) are independent of resource density because they represent a minimal energetic requirement for consumers in the $H$ and $F$ state, respectively.
+It follows that model predictions are robust only when $R \gg 0$, which holds for all cases that we explore.
 
-Equation~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
-\begin{align}
-\label{eq:ss}
-\begin{split}
-F^* &= (\sigma-\lambda)\frac{ \alpha  \lambda  \mu ^2  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
-H^* &= (\sigma-\lambda)\frac{ \alpha  \lambda ^2 \mu  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
-R^* &= (\sigma - \lambda)\frac{\mu  }{A}.
-\end{split}
-\end{align}
-where $A=(\lambda  \xi  \rho +\mu  \sigma )$ and $B=(\beta  \mu  \xi +\delta  \lambda  \xi -\lambda  \mu )$.
-The stability of this fixed point is determined by the Jacobian matrix $\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.
-The parameters in Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue of $\mathbf{J}$ is negative, so that the system is stable with respect to small perturbations from the fixed point.
-Because this fixed point is unique, it is the global attractor for all population trajectories for any initial condition where the resource and consumer densities are both nonzero.
-
-From Eq.~\eqref{eq:ss}, an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that the consumer and resource densities are positive.
+From the solution to the single internal fixed point (Eq.~\eqref{eq:ss}, see Methods), an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that the consumer and resource densities are positive.
 %In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
-The condition $\sigma = \lambda$ thus represents a transcritical (TC) bifurcation~\cite{Strogatz:2008wo} that demarcates a physical from an unphysical regime where all steady-state densities become negative after intersecting the trivial fixed point $(F^*,H^*,R^*)=(0,0,0)$.
+The condition $\sigma = \lambda$ thus represents a transcritical (TC) bifurcation~\citep{Strogatz:2008wo} that demarcates a physical from an unphysical regime where all steady-state densities become negative after intersecting the trivial fixed point $(F^*,H^*,R^*)=(0,0,0)$.
 The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
-As we will discuss below, this inequality is a natural consequence of allometric constraints~\cite{Kempes:2012hy} for organisms within empirically observed body size ranges. %(Fig.~\ref{fig:gvs})
+As we will discuss below, this inequality is a natural consequence of allometric constraints~\citep{Kempes:2012hy} for organisms within empirically observed body size ranges. %(Fig.~\ref{fig:gvs})
 
 In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).
-In continuous-time systems, a limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts~\cite{GuckHolmes}.
-This Hopf bifurcation is defined by ${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial of the Jacobian matrix~\cite{Gross:2004p2428}.
+In continuous-time systems, a limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts~\citep{GuckHolmes}.
+This Hopf bifurcation is defined by ${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial of the Jacobian matrix~\citep{Gross:2004p2428}.
 As the system parameters are tuned to be within the stable regime, but close to the Hopf bifurcation, the amplitude of the transient cycles becomes large.
-Given that ecological systems are constantly being perturbed~\cite{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase the extinction risk~\cite{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
+Given that ecological systems are constantly being perturbed~\citep{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase extinction risk~\citep{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
 %Thus the distance of a system from the Hopf bifurcation provides a measure of its persistence.
@@
 This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.
-\\
+
 % Whereas the relation between consumer growth rate $\lambda$ and the starvation rate $\sigma$ defines an absolute bound of biological feasibility---the TC bifurcation---$\sigma$ also determines the sensitivity of the consumer population to changes in resource density.
@@
 
-\noindent {\bf Role of allometry} \\
+% \noindent {\bf Role of allometry} \\
 %[Link Allometry stuff to our model - what does it provide to the story]
@@
 While there are no a priori constraints on the parameters in the NSM, most
-organisms correspond to restricted portions of the parameter space.  Here we
+organisms correspond to restricted portions of the parameter space.  We
 use allometric scaling relations to constrain the covariation of rates in a
-principled and biologically meaningful manner.  Allometric scaling relations
+principled and biologically meaningful manner (see Methods).  Allometric scaling relations
 highlight common constraints and average trends across large ranges in body
 size and species diversity. Many of these relations can be derived from a
-small set of assumptions and below we describe our framework to determine the
-covariation of timescales and rates across the range of mammals for each of
-the key parameters of our model (cf. ref.~\cite{Yodzis:1992hg}).  We are thereby
-able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.
+small set of assumptions and in the Methods we describe our framework to determine the
+covariation of timescales and rates across a range of body sizes for each of
+the key parameters of our model (cf. ref.~\citep{Yodzis:1992hg}).  
+% We are thereby able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.
 
@@
 metabolism, which can be used to describe a variety of organismal features
-\cite{Brown:2004wq}.  The scaling relation between an organism's metabolic
+\citep{Brown:2004wq}.
+We derive relationships for the rates of reproduction, starvation, recovery, and mortality based on first principles, and as a function of an organism's body size and metabolic rate (see Methods).
+Because we aim to explore the starvation-recovery dynamics as a function of an organism's body mass $M$, we parameterize these rates in terms of the \emph{percent} gain and loss of the asymptotic (maximum) body mass, $\epsilon M$, where different values of $\epsilon$ defines different states of the consumer (Fig.~\ref{fig:growth}).
+Although the rate equations \eqref{eq:system} are general and can in principle be used to explore the starvation recovery dynamics for most organisms, here we focus on allometric relationships for terrestrial-bound lower trophic level endotherms (see Supplementary Information for values), specifically herbivorous mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
+Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
+Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.\\
+
+% \noindent {\bf  Stabilizing effects of allometric constraints} \\
+% As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
+% Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
+% Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
+%Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
+% These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
+However, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation.
+% This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.\\
+
+% Previous studies used allometric constraints to explain the periodicity of cyclic populations \citep{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
+% However this relation seems to hold only for some species \citep{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\citep{Kendall:1999iy,Hogstedt:2005cr}.
+% Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\citep{Kendall:1998hl}.
+% However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.\\
+
+\begin{figure*}
+\centering
+\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometricComb3.eps}
+\caption{\small{ Probability of extinction for a consumer with ({\bf a}) $M=10^2$g, ({\bf b}) $M=10^4$g, and ({\bf c}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
+Note the change in scale for $M=10^4$ and $M=10^6$g.
+Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
+The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
+\end{figure*}
+
+%Extinction to the left and the right
+\noindent {\bf Extinction risk}\\
+As the allometric derivations of the NSM rate laws reveal (see Methods), starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
+Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
+Indeed, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation. 
+This suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
+% Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
+
+Higher rates of starvation result in a larger flux of the population to the hungry state.
+In this state, reproduction is absent, thus increasing the likelihood of extinction.
+From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
+We therefore examine the competing effects of cyclic dynamics vs. changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
+To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
+This procedure is repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for organisms with mass ranging from $10^2$ to $10^6$ grams.
+%the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
+In each replicate the initial densities are chosen to be $(XF^*,XH^*,R^*)$, with $X$ a random variable that is uniformly distributed in $[0,2]$.
+By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-7}$ to $10^{-3}$.
+%, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
+As expected, higher rates of extinction correlate with both high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if $\sigma$ is small.
+For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
+For high values of $\sigma$ and low values of $\rho$, increased extinction risk arises because of the decrease in the steady-state consumer population density (Figs. \ref{fig:fp}b, \ref{fig:ext}).
+This interplay creates an `extinction refuge', such that for a constrained range of $\sigma$ and $\rho$, extinction probabilities are minimized.
+
+We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge across a range of $M$ (Fig. \ref{fig:ext}a-c, white points).
+These values are close enough to the Hopf bifurcation to avoid low steady-state densities, and far enough away to avoid large-amplitude transient cycles.
+The feature that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
+Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that serve to minimize extinction risk.
+This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
+We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig.
+\ref{fig:ext}a-c), suggesting that the population dynamics of larger organisms are more sensitive to smaller changes in physiological rates controlling starvation and recovery.
+% OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
+To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
+
+%Results are insensitive to alpha but possibly sensitive to beta (eff parameter)
+
+%NOTES 7/8/16
+%Calder has a hypothesis that period ~ Mass^(1/4).
+%Can we get "transient period" from Jacobian?
+%Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.
+
+\noindent {\bf Discussion}\\
+Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\citep{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\citep{West:2002ud}.
+Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
+One important such example is the \emph{fasting endurance hypothesis}, which contends that larger body size, with consequent lower metabolic rates and increased ability to maintain more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability~\citep{Millar:1990p923}.
+Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's rule)~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ grams~\citep{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\citep{Clauset:2009fh}.
+These trends are thought to be driven by a combination of climate change and niche availability~\citep{Saarinen:2014br}; however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, have not been explored.
+We argue that the NSM provides a suitable framework to explore these issues.
+
+\noindent {\bf Energetic barriers to body size}
+The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}a).
+Similar predictions have been made for carnivore populations using alternative consumer-resource models \citep{DeLong:2012kw}.
+Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998}.
+The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size~\citep{Damuth:1987kr,allen2002,enquist1998}.
+%Previous studies have focused on t
+This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism, such that $N^*\propto M^{-3/4}/B_{0}$~\citep{allen2002,enquist1998}.
+This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \citep{Damuth:1987kr,allen2002,enquist1998}.
+Figure \ref{fig:mass}a shows that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly constant over this same range.
+This result is remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to energy equivalence.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.48\textwidth]{fig_FPAllometric2.eps}
+\caption{\small{ ({\bf a}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
+  body mass. Inset: Resource steady state $R^*$ as a function of consumer body mass.
+  ({\bf b}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
+  The data are from Damuth \citep{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see Supplementary Information and Refs.
+\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:mass}}
+\end{figure}
+
+Our model shows that the equivalence breaks down at large $M$ suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
+%, violating the energy equivalence.
+% set by deviations from constant efficiency obeyed by other populations.
+In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
+This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero, causing the timescales to become infinite and the rates $\mu$ and $\lambda$ to equal zero.
+The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$.
+This point predicts a strong upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
+Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is within an order of magnitude of the maximum body size observed in the North American mammalian fossil record~\citep{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\citep{Clauset:2009fh}.
+It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters.
+We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\citep{Kempes:2012hy,Kempes:2016}.
+% reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
+\\
+
+%Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end.
+%Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is close to the minimum observed mammalian body size of ca.\ 1.3-2.5g, a range that occurs as the recovery rate begins its decline.
+%In addition to known transport limitations~\citep{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.
+%This result mirrors other efforts \citep{Kempes:2012hy,Kempes:2016} where at a given scale multiple limitations constrain the smallest possibilities for life within a class of organisms.
+
+%The combined steady-state dynamics and allometric timescales of the NSM predict that larger mammals should outcompete smaller ones, and this suggests that the model may provide a framework with which to explore the energetic mechanisms behind Cope's rule.
+%, which we explore further below.
+
+\noindent{\bf A within-lineage driver of Cope's Rule}
+The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
+\ref{fig:mass}a, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \citep{tilman1981,dutkiewicz2009,barton2010}.
+Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but these absolute limits do not offer a within-lineage mechanism by which larger body sizes are selected for or against.
+We will now show that the NSM indeed provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time known as Cope's Rule.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
+\caption{\small{ Invasion feasibility for organisms with a proportional change in
+  mass $\chi$ against a population with a resident body mass $M$.  The blue
+  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
+  red region denotes values of $\chi$ that result in a mass that is below the
+  starvation threshold and are thus infeasible.
+  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
+  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\citep{Smith:2010p3442}.}\label{fig:invasion}}
+\end{figure}
+
+To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120$ times the mass of a male African elephant.
+Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
+
+We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by $^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$.
+If $\chi < 0$, individuals within the invading mutant population have fewer metabolic reserves, and if $\chi>0$, individuals have more metabolic reserves than the resident population.
+For the allowable values of $\chi$ the adjusted mass should exceed the minimal amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given Equation \ref{t1}, implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
+Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$ and the lower bound is mass-dependent.
+The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
+Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the body composition of starved foragers, is assumed to remain unchanged for the invader population (see Supplementary Information for detailed derivations of invader rates).
+%There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
+
+To assess the susceptibility of the resident population to invasion, we determine which consumer pushes the steady-state resource density to lower values for a given value of $\chi$, with the expectation that populations able to survive on lower resource densities have a competitive advantage \citep{tilman1981}.
+We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population (Fig.
+\ref{fig:invasion}).
+However, for $M> 1.748\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage from having more metabolic reserves to having less.
+
+%, and this is due to the changing
+%covariance between energetic rates as a function of modified energetic
+%reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
+
+The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
+This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca.
+$1.5\times10^7$g and the late Miocene \emph{Deinotherium} at ca.
+$1.74\times10^7$g ~\citep{Smith:2010p3442}.
+Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \citep{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals outlined by Alroy~\citep{Alroy:1998p1594}.
+%as well as  \citep{Alroy:1998p1594,Clauset:2009fh}.
+%This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
+While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general within-lineage mechanism for the evolution of larger body size among terrestrial mammals.
+
+%and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
+
+%One might be concerned a greater number of large mammals are currently not observed in the modern world given that larger mammals are less susceptible to extinction.
+%However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \citep{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also much more geographically widespread than today.
+%These results, combined with our findings, suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \citep{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
+
+%Closing
+The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations~\citep{Stearns:1989ip}.
+The NSM is a minimal and general model that incorporates the dynamics of starvation and recovery that are expected to occur in resource-limited environments.
+By incorporating allometric relations between the rates in the NSM, we found: (i) different organismal masses have distinct population dynamic regimes, (ii) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and (iii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of maximum body size.
+We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.
+Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.
+
+\section*{Methods}
+\small{
+{\bf Analytical solution to the NSM}
+Equation~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
+\begin{align}
+\label{eq:ss}
+\begin{split}
+F^* &= (\sigma-\lambda)\frac{ \alpha  \lambda  \mu ^2  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
+H^* &= (\sigma-\lambda)\frac{ \alpha  \lambda ^2 \mu  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
+R^* &= (\sigma - \lambda)\frac{\mu  }{A}.
+\end{split}
+\end{align}
+where $A=(\lambda  \xi  \rho +\mu  \sigma )$ and $B=(\beta  \mu  \xi +\delta  \lambda  \xi -\lambda  \mu )$.
+The stability of this fixed point is determined by the Jacobian matrix $\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.
+The parameters in Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue of $\mathbf{J}$ is negative, so that the system is stable with respect to small perturbations from the fixed point.
+Because this fixed point is unique, it is the global attractor for all population trajectories for any initial condition where the resource and consumer densities are both nonzero.
+
+{\bf Metabolic scaling relationships}
+The scaling relation between an organism's metabolic
 rate $B$ and its body mass $M$ at reproductive maturity is known to scale as
-$B = B_0 M^\eta$~\cite{West:2002it}, where the scaling exponent $\eta$ is
-typically close to $2/3$ or $3/4$ for metazoans (e.g., ref. \cite{Brown:2004wq}),
+$B = B_0 M^\eta$~\citep{West:2002it}, where the scaling exponent $\eta$ is
+typically close to $2/3$ or $3/4$ for metazoans (e.g., ref. \citep{Brown:2004wq}),
 and has taxonomic shifts for unicellular species between $\eta\approx 1$ in
 eukaryotes and $\eta\approx 1.76$ in bacteria
-\cite{DeLong:2010dy,Kempes:2012hy}.
+\citep{DeLong:2010dy,Kempes:2012hy}.
 %Justin adding an explanation for beta
@@
 metazoans
-\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. This relation is derived from the simple balance condition 
+\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. This relation is derived from the simple balance condition 
 $B_{0}m^{\eta}=E_{m}\dot{m}+B_{m}m\,,$
@@
 % \end{eqnarray}
-\cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy} where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
+\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy} where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
 the metabolic rate to support an existing unit of mass, and $m$ is the mass
 of the organism at any point in its development.  This balance has the
-general solution \cite{bettencourt,Kempes:2012hy}
+general solution \citep{bettencourt,Kempes:2012hy}
 \begin{eqnarray}
@@
 \end{eqnarray}
-where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically (see SI).
-We now use this solution to define the timescale for reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \cite{moses2008rmo} for a detailed presentation of these timescales).
+where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically (see Supplementary Information).
+We now use this solution to define the timescale for reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \citep{moses2008rmo} for a detailed presentation of these timescales).
 The time that it takes to reach a particular mass $\epsilon M$ is given by the timescale
@@
 where we will define values of $\epsilon$ to describe a set of rates within our model.
-For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$; \cite{West:2001bv}).
-The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$, this rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ \cite{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
+For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$; \citep{West:2001bv}).
+The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$, this rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
 
@@
 and recovery by the timescales required to reach, or return from, specific
-fractions of the replete-state mass (see SI, Table I for parameterizations).  We define
+fractions of the replete-state mass (see Supplementary Information, Table I for parameterizations).  We define
 $m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction of
@@
 \end{eqnarray}
-where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in the biosynthetic energetics during recovery (see SI).
+where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in the biosynthetic energetics during recovery (see Supplementary Information).
 It should be noted that more complicated ontogenetic models explicitly handle
-storage \cite{hou}, whereas this feature is implicitly covered by the body
+storage \citep{hou}, whereas this feature is implicitly covered by the body
 fat scaling in our framework.
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometricComb3.eps}
-\caption{\small{ Probability of extinction for a consumer with (\emph{A}) $M=10^2$g, (\emph{B}) $M=10^4$g, and (\emph{C}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
-Note the change in scale for $M=10^4$ and $M=10^6$g.
-Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
-The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
-\end{figure*}
 
@@
 body mass, which differs from $E_{m}$, the energy required to
-synthesis a unit of biomass \cite{hou}. Given the replete mass, $M$, of an organism, the
+synthesis a unit of biomass \citep{hou}. Given the replete mass, $M$, of an organism, the
 above energy balance prescribes the mass trajectory of a non-consuming
@@
 $\epsilon_{\mu}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
-mass has been shown to be roughly proportional to body mass~\cite{Folland:2008ij} in
+mass has been shown to be roughly proportional to body mass~\citep{Folland:2008ij} in
 mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant.
@@
 and $\mu=1/t_{\mu}$.
-
-Although the rate equations \eqref{eq:system} are general, here we focus on parameterizations for terrestrial-bound endotherms, specifically mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
-Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
-Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.\\
-
-% \noindent {\bf  Stabilizing effects of allometric constraints} \\
-% As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
-% Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
-% Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
-%Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
-% These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
-However, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation.
-% This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.\\
-
-% Previous studies used allometric constraints to explain the periodicity of cyclic populations \cite{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
-% However this relation seems to hold only for some species \cite{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\cite{Kendall:1999iy,Hogstedt:2005cr}.
-% Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\cite{Kendall:1998hl}.
-% However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.\\
-
-%Extinction to the left and the right
-\noindent {\bf Extinction risk}\\
-As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
-Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
-Indeed, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation. 
-This suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
-% Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
-
-Higher rates of starvation result in a larger flux of the population to the hungry state.
-In this state, reproduction is absent, thus increasing the likelihood of extinction.
-From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
-We therefore examine the competing effects of cyclic dynamics vs. changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
-To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
-This procedure is repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for organisms with mass ranging from $10^2$ to $10^6$ grams.
-%the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
-In each replicate the initial densities are chosen to be $(XF^*,XH^*,R^*)$, with $X$ a random variable that is uniformly distributed in $[0,2]$.
-By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-7}$ to $10^{-3}$.
-%, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
-As expected, higher rates of extinction correlate with both high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if $\sigma$ is small.
-For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
-For high values of $\sigma$ and low values of $\rho$, increased extinction risk arises because of the decrease in the steady-state consumer population density (Figs.
-\ref{fig:fp}\emph{B}, \ref{fig:ext}).
-This interplay creates an `extinction refuge', such that for a constrained range of $\sigma$ and $\rho$, extinction probabilities are minimized.
-
-We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge across a range of $M$ (Fig. \ref{fig:ext}\emph{A-C}, white points).
-These values are close enough to the Hopf bifurcation to avoid low steady-state densities, and far enough away to avoid large-amplitude transient cycles.
-The feature that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
-Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that serve to minimize extinction risk.
-This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
-We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig.
-\ref{fig:ext}\emph{A-C}), suggesting that the population dynamics of larger organisms are more sensitive to smaller changes in physiological rates controlling starvation and recovery.
-% OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
-
-%Results are insensitive to alpha but possibly sensitive to beta (eff parameter)
-
-%NOTES 7/8/16
-%Calder has a hypothesis that period ~ Mass^(1/4).
-%Can we get "transient period" from Jacobian?
-%Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.
-
-\noindent {\bf Dynamic and energetic barriers to body size} \\
-Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\cite{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\cite{West:2002ud}.
-Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
-One important such example is the \emph{fasting endurance hypothesis}, which contends that larger body size, with consequent lower metabolic rates and increased ability to maintain more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability~\cite{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's rule)~\cite{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ grams~\cite{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\cite{Clauset:2009fh}.
-These trends are thought to be driven by a combination of climate change and niche availability~\cite{Saarinen:2014br}; however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, have not been explored.
-We argue that the NSM provides a suitable framework to explore these issues.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.48\textwidth]{fig_FPAllometric2.eps}
-\caption{\small{ (\emph{A}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-  body mass. Inset: Resource steady state $R^*$ as a function of consumer body mass.
-  (\emph{B}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  The data are from Damuth \cite{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see SI and Refs.
-\cite{West:2001bv,hou,moses2008rmo}).}\label{fig:mass}}
-\end{figure}
-
-%Minimal mass bound
-%A theoretical lower bound on mammalian body size occurs when $\epsilon=0$ where mammals have no metabolic reserves and immediately starve; this occurs at a size of $M=0$, which is obviously infeasible.
-%Such prediction gives an extreme limit on size but does not account for the subtleties of starvation dynamics
-%that may limit body size (among other factors).
-The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}\emph{A}).
-Similar predictions have been made for carnivore populations using alternative consumer-resource models \cite{DeLong:2012kw}.
-Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \cite{Damuth:1987kr,allen2002,enquist1998}.
-The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size~\cite{Damuth:1987kr,allen2002,enquist1998}.
-%Previous studies have focused on t
-This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism, such that $N^*\propto M^{-3/4}/B_{0}$~\cite{allen2002,enquist1998}.
-This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \cite{Damuth:1987kr,allen2002,enquist1998}.
-Figure \ref{fig:mass}\emph{A} shows that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}\emph{B} shows that $F^{*}B$ is nearly constant over this same range.
-This result is remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to energy equivalence.
-
-Our model shows that the equivalence breaks down at the maximum observed body sizes for mammals, suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
-%, violating the energy equivalence.
-% set by deviations from constant efficiency obeyed by other populations.
-In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
-This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero, causing the timescales to become infinite and the rates $\mu$ and $\lambda$ to equal zero.
-The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$.
-This point predicts a strong upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
-Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is close to the maximum body size observed in the North American mammalian fossil record~\cite{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\cite{Clauset:2009fh}.
-It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters.
-We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\cite{Kempes:2012hy,Kempes:2016}.
-% reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
-
-%Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end.
-%Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is close to the minimum observed mammalian body size of ca.\ 1.3-2.5g, a range that occurs as the recovery rate begins its decline.
-%In addition to known transport limitations~\cite{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.
-%This result mirrors other efforts \cite{Kempes:2012hy,Kempes:2016} where at a given scale multiple limitations constrain the smallest possibilities for life within a class of organisms.
-
-%The combined steady-state dynamics and allometric timescales of the NSM predict that larger mammals should outcompete smaller ones, and this suggests that the model may provide a framework with which to explore the energetic mechanisms behind Cope's rule.
-%, which we explore further below.
-
-We contend that the NSM provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time.
-The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
-\ref{fig:mass}\emph{A}, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \cite{tilman1981,dutkiewicz2009,barton2010}.
-Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but does not offer a within-lineage mechanism by which larger body sizes are selected for.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
-\caption{\small{ Invasion feasibility for organisms with a proportional change in
-  mass $\chi$ against a population with a resident body mass $M$.  The blue
-  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
-  red region denotes values of $\chi$ that result in a mass that is below the
-  starvation threshold and are thus infeasible.
-  Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
-  The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\cite{Smith:2010p3442}.}\label{fig:invasion}}
-\end{figure}
-
-To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120$ times the mass of a male African elephant.
-Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
-We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by $^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$.
-For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given Equation \ref{t1}, implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
-Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
-The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
-Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the body composition of starved foragers, is assumed to remain unchanged for the invader population (see SI for detailed derivations of invader rates).
-%There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
-
-To assess the susceptibility of the resident population to invasion, we determine which consumer has a lower steady-state resource density for a given value of $\chi$, again with the expectation that populations able to survive on lower resource densities have a competitive advantage \cite{tilman1981}.
-We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population (Fig.
-\ref{fig:invasion}).
-However, for $M> 1.748\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage for higher values of $M$.
-
-%, and this is due to the changing
-%covariance between energetic rates as a function of modified energetic
-%reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
-
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
-This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca.
-$1.5\times10^7$g and the late Miocene \emph{Deinotherium} at ca.
-$1.74\times10^7$g ~\cite{Smith:2010p3442}.
-Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \cite{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals outlined by Alroy~\cite{Alroy:1998p1594}.
-%as well as  \cite{Alroy:1998p1594,Clauset:2009fh}.
-%This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
-While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\cite{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general within-lineage mechanism for the evolution of larger body size among terrestrial mammals.
-
-%and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
-
-%One might be concerned a greater number of large mammals are currently not observed in the modern world given that larger mammals are less susceptible to extinction.
-%However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also much more geographically widespread than today.
-%These results, combined with our findings, suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \cite{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
-
-%Closing
-The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations~\cite{Stearns:1989ip}.
-The NSM is a minimal and general model that incorporates the dynamics of starvation and recovery that are expected to occur in resource-limited environments.
-By incorporating allometric relations between the rates in the NSM, we found: (i) different organismal masses have distinct population dynamic regimes, (ii) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and (iii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of maximum body size.
-We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.
-Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.
+}
 
@@
 %
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+% SUPPLEMENTARY MATERIAL
+%%%%%%%%%%%%%%%%%%%%%%%%
+
 \clearpage
@@
 For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
-\cite{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
+\citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
 This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g.
-\cite{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \cite{kempes2014morphological}.
+\citep{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \citep{kempes2014morphological}.
 Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).
@@
 We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
-\begin{align}
-\begin{split}
-\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
-\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
-&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
-\end{split}
-\end{align}
-where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
-%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
-In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
-Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
-The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
-A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
-
-In the above equations $Y$ represents the yield coefficient (e.g., Refs.
-\cite{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
-If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
-This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
-In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment.
-
-Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
-As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
-Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
 \begin{align}
@@
 \end{align}
+where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
+In the above equations $Y$ represents the yield coefficient (e.g., Refs.
+\citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
+If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
+This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
+In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
+%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
+
+%\begin{align}
+%\begin{split}
+%\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+%\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
+%\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
+%&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
+%\end{split}
+%\end{align}
+%
+%In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
+Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
+The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
+%A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
+%
+%Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
+As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
+Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
+
 %where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
@@
 The above equations represent the system of equations presented in the main text.
+\\
 
@@
 Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs.
-\cite{West:2001bv,moses2008rmo} and the supplement for Ref.
-\cite{hou}).
-We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \cite{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
+\citep{West:2001bv,moses2008rmo} and the supplement for Ref.
+\citep{hou}).
+We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
 The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.
 For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs.
-\cite{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
+\citep{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
 We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).
@@
 This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref.
-\cite{peters1986ecological} for a review).
-We use $m_{0}=0.097M^{0.92}$ \cite{blueweiss1978relationships} which encompasses the widest range of body sizes \cite{peters1986ecological}.
+\citep{peters1986ecological} for a review).
+We use $m_{0}=0.097M^{0.92}$ \citep{blueweiss1978relationships} which encompasses the widest range of body sizes \citep{peters1986ecological}.
 
@@
 Refs.
-\cite{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \cite{moses2008rmo}.
-The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \cite{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \cite{stryer,hou}.
-For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \cite{hou}.
+\citep{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \citep{moses2008rmo}.
+The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \citep{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \citep{stryer,hou}.
+For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \citep{hou}.
 
-For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \cite{Dunbrack:1993ec}, $0.02M^{1.19}$ \cite{Lindstedt:1985hm}, and $0.026M^{1.14}$ \cite{Lindstedt:2002td}.
-We use the values from \cite{Lindstedt:1985hm} which falls in the middle of this range.
-Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \cite{Lindstedt:2002td}.
+For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \citep{Dunbrack:1993ec}, $0.02M^{1.19}$ \citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ \citep{Lindstedt:2002td}.
+We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range.
+Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \citep{Lindstedt:2002td}.
 
@@
 That is, we are interested in the relative rates of resource recovery and consumption by the total population.
-From \cite{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$).
-Considering an energy density of $18200$ (J g$^{-1}$) of grass \cite{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$).
+From \citep{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$).
+Considering an energy density of $18200$ (J g$^{-1}$) of grass \citep{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$).
 This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.
@@
 Given that our model considers transitions over entire stages of ontogeny or nutritional states, the value of $Y$ must represent yields integrated over entire life stages.
-Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \cite{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
+Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
 \begin{equation}
@@
 From the data in Ref.
-\cite{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
-It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \cite{damuth1987interspecific}.
+\citep{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
+It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{damuth1987interspecific}.
 We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above.
 However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.
-
+\\
 %For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
@@
      \hline
-   $\eta$ & $3/4$  &  (e.g. \cite{West:2001bv,moses2008rmo,hou}) \\
-   $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \cite{moses2008rmo,West:2001bv,hou} \\
-   $E_{m}^{\prime}$ & $7000$  & \cite{stryer,hou} \\
-   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \cite{hou}  \\
+   $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
+   $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
+   $E_{m}^{\prime}$ & $7000$  & \citep{stryer,hou} \\
+   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
 %   $a$ & $1.78\times10^{-6}$  \quad \quad \\
 %   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
-   $\gamma$ & $1.19$ & \cite{Lindstedt:1985hm} \\
-   $f_{0}$ & $0.02$ & \cite{Lindstedt:1985hm}\\
-   $\zeta$ & $1.00$  & \cite{Lindstedt:2002td} \\
-   $u_{0}$ & $0.38$  & \cite{Lindstedt:2002td} \\
+   $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
+   $f_{0}$ & $0.02$ & \citep{Lindstedt:1985hm}\\
+   $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
+   $u_{0}$ & $0.38$  & \citep{Lindstedt:2002td} \\
 
@@
 First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of body fat reserves.
-Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \cite{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
+Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
 \begin{equation}

### 2017-06-12 Manuscript/NatureComm/CoverLetter/YKR_cover.tex
cover letter

\documentclass[ucm,12pt]{ucletter}

\usepackage{setspace}
\usepackage{graphicx}

\setlength{\parindent}{5ex}
\linespread{1.2}

%\name{Justin D. Yeakel}
\telephone{(209)~285-9571}
\email{jyeakel@ucmerced.edu}

\begin{document}
\begin{letter}{
    \\ Nature Communications\\
    Nature Publishing Group\\
    The Macmillan Building\\
    4 Crinan Street\\ 
    London N1 9XW\\
    United Kingdom\\ \\
    \vspace{1mm}
    \centerline{\bf{Re: The Dynamics of Starvation and Recovery}} \\
}

\opening{To the Editorial Board at \emph{Nature Communications},}

Please find attached the manuscript entitled "The dynamics of starvation and recovery" co-authored by Justin Yeakel, Christopher Kempes, and Sid Redner, which we would like to submit for publication in \emph{Nature Communications}.

A longstanding question in ecology concerns how individuals invest energy in somatic growth and maintenance vs. reproduction. Specific solutions to this problem contribute greatly to the the diversity of species’ life histories, however a  solid connection between somatic energetics and reproduction with an understanding of how these processes influence population dynamics (and vice versa) is lacking. In the accompanying manuscript we introduce a principled mathematical Nutritional State-structured Model (NSM), which describes how reproduction, starvation, and recovery from starvation, influence the dynamics of resource and consumer populations. Within this framework we incorporate allometric relationships for all timescales and rates, and in so doing connect the NSM to realistic mammalian populations. Our macro-scale model makes a number of predictions concerning the dynamic implications of energetic rate laws for consumer populations and helps elucidate important limits and potential drivers of body size evolution. We make several contributions:

\begin{itemize}
  \item Across the observed body size range, terrestrial mammals occupy a region of parameter space where sustained cyclic dynamics are not permitted. Relevant to observations of cycles in natural populations, the NSM predicts  that small mammal populations should be more prone to — whereas large mammals are buffered against — large transient oscillations, and this is a consequence of the timescales of starvation vs. recovery, both of which are governed by metabolism.
  \item Even transient cyclic dynamics increase the probability of extinction, and we argue that mammals should have energetic rates that avoid parameter regions close to transient oscillatory regimes. Moreover, regions of parameter space with small steady-state population sizes also increase the extinction probability. Thus we define a ``refuge'' of parameter space where the extinction probability is minimized between high probabilities of extinction via transience and high probabilities of extinction via low steady state densities, and show that mammals have allometrically constrained rates of starvation and recovery that place them squarely within this lower-risk space.
  \item Steady-state population sizes in the NSM support the energy equivalence hypothesis (equal energy use across populations of species with different individual body size) for a wide range of body sizes. Our model predicts that this energy equivalence breaks down at the smallest and largest body sizes observed for mammals, suggesting that there are strong dynamical constraints on mammalian body size. 
  \item We further investigate dynamical constraints on the range of mammalian body sizes by considering the invasion of a steady-state population by mammals with altered amounts of endogenous energetic reserves. We show that mammals with greater amounts of endogenous reserves are able to invade up to the size of the largest recorded mammal (ca. $10^7$g) above which leaner individuals can invade, such that this maximum body size bound serves as an evolutionary stable state. This suggests that the dynamics of starvation and recovery described in the NSM may provide a within-lineage mechanism for the evolution of larger body size among terrestrial mammals, a well-documented phenomenon known as Cope’s rule. This result is also supported by our finding that the steady state resource abundance is decreasing for larger mammals, which, in the context of classic resource competition theory, suggests that larger mammals will outcompete smaller ones given a shared resource. 
\end{itemize}

Our theory is analytic, relatively simple, and synthesizes a variety of fundamental ecological theories (energy equivalence hypothesis, resource competition theory, Cope’s rule, the fasting endurance hypothesis) and shows how many of these are the natural consequence of a simple dynamical framework that incorporates the energetic consequences of starvation and recovery combined with allometric timescales. We believe that our findings will be of general interest to ecologists, evolutionary biologists, and paleoecologists, such that consideration for publication in a broad interest journal such as PNAS is warranted.

We appreciate your consideration of our manuscript.

Suggested Referees:
\begin{itemize}
\item[] Alan Hastings, UC Davis, amhastings@ucdavis.edu
\item[] Felisa Smith, University of New Mexico, fasmith@unm.edu
\item[] Richard Sibly, University of Reading, r.m.sibly@reading.ac.uk
\item[] John DeLong, University of Nebraska, jpdelong@unl.edu
\item[] Seth Finnegan, Integrative Biology, UC Berkeley, sethf@berkeley.edu
\item[] Van Savage, University of California Los Angeles, vsavage@ucla.edu
\item[] Pablo Marquet, Department of Ecology, Pontificia Universidad Cat\`olica de Chile, pmarquet@bio.puc.cl
\end{itemize}

\vspace{5mm}

\singlespacing
\closing{Sincerely,\\
\fromsig{\includegraphics[scale=0.2]{signature.jpg}}\\
\fromname{
Justin D. Yeakel,\\
Christopher Kempes,\\
Sidney Redner}
}

\end{letter}
\end{document}

### 2017-06-12 Manuscript/NatureComm/YKR_NatureComm.tex
cover letter

@@
 
-\noindent {\bf Energetic barriers to body size}
+% \noindent {\bf Energetic barriers to body size}
 The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}a).
@@
 
-\noindent{\bf A within-lineage driver of Cope's Rule}
+% \noindent{\bf A within-lineage driver of Cope's Rule}
 The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.

### 2017-06-15 Manuscript/NatureComm/YKR_NatureComm.tex
final changes for naturecomm first submission

@@
+\RequirePackage{lineno}
 \documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
@@
 %\usepackage{rotating}
-%\usepackage{lineno}
+\usepackage{lineno}
 %\usepackage[left]{lineno}
@@
 
-\usepackage{bibunits}
+% \usepackage{bibunits}
 
-%\linenumbers
-%\setlength\linenumbersep{3pt}
+\linenumbers
+\setlength\linenumbersep{3pt}
 
@@
 
-\begin{bibunit}[unsrt]
+% \begin{bibunit}[unsrt]
 
@@
 %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-While there are no a priori constraints on the parameters in the NSM, most
-organisms correspond to restricted portions of the parameter space.  We
+While there are no {\it a priori} constraints on the parameters in the NSM, we expect that each species should be restricted to a distinct portion of the parameter space.  We
 use allometric scaling relations to constrain the covariation of rates in a
@@
 We derive relationships for the rates of reproduction, starvation, recovery, and mortality based on first principles, and as a function of an organism's body size and metabolic rate (see Methods).
-Because we aim to explore the starvation-recovery dynamics as a function of an organism's body mass $M$, we parameterize these rates in terms of the \emph{percent} gain and loss of the asymptotic (maximum) body mass, $\epsilon M$, where different values of $\epsilon$ defines different states of the consumer (Fig.~\ref{fig:growth}).
+Because we aim to explore the starvation-recovery dynamics as a function of an organism's body mass $M$, we parameterize these rates in terms of the \emph{percent} gain and loss of the asymptotic (maximum) body mass, $\epsilon M$, where different values of $\epsilon$ define different states of the consumer (Fig.~\ref{fig:growth}; see Methods for derivations of allometrically constrained rate equations).
 Although the rate equations \eqref{eq:system} are general and can in principle be used to explore the starvation recovery dynamics for most organisms, here we focus on allometric relationships for terrestrial-bound lower trophic level endotherms (see Supplementary Information for values), specifically herbivorous mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
@@
 We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig.
-\ref{fig:ext}a-c), suggesting that the population dynamics of larger organisms are more sensitive to smaller changes in physiological rates controlling starvation and recovery.
+\ref{fig:ext}a-c), suggesting that the population dynamics of larger organisms are more sensitive to variability in physiological rates controlling starvation and recovery.
+This finding is in accordance with, and may serve as contributing support for, observations of increased extinction risk among larger mammals \cite{Liow:2008jx}.
 % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
@@
 
-\putbib[aa_starving3]
+\bibliography{aa_starving3}
 
-\end{bibunit}
+% \end{bibunit}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-
-\clearpage
-
-\begin{bibunit}[unsrt]
-{\bf Supporting Information for ``The dynamics of starvation and recovery''}\\
-
-{\bf Mechanisms of Starvation and Recovery}
-Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
+% 
+% 
+% \clearpage
+% 
+% \begin{bibunit}[unsrt]
+% {\bf Supporting Information for ``The dynamics of starvation and recovery''}\\
+% 
+% {\bf Mechanisms of Starvation and Recovery}
+% Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
 For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
@@
 Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).
-
-We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
-\begin{align}
-\begin{split}
-\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
-\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
-& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
-\end{split}
-\end{align}
-where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
+% 
+% We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
+% \begin{align}
+% \begin{split}
+% \dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+% \dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
+% \dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
+% & \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
+% \end{split}
+% \end{align}
+% where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
 In the above equations $Y$ represents the yield coefficient (e.g., Refs.
@@
 In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
-%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
-
-%\begin{align}
-%\begin{split}
-%\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-%\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
-%\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
-%&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
-%\end{split}
-%\end{align}
+% %(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
 % 
-%In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
+% 
+% 
+% 
+% %\begin{align}
+% %\begin{split}
+% %\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+% %\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
+% %\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
+% %&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
+% %\end{split}
+% %\end{align}
+% %
+% %In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
 Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
 The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
-%A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
-%
-%Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
+% %A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
+% %
+% %Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
 As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
 Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
-
-%where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
-
-We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
-%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
-\begin{align}
-\begin{split}
-\dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
-\dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
-\dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
-\end{split}
-\end{align}
-If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
-\begin{align}
-\begin{split}
-\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
-\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
-\end{split}
-\end{align}
-where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$.
+% 
+% 
+% 
+% 
+% 
+% %where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
+% 
+% We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
+% %(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
+% \begin{align}
+% \begin{split}
+% \dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
+% \dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
+% \dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
+% \end{split}
+% \end{align}
+% If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
+% \begin{align}
+% \begin{split}
+% \dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
+% \dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
+% \dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
+% \end{split}
+% \end{align}
+% where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$.
 The above equations represent the system of equations presented in the main text.
-\\
-
-{\bf Parameter Values and Estimates}
-All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
+% \\
+% 
+% {\bf Parameter Values and Estimates}
+% All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
 Below we outline previous measurements and simple estimates of the parameters.
-
-Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs.
+% 
+% Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs.
 \citep{West:2001bv,moses2008rmo} and the supplement for Ref.
@@
 We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).
-
-An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction.
+% 
+% An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction.
 This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref.
@@
 We use $m_{0}=0.097M^{0.92}$ \citep{blueweiss1978relationships} which encompasses the widest range of body sizes \citep{peters1986ecological}.
-
-The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g.
+% 
+% The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g.
 Refs.
@@
 For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \citep{hou}.
-
-For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \citep{Dunbrack:1993ec}, $0.02M^{1.19}$ \citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ \citep{Lindstedt:2002td}.
+% 
+% For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \citep{Dunbrack:1993ec}, $0.02M^{1.19}$ \citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ \citep{Lindstedt:2002td}.
 We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range.
 Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \citep{Lindstedt:2002td}.
-
-%We also connect the resource growth rate to the total metabolic rate of an organism.
+% 
+% %We also connect the resource growth rate to the total metabolic rate of an organism.
 That is, we are interested in the relative rates of resource recovery and consumption by the total population.
@@
 This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.
-
-Typically the value of $\xi=C/k$ should roughly be $2$.
+% 
+% Typically the value of $\xi=C/k$ should roughly be $2$.
 The value of $\rho$, $\lambda$, $\sigma$, and $\mu$ are all simple rates (note that we have not rescaled time in our non-dimensionalization) as defined in the maintext.
@@
 Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
-\begin{equation}
-B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-Similarly, the maintenance for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
-\begin{equation}
-B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.
-
-Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities.
+% \begin{equation}
+% B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+% \end{equation}
+% Similarly, the maintenance for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
+% \begin{equation}
+% B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+% \end{equation}
+% Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.
+% 
+% Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities.
 From the data in Ref.
@@
 However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.
-\\
-%For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
-
-%More complicated models of fecundity (which, for example, account for the average length of adulthood and the number of individuals produced over this span) could be employed.
+% \\
+% %For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
+% 
+% %More complicated models of fecundity (which, for example, account for the average length of adulthood and the number of individuals produced over this span) could be employed.
 However, the scaling of population growth rate has been studied in detail before and follows a relationship of $$ which matches the theory well for $\phi=.95$ and $=$.
-
-%In our calculations we include $20\%$ variation around this value which could account for differences in efficiency during
-
- \begin{table}[h]
-\caption{Parameter values for mammals}
-\label{param}
-    \begin{center}
-    \small
-     \begin{tabular}{ p{1.2cm} p{3.2cm} l p{2.2cm}|}
-     \hline
-     Parameter & Value & References  \\
-     \hline
-   $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
-   $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
-   $E_{m}^{\prime}$ & $7000$  & \citep{stryer,hou} \\
-   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
-%   $a$ & $1.78\times10^{-6}$  \quad \quad \\
-%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
-   $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
-   $f_{0}$ & $0.02$ & \citep{Lindstedt:1985hm}\\
-   $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
-   $u_{0}$ & $0.38$  & \citep{Lindstedt:2002td} \\
-
-   \hline
-    \end{tabular}
-    \end{center}
-   \end{table}
-
-{\bf Rate equations for invaders with modified body mass}
-We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
-Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
-For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (see main text), implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
-Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
-
-Although the starved state of invading organisms remains unchanged, the rate of starvation from the modified full state to the starved state, the rate of recovery from the starved state to the modified full state, and the maintenance rates of both, will be different, such that $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$, $\delta^\prime = \delta(M^\prime)$.
-Rates of starvation and recovery for the invading population are easily derived by adjusting the starting or ending state before and after starvation and recovery, leading to the following timescales:
-
-\begin{align}
-t_{\sigma^\prime} &= -\frac{M^{1-\eta}}{a^{\prime}}\ln \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
-t_{\rho^\prime} &= \ln \left(\frac{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}{1-( \epsilon_\lambda(\chi +1))^{1/4}}\right)\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}.
-\end{align}
-
-The maintenance rates for the invading population require more careful consideration.
-First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of body fat reserves.
-Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
-\begin{equation}
-B^{\prime}_{\lambda}=\int_{0}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-where
-\begin{equation}
-t_{\lambda^\prime} = \frac{M^{1-\eta} }{a(1-\eta)}\ln \left(\frac{1-(m_0/M)^{1-\eta}}{1-(\epsilon_\lambda (1+\chi))^{1-\eta}} \right).
-\end{equation}
-Note that we do not use this timescale to determine the reproductive rate of the invading consumer---which is assumed to remain the same as the resident population---but only to calulate the lifetime energy use.
-Similarly, the maintenance for hungry individuals $P^\prime_{H}=B_{0}(\epsilon_{\sigma}(1+\chi)M)^{3/4}/(\epsilon_{\sigma}(1+\chi)M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y^\prime_{H}=(1+\chi)ME_{d}/B^{\prime}_{\rho}$ where
-\begin{equation}
-B^{\prime}_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-Finally, we can calculate the maintenance of the invaders as
-
-\begin{align}
-  \delta^\prime &= P^\prime_{H}Y^\prime_{H}/\xi \\ \nonumber
-  \beta^\prime &= \left(\frac{\lambda_{\rm max}}{Y^\prime_{F}}+P^\prime_{F} \right)Y^\prime_{H}/\xi.
-\end{align}
-
-To determine whether or not the invader or resident population has an advantage, we compute $R^*(M)$ and $R^*(M^\prime=M(1+\chi))$ for values of $\chi \in (-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$, and the invading population is assumed to have an advantage over the resident population if $R^*(M^\prime)<R^*(M)$.
-
-\putbib[aa_starving_supplement]
-
-\end{bibunit}
+% 
+% %In our calculations we include $20\%$ variation around this value which could account for differences in efficiency during
+% 
+%  \begin{table}[h]
+% \caption{Parameter values for mammals}
+% \label{param}
+%     \begin{center}
+%     \small
+%      \begin{tabular}{ p{1.2cm} p{3.2cm} l p{2.2cm}|}
+%      \hline
+%      Parameter & Value & References  \\
+%      \hline
+%    $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
+%    $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
+%    $E_{m}^{\prime}$ & $7000$  & \citep{stryer,hou} \\
+%    $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
+% %   $a$ & $1.78\times10^{-6}$  \quad \quad \\
+% %   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
+%    $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
+%    $f_{0}$ & $0.02$ & \citep{Lindstedt:1985hm}\\
+%    $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
+%    $u_{0}$ & $0.38$  & \citep{Lindstedt:2002td} \\
+% 
+%    \hline
+%     \end{tabular}
+%     \end{center}
+%    \end{table}
+% 
+% 
+% {\bf Rate equations for invaders with modified body mass}
+% We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
+% Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
+% For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (see main text), implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
+% Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
+% 
+% Although the starved state of invading organisms remains unchanged, the rate of starvation from the modified full state to the starved state, the rate of recovery from the starved state to the modified full state, and the maintenance rates of both, will be different, such that $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$, $\delta^\prime = \delta(M^\prime)$.
+% Rates of starvation and recovery for the invading population are easily derived by adjusting the starting or ending state before and after starvation and recovery, leading to the following timescales:
+% 
+% \begin{align}
+% t_{\sigma^\prime} &= -\frac{M^{1-\eta}}{a^{\prime}}\ln \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
+% t_{\rho^\prime} &= \ln \left(\frac{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}{1-( \epsilon_\lambda(\chi +1))^{1/4}}\right)\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}.
+% \end{align}
+% 
+% 
+% The maintenance rates for the invading population require more careful consideration.
+% First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of body fat reserves.
+% Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
+% \begin{equation}
+% B^{\prime}_{\lambda}=\int_{0}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
+% \end{equation}
+% where
+% \begin{equation}
+% t_{\lambda^\prime} = \frac{M^{1-\eta} }{a(1-\eta)}\ln \left(\frac{1-(m_0/M)^{1-\eta}}{1-(\epsilon_\lambda (1+\chi))^{1-\eta}} \right).
+% \end{equation}
+% Note that we do not use this timescale to determine the reproductive rate of the invading consumer---which is assumed to remain the same as the resident population---but only to calulate the lifetime energy use.
+% Similarly, the maintenance for hungry individuals $P^\prime_{H}=B_{0}(\epsilon_{\sigma}(1+\chi)M)^{3/4}/(\epsilon_{\sigma}(1+\chi)M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y^\prime_{H}=(1+\chi)ME_{d}/B^{\prime}_{\rho}$ where
+% \begin{equation}
+% B^{\prime}_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
+% \end{equation}
+% Finally, we can calculate the maintenance of the invaders as
+% 
+% \begin{align}
+%   \delta^\prime &= P^\prime_{H}Y^\prime_{H}/\xi \\ \nonumber
+%   \beta^\prime &= \left(\frac{\lambda_{\rm max}}{Y^\prime_{F}}+P^\prime_{F} \right)Y^\prime_{H}/\xi.
+% \end{align}
+% 
+% To determine whether or not the invader or resident population has an advantage, we compute $R^*(M)$ and $R^*(M^\prime=M(1+\chi))$ for values of $\chi \in (-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$, and the invading population is assumed to have an advantage over the resident population if $R^*(M^\prime)<R^*(M)$.
+% 
+% \putbib[aa_starving_supplement]
+% 
+% 
+% \end{bibunit}
+% 
+%
 

### 2017-10-18 Manuscript/NatureComm/ReviewResponses.tex
new revision

\documentclass[ucm,12pt]{ucletter}

\setlength\parindent{1cm}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage{xcolor}

%\name{Justin D. Yeakel}
\telephone{(209)~285-9571}
\email{jyeakel@ucmerced.edu}

\newcounter{section}
\newcounter{subsection}[section]
\setcounter{secnumdepth}{3}
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
                                   {-3.5ex \@plus -1ex \@minus -.2ex}%
                                   {2.3ex \@plus.2ex}%
                                   {\normalfont\Large\bfseries}}
\newcommand\subsection{\@startsection{subsection}{2}{\z@}%
                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\normalfont\large\bfseries}}
\renewcommand \thesection{\@arabic\c@section}
\renewcommand\thesubsection{\thesection.\@arabic\c@subsection}

\makeatother

\begin{document}
\begin{letter}{
    Philosophical Transactions of the Royal Society B\\
    Publishing Section\\
    The Royal Society\\
    6-9 Carlton House Terrace\\
    London SW1Y 5AG\\
    England\\
    \centerline{\bf{Re: Eco-evolutionary dynamics and collective dispersal:}}\\
     \centerline{\bf{implications for salmon metapopulation robustness}}
}

\opening{To the editorial committee at Nature Communications,}

We thank both the Reviewers for their insightful comments and criticisms. Below, we address each one, point by point. We believe that the edits that we have made have increased the clarity of our contribution, in no small part due to the thought and detail exercised by the Reviewers. We hope that our responses satisfy the concerns listed below. Please do not hesitate to contact us if there are additional clarifications required.

\setstretch{1.2}

\section*{Reviewer I}
\subsection*{Major comments:}
\noindent \textcolor{cyan}{
{\bf Comment 1.1} stuff
}

\noindent {\bf Response 1.1} Stuff
\noindent To clarify, we have made the following edits to the text:\\
\noindent {\bf Line 111:} \emph{The partial derivative in Eq. 4 determines how the mean trait changes through time due to natural selection [ref], which is proportional to the change in mean fitness with respect to $\mu_i$.
We note that the derivative is dependent on both $\mu_i$ and $\mu_j$ due to the influence of the logarithm on the sum.}

\noindent \textcolor{cyan}{
{\bf Comment 1.2} Stuff
}

\noindent {\bf Response 1.2} Stuff

\vspace{5mm}

\singlespacing
\closing{Sincerely,\\
\fromsig{\includegraphics[scale=0.2]{signature.jpg}}\\
\fromname{
Justin D. Yeakel\\
Christopher P. Kempes\\
Sidney Redner}
}

\end{letter}
\end{document}

### 2017-10-18 NatureComm_REVISION/CoverLetter/YKR_cover.tex
new revision

--- 2017-06-12 Manuscript/NatureComm/CoverLetter/YKR_cover.tex
@@
   \item Across the observed body size range, terrestrial mammals occupy a region of parameter space where sustained cyclic dynamics are not permitted.
-Relevant to observations of cycles in natural populations, the NSM predicts  that small mammal populations should be more prone to — whereas large mammals are buffered against — large transient oscillations, and this is a consequence of the timescales of starvation vs.
-recovery, both of which are governed by metabolism.
   \item Even transient cyclic dynamics increase the probability of extinction, and we argue that mammals should have energetic rates that avoid parameter regions close to transient oscillatory regimes.
@@
 Thus we define a ``refuge'' of parameter space where the extinction probability is minimized between high probabilities of extinction via transience and high probabilities of extinction via low steady state densities, and show that mammals have allometrically constrained rates of starvation and recovery that place them squarely within this lower-risk space.
+  \item We show that this lower-risk space becomes smaller for mammals with larger body size, thus providing a mechanistic explanation for the slightly higher extinction risk observed among larger mammals over evolutionary time.
   \item Steady-state population sizes in the NSM support the energy equivalence hypothesis (equal energy use across populations of species with different individual body size) for a wide range of body sizes.
@@
   \item We further investigate dynamical constraints on the range of mammalian body sizes by considering the invasion of a steady-state population by mammals with altered amounts of endogenous energetic reserves.
-We show that mammals with greater amounts of endogenous reserves are able to invade up to the size of the largest recorded mammal (ca.
-$10^7$g) above which leaner individuals can invade, such that this maximum body size bound serves as an evolutionary stable state.
-This suggests that the dynamics of starvation and recovery described in the NSM may provide a within-lineage mechanism for the evolution of larger body size among terrestrial mammals, a well-documented phenomenon known as Cope’s rule.
-This result is also supported by our finding that the steady state resource abundance is decreasing for larger mammals, which, in the context of classic resource competition theory, suggests that larger mammals will outcompete smaller ones given a shared resource.
+We show that mammals with greater amounts of endogenous reserves are able to invade up to the size of the largest recorded mammal (predicted $1.748\times10^7$g compared to an observed $1.74\times10^7$g estimated for \emph{Deinotherium}) above which leaner individuals can invade.
+This suggests that the dynamics of starvation and recovery described in the NSM may provide a within-lineage mechanism for the evolution of larger body size among terrestrial mammals, a well-documented phenomenon known as Cope's rule.
 \end{itemize}
 
-Our theory is analytic, relatively simple, and synthesizes a variety of fundamental ecological theories (energy equivalence hypothesis, resource competition theory, Cope’s rule, the fasting endurance hypothesis) and shows how many of these are the natural consequence of a simple dynamical framework that incorporates the energetic consequences of starvation and recovery combined with allometric timescales.
-We believe that our findings will be of general interest to ecologists, evolutionary biologists, and paleoecologists, such that consideration for publication in a broad interest journal such as PNAS is warranted.
+Our theory is analytic, relatively simple, and synthesizes a variety of fundamental ecological theories (energy equivalence hypothesis, resource competition theory, Cope's rule, the fasting endurance hypothesis) and shows how many of these are the natural consequence of a simple dynamical framework that incorporates the energetic consequences of starvation and recovery combined with allometric timescales.
+We believe that our findings will be of general interest to ecologists, evolutionary biologists, and paleoecologists, such that consideration for publication in a broad interest journal such as \emph{Nature Communications} is warranted.
 
@@
 \begin{itemize}
-\item[] Alan Hastings, UC Davis, amhastings@ucdavis.edu
 \item[] Felisa Smith, University of New Mexico, fasmith@unm.edu

### 2017-10-18 NatureComm_REVISION/YKR_NatureComm-Revision-1.tex
new revision

--- 2017-06-15 Manuscript/NatureComm/YKR_NatureComm.tex
@@
 The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of its constituent individuals. 
-Of particular importance are the tradeoffs between reproduction and the dynamics of starvation and recovery. % in resource-limited environments
+Of particular importance is the interplay between reproduction and the dynamics of starvation and recovery. % in resource-limited environments
 We introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete, reproducing consumers, and undernourished, non-reproducing consumers.
@@
 % Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations. 
-We also show that our model provides a natural framework that predicts maximum body size for mammals by determining the relative stability of an otherwise homogeneous population to a mutant population with altered percent body fat, providing a principled mechanism for a within-lineage driver of Cope's rule.
+We also show that our model provides a natural framework that predicts
+maximum body size for mammals by determining the relative stability of an
+otherwise homogeneous population to a competing population with altered percent
+body fat, providing a principled mechanism for a within-lineage driver of
+Cope's rule.
 % For body masses $< 1.748\times10^7$g, individuals with increased energetic reserves can invade resident populations, and vice versa for body mass $> 1.748\times10^7$g, thus providing a principled mechanism for a within-lineage driver of Cope's rule.
@@
 Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\citep{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\citep{Kirk:1997cc}.
-Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate nutritional reserves~\citep{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
+Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate reserves~\citep{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
 
 Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.
-The data collected thus far has shown that diverse mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development (blastocyst implantation) until nutritional reserves can be accumulated~\citep{Mead:1989dt,Sandell:1990kw}.
+Many mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development until nutritional reserves can be accumulated~\citep{Mead:1989dt,Sandell:1990kw}.
 Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\citep{Bulik:1999eo,Trites:2003cc}.
 In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \citep{Glazier:2009hq}.
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the importance and universality of the fundamental tradeoff between somatic and reproductive investment.
-However the general dynamic implications of these constraints are unknown.
+The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the universality of the fundamental tradeoff between somatic and reproductive investment.
 
-Though straightforward conceptually, incorporating the energetic dynamics of individuals~\citep{Kooi2000} into a population-level framework~\citep{Kooi2000,Sousa:2010ez} presents numerous mathematical obstacles~\citep{Diekmann:2010da}.
-An alternative approach involves modeling the macroscale relations that guide somatic versus reproductive investment in a consumer-resource system.
-For example, macroscale Lotka-Volterra models assume that the growth rate of the consumer population depends on resource density, thus \emph{implicitly} incorporating the requirement of resource availability for reproduction~\citep{murdoch:2003}.
+Including individual energetic dynamics~\citep{Kooi2000} in a population-level framework~\citep{Kooi2000,Sousa:2010ez} is challenging~\citep{Diekmann:2010da}, and
+a common simplifying approach is provided in the classic Lotka-Volterra (LV) model, which assumes that consumer population growth rate depends linearly on resource
+  density~\citep{murdoch:2003}. Here, we introduce an alternative
+  approach---the Nutritional State-structured Model (NSM)---that accounts for
+  resource limitation through the consequences of starvation. In contrast to
+  the LV model, the NSM incorporates two consumer states:
+  hungry and full, with only the former susceptible to mortality and only the
+  latter possessing sufficient energetic reserves to reproduce.
+  Additionally, we incorporate allometrically derived constraints on reproduction \citep{Kempes:2012hy}, incorporating the timescales of starvation and recovery. 
+  As we shall show, our model makes several important predictions: 
+  (i) the dynamics are typically driven to a refuge far from cyclic behavior and extinction risk, 
+  (ii) steady state conditions of the NSM accurately predict measured biomass densities for mammals as well as Damuth's energetic equivalence principle, 
+  (iii) an allometrically constrained upper-bound for mammalian body size, and
+  (iv) a selective mechanism for the evolution of larger body size, known as Cope's rule.
 
-In this work, we adopt an alternative approach in which we \emph{explicitly} account for resource limitation and the subsequent effects of starvation.
-Namely, only individuals with sufficient energetic reserves can reproduce.
-Such a constraint leads to reproductive time lags due to some members of the population going hungry and then recovering.
-Additionally, we incorporate the idea that reproduction is strongly constrained allometrically \citep{Kempes:2012hy}, and is not generally linearly related to resource density.
-As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.\\
+% Though straightforward conceptually, incorporating the energetic dynamics of
+% individuals~\citep{Kooi2000} into a population-level
+% framework~\citep{Kooi2000,Sousa:2010ez} presents mathematical
+% obstacles~\citep{Diekmann:2010da}.  An alternative approach involves modeling
+% the macroscale relations that guide somatic versus reproductive investment in
+% a consumer-resource system.  {\color{red}For example, macroscale
+%   Lotka-Volterra models assume that the growth rate of the consumer
+%   population depends on resource density, thus incorporating the requirement
+%   of resource availability for reproduction~\citep{murdoch:2003}.
 
-\noindent {\bf Nutritional state-structured model (NSM)}\\
-We begin by defining a minimal Nutritional State-structured population Model (NSM), where the consumer population is partitioned into two states: (a) an energetically replete (full) state $F$, where the consumer reproduces at a constant rate $\lambda$ and does not die from starvation, and (b) an energetically deficient (hungry) state $H$, where the consumer does not reproduce but dies by starvation at rate $\mu$.
-The underlying resource $R$ evolves by logistic growth with an intrinsic growth rate $\alpha$ and a carrying capacity $C$.
-The rate at which consumers transition between states and consume resources is dependent on their overall abundance, the abundance of resources, the efficiency of converting resources into metabolism, and how that metabolism is partitioned between maintenance and growth purposes.
-In the Supplementary Information we provide a fully mechanistic model for each of these dynamics and constants, and show that the system produces a simple non-dimensional form which we describe below.
+% In this work, we adopt an alternative approach to account for resource
+% limitation through the consequences of starvation.  Namely, in contrast to
+% the Lotka-Volterra model, we incorporate two states: hungry individuals and
+% individuals with sufficient energetic reserves to reproduce.}  Such a
+% constraint leads to reproductive time lags due to some members of the
+% population going hungry and then recovering.  Additionally, we incorporate
+% the idea that reproduction is strongly constrained allometrically
+% \citep{Kempes:2012hy}, and is not generally linearly related to resource
+% density.
+% As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.\\
+
+\noindent \paragraph*{{\bf Nutritional state-structured model (NSM).}}
+We begin by defining a minimal Nutritional State-structured population Model
+(NSM), where the consumer population is partitioned into two states: (a) an
+energetically replete (full) state $F$, where the consumer reproduces at a
+constant rate $\lambda$ and does not die from starvation, and (b) an
+energetically deficient (hungry) state $H$, where the consumer does not
+reproduce but dies by starvation at rate $\mu$. The dynamics of the
+underlying resource $R$ is governed by logistic growth with an intrinsic
+growth rate $\alpha$ and a carrying capacity $C$. The rate at which consumers
+transition between states and consume resources is dependent on their number,
+the abundance of resources, the efficiency of converting resources into
+metabolism, and how that metabolism is partitioned between maintenance and
+growth purposes.  We provide a physiologically and energetically mechanistic
+model for each of these dynamics and constants (see the Supplementary Information
+(SI)), and show that the system produces a simple non-dimensional form which
+we describe below.
 
@@
 
-Consumers transition from the full state $F$ to the hungry state $H$ at a rate $\sigma$---the starvation rate---and also in proportion to the absence of resources $(1-R)$.
-Conversely, consumers recover from state $H$ to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a ratio between maximal resource consumption and the carrying capacity of the resource.
-%and $\approx 2$.
-Resources are eaten by the hungry consumers at rate $\rho R + \delta$, that accounts for their somatic growth ($\rho R$) and  maintenance ($\delta$).
-Full consumers eat resources at a constant rate $\beta$ that accounts for maximal maintenance and somatic growth (see Supplementary Information for mechanistic derivations of these rates from resource energetics).
+Consumers transition from the full state $F$ to the hungry state $H$ at a
+rate $\sigma$---the starvation rate---and also in proportion to the absence
+of resources $(1-R)$ (we assume a maximum density, the carrying capacity equal to 1).  Conversely, consumers recover from state $H$ to state
+$F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a
+ratio between maximal resource consumption and the carrying capacity of the
+resource. %and $\approx 2$.
+The resources that are eaten by hungry consumers (at rate $\rho R + \delta$)
+account for their somatic growth ($\rho R$) and maintenance ($\delta$).  Full
+consumers eat resources at a constant rate $\beta$ that accounts for maximal
+maintenance and somatic growth (see the SI for mechanistic
+derivations of these rates from resource energetics).
 %This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
@@
 
-In the mean-field approximation, in which the consumers and resources are perfectly mixed, their densities evolve according to the rate equations
+In the mean-field approximation, in which the consumers and resources are
+perfectly mixed, their densities are governed by the rate equations
 
@@
 
-This system of nondimensional equations follows from a set of first-principle relationships for resource consumption and growth (see Supplementary Information for a full derivation and the dimensional form).
+This system of nondimensional equations follows from a set of first-principle
+relationships for resource consumption and growth (see the SI for a full derivation and the dimensional form).
 Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
-This resembles the equation of motion for the predator density in the classic Lotka-Volterra model~\citep{murray2011mathematical}, except that the resource density does not appear in the growth term.
+This resembles the equation of motion for the predator density in the LV model~\citep{murray2011mathematical}, except that the resource density does not appear in the growth term.
 % As discussed above, the attributes of reproduction and mortality have been explicitly apportioned to the full and hungry consumers, respectively, so that the growth in the total density is decoupled from the resource density.
-The rate of reproduction is independent of resource density because it is assumed that the satiated state of the full consumer allows it to partition a constant amount of energy towards reproduction, whereas a starved consumer partitions no energy towards reproduction.
-The rate of reproduction for the total consumer density \emph{is} dependent on resource density, which determines the size of the full and starved portions of the consumer population.
-Similarly, the consumer maintenance terms ($\delta H$ and $\beta F$) are independent of resource density because they represent a minimal energetic requirement for consumers in the $H$ and $F$ state, respectively.
-It follows that model predictions are robust only when $R \gg 0$, which holds for all cases that we explore.
+The rate of reproduction is independent of resource density because it is
+assumed that the satiated state of the full consumer allows it to partition a
+constant amount of energy towards reproduction, whereas a starved consumer
+partitions no energy towards reproduction.  The rate of reproduction for the
+total consumer density \emph{is} dependent on resource density, which
+determines the size of the full and starved portions of the consumer
+population.  Similarly, the consumer maintenance terms ($\delta H$ and
+$\beta F$) are independent of resource density because they represent a
+minimal energetic requirement for consumers in the $H$ and $F$ state,
+respectively.  It follows that model predictions are robust only when $R$ is
+of the order of 1, which holds for all cases that we explore.
 
-From the solution to the single internal fixed point (Eq.~\eqref{eq:ss}, see Methods), an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that the consumer and resource densities are positive.
+\noindent \paragraph*{ {\bf Steady states of the NSM.}} From the solution to the single internal fixed point (Eq.~\eqref{eq:ss}, see Methods), an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that the consumer and resource densities are positive.
 %In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
@@
 If the starvation rate is low enough that the Hopf bifurcation is crossed, these oscillations become stable over time.
-This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.
+This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.\\
 
@@
 %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-While there are no {\it a priori} constraints on the parameters in the NSM, we expect that each species should be restricted to a distinct portion of the parameter space.  We
+\noindent {\bf Results \& Discussion}
+\noindent \paragraph*{{\bf Allometry minimizes extinction risk.}} While there are no {\it a priori} constraints on the parameters in the NSM, we expect that each species should be restricted to a distinct portion of the parameter space.
+We
 use allometric scaling relations to constrain the covariation of rates in a
@@
 Because we aim to explore the starvation-recovery dynamics as a function of an organism's body mass $M$, we parameterize these rates in terms of the \emph{percent} gain and loss of the asymptotic (maximum) body mass, $\epsilon M$, where different values of $\epsilon$ define different states of the consumer (Fig.~\ref{fig:growth}; see Methods for derivations of allometrically constrained rate equations).
-Although the rate equations \eqref{eq:system} are general and can in principle be used to explore the starvation recovery dynamics for most organisms, here we focus on allometric relationships for terrestrial-bound lower trophic level endotherms (see Supplementary Information for values), specifically herbivorous mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
+Although the rate equations \eqref{eq:system} are general and can in
+principle be used to explore the starvation recovery dynamics for most
+organisms, here we focus on allometric relationships for terrestrial-bound
+lower trophic level endotherms (see the SI for values), specifically herbivorous mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
 Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
-Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.\\
+Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
 
@@
 %Extinction to the left and the right
-\noindent {\bf Extinction risk}\\
+% \noindent \paragraph*{{\bf Extinction risk}} 
 As the allometric derivations of the NSM rate laws reveal (see Methods), starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
@@
 
+% \noindent \paragraph*{{\bf Extinction refuge.}} 
 We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge across a range of $M$ (Fig. \ref{fig:ext}a-c, white points).
@@
 This finding is in accordance with, and may serve as contributing support for, observations of increased extinction risk among larger mammals \cite{Liow:2008jx}.
+Moreover, larger body sizes decrease the steady state resource density, such that fluctuations for larger organisms will be more likely to drive resources to extinction.
 % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
@@
 
-\noindent {\bf Discussion}\\
-Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\citep{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\citep{West:2002ud}.
-Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
-One important such example is the \emph{fasting endurance hypothesis}, which contends that larger body size, with consequent lower metabolic rates and increased ability to maintain more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability~\citep{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size (known as Cope's rule)~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ grams~\citep{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\citep{Clauset:2009fh}.
-These trends are thought to be driven by a combination of climate change and niche availability~\citep{Saarinen:2014br}; however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, have not been explored.
-We argue that the NSM provides a suitable framework to explore these issues.
-
 % \noindent {\bf Energetic barriers to body size}
-The NSM correctly predicts that species with smaller masses have larger steady-state population densities (Fig.~\ref{fig:mass}a).
-Similar predictions have been made for carnivore populations using alternative consumer-resource models \citep{DeLong:2012kw}.
-Moreover, we show that the NSM provides independent theoretical support for the energy equivalence hypothesis and Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998}.
-The energy equivalence hypothesis argues that the total energy use, $B_{\rm tot}$, of a population is constant independent of species size~\citep{Damuth:1987kr,allen2002,enquist1998}.
-%Previous studies have focused on t
-This hypothesis is based on observations showing that the steady state abundance, $N^*$, of a species is proportional to the inverse of individual metabolism, such that $N^*\propto M^{-3/4}/B_{0}$~\citep{allen2002,enquist1998}.
-This relationship implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been shown to hold in both mammalian and vascular plant communities \citep{Damuth:1987kr,allen2002,enquist1998}.
-Figure \ref{fig:mass}a shows that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of organism sizes and Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly constant over this same range.
-This result is remarkable because it illustrates that the steady state values of the NSM combined with the derived timescales naturally give rise to energy equivalence.
+\noindent \paragraph*{{\bf Predicting energetic equivalence and body size limits.}} The NSM correctly predicts that species with smaller masses have larger
+steady-state population densities (Fig.~\ref{fig:mass}a).  Similar
+predictions have been made for carnivore populations using alternative
+consumer-resource models \citep{DeLong:2012kw}.  Moreover, we show that the
+NSM provides independent theoretical support for the energy equivalence
+hypothesis and Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998}.  The
+energy equivalence hypothesis argues that the total energy use,
+$B_{\rm tot}$, of a population is constant independent of species
+size~\citep{Damuth:1987kr,allen2002,enquist1998}. %Previous studies have focused on t
+This hypothesis is based on observations showing that the steady state
+abundance, $N^*$, of a species is proportional to the inverse of individual
+metabolism, such that
+$N^*\propto M^{-3/4}/B_{0}$~\citep{allen2002,enquist1998}.  This relationship
+implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been
+shown to hold in both mammalian and vascular plant communities
+\citep{Damuth:1987kr,allen2002,enquist1998}.  Figure \ref{fig:mass}a shows
+that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of
+organism sizes and Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly
+constant over this same range.  This result is remarkable because it
+illustrates that the steady state values of the NSM combined with the derived
+timescales naturally give rise to energy equivalence.
 
@@
   ({\bf b}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  The data are from Damuth \citep{Damuth:1987kr} and have been converted to total population metabolism using the allometric relationships for metabolic rate (please see Supplementary Information and Refs.
-\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:mass}}
+  The data are from Damuth \citep{Damuth:1987kr} and have been converted to
+  total population metabolism using the allometric relationships for
+  metabolic rate (see the SI and Refs.~\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:mass}}
 \end{figure}
 
-Our model shows that the equivalence breaks down at large $M$ suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
-%, violating the energy equivalence.
-% set by deviations from constant efficiency obeyed by other populations.
-In the framework of our model, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occur at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
+% \paragraph*{{\bf Maximum Body Size}}
+Our model shows that energetic equivalence breaks down at large $M$ suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.  
+%, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
+With respect to predicted steady state densities, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occurs at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
 This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero, causing the timescales to become infinite and the rates $\mu$ and $\lambda$ to equal zero.
-The $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$.
-This point predicts a strong upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
+A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120$ times the mass of a male African elephant.
+In contrast, the $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$. 
+This point predicts a more realistic upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
 Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is within an order of magnitude of the maximum body size observed in the North American mammalian fossil record~\citep{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\citep{Clauset:2009fh}.
@@
 % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
-\\
 
@@
 
+\noindent \paragraph*{{\bf A mechanism for Cope's rule}} Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\citep{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\citep{West:2002ud}.
+Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
+One important such example is the \emph{fasting endurance hypothesis}, which contends that larger body size, with consequent lower metabolic rates and increased ability to maintain more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability~\citep{Millar:1990p923}.
+Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size known as Cope's rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ grams~\citep{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\citep{Clauset:2009fh}.
+These trends are thought to be driven by a combination of climate change and niche availability~\citep{Saarinen:2014br}; however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, have not been explored.
+% We argue that the NSM provides a suitable framework to explore these issues.
+
 % \noindent{\bf A within-lineage driver of Cope's Rule}
+% \paragraph*{{\bf Body Size Competition}} 
 The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
 \ref{fig:mass}a, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \citep{tilman1981,dutkiewicz2009,barton2010}.
-Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but these absolute limits do not offer a within-lineage mechanism by which larger body sizes are selected for or against.
-We will now show that the NSM indeed provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time known as Cope's Rule.
+Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but these absolute limits do not offer a mechanism by which larger body sizes are selected for or against.
+% We will now show that the NSM indeed provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time known as Cope's Rule.
 
@@
 \includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
-\caption{\small{ Invasion feasibility for organisms with a proportional change in
-  mass $\chi$ against a population with a resident body mass $M$.  The blue
-  region denotes proportions of modified mass $\chi$ resulting in successful invasion.  The
-  red region denotes values of $\chi$ that result in a mass that is below the
-  starvation threshold and are thus infeasible.
+\caption{\small{Competitive outcomes for a resident species with body mass $M$ vs. a closely related competing species with modified body mass $M^\prime=M(1+\chi)$.  
+The blue region denotes proportions of modified mass $\chi$ resulting in exclusion of the resident species. 
+The red region denotes values of $\chi$ that result in a mass that is below the starvation threshold and are thus infeasible. 
 Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
@@
 
-To examine whether the NSM could provide such a mechanism, we begin by noting that a theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120$ times the mass of a male African elephant.
-Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
+% Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
 
-We directly assess the susceptibility of an otherwise homogeneous population to invasion by a mutated subset of the population (denoted by $^\prime$) where individuals have a modified proportion of body fat $M^\prime=M(1+\chi)$.
-If $\chi < 0$, individuals within the invading mutant population have fewer metabolic reserves, and if $\chi>0$, individuals have more metabolic reserves than the resident population.
+We directly assess competitive outcome between two closely related species: a resident species of mass $M$, and a competing species (denoted by $^\prime$) where individuals have a different proportion of body fat such that $M^\prime=M(1+\chi)$.
+If $\chi < 0$, individuals within the competing population have fewer metabolic reserves, and if $\chi>0$, individuals have more metabolic reserves than the resident species.
 For the allowable values of $\chi$ the adjusted mass should exceed the minimal amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given Equation \ref{t1}, implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
 Together these conditions imply that $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$ and the lower bound is mass-dependent.
-The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$.
-Importantly, $\epsilon_\sigma$, which determines the point along the growth curve that defines the body composition of starved foragers, is assumed to remain unchanged for the invader population (see Supplementary Information for detailed derivations of invader rates).
+The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$ (see the SI for detailed derivations of competitor rates).
+Importantly, $\epsilon_\sigma$, which determines the point along the growth
+curve that defines the body composition of starved foragers, is assumed to
+remain unchanged for the competing population.
 %There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
 
-To assess the susceptibility of the resident population to invasion, we determine which consumer pushes the steady-state resource density to lower values for a given value of $\chi$, with the expectation that populations able to survive on lower resource densities have a competitive advantage \citep{tilman1981}.
-We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the invader has an intrinsic competitive advantage over the resident population (Fig.
+To assess the susceptibility of the resident species to competitive exclusion, we determine which consumer pushes the steady-state resource density $R^*$ to lower values for a given value of $\chi$, with the expectation that a population capable of surviving on lower resource densities has a competitive advantage \citep{tilman1981}.
+We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the competitor has an intrinsic advantage over the resident species (Fig.
 \ref{fig:invasion}).
@@
 $1.74\times10^7$g ~\citep{Smith:2010p3442}.
-Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \citep{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals outlined by Alroy~\citep{Alroy:1998p1594}.
+Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \citep{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals explored by Alroy~\citep{Alroy:1998p1594}.
 %as well as  \citep{Alroy:1998p1594,Clauset:2009fh}.
 %This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
-While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general within-lineage mechanism for the evolution of larger body size among terrestrial mammals.
+While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general selective mechanism for the evolution of larger body size among terrestrial mammals.\\
+
+\noindent {\bf Conclusion} \\
+{\color{red} Complementary previous work has analyzed the tradeoffs associated with carnivory in terms of the energy expenditure on hunting and prey size, where a maximum carnivore size between 700 and 1,000 is predicted from considerations of energy intake going to zero \cite{Carbone:1999ju}.
+Similarly, it has been shown in the past that the largest body size within a particular lineage should be a function of the metabolic normalization constant and a critical death rate as supported by data.
+This approach, which consider the energetics associated with active metabolism, provides a complementary prediction for maximum size from energetics \cite{Okie:2013ju}, and also suggests that future connections between the scaling of fat and muscle mass should be systematically connected with $B_{0}$.
+}
 
@@
 The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations~\citep{Stearns:1989ip}.
-The NSM is a minimal and general model that incorporates the dynamics of starvation and recovery that are expected to occur in resource-limited environments.
-By incorporating allometric relations between the rates in the NSM, we found: (i) different organismal masses have distinct population dynamic regimes, (ii) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, and (iii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of maximum body size.
+The NSM is a general model that incorporates the dynamics of starvation and recovery that are expected to occur in resource-limited environments.
+By incorporating allometric relations between the rates in the NSM, we found: (i) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, (ii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of maximum body size, and (iii) a selective mechanism for the evolution of larger body sizes known as Cope's rule.
 We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.
@@
 \end{eqnarray}
-where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass, $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically (see Supplementary Information).
-We now use this solution to define the timescale for reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see \citep{moses2008rmo} for a detailed presentation of these timescales).
-The time that it takes to reach a particular mass $\epsilon M$ is given by the timescale
+where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass,
+$a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically
+(see the SI).  We now use this solution to define the timescale for
+reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see
+\citep{moses2008rmo} for a detailed presentation of these timescales). The
+time that it takes to reach a particular mass $\epsilon M$ is given by the
+timescale
 \begin{equation}
@@
 \end{equation}
-where we will define values of $\epsilon$ to describe a set of rates within our model.
-For the time to reproduce, $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$ is the fraction of the asymptotic mass where an organism is reproductively mature and should be close to one (typically $\epsilon_{\lambda}\approx0.95$; \citep{West:2001bv}).
-The growth rate is then given by $\lambda=\ln\left(\upsilon\right)/t_{\lambda}$ where $\upsilon$ is the number of offspring produced, and for any constant value of $\epsilon_{\lambda}$, this rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
+where we will define values of $\epsilon$ to describe a set of rates within
+our model.  The time to reproduce is given by
+$t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
+is the fraction of the asymptotic mass where an organism is reproductively
+mature and should be close to one (typically
+$\epsilon_{\lambda}\approx0.95\;$ \citep{West:2001bv}). The growth rate then
+is $\lambda=\ln\left(\nu\right)/t_{\lambda}$ where $\nu$ is the number of
+offspring produced, and for any constant value of $\epsilon_{\lambda}$, this
+rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$
+\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
 
@@
 and recovery by the timescales required to reach, or return from, specific
-fractions of the replete-state mass (see Supplementary Information, Table I for parameterizations).  We define
+fractions of the replete-state mass (see the SI, Table I, for parameterizations).  We define
 $m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction of
@@
 \end{eqnarray}
-where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in the biosynthetic energetics during recovery (see Supplementary Information).
-It should be noted that more complicated ontogenetic models explicitly handle
+where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in
+the biosynthetic energetics during recovery (see the SI). It should be noted that more complicated ontogenetic models explicitly handle
 storage \citep{hou}, whereas this feature is implicitly covered by the body

### 2017-10-18 NatureComm_REVISION/YKR_supp-Revision-1.tex
new revision

--- 2017-06-09 Manuscript/Nature-Submission/Revised-Supplement-3.tex
@@
-%\usepackage{scicite}
-%\usepackage{times}
-%\usepackage{graphicx}
-%\usepackage{verbatim}
-%\usepackage{subfigure}
-%\usepackage{tikz}
+\RequirePackage{lineno}
+\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
+%\usepackage[pdftex]{graphicx}
+
+\usepackage{amsmath,amsfonts,amssymb}
+\usepackage[english]{babel}
+\usepackage[latin1]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{color}
+\usepackage{float}
+\usepackage{verbatim}
+\usepackage{graphicx}
+\usepackage{bm}
+\usepackage{mathtools}
+\usepackage{stmaryrd}
+\usepackage{anyfontsize}
+
+%\usepackage{epstopdf}
 %\usepackage{array}
@@
 %\usepackage{multirow}
-%\usepackage{multicol}
+\usepackage{color}
 %\usepackage{multibox}
 %\usepackage{rotating}
-%\usepackage{layout}
-%\usepackage{epstopdf}
-%\usepackage{afterpage}
+%\usepackage{lineno}
 %\usepackage[left]{lineno}
+\usepackage[comma,sort&compress]{natbib}
+\bibpunct{}{}{,}{s}{}{;}
+%\usepackage{authblk}
+%\usepackage{multicol}
 
-%% PNAStmpl.tex
-%% Template file to use for PNAS articles prepared in LaTeX
-%% Version: Apr 14, 2008
+\bibliographystyle{naturemag}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% BASIC CLASS FILE
+% \usepackage{bibunits}
 
-\documentclass{pnastwo}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL GRAPHICS STYLE FILE
-
-\usepackage[pdftex]{graphicx}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL POSTSCRIPT FONT FILES
-
-\usepackage{pnastwof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% ADDITIONAL OPTIONAL STYLE FILES
-
-%\usepackage{amssymb,amsfonts,amsmath}
-\usepackage{amsmath}
-\usepackage{amssymb}
-
-\usepackage[font=large]{caption}
-\usepackage{subfigure}
-\usepackage{epstopdf}
-\usepackage{graphicx}
-\usepackage{verbatim}
-\usepackage{array}
-\usepackage{tabularx}
-\usepackage{multirow}
-\usepackage{color}
-%\usepackage{multibox}
-\usepackage{rotating}
-\usepackage{color}
-%\usepackage{lineno}
-\usepackage[left]{lineno}
-\usepackage[comma,sort&compress]{natbib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% OPTIONAL MACRO FILES
-
-\bibliographystyle{pnas}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\sid}[1]{\textcolor{red}{\bf [#1]}}
-
-%% Don't type in anything in the following section:
-%%%%%%%%%%%%
-%% For PNAS Only:
-\contributor{Submitted to Proceedings
-of the National Academy of Sciences of the United States of America}
-\url{www.pnas.org/cgi/doi/10.1073/pnas.0709640104}
-\copyrightyear{2011}
-\issuedate{Issue Date}
-\volume{Volume}
-\issuenumber{Issue Number}
-%%%%%%%%%%%%
+\linenumbers
+% \setlength\linenumbersep{3pt}
 
@@
 
-\linenumbers
-\setlength\linenumbersep{3pt}
+\author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
+  of California, Merced, Merced, CA 95340, USA}\affiliation{The Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author: jdyeakel@gmail.com}
 
-%\title{Ecological and evolutionary implications of starvation and body size}
+\author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
+  Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}
 
-%\title{Survival of the Stable: Dynamic Consequences of Starvation and Reproduction}
+\author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
+  Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}
 
-%\title{Size Constraints on Starvation \& Reproduction Dynamics}
-
-%\title{Eco-evolutionary Consequences of Starvation and Reproduction: Survival
-%of the Stable}
-
-\title{Supporting Information for ``The dynamics of starvation and recovery''}%: Eco-evolutionary feedbacks}
-
-\author
-{Justin D.
-Yeakel\affil{1}{School of Natural Science, University of California Merced, Merced, CA} \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}\affil{5}{To whom correspondence should be addressed: jdyeakel@gmail.com}, Christopher P.
-Kempes \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{4}{Contributed equally}, \and Sidney Redner \affil{2}{The Santa Fe Institute, Santa Fe, NM}\affil{3}{Department of Physics, Boston University, Boston MA}\affil{4}{Contributed equally}
-}
-
-\contributor{Submitted to Proceedings of the National Academy of Sciences
-of the United States of America}
+\title{ Supporting Information for ``The dynamics of starvation and recovery''}%: Eco-evolutionary feedbacks}
 
@@
 
-\begin{article}
+%%%%%%%%%%%%%%%%%%%%%%%%
+% SUPPLEMENTARY MATERIAL
+%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section*{Mechanisms of Starvation and Recovery}
+%New additions in red text
+
+\clearpage
+
+{\bf Mechanisms of Starvation and Recovery}
 Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
 For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
-\cite{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
+\citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
 This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g.
-\cite{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \cite{kempes2014morphological}.
+\citep{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \citep{kempes2014morphological}.
 Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).
@@
 In the above equations $Y$ represents the yield coefficient (e.g., Refs.
-\citenum{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
+\citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
 If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
@@
 The above equations represent the system of equations presented in the main text.
+\\
 
-\section*{Parameter Values and Estimates}
-
-%\subsection*{Standard synthesis and metabolic parameters}
-
+{\bf Parameter Values and Estimates}
 All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
@@
 Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs.
-\citenum{West:2001bv,moses2008rmo} and the supplement for Ref.
-\citenum{hou}).
-We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \cite{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
+\citep{West:2001bv,moses2008rmo} and the supplement for Ref.
+\citep{hou}).
+We make this assumption in the current paper, although alternate exponents, which are known to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
 The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.
 For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs.
-\citenum{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
+\citep{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
 We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).
@@
 This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref.
-\citenum{peters1986ecological} for a review).
-We use $m_{0}=0.097M^{0.92}$ \cite{blueweiss1978relationships} which encompasses the widest range of body sizes \cite{peters1986ecological}.
+\citep{peters1986ecological} for a review).
+We use $m_{0}=0.097M^{0.92}$ \citep{blueweiss1978relationships} which encompasses the widest range of body sizes \citep{peters1986ecological}.
 
@@
 Refs.
-\citenum{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \cite{moses2008rmo}.
-The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \cite{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \cite{stryer,hou}.
-For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \cite{hou}.
+\citep{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \citep{moses2008rmo}.
+The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \citep{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \citep{stryer,hou}.
+For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \citep{hou}.
 
-For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \cite{Dunbrack:1993ec}, $0.02M^{1.19}$ \cite{Lindstedt:1985hm}, and $0.026M^{1.14}$ \cite{Lindstedt:2002td}.
-We use the values from \cite{Lindstedt:1985hm} which falls in the middle of this range.
-Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \cite{Lindstedt:2002td}.
+For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \citep{Dunbrack:1993ec}, $0.02M^{1.19}$ \citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ \citep{Lindstedt:2002td}.
+We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range.
+Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \citep{Lindstedt:2002td}.
 
@@
 From the data in Ref.
-\citenum{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
-It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \cite{damuth1987interspecific}.
+\citep{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
+It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{damuth1987interspecific}.
 We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above.
 However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.
-
+\\
 %For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
@@
  \begin{table}[h]
+  {\color{red}
 \caption{Parameter values for mammals}
@@
     \begin{center}
-    \small
-     \begin{tabular}{ p{1.2cm} p{3.2cm} l p{2.2cm}|}
+    \footnotesize
+     \begin{tabular}{p{3.8cm} c p{2.2cm} p{1.4cm}}
      \hline
-     Parameter & Value & References  \\
+
+     Definition & Parameter & Value & References  \\
      \hline
-   $\eta$ & $3/4$  &  (e.g. \cite{West:2001bv,moses2008rmo,hou}) \\
-   $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \cite{moses2008rmo,West:2001bv,hou} \\
-   $E_{m}^{\prime}$ & $7000$  & \cite{stryer,hou} \\
-   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \cite{hou}  \\
+   Asymptotic adult mass & $M$ & (g) &  \\
+   Initial mass of an organism & $m_{0}$ & (g) &  \\
+   Metabolic rate scaling exponent & $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
+   Metabolic Normalization Constant & $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
+   Initial mass scaling exponent & $\upsilon$ & $0.92$ &  \citep{blueweiss1978relationships,peters1986ecological} \\
+   Initial mass scaling normalization constant & $n_{0}$ & $0.097$ (g$^{1-\upsilon}$) & \citep{blueweiss1978relationships,peters1986ecological}  \\   
+   Fat mass scaling exponent & $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
+   Fat scaling normalization constant & $f_{0}$ & $0.02$ (g$^{1-\eta}$) & \citep{Lindstedt:1985hm}\\
+   Muscle mass scaling exponent & $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
+   Muscle scaling normalization constantv& $u_{0}$ & $0.38$ (g$^{1-\zeta}$)  & \citep{Lindstedt:2002td} \\
+   Energy to synthesis a unit of mass & $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
+   Energy to synthesis a unit of mass during recovery & $E_{m}^{\prime}$ & $7000$ (J gram$^{-1}$) & \citep{stryer,hou} \\
+   Specific resource growth rate & $\alpha$ & $9.45\times10^{-9}$ (s$^{-1}$) & see text  \\
+   Fraction of asymptotic mass representing full state & $\epsilon_{\lambda}$ & $0.95$ & \citep{West:2001bv}  \\
+   Fraction of asymptotic mass representing starving state & $\epsilon_{\sigma}$ & $1-f_{0}M^{\gamma-1}$ & see text  \\
+   Fraction of asymptotic mass representing death & $\epsilon_{\mu}$ & $1-\frac{f_{0}M^{\gamma}+u_{0}M^{\zeta}}{M}$ & see text \\
+   Carrying capacity (maximum density) of resources & $C$ & (g m$^{-2}$) & \\
+   Half Saturation Constant & $k$ & (g m$^{-2}$) &   \\
+   Normalized carrying capacity & $\xi$ & $C/k\approx2$ &   \\
+   Reproductive fecundity & $\nu$ & $2$ & \citep{}  \\
+
 %   $a$ & $1.78\times10^{-6}$  \quad \quad \\
 %   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
-   $\gamma$ & $1.19$ & \cite{Lindstedt:1985hm} \\
-   $f_{0}$ & $0.02$ & \cite{Lindstedt:1985hm}\\
-   $\zeta$ & $1.00$  & \cite{Lindstedt:2002td} \\
-   $u_{0}$ & $0.38$  & \cite{Lindstedt:2002td} \\
 
@@
     \end{center}
+    }
    \end{table}
 
-\subsection*{Rate equations for invaders with modified body mass}
+{\bf Rate equations for invaders with modified body mass}
 We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
@@
 
-% \bibliography{aa_starving_supplement}
-\begin{thebibliography}{10}
+\bibliography{aa_starving_supplement}
 
-\bibitem{Kempes:2012hy}
-Kempes CP, Dutkiewicz S, Follows MJ
-\newblock (2012) {Growth, metabolic partitioning, and the size of
-  microorganisms.}
-\newblock \emph{PNAS} 109:495--500.
-
-\bibitem{kempes2014morphological}
-Kempes CP, Okegbe C, Mears-Clarke Z, Follows MJ, Dietrich LE
-\newblock (2014) Morphological optimization for access to dual oxidants in
-  biofilms.
-\newblock \emph{Proceedings of the National Academy of Sciences} 111:208--213.
-
-\bibitem{West:2001bv}
-West GB, Brown JH, Enquist BJ
-\newblock (2001) {A general model for ontogenetic growth}.
-\newblock \emph{Nature} 413:628--631.
-
-\bibitem{moses2008rmo}
-Moses ME, {et~al.}
-\newblock (2008) {Revisiting a model of ontogenetic growth: Estimating model
-  parameters from theory and data}.
-\newblock \emph{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}
-  171:632--645.
-
-\bibitem{hou}
-Hou C, {et~al.}
-\newblock (2008) {Energy uptake and allocation during ontogeny}.
-\newblock \emph{Science} 322:736--739.
-
-\bibitem{pirt}
-Pirt S
-\newblock (1965) The maintenance energy of bacteria in growing cultures.
-\newblock \emph{Proceedings of the Royal Society of London B: Biological
-  Sciences} 163:224--231.
-
-\bibitem{Heijnen}
-Heijnen J, Roels J
-\newblock (1981) A macroscopic model describing yield and maintenance
-  relationships in aerobic fermentation processes.
-\newblock \emph{Biotechnology and Bioengineering} 23:739--763.
-
-\bibitem{peters1986ecological}
-Peters RH
-\newblock (1986) \emph{The Ecological Implications of Body Size}
-\newblock (Cambridge University Press, Cambridge) Vol.{}~2.
-
-\bibitem{blueweiss1978relationships}
-Blueweiss L, {et~al.}
-\newblock (1978) Relationships between body size and some life history
-  parameters.
-\newblock \emph{Oecologia} 37:257--272.
-
-\bibitem{stryer}
-Stryer L
-\newblock (1995) \emph{{Biochemistry, Fourth Edition}}
-\newblock (W.H. Freeman and Company, New York), pp 608--611.
-
-\bibitem{Dunbrack:1993ec}
-Dunbrack RL, Ramsay MA
-\newblock (1993) {The Allometry of Mammalian Adaptations to Seasonal
-  Environments: A Critique of the Fasting Endurance Hypothesis}.
-\newblock \emph{Oikos} 66:336--342.
-
-\bibitem{Lindstedt:1985hm}
-Lindstedt SL, Boyce MS
-\newblock (1985) {Seasonality, Fasting Endurance, and Body Size in Mammals}.
-\newblock \emph{Am. Nat.} 125:873--878.
-
-\bibitem{Lindstedt:2002td}
-Lindstedt SL, Schaeffer PJ
-\newblock (2002) {Use of allometry in predicting anatomical and physiological
-  parameters of mammals.}
-\newblock \emph{Lab. Anim.} 36:1--19.
-
-\bibitem{estermann}
-Estermann BL, Wettstein HR, Sutter F, Kreuzer M
-\newblock (2001) Nutrient and energy conversion of grass-fed dairy and suckler
-  beef cattle kept indoors and on high altitude pasture.
-\newblock \emph{Animal Research} 50:477--493.
-
-\bibitem{michaletz2014convergence}
-Michaletz ST, Cheng D, Kerkhoff AJ, Enquist BJ
-\newblock (2014) Convergence of terrestrial plant production across global
-  climate gradients.
-\newblock \emph{Nature} 512:39--43.
-
-\bibitem{damuth1987interspecific}
-Damuth J
-\newblock (1987) Interspecific allometry of population density in mammals and
-  other animals: the independence of body mass and population energy-use.
-\newblock \emph{Biological Journal of the Linnean Society} 31:193--246.
-
-\end{thebibliography}
-
-%\let\cleardoublepage\clearpage
-
-%FIGURE 1
-
-\end{article}
+% \end{bibunit}
 

### 2017-10-18 NatureComm_REVISION/YKR_supp.tex
new revision

--- 2017-10-18 NatureComm_REVISION/YKR_supp-Revision-1.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%
-
-%New additions in red text
 
@@
 \citep{hou}).
-We make this assumption in the current paper, although alternate exponents, which are known to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
+We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
 The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.
@@
  \begin{table}[h]
-  {\color{red}
 \caption{Parameter values for mammals}
@@
     \begin{center}
-    \footnotesize
-     \begin{tabular}{p{3.8cm} c p{2.2cm} p{1.4cm}}
+    \small
+     \begin{tabular}{ p{1.2cm} p{3.2cm} l p{2.2cm}|}
      \hline
-
-     Definition & Parameter & Value & References  \\
+     Parameter & Value & References  \\
      \hline
-   Asymptotic adult mass & $M$ & (g) &  \\
-   Initial mass of an organism & $m_{0}$ & (g) &  \\
-   Metabolic rate scaling exponent & $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
-   Metabolic Normalization Constant & $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
-   Initial mass scaling exponent & $\upsilon$ & $0.92$ &  \citep{blueweiss1978relationships,peters1986ecological} \\
-   Initial mass scaling normalization constant & $n_{0}$ & $0.097$ (g$^{1-\upsilon}$) & \citep{blueweiss1978relationships,peters1986ecological}  \\   
-   Fat mass scaling exponent & $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
-   Fat scaling normalization constant & $f_{0}$ & $0.02$ (g$^{1-\eta}$) & \citep{Lindstedt:1985hm}\\
-   Muscle mass scaling exponent & $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
-   Muscle scaling normalization constantv& $u_{0}$ & $0.38$ (g$^{1-\zeta}$)  & \citep{Lindstedt:2002td} \\
-   Energy to synthesis a unit of mass & $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
-   Energy to synthesis a unit of mass during recovery & $E_{m}^{\prime}$ & $7000$ (J gram$^{-1}$) & \citep{stryer,hou} \\
-   Specific resource growth rate & $\alpha$ & $9.45\times10^{-9}$ (s$^{-1}$) & see text  \\
-   Fraction of asymptotic mass representing full state & $\epsilon_{\lambda}$ & $0.95$ & \citep{West:2001bv}  \\
-   Fraction of asymptotic mass representing starving state & $\epsilon_{\sigma}$ & $1-f_{0}M^{\gamma-1}$ & see text  \\
-   Fraction of asymptotic mass representing death & $\epsilon_{\mu}$ & $1-\frac{f_{0}M^{\gamma}+u_{0}M^{\zeta}}{M}$ & see text \\
-   Carrying capacity (maximum density) of resources & $C$ & (g m$^{-2}$) & \\
-   Half Saturation Constant & $k$ & (g m$^{-2}$) &   \\
-   Normalized carrying capacity & $\xi$ & $C/k\approx2$ &   \\
-   Reproductive fecundity & $\nu$ & $2$ & \citep{}  \\
-
+   $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
+   $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
+   $E_{m}^{\prime}$ & $7000$  & \citep{stryer,hou} \\
+   $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
 %   $a$ & $1.78\times10^{-6}$  \quad \quad \\
 %   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
+   $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
+   $f_{0}$ & $0.02$ & \citep{Lindstedt:1985hm}\\
+   $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
+   $u_{0}$ & $0.38$  & \citep{Lindstedt:2002td} \\
 
@@
     \end{center}
-    }
    \end{table}

### 2017-11-03 NatureComm_REVISION/YKR_NatureComm-Revision-1.tex
revision 3 figure changes

@@
 %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-\noindent {\bf Results \& Discussion}
-\noindent \paragraph*{{\bf Allometry minimizes extinction risk.}} While there are no {\it a priori} constraints on the parameters in the NSM, we expect that each species should be restricted to a distinct portion of the parameter space.
+\noindent {\bf Results}
+\noindent \paragraph*{{\bf The allometry of extinction risk.}} While there are no {\it a priori} constraints on the parameters in the NSM, we expect that each species should be restricted to a distinct portion of the parameter space.
 We
@@
 % \noindent {\bf Energetic barriers to body size}
-\noindent \paragraph*{{\bf Predicting energetic equivalence and body size limits.}} The NSM correctly predicts that species with smaller masses have larger
+\noindent \paragraph*{{\bf Predicting Damuth's Law and body size limits.}} The NSM correctly predicts that species with smaller masses have larger
 steady-state population densities (Fig.~\ref{fig:mass}a).  Similar
@@
 
-\noindent {\bf Conclusion} \\
+\noindent {\bf Discussion} \\
 {\color{red} Complementary previous work has analyzed the tradeoffs associated with carnivory in terms of the energy expenditure on hunting and prey size, where a maximum carnivore size between 700 and 1,000 is predicted from considerations of energy intake going to zero \cite{Carbone:1999ju}.

### 2017-11-03 NatureComm_REVISION/YKR_NatureComm-Revision-2.tex
revision 3 figure changes

--- 2017-11-03 NatureComm_REVISION/YKR_NatureComm-Revision-1.tex
@@
 Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\citep{Bulik:1999eo,Trites:2003cc}.
-In the extreme case of unicellular organisms, nutrition is unavoidably linked to reproduction because the nutritional state of the cell regulates all aspects of the cell cycle \citep{Glazier:2009hq}.
+In the extreme case of unicellular organisms, nutrition directly controls growth to a reproductive state \citep{Glazier:2009hq,Kempes:2012hy}.
 The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the universality of the fundamental tradeoff between somatic and reproductive investment.
@@
 consumer-resource models \citep{DeLong:2012kw}.  Moreover, we show that the
-NSM provides independent theoretical support for the energy equivalence
-hypothesis and Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998}.  The
-energy equivalence hypothesis argues that the total energy use,
-$B_{\rm tot}$, of a population is constant independent of species
-size~\citep{Damuth:1987kr,allen2002,enquist1998}. %Previous studies have focused on t
-This hypothesis is based on observations showing that the steady state
-abundance, $N^*$, of a species is proportional to the inverse of individual
-metabolism, such that
-$N^*\propto M^{-3/4}/B_{0}$~\citep{allen2002,enquist1998}.  This relationship
-implies that $B_{\rm tot}=N^*B(M)=Q$, where $Q$ is a constant, and has been
-shown to hold in both mammalian and vascular plant communities
-\citep{Damuth:1987kr,allen2002,enquist1998}.  Figure \ref{fig:mass}a shows
+NSM provides independent theoretical support for Damuth's Law and the energy equivalence
+hypothesis\citep{Damuth:1987kr,allen2002,enquist1998,pedersen2017shallow}.
+{\color{red} Damuth's law shows that the  abundances of species, $N^{*}$, follows $N^*\propto M^{-0.78}$.
+The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}.
+As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}, however the actually equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations.
+Figure \ref{fig:mass}a shows
 that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of
-organism sizes and Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly
-constant over this same range.  This result is remarkable because it
+organism sizes and that $F^{*}+H^{*}$ closely matches the best fit to Damuth's data.
+Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly constant over this same range.
+This result is remarkable because it
 illustrates that the steady state values of the NSM combined with the derived
-timescales naturally give rise to energy equivalence.
+timescales naturally give rise to Damuth's law.
+While the previous metabolic studies supporting Damuth's law provided arguments for the value of the exponent \citep{allen2002}, these studies are only able to infer the intercept from the data.
+Our model predicts not only the exponent but also this intercept by explicitly including the resource dynamics and the parameters adjusting growth and consumption.
+
+It should be noted the density relationships of individual clades follow more shallow scaling relationships than Damuth's law \cite{pedersen2017shallow}.
+In the context of our model, this suggests that future work may be able to anticipate these shifts from differences in the physiological parameters associated with each clade.}
 
@@
 This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero, causing the timescales to become infinite and the rates $\mu$ and $\lambda$ to equal zero.
-A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$, or $120$ times the mass of a male African elephant.
+A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass of a male African elephant.
 In contrast, the $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$. 
-This point predicts a more realistic upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$.
+This point predicts a more realistic upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$ (g).
 Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is within an order of magnitude of the maximum body size observed in the North American mammalian fossil record~\citep{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\citep{Clauset:2009fh}.
@@
 We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\citep{Kempes:2012hy,Kempes:2016}.
-% reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
+{\color{red} It has also been shown that models of the energetic allometry of hunting and resting combined with foraging time predicts a maximum carnivore size between $7\times10^{5}$ and $1.1\times10^{6}$ (g) \cite{Carbone:1999ju,carbone2007costs}.
+{\color{red} Similarly, other past work has shown that the maximum body size within a particular lineage should scale with the metabolic normalization constant and depend on a critical death found to be constant from data \cite{Okie:2013ju}.
+This complementary approach is based on the balance between death and growth, and in connection with our model here suggests that future connections between the scaling of fat and muscle mass should be systematically connected with $B_{0}$ when comparing lineages.}} % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
 
@@
 Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \citep{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals explored by Alroy~\citep{Alroy:1998p1594}.
+{\color{red} It should be noted that the model of \cite{Brown:1993p708} predicts an optimal mammal size at intermediate mammalian sizes using reproductive maximization.
+This model considers the transition between hungry and full individuals as well, but our coupling to resources and an explicit treatment of storage shows that the largest mammals have an advantage in terms resource competition.
+Another subtlety is that the distributions in \cite{Brown:1993p708} show that intermediate mammal sizes have the most number of species, and it is important to recognize that our efforts consider total biomass where recent work shows that many ecosystems are dominated by the biomass of the large \cite{}.}
 %as well as  \citep{Alroy:1998p1594,Clauset:2009fh}.
 %This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
+
 While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general selective mechanism for the evolution of larger body size among terrestrial mammals.\\
@@
 \noindent {\bf Discussion} \\
-{\color{red} Complementary previous work has analyzed the tradeoffs associated with carnivory in terms of the energy expenditure on hunting and prey size, where a maximum carnivore size between 700 and 1,000 is predicted from considerations of energy intake going to zero \cite{Carbone:1999ju}.
-Similarly, it has been shown in the past that the largest body size within a particular lineage should be a function of the metabolic normalization constant and a critical death rate as supported by data.
-This approach, which consider the energetics associated with active metabolism, provides a complementary prediction for maximum size from energetics \cite{Okie:2013ju}, and also suggests that future connections between the scaling of fat and muscle mass should be systematically connected with $B_{0}$.
-}
 
@@
 \end{equation}
-where we will define values of $\epsilon$ to describe a set of rates within
-our model.  The time to reproduce is given by
-$t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
+{\color{red} where we will define values of $\epsilon$ to describe a variety of timescales, and related rates, within
+our model.
+For example, the rate of reproduction is given by the timescale to go from the birth mass to the adult mass.
+The time to reproduce is given by Equation \ref{t1} as $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
 is the fraction of the asymptotic mass where an organism is reproductively
 mature and should be close to one (typically
-$\epsilon_{\lambda}\approx0.95\;$ \citep{West:2001bv}). The growth rate then
-is $\lambda=\ln\left(\nu\right)/t_{\lambda}$ where $\nu$ is the number of
-offspring produced, and for any constant value of $\epsilon_{\lambda}$, this
-rate will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$
-\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
+$\epsilon_{\lambda}\approx0.95\;$ \citep{West:2001bv}).
+Our reproductive rate, $\lambda$, is a specific rate, or the number of offspring produced per time per individual, defined as $\dot{F} = \lambda F$.
+In isolation this functional form gives population growth following $F\left(t\right) = F_{0}e^{\lambda t}$ which can be related to the reproductive timescale by assuming that when $t=t_{\lambda}$ it is also the case that $F=\nu F_{0}$, where $\nu-1$ is the number of offspring produced per reproductive cycle.
+Following this relationship the growth rate is given by $\lambda=\ln\left(\nu\right)/t_{\lambda}$, which is the standard relationship \cite{} and will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of $\epsilon_{\lambda}$
+\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.}
 

### 2017-11-03 NatureComm_REVISION/YKR_NatureComm-Revision-3.tex
revision 3 figure changes

--- 2017-11-03 NatureComm_REVISION/YKR_NatureComm-Revision-2.tex
@@
 Of particular importance is the interplay between reproduction and the dynamics of starvation and recovery. % in resource-limited environments
-We introduce a minimal nutritional state-structured model that incorporates two classes of consumer: nutritionally replete, reproducing consumers, and undernourished, non-reproducing consumers.
+We introduce a nutritional state-structured model that incorporates two classes of consumer: nutritionally replete, reproducing consumers, and undernourished, non-reproducing consumers.
 % that are susceptible to mortality
@@
 
-The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly influences how they invest reserves in uncertain environments.
+The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly impacts how they invest reserves in uncertain environments.
 Such behaviors are generally manifested as tradeoffs between investing in somatic maintenance and growth, or allocating energy towards reproduction~\citep{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
@@
 In the extreme case of unicellular organisms, nutrition directly controls growth to a reproductive state \citep{Glazier:2009hq,Kempes:2012hy}.
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the universality of the fundamental tradeoff between somatic and reproductive investment.
+The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the near-universality of the fundamental tradeoff between somatic and reproductive investment.
 
@@
   approach---the Nutritional State-structured Model (NSM)---that accounts for
-  resource limitation through the consequences of starvation. In contrast to
+  resource limitation via explicit starvation. In contrast to
   the LV model, the NSM incorporates two consumer states:
@@
   (i) the dynamics are typically driven to a refuge far from cyclic behavior and extinction risk, 
-  (ii) steady state conditions of the NSM accurately predict measured biomass densities for mammals as well as Damuth's energetic equivalence principle, 
-  (iii) an allometrically constrained upper-bound for mammalian body size, and
-  (iv) a selective mechanism for the evolution of larger body size, known as Cope's rule.
+  (ii) steady state conditions of the NSM accurately predict the measured biomass densities for mammals described by Damuth's law,
+  %  energetic equivalence principle, 
+  (iii) there is an allometrically constrained upper-bound for mammalian body size, and
+  (iv) the NSM provides a selective mechanism for the evolution of larger body size, known as Cope's rule.
 
@@
 \noindent \paragraph*{{\bf Nutritional state-structured model (NSM).}}
-We begin by defining a minimal Nutritional State-structured population Model
-(NSM), where the consumer population is partitioned into two states: (a) an
+We begin by defining the nutritional state-structured population model, where the consumer population is partitioned into two states: (a) an
 energetically replete (full) state $F$, where the consumer reproduces at a
@@
 reproduce but dies by starvation at rate $\mu$. The dynamics of the
-underlying resource $R$ is governed by logistic growth with an intrinsic
+underlying resource $R$ are governed by logistic growth with an intrinsic
 growth rate $\alpha$ and a carrying capacity $C$. The rate at which consumers
@@
 Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
+{\color{red} We note that including additional allometrically-scaled mortality terms to both $F$ and $H$ does not impact our findings (see SI for derivation).}
 
@@
 
-\begin{figure*}
+\begin{figure}
 \centering
-\includegraphics[width=0.8\textwidth]{fig_ExtinctionAllometricComb3.eps}
-\caption{\small{ Probability of extinction for a consumer with ({\bf a}) $M=10^2$g, ({\bf b}) $M=10^4$g, and ({\bf c}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
+\includegraphics[width=0.4\textwidth]{fig_ExtinctionAllometricComb4.eps}
+\caption{\small{ Probability of extinction for a consumer with ({\bf a}) $M=10^2$g and ({\bf b}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
 Note the change in scale for $M=10^4$ and $M=10^6$g.
@@
 The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
-\end{figure*}
+\end{figure}
 
@@
 To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
-This procedure is repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for organisms with mass ranging from $10^2$ to $10^6$ grams.
+This procedure was repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for organisms with mass ranging from $10^2$ to $10^6$ grams.
 %the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
-In each replicate the initial densities are chosen to be $(XF^*,XH^*,R^*)$, with $X$ a random variable that is uniformly distributed in $[0,2]$.
-By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-7}$ to $10^{-3}$.
+In each replicate the initial densities were chosen to be $(XF^*,XH^*,R^*)$, with $X$ a random variable uniformly distributed in $[0,2]$.
+By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-8}$ to $10^{-3}$.
 %, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
-As expected, higher rates of extinction correlate with both high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if $\sigma$ is small.
-For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from transient cycles with larger amplitudes as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
+Higher rates of extinction correspond to both high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if $\sigma$ is small.
+For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from higher-amplitude transient cycles as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
 For high values of $\sigma$ and low values of $\rho$, increased extinction risk arises because of the decrease in the steady-state consumer population density (Figs. \ref{fig:fp}b, \ref{fig:ext}).
@@
 % \noindent \paragraph*{{\bf Extinction refuge.}} 
-We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge across a range of $M$ (Fig. \ref{fig:ext}a-c, white points).
-These values are close enough to the Hopf bifurcation to avoid low steady-state densities, and far enough away to avoid large-amplitude transient cycles.
-The feature that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
-Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that serve to minimize extinction risk.
-This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
+We find that the allometrically constrained values of $\sigma$ and $\rho$, each representing different trajectories along the ontogenetic curve (Fig.
+\ref{fig:growth}), fall squarely within the extinction refuge across a range of $M$ (Fig.
+\ref{fig:ext}a,b, white points).
+These values are close enough to the Hopf bifurcation to avoid low steady-state densities, yet distant enough to avoid large-amplitude transient cycles.
+That allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving starvation and recovery rates within populations.
+Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that minimize extinction risk.
+This selection could occur via the tuning of body fat percentages, metabolic rates, and/or biomass maintenance efficiencies.
 We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig.
-\ref{fig:ext}a-c), suggesting that the population dynamics of larger organisms are more sensitive to variability in physiological rates controlling starvation and recovery.
-This finding is in accordance with, and may serve as contributing support for, observations of increased extinction risk among larger mammals \cite{Liow:2008jx}.
-Moreover, larger body sizes decrease the steady state resource density, such that fluctuations for larger organisms will be more likely to drive resources to extinction.
+\ref{fig:ext}b), suggesting that the population dynamics for larger organisms are more sensitive to variability in physiological rates.
+This finding is in accordance with, and may serve as contributing support for, observations of increased extinction risk among larger mammals \cite{Liow:2008jx}.\\
+% Moreover, larger body sizes decrease the steady state resource density, such that fluctuations for larger organisms will be more likely to drive resources to extinction.
 % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
+% To summarize, our finding that allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
 
@@
 % \noindent {\bf Energetic barriers to body size}
-\noindent \paragraph*{{\bf Predicting Damuth's Law and body size limits.}} The NSM correctly predicts that species with smaller masses have larger
-steady-state population densities (Fig.~\ref{fig:mass}a).  Similar
+\noindent \paragraph*{{\bf Predicting Damuth's Law and body size limits.}} The NSM correctly predicts that smaller species have larger
+steady-state population densities (Fig.~\ref{fig:mass}).  Similar
 predictions have been made for carnivore populations using alternative
 consumer-resource models \citep{DeLong:2012kw}.  Moreover, we show that the
-NSM provides independent theoretical support for Damuth's Law and the energy equivalence
-hypothesis\citep{Damuth:1987kr,allen2002,enquist1998,pedersen2017shallow}.
-{\color{red} Damuth's law shows that the  abundances of species, $N^{*}$, follows $N^*\propto M^{-0.78}$.
-The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}.
-As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}, however the actually equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations.
-Figure \ref{fig:mass}a shows
+NSM provides independent theoretical support for Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998,pedersen2017shallow}. 
+Damuth's law reveals that species abundances, $N^{*}$, follow $N^*\propto M^{-0.78}$. Figure \ref{fig:mass} shows
 that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of
-organism sizes and that $F^{*}+H^{*}$ closely matches the best fit to Damuth's data.
-Figure \ref{fig:mass}b shows that $F^{*}B$ is nearly constant over this same range.
-This result is remarkable because it
+organismal sizes and that $F^{*}+H^{*}$ closely matches the best fit to Damuth's data. This result is remarkable because it
 illustrates that the steady state values of the NSM combined with the derived
 timescales naturally give rise to Damuth's law.
-While the previous metabolic studies supporting Damuth's law provided arguments for the value of the exponent \citep{allen2002}, these studies are only able to infer the intercept from the data.
-Our model predicts not only the exponent but also this intercept by explicitly including the resource dynamics and the parameters adjusting growth and consumption.
-
-It should be noted the density relationships of individual clades follow more shallow scaling relationships than Damuth's law \cite{pedersen2017shallow}.
-In the context of our model, this suggests that future work may be able to anticipate these shifts from differences in the physiological parameters associated with each clade.}
+While the previous metabolic studies supporting Damuth's law provided arguments for the value of the exponent \citep{allen2002}, these studies are only able to infer the intercept from the data (see SI for a discussion of the energy equivalence hypothesis related to these metabolic arguments).
+Our model predicts not only the exponent but also the intercept by explicitly including the resource dynamics and the parameters adjusting growth and consumption.
+It should be noted that density relationships of individual clades follow a more shallow scaling relationship than predicted by Damuth's law \cite{pedersen2017shallow}.
+In the context of our model, this suggests that future work may be able to anticipate these shifts by accounting for differences in the physiological parameters associated with each clade.
 
@@
 \centering
-\includegraphics[width=0.48\textwidth]{fig_FPAllometric2.eps}
-\caption{\small{ ({\bf a}) Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-  body mass. Inset: Resource steady state $R^*$ as a function of consumer body mass.
-  ({\bf b}) Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
-  The data are from Damuth \citep{Damuth:1987kr} and have been converted to
-  total population metabolism using the allometric relationships for
-  metabolic rate (see the SI and Refs.~\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:mass}}
+\includegraphics[width=0.4\textwidth]{fig_FPAllometric2.eps}
+\caption{\small{Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
+  body mass along with the data from Damuth \citep{Damuth:1987kr}. Inset: Resource steady state $R^*$ as a function of consumer body mass.}\label{fig:mass}}
 \end{figure}
@@
 % \paragraph*{{\bf Maximum Body Size}}
-Our model shows that energetic equivalence breaks down at large $M$ suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.  
+%JUSTIN I REMOVED: Our model shows that energetic equivalence breaks down at large $M$ suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
 %, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
@@
 We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\citep{Kempes:2012hy,Kempes:2016}.
-{\color{red} It has also been shown that models of the energetic allometry of hunting and resting combined with foraging time predicts a maximum carnivore size between $7\times10^{5}$ and $1.1\times10^{6}$ (g) \cite{Carbone:1999ju,carbone2007costs}.
-{\color{red} Similarly, other past work has shown that the maximum body size within a particular lineage should scale with the metabolic normalization constant and depend on a critical death found to be constant from data \cite{Okie:2013ju}.
-This complementary approach is based on the balance between death and growth, and in connection with our model here suggests that future connections between the scaling of fat and muscle mass should be systematically connected with $B_{0}$ when comparing lineages.}} % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
+It has also been shown that models that incorporate the allometry of hunting and resting combined with foraging time predicts a maximum carnivore size between $7\times10^{5}$ and $1.1\times10^{6}$ (g) \cite{Carbone:1999ju,carbone2007costs}.
+Similarly, the maximum body size within a particular lineage has been shown to scale with the metabolic normalization constant \cite{Okie:2013ju}.
+%and depend on a critical death found to be constant from data
+This complementary approach is based on the balance between growth and mortality, and suggests that future connections between the scaling of fat and muscle mass should systematically be connected with $B_{0}$ when comparing lineages.
+% reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
+
+% {\color{red}
+% 
+% It should be noted that in our model all death is related to nutritional state, however the addition of a constant death term \cite{} would only adjust the values of $\lambda$ and $\mu$ without changing the form or or qualitative results of our model.
+We explored the sensitivity of our results to shifts in a constant death term (see SI) and find that additional sources of death do not shift our quantitative prediction of Damuth's law except for large sizes (Figure S\ref{}).
+We find that additional sources of death shifts the asymptotic behavior of our model and decreases the maximum possible mammal size.
+Within the range of reasonable death rates $M_{max}$ varies by $VALUE\%$.
+These results suggest that the largest mammals are most sensitive to additional sources of death perhaps putting them at greater extinction risk.
+% 
+% }
 
@@
 One important such example is the \emph{fasting endurance hypothesis}, which contends that larger body size, with consequent lower metabolic rates and increased ability to maintain more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability~\citep{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size known as Cope's rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ grams~\citep{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\citep{Clauset:2009fh}.
+Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size known as Cope's rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ (g)~\citep{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\citep{Clauset:2009fh}.
 These trends are thought to be driven by a combination of climate change and niche availability~\citep{Saarinen:2014br}; however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, have not been explored.
@@
 The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
-\ref{fig:mass}a, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \citep{tilman1981,dutkiewicz2009,barton2010}.
+\ref{fig:mass}, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \citep{tilman1981,dutkiewicz2009,barton2010}.
 Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but these absolute limits do not offer a mechanism by which larger body sizes are selected for or against.
@@
 Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
-The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$g~\citep{Smith:2010p3442}.}\label{fig:invasion}}
+The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$ (g)~\citep{Smith:2010p3442}.}\label{fig:invasion}}
 \end{figure}
@@
 We directly assess competitive outcome between two closely related species: a resident species of mass $M$, and a competing species (denoted by $^\prime$) where individuals have a different proportion of body fat such that $M^\prime=M(1+\chi)$.
-If $\chi < 0$, individuals within the competing population have fewer metabolic reserves, and if $\chi>0$, individuals have more metabolic reserves than the resident species.
-For the allowable values of $\chi$ the adjusted mass should exceed the minimal amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given Equation \ref{t1}, implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
+If $\chi < 0$, individuals within the competing population have fewer metabolic reserves, and if $\chi>0$, individuals have greater metabolic reserves than the resident species.
+For the allowable values of $\chi$ the adjusted mass should exceed the minimal amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given Eq.
+\ref{t1}, implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
 Together these conditions imply that $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$ and the lower bound is mass-dependent.
-The modified mass adjusts our model via the altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$ (see the SI for detailed derivations of competitor rates).
+The modified mass of the competitor leads to altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$ (see the SI for detailed derivations of competitor rates).
 Importantly, $\epsilon_\sigma$, which determines the point along the growth
@@
 To assess the susceptibility of the resident species to competitive exclusion, we determine which consumer pushes the steady-state resource density $R^*$ to lower values for a given value of $\chi$, with the expectation that a population capable of surviving on lower resource densities has a competitive advantage \citep{tilman1981}.
-We find that for $M\leq 1.748\times10^7$g, having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the competitor has an intrinsic advantage over the resident species (Fig.
+We find that for $M\leq 1.748\times10^7$ (g), having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the competitor has an intrinsic advantage over the resident species (Fig.
 \ref{fig:invasion}).
-However, for $M> 1.748\times10^7$g, leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage from having more metabolic reserves to having less.
+However, for $M> 1.748\times10^7$ (g), leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage.
 
@@
 
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$g thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
+The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$ (g) thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
 This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca.
-$1.5\times10^7$g and the late Miocene \emph{Deinotherium} at ca.
-$1.74\times10^7$g ~\citep{Smith:2010p3442}.
-Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \citep{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' for North American mammals explored by Alroy~\citep{Alroy:1998p1594}.
-{\color{red} It should be noted that the model of \cite{Brown:1993p708} predicts an optimal mammal size at intermediate mammalian sizes using reproductive maximization.
-This model considers the transition between hungry and full individuals as well, but our coupling to resources and an explicit treatment of storage shows that the largest mammals have an advantage in terms resource competition.
-Another subtlety is that the distributions in \cite{Brown:1993p708} show that intermediate mammal sizes have the most number of species, and it is important to recognize that our efforts consider total biomass where recent work shows that many ecosystems are dominated by the biomass of the large \cite{}.}
+$1.5\times10^7$ (g) and the late Miocene \emph{Deinotherium} at ca.
+$1.74\times10^7$ (g) ~\citep{Smith:2010p3442}.
+Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \citep{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' explored by Alroy~\citep{Alroy:1998p1594}.
+% {\color{red} We note that the model of \cite{Brown:1993p708} predicts
+
+An optimal size for mammals at intermediate body mass was predicted by Brown et al. based on reproductive maximization and the transition between hungry and full individuals \cite{Brown:1993p708}. 
+By coupling the NSM to resource dynamics as well as introducing an explicit treatment of storage, we show that species with larger body masses have an inherent competitive advantage for size classes up to $M_{\rm opt}= 1.748\times10^7$.
+% shows that the largest mammals have an advantage in terms resource competition. 
+% Moreover, the mass distributions in \cite{Brown:1993p708} show that intermediate mammal sizes have the most number of species, and it is important to recognize that our efforts consider total biomass where recent work shows that many ecosystems are dominated by the biomass of the large \cite{}.
 %as well as  \citep{Alroy:1998p1594,Clauset:2009fh}.
 %This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
-
-While the state of the environment, as well as the competitive landscape, will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general selective mechanism for the evolution of larger body size among terrestrial mammals.\\
+While the state of the environment as well as the competitive landscape will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general selective mechanism for the evolution of larger body size among terrestrial mammals.\\
 
@@
 metazoans
-\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}. This relation is derived from the simple balance condition 
+\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,savage2004effects,Kempes:2012hy}. This relation is derived from the simple balance condition 
 $B_{0}m^{\eta}=E_{m}\dot{m}+B_{m}m\,,$
@@
 % \end{eqnarray}
-\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy} where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
+\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,savage2004effects,Kempes:2012hy} where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
 the metabolic rate to support an existing unit of mass, and $m$ is the mass
@@
 \end{equation}
-{\color{red} where we will define values of $\epsilon$ to describe a variety of timescales, and related rates, within
+where we will define values of $\epsilon$ to describe a variety of timescales, and related rates, within
 our model.
@@
 In isolation this functional form gives population growth following $F\left(t\right) = F_{0}e^{\lambda t}$ which can be related to the reproductive timescale by assuming that when $t=t_{\lambda}$ it is also the case that $F=\nu F_{0}$, where $\nu-1$ is the number of offspring produced per reproductive cycle.
-Following this relationship the growth rate is given by $\lambda=\ln\left(\nu\right)/t_{\lambda}$, which is the standard relationship \cite{} and will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of $\epsilon_{\lambda}$
-\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.}
+Following this relationship the growth rate is given by $\lambda=\ln\left(\nu\right)/t_{\lambda}$, which is the standard relationship (e.g.
+\cite{savage2004effects}) and will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of $\epsilon_{\lambda}$
+\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
 

### 2017-11-03 NatureComm_REVISION/YKR_supp-Revision-1.tex
revision 3 figure changes

@@
 & \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
+\label{bigdynamics}
 \end{split}
@@
 \dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
+\label{reduceddynamics}
 \end{split}
@@
 
+{\color{red} \section{Sensitivity to additional death terms}
+
+It should be noted that our set of dynamics (Equations \ref{bigdynamics} and \ref{reduceddynamics}) could include a constant death term of the form $-d_{F}F$ and $-d_{H}H$ to represent death not directly linked to starvation.
+Adding terms of this form to our model would simply adjust the effective value of $\lambda$ and $\mu$, and we could rewrite Equation \ref{reduceddynamics} with $\lambda^{\prime}=\lambda-d$ and $\mu^{\prime}=\mu-d$.
+These substitutions would not alter the functional form of our model nor the steady-states and qualitative results, however the quantitive values could shift based on the size of $d$ relative to $\lambda$ and $\mu$.
+
+Survivorship has a well-known functional form which changes systematically with size \cite{}. Typically survivorship is defined using the Gompertz curve 
+\begin{equation}
+F=F_{0}e^{\left(q_{0}/c\right)\left(1-e^{ct}\right)}
+\label{gompertz}
+\end{equation}
+where the parameters have the following allometric dependencies on adult mass \cite{}.
+
+We are interested in the specific death rate of the form $\cdot{F}=-dF$, and using the derivative of Equation \ref{gompertz} we find that $d=e^{ct}q_{0}$.
+Our model considers the average rates over a population and lifecycle and the average death rate is given by
+\begin{eqnarray}
+\bar{d}&=&\frac{1}{t_{exp}}\int_{0}^{t_{exp}}e^{ct}q_{0} dt \\
+&=&\frac{q_{0}\left(e^{c t_{exp}}-1\right)}{ct_{exp}}
+\end{eqnarray}
+where $t_{exp}$ is the expected lifespan following the allometry of $t_{exp}=$ \cite{}. Given the allometries above we have that
+\begin{equation}
+d=
+\end{equation}
+which scales roughly like $SCALING$.
+In figure \ref{} we compare the value of $d$ to the reproductive and starvation mortality rates .
+The values of $d$ are orders of magnitude smaller than these rates at all scales, and adding this non-starvation based death rate to our model does not shift our results within numerical confidence.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.48\textwidth]{fig_FPAllometric2.eps}
+\caption{\small{ Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
+  The data are from Damuth \citep{Damuth:1987kr} and have been converted to
+  total population metabolism using the allometric relationships for
+  metabolic rate (e.g. Refs.~\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:equivalence}}
+\end{figure}
+
+}
+
+{\color{red} \section{NSM and the energy equivalence hypothesis}
+
+The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}.
+As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}, however the actually equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations.
+Figure \ref{fig:equivalence} shows that $F^{*}B$ is nearly constant over this same range.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.48\textwidth]{fig_FPAllometric2.eps}
+\caption{\small{ Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass.
+  The data are from Damuth \citep{Damuth:1987kr} and have been converted to
+  total population metabolism using the allometric relationships for
+  metabolic rate (e.g. Refs.~\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:equivalence}}
+\end{figure}
+
+}
+
 \bibliography{aa_starving_supplement}

### 2017-12-13 NatureComm_REVISION/Arxiv/YKR_Arxiv3.tex
editorial revisions

\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
%\usepackage[pdftex]{graphicx}

\usepackage{amsmath,amsfonts,amssymb}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage{float}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{stmaryrd}
\usepackage{anyfontsize}

%\usepackage{epstopdf}
%\usepackage{array}
%\usepackage{tabularx}
%\usepackage{multirow}
\usepackage{color}
%\usepackage{multibox}
%\usepackage{rotating}
%\usepackage{lineno}
%\usepackage[left]{lineno}
%\usepackage[comma,sort&compress]{natbib}
%\usepackage{authblk}
%\usepackage{multicol}

%\bibliographystyle{ieeetr}

\usepackage{bibunits}

%\linenumbers
%\setlength\linenumbersep{3pt}

\begin{document}

\author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
  of California, Merced, Merced, CA 95340, USA}

\author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
  Park Road, Santa Fe, NM 87501, USA}

\author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
  Road, Santa Fe, NM 87501, USA}

\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}

\begin{abstract} %250 words
The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of its constituent individuals. Of particular importance is the interplay between reproduction and the dynamics of starvation and recovery. To elucidate this interplay, we introduce a nutritional state-structured model that incorporates two classes of consumer: nutritionally replete, reproducing consumers, and undernourished, non-reproducing consumers. We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships and find that population dynamics are typically driven to a steady state. Moreover, these rates fall within a `refuge' in parameter space, where the probability of population extinction is minimized. We also show that our model provides a natural framework to predict maximum mammalian body size by determining the relative stability of an otherwise homogeneous population to a competing population with altered percent body fat. This framework provides a principled mechanism for a selective driver of Cope's rule.
\end{abstract}

\maketitle

\begin{bibunit}[unsrt]

  The behavioral ecology of all organisms is influenced by their energetic
  states, which directly impacts how they invest reserves in uncertain
  environments.  Such behaviors are generally manifested as tradeoffs between
  investing in somatic maintenance and growth, or allocating energy towards
  reproduction~\citep{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.  The timing of
  these behaviors responds to selective pressure, as the choice of the
  investment impacts future
  fitness~\citep{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.  The influence of
  resource limitation on an organism's ability to maintain its nutritional
  stores may lead to repeated delays or shifts in reproduction over the course
  of an organism's life.

  The balance between (a) somatic growth and maintenance, and (b) reproduction depends on resource availability~\citep{Morris:1987eo}.
  For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\citep{Tveraa:2003fq}.
  Many bird species invest differently in broods during periods of resource scarcity~\citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\citep{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
  Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\citep{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\citep{Kirk:1997cc}.
  Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate reserves~\citep{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.

  Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.
  Many mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development until nutritional reserves can be accumulated~\citep{Mead:1989dt,Sandell:1990kw}.
  Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\citep{Bulik:1999eo,Trites:2003cc}.
  In the extreme case of unicellular organisms, nutrition directly controls growth to a reproductive state \citep{Glazier:2009hq,Kempes:2012hy}. The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the near-universality of the fundamental tradeoff between somatic and reproductive investment.

  Including individual energetic dynamics~\citep{Kooi2000} in a
  population-level framework~\citep{Kooi2000,Sousa:2010ez} is
  challenging~\citep{Diekmann:2010da}.  A common simplifying approach is the
  classic Lotka-Volterra (LV) model, which assumes that consumer population
  growth rate depends linearly on resource density~\citep{murdoch:2003}. Here,
  we introduce an alternative approach---the Nutritional State-structured Model
  (NSM)---that accounts for resource limitation via explicit starvation. In
  contrast to the LV model, the NSM incorporates two consumer states: hungry
  and full, with only the former susceptible to mortality and only the latter
  possessing sufficient energetic reserves to reproduce.  Additionally, we
  incorporate allometrically derived constraints on the time scales for
  reproduction~\citep{Kempes:2012hy}, starvation, and recovery.  Our model
  makes several basic predictions: (i) the dynamics are typically driven to a
  refuge far from cyclic behavior and extinction risk, (ii) the steady-state
  conditions of the NSM accurately predict the measured biomass densities for
  mammals described by Damuth's law~\citep{Damuth:1987kr,allen2002,enquist1998,Pedersen:2017he},
    %  energetic equivalence principle, 
  (iii) there is an allometrically constrained upper-bound for mammalian body size, and
  (iv) the NSM provides a selective mechanism for the evolution of larger body size, known as Cope's rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}.

  % Though straightforward conceptually, incorporating the energetic dynamics of
  % individuals~\citep{Kooi2000} into a population-level
  % framework~\citep{Kooi2000,Sousa:2010ez} presents mathematical
  % obstacles~\citep{Diekmann:2010da}.  An alternative approach involves modeling
  % the macroscale relations that guide somatic versus reproductive investment in
  % a consumer-resource system.  {\color{red}For example, macroscale
  %   Lotka-Volterra models assume that the growth rate of the consumer
  %   population depends on resource density, thus incorporating the requirement
  %   of resource availability for reproduction~\citep{murdoch:2003}.

  % In this work, we adopt an alternative approach to account for resource
  % limitation through the consequences of starvation.  Namely, in contrast to
  % the Lotka-Volterra model, we incorporate two states: hungry individuals and
  % individuals with sufficient energetic reserves to reproduce.}  Such a
  % constraint leads to reproductive time lags due to some members of the
  % population going hungry and then recovering.  Additionally, we incorporate
  % the idea that reproduction is strongly constrained allometrically
  % \citep{Kempes:2012hy}, and is not generally linearly related to resource
  % density.
  % As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.\\

  \noindent \paragraph*{{\bf Nutritional state-structured model (NSM).}}
  We begin by defining the nutritional state-structured population model, where the consumer population is partitioned into two states: (a) an
  energetically replete (full) state $F$, where the consumer reproduces at a
  constant rate $\lambda$ and does not die from starvation, and (b) an
  energetically deficient (hungry) state $H$, where the consumer does not
  reproduce but dies by starvation at rate $\mu$. The dynamics of the
  underlying resource $R$ are governed by logistic growth with an intrinsic
  growth rate $\alpha$ and a carrying capacity $C$. The rate at which consumers
  transition between states and consume resources is dependent on their number,
  the abundance of resources, the efficiency of converting resources into
  metabolism, and how that metabolism is partitioned between maintenance and
  growth purposes.  We provide a physiologically and energetically mechanistic
  model for each of these dynamics and constants (see the Supplementary Information
  (SI)), and show that the system produces a simple non-dimensional form which
  we describe below.

  \begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{fig_FixedPoint2-eps-converted-to.pdf}
  \caption{\small{ The transcritical (TC; dashed line) and Hopf bifurcation (solid line) as a
    function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
    bifurcation conditions separate parameter space into unphysical (left of the TC), cyclic,
    and steady state dynamic regimes.  The colors show the steady state densities for the energetically replete consumers $F^*$.  
    % Steady state densities for the energetically deficient consumers $H^*$ (not shown)
    % scale with those for $F^*$.  
  }\label{fig:fp}}
  \end{figure}

  Consumers transition from the full state $F$ to the hungry state $H$ at a
  rate $\sigma$---the starvation rate---and also in proportion to the absence
  of resources $(1-R)$ (the maximum resource density has been non
  dimensionalized to 1; see SI).  Conversely, consumers recover from state $H$
  to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$
  represents a ratio between maximal resource consumption and the carrying
  capacity of the resource. %and $\approx 2$.
  The resources that are eaten by hungry consumers (at rate $\rho R + \delta$)
  account for their somatic growth ($\rho R$) and maintenance ($\delta$).  Full
  consumers eat resources at a constant rate $\beta$ that accounts for maximal
  maintenance and somatic growth (see the SI for mechanistic derivations of
  these rates from resource energetics).
  %This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
  The NSM represents an ecologically motivated fundamental extension of the
  idealized starving random walk model of foraging, which focuses on resource
  depletion, to include reproduction and resource
  replenishment~\citep{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is
  a more general formulation than previous models that incorporate
  starvation~\citep{Persson:1998hz}.

  In the mean-field approximation, in which the consumers and resources are
  perfectly mixed, their densities are governed by the rate equations

  \begin{align}
  \label{eq:system}
  \begin{split}
  \dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
  \dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
  \dot{R} &= \alpha \left(1-R\right)R -\left(\rho R+\delta\right)H-\beta F.
  \end{split}
  \end{align}

  This system of nondimensional equations follows from a set of first-principle
  relationships for resource consumption and growth (see the SI for a full derivation and the dimensional form).
  Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
  This resembles the equation of motion for the predator density in the LV model~\citep{murray2011mathematical}, except that the resource density does not appear in the growth term.
  % As discussed above, the attributes of reproduction and mortality have been explicitly apportioned to the full and hungry consumers, respectively, so that the growth in the total density is decoupled from the resource density.
  The rate of reproduction is independent of resource density because the full
  consumer partitions a constant amount of energy towards reproduction, whereas
  a hungry consumer partitions no energy towards reproduction.  Similarly, the
  consumer maintenance terms ($\delta H$ and $\beta F$) are also independent of
  resource density because they represent a minimal energetic requirement for
  consumers in the $H$ and $F$ state, respectively.

  % It follows that model predictions are robust only when $R$ is of the order of 1, which holds for all cases that we explore.

  \noindent \paragraph*{ {\bf Steady states of the NSM.}} From the single
  internal fixed point (Eq.~\eqref{eq:ss}, see Methods), an obvious constraint
  on the NSM is that the reproduction rate $\lambda$ must be less than the
  starvation rate $\sigma$, so that the consumer and resource densities are
  positive.
  %In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
  The condition $\sigma = \lambda$ represents a transcritical (TC)
  bifurcation~\citep{Strogatz:2008wo} that demarcates a physical from an
  unphysical (negative steady-state densities) regime.  The biological
  implication of the constraint $\lambda<\sigma$ has a simple
  interpretation---the rate at which a macroscopic organism loses mass due to
  lack of resources is generally much faster than the rate of reproduction.  As
  we will discuss below, this inequality is also a natural consequence of
  allometric constraints~\citep{Kempes:2012hy} for organisms within empirically
  observed body size ranges. %(Fig.~\ref{fig:gvs})

  In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).
  In continuous-time systems, a limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts~\citep{GuckHolmes}.
  This Hopf bifurcation is defined by ${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial of the Jacobian matrix~\citep{Gross:2004p2428}.
  As the system parameters are tuned to be within the stable regime, but close to the Hopf bifurcation, the amplitude of the transient cycles becomes large.
  Given that ecological systems are constantly being perturbed~\citep{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase extinction risk~\citep{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
  %Thus the distance of a system from the Hopf bifurcation provides a measure of its persistence.

  When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the
  consumers are driven to the hungry non-reproducing state.  Because
  reproduction is inhibited, there is a low steady-state consumer density and a
  high steady-state resource density.  However, if $\sigma/\lambda\to 1$ from
  above, the population is overloaded with energetically-replete (reproducing)
  individuals, thereby promoting transient oscillations between the consumer
  and resource densities (Fig.~\ref{fig:fp}).  If the starvation rate is low
  enough that the Hopf bifurcation is crossed, these oscillations become
  stable.  This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.\\

  % Whereas the relation between consumer growth rate $\lambda$ and the starvation rate $\sigma$ defines an absolute bound of biological feasibility---the TC bifurcation---$\sigma$ also determines the sensitivity of the consumer population to changes in resource density.
  % When $\sigma\gg\lambda$, the steady-state population density is small, thereby increasing the risk of stochastic extinction.
  % On the other hand, as $\sigma$ decreases, the system will ultimately be poised either near the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).
  % If the recovery rate $\rho$ is sufficiently small, the TC bifurcation is reached and the resource eventually is eliminated.
  % If $\rho$ exceeds a threshold value, cyclic dynamics will develop as the Hopf bifurcation is approached.

  \begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram-eps-converted-to.pdf}
  \caption{\small{ The growth trajectory over absolute time of an individual organism as a function of body mass.  
  Initial growth follows the black trajectory to an energetically replete reproductive adult mass of $m=\epsilon_\lambda M$ (see Methods). %which we assume is 95\% asymptotic mass $M$.  
  Starvation follows the red trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. 
  Recovery follows the green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires a longer time to reach $\epsilon_\lambda M$. %different energetics and 
  Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}\label{fig:growth}}
  \end{figure}

  % \noindent {\bf Role of allometry} \\
  %[Link Allometry stuff to our model - what does it provide to the story]
  %[Introduce biological and linked constraints]
  %[Allometries capture vast amounts of diversity via a single parameter --- body size]
  %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
  \noindent {\bf Results}
  \noindent \paragraph*{{\bf The allometry of extinction risk.}} While there
  are no {\it a priori} constraints on the parameters in the NSM, we expect
  that each species should be restricted to a distinct portion of the parameter
  space.  We use allometric scaling relations to constrain the covariation of
  rates in a principled and biologically meaningful manner (see Methods).
  Allometric scaling relations highlight common constraints and average trends
  across large ranges in body size and species diversity. Many of these
  relations can be derived from a small set of assumptions.  In the Methods we
  describe our framework to determine the covariation of timescales and rates
  across a range of body sizes for each of the key parameters of our model
  (cf.\ Ref.~\citep{Yodzis:1992hg}).
  % We are thereby able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.

  Nearly all of the rates described in the NSM are determined by consumer
  metabolism, which can be used to describe a variety of organismal features
  \citep{Brown:2004wq}.  We derive, from first principles, the relationships
  for the rates of reproduction, starvation, recovery, and mortality as a
  function of an organism's body size and metabolic rate (see Methods).
  Because we aim to explore the starvation-recovery dynamics as a function of
  an organism's body mass $M$, we parameterize these rates in terms of the
  \emph{percent} gain and loss of the asymptotic (maximum) body mass,
  $\epsilon M$, where different values of $\epsilon$ define different states of
  the consumer (Fig.~\ref{fig:growth}; see Methods for derivations of
  allometrically constrained rate equations).  Although the rate equations
  \eqref{eq:system} are general and can in principle be used to explore the
  starvation recovery dynamics for most organisms, here we focus on allometric
  relationships for terrestrial-bound lower-trophic level endotherms (see the
  SI for values), specifically herbivorous mammals, which range from a minimum
  of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of
  $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene
  Deinotheriinae).  Investigating other classes of organisms would simply
  involve altering the metabolic exponents and scalings associated with
  $\epsilon$. Moreover, we emphasize that our allometric equations (see
  Methods) describe mean relationships, and do not account for the (sometimes
  considerable) variance associated with individual species.  We note that
  including additional allometrically-scaled mortality terms to both $F$ and
  $H$ does not change the form of our model nor impact our quantitative findings
  (see SI for the derivation).

  % \noindent {\bf  Stabilizing effects of allometric constraints} \\
  % As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
  % Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
  % Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
  %Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
  % These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.  However, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation. 
  % This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.\\

  % Previous studies used allometric constraints to explain the periodicity of cyclic populations \citep{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
  % However this relation seems to hold only for some species \citep{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\citep{Kendall:1999iy,Hogstedt:2005cr}.
  % Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\citep{Kendall:1998hl}.
  % However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.\\

  \begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig_ExtinctionAllometricComb4-eps-converted-to.pdf}
  \caption{\small{ Probability of extinction for a consumer with ({\bf a}) $M=10^2$g and ({\bf b}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$. Note the change in scale in panel {\bf b}.  Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state. The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
  \end{figure}

  %Extinction to the left and the right
  % \noindent \paragraph*{{\bf Extinction risk}} 
  As the allometric derivations of the NSM rate laws reveal (see Methods),
  starvation and recovery rates are not independent parameters, and the
  biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp}
  is constrained via covarying parameters.  Given the parameters of terrestrial
  endotherms, we find that the starvation rate $\sigma$ and the recovery rate
  $\rho$ are constrained to lie within a small region of potential values for
  the known range of body sizes $M$.  Indeed, starvation and recovery rates
  across all values of $M$ fall squarely in the steady-state region at some
  distance from the Hopf bifurcation.  This suggests that cyclic population
  dynamics should be rare, particularly in resource-limited environments.
  % Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.

  Higher rates of starvation result in a larger flux of the population to the hungry state.
  In this state, reproduction is absent, thus increasing the likelihood of extinction.  From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
  We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
  To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
  This procedure was repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for organisms with mass ranging from $10^2$ to $10^6$ grams.
  %the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
  In each replicate the initial densities were chosen to be $(XF^*,XH^*,R^*)$,
  with $X$ a random variable uniformly distributed in $[0,2]$.  By allowing the
  rate of starvation to vary, we assessed extinction risk across a range of
  values for $\sigma$ and $\rho$ between ca.\ $10^{-8}$ to
  $10^{-3}$. %, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
  Higher rates of extinction correspond to both large $\sigma$ if $\rho$ is
  small, and large $\rho$ if $\sigma$ is small.  In the former case, increased
  extinction risk arises because of the decrease in the steady-state consumer
  population density (Figs. \ref{fig:fp}b, \ref{fig:ext}).  In the latter case,
  the increased extinction risk results from higher-amplitude transient cycles
  as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).  This
  interplay creates an `extinction refuge', such that for a constrained range
  of $\sigma$ and $\rho$, extinction probabilities are minimized.

  % \noindent \paragraph*{{\bf Extinction refuge.}} 
  We find that the allometrically constrained values of $\sigma$ and $\rho$,
  each representing different trajectories along the ontogenetic curve
  (Fig. \ref{fig:growth}), fall squarely within the extinction refuge across a
  range of $M$ (Fig. \ref{fig:ext}a,b, white points). These values are close
  enough to the Hopf bifurcation to avoid low steady-state densities, yet
  distant enough to avoid large-amplitude transient cycles.  Allometric values
  of $\sigma$ and $\rho$ fall within this relatively small window, which
  supports the possibility that a selective mechanism has constrained the
  physiological conditions driving starvation and recovery rates within
  populations.  Such a mechanism would select for organism physiology that
  generates appropriate $\sigma$ and $\rho$ values that minimize extinction
  risk.  This selection could occur via the tuning of body fat percentages,
  metabolic rates, and/or biomass maintenance efficiencies.  We also find that
  as body size increases, the size of the low extinction-risk parameter space
  shrinks (Fig.~\ref{fig:ext}b), suggesting that the population dynamics for
  larger organisms are more sensitive to variability in physiological rates.
  This finding is in accordance with, and may serve as contributing support for, observations of increased extinction risk among larger mammals \cite{Liow:2008jx}.\\
  % Moreover, larger body sizes decrease the steady state resource density, such that fluctuations for larger organisms will be more likely to drive resources to extinction.
  % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
  % To summarize, our finding that allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\

  %Results are insensitive to alpha but possibly sensitive to beta (eff parameter)

  %NOTES 7/8/16
  %Calder has a hypothesis that period ~ Mass^(1/4).
  %Can we get "transient period" from Jacobian?
  %Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.

  % \noindent {\bf Energetic barriers to body size}
  \noindent \paragraph*{{\bf Damuth's Law and body size limits.}} The NSM
  correctly predicts that smaller species have larger steady-state population
  densities (Fig.~\ref{fig:mass}).  Similar predictions have been made for
  carnivore populations using alternative consumer-resource models
  \citep{DeLong:2012kw}.  Moreover, we show that the NSM provides independent
  theoretical support for Damuth's Law
  \citep{Damuth:1987kr,allen2002,enquist1998,Pedersen:2017he}.  Damuth's
  law shows that species abundances, $N^{*}$, follow $N^*=0.01
  M^{-0.78}$ (g m$^{-2}$). Figure \ref{fig:mass} shows that both $F^{*}$ and $H^{*}$ scale
  as $M^{-\eta}$, with $\eta\approx 3/4$,  over a wide range of organismal sizes and that $F^{*}+H^{*}$
  closely matches the best fit to Damuth's data.  Remarkably, this result
  illustrates that the steady state values of the NSM combined with the derived
  timescales naturally give rise to Damuth's law. While the previous metabolic
  studies supporting Damuth's law provided arguments for the value of the
  exponent~\citep{allen2002}, these studies are only able to infer the
  normalization constant ($0.01$ g$^{1.78}$ m$^{-2}$ in the above equation) from the data (see SI for a discussion of the energy
  equivalence hypothesis related to these metabolic arguments). Our model
  predicts not only the exponent but also the normalization constant by
  explicitly including the resource dynamics and the parameters that determine
  growth and consumption. It should be noted that density relationships of
  individual clades follow a more shallow scaling relationship than predicted
  by Damuth's law~\cite{Pedersen:2017he}.  In the context of our model,
  this finding suggests that future work may be able to anticipate these shifts
  by accounting for differences in the physiological parameters associated with
  each clade.

  \begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig_FPAllometric2-eps-converted-to.pdf}
  \caption{\small{Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
    body mass along with the data from Damuth \citep{Damuth:1987kr}. Inset: Resource steady state $R^*$ as a function of consumer body mass.}\label{fig:mass}}
  \end{figure}

  % \paragraph*{{\bf Maximum Body Size}}
  %JUSTIN I REMOVED: Our model shows that energetic equivalence breaks down at large $M$ suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.  
  %, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
  With respect to predicted steady state densities, the total metabolic rate of
  $F$ and $H$ becomes infinite at a finite mass, and occurs at the same scale
  where the steady state resources vanish (Fig.~\ref{fig:mass}). This
  asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and
  $\epsilon_{\lambda}$ (see Fig.~\ref{fig:growth}) equal zero, causing the
  timescales (Eqn. \ref{t1}) to become infinite and the rates $\mu$ and $\lambda$ to equal
  zero.
  % A theoretical upper bound on mammalian body size is given by
  % $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic
  % reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass
  % of a male African elephant.
  The $\mu=0$ asymptote occurs first when
  $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to
  $(F^*,H^*,R^*)=(0,0,0)$.  This point predicts an upper bound on mammalian
  body size at $M_{\rm max}=6.54\times10^7$ (g).  Moreover, $M_{\rm max}$,
  which is entirely determined by the population-level consequences of
  energetic constraints, is within an order of magnitude of the maximum body
  size observed in the North American mammalian fossil
  record~\citep{Alroy:1998p1594}, as well as the mass predicted from an
  evolutionary model of body size evolution~\citep{Clauset:2009fh}.  We
  emphasize that the asymptotic behavior and predicted upper bound depend only
  on the scaling of body composition and are independent of the resource
  parameters.  The prediction of an asymptotic limit on mammalian size
  parallels work on microbial life where an upper and lower bound on bacterial
  size, and an upper bound on single cell eukaryotic size, is predicted from
  similar growth and energetic scaling
  relationships~\citep{Kempes:2012hy,Kempes:2016}.  It has also been shown that
  models that incorporate the allometry of hunting and resting combined with
  foraging time predicts a maximum carnivore size between $7\times10^{5}$ and
  $1.1\times10^{6}$ (g) \cite{Carbone:1999ju,Carbone:2007dz}.  Similarly, the
  maximum body size within a particular lineage has been shown to scale with
  the metabolic normalization constant
  \cite{Okie:2013ju}. %and depend on a critical death found to be constant from data
  This complementary approach is based on the balance between growth and
  mortality, and suggests that future connections between the scaling of fat
  and muscle mass should systematically be connected with $B_{0}$ when
  comparing
  lineages. % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here

  % {\color{red}
  % 
  % It should be noted that in our model all death is related to nutritional state, however the addition of a constant death term \cite{} would only adjust the values of $\lambda$ and $\mu$ without changing the form or or qualitative results of our model. We explored the sensitivity of our results to shifts in a constant death term (see SI) and find that additional sources of death do not shift our quantitative prediction of Damuth's law except for large sizes (Figure S\ref{}). We find that additional sources of death shifts the asymptotic behavior of our model and decreases the maximum possible mammal size. Within the range of reasonable death rates $M_{max}$ varies by $VALUE\%$. These results suggest that the largest mammals are most sensitive to additional sources of death perhaps putting them at greater extinction risk.
  % 
  % }

  %Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end.
  %Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is close to the minimum observed mammalian body size of ca.\ 1.3-2.5g, a range that occurs as the recovery rate begins its decline.
  %In addition to known transport limitations~\citep{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.
  %This result mirrors other efforts \citep{Kempes:2012hy,Kempes:2016} where at a given scale multiple limitations constrain the smallest possibilities for life within a class of organisms.

  %The combined steady-state dynamics and allometric timescales of the NSM predict that larger mammals should outcompete smaller ones, and this suggests that the model may provide a framework with which to explore the energetic mechanisms behind Cope's rule. %, which we explore further below.

  \noindent \paragraph*{{\bf A mechanism for Cope's rule}} Metabolite transport
  constraints are widely thought to place strict boundaries on biological
  scaling~\citep{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to
  specific predictions on the minimum possible body size for
  organisms~\citep{West:2002ud}.  Above this bound, a number of energetic and
  evolutionary mechanisms have been explored to assess the costs and benefits
  associated with larger body masses, particularly for mammals.  One important
  such example is the \emph{fasting endurance hypothesis}, which contends that
  larger body size, with consequent lower metabolic rates and increased ability
  to maintain more endogenous energetic reserves, may buffer organisms against
  environmental fluctuations in resource availability~\citep{Millar:1990p923}.
  Over evolutionary time, terrestrial mammalian lineages show a significant
  trend towards larger body size---Cope's
  rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}.
  It is thought that within-lineage drivers generate selection towards an
  optimal upper bound of roughly $10^7$ (g)~\citep{Alroy:1998p1594}, a value
  that is likely limited by higher extinction risk for large taxa over longer
  timescales~\citep{Clauset:2009fh}.  These trends are thought to be driven by
  a combination of climate change and niche
  availability~\citep{Saarinen:2014br}; however the underpinning energetic
  costs and benefits of larger body sizes, and how they influence dynamics over
  ecological timescales, have not been explored.
  % We argue that the NSM provides a suitable framework to explore these issues.

  % \noindent{\bf A within-lineage driver of Cope's Rule}
  % \paragraph*{{\bf Body Size Competition}} 
  The NSM predicts that the steady state resource density $R^{*}$ decreases
  with increasing body size of the consumer population (Fig.~\ref{fig:mass},
  inset), and classic resource competition theory predicts that the species
  surviving on the lowest resource abundance will outcompete others
  \citep{tilman1981,dutkiewicz2009,barton2010}. Thus, the combined NSM
  steady-state dynamics and allometric timescales (see Eq.~\eqref{t1}) predict
  that larger mammals have an intrinsic competitive advantage given a common
  resource.  
  % We will now show that the NSM indeed provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time known as Cope's Rule.

  \begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{fig_Invasion-eps-converted-to.pdf}
  \caption{\small{Competitive outcomes for a resident species with body mass
      $M$ vs. a closely related competing species with modified body mass
      $M^\prime=M(1+\chi)$.  The blue region denotes proportions of modified
      mass $\chi$ resulting in exclusion of the resident species.  The red
      region denotes values of $\chi$ that result in a mass that is below the
      starvation threshold and are thus infeasible.  Arrows point to the
      predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$,
      which may serve as an evolutionary attractor for body mass.  The black
      wedge points to the largest body mass known for terrestrial mammals
      (\emph{Deinotherium} spp.) at $1.74\times10^7$
      (g)~\citep{Smith:2010p3442}.}\label{fig:invasion}}
  \end{figure}

  % Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.

  However, the above resource relationships do not offer a mechanism for how
  body size is selected.  We directly assess competitive outcome between two
  closely related species: a resident species of mass $M$, and a competing
  species (denoted by $^\prime$) where individuals have a different proportion
  of body fat such that $M^\prime=M(1+\chi)$.  For $\chi < 0$, the competing
  individuals have fewer metabolic reserves than the resident species and vice
  versa for $\chi>0$.  For the allowable values of $\chi$ (see SI), the mass of
  the competitor $M'$ should exceed the minimal amount of body fat,
  $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be
  positive, which, given Eq.~\ref{t1}, implies that
  $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.  These
  conditions imply that $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$
  where the upper bound approximately equals $0.05$ and the lower bound is
  mass-dependent.  The modified mass of the competitor leads to altered rates
  of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the
  maintenance of both starving $\delta(M^\prime)$ and full consumers
  $\beta(M^\prime)$ (see the SI for derivations of competitor rates).
  Importantly, $\epsilon_\sigma$, which determines the point along the growth
  curve that defines the body composition of starved foragers, is assumed to
  remain unchanged for the competing population (see SI).
  %There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).

  To assess the susceptibility of the resident species to competitive
  exclusion, we determine which consumer pushes the steady-state resource
  density $R^*$ to lower values for a given value of $\chi$, with the
  expectation that a population capable of surviving on lower resource
  densities has a competitive advantage \citep{tilman1981}.  We find that for
  $M\leq 1.748\times10^7$ (g), having additional body fat ($\chi > 0$) results
  in a lower steady state resource density ($R^{\prime *}<R^*$), such that the
  competitor has an intrinsic advantage over the resident species
  (Fig. \ref{fig:invasion}).  However, for $M> 1.748\times10^7$ (g), leaner
  individuals ($\chi < 0$) have lower resource steady state densities.

  %, and this is due to the changing
  %covariance between energetic rates as a function of modified energetic
  %reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.

  The observed switch in susceptibility as a function of $\chi$ at
  $M_{\rm opt}= 1.748\times10^7$ (g) thus serves as an attractor, such that the
  NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if
  $M>M_{\rm opt}$.  This value is close to but smaller than the asymptotic
  upper bound for terrestrial mammal body size predicted by the NSM, and is remarkably close to independent estimates of the largest land
  mammals, the early Oligocene \emph{Indricotherium} at $\approx$
  $1.5\times10^7$ (g) and the late Miocene \emph{Deinotherium} at $\approx$
  $1.74\times10^7$ (g) ~\citep{Smith:2010p3442}.  Additionally, our calculation
  of $M_{\rm opt}$ as a function of mass-dependent physiological rates is
  similar to theoretical estimates of maximum body size \citep{Clauset:2009fh},
  and provides independent theoretical support for the observation of a
  `maximum body size attractor' explored by Alroy~\citep{Alroy:1998p1594}.
  % {\color{red} We note that the model of \cite{Brown:1993p708} predicts

  An optimal size for mammals at intermediate body mass was predicted by Brown et al.\ based on reproductive maximization and the transition between hungry and full individuals \cite{Brown:1993p708}. 
  By coupling the NSM to resource dynamics as well as introducing an explicit treatment of storage, we show that species with larger body masses have an inherent competitive advantage for size classes up to $M_{\rm opt}= 1.748\times10^7$ based on resource competition. Moreover, the mass distributions in Ref. \cite{Brown:1993p708} show that intermediate mammal sizes have the greatest species diversity, in contrast to our efforts, which consider total biomass and predict a much larger $M_{\rm opt}$.
  Compellingly, recent work shows that many communities can be dominated by the biomass of the large \cite{Hempson:2015hka}.
  % shows that the largest mammals have an advantage in terms resource competition. 
  %as well as  \citep{Alroy:1998p1594,Clauset:2009fh}.
  %This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
  While the state of the environment as well as the competitive landscape will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general selective mechanism for the evolution of larger body size among terrestrial mammals.\\

  \noindent {\bf Discussion} \\

  %and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.

  %One might be concerned a greater number of large mammals are currently not observed in the modern world given that larger mammals are less susceptible to extinction.
  %However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \citep{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also much more geographically widespread than today.
  %These results, combined with our findings, suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \citep{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.

  %Closing
  The energetics associated with somatic maintenance, growth, and reproduction
  are important elements that influence the dynamics of all
  populations~\citep{Stearns:1989ip}.  The NSM incorporates the dynamics of
  starvation and recovery that are expected to occur in resource-limited
  environments.  We found that incorporating allometrically-determined rates
  into the NSM predicts that: (i) extinction risk is minimized, (ii) the
  derived steady-states quantitatively reproduce Damuth's law, and (iii) the
  selective mechanism for the evolution of larger body sizes agrees with Cope's
  rule.  The NSM offers a means by which the dynamic consequences of energetic
  constraints can be assessed using macroscale interactions between and among
  species.

  %Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.

  \section*{Methods}
  \small{
  {\bf Analytical solution to the NSM}
  Equation~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
  \begin{align}
  \label{eq:ss}
  \begin{split}
  F^* &= (\sigma-\lambda)\frac{ \alpha  \lambda  \mu ^2  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
  H^* &= (\sigma-\lambda)\frac{ \alpha  \lambda ^2 \mu  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
  R^* &= (\sigma - \lambda)\frac{\mu  }{A}.
  \end{split}
  \end{align}
  where $A=(\lambda \xi \rho +\mu \sigma )$ and
  $B=(\beta \mu \xi +\delta \lambda \xi -\lambda \mu )$. The stability of this
  fixed point is determined by the Jacobian matrix $\bf J$, with
  $J_{ij}=\partial{\dot X_i}/\partial{X_j}$, when evaluated at the internal
  fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.  The parameters in
  Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue
  of $\mathbf{J}$ is negative, so that the system is stable with respect to
  small perturbations from the fixed point.  Because this fixed point is
  unique, it is the global attractor for all population trajectories for any
  initial condition where the resource and consumer densities are both nonzero.

  {\bf Metabolic scaling relationships} The scaling relation between an
  organism's metabolic rate $B$ and its body mass $M$ at reproductive maturity
  is known to scale as $B = B_0 M^\eta$, where the scaling exponent $\eta$ is
  typically close to $2/3$ or $3/4$ for metazoans (e.g.,
  Ref.~\citep{West:2002it,Brown:2004wq}), and has taxonomic shifts for
  unicellular species between $\eta\approx 1$ in eukaryotes and
  $\eta\approx 1.76$ in bacteria \citep{DeLong:2010dy,Kempes:2012hy}.
  %Justin adding an explanation for beta

  Several efforts have shown how a partitioning of $B$ between growth and
  maintenance purposes can be used to derive a general equation for both the
  growth trajectories and growth rates of organisms ranging from bacteria to
  metazoans
  \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Savage:2004ed,Kempes:2012hy}. This relation is derived from the simple balance condition 
  $B_{0}m^{\eta}=E_{m}\dot{m}+B_{m}m\,,$
  % \begin{eqnarray}
  % \label{balance}
  % B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
  % \end{eqnarray}
  \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Savage:2004ed,Kempes:2012hy} where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
  the metabolic rate to support an existing unit of mass, and $m$ is the mass
  of the organism at any point in its development.  This balance has the
  general solution \citep{bettencourt,Kempes:2012hy}
  \begin{eqnarray}
  \label{m1}
  \left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}},
  \end{eqnarray}
  where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass,
  $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically
  (see the SI).  We now use this solution to define the timescale for
  reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see
  \citep{moses2008rmo} for a detailed presentation of these timescales). The
  time that an organism takes to reach a particular mass $\epsilon M$ is given
  by the timescale
  \begin{equation}
  \label{t1}
  \tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)},
  \end{equation}
  where we define values of $\epsilon$ below to describe a variety of
  timescales, along with the rates related to $\tau$.  For example, the rate of
  reproduction is given by the timescale to go from the birth mass to the adult
  mass. The time to reproduce is given by Equation \ref{t1} as
  $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
  is the fraction of the asymptotic mass where an organism is reproductively
  mature and should be close to one (typically
  $\epsilon_{\lambda}\approx0.95\;$ \citep{West:2001bv}). Our reproductive
  rate, $\lambda$, is a specific rate, or the number of offspring produced per
  time per individual, defined as $\dot{F} = \lambda F$. In isolation this
  functional form gives the population growth
  $F\left(t\right) = F_{0}e^{\lambda t}$ which can be related to the
  reproductive timescale by assuming that when $t=t_{\lambda}$ it is also the
  case that $F=\nu F_{0}$, where $\nu-1$ is the number of offspring produced
  per reproductive cycle. Following this relationship the growth rate is given
  by $\lambda=\ln\left(\nu\right)/t_{\lambda}$, which is the standard
  relationship (e.g.,~\cite{Savage:2004ed}) and will scales as
  $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of
  $\epsilon_{\lambda}$
  \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.

  The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
  sufficient tissue to transition from the hungry to the full state.  Since
  only certain tissues can be digested for energy (for example the brain cannot
  be degraded to fuel metabolism), we define the rates for starvation, death,
  and recovery by the timescales required to reach, or return from, specific
  fractions of the replete-state mass (see the SI, Table I, for
  parameterizations).  We define $m_{\sigma}=\epsilon_{\sigma} M$, where
  $\epsilon_{\sigma}<1$ is the fraction of replete-state mass where
  reproduction ceases. This fraction will deviate from a constant if tissue
  composition systematically scales with adult mass.  For example, making use
  of the observation that body fat in mammals scales with overall body size
  according to $M_{\rm fat}=f_{0}M^{\gamma}$ and assuming that once this mass
  is fully digested the organism starves, this would imply that
  $\epsilon_{\sigma}=1-f_{0}M^{\gamma}/M$. It follows that the recovery
  timescale, $t_{\rho}$, is the time to go from mass
  $m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$
  (Fig. \ref{fig:growth}). Using Eqs.~\eqref{m1} and \eqref{t1} this timescale
  is given by simply considering the growth curve starting from a mass of
  $m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
  \begin{eqnarray}
  \label{rhotimescale}
  t_{\rho}=\ln\left[\frac{1-\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)^{1-\eta}}{1-\epsilon_\lambda^{1-\eta}}\right]\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}
  \end{eqnarray}
  where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in
  the biosynthetic energetics during recovery (see the SI). It should be noted that more complicated ontogenetic models explicitly handle
  storage \citep{hou}, whereas this feature is implicitly covered by the body
  fat scaling in our framework.

  To determine the starvation rate, $\sigma$, we are interested in the time
  required for an organism to go from a mature adult that reproduces at rate
  $\lambda$, to a reduced-mass hungry state where reproduction is impossible.
  For starving individuals we assume that an organism must meet its maintenance
  requirements by using the digestion of existing mass as the sole energy
  source.  This assumption implies the metabolic balance
  $\dot{m}E_{m}^{\prime}=-B_{m}m$ or $\dot{m}=-a^{\prime}m/M^{1-\eta}$
  % \begin{eqnarray}
  % \frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
  % \end{eqnarray}
  % or
  % \begin{eqnarray}
  % \frac{dm}{dt}=-\frac{a^{\prime}}{M^{1-\eta}}m
  % \end{eqnarray}
  where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
  body mass, which differs from $E_{m}$, the energy required to
  synthesis a unit of biomass \citep{hou}. Given the replete mass, $M$, of an organism, the
  above energy balance prescribes the mass trajectory of a non-consuming
  organism: $m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}$.
  % \begin{eqnarray}
  % \label{mt}
  % m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}.
  % \end{eqnarray}
  The timescale for starvation is
  given by the time it takes $m(t)$ to reach $\epsilon_{\sigma} M$, which gives
  \begin{equation}
  \label{eq:sigma}
  t_{\sigma}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\sigma}\right).
  \end{equation}
  The starvation rate is then $\sigma=1/t_{\sigma}$, which scales with
  replete-state mass as $1/M^{1-\eta}\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important
  feature is that $\sigma$ does not have a simple scaling dependence on
  $\lambda$, which is important for the dynamics that we
  later discuss.

  The time to death should follow a similar relation, but defined by a lower
  fraction of replete-state mass, $m_{\mu}=\epsilon_{\mu} M$ where $\epsilon_\mu < \epsilon_\sigma$.
  Suppose, for example, that an organism dies once it has digested all fat and
  muscle tissues, and that muscle tissue scales with body mass according to
  $M_{\rm musc}=u_{0}M^{\zeta}$.  This gives
  $\epsilon_{\mu}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
  mass has been shown to be roughly proportional to body mass~\citep{Folland:2008ij} in
  mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant. The time to go from starvation to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or $t_{\mu}=-M^{1-\eta}\ln\left(\epsilon_{\mu}\right)/a^{\prime}-t_{\sigma}$,
  % \begin{eqnarray}
  % \label{mutimescale}
  % t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
  % \end{eqnarray}
  and $\mu=1/t_{\mu}$.
  }

\def\bibfont{\footnotesize}
% \putbib[aa_starving3]

\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\providecommand{\bibinfo}[2]{#2}
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem{Martin:1987dl}
\bibinfo{author}{Martin, T.~E.}
\newblock \bibinfo{title}{{Food as a limit on breeding birds: A life-history
  perspective}}.
\newblock \emph{\bibinfo{journal}{Annu. Rev. Ecol. Syst.}}
  \textbf{\bibinfo{volume}{18}}, \bibinfo{pages}{453--487}
  (\bibinfo{year}{1987}).

\bibitem{Kirk:1997cc}
\bibinfo{author}{Kirk, K.~L.}
\newblock \bibinfo{title}{{Life-history responses to variable environments:
  Starvation and reproduction in planktonic rotifers}}.
\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{78}},
  \bibinfo{pages}{434--441} (\bibinfo{year}{1997}).

\bibitem{Kempes:2012hy}
\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Dutkiewicz, S.} \&
  \bibinfo{author}{Follows, M.~J.}
\newblock \bibinfo{title}{{Growth, metabolic partitioning, and the size of
  microorganisms}}.
\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
  \textbf{\bibinfo{volume}{109}}, \bibinfo{pages}{495--500}
  (\bibinfo{year}{2012}).

\bibitem{Mangel:1988uaa}
\bibinfo{author}{Mangel, M.} \& \bibinfo{author}{Clark, C.~W.}
\newblock \emph{\bibinfo{title}{{Dynamic Modeling in Behavioral Ecology}}}
  (\bibinfo{publisher}{Princeton University Press},
  \bibinfo{address}{Princeton}, \bibinfo{year}{1988}).

\bibitem{Mangel:2014kz}
\bibinfo{author}{Mangel, M.}
\newblock \bibinfo{title}{{Stochastic dynamic programming illuminates the link
  between environment, physiology, and evolution}}.
\newblock \emph{\bibinfo{journal}{B. Math. Biol.}}
  \textbf{\bibinfo{volume}{77}}, \bibinfo{pages}{857--877}
  (\bibinfo{year}{2014}).

\bibitem{Yeakel:2013hi}
\bibinfo{author}{Yeakel, J.~D.}, \bibinfo{author}{Dominy, N.~J.},
  \bibinfo{author}{Koch, P.~L.} \& \bibinfo{author}{Mangel, M.}
\newblock \bibinfo{title}{{Functional morphology, stable isotopes, and human
  evolution: a model of consilience}}.
\newblock \emph{\bibinfo{journal}{Evolution}} \textbf{\bibinfo{volume}{68}},
  \bibinfo{pages}{190--203} (\bibinfo{year}{2014}).

\bibitem{Morris:1987eo}
\bibinfo{author}{Morris, D.~W.}
\newblock \bibinfo{title}{{Optimal allocation of parental investment}}.
\newblock \emph{\bibinfo{journal}{Oikos}} \textbf{\bibinfo{volume}{49}},
  \bibinfo{pages}{332--339} (\bibinfo{year}{1987}).

\bibitem{Tveraa:2003fq}
\bibinfo{author}{Tveraa, T.}, \bibinfo{author}{Fauchald, P.},
  \bibinfo{author}{Henaug, C.} \& \bibinfo{author}{Yoccoz, N.~G.}
\newblock \bibinfo{title}{{An examination of a compensatory relationship
  between food limitation and predation in semi-domestic reindeer}}.
\newblock \emph{\bibinfo{journal}{Oecologia}} \textbf{\bibinfo{volume}{137}},
  \bibinfo{pages}{370--376} (\bibinfo{year}{2003}).

\bibitem{Daan:1988va}
\bibinfo{author}{Daan, S.}, \bibinfo{author}{Dijkstra, C.},
  \bibinfo{author}{Drent, R.} \& \bibinfo{author}{Meijer, T.}
\newblock \bibinfo{title}{{Food supply and the annual timing of avian
  reproduction}}.
\newblock In \bibinfo{editor}{Ouellet, H.} (ed.) \emph{\bibinfo{booktitle}{Acta
  XIX Congressus Internationalis Ornithologici, Volume I: Proceedings XIX
  International Ornithological Congress, 1986, Ottawa}},
  \bibinfo{pages}{392--407} (\bibinfo{publisher}{Proceedings XIX International
  Ornithological Congress}, \bibinfo{address}{Ottawa}, \bibinfo{year}{1989}).

\bibitem{Jacot:2009dv}
\bibinfo{author}{Jacot, A.}, \bibinfo{author}{Valcu, M.}, \bibinfo{author}{van
  Oers, K.} \& \bibinfo{author}{Kempenaers, B.}
\newblock \bibinfo{title}{{Experimental nest site limitation affects
  reproductive strategies and parental investment in a hole-nesting
  passerine}}.
\newblock \emph{\bibinfo{journal}{Animal Behaviour}}
  \textbf{\bibinfo{volume}{77}}, \bibinfo{pages}{1075--1083}
  (\bibinfo{year}{2009}).

\bibitem{Stearns:1989ip}
\bibinfo{author}{Stearns, S.~C.}
\newblock \bibinfo{title}{{Trade-offs in life-history evolution}}.
\newblock \emph{\bibinfo{journal}{Funct. Ecol.}} \textbf{\bibinfo{volume}{3}},
  \bibinfo{pages}{259} (\bibinfo{year}{1989}).

\bibitem{Barboza:2002in}
\bibinfo{author}{Barboza, P.} \& \bibinfo{author}{Jorde, D.}
\newblock \bibinfo{title}{{Intermittent fasting during winter and spring
  affects body composition and reproduction of a migratory duck}}.
\newblock \emph{\bibinfo{journal}{J Comp Physiol B}}
  \textbf{\bibinfo{volume}{172}}, \bibinfo{pages}{419--434}
  (\bibinfo{year}{2002}).

\bibitem{Threlkeld:1976ih}
\bibinfo{author}{Threlkeld, S.~T.}
\newblock \bibinfo{title}{{Starvation and the size structure of zooplankton
  communities}}.
\newblock \emph{\bibinfo{journal}{Freshwater Biol.}}
  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{489--496}
  (\bibinfo{year}{1976}).

\bibitem{Weber:1998jg}
\bibinfo{author}{Weber, T.~P.}, \bibinfo{author}{Ens, B.~J.} \&
  \bibinfo{author}{Houston, A.~I.}
\newblock \bibinfo{title}{{Optimal avian migration: A dynamic model of fuel
  stores and site use}}.
\newblock \emph{\bibinfo{journal}{Evolutionary Ecology}}
  \textbf{\bibinfo{volume}{12}}, \bibinfo{pages}{377--401}
  (\bibinfo{year}{1998}).

\bibitem{Mduma:1999cp}
\bibinfo{author}{Mduma, S. A.~R.}, \bibinfo{author}{Sinclair, A. R.~E.} \&
  \bibinfo{author}{Hilborn, R.}
\newblock \bibinfo{title}{{Food regulates the Serengeti wildebeest: a 40-year
  record}}.
\newblock \emph{\bibinfo{journal}{J. Anim. Ecol.}}
  \textbf{\bibinfo{volume}{68}}, \bibinfo{pages}{1101--1122}
  (\bibinfo{year}{1999}).

\bibitem{Moore:2014hi}
\bibinfo{author}{Moore, J.~W.}, \bibinfo{author}{Yeakel, J.~D.},
  \bibinfo{author}{Peard, D.}, \bibinfo{author}{Lough, J.} \&
  \bibinfo{author}{Beere, M.}
\newblock \bibinfo{title}{{Life-history diversity and its importance to
  population stability and persistence of a migratory fish: steelhead in two
  large North American watersheds}}.
\newblock \emph{\bibinfo{journal}{J. Anim. Ecol.}}
  \textbf{\bibinfo{volume}{83}}, \bibinfo{pages}{1035--1046}
  (\bibinfo{year}{2014}).

\bibitem{Mead:1989dt}
\bibinfo{author}{Mead, R.~A.}
\newblock \bibinfo{title}{{The Physiology and Evolution of Delayed Implantation
  in Carnivores}}.
\newblock In \bibinfo{editor}{Gittleman, J.~L.} (ed.)
  \emph{\bibinfo{booktitle}{Carnivore Behavior, Ecology, and Evolution}},
  \bibinfo{pages}{437--464} (\bibinfo{publisher}{Springer US},
  \bibinfo{address}{Ithaca}, \bibinfo{year}{1989}).

\bibitem{Sandell:1990kw}
\bibinfo{author}{Sandell, M.}
\newblock \bibinfo{title}{{The evolution of seasonal delayed implantation}}.
\newblock \emph{\bibinfo{journal}{Q Rev Biol}} \textbf{\bibinfo{volume}{65}},
  \bibinfo{pages}{23--42} (\bibinfo{year}{1990}).

\bibitem{Bulik:1999eo}
\bibinfo{author}{Bulik, C.~M.} \emph{et~al.}
\newblock \bibinfo{title}{{Fertility and reproduction in women with anorexia
  nervosa}}.
\newblock \emph{\bibinfo{journal}{J. Clin. Psychiat.}}
  \textbf{\bibinfo{volume}{60}}, \bibinfo{pages}{130--135}
  (\bibinfo{year}{1999}).

\bibitem{Trites:2003cc}
\bibinfo{author}{Trites, A.~W.} \& \bibinfo{author}{Donnelly, C.~P.}
\newblock \bibinfo{title}{{The decline of Steller sea lions Eumetopias jubatus
  in Alaska: a review of the nutritional stress hypothesis}}.
\newblock \emph{\bibinfo{journal}{Mammal Rev.}} \textbf{\bibinfo{volume}{33}},
  \bibinfo{pages}{3--28} (\bibinfo{year}{2003}).

\bibitem{Glazier:2009hq}
\bibinfo{author}{Glazier, D.~S.}
\newblock \bibinfo{title}{{Metabolic level and size scaling of rates of
  respiration and growth in unicellular organisms}}.
\newblock \emph{\bibinfo{journal}{Funct. Ecol.}} \textbf{\bibinfo{volume}{23}},
  \bibinfo{pages}{963--968} (\bibinfo{year}{2009}).

\bibitem{Kooi2000}
\bibinfo{author}{Kooijman, S. A. L.~M.}
\newblock \emph{\bibinfo{title}{{Dynamic Energy and Mass Budgets in Biological
  Systems}}} (\bibinfo{address}{Cambridge}, \bibinfo{year}{2000}).

\bibitem{Sousa:2010ez}
\bibinfo{author}{Sousa, T.}, \bibinfo{author}{Domingos, T.},
  \bibinfo{author}{Poggiale, J.~C.} \& \bibinfo{author}{Kooijman, S. A. L.~M.}
\newblock \bibinfo{title}{{Dynamic energy budget theory restores coherence in
  biology}}.
\newblock \emph{\bibinfo{journal}{Philos. T. Roy. Soc. B}}
  \textbf{\bibinfo{volume}{365}}, \bibinfo{pages}{3413--3428}
  (\bibinfo{year}{2010}).

\bibitem{Diekmann:2010da}
\bibinfo{author}{Diekmann, O.} \& \bibinfo{author}{Metz, J. A.~J.}
\newblock \bibinfo{title}{{How to lift a model for individual behaviour to the
  population level?}}
\newblock \emph{\bibinfo{journal}{Philos. T. Roy. Soc. B}}
  \textbf{\bibinfo{volume}{365}}, \bibinfo{pages}{3523--3530}
  (\bibinfo{year}{2010}).

\bibitem{murdoch:2003}
\bibinfo{author}{Murdoch, W.~W.}, \bibinfo{author}{Briggs, C.~J.} \&
  \bibinfo{author}{Nisbet, R.~M.}
\newblock \emph{\bibinfo{title}{{Consumer-resource Dynamics}}},
  vol.~\bibinfo{volume}{36} of \emph{\bibinfo{series}{Monographs in population
  biology}} (\bibinfo{publisher}{Princeton University Press},
  \bibinfo{address}{Princeton}, \bibinfo{year}{2003}).

\bibitem{Damuth:1987kr}
\bibinfo{author}{Damuth, J.}
\newblock \bibinfo{title}{{Interspecific allometry of population density in
  mammals and other animals: the independence of body mass and population
  energy-use}}.
\newblock \emph{\bibinfo{journal}{Biol. J. Linn. Soc.}}
  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
  (\bibinfo{year}{1987}).

\bibitem{allen2002}
\bibinfo{author}{Allen, A.~P.}, \bibinfo{author}{Brown, J.~H.} \&
  \bibinfo{author}{Gillooly, J.~F.}
\newblock \bibinfo{title}{{Global biodiversity, biochemical kinetics, and the
  energetic-equivalence rule}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{297}},
  \bibinfo{pages}{1545--1548} (\bibinfo{year}{2002}).

\bibitem{enquist1998}
\bibinfo{author}{Enquist, B.~J.}, \bibinfo{author}{Brown, J.~H.} \&
  \bibinfo{author}{West, G.~B.}
\newblock \bibinfo{title}{{Allometric scaling of plant energetics and
  population density}}.
\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{395}},
  \bibinfo{pages}{163--165} (\bibinfo{year}{1998}).

\bibitem{Pedersen:2017he}
\bibinfo{author}{Pedersen, R.~{\O}.}, \bibinfo{author}{Faurby, S.} \&
  \bibinfo{author}{Svenning, J.-C.}
\newblock \bibinfo{title}{{Shallow size{\textendash}density relations within
  mammal clades suggest greater intra-guild ecological impact of large-bodied
  species}}.
\newblock \emph{\bibinfo{journal}{J. Anim. Ecol.}}
  \textbf{\bibinfo{volume}{86}}, \bibinfo{pages}{1205--1213}
  (\bibinfo{year}{2017}).

\bibitem{Alroy:1998p1594}
\bibinfo{author}{Alroy, J.}
\newblock \bibinfo{title}{{Cope's rule and the dynamics of body mass evolution
  in North American fossil mammals}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{280}},
  \bibinfo{pages}{731--734} (\bibinfo{year}{1998}).

\bibitem{Clauset:2009fh}
\bibinfo{author}{Clauset, A.} \& \bibinfo{author}{Redner, S.}
\newblock \bibinfo{title}{{Evolutionary model of species body mass
  diversification}}.
\newblock \emph{\bibinfo{journal}{Phys. Rev. Lett.}}
  \textbf{\bibinfo{volume}{102}}, \bibinfo{pages}{038103}
  (\bibinfo{year}{2009}).

\bibitem{Smith:2010p3442}
\bibinfo{author}{Smith, F.~A.} \emph{et~al.}
\newblock \bibinfo{title}{{The evolution of maximum body size of terrestrial
  mammals}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{330}},
  \bibinfo{pages}{1216--1219} (\bibinfo{year}{2010}).

\bibitem{Saarinen:2014br}
\bibinfo{author}{Saarinen, J.~J.} \emph{et~al.}
\newblock \bibinfo{title}{{Patterns of maximum body size evolution in Cenozoic
  land mammals: Eco-evolutionary processes and abiotic forcing}}.
\newblock \emph{\bibinfo{journal}{Proc Biol Sci}}
  \textbf{\bibinfo{volume}{281}}, \bibinfo{pages}{20132049}
  (\bibinfo{year}{2014}).

\bibitem{Benichou:2014wu}
\bibinfo{author}{B{\'e}nichou, O.} \& \bibinfo{author}{Redner, S.}
\newblock \bibinfo{title}{{Depletion-controlled starvation of a diffusing
  forager.}}
\newblock \emph{\bibinfo{journal}{Phys. Rev. Lett.}}
  \textbf{\bibinfo{volume}{113}}, \bibinfo{pages}{238101}
  (\bibinfo{year}{2014}).

\bibitem{Benichou:2016wl}
\bibinfo{author}{B{\'e}nichou, O.}, \bibinfo{author}{Chupeau, M.} \&
  \bibinfo{author}{Redner, S.}
\newblock \bibinfo{title}{{Role of depletion on the dynamics of a diffusing
  forager}}.
\newblock \emph{\bibinfo{journal}{J Phys A-Math Theor}}
  \textbf{\bibinfo{volume}{49}}, \bibinfo{pages}{394003}
  (\bibinfo{year}{2016}).

\bibitem{Chupeau:2016jf}
\bibinfo{author}{Chupeau, M.}, \bibinfo{author}{B{\'e}nichou, O.} \&
  \bibinfo{author}{Redner, S.}
\newblock \bibinfo{title}{{Universality classes of foraging with resource
  renewal}}.
\newblock \emph{\bibinfo{journal}{Phys. Rev. E}} \textbf{\bibinfo{volume}{93}},
  \bibinfo{pages}{032403} (\bibinfo{year}{2016}).

\bibitem{Persson:1998hz}
\bibinfo{author}{Persson, L.}, \bibinfo{author}{Leonardsson, K.},
  \bibinfo{author}{De~Roos, A.~M.}, \bibinfo{author}{Gyllenberg, M.} \&
  \bibinfo{author}{Christensen, B.}
\newblock \bibinfo{title}{{Ontogenetic scaling of foraging rates and the
  dynamics of a size-structured consumer-resource model}}.
\newblock \emph{\bibinfo{journal}{Theor Popul Biol}}
  \textbf{\bibinfo{volume}{54}}, \bibinfo{pages}{270--293}
  (\bibinfo{year}{1998}).

\bibitem{murray2011mathematical}
\bibinfo{author}{Murray, J.~D.}
\newblock \emph{\bibinfo{title}{{Mathematical Biology: I. An Introduction}}},
  vol. \bibinfo{volume}{110} of \emph{\bibinfo{series}{Interdisciplinary
  Applied Mathematics}} (\bibinfo{publisher}{Springer New York},
  \bibinfo{year}{2011}).

\bibitem{Strogatz:2008wo}
\bibinfo{author}{Strogatz, S.~H.}
\newblock \emph{\bibinfo{title}{{Nonlinear Dynamics and Chaos: With
  Applications to Physics, Biology, Chemistry, and Engineering}}}.
\newblock Studies in nonlinearity (\bibinfo{publisher}{Westview Press},
  \bibinfo{address}{Boulder}, \bibinfo{year}{2008}).

\bibitem{GuckHolmes}
\bibinfo{author}{Guckenheimer, J.} \& \bibinfo{author}{Holmes, P.}
\newblock \emph{\bibinfo{title}{{Nonlinear Oscillations, Dynamical Systems, and
  Bifurcations of Vector Fields}}} (\bibinfo{publisher}{Springer},
  \bibinfo{address}{New York}, \bibinfo{year}{1983}).

\bibitem{Gross:2004p2428}
\bibinfo{author}{Gross, T.} \& \bibinfo{author}{Feudel, U.}
\newblock \bibinfo{title}{{Analytical search for bifurcation surfaces in
  parameter space}}.
\newblock \emph{\bibinfo{journal}{Physica D}} \textbf{\bibinfo{volume}{195}},
  \bibinfo{pages}{292--302} (\bibinfo{year}{2004}).

\bibitem{Hastings:2001jh}
\bibinfo{author}{Hastings, A.}
\newblock \bibinfo{title}{{Transient dynamics and persistence of ecological
  systems}}.
\newblock \emph{\bibinfo{journal}{Ecol. Lett.}} \textbf{\bibinfo{volume}{4}},
  \bibinfo{pages}{215--220} (\bibinfo{year}{2001}).

\bibitem{Neubert:1997wk}
\bibinfo{author}{Neubert, M.} \& \bibinfo{author}{Caswell, H.}
\newblock \bibinfo{title}{{Alternatives to resilience for measuring the
  responses of ecological systems to perturbations}}.
\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{78}},
  \bibinfo{pages}{653--665} (\bibinfo{year}{1997}).

\bibitem{Caswell:2005eo}
\bibinfo{author}{Caswell, H.} \& \bibinfo{author}{Neubert, M.~G.}
\newblock \bibinfo{title}{{Reactivity and transient dynamics of discrete-time
  ecological systems}}.
\newblock \emph{\bibinfo{journal}{J Differ Equ Appl}}
  \textbf{\bibinfo{volume}{11}}, \bibinfo{pages}{295--310}
  (\bibinfo{year}{2005}).

\bibitem{Neubert:2009td}
\bibinfo{author}{Neubert, M.} \& \bibinfo{author}{Caswell, H.}
\newblock \bibinfo{title}{{Detecting reactivity}}.
\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{90}},
  \bibinfo{pages}{2683--2688} (\bibinfo{year}{2009}).

\bibitem{Yodzis:1992hg}
\bibinfo{author}{Yodzis, P.} \& \bibinfo{author}{Innes, S.}
\newblock \bibinfo{title}{{Body size and consumer-resource dynamics}}.
\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{139}},
  \bibinfo{pages}{1151--1175} (\bibinfo{year}{1992}).

\bibitem{Brown:2004wq}
\bibinfo{author}{Brown, J.}, \bibinfo{author}{Gillooly, J.},
  \bibinfo{author}{Allen, A.}, \bibinfo{author}{Savage, V.} \&
  \bibinfo{author}{West, G.}
\newblock \bibinfo{title}{{Toward a metabolic theory of ecology}}.
\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{85}},
  \bibinfo{pages}{1771--1789} (\bibinfo{year}{2004}).

\bibitem{Liow:2008jx}
\bibinfo{author}{Liow, L.~H.} \emph{et~al.}
\newblock \bibinfo{title}{{Higher origination and extinction rates in larger
  mammals}}.
\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
  \textbf{\bibinfo{volume}{105}}, \bibinfo{pages}{6097--6102}
  (\bibinfo{year}{2008}).

\bibitem{DeLong:2012kw}
\bibinfo{author}{DeLong, J.~P.} \& \bibinfo{author}{Vasseur, D.~A.}
\newblock \bibinfo{title}{{A dynamic explanation of size{\textendash}density
  scaling in carnivores}}.
\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{93}},
  \bibinfo{pages}{470--476} (\bibinfo{year}{2012}).

\bibitem{Kempes:2016}
\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Wang, L.},
  \bibinfo{author}{Amend, J.~P.}, \bibinfo{author}{Doyle, J.} \&
  \bibinfo{author}{Hoehler, T.}
\newblock \bibinfo{title}{{Evolutionary tradeoffs in cellular composition
  across diverse bacteria}}.
\newblock \emph{\bibinfo{journal}{ISME J}} \textbf{\bibinfo{volume}{10}},
  \bibinfo{pages}{2145--2157} (\bibinfo{year}{2016}).

\bibitem{Carbone:1999ju}
\bibinfo{author}{Carbone, C.}, \bibinfo{author}{Mace, G.~M.},
  \bibinfo{author}{Roberts, S.~C.} \& \bibinfo{author}{Macdonald, D.~W.}
\newblock \bibinfo{title}{{Energetic constraints on the diet of terrestrial
  carnivores}}.
\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{402}},
  \bibinfo{pages}{286--288} (\bibinfo{year}{1999}).

\bibitem{Carbone:2007dz}
\bibinfo{author}{Carbone, C.}, \bibinfo{author}{Teacher, A.} \&
  \bibinfo{author}{Rowcliffe, J.~M.}
\newblock \bibinfo{title}{{The Costs of Carnivory}}.
\newblock \emph{\bibinfo{journal}{PLoS Biol}} \textbf{\bibinfo{volume}{5}},
  \bibinfo{pages}{e22} (\bibinfo{year}{2007}).

\bibitem{Okie:2013ju}
\bibinfo{author}{Okie, J.~G.} \emph{et~al.}
\newblock \bibinfo{title}{{Effects of allometry, productivity and lifestyle on
  rates and limits of body size evolution}}.
\newblock \emph{\bibinfo{journal}{Proc Biol Sci}}
  \textbf{\bibinfo{volume}{280}}, \bibinfo{pages}{20131007--20131007}
  (\bibinfo{year}{2013}).

\bibitem{Brown:1993p708}
\bibinfo{author}{Brown, J.}, \bibinfo{author}{Marquet, P.} \&
  \bibinfo{author}{Taper, M.}
\newblock \bibinfo{title}{{Evolution of body size: consequences of an energetic
  definition of fitness}}.
\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{142}},
  \bibinfo{pages}{573--584} (\bibinfo{year}{1993}).

\bibitem{West:1997cg}
\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Brown, J.~H.} \&
  \bibinfo{author}{Enquist, B.~J.}
\newblock \bibinfo{title}{{A general model for the origin of allometric scaling
  laws in biology}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{276}},
  \bibinfo{pages}{122--126} (\bibinfo{year}{1997}).

\bibitem{West:2002ud}
\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Woodruff, W.~H.} \&
  \bibinfo{author}{Brown, J.~H.}
\newblock \bibinfo{title}{{Allometric scaling of metabolic rate from molecules
  and mitochondria to cells and mammals.}}
\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
  \textbf{\bibinfo{volume}{99 Suppl 1}}, \bibinfo{pages}{2473--2478}
  (\bibinfo{year}{2002}).

\bibitem{Millar:1990p923}
\bibinfo{author}{Millar, J.} \& \bibinfo{author}{Hickling, G.}
\newblock \bibinfo{title}{{Fasting endurance and the evolution of mammalian
  body size}}.
\newblock \emph{\bibinfo{journal}{Funct. Ecol.}} \textbf{\bibinfo{volume}{4}},
  \bibinfo{pages}{5--12} (\bibinfo{year}{1990}).

\bibitem{tilman1981}
\bibinfo{author}{Tilman, D.}
\newblock \bibinfo{title}{{Tests of resource competition theory using four
  species of lake michigan algae}}.
\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{62}},
  \bibinfo{pages}{802--815} (\bibinfo{year}{1981}).

\bibitem{dutkiewicz2009}
\bibinfo{author}{Dutkiewicz, S.}, \bibinfo{author}{Follows, M.~J.} \&
  \bibinfo{author}{Bragg, J.~G.}
\newblock \bibinfo{title}{{Modeling the coupling of ocean ecology and
  biogeochemistry}}.
\newblock \emph{\bibinfo{journal}{Global Biogeochem. Cycles}}
  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{1--15} (\bibinfo{year}{2009}).

\bibitem{barton2010}
\bibinfo{author}{Barton, A.~D.}, \bibinfo{author}{Dutkiewicz, S.},
  \bibinfo{author}{Flierl, G.}, \bibinfo{author}{Bragg, J.} \&
  \bibinfo{author}{Follows, M.~J.}
\newblock \bibinfo{title}{{Patterns of diversity in marine phytoplankton}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{327}},
  \bibinfo{pages}{1509--1511} (\bibinfo{year}{2010}).

\bibitem{Hempson:2015hka}
\bibinfo{author}{Hempson, G.~P.}, \bibinfo{author}{Archibald, S.} \&
  \bibinfo{author}{Bond, W.~J.}
\newblock \bibinfo{title}{{A continent-wide assessment of the form and
  intensity of large mammal herbivory in Africa}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{350}},
  \bibinfo{pages}{1056--1061} (\bibinfo{year}{2015}).

\bibitem{West:2002it}
\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Woodruff, W.~H.} \&
  \bibinfo{author}{Brown, J.~H.}
\newblock \bibinfo{title}{{Allometric scaling of metabolic rate from molecules
  and mitochondria to cells and mammals.}}
\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
  \textbf{\bibinfo{volume}{99 Suppl 1}}, \bibinfo{pages}{2473--2478}
  (\bibinfo{year}{2002}).

\bibitem{DeLong:2010dy}
\bibinfo{author}{DeLong, J.~P.}, \bibinfo{author}{Okie, J.~G.},
  \bibinfo{author}{Moses, M.~E.}, \bibinfo{author}{Sibly, R.~M.} \&
  \bibinfo{author}{Brown, J.~H.}
\newblock \bibinfo{title}{{Shifts in metabolic scaling, production, and
  efficiency across major evolutionary transitions of life}}.
\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
  \textbf{\bibinfo{volume}{107}}, \bibinfo{pages}{12941--12945}
  (\bibinfo{year}{2010}).

\bibitem{West:2001bv}
\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Brown, J.~H.} \&
  \bibinfo{author}{Enquist, B.~J.}
\newblock \bibinfo{title}{{A general model for ontogenetic growth}}.
\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{413}},
  \bibinfo{pages}{628--631} (\bibinfo{year}{2001}).

\bibitem{moses2008rmo}
\bibinfo{author}{Moses, M.~E.} \emph{et~al.}
\newblock \bibinfo{title}{{Revisiting a Model of Ontogenetic Growth: Estimating
  Model Parameters from Theory and Data.}}
\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{171}},
  \bibinfo{pages}{632--645} (\bibinfo{year}{2008}).

\bibitem{gillooly2002esa}
\bibinfo{author}{Gillooly, J.~F.}, \bibinfo{author}{Charnov, E.~L.},
  \bibinfo{author}{West, G.~B.}, \bibinfo{author}{Savage, V.~M.} \&
  \bibinfo{author}{Brown, J.~H.}
\newblock \bibinfo{title}{{Effects of size and temperature on developmental
  time}}.
\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{417}},
  \bibinfo{pages}{70--73} (\bibinfo{year}{2002}).

\bibitem{hou}
\bibinfo{author}{Hou, C.} \emph{et~al.}
\newblock \bibinfo{title}{{Energy uptake and allocation during ontogeny}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{322}},
  \bibinfo{pages}{736--739} (\bibinfo{year}{2008}).

\bibitem{Savage:2004ed}
\bibinfo{author}{Savage, V.~M.}, \bibinfo{author}{Gillooly, J.~F.},
  \bibinfo{author}{Brown, J.~H.}, \bibinfo{author}{West, G.~B.} \&
  \bibinfo{author}{Charnov, E.~L.}
\newblock \bibinfo{title}{{Effects of Body Size and Temperature on Population
  Growth}}.
\newblock
  \emph{\bibinfo{journal}{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}}
  \textbf{\bibinfo{volume}{163}}, \bibinfo{pages}{429--441}
  (\bibinfo{year}{2004}).

\bibitem{bettencourt}
\bibinfo{author}{Bettencourt, L. M.~A.}, \bibinfo{author}{Lobo, J.},
  \bibinfo{author}{Helbing, D.}, \bibinfo{author}{Kuhnert, C.} \&
  \bibinfo{author}{West, G.~B.}
\newblock \bibinfo{title}{{Growth, innovation, scaling, and the pace of life in
  cities}}.
\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
  \textbf{\bibinfo{volume}{104}}, \bibinfo{pages}{7301--7306}
  (\bibinfo{year}{2007}).

\bibitem{Folland:2008ij}
\bibinfo{author}{Folland, J.~P.}, \bibinfo{author}{Mc~Cauley, T.~M.} \&
  \bibinfo{author}{Williams, A.~G.}
\newblock \bibinfo{title}{{Allometric scaling of strength measurements to body
  size}}.
\newblock \emph{\bibinfo{journal}{Eur J Appl Physiol}}
  \textbf{\bibinfo{volume}{102}}, \bibinfo{pages}{739--745}
  (\bibinfo{year}{2008}).

\end{thebibliography}

\end{bibunit}

% \bibliography{aa_starving3}

%
% \begin{acknowledgments}
%   We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
%   discussions and comments on the manuscript.  J.D.Y. was supported by
%   startup funds at the University of California, Merced, and an Omidyar
%   Postdoctoral Fellowship at the Santa Fe Institute.  C.P.K. was supported by
%   an Omidyar Postdoctoral Fellowship at the Santa Fe Institute.  S.R. was
%   supported by grants DMR-1608211 and 1623243 from the National Science
%   Foundation, and by the John Templeton Foundation, all at the Santa Fe
%   Institute.
% \end{acknowledgments}
%
\clearpage

\begin{bibunit}[unsrt]

% \addto\captionsenglish{\renewcommand{\figurename}{Figure S\!}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}%
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}%

\section*{Supporting Information for ``The dynamics of starvation and recovery''}

{\bf Mechanisms of Starvation and Recovery}
To understand the dynamics of starvation, recovery, reproduction, and resource competition, our framework partitions consumers into two classes: (a) a full class that is able to reproduce and, (b) a hungry class that experiences mortality at a given rate and is unable to reproduce. For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate, as dictated by body size, with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs. \citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective). This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g. \citep{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \citep{kempes2014morphological}. Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).

We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
\begin{align}
\begin{split}
\dot{F_{d}} &= \lambda_{\text{max}} F_{d} + \rho_{\text{max}}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{\text{max}}R_{d} H_{d}/k - \mu H_{d},  \\
\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
& \left[\left(\frac{\rho_{\text{max}}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)F_{d}\right].
\label{bigdynamics}
\end{split}
\end{align}
where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with the subscript $_{d}$ before introducing the non-dimensional form that is presented in the main text). In the above equations $Y$ represents the yield coefficient (e.g., Refs. \citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism$^{-1}$). If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$. This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density. In these units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)

%\begin{align}
%\begin{split}
%\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
%\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
%\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
%&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
%\end{split}
%\end{align}
%
%In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability. Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources. The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
%A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
%
%Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification. As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$. Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to

%where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.

We can formally non-dimensionalize this system by the rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
\begin{align}
\begin{split}
&\dot{F} = \frac{1}{s}\left[\lambda_{\text{max}} F + \rho_{\text{max}}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
&\dot{H} = \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{\text{max}}\frac{R}{qk} H - \mu H\right],  \\
& \dot{R} = \\
&\frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{\text{max}}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)F\right]\right].
\end{split}
\end{align}
If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
\begin{align}
\begin{split}
\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
\label{reduceddynamics}
\end{split}
\end{align}
where we have dropped the subscripts on $\lambda_{\text{max}}$ and $\rho_{\text{max}}$ for simplicity, and $\xi\equiv C/k$, $\delta\equiv Y_{H}kP_{H}/C$, and $\beta\equiv Y_{H}k\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)/C$. The above equations represent the system of equations presented in the main text.
\\

{\bf Parameter Values and Estimates}
All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies. Below we outline previous measurements and simple estimates of the parameters.

Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs. \citep{West:2001bv,moses2008rmo} and the supplement for Ref. \citep{hou}). We make this assumption in the current paper, although alternate exponents, which are known to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends. The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.  For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs. \citep{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate. We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).

An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction. This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref. \citep{peters1986ecological} for a review). We use $m_{0}=0.097M^{0.92}$ \citep{blueweiss1978relationships} which encompasses the widest range of body sizes \citep{peters1986ecological}.

The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g. Refs. \citep{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \citep{moses2008rmo}. The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \citep{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \citep{stryer,hou}. For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \citep{hou}.

For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ ~\citep{Dunbrack:1993ec}, $0.02M^{1.19}$ ~\citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ ~\citep{Lindstedt:2002td}. We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range. Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ ~\citep{Lindstedt:2002td}.

%We also connect the resource growth rate to the total metabolic rate of an organism. That is, we are interested in the relative rates of resource recovery and consumption by the total population. From \citep{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$). Considering an energy density of $18200$ (J g$^{-1}$) of grass \citep{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$). This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.

Typically the value of $\xi=C/k$ should roughly be $2$. The value of $\rho$, $\lambda$, $\sigma$, and $\mu$ are all simple rates (note that we have not rescaled time in our non-dimensionalization) as defined in the maintext. Given that our model considers transitions over entire stages of ontogeny or nutritional states, the value of $Y$ must represent yields integrated over entire life stages. Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
\begin{equation}
B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
\end{equation}
Similarly, the maintenance resource consumption rate for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
\begin{equation}
B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
\end{equation}
Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.

Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities. From the data in Ref. \citep{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally. It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with $B_{tot}$ for each type. As such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{damuth1987interspecific}. We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above. However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.
\\
%For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$

%More complicated models of fecundity (which, for example, account for the average length of adulthood and the number of individuals produced over this span) could be employed. However, the scaling of population growth rate has been studied in detail before and follows a relationship of $$ which matches the theory well for $\phi=.95$ and $=$.

%In our calculations we include $20\%$ variation around this value which could account for differences in efficiency during

 \begin{table}[h]
\caption{Parameter values for mammals}
\label{param}
    \begin{center}
    \footnotesize
     \begin{tabular}{p{3.8cm} c p{2.2cm} p{1.4cm}}
     \hline

     Definition & Parameter & Value & References  \\
     \hline
   Asymptotic adult mass & $M$ & (g) &  \\
   Initial mass of an organism & $m_{0}$ & (g) &  \\
   Metabolic rate scaling exponent & $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
   Metabolic Normalization Constant & $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
   Initial mass scaling exponent & $\upsilon$ & $0.92$ &  \citep{blueweiss1978relationships,peters1986ecological} \\
   Initial mass scaling normalization constant & $n_{0}$ & $0.097$ (g$^{1-\upsilon}$) & \citep{blueweiss1978relationships,peters1986ecological}  \\   
   Fat mass scaling exponent & $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
   Fat scaling normalization constant & $f_{0}$ & $0.02$ (g$^{1-\eta}$) & \citep{Lindstedt:1985hm}\\
   Muscle mass scaling exponent & $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
   Muscle scaling normalization constantv& $u_{0}$ & $0.38$ (g$^{1-\zeta}$)  & \citep{Lindstedt:2002td} \\
   Energy to synthesis a unit of mass & $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
   Energy to synthesis a unit of mass during recovery & $E_{m}^{\prime}$ & $7000$ (J gram$^{-1}$) & \citep{stryer,hou} \\
   Specific resource growth rate & $\alpha$ & $9.45\times10^{-9}$ (s$^{-1}$) & see text  \\
   Fraction of asymptotic mass representing full state & $\epsilon_{\lambda}$ & $0.95$ & \citep{West:2001bv}  \\
   Fraction of asymptotic mass representing starving state & $\epsilon_{\sigma}$ & $1-f_{0}M^{\gamma-1}$ & see text  \\
   Fraction of asymptotic mass representing death & $\epsilon_{\mu}$ & $1-\frac{f_{0}M^{\gamma}+u_{0}M^{\zeta}}{M}$ & see text \\
   Carrying capacity (maximum density) of resources & $C$ & (g m$^{-2}$) & \\
   Half Saturation Constant & $k$ & (g m$^{-2}$) &   \\
   Normalized carrying capacity & $\xi$ & $C/k\approx2$ &   \\
   Reproductive fecundity & $\nu$ & $2$ & \citep{}  \\

%   $a$ & $1.78\times10^{-6}$  \quad \quad \\
%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\

   \hline
    \end{tabular}
    \end{center}
   \end{table}

{\bf Rate equations for invaders with modified body mass}
We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (see main text), implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.

Although the starved state of invading organisms remains unchanged, the rate of starvation from the modified full state to the starved state, the rate of recovery from the starved state to the modified full state, and the maintenance rates of both, will be different, such that $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$, $\delta^\prime = \delta(M^\prime)$.
Rates of starvation and recovery for the invading population are easily derived by adjusting the starting or ending state before and after starvation and recovery, leading to the following timescales:

\begin{align}
t_{\sigma^\prime} &= -\frac{M^{1-\eta}}{a^{\prime}}\ln \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
t_{\rho^\prime} &= \ln \left(\frac{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}{1-( \epsilon_\lambda(\chi +1))^{1/4}}\right)\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}.
\end{align}

The maintenance rates for the invading population require more careful consideration.
First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of body fat reserves.
Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
\begin{equation}
B^{\prime}_{\lambda}=\int_{0}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
\end{equation}
where
\begin{equation}
t_{\lambda^\prime} = \frac{M^{1-\eta} }{a(1-\eta)}\ln \left(\frac{1-(m_0/M)^{1-\eta}}{1-(\epsilon_\lambda (1+\chi))^{1-\eta}} \right).
\end{equation}
Note that we do not use this timescale to determine the reproductive rate of the invading consumer---which is assumed to remain the same as the resident population---but only to calulate the lifetime energy use.
Similarly, the maintenance for hungry individuals $P^\prime_{H}=B_{0}(\epsilon_{\sigma}(1+\chi)M)^{3/4}/(\epsilon_{\sigma}(1+\chi)M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y^\prime_{H}=(1+\chi)ME_{d}/B^{\prime}_{\rho}$ where
\begin{equation}
B^{\prime}_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
\end{equation}
Finally, we can calculate the maintenance of the invaders as

\begin{align}
  \delta^\prime &= P^\prime_{H}Y^\prime_{H}/\xi \\ \nonumber
  \beta^\prime &= \left(\frac{\lambda_{\rm max}}{Y^\prime_{F}}+P^\prime_{F} \right)Y^\prime_{H}/\xi.
\end{align}

To determine whether or not the invader or resident population has an advantage, we compute $R^*(M)$ and $R^*(M^\prime=M(1+\chi))$ for values of $\chi \in (-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$, and the invading population is assumed to have an advantage over the resident population if $R^*(M^\prime)<R^*(M)$.

{\bf Sensitivity to additional death terms}

It should be noted that our set of dynamics (Equations \ref{bigdynamics} and \ref{reduceddynamics}) could include a constant death term of the form $-d_{F}F$ and $-d_{H}H$ to represent death not directly linked to starvation. Adding terms of this form to our model would simply adjust the effective value of $\lambda$ and $\mu$, and we could rewrite Equation \ref{reduceddynamics} with $\lambda^{\prime}=\lambda-d$ and $\mu^{\prime}=\mu-d$. These substitutions would not alter the functional form of our model nor the steady-states and qualitative results, however the quantitative values could shift based on the size of $d$ relative to $\lambda$ and $\mu$.

Survivorship has a well-known functional form which changes systematically with size (e.g. \cite{calder1984}). Typically survivorship is defined using the Gompertz curve 
\begin{equation}
F=F_{0}e^{\left(c_{0}/c_{1}\right)\left(1-e^{c_{1}t}\right)}
\label{gompertz}
\end{equation}
where the parameters have the following allometric dependencies on adult mass $c_{0}=a_{0}M^{b_{0}}$ and $c_{1}=a_{1}M^{b_{1}}$, with $a_{0}=1.88\times10^{-8}$ (s g$^{-b_{0}}$), $b_{0}=-0.56$, $a_{1}=1.45\times10^{-7}$ (s g$^{-b_{1}}$), and $b_{1}=-0.27$ (see \cite{calder1984} for a review).

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{mortality-rate-comparison-eps-converted-to.pdf}
\caption{\small{The rates of reproduction $\lambda$ (blue), starvation-based mortality $\mu$ (red), and survivorship-based death $\bar{d}$ (black) as a function of adult mass.}\label{fig:ratescomp}}
\end{figure}

We are interested in the specific death rate of the form $\dot{F}=-dF$, and using the derivative of Equation \ref{gompertz} we find that $d=c_{0}e^{c_{1}t}$. Our model considers the average rates over a population and lifecycle and the average death rate is given by 
\begin{eqnarray}
\bar{d}&=&\frac{1}{t_{\text{exp}}}\int_{0}^{t_{\text{exp}}}c_{0}e^{c_{1}t} dt \\
&=&\frac{c_{0}\left(e^{c_{1} t_{\text{exp}}}-1\right)}{c_{1}t_{\text{exp}}}
\end{eqnarray}
where $t_{\text{exp}}$ is the expected lifespan following the allometry of $t_{\text{exp}}=a_{2}M^{b_{2}}$ with $a_{2}=4.04\times10^{6}$ (s g$^{-b_{2}}$) and $b_{2}=0.30$ ~\cite{damuth1982analysis,calder1984}. Given the allometries above we have that
\begin{equation}
\bar{d}=\frac{a_{0} \left(e^{a_{1}a_{2}M^{b_{1}+b_{2}}}-1\right) M^{b_{0}-b_{1}-b_{2}}}{a_{1} a_{2}}
\end{equation}
which scales roughly like $M^{b_{0}}$ because $b_{1}$ and $b_{2}$ are close in value but opposite in sign. In Figure S\ref{fig:ratescomp} we compare the value of $\bar{d}$ to the reproductive, $\lambda$, and starvation-based mortality, $\mu$, rates. The values of $\bar{d}$ are orders of magnitude smaller than these rates for all mammalian masses, and thus, adding this non-starvation based death rate to our model does not shift our results within numerical confidence.

\begin{figure}[h!]
\centering
\includegraphics[width=0.4\textwidth]{fig_FPenergyequiv-eps-converted-to.pdf}
\caption{\small{ Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass ($F^*$ is shown in green and $H^*$ in orange).  The data are from Damuth \citep{Damuth:1987kr} and have been converted to
  total population metabolism using the allometric relationships for
  metabolic rate (e.g. Refs.~\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:equivalence}}
\end{figure}

{\bf NSM and the energy equivalence hypothesis}

The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}. As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}. However, the actual equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations. Figure S\ref{fig:equivalence} recasts the results of the NSM in terms of this hypothesis and shows that $F^{*}B$ is nearly constant over the same range of mammalian sizes up to the asymptotic behavior for the largest terrestrial mammals.

{\bf Application of NSM limits to aquatic mammals}
A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass of a male African elephant. We note this particular limit as it may have future relevance to considerations of the ultimate constraints on aquatic mammals.

% \putbib[aa_starving_supplement]

\def\bibfont{\footnotesize}

% \bibliography{aa_starving_supplement}
\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\providecommand{\bibinfo}[2]{#2}
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem{Kempes:2012hy}
\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Dutkiewicz, S.} \&
  \bibinfo{author}{Follows, M.~J.}
\newblock \bibinfo{title}{{Growth, metabolic partitioning, and the size of
  microorganisms.}}
\newblock \emph{\bibinfo{journal}{PNAS}} \textbf{\bibinfo{volume}{109}},
  \bibinfo{pages}{495--500} (\bibinfo{year}{2012}).

\bibitem{kempes2014morphological}
\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Okegbe, C.},
  \bibinfo{author}{Mears-Clarke, Z.}, \bibinfo{author}{Follows, M.~J.} \&
  \bibinfo{author}{Dietrich, L.~E.}
\newblock \bibinfo{title}{Morphological optimization for access to dual
  oxidants in biofilms}.
\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
  Sciences}} \textbf{\bibinfo{volume}{111}}, \bibinfo{pages}{208--213}
  (\bibinfo{year}{2014}).

\bibitem{West:2001bv}
\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Brown, J.~H.} \&
  \bibinfo{author}{Enquist, B.~J.}
\newblock \bibinfo{title}{{A general model for ontogenetic growth}}.
\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{413}},
  \bibinfo{pages}{628--631} (\bibinfo{year}{2001}).

\bibitem{moses2008rmo}
\bibinfo{author}{Moses, M.~E.} \emph{et~al.}
\newblock \bibinfo{title}{{Revisiting a model of ontogenetic growth: Estimating
  model parameters from theory and data}}.
\newblock
  \emph{\bibinfo{journal}{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}}
  \textbf{\bibinfo{volume}{171}}, \bibinfo{pages}{632--645}
  (\bibinfo{year}{2008}).

\bibitem{hou}
\bibinfo{author}{Hou, C.} \emph{et~al.}
\newblock \bibinfo{title}{{Energy uptake and allocation during ontogeny}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{322}},
  \bibinfo{pages}{736--739} (\bibinfo{year}{2008}).

\bibitem{pirt}
\bibinfo{author}{Pirt, S.}
\newblock \bibinfo{title}{The maintenance energy of bacteria in growing
  cultures}.
\newblock \emph{\bibinfo{journal}{Proceedings of the Royal Society of London B:
  Biological Sciences}} \textbf{\bibinfo{volume}{163}},
  \bibinfo{pages}{224--231} (\bibinfo{year}{1965}).

\bibitem{Heijnen}
\bibinfo{author}{Heijnen, J.} \& \bibinfo{author}{Roels, J.}
\newblock \bibinfo{title}{A macroscopic model describing yield and maintenance
  relationships in aerobic fermentation processes}.
\newblock \emph{\bibinfo{journal}{Biotechnology and Bioengineering}}
  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{739--763}
  (\bibinfo{year}{1981}).

\bibitem{peters1986ecological}
\bibinfo{author}{Peters, R.~H.}
\newblock \emph{\bibinfo{title}{The Ecological Implications of Body Size}},
  vol.~\bibinfo{volume}{2} (\bibinfo{publisher}{Cambridge University Press},
  \bibinfo{address}{Cambridge}, \bibinfo{year}{1986}).

\bibitem{blueweiss1978relationships}
\bibinfo{author}{Blueweiss, L.} \emph{et~al.}
\newblock \bibinfo{title}{Relationships between body size and some life history
  parameters}.
\newblock \emph{\bibinfo{journal}{Oecologia}} \textbf{\bibinfo{volume}{37}},
  \bibinfo{pages}{257--272} (\bibinfo{year}{1978}).

\bibitem{stryer}
\bibinfo{author}{Stryer, L.}
\newblock \emph{\bibinfo{title}{{Biochemistry, Fourth Edition}}}
  (\bibinfo{publisher}{W.H. Freeman and Company}, \bibinfo{address}{New York},
  \bibinfo{year}{1995}).

\bibitem{Dunbrack:1993ec}
\bibinfo{author}{Dunbrack, R.~L.} \& \bibinfo{author}{Ramsay, M.~A.}
\newblock \bibinfo{title}{{The Allometry of Mammalian Adaptations to Seasonal
  Environments: A Critique of the Fasting Endurance Hypothesis}}.
\newblock \emph{\bibinfo{journal}{Oikos}} \textbf{\bibinfo{volume}{66}},
  \bibinfo{pages}{336--342} (\bibinfo{year}{1993}).

\bibitem{Lindstedt:1985hm}
\bibinfo{author}{Lindstedt, S.~L.} \& \bibinfo{author}{Boyce, M.~S.}
\newblock \bibinfo{title}{{Seasonality, Fasting Endurance, and Body Size in
  Mammals}}.
\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{125}},
  \bibinfo{pages}{873--878} (\bibinfo{year}{1985}).

\bibitem{Lindstedt:2002td}
\bibinfo{author}{Lindstedt, S.~L.} \& \bibinfo{author}{Schaeffer, P.~J.}
\newblock \bibinfo{title}{{Use of allometry in predicting anatomical and
  physiological parameters of mammals.}}
\newblock \emph{\bibinfo{journal}{Lab. Anim.}} \textbf{\bibinfo{volume}{36}},
  \bibinfo{pages}{1--19} (\bibinfo{year}{2002}).

\bibitem{estermann}
\bibinfo{author}{Estermann, B.~L.}, \bibinfo{author}{Wettstein, H.-R.},
  \bibinfo{author}{Sutter, F.} \& \bibinfo{author}{Kreuzer, M.}
\newblock \bibinfo{title}{Nutrient and energy conversion of grass-fed dairy and
  suckler beef cattle kept indoors and on high altitude pasture}.
\newblock \emph{\bibinfo{journal}{Animal Research}}
  \textbf{\bibinfo{volume}{50}}, \bibinfo{pages}{477--493}
  (\bibinfo{year}{2001}).

\bibitem{michaletz2014convergence}
\bibinfo{author}{Michaletz, S.~T.}, \bibinfo{author}{Cheng, D.},
  \bibinfo{author}{Kerkhoff, A.~J.} \& \bibinfo{author}{Enquist, B.~J.}
\newblock \bibinfo{title}{Convergence of terrestrial plant production across
  global climate gradients}.
\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{512}},
  \bibinfo{pages}{39--43} (\bibinfo{year}{2014}).

\bibitem{damuth1987interspecific}
\bibinfo{author}{Damuth, J.}
\newblock \bibinfo{title}{Interspecific allometry of population density in
  mammals and other animals: the independence of body mass and population
  energy-use}.
\newblock \emph{\bibinfo{journal}{Biological Journal of the Linnean Society}}
  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
  (\bibinfo{year}{1987}).

\bibitem{calder1984}
\bibinfo{author}{Calder, W.~A.}
\newblock \emph{\bibinfo{title}{Size, function, and life history}}
  (\bibinfo{publisher}{Harvard University Press}, \bibinfo{year}{1984}).

\bibitem{damuth1982analysis}
\bibinfo{author}{Damuth, J.}
\newblock \bibinfo{title}{Analysis of the preservation of community structure
  in assemblages of fossil mammals}.
\newblock \emph{\bibinfo{journal}{Paleobiology}} \textbf{\bibinfo{volume}{8}},
  \bibinfo{pages}{434--446} (\bibinfo{year}{1982}).

\bibitem{allen2002}
\bibinfo{author}{Allen, A.~P.}, \bibinfo{author}{Brown, J.~H.} \&
  \bibinfo{author}{Gillooly, J.~F.}
\newblock \bibinfo{title}{{Global biodiversity, biochemical kinetics, and the
  energetic-equivalence rule}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{297}},
  \bibinfo{pages}{1545--1548} (\bibinfo{year}{2002}).

\bibitem{enquist1998}
\bibinfo{author}{Enquist, B.~J.}, \bibinfo{author}{Brown, J.~H.} \&
  \bibinfo{author}{West, G.~B.}
\newblock \bibinfo{title}{{Allometric scaling of plant energetics and
  population density}}.
\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{395}},
  \bibinfo{pages}{163--165} (\bibinfo{year}{1998}).

\bibitem{Damuth:1987kr}
\bibinfo{author}{Damuth, J.}
\newblock \bibinfo{title}{{Interspecific allometry of population density in
  mammals and other animals: the independence of body mass and population
  energy-use}}.
\newblock \emph{\bibinfo{journal}{Biol. J. Linn. Soc.}}
  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
  (\bibinfo{year}{1987}).

\end{thebibliography}

\end{bibunit}

\end{document}

### 2017-12-13 NatureComm_REVISION/YKR_NatureComm-R6.tex
editorial revisions

--- 2017-12-13 NatureComm_REVISION/Arxiv/YKR_Arxiv3.tex
@@
+% \RequirePackage{lineno}
 \documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
@@
 %\usepackage{rotating}
-%\usepackage{lineno}
+\usepackage{lineno}
 %\usepackage[left]{lineno}
-%\usepackage[comma,sort&compress]{natbib}
+\usepackage[comma,sort&compress]{natbib}
+\bibpunct{}{}{,}{s}{}{;}
 %\usepackage{authblk}
@@
 
-%\bibliographystyle{ieeetr}
-
-\usepackage{bibunits}
-
-%\linenumbers
-%\setlength\linenumbersep{3pt}
+\bibliographystyle{naturemag}
+
+% \usepackage{bibunits}
+
+% \linenumbers
+% \setlength\linenumbersep{3pt}
 
@@
 
-\author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
-  of California, Merced, Merced, CA 95340, USA}
-
-\author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
-  Park Road, Santa Fe, NM 87501, USA}
-
-\author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
-  Road, Santa Fe, NM 87501, USA}
-
-\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
+\author{Justin D. Yeakel,${}^{1,2,*}$, Christopher P. Kempes,${}^{2,*}$,Sidney Redner,${}^{2,*}$\\
+${}^{1}$School of Natural Sciences, University of California, Merced, Merced, CA 95340, USA\\
+${}^{2}$The Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA\\
+${}^*$Contributed equally\\
+Correspondence and requests for materials should be addressed to J.D.Y. (email: jdyeakel@gmail.com), C.P.K. (email: ckempes@gmail.com), S.R. (email: sredner@santafe.edu)
+} 
+% \affiliation{School of Natural Sciences, University
+%   of California, Merced, Merced, CA 95340, USA}\affiliation{The Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
+
+% \author{} 
+% \affiliation{The Santa Fe Institute, 1399 Hyde
+%   Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
+
+% \author{} 
+% \affiliation{The Santa Fe Institute, 1399 Hyde Park
+%   Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
+\title{The dynamics of starvation and recovery: predicting extinction risk, Damuth's Law, and Cope's rule}%: Eco-evolutionary feedbacks}
 
 \begin{abstract} %250 words
-The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of its constituent individuals.
+The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of their constituent individuals.
 Of particular importance is the interplay between reproduction and the dynamics of starvation and recovery.
-To elucidate this interplay, we introduce a nutritional state-structured model that incorporates two classes of consumer: nutritionally replete, reproducing consumers, and undernourished, non-reproducing consumers.
+To elucidate this interplay, we here introduce a nutritional state-structured model that incorporates two classes of consumer: nutritionally replete, reproducing consumers, and undernourished, non-reproducing consumers.
 We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships and find that population dynamics are typically driven to a steady state.
@@
 
-\begin{bibunit}[unsrt]
-
+% \begin{bibunit}[unsrt]
+
+\noindent {\bf Introduction}\\
 The behavioral ecology of all organisms is influenced by their energetic
@@
 (iii) there is an allometrically constrained upper-bound for mammalian body size, and
-  (iv) the NSM provides a selective mechanism for the evolution of larger body size, known as Cope's rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}.
+(iv) the NSM provides a selective mechanism for the evolution of larger body size, known as Cope's rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}.\\
 
@@
 % As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.\\
-
-  \noindent \paragraph*{{\bf Nutritional state-structured model (NSM).}}
+\noindent {\bf Results}\\
+\noindent {\bf Nutritional state-structured model (NSM)}\\
 We begin by defining the nutritional state-structured population model, where the consumer population is partitioned into two states: (a) an
@@
 growth purposes.  We provide a physiologically and energetically mechanistic
-  model for each of these dynamics and constants (see the Supplementary Information
-  (SI)), and show that the system produces a simple non-dimensional form which
+model for each of these dynamics and constants (see Methods), and show that the system produces a simple non-dimensional form which
 we describe below.
-
-  \begin{figure}
-  \centering
-  \includegraphics[width=0.45\textwidth]{fig_FixedPoint2-eps-converted-to.pdf}
-  \caption{\small{ The transcritical (TC; dashed line) and Hopf bifurcation (solid line) as a
-    function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-    bifurcation conditions separate parameter space into unphysical (left of the TC), cyclic,
-    and steady state dynamic regimes.  The colors show the steady state densities for the energetically replete consumers $F^*$.  
-    % Steady state densities for the energetically deficient consumers $H^*$ (not shown)
-    % scale with those for $F^*$.  
-  }\label{fig:fp}}
-  \end{figure}
 
@@
 of resources $(1-R)$ (the maximum resource density has been non
-  dimensionalized to 1; see SI).  Conversely, consumers recover from state $H$
+dimensionalized to 1; see Methods).  Conversely, consumers recover from state $H$
 to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$
@@
 consumers eat resources at a constant rate $\beta$ that accounts for maximal
-  maintenance and somatic growth (see the SI for mechanistic derivations of
+maintenance and somatic growth (see Methods for mechanistic derivations of
 these rates from resource energetics).
@@
 This system of nondimensional equations follows from a set of first-principle
-  relationships for resource consumption and growth (see the SI for a full derivation and the dimensional form).
-  Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
+relationships for resource consumption and growth (see Methods for a full derivation and the dimensional form).
+Notice that the total consumer density $F+H$ follows $\dot{F}+\dot{H}=\lambda F-\mu H$.
 This resembles the equation of motion for the predator density in the LV model~\citep{murray2011mathematical}, except that the resource density does not appear in the growth term.
@@
 resource density because they represent a minimal energetic requirement for
-  consumers in the $H$ and $F$ state, respectively.
+consumers in the $H$ and $F$ state, respectively.\\
 
@@
 
-  \noindent \paragraph*{ {\bf Steady states of the NSM.}} From the single
+\noindent {\bf Steady states of the NSM}\\ 
+From the single
 internal fixed point (Eq.~\eqref{eq:ss}, see Methods), an obvious constraint
@@
 
-  \begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram-eps-converted-to.pdf}
-  \caption{\small{ The growth trajectory over absolute time of an individual organism as a function of body mass.  
-  Initial growth follows the black trajectory to an energetically replete reproductive adult mass of $m=\epsilon_\lambda M$ (see Methods). %which we assume is 95\% asymptotic mass $M$.  
-  Starvation follows the red trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. 
-  Recovery follows the green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires a longer time to reach $\epsilon_\lambda M$.
-%different energetics and
-  Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}\label{fig:growth}}
-  \end{figure}
-
 % \noindent {\bf Role of allometry} \\
@@
 %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-  \noindent {\bf Results}
-  \noindent \paragraph*{{\bf The allometry of extinction risk.}} While there
+
+\noindent {\bf The allometry of extinction risk}\\
+While there
 are no {\it a priori} constraints on the parameters in the NSM, we expect
@@
 an organism's body mass $M$, we parameterize these rates in terms of the
-  \emph{percent} gain and loss of the asymptotic (maximum) body mass,
+percent gain and loss of the asymptotic (maximum) body mass,
 $\epsilon M$, where different values of $\epsilon$ define different states of
@@
 starvation recovery dynamics for most organisms, here we focus on allometric
-  relationships for terrestrial-bound lower-trophic level endotherms (see the
-  SI for values), specifically herbivorous mammals, which range from a minimum
+relationships for terrestrial-bound lower-trophic level endotherms (table \ref{param}), specifically herbivorous mammals, which range from a minimum
 of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of
@@
 $H$ does not change the form of our model nor impact our quantitative findings
-  (see SI for the derivation).
+(see Supplementary Note 1: Sensitivity to additional death terms).
 
@@
 % However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.\\
-
-  \begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{fig_ExtinctionAllometricComb4-eps-converted-to.pdf}
-  \caption{\small{ Probability of extinction for a consumer with ({\bf a}) $M=10^2$g and ({\bf b}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
-Note the change in scale in panel {\bf b}.
-Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
-The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
-  \end{figure}
 
@@
 larger organisms are more sensitive to variability in physiological rates.
-  This finding is in accordance with, and may serve as contributing support for, observations of increased extinction risk among larger mammals \cite{Liow:2008jx}.\\
+This finding is in accordance with, and may serve as contributing support for, observations of increased extinction risk among larger mammals \citep{Liow:2008jx}.\\
 % Moreover, larger body sizes decrease the steady state resource density, such that fluctuations for larger organisms will be more likely to drive resources to extinction.
@@
 % \noindent {\bf Energetic barriers to body size}
-  \noindent \paragraph*{{\bf Damuth's Law and body size limits.}} The NSM
+\noindent {\bf Damuth's Law and body size limits}\\ 
+The NSM
 correctly predicts that smaller species have larger steady-state population
@@
 illustrates that the steady state values of the NSM combined with the derived
-  timescales naturally give rise to Damuth's law. While the previous metabolic
-  studies supporting Damuth's law provided arguments for the value of the
-  exponent~\citep{allen2002}, these studies are only able to infer the
-  normalization constant ($0.01$ g$^{1.78}$ m$^{-2}$ in the above equation) from the data (see SI for a discussion of the energy
-  equivalence hypothesis related to these metabolic arguments). Our model
-  predicts not only the exponent but also the normalization constant by
-  explicitly including the resource dynamics and the parameters that determine
-  growth and consumption. It should be noted that density relationships of
+timescales naturally give rise to Damuth's law. 
+While the initial metabolic studies supporting Damuth's law provide arguments for the value of the exponent \citep{Damuth:1987kr,allen2002} (see Supplementary Note 2: NSM and the energy equivalence hypothesis), our model predicts not only the exponent but also the normalization constant by explicitly including the resource dynamics and the parameters that determine growth and consumption.
+These predictions are complementary to recent work that also predicts the exponent and normalization constant of density relationships from the detailed allometries of reproduction, capture area, conversion efficiency, and mortality within predator-prey dynamic models \citep{DeLong:2012kw,DeLong:2012fj}.
+It should be noted that density relationships of
 individual clades follow a more shallow scaling relationship than predicted
-  by Damuth's law~\cite{Pedersen:2017he}.  In the context of our model,
+by Damuth's law~\citep{Pedersen:2017he}.  In the context of our model,
 this finding suggests that future work may be able to anticipate these shifts
@@
 
-  \begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{fig_FPAllometric2-eps-converted-to.pdf}
-  \caption{\small{Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-    body mass along with the data from Damuth \citep{Damuth:1987kr}. Inset: Resource steady state $R^*$ as a function of consumer body mass.}\label{fig:mass}}
-  \end{figure}
-
 % \paragraph*{{\bf Maximum Body Size}}
@@
 %, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
-  With respect to predicted steady state densities, the total metabolic rate of
-  $F$ and $H$ becomes infinite at a finite mass, and occurs at the same scale
-  where the steady state resources vanish (Fig.~\ref{fig:mass}). This
-  asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and
-  $\epsilon_{\lambda}$ (see Fig.~\ref{fig:growth}) equal zero, causing the
+With respect to predicted steady state densities, the population sizes of F and H become infinite at a finite body size, where the steady state resources also vanish (Fig.~\ref{fig:mass}). 
+This asymptotic behavior is governed by body sizes at which $\epsilon_{\rm \mu}$ and
+$\epsilon_{\rm \lambda}$ (see Fig.~\ref{fig:growth}) equal zero, causing the
 timescales (Eqn. \ref{t1}) to become infinite and the rates $\mu$ and $\lambda$ to equal
@@
 The $\mu=0$ asymptote occurs first when
-  $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to
+$f_{\rm 0}M^{\gamma-1}+u_{\rm 0}M^{\zeta-1}=1$, and corresponds to
 $(F^*,H^*,R^*)=(0,0,0)$.  This point predicts an upper bound on mammalian
-  body size at $M_{\rm max}=6.54\times10^7$ (g).  Moreover, $M_{\rm max}$,
+body size at $M_{\rm \rm max}=6.54\times10^7$ (g).  Moreover, $M_{\rm \rm max}$,
 which is entirely determined by the population-level consequences of
@@
 foraging time predicts a maximum carnivore size between $7\times10^{5}$ and
-  $1.1\times10^{6}$ (g) \cite{Carbone:1999ju,Carbone:2007dz}.  Similarly, the
+$1.1\times10^{6}$ (g) \citep{Carbone:1999ju,Carbone:2007dz}.  Similarly, the
 maximum body size within a particular lineage has been shown to scale with
 the metabolic normalization constant
-  \cite{Okie:2013ju}. %and depend on a critical death found to be constant from data
+\citep{Okie:2013ju}. %and depend on a critical death found to be constant from data
 This complementary approach is based on the balance between growth and
 mortality, and suggests that future connections between the scaling of fat
-  and muscle mass should systematically be connected with $B_{0}$ when
+and muscle mass should systematically be connected with $B_{\rm 0}$ when
 comparing
-  lineages. % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
-
-  % {\color{red}
-  % 
-  % It should be noted that in our model all death is related to nutritional state, however the addition of a constant death term \cite{} would only adjust the values of $\lambda$ and $\mu$ without changing the form or or qualitative results of our model.
-We explored the sensitivity of our results to shifts in a constant death term (see SI) and find that additional sources of death do not shift our quantitative prediction of Damuth's law except for large sizes (Figure S\ref{}).
-We find that additional sources of death shifts the asymptotic behavior of our model and decreases the maximum possible mammal size.
-Within the range of reasonable death rates $M_{max}$ varies by $VALUE\%$.
-These results suggest that the largest mammals are most sensitive to additional sources of death perhaps putting them at greater extinction risk.
-  % 
-  % }
-
-  %Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end.
-  %Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is close to the minimum observed mammalian body size of ca.\ 1.3-2.5g, a range that occurs as the recovery rate begins its decline.
-  %In addition to known transport limitations~\citep{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.
-  %This result mirrors other efforts \citep{Kempes:2012hy,Kempes:2016} where at a given scale multiple limitations constrain the smallest possibilities for life within a class of organisms.
-
-  %The combined steady-state dynamics and allometric timescales of the NSM predict that larger mammals should outcompete smaller ones, and this suggests that the model may provide a framework with which to explore the energetic mechanisms behind Cope's rule.
-%, which we explore further below.
-
-  \noindent \paragraph*{{\bf A mechanism for Cope's rule}} Metabolite transport
-  constraints are widely thought to place strict boundaries on biological
+lineages.
+
+\noindent{\bf A mechanism for Cope's rule}\\
+Metabolite transport constraints are widely thought to place strict boundaries on biological
 scaling~\citep{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to
@@
 associated with larger body masses, particularly for mammals.  One important
-  such example is the \emph{fasting endurance hypothesis}, which contends that
+such example is the `fasting endurance hypothesis', which contends that
 larger body size, with consequent lower metabolic rates and increased ability
@@
 ecological timescales, have not been explored.
-  % We argue that the NSM provides a suitable framework to explore these issues.
-
-  % \noindent{\bf A within-lineage driver of Cope's Rule}
-  % \paragraph*{{\bf Body Size Competition}} 
+
 The NSM predicts that the steady state resource density $R^{*}$ decreases
@@
 resource.
-  % We will now show that the NSM indeed provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time known as Cope's Rule.
-
-  \begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{fig_Invasion-eps-converted-to.pdf}
-  \caption{\small{Competitive outcomes for a resident species with body mass
-      $M$ vs. a closely related competing species with modified body mass
-      $M^\prime=M(1+\chi)$.  The blue region denotes proportions of modified
-      mass $\chi$ resulting in exclusion of the resident species.  The red
-      region denotes values of $\chi$ that result in a mass that is below the
-      starvation threshold and are thus infeasible.  Arrows point to the
-      predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$,
-      which may serve as an evolutionary attractor for body mass.  The black
-      wedge points to the largest body mass known for terrestrial mammals
-      (\emph{Deinotherium} spp.) at $1.74\times10^7$
-      (g)~\citep{Smith:2010p3442}.}\label{fig:invasion}}
-  \end{figure}
-
-  % Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
 
@@
 individuals have fewer metabolic reserves than the resident species and vice
-  versa for $\chi>0$.  For the allowable values of $\chi$ (see SI), the mass of
+versa for $\chi>0$.  For the allowable values of $\chi$ (see Methods), the mass of
 the competitor $M'$ should exceed the minimal amount of body fat,
-  $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be
+$1+\chi>\epsilon_{\rm \sigma}$, and the adjusted time to reproduce must be
 positive, which, given Eq.~\ref{t1}, implies that
-  $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.  These
-  conditions imply that $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$
+$1-\epsilon_{\rm \lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.  These
+conditions imply that $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\rm \lambda}-1)$
 where the upper bound approximately equals $0.05$ and the lower bound is
@@
 maintenance of both starving $\delta(M^\prime)$ and full consumers
-  $\beta(M^\prime)$ (see the SI for derivations of competitor rates).
+$\beta(M^\prime)$ (see Methods for derivations of competitor rates).
 Importantly, $\epsilon_\sigma$, which determines the point along the growth
 curve that defines the body composition of starved foragers, is assumed to
-  remain unchanged for the competing population (see SI).
-  %There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
+remain unchanged for the competing population.
 
@@
 
-  %, and this is due to the changing
-  %covariance between energetic rates as a function of modified energetic
-  %reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
-
 The observed switch in susceptibility as a function of $\chi$ at
-  $M_{\rm opt}= 1.748\times10^7$ (g) thus serves as an attractor, such that the
-  NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if
-  $M>M_{\rm opt}$.  This value is close to but smaller than the asymptotic
+$M_{\rm \rm opt}= 1.748\times10^7$ (g) thus serves as an attractor, such that the
+NSM predicts organismal mass to increase if $M<M_{\rm \rm opt}$ and decrease if
+$M>M_{\rm \rm opt}$.  This value is close to but smaller than the asymptotic
 upper bound for terrestrial mammal body size predicted by the NSM, and is remarkably close to independent estimates of the largest land
@@
 $1.74\times10^7$ (g) ~\citep{Smith:2010p3442}.  Additionally, our calculation
-  of $M_{\rm opt}$ as a function of mass-dependent physiological rates is
+of $M_{\rm \rm opt}$ as a function of mass-dependent physiological rates is
 similar to theoretical estimates of maximum body size \citep{Clauset:2009fh},
@@
 `maximum body size attractor' explored by Alroy~\citep{Alroy:1998p1594}.
-  % {\color{red} We note that the model of \cite{Brown:1993p708} predicts
-
-  An optimal size for mammals at intermediate body mass was predicted by Brown et al.\ based on reproductive maximization and the transition between hungry and full individuals \cite{Brown:1993p708}. 
-  By coupling the NSM to resource dynamics as well as introducing an explicit treatment of storage, we show that species with larger body masses have an inherent competitive advantage for size classes up to $M_{\rm opt}= 1.748\times10^7$ based on resource competition.
+
+An optimal size for mammals at intermediate body mass was predicted by Brown et al. \citep{Brown:1993p708} based on reproductive maximization and the transition between hungry and full individuals. 
+By coupling the NSM to resource dynamics as well as introducing an explicit treatment of storage, we show that species with larger body masses have an inherent competitive advantage for size classes up to $M_{\rm \rm opt}= 1.748\times10^7$ based on resource competition.
 Moreover, the mass distributions in Ref.
-\cite{Brown:1993p708} show that intermediate mammal sizes have the greatest species diversity, in contrast to our efforts, which consider total biomass and predict a much larger $M_{\rm opt}$.
-  Compellingly, recent work shows that many communities can be dominated by the biomass of the large \cite{Hempson:2015hka}.
-  % shows that the largest mammals have an advantage in terms resource competition. 
-  %as well as  \citep{Alroy:1998p1594,Clauset:2009fh}.
-  %This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
-  While the state of the environment as well as the competitive landscape will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general selective mechanism for the evolution of larger body size among terrestrial mammals.\\
+\citep{Brown:1993p708} show that intermediate mammal sizes have the greatest species diversity, in contrast to our efforts, which consider total biomass and predict a much larger $M_{\rm \rm opt}$.
+Compellingly, recent work shows that many communities can be dominated by the biomass of the large \citep{Hempson:2015hka}.
+While the state of the environment as well as the competitive landscape will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general selective mechanism for the evolution of larger body size among terrestrial mammals (cf.
+Supplementary Note 3: Application of NSM limits to aquatic mammals).\\
 
 \noindent {\bf Discussion} \\
-
-  %and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
-
-  %One might be concerned a greater number of large mammals are currently not observed in the modern world given that larger mammals are less susceptible to extinction.
-  %However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \citep{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also much more geographically widespread than today.
-  %These results, combined with our findings, suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \citep{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
 
@@
 constraints can be assessed using macroscale interactions between and among
-  species.
-
-  %Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.
+species.\\
 
@@
 \small{
-  {\bf Analytical solution to the NSM}
-  Equation~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
+\noindent {\bf Mechanisms of Starvation and Recovery}\\
+To understand the dynamics of starvation, recovery, reproduction, and resource competition, our framework partitions consumers into two classes: (a) a full class that is able to reproduce and, (b) a hungry class that experiences mortality at a given rate and is unable to reproduce.
+For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate, as dictated by body size, with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
+\citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
+This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis  \citep{West:2001bv,moses2008rmo,hou,Kempes:2012hy} and predictions for the steady-state resource abundance in communities of cells \citep{kempes2014morphological}.
+Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).
+
+We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
+\begin{align}
+\begin{split}
+\dot{F_{\rm d}} &= \lambda_{\rm \text{max}} F_{\rm d} + \rho_{\rm \text{max}}R_{\rm d}H_{\rm d}/k - \sigma \left(1-\frac{R_{\rm d}}{C}\right)F_{\rm d},  \\
+\dot{H_{\rm d}} &= \sigma \left(1-\frac{R_{\rm d}}{C}\right)F_{\rm d} - \rho_{\rm \text{max}}R_{\rm d} H_{\rm d}/k - \mu H_{\rm d},  \\
+\dot{R_{\rm d}} &= \alpha R_{\rm d}\left(1-\frac{R_{\rm d}}{C}\right) -\\
+& \left[\left(\frac{\rho_{\rm \text{max}}R_{\rm d}}{Y_{\rm H}k}+P_{\rm H}\right)H_{\rm d}+\left(\frac{\lambda_{\rm \text{max}}}{Y_{\rm F}}+P_{\rm F}\right)F_{\rm d}\right].
+\label{bigdynamics}
+\end{split}
+\end{align}
+where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with the subscript $_{\rm d}$ before introducing the non-dimensional form that is presented in the main text).
+In the above equations $Y$ represents the yield coefficient \citep{pirt,Heijnen} which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism$^{-1}$).
+If we pick $F_{\rm d}$ and $H_{\rm d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{\rm d}}$, such as $\frac{\rho\left(R_{\rm d}\right)}{Y}H_{\rm d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{\rm d}}$.
+This choice also gives $R_{\rm d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
+In these units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{\rm d}$, $C$ is the carrying capacity, or maximum density, of $R_{\rm d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
+
+We can formally non-dimensionalize this system by the rescaling of $F=fF_{\rm d}$, $H=fH_{\rm d}$, $R=qR_{\rm d}$, $t=st_{\rm d}$, in which case our system of equations becomes
+%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
+\begin{align}
+\begin{split}
+&\dot{F} = \frac{1}{s}\left[\lambda_{\rm \text{max}} F + \rho_{\rm \text{max}}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
+&\dot{H} = \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{\rm \text{max}}\frac{R}{qk} H - \mu H\right],  \\
+& \dot{R} = \\
+&\frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{\rm \text{max}}R}{Y_{\rm H}kq}+P_{\rm H}\right)H+\left(\frac{\lambda_{\rm \text{max}}}{Y_{\rm F}}+P_{\rm F}\right)F\right]\right].
+\end{split}
+\end{align}
+If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{\rm H}k$, then we are left with
+\begin{align}
+\begin{split}
+\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
+\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
+\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
+\label{reduceddynamics}
+\end{split}
+\end{align}
+where we have dropped the subscripts on $\lambda_{\rm \text{max}}$ and $\rho_{\rm \text{max}}$ for simplicity, and $\xi\equiv C/k$, $\delta\equiv Y_{\rm H}kP_{\rm H}/C$, and $\beta\equiv Y_{\rm H}k\left(\frac{\lambda_{\rm \text{max}}}{Y_{\rm F}}+P_{\rm F}\right)/C$.
+he above equations represent the system of equations presented in the main text.\\
+
+\noindent {\bf Analytical solution to the NSM}\\
+Eq.~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
 \begin{align}
@@
 fixed point is determined by the Jacobian matrix $\bf J$, with
-  $J_{ij}=\partial{\dot X_i}/\partial{X_j}$, when evaluated at the internal
+$J_{\rm ij}=\partial{\dot X_i}/\partial{X_j}$, when evaluated at the internal
 fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.  The parameters in
@@
 unique, it is the global attractor for all population trajectories for any
-  initial condition where the resource and consumer densities are both nonzero.
-
-  {\bf Metabolic scaling relationships} The scaling relation between an
+initial condition where the resource and consumer densities are both nonzero.\\
+
+\noindent {\bf Metabolic scaling relationships}\\ 
+The scaling relation between an
 organism's metabolic rate $B$ and its body mass $M$ at reproductive maturity
 is known to scale as $B = B_0 M^\eta$, where the scaling exponent $\eta$ is
-  typically close to $2/3$ or $3/4$ for metazoans (e.g.,
-  Ref.~\citep{West:2002it,Brown:2004wq}), and has taxonomic shifts for
+typically close to $2/3$ or $3/4$ for metazoans \citep{West:2002ud,Brown:2004wq}, and has taxonomic shifts for
 unicellular species between $\eta\approx 1$ in eukaryotes and
 $\eta\approx 1.76$ in bacteria \citep{DeLong:2010dy,Kempes:2012hy}.
-  %Justin adding an explanation for beta
 
@@
 \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Savage:2004ed,Kempes:2012hy}. This relation is derived from the simple balance condition 
-  $B_{0}m^{\eta}=E_{m}\dot{m}+B_{m}m\,,$
-  % \begin{eqnarray}
-  % \label{balance}
-  % B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
-  % \end{eqnarray}
-  \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Savage:2004ed,Kempes:2012hy} where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
+$B_{\rm 0}m^{\eta}=E_{\rm m}\dot{m}+B_{\rm m}m\,,$
+\citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Savage:2004ed,Kempes:2012hy} where $E_{\rm m}$ is the energy needed to synthesize a unit of mass, $B_{\rm m}$ is
 the metabolic rate to support an existing unit of mass, and $m$ is the mass
@@
 \label{m1}
-  \left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}},
+\left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{\rm 0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}},
 \end{eqnarray}
-  where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass,
-  $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically
-  (see the SI).  We now use this solution to define the timescale for
+where, for $\eta<1$, $M=(B_{\rm 0}/B_{\rm m})^{1/(1-\eta)}$ is the asymptotic mass,
+$a=B_{\rm 0}/E_{\rm m}$, and $m_0$ is mass at birth, itself varying allometrically.  We now use this solution to define the timescale for
 reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see
@@
 \label{t1}
-  \tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)},
+\tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{\rm 0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)},
 \end{equation}
@@
 reproduction is given by the timescale to go from the birth mass to the adult
-  mass. The time to reproduce is given by Equation \ref{t1} as
-  $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
+mass. The time to reproduce is given by Eq. \ref{t1} as
+$t_{\rm \lambda}=\tau\left(\epsilon_{\rm \lambda}\right)$, where $\epsilon_{\rm \lambda}$
 is the fraction of the asymptotic mass where an organism is reproductively
 mature and should be close to one (typically
-  $\epsilon_{\lambda}\approx0.95\;$ \citep{West:2001bv}). Our reproductive
+$\epsilon_{\rm \lambda}\approx0.95\;$ \citep{West:2001bv}). Our reproductive
 rate, $\lambda$, is a specific rate, or the number of offspring produced per
@@
 functional form gives the population growth
-  $F\left(t\right) = F_{0}e^{\lambda t}$ which can be related to the
-  reproductive timescale by assuming that when $t=t_{\lambda}$ it is also the
-  case that $F=\nu F_{0}$, where $\nu-1$ is the number of offspring produced
+$F\left(t\right) = F_{\rm 0}e^{\lambda t}$ which can be related to the
+reproductive timescale by assuming that when $t=t_{\rm \lambda}$ it is also the
+case that $F=\nu F_{\rm 0}$, where $\nu-1$ is the number of offspring produced
 per reproductive cycle. Following this relationship the growth rate is given
-  by $\lambda=\ln\left(\nu\right)/t_{\lambda}$, which is the standard
-  relationship (e.g.,~\cite{Savage:2004ed}) and will scales as
-  $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of
-  $\epsilon_{\lambda}$
+by $\lambda=\ln\left(\nu\right)/t_{\rm \lambda}$, which is the standard
+relationship \citep{Savage:2004ed} and will scales as
+$\lambda\propto M^{\eta-1}$ for $M\gg m_{\rm 0}$ for any constant value of
+$\epsilon_{\rm \lambda}$
 \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
@@
 and recovery by the timescales required to reach, or return from, specific
-  fractions of the replete-state mass (see the SI, Table I, for
-  parameterizations).  We define $m_{\sigma}=\epsilon_{\sigma} M$, where
-  $\epsilon_{\sigma}<1$ is the fraction of replete-state mass where
+fractions of the replete-state mass (table \ref{params}).  We define $m_{\rm \sigma}=\epsilon_{\rm \sigma} M$, where
+$\epsilon_{\rm \sigma}<1$ is the fraction of replete-state mass where
 reproduction ceases. This fraction will deviate from a constant if tissue
@@
 of the observation that body fat in mammals scales with overall body size
-  according to $M_{\rm fat}=f_{0}M^{\gamma}$ and assuming that once this mass
+according to $M_{\rm \rm fat}=f_{\rm 0}M^{\gamma}$ and assuming that once this mass
 is fully digested the organism starves, this would imply that
-  $\epsilon_{\sigma}=1-f_{0}M^{\gamma}/M$. It follows that the recovery
-  timescale, $t_{\rho}$, is the time to go from mass
-  $m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$
+$\epsilon_{\rm \sigma}=1-f_{\rm 0}M^{\gamma}/M$. It follows that the recovery
+timescale, $t_{\rm \rho}$, is the time to go from mass
+$m=\epsilon_{\rm \sigma} \epsilon_{\rm \lambda} M$ to $m=\epsilon_{\rm \lambda}M$
 (Fig. \ref{fig:growth}). Using Eqs.~\eqref{m1} and \eqref{t1} this timescale
 is given by simply considering the growth curve starting from a mass of
-  $m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
+$m_{\rm 0}^{\prime}=\epsilon_{\rm \sigma}\epsilon_{\rm \lambda}M$, in which case
 \begin{eqnarray}
 \label{rhotimescale}
-  t_{\rho}=\ln\left[\frac{1-\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)^{1-\eta}}{1-\epsilon_\lambda^{1-\eta}}\right]\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}
+t_{\rm \rho}=\ln\left[\frac{1-\left(\epsilon_{\rm \sigma}\epsilon_{\rm \lambda}\right)^{1-\eta}}{1-\epsilon_\lambda^{1-\eta}}\right]\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}
 \end{eqnarray}
-  where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in
-  the biosynthetic energetics during recovery (see the SI). It should be noted that more complicated ontogenetic models explicitly handle
+where $a^{\prime}=B_{\rm 0}/E_{\rm m}^{\prime}$ accounts for possible deviations in
+the biosynthetic energetics during recovery. It should be noted that more complicated ontogenetic models explicitly handle
 storage \citep{hou}, whereas this feature is implicitly covered by the body
@@
 source.  This assumption implies the metabolic balance
-  $\dot{m}E_{m}^{\prime}=-B_{m}m$ or $\dot{m}=-a^{\prime}m/M^{1-\eta}$
-  % \begin{eqnarray}
-  % \frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
-  % \end{eqnarray}
-  % or
-  % \begin{eqnarray}
-  % \frac{dm}{dt}=-\frac{a^{\prime}}{M^{1-\eta}}m
-  % \end{eqnarray}
-  where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
-  body mass, which differs from $E_{m}$, the energy required to
+$\dot{m}E_{\rm m}^{\prime}=-B_{\rm m}m$ or $\dot{m}=-a^{\prime}m/M^{1-\eta}$,
+where $E_{\rm m}^{\prime}$ is the amount of energy stored in a unit of existing
+body mass, which differs from $E_{\rm m}$, the energy required to
 synthesis a unit of biomass \citep{hou}. Given the replete mass, $M$, of an organism, the
@@
 organism: $m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}$.
-  % \begin{eqnarray}
-  % \label{mt}
-  % m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}.
-  % \end{eqnarray}
 The timescale for starvation is
-  given by the time it takes $m(t)$ to reach $\epsilon_{\sigma} M$, which gives
+given by the time it takes $m(t)$ to reach $\epsilon_{\rm \sigma} M$, which gives
 \begin{equation}
 \label{eq:sigma}
-  t_{\sigma}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\sigma}\right).
+t_{\rm \sigma}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\rm \sigma}\right).
 \end{equation}
-  The starvation rate is then $\sigma=1/t_{\sigma}$, which scales with
-  replete-state mass as $1/M^{1-\eta}\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important
+The starvation rate is then $\sigma=1/t_{\rm \sigma}$, which scales with
+replete-state mass as $1/M^{1-\eta}\ln\left(1-f_{\rm 0}M^{\gamma}/M\right)$.  An important
 feature is that $\sigma$ does not have a simple scaling dependence on
@@
 The time to death should follow a similar relation, but defined by a lower
-  fraction of replete-state mass, $m_{\mu}=\epsilon_{\mu} M$ where $\epsilon_\mu < \epsilon_\sigma$.
+fraction of replete-state mass, $m_{\rm \mu}=\epsilon_{\rm \mu} M$ where $\epsilon_\mu < \epsilon_\sigma$.
 Suppose, for example, that an organism dies once it has digested all fat and
 muscle tissues, and that muscle tissue scales with body mass according to
-  $M_{\rm musc}=u_{0}M^{\zeta}$.  This gives
-  $\epsilon_{\mu}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
+$M_{\rm \rm musc}=u_{\rm 0}M^{\zeta}$.  This gives
+$\epsilon_{\rm \mu}=1-\left(f_{\rm 0}M^{\gamma}+u_{\rm 0}M^{\zeta}\right)/M$. Muscle
 mass has been shown to be roughly proportional to body mass~\citep{Folland:2008ij} in
-  mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant.
-The time to go from starvation to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or $t_{\mu}=-M^{1-\eta}\ln\left(\epsilon_{\mu}\right)/a^{\prime}-t_{\sigma}$,
-  % \begin{eqnarray}
-  % \label{mutimescale}
-  % t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
-  % \end{eqnarray}
-  and $\mu=1/t_{\mu}$.
-  }
-
-\def\bibfont{\footnotesize}
-% \putbib[aa_starving3]
-
-\begin{thebibliography}{10}
-\expandafter\ifx\csname url\endcsname\relax
-  \def\url#1{\texttt{#1}}\fi
-\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
-\providecommand{\bibinfo}[2]{#2}
-\providecommand{\eprint}[2][]{\url{#2}}
-
-\bibitem{Martin:1987dl}
-\bibinfo{author}{Martin, T.~E.}
-\newblock \bibinfo{title}{{Food as a limit on breeding birds: A life-history
-  perspective}}.
-\newblock \emph{\bibinfo{journal}{Annu. Rev. Ecol. Syst.}}
-  \textbf{\bibinfo{volume}{18}}, \bibinfo{pages}{453--487}
-  (\bibinfo{year}{1987}).
-
-\bibitem{Kirk:1997cc}
-\bibinfo{author}{Kirk, K.~L.}
-\newblock \bibinfo{title}{{Life-history responses to variable environments:
-  Starvation and reproduction in planktonic rotifers}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{78}},
-  \bibinfo{pages}{434--441} (\bibinfo{year}{1997}).
-
-\bibitem{Kempes:2012hy}
-\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Dutkiewicz, S.} \&
-  \bibinfo{author}{Follows, M.~J.}
-\newblock \bibinfo{title}{{Growth, metabolic partitioning, and the size of
-  microorganisms}}.
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{109}}, \bibinfo{pages}{495--500}
-  (\bibinfo{year}{2012}).
-
-\bibitem{Mangel:1988uaa}
-\bibinfo{author}{Mangel, M.} \& \bibinfo{author}{Clark, C.~W.}
-\newblock \emph{\bibinfo{title}{{Dynamic Modeling in Behavioral Ecology}}}
-  (\bibinfo{publisher}{Princeton University Press},
-  \bibinfo{address}{Princeton}, \bibinfo{year}{1988}).
-
-\bibitem{Mangel:2014kz}
-\bibinfo{author}{Mangel, M.}
-\newblock \bibinfo{title}{{Stochastic dynamic programming illuminates the link
-  between environment, physiology, and evolution}}.
-\newblock \emph{\bibinfo{journal}{B. Math. Biol.}}
-  \textbf{\bibinfo{volume}{77}}, \bibinfo{pages}{857--877}
-  (\bibinfo{year}{2014}).
-
-\bibitem{Yeakel:2013hi}
-\bibinfo{author}{Yeakel, J.~D.}, \bibinfo{author}{Dominy, N.~J.},
-  \bibinfo{author}{Koch, P.~L.} \& \bibinfo{author}{Mangel, M.}
-\newblock \bibinfo{title}{{Functional morphology, stable isotopes, and human
-  evolution: a model of consilience}}.
-\newblock \emph{\bibinfo{journal}{Evolution}} \textbf{\bibinfo{volume}{68}},
-  \bibinfo{pages}{190--203} (\bibinfo{year}{2014}).
-
-\bibitem{Morris:1987eo}
-\bibinfo{author}{Morris, D.~W.}
-\newblock \bibinfo{title}{{Optimal allocation of parental investment}}.
-\newblock \emph{\bibinfo{journal}{Oikos}} \textbf{\bibinfo{volume}{49}},
-  \bibinfo{pages}{332--339} (\bibinfo{year}{1987}).
-
-\bibitem{Tveraa:2003fq}
-\bibinfo{author}{Tveraa, T.}, \bibinfo{author}{Fauchald, P.},
-  \bibinfo{author}{Henaug, C.} \& \bibinfo{author}{Yoccoz, N.~G.}
-\newblock \bibinfo{title}{{An examination of a compensatory relationship
-  between food limitation and predation in semi-domestic reindeer}}.
-\newblock \emph{\bibinfo{journal}{Oecologia}} \textbf{\bibinfo{volume}{137}},
-  \bibinfo{pages}{370--376} (\bibinfo{year}{2003}).
-
-\bibitem{Daan:1988va}
-\bibinfo{author}{Daan, S.}, \bibinfo{author}{Dijkstra, C.},
-  \bibinfo{author}{Drent, R.} \& \bibinfo{author}{Meijer, T.}
-\newblock \bibinfo{title}{{Food supply and the annual timing of avian
-  reproduction}}.
-\newblock In \bibinfo{editor}{Ouellet, H.} (ed.) \emph{\bibinfo{booktitle}{Acta
-  XIX Congressus Internationalis Ornithologici, Volume I: Proceedings XIX
-  International Ornithological Congress, 1986, Ottawa}},
-  \bibinfo{pages}{392--407} (\bibinfo{publisher}{Proceedings XIX International
-  Ornithological Congress}, \bibinfo{address}{Ottawa}, \bibinfo{year}{1989}).
-
-\bibitem{Jacot:2009dv}
-\bibinfo{author}{Jacot, A.}, \bibinfo{author}{Valcu, M.}, \bibinfo{author}{van
-  Oers, K.} \& \bibinfo{author}{Kempenaers, B.}
-\newblock \bibinfo{title}{{Experimental nest site limitation affects
-  reproductive strategies and parental investment in a hole-nesting
-  passerine}}.
-\newblock \emph{\bibinfo{journal}{Animal Behaviour}}
-  \textbf{\bibinfo{volume}{77}}, \bibinfo{pages}{1075--1083}
-  (\bibinfo{year}{2009}).
-
-\bibitem{Stearns:1989ip}
-\bibinfo{author}{Stearns, S.~C.}
-\newblock \bibinfo{title}{{Trade-offs in life-history evolution}}.
-\newblock \emph{\bibinfo{journal}{Funct. Ecol.}} \textbf{\bibinfo{volume}{3}},
-  \bibinfo{pages}{259} (\bibinfo{year}{1989}).
-
-\bibitem{Barboza:2002in}
-\bibinfo{author}{Barboza, P.} \& \bibinfo{author}{Jorde, D.}
-\newblock \bibinfo{title}{{Intermittent fasting during winter and spring
-  affects body composition and reproduction of a migratory duck}}.
-\newblock \emph{\bibinfo{journal}{J Comp Physiol B}}
-  \textbf{\bibinfo{volume}{172}}, \bibinfo{pages}{419--434}
-  (\bibinfo{year}{2002}).
-
-\bibitem{Threlkeld:1976ih}
-\bibinfo{author}{Threlkeld, S.~T.}
-\newblock \bibinfo{title}{{Starvation and the size structure of zooplankton
-  communities}}.
-\newblock \emph{\bibinfo{journal}{Freshwater Biol.}}
-  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{489--496}
-  (\bibinfo{year}{1976}).
-
-\bibitem{Weber:1998jg}
-\bibinfo{author}{Weber, T.~P.}, \bibinfo{author}{Ens, B.~J.} \&
-  \bibinfo{author}{Houston, A.~I.}
-\newblock \bibinfo{title}{{Optimal avian migration: A dynamic model of fuel
-  stores and site use}}.
-\newblock \emph{\bibinfo{journal}{Evolutionary Ecology}}
-  \textbf{\bibinfo{volume}{12}}, \bibinfo{pages}{377--401}
-  (\bibinfo{year}{1998}).
-
-\bibitem{Mduma:1999cp}
-\bibinfo{author}{Mduma, S. A.~R.}, \bibinfo{author}{Sinclair, A. R.~E.} \&
-  \bibinfo{author}{Hilborn, R.}
-\newblock \bibinfo{title}{{Food regulates the Serengeti wildebeest: a 40-year
-  record}}.
-\newblock \emph{\bibinfo{journal}{J. Anim. Ecol.}}
-  \textbf{\bibinfo{volume}{68}}, \bibinfo{pages}{1101--1122}
-  (\bibinfo{year}{1999}).
-
-\bibitem{Moore:2014hi}
-\bibinfo{author}{Moore, J.~W.}, \bibinfo{author}{Yeakel, J.~D.},
-  \bibinfo{author}{Peard, D.}, \bibinfo{author}{Lough, J.} \&
-  \bibinfo{author}{Beere, M.}
-\newblock \bibinfo{title}{{Life-history diversity and its importance to
-  population stability and persistence of a migratory fish: steelhead in two
-  large North American watersheds}}.
-\newblock \emph{\bibinfo{journal}{J. Anim. Ecol.}}
-  \textbf{\bibinfo{volume}{83}}, \bibinfo{pages}{1035--1046}
-  (\bibinfo{year}{2014}).
-
-\bibitem{Mead:1989dt}
-\bibinfo{author}{Mead, R.~A.}
-\newblock \bibinfo{title}{{The Physiology and Evolution of Delayed Implantation
-  in Carnivores}}.
-\newblock In \bibinfo{editor}{Gittleman, J.~L.} (ed.)
-  \emph{\bibinfo{booktitle}{Carnivore Behavior, Ecology, and Evolution}},
-  \bibinfo{pages}{437--464} (\bibinfo{publisher}{Springer US},
-  \bibinfo{address}{Ithaca}, \bibinfo{year}{1989}).
-
-\bibitem{Sandell:1990kw}
-\bibinfo{author}{Sandell, M.}
-\newblock \bibinfo{title}{{The evolution of seasonal delayed implantation}}.
-\newblock \emph{\bibinfo{journal}{Q Rev Biol}} \textbf{\bibinfo{volume}{65}},
-  \bibinfo{pages}{23--42} (\bibinfo{year}{1990}).
-
-\bibitem{Bulik:1999eo}
-\bibinfo{author}{Bulik, C.~M.} \emph{et~al.}
-\newblock \bibinfo{title}{{Fertility and reproduction in women with anorexia
-  nervosa}}.
-\newblock \emph{\bibinfo{journal}{J. Clin. Psychiat.}}
-  \textbf{\bibinfo{volume}{60}}, \bibinfo{pages}{130--135}
-  (\bibinfo{year}{1999}).
-
-\bibitem{Trites:2003cc}
-\bibinfo{author}{Trites, A.~W.} \& \bibinfo{author}{Donnelly, C.~P.}
-\newblock \bibinfo{title}{{The decline of Steller sea lions Eumetopias jubatus
-  in Alaska: a review of the nutritional stress hypothesis}}.
-\newblock \emph{\bibinfo{journal}{Mammal Rev.}} \textbf{\bibinfo{volume}{33}},
-  \bibinfo{pages}{3--28} (\bibinfo{year}{2003}).
-
-\bibitem{Glazier:2009hq}
-\bibinfo{author}{Glazier, D.~S.}
-\newblock \bibinfo{title}{{Metabolic level and size scaling of rates of
-  respiration and growth in unicellular organisms}}.
-\newblock \emph{\bibinfo{journal}{Funct. Ecol.}} \textbf{\bibinfo{volume}{23}},
-  \bibinfo{pages}{963--968} (\bibinfo{year}{2009}).
-
-\bibitem{Kooi2000}
-\bibinfo{author}{Kooijman, S. A. L.~M.}
-\newblock \emph{\bibinfo{title}{{Dynamic Energy and Mass Budgets in Biological
-  Systems}}} (\bibinfo{address}{Cambridge}, \bibinfo{year}{2000}).
-
-\bibitem{Sousa:2010ez}
-\bibinfo{author}{Sousa, T.}, \bibinfo{author}{Domingos, T.},
-  \bibinfo{author}{Poggiale, J.~C.} \& \bibinfo{author}{Kooijman, S. A. L.~M.}
-\newblock \bibinfo{title}{{Dynamic energy budget theory restores coherence in
-  biology}}.
-\newblock \emph{\bibinfo{journal}{Philos. T. Roy. Soc. B}}
-  \textbf{\bibinfo{volume}{365}}, \bibinfo{pages}{3413--3428}
-  (\bibinfo{year}{2010}).
-
-\bibitem{Diekmann:2010da}
-\bibinfo{author}{Diekmann, O.} \& \bibinfo{author}{Metz, J. A.~J.}
-\newblock \bibinfo{title}{{How to lift a model for individual behaviour to the
-  population level?}}
-\newblock \emph{\bibinfo{journal}{Philos. T. Roy. Soc. B}}
-  \textbf{\bibinfo{volume}{365}}, \bibinfo{pages}{3523--3530}
-  (\bibinfo{year}{2010}).
-
-\bibitem{murdoch:2003}
-\bibinfo{author}{Murdoch, W.~W.}, \bibinfo{author}{Briggs, C.~J.} \&
-  \bibinfo{author}{Nisbet, R.~M.}
-\newblock \emph{\bibinfo{title}{{Consumer-resource Dynamics}}},
-  vol.~\bibinfo{volume}{36} of \emph{\bibinfo{series}{Monographs in population
-  biology}} (\bibinfo{publisher}{Princeton University Press},
-  \bibinfo{address}{Princeton}, \bibinfo{year}{2003}).
-
-\bibitem{Damuth:1987kr}
-\bibinfo{author}{Damuth, J.}
-\newblock \bibinfo{title}{{Interspecific allometry of population density in
-  mammals and other animals: the independence of body mass and population
-  energy-use}}.
-\newblock \emph{\bibinfo{journal}{Biol. J. Linn. Soc.}}
-  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
-  (\bibinfo{year}{1987}).
-
-\bibitem{allen2002}
-\bibinfo{author}{Allen, A.~P.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{Gillooly, J.~F.}
-\newblock \bibinfo{title}{{Global biodiversity, biochemical kinetics, and the
-  energetic-equivalence rule}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{297}},
-  \bibinfo{pages}{1545--1548} (\bibinfo{year}{2002}).
-
-\bibitem{enquist1998}
-\bibinfo{author}{Enquist, B.~J.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{West, G.~B.}
-\newblock \bibinfo{title}{{Allometric scaling of plant energetics and
-  population density}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{395}},
-  \bibinfo{pages}{163--165} (\bibinfo{year}{1998}).
-
-\bibitem{Pedersen:2017he}
-\bibinfo{author}{Pedersen, R.~{\O}.}, \bibinfo{author}{Faurby, S.} \&
-  \bibinfo{author}{Svenning, J.-C.}
-\newblock \bibinfo{title}{{Shallow size{\textendash}density relations within
-  mammal clades suggest greater intra-guild ecological impact of large-bodied
-  species}}.
-\newblock \emph{\bibinfo{journal}{J. Anim. Ecol.}}
-  \textbf{\bibinfo{volume}{86}}, \bibinfo{pages}{1205--1213}
-  (\bibinfo{year}{2017}).
-
-\bibitem{Alroy:1998p1594}
-\bibinfo{author}{Alroy, J.}
-\newblock \bibinfo{title}{{Cope's rule and the dynamics of body mass evolution
-  in North American fossil mammals}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{280}},
-  \bibinfo{pages}{731--734} (\bibinfo{year}{1998}).
-
-\bibitem{Clauset:2009fh}
-\bibinfo{author}{Clauset, A.} \& \bibinfo{author}{Redner, S.}
-\newblock \bibinfo{title}{{Evolutionary model of species body mass
-  diversification}}.
-\newblock \emph{\bibinfo{journal}{Phys. Rev. Lett.}}
-  \textbf{\bibinfo{volume}{102}}, \bibinfo{pages}{038103}
-  (\bibinfo{year}{2009}).
-
-\bibitem{Smith:2010p3442}
-\bibinfo{author}{Smith, F.~A.} \emph{et~al.}
-\newblock \bibinfo{title}{{The evolution of maximum body size of terrestrial
-  mammals}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{330}},
-  \bibinfo{pages}{1216--1219} (\bibinfo{year}{2010}).
-
-\bibitem{Saarinen:2014br}
-\bibinfo{author}{Saarinen, J.~J.} \emph{et~al.}
-\newblock \bibinfo{title}{{Patterns of maximum body size evolution in Cenozoic
-  land mammals: Eco-evolutionary processes and abiotic forcing}}.
-\newblock \emph{\bibinfo{journal}{Proc Biol Sci}}
-  \textbf{\bibinfo{volume}{281}}, \bibinfo{pages}{20132049}
-  (\bibinfo{year}{2014}).
-
-\bibitem{Benichou:2014wu}
-\bibinfo{author}{B{\'e}nichou, O.} \& \bibinfo{author}{Redner, S.}
-\newblock \bibinfo{title}{{Depletion-controlled starvation of a diffusing
-  forager.}}
-\newblock \emph{\bibinfo{journal}{Phys. Rev. Lett.}}
-  \textbf{\bibinfo{volume}{113}}, \bibinfo{pages}{238101}
-  (\bibinfo{year}{2014}).
-
-\bibitem{Benichou:2016wl}
-\bibinfo{author}{B{\'e}nichou, O.}, \bibinfo{author}{Chupeau, M.} \&
-  \bibinfo{author}{Redner, S.}
-\newblock \bibinfo{title}{{Role of depletion on the dynamics of a diffusing
-  forager}}.
-\newblock \emph{\bibinfo{journal}{J Phys A-Math Theor}}
-  \textbf{\bibinfo{volume}{49}}, \bibinfo{pages}{394003}
-  (\bibinfo{year}{2016}).
-
-\bibitem{Chupeau:2016jf}
-\bibinfo{author}{Chupeau, M.}, \bibinfo{author}{B{\'e}nichou, O.} \&
-  \bibinfo{author}{Redner, S.}
-\newblock \bibinfo{title}{{Universality classes of foraging with resource
-  renewal}}.
-\newblock \emph{\bibinfo{journal}{Phys. Rev. E}} \textbf{\bibinfo{volume}{93}},
-  \bibinfo{pages}{032403} (\bibinfo{year}{2016}).
-
-\bibitem{Persson:1998hz}
-\bibinfo{author}{Persson, L.}, \bibinfo{author}{Leonardsson, K.},
-  \bibinfo{author}{De~Roos, A.~M.}, \bibinfo{author}{Gyllenberg, M.} \&
-  \bibinfo{author}{Christensen, B.}
-\newblock \bibinfo{title}{{Ontogenetic scaling of foraging rates and the
-  dynamics of a size-structured consumer-resource model}}.
-\newblock \emph{\bibinfo{journal}{Theor Popul Biol}}
-  \textbf{\bibinfo{volume}{54}}, \bibinfo{pages}{270--293}
-  (\bibinfo{year}{1998}).
-
-\bibitem{murray2011mathematical}
-\bibinfo{author}{Murray, J.~D.}
-\newblock \emph{\bibinfo{title}{{Mathematical Biology: I. An Introduction}}},
-  vol. \bibinfo{volume}{110} of \emph{\bibinfo{series}{Interdisciplinary
-  Applied Mathematics}} (\bibinfo{publisher}{Springer New York},
-  \bibinfo{year}{2011}).
-
-\bibitem{Strogatz:2008wo}
-\bibinfo{author}{Strogatz, S.~H.}
-\newblock \emph{\bibinfo{title}{{Nonlinear Dynamics and Chaos: With
-  Applications to Physics, Biology, Chemistry, and Engineering}}}.
-\newblock Studies in nonlinearity (\bibinfo{publisher}{Westview Press},
-  \bibinfo{address}{Boulder}, \bibinfo{year}{2008}).
-
-\bibitem{GuckHolmes}
-\bibinfo{author}{Guckenheimer, J.} \& \bibinfo{author}{Holmes, P.}
-\newblock \emph{\bibinfo{title}{{Nonlinear Oscillations, Dynamical Systems, and
-  Bifurcations of Vector Fields}}} (\bibinfo{publisher}{Springer},
-  \bibinfo{address}{New York}, \bibinfo{year}{1983}).
-
-\bibitem{Gross:2004p2428}
-\bibinfo{author}{Gross, T.} \& \bibinfo{author}{Feudel, U.}
-\newblock \bibinfo{title}{{Analytical search for bifurcation surfaces in
-  parameter space}}.
-\newblock \emph{\bibinfo{journal}{Physica D}} \textbf{\bibinfo{volume}{195}},
-  \bibinfo{pages}{292--302} (\bibinfo{year}{2004}).
-
-\bibitem{Hastings:2001jh}
-\bibinfo{author}{Hastings, A.}
-\newblock \bibinfo{title}{{Transient dynamics and persistence of ecological
-  systems}}.
-\newblock \emph{\bibinfo{journal}{Ecol. Lett.}} \textbf{\bibinfo{volume}{4}},
-  \bibinfo{pages}{215--220} (\bibinfo{year}{2001}).
-
-\bibitem{Neubert:1997wk}
-\bibinfo{author}{Neubert, M.} \& \bibinfo{author}{Caswell, H.}
-\newblock \bibinfo{title}{{Alternatives to resilience for measuring the
-  responses of ecological systems to perturbations}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{78}},
-  \bibinfo{pages}{653--665} (\bibinfo{year}{1997}).
-
-\bibitem{Caswell:2005eo}
-\bibinfo{author}{Caswell, H.} \& \bibinfo{author}{Neubert, M.~G.}
-\newblock \bibinfo{title}{{Reactivity and transient dynamics of discrete-time
-  ecological systems}}.
-\newblock \emph{\bibinfo{journal}{J Differ Equ Appl}}
-  \textbf{\bibinfo{volume}{11}}, \bibinfo{pages}{295--310}
-  (\bibinfo{year}{2005}).
-
-\bibitem{Neubert:2009td}
-\bibinfo{author}{Neubert, M.} \& \bibinfo{author}{Caswell, H.}
-\newblock \bibinfo{title}{{Detecting reactivity}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{90}},
-  \bibinfo{pages}{2683--2688} (\bibinfo{year}{2009}).
-
-\bibitem{Yodzis:1992hg}
-\bibinfo{author}{Yodzis, P.} \& \bibinfo{author}{Innes, S.}
-\newblock \bibinfo{title}{{Body size and consumer-resource dynamics}}.
-\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{139}},
-  \bibinfo{pages}{1151--1175} (\bibinfo{year}{1992}).
-
-\bibitem{Brown:2004wq}
-\bibinfo{author}{Brown, J.}, \bibinfo{author}{Gillooly, J.},
-  \bibinfo{author}{Allen, A.}, \bibinfo{author}{Savage, V.} \&
-  \bibinfo{author}{West, G.}
-\newblock \bibinfo{title}{{Toward a metabolic theory of ecology}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{85}},
-  \bibinfo{pages}{1771--1789} (\bibinfo{year}{2004}).
-
-\bibitem{Liow:2008jx}
-\bibinfo{author}{Liow, L.~H.} \emph{et~al.}
-\newblock \bibinfo{title}{{Higher origination and extinction rates in larger
-  mammals}}.
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{105}}, \bibinfo{pages}{6097--6102}
-  (\bibinfo{year}{2008}).
-
-\bibitem{DeLong:2012kw}
-\bibinfo{author}{DeLong, J.~P.} \& \bibinfo{author}{Vasseur, D.~A.}
-\newblock \bibinfo{title}{{A dynamic explanation of size{\textendash}density
-  scaling in carnivores}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{93}},
-  \bibinfo{pages}{470--476} (\bibinfo{year}{2012}).
-
-\bibitem{Kempes:2016}
-\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Wang, L.},
-  \bibinfo{author}{Amend, J.~P.}, \bibinfo{author}{Doyle, J.} \&
-  \bibinfo{author}{Hoehler, T.}
-\newblock \bibinfo{title}{{Evolutionary tradeoffs in cellular composition
-  across diverse bacteria}}.
-\newblock \emph{\bibinfo{journal}{ISME J}} \textbf{\bibinfo{volume}{10}},
-  \bibinfo{pages}{2145--2157} (\bibinfo{year}{2016}).
-
-\bibitem{Carbone:1999ju}
-\bibinfo{author}{Carbone, C.}, \bibinfo{author}{Mace, G.~M.},
-  \bibinfo{author}{Roberts, S.~C.} \& \bibinfo{author}{Macdonald, D.~W.}
-\newblock \bibinfo{title}{{Energetic constraints on the diet of terrestrial
-  carnivores}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{402}},
-  \bibinfo{pages}{286--288} (\bibinfo{year}{1999}).
-
-\bibitem{Carbone:2007dz}
-\bibinfo{author}{Carbone, C.}, \bibinfo{author}{Teacher, A.} \&
-  \bibinfo{author}{Rowcliffe, J.~M.}
-\newblock \bibinfo{title}{{The Costs of Carnivory}}.
-\newblock \emph{\bibinfo{journal}{PLoS Biol}} \textbf{\bibinfo{volume}{5}},
-  \bibinfo{pages}{e22} (\bibinfo{year}{2007}).
-
-\bibitem{Okie:2013ju}
-\bibinfo{author}{Okie, J.~G.} \emph{et~al.}
-\newblock \bibinfo{title}{{Effects of allometry, productivity and lifestyle on
-  rates and limits of body size evolution}}.
-\newblock \emph{\bibinfo{journal}{Proc Biol Sci}}
-  \textbf{\bibinfo{volume}{280}}, \bibinfo{pages}{20131007--20131007}
-  (\bibinfo{year}{2013}).
-
-\bibitem{Brown:1993p708}
-\bibinfo{author}{Brown, J.}, \bibinfo{author}{Marquet, P.} \&
-  \bibinfo{author}{Taper, M.}
-\newblock \bibinfo{title}{{Evolution of body size: consequences of an energetic
-  definition of fitness}}.
-\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{142}},
-  \bibinfo{pages}{573--584} (\bibinfo{year}{1993}).
-
-\bibitem{West:1997cg}
-\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{Enquist, B.~J.}
-\newblock \bibinfo{title}{{A general model for the origin of allometric scaling
-  laws in biology}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{276}},
-  \bibinfo{pages}{122--126} (\bibinfo{year}{1997}).
-
-\bibitem{West:2002ud}
-\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Woodruff, W.~H.} \&
-  \bibinfo{author}{Brown, J.~H.}
-\newblock \bibinfo{title}{{Allometric scaling of metabolic rate from molecules
-  and mitochondria to cells and mammals.}}
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{99 Suppl 1}}, \bibinfo{pages}{2473--2478}
-  (\bibinfo{year}{2002}).
-
-\bibitem{Millar:1990p923}
-\bibinfo{author}{Millar, J.} \& \bibinfo{author}{Hickling, G.}
-\newblock \bibinfo{title}{{Fasting endurance and the evolution of mammalian
-  body size}}.
-\newblock \emph{\bibinfo{journal}{Funct. Ecol.}} \textbf{\bibinfo{volume}{4}},
-  \bibinfo{pages}{5--12} (\bibinfo{year}{1990}).
-
-\bibitem{tilman1981}
-\bibinfo{author}{Tilman, D.}
-\newblock \bibinfo{title}{{Tests of resource competition theory using four
-  species of lake michigan algae}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{62}},
-  \bibinfo{pages}{802--815} (\bibinfo{year}{1981}).
-
-\bibitem{dutkiewicz2009}
-\bibinfo{author}{Dutkiewicz, S.}, \bibinfo{author}{Follows, M.~J.} \&
-  \bibinfo{author}{Bragg, J.~G.}
-\newblock \bibinfo{title}{{Modeling the coupling of ocean ecology and
-  biogeochemistry}}.
-\newblock \emph{\bibinfo{journal}{Global Biogeochem. Cycles}}
-  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{1--15} (\bibinfo{year}{2009}).
-
-\bibitem{barton2010}
-\bibinfo{author}{Barton, A.~D.}, \bibinfo{author}{Dutkiewicz, S.},
-  \bibinfo{author}{Flierl, G.}, \bibinfo{author}{Bragg, J.} \&
-  \bibinfo{author}{Follows, M.~J.}
-\newblock \bibinfo{title}{{Patterns of diversity in marine phytoplankton}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{327}},
-  \bibinfo{pages}{1509--1511} (\bibinfo{year}{2010}).
-
-\bibitem{Hempson:2015hka}
-\bibinfo{author}{Hempson, G.~P.}, \bibinfo{author}{Archibald, S.} \&
-  \bibinfo{author}{Bond, W.~J.}
-\newblock \bibinfo{title}{{A continent-wide assessment of the form and
-  intensity of large mammal herbivory in Africa}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{350}},
-  \bibinfo{pages}{1056--1061} (\bibinfo{year}{2015}).
-
-\bibitem{West:2002it}
-\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Woodruff, W.~H.} \&
-  \bibinfo{author}{Brown, J.~H.}
-\newblock \bibinfo{title}{{Allometric scaling of metabolic rate from molecules
-  and mitochondria to cells and mammals.}}
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{99 Suppl 1}}, \bibinfo{pages}{2473--2478}
-  (\bibinfo{year}{2002}).
-
-\bibitem{DeLong:2010dy}
-\bibinfo{author}{DeLong, J.~P.}, \bibinfo{author}{Okie, J.~G.},
-  \bibinfo{author}{Moses, M.~E.}, \bibinfo{author}{Sibly, R.~M.} \&
-  \bibinfo{author}{Brown, J.~H.}
-\newblock \bibinfo{title}{{Shifts in metabolic scaling, production, and
-  efficiency across major evolutionary transitions of life}}.
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{107}}, \bibinfo{pages}{12941--12945}
-  (\bibinfo{year}{2010}).
-
-\bibitem{West:2001bv}
-\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{Enquist, B.~J.}
-\newblock \bibinfo{title}{{A general model for ontogenetic growth}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{413}},
-  \bibinfo{pages}{628--631} (\bibinfo{year}{2001}).
-
-\bibitem{moses2008rmo}
-\bibinfo{author}{Moses, M.~E.} \emph{et~al.}
-\newblock \bibinfo{title}{{Revisiting a Model of Ontogenetic Growth: Estimating
-  Model Parameters from Theory and Data.}}
-\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{171}},
-  \bibinfo{pages}{632--645} (\bibinfo{year}{2008}).
-
-\bibitem{gillooly2002esa}
-\bibinfo{author}{Gillooly, J.~F.}, \bibinfo{author}{Charnov, E.~L.},
-  \bibinfo{author}{West, G.~B.}, \bibinfo{author}{Savage, V.~M.} \&
-  \bibinfo{author}{Brown, J.~H.}
-\newblock \bibinfo{title}{{Effects of size and temperature on developmental
-  time}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{417}},
-  \bibinfo{pages}{70--73} (\bibinfo{year}{2002}).
-
-\bibitem{hou}
-\bibinfo{author}{Hou, C.} \emph{et~al.}
-\newblock \bibinfo{title}{{Energy uptake and allocation during ontogeny}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{322}},
-  \bibinfo{pages}{736--739} (\bibinfo{year}{2008}).
-
-\bibitem{Savage:2004ed}
-\bibinfo{author}{Savage, V.~M.}, \bibinfo{author}{Gillooly, J.~F.},
-  \bibinfo{author}{Brown, J.~H.}, \bibinfo{author}{West, G.~B.} \&
-  \bibinfo{author}{Charnov, E.~L.}
-\newblock \bibinfo{title}{{Effects of Body Size and Temperature on Population
-  Growth}}.
-\newblock
-  \emph{\bibinfo{journal}{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}}
-  \textbf{\bibinfo{volume}{163}}, \bibinfo{pages}{429--441}
-  (\bibinfo{year}{2004}).
-
-\bibitem{bettencourt}
-\bibinfo{author}{Bettencourt, L. M.~A.}, \bibinfo{author}{Lobo, J.},
-  \bibinfo{author}{Helbing, D.}, \bibinfo{author}{Kuhnert, C.} \&
-  \bibinfo{author}{West, G.~B.}
-\newblock \bibinfo{title}{{Growth, innovation, scaling, and the pace of life in
-  cities}}.
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{104}}, \bibinfo{pages}{7301--7306}
-  (\bibinfo{year}{2007}).
-
-\bibitem{Folland:2008ij}
-\bibinfo{author}{Folland, J.~P.}, \bibinfo{author}{Mc~Cauley, T.~M.} \&
-  \bibinfo{author}{Williams, A.~G.}
-\newblock \bibinfo{title}{{Allometric scaling of strength measurements to body
-  size}}.
-\newblock \emph{\bibinfo{journal}{Eur J Appl Physiol}}
-  \textbf{\bibinfo{volume}{102}}, \bibinfo{pages}{739--745}
-  (\bibinfo{year}{2008}).
-
-\end{thebibliography}
-
-\end{bibunit}
-
-% \bibliography{aa_starving3}
-
-%
-% \begin{acknowledgments}
-%   We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
-%   discussions and comments on the manuscript.  J.D.Y. was supported by
-%   startup funds at the University of California, Merced, and an Omidyar
-%   Postdoctoral Fellowship at the Santa Fe Institute.  C.P.K. was supported by
-%   an Omidyar Postdoctoral Fellowship at the Santa Fe Institute.  S.R. was
-%   supported by grants DMR-1608211 and 1623243 from the National Science
-%   Foundation, and by the John Templeton Foundation, all at the Santa Fe
-%   Institute.
-% \end{acknowledgments}
-%
-\clearpage
-
-\begin{bibunit}[unsrt]
-
-% \addto\captionsenglish{\renewcommand{\figurename}{Figure S\!}}
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}%
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}%
-
-\section*{Supporting Information for ``The dynamics of starvation and recovery''}
-
-{\bf Mechanisms of Starvation and Recovery}
-To understand the dynamics of starvation, recovery, reproduction, and resource competition, our framework partitions consumers into two classes: (a) a full class that is able to reproduce and, (b) a hungry class that experiences mortality at a given rate and is unable to reproduce.
-For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate, as dictated by body size, with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
-\citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
-This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g.
-\citep{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \citep{kempes2014morphological}.
-Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).
-
-We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
-\begin{align}
-\begin{split}
-\dot{F_{d}} &= \lambda_{\text{max}} F_{d} + \rho_{\text{max}}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{\text{max}}R_{d} H_{d}/k - \mu H_{d},  \\
-\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
-& \left[\left(\frac{\rho_{\text{max}}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)F_{d}\right].
-\label{bigdynamics}
-\end{split}
-\end{align}
-where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with the subscript $_{d}$ before introducing the non-dimensional form that is presented in the main text).
-In the above equations $Y$ represents the yield coefficient (e.g., Refs.
-\citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism$^{-1}$).
-If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
-This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
-In these units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
-%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
-
-%\begin{align}
-%\begin{split}
-%\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-%\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
-%\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
-%&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
-%\end{split}
-%\end{align}
-%
-%In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
-Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
-The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
-%A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
-%
-%Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
-As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
-Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
-
-%where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
-
-We can formally non-dimensionalize this system by the rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
-%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
-\begin{align}
-\begin{split}
-&\dot{F} = \frac{1}{s}\left[\lambda_{\text{max}} F + \rho_{\text{max}}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
-&\dot{H} = \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{\text{max}}\frac{R}{qk} H - \mu H\right],  \\
-& \dot{R} = \\
-&\frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{\text{max}}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)F\right]\right].
-\end{split}
-\end{align}
-If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
-\begin{align}
-\begin{split}
-\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
-\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
-\label{reduceddynamics}
-\end{split}
-\end{align}
-where we have dropped the subscripts on $\lambda_{\text{max}}$ and $\rho_{\text{max}}$ for simplicity, and $\xi\equiv C/k$, $\delta\equiv Y_{H}kP_{H}/C$, and $\beta\equiv Y_{H}k\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)/C$.
-The above equations represent the system of equations presented in the main text.
-\\
-
-{\bf Parameter Values and Estimates}
+mammals and thus $\epsilon_{\rm \mu}$ is merely $\epsilon_{\rm \sigma}$ minus a constant.
+The time to go from starvation to death is the total time to reach $\epsilon_{\rm \mu}M$ minus the time to starve, or $t_{\rm \mu}=-M^{1-\eta}\ln\left(\epsilon_{\rm \mu}\right)/a^{\prime}-t_{\rm \sigma}$, and $\mu=1/t_{\rm \mu}$.\\
+
+\noindent {\bf Parameter Values and Estimates}\\
 All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
@@
 
-Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs.
-\citep{West:2001bv,moses2008rmo} and the supplement for Ref.
-\citep{hou}).
+Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$  \citep{West:2001bv,moses2008rmo,hou}.
 We make this assumption in the current paper, although alternate exponents, which are known to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
-The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.
+The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{\rm 0}$, given a set of data.
 For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs.
@@
 
-An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction.
-This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref.
+An important feature of our framework is the starting size, $m_{\rm 0}$, of a mammal which adjusts the overall timescales for reproduction.
+This starting size is known to follow an allometric relationship with adult mass of the form $m_{\rm 0}=n_{\rm 0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref.
 \citep{peters1986ecological} for a review).
-We use $m_{0}=0.097M^{0.92}$ \citep{blueweiss1978relationships} which encompasses the widest range of body sizes \citep{peters1986ecological}.
-
-The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g.
-Refs.
-\citep{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \citep{moses2008rmo}.
+We use $m_{\rm 0}=0.097M^{0.92}$ \citep{blueweiss1978relationships} which encompasses the widest range of body sizes \citep{peters1986ecological}.
+
+The energy to synthesize a unit of biomass, $E_{\rm m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) \citep{West:2001bv,moses2008rmo,hou} in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \citep{moses2008rmo}.
 The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \citep{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \citep{stryer,hou}.
@@
 
-For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ ~\citep{Dunbrack:1993ec}, $0.02M^{1.19}$ ~\citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ ~\citep{Lindstedt:2002td}.
+For the scaling of body composition it has been shown that fat mass follows $M_{\rm \rm fat}=f_{\rm 0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ ~\citep{Dunbrack:1993ec}, $0.02M^{1.19}$ ~\citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ ~\citep{Lindstedt:2002td}.
 We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range.
-Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ ~\citep{Lindstedt:2002td}.
-
-%We also connect the resource growth rate to the total metabolic rate of an organism.
-That is, we are interested in the relative rates of resource recovery and consumption by the total population.
-From \citep{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$).
-Considering an energy density of $18200$ (J g$^{-1}$) of grass \citep{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$).
-This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.
+Similarly, the muscle mass follows $M_{\rm \rm musc}=u_{\rm 0}M^{\zeta}$ with $u_{\rm 0}=0.383$ and $\zeta=1.00$ ~\citep{Lindstedt:2002td}.
 
@@
 Given that our model considers transitions over entire stages of ontogeny or nutritional states, the value of $Y$ must represent yields integrated over entire life stages.
-Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
+Given an energy density of $E_{\rm d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value is given by $P_{\rm F}=B_{\rm 0}M^{3/4}/ME_{\rm d}$, and the yield for a full organism will be given by $Y_{\rm F}=ME_{\rm d}/B_{\rm \lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\rm \lambda}$ is the lifetime energy use for reaching maturity given by
 \begin{equation}
-B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+B_{\rm \lambda}=\int_{\rm 0}^{t_{\rm \lambda}}B_{\rm 0}m\left(t\right)^{\eta}dt.
 \end{equation}
-Similarly, the maintenance resource consumption rate for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
+Similarly, the maintenance resource consumption rate for hungry individuals is $P_{\rm H}=B_{\rm 0}(\epsilon_{\rm \sigma}M)^{3/4}/(\epsilon_{\rm \sigma}M)E_{\rm d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{\rm H}=ME_{\rm d}/B_{\rm \rho}$ where
 \begin{equation}
-B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+B_{\rm \rho}=\int_{\rm \tau\left(\epsilon_{\rm \sigma}\epsilon_{\rm \lambda}\right)}^{t_{\rm \lambda}}B_{\rm 0}m\left(t\right)^{\eta}dt.
 \end{equation}
@@
 \citep{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
-It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with $B_{tot}$ for each type.
-As such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{damuth1987interspecific}.
+It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with $B_{\rm tot}$ for each type.
+As such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{Damuth:1987kr}.
 We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above.
-However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.
-\\
-%For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
-
-%More complicated models of fecundity (which, for example, account for the average length of adulthood and the number of individuals produced over this span) could be employed.
-However, the scaling of population growth rate has been studied in detail before and follows a relationship of $$ which matches the theory well for $\phi=.95$ and $=$.
-
-%In our calculations we include $20\%$ variation around this value which could account for differences in efficiency during
+However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.\\
+
+\noindent {\bf Rate equations for competitors with modified body mass}\\
+A resident population with mass $M$ competes for resources with a closely-related species with an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm \rm min} <0$ and $\chi_{\rm \rm max}>0$, where $\chi<0$ denotes a leaner competitor and $\chi > 0$ denotes a competitor with additional reserves of body fat.
+Importantly, we assume that the competing and resident individuals have the same proportion of non-fat tissues.
+For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\rm \sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (Eq.
+\ref{t1}), implies that $1-\epsilon_{\rm \lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
+Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\rm \lambda}-1)$ where the upper bound approximately equals $0.05$.
+
+Although the starved state of competing organisms remains unchanged, the rate of starvation from the modified full state to the starved state, the rate of recovery from the starved state to the modified full state, and the maintenance rates of both, will be different, such that $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$, $\delta^\prime = \delta(M^\prime)$.
+Rates of starvation and recovery for the competing population are derived by adjusting the starting or ending state before and after starvation and recovery, leading to the following timescales:
+
+\begin{align}
+t_{\rm \sigma^\prime} &= -\frac{M^{1-\eta}}{a^{\prime}}\ln \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
+t_{\rm \rho^\prime} &= \ln \left(\frac{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}{1-( \epsilon_\lambda(\chi +1))^{1/4}}\right)\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}.
+\end{align}
+
+The maintenance rates for the competing population require more careful consideration.
+First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of body fat reserves.
+Given an energy density of $E_{\rm d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value of the invading population is given by $P_{\rm F}=B_{\rm 0}(1+\chi)M^{3/4}/(1+\chi)ME_{\rm d}$, and the yield for a full organism will be given by $Y_{\rm F}=(1+\chi)ME_{\rm d}/B^{\prime}_{\rm \lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\rm \lambda}$ is the lifetime energy use for the competing population reaching maturity given by
+\begin{equation}
+B^{\prime}_{\rm \lambda}=\int_{\rm 0}^{t_{\rm \lambda^\prime}}B_{\rm 0}m\left(t\right)^{\eta}dt.
+\end{equation}
+where
+\begin{equation}
+t_{\rm \lambda^\prime} = \frac{M^{1-\eta} }{a(1-\eta)}\ln \left(\frac{1-(m_0/M)^{1-\eta}}{1-(\epsilon_\lambda (1+\chi))^{1-\eta}} \right).
+\end{equation}
+Note that we do not use this timescale to determine the reproductive rate of the competitor---which is assumed to remain the same as the resident population---but only to calculate the lifetime energy use.
+Similarly, the maintenance for hungry individuals $P^\prime_{\rm H}=B_{\rm 0}(\epsilon_{\rm \sigma}(1+\chi)M)^{3/4}/(\epsilon_{\rm \sigma}(1+\chi)M)E_{\rm d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y^\prime_{\rm H}=(1+\chi)ME_{\rm d}/B^{\prime}_{\rm \rho}$ where
+\begin{equation}
+B^{\prime}_{\rm \rho}=\int_{\rm \tau\left(\epsilon_{\rm \sigma}\epsilon_{\rm \lambda}\right)}^{t_{\rm \lambda^\prime}}B_{\rm 0}m\left(t\right)^{\eta}dt.
+\end{equation}
+Finally, we can calculate the maintenance of the competitors as
+
+\begin{align}
+  \delta^\prime &= P^\prime_{\rm H}Y^\prime_{\rm H}/\xi \\ \nonumber
+  \beta^\prime &= \left(\frac{\lambda_{\rm \rm max}}{Y^\prime_{\rm F}}+P^\prime_{\rm F} \right)Y^\prime_{\rm H}/\xi.
+\end{align}
+
+To determine whether or not the competitor or resident population has an advantage, we compute $R^*(M)$ and $R^*(M^\prime=M(1+\chi))$ for values of $\chi \in (-f_0M^{\gamma-1},1/\epsilon_{\rm \lambda}-1)$, and the competing population is assumed to have an advantage over the resident population if $R^*(M^\prime)<R^*(M)$.
+}\\
+
+\noindent {\bf Acknowledgments}\\
+We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, Seth Newsome, and three anonymous reviewers for helpful
+discussions and comments on the manuscript.  J.D.Y. was supported by
+startup funds at the University of California, Merced, and an Omidyar
+Postdoctoral Fellowship at the Santa Fe Institute.  C.P.K. was supported by
+an Omidyar Postdoctoral Fellowship at the Santa Fe Institute.  S.R. was
+supported by grants DMR-1608211 and 1623243 from the National Science
+Foundation, and by the John Templeton Foundation, all at the Santa Fe
+Institute.\\
+
+\noindent {\bf Data Availability}\\
+No unpublished datasets are used in this publication.\\
+
+\noindent {\bf Author Contributions}\\
+All authors contributed equally to the generation of ideas, analyses, and composition of the manuscript.\\
+
+\noindent {\bf Competing Interest Statement}\\
+The authors declare no competing interests.\\
+
+\def\bibfont{\footnotesize}
+\renewcommand{\bibsection}{\noindent {\bf References}}
+
+\bibliography{aa_starvingfull}
+
+\clearpage
+
+\noindent {\bf Tables}\\
 
@@
    Asymptotic adult mass & $M$ & (g) &  \\
-   Initial mass of an organism & $m_{0}$ & (g) &  \\
-   Metabolic rate scaling exponent & $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
-   Metabolic Normalization Constant & $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
+   Initial mass of an organism & $m_{\rm 0}$ & (g) &  \\
+   Metabolic rate scaling exponent & $\eta$ & $3/4$  &  \citep{West:2001bv,moses2008rmo,hou} \\
+   Metabolic Normalization Constant & $B_{\rm 0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
    Initial mass scaling exponent & $\upsilon$ & $0.92$ &  \citep{blueweiss1978relationships,peters1986ecological} \\
-   Initial mass scaling normalization constant & $n_{0}$ & $0.097$ (g$^{1-\upsilon}$) & \citep{blueweiss1978relationships,peters1986ecological}  \\   
+   Initial mass scaling normalization constant & $n_{\rm 0}$ & $0.097$ (g$^{1-\upsilon}$) & \citep{blueweiss1978relationships,peters1986ecological}  \\   
    Fat mass scaling exponent & $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
-   Fat scaling normalization constant & $f_{0}$ & $0.02$ (g$^{1-\eta}$) & \citep{Lindstedt:1985hm}\\
+   Fat scaling normalization constant & $f_{\rm 0}$ & $0.02$ (g$^{1-\eta}$) & \citep{Lindstedt:1985hm}\\
    Muscle mass scaling exponent & $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
-   Muscle scaling normalization constantv& $u_{0}$ & $0.38$ (g$^{1-\zeta}$)  & \citep{Lindstedt:2002td} \\
-   Energy to synthesis a unit of mass & $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
-   Energy to synthesis a unit of mass during recovery & $E_{m}^{\prime}$ & $7000$ (J gram$^{-1}$) & \citep{stryer,hou} \\
+   Muscle scaling normalization constantv& $u_{\rm 0}$ & $0.38$ (g$^{1-\zeta}$)  & \citep{Lindstedt:2002td} \\
+   Energy to synthesis a unit of mass & $E_{\rm m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
+   Energy to synthesis a unit of mass during recovery & $E_{\rm m}^{\prime}$ & $7000$ (J gram$^{-1}$) & \citep{stryer,hou} \\
    Specific resource growth rate & $\alpha$ & $9.45\times10^{-9}$ (s$^{-1}$) & see text  \\
-   Fraction of asymptotic mass representing full state & $\epsilon_{\lambda}$ & $0.95$ & \citep{West:2001bv}  \\
-   Fraction of asymptotic mass representing starving state & $\epsilon_{\sigma}$ & $1-f_{0}M^{\gamma-1}$ & see text  \\
-   Fraction of asymptotic mass representing death & $\epsilon_{\mu}$ & $1-\frac{f_{0}M^{\gamma}+u_{0}M^{\zeta}}{M}$ & see text \\
+   Fraction of asymptotic mass representing full state & $\epsilon_{\rm \lambda}$ & $0.95$ & \citep{West:2001bv}  \\
+   Fraction of asymptotic mass representing starving state & $\epsilon_{\rm \sigma}$ & $1-f_{\rm 0}M^{\gamma-1}$ & see text  \\
+   Fraction of asymptotic mass representing death & $\epsilon_{\rm \mu}$ & $1-\frac{f_{\rm 0}M^{\gamma}+u_{\rm 0}M^{\zeta}}{M}$ & see text \\
    Carrying capacity (maximum density) of resources & $C$ & (g m$^{-2}$) & \\
@@
    Normalized carrying capacity & $\xi$ & $C/k\approx2$ &   \\
-   Reproductive fecundity & $\nu$ & $2$ & \citep{}  \\
-
-%   $a$ & $1.78\times10^{-6}$  \quad \quad \\
-%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
-
+   Reproductive fecundity & $\nu$ & $2$ &  \\ 
  \hline
@@
 
-{\bf Rate equations for invaders with modified body mass}
-We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
-Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
-For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (see main text), implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
-Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
-
-Although the starved state of invading organisms remains unchanged, the rate of starvation from the modified full state to the starved state, the rate of recovery from the starved state to the modified full state, and the maintenance rates of both, will be different, such that $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$, $\delta^\prime = \delta(M^\prime)$.
-Rates of starvation and recovery for the invading population are easily derived by adjusting the starting or ending state before and after starvation and recovery, leading to the following timescales:
-
-\begin{align}
-t_{\sigma^\prime} &= -\frac{M^{1-\eta}}{a^{\prime}}\ln \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
-t_{\rho^\prime} &= \ln \left(\frac{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}{1-( \epsilon_\lambda(\chi +1))^{1/4}}\right)\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}.
-\end{align}
-
-The maintenance rates for the invading population require more careful consideration.
-First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of body fat reserves.
-Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
-\begin{equation}
-B^{\prime}_{\lambda}=\int_{0}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-where
-\begin{equation}
-t_{\lambda^\prime} = \frac{M^{1-\eta} }{a(1-\eta)}\ln \left(\frac{1-(m_0/M)^{1-\eta}}{1-(\epsilon_\lambda (1+\chi))^{1-\eta}} \right).
-\end{equation}
-Note that we do not use this timescale to determine the reproductive rate of the invading consumer---which is assumed to remain the same as the resident population---but only to calulate the lifetime energy use.
-Similarly, the maintenance for hungry individuals $P^\prime_{H}=B_{0}(\epsilon_{\sigma}(1+\chi)M)^{3/4}/(\epsilon_{\sigma}(1+\chi)M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y^\prime_{H}=(1+\chi)ME_{d}/B^{\prime}_{\rho}$ where
-\begin{equation}
-B^{\prime}_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-Finally, we can calculate the maintenance of the invaders as
-
-\begin{align}
-  \delta^\prime &= P^\prime_{H}Y^\prime_{H}/\xi \\ \nonumber
-  \beta^\prime &= \left(\frac{\lambda_{\rm max}}{Y^\prime_{F}}+P^\prime_{F} \right)Y^\prime_{H}/\xi.
-\end{align}
-
-To determine whether or not the invader or resident population has an advantage, we compute $R^*(M)$ and $R^*(M^\prime=M(1+\chi))$ for values of $\chi \in (-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$, and the invading population is assumed to have an advantage over the resident population if $R^*(M^\prime)<R^*(M)$.
-
-{\bf Sensitivity to additional death terms}
-
-It should be noted that our set of dynamics (Equations \ref{bigdynamics} and \ref{reduceddynamics}) could include a constant death term of the form $-d_{F}F$ and $-d_{H}H$ to represent death not directly linked to starvation.
-Adding terms of this form to our model would simply adjust the effective value of $\lambda$ and $\mu$, and we could rewrite Equation \ref{reduceddynamics} with $\lambda^{\prime}=\lambda-d$ and $\mu^{\prime}=\mu-d$.
-These substitutions would not alter the functional form of our model nor the steady-states and qualitative results, however the quantitative values could shift based on the size of $d$ relative to $\lambda$ and $\mu$.
-
-Survivorship has a well-known functional form which changes systematically with size (e.g. \cite{calder1984}). Typically survivorship is defined using the Gompertz curve 
-\begin{equation}
-F=F_{0}e^{\left(c_{0}/c_{1}\right)\left(1-e^{c_{1}t}\right)}
-\label{gompertz}
-\end{equation}
-where the parameters have the following allometric dependencies on adult mass $c_{0}=a_{0}M^{b_{0}}$ and $c_{1}=a_{1}M^{b_{1}}$, with $a_{0}=1.88\times10^{-8}$ (s g$^{-b_{0}}$), $b_{0}=-0.56$, $a_{1}=1.45\times10^{-7}$ (s g$^{-b_{1}}$), and $b_{1}=-0.27$ (see \cite{calder1984} for a review).
-
-\begin{figure}
+\clearpage
+
+\noindent {\bf Figure Legends}\\
+
+\begin{figure*}[h!]
 \centering
-\includegraphics[width=0.4\textwidth]{mortality-rate-comparison-eps-converted-to.pdf}
-\caption{\small{The rates of reproduction $\lambda$ (blue), starvation-based mortality $\mu$ (red), and survivorship-based death $\bar{d}$ (black) as a function of adult mass.}\label{fig:ratescomp}}
-\end{figure}
-
-We are interested in the specific death rate of the form $\dot{F}=-dF$, and using the derivative of Equation \ref{gompertz} we find that $d=c_{0}e^{c_{1}t}$.
-Our model considers the average rates over a population and lifecycle and the average death rate is given by
-\begin{eqnarray}
-\bar{d}&=&\frac{1}{t_{\text{exp}}}\int_{0}^{t_{\text{exp}}}c_{0}e^{c_{1}t} dt \\
-&=&\frac{c_{0}\left(e^{c_{1} t_{\text{exp}}}-1\right)}{c_{1}t_{\text{exp}}}
-\end{eqnarray}
-where $t_{\text{exp}}$ is the expected lifespan following the allometry of $t_{\text{exp}}=a_{2}M^{b_{2}}$ with $a_{2}=4.04\times10^{6}$ (s g$^{-b_{2}}$) and $b_{2}=0.30$ ~\cite{damuth1982analysis,calder1984}.
-Given the allometries above we have that
-\begin{equation}
-\bar{d}=\frac{a_{0} \left(e^{a_{1}a_{2}M^{b_{1}+b_{2}}}-1\right) M^{b_{0}-b_{1}-b_{2}}}{a_{1} a_{2}}
-\end{equation}
-which scales roughly like $M^{b_{0}}$ because $b_{1}$ and $b_{2}$ are close in value but opposite in sign.
-In Figure S\ref{fig:ratescomp} we compare the value of $\bar{d}$ to the reproductive, $\lambda$, and starvation-based mortality, $\mu$, rates.
-The values of $\bar{d}$ are orders of magnitude smaller than these rates for all mammalian masses, and thus, adding this non-starvation based death rate to our model does not shift our results within numerical confidence.
-
-\begin{figure}[h!]
+% \includegraphics[width=0.45\textwidth]{fig_FixedPoint2.eps}
+\caption{\small{ The transcritical (TC; dashed line) and Hopf bifurcation (solid line) as a
+  function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
+  bifurcation conditions separate parameter space into unphysical (left of the TC), cyclic,
+  and steady state dynamic regimes.  The colors show the steady state densities for the energetically replete consumers $F^*$.  
+  % Steady state densities for the energetically deficient consumers $H^*$ (not shown)
+  % scale with those for $F^*$.  
+}\label{fig:fp}}
+\end{figure*}
+
+\begin{figure*}[h!]
 \centering
-\includegraphics[width=0.4\textwidth]{fig_FPenergyequiv-eps-converted-to.pdf}
-\caption{\small{ Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass ($F^*$ is shown in green and $H^*$ in orange).
-The data are from Damuth \citep{Damuth:1987kr} and have been converted to
-  total population metabolism using the allometric relationships for
-  metabolic rate (e.g. Refs.~\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:equivalence}}
-\end{figure}
-
-{\bf NSM and the energy equivalence hypothesis}
-
-The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}.
-As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}.
-However, the actual equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations.
-Figure S\ref{fig:equivalence} recasts the results of the NSM in terms of this hypothesis and shows that $F^{*}B$ is nearly constant over the same range of mammalian sizes up to the asymptotic behavior for the largest terrestrial mammals.
-
-{\bf Application of NSM limits to aquatic mammals}
-A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass of a male African elephant.
-We note this particular limit as it may have future relevance to considerations of the ultimate constraints on aquatic mammals.
-
-% \putbib[aa_starving_supplement]
-
-\def\bibfont{\footnotesize}
-
-% \bibliography{aa_starving_supplement}
-\begin{thebibliography}{10}
-\expandafter\ifx\csname url\endcsname\relax
-  \def\url#1{\texttt{#1}}\fi
-\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
-\providecommand{\bibinfo}[2]{#2}
-\providecommand{\eprint}[2][]{\url{#2}}
-
-\bibitem{Kempes:2012hy}
-\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Dutkiewicz, S.} \&
-  \bibinfo{author}{Follows, M.~J.}
-\newblock \bibinfo{title}{{Growth, metabolic partitioning, and the size of
-  microorganisms.}}
-\newblock \emph{\bibinfo{journal}{PNAS}} \textbf{\bibinfo{volume}{109}},
-  \bibinfo{pages}{495--500} (\bibinfo{year}{2012}).
-
-\bibitem{kempes2014morphological}
-\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Okegbe, C.},
-  \bibinfo{author}{Mears-Clarke, Z.}, \bibinfo{author}{Follows, M.~J.} \&
-  \bibinfo{author}{Dietrich, L.~E.}
-\newblock \bibinfo{title}{Morphological optimization for access to dual
-  oxidants in biofilms}.
-\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
-  Sciences}} \textbf{\bibinfo{volume}{111}}, \bibinfo{pages}{208--213}
-  (\bibinfo{year}{2014}).
-
-\bibitem{West:2001bv}
-\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{Enquist, B.~J.}
-\newblock \bibinfo{title}{{A general model for ontogenetic growth}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{413}},
-  \bibinfo{pages}{628--631} (\bibinfo{year}{2001}).
-
-\bibitem{moses2008rmo}
-\bibinfo{author}{Moses, M.~E.} \emph{et~al.}
-\newblock \bibinfo{title}{{Revisiting a model of ontogenetic growth: Estimating
-  model parameters from theory and data}}.
-\newblock
-  \emph{\bibinfo{journal}{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}}
-  \textbf{\bibinfo{volume}{171}}, \bibinfo{pages}{632--645}
-  (\bibinfo{year}{2008}).
-
-\bibitem{hou}
-\bibinfo{author}{Hou, C.} \emph{et~al.}
-\newblock \bibinfo{title}{{Energy uptake and allocation during ontogeny}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{322}},
-  \bibinfo{pages}{736--739} (\bibinfo{year}{2008}).
-
-\bibitem{pirt}
-\bibinfo{author}{Pirt, S.}
-\newblock \bibinfo{title}{The maintenance energy of bacteria in growing
-  cultures}.
-\newblock \emph{\bibinfo{journal}{Proceedings of the Royal Society of London B:
-  Biological Sciences}} \textbf{\bibinfo{volume}{163}},
-  \bibinfo{pages}{224--231} (\bibinfo{year}{1965}).
-
-\bibitem{Heijnen}
-\bibinfo{author}{Heijnen, J.} \& \bibinfo{author}{Roels, J.}
-\newblock \bibinfo{title}{A macroscopic model describing yield and maintenance
-  relationships in aerobic fermentation processes}.
-\newblock \emph{\bibinfo{journal}{Biotechnology and Bioengineering}}
-  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{739--763}
-  (\bibinfo{year}{1981}).
-
-\bibitem{peters1986ecological}
-\bibinfo{author}{Peters, R.~H.}
-\newblock \emph{\bibinfo{title}{The Ecological Implications of Body Size}},
-  vol.~\bibinfo{volume}{2} (\bibinfo{publisher}{Cambridge University Press},
-  \bibinfo{address}{Cambridge}, \bibinfo{year}{1986}).
-
-\bibitem{blueweiss1978relationships}
-\bibinfo{author}{Blueweiss, L.} \emph{et~al.}
-\newblock \bibinfo{title}{Relationships between body size and some life history
-  parameters}.
-\newblock \emph{\bibinfo{journal}{Oecologia}} \textbf{\bibinfo{volume}{37}},
-  \bibinfo{pages}{257--272} (\bibinfo{year}{1978}).
-
-\bibitem{stryer}
-\bibinfo{author}{Stryer, L.}
-\newblock \emph{\bibinfo{title}{{Biochemistry, Fourth Edition}}}
-  (\bibinfo{publisher}{W.H. Freeman and Company}, \bibinfo{address}{New York},
-  \bibinfo{year}{1995}).
-
-\bibitem{Dunbrack:1993ec}
-\bibinfo{author}{Dunbrack, R.~L.} \& \bibinfo{author}{Ramsay, M.~A.}
-\newblock \bibinfo{title}{{The Allometry of Mammalian Adaptations to Seasonal
-  Environments: A Critique of the Fasting Endurance Hypothesis}}.
-\newblock \emph{\bibinfo{journal}{Oikos}} \textbf{\bibinfo{volume}{66}},
-  \bibinfo{pages}{336--342} (\bibinfo{year}{1993}).
-
-\bibitem{Lindstedt:1985hm}
-\bibinfo{author}{Lindstedt, S.~L.} \& \bibinfo{author}{Boyce, M.~S.}
-\newblock \bibinfo{title}{{Seasonality, Fasting Endurance, and Body Size in
-  Mammals}}.
-\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{125}},
-  \bibinfo{pages}{873--878} (\bibinfo{year}{1985}).
-
-\bibitem{Lindstedt:2002td}
-\bibinfo{author}{Lindstedt, S.~L.} \& \bibinfo{author}{Schaeffer, P.~J.}
-\newblock \bibinfo{title}{{Use of allometry in predicting anatomical and
-  physiological parameters of mammals.}}
-\newblock \emph{\bibinfo{journal}{Lab. Anim.}} \textbf{\bibinfo{volume}{36}},
-  \bibinfo{pages}{1--19} (\bibinfo{year}{2002}).
-
-\bibitem{estermann}
-\bibinfo{author}{Estermann, B.~L.}, \bibinfo{author}{Wettstein, H.-R.},
-  \bibinfo{author}{Sutter, F.} \& \bibinfo{author}{Kreuzer, M.}
-\newblock \bibinfo{title}{Nutrient and energy conversion of grass-fed dairy and
-  suckler beef cattle kept indoors and on high altitude pasture}.
-\newblock \emph{\bibinfo{journal}{Animal Research}}
-  \textbf{\bibinfo{volume}{50}}, \bibinfo{pages}{477--493}
-  (\bibinfo{year}{2001}).
-
-\bibitem{michaletz2014convergence}
-\bibinfo{author}{Michaletz, S.~T.}, \bibinfo{author}{Cheng, D.},
-  \bibinfo{author}{Kerkhoff, A.~J.} \& \bibinfo{author}{Enquist, B.~J.}
-\newblock \bibinfo{title}{Convergence of terrestrial plant production across
-  global climate gradients}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{512}},
-  \bibinfo{pages}{39--43} (\bibinfo{year}{2014}).
-
-\bibitem{damuth1987interspecific}
-\bibinfo{author}{Damuth, J.}
-\newblock \bibinfo{title}{Interspecific allometry of population density in
-  mammals and other animals: the independence of body mass and population
-  energy-use}.
-\newblock \emph{\bibinfo{journal}{Biological Journal of the Linnean Society}}
-  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
-  (\bibinfo{year}{1987}).
-
-\bibitem{calder1984}
-\bibinfo{author}{Calder, W.~A.}
-\newblock \emph{\bibinfo{title}{Size, function, and life history}}
-  (\bibinfo{publisher}{Harvard University Press}, \bibinfo{year}{1984}).
-
-\bibitem{damuth1982analysis}
-\bibinfo{author}{Damuth, J.}
-\newblock \bibinfo{title}{Analysis of the preservation of community structure
-  in assemblages of fossil mammals}.
-\newblock \emph{\bibinfo{journal}{Paleobiology}} \textbf{\bibinfo{volume}{8}},
-  \bibinfo{pages}{434--446} (\bibinfo{year}{1982}).
-
-\bibitem{allen2002}
-\bibinfo{author}{Allen, A.~P.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{Gillooly, J.~F.}
-\newblock \bibinfo{title}{{Global biodiversity, biochemical kinetics, and the
-  energetic-equivalence rule}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{297}},
-  \bibinfo{pages}{1545--1548} (\bibinfo{year}{2002}).
-
-\bibitem{enquist1998}
-\bibinfo{author}{Enquist, B.~J.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{West, G.~B.}
-\newblock \bibinfo{title}{{Allometric scaling of plant energetics and
-  population density}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{395}},
-  \bibinfo{pages}{163--165} (\bibinfo{year}{1998}).
-
-\bibitem{Damuth:1987kr}
-\bibinfo{author}{Damuth, J.}
-\newblock \bibinfo{title}{{Interspecific allometry of population density in
-  mammals and other animals: the independence of body mass and population
-  energy-use}}.
-\newblock \emph{\bibinfo{journal}{Biol. J. Linn. Soc.}}
-  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
-  (\bibinfo{year}{1987}).
-
-\end{thebibliography}
-
-\end{bibunit}
+% \includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.eps}
+\caption{\small{ The growth trajectory over absolute time of an individual organism as a function of body mass.  
+Initial growth follows the black trajectory to an energetically replete reproductive adult mass of $m=\epsilon_\lambda M$ (see Methods). %which we assume is 95\% asymptotic mass $M$.  
+Starvation follows the red trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. 
+Recovery follows the green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires a longer time to reach $\epsilon_\lambda M$.
+%different energetics and
+Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}\label{fig:growth}}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+% \includegraphics[width=0.4\textwidth]{fig_ExtinctionAllometricComb4.eps}
+\caption{\small{ Probability of extinction for a consumer with ({\bf a}) $M=10^2$g and ({\bf b}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
+Note the change in scale in panel {\bf b}.
+Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
+The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+% \includegraphics[width=0.4\textwidth]{fig_FPAllometric2.eps}
+\caption{\small{Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
+  body mass along with the data from Damuth \citep{Damuth:1987kr}. Inset: Resource steady state $R^*$ as a function of consumer body mass.}\label{fig:mass}}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+% \includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
+\caption{\small{Competitive outcomes for a resident species with body mass
+    $M$ vs. a closely related competing species with modified body mass
+    $M^\prime=M(1+\chi)$.  The blue region denotes proportions of modified
+    mass $\chi$ resulting in exclusion of the resident species.  The red
+    region denotes values of $\chi$ that result in a mass that is below the
+    starvation threshold and are thus infeasible.  Arrows point to the
+    predicted optimal mass from our model $M_{\rm \rm opt}=1.748\times 10^7$,
+    which may serve as an evolutionary attractor for body mass.  The black
+    wedge points to the largest body mass known for terrestrial mammals
+    (\emph{Deinotherium} spp.) at $1.74\times10^7$
+    (g)~\citep{Smith:2010p3442}.}\label{fig:invasion}}
+\end{figure*}
 

### 2017-12-13 NatureComm_REVISION/YKR_NatureComm-Revision-3.tex
editorial revisions

@@
 Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
-{\color{red} We note that including additional allometrically-scaled mortality terms to both $F$ and $H$ does not impact our findings (see SI for derivation).}
+{\color{red} We note that including additional allometrically-scaled mortality terms to both $F$ and $H$ does not change the form of our model nor impact our quantitative findings (see SI for derivation).}
 
@@
 Following this relationship the growth rate is given by $\lambda=\ln\left(\nu\right)/t_{\lambda}$, which is the standard relationship (e.g.
-\cite{savage2004effects}) and will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of $\epsilon_{\lambda}$
+\cite{savage2004effects}) and will scales as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of $\epsilon_{\lambda}$
 \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.

### 2017-12-13 NatureComm_REVISION/YKR_NatureComm-Revision-4.tex
editorial revisions

--- 2017-11-03 NatureComm_REVISION/YKR_NatureComm-Revision-3.tex
@@
 Of particular importance is the interplay between reproduction and the dynamics of starvation and recovery. % in resource-limited environments
-We introduce a nutritional state-structured model that incorporates two classes of consumer: nutritionally replete, reproducing consumers, and undernourished, non-reproducing consumers.
+To elucidate this interplay, we introduce a nutritional state-structured model that incorporates two classes of consumer: nutritionally replete, reproducing consumers, and undernourished, non-reproducing consumers.
 % that are susceptible to mortality
@@
 %subject to these constraints between body size and a variety of traits
-Moreover, we find that these rates fall within a `refuge' in parameter space, where the probability of extinction of the consumer population is minimized. 
+Moreover, these rates fall within a `refuge' in parameter space, where the probability of population extinction is minimized. 
 % Thus we identify a potential mechanism that may both drive and constrain the dynamics of animal populations. 
-We also show that our model provides a natural framework that predicts
-maximum body size for mammals by determining the relative stability of an
+We also show that our model provides a natural framework to predict
+maximum mammalian body size by determining the relative stability of an
 otherwise homogeneous population to a competing population with altered percent
-body fat, providing a principled mechanism for a within-lineage driver of
+body fat. This framework provides a principled mechanism for a within-lineage driver of
 Cope's rule.
@@
 
-The behavioral ecology of all organisms is influenced by the energetic state of individuals, which directly impacts how they invest reserves in uncertain environments.
-Such behaviors are generally manifested as tradeoffs between investing in somatic maintenance and growth, or allocating energy towards reproduction~\citep{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.
-The timing of these behaviors responds to selective pressure, as the choice of the investment impacts future fitness~\citep{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.
-The influence of resource limitation on an organism's ability to maintain its nutritional stores may lead to repeated delays or shifts in reproduction over the course of an organism's life.
+The behavioral ecology of all organisms is influenced by their energetic
+states, which directly impacts how they invest reserves in uncertain
+environments.  Such behaviors are generally manifested as tradeoffs between
+investing in somatic maintenance and growth, or allocating energy towards
+reproduction~\citep{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.  The timing of
+these behaviors responds to selective pressure, as the choice of the
+investment impacts future
+fitness~\citep{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.  The influence of
+resource limitation on an organism's ability to maintain its nutritional
+stores may lead to repeated delays or shifts in reproduction over the course
+of an organism's life.
 
@@
 For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\citep{Tveraa:2003fq}.
-Many bird species invest differently in broods during periods of resource scarcity compared to normal periods~\citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\citep{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
+Many bird species invest differently in broods during periods of resource scarcity~\citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\citep{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
 Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\citep{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\citep{Kirk:1997cc}.
@@
 
-Including individual energetic dynamics~\citep{Kooi2000} in a population-level framework~\citep{Kooi2000,Sousa:2010ez} is challenging~\citep{Diekmann:2010da}, and
-a common simplifying approach is provided in the classic Lotka-Volterra (LV) model, which assumes that consumer population growth rate depends linearly on resource
-  density~\citep{murdoch:2003}. Here, we introduce an alternative
-  approach---the Nutritional State-structured Model (NSM)---that accounts for
-  resource limitation via explicit starvation. In contrast to
-  the LV model, the NSM incorporates two consumer states:
-  hungry and full, with only the former susceptible to mortality and only the
-  latter possessing sufficient energetic reserves to reproduce.
-  Additionally, we incorporate allometrically derived constraints on reproduction \citep{Kempes:2012hy}, incorporating the timescales of starvation and recovery. 
-  As we shall show, our model makes several important predictions: 
-  (i) the dynamics are typically driven to a refuge far from cyclic behavior and extinction risk, 
-  (ii) steady state conditions of the NSM accurately predict the measured biomass densities for mammals described by Damuth's law,
+Including individual energetic dynamics~\citep{Kooi2000} in a
+population-level framework~\citep{Kooi2000,Sousa:2010ez} is
+challenging~\citep{Diekmann:2010da}.  A common simplifying approach is the
+classic Lotka-Volterra (LV) model, which assumes that consumer population
+growth rate depends linearly on resource density~\citep{murdoch:2003}. Here,
+we introduce an alternative approach---the Nutritional State-structured Model
+(NSM)---that accounts for resource limitation via explicit starvation. In
+contrast to the LV model, the NSM incorporates two consumer states: hungry
+and full, with only the former susceptible to mortality and only the latter
+possessing sufficient energetic reserves to reproduce.  Additionally, we
+incorporate allometrically derived constraints on the time scales for
+reproduction~\citep{Kempes:2012hy}, starvation, and recovery.  Our model
+makes several basic predictions: (i) the dynamics are typically driven to a
+refuge far from cyclic behavior and extinction risk, (ii) the steady-state
+conditions of the NSM accurately predict the measured biomass densities for
+mammals described by Damuth's law \citep{Damuth:1987kr,allen2002,enquist1998,pedersen2017shallow},
   %  energetic equivalence principle, 
   (iii) there is an allometrically constrained upper-bound for mammalian body size, and
-  (iv) the NSM provides a selective mechanism for the evolution of larger body size, known as Cope's rule.
+  (iv) the NSM provides a selective mechanism for the evolution of larger body size, known as Cope's rule ~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}.
 
@@
 \centering
-\includegraphics[width=0.4\textwidth]{fig_FixedPoint2.eps}
+\includegraphics[width=0.45\textwidth]{fig_FixedPoint2.eps}
 \caption{\small{ The transcritical (TC; dashed line) and Hopf bifurcation (solid line) as a
@@
 rate $\sigma$---the starvation rate---and also in proportion to the absence
-of resources $(1-R)$ (we assume a maximum density, the carrying capacity equal to 1).  Conversely, consumers recover from state $H$ to state
-$F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$ represents a
-ratio between maximal resource consumption and the carrying capacity of the
-resource. %and $\approx 2$.
+of resources $(1-R)$ (the maximum resource density has been non
+dimensionalized to 1; see SI).  Conversely, consumers recover from state $H$
+to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$
+represents a ratio between maximal resource consumption and the carrying
+capacity of the resource. %and $\approx 2$.
 The resources that are eaten by hungry consumers (at rate $\rho R + \delta$)
@@
 consumers eat resources at a constant rate $\beta$ that accounts for maximal
-maintenance and somatic growth (see the SI for mechanistic
-derivations of these rates from resource energetics).
+maintenance and somatic growth (see the SI for mechanistic derivations of
+these rates from resource energetics).
 %This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
-The NSM represents an ecologically motivated fundamental extension of the idealized starving random walk model of foraging, which focuses on resource depletion, to include reproduction and resource replenishment~\citep{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is a more general formulation than previous models incorporating starvation~\citep{Persson:1998hz}.
+The NSM represents an ecologically motivated fundamental extension of the
+idealized starving random walk model of foraging, which focuses on resource
+depletion, to include reproduction and resource
+replenishment~\citep{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is
+a more general formulation than previous models that incorporate
+starvation~\citep{Persson:1998hz}.
 
@@
 % As discussed above, the attributes of reproduction and mortality have been explicitly apportioned to the full and hungry consumers, respectively, so that the growth in the total density is decoupled from the resource density.
-The rate of reproduction is independent of resource density because it is
-assumed that the satiated state of the full consumer allows it to partition a
-constant amount of energy towards reproduction, whereas a starved consumer
-partitions no energy towards reproduction.  The rate of reproduction for the
-total consumer density \emph{is} dependent on resource density, which
-determines the size of the full and starved portions of the consumer
-population.  Similarly, the consumer maintenance terms ($\delta H$ and
-$\beta F$) are independent of resource density because they represent a
-minimal energetic requirement for consumers in the $H$ and $F$ state,
-respectively.  It follows that model predictions are robust only when $R$ is
-of the order of 1, which holds for all cases that we explore.
+The rate of reproduction is independent of resource density because the full
+consumer partitions a constant amount of energy towards reproduction, whereas
+a hungry consumer partitions no energy towards reproduction.  Similarly, the
+consumer maintenance terms ($\delta H$ and $\beta F$) are also independent of
+resource density because they represent a minimal energetic requirement for
+consumers in the $H$ and $F$ state, respectively.
 
-\noindent \paragraph*{ {\bf Steady states of the NSM.}} From the solution to the single internal fixed point (Eq.~\eqref{eq:ss}, see Methods), an obvious constraint on the NSM is that the reproduction rate $\lambda$ must be less than the starvation rate $\sigma$, so that the consumer and resource densities are positive.
+% It follows that model predictions are robust only when $R$ is of the order of 1, which holds for all cases that we explore.
+
+\noindent \paragraph*{ {\bf Steady states of the NSM.}} From the single
+internal fixed point (Eq.~\eqref{eq:ss}, see Methods), an obvious constraint
+on the NSM is that the reproduction rate $\lambda$ must be less than the
+starvation rate $\sigma$, so that the consumer and resource densities are
+positive.
 %In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
-The condition $\sigma = \lambda$ thus represents a transcritical (TC) bifurcation~\citep{Strogatz:2008wo} that demarcates a physical from an unphysical regime where all steady-state densities become negative after intersecting the trivial fixed point $(F^*,H^*,R^*)=(0,0,0)$.
-The biological implication of the constraint $\lambda<\sigma$ has a simple interpretation---the rate at which a macroscopic organism loses mass due to lack of resources is generally much faster than the rate of reproduction.
-As we will discuss below, this inequality is a natural consequence of allometric constraints~\citep{Kempes:2012hy} for organisms within empirically observed body size ranges. %(Fig.~\ref{fig:gvs})
+The condition $\sigma = \lambda$ represents a transcritical (TC)
+bifurcation~\citep{Strogatz:2008wo} that demarcates a physical from an
+unphysical (negative steady-state densities) regime.  The biological
+implication of the constraint $\lambda<\sigma$ has a simple
+interpretation---the rate at which a macroscopic organism loses mass due to
+lack of resources is generally much faster than the rate of reproduction.  As
+we will discuss below, this inequality is also a natural consequence of
+allometric constraints~\citep{Kempes:2012hy} for organisms within empirically
+observed body size ranges. %(Fig.~\ref{fig:gvs})
 
@@
 
-When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the consumers are driven to the hungry non-reproducing state.
-Because reproduction is inhibited, there is a low steady-state consumer density and a high steady-state resource density.
-However, if $\sigma/\lambda\to 1$ from above, the population is overloaded with energetically-replete (reproducing) individuals, thereby promoting transient oscillations between the consumer and resource densities (Fig.~\ref{fig:fp}).
-If the starvation rate is low enough that the Hopf bifurcation is crossed, these oscillations become stable over time.
+When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the
+consumers are driven to the hungry non-reproducing state.  Because
+reproduction is inhibited, there is a low steady-state consumer density and a
+high steady-state resource density.  However, if $\sigma/\lambda\to 1$ from
+above, the population is overloaded with energetically-replete (reproducing)
+individuals, thereby promoting transient oscillations between the consumer
+and resource densities (Fig.~\ref{fig:fp}).  If the starvation rate is low
+enough that the Hopf bifurcation is crossed, these oscillations become
+stable.
 This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.\\
@@
 \centering
-\includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.eps}
+\includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.pdf}
 \caption{\small{ The growth trajectory over absolute time of an individual organism as a function of body mass.  
-Initial growth follows the black trajectory to an energetically replete reproductive adult mass $m=\epsilon_\lambda M$. %which we assume is 95\% asymptotic mass $M$.  
+Initial growth follows the black trajectory to an energetically replete reproductive adult mass of $m=\epsilon_\lambda M$ (see Methods). %which we assume is 95\% asymptotic mass $M$.  
 Starvation follows the red trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. 
@@
 \noindent {\bf Results}
-\noindent \paragraph*{{\bf The allometry of extinction risk.}} While there are no {\it a priori} constraints on the parameters in the NSM, we expect that each species should be restricted to a distinct portion of the parameter space.
-We
-use allometric scaling relations to constrain the covariation of rates in a
-principled and biologically meaningful manner (see Methods).  Allometric scaling relations
-highlight common constraints and average trends across large ranges in body
-size and species diversity. Many of these relations can be derived from a
-small set of assumptions and in the Methods we describe our framework to determine the
-covariation of timescales and rates across a range of body sizes for each of
-the key parameters of our model (cf. ref.~\citep{Yodzis:1992hg}).  
+\noindent \paragraph*{{\bf The allometry of extinction risk.}} While there
+are no {\it a priori} constraints on the parameters in the NSM, we expect
+that each species should be restricted to a distinct portion of the parameter
+space.  We use allometric scaling relations to constrain the covariation of
+rates in a principled and biologically meaningful manner (see Methods).
+Allometric scaling relations highlight common constraints and average trends
+across large ranges in body size and species diversity. Many of these
+relations can be derived from a small set of assumptions.  In the Methods we
+describe our framework to determine the covariation of timescales and rates
+across a range of body sizes for each of the key parameters of our model
+(cf.\ Ref.~\citep{Yodzis:1992hg}).
 % We are thereby able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.
@@
 metabolism, which can be used to describe a variety of organismal features
-\citep{Brown:2004wq}.
-We derive relationships for the rates of reproduction, starvation, recovery, and mortality based on first principles, and as a function of an organism's body size and metabolic rate (see Methods).
-Because we aim to explore the starvation-recovery dynamics as a function of an organism's body mass $M$, we parameterize these rates in terms of the \emph{percent} gain and loss of the asymptotic (maximum) body mass, $\epsilon M$, where different values of $\epsilon$ define different states of the consumer (Fig.~\ref{fig:growth}; see Methods for derivations of allometrically constrained rate equations).
-Although the rate equations \eqref{eq:system} are general and can in
-principle be used to explore the starvation recovery dynamics for most
-organisms, here we focus on allometric relationships for terrestrial-bound
-lower trophic level endotherms (see the SI for values), specifically herbivorous mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
-Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$.
-Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.
-{\color{red} We note that including additional allometrically-scaled mortality terms to both $F$ and $H$ does not impact our findings (see SI for derivation).}
+\citep{Brown:2004wq}.  We derive, from first principles, the relationships
+for the rates of reproduction, starvation, recovery, and mortality as a
+function of an organism's body size and metabolic rate (see Methods).
+Because we aim to explore the starvation-recovery dynamics as a function of
+an organism's body mass $M$, we parameterize these rates in terms of the
+\emph{percent} gain and loss of the asymptotic (maximum) body mass,
+$\epsilon M$, where different values of $\epsilon$ define different states of
+the consumer (Fig.~\ref{fig:growth}; see Methods for derivations of
+allometrically constrained rate equations).  Although the rate equations
+\eqref{eq:system} are general and can in principle be used to explore the
+starvation recovery dynamics for most organisms, here we focus on allometric
+relationships for terrestrial-bound lower-trophic level endotherms (see the
+SI for values), specifically herbivorous mammals, which range from a minimum
+of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of
+$M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene
+Deinotheriinae).  Investigating other classes of organisms would simply
+involve altering the metabolic exponents and scalings associated with
+$\epsilon$. Moreover, we emphasize that our allometric equations (see
+Methods) describe mean relationships, and do not account for the (sometimes
+considerable) variance associated with individual species.  We note that
+including additional allometrically-scaled mortality terms to both $F$ and
+$H$ does not change the form of our model nor impact our quantitative findings
+(see SI for the derivation).
 
@@
 % \noindent \paragraph*{{\bf Extinction risk}} 
-As the allometric derivations of the NSM rate laws reveal (see Methods), starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
-Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
-Indeed, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation. 
-This suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.
+As the allometric derivations of the NSM rate laws reveal (see Methods),
+starvation and recovery rates are not independent parameters, and the
+biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp}
+is constrained via covarying parameters.  Given the parameters of terrestrial
+endotherms, we find that the starvation rate $\sigma$ and the recovery rate
+$\rho$ are constrained to lie within a small region of potential values for
+the known range of body sizes $M$.  Indeed, starvation and recovery rates
+across all values of $M$ fall squarely in the steady-state region at some
+distance from the Hopf bifurcation.  This suggests that cyclic population
+dynamics should be rare, particularly in resource-limited environments.
 % Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
@@
 From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
-We therefore examine the competing effects of cyclic dynamics vs. changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
+We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
 To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
@@
 %the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
-In each replicate the initial densities were chosen to be $(XF^*,XH^*,R^*)$, with $X$ a random variable uniformly distributed in $[0,2]$.
-By allowing the rate of starvation to vary, we assessed extinction risk across a range of values for $\sigma$ and $\rho$ between ca.\ $10^{-8}$ to $10^{-3}$.
-%, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
-Higher rates of extinction correspond to both high values of $\sigma$ if $\rho$ is small, and high values of $\rho$ if $\sigma$ is small.
-For low values of $\sigma$ and high values of $\rho$, the increased extinction risk results from higher-amplitude transient cycles as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).
-For high values of $\sigma$ and low values of $\rho$, increased extinction risk arises because of the decrease in the steady-state consumer population density (Figs. \ref{fig:fp}b, \ref{fig:ext}).
-This interplay creates an `extinction refuge', such that for a constrained range of $\sigma$ and $\rho$, extinction probabilities are minimized.
+In each replicate the initial densities were chosen to be $(XF^*,XH^*,R^*)$,
+with $X$ a random variable uniformly distributed in $[0,2]$.  By allowing the
+rate of starvation to vary, we assessed extinction risk across a range of
+values for $\sigma$ and $\rho$ between ca.\ $10^{-8}$ to
+$10^{-3}$. %, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
+Higher rates of extinction correspond to both large $\sigma$ if $\rho$ is
+small, and large $\rho$ if $\sigma$ is small.  In the former case, increased
+extinction risk arises because of the decrease in the steady-state consumer
+population density (Figs. \ref{fig:fp}b, \ref{fig:ext}).  In the latter case,
+the increased extinction risk results from higher-amplitude transient cycles
+as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).  This
+interplay creates an `extinction refuge', such that for a constrained range
+of $\sigma$ and $\rho$, extinction probabilities are minimized.
 
 % \noindent \paragraph*{{\bf Extinction refuge.}} 
-We find that the allometrically constrained values of $\sigma$ and $\rho$, each representing different trajectories along the ontogenetic curve (Fig.
-\ref{fig:growth}), fall squarely within the extinction refuge across a range of $M$ (Fig.
-\ref{fig:ext}a,b, white points).
-These values are close enough to the Hopf bifurcation to avoid low steady-state densities, yet distant enough to avoid large-amplitude transient cycles.
-That allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions driving starvation and recovery rates within populations.
-Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that minimize extinction risk.
-This selection could occur via the tuning of body fat percentages, metabolic rates, and/or biomass maintenance efficiencies.
-We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig.
-\ref{fig:ext}b), suggesting that the population dynamics for larger organisms are more sensitive to variability in physiological rates.
+We find that the allometrically constrained values of $\sigma$ and $\rho$,
+each representing different trajectories along the ontogenetic curve
+(Fig. \ref{fig:growth}), fall squarely within the extinction refuge across a
+range of $M$ (Fig. \ref{fig:ext}a,b, white points). These values are close
+enough to the Hopf bifurcation to avoid low steady-state densities, yet
+distant enough to avoid large-amplitude transient cycles.  Allometric values
+of $\sigma$ and $\rho$ fall within this relatively small window, which
+supports the possibility that a selective mechanism has constrained the
+physiological conditions driving starvation and recovery rates within
+populations.  Such a mechanism would select for organism physiology that
+generates appropriate $\sigma$ and $\rho$ values that minimize extinction
+risk.  This selection could occur via the tuning of body fat percentages,
+metabolic rates, and/or biomass maintenance efficiencies.  We also find that
+as body size increases, the size of the low extinction-risk parameter space
+shrinks (Fig.~\ref{fig:ext}b), suggesting that the population dynamics for
+larger organisms are more sensitive to variability in physiological rates.
 This finding is in accordance with, and may serve as contributing support for, observations of increased extinction risk among larger mammals \cite{Liow:2008jx}.\\
@@
 % \noindent {\bf Energetic barriers to body size}
-\noindent \paragraph*{{\bf Predicting Damuth's Law and body size limits.}} The NSM correctly predicts that smaller species have larger
-steady-state population densities (Fig.~\ref{fig:mass}).  Similar
-predictions have been made for carnivore populations using alternative
-consumer-resource models \citep{DeLong:2012kw}.  Moreover, we show that the
-NSM provides independent theoretical support for Damuth's Law \citep{Damuth:1987kr,allen2002,enquist1998,pedersen2017shallow}. 
-Damuth's law reveals that species abundances, $N^{*}$, follow $N^*\propto M^{-0.78}$. Figure \ref{fig:mass} shows
-that both $F^{*}$ and $H^{*}$ scale as $M^{-\eta}$ over a wide range of
-organismal sizes and that $F^{*}+H^{*}$ closely matches the best fit to Damuth's data. This result is remarkable because it
+\noindent \paragraph*{{\bf Damuth's Law and body size limits.}} The NSM
+correctly predicts that smaller species have larger steady-state population
+densities (Fig.~\ref{fig:mass}).  Similar predictions have been made for
+carnivore populations using alternative consumer-resource models
+\citep{DeLong:2012kw}.  Moreover, we show that the NSM provides independent
+theoretical support for Damuth's Law
+\citep{Damuth:1987kr,allen2002,enquist1998,pedersen2017shallow}.  Damuth's
+law shows that species abundances, $N^{*}$, follow $N^*=0.01
+M^{-0.78}$ (g m$^{-2}$). Figure \ref{fig:mass} shows that both $F^{*}$ and $H^{*}$ scale
+as $M^{-\eta}$, with $\eta\approx 3/4$,  over a wide range of organismal sizes and that $F^{*}+H^{*}$
+closely matches the best fit to Damuth's data.  Remarkably, this result
 illustrates that the steady state values of the NSM combined with the derived
-timescales naturally give rise to Damuth's law.
-While the previous metabolic studies supporting Damuth's law provided arguments for the value of the exponent \citep{allen2002}, these studies are only able to infer the intercept from the data (see SI for a discussion of the energy equivalence hypothesis related to these metabolic arguments).
-Our model predicts not only the exponent but also the intercept by explicitly including the resource dynamics and the parameters adjusting growth and consumption.
-It should be noted that density relationships of individual clades follow a more shallow scaling relationship than predicted by Damuth's law \cite{pedersen2017shallow}.
-In the context of our model, this suggests that future work may be able to anticipate these shifts by accounting for differences in the physiological parameters associated with each clade.
+timescales naturally give rise to Damuth's law. While the previous metabolic
+studies supporting Damuth's law provided arguments for the value of the
+exponent~\citep{allen2002}, these studies are only able to infer the
+normalization constant ($0.01$ g$^{1.78}$ m$^{-2}$ in the above equation) from the data (see SI for a discussion of the energy
+equivalence hypothesis related to these metabolic arguments). Our model
+predicts not only the exponent but also the normalization constant by
+explicitly including the resource dynamics and the parameters that determine
+growth and consumption. It should be noted that density relationships of
+individual clades follow a more shallow scaling relationship than predicted
+by Damuth's law~\cite{pedersen2017shallow}.  In the context of our model,
+this finding suggests that future work may be able to anticipate these shifts
+by accounting for differences in the physiological parameters associated with
+each clade.
 
@@
 %, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
-With respect to predicted steady state densities, the total metabolic rate of $F$ and $H$ becomes infinite at a finite mass, and occurs at the same scale where the steady state resources vanish (Fig.~\ref{fig:mass}).
-This asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and $\epsilon_{\lambda}$ equal zero, causing the timescales to become infinite and the rates $\mu$ and $\lambda$ to equal zero.
-A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass of a male African elephant.
-In contrast, the $\mu=0$ asymptote occurs first when $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to $(F^*,H^*,R^*)=(0,0,0)$. 
-This point predicts a more realistic upper bound on mammalian body size and occurs at $M_{\rm max}=6.54\times10^7$ (g).
-Moreover, $M_{\rm max}$, which is entirely determined by the population-level consequences of energetic constraints, is within an order of magnitude of the maximum body size observed in the North American mammalian fossil record~\citep{Alroy:1998p1594}, as well as the mass predicted from an evolutionary model of body size evolution~\citep{Clauset:2009fh}.
-It should be noted that the asymptotic behavior and predicted upper bound depend only on the scaling of body composition and are independent of the resource parameters.
-We also note that the prediction of an asymptotic limit on mammalian size parallels work on microbial life where an upper and lower bound on bacterial size, and an upper bound on single cell eukaryotic size, is predicted from similar growth and energetic scaling relationships~\citep{Kempes:2012hy,Kempes:2016}.
-It has also been shown that models that incorporate the allometry of hunting and resting combined with foraging time predicts a maximum carnivore size between $7\times10^{5}$ and $1.1\times10^{6}$ (g) \cite{Carbone:1999ju,carbone2007costs}.
-Similarly, the maximum body size within a particular lineage has been shown to scale with the metabolic normalization constant \cite{Okie:2013ju}.
-%and depend on a critical death found to be constant from data
-This complementary approach is based on the balance between growth and mortality, and suggests that future connections between the scaling of fat and muscle mass should systematically be connected with $B_{0}$ when comparing lineages.
-% reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
+With respect to predicted steady state densities, the total metabolic rate of
+$F$ and $H$ becomes infinite at a finite mass, and occurs at the same scale
+where the steady state resources vanish (Fig.~\ref{fig:mass}). This
+asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and
+$\epsilon_{\lambda}$ (see Fig.~\ref{fig:growth}) equal zero, causing the
+timescales (Eqn. \ref{t1}) to become infinite and the rates $\mu$ and $\lambda$ to equal
+zero.
+% A theoretical upper bound on mammalian body size is given by
+% $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic
+% reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass
+% of a male African elephant.
+The $\mu=0$ asymptote occurs first when
+$f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to
+$(F^*,H^*,R^*)=(0,0,0)$.  This point predicts an upper bound on mammalian
+body size at $M_{\rm max}=6.54\times10^7$ (g).  Moreover, $M_{\rm max}$,
+which is entirely determined by the population-level consequences of
+energetic constraints, is within an order of magnitude of the maximum body
+size observed in the North American mammalian fossil
+record~\citep{Alroy:1998p1594}, as well as the mass predicted from an
+evolutionary model of body size evolution~\citep{Clauset:2009fh}.  We
+emphasize that the asymptotic behavior and predicted upper bound depend only
+on the scaling of body composition and are independent of the resource
+parameters.  The prediction of an asymptotic limit on mammalian size
+parallels work on microbial life where an upper and lower bound on bacterial
+size, and an upper bound on single cell eukaryotic size, is predicted from
+similar growth and energetic scaling
+relationships~\citep{Kempes:2012hy,Kempes:2016}.  It has also been shown that
+models that incorporate the allometry of hunting and resting combined with
+foraging time predicts a maximum carnivore size between $7\times10^{5}$ and
+$1.1\times10^{6}$ (g) \cite{Carbone:1999ju,carbone2007costs}.  Similarly, the
+maximum body size within a particular lineage has been shown to scale with
+the metabolic normalization constant
+\cite{Okie:2013ju}. %and depend on a critical death found to be constant from data
+This complementary approach is based on the balance between growth and
+mortality, and suggests that future connections between the scaling of fat
+and muscle mass should systematically be connected with $B_{0}$ when
+comparing
+lineages. % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
 
@@
 
-\noindent \paragraph*{{\bf A mechanism for Cope's rule}} Metabolite transport constraints are widely thought to place strict boundaries on biological scaling~\citep{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to specific predictions on the minimum possible body size for organisms~\citep{West:2002ud}.
-Above this bound, a number of energetic and evolutionary mechanisms have been explored to assess the costs and benefits associated with larger body masses, particularly for mammals.
-One important such example is the \emph{fasting endurance hypothesis}, which contends that larger body size, with consequent lower metabolic rates and increased ability to maintain more endogenous energetic reserves, may buffer organisms against environmental fluctuations in resource availability~\citep{Millar:1990p923}.
-Over evolutionary time, terrestrial mammalian lineages show a significant trend towards larger body size known as Cope's rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}, and it is thought that within-lineage drivers generate selection towards an optimal upper bound of roughly $10^7$ (g)~\citep{Alroy:1998p1594}, a value that is likely limited by higher extinction risk for large taxa over longer timescales~\citep{Clauset:2009fh}.
-These trends are thought to be driven by a combination of climate change and niche availability~\citep{Saarinen:2014br}; however the underpinning energetic costs and benefits of larger body sizes, and how they influence dynamics over ecological timescales, have not been explored.
+\noindent \paragraph*{{\bf A mechanism for Cope's rule}} Metabolite transport
+constraints are widely thought to place strict boundaries on biological
+scaling~\citep{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to
+specific predictions on the minimum possible body size for
+organisms~\citep{West:2002ud}.  Above this bound, a number of energetic and
+evolutionary mechanisms have been explored to assess the costs and benefits
+associated with larger body masses, particularly for mammals.  One important
+such example is the \emph{fasting endurance hypothesis}, which contends that
+larger body size, with consequent lower metabolic rates and increased ability
+to maintain more endogenous energetic reserves, may buffer organisms against
+environmental fluctuations in resource availability~\citep{Millar:1990p923}.
+Over evolutionary time, terrestrial mammalian lineages show a significant
+trend towards larger body size---Cope's
+rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}.
+It is thought that within-lineage drivers generate selection towards an
+optimal upper bound of roughly $10^7$ (g)~\citep{Alroy:1998p1594}, a value
+that is likely limited by higher extinction risk for large taxa over longer
+timescales~\citep{Clauset:2009fh}.  These trends are thought to be driven by
+a combination of climate change and niche
+availability~\citep{Saarinen:2014br}; however the underpinning energetic
+costs and benefits of larger body sizes, and how they influence dynamics over
+ecological timescales, have not been explored.
 % We argue that the NSM provides a suitable framework to explore these issues.
@@
 % \paragraph*{{\bf Body Size Competition}} 
-The NSM predicts that the steady state resource density $R^{*}$ decreases with increasing body size of the consumer population (Fig.
-\ref{fig:mass}, inset), and classic resource competition theory predicts that the species surviving on the lowest resource abundance will outcompete others \citep{tilman1981,dutkiewicz2009,barton2010}.
-Thus, the combined NSM steady-state dynamics and allometric timescales predict that larger mammals have an intrinsic competitive advantage given a common resource, but these absolute limits do not offer a mechanism by which larger body sizes are selected for or against.
+The NSM predicts that the steady state resource density $R^{*}$ decreases
+with increasing body size of the consumer population (Fig.~\ref{fig:mass},
+inset), and classic resource competition theory predicts that the species
+surviving on the lowest resource abundance will outcompete others
+\citep{tilman1981,dutkiewicz2009,barton2010}. Thus, the combined NSM
+steady-state dynamics and allometric timescales (see Eq.~\eqref{t1}) predict
+that larger mammals have an intrinsic competitive advantage given a common
+resource.  
 % We will now show that the NSM indeed provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time known as Cope's Rule.
@@
 \includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
-\caption{\small{Competitive outcomes for a resident species with body mass $M$ vs. a closely related competing species with modified body mass $M^\prime=M(1+\chi)$.  
-The blue region denotes proportions of modified mass $\chi$ resulting in exclusion of the resident species. 
-The red region denotes values of $\chi$ that result in a mass that is below the starvation threshold and are thus infeasible. 
-Arrows point to the predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$, which may serve as an evolutionary attractor for body mass.
-The black wedge points to the largest body mass known for terrestrial mammals (\emph{Deinotherium} spp.) at $1.74\times10^7$ (g)~\citep{Smith:2010p3442}.}\label{fig:invasion}}
+\caption{\small{Competitive outcomes for a resident species with body mass
+    $M$ vs. a closely related competing species with modified body mass
+    $M^\prime=M(1+\chi)$.  The blue region denotes proportions of modified
+    mass $\chi$ resulting in exclusion of the resident species.  The red
+    region denotes values of $\chi$ that result in a mass that is below the
+    starvation threshold and are thus infeasible.  Arrows point to the
+    predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$,
+    which may serve as an evolutionary attractor for body mass.  The black
+    wedge points to the largest body mass known for terrestrial mammals
+    (\emph{Deinotherium} spp.) at $1.74\times10^7$
+    (g)~\citep{Smith:2010p3442}.}\label{fig:invasion}}
 \end{figure}
@@
 
-We directly assess competitive outcome between two closely related species: a resident species of mass $M$, and a competing species (denoted by $^\prime$) where individuals have a different proportion of body fat such that $M^\prime=M(1+\chi)$.
-If $\chi < 0$, individuals within the competing population have fewer metabolic reserves, and if $\chi>0$, individuals have greater metabolic reserves than the resident species.
-For the allowable values of $\chi$ the adjusted mass should exceed the minimal amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given Eq.
-\ref{t1}, implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
-Together these conditions imply that $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$ and the lower bound is mass-dependent.
-The modified mass of the competitor leads to altered rates of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the maintenance of both starving $\delta(M^\prime)$ and full consumers $\beta(M^\prime)$ (see the SI for detailed derivations of competitor rates).
+However, the above resource relationships do not offer a mechanism for how
+body size is selected.  We directly assess competitive outcome between two
+closely related species: a resident species of mass $M$, and a competing
+species (denoted by $^\prime$) where individuals have a different proportion
+of body fat such that $M^\prime=M(1+\chi)$.  For $\chi < 0$, the competing
+individuals have fewer metabolic reserves than the resident species and vice
+versa for $\chi>0$.  For the allowable values of $\chi$ (see SI), the mass of
+the competitor $M'$ should exceed the minimal amount of body fat,
+$1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be
+positive, which, given Eq.~\ref{t1}, implies that
+$1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.  These
+conditions imply that $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$
+where the upper bound approximately equals $0.05$ and the lower bound is
+mass-dependent.  The modified mass of the competitor leads to altered rates
+of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the
+maintenance of both starving $\delta(M^\prime)$ and full consumers
+$\beta(M^\prime)$ (see the SI for derivations of competitor rates).
 Importantly, $\epsilon_\sigma$, which determines the point along the growth
 curve that defines the body composition of starved foragers, is assumed to
-remain unchanged for the competing population.
+remain unchanged for the competing population (see SI).
 %There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
 
-To assess the susceptibility of the resident species to competitive exclusion, we determine which consumer pushes the steady-state resource density $R^*$ to lower values for a given value of $\chi$, with the expectation that a population capable of surviving on lower resource densities has a competitive advantage \citep{tilman1981}.
-We find that for $M\leq 1.748\times10^7$ (g), having additional body fat ($\chi > 0$) results in a lower steady state resource density ($R^{\prime *}<R^*$), such that the competitor has an intrinsic advantage over the resident species (Fig.
-\ref{fig:invasion}).
-However, for $M> 1.748\times10^7$ (g), leaner individuals ($\chi < 0$) have lower resource steady state densities, switching the advantage.
+To assess the susceptibility of the resident species to competitive
+exclusion, we determine which consumer pushes the steady-state resource
+density $R^*$ to lower values for a given value of $\chi$, with the
+expectation that a population capable of surviving on lower resource
+densities has a competitive advantage \citep{tilman1981}.  We find that for
+$M\leq 1.748\times10^7$ (g), having additional body fat ($\chi > 0$) results
+in a lower steady state resource density ($R^{\prime *}<R^*$), such that the
+competitor has an intrinsic advantage over the resident species
+(Fig. \ref{fig:invasion}).  However, for $M> 1.748\times10^7$ (g), leaner
+individuals ($\chi < 0$) have lower resource steady state densities.
 
@@
 
-The observed switch in susceptibility as a function of $\chi$ at $M_{\rm opt}= 1.748\times10^7$ (g) thus serves as an attractor, such that the NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if $M>M_{\rm opt}$.
-This value is close to but smaller than the asymptotic upper bound for terrestrial mammal body size predicted by the NSM, however it is remarkably close to independent estimates of the largest land mammals, the early Oligocene \emph{Indricotherium} at ca.
-$1.5\times10^7$ (g) and the late Miocene \emph{Deinotherium} at ca.
-$1.74\times10^7$ (g) ~\citep{Smith:2010p3442}.
-Additionally, our calculation of $M_{\rm opt}$ as a function of mass-dependent physiological rates is similar to theoretical estimates of maximum body size \citep{Clauset:2009fh}, and provides independent theoretical support for the observation of a `maximum body size attractor' explored by Alroy~\citep{Alroy:1998p1594}.
+The observed switch in susceptibility as a function of $\chi$ at
+$M_{\rm opt}= 1.748\times10^7$ (g) thus serves as an attractor, such that the
+NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if
+$M>M_{\rm opt}$.  This value is close to but smaller than the asymptotic
+upper bound for terrestrial mammal body size predicted by the NSM.  This
+optimum it is remarkably close to independent estimates of the largest land
+mammals, the early Oligocene \emph{Indricotherium} at $\approx$
+$1.5\times10^7$ (g) and the late Miocene \emph{Deinotherium} at $\approx$
+$1.74\times10^7$ (g) ~\citep{Smith:2010p3442}.  Additionally, our calculation
+of $M_{\rm opt}$ as a function of mass-dependent physiological rates is
+similar to theoretical estimates of maximum body size \citep{Clauset:2009fh},
+and provides independent theoretical support for the observation of a
+`maximum body size attractor' explored by Alroy~\citep{Alroy:1998p1594}.
 % {\color{red} We note that the model of \cite{Brown:1993p708} predicts
 
-An optimal size for mammals at intermediate body mass was predicted by Brown et al. based on reproductive maximization and the transition between hungry and full individuals \cite{Brown:1993p708}. 
+An optimal size for mammals at intermediate body mass was predicted by Brown et al.\ based on reproductive maximization and the transition between hungry and full individuals \cite{Brown:1993p708}. 
 By coupling the NSM to resource dynamics as well as introducing an explicit treatment of storage, we show that species with larger body masses have an inherent competitive advantage for size classes up to $M_{\rm opt}= 1.748\times10^7$.
@@
 %Closing
-The energetics associated with somatic maintenance, growth, and reproduction are important elements that influence the dynamics of all populations~\citep{Stearns:1989ip}.
-The NSM is a general model that incorporates the dynamics of starvation and recovery that are expected to occur in resource-limited environments.
-By incorporating allometric relations between the rates in the NSM, we found: (i) allometrically-determined rates of starvation and recovery appear to minimize extinction risk, (ii) the dynamic consequences of these rates may introduce additional drivers and hard boundaries on the evolution of maximum body size, and (iii) a selective mechanism for the evolution of larger body sizes known as Cope's rule.
-We suggest that the NSM offers a means by which the dynamic consequences of energetic constraints can be assessed using macroscale interactions between and among species.
-Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.
+The energetics associated with somatic maintenance, growth, and reproduction
+are important elements that influence the dynamics of all
+populations~\citep{Stearns:1989ip}.  The NSM incorporates the dynamics of
+starvation and recovery that are expected to occur in resource-limited
+environments.  We found that incorporating allometrically-determined rates
+into the NSM predicts that: (i) extinction risk is minimized, (ii) the
+derived steady-states quantitatively reproduce Damuth's law, and (iii) the
+selective mechanism for the evolution of larger body sizes agrees with Cope's
+rule.  The NSM offers a means by which the dynamic consequences of energetic
+constraints can be assessed using macroscale interactions between and among
+species.
+
+%Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.
 
@@
 \end{align}
-where $A=(\lambda  \xi  \rho +\mu  \sigma )$ and $B=(\beta  \mu  \xi +\delta  \lambda  \xi -\lambda  \mu )$.
-The stability of this fixed point is determined by the Jacobian matrix $\bf J$, where each matrix element $J_{ij}=\partial{\dot X_i}/\partial{X_j}$ when evaluated at the internal fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.
-The parameters in Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue of $\mathbf{J}$ is negative, so that the system is stable with respect to small perturbations from the fixed point.
-Because this fixed point is unique, it is the global attractor for all population trajectories for any initial condition where the resource and consumer densities are both nonzero.
+where $A=(\lambda \xi \rho +\mu \sigma )$ and
+$B=(\beta \mu \xi +\delta \lambda \xi -\lambda \mu )$. The stability of this
+fixed point is determined by the Jacobian matrix $\bf J$, with
+$J_{ij}=\partial{\dot X_i}/\partial{X_j}$, when evaluated at the internal
+fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.  The parameters in
+Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue
+of $\mathbf{J}$ is negative, so that the system is stable with respect to
+small perturbations from the fixed point.  Because this fixed point is
+unique, it is the global attractor for all population trajectories for any
+initial condition where the resource and consumer densities are both nonzero.
 
-{\bf Metabolic scaling relationships}
-The scaling relation between an organism's metabolic
-rate $B$ and its body mass $M$ at reproductive maturity is known to scale as
-$B = B_0 M^\eta$~\citep{West:2002it}, where the scaling exponent $\eta$ is
-typically close to $2/3$ or $3/4$ for metazoans (e.g., ref. \citep{Brown:2004wq}),
-and has taxonomic shifts for unicellular species between $\eta\approx 1$ in
-eukaryotes and $\eta\approx 1.76$ in bacteria
-\citep{DeLong:2010dy,Kempes:2012hy}.
+{\bf Metabolic scaling relationships} The scaling relation between an
+organism's metabolic rate $B$ and its body mass $M$ at reproductive maturity
+is known to scale as $B = B_0 M^\eta$, where the scaling exponent $\eta$ is
+typically close to $2/3$ or $3/4$ for metazoans (e.g.,
+Ref.~\citep{West:2002it,Brown:2004wq}), and has taxonomic shifts for
+unicellular species between $\eta\approx 1$ in eukaryotes and
+$\eta\approx 1.76$ in bacteria \citep{DeLong:2010dy,Kempes:2012hy}.
 %Justin adding an explanation for beta
@@
 \citep{moses2008rmo} for a detailed presentation of these timescales). The
-time that it takes to reach a particular mass $\epsilon M$ is given by the
-timescale
+time that an organism takes to reach a particular mass $\epsilon M$ is given
+by the timescale
 \begin{equation}
@@
 \end{equation}
-where we will define values of $\epsilon$ to describe a variety of timescales, and related rates, within
-our model.
-For example, the rate of reproduction is given by the timescale to go from the birth mass to the adult mass.
-The time to reproduce is given by Equation \ref{t1} as $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
+where we define values of $\epsilon$ below to describe a variety of
+timescales, along with the rates related to $\tau$.  For example, the rate of
+reproduction is given by the timescale to go from the birth mass to the adult
+mass. The time to reproduce is given by Equation \ref{t1} as
+$t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
 is the fraction of the asymptotic mass where an organism is reproductively
 mature and should be close to one (typically
-$\epsilon_{\lambda}\approx0.95\;$ \citep{West:2001bv}).
-Our reproductive rate, $\lambda$, is a specific rate, or the number of offspring produced per time per individual, defined as $\dot{F} = \lambda F$.
-In isolation this functional form gives population growth following $F\left(t\right) = F_{0}e^{\lambda t}$ which can be related to the reproductive timescale by assuming that when $t=t_{\lambda}$ it is also the case that $F=\nu F_{0}$, where $\nu-1$ is the number of offspring produced per reproductive cycle.
-Following this relationship the growth rate is given by $\lambda=\ln\left(\nu\right)/t_{\lambda}$, which is the standard relationship (e.g.
-\cite{savage2004effects}) and will scale as $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of $\epsilon_{\lambda}$
+$\epsilon_{\lambda}\approx0.95\;$ \citep{West:2001bv}). Our reproductive
+rate, $\lambda$, is a specific rate, or the number of offspring produced per
+time per individual, defined as $\dot{F} = \lambda F$. In isolation this
+functional form gives the population growth
+$F\left(t\right) = F_{0}e^{\lambda t}$ which can be related to the
+reproductive timescale by assuming that when $t=t_{\lambda}$ it is also the
+case that $F=\nu F_{0}$, where $\nu-1$ is the number of offspring produced
+per reproductive cycle. Following this relationship the growth rate is given
+by $\lambda=\ln\left(\nu\right)/t_{\lambda}$, which is the standard
+relationship (e.g.,~\cite{savage2004effects}) and will scales as
+$\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of
+$\epsilon_{\lambda}$
 \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
@@
 and recovery by the timescales required to reach, or return from, specific
-fractions of the replete-state mass (see the SI, Table I, for parameterizations).  We define
-$m_{\sigma}=\epsilon_{\sigma} M$, where $\epsilon_{\sigma}<1$ is the fraction of
-replete-state mass where reproduction ceases. This fraction will deviate from a constant
-if tissue composition systematically scales with adult mass.  For example,
-making use of the observation that body fat in mammals scales with overall
-body size according to $M_{\rm fat}=f_{0}M^{\gamma}$ and assuming that once
-this mass is fully digested the organism starves, this would imply that
-$\epsilon_{\sigma}=1-f_{0}M^{\gamma}/M$. It follows that the
-recovery timescale, $t_{\rho}$, is the time to go from
-$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$ (Fig.
-\ref{fig:growth}).
-Using Eqs.~\eqref{m1} and \eqref{t1} this timescale is given by simply considering an adjusted starting mass of $m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
+fractions of the replete-state mass (see the SI, Table I, for
+parameterizations).  We define $m_{\sigma}=\epsilon_{\sigma} M$, where
+$\epsilon_{\sigma}<1$ is the fraction of replete-state mass where
+reproduction ceases. This fraction will deviate from a constant if tissue
+composition systematically scales with adult mass.  For example, making use
+of the observation that body fat in mammals scales with overall body size
+according to $M_{\rm fat}=f_{0}M^{\gamma}$ and assuming that once this mass
+is fully digested the organism starves, this would imply that
+$\epsilon_{\sigma}=1-f_{0}M^{\gamma}/M$. It follows that the recovery
+timescale, $t_{\rho}$, is the time to go from mass
+$m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$
+(Fig. \ref{fig:growth}). Using Eqs.~\eqref{m1} and \eqref{t1} this timescale
+is given by simply considering the growth curve starting from a mass of
+$m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
 \begin{eqnarray}
@@
 required for an organism to go from a mature adult that reproduces at rate
-$\lambda$, to a
-reduced-mass hungry state where reproduction is impossible.
-For starving individuals we assume that an organism must meet its maintenance requirements by using the digestion of existing mass as the sole energy source.
-This assumption implies the following simple metabolic balance
+$\lambda$, to a reduced-mass hungry state where reproduction is impossible.
+For starving individuals we assume that an organism must meet its maintenance
+requirements by using the digestion of existing mass as the sole energy
+source.  This assumption implies the metabolic balance
 $\dot{m}E_{m}^{\prime}=-B_{m}m$ or $\dot{m}=-a^{\prime}m/M^{1-\eta}$
@@
 above energy balance prescribes the mass trajectory of a non-consuming
-organism:$m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}$.
+organism: $m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}$.
 % \begin{eqnarray}

### 2017-12-13 NatureComm_REVISION/YKR_NatureComm-Revision-5.tex
editorial revisions

--- 2017-12-13 NatureComm_REVISION/Arxiv/YKR_Arxiv3.tex
@@
 
-%\bibliographystyle{ieeetr}
-
-\usepackage{bibunits}
-
-%\linenumbers
-%\setlength\linenumbersep{3pt}
+\bibliographystyle{naturemag}
+
+% \usepackage{bibunits}
+
+\linenumbers
+\setlength\linenumbersep{3pt}
 
@@
 \author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
-  of California, Merced, Merced, CA 95340, USA}
+  of California, Merced, Merced, CA 95340, USA}\affiliation{The Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
 
 \author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
-  Park Road, Santa Fe, NM 87501, USA}
+  Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
 
 \author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
-  Road, Santa Fe, NM 87501, USA}
-
-\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
+  Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
+
+\title{The dynamics of starvation and recovery: predicting extinction risk, Damuth's Law, and Cope's Rule}%: Eco-evolutionary feedbacks}
 
@@
 \centering
-  \includegraphics[width=0.45\textwidth]{fig_FixedPoint2-eps-converted-to.pdf}
+\includegraphics[width=0.45\textwidth]{fig_FixedPoint2.eps}
 \caption{\small{ The transcritical (TC; dashed line) and Hopf bifurcation (solid line) as a
@@
 relationships for resource consumption and growth (see the SI for a full derivation and the dimensional form).
-  Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
+Notice that the total consumer density $F+H$ follows $\dot{F}+\dot{H}=\lambda F-\mu H$.
 This resembles the equation of motion for the predator density in the LV model~\citep{murray2011mathematical}, except that the resource density does not appear in the growth term.
@@
 \centering
-  \includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram-eps-converted-to.pdf}
+\includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram.eps}
 \caption{\small{ The growth trajectory over absolute time of an individual organism as a function of body mass.  
@@
 \centering
-  \includegraphics[width=0.4\textwidth]{fig_ExtinctionAllometricComb4-eps-converted-to.pdf}
+\includegraphics[width=0.4\textwidth]{fig_ExtinctionAllometricComb4.eps}
 \caption{\small{ Probability of extinction for a consumer with ({\bf a}) $M=10^2$g and ({\bf b}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
@@
 \centering
-  \includegraphics[width=0.4\textwidth]{fig_FPAllometric2-eps-converted-to.pdf}
+\includegraphics[width=0.4\textwidth]{fig_FPAllometric2.eps}
 \caption{\small{Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
@@
 \centering
-  \includegraphics[width=0.4\textwidth]{fig_Invasion-eps-converted-to.pdf}
+\includegraphics[width=0.4\textwidth]{fig_Invasion.eps}
 \caption{\small{Competitive outcomes for a resident species with body mass
@@
 \def\bibfont{\footnotesize}
-% \putbib[aa_starving3]
+
+% \bibliography{aa_starving3}
 
@@
 
-\end{bibunit}
+% \end{bibunit}
 
@@
 %
-\clearpage
-
-\begin{bibunit}[unsrt]
-
-% \addto\captionsenglish{\renewcommand{\figurename}{Figure S\!}}
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}%
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}%
-
-\section*{Supporting Information for ``The dynamics of starvation and recovery''}
-
-{\bf Mechanisms of Starvation and Recovery}
-To understand the dynamics of starvation, recovery, reproduction, and resource competition, our framework partitions consumers into two classes: (a) a full class that is able to reproduce and, (b) a hungry class that experiences mortality at a given rate and is unable to reproduce.
-For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate, as dictated by body size, with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+% SUPPLEMENTARY MATERIAL
+%%%%%%%%%%%%%%%%%%%%%%%%
+% 
+% 
+% \clearpage
+% 
+% \begin{bibunit}[unsrt]
+% {\bf Supporting Information for ``The dynamics of starvation and recovery''}\\
+% 
+% {\bf Mechanisms of Starvation and Recovery}
+% Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
+For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
 \citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
@@
 Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).
-
-We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
-\begin{align}
-\begin{split}
-\dot{F_{d}} &= \lambda_{\text{max}} F_{d} + \rho_{\text{max}}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{\text{max}}R_{d} H_{d}/k - \mu H_{d},  \\
-\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
-& \left[\left(\frac{\rho_{\text{max}}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)F_{d}\right].
-\label{bigdynamics}
-\end{split}
-\end{align}
-where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with the subscript $_{d}$ before introducing the non-dimensional form that is presented in the main text).
+% 
+% We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
+% \begin{align}
+% \begin{split}
+% \dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+% \dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
+% \dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
+% & \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
+% \end{split}
+% \end{align}
+% where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
 In the above equations $Y$ represents the yield coefficient (e.g., Refs.
-\citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism$^{-1}$).
+\citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
 If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
 This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
-In these units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
-%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
-
-%\begin{align}
-%\begin{split}
-%\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-%\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
-%\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
-%&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
-%\end{split}
-%\end{align}
-%
-%In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
+In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
+% %(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
+% 
+% 
+% 
+% 
+% %\begin{align}
+% %\begin{split}
+% %\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+% %\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
+% %\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
+% %&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
+% %\end{split}
+% %\end{align}
+% %
+% %In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
 Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
 The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
-%A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
-%
-%Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
+% %A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
+% %
+% %Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
 As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
 Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
-
-%where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
-
-We can formally non-dimensionalize this system by the rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
-%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
-\begin{align}
-\begin{split}
-&\dot{F} = \frac{1}{s}\left[\lambda_{\text{max}} F + \rho_{\text{max}}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
-&\dot{H} = \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{\text{max}}\frac{R}{qk} H - \mu H\right],  \\
-& \dot{R} = \\
-&\frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{\text{max}}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)F\right]\right].
-\end{split}
-\end{align}
-If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
-\begin{align}
-\begin{split}
-\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
-\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
-\label{reduceddynamics}
-\end{split}
-\end{align}
-where we have dropped the subscripts on $\lambda_{\text{max}}$ and $\rho_{\text{max}}$ for simplicity, and $\xi\equiv C/k$, $\delta\equiv Y_{H}kP_{H}/C$, and $\beta\equiv Y_{H}k\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)/C$.
+% 
+% 
+% 
+% 
+% 
+% %where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
+% 
+% We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
+% %(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
+% \begin{align}
+% \begin{split}
+% \dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
+% \dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
+% \dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
+% \end{split}
+% \end{align}
+% If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
+% \begin{align}
+% \begin{split}
+% \dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
+% \dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
+% \dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
+% \end{split}
+% \end{align}
+% where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$.
 The above equations represent the system of equations presented in the main text.
-\\
-
-{\bf Parameter Values and Estimates}
-All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
+% \\
+% 
+% {\bf Parameter Values and Estimates}
+% All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
 Below we outline previous measurements and simple estimates of the parameters.
-
-Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs.
+% 
+% Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs.
 \citep{West:2001bv,moses2008rmo} and the supplement for Ref.
 \citep{hou}).
-We make this assumption in the current paper, although alternate exponents, which are known to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
+We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
 The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.
@@
 For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \citep{hou}.
-
-For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ ~\citep{Dunbrack:1993ec}, $0.02M^{1.19}$ ~\citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ ~\citep{Lindstedt:2002td}.
+% 
+% For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \citep{Dunbrack:1993ec}, $0.02M^{1.19}$ \citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ \citep{Lindstedt:2002td}.
 We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range.
-Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ ~\citep{Lindstedt:2002td}.
-
-%We also connect the resource growth rate to the total metabolic rate of an organism.
+Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \citep{Lindstedt:2002td}.
+% 
+% %We also connect the resource growth rate to the total metabolic rate of an organism.
 That is, we are interested in the relative rates of resource recovery and consumption by the total population.
@@
 Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
-\begin{equation}
-B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-Similarly, the maintenance resource consumption rate for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
-\begin{equation}
-B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.
-
-Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities.
+% \begin{equation}
+% B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+% \end{equation}
+% Similarly, the maintenance for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
+% \begin{equation}
+% B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
+% \end{equation}
+% Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.
+% 
+% Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities.
 From the data in Ref.
 \citep{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
-It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with $B_{tot}$ for each type.
-As such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{damuth1987interspecific}.
+It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{damuth1987interspecific}.
 We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above.
 However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.
-\\
-%For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
-
-%More complicated models of fecundity (which, for example, account for the average length of adulthood and the number of individuals produced over this span) could be employed.
+% \\
+% %For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
+% 
+% %More complicated models of fecundity (which, for example, account for the average length of adulthood and the number of individuals produced over this span) could be employed.
 However, the scaling of population growth rate has been studied in detail before and follows a relationship of $$ which matches the theory well for $\phi=.95$ and $=$.
-
-%In our calculations we include $20\%$ variation around this value which could account for differences in efficiency during
-
- \begin{table}[h]
-\caption{Parameter values for mammals}
-\label{param}
-    \begin{center}
-    \footnotesize
-     \begin{tabular}{p{3.8cm} c p{2.2cm} p{1.4cm}}
-     \hline
-
-     Definition & Parameter & Value & References  \\
-     \hline
-   Asymptotic adult mass & $M$ & (g) &  \\
-   Initial mass of an organism & $m_{0}$ & (g) &  \\
-   Metabolic rate scaling exponent & $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
-   Metabolic Normalization Constant & $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
-   Initial mass scaling exponent & $\upsilon$ & $0.92$ &  \citep{blueweiss1978relationships,peters1986ecological} \\
-   Initial mass scaling normalization constant & $n_{0}$ & $0.097$ (g$^{1-\upsilon}$) & \citep{blueweiss1978relationships,peters1986ecological}  \\   
-   Fat mass scaling exponent & $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
-   Fat scaling normalization constant & $f_{0}$ & $0.02$ (g$^{1-\eta}$) & \citep{Lindstedt:1985hm}\\
-   Muscle mass scaling exponent & $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
-   Muscle scaling normalization constantv& $u_{0}$ & $0.38$ (g$^{1-\zeta}$)  & \citep{Lindstedt:2002td} \\
-   Energy to synthesis a unit of mass & $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
-   Energy to synthesis a unit of mass during recovery & $E_{m}^{\prime}$ & $7000$ (J gram$^{-1}$) & \citep{stryer,hou} \\
-   Specific resource growth rate & $\alpha$ & $9.45\times10^{-9}$ (s$^{-1}$) & see text  \\
-   Fraction of asymptotic mass representing full state & $\epsilon_{\lambda}$ & $0.95$ & \citep{West:2001bv}  \\
-   Fraction of asymptotic mass representing starving state & $\epsilon_{\sigma}$ & $1-f_{0}M^{\gamma-1}$ & see text  \\
-   Fraction of asymptotic mass representing death & $\epsilon_{\mu}$ & $1-\frac{f_{0}M^{\gamma}+u_{0}M^{\zeta}}{M}$ & see text \\
-   Carrying capacity (maximum density) of resources & $C$ & (g m$^{-2}$) & \\
-   Half Saturation Constant & $k$ & (g m$^{-2}$) &   \\
-   Normalized carrying capacity & $\xi$ & $C/k\approx2$ &   \\
-   Reproductive fecundity & $\nu$ & $2$ & \citep{}  \\
-
-%   $a$ & $1.78\times10^{-6}$  \quad \quad \\
-%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
-
-   \hline
-    \end{tabular}
-    \end{center}
-   \end{table}
-
-{\bf Rate equations for invaders with modified body mass}
-We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
-Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
-For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (see main text), implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
-Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
-
-Although the starved state of invading organisms remains unchanged, the rate of starvation from the modified full state to the starved state, the rate of recovery from the starved state to the modified full state, and the maintenance rates of both, will be different, such that $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$, $\delta^\prime = \delta(M^\prime)$.
-Rates of starvation and recovery for the invading population are easily derived by adjusting the starting or ending state before and after starvation and recovery, leading to the following timescales:
-
-\begin{align}
-t_{\sigma^\prime} &= -\frac{M^{1-\eta}}{a^{\prime}}\ln \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
-t_{\rho^\prime} &= \ln \left(\frac{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}{1-( \epsilon_\lambda(\chi +1))^{1/4}}\right)\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}.
-\end{align}
-
-The maintenance rates for the invading population require more careful consideration.
-First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of body fat reserves.
-Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
-\begin{equation}
-B^{\prime}_{\lambda}=\int_{0}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-where
-\begin{equation}
-t_{\lambda^\prime} = \frac{M^{1-\eta} }{a(1-\eta)}\ln \left(\frac{1-(m_0/M)^{1-\eta}}{1-(\epsilon_\lambda (1+\chi))^{1-\eta}} \right).
-\end{equation}
-Note that we do not use this timescale to determine the reproductive rate of the invading consumer---which is assumed to remain the same as the resident population---but only to calulate the lifetime energy use.
-Similarly, the maintenance for hungry individuals $P^\prime_{H}=B_{0}(\epsilon_{\sigma}(1+\chi)M)^{3/4}/(\epsilon_{\sigma}(1+\chi)M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y^\prime_{H}=(1+\chi)ME_{d}/B^{\prime}_{\rho}$ where
-\begin{equation}
-B^{\prime}_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-Finally, we can calculate the maintenance of the invaders as
-
-\begin{align}
-  \delta^\prime &= P^\prime_{H}Y^\prime_{H}/\xi \\ \nonumber
-  \beta^\prime &= \left(\frac{\lambda_{\rm max}}{Y^\prime_{F}}+P^\prime_{F} \right)Y^\prime_{H}/\xi.
-\end{align}
-
-To determine whether or not the invader or resident population has an advantage, we compute $R^*(M)$ and $R^*(M^\prime=M(1+\chi))$ for values of $\chi \in (-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$, and the invading population is assumed to have an advantage over the resident population if $R^*(M^\prime)<R^*(M)$.
-
-{\bf Sensitivity to additional death terms}
-
-It should be noted that our set of dynamics (Equations \ref{bigdynamics} and \ref{reduceddynamics}) could include a constant death term of the form $-d_{F}F$ and $-d_{H}H$ to represent death not directly linked to starvation.
-Adding terms of this form to our model would simply adjust the effective value of $\lambda$ and $\mu$, and we could rewrite Equation \ref{reduceddynamics} with $\lambda^{\prime}=\lambda-d$ and $\mu^{\prime}=\mu-d$.
-These substitutions would not alter the functional form of our model nor the steady-states and qualitative results, however the quantitative values could shift based on the size of $d$ relative to $\lambda$ and $\mu$.
-
-Survivorship has a well-known functional form which changes systematically with size (e.g. \cite{calder1984}). Typically survivorship is defined using the Gompertz curve 
-\begin{equation}
-F=F_{0}e^{\left(c_{0}/c_{1}\right)\left(1-e^{c_{1}t}\right)}
-\label{gompertz}
-\end{equation}
-where the parameters have the following allometric dependencies on adult mass $c_{0}=a_{0}M^{b_{0}}$ and $c_{1}=a_{1}M^{b_{1}}$, with $a_{0}=1.88\times10^{-8}$ (s g$^{-b_{0}}$), $b_{0}=-0.56$, $a_{1}=1.45\times10^{-7}$ (s g$^{-b_{1}}$), and $b_{1}=-0.27$ (see \cite{calder1984} for a review).
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{mortality-rate-comparison-eps-converted-to.pdf}
-\caption{\small{The rates of reproduction $\lambda$ (blue), starvation-based mortality $\mu$ (red), and survivorship-based death $\bar{d}$ (black) as a function of adult mass.}\label{fig:ratescomp}}
-\end{figure}
-
-We are interested in the specific death rate of the form $\dot{F}=-dF$, and using the derivative of Equation \ref{gompertz} we find that $d=c_{0}e^{c_{1}t}$.
-Our model considers the average rates over a population and lifecycle and the average death rate is given by
-\begin{eqnarray}
-\bar{d}&=&\frac{1}{t_{\text{exp}}}\int_{0}^{t_{\text{exp}}}c_{0}e^{c_{1}t} dt \\
-&=&\frac{c_{0}\left(e^{c_{1} t_{\text{exp}}}-1\right)}{c_{1}t_{\text{exp}}}
-\end{eqnarray}
-where $t_{\text{exp}}$ is the expected lifespan following the allometry of $t_{\text{exp}}=a_{2}M^{b_{2}}$ with $a_{2}=4.04\times10^{6}$ (s g$^{-b_{2}}$) and $b_{2}=0.30$ ~\cite{damuth1982analysis,calder1984}.
-Given the allometries above we have that
-\begin{equation}
-\bar{d}=\frac{a_{0} \left(e^{a_{1}a_{2}M^{b_{1}+b_{2}}}-1\right) M^{b_{0}-b_{1}-b_{2}}}{a_{1} a_{2}}
-\end{equation}
-which scales roughly like $M^{b_{0}}$ because $b_{1}$ and $b_{2}$ are close in value but opposite in sign.
-In Figure S\ref{fig:ratescomp} we compare the value of $\bar{d}$ to the reproductive, $\lambda$, and starvation-based mortality, $\mu$, rates.
-The values of $\bar{d}$ are orders of magnitude smaller than these rates for all mammalian masses, and thus, adding this non-starvation based death rate to our model does not shift our results within numerical confidence.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.4\textwidth]{fig_FPenergyequiv-eps-converted-to.pdf}
-\caption{\small{ Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass ($F^*$ is shown in green and $H^*$ in orange).
-The data are from Damuth \citep{Damuth:1987kr} and have been converted to
-  total population metabolism using the allometric relationships for
-  metabolic rate (e.g. Refs.~\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:equivalence}}
-\end{figure}
-
-{\bf NSM and the energy equivalence hypothesis}
-
-The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}.
-As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}.
-However, the actual equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations.
-Figure S\ref{fig:equivalence} recasts the results of the NSM in terms of this hypothesis and shows that $F^{*}B$ is nearly constant over the same range of mammalian sizes up to the asymptotic behavior for the largest terrestrial mammals.
-
-{\bf Application of NSM limits to aquatic mammals}
-A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass of a male African elephant.
-We note this particular limit as it may have future relevance to considerations of the ultimate constraints on aquatic mammals.
-
+% 
+% %In our calculations we include $20\%$ variation around this value which could account for differences in efficiency during
+% 
+%  \begin{table}[h]
+% \caption{Parameter values for mammals}
+% \label{param}
+%     \begin{center}
+%     \small
+%      \begin{tabular}{ p{1.2cm} p{3.2cm} l p{2.2cm}|}
+%      \hline
+%      Parameter & Value & References  \\
+%      \hline
+%    $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
+%    $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
+%    $E_{m}^{\prime}$ & $7000$  & \citep{stryer,hou} \\
+%    $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
+% %   $a$ & $1.78\times10^{-6}$  \quad \quad \\
+% %   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
+%    $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
+%    $f_{0}$ & $0.02$ & \citep{Lindstedt:1985hm}\\
+%    $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
+%    $u_{0}$ & $0.38$  & \citep{Lindstedt:2002td} \\
+% 
+%    \hline
+%     \end{tabular}
+%     \end{center}
+%    \end{table}
+% 
+% 
+% {\bf Rate equations for invaders with modified body mass}
+% We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
+% Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
+% For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (see main text), implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
+% Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
+% 
+% Although the starved state of invading organisms remains unchanged, the rate of starvation from the modified full state to the starved state, the rate of recovery from the starved state to the modified full state, and the maintenance rates of both, will be different, such that $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$, $\delta^\prime = \delta(M^\prime)$.
+% Rates of starvation and recovery for the invading population are easily derived by adjusting the starting or ending state before and after starvation and recovery, leading to the following timescales:
+% 
+% \begin{align}
+% t_{\sigma^\prime} &= -\frac{M^{1-\eta}}{a^{\prime}}\ln \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
+% t_{\rho^\prime} &= \ln \left(\frac{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}{1-( \epsilon_\lambda(\chi +1))^{1/4}}\right)\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}.
+% \end{align}
+% 
+% 
+% The maintenance rates for the invading population require more careful consideration.
+% First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of body fat reserves.
+% Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
+% \begin{equation}
+% B^{\prime}_{\lambda}=\int_{0}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
+% \end{equation}
+% where
+% \begin{equation}
+% t_{\lambda^\prime} = \frac{M^{1-\eta} }{a(1-\eta)}\ln \left(\frac{1-(m_0/M)^{1-\eta}}{1-(\epsilon_\lambda (1+\chi))^{1-\eta}} \right).
+% \end{equation}
+% Note that we do not use this timescale to determine the reproductive rate of the invading consumer---which is assumed to remain the same as the resident population---but only to calulate the lifetime energy use.
+% Similarly, the maintenance for hungry individuals $P^\prime_{H}=B_{0}(\epsilon_{\sigma}(1+\chi)M)^{3/4}/(\epsilon_{\sigma}(1+\chi)M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y^\prime_{H}=(1+\chi)ME_{d}/B^{\prime}_{\rho}$ where
+% \begin{equation}
+% B^{\prime}_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
+% \end{equation}
+% Finally, we can calculate the maintenance of the invaders as
+% 
+% \begin{align}
+%   \delta^\prime &= P^\prime_{H}Y^\prime_{H}/\xi \\ \nonumber
+%   \beta^\prime &= \left(\frac{\lambda_{\rm max}}{Y^\prime_{F}}+P^\prime_{F} \right)Y^\prime_{H}/\xi.
+% \end{align}
+% 
+% To determine whether or not the invader or resident population has an advantage, we compute $R^*(M)$ and $R^*(M^\prime=M(1+\chi))$ for values of $\chi \in (-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$, and the invading population is assumed to have an advantage over the resident population if $R^*(M^\prime)<R^*(M)$.
+% 
 % \putbib[aa_starving_supplement]
-
-\def\bibfont{\footnotesize}
-
-% \bibliography{aa_starving_supplement}
-\begin{thebibliography}{10}
-\expandafter\ifx\csname url\endcsname\relax
-  \def\url#1{\texttt{#1}}\fi
-\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
-\providecommand{\bibinfo}[2]{#2}
-\providecommand{\eprint}[2][]{\url{#2}}
-
-\bibitem{Kempes:2012hy}
-\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Dutkiewicz, S.} \&
-  \bibinfo{author}{Follows, M.~J.}
-\newblock \bibinfo{title}{{Growth, metabolic partitioning, and the size of
-  microorganisms.}}
-\newblock \emph{\bibinfo{journal}{PNAS}} \textbf{\bibinfo{volume}{109}},
-  \bibinfo{pages}{495--500} (\bibinfo{year}{2012}).
-
-\bibitem{kempes2014morphological}
-\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Okegbe, C.},
-  \bibinfo{author}{Mears-Clarke, Z.}, \bibinfo{author}{Follows, M.~J.} \&
-  \bibinfo{author}{Dietrich, L.~E.}
-\newblock \bibinfo{title}{Morphological optimization for access to dual
-  oxidants in biofilms}.
-\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
-  Sciences}} \textbf{\bibinfo{volume}{111}}, \bibinfo{pages}{208--213}
-  (\bibinfo{year}{2014}).
-
-\bibitem{West:2001bv}
-\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{Enquist, B.~J.}
-\newblock \bibinfo{title}{{A general model for ontogenetic growth}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{413}},
-  \bibinfo{pages}{628--631} (\bibinfo{year}{2001}).
-
-\bibitem{moses2008rmo}
-\bibinfo{author}{Moses, M.~E.} \emph{et~al.}
-\newblock \bibinfo{title}{{Revisiting a model of ontogenetic growth: Estimating
-  model parameters from theory and data}}.
-\newblock
-  \emph{\bibinfo{journal}{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}}
-  \textbf{\bibinfo{volume}{171}}, \bibinfo{pages}{632--645}
-  (\bibinfo{year}{2008}).
-
-\bibitem{hou}
-\bibinfo{author}{Hou, C.} \emph{et~al.}
-\newblock \bibinfo{title}{{Energy uptake and allocation during ontogeny}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{322}},
-  \bibinfo{pages}{736--739} (\bibinfo{year}{2008}).
-
-\bibitem{pirt}
-\bibinfo{author}{Pirt, S.}
-\newblock \bibinfo{title}{The maintenance energy of bacteria in growing
-  cultures}.
-\newblock \emph{\bibinfo{journal}{Proceedings of the Royal Society of London B:
-  Biological Sciences}} \textbf{\bibinfo{volume}{163}},
-  \bibinfo{pages}{224--231} (\bibinfo{year}{1965}).
-
-\bibitem{Heijnen}
-\bibinfo{author}{Heijnen, J.} \& \bibinfo{author}{Roels, J.}
-\newblock \bibinfo{title}{A macroscopic model describing yield and maintenance
-  relationships in aerobic fermentation processes}.
-\newblock \emph{\bibinfo{journal}{Biotechnology and Bioengineering}}
-  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{739--763}
-  (\bibinfo{year}{1981}).
-
-\bibitem{peters1986ecological}
-\bibinfo{author}{Peters, R.~H.}
-\newblock \emph{\bibinfo{title}{The Ecological Implications of Body Size}},
-  vol.~\bibinfo{volume}{2} (\bibinfo{publisher}{Cambridge University Press},
-  \bibinfo{address}{Cambridge}, \bibinfo{year}{1986}).
-
-\bibitem{blueweiss1978relationships}
-\bibinfo{author}{Blueweiss, L.} \emph{et~al.}
-\newblock \bibinfo{title}{Relationships between body size and some life history
-  parameters}.
-\newblock \emph{\bibinfo{journal}{Oecologia}} \textbf{\bibinfo{volume}{37}},
-  \bibinfo{pages}{257--272} (\bibinfo{year}{1978}).
-
-\bibitem{stryer}
-\bibinfo{author}{Stryer, L.}
-\newblock \emph{\bibinfo{title}{{Biochemistry, Fourth Edition}}}
-  (\bibinfo{publisher}{W.H. Freeman and Company}, \bibinfo{address}{New York},
-  \bibinfo{year}{1995}).
-
-\bibitem{Dunbrack:1993ec}
-\bibinfo{author}{Dunbrack, R.~L.} \& \bibinfo{author}{Ramsay, M.~A.}
-\newblock \bibinfo{title}{{The Allometry of Mammalian Adaptations to Seasonal
-  Environments: A Critique of the Fasting Endurance Hypothesis}}.
-\newblock \emph{\bibinfo{journal}{Oikos}} \textbf{\bibinfo{volume}{66}},
-  \bibinfo{pages}{336--342} (\bibinfo{year}{1993}).
-
-\bibitem{Lindstedt:1985hm}
-\bibinfo{author}{Lindstedt, S.~L.} \& \bibinfo{author}{Boyce, M.~S.}
-\newblock \bibinfo{title}{{Seasonality, Fasting Endurance, and Body Size in
-  Mammals}}.
-\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{125}},
-  \bibinfo{pages}{873--878} (\bibinfo{year}{1985}).
-
-\bibitem{Lindstedt:2002td}
-\bibinfo{author}{Lindstedt, S.~L.} \& \bibinfo{author}{Schaeffer, P.~J.}
-\newblock \bibinfo{title}{{Use of allometry in predicting anatomical and
-  physiological parameters of mammals.}}
-\newblock \emph{\bibinfo{journal}{Lab. Anim.}} \textbf{\bibinfo{volume}{36}},
-  \bibinfo{pages}{1--19} (\bibinfo{year}{2002}).
-
-\bibitem{estermann}
-\bibinfo{author}{Estermann, B.~L.}, \bibinfo{author}{Wettstein, H.-R.},
-  \bibinfo{author}{Sutter, F.} \& \bibinfo{author}{Kreuzer, M.}
-\newblock \bibinfo{title}{Nutrient and energy conversion of grass-fed dairy and
-  suckler beef cattle kept indoors and on high altitude pasture}.
-\newblock \emph{\bibinfo{journal}{Animal Research}}
-  \textbf{\bibinfo{volume}{50}}, \bibinfo{pages}{477--493}
-  (\bibinfo{year}{2001}).
-
-\bibitem{michaletz2014convergence}
-\bibinfo{author}{Michaletz, S.~T.}, \bibinfo{author}{Cheng, D.},
-  \bibinfo{author}{Kerkhoff, A.~J.} \& \bibinfo{author}{Enquist, B.~J.}
-\newblock \bibinfo{title}{Convergence of terrestrial plant production across
-  global climate gradients}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{512}},
-  \bibinfo{pages}{39--43} (\bibinfo{year}{2014}).
-
-\bibitem{damuth1987interspecific}
-\bibinfo{author}{Damuth, J.}
-\newblock \bibinfo{title}{Interspecific allometry of population density in
-  mammals and other animals: the independence of body mass and population
-  energy-use}.
-\newblock \emph{\bibinfo{journal}{Biological Journal of the Linnean Society}}
-  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
-  (\bibinfo{year}{1987}).
-
-\bibitem{calder1984}
-\bibinfo{author}{Calder, W.~A.}
-\newblock \emph{\bibinfo{title}{Size, function, and life history}}
-  (\bibinfo{publisher}{Harvard University Press}, \bibinfo{year}{1984}).
-
-\bibitem{damuth1982analysis}
-\bibinfo{author}{Damuth, J.}
-\newblock \bibinfo{title}{Analysis of the preservation of community structure
-  in assemblages of fossil mammals}.
-\newblock \emph{\bibinfo{journal}{Paleobiology}} \textbf{\bibinfo{volume}{8}},
-  \bibinfo{pages}{434--446} (\bibinfo{year}{1982}).
-
-\bibitem{allen2002}
-\bibinfo{author}{Allen, A.~P.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{Gillooly, J.~F.}
-\newblock \bibinfo{title}{{Global biodiversity, biochemical kinetics, and the
-  energetic-equivalence rule}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{297}},
-  \bibinfo{pages}{1545--1548} (\bibinfo{year}{2002}).
-
-\bibitem{enquist1998}
-\bibinfo{author}{Enquist, B.~J.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{West, G.~B.}
-\newblock \bibinfo{title}{{Allometric scaling of plant energetics and
-  population density}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{395}},
-  \bibinfo{pages}{163--165} (\bibinfo{year}{1998}).
-
-\bibitem{Damuth:1987kr}
-\bibinfo{author}{Damuth, J.}
-\newblock \bibinfo{title}{{Interspecific allometry of population density in
-  mammals and other animals: the independence of body mass and population
-  energy-use}}.
-\newblock \emph{\bibinfo{journal}{Biol. J. Linn. Soc.}}
-  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
-  (\bibinfo{year}{1987}).
-
-\end{thebibliography}
-
-\end{bibunit}
+% 
+% 
+% \end{bibunit}
+% 
+%
 

### 2017-12-13 NatureComm_REVISION/YKR_supp-R2.tex
editorial revisions

--- 2017-12-13 NatureComm_REVISION/Arxiv/YKR_Arxiv3.tex
@@
-\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
+% \RequirePackage{lineno}
+\documentclass[onecolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
 %\usepackage[pdftex]{graphicx}
@@
 
-%\bibliographystyle{ieeetr}
+\bibliographystyle{naturemag}
 
-\usepackage{bibunits}
+% \usepackage{bibunits}
 
-%\linenumbers
-%\setlength\linenumbersep{3pt}
+% \linenumbers
+%  \setlength\linenumbersep{3pt}
+
+\addto\captionsenglish{\renewcommand{\figurename}{Supplementary Figure\!}}
 
@@
 
-\author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
-  of California, Merced, Merced, CA 95340, USA}
+% 
+% \author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
+%   of California, Merced, Merced, CA 95340, USA}\affiliation{The Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
+% 
+% \author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
+%   Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
+% 
+% \author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
+%   Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
+% 
+% \title{ Supporting Information for ``The dynamics of starvation and recovery''}%: Eco-evolutionary feedbacks}
+% 
+% 
+% 
+% \maketitle
+%
 
-\author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
-  Park Road, Santa Fe, NM 87501, USA}
+%%%%%%%%%%%%%%%%%%%%%%%%
+% SUPPLEMENTARY MATERIAL
+%%%%%%%%%%%%%%%%%%%%%%%%
 
-\author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
-  Road, Santa Fe, NM 87501, USA}
+%New additions in red text
 
-\title{The dynamics of starvation and recovery}%: Eco-evolutionary feedbacks}
+% \clearpage
 
-\begin{abstract} %250 words
-The eco-evolutionary dynamics of species are fundamentally linked to the energetic constraints of its constituent individuals.
-Of particular importance is the interplay between reproduction and the dynamics of starvation and recovery.
-To elucidate this interplay, we introduce a nutritional state-structured model that incorporates two classes of consumer: nutritionally replete, reproducing consumers, and undernourished, non-reproducing consumers.
-We obtain strong constraints on starvation and recovery rates by deriving allometric scaling relationships and find that population dynamics are typically driven to a steady state.
-Moreover, these rates fall within a `refuge' in parameter space, where the probability of population extinction is minimized.
-We also show that our model provides a natural framework to predict maximum mammalian body size by determining the relative stability of an otherwise homogeneous population to a competing population with altered percent body fat.
-This framework provides a principled mechanism for a selective driver of Cope's rule.
-\end{abstract}
+\noindent {\bf Supplementary Note 1: Sensitivity to additional death terms}
 
-\maketitle
-
-\begin{bibunit}[unsrt]
-
-  The behavioral ecology of all organisms is influenced by their energetic
-  states, which directly impacts how they invest reserves in uncertain
-  environments.  Such behaviors are generally manifested as tradeoffs between
-  investing in somatic maintenance and growth, or allocating energy towards
-  reproduction~\citep{Martin:1987dl,Kirk:1997cc,Kempes:2012hy}.  The timing of
-  these behaviors responds to selective pressure, as the choice of the
-  investment impacts future
-  fitness~\citep{Mangel:1988uaa,Mangel:2014kz,Yeakel:2013hi}.  The influence of
-  resource limitation on an organism's ability to maintain its nutritional
-  stores may lead to repeated delays or shifts in reproduction over the course
-  of an organism's life.
-
-  The balance between (a) somatic growth and maintenance, and (b) reproduction depends on resource availability~\citep{Morris:1987eo}.
-  For example, reindeer invest less in calves born after harsh winters (when the mother's energetic state is depleted) than in calves born after moderate winters~\citep{Tveraa:2003fq}.
-  Many bird species invest differently in broods during periods of resource scarcity~\citep{Daan:1988va,Jacot:2009dv}, sometimes delaying or even foregoing reproduction for a breeding season~\citep{Martin:1987dl,Stearns:1989ip,Barboza:2002in}.
-  Even freshwater and marine zooplankton have been observed to avoid reproduction under nutritional stress~\citep{Threlkeld:1976ih}, and those that do reproduce have lower survival rates~\citep{Kirk:1997cc}.
-  Organisms may also separate maintenance and growth from reproduction over space and time: many salmonids, birds, and some mammals return to migratory breeding grounds to reproduce after one or multiple seasons in resource-rich environments where they accumulate reserves~\citep{Weber:1998jg,Mduma:1999cp,Moore:2014hi}.
-
-  Physiology also plays an important role in regulating reproductive expenditures during periods of resource limitation.
-  Many mammals (47 species in 10 families) exhibit delayed implantation, whereby females postpone fetal development until nutritional reserves can be accumulated~\citep{Mead:1989dt,Sandell:1990kw}.
-  Many other species (including humans) suffer irregular menstrual cycling and higher abortion rates during periods of nutritional stress~\citep{Bulik:1999eo,Trites:2003cc}.
-  In the extreme case of unicellular organisms, nutrition directly controls growth to a reproductive state \citep{Glazier:2009hq,Kempes:2012hy}.
-The existence of so many independently evolved mechanisms across such a diverse suite of organisms highlights the near-universality of the fundamental tradeoff between somatic and reproductive investment.
-
-  Including individual energetic dynamics~\citep{Kooi2000} in a
-  population-level framework~\citep{Kooi2000,Sousa:2010ez} is
-  challenging~\citep{Diekmann:2010da}.  A common simplifying approach is the
-  classic Lotka-Volterra (LV) model, which assumes that consumer population
-  growth rate depends linearly on resource density~\citep{murdoch:2003}. Here,
-  we introduce an alternative approach---the Nutritional State-structured Model
-  (NSM)---that accounts for resource limitation via explicit starvation. In
-  contrast to the LV model, the NSM incorporates two consumer states: hungry
-  and full, with only the former susceptible to mortality and only the latter
-  possessing sufficient energetic reserves to reproduce.  Additionally, we
-  incorporate allometrically derived constraints on the time scales for
-  reproduction~\citep{Kempes:2012hy}, starvation, and recovery.  Our model
-  makes several basic predictions: (i) the dynamics are typically driven to a
-  refuge far from cyclic behavior and extinction risk, (ii) the steady-state
-  conditions of the NSM accurately predict the measured biomass densities for
-  mammals described by Damuth's law~\citep{Damuth:1987kr,allen2002,enquist1998,Pedersen:2017he},
-    %  energetic equivalence principle, 
-  (iii) there is an allometrically constrained upper-bound for mammalian body size, and
-  (iv) the NSM provides a selective mechanism for the evolution of larger body size, known as Cope's rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}.
-
-  % Though straightforward conceptually, incorporating the energetic dynamics of
-  % individuals~\citep{Kooi2000} into a population-level
-  % framework~\citep{Kooi2000,Sousa:2010ez} presents mathematical
-  % obstacles~\citep{Diekmann:2010da}.  An alternative approach involves modeling
-  % the macroscale relations that guide somatic versus reproductive investment in
-  % a consumer-resource system.  {\color{red}For example, macroscale
-  %   Lotka-Volterra models assume that the growth rate of the consumer
-  %   population depends on resource density, thus incorporating the requirement
-  %   of resource availability for reproduction~\citep{murdoch:2003}.
-
-  % In this work, we adopt an alternative approach to account for resource
-  % limitation through the consequences of starvation.  Namely, in contrast to
-  % the Lotka-Volterra model, we incorporate two states: hungry individuals and
-  % individuals with sufficient energetic reserves to reproduce.}  Such a
-  % constraint leads to reproductive time lags due to some members of the
-  % population going hungry and then recovering.  Additionally, we incorporate
-  % the idea that reproduction is strongly constrained allometrically
-  % \citep{Kempes:2012hy}, and is not generally linearly related to resource
-  % density.
-  % As we shall show, these constraints influence the ensuing population dynamics in dramatic ways.\\
-
-  \noindent \paragraph*{{\bf Nutritional state-structured model (NSM).}}
-  We begin by defining the nutritional state-structured population model, where the consumer population is partitioned into two states: (a) an
-  energetically replete (full) state $F$, where the consumer reproduces at a
-  constant rate $\lambda$ and does not die from starvation, and (b) an
-  energetically deficient (hungry) state $H$, where the consumer does not
-  reproduce but dies by starvation at rate $\mu$. The dynamics of the
-  underlying resource $R$ are governed by logistic growth with an intrinsic
-  growth rate $\alpha$ and a carrying capacity $C$. The rate at which consumers
-  transition between states and consume resources is dependent on their number,
-  the abundance of resources, the efficiency of converting resources into
-  metabolism, and how that metabolism is partitioned between maintenance and
-  growth purposes.  We provide a physiologically and energetically mechanistic
-  model for each of these dynamics and constants (see the Supplementary Information
-  (SI)), and show that the system produces a simple non-dimensional form which
-  we describe below.
-
-  \begin{figure}
-  \centering
-  \includegraphics[width=0.45\textwidth]{fig_FixedPoint2-eps-converted-to.pdf}
-  \caption{\small{ The transcritical (TC; dashed line) and Hopf bifurcation (solid line) as a
-    function of the starvation rate $\sigma$ and recovery rate $\rho$ for a 100g consumer.  These
-    bifurcation conditions separate parameter space into unphysical (left of the TC), cyclic,
-    and steady state dynamic regimes.  The colors show the steady state densities for the energetically replete consumers $F^*$.  
-    % Steady state densities for the energetically deficient consumers $H^*$ (not shown)
-    % scale with those for $F^*$.  
-  }\label{fig:fp}}
-  \end{figure}
-
-  Consumers transition from the full state $F$ to the hungry state $H$ at a
-  rate $\sigma$---the starvation rate---and also in proportion to the absence
-  of resources $(1-R)$ (the maximum resource density has been non
-  dimensionalized to 1; see SI).  Conversely, consumers recover from state $H$
-  to state $F$ at rate $\xi \rho$ and in proportion to $R$, where $\xi$
-  represents a ratio between maximal resource consumption and the carrying
-  capacity of the resource. %and $\approx 2$.
-  The resources that are eaten by hungry consumers (at rate $\rho R + \delta$)
-  account for their somatic growth ($\rho R$) and maintenance ($\delta$).  Full
-  consumers eat resources at a constant rate $\beta$ that accounts for maximal
-  maintenance and somatic growth (see the SI for mechanistic derivations of
-  these rates from resource energetics).
-  %This inequality accounts for hungry consumers requiring more resources to replace lost body tissue.
-  The NSM represents an ecologically motivated fundamental extension of the
-  idealized starving random walk model of foraging, which focuses on resource
-  depletion, to include reproduction and resource
-  replenishment~\citep{Benichou:2014wu,Benichou:2016wl,Chupeau:2016jf}, and is
-  a more general formulation than previous models that incorporate
-  starvation~\citep{Persson:1998hz}.
-
-  In the mean-field approximation, in which the consumers and resources are
-  perfectly mixed, their densities are governed by the rate equations
-
-  \begin{align}
-  \label{eq:system}
-  \begin{split}
-  \dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
-  \dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-  \dot{R} &= \alpha \left(1-R\right)R -\left(\rho R+\delta\right)H-\beta F.
-  \end{split}
-  \end{align}
-
-  This system of nondimensional equations follows from a set of first-principle
-  relationships for resource consumption and growth (see the SI for a full derivation and the dimensional form).
-  Notice that the total consumer density $F+H$ evolves according to $\dot{F}+\dot{H}=\lambda F-\mu H$.
-  This resembles the equation of motion for the predator density in the LV model~\citep{murray2011mathematical}, except that the resource density does not appear in the growth term.
-  % As discussed above, the attributes of reproduction and mortality have been explicitly apportioned to the full and hungry consumers, respectively, so that the growth in the total density is decoupled from the resource density.
-  The rate of reproduction is independent of resource density because the full
-  consumer partitions a constant amount of energy towards reproduction, whereas
-  a hungry consumer partitions no energy towards reproduction.  Similarly, the
-  consumer maintenance terms ($\delta H$ and $\beta F$) are also independent of
-  resource density because they represent a minimal energetic requirement for
-  consumers in the $H$ and $F$ state, respectively.
-
-  % It follows that model predictions are robust only when $R$ is of the order of 1, which holds for all cases that we explore.
-
-  \noindent \paragraph*{ {\bf Steady states of the NSM.}} From the single
-  internal fixed point (Eq.~\eqref{eq:ss}, see Methods), an obvious constraint
-  on the NSM is that the reproduction rate $\lambda$ must be less than the
-  starvation rate $\sigma$, so that the consumer and resource densities are
-  positive.
-  %In fact, when the resource density $R=0$, the rate equation for $F$ gives exponential growth of $F$ for $\lambda>\sigma$.
-  The condition $\sigma = \lambda$ represents a transcritical (TC)
-  bifurcation~\citep{Strogatz:2008wo} that demarcates a physical from an
-  unphysical (negative steady-state densities) regime.  The biological
-  implication of the constraint $\lambda<\sigma$ has a simple
-  interpretation---the rate at which a macroscopic organism loses mass due to
-  lack of resources is generally much faster than the rate of reproduction.  As
-  we will discuss below, this inequality is also a natural consequence of
-  allometric constraints~\citep{Kempes:2012hy} for organisms within empirically
-  observed body size ranges. %(Fig.~\ref{fig:gvs})
-
-  In the physical regime of $\lambda<\sigma$, the fixed point \eqref{eq:ss} may either be a stable node or a limit cycle (Fig.~\ref{fig:fp}).
-  In continuous-time systems, a limit cycle arises when a pair of complex conjugate eigenvalues crosses the imaginary axis to attain positive real parts~\citep{GuckHolmes}.
-  This Hopf bifurcation is defined by ${\rm Det}({\bf S}) = 0$, with $\bf S$ the Sylvester matrix, which is composed of the coefficients of the characteristic polynomial of the Jacobian matrix~\citep{Gross:2004p2428}.
-  As the system parameters are tuned to be within the stable regime, but close to the Hopf bifurcation, the amplitude of the transient cycles becomes large.
-  Given that ecological systems are constantly being perturbed~\citep{Hastings:2001jh}, the onset of transient cycles, even though they decay with time in the mean-field description, can increase extinction risk~\citep{Neubert:1997wk,Caswell:2005eo,Neubert:2009td}.
-  %Thus the distance of a system from the Hopf bifurcation provides a measure of its persistence.
-
-  When the starvation rate $\sigma\gg\lambda$, a substantial fraction of the
-  consumers are driven to the hungry non-reproducing state.  Because
-  reproduction is inhibited, there is a low steady-state consumer density and a
-  high steady-state resource density.  However, if $\sigma/\lambda\to 1$ from
-  above, the population is overloaded with energetically-replete (reproducing)
-  individuals, thereby promoting transient oscillations between the consumer
-  and resource densities (Fig.~\ref{fig:fp}).  If the starvation rate is low
-  enough that the Hopf bifurcation is crossed, these oscillations become
-  stable.
-This threshold occurs at higher values of the starvation rate as the recovery rate $\rho$ increases, such that the range of parameter space giving rise to cyclic dynamics also increases with higher recovery rates.\\
-
-  % Whereas the relation between consumer growth rate $\lambda$ and the starvation rate $\sigma$ defines an absolute bound of biological feasibility---the TC bifurcation---$\sigma$ also determines the sensitivity of the consumer population to changes in resource density.
-  % When $\sigma\gg\lambda$, the steady-state population density is small, thereby increasing the risk of stochastic extinction.
-  % On the other hand, as $\sigma$ decreases, the system will ultimately be poised either near the TC or the Hopf bifurcation (Fig.~\ref{fig:fp}).
-  % If the recovery rate $\rho$ is sufficiently small, the TC bifurcation is reached and the resource eventually is eliminated.
-  % If $\rho$ exceeds a threshold value, cyclic dynamics will develop as the Hopf bifurcation is approached.
-
-  \begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{Growth-trajectory-diagram-eps-converted-to.pdf}
-  \caption{\small{ The growth trajectory over absolute time of an individual organism as a function of body mass.  
-  Initial growth follows the black trajectory to an energetically replete reproductive adult mass of $m=\epsilon_\lambda M$ (see Methods). %which we assume is 95\% asymptotic mass $M$.  
-  Starvation follows the red trajectory to $m = \epsilon_\sigma \epsilon_\lambda  M$. 
-  Recovery follows the green curve to the replete adult mass, where this trajectory differs from the original growth because only fat is being regrown which requires a longer time to reach $\epsilon_\lambda M$.
-%different energetics and
-  Alternatively, death from starvation follows the blue trajectory to $m=\epsilon_\mu \epsilon_\lambda  M$.}\label{fig:growth}}
-  \end{figure}
-
-  % \noindent {\bf Role of allometry} \\
-  %[Link Allometry stuff to our model - what does it provide to the story]
-  %[Introduce biological and linked constraints]
-  %[Allometries capture vast amounts of diversity via a single parameter --- body size]
-  %[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
-  \noindent {\bf Results}
-  \noindent \paragraph*{{\bf The allometry of extinction risk.}} While there
-  are no {\it a priori} constraints on the parameters in the NSM, we expect
-  that each species should be restricted to a distinct portion of the parameter
-  space.  We use allometric scaling relations to constrain the covariation of
-  rates in a principled and biologically meaningful manner (see Methods).
-  Allometric scaling relations highlight common constraints and average trends
-  across large ranges in body size and species diversity. Many of these
-  relations can be derived from a small set of assumptions.  In the Methods we
-  describe our framework to determine the covariation of timescales and rates
-  across a range of body sizes for each of the key parameters of our model
-  (cf.\ Ref.~\citep{Yodzis:1992hg}).
-  % We are thereby able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.
-
-  Nearly all of the rates described in the NSM are determined by consumer
-  metabolism, which can be used to describe a variety of organismal features
-  \citep{Brown:2004wq}.  We derive, from first principles, the relationships
-  for the rates of reproduction, starvation, recovery, and mortality as a
-  function of an organism's body size and metabolic rate (see Methods).
-  Because we aim to explore the starvation-recovery dynamics as a function of
-  an organism's body mass $M$, we parameterize these rates in terms of the
-  \emph{percent} gain and loss of the asymptotic (maximum) body mass,
-  $\epsilon M$, where different values of $\epsilon$ define different states of
-  the consumer (Fig.~\ref{fig:growth}; see Methods for derivations of
-  allometrically constrained rate equations).  Although the rate equations
-  \eqref{eq:system} are general and can in principle be used to explore the
-  starvation recovery dynamics for most organisms, here we focus on allometric
-  relationships for terrestrial-bound lower-trophic level endotherms (see the
-  SI for values), specifically herbivorous mammals, which range from a minimum
-  of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of
-  $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene
-  Deinotheriinae).  Investigating other classes of organisms would simply
-  involve altering the metabolic exponents and scalings associated with
-  $\epsilon$. Moreover, we emphasize that our allometric equations (see
-  Methods) describe mean relationships, and do not account for the (sometimes
-  considerable) variance associated with individual species.  We note that
-  including additional allometrically-scaled mortality terms to both $F$ and
-  $H$ does not change the form of our model nor impact our quantitative findings
-  (see SI for the derivation).
-
-  % \noindent {\bf  Stabilizing effects of allometric constraints} \\
-  % As the allometric derivations of the NSM rate laws reveal, starvation and recovery rates are not independent parameters, and the biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp} is constrained via covarying parameters.
-  % Given the parameters of terrestrial endotherms, we find that the starvation rate $\sigma$ and the recovery rate $\rho$ are constrained to lie within a small region of potential values for the known range of body sizes $M$.
-  % Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
-  %Incorporating uncertainty in allometric parameters (20\% variation around the mean; Fig.~\ref{fig:hopf}), we find that, for larger $M$, the distance to the TC and Hopf bifurcation decreases.
-  % These results suggest that small mammals are marginally less prone to population oscillations---both stable limit cycles and transient cycles---than mammals with larger body size.
-However, starvation and recovery rates across all values of $M$ fall squarely in the steady state region at some distance from the Hopf bifurcation.
-  % This result suggests that cyclic population dynamics should be rare, particularly in environments where resources are limiting.\\
-
-  % Previous studies used allometric constraints to explain the periodicity of cyclic populations \citep{CalderIII:1983jd,Peterson:1984hj,Krukonis:1991fk}, suggesting a period $\propto M^{0.25}$.
-  % However this relation seems to hold only for some species \citep{Hendriks:2012fc}, and potential drivers of variation and systematically different behavior range from predator and/or prey lifespans to competitive dynamics~\citep{Kendall:1999iy,Hogstedt:2005cr}.
-  % Statistically significant support for the existence of population cycles among mammals is relatively rare, though predominantly based on time series for small mammals~\citep{Kendall:1998hl}.
-  % However the longer gestational times and increased difficulty in collecting adequate data precludes obtaining similar-quality information for larger organisms.\\
-
-  \begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{fig_ExtinctionAllometricComb4-eps-converted-to.pdf}
-  \caption{\small{ Probability of extinction for a consumer with ({\bf a}) $M=10^2$g and ({\bf b}) $M=10^6$g as a function of the starvation rate $\sigma$ and recovery rate $\rho$, where the initial density is given as $(XF^*,XH^*,R^*)$, where $X$ is a random uniform variable in $[0,2]$.
-Note the change in scale in panel {\bf b}.
-Extinction is defined as the population trajectory falling below $0.2\times$ the allometrically constrained steady state.
-The white points denote the allometrically constrained starvation and recovery rate.}\label{fig:ext}}
-  \end{figure}
-
-  %Extinction to the left and the right
-  % \noindent \paragraph*{{\bf Extinction risk}} 
-  As the allometric derivations of the NSM rate laws reveal (see Methods),
-  starvation and recovery rates are not independent parameters, and the
-  biologically relevant portion of the phase space shown in Fig.~\ref{fig:fp}
-  is constrained via covarying parameters.  Given the parameters of terrestrial
-  endotherms, we find that the starvation rate $\sigma$ and the recovery rate
-  $\rho$ are constrained to lie within a small region of potential values for
-  the known range of body sizes $M$.  Indeed, starvation and recovery rates
-  across all values of $M$ fall squarely in the steady-state region at some
-  distance from the Hopf bifurcation.  This suggests that cyclic population
-  dynamics should be rare, particularly in resource-limited environments.
-  % Moreover, the dynamics for all mammalian body sizes are confined to the steady-state regime of the NSM and that limit-cycle behavior is precluded.
-
-  Higher rates of starvation result in a larger flux of the population to the hungry state.
-  In this state, reproduction is absent, thus increasing the likelihood of extinction.
-From the perspective of population survival, it is the rate of starvation relative to the rate of recovery that determines the long-term dynamics of the various species (Fig.~\ref{fig:fp}).
-  We therefore examine the competing effects of cyclic dynamics vs.\ changes in steady-state density on extinction risk, both as functions of $\sigma$ and $\rho$.
-  To this end, we computed the probability of extinction, where we define extinction as a population trajectory falling below one fifth of the allometrically constrained steady state at any time between $t=10^8$ and $t=10^{10}$.
-  This procedure was repeated for 50 replicates of the continuous-time system shown in Eq.~\ref{eq:system} for organisms with mass ranging from $10^2$ to $10^6$ grams.
-  %the initial condition isdistributed around the steady state (Eq.~\ref{eq:ss}).  Specifically
-  In each replicate the initial densities were chosen to be $(XF^*,XH^*,R^*)$,
-  with $X$ a random variable uniformly distributed in $[0,2]$.  By allowing the
-  rate of starvation to vary, we assessed extinction risk across a range of
-  values for $\sigma$ and $\rho$ between ca.\ $10^{-8}$ to
-  $10^{-3}$. %, thus examining a horizontal cross-section of Fig.~\ref{fig:fp}.
-  Higher rates of extinction correspond to both large $\sigma$ if $\rho$ is
-  small, and large $\rho$ if $\sigma$ is small.  In the former case, increased
-  extinction risk arises because of the decrease in the steady-state consumer
-  population density (Figs. \ref{fig:fp}b, \ref{fig:ext}).  In the latter case,
-  the increased extinction risk results from higher-amplitude transient cycles
-  as the system nears the Hopf bifurcation (Fig.~\ref{fig:ext}).  This
-  interplay creates an `extinction refuge', such that for a constrained range
-  of $\sigma$ and $\rho$, extinction probabilities are minimized.
-
-  % \noindent \paragraph*{{\bf Extinction refuge.}} 
-  We find that the allometrically constrained values of $\sigma$ and $\rho$,
-  each representing different trajectories along the ontogenetic curve
-  (Fig. \ref{fig:growth}), fall squarely within the extinction refuge across a
-  range of $M$ (Fig. \ref{fig:ext}a,b, white points). These values are close
-  enough to the Hopf bifurcation to avoid low steady-state densities, yet
-  distant enough to avoid large-amplitude transient cycles.  Allometric values
-  of $\sigma$ and $\rho$ fall within this relatively small window, which
-  supports the possibility that a selective mechanism has constrained the
-  physiological conditions driving starvation and recovery rates within
-  populations.  Such a mechanism would select for organism physiology that
-  generates appropriate $\sigma$ and $\rho$ values that minimize extinction
-  risk.  This selection could occur via the tuning of body fat percentages,
-  metabolic rates, and/or biomass maintenance efficiencies.  We also find that
-  as body size increases, the size of the low extinction-risk parameter space
-  shrinks (Fig.~\ref{fig:ext}b), suggesting that the population dynamics for
-  larger organisms are more sensitive to variability in physiological rates.
-  This finding is in accordance with, and may serve as contributing support for, observations of increased extinction risk among larger mammals \cite{Liow:2008jx}.\\
-  % Moreover, larger body sizes decrease the steady state resource density, such that fluctuations for larger organisms will be more likely to drive resources to extinction.
-  % OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
-  % To summarize, our finding that allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\
-
-  %Results are insensitive to alpha but possibly sensitive to beta (eff parameter)
-
-  %NOTES 7/8/16
-  %Calder has a hypothesis that period ~ Mass^(1/4).
-  %Can we get "transient period" from Jacobian?
-  %Lots of food web lit on stabilizing effects of body size! Brose, Petchey for instance. Good for discussion.
-
-  % \noindent {\bf Energetic barriers to body size}
-  \noindent \paragraph*{{\bf Damuth's Law and body size limits.}} The NSM
-  correctly predicts that smaller species have larger steady-state population
-  densities (Fig.~\ref{fig:mass}).  Similar predictions have been made for
-  carnivore populations using alternative consumer-resource models
-  \citep{DeLong:2012kw}.  Moreover, we show that the NSM provides independent
-  theoretical support for Damuth's Law
-  \citep{Damuth:1987kr,allen2002,enquist1998,Pedersen:2017he}.  Damuth's
-  law shows that species abundances, $N^{*}$, follow $N^*=0.01
-  M^{-0.78}$ (g m$^{-2}$). Figure \ref{fig:mass} shows that both $F^{*}$ and $H^{*}$ scale
-  as $M^{-\eta}$, with $\eta\approx 3/4$,  over a wide range of organismal sizes and that $F^{*}+H^{*}$
-  closely matches the best fit to Damuth's data.  Remarkably, this result
-  illustrates that the steady state values of the NSM combined with the derived
-  timescales naturally give rise to Damuth's law. While the previous metabolic
-  studies supporting Damuth's law provided arguments for the value of the
-  exponent~\citep{allen2002}, these studies are only able to infer the
-  normalization constant ($0.01$ g$^{1.78}$ m$^{-2}$ in the above equation) from the data (see SI for a discussion of the energy
-  equivalence hypothesis related to these metabolic arguments). Our model
-  predicts not only the exponent but also the normalization constant by
-  explicitly including the resource dynamics and the parameters that determine
-  growth and consumption. It should be noted that density relationships of
-  individual clades follow a more shallow scaling relationship than predicted
-  by Damuth's law~\cite{Pedersen:2017he}.  In the context of our model,
-  this finding suggests that future work may be able to anticipate these shifts
-  by accounting for differences in the physiological parameters associated with
-  each clade.
-
-  \begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{fig_FPAllometric2-eps-converted-to.pdf}
-  \caption{\small{Consumer steady states $F^*$ (green) and $H^*$ (orange) as a function of
-    body mass along with the data from Damuth \citep{Damuth:1987kr}. Inset: Resource steady state $R^*$ as a function of consumer body mass.}\label{fig:mass}}
-  \end{figure}
-
-  % \paragraph*{{\bf Maximum Body Size}}
-  %JUSTIN I REMOVED: Our model shows that energetic equivalence breaks down at large $M$ suggesting that this maximum is a hard limit where deviations outside of this range are energetically suboptimal.
-  %, violating the energy equivalence. % set by deviations from constant efficiency obeyed by other populations.
-  With respect to predicted steady state densities, the total metabolic rate of
-  $F$ and $H$ becomes infinite at a finite mass, and occurs at the same scale
-  where the steady state resources vanish (Fig.~\ref{fig:mass}). This
-  asymptotic behavior is governed by body sizes at which $\epsilon_{\mu}$ and
-  $\epsilon_{\lambda}$ (see Fig.~\ref{fig:growth}) equal zero, causing the
-  timescales (Eqn. \ref{t1}) to become infinite and the rates $\mu$ and $\lambda$ to equal
-  zero.
-  % A theoretical upper bound on mammalian body size is given by
-  % $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic
-  % reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass
-  % of a male African elephant.
-  The $\mu=0$ asymptote occurs first when
-  $f_{0}M^{\gamma-1}+u_{0}M^{\zeta-1}=1$, and corresponds to
-  $(F^*,H^*,R^*)=(0,0,0)$.  This point predicts an upper bound on mammalian
-  body size at $M_{\rm max}=6.54\times10^7$ (g).  Moreover, $M_{\rm max}$,
-  which is entirely determined by the population-level consequences of
-  energetic constraints, is within an order of magnitude of the maximum body
-  size observed in the North American mammalian fossil
-  record~\citep{Alroy:1998p1594}, as well as the mass predicted from an
-  evolutionary model of body size evolution~\citep{Clauset:2009fh}.  We
-  emphasize that the asymptotic behavior and predicted upper bound depend only
-  on the scaling of body composition and are independent of the resource
-  parameters.  The prediction of an asymptotic limit on mammalian size
-  parallels work on microbial life where an upper and lower bound on bacterial
-  size, and an upper bound on single cell eukaryotic size, is predicted from
-  similar growth and energetic scaling
-  relationships~\citep{Kempes:2012hy,Kempes:2016}.  It has also been shown that
-  models that incorporate the allometry of hunting and resting combined with
-  foraging time predicts a maximum carnivore size between $7\times10^{5}$ and
-  $1.1\times10^{6}$ (g) \cite{Carbone:1999ju,Carbone:2007dz}.  Similarly, the
-  maximum body size within a particular lineage has been shown to scale with
-  the metabolic normalization constant
-  \cite{Okie:2013ju}. %and depend on a critical death found to be constant from data
-  This complementary approach is based on the balance between growth and
-  mortality, and suggests that future connections between the scaling of fat
-  and muscle mass should systematically be connected with $B_{0}$ when
-  comparing
-  lineages. % reasons or from dissimilar scaling of cellular composition which is analogous to the body fat composition employed here
-
-  % {\color{red}
-  % 
-  % It should be noted that in our model all death is related to nutritional state, however the addition of a constant death term \cite{} would only adjust the values of $\lambda$ and $\mu$ without changing the form or or qualitative results of our model.
-We explored the sensitivity of our results to shifts in a constant death term (see SI) and find that additional sources of death do not shift our quantitative prediction of Damuth's law except for large sizes (Figure S\ref{}).
-We find that additional sources of death shifts the asymptotic behavior of our model and decreases the maximum possible mammal size.
-Within the range of reasonable death rates $M_{max}$ varies by $VALUE\%$.
-These results suggest that the largest mammals are most sensitive to additional sources of death perhaps putting them at greater extinction risk.
-  % 
-  % }
-
-  %Significant deviations from constant energy use occur at $M \lesssim 1$ at the small end of the mammalian range and $M\approx 6.5*10^7$ at the large end.
-  %Compellingly, this dynamic bound, which is determined by the rate of energetic recovery, is close to the minimum observed mammalian body size of ca.\ 1.3-2.5g, a range that occurs as the recovery rate begins its decline.
-  %In addition to known transport limitations~\citep{West:2002ud}, we suggest that an additional constraint of lower body size stems from the dynamics of starvation.
-  %This result mirrors other efforts \citep{Kempes:2012hy,Kempes:2016} where at a given scale multiple limitations constrain the smallest possibilities for life within a class of organisms.
-
-  %The combined steady-state dynamics and allometric timescales of the NSM predict that larger mammals should outcompete smaller ones, and this suggests that the model may provide a framework with which to explore the energetic mechanisms behind Cope's rule.
-%, which we explore further below.
-
-  \noindent \paragraph*{{\bf A mechanism for Cope's rule}} Metabolite transport
-  constraints are widely thought to place strict boundaries on biological
-  scaling~\citep{Brown:1993p708,West:1997cg,Brown:2004wq} and thereby lead to
-  specific predictions on the minimum possible body size for
-  organisms~\citep{West:2002ud}.  Above this bound, a number of energetic and
-  evolutionary mechanisms have been explored to assess the costs and benefits
-  associated with larger body masses, particularly for mammals.  One important
-  such example is the \emph{fasting endurance hypothesis}, which contends that
-  larger body size, with consequent lower metabolic rates and increased ability
-  to maintain more endogenous energetic reserves, may buffer organisms against
-  environmental fluctuations in resource availability~\citep{Millar:1990p923}.
-  Over evolutionary time, terrestrial mammalian lineages show a significant
-  trend towards larger body size---Cope's
-  rule~\citep{Alroy:1998p1594,Clauset:2009fh,Smith:2010p3442,Saarinen:2014br}.
-  It is thought that within-lineage drivers generate selection towards an
-  optimal upper bound of roughly $10^7$ (g)~\citep{Alroy:1998p1594}, a value
-  that is likely limited by higher extinction risk for large taxa over longer
-  timescales~\citep{Clauset:2009fh}.  These trends are thought to be driven by
-  a combination of climate change and niche
-  availability~\citep{Saarinen:2014br}; however the underpinning energetic
-  costs and benefits of larger body sizes, and how they influence dynamics over
-  ecological timescales, have not been explored.
-  % We argue that the NSM provides a suitable framework to explore these issues.
-
-  % \noindent{\bf A within-lineage driver of Cope's Rule}
-  % \paragraph*{{\bf Body Size Competition}} 
-  The NSM predicts that the steady state resource density $R^{*}$ decreases
-  with increasing body size of the consumer population (Fig.~\ref{fig:mass},
-  inset), and classic resource competition theory predicts that the species
-  surviving on the lowest resource abundance will outcompete others
-  \citep{tilman1981,dutkiewicz2009,barton2010}. Thus, the combined NSM
-  steady-state dynamics and allometric timescales (see Eq.~\eqref{t1}) predict
-  that larger mammals have an intrinsic competitive advantage given a common
-  resource.  
-  % We will now show that the NSM indeed provides a mechanistic understanding of the energetic dynamics that give rise to both observed limitations on mammalian body size, as well as the observed trend towards larger body size over evolutionary time known as Cope's Rule.
-
-  \begin{figure}
-  \centering
-  \includegraphics[width=0.4\textwidth]{fig_Invasion-eps-converted-to.pdf}
-  \caption{\small{Competitive outcomes for a resident species with body mass
-      $M$ vs. a closely related competing species with modified body mass
-      $M^\prime=M(1+\chi)$.  The blue region denotes proportions of modified
-      mass $\chi$ resulting in exclusion of the resident species.  The red
-      region denotes values of $\chi$ that result in a mass that is below the
-      starvation threshold and are thus infeasible.  Arrows point to the
-      predicted optimal mass from our model $M_{\rm opt}=1.748\times 10^7$,
-      which may serve as an evolutionary attractor for body mass.  The black
-      wedge points to the largest body mass known for terrestrial mammals
-      (\emph{Deinotherium} spp.) at $1.74\times10^7$
-      (g)~\citep{Smith:2010p3442}.}\label{fig:invasion}}
-  \end{figure}
-
-  % Next we examine to what extent a more realistic upper bound to body mass may serve as an evolutionary attractor, thus providing a suitable within-lineage mechanism for Cope's rule.
-
-  However, the above resource relationships do not offer a mechanism for how
-  body size is selected.  We directly assess competitive outcome between two
-  closely related species: a resident species of mass $M$, and a competing
-  species (denoted by $^\prime$) where individuals have a different proportion
-  of body fat such that $M^\prime=M(1+\chi)$.  For $\chi < 0$, the competing
-  individuals have fewer metabolic reserves than the resident species and vice
-  versa for $\chi>0$.  For the allowable values of $\chi$ (see SI), the mass of
-  the competitor $M'$ should exceed the minimal amount of body fat,
-  $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be
-  positive, which, given Eq.~\ref{t1}, implies that
-  $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.  These
-  conditions imply that $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$
-  where the upper bound approximately equals $0.05$ and the lower bound is
-  mass-dependent.  The modified mass of the competitor leads to altered rates
-  of starvation $\sigma(M^\prime)$, recovery $\rho(M^\prime)$, and the
-  maintenance of both starving $\delta(M^\prime)$ and full consumers
-  $\beta(M^\prime)$ (see the SI for derivations of competitor rates).
-  Importantly, $\epsilon_\sigma$, which determines the point along the growth
-  curve that defines the body composition of starved foragers, is assumed to
-  remain unchanged for the competing population (see SI).
-  %There is no internal fixed point corresponding to a state where both original residents and invaders coexist (except for the trivial state $\chi=0$).
-
-  To assess the susceptibility of the resident species to competitive
-  exclusion, we determine which consumer pushes the steady-state resource
-  density $R^*$ to lower values for a given value of $\chi$, with the
-  expectation that a population capable of surviving on lower resource
-  densities has a competitive advantage \citep{tilman1981}.  We find that for
-  $M\leq 1.748\times10^7$ (g), having additional body fat ($\chi > 0$) results
-  in a lower steady state resource density ($R^{\prime *}<R^*$), such that the
-  competitor has an intrinsic advantage over the resident species
-  (Fig. \ref{fig:invasion}).  However, for $M> 1.748\times10^7$ (g), leaner
-  individuals ($\chi < 0$) have lower resource steady state densities.
-
-  %, and this is due to the changing
-  %covariance between energetic rates as a function of modified energetic
-  %reserves \sid{I don't understand the phrase after the comma AND STILL DON'T}.
-
-  The observed switch in susceptibility as a function of $\chi$ at
-  $M_{\rm opt}= 1.748\times10^7$ (g) thus serves as an attractor, such that the
-  NSM predicts organismal mass to increase if $M<M_{\rm opt}$ and decrease if
-  $M>M_{\rm opt}$.  This value is close to but smaller than the asymptotic
-  upper bound for terrestrial mammal body size predicted by the NSM, and is remarkably close to independent estimates of the largest land
-  mammals, the early Oligocene \emph{Indricotherium} at $\approx$
-  $1.5\times10^7$ (g) and the late Miocene \emph{Deinotherium} at $\approx$
-  $1.74\times10^7$ (g) ~\citep{Smith:2010p3442}.  Additionally, our calculation
-  of $M_{\rm opt}$ as a function of mass-dependent physiological rates is
-  similar to theoretical estimates of maximum body size \citep{Clauset:2009fh},
-  and provides independent theoretical support for the observation of a
-  `maximum body size attractor' explored by Alroy~\citep{Alroy:1998p1594}.
-  % {\color{red} We note that the model of \cite{Brown:1993p708} predicts
-
-  An optimal size for mammals at intermediate body mass was predicted by Brown et al.\ based on reproductive maximization and the transition between hungry and full individuals \cite{Brown:1993p708}. 
-  By coupling the NSM to resource dynamics as well as introducing an explicit treatment of storage, we show that species with larger body masses have an inherent competitive advantage for size classes up to $M_{\rm opt}= 1.748\times10^7$ based on resource competition.
-Moreover, the mass distributions in Ref.
-\cite{Brown:1993p708} show that intermediate mammal sizes have the greatest species diversity, in contrast to our efforts, which consider total biomass and predict a much larger $M_{\rm opt}$.
-  Compellingly, recent work shows that many communities can be dominated by the biomass of the large \cite{Hempson:2015hka}.
-  % shows that the largest mammals have an advantage in terms resource competition. 
-  %as well as  \citep{Alroy:1998p1594,Clauset:2009fh}.
-  %This value is only XX\% the estimated size of Indricotheriinae, a group that includes the largest known terrestrial mammal.
-  While the state of the environment as well as the competitive landscape will determine whether specific body sizes are selected for or against~\citep{Saarinen:2014br}, we propose that the dynamics of starvation and recovery described in the NSM provide a general selective mechanism for the evolution of larger body size among terrestrial mammals.\\
-
-  \noindent {\bf Discussion} \\
-
-  %and we suggest that the starvation dynamics described here act in concert with these other factors is the only driver of body size evolution.
-
-  %One might be concerned a greater number of large mammals are currently not observed in the modern world given that larger mammals are less susceptible to extinction.
-  %However, recent research suggests that the pleistocene may have been much more populated with a significant diversity of very large mammals \citep{Doughty:2013kd,Doughty:2015hy,Doughty:2015je} which were also much more geographically widespread than today.
-  %These results, combined with our findings, suggest that the modern diversity of mammals may not represent a true steady state the current distribution of nutrients and large seeds may be very different from the past \citep{Doughty:2013kd,Doughty:2015hy,Doughty:2015je}.
-
-  %Closing
-  The energetics associated with somatic maintenance, growth, and reproduction
-  are important elements that influence the dynamics of all
-  populations~\citep{Stearns:1989ip}.  The NSM incorporates the dynamics of
-  starvation and recovery that are expected to occur in resource-limited
-  environments.  We found that incorporating allometrically-determined rates
-  into the NSM predicts that: (i) extinction risk is minimized, (ii) the
-  derived steady-states quantitatively reproduce Damuth's law, and (iii) the
-  selective mechanism for the evolution of larger body sizes agrees with Cope's
-  rule.  The NSM offers a means by which the dynamic consequences of energetic
-  constraints can be assessed using macroscale interactions between and among
-  species.
-
-  %Future efforts will involve exploring the consequences of these dynamics in a spatially explicit framework, thus incorporating elements such as movement costs and spatial heterogeneity, which may elucidate additional tradeoffs associated with the dynamics of starvation and recovery.
-
-  \section*{Methods}
-  \small{
-  {\bf Analytical solution to the NSM}
-  Equation~\eqref{eq:system} has three fixed points: two trivial fixed points at $(F^*,H^*,R^*)=(0,0,0)$ and $(0,0,1)$, and one non-trivial, internal fixed point at
-  \begin{align}
-  \label{eq:ss}
-  \begin{split}
-  F^* &= (\sigma-\lambda)\frac{ \alpha  \lambda  \mu ^2  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
-  H^* &= (\sigma-\lambda)\frac{ \alpha  \lambda ^2 \mu  (\mu +\xi  \rho )}{A (\lambda  \rho  B+\mu  \sigma  (\beta  \mu +\lambda  (\delta +\rho )))}, \\
-  R^* &= (\sigma - \lambda)\frac{\mu  }{A}.
-  \end{split}
-  \end{align}
-  where $A=(\lambda \xi \rho +\mu \sigma )$ and
-  $B=(\beta \mu \xi +\delta \lambda \xi -\lambda \mu )$. The stability of this
-  fixed point is determined by the Jacobian matrix $\bf J$, with
-  $J_{ij}=\partial{\dot X_i}/\partial{X_j}$, when evaluated at the internal
-  fixed point, and $\mathbf{X}$ is the vector $(F,H,R)$.  The parameters in
-  Eq.~\eqref{eq:system} are such that the real part of the largest eigenvalue
-  of $\mathbf{J}$ is negative, so that the system is stable with respect to
-  small perturbations from the fixed point.  Because this fixed point is
-  unique, it is the global attractor for all population trajectories for any
-  initial condition where the resource and consumer densities are both nonzero.
-
-  {\bf Metabolic scaling relationships} The scaling relation between an
-  organism's metabolic rate $B$ and its body mass $M$ at reproductive maturity
-  is known to scale as $B = B_0 M^\eta$, where the scaling exponent $\eta$ is
-  typically close to $2/3$ or $3/4$ for metazoans (e.g.,
-  Ref.~\citep{West:2002it,Brown:2004wq}), and has taxonomic shifts for
-  unicellular species between $\eta\approx 1$ in eukaryotes and
-  $\eta\approx 1.76$ in bacteria \citep{DeLong:2010dy,Kempes:2012hy}.
-  %Justin adding an explanation for beta
-
-  Several efforts have shown how a partitioning of $B$ between growth and
-  maintenance purposes can be used to derive a general equation for both the
-  growth trajectories and growth rates of organisms ranging from bacteria to
-  metazoans
-  \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Savage:2004ed,Kempes:2012hy}. This relation is derived from the simple balance condition 
-  $B_{0}m^{\eta}=E_{m}\dot{m}+B_{m}m\,,$
-  % \begin{eqnarray}
-  % \label{balance}
-  % B_{0}m^{\eta}=E_{m}\frac{dm}{dt}+B_{m}m\,,
-  % \end{eqnarray}
-  \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Savage:2004ed,Kempes:2012hy} where $E_{m}$ is the energy needed to synthesize a unit of mass, $B_{m}$ is
-  the metabolic rate to support an existing unit of mass, and $m$ is the mass
-  of the organism at any point in its development.  This balance has the
-  general solution \citep{bettencourt,Kempes:2012hy}
-  \begin{eqnarray}
-  \label{m1}
-  \left(\frac{m\left(t\right)}{M}\right)^{1-\eta}\!=1\!-\!\left[1\!-\!\left(\frac{m_{0}}{M}\right)^{1\!-\!\eta}\right]e^{-a\left(1\!-\!\eta\right)t/M^{1-\eta}},
-  \end{eqnarray}
-  where, for $\eta<1$, $M=(B_{0}/B_{m})^{1/(1-\eta)}$ is the asymptotic mass,
-  $a=B_{0}/E_{m}$, and $m_0$ is mass at birth, itself varying allometrically
-  (see the SI).  We now use this solution to define the timescale for
-  reproduction and recovery from starvation (Fig.~\ref{fig:growth}; see
-  \citep{moses2008rmo} for a detailed presentation of these timescales). The
-  time that an organism takes to reach a particular mass $\epsilon M$ is given
-  by the timescale
-  \begin{equation}
-  \label{t1}
-  \tau\left(\epsilon\right) = \ln\left[\frac{1-\left(m_{0}/M\right)^{1-\eta}}{1-\epsilon^{1-\eta}}\right]\frac{M^{1-\eta}}{a\left(1-\eta\right)},
-  \end{equation}
-  where we define values of $\epsilon$ below to describe a variety of
-  timescales, along with the rates related to $\tau$.  For example, the rate of
-  reproduction is given by the timescale to go from the birth mass to the adult
-  mass. The time to reproduce is given by Equation \ref{t1} as
-  $t_{\lambda}=\tau\left(\epsilon_{\lambda}\right)$, where $\epsilon_{\lambda}$
-  is the fraction of the asymptotic mass where an organism is reproductively
-  mature and should be close to one (typically
-  $\epsilon_{\lambda}\approx0.95\;$ \citep{West:2001bv}). Our reproductive
-  rate, $\lambda$, is a specific rate, or the number of offspring produced per
-  time per individual, defined as $\dot{F} = \lambda F$. In isolation this
-  functional form gives the population growth
-  $F\left(t\right) = F_{0}e^{\lambda t}$ which can be related to the
-  reproductive timescale by assuming that when $t=t_{\lambda}$ it is also the
-  case that $F=\nu F_{0}$, where $\nu-1$ is the number of offspring produced
-  per reproductive cycle. Following this relationship the growth rate is given
-  by $\lambda=\ln\left(\nu\right)/t_{\lambda}$, which is the standard
-  relationship (e.g.,~\cite{Savage:2004ed}) and will scales as
-  $\lambda\propto M^{\eta-1}$ for $M\gg m_{0}$ for any constant value of
-  $\epsilon_{\lambda}$
-  \citep{West:2001bv,moses2008rmo,gillooly2002esa,hou,Kempes:2012hy}.
-
-  The rate of recovery $\rho = 1/t_\rho$ requires that an organism accrues
-  sufficient tissue to transition from the hungry to the full state.  Since
-  only certain tissues can be digested for energy (for example the brain cannot
-  be degraded to fuel metabolism), we define the rates for starvation, death,
-  and recovery by the timescales required to reach, or return from, specific
-  fractions of the replete-state mass (see the SI, Table I, for
-  parameterizations).  We define $m_{\sigma}=\epsilon_{\sigma} M$, where
-  $\epsilon_{\sigma}<1$ is the fraction of replete-state mass where
-  reproduction ceases. This fraction will deviate from a constant if tissue
-  composition systematically scales with adult mass.  For example, making use
-  of the observation that body fat in mammals scales with overall body size
-  according to $M_{\rm fat}=f_{0}M^{\gamma}$ and assuming that once this mass
-  is fully digested the organism starves, this would imply that
-  $\epsilon_{\sigma}=1-f_{0}M^{\gamma}/M$. It follows that the recovery
-  timescale, $t_{\rho}$, is the time to go from mass
-  $m=\epsilon_{\sigma} \epsilon_{\lambda} M$ to $m=\epsilon_{\lambda}M$
-  (Fig. \ref{fig:growth}). Using Eqs.~\eqref{m1} and \eqref{t1} this timescale
-  is given by simply considering the growth curve starting from a mass of
-  $m_{0}^{\prime}=\epsilon_{\sigma}\epsilon_{\lambda}M$, in which case
-  \begin{eqnarray}
-  \label{rhotimescale}
-  t_{\rho}=\ln\left[\frac{1-\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)^{1-\eta}}{1-\epsilon_\lambda^{1-\eta}}\right]\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}
-  \end{eqnarray}
-  where $a^{\prime}=B_{0}/E_{m}^{\prime}$ accounts for possible deviations in
-  the biosynthetic energetics during recovery (see the SI). It should be noted that more complicated ontogenetic models explicitly handle
-  storage \citep{hou}, whereas this feature is implicitly covered by the body
-  fat scaling in our framework.
-
-  To determine the starvation rate, $\sigma$, we are interested in the time
-  required for an organism to go from a mature adult that reproduces at rate
-  $\lambda$, to a reduced-mass hungry state where reproduction is impossible.
-  For starving individuals we assume that an organism must meet its maintenance
-  requirements by using the digestion of existing mass as the sole energy
-  source.  This assumption implies the metabolic balance
-  $\dot{m}E_{m}^{\prime}=-B_{m}m$ or $\dot{m}=-a^{\prime}m/M^{1-\eta}$
-  % \begin{eqnarray}
-  % \frac{dm}{dt}E_{m}^{\prime}=-B_{m}m
-  % \end{eqnarray}
-  % or
-  % \begin{eqnarray}
-  % \frac{dm}{dt}=-\frac{a^{\prime}}{M^{1-\eta}}m
-  % \end{eqnarray}
-  where $E_{m}^{\prime}$ is the amount of energy stored in a unit of existing
-  body mass, which differs from $E_{m}$, the energy required to
-  synthesis a unit of biomass \citep{hou}. Given the replete mass, $M$, of an organism, the
-  above energy balance prescribes the mass trajectory of a non-consuming
-  organism: $m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}$.
-  % \begin{eqnarray}
-  % \label{mt}
-  % m\left(t\right)=Me^{-a^{\prime}t/M^{1-\eta}}.
-  % \end{eqnarray}
-  The timescale for starvation is
-  given by the time it takes $m(t)$ to reach $\epsilon_{\sigma} M$, which gives
-  \begin{equation}
-  \label{eq:sigma}
-  t_{\sigma}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\sigma}\right).
-  \end{equation}
-  The starvation rate is then $\sigma=1/t_{\sigma}$, which scales with
-  replete-state mass as $1/M^{1-\eta}\ln\left(1-f_{0}M^{\gamma}/M\right)$.  An important
-  feature is that $\sigma$ does not have a simple scaling dependence on
-  $\lambda$, which is important for the dynamics that we
-  later discuss.
-
-  The time to death should follow a similar relation, but defined by a lower
-  fraction of replete-state mass, $m_{\mu}=\epsilon_{\mu} M$ where $\epsilon_\mu < \epsilon_\sigma$.
-  Suppose, for example, that an organism dies once it has digested all fat and
-  muscle tissues, and that muscle tissue scales with body mass according to
-  $M_{\rm musc}=u_{0}M^{\zeta}$.  This gives
-  $\epsilon_{\mu}=1-\left(f_{0}M^{\gamma}+u_{0}M^{\zeta}\right)/M$. Muscle
-  mass has been shown to be roughly proportional to body mass~\citep{Folland:2008ij} in
-  mammals and thus $\epsilon_{\mu}$ is merely $\epsilon_{\sigma}$ minus a constant.
-The time to go from starvation to death is the total time to reach $\epsilon_{\mu}M$ minus the time to starve, or $t_{\mu}=-M^{1-\eta}\ln\left(\epsilon_{\mu}\right)/a^{\prime}-t_{\sigma}$,
-  % \begin{eqnarray}
-  % \label{mutimescale}
-  % t_{\mu}=-\frac{M^{1-\eta}}{a^{\prime}}\ln\left(\epsilon_{\mu}\right)-t_{\sigma},
-  % \end{eqnarray}
-  and $\mu=1/t_{\mu}$.
-  }
-
-\def\bibfont{\footnotesize}
-% \putbib[aa_starving3]
-
-\begin{thebibliography}{10}
-\expandafter\ifx\csname url\endcsname\relax
-  \def\url#1{\texttt{#1}}\fi
-\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
-\providecommand{\bibinfo}[2]{#2}
-\providecommand{\eprint}[2][]{\url{#2}}
-
-\bibitem{Martin:1987dl}
-\bibinfo{author}{Martin, T.~E.}
-\newblock \bibinfo{title}{{Food as a limit on breeding birds: A life-history
-  perspective}}.
-\newblock \emph{\bibinfo{journal}{Annu. Rev. Ecol. Syst.}}
-  \textbf{\bibinfo{volume}{18}}, \bibinfo{pages}{453--487}
-  (\bibinfo{year}{1987}).
-
-\bibitem{Kirk:1997cc}
-\bibinfo{author}{Kirk, K.~L.}
-\newblock \bibinfo{title}{{Life-history responses to variable environments:
-  Starvation and reproduction in planktonic rotifers}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{78}},
-  \bibinfo{pages}{434--441} (\bibinfo{year}{1997}).
-
-\bibitem{Kempes:2012hy}
-\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Dutkiewicz, S.} \&
-  \bibinfo{author}{Follows, M.~J.}
-\newblock \bibinfo{title}{{Growth, metabolic partitioning, and the size of
-  microorganisms}}.
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{109}}, \bibinfo{pages}{495--500}
-  (\bibinfo{year}{2012}).
-
-\bibitem{Mangel:1988uaa}
-\bibinfo{author}{Mangel, M.} \& \bibinfo{author}{Clark, C.~W.}
-\newblock \emph{\bibinfo{title}{{Dynamic Modeling in Behavioral Ecology}}}
-  (\bibinfo{publisher}{Princeton University Press},
-  \bibinfo{address}{Princeton}, \bibinfo{year}{1988}).
-
-\bibitem{Mangel:2014kz}
-\bibinfo{author}{Mangel, M.}
-\newblock \bibinfo{title}{{Stochastic dynamic programming illuminates the link
-  between environment, physiology, and evolution}}.
-\newblock \emph{\bibinfo{journal}{B. Math. Biol.}}
-  \textbf{\bibinfo{volume}{77}}, \bibinfo{pages}{857--877}
-  (\bibinfo{year}{2014}).
-
-\bibitem{Yeakel:2013hi}
-\bibinfo{author}{Yeakel, J.~D.}, \bibinfo{author}{Dominy, N.~J.},
-  \bibinfo{author}{Koch, P.~L.} \& \bibinfo{author}{Mangel, M.}
-\newblock \bibinfo{title}{{Functional morphology, stable isotopes, and human
-  evolution: a model of consilience}}.
-\newblock \emph{\bibinfo{journal}{Evolution}} \textbf{\bibinfo{volume}{68}},
-  \bibinfo{pages}{190--203} (\bibinfo{year}{2014}).
-
-\bibitem{Morris:1987eo}
-\bibinfo{author}{Morris, D.~W.}
-\newblock \bibinfo{title}{{Optimal allocation of parental investment}}.
-\newblock \emph{\bibinfo{journal}{Oikos}} \textbf{\bibinfo{volume}{49}},
-  \bibinfo{pages}{332--339} (\bibinfo{year}{1987}).
-
-\bibitem{Tveraa:2003fq}
-\bibinfo{author}{Tveraa, T.}, \bibinfo{author}{Fauchald, P.},
-  \bibinfo{author}{Henaug, C.} \& \bibinfo{author}{Yoccoz, N.~G.}
-\newblock \bibinfo{title}{{An examination of a compensatory relationship
-  between food limitation and predation in semi-domestic reindeer}}.
-\newblock \emph{\bibinfo{journal}{Oecologia}} \textbf{\bibinfo{volume}{137}},
-  \bibinfo{pages}{370--376} (\bibinfo{year}{2003}).
-
-\bibitem{Daan:1988va}
-\bibinfo{author}{Daan, S.}, \bibinfo{author}{Dijkstra, C.},
-  \bibinfo{author}{Drent, R.} \& \bibinfo{author}{Meijer, T.}
-\newblock \bibinfo{title}{{Food supply and the annual timing of avian
-  reproduction}}.
-\newblock In \bibinfo{editor}{Ouellet, H.} (ed.) \emph{\bibinfo{booktitle}{Acta
-  XIX Congressus Internationalis Ornithologici, Volume I: Proceedings XIX
-  International Ornithological Congress, 1986, Ottawa}},
-  \bibinfo{pages}{392--407} (\bibinfo{publisher}{Proceedings XIX International
-  Ornithological Congress}, \bibinfo{address}{Ottawa}, \bibinfo{year}{1989}).
-
-\bibitem{Jacot:2009dv}
-\bibinfo{author}{Jacot, A.}, \bibinfo{author}{Valcu, M.}, \bibinfo{author}{van
-  Oers, K.} \& \bibinfo{author}{Kempenaers, B.}
-\newblock \bibinfo{title}{{Experimental nest site limitation affects
-  reproductive strategies and parental investment in a hole-nesting
-  passerine}}.
-\newblock \emph{\bibinfo{journal}{Animal Behaviour}}
-  \textbf{\bibinfo{volume}{77}}, \bibinfo{pages}{1075--1083}
-  (\bibinfo{year}{2009}).
-
-\bibitem{Stearns:1989ip}
-\bibinfo{author}{Stearns, S.~C.}
-\newblock \bibinfo{title}{{Trade-offs in life-history evolution}}.
-\newblock \emph{\bibinfo{journal}{Funct. Ecol.}} \textbf{\bibinfo{volume}{3}},
-  \bibinfo{pages}{259} (\bibinfo{year}{1989}).
-
-\bibitem{Barboza:2002in}
-\bibinfo{author}{Barboza, P.} \& \bibinfo{author}{Jorde, D.}
-\newblock \bibinfo{title}{{Intermittent fasting during winter and spring
-  affects body composition and reproduction of a migratory duck}}.
-\newblock \emph{\bibinfo{journal}{J Comp Physiol B}}
-  \textbf{\bibinfo{volume}{172}}, \bibinfo{pages}{419--434}
-  (\bibinfo{year}{2002}).
-
-\bibitem{Threlkeld:1976ih}
-\bibinfo{author}{Threlkeld, S.~T.}
-\newblock \bibinfo{title}{{Starvation and the size structure of zooplankton
-  communities}}.
-\newblock \emph{\bibinfo{journal}{Freshwater Biol.}}
-  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{489--496}
-  (\bibinfo{year}{1976}).
-
-\bibitem{Weber:1998jg}
-\bibinfo{author}{Weber, T.~P.}, \bibinfo{author}{Ens, B.~J.} \&
-  \bibinfo{author}{Houston, A.~I.}
-\newblock \bibinfo{title}{{Optimal avian migration: A dynamic model of fuel
-  stores and site use}}.
-\newblock \emph{\bibinfo{journal}{Evolutionary Ecology}}
-  \textbf{\bibinfo{volume}{12}}, \bibinfo{pages}{377--401}
-  (\bibinfo{year}{1998}).
-
-\bibitem{Mduma:1999cp}
-\bibinfo{author}{Mduma, S. A.~R.}, \bibinfo{author}{Sinclair, A. R.~E.} \&
-  \bibinfo{author}{Hilborn, R.}
-\newblock \bibinfo{title}{{Food regulates the Serengeti wildebeest: a 40-year
-  record}}.
-\newblock \emph{\bibinfo{journal}{J. Anim. Ecol.}}
-  \textbf{\bibinfo{volume}{68}}, \bibinfo{pages}{1101--1122}
-  (\bibinfo{year}{1999}).
-
-\bibitem{Moore:2014hi}
-\bibinfo{author}{Moore, J.~W.}, \bibinfo{author}{Yeakel, J.~D.},
-  \bibinfo{author}{Peard, D.}, \bibinfo{author}{Lough, J.} \&
-  \bibinfo{author}{Beere, M.}
-\newblock \bibinfo{title}{{Life-history diversity and its importance to
-  population stability and persistence of a migratory fish: steelhead in two
-  large North American watersheds}}.
-\newblock \emph{\bibinfo{journal}{J. Anim. Ecol.}}
-  \textbf{\bibinfo{volume}{83}}, \bibinfo{pages}{1035--1046}
-  (\bibinfo{year}{2014}).
-
-\bibitem{Mead:1989dt}
-\bibinfo{author}{Mead, R.~A.}
-\newblock \bibinfo{title}{{The Physiology and Evolution of Delayed Implantation
-  in Carnivores}}.
-\newblock In \bibinfo{editor}{Gittleman, J.~L.} (ed.)
-  \emph{\bibinfo{booktitle}{Carnivore Behavior, Ecology, and Evolution}},
-  \bibinfo{pages}{437--464} (\bibinfo{publisher}{Springer US},
-  \bibinfo{address}{Ithaca}, \bibinfo{year}{1989}).
-
-\bibitem{Sandell:1990kw}
-\bibinfo{author}{Sandell, M.}
-\newblock \bibinfo{title}{{The evolution of seasonal delayed implantation}}.
-\newblock \emph{\bibinfo{journal}{Q Rev Biol}} \textbf{\bibinfo{volume}{65}},
-  \bibinfo{pages}{23--42} (\bibinfo{year}{1990}).
-
-\bibitem{Bulik:1999eo}
-\bibinfo{author}{Bulik, C.~M.} \emph{et~al.}
-\newblock \bibinfo{title}{{Fertility and reproduction in women with anorexia
-  nervosa}}.
-\newblock \emph{\bibinfo{journal}{J. Clin. Psychiat.}}
-  \textbf{\bibinfo{volume}{60}}, \bibinfo{pages}{130--135}
-  (\bibinfo{year}{1999}).
-
-\bibitem{Trites:2003cc}
-\bibinfo{author}{Trites, A.~W.} \& \bibinfo{author}{Donnelly, C.~P.}
-\newblock \bibinfo{title}{{The decline of Steller sea lions Eumetopias jubatus
-  in Alaska: a review of the nutritional stress hypothesis}}.
-\newblock \emph{\bibinfo{journal}{Mammal Rev.}} \textbf{\bibinfo{volume}{33}},
-  \bibinfo{pages}{3--28} (\bibinfo{year}{2003}).
-
-\bibitem{Glazier:2009hq}
-\bibinfo{author}{Glazier, D.~S.}
-\newblock \bibinfo{title}{{Metabolic level and size scaling of rates of
-  respiration and growth in unicellular organisms}}.
-\newblock \emph{\bibinfo{journal}{Funct. Ecol.}} \textbf{\bibinfo{volume}{23}},
-  \bibinfo{pages}{963--968} (\bibinfo{year}{2009}).
-
-\bibitem{Kooi2000}
-\bibinfo{author}{Kooijman, S. A. L.~M.}
-\newblock \emph{\bibinfo{title}{{Dynamic Energy and Mass Budgets in Biological
-  Systems}}} (\bibinfo{address}{Cambridge}, \bibinfo{year}{2000}).
-
-\bibitem{Sousa:2010ez}
-\bibinfo{author}{Sousa, T.}, \bibinfo{author}{Domingos, T.},
-  \bibinfo{author}{Poggiale, J.~C.} \& \bibinfo{author}{Kooijman, S. A. L.~M.}
-\newblock \bibinfo{title}{{Dynamic energy budget theory restores coherence in
-  biology}}.
-\newblock \emph{\bibinfo{journal}{Philos. T. Roy. Soc. B}}
-  \textbf{\bibinfo{volume}{365}}, \bibinfo{pages}{3413--3428}
-  (\bibinfo{year}{2010}).
-
-\bibitem{Diekmann:2010da}
-\bibinfo{author}{Diekmann, O.} \& \bibinfo{author}{Metz, J. A.~J.}
-\newblock \bibinfo{title}{{How to lift a model for individual behaviour to the
-  population level?}}
-\newblock \emph{\bibinfo{journal}{Philos. T. Roy. Soc. B}}
-  \textbf{\bibinfo{volume}{365}}, \bibinfo{pages}{3523--3530}
-  (\bibinfo{year}{2010}).
-
-\bibitem{murdoch:2003}
-\bibinfo{author}{Murdoch, W.~W.}, \bibinfo{author}{Briggs, C.~J.} \&
-  \bibinfo{author}{Nisbet, R.~M.}
-\newblock \emph{\bibinfo{title}{{Consumer-resource Dynamics}}},
-  vol.~\bibinfo{volume}{36} of \emph{\bibinfo{series}{Monographs in population
-  biology}} (\bibinfo{publisher}{Princeton University Press},
-  \bibinfo{address}{Princeton}, \bibinfo{year}{2003}).
-
-\bibitem{Damuth:1987kr}
-\bibinfo{author}{Damuth, J.}
-\newblock \bibinfo{title}{{Interspecific allometry of population density in
-  mammals and other animals: the independence of body mass and population
-  energy-use}}.
-\newblock \emph{\bibinfo{journal}{Biol. J. Linn. Soc.}}
-  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
-  (\bibinfo{year}{1987}).
-
-\bibitem{allen2002}
-\bibinfo{author}{Allen, A.~P.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{Gillooly, J.~F.}
-\newblock \bibinfo{title}{{Global biodiversity, biochemical kinetics, and the
-  energetic-equivalence rule}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{297}},
-  \bibinfo{pages}{1545--1548} (\bibinfo{year}{2002}).
-
-\bibitem{enquist1998}
-\bibinfo{author}{Enquist, B.~J.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{West, G.~B.}
-\newblock \bibinfo{title}{{Allometric scaling of plant energetics and
-  population density}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{395}},
-  \bibinfo{pages}{163--165} (\bibinfo{year}{1998}).
-
-\bibitem{Pedersen:2017he}
-\bibinfo{author}{Pedersen, R.~{\O}.}, \bibinfo{author}{Faurby, S.} \&
-  \bibinfo{author}{Svenning, J.-C.}
-\newblock \bibinfo{title}{{Shallow size{\textendash}density relations within
-  mammal clades suggest greater intra-guild ecological impact of large-bodied
-  species}}.
-\newblock \emph{\bibinfo{journal}{J. Anim. Ecol.}}
-  \textbf{\bibinfo{volume}{86}}, \bibinfo{pages}{1205--1213}
-  (\bibinfo{year}{2017}).
-
-\bibitem{Alroy:1998p1594}
-\bibinfo{author}{Alroy, J.}
-\newblock \bibinfo{title}{{Cope's rule and the dynamics of body mass evolution
-  in North American fossil mammals}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{280}},
-  \bibinfo{pages}{731--734} (\bibinfo{year}{1998}).
-
-\bibitem{Clauset:2009fh}
-\bibinfo{author}{Clauset, A.} \& \bibinfo{author}{Redner, S.}
-\newblock \bibinfo{title}{{Evolutionary model of species body mass
-  diversification}}.
-\newblock \emph{\bibinfo{journal}{Phys. Rev. Lett.}}
-  \textbf{\bibinfo{volume}{102}}, \bibinfo{pages}{038103}
-  (\bibinfo{year}{2009}).
-
-\bibitem{Smith:2010p3442}
-\bibinfo{author}{Smith, F.~A.} \emph{et~al.}
-\newblock \bibinfo{title}{{The evolution of maximum body size of terrestrial
-  mammals}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{330}},
-  \bibinfo{pages}{1216--1219} (\bibinfo{year}{2010}).
-
-\bibitem{Saarinen:2014br}
-\bibinfo{author}{Saarinen, J.~J.} \emph{et~al.}
-\newblock \bibinfo{title}{{Patterns of maximum body size evolution in Cenozoic
-  land mammals: Eco-evolutionary processes and abiotic forcing}}.
-\newblock \emph{\bibinfo{journal}{Proc Biol Sci}}
-  \textbf{\bibinfo{volume}{281}}, \bibinfo{pages}{20132049}
-  (\bibinfo{year}{2014}).
-
-\bibitem{Benichou:2014wu}
-\bibinfo{author}{B{\'e}nichou, O.} \& \bibinfo{author}{Redner, S.}
-\newblock \bibinfo{title}{{Depletion-controlled starvation of a diffusing
-  forager.}}
-\newblock \emph{\bibinfo{journal}{Phys. Rev. Lett.}}
-  \textbf{\bibinfo{volume}{113}}, \bibinfo{pages}{238101}
-  (\bibinfo{year}{2014}).
-
-\bibitem{Benichou:2016wl}
-\bibinfo{author}{B{\'e}nichou, O.}, \bibinfo{author}{Chupeau, M.} \&
-  \bibinfo{author}{Redner, S.}
-\newblock \bibinfo{title}{{Role of depletion on the dynamics of a diffusing
-  forager}}.
-\newblock \emph{\bibinfo{journal}{J Phys A-Math Theor}}
-  \textbf{\bibinfo{volume}{49}}, \bibinfo{pages}{394003}
-  (\bibinfo{year}{2016}).
-
-\bibitem{Chupeau:2016jf}
-\bibinfo{author}{Chupeau, M.}, \bibinfo{author}{B{\'e}nichou, O.} \&
-  \bibinfo{author}{Redner, S.}
-\newblock \bibinfo{title}{{Universality classes of foraging with resource
-  renewal}}.
-\newblock \emph{\bibinfo{journal}{Phys. Rev. E}} \textbf{\bibinfo{volume}{93}},
-  \bibinfo{pages}{032403} (\bibinfo{year}{2016}).
-
-\bibitem{Persson:1998hz}
-\bibinfo{author}{Persson, L.}, \bibinfo{author}{Leonardsson, K.},
-  \bibinfo{author}{De~Roos, A.~M.}, \bibinfo{author}{Gyllenberg, M.} \&
-  \bibinfo{author}{Christensen, B.}
-\newblock \bibinfo{title}{{Ontogenetic scaling of foraging rates and the
-  dynamics of a size-structured consumer-resource model}}.
-\newblock \emph{\bibinfo{journal}{Theor Popul Biol}}
-  \textbf{\bibinfo{volume}{54}}, \bibinfo{pages}{270--293}
-  (\bibinfo{year}{1998}).
-
-\bibitem{murray2011mathematical}
-\bibinfo{author}{Murray, J.~D.}
-\newblock \emph{\bibinfo{title}{{Mathematical Biology: I. An Introduction}}},
-  vol. \bibinfo{volume}{110} of \emph{\bibinfo{series}{Interdisciplinary
-  Applied Mathematics}} (\bibinfo{publisher}{Springer New York},
-  \bibinfo{year}{2011}).
-
-\bibitem{Strogatz:2008wo}
-\bibinfo{author}{Strogatz, S.~H.}
-\newblock \emph{\bibinfo{title}{{Nonlinear Dynamics and Chaos: With
-  Applications to Physics, Biology, Chemistry, and Engineering}}}.
-\newblock Studies in nonlinearity (\bibinfo{publisher}{Westview Press},
-  \bibinfo{address}{Boulder}, \bibinfo{year}{2008}).
-
-\bibitem{GuckHolmes}
-\bibinfo{author}{Guckenheimer, J.} \& \bibinfo{author}{Holmes, P.}
-\newblock \emph{\bibinfo{title}{{Nonlinear Oscillations, Dynamical Systems, and
-  Bifurcations of Vector Fields}}} (\bibinfo{publisher}{Springer},
-  \bibinfo{address}{New York}, \bibinfo{year}{1983}).
-
-\bibitem{Gross:2004p2428}
-\bibinfo{author}{Gross, T.} \& \bibinfo{author}{Feudel, U.}
-\newblock \bibinfo{title}{{Analytical search for bifurcation surfaces in
-  parameter space}}.
-\newblock \emph{\bibinfo{journal}{Physica D}} \textbf{\bibinfo{volume}{195}},
-  \bibinfo{pages}{292--302} (\bibinfo{year}{2004}).
-
-\bibitem{Hastings:2001jh}
-\bibinfo{author}{Hastings, A.}
-\newblock \bibinfo{title}{{Transient dynamics and persistence of ecological
-  systems}}.
-\newblock \emph{\bibinfo{journal}{Ecol. Lett.}} \textbf{\bibinfo{volume}{4}},
-  \bibinfo{pages}{215--220} (\bibinfo{year}{2001}).
-
-\bibitem{Neubert:1997wk}
-\bibinfo{author}{Neubert, M.} \& \bibinfo{author}{Caswell, H.}
-\newblock \bibinfo{title}{{Alternatives to resilience for measuring the
-  responses of ecological systems to perturbations}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{78}},
-  \bibinfo{pages}{653--665} (\bibinfo{year}{1997}).
-
-\bibitem{Caswell:2005eo}
-\bibinfo{author}{Caswell, H.} \& \bibinfo{author}{Neubert, M.~G.}
-\newblock \bibinfo{title}{{Reactivity and transient dynamics of discrete-time
-  ecological systems}}.
-\newblock \emph{\bibinfo{journal}{J Differ Equ Appl}}
-  \textbf{\bibinfo{volume}{11}}, \bibinfo{pages}{295--310}
-  (\bibinfo{year}{2005}).
-
-\bibitem{Neubert:2009td}
-\bibinfo{author}{Neubert, M.} \& \bibinfo{author}{Caswell, H.}
-\newblock \bibinfo{title}{{Detecting reactivity}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{90}},
-  \bibinfo{pages}{2683--2688} (\bibinfo{year}{2009}).
-
-\bibitem{Yodzis:1992hg}
-\bibinfo{author}{Yodzis, P.} \& \bibinfo{author}{Innes, S.}
-\newblock \bibinfo{title}{{Body size and consumer-resource dynamics}}.
-\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{139}},
-  \bibinfo{pages}{1151--1175} (\bibinfo{year}{1992}).
-
-\bibitem{Brown:2004wq}
-\bibinfo{author}{Brown, J.}, \bibinfo{author}{Gillooly, J.},
-  \bibinfo{author}{Allen, A.}, \bibinfo{author}{Savage, V.} \&
-  \bibinfo{author}{West, G.}
-\newblock \bibinfo{title}{{Toward a metabolic theory of ecology}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{85}},
-  \bibinfo{pages}{1771--1789} (\bibinfo{year}{2004}).
-
-\bibitem{Liow:2008jx}
-\bibinfo{author}{Liow, L.~H.} \emph{et~al.}
-\newblock \bibinfo{title}{{Higher origination and extinction rates in larger
-  mammals}}.
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{105}}, \bibinfo{pages}{6097--6102}
-  (\bibinfo{year}{2008}).
-
-\bibitem{DeLong:2012kw}
-\bibinfo{author}{DeLong, J.~P.} \& \bibinfo{author}{Vasseur, D.~A.}
-\newblock \bibinfo{title}{{A dynamic explanation of size{\textendash}density
-  scaling in carnivores}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{93}},
-  \bibinfo{pages}{470--476} (\bibinfo{year}{2012}).
-
-\bibitem{Kempes:2016}
-\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Wang, L.},
-  \bibinfo{author}{Amend, J.~P.}, \bibinfo{author}{Doyle, J.} \&
-  \bibinfo{author}{Hoehler, T.}
-\newblock \bibinfo{title}{{Evolutionary tradeoffs in cellular composition
-  across diverse bacteria}}.
-\newblock \emph{\bibinfo{journal}{ISME J}} \textbf{\bibinfo{volume}{10}},
-  \bibinfo{pages}{2145--2157} (\bibinfo{year}{2016}).
-
-\bibitem{Carbone:1999ju}
-\bibinfo{author}{Carbone, C.}, \bibinfo{author}{Mace, G.~M.},
-  \bibinfo{author}{Roberts, S.~C.} \& \bibinfo{author}{Macdonald, D.~W.}
-\newblock \bibinfo{title}{{Energetic constraints on the diet of terrestrial
-  carnivores}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{402}},
-  \bibinfo{pages}{286--288} (\bibinfo{year}{1999}).
-
-\bibitem{Carbone:2007dz}
-\bibinfo{author}{Carbone, C.}, \bibinfo{author}{Teacher, A.} \&
-  \bibinfo{author}{Rowcliffe, J.~M.}
-\newblock \bibinfo{title}{{The Costs of Carnivory}}.
-\newblock \emph{\bibinfo{journal}{PLoS Biol}} \textbf{\bibinfo{volume}{5}},
-  \bibinfo{pages}{e22} (\bibinfo{year}{2007}).
-
-\bibitem{Okie:2013ju}
-\bibinfo{author}{Okie, J.~G.} \emph{et~al.}
-\newblock \bibinfo{title}{{Effects of allometry, productivity and lifestyle on
-  rates and limits of body size evolution}}.
-\newblock \emph{\bibinfo{journal}{Proc Biol Sci}}
-  \textbf{\bibinfo{volume}{280}}, \bibinfo{pages}{20131007--20131007}
-  (\bibinfo{year}{2013}).
-
-\bibitem{Brown:1993p708}
-\bibinfo{author}{Brown, J.}, \bibinfo{author}{Marquet, P.} \&
-  \bibinfo{author}{Taper, M.}
-\newblock \bibinfo{title}{{Evolution of body size: consequences of an energetic
-  definition of fitness}}.
-\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{142}},
-  \bibinfo{pages}{573--584} (\bibinfo{year}{1993}).
-
-\bibitem{West:1997cg}
-\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{Enquist, B.~J.}
-\newblock \bibinfo{title}{{A general model for the origin of allometric scaling
-  laws in biology}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{276}},
-  \bibinfo{pages}{122--126} (\bibinfo{year}{1997}).
-
-\bibitem{West:2002ud}
-\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Woodruff, W.~H.} \&
-  \bibinfo{author}{Brown, J.~H.}
-\newblock \bibinfo{title}{{Allometric scaling of metabolic rate from molecules
-  and mitochondria to cells and mammals.}}
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{99 Suppl 1}}, \bibinfo{pages}{2473--2478}
-  (\bibinfo{year}{2002}).
-
-\bibitem{Millar:1990p923}
-\bibinfo{author}{Millar, J.} \& \bibinfo{author}{Hickling, G.}
-\newblock \bibinfo{title}{{Fasting endurance and the evolution of mammalian
-  body size}}.
-\newblock \emph{\bibinfo{journal}{Funct. Ecol.}} \textbf{\bibinfo{volume}{4}},
-  \bibinfo{pages}{5--12} (\bibinfo{year}{1990}).
-
-\bibitem{tilman1981}
-\bibinfo{author}{Tilman, D.}
-\newblock \bibinfo{title}{{Tests of resource competition theory using four
-  species of lake michigan algae}}.
-\newblock \emph{\bibinfo{journal}{Ecology}} \textbf{\bibinfo{volume}{62}},
-  \bibinfo{pages}{802--815} (\bibinfo{year}{1981}).
-
-\bibitem{dutkiewicz2009}
-\bibinfo{author}{Dutkiewicz, S.}, \bibinfo{author}{Follows, M.~J.} \&
-  \bibinfo{author}{Bragg, J.~G.}
-\newblock \bibinfo{title}{{Modeling the coupling of ocean ecology and
-  biogeochemistry}}.
-\newblock \emph{\bibinfo{journal}{Global Biogeochem. Cycles}}
-  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{1--15} (\bibinfo{year}{2009}).
-
-\bibitem{barton2010}
-\bibinfo{author}{Barton, A.~D.}, \bibinfo{author}{Dutkiewicz, S.},
-  \bibinfo{author}{Flierl, G.}, \bibinfo{author}{Bragg, J.} \&
-  \bibinfo{author}{Follows, M.~J.}
-\newblock \bibinfo{title}{{Patterns of diversity in marine phytoplankton}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{327}},
-  \bibinfo{pages}{1509--1511} (\bibinfo{year}{2010}).
-
-\bibitem{Hempson:2015hka}
-\bibinfo{author}{Hempson, G.~P.}, \bibinfo{author}{Archibald, S.} \&
-  \bibinfo{author}{Bond, W.~J.}
-\newblock \bibinfo{title}{{A continent-wide assessment of the form and
-  intensity of large mammal herbivory in Africa}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{350}},
-  \bibinfo{pages}{1056--1061} (\bibinfo{year}{2015}).
-
-\bibitem{West:2002it}
-\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Woodruff, W.~H.} \&
-  \bibinfo{author}{Brown, J.~H.}
-\newblock \bibinfo{title}{{Allometric scaling of metabolic rate from molecules
-  and mitochondria to cells and mammals.}}
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{99 Suppl 1}}, \bibinfo{pages}{2473--2478}
-  (\bibinfo{year}{2002}).
-
-\bibitem{DeLong:2010dy}
-\bibinfo{author}{DeLong, J.~P.}, \bibinfo{author}{Okie, J.~G.},
-  \bibinfo{author}{Moses, M.~E.}, \bibinfo{author}{Sibly, R.~M.} \&
-  \bibinfo{author}{Brown, J.~H.}
-\newblock \bibinfo{title}{{Shifts in metabolic scaling, production, and
-  efficiency across major evolutionary transitions of life}}.
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{107}}, \bibinfo{pages}{12941--12945}
-  (\bibinfo{year}{2010}).
-
-\bibitem{West:2001bv}
-\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Brown, J.~H.} \&
-  \bibinfo{author}{Enquist, B.~J.}
-\newblock \bibinfo{title}{{A general model for ontogenetic growth}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{413}},
-  \bibinfo{pages}{628--631} (\bibinfo{year}{2001}).
-
-\bibitem{moses2008rmo}
-\bibinfo{author}{Moses, M.~E.} \emph{et~al.}
-\newblock \bibinfo{title}{{Revisiting a Model of Ontogenetic Growth: Estimating
-  Model Parameters from Theory and Data.}}
-\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{171}},
-  \bibinfo{pages}{632--645} (\bibinfo{year}{2008}).
-
-\bibitem{gillooly2002esa}
-\bibinfo{author}{Gillooly, J.~F.}, \bibinfo{author}{Charnov, E.~L.},
-  \bibinfo{author}{West, G.~B.}, \bibinfo{author}{Savage, V.~M.} \&
-  \bibinfo{author}{Brown, J.~H.}
-\newblock \bibinfo{title}{{Effects of size and temperature on developmental
-  time}}.
-\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{417}},
-  \bibinfo{pages}{70--73} (\bibinfo{year}{2002}).
-
-\bibitem{hou}
-\bibinfo{author}{Hou, C.} \emph{et~al.}
-\newblock \bibinfo{title}{{Energy uptake and allocation during ontogeny}}.
-\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{322}},
-  \bibinfo{pages}{736--739} (\bibinfo{year}{2008}).
-
-\bibitem{Savage:2004ed}
-\bibinfo{author}{Savage, V.~M.}, \bibinfo{author}{Gillooly, J.~F.},
-  \bibinfo{author}{Brown, J.~H.}, \bibinfo{author}{West, G.~B.} \&
-  \bibinfo{author}{Charnov, E.~L.}
-\newblock \bibinfo{title}{{Effects of Body Size and Temperature on Population
-  Growth}}.
-\newblock
-  \emph{\bibinfo{journal}{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}}
-  \textbf{\bibinfo{volume}{163}}, \bibinfo{pages}{429--441}
-  (\bibinfo{year}{2004}).
-
-\bibitem{bettencourt}
-\bibinfo{author}{Bettencourt, L. M.~A.}, \bibinfo{author}{Lobo, J.},
-  \bibinfo{author}{Helbing, D.}, \bibinfo{author}{Kuhnert, C.} \&
-  \bibinfo{author}{West, G.~B.}
-\newblock \bibinfo{title}{{Growth, innovation, scaling, and the pace of life in
-  cities}}.
-\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
-  \textbf{\bibinfo{volume}{104}}, \bibinfo{pages}{7301--7306}
-  (\bibinfo{year}{2007}).
-
-\bibitem{Folland:2008ij}
-\bibinfo{author}{Folland, J.~P.}, \bibinfo{author}{Mc~Cauley, T.~M.} \&
-  \bibinfo{author}{Williams, A.~G.}
-\newblock \bibinfo{title}{{Allometric scaling of strength measurements to body
-  size}}.
-\newblock \emph{\bibinfo{journal}{Eur J Appl Physiol}}
-  \textbf{\bibinfo{volume}{102}}, \bibinfo{pages}{739--745}
-  (\bibinfo{year}{2008}).
-
-\end{thebibliography}
-
-\end{bibunit}
-
-% \bibliography{aa_starving3}
-
-%
-% \begin{acknowledgments}
-%   We thank Luis Bettencourt, Jean Philippe Gibert, Eric Libby, and Seth Newsome for helpful
-%   discussions and comments on the manuscript.  J.D.Y. was supported by
-%   startup funds at the University of California, Merced, and an Omidyar
-%   Postdoctoral Fellowship at the Santa Fe Institute.  C.P.K. was supported by
-%   an Omidyar Postdoctoral Fellowship at the Santa Fe Institute.  S.R. was
-%   supported by grants DMR-1608211 and 1623243 from the National Science
-%   Foundation, and by the John Templeton Foundation, all at the Santa Fe
-%   Institute.
-% \end{acknowledgments}
-%
-\clearpage
-
-\begin{bibunit}[unsrt]
-
-% \addto\captionsenglish{\renewcommand{\figurename}{Figure S\!}}
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}%
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}%
-
-\section*{Supporting Information for ``The dynamics of starvation and recovery''}
-
-{\bf Mechanisms of Starvation and Recovery}
-To understand the dynamics of starvation, recovery, reproduction, and resource competition, our framework partitions consumers into two classes: (a) a full class that is able to reproduce and, (b) a hungry class that experiences mortality at a given rate and is unable to reproduce.
-For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate, as dictated by body size, with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
-\citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
-This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g.
-\citep{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \citep{kempes2014morphological}.
-Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).
-
-We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
-\begin{align}
-\begin{split}
-\dot{F_{d}} &= \lambda_{\text{max}} F_{d} + \rho_{\text{max}}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{\text{max}}R_{d} H_{d}/k - \mu H_{d},  \\
-\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
-& \left[\left(\frac{\rho_{\text{max}}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)F_{d}\right].
-\label{bigdynamics}
-\end{split}
-\end{align}
-where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with the subscript $_{d}$ before introducing the non-dimensional form that is presented in the main text).
-In the above equations $Y$ represents the yield coefficient (e.g., Refs.
-\citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism$^{-1}$).
-If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
-This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
-In these units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
-%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
-
-%\begin{align}
-%\begin{split}
-%\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-%\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
-%\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
-%&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
-%\end{split}
-%\end{align}
-%
-%In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability.
-Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources.
-The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
-%A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
-%
-%Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification.
-As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$.
-Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
-
-%where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
-
-We can formally non-dimensionalize this system by the rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
-%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
-\begin{align}
-\begin{split}
-&\dot{F} = \frac{1}{s}\left[\lambda_{\text{max}} F + \rho_{\text{max}}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
-&\dot{H} = \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{\text{max}}\frac{R}{qk} H - \mu H\right],  \\
-& \dot{R} = \\
-&\frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{\text{max}}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)F\right]\right].
-\end{split}
-\end{align}
-If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
-\begin{align}
-\begin{split}
-\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
-\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
-\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
-\label{reduceddynamics}
-\end{split}
-\end{align}
-where we have dropped the subscripts on $\lambda_{\text{max}}$ and $\rho_{\text{max}}$ for simplicity, and $\xi\equiv C/k$, $\delta\equiv Y_{H}kP_{H}/C$, and $\beta\equiv Y_{H}k\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)/C$.
-The above equations represent the system of equations presented in the main text.
-\\
-
-{\bf Parameter Values and Estimates}
-All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies.
-Below we outline previous measurements and simple estimates of the parameters.
-
-Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs.
-\citep{West:2001bv,moses2008rmo} and the supplement for Ref.
-\citep{hou}).
-We make this assumption in the current paper, although alternate exponents, which are known to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends.
-The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.
-For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs.
-\citep{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate.
-We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).
-
-An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction.
-This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref.
-\citep{peters1986ecological} for a review).
-We use $m_{0}=0.097M^{0.92}$ \citep{blueweiss1978relationships} which encompasses the widest range of body sizes \citep{peters1986ecological}.
-
-The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g.
-Refs.
-\citep{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \citep{moses2008rmo}.
-The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \citep{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \citep{stryer,hou}.
-For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \citep{hou}.
-
-For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ ~\citep{Dunbrack:1993ec}, $0.02M^{1.19}$ ~\citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ ~\citep{Lindstedt:2002td}.
-We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range.
-Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ ~\citep{Lindstedt:2002td}.
-
-%We also connect the resource growth rate to the total metabolic rate of an organism.
-That is, we are interested in the relative rates of resource recovery and consumption by the total population.
-From \citep{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$).
-Considering an energy density of $18200$ (J g$^{-1}$) of grass \citep{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$).
-This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.
-
-Typically the value of $\xi=C/k$ should roughly be $2$.
-The value of $\rho$, $\lambda$, $\sigma$, and $\mu$ are all simple rates (note that we have not rescaled time in our non-dimensionalization) as defined in the maintext.
-Given that our model considers transitions over entire stages of ontogeny or nutritional states, the value of $Y$ must represent yields integrated over entire life stages.
-Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
-\begin{equation}
-B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-Similarly, the maintenance resource consumption rate for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
-\begin{equation}
-B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.
-
-Finally, the value of $\alpha$ can be roughly estimated by the NPP divided by the corresponding biomass densities.
-From the data in Ref.
-\citep{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
-It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with $B_{tot}$ for each type.
-As such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{damuth1987interspecific}.
-We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above.
-However, two points are important to note here: first, our framework predicts the overall scaling of $F^{*}$ and $H^{*}$ independently of $\alpha$ and this correctly matches data, and second, both the asymptotic behavior and slope of $F^{*}$ and $H^{*}$ are independent of $\alpha$, such that our prediction of the maximum mammal size does not depend on $\alpha$.
-\\
-%For the growth rate $\lambda$ we consider the standard model of $\ln\left(\upsilon\right)/t_{\lambda}$
-
-%More complicated models of fecundity (which, for example, account for the average length of adulthood and the number of individuals produced over this span) could be employed.
-However, the scaling of population growth rate has been studied in detail before and follows a relationship of $$ which matches the theory well for $\phi=.95$ and $=$.
-
-%In our calculations we include $20\%$ variation around this value which could account for differences in efficiency during
-
- \begin{table}[h]
-\caption{Parameter values for mammals}
-\label{param}
-    \begin{center}
-    \footnotesize
-     \begin{tabular}{p{3.8cm} c p{2.2cm} p{1.4cm}}
-     \hline
-
-     Definition & Parameter & Value & References  \\
-     \hline
-   Asymptotic adult mass & $M$ & (g) &  \\
-   Initial mass of an organism & $m_{0}$ & (g) &  \\
-   Metabolic rate scaling exponent & $\eta$ & $3/4$  &  (e.g. \citep{West:2001bv,moses2008rmo,hou}) \\
-   Metabolic Normalization Constant & $B_{0}$ & $0.047$ (W g$^{-0.75}$)    & \citep{hou}  \\
-   Initial mass scaling exponent & $\upsilon$ & $0.92$ &  \citep{blueweiss1978relationships,peters1986ecological} \\
-   Initial mass scaling normalization constant & $n_{0}$ & $0.097$ (g$^{1-\upsilon}$) & \citep{blueweiss1978relationships,peters1986ecological}  \\   
-   Fat mass scaling exponent & $\gamma$ & $1.19$ & \citep{Lindstedt:1985hm} \\
-   Fat scaling normalization constant & $f_{0}$ & $0.02$ (g$^{1-\eta}$) & \citep{Lindstedt:1985hm}\\
-   Muscle mass scaling exponent & $\zeta$ & $1.00$  & \citep{Lindstedt:2002td} \\
-   Muscle scaling normalization constantv& $u_{0}$ & $0.38$ (g$^{1-\zeta}$)  & \citep{Lindstedt:2002td} \\
-   Energy to synthesis a unit of mass & $E_{m}$ & $5774$ (J gram$^{-1}$)  &  \citep{moses2008rmo,West:2001bv,hou} \\
-   Energy to synthesis a unit of mass during recovery & $E_{m}^{\prime}$ & $7000$ (J gram$^{-1}$) & \citep{stryer,hou} \\
-   Specific resource growth rate & $\alpha$ & $9.45\times10^{-9}$ (s$^{-1}$) & see text  \\
-   Fraction of asymptotic mass representing full state & $\epsilon_{\lambda}$ & $0.95$ & \citep{West:2001bv}  \\
-   Fraction of asymptotic mass representing starving state & $\epsilon_{\sigma}$ & $1-f_{0}M^{\gamma-1}$ & see text  \\
-   Fraction of asymptotic mass representing death & $\epsilon_{\mu}$ & $1-\frac{f_{0}M^{\gamma}+u_{0}M^{\zeta}}{M}$ & see text \\
-   Carrying capacity (maximum density) of resources & $C$ & (g m$^{-2}$) & \\
-   Half Saturation Constant & $k$ & (g m$^{-2}$) &   \\
-   Normalized carrying capacity & $\xi$ & $C/k\approx2$ &   \\
-   Reproductive fecundity & $\nu$ & $2$ & \citep{}  \\
-
-%   $a$ & $1.78\times10^{-6}$  \quad \quad \\
-%   $\lambda_{0}$ & $3.39\times10^{-7}$ (s$^{-1}$ gram$^{1-\eta}$) \quad \quad \\
-
-   \hline
-    \end{tabular}
-    \end{center}
-   \end{table}
-
-{\bf Rate equations for invaders with modified body mass}
-We allow an invading subset of the resident population with mass $M$ to have an altered mass $M^\prime = M(1+\chi)$ where $\chi$ varies between $\chi_{\rm min} <0$ and $\chi_{\rm max}>0$, where $\chi<0$ denotes a leaner invader and $\chi > 0$ denotes an invader with additional reserves of body fat.
-Importantly, we assume that the invading and resident individuals have the same proportion of non-fat tissues.
-For the allowable values of $\chi$ the adjusted mass should exceed the amount of body fat, $1+\chi>\epsilon_{\sigma}$, and the adjusted time to reproduce must be positive, which given our solution for $\tau(\epsilon)$ (see main text), implies that $1-\epsilon_{\lambda}^{1-\eta}\left(1+\chi\right)^{1-\eta}>0$.
-Together these conditions imply that  $\chi\in(-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$ where the upper bound approximately equals $0.05$.
-
-Although the starved state of invading organisms remains unchanged, the rate of starvation from the modified full state to the starved state, the rate of recovery from the starved state to the modified full state, and the maintenance rates of both, will be different, such that $\sigma^\prime = \sigma(M^\prime)$, $\rho^\prime = \rho(M^\prime)$, $\beta^\prime = \beta(M^\prime)$, $\delta^\prime = \delta(M^\prime)$.
-Rates of starvation and recovery for the invading population are easily derived by adjusting the starting or ending state before and after starvation and recovery, leading to the following timescales:
-
-\begin{align}
-t_{\sigma^\prime} &= -\frac{M^{1-\eta}}{a^{\prime}}\ln \left(\frac{\epsilon_\sigma}{\chi +1}\right), \\ \nonumber
-t_{\rho^\prime} &= \ln \left(\frac{1-(\epsilon_\lambda \epsilon_\sigma)^{1/4}}{1-( \epsilon_\lambda(\chi +1))^{1/4}}\right)\frac{M^{1-\eta}}{a^{\prime}\left(1-\eta\right)}.
-\end{align}
-
-The maintenance rates for the invading population require more careful consideration.
-First, we must recalculate the yields $Y$, as they must now be integrated over life stages that have also been slightly modified by the addition or subtraction of body fat reserves.
-Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value of the invading population is given by $P_{F}=B_{0}(1+\chi)M^{3/4}/(1+\chi)ME_{d}$, and the yield for a full organism will be given by $Y_{F}=(1+\chi)ME_{d}/B^{\prime}_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$) where $B^{\prime}_{\lambda}$ is the lifetime energy use for the invading population reaching maturity given by
-\begin{equation}
-B^{\prime}_{\lambda}=\int_{0}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-where
-\begin{equation}
-t_{\lambda^\prime} = \frac{M^{1-\eta} }{a(1-\eta)}\ln \left(\frac{1-(m_0/M)^{1-\eta}}{1-(\epsilon_\lambda (1+\chi))^{1-\eta}} \right).
-\end{equation}
-Note that we do not use this timescale to determine the reproductive rate of the invading consumer---which is assumed to remain the same as the resident population---but only to calulate the lifetime energy use.
-Similarly, the maintenance for hungry individuals $P^\prime_{H}=B_{0}(\epsilon_{\sigma}(1+\chi)M)^{3/4}/(\epsilon_{\sigma}(1+\chi)M)E_{d}$ and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y^\prime_{H}=(1+\chi)ME_{d}/B^{\prime}_{\rho}$ where
-\begin{equation}
-B^{\prime}_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda^\prime}}B_{0}m\left(t\right)^{\eta}dt.
-\end{equation}
-Finally, we can calculate the maintenance of the invaders as
-
-\begin{align}
-  \delta^\prime &= P^\prime_{H}Y^\prime_{H}/\xi \\ \nonumber
-  \beta^\prime &= \left(\frac{\lambda_{\rm max}}{Y^\prime_{F}}+P^\prime_{F} \right)Y^\prime_{H}/\xi.
-\end{align}
-
-To determine whether or not the invader or resident population has an advantage, we compute $R^*(M)$ and $R^*(M^\prime=M(1+\chi))$ for values of $\chi \in (-f_0M^{\gamma-1},1/\epsilon_{\lambda}-1)$, and the invading population is assumed to have an advantage over the resident population if $R^*(M^\prime)<R^*(M)$.
-
-{\bf Sensitivity to additional death terms}
-
-It should be noted that our set of dynamics (Equations \ref{bigdynamics} and \ref{reduceddynamics}) could include a constant death term of the form $-d_{F}F$ and $-d_{H}H$ to represent death not directly linked to starvation.
-Adding terms of this form to our model would simply adjust the effective value of $\lambda$ and $\mu$, and we could rewrite Equation \ref{reduceddynamics} with $\lambda^{\prime}=\lambda-d$ and $\mu^{\prime}=\mu-d$.
+It should be noted that our set of dynamics (Eqs.
+2 and 4, main text) could include a constant death term of the form $-d_{F}F$ and $-d_{H}H$ to represent death not directly linked to starvation.
+Adding terms of this form to our model would simply adjust the effective value of $\lambda$ and $\mu$, and we could rewrite Eq.
+4 with $\lambda^{\prime}=\lambda-d$ and $\mu^{\prime}=\mu-d$.
 These substitutions would not alter the functional form of our model nor the steady-states and qualitative results, however the quantitative values could shift based on the size of $d$ relative to $\lambda$ and $\mu$.
 
-Survivorship has a well-known functional form which changes systematically with size (e.g. \cite{calder1984}). Typically survivorship is defined using the Gompertz curve 
+Survivorship has a well-known functional form which changes systematically with size (e.g. \citep{calder1984}). Typically survivorship is defined using the Gompertz curve 
 \begin{equation}
@@
 \end{equation}
-where the parameters have the following allometric dependencies on adult mass $c_{0}=a_{0}M^{b_{0}}$ and $c_{1}=a_{1}M^{b_{1}}$, with $a_{0}=1.88\times10^{-8}$ (s g$^{-b_{0}}$), $b_{0}=-0.56$, $a_{1}=1.45\times10^{-7}$ (s g$^{-b_{1}}$), and $b_{1}=-0.27$ (see \cite{calder1984} for a review).
+where the parameters have the following allometric dependencies on adult mass $c_{0}=a_{0}M^{b_{0}}$ and $c_{1}=a_{1}M^{b_{1}}$, with $a_{0}=1.88\times10^{-8}$ (s g$^{-b_{0}}$), $b_{0}=-0.56$, $a_{1}=1.45\times10^{-7}$ (s g$^{-b_{1}}$), and $b_{1}=-0.27$ (see \citep{calder1984} for a review).
 
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{mortality-rate-comparison-eps-converted-to.pdf}
-\caption{\small{The rates of reproduction $\lambda$ (blue), starvation-based mortality $\mu$ (red), and survivorship-based death $\bar{d}$ (black) as a function of adult mass.}\label{fig:ratescomp}}
-\end{figure}
-
-We are interested in the specific death rate of the form $\dot{F}=-dF$, and using the derivative of Equation \ref{gompertz} we find that $d=c_{0}e^{c_{1}t}$.
+We are interested in the specific death rate of the form $\dot{F}=-dF$, and using the derivative of Eq.
+\ref{gompertz} we find that $d=c_{0}e^{c_{1}t}$.
 Our model considers the average rates over a population and lifecycle and the average death rate is given by
@@
 \end{eqnarray}
-where $t_{\text{exp}}$ is the expected lifespan following the allometry of $t_{\text{exp}}=a_{2}M^{b_{2}}$ with $a_{2}=4.04\times10^{6}$ (s g$^{-b_{2}}$) and $b_{2}=0.30$ ~\cite{damuth1982analysis,calder1984}.
+where $t_{\text{exp}}$ is the expected lifespan following the allometry of $t_{\text{exp}}=a_{2}M^{b_{2}}$ with $a_{2}=4.04\times10^{6}$ (s g$^{-b_{2}}$) and $b_{2}=0.30$ ~\citep{damuth1982analysis,calder1984}.
 Given the allometries above we have that
@@
 which scales roughly like $M^{b_{0}}$ because $b_{1}$ and $b_{2}$ are close in value but opposite in sign.
-In Figure S\ref{fig:ratescomp} we compare the value of $\bar{d}$ to the reproductive, $\lambda$, and starvation-based mortality, $\mu$, rates.
-The values of $\bar{d}$ are orders of magnitude smaller than these rates for all mammalian masses, and thus, adding this non-starvation based death rate to our model does not shift our results within numerical confidence.
+In Supplementary Figure \ref{fig:ratescomp} we compare the value of $\bar{d}$ to the reproductive, $\lambda$, and starvation-based mortality, $\mu$, rates.
+The values of $\bar{d}$ are orders of magnitude smaller than these rates for all mammalian masses, and thus, adding this non-starvation based death rate to our model does not shift our results within numerical confidence.\\
+
+\noindent {\bf Supplementary Note 2: NSM and the energy equivalence hypothesis}
+
+The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}.
+As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}.
+However, the actual equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations.
+Supplementary Figure \ref{fig:equivalence} recasts the results of the NSM in terms of this hypothesis and shows that $F^{*}B$ is nearly constant over the same range of mammalian sizes up to the asymptotic behavior for the largest terrestrial mammals.\\
+
+\noindent {\bf Supplementary Note 3: Application of NSM limits to aquatic mammals}
+A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass of a male African elephant.
+We note this particular limit as it may have future relevance to considerations of the ultimate constraints on aquatic mammals.\\
+
+\clearpage
+
+\noindent {\bf Supplementary Figures}\\
 
@@
 \centering
-\includegraphics[width=0.4\textwidth]{fig_FPenergyequiv-eps-converted-to.pdf}
+\includegraphics[width=0.4\textwidth]{mortality-rate-comparison.eps}
+\caption{\small{The rates of reproduction $\lambda$ (blue), starvation-based mortality $\mu$ (red), and survivorship-based death $\bar{d}$ (black) as a function of adult mass.}\label{fig:ratescomp}}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.4\textwidth]{fig_FPenergyequiv.eps}
 \caption{\small{ Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass ($F^*$ is shown in green and $H^*$ in orange).
@@
 
-{\bf NSM and the energy equivalence hypothesis}
+\clearpage
 
-The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}.
-As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}.
-However, the actual equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations.
-Figure S\ref{fig:equivalence} recasts the results of the NSM in terms of this hypothesis and shows that $F^{*}B$ is nearly constant over the same range of mammalian sizes up to the asymptotic behavior for the largest terrestrial mammals.
-
-{\bf Application of NSM limits to aquatic mammals}
-A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass of a male African elephant.
-We note this particular limit as it may have future relevance to considerations of the ultimate constraints on aquatic mammals.
-
-% \putbib[aa_starving_supplement]
-
+\noindent {\bf Supplementary References}\\
 \def\bibfont{\footnotesize}
@@
 \newblock
-  \emph{\bibinfo{journal}{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}}
+  \emph{\bibinfo{journal}{American Naturalist}}
   \textbf{\bibinfo{volume}{171}}, \bibinfo{pages}{632--645}
@@
 \bibinfo{author}{Dunbrack, R.~L.} \& \bibinfo{author}{Ramsay, M.~A.}
-\newblock \bibinfo{title}{{The Allometry of Mammalian Adaptations to Seasonal
-  Environments: A Critique of the Fasting Endurance Hypothesis}}.
+\newblock \bibinfo{title}{{The allometry of mammalian adaptations to seasonal
+  environments: A critique of the fasting endurance hypothesis}}.
 \newblock \emph{\bibinfo{journal}{Oikos}} \textbf{\bibinfo{volume}{66}},
@@
 \bibinfo{author}{Lindstedt, S.~L.} \& \bibinfo{author}{Boyce, M.~S.}
-\newblock \bibinfo{title}{{Seasonality, Fasting Endurance, and Body Size in
-  Mammals}}.
+\newblock \bibinfo{title}{{Seasonality, fasting endurance, and body size in
+  mammals}}.
 \newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{125}},

### 2017-12-13 NatureComm_REVISION/YKR_supp-Revision-1.tex
editorial revisions

--- 2017-10-18 NatureComm_REVISION/YKR_supp-Revision-1.tex
@@
 \linenumbers
-% \setlength\linenumbersep{3pt}
+ \setlength\linenumbersep{3pt}
+
+\addto\captionsenglish{\renewcommand{\figurename}{Figure S\!}}
 
@@
 \author{Justin D. Yeakel} \affiliation{School of Natural Sciences, University
-  of California, Merced, Merced, CA 95340, USA}\affiliation{The Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author: jdyeakel@gmail.com}
+  of California, Merced, Merced, CA 95340, USA}\affiliation{The Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
 
 \author{Christopher P. Kempes} \affiliation{The Santa Fe Institute, 1399 Hyde
-  Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}
+  Park Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
 
 \author{Sidney Redner} \affiliation{The Santa Fe Institute, 1399 Hyde Park
-  Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}
+  Road, Santa Fe, NM 87501, USA}\affiliation{Contributed equally}\affiliation{Corresponding author}
 
@@
 {\bf Mechanisms of Starvation and Recovery}
-Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce.
-For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
+To understand the dynamics of starvation, recovery, reproduction, and resource competition, our framework partitions consumers into two classes: (a) a full class that is able to reproduce and, (b) a hungry class that experiences mortality at a given rate and is unable to reproduce.
+For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate, as dictated by body size, with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs.
 \citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective).
@@
 \begin{split}
-\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
-\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
+\dot{F_{d}} &= \lambda_{\text{max}} F_{d} + \rho_{\text{max}}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
+\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{\text{max}}R_{d} H_{d}/k - \mu H_{d},  \\
 \dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
-& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
+& \left[\left(\frac{\rho_{\text{max}}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)F_{d}\right].
+\label{bigdynamics}
 \end{split}
 \end{align}
-where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text).
+where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with the subscript $_{d}$ before introducing the non-dimensional form that is presented in the main text).
 In the above equations $Y$ represents the yield coefficient (e.g., Refs.
-\citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism).
+\citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism$^{-1}$).
 If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$.
 This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density.
-In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
+In these units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
 %(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
@@
 
-We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
+We can formally non-dimensionalize this system by the rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
 %(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
@@
 \begin{split}
-\dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
-\dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
-\dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
+&\dot{F} = \frac{1}{s}\left[\lambda_{\text{max}} F + \rho_{\text{max}}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
+&\dot{H} = \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{\text{max}}\frac{R}{qk} H - \mu H\right],  \\
+& \dot{R} = \\
+&\frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{\text{max}}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)F\right]\right].
 \end{split}
@@
 \dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
+\label{reduceddynamics}
 \end{split}
 \end{align}
-where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$.
+where we have dropped the subscripts on $\lambda_{\text{max}}$ and $\rho_{\text{max}}$ for simplicity, and $\xi\equiv C/k$, $\delta\equiv Y_{H}kP_{H}/C$, and $\beta\equiv Y_{H}k\left(\frac{\lambda_{\text{max}}}{Y_{F}}+P_{F}\right)/C$.
 The above equations represent the system of equations presented in the main text.
@@
 
-For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \citep{Dunbrack:1993ec}, $0.02M^{1.19}$ \citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ \citep{Lindstedt:2002td}.
+For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ ~\citep{Dunbrack:1993ec}, $0.02M^{1.19}$ ~\citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ ~\citep{Lindstedt:2002td}.
 We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range.
-Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \citep{Lindstedt:2002td}.
+Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ ~\citep{Lindstedt:2002td}.
 
@@
 \end{equation}
-Similarly, the maintenance for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
+Similarly, the maintenance resource consumption rate for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
 \begin{equation}
@@
 \citep{michaletz2014convergence} we estimate the value of $\alpha$ to range between $2.81\times10^{-10}$ (s$^{-1}$) and $2.19\times10^{-8}$ (s$^{-1}$) globally.
-It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with the $B_{tot}$ for each type, and as such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{damuth1987interspecific}.
+It should be noted that the value of $\alpha$ sets the overall scale of the $F^{*}$ and $H^{*}$ steady states along with $B_{tot}$ for each type.
+As such, we use $\alpha$ as our fit parameter to match these steady states with the data from Damuth \citep{damuth1987interspecific}.
 We find that the best fit is $\alpha=9.45\times10^{-9}$ (s$^{-1}$) which compares well with the calculated range above.
@@
 
-\bibliography{aa_starving_supplement}
+{\bf Sensitivity to additional death terms}
+
+It should be noted that our set of dynamics (Equations \ref{bigdynamics} and \ref{reduceddynamics}) could include a constant death term of the form $-d_{F}F$ and $-d_{H}H$ to represent death not directly linked to starvation.
+Adding terms of this form to our model would simply adjust the effective value of $\lambda$ and $\mu$, and we could rewrite Equation \ref{reduceddynamics} with $\lambda^{\prime}=\lambda-d$ and $\mu^{\prime}=\mu-d$.
+These substitutions would not alter the functional form of our model nor the steady-states and qualitative results, however the quantitative values could shift based on the size of $d$ relative to $\lambda$ and $\mu$.
+
+Survivorship has a well-known functional form which changes systematically with size (e.g. \cite{calder1984}). Typically survivorship is defined using the Gompertz curve 
+\begin{equation}
+F=F_{0}e^{\left(c_{0}/c_{1}\right)\left(1-e^{c_{1}t}\right)}
+\label{gompertz}
+\end{equation}
+where the parameters have the following allometric dependencies on adult mass $c_{0}=a_{0}M^{b_{0}}$ and $c_{1}=a_{1}M^{b_{1}}$, with $a_{0}=1.88\times10^{-8}$ (s g$^{-b_{0}}$), $b_{0}=-0.56$, $a_{1}=1.45\times10^{-7}$ (s g$^{-b_{1}}$), and $b_{1}=-0.27$ (see \cite{calder1984} for a review).
+
+\begin{figure}
+\centering
+\includegraphics[width=0.4\textwidth]{mortality-rate-comparison.eps}
+\caption{\small{The rates of reproduction $\lambda$ (blue), starvation-based mortality $\mu$ (red), and survivorship-based death $\bar{d}$ (black) as a function of adult mass.}\label{fig:ratescomp}}
+\end{figure}
+
+We are interested in the specific death rate of the form $\dot{F}=-dF$, and using the derivative of Equation \ref{gompertz} we find that $d=c_{0}e^{c_{1}t}$.
+Our model considers the average rates over a population and lifecycle and the average death rate is given by
+\begin{eqnarray}
+\bar{d}&=&\frac{1}{t_{\text{exp}}}\int_{0}^{t_{\text{exp}}}c_{0}e^{c_{1}t} dt \\
+&=&\frac{c_{0}\left(e^{c_{1} t_{\text{exp}}}-1\right)}{c_{1}t_{\text{exp}}}
+\end{eqnarray}
+where $t_{\text{exp}}$ is the expected lifespan following the allometry of $t_{\text{exp}}=a_{2}M^{b_{2}}$ with $a_{2}=4.04\times10^{6}$ (s g$^{-b_{2}}$) and $b_{2}=0.30$ ~\cite{damuth1982analysis,calder1984}.
+Given the allometries above we have that
+\begin{equation}
+\bar{d}=\frac{a_{0} \left(e^{a_{1}a_{2}M^{b_{1}+b_{2}}}-1\right) M^{b_{0}-b_{1}-b_{2}}}{a_{1} a_{2}}
+\end{equation}
+which scales roughly like $M^{b_{0}}$ because $b_{1}$ and $b_{2}$ are close in value but opposite in sign.
+In Figure S\ref{fig:ratescomp} we compare the value of $\bar{d}$ to the reproductive, $\lambda$, and starvation-based mortality, $\mu$, rates.
+The values of $\bar{d}$ are orders of magnitude smaller than these rates for all mammalian masses, and thus, adding this non-starvation based death rate to our model does not shift our results within numerical confidence.
+
+{\bf NSM and the energy equivalence hypothesis}
+
+The energy equivalence hypothesis is based on the observation that if one assumes that the total metabolism of an ecosystem $B_{\rm tot}$ is equally partitioned between all species ($B_{i}$, the total metabolism of one species, is a constant), then the abundances should follow $N\left(M\right)B\left(M\right)=B_{i}$ implying that $N\left(M\right)\propto M^{-\eta}$, where $\eta$ is the metabolic scaling exponent \citep{allen2002,enquist1998}.
+As $\eta \approx 3/4$ this hypothesis is consistent with Damuth's law \citep{allen2002}.
+However, the actual equivalence of energy usage of diverse species has not been measured at the population level for a variety of whole populations.
+Figure S\ref{fig:equivalence} recasts the results of the NSM in terms of this hypothesis and shows that $F^{*}B$ is nearly constant over the same range of mammalian sizes up to the asymptotic behavior for the largest terrestrial mammals.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.4\textwidth]{fig_FPenergyequiv.eps}
+\caption{\small{ Total energetic use $B_{\rm tot}$ of consumer populations at the steady state as a function of body mass ($F^*$ is shown in green and $H^*$ in orange).
+The data are from Damuth \citep{Damuth:1987kr} and have been converted to
+  total population metabolism using the allometric relationships for
+  metabolic rate (e.g. Refs.~\citep{West:2001bv,hou,moses2008rmo}).}\label{fig:equivalence}}
+\end{figure}
+
+{\bf Application of NSM limits to aquatic mammals}
+A theoretical upper bound on mammalian body size is given by $\epsilon_\sigma=0$, where mammals are entirely composed of metabolic reserves, and this occurs at $M=8.3\times 10^8$ (g), or $120$ times the mass of a male African elephant.
+We note this particular limit as it may have future relevance to considerations of the ultimate constraints on aquatic mammals.
+
+\def\bibfont{\footnotesize}
+
+% \bibliography{aa_starving_supplement}
+\begin{thebibliography}{10}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{Kempes:2012hy}
+\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Dutkiewicz, S.} \&
+  \bibinfo{author}{Follows, M.~J.}
+\newblock \bibinfo{title}{{Growth, metabolic partitioning, and the size of
+  microorganisms.}}
+\newblock \emph{\bibinfo{journal}{PNAS}} \textbf{\bibinfo{volume}{109}},
+  \bibinfo{pages}{495--500} (\bibinfo{year}{2012}).
+
+\bibitem{kempes2014morphological}
+\bibinfo{author}{Kempes, C.~P.}, \bibinfo{author}{Okegbe, C.},
+  \bibinfo{author}{Mears-Clarke, Z.}, \bibinfo{author}{Follows, M.~J.} \&
+  \bibinfo{author}{Dietrich, L.~E.}
+\newblock \bibinfo{title}{Morphological optimization for access to dual
+  oxidants in biofilms}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{111}}, \bibinfo{pages}{208--213}
+  (\bibinfo{year}{2014}).
+
+\bibitem{West:2001bv}
+\bibinfo{author}{West, G.~B.}, \bibinfo{author}{Brown, J.~H.} \&
+  \bibinfo{author}{Enquist, B.~J.}
+\newblock \bibinfo{title}{{A general model for ontogenetic growth}}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{413}},
+  \bibinfo{pages}{628--631} (\bibinfo{year}{2001}).
+
+\bibitem{moses2008rmo}
+\bibinfo{author}{Moses, M.~E.} \emph{et~al.}
+\newblock \bibinfo{title}{{Revisiting a model of ontogenetic growth: Estimating
+  model parameters from theory and data}}.
+\newblock
+  \emph{\bibinfo{journal}{http://dx.doi.org.proxy.lib.sfu.ca/10.1086/679735}}
+  \textbf{\bibinfo{volume}{171}}, \bibinfo{pages}{632--645}
+  (\bibinfo{year}{2008}).
+
+\bibitem{hou}
+\bibinfo{author}{Hou, C.} \emph{et~al.}
+\newblock \bibinfo{title}{{Energy uptake and allocation during ontogeny}}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{322}},
+  \bibinfo{pages}{736--739} (\bibinfo{year}{2008}).
+
+\bibitem{pirt}
+\bibinfo{author}{Pirt, S.}
+\newblock \bibinfo{title}{The maintenance energy of bacteria in growing
+  cultures}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the Royal Society of London B:
+  Biological Sciences}} \textbf{\bibinfo{volume}{163}},
+  \bibinfo{pages}{224--231} (\bibinfo{year}{1965}).
+
+\bibitem{Heijnen}
+\bibinfo{author}{Heijnen, J.} \& \bibinfo{author}{Roels, J.}
+\newblock \bibinfo{title}{A macroscopic model describing yield and maintenance
+  relationships in aerobic fermentation processes}.
+\newblock \emph{\bibinfo{journal}{Biotechnology and Bioengineering}}
+  \textbf{\bibinfo{volume}{23}}, \bibinfo{pages}{739--763}
+  (\bibinfo{year}{1981}).
+
+\bibitem{peters1986ecological}
+\bibinfo{author}{Peters, R.~H.}
+\newblock \emph{\bibinfo{title}{The Ecological Implications of Body Size}},
+  vol.~\bibinfo{volume}{2} (\bibinfo{publisher}{Cambridge University Press},
+  \bibinfo{address}{Cambridge}, \bibinfo{year}{1986}).
+
+\bibitem{blueweiss1978relationships}
+\bibinfo{author}{Blueweiss, L.} \emph{et~al.}
+\newblock \bibinfo{title}{Relationships between body size and some life history
+  parameters}.
+\newblock \emph{\bibinfo{journal}{Oecologia}} \textbf{\bibinfo{volume}{37}},
+  \bibinfo{pages}{257--272} (\bibinfo{year}{1978}).
+
+\bibitem{stryer}
+\bibinfo{author}{Stryer, L.}
+\newblock \emph{\bibinfo{title}{{Biochemistry, Fourth Edition}}}
+  (\bibinfo{publisher}{W.H. Freeman and Company}, \bibinfo{address}{New York},
+  \bibinfo{year}{1995}).
+
+\bibitem{Dunbrack:1993ec}
+\bibinfo{author}{Dunbrack, R.~L.} \& \bibinfo{author}{Ramsay, M.~A.}
+\newblock \bibinfo{title}{{The Allometry of Mammalian Adaptations to Seasonal
+  Environments: A Critique of the Fasting Endurance Hypothesis}}.
+\newblock \emph{\bibinfo{journal}{Oikos}} \textbf{\bibinfo{volume}{66}},
+  \bibinfo{pages}{336--342} (\bibinfo{year}{1993}).
+
+\bibitem{Lindstedt:1985hm}
+\bibinfo{author}{Lindstedt, S.~L.} \& \bibinfo{author}{Boyce, M.~S.}
+\newblock \bibinfo{title}{{Seasonality, Fasting Endurance, and Body Size in
+  Mammals}}.
+\newblock \emph{\bibinfo{journal}{Am. Nat.}} \textbf{\bibinfo{volume}{125}},
+  \bibinfo{pages}{873--878} (\bibinfo{year}{1985}).
+
+\bibitem{Lindstedt:2002td}
+\bibinfo{author}{Lindstedt, S.~L.} \& \bibinfo{author}{Schaeffer, P.~J.}
+\newblock \bibinfo{title}{{Use of allometry in predicting anatomical and
+  physiological parameters of mammals.}}
+\newblock \emph{\bibinfo{journal}{Lab. Anim.}} \textbf{\bibinfo{volume}{36}},
+  \bibinfo{pages}{1--19} (\bibinfo{year}{2002}).
+
+\bibitem{estermann}
+\bibinfo{author}{Estermann, B.~L.}, \bibinfo{author}{Wettstein, H.-R.},
+  \bibinfo{author}{Sutter, F.} \& \bibinfo{author}{Kreuzer, M.}
+\newblock \bibinfo{title}{Nutrient and energy conversion of grass-fed dairy and
+  suckler beef cattle kept indoors and on high altitude pasture}.
+\newblock \emph{\bibinfo{journal}{Animal Research}}
+  \textbf{\bibinfo{volume}{50}}, \bibinfo{pages}{477--493}
+  (\bibinfo{year}{2001}).
+
+\bibitem{michaletz2014convergence}
+\bibinfo{author}{Michaletz, S.~T.}, \bibinfo{author}{Cheng, D.},
+  \bibinfo{author}{Kerkhoff, A.~J.} \& \bibinfo{author}{Enquist, B.~J.}
+\newblock \bibinfo{title}{Convergence of terrestrial plant production across
+  global climate gradients}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{512}},
+  \bibinfo{pages}{39--43} (\bibinfo{year}{2014}).
+
+\bibitem{damuth1987interspecific}
+\bibinfo{author}{Damuth, J.}
+\newblock \bibinfo{title}{Interspecific allometry of population density in
+  mammals and other animals: the independence of body mass and population
+  energy-use}.
+\newblock \emph{\bibinfo{journal}{Biological Journal of the Linnean Society}}
+  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
+  (\bibinfo{year}{1987}).
+
+\bibitem{calder1984}
+\bibinfo{author}{Calder, W.~A.}
+\newblock \emph{\bibinfo{title}{Size, function, and life history}}
+  (\bibinfo{publisher}{Harvard University Press}, \bibinfo{year}{1984}).
+
+\bibitem{damuth1982analysis}
+\bibinfo{author}{Damuth, J.}
+\newblock \bibinfo{title}{Analysis of the preservation of community structure
+  in assemblages of fossil mammals}.
+\newblock \emph{\bibinfo{journal}{Paleobiology}} \textbf{\bibinfo{volume}{8}},
+  \bibinfo{pages}{434--446} (\bibinfo{year}{1982}).
+
+\bibitem{allen2002}
+\bibinfo{author}{Allen, A.~P.}, \bibinfo{author}{Brown, J.~H.} \&
+  \bibinfo{author}{Gillooly, J.~F.}
+\newblock \bibinfo{title}{{Global biodiversity, biochemical kinetics, and the
+  energetic-equivalence rule}}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{297}},
+  \bibinfo{pages}{1545--1548} (\bibinfo{year}{2002}).
+
+\bibitem{enquist1998}
+\bibinfo{author}{Enquist, B.~J.}, \bibinfo{author}{Brown, J.~H.} \&
+  \bibinfo{author}{West, G.~B.}
+\newblock \bibinfo{title}{{Allometric scaling of plant energetics and
+  population density}}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{395}},
+  \bibinfo{pages}{163--165} (\bibinfo{year}{1998}).
+
+\bibitem{Damuth:1987kr}
+\bibinfo{author}{Damuth, J.}
+\newblock \bibinfo{title}{{Interspecific allometry of population density in
+  mammals and other animals: the independence of body mass and population
+  energy-use}}.
+\newblock \emph{\bibinfo{journal}{Biol. J. Linn. Soc.}}
+  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{193--246}
+  (\bibinfo{year}{1987}).
+
+\end{thebibliography}
 
